### 2016-02-24 deal84.tex
initial version of 8.4 paper

\documentclass{ansarticle-preprint}
%\usepackage{ucs}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
%\usepackage{cite}
\usepackage{anslistings}

\usepackage{fontenc}
\usepackage{graphicx}

\hypersetup{
  pdfauthor={Wolfgang Bangerth, Timo Heister, Luca Heltai, Guido Kanschat,
    Martin Kronbichler, Matthias Maier, Bruno Turcksin
  %, Toby D. Young
  },
  pdftitle={The deal.II Library, Version 8.4.0, 2016},
}

\newcommand{\specialword}[1]{\texttt{#1}}
\newcommand{\dealii}{{\specialword{deal.II}}}
\newcommand{\pfrst}{{\specialword{p4est}}}
\newcommand{\trilinos}{{\specialword{Trilinos}}}
\newcommand{\aspect}{\specialword{Aspect}}
\newcommand{\petsc}{\specialword{PETSc}}
\newcommand{\cmake}{{\specialword{CMake}}}
\newcommand{\autoconf}{{\specialword{autoconf}}}

\title{The \dealii{} Library, Version 8.4}
\author[1]{Wolfgang Bangerth}
\affil[1]{Department of Mathematics, Texas A\&M University, College Station,
    TX 77843, USA,
    {\texttt{bangerth@math.tamu.edu}}}
\author[2]{Denis Davydov}
\affil[2]{???,
  {\texttt{???}}}
\author[3]{Timo Heister}
\affil[3]{Mathematical Sciences,
  O-110 Martin Hall.
  Clemson University.
  Clemson, SC 29634, USA,
  {\texttt{heister@clemson.edu}}}
\author[4]{Luca Heltai}
\affil[4]{SISSA - International School for Advanced Studies, Via
  Bonomea 265, 34136 Trieste, Italy,
  {\texttt{luca.heltai@sissa.it}}}
\author[5]{Guido Kanschat}
\affil[5]{Interdisciplinary Center for Scientific Computing (IWR),
  Universit{\"a}t Heidelberg, Im Neuenheimer Feld 368, 69120 Heidelberg, Germany,
  {\texttt{kanschat@uni-heidelberg.de}}}
\author[6]{Martin Kronbichler}
\affil[6]{Institute for Computational Mechanics, Technische
  Universit{\"a}t M{\"u}nchen, Boltzmannstr.~15, 85748 Garching b. Mænchen,
  Germany,
  {\texttt{kronbichler@lnm.mw.tum.de}}}
\author[7]{Matthias Maier}
\affil[7]{School of Mathematics, University of Minnesota, 127 Vincent Hall,
  206 Church Street SE, Minneapolis, MN 55455, USA,
  {\texttt{msmaier@umn.edu}}}
\author[8]{Bruno Turcksin}
\affil[8]{Department of Mathematics, Texas A\&M University, College Station,
  TX 77843, USA,
  {\texttt{turcksin@math.tamu.edu}}}
\author[9]{David Wells}
\affil[9]{Department of Mathematical Sciences, Rensselaer Polytechnic Institute,
Troy, NY 12180, USA
  {\texttt{wellsd2@rpi.edu}}}
%\author[8]{Toby~D.~Young}
%\affil[8]{Institute of Fundamental Technological Research of the Polish
%  Academy of Sciences, ul. Pawi{\'n}skiego 5b, Warsaw 02-106, Poland,
%  {\texttt{tyoung@ippt.pan.pl}}}

\renewcommand{\labelitemi}{--}

\begin{document}
\maketitle

\begin{abstract}
  This paper provides an overview of the new features of the finite element
  library \dealii{} version 8.4.0.
\end{abstract}

\section{Overview}

\dealii{} version 8.4.0 was released February 29, 2016. This paper provides an
overview of the new features of this release and serves as a citable
reference for the \dealii{} software library version 8.4.0. \dealii{} is an
object-oriented finite element library used around the world in the
development of finite element solvers. It is available for free under the
GNU Lesser General Public License (LGPL) from the \dealii{} homepage at
\url{http://www.dealii.org/}.

The major changes of this release are:
\begin{itemize}
\item Preliminary support for parallel, adaptive, geometric multigrid
\item Parallel triangulations can now be partitioned in ways that allow
  weighting cells differently
\item Improved support for mixed-type arithmetic throughout the library
\item A new triangulation type that supports parallel computations
  but ensures that the entire mesh is available on every processor
\item An implementation of the Rannacher-Turek element, as well as an
  element that extends the usual Q(p) elements by bubble functions
\item Second and third derivatives of finite element fields are now
  computed exactly.
\item The various \emph{Concepts}, or requirements on template parameters
  in the library, are now consistently labeled and documented as such.
\item The interface between finite elements, quadrature, mapping, and the
  FEValues class has been rewritten. It is now much better documented
\item  More than 140 other features and bugfixes.
\end{itemize}
Some of these will be detailed in the following section.
Information on how to cite \dealii{} is provided in Section \ref{sec:cite}.

\section{Significant changes to the library}

This release of \dealii{} contains a number of large and significant changes
that will be discussed in the following sections. It of course also contains a
vast number of smaller changes and added functionality; the details of these
can be found
\href{https://www.dealii.org/8.4.0/doxygen/deal.II/changes_between_8_3_0_and_8_4.html}{in the file that lists all changes for this release}
and that is linked to from the web site of each release as well as the
release announcement.

\subsection{Preliminary support for parallel, adaptive, geometric
  multigrid}

\marginpar{Martin/Timo -- can you write this section?}
...

\subsection{Parallel triangulations can now be partitioned with weights}

Previously, partitioning a parallel mesh (represented by the
\texttt{parallel::distributed::Triangulation} class) between
processors assumed that every cell should be weighted equally. On the
other hand, the \pfrst{} library which manages the partitioning
process, allows for attaching weights for each cell and thereby
enables ways in which not the number of cells per MPI process is
equilibrated, but the sum of weights on the cells managed by each
process. \dealii{} now also supports this feature.

The implementation of this mechanism is based on a callback mechanism,
in the form of the signal-slot design pattern. User codes can register
functions that will be called upon mesh refinement and coarsening,
returning a weight for each cell. These weights will be added after
all slots (i.e., callback functions) connected to the signal have been
called.

The mechanism chosen has the advantage that all parties that use a
triangulation, for example multiple \texttt{DoFHandler} objects or a
scheme that tracks particles that are advected along with a flow field
and stored them per-cell, can indicate their computational needs for
each cell, and there no central place in a user code (other than the
triangulation itself) that has to collect these needs and forward this
information.

\subsection{Improved support for mixed-type arithmetic throughout the
  library}

When evaluating finite element fields or their derivatives at
a points $\mathbf x_q$, one typically has to do an operation of the form
\begin{align*}
  u_h(\mathbf x_q) &= \sum_{j} U_j \varphi_j(\mathbf x_q),
  \\
  \nabla u_h(\mathbf x_q) &= \sum_{j} U_j \nabla\varphi_j(\mathbf x_q).
\end{align*}
The value or derivatives of shape functions, $\varphi_j(\mathbf x_q)$
or $\nabla\varphi_j(\mathbf x_q)$ are internally evaluated by classes
derived from the \texttt{FiniteElement} and \texttt{Mapping}, and are
computed as scalars or tensors of type \texttt{double}. On the other
hand, the expansion coefficients of the field, $U_j$ are stored in
vectors of types chosen by the user; their underlying representation
can be \texttt{double}, but also \texttt{float}, \texttt{long double},
\texttt{PetscScalar}, or \texttt{std::complex<float>}. It could also
be an autodifferentiation type.

To facilitate the correct typing of the computed quantity, \dealii{}
now contains mechanisms by which one can evaluate the \textit{type} of
the product of two values, and this type is now consistently used
throughout the library in expressions such as those above. Thus, as an
example, if the user stores the expansion coefficients in a
\texttt{Vector<long double>}, then the gradients
$\nabla\varphi_j(\mathbf x_q)$ will be computed as objects of type
\texttt{Tensor<1,dim,long double>}. Likewise, if a user uses a PETSc
vector, and PETSc was configured with complex scalar types, then the
second derivatives of the solution field will be computed as
\texttt{SymmetricTensor<2,dim,PetscScalar>}, which should equal
\texttt{SymmetricTensor<2,dim,std::complex<double> >}.

These improvements make type-correct computations possible in many
places. In particular, this enables the use of complex-valued solution
vectors in many more places. On the other hand, many but not all
places have learned what actually to do with complex numbers. This is,
in particular, true for the \texttt{DataOut} class that generates an
intermediate data format that can then be written to graphical output
files for visualization. Since no format we are aware of supports
complex numbers, future versions still need to learn how to separate
real and imaginary parts of complex numbers, and output them as two
components.

\subsection{A new ``shared'' triangulation type for parallel computations}

\marginpar{Denis -- can you write this section?}
...

\subsection{Second and third derivatives of finite element fields are now
  computed exactly}

Second derivatives of solution fields, i.e., 
\begin{align*}
  \nabla^2 u_h(\mathbf x_q) &= \sum_{j} U_j \nabla^2\varphi_j(\mathbf x_q)
\end{align*}
were previously computed by finite differencing of first
derivatives. The reason for this approach is that in order to compute
the second derivatives of shape functions, $\nabla^2\varphi_j(\mathbf
x_q)$ one needs (at least) derivatives of the inverse of the Jacobian of the
mapping. This is easy to see because, for the usual $Q_p$ Lagrange
elements, one has that $\nabla\varphi_j(\mathbf
x_q)=J^{-1}\hat\nabla\hat\varphi_j(\hat{\mathbf x}_q)$, where quantities
with a hat refer to coordinates and functions on the reference cell,
and $J$ is the Jacobian of the mapping from reference to real
cell. Thus, the second derivatives satisfy
\begin{align*}
  \nabla^2\varphi_j(\mathbf x_q)
  =
  J^{-1}\hat\nabla\left[J^{-1}\hat\nabla\hat\varphi_j(\hat{\mathbf x}_q)\right]
  =
  J^{-1} J^{-1}\hat\nabla^2\hat\varphi_j(\hat{\mathbf x}_q)
  +
  J^{-1}\left(\hat\nabla J^{-1}\right)\hat\nabla\hat\varphi_j(\hat{\mathbf x}_q),
\end{align*}
where in the last expression, matrices and tensors of rank 1 and 3
have to be appropriately contracted.  Here, the difficulty lies in
computing the derivative $\hat\nabla J^{-1}$: for the usual mappings
on quadrilaterals and hexahedra, $J$ is in general a polynomial in the
reference coordinates $\hat {\mathbf x}$, so $J^{-1}$ is a rational
function. It is possible to compute the derivatives of this object,
but they are difficult and cumbersome to evaluate, especially for
higher order mappings.

The key to computing second (and higher) derivatives of shape
functions is to recognize that $\hat\nabla J^{-1}$ can be expressed
more conveniently by observing that
\begin{align*}
  0 &= \hat\nabla I
  \\
  &= \hat\nabla (JJ^{-1})
  \\
  &= J (\hat\nabla J^{-1}) + (\hat\nabla J) J^{-1},
\end{align*}
and consequently $\hat\nabla J^{-1} = - J^{-1} (\hat\nabla J)
J^{-1}$. Here, $J^{-1}$ is a matrix that is already available from
computing first derivatives, and $\hat\nabla J$ is an easily computed
rank-3 tensor with polynomial entries. Using this approach, we have
\begin{align*}
  \nabla^2\varphi_j(\mathbf x_q)
  =
  J^{-1} J^{-1}\hat\nabla^2\hat\varphi_j(\hat{\mathbf x}_q)
  -
  J^{-1} J^{-1} (\hat\nabla J)
  J^{-1}\hat\nabla\hat\varphi_j(\hat{\mathbf x}_q),
\end{align*}
again with an appropriate set of contractions over the indices of the
objects on the right.

This, and corresponding extensions to compute third derivatives, have
now been implemented in several of the finite element and mapping
classes by Maien Hamed, and are available through the
\texttt{FEValues} interface to shape functions and their derivatives.

\subsection{Incompatible changes}

\subsubsection{Revision of the interface between finite elements,
  quadratures, and mappings}

Finite element classes describe shape functions as continuous (as
opposed to discrete) objects on the reference cell. On the other hand,
in actual practice, one only needs information about shape functions
at finitely many quadrature points, and these are typically the same
on every cell in a loop over all cells. Furthermore, the evaluation of
shape functions at quadrature points then also needs to be mapped to
the real cell, typically using a polynomial mapping.

To facilitate this complex interplay, \dealii{} has three class
hierarchies with base classes \texttt{FiniteElement} (for the
description of shape functions on the reference cell),
\texttt{Quadrature} (for the locations and weights of quadrature
points on the reference cell), and \texttt{Mapping} (for the
description of mappings from the reference to the real cell). In
almost all cases, users create an object of a derived class for each
of these categories, but they never access any of the members of these
classes and instead leave this task to the \texttt{FEValues} class
(and \texttt{FEFaceValues}, \texttt{FESubfaceValues}) that presents
all one typically needs: the mapped values, gradients, and
higher order derivatives of shape functions at the quadrature points
of the real cell.

The interplay between these classes in the \texttt{FEValues} interface
is one of the oldest parts of the library. It was, at the time, not
designed based on specifications we explicitly or implicitly knew this
class had to satisfy, but instead organically grew to its current
state. These interfaces have now been fundamentally rewritten: some
functions have been replaced; several others had their argument lists
shuffled, sorted, and made more uniform; and everything has generally
been far better documented.

None of the changes in this arena is visible to the average user. The
only user codes that are affected in an incompatible way are those
that implement finite element or mapping classes.

\subsubsection{Other incompatible changes}

The
\href{https://www.dealii.org/8.4.0/doxygen/deal.II/changes_between_8_3_0_and_8_4.html}{file
  that lists all changes for this release} lists another 18
incompatible changes, but none of these should in fact be visible in
typical user codes. Some remove previously deprecated classes and
functions, and the majority change internal interfaces that are not
typically used in user codes.

\section{How to cite \dealii{}}\label{sec:cite}

In order to justify the work the developers of \dealii{} put into this
software, we ask that papers using the library reference one of the
\dealii{} papers. This helps us justify the effort we put into it.

There are various ways to reference \dealii{}. To acknowledge the use of a
particular version of the library, reference the present document. For up
to date information and bibtex snippets for this document see:

\bigskip

\begin{center}
 \url{https://www.dealii.org/publications.html}
\end{center}

\bigskip

% \begin{minipage}{0.9\textwidth}%no page break in here please
% \begin{verbatim}
% @article{dealII82,
%   title = {The {\tt deal.{I}{I}} Library, Version 8.2},
%   author = {W. Bangerth and T. Heister and L. Heltai
%    and G. Kanschat and M. Kronbichler and M. Maier
%    and B. Turcksin and T. D. Young},
%   journal = {preprint},
%   year = {2015},
% }
% \end{verbatim} %  journal = {arXiv preprint \url{http://arxiv.org/abs/TODO}},
% \end{minipage}

The original \texttt{\dealii{}} paper containing an overview of its
architecture is \cite{BangerthHartmannKanschat2007}. If you rely on specific
features of the library, please consider citing any of the following:
\begin{itemize}
 \item For geometric multigrid: \cite{Kanschat2004,JanssenKanschat2011};
 \item For distributed parallel computing: \cite{BangerthBursteddeHeisterKronbichler11};
 \item For $hp$ adaptivity: \cite{BangerthKayserHerold2007};
 \item For matrix-free and fast assembly techniques:
   \cite{KronbichlerKormann2012};
 \item For computations on lower-dimensional manifolds:
   \cite{DeSimoneHeltaiManigrasso2009};
 \item For integration with CAD files and tools:
   \cite{HeltaiMola2015};
 \item For \texttt{LinearOperator} and \texttt{PackagedOperation} facilities:
   \cite{MaierBardelloniHeltai-2015-a}.
 \item For uses of the \texttt{WorkStream} interface:
   \cite{TKB16}.
\end{itemize}

\dealii{} can interface with many other libraries:
\begin{itemize}
\item ARPACK \cite{arpack}
\item BLAS, LAPACK
\item HDF5 \cite{hdf5}
\item METIS \cite{karypis1998fast}
\item MUMPS \cite{ADE00,MUMPS:1,MUMPS:2,mumps-web-page}
\item muparser \cite{muparser-web-page}
\item NetCDF \cite{rew1990netcdf}
\item OpenCASCADE \cite{opencascade-web-page}
\item p4est \cite{p4est}
\item PETSc \cite{petsc-user-ref,petsc-web-page}
\item SLEPc \cite{Hernandez:2005:SSF}
\item Threading Building Blocks \cite{Rei07}
\item Trilinos \cite{trilinos,trilinos-web-page}
\item UMFPACK \cite{umfpack}
\end{itemize}
Please consider citing the appropriate references if you use interfaces to these
libraries.

Older releases of \dealii{} can be cited as \cite{dealII80,dealII81,dealII82,dealII83}.

\nocite{BangerthKanschat1999}

\section{Acknowledgments}

\dealii{} is a world-wide project with dozens of contributors around the
globe. Other than the authors of this paper, the following people contributed code to
this release:
%
% get this from the changes.h file using
%   egrep '^ *\(.*201[56789]' 8.3.0-vs-8.4.0.h | perl -p -e 's/201\d.*//g; s/, */\n/g; s/^ *\(?//g;' | sort | uniq
% then sort by last name and remove the authors of this paper
%
Daniel Arndt,
Mauro Bardelloni,
Alistair Bentley,
Andrea Bonito,
Claire Bruna-Rosso,
Krzysztof Bzowski,
Praveen Chandrashekar,
Conrad Clevenger,
Patrick Esser,
Rene Gassmoeller,
Arezou Ghesmati,
Maien Hamed,
Alexander Grayver,
Lukas Korous,
Aslan Kosakian,
Adam Kosik,
Konstantin Ladutenko,
Jean-Paul Pelteret,
Lei Qiao,
Gennadiy Rishin,
Angel Rodriguez,
Alberto Sartori,
Daniel Shapero,
Jason Sheldon,
Jan Stebel,
Florian Sonner,
Zhen Tao,
Heikki Virtanen,
Daniel Weygand

Their contributions are much appreciated!

\dealii{} and its developers are financially supported through a
variety of funding sources. W.~Bangerth and B.~Turcksin were partially
supported by the National Science Foundation under award OCI-1148116
as part of the Software Infrastructure for Sustained Innovation (SI2)
program; and by the Computational Infrastructure in Geodynamics initiative
(CIG), through the National Science Foundation under Award
No.~EAR-0949446 and The University of California -- Davis.

L.~Heltai was partially supported by the project OpenViewSHIP,
``Sviluppo di un ecosistema computazionale per la progettazione
idrodinamica del sistema elica-carena'', financed by Regione FVG - PAR
FSC 2007-2013, Fondo per lo Sviluppo e la Coesione, and by the project
TRIM ``Tecnologia e Ricerca Industriale per la Mobilit\`a Marina'',
CTN01-00176-163601, funded by MIUR -  Ministero dell'Istruzione,
dell'Universit\`a e della Ricerca.

T.~Heister was partially supported by the Computational Infrastructure in
Geodynamics initiative (CIG), through the National Science Foundation under Award
No. EAR-0949446 and The University of California -- Davis.

The Interdisciplinary Center for Scientific Computing (IWR) at Heidelberg University has provided
hosting services for the \dealii{} web page and the SVN archive.

%\marginpar{Everyone's funding, please}

\bibliography{deal84}{}
\bibliographystyle{abbrv}

\end{document}

### 2016-02-24 deal84.tex
update title
remove preliminary multigrid entries
as discussed with Guido and Martin
minor edits

@@
 
-\title{The \dealii{} Library, Version 8.4}
+\title{The \dealii{} Library, Version 8.4.0}
 \author[1]{Wolfgang Bangerth}
@@
 \begin{itemize}
-\item Preliminary support for parallel, adaptive, geometric multigrid
 \item Parallel triangulations can now be partitioned in ways that allow
@@
 release announcement.
-
-\subsection{Preliminary support for parallel, adaptive, geometric
-  multigrid}
-
-\marginpar{Martin/Timo -- can you write this section?}
-...
 
@@
 can be \texttt{double}, but also \texttt{float}, \texttt{long double},
-\texttt{PetscScalar}, or \texttt{std::complex<float>}. It could also
+\texttt{PetscScalar}, or \texttt{std::complex<float>} for example. It could also
 be an autodifferentiation type.
@@
 T.~Heister was partially supported by the Computational Infrastructure in
-Geodynamics initiative (CIG), through the National Science Foundation under Award
-No. EAR-0949446 and The University of California -- Davis.
+Geodynamics initiative (CIG), through the National Science Foundation under Award No. EAR-0949446 and The University of California—Davis, and National Science Foundation grant DMS1522191.
 

### 2016-02-25 deal84.tex
mention MSVC

@@
 \item Parallel triangulations can now be partitioned in ways that allow
-  weighting cells differently
-\item Improved support for mixed-type arithmetic throughout the library
+  weighting cells differently.
+\item Improved support for mixed-type arithmetic throughout the library.
 \item A new triangulation type that supports parallel computations
-  but ensures that the entire mesh is available on every processor
+  but ensures that the entire mesh is available on every processor.
 \item An implementation of the Rannacher-Turek element, as well as an
-  element that extends the usual Q(p) elements by bubble functions
+  element that extends the usual Q(p) elements by bubble functions.
 \item Second and third derivatives of finite element fields are now
@@
 \item The interface between finite elements, quadrature, mapping, and the
-  FEValues class has been rewritten. It is now much better documented
+  FEValues class has been rewritten. It is now much better documented.
+\item Initial support for compiling with Visual C++ 2013 and 2015 under Windows
+  has been added.
 \item  More than 140 other features and bugfixes.
@@
 \texttt{FEValues} interface to shape functions and their derivatives.
+
+\subsection{Visual C++ support}
+
+The library can now be compiled under Windows with Visual C++ 2013 and
+2015. The support is still experimental for the following reasons: First, we
+currently only support static linking. This will slow down linking of
+application code immensly.  Second, only a minimal testsuite is working, which
+is mainly because static linking of thousands of test executables is not
+viable. Therefore, we can not exclude the possibility of subtle bugs in the
+library.  Finally, there is of course limited support for external packages.
 

### 2016-02-28 deal84.tex
add a section on shared tria

@@
 \author[2]{Denis Davydov}
-\affil[2]{???,
-  {\texttt{???}}}
+\affil[2]{Chair of Applied Mechanics, University of
+    Erlangen-Nuremberg, Egerlandstr. 5, 91058 Erlangen, Germany,
+  {\texttt{denis.davydov@fau.de}}}
 \author[3]{Timo Heister}
@@
 
-\marginpar{Denis -- can you write this section?}
-...
+Previously, in order to use \texttt{Triangulation<dim,spacedim>} class with MPI one had to manually
+partition triangulation (e.g.
+via \texttt{METIS}), calculate and store index sets of locally owned and locally relevant degrees-of-freedom (DoF), querry \texttt{subdomain\_id}  of a cell during assembly etc.
+In order to facilitate the usage of \texttt{Triangulation} class with MPI  
+and to make its behaviour in this context consistent with \texttt{parallel::distributed::Triangulation}, a new class
+\texttt{parallel::shared::Triangulation} was introduced. 
+It extends \texttt{Triangulation} class to automatically partition triangulation when run with MPI.
+Identical functionality between \texttt{shared} and \texttt{distributed} triangulation  
+(e.g. locally owned and relevant DoFs, mpi communicator, etc)
+is grouped in the parent class \texttt{parallel::Triangulation}.
+The main difference between the two parallel triangulations is that in the case of 
+\texttt{parallel::shared::Triangulation} each process stores all cells of the triangulation.
+Consequently, by default there are no artificial cells. 
+That is, cells which are attributed to the current processor are marked as locally owned 
+(\texttt{cell->is\_locally\_owned()} returns \texttt{true}) 
+and the rest are ghost cells.
+This behaviour can be altered via an additional boolean flag provided to the constructor of the class. 
+In this case, a set of ghost cells will consist of a halo layer of cells around locally owned cells.
+Cells which are neither ghost nor locally owned are marked as artificial.
+This is consistent with the behaviour of \texttt{parallel::distributed::Triangulation},
+although in the latter case the size of the set of artificial cells will be much smaller.
+The introduction of \texttt{parallel::shared::Triangulation} class together with the 
+optional artificial cells and parent \texttt{parallel::Triangulation} class
+facilitates writing
+algorithms that are indifferent to the way triangulation is stored in the MPI context. 
+For example, \texttt{DoFTools::locally\_relevant\_dofs()} will now return a subset of 
+all DoFs for both triangulations, granted that artificial cells are enabled in 
+\texttt{parallel::shared::Triangulation}.
+Assmebly routines can now use \texttt{cell->is\_locally\_owned()} 
+for both triangulations.
+Currently, only standard (non hp) \texttt{DoFHandler} is supported.
 

### 2016-02-29 8.0/deal80.tex
Add previous release papers.

\documentclass{siamltex}

\usepackage{ucs}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{cite}

\usepackage{fontenc}
\usepackage[pdftex]{graphicx}

\usepackage[pdftex,
            pdfauthor={Wolfgang Bangerth, Timo Heister, Luca Heltai, Guido Kanschat, Martin Kronbichler, Matthias Maier, Toby D. Young},
            pdftitle={The deal.II Library, Version 8.0, 2013},
            colorlinks,linkcolor=blue,urlcolor=blue,citecolor=blue]{hyperref}

\newcommand{\specialword}[1]{\texttt{#1}}            
\newcommand{\dealii}{{\specialword{deal.II}}}
\newcommand{\pfrst}{{\specialword{p4est}}}
\newcommand{\trilinos}{{\specialword{Trilinos}}}
\newcommand{\aspect}{\specialword{Aspect}}
\newcommand{\petsc}{\specialword{PETSc}}
\newcommand{\cmake}{{\specialword{CMake}}}
\newcommand{\autoconf}{{\specialword{autoconf}}}

\title{The \dealii{} Library, Version 8.0}
\author{
  Wolfgang Bangerth%
  \thanks{Department of Mathematics, Texas A\&M University, College Station, 
    TX 77843, USA,
    \href{mailto:bangerth@math.tamu.edu}{\texttt{bangerth@math.tamu.edu}}}
  \and
  Timo Heister%
  \thanks{Mathematical Sciences,
O-110 Martin Hall.
Clemson University.
Clemson, SC 29634, USA,
    \href{mailto:heister@clemson.edu}{\texttt{heister@clemson.edu}}}
  \and
  Luca Heltai%
  \thanks{SISSA - International School for Advanced Studies, Via
    Bonomea 265, 34136 Trieste, Italy,
    \href{mailto:luca.heltai@sissa.it}{\texttt{luca.heltai@sissa.it}}}
  \and
  Guido Kanschat%
  \thanks{Interdisciplinary Center for Scientific Computing, Heidelberg
          University, Im Neuenheimer Feld 368, 69120 Heidelberg, Germany,
          \href{mailto:kanschat@uni-heidelberg.de}{\texttt{kanschat@uni-heidelberg.de}}}
  \and
  Martin Kronbichler%
  \thanks{Institute for Computational Mechanics, Technische Universit\"at M\"unchen, Boltzmannstr.~15, 85748 Garching b. M\"unchen, Germany, \href{mailto:kronbichler@lnm.mw.tum.de}{\texttt{kronbichler@lnm.mw.tum.de}}}
  \and
  Matthias Maier%
  \thanks{Institute of Applied Mathematics, Heidelberg
          University, Im Neuenheimer Feld 293/294, 69120 Heidelberg, Germany,
          \href{mailto:matthias.maier@iwr.uni-heidelberg.de}{\texttt{matthias.maier@iwr.uni-heidelberg.de}}}
  \and
  Bruno Turcksin%
  \thanks{Department of Mathematics, Texas A\&M University, College Station, 
    TX 77843, USA,
    \href{mailto:turcksin@math.tamu.edu}{\texttt{turcksin@math.tamu.edu}}}
  \and
  Toby~D.~Young%
  \thanks{Institute of Fundamental Technological Research of the Polish Academy of Sciences, ul. Pawi\'nskiego 5b, Warsaw 02-106, Poland,
    \href{mailto:tyoung@ippt.pan.pl}{\texttt{tyoung@ippt.pan.pl}}}
}

\renewcommand{\labelitemi}{--}

\begin{document}
\maketitle

\begin{abstract}
  This paper provides an overview of the new features of the finite element library \dealii{} version 8.0.
\end{abstract}

\section{Overview}

\dealii{} version 8.0 was released July 24, 2013. This paper provides an
overview of the new features of this release and serves as a citable web
reference for the \dealii{} software library version 8.0. \dealii{} is an
object-oriented finite element library used around the world in the
development of finite element solvers. It is available for free under the
GNU Lesser General Public License (LGPL) from the \dealii{} homepage at
\url{http://www.dealii.org/}.

Version 8.0 is a major release. It has numerous significant features along
with the usual set of bug fixes and documentation updates. In particular,
it has the following noteworthy large changes:
\begin{itemize}
  \item The configuration and build system has been switched to \cmake{},
    providing better support for a wide variety of platforms, better
    integration with IDEs such as Eclipse, and many other advantages. See Section \ref{sec:cmake}.

\item \dealii{} now supports 64-bit integers for degrees of freedom indices for
  problems with more than 2 billion unknowns and has been tested on problems
  of up to 27 billion unknowns. See Section \ref{sec:64bit}.

\item \dealii{} is now licensed under the GNU Lesser General Public License
  version 2.1 or later (LGPL-2.1+), see Section \ref{sec:license} for details.

\item There is the usual set of dozens or hundreds of new small features and
  bugfixes, some of which are described in \ref{sec:other}.
\end{itemize}
Information on how to cite \dealii{} is provided in Section \ref{sec:cite}.

\section{Important Changes}

\subsection{Build System}\label{sec:cmake}

With the release of version 8.0 and after more than 12 years with an
\autoconf{} and custom makefiles setup, \dealii{} switched its build system
over to \cmake{}. This is a major rewrite which changes over $25,000$ lines
of code (10,000 insertions and 15,000 deletions) together with a major
reorganization and cleanup of the code base. The rewrite addressed quite a
number of shortcomings with the old build system, that made a rewrite
highly desirable (if not necessary):
\begin{itemize}
  \item Over time, with the addition of an increasing number of external
    libraries, \dealii{} can optionally interface with, the configuration
    happened to be highly heterogeneous---almost every external dependency
    had to be set up with different command line options with completely
    different internal setup logic in the \autoconf{} files.
  \item Nowadays, \dealii{} is used in highly different environments
    ranging from installations on laptops to computing clusters. The old
    build system was not suited very well for quite a number of them,
    e.\,g. distributional needs or static linkage.
  \item The old build system strongly depended on the availability of a
    \textsc{UNIX} or \textsc{GNU/Linux} environment due to the fact that
    \textsc{GNU Make}, \textsc{Bash} and \text{Perl} were mandatory build
    time dependencies,
    which made support for other platforms (and especially Microsoft
    Windows) difficult.
\end{itemize}
Arguably, it would have been possible to address all this shortcomings in
the current build system, or by changing to another alternative, but there
are a number of reasons that make \cmake{} a particularly good choice as
build system for \dealii{}:
\begin{itemize}
  \item \cmake{} readily provides elaborate support for dependency
    resolution, incremental rebuilds and parallel builds, which were
    previously implemented mainly by hand.
  \item \cmake{} is a meta build system. It provides support for
    \emph{native toolchains} for \textsc{Linux}, \textsc{BSD},
    \textsc{Darwin}, and \textsc{Windows} platforms with minimal external
    dependencies, which drastically increases portability.
  \item \cmake{} supports a variety of commonly used IDEs for different
    platforms such as \textsc{Eclipse}, \textsc{KDevelop}, \textsc{Xcode},
    \textsc{CodeBlocks} or \textsc{MSVC} where it can directly produce a
    corresponding project.
\end{itemize}
The fact that the new build system was rewritten from scratch, offered the
possibility to reformulate and respect some major design criteria. Amongst
them are
\begin{itemize}
  \item Full support for all 16 external dependencies in various
    setups---either compiled and/or installed by the user, or provided by
    software distributions such as Linux distributions or Mac Ports.
  \item Feature auto detection to ease the setup for the average user,
    while at the same time providing full user override of configuration
    options to support almost any setup case.
  \item Be (almost) readily packageable for major distributions (e.\,g.
    Debian Linux, or Ubuntu).
  \item Provide a project configuration so that user projects based on
    \dealii{} only need a minimalistic \textsc{CMakeLists.txt} file on
    client side.
\end{itemize}

\subsection{64-bit Indices}\label{sec:64bit}

Up to version 7.3 of \dealii{}, the indices of degrees of freedom (which
are uniquely defined) were stored using the \emph{unsigned int} datatype. This
limited the total number of degrees of freedom in a computation to about four billion. When \dealii{}
was used in conjunction with \petsc{} or \trilinos{}, this number was in
fact only half as large since both of these libraries use \emph{signed
integer} indices (leaving only 31 bits for the indices). Furthermore, because an individual processor needs at
least on the order of $10^5$ unknowns to have a suitable amount of work,
\dealii{} could not scale to more than about $\frac{2\cdot 10^9}{10^5}=20,000$
processors. Even though this number may seem large, an increasing number of
supercomputers offer a much larger number of processors -- the fastest
supercomputers today are almost one hundred times larger. It is also
important to note that using simulations with more than two billions of
unknowns can be run using only a couple of thousand processors. This number
is easily reached by clusters at typical universities in the West.

To increase the number of unknowns that can be solved using \dealii{}, the
degrees of freedom can now be stored using \emph{unsigned long long int} (a
data type typically 64 bit wide) and the new 64 bit capabilities of Epetra
(Trilinos 11.2 and later) are employed (PETSc could already be compiled to use 64
bit indices for a long time). The new version of \dealii{} is completely
backward compatible with the previous version and can be compiled to use 32
or 64 bit indices indifferently. This was done by using a typedef
\emph{types::global\_dof\_index} to switch between \emph{unsigned int} and
\emph{unsigned long long int} indices. The switch between 32 and 64 bit
indices is done using the cmake option \hbox{\tt
  -DDEAL\_II\_WITH\_64BIT\_INDICES=OFF/ON}. By default, this option is turned
off to keep memory consumption at moderate levels.

Using 64 bit indices only makes sense for parallel computations for which
\dealii{} relies on either \petsc{} or \trilinos{}.  To use the 64 bit indices
with the former, \petsc{} must be configured with {\tt
  --with-64-bit-indices}. For the latter, \trilinos{} 11.2 or later must be
used (by the default \trilinos{} is compiled with 32 and 64 bit capabilities) and
uses the facilities laid out in \cite{trilinos64}.

To write a code compatible for 32 and 64 bit indices,
\emph{types::global\_dof\_index} needs to be used in all places where the
global indices of degrees of freedom are referenced. There are of course many
of these places: the library contained some 20,000 occurrences of the text
\texttt{unsigned int}, all of which needed to be examined, and some 1,600
places had to be modified. However, the library now compiles and runs with 64
bit indices and the adapted version of the step-40 tutorial program was run on
3600 processors; this computation had 1.47 billions of active cells, 5.898
billions of degrees of freedom, and was solved in 771 seconds.

\subsection{License Change}\label{sec:license}

\dealii{} used to be distributed under the Q Public License (QPL), a
license that was popular in the late 1990s mainly due to its utilization by
the Qt Application Framework.

The QPL, as many other open source licenses, puts restrictions on any use;
in particular, programs that are developed to use \dealii{} needed to be
open source as well. In practice, this puts some restrictions on commercial
use because companies are unwilling to engage in a model where they are
required to deliver their source code to customers who may then distribute
it further.

This license was chosen in the early days of \dealii{} out of a feeling that
whoever uses the library should be required to show the same open source
spirit. Abstractly, many will probably agree with this sentiment. However, in
more than 10 years of reality, it has only led to companies walking away from
using \dealii{} for commercial products. As a consequence, the project has
seen no gain, only losses from this choice of license.

We have therefore decided to change the license under which \dealii{} is
distributed to the GNU Lesser General Public License (LGPL), version 2.1 or
later at the discretion of the user. Among the advantages of this choice
are:
\begin{itemize}
  \item The LGPL allows proprietary software to dynamically link against
    \dealii{} without the need to redistribute this software under a
    restrictive license. On the other hand, the LGPL allows us to leave the
    full ``copyleft'' of the library itself intact.
  \item It avoids licensing incompatibilities with (external) dependencies
    that are either licensed under (L)GPL or a BSD style license---which
    would otherwise make any binary distribution of \dealii{} impossible,
    for example for inclusion in the package management system of many
    operating systems. This is especially important given the large number
    of external packages \dealii{} interacts with these days.
\end{itemize}
A further account of these considerations can be found in \cite{BH13}.

\subsection{Other Changes}\label{sec:other}

The \dealii{} release 8.0 also includes improvements in the following areas:

\begin{itemize}
\item Improvements to \dealii{} threading support: The WorkStream class used
  in \dealii{} to parallelize assembly loops now uses thread local memory,
  which gives considerably better utilization of caches on multi-core
  architectures. In addition, a few more linear algebra functions have been
  parallelized with threads. The new version also allows to directly limit the
  number of threads used by the task scheduler in Threading Building Blocks at
  run time.
\item \dealii{}'s own iterative solvers have been revised. CG and BiCGStab now
  use less auxiliary vectors, and GMRES avoids unnecessary work due to
  re-orthogonalization when it is not necessary. This makes behavior more
  similar to external linear algebra in PETSc and Trilinos (but those remain
  better adapted to large-scale parallel computers by reducing global
  communication).
\item The interfaces to parallel linear algebra objects of PETSc and Trilinos
  objects have been further unified to allow for easily exchanging one variant
  with the other.
\item The generic template instantiation mechanism in \dealii{} has been
  applied to several new places to pre-compile code for all vector types in
  ErrorEstimator, VectorTools, FETools, ConstraintMatrix.
\end{itemize}

\section{How to cite \dealii{}}\label{sec:cite}

In order to justify the work the developers of \dealii{} put into this
software, we ask that papers using the library reference one of the \dealii{}
papers. This helps us justify the effort we put into it.

There are various ways to reference \dealii{}. To acknowledge the use of a
particular version of the library, reference the present document as follows:
\begin{verbatim}
@article{dealII80,
  title = {The {\tt deal.{I}{I}} Library, Version 8.0},
  author = {W. Bangerth and T. Heister and L. Heltai and G. Kanschat
   and M. Kronbichler and M. Maier and B. Turcksin and T. D. Young},
  journal = {arXiv preprint \url{http://arxiv.org/abs/1312.2266v3}},
  year = {2013},
}
\end{verbatim}

The original \texttt{\dealii{}} paper containing an overview of its
architecture is \cite{BangerthHartmannKanschat2007}. If you rely on specific
features of the library, please consider citing any of the following:
\begin{itemize}
 \item For geometric multigrid: \cite{Kanschat2004,JanssenKanschat2011};
 \item For distributed parallel computing: \cite{BangerthBursteddeHeisterKronbichler11};
 \item For $hp$ adaptivity: \cite{BangerthKayserHerold2007};
 \item For matrix-free and fast assembly techniques:
   \cite{KronbichlerKormann2012};
 \item For computations on lower-dimensional manifolds:
   \cite{DeSimoneHeltaiManigrasso2009}.
\end{itemize}

\nocite{BangerthKanschat1999}

\section{Acknowledgements}

\dealii{} is a world-wide project with dozens of contributors around the
globe. Other than the authors of this paper, the following people contributed code to
this release: 
  Fahad Alrashed,
  Juan Carlos Araujo Cabarcas,
  Denis Davydov,
  Ashkan Dorostkar,
  J{\"o}rg Frohne,
  Felix Gruber,
  Scott Miller,
  Spencer Patty,
  Martin Steigemann,
  Kainan Wang,
  Christian W{\"u}lker.
Their contributions are much appreciated!

\dealii{} and its developers are financially supported through a
variety of funding sources. W.~Bangerth and B.~Turcksin were partially
supported by the National Science Foundation under award OCI-1148116
as part of the Software Infrastructure for Sustained Innovation (SI2)
program; by the Computational Infrastructure in Geodynamics initiative
(CIG), through the National Science Foundation under Award
No.~EAR-0949446 and The University of California -- Davis; and through
Award No.~KUS-C1-016-04, made by King Abdullah University of Science
and Technology (KAUST).

L.~Heltai was partially supported by the project OpenSHIP,
``Simulazioni di fluidodinamica computazionale (CFD) di alta qualita
per le previsioni di prestazioni idrodinamiche del sistema
carena-elica in ambiente OpenSOURCE'', financed by Regione FVG - POR
FESR 2007–2013 Obiettivo competitivita regionale e occupazione.

T.~Heister was partially supported by the Computational Infrastructure in
Geodynamics initiative (CIG), through the National Science Foundation under Award
No. EAR-0949446 and The University of California -- Davis and through Award No. KUS-
C1-016-04, made by King Abdullah University of Science and Technology (KAUST).

\bibliography{deal80}{}
\bibliographystyle{plain}

\end{document}

### 2016-02-29 8.1/deal81.tex
Add previous release papers.

\documentclass{siamltex}

\usepackage{ucs}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{cite}

\usepackage{fontenc}
\usepackage[pdftex]{graphicx}

\usepackage[pdftex,
            pdfauthor={Wolfgang Bangerth, Timo Heister, Luca Heltai, Guido Kanschat, Martin Kronbichler, Matthias Maier, Toby D. Young},
            pdftitle={The deal.II Library, Version 8.1, 2013},
            colorlinks,linkcolor=blue,urlcolor=blue,citecolor=blue]{hyperref}

\newcommand{\specialword}[1]{\texttt{#1}}
\newcommand{\dealii}{{\specialword{deal.II}}}
\newcommand{\pfrst}{{\specialword{p4est}}}
\newcommand{\trilinos}{{\specialword{Trilinos}}}
\newcommand{\aspect}{\specialword{Aspect}}
\newcommand{\petsc}{\specialword{PETSc}}
\newcommand{\cmake}{{\specialword{CMake}}}
\newcommand{\autoconf}{{\specialword{autoconf}}}

\title{The \dealii{} Library, Version 8.1}
\author{
  Wolfgang Bangerth%
  \thanks{Department of Mathematics, Texas A\&M University, College Station, 
    TX 77843, USA,
    \href{mailto:bangerth@math.tamu.edu}{\texttt{bangerth@math.tamu.edu}}}
  \and
  Timo Heister%
  \thanks{Mathematical Sciences,
O-110 Martin Hall.
Clemson University.
Clemson, SC 29634, USA,
    \href{mailto:heister@clemson.edu}{\texttt{heister@clemson.edu}}}
  \and
  Luca Heltai%
  \thanks{SISSA - International School for Advanced Studies, Via
    Bonomea 265, 34136 Trieste, Italy,
    \href{mailto:luca.heltai@sissa.it}{\texttt{luca.heltai@sissa.it}}}
  \and
  Guido Kanschat%
  \thanks{Interdisciplinary Center for Scientific Computing, Heidelberg
          University, Im Neuenheimer Feld 368, 69120 Heidelberg, Germany,
          \href{mailto:kanschat@uni-heidelberg.de}{\texttt{kanschat@uni-heidelberg.de}}}
  \and
  Martin Kronbichler%
  \thanks{Institute for Computational Mechanics, Technische Universit\"at M\"unchen, Boltzmannstr.~15, 85748 Garching b. M\"unchen, Germany, \href{mailto:kronbichler@lnm.mw.tum.de}{\texttt{kronbichler@lnm.mw.tum.de}}}
  \and
  Matthias Maier%
  \thanks{Institute of Applied Mathematics, Heidelberg
          University, Im Neuenheimer Feld 293/294, 69120 Heidelberg, Germany,
          \href{mailto:matthias.maier@iwr.uni-heidelberg.de}{\texttt{matthias.maier@iwr.uni-heidelberg.de}}}
  \and
  Bruno Turcksin%
  \thanks{Department of Mathematics, Texas A\&M University, College Station, 
    TX 77843, USA,
    \href{mailto:turcksin@math.tamu.edu}{\texttt{turcksin@math.tamu.edu}}}
  \and
  Toby~D.~Young%
  \thanks{Institute of Fundamental Technological Research of the Polish Academy of Sciences, ul. Pawi\'nskiego 5b, Warsaw 02-106, Poland,
    \href{mailto:tyoung@ippt.pan.pl}{\texttt{tyoung@ippt.pan.pl}}}
}

\renewcommand{\labelitemi}{--}

\begin{document}
\maketitle

\begin{abstract}
  This paper provides an overview of the new features of the finite element library \dealii{} version 8.1.
\end{abstract}

\section{Overview}

\dealii{} version 8.1 was released December 24, 2013. This paper provides an
overview of the new features of this release and serves as a citable
reference for the \dealii{} software library version 8.1. \dealii{} is an
object-oriented finite element library used around the world in the
development of finite element solvers. It is available for free under the
GNU Lesser General Public License (LGPL) from the \dealii{} homepage at
\url{http://www.dealii.org/}.

Version 8.1 contains,  along
with the usual set of bug fixes and documentation updates, the following noteworthy changes:
\begin{itemize}
\item Three new tutorial programs (see Section \ref{sec:step26}, \ref{sec:step42}, and \ref{sec:step51});
\item Improved support for multicore parallelization on shared memory machines (Section \ref{sec:threading});
\item The testsuite was ported to CTest/CDash (Section \ref{sec:testsuite});
\item Post-install tests (see Section \ref{sec:installtests}).
\end{itemize}
Information on how to cite \dealii{} is provided in Section \ref{sec:cite}.

\section{Changes to the governance structure}

\dealii{} has been a project with contributors from around the world for a
long time already. However, there has never been a formalized way to recognize
contributions other than by listing the authors in a file that is part of
the documentation.

Starting with this release, \dealii{} now has a more open governance structure
that we hope will more accurately reflect the extensive contributions of many
participants in this project: in addition to listing all contributors as
before, \dealii{} is now governed by a council of \textit{developers} --
currently Luca Heltai, Martin Kronbichler, Matthias Maier, Bruno Turcksin, and
Toby Young -- along with \textit{principal developers} who are responsible for
running the technical infrastructure -- currently Wolfgang
Bangerth, Timo Heister, and Guido Kanschat. It is our hope that this new
structure is a more adequate representation of many contributors' long-term
effort for this project. It is our intention that this structure remains open
to others and that this serves as motivation for others to participate!

\section{Changes to the library}

\subsection{A new tutorial: Elasto-plastic contact problem (step-42)}
\label{sec:step42}

The first of the three new tutorial programs, step-42 (written by Jörg Frohne,
Timo Heister, and Wolfgang Bangerth), shows how to solve an elasto-plastic
contact problem. The program is an extension of step-41 to a much more complex
equation (nonlinear elasto-plasticity) and also demonstrates how to compute
solutions for such problems in parallel.

The example shows how to solve a 3d, non-linear problem with a semi-smooth
Newton method coupled with an active set strategy on adaptively refined meshes
and scales well up to at least 1024 cores and 1 billion unknowns. An
accompanying paper \cite{FHB13} explains the details and design of the
algorithms behind this tutorial. It also shows scalability results for
parallel computations.

\subsection{A new tutorial: A hybridizable discontinuous Galerkin method (step-51)}
\label{sec:step51}

Step-51, written by Martin Kronbichler and Scott Miller, shows how to
implement a hybridizable discontinuous Galerkin method (HDG) in \dealii{}. An
HDG method is a special DG method which features a reduced number of
globally coupled degrees of freedom compared to other DG schemes for implicit
systems. This is achieved by
introducing a new variable for the numerical trace, i.e., the operator that
connects the solution of the subproblems on each element. The linear system to
be solved globally then only consists of degrees of freedom in the trace variable,
whereas the DG contributions interior to cells are eliminated during assembly by static
condensation. The tutorial program shows how to implement this concept in \dealii{},
where the class \texttt{FE\_FaceQ} represents the trace element for the usual
DG element \texttt{FE\_DGQ} applied to the scalar convection-diffusion
equation written as a first order system.

The tutorial program includes a few
practical aspects of an HDG implementation, namely the realization of
superconvergent post-processing and the utilization of the parallel character
of the assembly and static condensation process with the \texttt{WorkStream}
class. The tutorial also contains an extensive discussion of efficiency of
this approach compared to standard continuous finite elements. The results
show that the HDG method is competitive to continuous elements for medium
order between about 3 and 6 in terms of solution time on diffusion-dominated
systems. For convection-dominated problems, the method inherits the superior
stability properties of low to medium order DG methods, which allows for good
solution quality already on coarser meshes.

Furthermore, a new element \texttt{FE\_FaceP} has been added to the \dealii{}
library that can used for hybridization of the discontinuous elements of
complete degree $p$, \texttt{FE\_DGP}. \texttt{FE\_DGP} and \texttt{FE\_FaceP}
of degree $p$ have fewer degrees of freedom per cell compared to the elements
\texttt{FE\_DGQ} and \texttt{FE\_FaceQ} of tensor product degree $p$.

\subsection{A new tutorial: An adaptive mesh solver for the heat
  equation (step-26)}
\label{sec:step26}

\dealii{} has long had tutorial programs for the wave equation, but none for
the conceptually simpler heat equation. Furthermore, there was no simple
tutorial program that showed how one can solve time dependent problems where
the mesh changes every few time steps.

The new step-26 tutorial program closes this gap. It is a rather simple
program -- the program has just 151 lines with a semicolon -- and as such
serves as a gentle introduction to these topics.

\subsection{Improved WorkStream implementation}
\label{sec:threading}
The \texttt{WorkStream} namespace has already in the past contained the
functions that are used to parallelize (using multithreading) many of the loops over all cells or
faces that one encounters in finite element computations. Extensive
documentation on how to use this class and the rationale behind it is provided
in the ``Parallel computing with multiple processors accessing shared memory''
module.

This class has been significantly revamped in an effort to improve scaling to
large multicore workstations. In particular, we now use thread-local variables
in some places to improve cache performance. In addition, the namespace has
obtained another implementation of \texttt{WorkStream::run} that does not just
take a sequence of cells (or other objects to work on), but such a sequence
that has been colored in a way to indicate which cells will conflict with
others when writing into the global matrix, right hand side, or other
object. Using this function instead of the previously existing one obviously
requires changes to existing code bases, but can provide significant speedups
in some circumstances when using significant numbers of threads (say, above 16
or 32). The details of these algorithms have been documented in \cite{TKB13}.

As part of the changes necessary to allow this, \dealii{} has also gained a
set of generic functions that provide graph coloring algorithms. The are
located in namespace \texttt{GraphColoring}.

\subsection{Testsuite ported to CTest/CDash}
\label{sec:testsuite}

With this minor release \dealii{} is now fully ported to \cmake. The last
remaining step was the migration of the testsuite to \textsc{CTest} as test
driver and \textsc{CDash} as web front end. This involved porting over 3000
tests to a new directory structure. Furthermore, the regression and build
tests are no longer independent testsuites, but a combined one; and tests
are by default run against both the debug and release versions of the
library.

The port of the testsuite was also motivated by the fact that, in order to
provide official support for at least three major compiler brands
(\textsc{GCC}, \textsc{Clang} and \textsc{ICC}) on multiple platforms such
as Linux, MAC OSX and BSD, it is highly necessary to have regular
regression tests available for these platforms. Hence, in addition to the
regression tester already present (whose main purpose is to test every
single subversion revision) dedicated nightly build tests for a big variety
of above compilers and platforms (and external dependencies) have been set
up.

The testsuite is set up in the build directory via \texttt{make
setup\_tests}. After that, tests can be run by invoking the test driver
\texttt{ctest} with suitable options. In order to submit test results a
\texttt{CTest} script is available that can be passed as an option to the
driver, e.\,g. by \texttt{ctest -S ../tests/run\_testsuite.cmake} (or if
just the build tests should be run, \texttt{\dots/run\_buildtest.cmake}).
The scripts run a configure, build and test stage as necessary and submit
the results to a \texttt{CDash} instance. Public, anonymous submission
of test results to the \textsc{CDash} site is possible in order to
easily upload and share regression test results.

\subsection{Post-install tests}
\label{sec:installtests}

Every installation now ships with a small collection of tests that can be executed by calling \texttt{make test} in the build directory. This is used to identify common configuration problems, bugs in external dependencies, problems in the MPI wrappers, and to check the correct setup of some of the external packages.
Running the tests is now part of the installation instructions and we especially recommend to run the tests on new machines.

\subsection{Other Changes}\label{sec:other}

The \dealii{} release 8.1 also includes improvements in the following areas:

\begin{itemize}
\item The handling of periodic boundary conditions has been improved and extended
  to distributed triangulations.
\item \dealii{}'s output capabilities have been extended in several
  aspects. In particular, it is now possible to merge output from vectors
  belonging to several different \texttt{DoFHandler} objects (a common case
  when solving multiphysics problems) within a single file instead of forcing
  the user to join the data in one \texttt{DoFHandler} for output.
\item Higher order polynomial boundary descriptions now use support points of
  the Gauss-Lobatto quadrature formula instead of equidistant ones, which
  gives much better stability at high orders and more accuracy for some curved
  boundary shapes due to superconvergence effects.
\end{itemize}

\section{How to cite \dealii{}}\label{sec:cite}

In order to justify the work the developers of \dealii{} put into this
software, we ask that papers using the library reference one of the \dealii{}
papers. This helps us justify the effort we put into it.

There are various ways to reference \dealii{}. To acknowledge the use of a
particular version of the library, reference the present document as follows:
\begin{verbatim}
@article{dealII81,
  title = {The {\tt deal.{I}{I}} Library, Version 8.1},
  author = {W. Bangerth and T. Heister and L. Heltai and G. Kanschat
   and M. Kronbichler and M. Maier and B. Turcksin and T. D. Young},
  journal = {arXiv preprint \url{http://arxiv.org/abs/1312.2266v4}},
  year = {2013},
}
\end{verbatim}

The original \texttt{\dealii{}} paper containing an overview of its
architecture is \cite{BangerthHartmannKanschat2007}. If you rely on specific
features of the library, please consider citing any of the following:
\begin{itemize}
 \item For geometric multigrid: \cite{Kanschat2004,JanssenKanschat2011};
 \item For distributed parallel computing: \cite{BangerthBursteddeHeisterKronbichler11};
 \item For $hp$ adaptivity: \cite{BangerthKayserHerold2007};
 \item For matrix-free and fast assembly techniques:
   \cite{KronbichlerKormann2012};
 \item For computations on lower-dimensional manifolds:
   \cite{DeSimoneHeltaiManigrasso2009}.
\end{itemize}

\nocite{BangerthKanschat1999}

\section{Acknowledgements}

\dealii{} is a world-wide project with dozens of contributors around the
globe. Other than the authors of this paper, the following people contributed code to
this release: 
  Fahad Alrashed,
  Daniel Arndt,
  Juan Carlos Araujo Cabarcas,
  Krzyszof Bzowski,
  Francesco Cattoglio,
  Denis Davydov,
  David Emerson,
  Armin Ghajar Jazi,
  Eric Heien,
  Tobin Isaac, 
  Oleh Krehel,
  Craig Michoski,
  Scott Miller,
  Jean-Paul Pelteret,
  Andreas Putz,
  Mayank Sabharwal,
  Martin Steigemann.
Their contributions are much appreciated!

\dealii{} and its developers are financially supported through a
variety of funding sources. W.~Bangerth and B.~Turcksin were partially
supported by the National Science Foundation under award OCI-1148116
as part of the Software Infrastructure for Sustained Innovation (SI2)
program; by the Computational Infrastructure in Geodynamics initiative
(CIG), through the National Science Foundation under Award
No.~EAR-0949446 and The University of California -- Davis; and through
Award No.~KUS-C1-016-04, made by King Abdullah University of Science
and Technology (KAUST).

L.~Heltai was partially supported by the project OpenSHIP,
``Simulazioni di fluidodinamica computazionale (CFD) di alta qualit\`a
per le previsioni di prestazioni idrodinamiche del sistema
carena-elica in ambiente OpenSOURCE'', financed by Regione FVG - POR
FESR 2007–2013 Obiettivo competitivit\`a regionale e occupazione.

T.~Heister was partially supported by the Computational Infrastructure in
Geodynamics initiative (CIG), through the National Science Foundation under Award
No. EAR-0949446 and The University of California -- Davis and through Award No. KUS-
C1-016-04, made by King Abdullah University of Science and Technology (KAUST).

The Interdisciplinary Center for Scientific Computing (IWR) at Heidelberg University has provided 
hosting services for the deal.II web page and the SVN archive.

%\marginpar{Everyone's funding, please}

\bibliography{deal81}{}
\bibliographystyle{plain}

\end{document}

### 2016-02-29 8.2/deal82.tex
Add previous release papers.

\documentclass{ansarticle}
%\usepackage{ucs}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
%\usepackage{cite}
\usepackage{anslistings}

\usepackage{fontenc}
\usepackage{graphicx}

\hypersetup{
  pdfauthor={Wolfgang Bangerth, Timo Heister, Luca Heltai, Guido Kanschat, Martin Kronbichler, Matthias Maier, Bruno Turcksin, Toby D. Young},
  pdftitle={The deal.II Library, Version 8.2, 2015},
}

\newcommand{\specialword}[1]{\texttt{#1}}
\newcommand{\dealii}{{\specialword{deal.II}}}
\newcommand{\pfrst}{{\specialword{p4est}}}
\newcommand{\trilinos}{{\specialword{Trilinos}}}
\newcommand{\aspect}{\specialword{Aspect}}
\newcommand{\petsc}{\specialword{PETSc}}
\newcommand{\cmake}{{\specialword{CMake}}}
\newcommand{\autoconf}{{\specialword{autoconf}}}

\title{The \dealii{} Library, Version 8.2}
\author[1]{Wolfgang Bangerth}
\affil[1]{Department of Mathematics, Texas A\&M University, College Station, 
    TX 77843, USA,
    %\href{mailto:bangerth@math.tamu.edu}
    {\texttt{bangerth@math.tamu.edu}}}
\author[2]{Timo Heister}
\affil[2]{Mathematical Sciences,
O-110 Martin Hall.
Clemson University.
Clemson, SC 29634, USA,
%\href{mailto:heister@clemson.edu}
{\texttt{heister@clemson.edu}}}
\author[3]{Luca Heltai}
\affil[3]{SISSA - International School for Advanced Studies, Via
  Bonomea 265, 34136 Trieste, Italy,
  %\href{mailto:luca.heltai@sissa.it}
  {\texttt{luca.heltai@sissa.it}}}
\author[4]{Guido Kanschat}
\affil[4]{Interdisciplinary Center for Scientific Computing (IWR),
  Universität Heidelberg, Im Neuenheimer Feld 368, 69120 Heidelberg, Germany,
  %\href{mailto:kanschat@uni-heidelberg.de}
  {\texttt{kanschat@uni-heidelberg.de}}}
\author[5]{Martin Kronbichler}
\affil[5]{Institute for Computational Mechanics, Technische
  Universität München, Boltzmannstr.~15, 85748 Garching b. München,
  Germany,
  %\href{mailto:kronbichler@lnm.mw.tum.de}
  {\texttt{kronbichler@lnm.mw.tum.de}}}
\author[6]{Matthias Maier}
\affil[6]{Institute of Applied Mathematics, Heidelberg
  University, Im Neuenheimer Feld 293/294, 69120 Heidelberg, Germany,
  %\href{mailto:matthias.maier@iwr.uni-heidelberg.de}
  {\texttt{matthias.maier@iwr.uni-heidelberg.de}}}
\author[7]{Bruno Turcksin}
\affil[7]{Department of Mathematics, Texas A\&M University, College Station, 
  TX 77843, USA,
  %\href{mailto:turcksin@math.tamu.edu}
  {\texttt{turcksin@math.tamu.edu}}}
\author[8]{Toby~D.~Young}
\affil[8]{Institute of Fundamental Technological Research of the Polish Academy of Sciences, ul. Pawińskiego 5b, Warsaw 02-106, Poland,
  %\href{mailto:tyoung@ippt.pan.pl}
  {\texttt{tyoung@ippt.pan.pl}}}

\renewcommand{\labelitemi}{--}

\begin{document}
\maketitle

\begin{abstract}
  This paper provides an overview of the new features of the finite element library \dealii{} version 8.2.
\end{abstract}

\section{Overview}

\dealii{} version 8.2 was released January 1, 2015. This paper provides an
overview of the new features of this release and serves as a citable
reference for the \dealii{} software library version 8.2. \dealii{} is an
object-oriented finite element library used around the world in the
development of finite element solvers. It is available for free under the
GNU Lesser General Public License (LGPL) from the \dealii{} homepage at
\url{http://www.dealii.org/}.

Version 8.2 contains,  along
with the usual set of new functions, bug fixes and documentation updates, the following noteworthy changes:
\begin{itemize}
\item Comprehensive support for geometries described by arbitrary manifolds
  and meshes that respect this description not only on the boundary but
  also internally
\item Support for geometries imported from CAD using the OpenCASCADE library
\item Three new tutorial programs on complex geometries, CAD geometries, and
  time stepping methods
\item Support for users wanting to use C++11 features
\item Improvements to multithreading support
\item Vectorization of many vector operations using OpenMP SIMD directives
\end{itemize}
Some of these will be detailed in the following section.
Information on how to cite \dealii{} is provided in Section \ref{sec:cite}.

\section{Significant changes to the library}

This release of deal.II contains a number of large and significant changes
that will be discussed in the following sections. It of course also contains a
vast number of smaller changes and added functionality; the details of these
can be found in the file that lists all changes for this release and that is
linked to from the web site of each release as well as the release
announcement.

\subsection{Supporting complex geometries through manifolds}

Complex geometries can be described using the concept of a manifold, a
topological space that resembles Euclidean space near each point. In
mathematics, particularly topology, one describes a manifold using an
atlas. An atlas consists of individual charts that, roughly speaking,
describe individual regions of the manifold, by means of providing a
local coordinate system near each point. (Such coordinate systems may not
exist \textit{globally}, as in the case of the surface of a sphere, but it
must be possible to cover the entire manifold with charts that are valid in
some local neighborhood of every point.) In \dealii{} these concepts
have been formalized in the abstract base class \verb|Manifold| that
defines an interface by which the \verb|Triangulation| and
\verb|Mapping| classes can query geometric information about the
domain.

Whenever a new point is required (for example upon mesh refinement, or
when integrating over curved manifolds using higher order mappings), new points are typically
obtained by providing a local coordinate system on the manifold,
identifying existing points in the local coordinate system (pulling
them back using the local chart to obtain their local coordinates),
finding the new point in the local coordinate system by weighted sums of
the existing points, and transforming the point back to the real space
(pushing it forward using the local chart).

A new identifier (the \verb|manifold_id|) has been introduced to
describe geometrical features of complex domains: every entity of a mesh
(i.e., cells, faces, and edges) now have a \verb|manifold_id| associated with
them. This extends the existing behavior: Previous releases of
the library allowed only boundaries to be curved, and used the
\verb|boundary_id| of the faces of a triangulation to identify how to
add new points upon mesh refinement.  In contrast, the new \verb|manifold_id|
also exists for cells and interior faces. Upon refinement, the
\verb|Triangulation| class queries each object (hex, quad or line)
where new points should be placed. Attaching a custom \verb|Manifold|
to any such object ensures that new points are created consistently
with the underlying geometry.

We provide a series of classes that implement this mechanism for
commonly used geometries:
\begin{itemize}
\item \verb|FlatManifold|: in the simplest case, the
  objects that make up the \verb|Triangulation| are straight line
  segments, a bi-linear surface or a tri-linear volume. New vertices
  are then simply put into the middle of the old ones (where
  ``middle'' means a suitable average of the locations of the
  pre-existing vertices). This is the default manifold associated to
  each object of the \verb|Triangulation|;
\item \verb|SphericalManifold|: you can use this manifold object to
  describe any sphere, circle, hypersphere or hyperdisc in two or
  three dimensions, both as a co-dimension one manifold descriptor or
  as co-dimension zero manifold descriptor;
\item \verb|CylindricalManifold|: in three dimensions, points are
  transformed using a cylindrical coordinate system along an
  arbitrarily oriented cylinder.
\end{itemize}

More general specialisations can be obtained by using the
\verb|ChartManifold| partial specialisation: this class is intended to
directly implement the topological concept of a chart, by explicitly
providing \emph{pull back} and \emph{push forward} expressions. Classes
derived from \verb|ChartManifold| only have to overload the \verb|pull_back|
and \verb|push_forward| functions, while all other functions are then defined
in terms of these two functions. \verb|ChartManifold| assumes that the
manifold can be represented by a single, globally valid chart.
The
class \verb|FunctionManifold| further implements this using explicit \dealii{}
\verb|Function| classes, and the new tutorial \verb|step-53| provides a
concrete example on how to use these classes to work with complex
geometries.

\subsection{Interfacing with CAD geometries}

If you have OpenCASCADE (\url{http://www.opencascade.org/})
installed\cite{opencascade-web-page}, you can now use existing IGES or
STEP CAD files to describe the boundary of your geometry to \dealii{}.

The \dealii{} library allows the user to specify the geometry of the
analysis in one of two possible ways: i) by creating a coarse grid
internally (for simple geometries, such as boxes, cylinders, shells,
etc.) or ii) by reading a mesh input file in one of the supported
formats via the \verb|GridIn| class.

Along with the default boundary descriptor \verb|StraightBoundary|
which implements the standard behavior of \dealii{}, some elementary
boundary descriptors are also included in the library to describe
circles or spheres (\verb|HyperBallBoundary|), cylinders
(\verb|CylinderBoundary|), half circles or half spheres
(\verb|HalfHyperBallBoundary|), and so on. In general, \dealii{}
provides a curved boundary descriptor for each one of the elementary
mesh that can be created from within the library itself.

Real world geometries, however, can rarely be described by straight
lines, balls or cylinders, and much more complex structures are
required if one wants to tackle industrial problems. The industry
standard for the design of any geometry is based on Computer Aided
Design (CAD) tools, which rely on Non Uniform Rational B-Splines
(NURBS) descriptions of curves and surfaces. Virtually any object that
surrounds us has been modeled using some NURBS patch in a CAD tool.

The namespace \verb|dealii::OpenCASCADE| contains all utilities and
wrappers that are needed to manipulate CAD files. Most of the
functions in the \verb|dealii::OpenCASCADE| namespace deal with
\verb|TopoDS_Shape| objects: the default topological object of the
\verb|OpenCASCADE| library, and provide interfaces to deal.II objects,
like \verb|Triangulation|, \verb|Manifold| and \verb|Boundary|.

The new step-54 tutorial program of the \dealii{} library shows an example
application of the functions and classes in the
\verb|deal.II::OpenCASCADE| namespace. In step-54, a CAD shape and a
coarse surface triangulation are imported from external files, and the
resulting triangulation is then refined on the CAD surface using
different projection strategies. A detailed description of these
techniques is available in~\cite{HeltaiMola2015}.

\subsection{A new tutorial: Support for time stepping methods (step-52)}
The goal of this tutorial is to show how to use the different Runge--Kutta
methods of the new TimeStepping class. The Runge--Kutta methods used in this
tutorial include explicit methods, implicit methods, and embedded methods. In
step-52, the time dependent neutron diffusion equation is solved on a uniform
mesh. To compare the different time stepping methods, the solution of the
problem is chosen such that its spatial component can be exactly represented on
the uniform mesh by quadratic continuous finite elements. Moreover, at the
end of the last time step the exact solution is zero. Therefore, the different
time discretization methods can be compared by simply looking at the discretized
solutions.

\subsection{Support for some C++11 features}
The C++11 standard \cite{cpp11} provides many convenient features that can
help make writing finite element codes simpler. \dealii{} does not use any of
these features itself to ensure that it can be used with compilers that do not
implement C++11 yet.

However, this does not prevent users from using C++11
features in their own codes. Among these, range-based for loops are
particularly useful given that \dealii{} programs often contain many loops
over all cells and that writing these loops is frequently quite verbose:
\begin{c++}
  Triangulation<dim> triangulation;
  ...
  typename Triangulation<dim>::active_cell_iterator
    cell = triangulation.begin_active(),
    endc = triangulation.end();
  for (; cell!=endc; ++cell)
    cell->set_refine_flag();
\end{c++}
Code such as this can be significantly simplified by combining C++11's
range-based for loops and \texttt{auto}-typed variables. To support this,
\dealii{}'s container classes (i.e., the triangulations and DoF handler
classes) provide member functions that return objects that denote the entire
set of cells, active cells, or cells on a particular level. This allows to
write the same loop as above in the following, much shorter way:
\begin{c++}
  Triangulation<dim> triangulation;
  ...
  for (auto cell : triangulation.active_cell_iterators())
    cell->set_refine_flag();
\end{c++}
The function
\texttt{Triangulation::active\_cell\_iterators()}, 
and equivalents in the
\texttt{DoFHandler} and
\texttt{hp::DoFHandler} classes, were added in \dealii{} 8.2
to support this idiom.
Other variants of these functions provide iterator ranges
for all cells (not just the active ones) and for cells on individual
levels. A new documentation module elaborates on support for C++11 in
\dealii{}.

\section{How to cite \dealii{}}\label{sec:cite}

In order to justify the work the developers of \dealii{} put into this
software, we ask that papers using the library reference one of the \dealii{}
papers. This helps us justify the effort we put into it.

There are various ways to reference \dealii{}. To acknowledge the use of a
particular version of the library, reference the present document. For
up to date information and bibtex snippets for this document see:

\bigskip

\begin{center}
 \url{https://www.dealii.org/publications.html}
\end{center}

\bigskip

% \begin{minipage}{0.9\textwidth}%no page break in here please
% \begin{verbatim}
% @article{dealII82,
%   title = {The {\tt deal.{I}{I}} Library, Version 8.2},
%   author = {W. Bangerth and T. Heister and L. Heltai 
%    and G. Kanschat and M. Kronbichler and M. Maier
%    and B. Turcksin and T. D. Young},
%   journal = {preprint},
%   year = {2015},
% }
% \end{verbatim} %  journal = {arXiv preprint \url{http://arxiv.org/abs/TODO}},
% \end{minipage}

The original \texttt{\dealii{}} paper containing an overview of its
architecture is \cite{BangerthHartmannKanschat2007}. If you rely on specific
features of the library, please consider citing any of the following:
\begin{itemize}
 \item For geometric multigrid: \cite{Kanschat2004,JanssenKanschat2011};
 \item For distributed parallel computing: \cite{BangerthBursteddeHeisterKronbichler11};
 \item For $hp$ adaptivity: \cite{BangerthKayserHerold2007};
 \item For matrix-free and fast assembly techniques:
   \cite{KronbichlerKormann2012};
 \item For computations on lower-dimensional manifolds:
   \cite{DeSimoneHeltaiManigrasso2009};
 \item For integration with CAD files and tools:
   \cite{HeltaiMola2015}.
\end{itemize}

\dealii{} can interface with many other libraries:
\begin{itemize}
\item ARPACK \cite{arpack}
\item BLAS, LAPACK
\item HDF5 \cite{hdf5}
\item METIS \cite{karypis1998fast}
\item MUMPS \cite{ADE00,MUMPS:1,MUMPS:2,mumps-web-page}
\item muparser \cite{muparser-web-page}
\item NetCDF \cite{rew1990netcdf}
\item OpenCASCADE \cite{opencascade-web-page}
\item p4est \cite{p4est}
\item PETSc \cite{petsc-user-ref,petsc-web-page}
\item SLEPc \cite{Hernandez:2005:SSF}
\item Threading Building Blocks \cite{Rei07}
\item Trilinos \cite{trilinos,trilinos-web-page}
\item UMFPACK \cite{umfpack}
\end{itemize}
Please consider citing the appropriate references if you use interfaces to these
libraries.

Older releases of \dealii{} can be cited as \cite{dealII80,dealII81}.

\nocite{BangerthKanschat1999}

\section{Acknowledgements}

\dealii{} is a world-wide project with dozens of contributors around the
globe. Other than the authors of this paper, the following people contributed code to
this release:
Alexander Grayver,
Andrea Mola,
Andrew Baker,
Angela Klewinghaus,
Arezou Ghesmati,
Ben Thompson,
Christoph Heiniger,
Damien Lebrun-Grandie,
Daniel Arndt,
David Wells,
Denis Davydov,
Fahad Alrashed,
Giorgos Kourakos,
Jean-Paul Pelteret,
Kainan Wang,
Kevin Drzycimski,
Krysztof Bzowski,
Lukas Korous,
Manuel Quezada de Luna,
Markus Bürg,
Martin Steigemann,
Mayank Sabharwal,
Michal Wichrowski,
Mihai Alexe,
Minh Do-Quang,
Shiva Rudraraju,
Uwe Köcher,
Valentin Zingan,
Their contributions are much appreciated!

\dealii{} and its developers are financially supported through a
variety of funding sources. W.~Bangerth and B.~Turcksin were partially
supported by the National Science Foundation under award OCI-1148116
as part of the Software Infrastructure for Sustained Innovation (SI2)
program; by the Computational Infrastructure in Geodynamics initiative
(CIG), through the National Science Foundation under Award
No.~EAR-0949446 and The University of California -- Davis; and through
Award No.~KUS-C1-016-04, made by King Abdullah University of Science
and Technology (KAUST).

L.~Heltai was partially supported by the project OpenViewSHIP,
``Sviluppo di un ecosistema computazionale per la progettazione
idrodinamica del sistema elica-carena'', financed by Regione FVG - PAR
FSC 2007-2013, Fondo per lo Sviluppo e la Coesione.

T.~Heister was partially supported by the Computational Infrastructure in
Geodynamics initiative (CIG), through the National Science Foundation under Award
No. EAR-0949446 and The University of California -- Davis and through Award No. KUS-
C1-016-04, made by King Abdullah University of Science and Technology (KAUST).

The Interdisciplinary Center for Scientific Computing (IWR) at Heidelberg University has provided 
hosting services for the deal.II web page and the SVN archive.

%\marginpar{Everyone's funding, please}

\bibliography{deal82}{}
\bibliographystyle{plain}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2016-02-29 8.3/deal83.tex
Add previous release papers.

\documentclass{ansarticle-preprint}
%\usepackage{ucs}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
%\usepackage{cite}
\usepackage{anslistings}

\usepackage{fontenc}
\usepackage{graphicx}

\hypersetup{
  pdfauthor={Wolfgang Bangerth, Timo Heister, Luca Heltai, Guido Kanschat,
    Martin Kronbichler, Matthias Maier, Bruno Turcksin
  %, Toby D. Young
  },
  pdftitle={The deal.II Library, Version 8.3, 2015},
}

\newcommand{\specialword}[1]{\texttt{#1}}
\newcommand{\dealii}{{\specialword{deal.II}}}
\newcommand{\pfrst}{{\specialword{p4est}}}
\newcommand{\trilinos}{{\specialword{Trilinos}}}
\newcommand{\aspect}{\specialword{Aspect}}
\newcommand{\petsc}{\specialword{PETSc}}
\newcommand{\cmake}{{\specialword{CMake}}}
\newcommand{\autoconf}{{\specialword{autoconf}}}

\title{The \dealii{} Library, Version 8.3}
\author[1]{Wolfgang Bangerth}
\affil[1]{Department of Mathematics, Texas A\&M University, College Station,
    TX 77843, USA,
    %\href{mailto:bangerth@math.tamu.edu}
    {\texttt{bangerth@math.tamu.edu}}}
\author[2]{Timo Heister}
\affil[2]{Mathematical Sciences,
O-110 Martin Hall.
Clemson University.
Clemson, SC 29634, USA,
%\href{mailto:heister@clemson.edu}
{\texttt{heister@clemson.edu}}}
\author[3]{Luca Heltai}
\affil[3]{SISSA - International School for Advanced Studies, Via
  Bonomea 265, 34136 Trieste, Italy,
  %\href{mailto:luca.heltai@sissa.it}
  {\texttt{luca.heltai@sissa.it}}}
\author[4]{Guido Kanschat}
\affil[4]{Interdisciplinary Center for Scientific Computing (IWR),
  Universität Heidelberg, Im Neuenheimer Feld 368, 69120 Heidelberg, Germany,
  %\href{mailto:kanschat@uni-heidelberg.de}
  {\texttt{kanschat@uni-heidelberg.de}}}
\author[5]{Martin Kronbichler}
\affil[5]{Institute for Computational Mechanics, Technische
  Universität München, Boltzmannstr.~15, 85748 Garching b. München,
  Germany,
  %\href{mailto:kronbichler@lnm.mw.tum.de}
  {\texttt{kronbichler@lnm.mw.tum.de}}}
\author[6]{Matthias Maier}
\affil[6]{School of Mathematics, University of Minnesota, 127 Vincent Hall,
  206 Church Street SE, Minneapolis, MN 55455, USA,
  %\href{mailto:matthias.maier@iwr.uni-heidelberg.de}
  {\texttt{msmaier@umn.edu}}}
\author[7]{Bruno Turcksin}
\affil[7]{Department of Mathematics, Texas A\&M University, College Station,
  TX 77843, USA,
  %\href{mailto:turcksin@math.tamu.edu}
  {\texttt{turcksin@math.tamu.edu}}}
%\author[8]{Toby~D.~Young}
%\affil[8]{Institute of Fundamental Technological Research of the Polish
%  Academy of Sciences, ul. Pawińskiego 5b, Warsaw 02-106, Poland,
%  %\href{mailto:tyoung@ippt.pan.pl}
%  {\texttt{tyoung@ippt.pan.pl}}}

\renewcommand{\labelitemi}{--}

\begin{document}
\maketitle

\begin{abstract}
  This paper provides an overview of the new features of the finite element
  library \dealii{} version 8.3.
\end{abstract}

\section{Overview}

\dealii{} version 8.3 was released August 1, 2015. This paper provides an
overview of the new features of this release and serves as a citable
reference for the \dealii{} software library version 8.3. \dealii{} is an
object-oriented finite element library used around the world in the
development of finite element solvers. It is available for free under the
GNU Lesser General Public License (LGPL) from the \dealii{} homepage at
\url{http://www.dealii.org/}.

The major changes of this release are:
\begin{itemize}
\item Improved handling of parallel distributed meshes, including a better
  numbering of cells on coarse meshes. For meshes with many coarse cells, the
  previous Cuthill-McKee numbering would lead to very skinny partitions if the
  processor number was similar to the number of Cuthill-McKee layers. Instead,
  a hierarchical concept inspired by the Morton numbering has been applied to
  define close cells in terms of mesh connectivity, see Fig.~\ref{fig:numbering}. This leads to much better
  volume-to-surface ratios independently of the number of processors and
  combines nicely with the partitioning of additional child cells generated by
  p4est \cite{p4est} and speeds up certain iterative solvers.
\begin{figure}
\centering
\includegraphics[height=6cm]{figures/83_cuthill_mckee_ordering.png}\quad
\includegraphics[height=6cm]{figures/83_hierarchical_ordering.png}
\caption{Comparison of old coarse-cell numbering by owner based on Cuthill-McKee (left) to the new hierarchical concept (right) on an unstructured mesh with 1100 elements on 17 MPI ranks.}
\label{fig:numbering}
\end{figure}
\item New abstract C++11 interface to linear operators.
\item All examples have been changed to use the new DynamicSparsityPattern.
\item Improved support for periodic boundary conditions with arbitrary
  orientations.
  %\item Initial support for isogeometric analysis.
\item New quadrature formulas.
\item Full conversion to the new manifold mechanism (manifold\_id) for boundary
  descriptions: \dealii{} previously utilized a mechanism in which one could
  ask ``boundary'' objects for new points on the boundary, but new points
  inside cells were always created by just averaging surrounding vertex
  locations. A previous version of \dealii{} introduced ``manifold'' objects
  for this, and they are now used consistently throughout the library.
\item Better support for complex-valued problems by doing internal arithmetic in
  the correct data types, rather than defaulting to double precision. In
  particular, we now correctly deduce and store the resulting type when
  multiplying pieces of data with different type (such as shape functions
  evaluated at quadrature points -- which are represented in \texttt{double}
  precision -- and the values of degrees of freedom -- which are represented
  in ways chosen by the user, including complex-valued data types).
\item An implementation of Bernstein polynomial-based finite elements.
\item Interface to the new algebraic multilevel package MueLu of the
  Trilinos project \cite{trilinos-web-page}.
\item More descriptive exception messages in many places for improved user
  productivity in code development.
\item  More than 140 other features and bugfixes.
\end{itemize}
Some of these will be detailed in the following section.
Information on how to cite \dealii{} is provided in Section \ref{sec:cite}.

\section{Significant changes to the library}

This release of \dealii{} contains a number of large and significant changes
that will be discussed in the following sections. It of course also contains a
vast number of smaller changes and added functionality; the details of these
can be found
\href{https://www.dealii.org/8.3.0/doxygen/deal.II/changes_between_8_2_1_and_8_3.html}{in the file that lists all changes for this release}
and that is linked to from the web site of each release as well as the
release announcement.

\subsection{Abstract concepts of linear operators}

\dealii{} now offers a versatile mechanism for storing the concept of a
linear operator as well as partially applied computations. The mechanism is
inspired by the \emph{expression template} optimization strategy, where C++
templates are used to create an arithmetic expression at compile time that can
then be efficiently be evaluated at runtime. In contrast to the template
approach, the implementation in \dealii{} utilizes C++11 features%
\footnotemark\footnotetext{Consequently, the classes outlined here are only
available in \dealii{} if the compiler in use supports C++11.}
such as \emph{lambda expressions} and \texttt{std::function} objects that
avoid the majority of the ``template overhead'' that usually comes with a
pure template solution. In particular, it avoids the often very lengthy and
cumbersome error messages associated with programming expression templates.

Our implementation only requires a minimal vector and matrix interface,
that all of \dealii's concrete vector and matrix types adhere to. For
example, in essence a matrix only has to provide a \texttt{vmult} method
for matrix-vector multiplication. This makes the interface fully
transparent with respect to the concrete matrix, or vector type that is
used, but naturally rules out further low-level optimization approaches
that can be used for non-distributed linear algebra (such as loop fusion).

The mechanism is centered around two classes, \texttt{LinearOperator} and
\texttt{PackagedOperation}. Both essentially consist of several
\texttt{std::function} objects that store the knowledge of how to apply a
linear operator, or a packaged operation, as well as how to initialize
vectors of range and domain space:
\begin{c++}
template <typename Range, typename Domain> class LinearOperator
{
public:
  std::function<void(Range &v, const Domain &u)> vmult;
  // ...
  std::function<void(Range &v, bool fast)> reinit_range_vector;
  std::function<void(Domain &v, bool fast)> reinit_domain_vector;
};

template <typename Range> class PackagedOperation
{
public:
  std::function<void(Range &v)> apply;
  // ...;
  std::function<void(Range &v, bool fast)> reinit_vector;
};
 \end{c++}

The primary purpose of the linear operator concept is to provide syntactic
sugar for complex matrix-vector operations and free the user from having to
create, set up, and handle intermediate storage locations by hand. As an
example, consider the operation $(A+k\,B)\,C$, where $A$, $B$ and $C$ denote
(possible different) matrices (or, in fact, other linear operators). In order
to construct a \verb|LinearOperator|
object \texttt{op} that encapsulates knowledge of this operation,
\dealii{} now allows one to write:

\begin{c++}
  dealii::SparseMatrix<double> A, B, C;
  const double k = ...;
  // Setup and assembly of matrices
  .
  const auto op_a = linear_operator(A);
  const auto op_b = linear_operator(B);
  const auto op_c = linear_operator(C);

  const auto op = (op_a + k * op_b) * op_c;
\end{c++}
Now, \texttt{op} can be used as a matrix object in its own right for
further computation. This includes passing it to any of \dealii{}'s linear
solver classes as the operator for which to solve a linear system. To make
this work, \dealii{} appropriately overloads arithmetic operators to allow
composition of linear operators as above.

The \texttt{PackagedOperation} class allows lazy evaluation of expressions
involving vectors and linear operators. This is done by storing the
computational expression and only performing the computation when either
the object is implicitly converted to a vector object, or if
\texttt{apply()} is invoked by hand. This avoids unnecessary temporary
storage of intermediate results. As an example consider the addition of
multiple vectors:
\begin{c++}
   dealii::Vector<double> a, b, c, d;
   // ..
   dealii::Vector<double> result = a + b - c + d;
\end{c++}

\texttt{operator+} (with two vector arguments) is implemented to return a
\texttt{PackagedOperation} whose \texttt{apply} function is implemented the
following way:
\begin{c++}
  apply = [&a, &b](Range &x) {
                                x = a;
                                x += b;
                             };
\end{c++}
Similarly, addition and subtraction of mixed vector and packaged operation
types are implemented. If implemented in the most straightforward way,
executing these additions would require the use of multiple temporary
vectors, along with the necessary memory allocation and de-allocation. On
the other hand, creating a \texttt{PackagedOperation} for the expression
\texttt{a + b - c + d} and converting it to a vector results in code
equivalent to the following:
\begin{c++}
   dealii::Vector<double> a, b, c, d;
   // ..
   dealii::Vector<double> result = a;
   result += b;
   result -= c;
   result += d;
\end{c++}
Consequently, this operation avoids the use of any intermediate storage.

Further, in addition to pure vector operations, scalar multiplication
(and thus all vector space operations) and application of a
\texttt{LinearOperator} to a vector can be expressed within a
\texttt{PackagedOperation}. For example, a residual can be expressed as:
\begin{c++}
  dealii::Vector<double> residual =  b - op_a * x;
\end{c++}
In all of these cases, \texttt{PackagedOperation} only stores references to
vector and linear operator arguments. As a consequence, creating a
\texttt{PackagedOperation} object and evaluating it multiple times will always
use the then-current values of the arguments, not the values at the time the
packaged operation was created.

\subsection{Initial support for iso-geometric analysis}

Nonuniform rational B-splines (NURBS) are the basis of most modern CAD packages.
Isogeometric analysis is a computational technique that aims at integrating
finite element analysis (FEA) into conventional NURBS-based CAD design
tools by directly using NURBS basis functions in
the FEA application~\cite{Cottrell2009}. The preliminary
implementation of the iso-geometric analysis concept in the \dealii{} library
is based on two main classes: a new finite element based on Bernstein
polynomials called \verb|FE_Bernstein| and a general mapping called
\verb|MappingFEField|, which generalizes the iso-parametric model to allow
the description of the geometry using arbitrary finite element displacement or
location fields.

The scalar Bernstein finite element \verb|FE_Bernstein|, in analogy
with \verb|FE_Q|, yields the finite element space of continuous,
piecewise Bernstein polynomials of degree $p$ in each coordinate
direction~\cite{Cottrell2009}. This class is realized using tensor
product polynomials of Bernstein basis polynomials.

The \verb|MappingFEField| is a generalization of the
\verb|MappingQEulerian| class, for arbitrary vector finite elements. The
two main notable differences are that this class interprets the degrees of
freedom of the underlying \verb|DoFHandler| as absolute positions
instead of displacements; and it
allows for arbitrary finite element types.
This class effectively decouples topology from geometry, by relegating all
geometrical information to the components of a FiniteElement vector field.
The underlying \verb|Triangulation| class is only used to infer topological
and connectivity information between cells.

(This feature was primarily developed by Marco Tezzele.)

\subsection{New quadrature formulas}

Two specialized quadrature formulas have been added to the library.

\subsubsection{Telles quadrature formulas}

Many applications require the integration of functions which might be
singular in some a-priori known point locations. A common strategy in such
cases, which was already implemented in the library, is the use of
Lachat-Watson quadrature formulas~\cite{Lachat1976}. These formulas are
fairly accurate but very expensive. A new quadrature formula for singular
integrals was introduced in \dealii{} that implements the quadrature formula
developed by Telles in~\cite{Telles-2005-a}.

The main idea of this quadrature formula is to create an iso-morphism of a
standard Gauss quadrature formula according to a nonlinear transformation
that has both first and second order derivatives equal to zero at the
singularity point.  Such a transformation removes the singularity of the
integrand and drops the approximation error from about the $10 \%$ to
$0.3\%$ for a large class of singular integrands~\cite{Telles-2005-a}. These
quadrature formulas are particularly relevant for approximation of Boundary
Integral Equations (see, e.g.,~\cite{GiulianiMolaHeltai-2015-a}).

(This feature was primarily developed by Nicola Giuliani.)

\subsubsection{Gauss-Chebyshev quadrature formulas}

Gauss-Chebyshev quadrature formulas~\cite{AbramowitzStegun} are used to
approximate integrals of functions on the interval $[-1,1]$ with weight
$w(x)=(1-x^2)^{-1/2}$.  In \dealii{} the reference interval is $[0,1]$, hence
our implementation of Gauss-Chebyshev formulas are mapped to this interval,
and the weight is changed to $w(x)=1/\sqrt{x(1-x)}$.

As for the other Gaussian quadrature formulas, three flavors of
Gauss-Chebyshev quadrature rules have been implemented for any order:
\begin{description}
\item[Gauss-Chebyshev (GC)] formulas with $N$ points that integrate
  exactly polynomials up to the order $2N-1$;
\item[Gauss-Radau-Chebyshev (GRC)] formulas with $N$ points where one of
  the two endpoints of the interval can be specified as a constrained
  quadrature point; these formulas are exact up to the order $2N-2$;
\item[Gauss-Lobatto-Chebyshev (GLC)] formulas with $N$ points where two
  quadrature nodes are constrained to lie on the boundary points of the
  interval; these formulas exactly integrate polynomials up to the order $2N-3$.
\end{description}

One important application of the GLC quadrature formulas is in the context
of Spectral Element Methods (SEM) using Lagrangian interpolants on GLC
points as basis functions (this is a good choice for the composite
interpolation of ``very regular'' continuous functions). This pair of basis
functions and quadrature formulas generates diagonal mass matrices, that
can be efficiently applied both in direct and in inverse form. For saddle
point problems, one common choice in SEM~\cite{MadayPatera} is the
$\mathbb{Q}_N-\mathbb{Q}_{N-2}$ (inf-sup stable) couple based on Lagrangian
interpolants on $N+1$ GLC and $N-1$ GC nodes respectively.

(This feature was primarily developed by Giuseppe Pitton.)

\subsection{Better error messages}

\dealii{} uses assertions extensively to validate that input arguments to
functions are within their allowed ranges, are mutually consistent, and
satisfy appropriate preconditions. We also use assertions to check for
internal conditions as well as for postconditions of many functions. All told,
the sources of \dealii{} contain around 9,000 assertions.

Whenever an assertion is failed, \dealii{} aborts the program with an error
message that includes the failing condition, its location, the surrounding
function's name, the call stack, and additional information specific to the
exception class associated with this particular error. For example, trying to
read an element of a vector that does not exist would trigger the exception in
the following code:
\begin{c++}
template <typename Number>
inline
Number Vector<Number>::operator() (const size_type i) const
{
  Assert (i<vec_size, ExcIndexRange(i,0,vec_size));
  return val[i];
}
\end{c++}
The error message produced from this code would look like this:
\begin{verbatim}
--------------------------------------------------------
An error occurred in line <1222> of file <.../include/deal.II/lac/vector.h> in function
    Number& dealii::Vector<number>::operator()(dealii::Vector<number>::size_type)
      [with Number = double, dealii::Vector<number>::size_type = unsigned int]
The violated condition was:
    i<vec_size
The name and call sequence of the exception was:
    ExcIndexRangeType<size_type>(i,0,vec_size)
Additional Information:
Index 123456 is not in the half-open range [0,679).

Stacktrace:
-----------
#0  ./step-22: dealii::Vector<double>::operator()(unsigned int)
#1  ./step-22: Step22::StokesProblem<2>::assemble_system()
#2  ./step-22: Step22::StokesProblem<2>::run()
#3  ./step-22: main
--------------------------------------------------------
\end{verbatim}

In this case, the error message does state the index that is being accessed,
as well as the size of the vector, both under ``Additional Information''. The
error message may not be overly clear, but it helps illustrate the values of
variables that participated in the failed condition \verb|i<vec_size|.

However, many error messages produced by \dealii{} either do not produce any
additional information at all, or very little that is understandable to a
novice user. Observing users deal with errors in the courses we teach has
shown that poor error messages are a major obstacle to quickly identifying the
causes of problems. Ideally, error messages would attempt to explain what is
going wrong, what are typical causes for this, and what possible solutions
are.

In an effort to improve this situation, we have rewritten and augmented the
error messages produced by several dozen classes of assertions. Going through
the remainder of all error messages is a longer-term project.

\subsection{Incompatible changes}

This version of \dealii{} contains a number of changes that are
incompatible with previous versions. A complete list can be found
\href{https://www.dealii.org/8.3.0/doxygen/deal.II/changes_between_8_2_1_and_8_3.html}{in
  the file that lists all changes for this release}. Most of these
changes are corner cases that are not likely to affect a
significant fraction of users. In particular, many of these changes
remove functions and classes that were already deprecated (and
produced a compiler warning); often these functions had been marked as
depracted for several years already.

The only change that has the potential to be of more general interest
is a clarification of the \texttt{Tensor<1,dim>} and
\texttt{Point<dim>} classes. As they both represent vectors in a
\texttt{dim}-dimensional space, they were previously often used
interchangeably. This has been changed in many places where we
semantically now associate \texttt{Point<dim>} with a \textit{point}
in space, i.e., a vector anchored at the origin; and where we
semantically associate \texttt{Tensor<1,dim>} with vectors anchored
anywhere.

An example is that \texttt{operator-(Point<dim>, Point<dim>)} now
returns a \texttt{Tensor<1,dim>}. Likewise, normal vectors to cells
are also represented by this latter data type. Overall, there are few
resulting incompatibilities, with the exception being assigning an
expression that previously was a \texttt{Point<dim>} but now is a
\texttt{Tensor<1,dim>} (such as the difference between two points) to
a variable of type \texttt{Point<dim>}. In most contexts, however,
everything will continue to work because a \texttt{Point<dim>} can
automatically be (down-)cast to a \texttt{Tensor<1,dim>}.

\section{How to cite \dealii{}}\label{sec:cite}

In order to justify the work the developers of \dealii{} put into this
software, we ask that papers using the library reference one of the
\dealii{} papers. This helps us justify the effort we put into it.

There are various ways to reference \dealii{}. To acknowledge the use of a
particular version of the library, reference the present document. For up
to date information and bibtex snippets for this document see:

\bigskip

\begin{center}
 \url{https://www.dealii.org/publications.html}
\end{center}

\bigskip

% \begin{minipage}{0.9\textwidth}%no page break in here please
% \begin{verbatim}
% @article{dealII82,
%   title = {The {\tt deal.{I}{I}} Library, Version 8.2},
%   author = {W. Bangerth and T. Heister and L. Heltai
%    and G. Kanschat and M. Kronbichler and M. Maier
%    and B. Turcksin and T. D. Young},
%   journal = {preprint},
%   year = {2015},
% }
% \end{verbatim} %  journal = {arXiv preprint \url{http://arxiv.org/abs/TODO}},
% \end{minipage}

The original \texttt{\dealii{}} paper containing an overview of its
architecture is \cite{BangerthHartmannKanschat2007}. If you rely on specific
features of the library, please consider citing any of the following:
\begin{itemize}
 \item For geometric multigrid: \cite{Kanschat2004,JanssenKanschat2011};
 \item For distributed parallel computing: \cite{BangerthBursteddeHeisterKronbichler11};
 \item For $hp$ adaptivity: \cite{BangerthKayserHerold2007};
 \item For matrix-free and fast assembly techniques:
   \cite{KronbichlerKormann2012};
 \item For computations on lower-dimensional manifolds:
   \cite{DeSimoneHeltaiManigrasso2009};
 \item For integration with CAD files and tools:
   \cite{HeltaiMola2015};
 \item For \texttt{LinearOperator} and \texttt{PackagedOperation} facilities:
   \cite{MaierBardelloniHeltai-2015-a}.
\end{itemize}

\dealii{} can interface with many other libraries:
\begin{itemize}
\item ARPACK \cite{arpack}
\item BLAS, LAPACK
\item HDF5 \cite{hdf5}
\item METIS \cite{karypis1998fast}
\item MUMPS \cite{ADE00,MUMPS:1,MUMPS:2,mumps-web-page}
\item muparser \cite{muparser-web-page}
\item NetCDF \cite{rew1990netcdf}
\item OpenCASCADE \cite{opencascade-web-page}
\item p4est \cite{p4est}
\item PETSc \cite{petsc-user-ref,petsc-web-page}
\item SLEPc \cite{Hernandez:2005:SSF}
\item Threading Building Blocks \cite{Rei07}
\item Trilinos \cite{trilinos,trilinos-web-page}
\item UMFPACK \cite{umfpack}
\end{itemize}
Please consider citing the appropriate references if you use interfaces to these
libraries.

Older releases of \dealii{} can be cited as \cite{dealII80,dealII81,dealII82}.

\nocite{BangerthKanschat1999}

\section{Acknowledgments}

\dealii{} is a world-wide project with dozens of contributors around the
globe. Other than the authors of this paper, the following people contributed code to
this release:
%
% get this from the changes.h file using
%   egrep '^ *\(.*2015' 8.2.1-vs-8.3.0.h | perl -p -e 's/2015.*//g; s/,/\n/g; s/^ *\(?//g;' | sort | uniq
% then sort by last name and remove the authors of this paper
%
Daniel Arndt,
Mauro Bardelloni,
Praveen C,
Denis Davydov,
Anton Evgrafov,
Arezou Ghesmati,
Nicola Giuliani,
Lukas Korous,
Ross Kynch,
Giuseppe Pitton,
Fernando Posada,
Lei Qiao,
Angel Rodriguez,
Ian Rose,
Florian Sonner,
Martin Steigemann,
Zhen Tao,
Marco Tezzele,
David Wells.

Their contributions are much appreciated!

\dealii{} and its developers are financially supported through a
variety of funding sources. W.~Bangerth and B.~Turcksin were partially
supported by the National Science Foundation under award OCI-1148116
as part of the Software Infrastructure for Sustained Innovation (SI2)
program; and by the Computational Infrastructure in Geodynamics initiative
(CIG), through the National Science Foundation under Award
No.~EAR-0949446 and The University of California -- Davis.

L.~Heltai was partially supported by the project OpenViewSHIP,
``Sviluppo di un ecosistema computazionale per la progettazione
idrodinamica del sistema elica-carena'', financed by Regione FVG - PAR
FSC 2007-2013, Fondo per lo Sviluppo e la Coesione, and by the project
TRIM ``Tecnologia e Ricerca Industriale per la Mobilit\`a Marina'',
CTN01-00176-163601, funded by MIUR -  Ministero dell'Istruzione,
dell'Universit\`a e della Ricerca.

T.~Heister was partially supported by the Computational Infrastructure in
Geodynamics initiative (CIG), through the National Science Foundation under Award
No. EAR-0949446 and The University of California -- Davis.

The Interdisciplinary Center for Scientific Computing (IWR) at Heidelberg University has provided
hosting services for the \dealii{} web page and the SVN archive.

%\marginpar{Everyone's funding, please}

\bibliography{deal83}{}
\bibliographystyle{abbrv}

\end{document}

### 2016-02-29 make_offline_doc.sh
Move more files from the svn repo here.

@@ -1,0 +1,33 @@
+#!/bin/bash
+
+if [ "$#" -ne 3 ]; then
+    echo "Offline documentation generator for deal.II"
+    echo "Usage:"
+    echo "  ./make_offline_doc.sh <major> <minor> <patch>"
+    echo "Execute in the base directory of a clean checkout for the release."
+    echo "For an RC use: './make_offline_doc.sh 8 3 0-rc1'"
+    exit 1
+fi
+
+MAJOR=$1
+MINOR=$2
+PATCH=$3
+
+echo release $MAJOR.$MINOR.$PATCH
+
+# Generate documentation
+rm -rf builddoc
+mkdir builddoc
+cd builddoc
+cmake -DDEAL_II_WITH_MPI=OFF -DCMAKE_BUILD_TYPE=Debug -DDEAL_II_COMPONENT_DOCUMENTATION=ON -DCMAKE_INSTALL_PREFIX=`pwd`/../installeddoc/ ../
+make documentation -j 8
+
+cd ../installeddoc
+
+cd doc/doxygen/deal.II
+../../../../contrib/utilities/makeofflinedoc.sh
+
+cd ../../..
+tar czf ../dealii-$MAJOR.$MINOR.$PATCH-offline_documentation.tar.gz doc
+cd ..
+

### 2016-03-01 8.4/deal84.tex
Move 8.4 files to their own directory.
More funding

--- 2016-02-28 deal84.tex
@@
 
+D.~Davydov was supported by the ERC Advanced Grant MOCOPOLY and the Competence Network for Technical and Scientific High Performance Computing in Bavaria (KONWIHR).
+
 L.~Heltai was partially supported by the project OpenViewSHIP,
@@
 
+G.~Kanschat and M.~Kronbichler were partially supported by the German Research
+Foundation (DFG) through the project ExaDG. M.~Kronbichler also acknowledges
+the Gauss Centre for Supercomputing e.V.~for funding algorithm enhancements by
+providing computing time on the GCS Supercomputer SuperMUC at Leibniz
+Supercomputing Centre (LRZ) through project id \emph{pr83te}.
+
 The Interdisciplinary Center for Scientific Computing (IWR) at Heidelberg University has provided

### 2016-03-08 8.4/deal84.tex
Update affiliation

@@
 \author[6]{Martin Kronbichler}
-\affil[6]{Institute for Computational Mechanics, Technische
-  Universit{\"a}t M{\"u}nchen, Boltzmannstr.~15, 85748 Garching b. Mænchen,
+\affil[6]{Institute for Computational Mechanics, Technical University of
+  Munich, Boltzmannstr.~15, 85748 Garching,
   Germany,
@@
 providing computing time on the GCS Supercomputer SuperMUC at Leibniz
-Supercomputing Centre (LRZ) through project id \emph{pr83te}.
+Supercomputing Centre (LRZ) through project id pr83te.
 

### 2016-03-09 8.4/deal84.tex
Revisit and revise.

@@
 
-\dealii{} version 8.4.0 was released February 29, 2016. This paper provides an
+\dealii{} version 8.4.0 was released March 10, 2016. This paper provides an
 overview of the new features of this release and serves as a citable
@@
   FEValues class has been rewritten. It is now much better documented.
-\item Initial support for compiling with Visual C++ 2013 and 2015 under Windows
+\item Initial support for compiling with Visual C++ 2013 and 2015 under
+  Microsoft Windows
   has been added.
@@
 \end{itemize}
-Some of these will be detailed in the following section.
+The more important ones of these will be detailed in the following section.
 Information on how to cite \dealii{} is provided in Section \ref{sec:cite}.
@@
 
-Previously, partitioning a parallel mesh (represented by the
-\texttt{parallel::distributed::Triangulation} class) between
+Previously, partitioning a parallel mesh (represented by objects of class
+\texttt{parallel::}\-\texttt{distributed::}\-\texttt{Triangulation}) between
 processors assumed that every cell should be weighted equally. On the
 other hand, the \pfrst{} library which manages the partitioning
-process, allows for attaching weights for each cell and thereby
+process, allows for attaching weights to each cell and thereby
 enables ways in which not the number of cells per MPI process is
@@
 functions that will be called upon mesh refinement and coarsening,
-returning a weight for each cell. These weights will be added after
-all slots (i.e., callback functions) connected to the signal have been
-called.
+returning a weight for each cell. These weights will be added up over
+all slots (i.e., callback functions) connected to the signal.
 
@@
 scheme that tracks particles that are advected along with a flow field
-and stored them per-cell, can indicate their computational needs for
-each cell, and there no central place in a user code (other than the
+and stores them per-cell, can indicate their computational needs for
+each cell. In particular, there is no central place in a user code (other than
+the 
 triangulation itself) that has to collect these needs and forward this
@@
 hand, the expansion coefficients of the field, $U_j$ are stored in
-vectors of types chosen by the user; their underlying representation
-can be \texttt{double}, but also \texttt{float}, \texttt{long double},
-\texttt{PetscScalar}, or \texttt{std::complex<float>} for example. It could also
+vectors over scalar types chosen by the user; their underlying representation
+could be \texttt{double}, but also \texttt{float}, \texttt{long double},
+\texttt{PetscScalar}, or \texttt{std::complex<float>}. It could also
 be an autodifferentiation type.
@@
 \texttt{Vector<long double>}, then the gradients
-$\nabla\varphi_j(\mathbf x_q)$ will be computed as objects of type
+$\nabla u_h(\mathbf x_q)$ will be computed as objects of type
 \texttt{Tensor<1,dim,long double>}. Likewise, if a user uses a PETSc
@@
 places. In particular, this enables the use of complex-valued solution
-vectors in many more places. On the other hand, many but not all
+vectors in many more places than before. On the other hand, many but not all
 places have learned what actually to do with complex numbers. This is,
@@
 
-Previously, in order to use \texttt{Triangulation<dim,spacedim>} class with MPI one had to manually
-partition triangulation (e.g.
-via \texttt{METIS}), calculate and store index sets of locally owned and locally relevant degrees-of-freedom (DoF), querry \texttt{subdomain\_id}  of a cell during assembly etc.
-In order to facilitate the usage of \texttt{Triangulation} class with MPI
-and to make its behaviour in this context consistent with \texttt{parallel::distributed::Triangulation}, a new class
-\texttt{parallel::shared::Triangulation} was introduced.
-It extends \texttt{Triangulation} class to automatically partition triangulation when run with MPI.
-Identical functionality between \texttt{shared} and \texttt{distributed} triangulation
-(e.g. locally owned and relevant DoFs, mpi communicator, etc)
-is grouped in the parent class \texttt{parallel::Triangulation}.
-The main difference between the two parallel triangulations is that in the case of
+\dealii{} already has two types of triangulations: the
+\texttt{Triangulation<dim,spacedim>} class works entirely locally, whereas
+\texttt{parallel::distributed::Triangulation<dim,spacedim>} builds on the
+former, but only stores the local partition that corresponds to a globally
+distributed triangulation managed across an MPI network. One \textit{could},
+however, use the former also for parallel computations in situations where one
+needs access to \textit{all} cells, not just the subset of cells that
+correspond to the partition owned by the current processor. This required
+building the same triangulation on all processors, then
+manually partitioning it (e.g., via \texttt{METIS}), calculating and storing
+index sets of locally owned and locally relevant degrees of freedom (DoFs),
+querying the \texttt{subdomain\_id}  of a cell during assembly, etc.
+
+In order to simplify this usage of the \texttt{Triangulation} class with MPI
+and to make its behavior in this context consistent with \texttt{parallel::distributed::Triangulation}, a new class
+\texttt{parallel::shared::Triangulation} has been introduced.
+It extends the \texttt{Triangulation} class to automatically partition the triangulation when run with MPI.
+Shared functionality between the \texttt{shared} and \texttt{distributed}
+triangulation classes
+(e.g., locally owned and relevant DoFs, MPI communicators, etc)
+is now grouped in a common parent class \texttt{parallel::Triangulation}.
+The main difference between the two classes is that in the case of
 \texttt{parallel::shared::Triangulation} each process stores all cells of the triangulation.
@@
 and the rest are ghost cells.
-This behaviour can be altered via an additional boolean flag provided to the constructor of the class.
-In this case, a set of ghost cells will consist of a halo layer of cells around locally owned cells.
+This behavior can be altered via an additional boolean flag provided to the constructor of the class.
+In this case, the set of ghost cells will consist of a halo layer of cells around locally owned cells.
 Cells which are neither ghost nor locally owned are marked as artificial.
-This is consistent with the behaviour of \texttt{parallel::distributed::Triangulation},
-although in the latter case the size of the set of artificial cells will be much smaller.
-The introduction of \texttt{parallel::shared::Triangulation} class together with the
+This is consistent with the behavior of \texttt{parallel::distributed::Triangulation},
+although in the latter case the size of the set of artificial cells will be
+much smaller.
+
+The introduction of the \texttt{parallel::shared::Triangulation} class together with the
 optional artificial cells and parent \texttt{parallel::Triangulation} class
 facilitates writing
-algorithms that are indifferent to the way triangulation is stored in the MPI context.
-For example, \texttt{DoFTools::locally\_relevant\_dofs()} will now return a subset of
-all DoFs for both triangulations, granted that artificial cells are enabled in
-\texttt{parallel::shared::Triangulation}.
-Assmebly routines can now use \texttt{cell->is\_locally\_owned()}
+algorithms that are indifferent to the way a triangulation is stored in the MPI context.
+For example, the function \texttt{DoFTools::locally\_active\_dofs()} will
+return the appropriate subset of
+all DoF indices for both triangulations.
+Assmebly routines can use predicates such as \texttt{cell->is\_locally\_owned()}
 for both triangulations.
-Currently, only standard (non hp) \texttt{DoFHandler} is supported.
+
+Based on the new triangulation class, the (non-hp) \texttt{DoFHandler} manages
+degrees of freedom in the same way as it has already done for a long time for
+sequential and parallel distributed triangulations.
 
@@
   +
-  J^{-1}\left(\hat\nabla J^{-1}\right)\hat\nabla\hat\varphi_j(\hat{\mathbf x}_q),
+  J^{-1}\left(\hat\nabla [J^{-1}]\right)\hat\nabla\hat\varphi_j(\hat{\mathbf x}_q),
 \end{align*}
@@
 have to be appropriately contracted.  Here, the difficulty lies in
-computing the derivative $\hat\nabla J^{-1}$: for the usual mappings
+computing the derivative $\hat\nabla [J^{-1}]$: for the usual mappings
 on quadrilaterals and hexahedra, $J$ is in general a polynomial in the
@@
 The key to computing second (and higher) derivatives of shape
-functions is to recognize that $\hat\nabla J^{-1}$ can be expressed
+functions is to recognize that $\hat\nabla [J^{-1}]$ can be expressed
 more conveniently by observing that
@@
   \\
-  &= J (\hat\nabla J^{-1}) + (\hat\nabla J) J^{-1},
+  &= J (\hat\nabla [J^{-1}]) + (\hat\nabla J) J^{-1},
 \end{align*}
-and consequently $\hat\nabla J^{-1} = - J^{-1} (\hat\nabla J)
+and consequently $\hat\nabla [J^{-1}] = - J^{-1} (\hat\nabla J)
 J^{-1}$. Here, $J^{-1}$ is a matrix that is already available from
@@
 To facilitate this complex interplay, \dealii{} has three class
-hierarchies with base classes \texttt{FiniteElement} (for the
+hierarchies rooted in the base classes \texttt{FiniteElement} (for the
 description of shape functions on the reference cell),

### 2016-05-03 8.4/deal84.tex
change title to 8.4 but use 8.4.1 release date

@@
   },
-  pdftitle={The deal.II Library, Version 8.4.0, 2016},
+  pdftitle={The deal.II Library, Version 8.4, 2016},
 }
@@
 
-\title{The \dealii{} Library, Version 8.4.0}
+\title{The \dealii{} Library, Version 8.4}
 \author[1]{Wolfgang Bangerth}
@@
   This paper provides an overview of the new features of the finite element
-  library \dealii{} version 8.4.0.
+  library \dealii{} version 8.4.
 \end{abstract}
@@
 
-\dealii{} version 8.4.0 was released March 10, 2016. This paper provides an
+\dealii{} version 8.4.1 was released April 7, 2016. This paper provides an
 overview of the new features of this release and serves as a citable
-reference for the \dealii{} software library version 8.4.0. \dealii{} is an
+reference for the \dealii{} software library version 8.4. \dealii{} is an
 object-oriented finite element library used around the world in the

### 2016-05-03 8.4/deal84.tex
add link to changes

@@
 can be found
-\href{https://www.dealii.org/8.4.0/doxygen/deal.II/changes_between_8_3_0_and_8_4.html}{in the file that lists all changes for this release}
+\href{https://www.dealii.org/8.4.0/doxygen/deal.II/changes_between_8_3_and_8_4.html}{in the file that lists all changes for this release} (see \cite{changes84})
 and that is linked to from the web site of each release as well as the
@@
 The
-\href{https://www.dealii.org/8.4.0/doxygen/deal.II/changes_between_8_3_0_and_8_4.html}{file
-  that lists all changes for this release} lists another 18
+\href{https://www.dealii.org/8.4.0/doxygen/deal.II/changes_between_8_3_and_8_4.html}{file
+  that lists all changes for this release} (see \cite{changes84}) lists another 18
 incompatible changes, but none of these should in fact be visible in

### 2017-03-02 8.5/deal85.tex
initial paper
update

--- 2016-05-03 8.4/deal84.tex
@@
 \hypersetup{
-  pdfauthor={Wolfgang Bangerth, Timo Heister, Luca Heltai, Guido Kanschat,
-    Martin Kronbichler, Matthias Maier, Bruno Turcksin
-  %, Toby D. Young
-  },
-  pdftitle={The deal.II Library, Version 8.4, 2016},
+  pdfauthor={Wolfgang Bangerth, Timo Heister},
+  pdftitle={The deal.II Library, Version 8.5, 2016},
 }
@@
 
-\title{The \dealii{} Library, Version 8.4}
+\title{The \dealii{} Library, Version 8.5}
 \author[1]{Wolfgang Bangerth}
-\affil[1]{Department of Mathematics, Texas A\&M University, College Station,
-    TX 77843, USA,
+\affil[1]{xxx,
     {\texttt{bangerth@math.tamu.edu}}}    
-\author[2]{Denis Davydov}
-\affil[2]{Chair of Applied Mechanics, University of
-    Erlangen-Nuremberg, Egerlandstr. 5, 91058 Erlangen, Germany,
-  {\texttt{denis.davydov@fau.de}}}
-\author[3]{Timo Heister}
-\affil[3]{Mathematical Sciences,
+\author[2]{Timo Heister}
+\affil[2]{Mathematical Sciences,
   O-110 Martin Hall.
@@
   {\texttt{heister@clemson.edu}}}
-\author[4]{Luca Heltai}
-\affil[4]{SISSA - International School for Advanced Studies, Via
-  Bonomea 265, 34136 Trieste, Italy,
-  {\texttt{luca.heltai@sissa.it}}}
-\author[5]{Guido Kanschat}
-\affil[5]{Interdisciplinary Center for Scientific Computing (IWR),
-  Universit{\"a}t Heidelberg, Im Neuenheimer Feld 368, 69120 Heidelberg, Germany,
-  {\texttt{kanschat@uni-heidelberg.de}}}
-\author[6]{Martin Kronbichler}
-\affil[6]{Institute for Computational Mechanics, Technical University of
-  Munich, Boltzmannstr.~15, 85748 Garching,
-  Germany,
-  {\texttt{kronbichler@lnm.mw.tum.de}}}
-\author[7]{Matthias Maier}
-\affil[7]{School of Mathematics, University of Minnesota, 127 Vincent Hall,
-  206 Church Street SE, Minneapolis, MN 55455, USA,
-  {\texttt{msmaier@umn.edu}}}
-\author[8]{Bruno Turcksin}
-\affil[8]{Department of Mathematics, Texas A\&M University, College Station,
-  TX 77843, USA,
-  {\texttt{turcksin@math.tamu.edu}}}
-\author[9]{David Wells}
-\affil[9]{Department of Mathematical Sciences, Rensselaer Polytechnic Institute,
-Troy, NY 12180, USA
-  {\texttt{wellsd2@rpi.edu}}}
-%\author[8]{Toby~D.~Young}
-%\affil[8]{Institute of Fundamental Technological Research of the Polish
-%  Academy of Sciences, ul. Pawi{\'n}skiego 5b, Warsaw 02-106, Poland,
-%  {\texttt{tyoung@ippt.pan.pl}}}
 
@@
   This paper provides an overview of the new features of the finite element
-  library \dealii{} version 8.4.
+  library \dealii{} version 8.5.
 \end{abstract}
@@
 
-\dealii{} version 8.4.1 was released April 7, 2016. This paper provides an
+\dealii{} version 8.4.0 was released March 11, 2016. This paper provides an
 overview of the new features of this release and serves as a citable
@@
  \item For \texttt{LinearOperator} and \texttt{PackagedOperation} facilities:
-   \cite{MaierBardelloniHeltai-2015-a}.
+   \cite{MaierBardelloniHeltai-2016-a,MaierBardelloniHeltai-2016-b}.
  \item For uses of the \texttt{WorkStream} interface:
@@
 
-\bibliography{deal84}{}
+\bibliography{paper}{}
 \bibliographystyle{abbrv}

### 2017-03-02 8.5/deal85.tex
cleanup old changes, add 8.4 citation

@@
 
-\dealii{} version 8.4.0 was released March 11, 2016. This paper provides an
+\dealii{} version 8.5.0 was released March 11, 2017. This paper provides an
 overview of the new features of this release and serves as a citable
-reference for the \dealii{} software library version 8.4. \dealii{} is an
+reference for the \dealii{} software library version 8.5. \dealii{} is an
 object-oriented finite element library used around the world in the
@@
 \begin{itemize}
-\item Parallel triangulations can now be partitioned in ways that allow
-  weighting cells differently.
-\item Improved support for mixed-type arithmetic throughout the library.
-\item A new triangulation type that supports parallel computations
-  but ensures that the entire mesh is available on every processor.
-\item An implementation of the Rannacher-Turek element, as well as an
-  element that extends the usual Q(p) elements by bubble functions.
-\item Second and third derivatives of finite element fields are now
-  computed exactly.
-\item The various \emph{Concepts}, or requirements on template parameters
-  in the library, are now consistently labeled and documented as such.
-\item The interface between finite elements, quadrature, mapping, and the
-  FEValues class has been rewritten. It is now much better documented.
-\item Initial support for compiling with Visual C++ 2013 and 2015 under
-  Microsoft Windows
-  has been added.
+\item 
 \item More than 140 other features and bugfixes.
@@
 
-\subsection{Parallel triangulations can now be partitioned with weights}
+\subsection{important thing 1}
 
-Previously, partitioning a parallel mesh (represented by objects of class
-\texttt{parallel::}\-\texttt{distributed::}\-\texttt{Triangulation}) between
-processors assumed that every cell should be weighted equally. On the
-other hand, the \pfrst{} library which manages the partitioning
-process, allows for attaching weights to each cell and thereby
-enables ways in which not the number of cells per MPI process is
-equilibrated, but the sum of weights on the cells managed by each
-process. \dealii{} now also supports this feature.
-
-The implementation of this mechanism is based on a callback mechanism,
-in the form of the signal-slot design pattern. User codes can register
-functions that will be called upon mesh refinement and coarsening,
-returning a weight for each cell. These weights will be added up over
-all slots (i.e., callback functions) connected to the signal.
-
-The mechanism chosen has the advantage that all parties that use a
-triangulation, for example multiple \texttt{DoFHandler} objects or a
-scheme that tracks particles that are advected along with a flow field
-and stores them per-cell, can indicate their computational needs for
-each cell. In particular, there is no central place in a user code (other than
-the 
-triangulation itself) that has to collect these needs and forward this
-information.
-
-\subsection{Improved support for mixed-type arithmetic throughout the
-  library}
-
-When evaluating finite element fields or their derivatives at
-a points $\mathbf x_q$, one typically has to do an operation of the form
-\begin{align*}
-  u_h(\mathbf x_q) &= \sum_{j} U_j \varphi_j(\mathbf x_q),
-  \\
-  \nabla u_h(\mathbf x_q) &= \sum_{j} U_j \nabla\varphi_j(\mathbf x_q).
-\end{align*}
-The value or derivatives of shape functions, $\varphi_j(\mathbf x_q)$
-or $\nabla\varphi_j(\mathbf x_q)$ are internally evaluated by classes
-derived from the \texttt{FiniteElement} and \texttt{Mapping}, and are
-computed as scalars or tensors of type \texttt{double}. On the other
-hand, the expansion coefficients of the field, $U_j$ are stored in
-vectors over scalar types chosen by the user; their underlying representation
-could be \texttt{double}, but also \texttt{float}, \texttt{long double},
-\texttt{PetscScalar}, or \texttt{std::complex<float>}. It could also
-be an autodifferentiation type.
-
-To facilitate the correct typing of the computed quantity, \dealii{}
-now contains mechanisms by which one can evaluate the \textit{type} of
-the product of two values, and this type is now consistently used
-throughout the library in expressions such as those above. Thus, as an
-example, if the user stores the expansion coefficients in a
-\texttt{Vector<long double>}, then the gradients
-$\nabla u_h(\mathbf x_q)$ will be computed as objects of type
-\texttt{Tensor<1,dim,long double>}. Likewise, if a user uses a PETSc
-vector, and PETSc was configured with complex scalar types, then the
-second derivatives of the solution field will be computed as
-\texttt{SymmetricTensor<2,dim,PetscScalar>}, which should equal
-\texttt{SymmetricTensor<2,dim,std::complex<double> >}.
-
-These improvements make type-correct computations possible in many
-places. In particular, this enables the use of complex-valued solution
-vectors in many more places than before. On the other hand, many but not all
-places have learned what actually to do with complex numbers. This is,
-in particular, true for the \texttt{DataOut} class that generates an
-intermediate data format that can then be written to graphical output
-files for visualization. Since no format we are aware of supports
-complex numbers, future versions still need to learn how to separate
-real and imaginary parts of complex numbers, and output them as two
-components.
-
-\subsection{A new ``shared'' triangulation type for parallel computations}
-
-\dealii{} already has two types of triangulations: the
-\texttt{Triangulation<dim,spacedim>} class works entirely locally, whereas
-\texttt{parallel::distributed::Triangulation<dim,spacedim>} builds on the
-former, but only stores the local partition that corresponds to a globally
-distributed triangulation managed across an MPI network. One \textit{could},
-however, use the former also for parallel computations in situations where one
-needs access to \textit{all} cells, not just the subset of cells that
-correspond to the partition owned by the current processor. This required
-building the same triangulation on all processors, then
-manually partitioning it (e.g., via \texttt{METIS}), calculating and storing
-index sets of locally owned and locally relevant degrees of freedom (DoFs),
-querying the \texttt{subdomain\_id}  of a cell during assembly, etc.
-
-In order to simplify this usage of the \texttt{Triangulation} class with MPI
-and to make its behavior in this context consistent with \texttt{parallel::distributed::Triangulation}, a new class
-\texttt{parallel::shared::Triangulation} has been introduced.
-It extends the \texttt{Triangulation} class to automatically partition the triangulation when run with MPI.
-Shared functionality between the \texttt{shared} and \texttt{distributed}
-triangulation classes
-(e.g., locally owned and relevant DoFs, MPI communicators, etc)
-is now grouped in a common parent class \texttt{parallel::Triangulation}.
-The main difference between the two classes is that in the case of
-\texttt{parallel::shared::Triangulation} each process stores all cells of the triangulation.
-Consequently, by default there are no artificial cells.
-That is, cells which are attributed to the current processor are marked as locally owned
-(\texttt{cell->is\_locally\_owned()} returns \texttt{true})
-and the rest are ghost cells.
-This behavior can be altered via an additional boolean flag provided to the constructor of the class.
-In this case, the set of ghost cells will consist of a halo layer of cells around locally owned cells.
-Cells which are neither ghost nor locally owned are marked as artificial.
-This is consistent with the behavior of \texttt{parallel::distributed::Triangulation},
-although in the latter case the size of the set of artificial cells will be
-much smaller.
-
-The introduction of the \texttt{parallel::shared::Triangulation} class together with the
-optional artificial cells and parent \texttt{parallel::Triangulation} class
-facilitates writing
-algorithms that are indifferent to the way a triangulation is stored in the MPI context.
-For example, the function \texttt{DoFTools::locally\_active\_dofs()} will
-return the appropriate subset of
-all DoF indices for both triangulations.
-Assmebly routines can use predicates such as \texttt{cell->is\_locally\_owned()}
-for both triangulations.
-
-Based on the new triangulation class, the (non-hp) \texttt{DoFHandler} manages
-degrees of freedom in the same way as it has already done for a long time for
-sequential and parallel distributed triangulations.
-
-\subsection{Second and third derivatives of finite element fields are now
-  computed exactly}
-
-Second derivatives of solution fields, i.e.,
-\begin{align*}
-  \nabla^2 u_h(\mathbf x_q) &= \sum_{j} U_j \nabla^2\varphi_j(\mathbf x_q)
-\end{align*}
-were previously computed by finite differencing of first
-derivatives. The reason for this approach is that in order to compute
-the second derivatives of shape functions, $\nabla^2\varphi_j(\mathbf
-x_q)$ one needs (at least) derivatives of the inverse of the Jacobian of the
-mapping. This is easy to see because, for the usual $Q_p$ Lagrange
-elements, one has that $\nabla\varphi_j(\mathbf
-x_q)=J^{-1}\hat\nabla\hat\varphi_j(\hat{\mathbf x}_q)$, where quantities
-with a hat refer to coordinates and functions on the reference cell,
-and $J$ is the Jacobian of the mapping from reference to real
-cell. Thus, the second derivatives satisfy
-\begin{align*}
-  \nabla^2\varphi_j(\mathbf x_q)
-  =
-  J^{-1}\hat\nabla\left[J^{-1}\hat\nabla\hat\varphi_j(\hat{\mathbf x}_q)\right]
-  =
-  J^{-1} J^{-1}\hat\nabla^2\hat\varphi_j(\hat{\mathbf x}_q)
-  +
-  J^{-1}\left(\hat\nabla [J^{-1}]\right)\hat\nabla\hat\varphi_j(\hat{\mathbf x}_q),
-\end{align*}
-where in the last expression, matrices and tensors of rank 1 and 3
-have to be appropriately contracted.  Here, the difficulty lies in
-computing the derivative $\hat\nabla [J^{-1}]$: for the usual mappings
-on quadrilaterals and hexahedra, $J$ is in general a polynomial in the
-reference coordinates $\hat {\mathbf x}$, so $J^{-1}$ is a rational
-function. It is possible to compute the derivatives of this object,
-but they are difficult and cumbersome to evaluate, especially for
-higher order mappings.
-
-The key to computing second (and higher) derivatives of shape
-functions is to recognize that $\hat\nabla [J^{-1}]$ can be expressed
-more conveniently by observing that
-\begin{align*}
-  0 &= \hat\nabla I
-  \\
-  &= \hat\nabla (JJ^{-1})
-  \\
-  &= J (\hat\nabla [J^{-1}]) + (\hat\nabla J) J^{-1},
-\end{align*}
-and consequently $\hat\nabla [J^{-1}] = - J^{-1} (\hat\nabla J)
-J^{-1}$. Here, $J^{-1}$ is a matrix that is already available from
-computing first derivatives, and $\hat\nabla J$ is an easily computed
-rank-3 tensor with polynomial entries. Using this approach, we have
-\begin{align*}
-  \nabla^2\varphi_j(\mathbf x_q)
-  =
-  J^{-1} J^{-1}\hat\nabla^2\hat\varphi_j(\hat{\mathbf x}_q)
-  -
-  J^{-1} J^{-1} (\hat\nabla J)
-  J^{-1}\hat\nabla\hat\varphi_j(\hat{\mathbf x}_q),
-\end{align*}
-again with an appropriate set of contractions over the indices of the
-objects on the right.
-
-This, and corresponding extensions to compute third derivatives, have
-now been implemented in several of the finite element and mapping
-classes by Maien Hamed, and are available through the
-\texttt{FEValues} interface to shape functions and their derivatives.
-
-\subsection{Visual C++ support}
-
-The library can now be compiled under Windows with Visual C++ 2013 and
-2015. The support is still experimental for the following reasons: First, we
-currently only support static linking. This will slow down linking of
-application code immensly.  Second, only a minimal testsuite is working, which
-is mainly because static linking of thousands of test executables is not
-viable. Therefore, we can not exclude the possibility of subtle bugs in the
-library.  Finally, there is of course limited support for external packages.
+\subsection{important thing 2}
 
@@
 
-\subsubsection{Revision of the interface between finite elements,
-  quadratures, and mappings}
-
-Finite element classes describe shape functions as continuous (as
-opposed to discrete) objects on the reference cell. On the other hand,
-in actual practice, one only needs information about shape functions
-at finitely many quadrature points, and these are typically the same
-on every cell in a loop over all cells. Furthermore, the evaluation of
-shape functions at quadrature points then also needs to be mapped to
-the real cell, typically using a polynomial mapping.
-
-To facilitate this complex interplay, \dealii{} has three class
-hierarchies rooted in the base classes \texttt{FiniteElement} (for the
-description of shape functions on the reference cell),
-\texttt{Quadrature} (for the locations and weights of quadrature
-points on the reference cell), and \texttt{Mapping} (for the
-description of mappings from the reference to the real cell). In
-almost all cases, users create an object of a derived class for each
-of these categories, but they never access any of the members of these
-classes and instead leave this task to the \texttt{FEValues} class
-(and \texttt{FEFaceValues}, \texttt{FESubfaceValues}) that presents
-all one typically needs: the mapped values, gradients, and
-higher order derivatives of shape functions at the quadrature points
-of the real cell.
-
-The interplay between these classes in the \texttt{FEValues} interface
-is one of the oldest parts of the library. It was, at the time, not
-designed based on specifications we explicitly or implicitly knew this
-class had to satisfy, but instead organically grew to its current
-state. These interfaces have now been fundamentally rewritten: some
-functions have been replaced; several others had their argument lists
-shuffled, sorted, and made more uniform; and everything has generally
-been far better documented.
-
-None of the changes in this arena is visible to the average user. The
-only user codes that are affected in an incompatible way are those
-that implement finite element or mapping classes.
+\subsubsection{incompatible change 1}
 
@@
 
-Older releases of \dealii{} can be cited as \cite{dealII80,dealII81,dealII82,dealII83}.
+Older releases of \dealii{} can be cited as \cite{dealII80,dealII81,dealII82,dealII83,dealII84}.
 
@@
 %
-Daniel Arndt,
-Mauro Bardelloni,
-Alistair Bentley,
-Andrea Bonito,
-Claire Bruna-Rosso,
-Krzysztof Bzowski,
-Praveen Chandrashekar,
-Conrad Clevenger,
-Patrick Esser,
-Rene Gassmoeller,
-Arezou Ghesmati,
-Maien Hamed,
-Alexander Grayver,
-Lukas Korous,
-Aslan Kosakian,
-Adam Kosik,
-Konstantin Ladutenko,
-Jean-Paul Pelteret,
-Lei Qiao,
-Gennadiy Rishin,
-Angel Rodriguez,
-Alberto Sartori,
-Daniel Shapero,
-Jason Sheldon,
-Jan Stebel,
-Florian Sonner,
-Zhen Tao,
-Heikki Virtanen,
-Daniel Weygand
 
@@
 
+\bigskip
+
 \dealii{} and its developers are financially supported through a
-variety of funding sources. W.~Bangerth and B.~Turcksin were partially
-supported by the National Science Foundation under award OCI-1148116
-as part of the Software Infrastructure for Sustained Innovation (SI2)
-program; and by the Computational Infrastructure in Geodynamics initiative
-(CIG), through the National Science Foundation under Award
-No.~EAR-0949446 and The University of California -- Davis.
+variety of funding sources:
 
-D.~Davydov was supported by the ERC Advanced Grant MOCOPOLY and the Competence Network for Technical and Scientific High Performance Computing in Bavaria (KONWIHR).
-
-L.~Heltai was partially supported by the project OpenViewSHIP,
-``Sviluppo di un ecosistema computazionale per la progettazione
-idrodinamica del sistema elica-carena'', financed by Regione FVG - PAR
-FSC 2007-2013, Fondo per lo Sviluppo e la Coesione, and by the project
-TRIM ``Tecnologia e Ricerca Industriale per la Mobilit\`a Marina'',
-CTN01-00176-163601, funded by MIUR -  Ministero dell'Istruzione,
-dell'Universit\`a e della Ricerca.
+% TODO: add your funding here
 
@@
 
-G.~Kanschat and M.~Kronbichler were partially supported by the German Research
-Foundation (DFG) through the project ExaDG. M.~Kronbichler also acknowledges
-the Gauss Centre for Supercomputing e.V.~for funding algorithm enhancements by
-providing computing time on the GCS Supercomputer SuperMUC at Leibniz
-Supercomputing Centre (LRZ) through project id pr83te.
-
 The Interdisciplinary Center for Scientific Computing (IWR) at Heidelberg University has provided
 hosting services for the \dealii{} web page and the SVN archive.
-
-%\marginpar{Everyone's funding, please}
 

### 2017-03-02 8.5/deal85.tex
Add my complete address.
List major news.

@@
 
-\hypersetup{
-  pdfauthor={Wolfgang Bangerth, Timo Heister},
-  pdftitle={The deal.II Library, Version 8.5, 2016},
-}
-
 \newcommand{\specialword}[1]{\texttt{#1}}
@@
 
+%
+% Author list -- please add yourself in both places below (in
+%                alphabetical order) if you think that your
+%                contributions to the last release warrant this
+%
+
+\hypersetup{
+  pdfauthor={
+    Wolfgang Bangerth,
+    Timo Heister
+  },
+  pdftitle={The deal.II Library, Version 8.5, 2017},
+}
+
 \title{The \dealii{} Library, Version 8.5}
+
 \author[1]{Wolfgang Bangerth}
-\affil[1]{xxx,
-    {\texttt{bangerth@math.tamu.edu}}}    
+\affil[1]{Department of Mathematics, Colorado State University, Fort
+  Collins, CO 80523-1874, USA.
+    {\texttt{bangerth@colostate.edu}}}
+
 \author[2]{Timo Heister}
 \affil[2]{Mathematical Sciences,
-  O-110 Martin Hall.
-  Clemson University.
-  Clemson, SC 29634, USA,
+  O-110 Martin Hall,
+  Clemson University,
+  Clemson, SC 29634, USA.
   {\texttt{heister@clemson.edu}}}
@@
 
+\marginpar{Update date}
 \dealii{} version 8.5.0 was released March 11, 2017. This paper provides an
@@
 \begin{itemize}
-\item 
-\item More than 140 other features and bugfixes.
+\item CellDataStorage class
+
+\item MappingManifold
+
+\item Trilinos linear operators?
+
+\item Physics module
+
+\item FE\_Enriched
+
+\item FESeries namespace, use in step-27
+
+\item step-57, step-55, step-56
+
+\item step-44 improvements
+
+\item step-37 parallel improvements
+
+\item New code gallery codes demonstrating (i) quasi-static visco-elastic
+  material behavior, (ii) multiphase Navier-Stokes flow, (iii) the
+  evolution of global-scale topography on planetary bodies, (iv)
+  goal-oriented elastoplasticity.
+
+\item More than 230 other features and bugfixes.
 \end{itemize}
@@
 can be found
-\href{https://www.dealii.org/8.4.0/doxygen/deal.II/changes_between_8_3_and_8_4.html}{in the file that lists all changes for this release} (see \cite{changes84})
+\href{https://www.dealii.org/8.5.0/doxygen/deal.II/changes_between_8_4_and_8_5.html}{in the file that lists all changes for this release} (see \cite{changes84})
 and that is linked to from the web site of each release as well as the
@@
 
+Other things we could talk about:
+PVS static checker, coverity, ...
+
+GridReordering rewrite
+
+No longer use features of C++ deprecated with C++11/14/17
+Move operators
+
+Python bindings
+
 \subsection{Incompatible changes}
@@
 
+Switch default of Lagrange elements to Gauss-Lobatto
+
 \subsubsection{Other incompatible changes}
 
+No long double instantiations any more
+
+ParameterGUI moved to separate repo
+
 The
-\href{https://www.dealii.org/8.4.0/doxygen/deal.II/changes_between_8_3_and_8_4.html}{file
-  that lists all changes for this release} (see \cite{changes84}) lists another 18
+\href{https://www.dealii.org/8.5.0/doxygen/deal.II/changes_between_8_4_and_8_5.html}{file
+  that lists all changes for this release} (see \cite{changes84})
+lists another around 20
 incompatible changes, but none of these should in fact be visible in

### 2017-03-03 8.5/paper.tex
List major changes.

--- 2017-03-02 8.5/deal85.tex
@@
 The major changes of this release are:
+\marginpar{Add one sentence to each of the following.}
 \begin{itemize}
@@
 
-\item MappingManifold
+\item The \texttt{MappingManifold} class provides mappings between the
+  reference cell and a mesh cell that is ``exact'' in the sense that
+  the edges, faces, and the interior of the mesh cell matches an
+  underlying manifold. This is opposed to using the usual polynomial
+  approximations of the manifold.
 
@@
 
-\item step-57, step-55, step-56
+\item New tutorial programs step-55, step-56, and step-57,
+  demonstrating parallel solvers for the Stokes equations, a multigrid
+  method for the Stokes equation, and methods for the nonlinear
+  Navier-Stokes equations, respectively. In addition, the step-37
+  tutorial program has been rewritten to support parallel
+  computations, and the step-44 tutorial now uses the facilities of
+  the physics module mentioned above.
 
-\item step-44 improvements
-
-\item step-37 parallel improvements
-
-\item New code gallery codes demonstrating (i) quasi-static visco-elastic
-  material behavior, (ii) multiphase Navier-Stokes flow, (iii) the
-  evolution of global-scale topography on planetary bodies, (iv)
-  goal-oriented elastoplasticity.
+\item New code gallery programs demonstrating (i) quasi-static
+  visco-elastic material behavior, (ii) multiphase Navier-Stokes flow,
+  (iii) the evolution of global-scale topography on planetary bodies,
+  (iv) goal-oriented elastoplasticity.
 

### 2017-03-03 8.5/paper.tex
List a few topics worth talking about.

@@
 \usepackage{anslistings}
+\usepackage{multicol}
 
@@
 
-\subsection{important thing 1}
+\subsection{The \texttt{CellDataStorage} class}
 
-\subsection{important thing 2}
+\subsection{The \texttt{MappingManifold} class}
 
-Other things we could talk about:
-PVS static checker, coverity, ...
+\subsection{Linear operators with Trilinos?}
 
-GridReordering rewrite
+\subsection{The physics module?}
 
-No longer use features of C++ deprecated with C++11/14/17
-Move operators
+\subsection{The \texttt{FE\_Enriched} class}
 
-Python bindings
+\subsection{The \texttt{FE\_Series} namespace?}
 
@@
 
-No long double instantiations any more
+No \texttt{long double} instantiations any more
 
@@
 to date information and bibtex snippets for this document see:
-
-\bigskip
-
 \begin{center}
@@
 \end{center}
-
-\bigskip
 
@@
 \dealii{} can interface with many other libraries:
+\begin{multicols}{2}
 \begin{itemize}
@@
 \end{itemize}
+\end{multicols}
 Please consider citing the appropriate references if you use interfaces to these

### 2017-03-03 8.5/paper.tex
Add people who have contributed code; add funding sources.

@@
 %
-% get this from the changes.h file using
-%   egrep '^ *\(.*201[56789]' 8.3.0-vs-8.4.0.h | perl -p -e 's/201\d.*//g; s/, */\n/g; s/^ *\(?//g;' | sort | uniq
-% then sort by last name and remove the authors of this paper
+% get this from the changes/*/* files using the command listed in the
+% release-tasks paper and remove the authors of this paper
 %
+Rajat  Arora, 
+Mauro  Bardelloni, 
+Conrad  Clevenger, 
+Sam  Cox, 
+Juliane  Dannberg, 
+Ren{\'e}  Gassm{\"o}ller, 
+Joscha  Gedicke, 
+Sebastian  Gonzalez-Pintor, 
+Ryan  Grove, 
+Michael  Harmon, 
+Daniel  Jodlbauer, 
+Guido  Kanschat, 
+Justin  Kauffman, 
+Paul  Kuberry, 
+Dustin  Kumor, 
+Konstantin  Ladutenko, 
+Andrew  McBride, 
+Mathias  Mentler, 
+Andrea  Mola, 
+Dragan  Nikolic, 
+Vaibhav  Palkar, 
+Spencer  Patty, 
+Jonathan  Perry-Houts, 
+Giuseppe  Pitton, 
+Ce  Qin, 
+Jonathan  Robey, 
+Mayank  Sabharwal, 
+Ali  Samii, 
+Alberto  Sartori, 
+Daniel  Shapero, 
+Martin  Steigemann, 
+Jihuan  Tian, 
+Jaeryun  Yim, 
+Toby  Young, 
+Zhao, Liang.
 
@@
 
-% TODO: add your funding here
+\marginpar{Add your funding here as appropriate}
+W.~Bangerth was partially
+supported by the National Science Foundation under award OCI-1148116
+as part of the Software Infrastructure for Sustained Innovation (SI2)
+program; and by the Computational Infrastructure in Geodynamics initiative
+(CIG), through the National Science Foundation under Awards
+No.~EAR-0949446 and EAR-1550901 and The University of California -- Davis.
 
 T.~Heister was partially supported by the Computational Infrastructure in
-Geodynamics initiative (CIG), through the National Science Foundation under Award No. EAR-0949446 and The University of California—Davis, and National Science Foundation grant DMS1522191.
+Geodynamics initiative (CIG), through the National Science Foundation
+under Award No. EAR-0949446 and The University of California -- Davis, and National Science Foundation grant DMS1522191.
 

### 2017-03-03 8.5/paper.tex
add myself to the authors

@@
     Wolfgang Bangerth,
+    Denis Davydov
     Timo Heister
@@
 
-\author[2]{Timo Heister}
-\affil[2]{Mathematical Sciences,
+\author[2]{Denis Davydov}
+\affil[2]{Chair of Applied Mechanics, University of
+     Erlangen-Nuremberg, Egerlandstr.\ 5, 91058 Erlangen, Germany.
+    {\texttt{denis.davydov@fau.de}}}
+
+\author[3]{Timo Heister}
+\affil[3]{Mathematical Sciences,
   O-110 Martin Hall,

### 2017-03-03 8.5/paper.tex
cite PUM paper
add Matrix-free operators subsection; mark a couple to be written

@@
 
-\subsection{The \texttt{FE\_Series} namespace?}
+TODO: Denis
+
+\subsection{The \texttt{FE\_Series} namespace}
+
+TODO: Denis
+
+\subsection{Matrix-free operators}
+
+Facilitate the usage of matrix-free method by providing a \verb!MatrixFreeOperator::Base! class,
+which implements various interface to matrix-vector products, including necessary operations when used in
+the context of the geometric multigrids, methods needed for usage within the linear operator as well as with Jacobi preconditioner. 
+The derived classes only need to implement \verb!apply_add()! method that is
+used in the \verb!vmult()! functions, and a method to compute the diagonal entries of the underlying matrix. 
+The \verb!MatrixFreeOperator! namespace contains implementations of \verb!MatrixFreeOperators::LaplaceOperator! and 
+\verb!MatrixFreeOperators::MassOperator!.
+
+TODO: Mention updated Step-37.
 
@@
  \item For $hp$ adaptivity: \cite{BangerthKayserHerold2007};
+  \item For $PUM$ and enrichment of the FE space: \cite{Davydov2016};
  \item For matrix-free and fast assembly techniques:

### 2017-03-03 8.5/paper.tex
add myself, update section
Add myself to list of authors; add placeholders for TODOs

@@
     Wolfgang Bangerth,
-    Denis Davydov
-    Timo Heister
+    Denis Davydov,
+    Timo Heister,
+    Matthias Maier,
+    Jean-Paul Pelteret
   },
@@
   {\texttt{heister@clemson.edu}}}
+
+\author[4]{Matthias Maier}
+\affil[4]{School of Mathematics,
+  University of Minnesota,
+  127 Vincent Hall, 206 Church Street SE,
+  Minneapolis, MN 55455, USA,
+  {\texttt{msmaier@umn.edu}}}
+
+\author[2]{Jean-Paul Pelteret}
+\affil[2]{Chair of Applied Mechanics, 
+  University of Erlangen-Nuremberg, 
+  Egerlandstr.\ 5, 
+  91058 Erlangen, 
+  Germany.
+  {\texttt{jean-paul.pelteret@fau.de}}}
 
@@
 \item CellDataStorage class
+  TODO: Denis and Jean-Paul
 
@@
 \item Trilinos linear operators?
+  TODO: Jean-Paul and Matthias
 
 \item Physics module
+  TODO: Denis and Jean-Paul
 
@@
 
+TODO: Denis and Jean-Paul
+
 \subsection{The \texttt{MappingManifold} class}
 
-\subsection{Linear operators with Trilinos?}
+\subsection{TODO: Linear operators with Payload and Trilinos}
+
+TODO: Jean-Paul and Matthias
 
 \subsection{The physics module?}
+
+TODO: Jean-Paul
 
@@
 
+J-P.~Pelteret was supported by the European Research Council (ERC) through the Advanced Grant 289049 MOCOPOLY.
+
 The Interdisciplinary Center for Scientific Computing (IWR) at Heidelberg University has provided

### 2017-03-03 8.5/paper.tex
Added section on Physics module.

@@
 
-\author[2]{Jean-Paul Pelteret}
-\affil[2]{Chair of Applied Mechanics, 
+\author[5]{Jean-Paul Pelteret}
+\affil[5]{Chair of Applied Mechanics, 
   University of Erlangen-Nuremberg, 
@@
 
-\item Physics module
-  TODO: Denis and Jean-Paul
+\item A dedicated physics module has been created to provide some standard definitions and operations used in continuum physics.
 
@@
 
-\subsection{The physics module?}
+\subsection{The physics module}
 
-TODO: Jean-Paul
+A dedicated physics module has been created to facilitate the implementation of functions and classes that relate to continuum physics, physical fields and material constitutive laws.
+To date, it includes transformations of scalar or tensorial quantities between any two configurations, and some definitions typically utilized in both linear and finite-strain nonlinear elasticity.
+
+The \verb!Physics::Transformations!
+namespace offers push-forward and pull-back operations in the context of contravariant, covariant and Piola transformations, as well as rotation operations for Euclidean space.
+In the \verb!Physics::Elasticity::Kinematics! namespace, a selection of deformation, strain tensors and strain rate tensors are defined.
+The \verb!Physics::Elasticity::StandardTensors!
+class provides some frequently used second and fourth order metric tensors, and defines some referential and spatial projection operators and tensor derivatives as commonly required in the definition of material laws.
 

### 2017-03-03 8.5/paper.tex
Added TODO for Luca.

@@
     Timo Heister,
+    Luca Heltai,
     Matthias Maier,
@@
 
-\author[4]{Matthias Maier}
-\affil[4]{School of Mathematics,
+\author[4]{Luca Heltai}
+\affil[4]{SISSA,
+  International School for Advanced Studies,
+  Via Bonomea 265, 
+  34136, Trieste, Italy
+{\texttt{luca.heltai@sissa.it}}}
+
+\author[5]{Matthias Maier}
+\affil[5]{School of Mathematics,
   University of Minnesota,
@@
 
-\author[5]{Jean-Paul Pelteret}
-\affil[5]{Chair of Applied Mechanics, 
+\author[6]{Jean-Paul Pelteret}
+\affil[6]{Chair of Applied Mechanics, 
   University of Erlangen-Nuremberg, 
@@
   approximations of the manifold.
+  TODO: Luca
 
@@
 \subsection{The \texttt{MappingManifold} class}
+
+TODO: Luca
 

### 2017-03-03 8.5/paper.tex
List updates to the GMG framework

--- 2017-03-03 8.5/paper.tex
@@
 \usepackage{multicol}
+
+\usepackage{pgfplots}
+\usepackage{pgfplotstable}
 
@@
     Wolfgang Bangerth,
-    Denis Davydov,
-    Timo Heister,
-    Matthias Maier,
-    Jean-Paul Pelteret
+    Denis Davydov
+    Timo Heister
+    Martin Kronbichler
   },
@@
 
-\author[4]{Matthias Maier}
-\affil[4]{School of Mathematics,
-  University of Minnesota,
-  127 Vincent Hall, 206 Church Street SE,
-  Minneapolis, MN 55455, USA,
-  {\texttt{msmaier@umn.edu}}}
-
-\author[2]{Jean-Paul Pelteret}
-\affil[2]{Chair of Applied Mechanics, 
-  University of Erlangen-Nuremberg, 
-  Egerlandstr.\ 5, 
-  91058 Erlangen, 
-  Germany.
-  {\texttt{jean-paul.pelteret@fau.de}}}
+\author[4]{Martin Kronbichler}
+\affil[4]{Institute for Computational Mechanics,
+  Technical University of Munich,
+  Boltzmannstr.~15, 85748 Garching, Germany.
+  {\texttt{kronbichler@lnm.mw.tum.de}}}
 
@@
 \item CellDataStorage class
-  TODO: Denis and Jean-Paul
 
@@
 \item Trilinos linear operators?
-  TODO: Jean-Paul and Matthias
 
 \item Physics module
-  TODO: Denis and Jean-Paul
 
@@
 
-TODO: Denis and Jean-Paul
-
 \subsection{The \texttt{MappingManifold} class}
 
-\subsection{TODO: Linear operators with Payload and Trilinos}
-
-TODO: Jean-Paul and Matthias
+\subsection{Linear operators with Trilinos?}
 
@@
 
-TODO: Jean-Paul
+\subsection{Scalability of geometric multigrid framework}
+
+For the new release, the geometric multigrid in \dealii{} have been thoroughly
+overhauled regarding their scalability on large-scale parallel computers. To
+this end, the geometric multigrid algorithm based on the fast matrix-free
+kernels from \cite{KronbichlerKormann2012} have been benchmarked up to
+147,456 cores. Several scalability bottlenecks have been removed, including
+unnecessary inner products inside the Chebyshev smoother and
+$\mathcal O(n_\text{levels})$ global communication steps during the
+restriction process rather than only the single global communication step that
+is necessary when going to the coarest grid. New matrix-free transfer
+implementations called \texttt{MGTransferMatrixFree} were devised that replace
+the matrix-based \texttt{MGTransferPrebuilt}. Besides better scalability than
+the Trilinos Epetra matrices underlying the latter, the matrix-free transfer
+is also a much better for high-order elements with complexity per degree of
+freedom of $\mathcal O(d p)$ in the polynomial degree $p$ in $d$ dimensions
+rather than $\mathcal O(p^d)$ for the matrix-based approach.
+
+\begin{figure}
+\pgfplotstableread{
+nprocs     fem256k    fem2m     fem16m    fem128m  fem1g    fem8g
+16         0.640934   4.97684   nan       nan      nan      nan
+32         0.325741   2.51771   nan       nan      nan      nan
+64         0.1645     1.2823    nan       nan      nan      nan
+128        0.090898   0.658832  5.00366   nan      nan      nan
+256        0.059922   0.339999  2.56216   nan      nan      nan
+512        0.0455449  0.176482  1.29986   nan      nan      nan
+1024       0.0368049  0.099691  0.67364   6.48155  nan      nan
+2048       0.0348921  0.069573  0.356066  2.59601  nan      nan
+4096       0.0367949  0.056833  0.19293   1.3251   nan      nan
+8192       0.033958   0.045350  0.110485  0.790214 5.50143  nan
+16384      0.0379629  0.049351  0.099904  0.424692 2.81479  nan
+32768      0.0461671  0.051276  0.077546  0.229114 1.53091  nan
+65536      0.0466189  0.058941  0.075194  0.127353 0.819    6.2909
+147456     nan        nan       nan       0.087949 0.43213  2.90856
+}\scalinglarge
+\pgfplotstableread{
+nprocs    newdg256k    newdg2m   newdg16m olddg256k  olddg2m   olddg16m
+28        1.4398       12.2014   nan      1.458725   12.25674  nan
+56        0.6987        6.2015   nan      0.721993   6.266567  nan
+112       0.3352        3.1782   nan      0.353412   3.263585  nan
+224       0.1888        1.5687   12.8235  0.183643   1.611967  13.03189
+448       0.0853        0.7686   6.5649   0.105388   0.805367  6.700762
+896       0.0478        0.3537   3.3685   0.063270   0.383875  3.446427
+1792      0.0317        0.1717   1.7061   0.046065   0.199738  1.762310
+3584      0.0235        0.1079   0.8580   0.042032   0.123113  0.902782
+7168      0.0207        0.0668   0.4328   0.037678   0.091822  0.462007
+14336     0.0183699     0.045095 0.23276  0.038804   0.071644  0.288911
+}\scalingHSW
+\centering
+\definecolor{gnuplot@green}{RGB}{0,158,115}
+\begin{tikzpicture}
+    \begin{loglogaxis}[
+      title style={at={(0.5,0.965)},anchor=north,draw=black,fill=white,font=\scriptsize\bf},
+      title={strong and weak scaling, continuous $\mathcal Q_3$ elements},
+      width=0.53\textwidth,
+      height=0.5\textwidth,
+      xlabel={Number of cores},
+      ylabel={Solver time [s]},
+      x label style={at={(0.5,0.02)}},
+      y label style={at={(0.05,0.5)}},
+      xtick={32,128,512,2048,8192,32768,147456},
+      xticklabels={32,128,512,2048,8192,32k,147k},
+      tick label style={font=\scriptsize},
+      label style={font=\scriptsize},
+      legend style={font=\scriptsize},
+      legend pos=south west,
+      ymin=5e-3, ymax=15,
+      xmin=8, xmax=147456,
+      grid
+      ]
+      \addplot table[x={nprocs}, y={fem8g}] {\scalinglarge};
+      \addlegendentry{8B cells};
+      \addplot table[x={nprocs}, y={fem1g}] {\scalinglarge};
+      \addlegendentry{1B cells};
+      \addplot table[x={nprocs}, y={fem128m}] {\scalinglarge};
+      \addlegendentry{128M cells};
+      \addplot table[x={nprocs}, y={fem16m}] {\scalinglarge};
+      \addlegendentry{16M cells};
+      \addplot[gnuplot@green,mark=diamond*,mark options={fill=gnuplot@green!40}] table[x={nprocs}, y={fem2m}] {\scalinglarge};
+      \addlegendentry{2M cells};
+      \addplot[dashed,black] coordinates {
+        (8,10)
+        (147456,5/9168)
+      };
+      \addlegendentry{linear scaling};
+      \addplot[dashed,black] coordinates {
+        (16,8*5)
+        (147456,8*5/9168)
+      };
+      \addplot[dashed,black] coordinates {
+        (16,64*5)
+        (147456,64*5/9168)
+      };
+      \addplot[dashed,black] coordinates {
+        (16,5*512)
+        (147456,5*512/9168)
+      };
+      \addplot[dashed,black] coordinates {
+        (16,5*4096)
+        (147456,5*4096/9168)
+      };
+    \end{loglogaxis}
+  \end{tikzpicture}
+  \begin{tikzpicture}
+    \begin{loglogaxis}[
+      title style={at={(1,0.965)},anchor=north east,draw=black,fill=white,font=\scriptsize\bf},
+      title={$256^3$ mesh, discontinuous $\mathcal Q_3$ elements},
+      width=0.48\textwidth,
+      height=0.5\textwidth,
+      xlabel={Number of cores},
+      x label style={at={(0.5,0.02)}},
+      xtick={56,224,896,3584,14336},
+      xticklabels={56,224,896,3584,14336},
+      tick label style={font=\scriptsize},
+      label style={font=\scriptsize},
+      legend style={font=\scriptsize},
+      legend pos=south west,
+      xmin=28, xmax=14336,
+      ymin=5e-3, ymax=15,
+      grid
+      ]
+      \addplot[blue,mark=*,densely dashed] table[x={nprocs}, y={olddg16m}] {\scalingHSW};
+      \addlegendentry{old, 16M cells};
+      \addplot[blue,mark=o] table[x={nprocs}, y={newdg16m}] {\scalingHSW};
+      \addlegendentry{new, 16M cells};
+      \addplot[red,mark=square*,densely dashed] table[x={nprocs}, y={olddg2m}] {\scalingHSW};
+      \addlegendentry{old, 2M cells};
+      \addplot[red,mark=square] table[x={nprocs}, y={newdg2m}] {\scalingHSW};
+      \addlegendentry{new, 2M cells};
+      \addplot[gnuplot@green,mark=diamond*,densely dashed] table[x={nprocs}, y={olddg256k}] {\scalingHSW};
+      \addlegendentry{old, 256k cells};
+      \addplot[gnuplot@green,mark=diamond] table[x={nprocs}, y={newdg256k}] {\scalingHSW};
+      \addlegendentry{new, 256k cells};
+    \end{loglogaxis}
+  \end{tikzpicture}
+  \caption{Scaling of multigrid algorithms on SuperMUC.}
+\label{fig:scaling_mg}
+\end{figure}
+
+The scalability of the improved geometric multigrid framework is shown in
+Fig.~\ref{fig:scaling_mg}, including a combined strong and weak scalability
+plot in the left panel using continuous $\mathcal Q_3$ elements on 57 million
+to 232 billion degrees of freedom for the Laplacian. Along each line, the same
+problem size is solved with an increasing number of cores, whereas different
+lines always start out at 3.5 million degrees of freedom per core. Almost
+ideal scalability down to approximately 0.1 seconds can be observed also on
+147k cores. The right panel of Fig.~\ref{fig:scaling_mg} shows the effect of
+the aforementioned algorithmic improvements on a setup with discontinuous DG
+elements, clearly improving the latency of the multigrid V-cycle. The improved
+algorithms are shown in the updated step-37 tutorial program.
 
@@
 
-TODO: Mention updated Step-37.
+The framework was also included in the updated step-37 tutorial program.
 
@@
 
-J-P.~Pelteret was supported by the European Research Council (ERC) through the Advanced Grant 289049 MOCOPOLY.
+M.~Kronbichler was partially supported by the German Research Foundation (DFG)
+under the project ``High-order discontinuous Galerkin for the exa-scale''
+(ExaDG) within the priority program ``Software for Exascale Computing''
+(SPPEXA), grant agreement no.~KR4661/2-1, the Bayerisches Kompetenznetzwerk
+f\"ur Technisch-Wissenschaftliches Hoch- und H\"ochstleistungsrechnen
+(KONWIHR), and the Gauss Centre for Supercomputing e.V.~by providing computing
+time on the GCS Supercomputer SuperMUC at Leibniz Supercomputing Centre (LRZ)
+through project id pr83te.
 

### 2017-03-03 8.5/paper.tex

@@
     Wolfgang Bangerth,
-    Denis Davydov
-    Timo Heister
-    Martin Kronbichler
+    Denis Davydov,
+    Timo Heister,
+    Luca Heltai,
+    Martin Kronbichler,
+    Matthias Maier,
+    Jean-Paul Pelteret
   },
@@
 
-\author[4]{Martin Kronbichler}
-\affil[4]{Institute for Computational Mechanics,
+\author[4]{Luca Heltai}
+\affil[4]{SISSA,
+  International School for Advanced Studies,
+  Via Bonomea 265,
+  34136, Trieste, Italy
+{\texttt{luca.heltai@sissa.it}}}
+
+\author[5]{Martin Kronbichler}
+\affil[5]{Institute for Computational Mechanics,
   Technical University of Munich,
@@
   {\texttt{kronbichler@lnm.mw.tum.de}}}
+
+\author[6]{Matthias Maier}
+\affil[6]{School of Mathematics,
+  University of Minnesota,
+  127 Vincent Hall, 206 Church Street SE,
+  Minneapolis, MN 55455, USA,
+  {\texttt{msmaier@umn.edu}}}
+
+\author[7]{Jean-Paul Pelteret}
+\affil[7]{Chair of Applied Mechanics,
+  University of Erlangen-Nuremberg,
+  Egerlandstr.\ 5,
+  91058 Erlangen,
+  Germany.
+  {\texttt{jean-paul.pelteret@fau.de}}}
 
@@
   approximations of the manifold.
+  TODO: Luca
 
 \item Trilinos linear operators?
+  TODO: Jean-Paul and Matthias
 
-\item Physics module
+\item A dedicated physics module has been created to provide some standard definitions and operations used in continuum physics.
 
@@
 
+TODO: Denis and Jean-Paul
+
 \subsection{The \texttt{MappingManifold} class}
 
-\subsection{Linear operators with Trilinos?}
+TODO: Luca
 
-\subsection{The physics module?}
+\subsection{TODO: Linear operators with Payload and Trilinos}
+
+TODO: Jean-Paul and Matthias
+
+\subsection{The physics module}
+
+A dedicated physics module has been created to facilitate the implementation of functions and classes that relate to continuum physics, physical fields and material constitutive laws.
+To date, it includes transformations of scalar or tensorial quantities between any two configurations, and some definitions typically utilized in both linear and finite-strain nonlinear elasticity.
+
+The \verb!Physics::Transformations!
+namespace offers push-forward and pull-back operations in the context of contravariant, covariant and Piola transformations, as well as rotation operations for Euclidean space.
+In the \verb!Physics::Elasticity::Kinematics! namespace, a selection of deformation, strain tensors and strain rate tensors are defined.
+The \verb!Physics::Elasticity::StandardTensors!
+class provides some frequently used second and fourth order metric tensors, and defines some referential and spatial projection operators and tensor derivatives as commonly required in the definition of material laws.
 

### 2017-03-03 8.5/paper.tex
Improve text

@@
 
-For the new release, the geometric multigrid in \dealii{} have been thoroughly
-overhauled regarding their scalability on large-scale parallel computers. To
-this end, the geometric multigrid algorithm based on the fast matrix-free
-kernels from \cite{KronbichlerKormann2012} have been benchmarked up to
-147,456 cores. Several scalability bottlenecks have been removed, including
-unnecessary inner products inside the Chebyshev smoother and
-$\mathcal O(n_\text{levels})$ global communication steps during the
-restriction process rather than only the single global communication step that
-is necessary when going to the coarest grid. New matrix-free transfer
-implementations called \texttt{MGTransferMatrixFree} were devised that replace
-the matrix-based \texttt{MGTransferPrebuilt}. Besides better scalability than
-the Trilinos Epetra matrices underlying the latter, the matrix-free transfer
-is also a much better for high-order elements with complexity per degree of
-freedom of $\mathcal O(d p)$ in the polynomial degree $p$ in $d$ dimensions
-rather than $\mathcal O(p^d)$ for the matrix-based approach.
+For the new release, the geometric multigrid facilities in \dealii{} have been
+thoroughly overhauled regarding their scalability on large-scale parallel
+computers. During this process, a geometric multigrid implementation based on
+the fast matrix-free kernels from \cite{KronbichlerKormann2012} has been
+benchmarked up to 147,456 cores. The fast matrix-vector product revealed
+several scalability bottlenecks, including unnecessary inner products inside
+the Chebyshev smoother and $\mathcal O(n_\text{levels})$ global communication
+steps during the restriction process rather than only the single global
+communication step that is necessary when going to the coarest grid. New
+matrix-free transfer implementations called \texttt{MGTransferMatrixFree} were
+devised that can replace the matrix-based \texttt{MGTransferPrebuilt} class
+for tensor product elements. Besides better scalability than the Trilinos
+Epetra matrices underlying the latter, the matrix-free transfer is also a much
+better for high-order elements with complexity per degree of freedom of
+$\mathcal O(d p)$ in the polynomial degree $p$ in $d$ dimensions rather than
+$\mathcal O(p^d)$ for the matrix-based approach.
 
@@
 The scalability of the improved geometric multigrid framework is shown in
-Fig.~\ref{fig:scaling_mg}, including a combined strong and weak scalability
-plot in the left panel using continuous $\mathcal Q_3$ elements on 57 million
-to 232 billion degrees of freedom for the Laplacian. Along each line, the same
-problem size is solved with an increasing number of cores, whereas different
-lines always start out at 3.5 million degrees of freedom per core. Almost
-ideal scalability down to approximately 0.1 seconds can be observed also on
-147k cores. The right panel of Fig.~\ref{fig:scaling_mg} shows the effect of
-the aforementioned algorithmic improvements on a setup with discontinuous DG
-elements, clearly improving the latency of the multigrid V-cycle. The improved
-algorithms are shown in the updated step-37 tutorial program.
+Fig.~\ref{fig:scaling_mg}, including a combined strong and weak scaling plot
+in the left panel using continuous $\mathcal Q_3$ elements with 57~million to
+232~billion degrees of freedom for discretizing the Laplacian. Along each
+line, the same problem size is solved with an increasing number of cores,
+whereas different lines are a factor of eight apart and always start at
+3.5~million degrees of freedom per core. Almost ideal scalability down to
+approximately 0.1~seconds can be observed also on 147k~cores. The right panel
+of Fig.~\ref{fig:scaling_mg} shows the effect of the aforementioned
+algorithmic improvements on a setup with discontinuous DG elements, clearly
+improving the latency of the multigrid V-cycle. The improved algorithms are
+shown in the updated step-37 tutorial program.
 

### 2017-03-03 8.5/paper.tex
Add contribution for CellStorageData

@@
 \begin{itemize}
-\item CellDataStorage class
-  TODO: Denis and Jean-Paul
+\item The \texttt{CellDataStorage} class provides a mechanism to store user-defined data on each cell.
+This data is treated as a first-class citizen to \dealii{} and, when used in conjunction with other classes, can be shipped to other MPI cores during mesh refinement and repartitioning.
 
@@
 
-\subsection{The \texttt{CellDataStorage} class}
+\subsection{The \texttt{CellDataStorage} class and friends}
 
-TODO: Denis and Jean-Paul
+The \texttt{CellDataStorage} class is an integrated mechanism to safely store user-defined data, such as that required at each computation point, within each cell.
+Although the same was previously achieved through the use of a cell \texttt{user\_pointer}, it required users to manage this data themselves.
+As this data is now treated as a first-class citizen to \dealii{}, through the abstract \texttt{TransferableQuadraturePointData} we have provided a generic interface to facilitate several low-level operations that were previously tedious or technically challenging to implement.
+For instance, the \texttt{parallel::distributed::ContinuousQuadratureDataTransfer} class assists in the transfer of arbitrary data (that is continuous within a cell) stored at quadrature points when performing h-adaptive refinement of \texttt{parallel::distributed::Triangulation}.
+Not only does it perform an $\mathcal{L}^2$ projection of the specified user data between quadrature points, but it also ships the data automatically between MPI processes.
 

### 2017-03-03 8.5/paper.tex
FE_Enriched

@@
 
-TODO: Denis
+\verb!FE_Enriched!
+finite element implements a partition of unity finite element method (PUM) by Babuska and Melenk which enriches a standard finite element with an enrichment function multiplied with another (usually linear) finite element.
+This allows to include in the finite element space a priori knowledge about the partial differential equation being solved which in turn improves the local approximation properties of the spaces.
+The user can also use enriched and non-enriched finite elements in different parts of the domain.
+\verb|!DoFTools::make_hanging_node_constraints()| function can automatically make the resulting space $C^0$ continuous.
+Existing \verb|SolutionTransfer| class can be used to transfer the solution during $h$\,-adaptive refinement from a coarse to a fine mesh under the condition that all child elements are also enriched.
 

### 2017-03-03 8.5/paper.tex
FESeries

@@
 
-\subsection{The \texttt{FE\_Series} namespace}
+\subsection{The \texttt{FESeries} namespace}
 
-TODO: Denis
+\verb|FESeries| namespace offers functions to calculate expansion series of the solution on the reference element.
+Coefficients of expansion are often used to estimate local smoothness of the underlying finite element field to decide on h- or p-adaptive refinement strategy.
+Specifically, \verb|FESeries::Legendre| calculates expansion of a scalar FE field into series of Legendre functions on a reference element, whereas \verb|FESeries::Fourier| calculates Fourier coefficients.
+The user has to specify the required number of coefficients in each direction as well as provide a collection of finite elements and quadrature rules, that will be used for calculation of the transformation matrices.
+Obviously, those matrices are calculated only once, namely the first time coefficients are requested on a cell with the given finite element.
+The step-27 has been updated to use \verb|FESeries::Fourier|.
 

### 2017-03-03 8.5/paper.tex
funding for myself

@@
 
-\item Trilinos linear operators?
-  TODO: Jean-Paul and Matthias
+\item The \texttt{LinearOperator} class has been extended to fully support {\trilinos} solvers.
 
@@
 
-\subsection{TODO: Linear operators with Payload and Trilinos}
+\subsection{Extension of the \texttt{LinearOperator} class}
 
-TODO: Jean-Paul and Matthias
+By introduction of an arbitrary \texttt{Payload} base class, the \texttt{LinearOperator} class can be extended to seamlessly exploit the native features and operations offered by external linear algebra libraries.
+We have thus developed a \texttt{TrilinosPayload} class, through which full support for the suite of {\trilinos} parallel iterative solvers and preconditioners has been achieved.
+In the case of the construction of an \texttt{inverse\_operator}, both \dealii's built in solvers and those offered by {\trilinos} can be selected.
+This has been achieved by using the \texttt{Epetra\_Operator} as the basis for the \texttt{TrilinosPayload}, for which the result of both standard and composite operations involving forward (\texttt{Apply()}) and inverse (\texttt{ApplyInverse()}) matrix-vector multiplication are collated using lambda functions.
+It is envisaged that, in the future, similar can be done for the {\petsc} iterative solvers.
+
+An additional feature of the \texttt{LinearOperator} suite is the defintion of a \texttt{schur\_complement} operator, and its associated condensation and post-processing \texttt{PackagedOperation}s.
+An operator representing the Schur complement of a block system can be declared and, though the delayed evaluation offered by \texttt{PackagedOperation}s, reused on any number of vector systems.
 
@@
 
+D.
+Davydov was supported by the European Research Council (ERC) through the Advanced Grant 289049 MOCOPOLY and the Competence Network for Technical and Scientific High Performance Computing in Bavaria (KONWIHR).
+
 T.~Heister was partially supported by the Computational Infrastructure in

### 2017-03-03 8.5/paper.tex
Reference the correct changes file.

@@
 can be found
-\href{https://www.dealii.org/8.5.0/doxygen/deal.II/changes_between_8_4_and_8_5.html}{in the file that lists all changes for this release} (see \cite{changes84})
+\href{https://www.dealii.org/8.5.0/doxygen/deal.II/changes_between_8_4_and_8_5.html}{in the file that lists all changes for this release} (see \cite{changes85})
 and that is linked to from the web site of each release as well as the
@@
 \href{https://www.dealii.org/8.5.0/doxygen/deal.II/changes_between_8_4_and_8_5.html}{file
-  that lists all changes for this release} (see \cite{changes84})
+  that lists all changes for this release} (see \cite{changes85})
 lists another around 20

### 2017-03-04 8.5/paper.tex
Extend text

@@
 
+\item Various improvements for high-order elements, including a switch of
+  support points in \texttt{FE\_Q} and \texttt{FE\_DGQ} to Gauss--Lobatto
+  support points, more stable evaluation of Legendre polynomials, and several
+  bugfixes for high-order polynomial mappings defined through the
+  \texttt{MappingQ} class.
+
 \item More than 230 other features and bugfixes.
@@
 thoroughly overhauled regarding their scalability on large-scale parallel
-computers. During this process, a geometric multigrid implementation based on
+systems. During this process, a geometric multigrid implementation based on
 the fast matrix-free kernels from \cite{KronbichlerKormann2012} has been
-benchmarked up to 147,456 cores. The fast matrix-vector product revealed
-several scalability bottlenecks, including unnecessary inner products inside
-the Chebyshev smoother and $\mathcal O(n_\text{levels})$ global communication
-steps during the restriction process rather than only the single global
-communication step that is necessary when going to the coarest grid. New
+benchmarked on up to 147,456 cores. The fast matrix-vector products revealed
+several scalability bottlenecks in the other multigrid components, including
+unnecessary inner products inside the Chebyshev smoother and
+$\mathcal O(n_\text{levels})$ global communication steps during the
+restriction process rather than the single unavoidable global communication
+step inherent to going to the coarsest grid and the coarse solve. New
 matrix-free transfer implementations called \texttt{MGTransferMatrixFree} were
@@
 Epetra matrices underlying the latter, the matrix-free transfer is also a much
-better for high-order elements with complexity per degree of freedom of
+better for high-order elements with a complexity per degree of freedom of
 $\mathcal O(d p)$ in the polynomial degree $p$ in $d$ dimensions rather than
-$\mathcal O(p^d)$ for the matrix-based approach.
+$\mathcal O(p^d)$ for the matrices.
 
@@
   \end{tikzpicture}
-  \caption{Scaling of multigrid algorithms on SuperMUC.}
+  \caption{Scaling of \dealii{}'s geometric multigrid algorithms on SuperMUC.}
 \label{fig:scaling_mg}
@@
 whereas different lines are a factor of eight apart and always start at
-3.5~million degrees of freedom per core. Almost ideal scalability down to
-approximately 0.1~seconds can be observed also on 147k~cores. The right panel
-of Fig.~\ref{fig:scaling_mg} shows the effect of the aforementioned
+3.5~million degrees of freedom per core with an absolute performance of around
+650,000 degrees of freedom per core and second. Almost ideal scalability down
+to approximately 0.1~seconds can be observed also on 147k~cores. The right
+panel of Fig.~\ref{fig:scaling_mg} shows the effect of the aforementioned
 algorithmic improvements on a setup with discontinuous DG elements, clearly
-improving the latency of the multigrid V-cycle. The improved algorithms are
-shown in the updated step-37 tutorial program.
+improving the latency of the multigrid V-cycle. The updated step-37 tutorial
+programs presents the updated algorithms and the MPI-parallel multigrid
+setting with matrix-free operator evaluation.
 
@@
 The user can also use enriched and non-enriched finite elements in different parts of the domain.
-\verb|!DoFTools::make_hanging_node_constraints()| function can automatically make the resulting space $C^0$ continuous.
+\verb|DoFTools::make_hanging_node_constraints()| function can automatically make the resulting space $C^0$ continuous.
 Existing \verb|SolutionTransfer| class can be used to transfer the solution during $h$\,-adaptive refinement from a coarse to a fine mesh under the condition that all child elements are also enriched.
@@
 
-Facilitate the usage of matrix-free method by providing a \verb!MatrixFreeOperator::Base! class,
-which implements various interface to matrix-vector products, including necessary operations when used in
-the context of the geometric multigrids, methods needed for usage within the linear operator as well as with Jacobi preconditioner.
-The derived classes only need to implement \verb!apply_add()! method that is
+In order to facilitate the usage of matrix-free methods, a \verb!MatrixFreeOperator::Base! class has been introduced,
+implementing various interfaces to matrix-vector products and the necessary operations for the interface residuals according to \cite{JanssenKanschat2011} in
+the context of the geometric multigrids. Furthermore, the class is compatible with the linear operator framework and provides an interface to a Jacobi preconditioner.
+The derived classes only need to implement the \verb!apply_add()! method that is
 used in the \verb!vmult()! functions, and a method to compute the diagonal entries of the underlying matrix.
@@
 \verb!MatrixFreeOperators::MassOperator!.
-
-The framework was also included in the updated step-37 tutorial program.
+The updated step-37 tutorial program makes use of these facilities and
+explains their usage in detail. Using the matrix-free mass operator,
+\verb!VectorTools::project! has become much faster than the previous
+matrix-based approach and also works in parallel with MPI.
 
@@
 
-Switch default of Lagrange elements to Gauss-Lobatto
+High-order Lagrange elements, both continuous \verb!FE_Q! and discontinuous
+\verb!FE_DGQ! types, now use the nodal points of the Gauss--Lobatto
+quadrature formula as support points, rather than the previous equidistant
+ones. For cubic polynomials and higher, the point distribution has changed and
+thus the entries in solution vectors will look different now. Note, however,
+that using the Gauss--Lobatto points as nodal points results in much more
+stable interpolation, including better iteration counts in most iterative
+solvers.
 

### 2017-03-05 8.5/paper.tex
add GSL
Small text improvement
add myself

@@
 \usepackage{graphicx}
+
+\pgfplotsset{compat=1.12}
 
@@
   pdfauthor={
+    Daniel Arndt,
     Wolfgang Bangerth,
@@
 
-\author[1]{Wolfgang Bangerth}
-\affil[1]{Department of Mathematics, Colorado State University, Fort
+\author[1]{Daniel Arndt}
+\affil[1]{Interdisciplinary Center for Scientific Computing,
+    Heidelberg University, Im Neuenheimer Feld 205, 69120 Heidelberg, Germany.
+    {\texttt{daniel.arndt@iwr.uni-heidelberg.de}}}
+
+\author[2]{Wolfgang Bangerth}
+\affil[2]{Department of Mathematics, Colorado State University, Fort
   Collins, CO 80523-1874, USA.
@@
 
-\author[2]{Denis Davydov}
-\affil[2]{Chair of Applied Mechanics, University of
+\author[3]{Denis Davydov}
+\affil[3]{Chair of Applied Mechanics, University of
      Erlangen-Nuremberg, Egerlandstr.\ 5, 91058 Erlangen, Germany.
@@
 
-\author[3]{Timo Heister}
-\affil[3]{Mathematical Sciences,
+\author[4]{Timo Heister}
+\affil[4]{Mathematical Sciences,
   O-110 Martin Hall,
@@
 
-\author[4]{Luca Heltai}
-\affil[4]{SISSA,
+\author[5]{Luca Heltai}
+\affil[5]{SISSA,
   International School for Advanced Studies,
@@
 
-\author[5]{Martin Kronbichler}
-\affil[5]{Institute for Computational Mechanics,
+\author[6]{Martin Kronbichler}
+\affil[6]{Institute for Computational Mechanics,
   Technical University of Munich,
@@
 
-\author[6]{Matthias Maier}
-\affil[6]{School of Mathematics,
+\author[7]{Matthias Maier}
+\affil[7]{School of Mathematics,
   University of Minnesota,
@@
 
-\author[7]{Jean-Paul Pelteret}
-\affil[7]{Chair of Applied Mechanics,
+\author[8]{Jean-Paul Pelteret}
+\affil[8]{Chair of Applied Mechanics,
   University of Erlangen-Nuremberg,
@@
 \verb!VectorTools::project! has become much faster than the previous
-matrix-based approach and also works in parallel with MPI.
+matrix-based approach for elements supported by \verb!MatrixFree! and also
+works in parallel with MPI.
 
@@
 \item BLAS, LAPACK
+\item GSL \cite{gsl2016}
 \item HDF5 \cite{hdf5}
@@
 \marginpar{Add your funding here as appropriate}
+D.~Arndt was supported by the German Research Foundation (DFG) under the
+project ``High-order discontinuous Galerkin for the exa-scale'' (ExaDG) within the
+priority program ``Software for Exascale Computing'' (SPPEXA).
+
 W.~Bangerth was partially
@@
 
-D.
-Davydov was supported by the European Research Council (ERC) through the Advanced Grant 289049 MOCOPOLY and the Competence Network for Technical and Scientific High Performance Computing in Bavaria (KONWIHR).
+D.~Davydov was supported by the European Research Council (ERC) through the Advanced Grant 289049 MOCOPOLY and the Competence Network for Technical and Scientific High Performance Computing in Bavaria (KONWIHR).
 

### 2017-03-05 8.5/paper.tex
Improve text

@@
   support points in \texttt{FE\_Q} and \texttt{FE\_DGQ} to Gauss--Lobatto
-  support points, more stable evaluation of Legendre polynomials, and several
+  support points, stable evaluation of high-order Legendre polynomials, and several
   bugfixes for high-order polynomial mappings defined through the
@@
       ylabel={Solver time [s]},
-      x label style={at={(0.5,0.02)}},
-      y label style={at={(0.05,0.5)}},
       xtick={32,128,512,2048,8192,32768,147456},
@@
       title style={at={(1,0.965)},anchor=north east,draw=black,fill=white,font=\scriptsize\bf},
-      title={$256^3$ mesh, discontinuous $\mathcal Q_3$ elements},
+      title={discontinuous $\mathcal Q_3$ elements},
       width=0.48\textwidth,
@@
       xlabel={Number of cores},
-      x label style={at={(0.5,0.02)}},
       xtick={56,224,896,3584,14336},
@@
 quadrature formula as support points, rather than the previous equidistant
-ones. For cubic polynomials and higher, the point distribution has changed and
-thus the entries in solution vectors will look different now. Note, however,
-that using the Gauss--Lobatto points as nodal points results in much more
+ones. For cubic polynomials and higher, the point distribution has thus changed and
+thus the entries in solution vectors will look different as compared to the previous version. Note, however,
+that using the Gauss--Lobatto points as nodal points results in a much more
 stable interpolation, including better iteration counts in most iterative

### 2017-03-06 8.5/paper.tex
add information about static code analyzers and small style fixes

@@
 
-\item FE\_Enriched
+\item \texttt{FE\_Enriched}
 
-\item FESeries namespace, use in step-27
+\item \texttt{FESeries} namespace, use in step-27
 
@@
 
-\item New code gallery programs demonstrating (i) quasi-static
-  visco-elastic material behavior, (ii) multiphase Navier-Stokes flow,
-  (iii) the evolution of global-scale topography on planetary bodies,
-  (iv) goal-oriented elastoplasticity.
+\item New code gallery programs demonstrating
+  \begin{enumerate}
+    \item[(i)] quasi-static visco-elastic material behavior, 
+    \item[(ii)] multiphase Navier-Stokes flow,
+    \item[(iii)] the evolution of global-scale topography on planetary bodies,
+    \item[(iv)] goal-oriented elastoplasticity.
+  \end{enumerate}
 
@@
 
-\item More than 230 other features and bugfixes.
+\item More than 230 other features and bugfixes.\\
+  We realized that static code analyzers are valuable tools to improve and maintain the quality of the code in our library 
+  on top off build and regression tests for a variety of setups using \href{https://cdash.kyomu.43-1.org/index.php?project=deal.II}{CDash}.
+  This release was tested with \href{http://cppcheck.sourceforge.net/}{Cppcheck}, 
+  \href{https://www.viva64.com/en/pvs-studio/}{PVS-Studio} and \href{https://scan.coverity.com/}{Coverity-Scan}. 
+  For the latter the current status can be observed \href{https://scan.coverity.com/projects/deal-ii}{online}.
 \end{itemize}
@@
       title={strong and weak scaling, continuous $\mathcal Q_3$ elements},
-      width=0.53\textwidth,
+      width=0.52\textwidth,
       height=0.5\textwidth,
@@
 globe. Other than the authors of this paper, the following people contributed code to
-this release:
+this release:\\
 %

### 2017-03-09 8.5/paper.tex
unify punctuation

@@
   Via Bonomea 265,
-  34136, Trieste, Italy
+  34136, Trieste, Italy.
 {\texttt{luca.heltai@sissa.it}}}
@@
   127 Vincent Hall, 206 Church Street SE,
-  Minneapolis, MN 55455, USA,
+  Minneapolis, MN 55455, USA.
   {\texttt{msmaier@umn.edu}}}

### 2017-03-09 8.5/paper.tex
Reformat and improve introduction

@@
 \begin{itemize}
-\item The \texttt{CellDataStorage} class provides a mechanism to store user-defined data on each cell.
-This data is treated as a first-class citizen to \dealii{} and, when used in conjunction with other classes, can be shipped to other MPI cores during mesh refinement and repartitioning.
+\item The \texttt{CellDataStorage} class provides a mechanism to store
+  user-defined data on each cell. This data is treated as a first-class
+  citizen to \dealii{} and, when used in conjunction with other classes,
+  can be shipped to other MPI cores during mesh refinement and
+  repartitioning.
 
@@
 
-\item The \texttt{LinearOperator} class has been extended to fully support {\trilinos} solvers.
+\item The \texttt{LinearOperator} class has been extended by a generic
+  ``payload'' mechanism that allows to attach arbitrary additional
+  information to a \texttt{LinearOperator}. The mechanism was used to
+  implement a \texttt{TrilinosPayload} class to have full support for using
+  the suite of {\trilinos} iterative solvers and preconditioners with a
+  \texttt{LinearOperator}.
 
-\item A dedicated physics module has been created to provide some standard definitions and operations used in continuum physics.
+\item A dedicated physics module has been created to provide some standard
+  definitions and operations used in continuum mechanics.
 
-\item \texttt{FE\_Enriched}
+\item TODO \texttt{FE\_Enriched}
 
-\item \texttt{FESeries} namespace, use in step-27
+\item TODO \texttt{FESeries} namespace, use in step-27
 
@@
   support points in \texttt{FE\_Q} and \texttt{FE\_DGQ} to Gauss--Lobatto
-  support points, stable evaluation of high-order Legendre polynomials, and several
-  bugfixes for high-order polynomial mappings defined through the
+  support points, stable evaluation of high-order Legendre polynomials, and
+  several bugfixes for high-order polynomial mappings defined through the
   \texttt{MappingQ} class.
 
-\item More than 230 other features and bugfixes.\\
-  We realized that static code analyzers are valuable tools to improve and maintain the quality of the code in our library 
-  on top off build and regression tests for a variety of setups using \href{https://cdash.kyomu.43-1.org/index.php?project=deal.II}{CDash}.
-  This release was tested with \href{http://cppcheck.sourceforge.net/}{Cppcheck}, 
-  \href{https://www.viva64.com/en/pvs-studio/}{PVS-Studio} and \href{https://scan.coverity.com/}{Coverity-Scan}. 
-  For the latter the current status can be observed \href{https://scan.coverity.com/projects/deal-ii}{online}.
+\item We realized that static code analyzers are valuable tools to improve
+  and maintain the quality of the code in our library on top off build and
+  regression tests for a variety of setups using
+  \href{https://cdash.kyomu.43-1.org/index.php?project=deal.II}{CDash}.
+  This release was tested with
+  \href{http://cppcheck.sourceforge.net/}{Cppcheck},
+  \href{https://www.viva64.com/en/pvs-studio/}{PVS-Studio} and
+  \href{https://scan.coverity.com/}{Coverity-Scan}. For the latter the
+  current status can be observed
+  \href{https://scan.coverity.com/projects/deal-ii}{online}.
+
+  \item More than 230 other features and bugfixes.
 \end{itemize}
@@
 
-By introduction of an arbitrary \texttt{Payload} base class, the \texttt{LinearOperator} class can be extended to seamlessly exploit the native features and operations offered by external linear algebra libraries.
-We have thus developed a \texttt{TrilinosPayload} class, through which full support for the suite of {\trilinos} parallel iterative solvers and preconditioners has been achieved.
-In the case of the construction of an \texttt{inverse\_operator}, both \dealii's built in solvers and those offered by {\trilinos} can be selected.
-This has been achieved by using the \texttt{Epetra\_Operator} as the basis for the \texttt{TrilinosPayload}, for which the result of both standard and composite operations involving forward (\texttt{Apply()}) and inverse (\texttt{ApplyInverse()}) matrix-vector multiplication are collated using lambda functions.
-It is envisaged that, in the future, similar can be done for the {\petsc} iterative solvers.
+By introduction of an arbitrary \texttt{Payload} base class, the
+\texttt{LinearOperator} class can be extended to seamlessly exploit the
+native features and operations offered by external linear algebra
+libraries. We have thus developed a \texttt{TrilinosPayload} class, through
+which full support for the suite of {\trilinos} parallel iterative solvers
+and preconditioners has been achieved. In the case of the construction of
+an \texttt{inverse\_operator}, both \dealii's built in solvers and those
+offered by {\trilinos} can be selected. This has been achieved by using the
+\texttt{Epetra\_Operator} as the basis for the \texttt{TrilinosPayload},
+for which the result of both standard and composite operations involving
+forward (\texttt{Apply()}) and inverse (\texttt{ApplyInverse()})
+matrix-vector multiplication are collated using lambda functions. It is
+envisaged that, in the future, similar can be done for the {\petsc}
+iterative solvers.
 
-An additional feature of the \texttt{LinearOperator} suite is the defintion of a \texttt{schur\_complement} operator, and its associated condensation and post-processing \texttt{PackagedOperation}s.
-An operator representing the Schur complement of a block system can be declared and, though the delayed evaluation offered by \texttt{PackagedOperation}s, reused on any number of vector systems.
+An additional feature of the \texttt{LinearOperator} suite is the defintion
+of a \texttt{schur\_complement} operator, and its associated condensation
+and post-processing \texttt{PackagedOperation}s. An operator representing
+the Schur complement of a block system can be declared and, though the
+delayed evaluation offered by \texttt{PackagedOperation}s, reused on any
+number of vector systems.
 

### 2017-03-09 8.5/paper.tex
Restructure and reword Section 2.3 LinearOperator

@@
 
-\item The \texttt{LinearOperator} class has been extended by a generic
+\item the \texttt{LinearOperator} class has been extended by a generic
   ``payload'' mechanism that allows to attach arbitrary additional
@@
 can be found
-\href{https://www.dealii.org/8.5.0/doxygen/deal.II/changes_between_8_4_and_8_5.html}{in the file that lists all changes for this release} (see \cite{changes85})
-and that is linked to from the web site of each release as well as the
-release announcement.
+\href{https://www.dealii.org/8.5.0/doxygen/deal.II/changes_between_8_4_and_8_5.html}{in
+the file that lists all changes for this release}; see \cite{changes85}.
+(The file is also linked to from the web site of each release as well as
+the release announcement.)
 
@@
 
-The \texttt{CellDataStorage} class is an integrated mechanism to safely store user-defined data, such as that required at each computation point, within each cell.
-Although the same was previously achieved through the use of a cell \texttt{user\_pointer}, it required users to manage this data themselves.
-As this data is now treated as a first-class citizen to \dealii{}, through the abstract \texttt{TransferableQuadraturePointData} we have provided a generic interface to facilitate several low-level operations that were previously tedious or technically challenging to implement.
-For instance, the \texttt{parallel::distributed::ContinuousQuadratureDataTransfer} class assists in the transfer of arbitrary data (that is continuous within a cell) stored at quadrature points when performing h-adaptive refinement of \texttt{parallel::distributed::Triangulation}.
-Not only does it perform an $\mathcal{L}^2$ projection of the specified user data between quadrature points, but it also ships the data automatically between MPI processes.
+The \texttt{CellDataStorage} class is an integrated mechanism to safely
+store user-defined data, such as that required at each computation point,
+within each cell. Although the same was previously achieved through the use
+of a cell \texttt{user\_pointer}, it required users to manage this data
+themselves. This data is now treated as a first-class citizen to \dealii{}.
+
+Through the abstract \texttt{TransferableQuadraturePointData} we have
+provided a generic interface to facilitate several low-level operations
+that were previously tedious or technically challenging to implement. For
+instance, the
+\texttt{parallel::distributed::ContinuousQuadratureDataTransfer} class
+assists in the transfer of arbitrary data (that is continuous within a
+cell) stored at quadrature points when performing h-adaptive refinement of
+\texttt{parallel::distributed::Triangulation}. Not only does it perform an
+$\mathcal{L}^2$-projection of the specified user data between quadrature
+points, but it also ships the data automatically between MPI processes.
 
@@
 
-By introduction of an arbitrary \texttt{Payload} base class, the
-\texttt{LinearOperator} class can be extended to seamlessly exploit the
-native features and operations offered by external linear algebra
-libraries. We have thus developed a \texttt{TrilinosPayload} class, through
-which full support for the suite of {\trilinos} parallel iterative solvers
-and preconditioners has been achieved. In the case of the construction of
-an \texttt{inverse\_operator}, both \dealii's built in solvers and those
-offered by {\trilinos} can be selected. This has been achieved by using the
-\texttt{Epetra\_Operator} as the basis for the \texttt{TrilinosPayload},
-for which the result of both standard and composite operations involving
-forward (\texttt{Apply()}) and inverse (\texttt{ApplyInverse()})
-matrix-vector multiplication are collated using lambda functions. It is
-envisaged that, in the future, similar can be done for the {\petsc}
+The \texttt{LinearOperator} class has been extended by a generic
+``payload'' mechanism that allows to attach arbitrary additional
+information to a \texttt{LinearOperator}. This was achieved by introducing
+a generic \texttt{Payload} base class. The main use case of the new
+mechanism is to extend the \texttt{LinearOperator} class to seamlessly
+exploit the native features and operations offered by external linear
+algebra libraries. We have thus developed a \texttt{TrilinosPayload} class,
+that provides full support for the suite of {\trilinos} parallel iterative
+solvers and preconditioners.
+
+A particular interesting case is the construction of an
+\texttt{inverse\_operator}. Now both, \dealii's built in solvers and those
+offered by {\trilinos}, can be selected. This has been achieved by using
+the \texttt{Epetra\_Operator} as the basis for the
+\texttt{TrilinosPayload}, for which the result of standard and composite
+operations involving forward (\texttt{Apply()}) and inverse
+(\texttt{ApplyInverse()}) matrix-vector multiplication are collated using
+lambda functions.
+
+It is envisaged that, in the future, similar can be done for the {\petsc}
 iterative solvers.
@@
 An additional feature of the \texttt{LinearOperator} suite is the defintion
-of a \texttt{schur\_complement} operator, and its associated condensation
+of a \texttt{schur\_complement} operator and its associated condensation
 and post-processing \texttt{PackagedOperation}s. An operator representing

### 2017-03-09 8.5/paper.tex
Reformat and slightly reword section 2.4

@@
 
-A dedicated physics module has been created to facilitate the implementation of functions and classes that relate to continuum physics, physical fields and material constitutive laws.
-To date, it includes transformations of scalar or tensorial quantities between any two configurations, and some definitions typically utilized in both linear and finite-strain nonlinear elasticity.
+A dedicated physics module has been created to facilitate the
+implementation of functions and classes that relate to continuum mechanics,
+physical fields and material constitutive laws. To date, it includes
+transformations of scalar or tensorial quantities between any two
+configurations,
+\marginpar{which configurations?}
+and some definitions typically utilized in both linear and finite-strain
+nonlinear elasticity.
 
-The \verb!Physics::Transformations!
-namespace offers push-forward and pull-back operations in the context of contravariant, covariant and Piola transformations, as well as rotation operations for Euclidean space.
-In the \verb!Physics::Elasticity::Kinematics! namespace, a selection of deformation, strain tensors and strain rate tensors are defined.
-The \verb!Physics::Elasticity::StandardTensors!
-class provides some frequently used second and fourth order metric tensors, and defines some referential and spatial projection operators and tensor derivatives as commonly required in the definition of material laws.
+The \verb!Physics::Transformations! namespace offers push-forward and
+pull-back operations in the context of contravariant, covariant and Piola
+transformations, as well as rotation operations for the Euclidean space. In
+the \verb!Physics::Elasticity::Kinematics! namespace, a selection of
+deformation, strain tensors and strain rate tensors are defined. The
+\verb!Physics::Elasticity::StandardTensors! class provides some frequently
+used second and fourth order metric tensors, and defines a number of
+referential and spatial projection operators and tensor derivatives that
+are commonly required in the definition of material laws.
 

### 2017-03-09 8.5/paper.tex
Reformat Sections 2.5 and 2.6, fix two typos

@@
 
-\verb!FE_Enriched!
-finite element implements a partition of unity finite element method (PUM) by Babuska and Melenk which enriches a standard finite element with an enrichment function multiplied with another (usually linear) finite element.
-This allows to include in the finite element space a priori knowledge about the partial differential equation being solved which in turn improves the local approximation properties of the spaces.
-The user can also use enriched and non-enriched finite elements in different parts of the domain.
-\verb|DoFTools::make_hanging_node_constraints()| function can automatically make the resulting space $C^0$ continuous.
-Existing \verb|SolutionTransfer| class can be used to transfer the solution during $h$\,-adaptive refinement from a coarse to a fine mesh under the condition that all child elements are also enriched.
+The \verb!FE_Enriched! finite element implements a partition of unity
+finite element method (PUM) by Babuska and Melenk which enriches a standard
+finite element with an enrichment function multiplied with another (usually
+linear) finite element. This allows to include in the finite element space
+a priori knowledge about the partial differential equation being solved
+which in turn improves the local approximation properties of the spaces.
+The user can also use enriched and non-enriched finite elements in
+different parts of the domain.
+\verb|DoFTools::make_hanging_node_constraints()| function can automatically
+make the resulting space $C^0$ continuous. Existing \verb|SolutionTransfer|
+class can be used to transfer the solution during $h$\,-adaptive refinement
+from a coarse to a fine mesh under the condition that all child elements
+are also enriched.
 
@@
 
-\verb|FESeries| namespace offers functions to calculate expansion series of the solution on the reference element.
-Coefficients of expansion are often used to estimate local smoothness of the underlying finite element field to decide on h- or p-adaptive refinement strategy.
-Specifically, \verb|FESeries::Legendre| calculates expansion of a scalar FE field into series of Legendre functions on a reference element, whereas \verb|FESeries::Fourier| calculates Fourier coefficients.
-The user has to specify the required number of coefficients in each direction as well as provide a collection of finite elements and quadrature rules, that will be used for calculation of the transformation matrices.
-Obviously, those matrices are calculated only once, namely the first time coefficients are requested on a cell with the given finite element.
-The step-27 has been updated to use \verb|FESeries::Fourier|.
+The \verb|FESeries| namespace offers functions to calculate expansion
+series of the solution on the reference element. Coefficients of expansion
+are often used to estimate local smoothness of the underlying finite
+element field to decide on h- or p-adaptive refinement strategy.
+Specifically, \verb|FESeries::Legendre| calculates expansion of a scalar FE
+field into series of Legendre functions on a reference element, whereas
+\verb|FESeries::Fourier| calculates Fourier coefficients. The user has to
+specify the required number of coefficients in each direction as well as
+provide a collection of finite elements and quadrature rules, that will be
+used for calculation of the transformation matrices. Obviously, those
+matrices are calculated only once, namely the first time coefficients are
+requested on a cell with the given finite element. The step-27 has been
+updated to use \verb|FESeries::Fourier|.
 

### 2017-03-09 8.5/paper.tex
Finish reformatting Section 2, rephrase section about incompatible changes

@@
 
-In order to facilitate the usage of matrix-free methods, a \verb!MatrixFreeOperator::Base! class has been introduced,
-implementing various interfaces to matrix-vector products and the necessary operations for the interface residuals according to \cite{JanssenKanschat2011} in
-the context of the geometric multigrids. Furthermore, the class is compatible with the linear operator framework and provides an interface to a Jacobi preconditioner.
-The derived classes only need to implement the \verb!apply_add()! method that is
-used in the \verb!vmult()! functions, and a method to compute the diagonal entries of the underlying matrix.
-The \verb!MatrixFreeOperator! namespace contains implementations of \verb!MatrixFreeOperators::LaplaceOperator! and
-\verb!MatrixFreeOperators::MassOperator!.
-The updated step-37 tutorial program makes use of these facilities and
-explains their usage in detail. Using the matrix-free mass operator,
-\verb!VectorTools::project! has become much faster than the previous
-matrix-based approach for elements supported by \verb!MatrixFree! and also
-works in parallel with MPI.
+In order to facilitate the usage of matrix-free methods, a
+\verb!MatrixFreeOperator::Base! class has been introduced, implementing
+various interfaces to matrix-vector products and the necessary operations
+for the interface residuals according to \cite{JanssenKanschat2011} in the
+context of the geometric multigrids. Furthermore, the class is compatible
+with the linear operator framework and provides an interface to a Jacobi
+preconditioner. The derived classes only need to implement the
+\verb!apply_add()! method that is used in the \verb!vmult()! functions, and
+a method to compute the diagonal entries of the underlying matrix. The
+\verb!MatrixFreeOperator! namespace contains implementations of
+\verb!MatrixFreeOperators::LaplaceOperator! and
+\verb!MatrixFreeOperators::MassOperator!. The updated step-37 tutorial
+program makes use of these facilities and explains their usage in detail.
+Using the matrix-free mass operator, \verb!VectorTools::project! has become
+much faster than the previous matrix-based approach for elements supported
+by \verb!MatrixFree! and also works in parallel with MPI.
 
@@
 
-\subsubsection{incompatible change 1}
-
-High-order Lagrange elements, both continuous \verb!FE_Q! and discontinuous
-\verb!FE_DGQ! types, now use the nodal points of the Gauss--Lobatto
-quadrature formula as support points, rather than the previous equidistant
-ones. For cubic polynomials and higher, the point distribution has thus changed and
-thus the entries in solution vectors will look different as compared to the previous version. Note, however,
-that using the Gauss--Lobatto points as nodal points results in a much more
-stable interpolation, including better iteration counts in most iterative
-solvers.
-
-\subsubsection{Other incompatible changes}
-
-No \texttt{long double} instantiations any more
-
-ParameterGUI moved to separate repo
-
-The
-\href{https://www.dealii.org/8.5.0/doxygen/deal.II/changes_between_8_4_and_8_5.html}{file
-  that lists all changes for this release} (see \cite{changes85})
-lists another around 20
-incompatible changes, but none of these should in fact be visible in
-typical user codes. Some remove previously deprecated classes and
-functions, and the majority change internal interfaces that are not
-typically used in user codes.
+The 8.5.0 release includes around 20
+\href{https://www.dealii.org/8.5.0/doxygen/deal.II/changes_between_8_4_and_8_5.html}{
+incompatible changes}; see \cite{changes85}. The majority of these changes
+should not be visible to typical user codes; some remove previously
+deprecated classes and functions, and the majority change internal
+interfaces that are not typically used in user codes. However, a number of
+incompatible changes a worth mentioning.
+\begin{itemize}
+  \item High-order Lagrange elements, both continuous \verb!FE_Q! and
+    discontinuous \verb!FE_DGQ! types, now use the nodal points of the
+    Gauss--Lobatto quadrature formula as support points, rather than the
+    previous equidistant ones. For cubic polynomials and higher, the point
+    distribution has thus changed and, consequently, the entries in
+    solution vectors will look different as compared to the previous
+    version. Note, however, that using the Gauss--Lobatto points as nodal
+    points results in a much more stable interpolation, including better
+    iteration counts in most iterative solvers.
+  \item
+    The library does not instantiate template versions with \texttt{long
+    double} any more. If you need template versions with \texttt{long
+    double} make sure to include the corresponding \texttt{templates.h}
+    header file.
+  \item
+    The \texttt{ParameterGUI} has been moved to a separate repository.
+\end{itemize}
 

### 2017-03-09 8.5/paper.tex
check list of contributors, remove a comma

@@
 Toby  Young,
-Zhao, Liang.
+Zhao  Liang.
 

### 2017-03-10 8.5/paper.tex
Add note; fix typos

@@
 \item New code gallery programs demonstrating
+\marginpar{Isn't the code gallery itself new for 8.5?}
   \begin{enumerate}
@@
 
-It is envisaged that, in the future, similar can be done for the {\petsc}
+We envisage that, in the future, similar can be done for the {\petsc}
 iterative solvers.
@@
 and post-processing \texttt{PackagedOperation}s. An operator representing
-the Schur complement of a block system can be declared and, though the
+the Schur complement of a block system can be declared and, through the
 delayed evaluation offered by \texttt{PackagedOperation}s, reused on any

### 2017-03-10 8.5/paper.tex
Abbreviated some points under "major changes"

@@
 \begin{itemize}
-\item The \texttt{CellDataStorage} class provides a mechanism to store
-  user-defined data on each cell. This data is treated as a first-class
-  citizen to \dealii{} and, when used in conjunction with other classes,
-  can be shipped to other MPI cores during mesh refinement and
-  repartitioning.
+\item The \texttt{CellDataStorage} class provides a safe mechanism to store
+  and communicate user-defined data on each cell.
 
@@
 
-\item the \texttt{LinearOperator} class has been extended by a generic
-  ``payload'' mechanism that allows to attach arbitrary additional
-  information to a \texttt{LinearOperator}. The mechanism was used to
-  implement a \texttt{TrilinosPayload} class to have full support for using
-  the suite of {\trilinos} iterative solvers and preconditioners with a
-  \texttt{LinearOperator}.
+\item The \texttt{LinearOperator} class has been extended by a generic
+  ``payload'' mechanism that allows the attachment of arbitrary additional
+  information to a \texttt{LinearOperator}.
 

### 2017-03-10 8.5/paper.tex
Amend discussion on linear operator extensions

@@
 The \texttt{LinearOperator} class has been extended by a generic
-``payload'' mechanism that allows to attach arbitrary additional
+``payload'' mechanism that allows for the attachment of arbitrary additional
 information to a \texttt{LinearOperator}. This was achieved by introducing
@@
 exploit the native features and operations offered by external linear
-algebra libraries. We have thus developed a \texttt{TrilinosPayload} class,
+algebra libraries. We have thus developed a \texttt{TrilinosPayload} class
 that provides full support for the suite of {\trilinos} parallel iterative
@@
 
-A particular interesting case is the construction of an
-\texttt{inverse\_operator}. Now both, \dealii's built in solvers and those
-offered by {\trilinos}, can be selected. This has been achieved by using
+A particularly interesting case is the construction of an
+\texttt{inverse\_operator}. Whereas previously only \dealii's built-in solvers
+were compatible with this operator, now both them and those
+offered by {\trilinos} can be selected. This has been achieved by using
 the \texttt{Epetra\_Operator} as the basis for the

### 2017-03-10 8.5/paper.tex
Added more details to physics module, chiefly the sec on transformations

@@
 transformations of scalar or tensorial quantities between any two
-configurations,
-\marginpar{which configurations?}
-and some definitions typically utilized in both linear and finite-strain
-nonlinear elasticity.
+configurations (by user-specification of a linear map $\mathbf{F}$), and some 
+definitions typically utilized in both linear and finite-strain nonlinear 
+elasticity.
 
@@
 pull-back operations in the context of contravariant, covariant and Piola
-transformations, as well as rotation operations for the Euclidean space. In
-the \verb!Physics::Elasticity::Kinematics! namespace, a selection of
+transformations, as well as rotation operations for the Euclidean space. 
+Although these transformations are defined in a general manner, one typical
+use of them in finite-strain elasticity would be the determination of the 
+Cauchy stress tensor $\boldsymbol{\sigma} = \boldsymbol{\sigma}\left(\mathbf{x}\right)$ 
+defined at the spatial position $\mathbf{x} \in \mathcal{B}$ from its
+fully referential counterpart, namely the Piola--Kirchhoff stress tensor 
+$\mathbf{S} = \mathbf{S}\left(\mathbf{X}\right)$ computed at the material 
+coordinate $\mathbf{X} \in \mathcal{B}_{0}$.
+By choosing
+$\mathbf{F} \left(\mathbf{X}\right) = \dfrac{\partial \mathbf{x}\left(\mathbf{X}\right)}{\partial \mathbf{X}}$,
+this is achieved through the action of the Piola push-forward
+$\boldsymbol{\sigma} = \dfrac{1}{\det \mathbf{F}} \, \mathbf{F} \cdot \mathbf{S} \cdot \mathbf{F}^{T}$.
+
+In the \verb!Physics::Elasticity::Kinematics! namespace, a selection of
 deformation, strain tensors and strain rate tensors are defined. The

### 2017-03-16 8.5/paper.tex
add a blurb about tutorials

@@
 \subsection{The physics module}
+\label{sec:physics}
 
@@
 
+\subsection{New and updated tutorial programs}
+
+\begin{itemize}
+ \item {\bf step-55} is a new tutorial that explains how to solve the Stokes 
+ equations efficiently in parallel. It is a good introduction to solving
+ systems of PDEs in parallel, discusses optimal block 
+ preconditioners, and demonstrates other aspects like error computation.
+ Inverses of individual blocks of the linear system are approximated with an algebraic 
+ multigrid preconditioner.
+
+ \item {\bf step-56} is a new tutorial that shows how to apply geometric multigrid
+ preconditioners on a subset of a system of PDEs. The problem solved here is the 
+ Stokes equations, like in step-55.
+
+ \item {\bf step-57} is a new tutorial that solves the stationary Navier-Stokes equations.
+ The nonlinear system is solved using Newton's method on a sequence of adaptively refined
+ grids. The preconditioner is again built on a block factorization of the saddle point
+ system like in step-55 and step-56, but the non-symmetric terms stemming from the
+ nonlinear convective part requires more sophisticated solvers. The benchmark problem,
+ flow in the 2d lid-driven cavity, requires a continuation method for high Reynold's 
+ numbers.
+
+ \item {\bf step-37}, discussing matrix-free computations of the 
+ Poisson equation, has been rewritten to support parallel computations.
+
+ \item {\bf step-44} now uses the physics module, discussed in Section~\ref{sec:physics},
+ to solve a solid mechanics problem.
+ \end{itemize}
+
 \subsection{Matrix-free operators}

### 2017-03-18 8.5/paper.tex
Added draft blurp on MappingManifold

@@
 
-TODO: Luca
+The \texttt{MappingManifold} class implements the same functionality
+of the \texttt{Mapping} class for \emph{manifold conforming}
+mappings. This class computes the transformation between the
+reference and real cell by exploiting the geometrical information
+coming from the underlying \texttt{Manifold} object.
+
+Quadrature points computed using this mapping lie on the exact
+geometrical objects, and tangent and normal vectors computed using
+this class are tangent and normal to the underlying geometry. This is
+in constrast with the \texttt{MappingQ} and \texttt{MappingQGeneric}
+classes, which approximate the geometry using a polynomial of some
+order, and then compute the normals and tangents using the
+approximated surface.
+
+Currently only first order transformations are implemented, and an
+assertion is thrown if the user requests any higher order
+transformations, for example, the hessian of the mapping.
 

### 2017-03-19 8.5/paper.tex
older pgfplots version
Sorry, my version is much older and I am too lazy to update.
spell checking, language fixes, etc.

@@
 
-\pgfplotsset{compat=1.12}
+\pgfplotsset{compat=1.9}
 
@@
 
-\item We realized that static code analyzers are valuable tools to improve
-  and maintain the quality of the code in our library on top off build and
+\item Static code analyzers are valuable tools to improve
+  and maintain the quality of the code in our library in addition to build and
   regression tests for a variety of setups using
@@
 The \texttt{CellDataStorage} class is an integrated mechanism to safely
-store user-defined data, such as that required at each computation point,
-within each cell. Although the same was previously achieved through the use
+store user-defined data within each cell, such as a coefficient at each quadrature point.
+Although the same was previously achieved through the use
 of a cell \texttt{user\_pointer}, it required users to manage this data
@@
 this class are tangent and normal to the underlying geometry. This is
-in constrast with the \texttt{MappingQ} and \texttt{MappingQGeneric}
+in contrast with the \texttt{MappingQ} and \texttt{MappingQGeneric}
 classes, which approximate the geometry using a polynomial of some
@@
 assertion is thrown if the user requests any higher order
-transformations, for example, the hessian of the mapping.
+transformations, for example, the Hessian of the mapping.
 
@@
 
-An additional feature of the \texttt{LinearOperator} suite is the defintion
+An additional feature of the \texttt{LinearOperator} suite is the definition
 of a \texttt{schur\_complement} operator and its associated condensation
@@
 finite element with an enrichment function multiplied with another (usually
-linear) finite element. This allows to include in the finite element space
+linear) finite element. This allows including 
 a priori knowledge about the partial differential equation being solved
+in the finite element space,
 which in turn improves the local approximation properties of the spaces.
@@
 \verb|DoFTools::make_hanging_node_constraints()| function can automatically
-make the resulting space $C^0$ continuous. Existing \verb|SolutionTransfer|
+make the resulting space $C^0$ continuous. The existing \verb|SolutionTransfer|
 class can be used to transfer the solution during $h$\,-adaptive refinement
@@
 
-The 8.5.0 release includes around 20
-\href{https://www.dealii.org/8.5.0/doxygen/deal.II/changes_between_8_4_and_8_5.html}{
+The 8.5 release includes around 20
+\href{https://www.dealii.org/developer/doxygen/deal.II/changes_between_8_4_and_8_5.html}{
 incompatible changes}; see \cite{changes85}. The majority of these changes
@@
 interfaces that are not typically used in user codes. However, a number of
-incompatible changes a worth mentioning.
+incompatible changes are worth mentioning.
 \begin{itemize}
@@
 
-There are various ways to reference \dealii{}. To acknowledge the use of a
-particular version of the library, reference the present document. For up
+There are various ways to reference \dealii{}. To acknowledge the use of the
+current version of the library, \textbf{please reference the present document}. For up
 to date information and bibtex snippets for this document see:

### 2017-03-19 8.5/paper.tex
Go through the paper one more time.

@@
 The major changes of this release are:
-\marginpar{Add one sentence to each of the following.}
 \begin{itemize}
-\item The \texttt{CellDataStorage} class provides a safe mechanism to store
+\item The \texttt{CellDataStorage} class provides a mechanism to store
   and communicate user-defined data on each cell.
@@
 \item The \texttt{MappingManifold} class provides mappings between the
-  reference cell and a mesh cell that is ``exact'' in the sense that
-  the edges, faces, and the interior of the mesh cell matches an
-  underlying manifold. This is opposed to using the usual polynomial
-  approximations of the manifold.
-  TODO: Luca
+  reference cell and a mesh cell that is ``exact'', rather than the
+  usual polynomial approximations of a manifold.
 
@@
 
-\item TODO \texttt{FE\_Enriched}
+\item The \texttt{FE\_Enriched} class implements the operation of
+  enriching the finite element space of an underlying element.
 
-\item TODO \texttt{FESeries} namespace, use in step-27
+\item The \texttt{FESeries} namespace provides expansions of a finite
+  element solution in terms of different, hierarchical bases.
 
-\item New tutorial programs step-55, step-56, and step-57,
-  demonstrating parallel solvers for the Stokes equations, a multigrid
-  method for the Stokes equation, and methods for the nonlinear
-  Navier-Stokes equations, respectively. In addition, the step-37
-  tutorial program has been rewritten to support parallel
-  computations, and the step-44 tutorial now uses the facilities of
-  the physics module mentioned above.
+\item New tutorial programs step-55, step-56, and step-57; as well as
+  updates to step-27, step-37, and step-44.
 
@@
 \item Various improvements for high-order elements, including a switch of
-  support points in \texttt{FE\_Q} and \texttt{FE\_DGQ} to Gauss--Lobatto
+  support points in \texttt{FE\_Q} and \texttt{FE\_DGQ} to Gauss-Lobatto
   support points, stable evaluation of high-order Legendre polynomials, and
@@
   \href{https://www.viva64.com/en/pvs-studio/}{PVS-Studio} and
-  \href{https://scan.coverity.com/}{Coverity-Scan}. For the latter the
-  current status can be observed
-  \href{https://scan.coverity.com/projects/deal-ii}{online}.
+  \href{https://scan.coverity.com/}{Coverity-Scan}.
 
@@
 \end{itemize}
-The more important ones of these will be detailed in the following section.
-Information on how to cite \dealii{} is provided in Section \ref{sec:cite}.
+The more important ones of these changes will be detailed in the
+following section.  Information on how to cite \dealii{} is provided
+in Section \ref{sec:cite}.
 
@@
 \href{https://www.dealii.org/8.5.0/doxygen/deal.II/changes_between_8_4_and_8_5.html}{in
-the file that lists all changes for this release}; see \cite{changes85}.
+the file that lists all changes for this release}, see \cite{changes85}.
 (The file is also linked to from the web site of each release as well as
@@
 
-The \texttt{CellDataStorage} class is an integrated mechanism to safely
+The \texttt{CellDataStorage} class is an integrated mechanism to
 store user-defined data within each cell, such as a coefficient at each quadrature point.
@@
 of a cell \texttt{user\_pointer}, it required users to manage this data
-themselves. This data is now treated as a first-class citizen to \dealii{}.
+themselves, including transfer of this data if ownership of a cell is
+transferred from one processor to another in parallel
+computations. Instead, this data is now treated as a first-class
+citizen in \dealii{}.
 
-Through the abstract \texttt{TransferableQuadraturePointData} we have
-provided a generic interface to facilitate several low-level operations
-that were previously tedious or technically challenging to implement. For
-instance, the
+\texttt{CellDataStorage} can work with arbitrary kinds of data defined
+to live on a cell as long as the corresponding data type implements a
+certain interface. As an example,
+the abstract \texttt{TransferableQuadraturePointData} class provides
+a generic interface to facilitate several low-level operations
+that were previously tedious or technically challenging to implement.
+\marginpar{Denis: It is not clear to me how \texttt{CellDataStorage} and
+  \texttt{p::d::ContinuousQuadratureDataTransfer} relate to each
+  other. Can you clarify the paragraph?}
+For instance, the
 \texttt{parallel::distributed::ContinuousQuadratureDataTransfer} class
 assists in the transfer of arbitrary data (that is continuous within a
-cell) stored at quadrature points when performing h-adaptive refinement of
+cell) stored at quadrature points when performing $h$-adaptive refinement of
 \texttt{parallel::distributed::Triangulation}. Not only does it perform an
-$\mathcal{L}^2$-projection of the specified user data between quadrature
+$L^2$-projection of the specified user data between quadrature
 points, but it also ships the data automatically between MPI processes.
@@
 
-The \texttt{MappingManifold} class implements the same functionality
-of the \texttt{Mapping} class for \emph{manifold conforming}
-mappings. This class computes the transformation between the
+The \texttt{MappingManifold} class implements the functionality
+of the \texttt{Mapping} interface for \emph{manifold conforming}
+mappings. In other words, instead of using polynomial approximations
+of the objects that describe the boundaries (or interiors) of cells,
+this class computes the transformation between the
 reference and real cell by exploiting the geometrical information
@@
 
-Quadrature points computed using this mapping lie on the exact
-geometrical objects, and tangent and normal vectors computed using
-this class are tangent and normal to the underlying geometry. This is
+When using this class,
+quadrature points lie on the \textit{exact}
+geometrical objects, and tangent and normal vectors computed
+are tangent and normal to the underlying geometry. This is
 in contrast with the \texttt{MappingQ} and \texttt{MappingQGeneric}
@@
 
-Currently only first order transformations are implemented, and an
-assertion is thrown if the user requests any higher order
-transformations, for example, the Hessian of the mapping.
+The class currently only implements the information that relates to
+the mapping itself, as well to its derivatives (such as the Jacobian
+of the mapping, or the determinant thereof). Information related to
+higher order derivatives -- such as the Hessian of the mapping -- will
+result in an exception.
 
@@
 
-The \texttt{LinearOperator} class has been extended by a generic
+We have extended the \texttt{LinearOperator} class by a generic
 ``payload'' mechanism that allows for the attachment of arbitrary additional
@@
 \texttt{inverse\_operator}. Whereas previously only \dealii's built-in solvers
-were compatible with this operator, now both them and those
-offered by {\trilinos} can be selected. This has been achieved by using
+were compatible with this operator, one can now also select those
+offered by {\trilinos}. This has been achieved by using
 the \texttt{Epetra\_Operator} as the basis for the
@@
 
-We envisage that, in the future, similar can be done for the {\petsc}
+We envisage that, in the future, similar extensions can be implemented for the {\petsc}
 iterative solvers.
 
-An additional feature of the \texttt{LinearOperator} suite is the definition
+Another additional in this release related to the \texttt{LinearOperator} suite is the definition
 of a \texttt{schur\_complement} operator and its associated condensation
@@
 
-A dedicated physics module has been created to facilitate the
+We have created a dedicated physics module to facilitate the
 implementation of functions and classes that relate to continuum mechanics,
@@
 Cauchy stress tensor $\boldsymbol{\sigma} = \boldsymbol{\sigma}\left(\mathbf{x}\right)$ 
-defined at the spatial position $\mathbf{x} \in \mathcal{B}$ from its
-fully referential counterpart, namely the Piola--Kirchhoff stress tensor 
+defined at a spatial position $\mathbf{x} \in \mathcal{B}$ from its
+fully referential counterpart, namely the Piola-Kirchhoff stress tensor 
 $\mathbf{S} = \mathbf{S}\left(\mathbf{X}\right)$ computed at the material 
@@
 In the \verb!Physics::Elasticity::Kinematics! namespace, a selection of
-deformation, strain tensors and strain rate tensors are defined. The
+deformation, strain and strain rate tensors are defined. The
 \verb!Physics::Elasticity::StandardTensors! class provides some frequently
@@
 are commonly required in the definition of material laws.
+
+The updated step-44 tutorial program demonstrates the
+use of the physical module in the context of a solid mechanics problem.
 
@@
 unnecessary inner products inside the Chebyshev smoother and
-$\mathcal O(n_\text{levels})$ global communication steps during the
+${\mathcal O}(n_\text{levels})$ global communication steps during the
 restriction process rather than the single unavoidable global communication
-step inherent to going to the coarsest grid and the coarse solve. New
-matrix-free transfer implementations called \texttt{MGTransferMatrixFree} were
-devised that can replace the matrix-based \texttt{MGTransferPrebuilt} class
+step inherent to going to the coarsest grid and the coarse solve. We
+implemented new
+matrix-free transfer implementations called \texttt{MGTransferMatrixFree}
+that can replace the matrix-based \texttt{MGTransferPrebuilt} class
 for tensor product elements. Besides better scalability than the Trilinos
 Epetra matrices underlying the latter, the matrix-free transfer is also a much
-better for high-order elements with a complexity per degree of freedom of
+faster for high-order elements with a complexity per degree of freedom of
 $\mathcal O(d p)$ in the polynomial degree $p$ in $d$ dimensions rather than
@@
   \end{tikzpicture}
-  \caption{Scaling of \dealii{}'s geometric multigrid algorithms on SuperMUC.}
+  \caption{Scaling of \dealii{}'s geometric multigrid algorithms on
+    SuperMUC. Each line corresponds to a strong scaling experiment,
+    increasing the number of processor cores for a fixed-size
+    problem. Comparing corresponding data points on different lines
+    yields weak scaling information.}
 \label{fig:scaling_mg}
@@
 algorithmic improvements on a setup with discontinuous DG elements, clearly
-improving the latency of the multigrid V-cycle. The updated step-37 tutorial
+improving the latency of the multigrid V-cycle.
+
+The updated step-37 tutorial
 programs presents the updated algorithms and the MPI-parallel multigrid
 setting with matrix-free operator evaluation.
+
+\subsection{Matrix-free operators}
+
+In order to facilitate the usage of matrix-free methods, a
+\verb!MatrixFreeOperator::Base! class has been introduced, implementing
+functionality for matrix-vector products and the necessary operations
+for the interface residuals for multigrid on adaptively refined meshes
+(see \cite{JanssenKanschat2011}). Furthermore, the class is compatible
+with the linear operator framework and provides an interface to a Jacobi
+preconditioner. Derived classes only need to implement the
+\verb!apply_add()! method that is used in the \verb!vmult()! functions, and
+a method to compute the diagonal entries of the underlying matrix. The
+\verb!MatrixFreeOperator! namespace contains implementations of
+\verb!MatrixFreeOperators::LaplaceOperator! and
+\verb!MatrixFreeOperators::MassOperator!.
+
+The updated step-37 tutorial
+program makes use of these facilities and explains their usage in detail.
+Using the matrix-free mass operator, \verb!VectorTools::project! has become
+much faster than the previous matrix-based approach for elements supported
+by \verb!MatrixFree! and also works for parallel computations based on MPI.
 
@@
 which in turn improves the local approximation properties of the spaces.
-The user can also use enriched and non-enriched finite elements in
+Programs can also use enriched and non-enriched finite elements in
 different parts of the domain.
+
+The
 \verb|DoFTools::make_hanging_node_constraints()| function can automatically
@@
 are often used to estimate local smoothness of the underlying finite
-element field to decide on h- or p-adaptive refinement strategy.
-Specifically, \verb|FESeries::Legendre| calculates expansion of a scalar FE
-field into series of Legendre functions on a reference element, whereas
-\verb|FESeries::Fourier| calculates Fourier coefficients. The user has to
+element field to decide on a $h$- or $p$-adaptive refinement strategy.
+Specifically, \verb|FESeries::Legendre| calculates expansion of a
+scalar finite element
+field into series of Legendre functions on the reference element, whereas
+\verb|FESeries::Fourier| calculates Fourier coefficients. Programs
+using this functionality have to
 specify the required number of coefficients in each direction as well as
-provide a collection of finite elements and quadrature rules, that will be
-used for calculation of the transformation matrices. Obviously, those
-matrices are calculated only once, namely the first time coefficients are
-requested on a cell with the given finite element. The step-27 has been
-updated to use \verb|FESeries::Fourier|.
+provide a collection of finite elements and quadrature rules; these will be
+used for the calculation of the transformation matrices.
+
+The updated step-27 tutorial program demonstrates the
+use of \verb|FESeries::Fourier|.
 
@@
 
+In addition to the update tutorial programs mentioned in the previous
+section, this release of \dealii{} includes three new tutorials:
 \begin{itemize}
- \item {\bf step-55} is a new tutorial that explains how to solve the Stokes 
+ \item {\bf step-55} explains how to solve the Stokes 
  equations efficiently in parallel. It is a good introduction to solving
@@
 
- \item {\bf step-56} is a new tutorial that shows how to apply geometric multigrid
+ \item {\bf step-56} shows how to apply geometric multigrid
  preconditioners on a subset of a system of PDEs. The problem solved here is the 
@@
 
- \item {\bf step-57} is a new tutorial that solves the stationary Navier-Stokes equations.
+ \item {\bf step-57} solves the stationary Navier-Stokes equations.
  The nonlinear system is solved using Newton's method on a sequence of adaptively refined
@@
  numbers.
-
- \item {\bf step-37}, discussing matrix-free computations of the 
- Poisson equation, has been rewritten to support parallel computations.
-
- \item {\bf step-44} now uses the physics module, discussed in Section~\ref{sec:physics},
- to solve a solid mechanics problem.
  \end{itemize}
-
-\subsection{Matrix-free operators}
-
-In order to facilitate the usage of matrix-free methods, a
-\verb!MatrixFreeOperator::Base! class has been introduced, implementing
-various interfaces to matrix-vector products and the necessary operations
-for the interface residuals according to \cite{JanssenKanschat2011} in the
-context of the geometric multigrids. Furthermore, the class is compatible
-with the linear operator framework and provides an interface to a Jacobi
-preconditioner. The derived classes only need to implement the
-\verb!apply_add()! method that is used in the \verb!vmult()! functions, and
-a method to compute the diagonal entries of the underlying matrix. The
-\verb!MatrixFreeOperator! namespace contains implementations of
-\verb!MatrixFreeOperators::LaplaceOperator! and
-\verb!MatrixFreeOperators::MassOperator!. The updated step-37 tutorial
-program makes use of these facilities and explains their usage in detail.
-Using the matrix-free mass operator, \verb!VectorTools::project! has become
-much faster than the previous matrix-based approach for elements supported
-by \verb!MatrixFree! and also works in parallel with MPI.
 
@@
 deprecated classes and functions, and the majority change internal
-interfaces that are not typically used in user codes. However, a number of
-incompatible changes are worth mentioning.
+interfaces that are not usually used in external applications. However, three
+incompatible changes are worth mentioning:
 \begin{itemize}
@@
     discontinuous \verb!FE_DGQ! types, now use the nodal points of the
-    Gauss--Lobatto quadrature formula as support points, rather than the
+    Gauss-Lobatto quadrature formula as support points, rather than the
     previous equidistant ones. For cubic polynomials and higher, the point
     distribution has thus changed and, consequently, the entries in
-    solution vectors will look different as compared to the previous
-    version. Note, however, that using the Gauss--Lobatto points as nodal
+    solution vectors will be different compared to previous
+    versions of \dealii{}. Note, however, that using the Gauss-Lobatto points as nodal
     points results in a much more stable interpolation, including better
@@
   \item
-    The library does not instantiate template versions with \texttt{long
-    double} any more. If you need template versions with \texttt{long
-    double} make sure to include the corresponding \texttt{templates.h}
-    header file.
+    The library no longer instantiates template classes with \texttt{long
+    double}. These were rarely used, but took up a significant
+    fraction of compile and link time, as well as library
+    size. Application programs can, however, still instantiate all
+    template classes with \texttt{long
+    double} as long as they include the corresponding \texttt{.templates.h}
+    header files.
   \item

### 2017-03-19 8.5/paper.tex
Add one name.

@@
 Juliane  Dannberg,
+Niklas Fehn,
 Ren{\'e}  Gassm{\"o}ller,

### 2017-03-20 8.5/paper.tex
clarify CellDataStorage

@@
 
-The \texttt{CellDataStorage} class is an integrated mechanism to
-store user-defined data within each cell, such as a coefficient at each quadrature point.
+The \texttt{CellDataStorage} templated class is an integrated mechanism to
+store a vector of user-defined data within each cell,
+such as a coefficient at each quadrature point.
 Although the same was previously achieved through the use
-of a cell \texttt{user\_pointer}, it required users to manage this data
+of a cell's \texttt{user\_pointer}, it required users to manage this data
 themselves, including transfer of this data if ownership of a cell is
@@
 that were previously tedious or technically challenging to implement.
-\marginpar{Denis: It is not clear to me how \texttt{CellDataStorage} and
-  \texttt{p::d::ContinuousQuadratureDataTransfer} relate to each
-  other. Can you clarify the paragraph?}
-For instance, the
-\texttt{parallel::distributed::ContinuousQuadratureDataTransfer} class
-assists in the transfer of arbitrary data (that is continuous within a
-cell) stored at quadrature points when performing $h$-adaptive refinement of
+For instance, when user's custom data type is derived from \texttt{TransferableQuadraturePointData}
+and stored at quadrature points using \texttt{CellDataStorage},
+the \texttt{parallel::distributed::ContinuousQuadratureDataTransfer} class
+can transfer this data (assuming that it is continuous within a cell)
+when performing $h$-adaptive refinement of
 \texttt{parallel::distributed::Triangulation}. Not only does it perform an

### 2017-03-20 8.5/paper.tex
Minor edits.

@@
 certain interface. As an example,
-the abstract \texttt{TransferableQuadraturePointData} class provides
+the abstract class \texttt{TransferableQuadraturePointData} provides
 a generic interface to facilitate several low-level operations
 that were previously tedious or technically challenging to implement.
-For instance, when user's custom data type is derived from \texttt{TransferableQuadraturePointData}
-and stored at quadrature points using \texttt{CellDataStorage},
-the \texttt{parallel::distributed::ContinuousQuadratureDataTransfer} class
-can transfer this data (assuming that it is continuous within a cell)
-when performing $h$-adaptive refinement of
-\texttt{parallel::distributed::Triangulation}. Not only does it perform an
-$L^2$-projection of the specified user data between quadrature
-points, but it also ships the data automatically between MPI processes.
+For instance, when a user's custom data type is derived from \texttt{TransferableQuadraturePointData}
+-- a data type that stores information at the quadrature points of cells using
+\texttt{CellDataStorage} --,
+then the \texttt{parallel::distributed::ContinuousQuadratureDataTransfer} class
+can transfer this data during
+$h$-adaptive refinement from parent to children and the other way around
+(interpreting the data on quadrature points as discrete representations of an
+underlying continuous field),
+as well as from one processor to another during repartitioning.
 

### 2017-03-20 8.5/paper.tex
Add myself to the release author list.

@@
     Matthias Maier,
-    Jean-Paul Pelteret
+    Jean-Paul Pelteret,
+    David Wells
   },
@@
   {\texttt{jean-paul.pelteret@fau.de}}}
+
+\author[9]{David Wells}
+\affil[9]{Department of Mathematical Sciences, Rensselaer Polytechnic Institute,
+Troy, NY 12180, USA
+  {\texttt{wellsd2@rpi.edu}}}
 
@@
 
+D.~Wells was supported by the National Science Foundation (NSF) through Grant
+DMS-1344962.
+
 The Interdisciplinary Center for Scientific Computing (IWR) at Heidelberg University has provided

### 2017-03-20 8.5/paper.tex
Make Reynolds not possessive.
edit the edits

@@
 
-\texttt{CellDataStorage} can work with arbitrary kinds of data defined
-to live on a cell as long as the corresponding data type implements a
-certain interface. As an example,
-the abstract class \texttt{TransferableQuadraturePointData} provides
-a generic interface to facilitate several low-level operations
-that were previously tedious or technically challenging to implement.
-For instance, when a user's custom data type is derived from \texttt{TransferableQuadraturePointData}
--- a data type that stores information at the quadrature points of cells using
-\texttt{CellDataStorage} --,
+\texttt{CellDataStorage} can work with arbitrary kinds of data defined to live on a cell
+as long as the corresponding data type implements a default constructor.
+Additionally, when a user's custom data type is derived from \texttt{TransferableQuadraturePointData}
+and thereby implements certain interface,
 then the \texttt{parallel::distributed::ContinuousQuadratureDataTransfer} class
-can transfer this data during
-$h$-adaptive refinement from parent to children and the other way around
+can transfer this data during $h$-adaptive refinement from parent cell to children
 (interpreting the data on quadrature points as discrete representations of an
@@
  nonlinear convective part requires more sophisticated solvers. The benchmark problem,
- flow in the 2d lid-driven cavity, requires a continuation method for high Reynold's 
+ flow in the 2d lid-driven cavity, requires a continuation method for high Reynolds
  numbers.

### 2017-03-20 8.5/paper.tex
Add myself.

@@
     Jean-Paul Pelteret,
+    Bruno Turcksin,
     David Wells
@@
 
-\author[9]{David Wells}
-\affil[9]{Department of Mathematical Sciences, Rensselaer Polytechnic Institute,
+\author[9]{Bruno Turcksin\footnote{
+  This manuscript has been authored by UT-Battelle, LLC under Contract No.
+  DE-AC05-00OR22725 with the U.S. Department of Energy. The United States
+  Government retains and the publisher, by accepting the article for
+  publication, acknowledges that the United States Government retains a
+  non-exclusive, paid-up, irrevocable, worldwide license to publish or reproduce
+  the published form of this manuscript, or allow others to do so, for United
+  States Government purposes. The Department of Energy will provide public
+  access to these results of federally sponsored research in accordance with the
+  DOE Public Access Plan (http://energy.gov/downloads/doe-public-access-plan).}}
+\affil[9]{Computational Engineering and Energy Sciences Group,
+  Computional Sciences and Engineering Division,
+  Oak Ridge National Laboratory, 1 Bethel Valley Rd., 
+  TN 37831, USA
+  {\texttt{turcksinbr@ornl.gov}}}
+
+\author[10]{David Wells}
+\affil[10]{Department of Mathematical Sciences, Rensselaer Polytechnic Institute,
 Troy, NY 12180, USA
@@
 
+B.~Turcksin: This material is based upon work supported by the U.S. Department of
+Energy, Office of Science, under contract number DE-AC05-00OR22725.
+
 D.~Wells was supported by the National Science Foundation (NSF) through Grant

### 2017-03-21 8.5/paper.tex
Minor edits.

@@
 Additionally, when a user's custom data type is derived from \texttt{TransferableQuadraturePointData}
-and thereby implements certain interface,
+and thereby implements certain interfaces,
 then the \texttt{parallel::distributed::ContinuousQuadratureDataTransfer} class
@@
 
-\marginpar{Add your funding here as appropriate}
 D.~Arndt was supported by the German Research Foundation (DFG) under the

### 2017-03-21 8.5/paper.tex
Move the discussion of the code gallery.

@@
 \item New tutorial programs step-55, step-56, and step-57; as well as
-  updates to step-27, step-37, and step-44.
-
-\item New code gallery programs demonstrating
-\marginpar{Isn't the code gallery itself new for 8.5?}
-  \begin{enumerate}
-    \item[(i)] quasi-static visco-elastic material behavior,
-    \item[(ii)] multiphase Navier-Stokes flow,
-    \item[(iii)] the evolution of global-scale topography on planetary bodies,
-    \item[(iv)] goal-oriented elastoplasticity.
-  \end{enumerate}
+  updates to step-27, step-37, and step-44. In addition, the separate code
+  gallery of \dealii{} has gained a number of new entries.
 
@@
 
+In addition to tutorials, \dealii{} has a separate ``code gallery'' that
+consists of programs shared by users as examples of what can be
+done with \dealii{}. While not part of the release process, it is nonetheless
+worth mentioning that the set of new programs since the last release covers
+the following topics:
+\marginpar{Add the muscle elasticity example if it is merged in time.}
+  \begin{itemize}
+    \item Auasi-static visco-elastic material behavior;
+    \item Multiphase Navier-Stokes flow;
+    \item The evolution of global-scale topography on planetary bodies;
+    \item Goal-oriented elastoplasticity.
+  \end{itemize}
+
 \subsection{Incompatible changes}

### 2017-03-23 8.5/paper.tex
Update number of changes.
Add further points on changes to step-44 tutorial.
Also, fix and add to code-gallery example description

@@
 
-  \item More than 230 other features and bugfixes.
+  \item More than 240 other features and bugfixes.
 \end{itemize}
@@
 The updated step-44 tutorial program demonstrates the
-use of the physical module in the context of a solid mechanics problem.
+use of the physics module in the context of a solid mechanics problem.
+Additional updates to this tutorial include application of the new 
+\texttt{CellDataStorage} class to store and retrieve local quadrature point 
+data. An alternative approach to solving the linear system, previously 
+performed by global static condensation of two of the three field variables, 
+has been implemented using the \texttt{LinearOperator} class.
 
@@
   \begin{itemize}
-    \item Auasi-static visco-elastic material behavior;
+    \item Quasi-static quasi-incompressible visco-elastic material behavior;
     \item Multiphase Navier-Stokes flow;

### 2017-03-23 8.5/paper.tex
Minor cleanup

@@
 
+\item Various improvements for high-order elements, including a switch of
+  support points in \texttt{FE\_Q} and \texttt{FE\_DGQ} to Gauss-Lobatto
+  support points, stable evaluation of high-order Legendre polynomials, and
+  several bugfixes for high-order polynomial mappings defined through the
+  \texttt{MappingQ} class.
+
 \item The \texttt{LinearOperator} class has been extended by a generic
@@
   gallery of \dealii{} has gained a number of new entries.
-
-\item Various improvements for high-order elements, including a switch of
-  support points in \texttt{FE\_Q} and \texttt{FE\_DGQ} to Gauss-Lobatto
-  support points, stable evaluation of high-order Legendre polynomials, and
-  several bugfixes for high-order polynomial mappings defined through the
-  \texttt{MappingQ} class.
 
@@
 
-In addition to the update tutorial programs mentioned in the previous
+In addition to the updated tutorial programs mentioned in the previous
 section, this release of \dealii{} includes three new tutorials:
@@
     discontinuous \verb!FE_DGQ! types, now use the nodal points of the
-    Gauss-Lobatto quadrature formula as support points, rather than the
+    Gauss-Lobatto quadrature formula as support points by default, rather than the
     previous equidistant ones. For cubic polynomials and higher, the point

### 2017-03-24 8.5/paper.tex
update authors

@@
 %
+
   Rajat Arora,
@@
   Sam Cox,
+  Toby D. Young,
   Juliane Dannberg,
+  Nicola Demo,
+  Patrick Esser,
   Niklas Fehn,
-Ren{\'e}  Gassm{\"o}ller,
+  Rene Gassmoeller,
   Joscha Gedicke,
+  Nicola Giuliani,
   Sebastian Gonzalez-Pintor,
@@
   Justin Kauffman,
+  Eldar Khattatov ,
+  Uwe Koecher,
+  Alex Kokomov,
   Paul Kuberry,
@@
   Konstantin Ladutenko,
+  Karl Ljungkvist,
   Andrew McBride,
@@
   Jaeryun Yim,
-Toby  Young,
-Zhao  Liang.
+  Liang Zhao
 

### 2017-04-07 8.5/paper.tex
update date
remove marginpar
Add missing punctuation in affiliation addresses.

@@
   Oak Ridge National Laboratory, 1 Bethel Valley Rd.,
-  TN 37831, USA
+  TN 37831, USA.
   {\texttt{turcksinbr@ornl.gov}}}
@@
 \affil[10]{Department of Mathematical Sciences, Rensselaer Polytechnic Institute,
-Troy, NY 12180, USA
+Troy, NY 12180, USA.
   {\texttt{wellsd2@rpi.edu}}}
@@
 
-\marginpar{Update date}
-\dealii{} version 8.5.0 was released March 11, 2017. This paper provides an
+\dealii{} version 8.5.0 was released April 6, 2017. This paper provides an
 overview of the new features of this release and serves as a citable
@@
 the following topics:
-\marginpar{Add the muscle elasticity example if it is merged in time.}
+%\marginpar{Add the muscle elasticity example if it is merged in time.}
   \begin{itemize}

### 2017-04-07 8.5/paper.tex
Fix links.

@@
 can be found
-\href{https://www.dealii.org/8.5.0/doxygen/deal.II/changes_between_8_4_and_8_5.html}{in
+\href{https://www.dealii.org/8.5.0/doxygen/deal.II/changes_between_8_4_2_and_8_5_0.html}{in
 the file that lists all changes for this release}, see \cite{changes85}.
@@
 The 8.5 release includes around 20
-\href{https://www.dealii.org/developer/doxygen/deal.II/changes_between_8_4_and_8_5.html}{
+\href{https://www.dealii.org/8.5.0/doxygen/deal.II/changes_between_8_4_2_and_8_5_0.html}{
 incompatible changes}; see \cite{changes85}. The majority of these changes

### 2018-04-04 9.0/paper.tex
9.0 template

--- 2017-04-07 8.5/paper.tex
@@
   pdfauthor={
-    Daniel Arndt,
-    Wolfgang Bangerth,
-    Denis Davydov,
     Timo Heister,
-    Luca Heltai,
-    Martin Kronbichler,
-    Matthias Maier,
-    Jean-Paul Pelteret,
-    Bruno Turcksin,
-    David Wells
   },
-  pdftitle={The deal.II Library, Version 8.5, 2017},
+  pdftitle={The deal.II Library, Version 9.0, 2018},
 }
 
-\title{The \dealii{} Library, Version 8.5}
+\title{The \dealii{} Library, Version 9.0}
 
-\author[1]{Daniel Arndt}
-\affil[1]{Interdisciplinary Center for Scientific Computing,
-    Heidelberg University, Im Neuenheimer Feld 205, 69120 Heidelberg, Germany.
-    {\texttt{daniel.arndt@iwr.uni-heidelberg.de}}}
-
-\author[2]{Wolfgang Bangerth}
-\affil[2]{Department of Mathematics, Colorado State University, Fort
-  Collins, CO 80523-1874, USA.
-    {\texttt{bangerth@colostate.edu}}}
-
-\author[3]{Denis Davydov}
-\affil[3]{Chair of Applied Mechanics, University of
-     Erlangen-Nuremberg, Egerlandstr.\ 5, 91058 Erlangen, Germany.
-    {\texttt{denis.davydov@fau.de}}}
+% \author[1]{Daniel Arndt}
+% \affil[1]{Interdisciplinary Center for Scientific Computing,
+%     Heidelberg University, Im Neuenheimer Feld 205, 69120 Heidelberg, Germany.
+%     {\texttt{daniel.arndt@iwr.uni-heidelberg.de}}}
+% 
+% \author[2]{Wolfgang Bangerth}
+% \affil[2]{Department of Mathematics, Colorado State University, Fort
+%   Collins, CO 80523-1874, USA.
+%     {\texttt{bangerth@colostate.edu}}}
+% 
+% \author[3]{Denis Davydov}
+% \affil[3]{Chair of Applied Mechanics, University of
+%      Erlangen-Nuremberg, Egerlandstr.\ 5, 91058 Erlangen, Germany.
+%     {\texttt{denis.davydov@fau.de}}}
 
@@
   {\texttt{heister@clemson.edu}}}
-
-\author[5]{Luca Heltai}
-\affil[5]{SISSA,
-  International School for Advanced Studies,
-  Via Bonomea 265,
-  34136, Trieste, Italy.
-{\texttt{luca.heltai@sissa.it}}}
-
-\author[6]{Martin Kronbichler}
-\affil[6]{Institute for Computational Mechanics,
-  Technical University of Munich,
-  Boltzmannstr.~15, 85748 Garching, Germany.
-  {\texttt{kronbichler@lnm.mw.tum.de}}}
-
-\author[7]{Matthias Maier}
-\affil[7]{School of Mathematics,
-  University of Minnesota,
-  127 Vincent Hall, 206 Church Street SE,
-  Minneapolis, MN 55455, USA.
-  {\texttt{msmaier@umn.edu}}}
-
-\author[8]{Jean-Paul Pelteret}
-\affil[8]{Chair of Applied Mechanics,
-  University of Erlangen-Nuremberg,
-  Egerlandstr.\ 5,
-  91058 Erlangen,
-  Germany.
-  {\texttt{jean-paul.pelteret@fau.de}}}
-
-\author[9]{Bruno Turcksin\footnote{
-  This manuscript has been authored by UT-Battelle, LLC under Contract No.
-  DE-AC05-00OR22725 with the U.S. Department of Energy. The United States
-  Government retains and the publisher, by accepting the article for
-  publication, acknowledges that the United States Government retains a
-  non-exclusive, paid-up, irrevocable, worldwide license to publish or reproduce
-  the published form of this manuscript, or allow others to do so, for United
-  States Government purposes. The Department of Energy will provide public
-  access to these results of federally sponsored research in accordance with the
-  DOE Public Access Plan (http://energy.gov/downloads/doe-public-access-plan).}}
-\affil[9]{Computational Engineering and Energy Sciences Group,
-  Computional Sciences and Engineering Division,
-  Oak Ridge National Laboratory, 1 Bethel Valley Rd.,
-  TN 37831, USA.
-  {\texttt{turcksinbr@ornl.gov}}}
-
-\author[10]{David Wells}
-\affil[10]{Department of Mathematical Sciences, Rensselaer Polytechnic Institute,
-Troy, NY 12180, USA.
-  {\texttt{wellsd2@rpi.edu}}}
+% 
+% \author[5]{Luca Heltai}
+% \affil[5]{SISSA,
+%   International School for Advanced Studies,
+%   Via Bonomea 265,
+%   34136, Trieste, Italy.
+% {\texttt{luca.heltai@sissa.it}}}
+% 
+% \author[6]{Martin Kronbichler}
+% \affil[6]{Institute for Computational Mechanics,
+%   Technical University of Munich,
+%   Boltzmannstr.~15, 85748 Garching, Germany.
+%   {\texttt{kronbichler@lnm.mw.tum.de}}}
+% 
+% \author[7]{Matthias Maier}
+% \affil[7]{School of Mathematics,
+%   University of Minnesota,
+%   127 Vincent Hall, 206 Church Street SE,
+%   Minneapolis, MN 55455, USA.
+%   {\texttt{msmaier@umn.edu}}}
+% 
+% \author[8]{Jean-Paul Pelteret}
+% \affil[8]{Chair of Applied Mechanics,
+%   University of Erlangen-Nuremberg,
+%   Egerlandstr.\ 5,
+%   91058 Erlangen,
+%   Germany.
+%   {\texttt{jean-paul.pelteret@fau.de}}}
+% 
+% \author[9]{Bruno Turcksin\footnote{
+%   This manuscript has been authored by UT-Battelle, LLC under Contract No.
+%   DE-AC05-00OR22725 with the U.S. Department of Energy. The United States
+%   Government retains and the publisher, by accepting the article for
+%   publication, acknowledges that the United States Government retains a
+%   non-exclusive, paid-up, irrevocable, worldwide license to publish or reproduce
+%   the published form of this manuscript, or allow others to do so, for United
+%   States Government purposes. The Department of Energy will provide public
+%   access to these results of federally sponsored research in accordance with the
+%   DOE Public Access Plan (http://energy.gov/downloads/doe-public-access-plan).}}
+% \affil[9]{Computational Engineering and Energy Sciences Group,
+%   Computional Sciences and Engineering Division,
+%   Oak Ridge National Laboratory, 1 Bethel Valley Rd.,
+%   TN 37831, USA.
+%   {\texttt{turcksinbr@ornl.gov}}}
+% 
+% \author[10]{David Wells}
+% \affil[10]{Department of Mathematical Sciences, Rensselaer Polytechnic Institute,
+% Troy, NY 12180, USA.
+%   {\texttt{wellsd2@rpi.edu}}}
 
@@
   This paper provides an overview of the new features of the finite element
-  library \dealii{} version 8.5.
+  library \dealii{} version 9.0.
 \end{abstract}
@@
 
-\dealii{} version 8.5.0 was released April 6, 2017. This paper provides an
+\dealii{} version 9.0.0 was released April X, 2018. This paper provides an
 overview of the new features of this release and serves as a citable
-reference for the \dealii{} software library version 8.5. \dealii{} is an
+reference for the \dealii{} software library version 9.0. \dealii{} is an
 object-oriented finite element library used around the world in the
@@
 \begin{itemize}
-\item The \texttt{CellDataStorage} class provides a mechanism to store
-  and communicate user-defined data on each cell.
+\item The \texttt{CellDataStorage} ...
 
-\item The \texttt{MappingManifold} class provides mappings between the
-  reference cell and a mesh cell that is ``exact'', rather than the
-  usual polynomial approximations of a manifold.
-
-\item Various improvements for high-order elements, including a switch of
-  support points in \texttt{FE\_Q} and \texttt{FE\_DGQ} to Gauss-Lobatto
-  support points, stable evaluation of high-order Legendre polynomials, and
-  several bugfixes for high-order polynomial mappings defined through the
-  \texttt{MappingQ} class.
-
-\item The \texttt{LinearOperator} class has been extended by a generic
-  ``payload'' mechanism that allows the attachment of arbitrary additional
-  information to a \texttt{LinearOperator}.
-
-\item A dedicated physics module has been created to provide some standard
-  definitions and operations used in continuum mechanics.
-
-\item The \texttt{FE\_Enriched} class implements the operation of
-  enriching the finite element space of an underlying element.
-
-\item The \texttt{FESeries} namespace provides expansions of a finite
-  element solution in terms of different, hierarchical bases.
-
-\item New tutorial programs step-55, step-56, and step-57; as well as
-  updates to step-27, step-37, and step-44. In addition, the separate code
+\item New tutorial programs XYZ; as well as
+  updates to XYZ. In addition, the separate code
   gallery of \dealii{} has gained a number of new entries.
 
-\item Static code analyzers are valuable tools to improve
-  and maintain the quality of the code in our library in addition to build and
-  regression tests for a variety of setups using
-  \href{https://cdash.kyomu.43-1.org/index.php?project=deal.II}{CDash}.
-  This release was tested with
-  \href{http://cppcheck.sourceforge.net/}{Cppcheck},
-  \href{https://www.viva64.com/en/pvs-studio/}{PVS-Studio} and
-  \href{https://scan.coverity.com/}{Coverity-Scan}.
-
-  \item More than 240 other features and bugfixes.
+  \item More than XYZ other features and bugfixes.
 \end{itemize}
@@
 can be found
-\href{https://www.dealii.org/8.5.0/doxygen/deal.II/changes_between_8_4_and_8_5.html}{in
-the file that lists all changes for this release}, see \cite{changes85}.
+\href{https://www.dealii.org/8.5.0/doxygen/deal.II/changes_between_8_4_2_and_8_5_0.html}{in
+the file that lists all changes for this release}, see \cite{changes90}.
 (The file is also linked to from the web site of each release as well as
@@
 
-\subsection{The \texttt{CellDataStorage} class and friends}
+\subsection{Feature 1}
 
-The \texttt{CellDataStorage} templated class is an integrated mechanism to
-store a vector of user-defined data within each cell,
-such as a coefficient at each quadrature point.
-Although the same was previously achieved through the use
-of a cell's \texttt{user\_pointer}, it required users to manage this data
-themselves, including transfer of this data if ownership of a cell is
-transferred from one processor to another in parallel
-computations. Instead, this data is now treated as a first-class
-citizen in \dealii{}.
-
-\texttt{CellDataStorage} can work with arbitrary kinds of data defined to live on a cell
-as long as the corresponding data type implements a default constructor.
-Additionally, when a user's custom data type is derived from \texttt{TransferableQuadraturePointData}
-and thereby implements certain interfaces,
-then the \texttt{parallel::distributed::ContinuousQuadratureDataTransfer} class
-can transfer this data during $h$-adaptive refinement from parent cell to children
-(interpreting the data on quadrature points as discrete representations of an
-underlying continuous field),
-as well as from one processor to another during repartitioning.
-
-\subsection{The \texttt{MappingManifold} class}
-
-The \texttt{MappingManifold} class implements the functionality
-of the \texttt{Mapping} interface for \emph{manifold conforming}
-mappings. In other words, instead of using polynomial approximations
-of the objects that describe the boundaries (or interiors) of cells,
-this class computes the transformation between the
-reference and real cell by exploiting the geometrical information
-coming from the underlying \texttt{Manifold} object.
-
-When using this class,
-quadrature points lie on the \textit{exact}
-geometrical objects, and tangent and normal vectors computed
-are tangent and normal to the underlying geometry. This is
-in contrast with the \texttt{MappingQ} and \texttt{MappingQGeneric}
-classes, which approximate the geometry using a polynomial of some
-order, and then compute the normals and tangents using the
-approximated surface.
-
-The class currently only implements the information that relates to
-the mapping itself, as well to its derivatives (such as the Jacobian
-of the mapping, or the determinant thereof). Information related to
-higher order derivatives -- such as the Hessian of the mapping -- will
-result in an exception.
-
-\subsection{Extension of the \texttt{LinearOperator} class}
-
-We have extended the \texttt{LinearOperator} class by a generic
-``payload'' mechanism that allows for the attachment of arbitrary additional
-information to a \texttt{LinearOperator}. This was achieved by introducing
-a generic \texttt{Payload} base class. The main use case of the new
-mechanism is to extend the \texttt{LinearOperator} class to seamlessly
-exploit the native features and operations offered by external linear
-algebra libraries. We have thus developed a \texttt{TrilinosPayload} class
-that provides full support for the suite of {\trilinos} parallel iterative
-solvers and preconditioners.
-
-A particularly interesting case is the construction of an
-\texttt{inverse\_operator}. Whereas previously only \dealii's built-in solvers
-were compatible with this operator, one can now also select those
-offered by {\trilinos}. This has been achieved by using
-the \texttt{Epetra\_Operator} as the basis for the
-\texttt{TrilinosPayload}, for which the result of standard and composite
-operations involving forward (\texttt{Apply()}) and inverse
-(\texttt{ApplyInverse()}) matrix-vector multiplication are collated using
-lambda functions.
-
-We envisage that, in the future, similar extensions can be implemented for the {\petsc}
-iterative solvers.
-
-Another additional in this release related to the \texttt{LinearOperator} suite is the definition
-of a \texttt{schur\_complement} operator and its associated condensation
-and post-processing \texttt{PackagedOperation}s. An operator representing
-the Schur complement of a block system can be declared and, through the
-delayed evaluation offered by \texttt{PackagedOperation}s, reused on any
-number of vector systems.
-
-\subsection{The physics module}
-\label{sec:physics}
-
-We have created a dedicated physics module to facilitate the
-implementation of functions and classes that relate to continuum mechanics,
-physical fields and material constitutive laws. To date, it includes
-transformations of scalar or tensorial quantities between any two
-configurations (by user-specification of a linear map $\mathbf{F}$), and some
-definitions typically utilized in both linear and finite-strain nonlinear
-elasticity.
-
-The \verb!Physics::Transformations! namespace offers push-forward and
-pull-back operations in the context of contravariant, covariant and Piola
-transformations, as well as rotation operations for the Euclidean space.
-Although these transformations are defined in a general manner, one typical
-use of them in finite-strain elasticity would be the determination of the
-Cauchy stress tensor $\boldsymbol{\sigma} = \boldsymbol{\sigma}\left(\mathbf{x}\right)$
-defined at a spatial position $\mathbf{x} \in \mathcal{B}$ from its
-fully referential counterpart, namely the Piola-Kirchhoff stress tensor
-$\mathbf{S} = \mathbf{S}\left(\mathbf{X}\right)$ computed at the material
-coordinate $\mathbf{X} \in \mathcal{B}_{0}$.
-By choosing
-$\mathbf{F} \left(\mathbf{X}\right) = \dfrac{\partial \mathbf{x}\left(\mathbf{X}\right)}{\partial \mathbf{X}}$,
-this is achieved through the action of the Piola push-forward
-$\boldsymbol{\sigma} = \dfrac{1}{\det \mathbf{F}} \, \mathbf{F} \cdot \mathbf{S} \cdot \mathbf{F}^{T}$.
-
-In the \verb!Physics::Elasticity::Kinematics! namespace, a selection of
-deformation, strain and strain rate tensors are defined. The
-\verb!Physics::Elasticity::StandardTensors! class provides some frequently
-used second and fourth order metric tensors, and defines a number of
-referential and spatial projection operators and tensor derivatives that
-are commonly required in the definition of material laws.
-
-The updated step-44 tutorial program demonstrates the
-use of the physics module in the context of a solid mechanics problem.
-Additional updates to this tutorial include application of the new 
-\texttt{CellDataStorage} class to store and retrieve local quadrature point 
-data. An alternative approach to solving the linear system, previously 
-performed by global static condensation of two of the three field variables, 
-has been implemented using the \texttt{LinearOperator} class.
-
-\subsection{Scalability of geometric multigrid framework}
-
-For the new release, the geometric multigrid facilities in \dealii{} have been
-thoroughly overhauled regarding their scalability on large-scale parallel
-systems. During this process, a geometric multigrid implementation based on
-the fast matrix-free kernels from \cite{KronbichlerKormann2012} has been
-benchmarked on up to 147,456 cores. The fast matrix-vector products revealed
-several scalability bottlenecks in the other multigrid components, including
-unnecessary inner products inside the Chebyshev smoother and
-${\mathcal O}(n_\text{levels})$ global communication steps during the
-restriction process rather than the single unavoidable global communication
-step inherent to going to the coarsest grid and the coarse solve. We
-implemented new
-matrix-free transfer implementations called \texttt{MGTransferMatrixFree}
-that can replace the matrix-based \texttt{MGTransferPrebuilt} class
-for tensor product elements. Besides better scalability than the Trilinos
-Epetra matrices underlying the latter, the matrix-free transfer is also a much
-faster for high-order elements with a complexity per degree of freedom of
-$\mathcal O(d p)$ in the polynomial degree $p$ in $d$ dimensions rather than
-$\mathcal O(p^d)$ for the matrices.
-
-\begin{figure}
-\pgfplotstableread{
-nprocs     fem256k    fem2m     fem16m    fem128m  fem1g    fem8g
-16         0.640934   4.97684   nan       nan      nan      nan
-32         0.325741   2.51771   nan       nan      nan      nan
-64         0.1645     1.2823    nan       nan      nan      nan
-128        0.090898   0.658832  5.00366   nan      nan      nan
-256        0.059922   0.339999  2.56216   nan      nan      nan
-512        0.0455449  0.176482  1.29986   nan      nan      nan
-1024       0.0368049  0.099691  0.67364   6.48155  nan      nan
-2048       0.0348921  0.069573  0.356066  2.59601  nan      nan
-4096       0.0367949  0.056833  0.19293   1.3251   nan      nan
-8192       0.033958   0.045350  0.110485  0.790214 5.50143  nan
-16384      0.0379629  0.049351  0.099904  0.424692 2.81479  nan
-32768      0.0461671  0.051276  0.077546  0.229114 1.53091  nan
-65536      0.0466189  0.058941  0.075194  0.127353 0.819    6.2909
-147456     nan        nan       nan       0.087949 0.43213  2.90856
-}\scalinglarge
-\pgfplotstableread{
-nprocs    newdg256k    newdg2m   newdg16m olddg256k  olddg2m   olddg16m
-28        1.4398       12.2014   nan      1.458725   12.25674  nan
-56        0.6987        6.2015   nan      0.721993   6.266567  nan
-112       0.3352        3.1782   nan      0.353412   3.263585  nan
-224       0.1888        1.5687   12.8235  0.183643   1.611967  13.03189
-448       0.0853        0.7686   6.5649   0.105388   0.805367  6.700762
-896       0.0478        0.3537   3.3685   0.063270   0.383875  3.446427
-1792      0.0317        0.1717   1.7061   0.046065   0.199738  1.762310
-3584      0.0235        0.1079   0.8580   0.042032   0.123113  0.902782
-7168      0.0207        0.0668   0.4328   0.037678   0.091822  0.462007
-14336     0.0183699     0.045095 0.23276  0.038804   0.071644  0.288911
-}\scalingHSW
-\centering
-\definecolor{gnuplot@green}{RGB}{0,158,115}
-\begin{tikzpicture}
-    \begin{loglogaxis}[
-      title style={at={(0.5,0.965)},anchor=north,draw=black,fill=white,font=\scriptsize\bf},
-      title={strong and weak scaling, continuous $\mathcal Q_3$ elements},
-      width=0.52\textwidth,
-      height=0.5\textwidth,
-      xlabel={Number of cores},
-      ylabel={Solver time [s]},
-      xtick={32,128,512,2048,8192,32768,147456},
-      xticklabels={32,128,512,2048,8192,32k,147k},
-      tick label style={font=\scriptsize},
-      label style={font=\scriptsize},
-      legend style={font=\scriptsize},
-      legend pos=south west,
-      ymin=5e-3, ymax=15,
-      xmin=8, xmax=147456,
-      grid
-      ]
-      \addplot table[x={nprocs}, y={fem8g}] {\scalinglarge};
-      \addlegendentry{8B cells};
-      \addplot table[x={nprocs}, y={fem1g}] {\scalinglarge};
-      \addlegendentry{1B cells};
-      \addplot table[x={nprocs}, y={fem128m}] {\scalinglarge};
-      \addlegendentry{128M cells};
-      \addplot table[x={nprocs}, y={fem16m}] {\scalinglarge};
-      \addlegendentry{16M cells};
-      \addplot[gnuplot@green,mark=diamond*,mark options={fill=gnuplot@green!40}] table[x={nprocs}, y={fem2m}] {\scalinglarge};
-      \addlegendentry{2M cells};
-      \addplot[dashed,black] coordinates {
-        (8,10)
-        (147456,5/9168)
-      };
-      \addlegendentry{linear scaling};
-      \addplot[dashed,black] coordinates {
-        (16,8*5)
-        (147456,8*5/9168)
-      };
-      \addplot[dashed,black] coordinates {
-        (16,64*5)
-        (147456,64*5/9168)
-      };
-      \addplot[dashed,black] coordinates {
-        (16,5*512)
-        (147456,5*512/9168)
-      };
-      \addplot[dashed,black] coordinates {
-        (16,5*4096)
-        (147456,5*4096/9168)
-      };
-    \end{loglogaxis}
-  \end{tikzpicture}
-  \begin{tikzpicture}
-    \begin{loglogaxis}[
-      title style={at={(1,0.965)},anchor=north east,draw=black,fill=white,font=\scriptsize\bf},
-      title={discontinuous $\mathcal Q_3$ elements},
-      width=0.48\textwidth,
-      height=0.5\textwidth,
-      xlabel={Number of cores},
-      xtick={56,224,896,3584,14336},
-      xticklabels={56,224,896,3584,14336},
-      tick label style={font=\scriptsize},
-      label style={font=\scriptsize},
-      legend style={font=\scriptsize},
-      legend pos=south west,
-      xmin=28, xmax=14336,
-      ymin=5e-3, ymax=15,
-      grid
-      ]
-      \addplot[blue,mark=*,densely dashed] table[x={nprocs}, y={olddg16m}] {\scalingHSW};
-      \addlegendentry{old, 16M cells};
-      \addplot[blue,mark=o] table[x={nprocs}, y={newdg16m}] {\scalingHSW};
-      \addlegendentry{new, 16M cells};
-      \addplot[red,mark=square*,densely dashed] table[x={nprocs}, y={olddg2m}] {\scalingHSW};
-      \addlegendentry{old, 2M cells};
-      \addplot[red,mark=square] table[x={nprocs}, y={newdg2m}] {\scalingHSW};
-      \addlegendentry{new, 2M cells};
-      \addplot[gnuplot@green,mark=diamond*,densely dashed] table[x={nprocs}, y={olddg256k}] {\scalingHSW};
-      \addlegendentry{old, 256k cells};
-      \addplot[gnuplot@green,mark=diamond] table[x={nprocs}, y={newdg256k}] {\scalingHSW};
-      \addlegendentry{new, 256k cells};
-    \end{loglogaxis}
-  \end{tikzpicture}
-  \caption{Scaling of \dealii{}'s geometric multigrid algorithms on
-    SuperMUC. Each line corresponds to a strong scaling experiment,
-    increasing the number of processor cores for a fixed-size
-    problem. Comparing corresponding data points on different lines
-    yields weak scaling information.}
-\label{fig:scaling_mg}
-\end{figure}
-
-The scalability of the improved geometric multigrid framework is shown in
-Fig.~\ref{fig:scaling_mg}, including a combined strong and weak scaling plot
-in the left panel using continuous $\mathcal Q_3$ elements with 57~million to
-232~billion degrees of freedom for discretizing the Laplacian. Along each
-line, the same problem size is solved with an increasing number of cores,
-whereas different lines are a factor of eight apart and always start at
-3.5~million degrees of freedom per core with an absolute performance of around
-650,000 degrees of freedom per core and second. Almost ideal scalability down
-to approximately 0.1~seconds can be observed also on 147k~cores. The right
-panel of Fig.~\ref{fig:scaling_mg} shows the effect of the aforementioned
-algorithmic improvements on a setup with discontinuous DG elements, clearly
-improving the latency of the multigrid V-cycle.
-
-The updated step-37 tutorial
-programs presents the updated algorithms and the MPI-parallel multigrid
-setting with matrix-free operator evaluation.
-
-\subsection{Matrix-free operators}
-
-In order to facilitate the usage of matrix-free methods, a
-\verb!MatrixFreeOperator::Base! class has been introduced, implementing
-functionality for matrix-vector products and the necessary operations
-for the interface residuals for multigrid on adaptively refined meshes
-(see \cite{JanssenKanschat2011}). Furthermore, the class is compatible
-with the linear operator framework and provides an interface to a Jacobi
-preconditioner. Derived classes only need to implement the
-\verb!apply_add()! method that is used in the \verb!vmult()! functions, and
-a method to compute the diagonal entries of the underlying matrix. The
-\verb!MatrixFreeOperator! namespace contains implementations of
-\verb!MatrixFreeOperators::LaplaceOperator! and
-\verb!MatrixFreeOperators::MassOperator!.
-
-The updated step-37 tutorial
-program makes use of these facilities and explains their usage in detail.
-Using the matrix-free mass operator, \verb!VectorTools::project! has become
-much faster than the previous matrix-based approach for elements supported
-by \verb!MatrixFree! and also works for parallel computations based on MPI.
-
-\subsection{The \texttt{FE\_Enriched} class}
-
-The \verb!FE_Enriched! finite element implements a partition of unity
-finite element method (PUM) by Babuska and Melenk which enriches a standard
-finite element with an enrichment function multiplied with another (usually
-linear) finite element. This allows including
-a priori knowledge about the partial differential equation being solved
-in the finite element space,
-which in turn improves the local approximation properties of the spaces.
-Programs can also use enriched and non-enriched finite elements in
-different parts of the domain.
-
-The
-\verb|DoFTools::make_hanging_node_constraints()| function can automatically
-make the resulting space $C^0$ continuous. The existing \verb|SolutionTransfer|
-class can be used to transfer the solution during $h$\,-adaptive refinement
-from a coarse to a fine mesh under the condition that all child elements
-are also enriched.
-
-\subsection{The \texttt{FESeries} namespace}
-
-The \verb|FESeries| namespace offers functions to calculate expansion
-series of the solution on the reference element. Coefficients of expansion
-are often used to estimate local smoothness of the underlying finite
-element field to decide on a $h$- or $p$-adaptive refinement strategy.
-Specifically, \verb|FESeries::Legendre| calculates expansion of a
-scalar finite element
-field into series of Legendre functions on the reference element, whereas
-\verb|FESeries::Fourier| calculates Fourier coefficients. Programs
-using this functionality have to
-specify the required number of coefficients in each direction as well as
-provide a collection of finite elements and quadrature rules; these will be
-used for the calculation of the transformation matrices.
-
-The updated step-27 tutorial program demonstrates the
-use of \verb|FESeries::Fourier|.
+\subsection{Feature 2}
 
 \subsection{New and updated tutorial programs}
-
+\marginpar{OLD}
 In addition to the updated tutorial programs mentioned in the previous
-section, this release of \dealii{} includes three new tutorials:
+section, this release of \dealii{} includes three new tutorials TODO: old:
 \begin{itemize}
@@
 the following topics:
-%\marginpar{Add the muscle elasticity example if it is merged in time.}
+\marginpar{OLD}
   \begin{itemize}
@@
 
-The 8.5 release includes around 20
-\href{https://www.dealii.org/developer/doxygen/deal.II/changes_between_8_4_and_8_5.html}{
-incompatible changes}; see \cite{changes85}. The majority of these changes
+\marginpar{OLD}
+The 9.0 release includes around XYZ
+\href{https://www.dealii.org/8.5.0/doxygen/deal.II/changes_between_8_4_2_and_8_5_0.html}{
+incompatible changes}; see \cite{changes90}. The majority of these changes
 should not be visible to typical user codes; some remove previously
@@
 
+\marginpar{update}
 \dealii{} can interface with many other libraries:
@@
 
-Older releases of \dealii{} can be cited as \cite{dealII80,dealII81,dealII82,dealII83,dealII84}.
+Older releases of \dealii{} can be cited as \cite{dealII80,dealII81,dealII82,dealII83,dealII84,dealII85}.
 
@@
 %
-
+\marginpar{update}
   Rajat Arora,
@@
 variety of funding sources:
+\marginpar{update}
 
-D.~Arndt was supported by the German Research Foundation (DFG) under the
-project ``High-order discontinuous Galerkin for the exa-scale'' (ExaDG) within the
-priority program ``Software for Exascale Computing'' (SPPEXA).
-
-W.~Bangerth was partially
-supported by the National Science Foundation under award OCI-1148116
-as part of the Software Infrastructure for Sustained Innovation (SI2)
-program; and by the Computational Infrastructure in Geodynamics initiative
-(CIG), through the National Science Foundation under Awards
-No.~EAR-0949446 and EAR-1550901 and The University of California -- Davis.
-
-D.~Davydov was supported by the European Research Council (ERC) through the Advanced Grant 289049 MOCOPOLY and the Competence Network for Technical and Scientific High Performance Computing in Bavaria (KONWIHR).
+% D.~Arndt was supported by the German Research Foundation (DFG) under the
+% project ``High-order discontinuous Galerkin for the exa-scale'' (ExaDG) within the
+% priority program ``Software for Exascale Computing'' (SPPEXA).
+% 
+% W.~Bangerth was partially
+% supported by the National Science Foundation under award OCI-1148116
+% as part of the Software Infrastructure for Sustained Innovation (SI2)
+% program; and by the Computational Infrastructure in Geodynamics initiative
+% (CIG), through the National Science Foundation under Awards
+% No.~EAR-0949446 and EAR-1550901 and The University of California -- Davis.
+% 
+% D.~Davydov was supported by the European Research Council (ERC) through the Advanced Grant 289049 MOCOPOLY and the Competence Network for Technical and Scientific High Performance Computing in Bavaria (KONWIHR).
 
@@
 Geodynamics initiative (CIG), through the National Science Foundation
-under Award No. EAR-0949446 and The University of California -- Davis, and National Science Foundation grant DMS1522191.
-
-M.~Kronbichler was partially supported by the German Research Foundation (DFG)
-under the project ``High-order discontinuous Galerkin for the exa-scale''
-(ExaDG) within the priority program ``Software for Exascale Computing''
-(SPPEXA), grant agreement no.~KR4661/2-1, the Bayerisches Kompetenznetzwerk
-f\"ur Technisch-Wissenschaftliches Hoch- und H\"ochstleistungsrechnen
-(KONWIHR), and the Gauss Centre for Supercomputing e.V.~by providing computing
-time on the GCS Supercomputer SuperMUC at Leibniz Supercomputing Centre (LRZ)
-through project id pr83te.
-
-J-P.~Pelteret was supported by the European Research Council (ERC) through the Advanced Grant 289049 MOCOPOLY.
-
-B.~Turcksin: This material is based upon work supported by the U.S. Department of
-Energy, Office of Science, under contract number DE-AC05-00OR22725.
-
-D.~Wells was supported by the National Science Foundation (NSF) through Grant
-DMS-1344962.
+under Award No. EAR-0949446 and EAR-1550901 and The University of California -- Davis, and National Science Foundation grant DMS-1522191.
+% 
+% M.~Kronbichler was partially supported by the German Research Foundation (DFG)
+% under the project ``High-order discontinuous Galerkin for the exa-scale''
+% (ExaDG) within the priority program ``Software for Exascale Computing''
+% (SPPEXA), grant agreement no.~KR4661/2-1, the Bayerisches Kompetenznetzwerk
+% f\"ur Technisch-Wissenschaftliches Hoch- und H\"ochstleistungsrechnen
+% (KONWIHR), and the Gauss Centre for Supercomputing e.V.~by providing computing
+% time on the GCS Supercomputer SuperMUC at Leibniz Supercomputing Centre (LRZ)
+% through project id pr83te.
+% 
+% J-P.~Pelteret was supported by the European Research Council (ERC) through the Advanced Grant 289049 MOCOPOLY.
+% 
+% B.~Turcksin: This material is based upon work supported by the U.S. Department of
+% Energy, Office of Science, under contract number DE-AC05-00OR22725.
+% 
+% 
+% D.~Wells was supported by the National Science Foundation (NSF) through Grant
+% DMS-1344962.
 

### 2018-04-04 9.0/paper.tex
add

@@
 \begin{itemize}
-\item The \texttt{CellDataStorage} ...
+\item Manifold smoothing: The manifold smoothing algorithms applied in the Triangulation class and MappingQGeneric have been changed from the old Laplace-style smoothing to a transfinite interpolation that linearly blends between the descriptions on the faces around a cell.
+The old transformation introduced boundary layers inside cells that prevented convergence rates from exceeding 3.5 in the global L2 errors on typical settings.
+This change also considerably improves mesh quality on settings where curved descriptions are only applied to the boundary rather than the whole volume.
+
+\item C++11 is now required and used in many places, move semantics, ... (...)
+
+\item
 

### 2018-04-04 9.0/paper.tex
list of changes url

@@
 \dealii{} version 9.0.0 was released April X, 2018. This paper provides an
-overview of the new features of this release and serves as a citable
+overview of the new features of this major release and serves as a citable
 reference for the \dealii{} software library version 9.0. \dealii{} is an
@@
 can be found
-\href{https://www.dealii.org/8.5.0/doxygen/deal.II/changes_between_8_4_2_and_8_5_0.html}{in
-the file that lists all changes for this release}, see \cite{changes90}.
+\href{https://www.dealii.org/developer/doxygen/deal.II/changes_between_8_5_and_9_0.html}{
+in the file that lists all changes for this release}, see \cite{changes90}.
 (The file is also linked to from the web site of each release as well as
@@
 The 9.0 release includes around XYZ
-\href{https://www.dealii.org/8.5.0/doxygen/deal.II/changes_between_8_4_2_and_8_5_0.html}{
+\href{https://www.dealii.org/developer/doxygen/deal.II/changes_between_8_5_and_9_0.html}{
 incompatible changes}; see \cite{changes90}. The majority of these changes

### 2018-04-04 9.0/paper.tex
add myself, add a couple of points

@@
 % 
-% \author[7]{Matthias Maier}
-% \affil[7]{School of Mathematics,
-%   University of Minnesota,
-%   127 Vincent Hall, 206 Church Street SE,
-%   Minneapolis, MN 55455, USA.
-%   {\texttt{msmaier@umn.edu}}}
+\author[7]{Matthias Maier}
+\affil[7]{School of Mathematics,
+  University of Minnesota,
+  127 Vincent Hall, 206 Church Street SE,
+  Minneapolis, MN 55455, USA.
+  {\texttt{msmaier@umn.edu}}}
 % 
@@
 \begin{itemize}
-\item Manifold smoothing: The manifold smoothing algorithms applied in the Triangulation class and MappingQGeneric have been changed from the old Laplace-style smoothing to a transfinite interpolation that linearly blends between the descriptions on the faces around a cell.
-The old transformation introduced boundary layers inside cells that prevented convergence rates from exceeding 3.5 in the global L2 errors on typical settings.
-This change also considerably improves mesh quality on settings where curved descriptions are only applied to the boundary rather than the whole volume.
-
-\item C++11 is now required and used in many places, move semantics, ... (...)
+\item
+  Manifold smoothing: The manifold smoothing algorithms applied in the
+  Triangulation class and MappingQGeneric have been changed from the old
+  Laplace-style smoothing to a transfinite interpolation that linearly
+  blends between the descriptions on the faces around a cell. The old
+  transformation introduced boundary layers inside cells that prevented
+  convergence rates from exceeding 3.5 in the global L2 errors on typical
+  settings. This change also considerably improves mesh quality on settings
+  where curved descriptions are only applied to the boundary rather than
+  the whole volume.
 
 \item
+  C++11 is now required and used in many places, move semantics, ... (...)
+
+\item
+  Improved support for LinearOperator (Trilinos payload, Schur complement,
+  iterative inverse), removal of old wrapper classes
+
+\item
+  Initial support for generalized support points: Internal interpolation
+  rewritten (support for $H(\text{curl})$ and $H(\text{div})$ conforming
+  elements).
 

### 2018-04-04 9.0/paper.tex
Add myself to the authors

@@
   pdfauthor={
+    Daniel Arndt,
     Timo Heister,
@@
 
-% \author[1]{Daniel Arndt}
-% \affil[1]{Interdisciplinary Center for Scientific Computing,
-%     Heidelberg University, Im Neuenheimer Feld 205, 69120 Heidelberg, Germany.
-%     {\texttt{daniel.arndt@iwr.uni-heidelberg.de}}}
-% 
+\author[1]{Daniel Arndt}
+\affil[1]{Interdisciplinary Center for Scientific Computing,
+  Heidelberg University,
+  Im Neuenheimer Feld 205,
+  69120 Heidelberg, Germany.
+  {\texttt{daniel.arndt@iwr.uni-heidelberg.de}}}
+
 % \author[2]{Wolfgang Bangerth}

### 2018-04-04 9.0/paper.tex
Add list of major changes.

@@
 % 
-% \author[9]{Bruno Turcksin\footnote{
-%   This manuscript has been authored by UT-Battelle, LLC under Contract No.
-%   DE-AC05-00OR22725 with the U.S. Department of Energy. The United States
-%   Government retains and the publisher, by accepting the article for
-%   publication, acknowledges that the United States Government retains a
-%   non-exclusive, paid-up, irrevocable, worldwide license to publish or reproduce
-%   the published form of this manuscript, or allow others to do so, for United
-%   States Government purposes. The Department of Energy will provide public
-%   access to these results of federally sponsored research in accordance with the
-%   DOE Public Access Plan (http://energy.gov/downloads/doe-public-access-plan).}}
-% \affil[9]{Computational Engineering and Energy Sciences Group,
-%   Computional Sciences and Engineering Division,
-%   Oak Ridge National Laboratory, 1 Bethel Valley Rd.,
-%   TN 37831, USA.
-%   {\texttt{turcksinbr@ornl.gov}}}
+ \author[9]{Bruno Turcksin\footnote{
+   This manuscript has been authored by UT-Battelle, LLC under Contract No.
+   DE-AC05-00OR22725 with the U.S. Department of Energy. The United States
+   Government retains and the publisher, by accepting the article for
+   publication, acknowledges that the United States Government retains a
+   non-exclusive, paid-up, irrevocable, worldwide license to publish or reproduce
+   the published form of this manuscript, or allow others to do so, for United
+   States Government purposes. The Department of Energy will provide public
+   access to these results of federally sponsored research in accordance with the
+   DOE Public Access Plan (http://energy.gov/downloads/doe-public-access-plan).}}
+ \affil[9]{Computational Engineering and Energy Sciences Group,
+   Computional Sciences and Engineering Division,
+   Oak Ridge National Laboratory, 1 Bethel Valley Rd.,
+   TN 37831, USA.
+   {\texttt{turcksinbr@ornl.gov}}}
 % 
@@
 \item 
+  Support for GPU
+
+\item
+  Increase support for dense solvers (LAPACK changes and new ScaLAPACK
+  interface).
+
+\item
+  Support for Sundials
+
+\item 
+  Support for Particles
+
+\item 
+  Support for automatically differentiable numbers
+
+\item 
+  Support for complex-values vectors is now at the same level as real-valued
+  vectors.
+
+\item
   Initial support for generalized support points: Internal interpolation
@@
 
-\item New tutorial programs XYZ; as well as
-  updates to XYZ. In addition, the separate code
+\item New python tutorial program tutorial-1; as well as
+  updates to step-37. In addition, the separate code
   gallery of \dealii{} has gained a number of new entries.
 
-  \item More than XYZ other features and bugfixes.
+  \item More than 330 other features and bugfixes.
 \end{itemize}

### 2018-04-04 9.0/paper.tex
Add funding; remove svn archive

@@
 
-% D.~Arndt was supported by the German Research Foundation (DFG) under the
-% project ``High-order discontinuous Galerkin for the exa-scale'' (ExaDG) within the
-% priority program ``Software for Exascale Computing'' (SPPEXA).
-% 
+D.~Arndt was partially supported by the German Research Foundation (DFG) under the
+project ``High-order discontinuous Galerkin for the exa-scale'' (ExaDG) within the
+priority program ``Software for Exascale Computing'' (SPPEXA).
+
 % W.~Bangerth was partially
@@
 The Interdisciplinary Center for Scientific Computing (IWR) at Heidelberg University has provided
-hosting services for the \dealii{} web page and the SVN archive.
+hosting services for the \dealii{} web page.
 

### 2018-04-05 9.0/paper.tex
add myself, fix trailing whitespace and one bibtex reference
Some small changes.

@@
     Daniel Arndt,
+    Denis Davydov,
     Timo Heister,
+    Luca Heltai,
+    Matthias Maier
   },
@@
 
-% \author[2]{Wolfgang Bangerth}
-% \affil[2]{Department of Mathematics, Colorado State University, Fort
-%   Collins, CO 80523-1874, USA.
-%     {\texttt{bangerth@colostate.edu}}}
-% 
-% \author[3]{Denis Davydov}
-% \affil[3]{Chair of Applied Mechanics, University of
-%      Erlangen-Nuremberg, Egerlandstr.\ 5, 91058 Erlangen, Germany.
-%     {\texttt{denis.davydov@fau.de}}}
+\author[2]{Wolfgang Bangerth}
+\affil[2]{Department of Mathematics, Colorado State University, Fort
+  Collins, CO 80523-1874, USA.
+    {\texttt{bangerth@colostate.edu}}}
+
+\author[3]{Denis Davydov}
+\affil[3]{Chair of Applied Mechanics, University of
+     Erlangen-Nuremberg, Egerlandstr.\ 5, 91058 Erlangen, Germany.
+    {\texttt{denis.davydov@fau.de}}}
 
@@
 %
-% \author[5]{Luca Heltai}
-% \affil[5]{SISSA,
-%   International School for Advanced Studies,
-%   Via Bonomea 265,
-%   34136, Trieste, Italy.
-% {\texttt{luca.heltai@sissa.it}}}
+\author[5]{Luca Heltai}
+\affil[5]{SISSA,
+  International School for Advanced Studies,
+  Via Bonomea 265,
+  34136, Trieste, Italy.
+{\texttt{luca.heltai@sissa.it}}}
 %
@@
   {\texttt{msmaier@umn.edu}}}
-% 
-% \author[8]{Jean-Paul Pelteret}
-% \affil[8]{Chair of Applied Mechanics,
-%   University of Erlangen-Nuremberg,
-%   Egerlandstr.\ 5,
-%   91058 Erlangen,
-%   Germany.
-%   {\texttt{jean-paul.pelteret@fau.de}}}
-% 
+
+\author[8]{Jean-Paul Pelteret}
+\affil[8]{Chair of Applied Mechanics,
+  University of Erlangen-Nuremberg,
+  Egerlandstr.\ 5,
+  91058 Erlangen,
+  Germany.
+  {\texttt{jean-paul.pelteret@fau.de}}}
+
  \author[9]{Bruno Turcksin\footnote{
@@
 \item
-  Support for Sundials
+  Support for Sundials and ROL
 
@@
   vectors.
+
+\item 
+  Support for Nanoflann
+
+\item 
+  Support for Assimp
+
+\item 
+  Initial support for GMHS
 
@@
   \end{itemize}
+
+\subsection{New non-standard quadrature rules}
+
+\begin{itemize}
+\item QSimplex
+\item QDuffy
+\item QTrianglePolar
+\item QSplit
+\end{itemize}
+
+Mostly used for BEM problems.
 
@@
 \item ARPACK \cite{arpack}
+\item Assimp \cite{assimp}
 \item BLAS, LAPACK
@@
 \item SLEPc \cite{Hernandez:2005:SSF}
+\item SUNDIALS \cite{sundials}
 \item Threading Building Blocks \cite{Rei07}
@@
 %
-% D.~Davydov was supported by the European Research Council (ERC) through the Advanced Grant 289049 MOCOPOLY and the Competence Network for Technical and Scientific High Performance Computing in Bavaria (KONWIHR).
+D.~Davydov was supported by the German Science Foundation
+(Deutsche Forschungs-Gemeinschaft, DFG), grant DA 1664/2-1.
 
@@
 \end{document}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2018-04-05 9.0/paper.tex
Extend the list of changes.

@@
     Luca Heltai,
+    Martin Kronbichler
     Matthias Maier
@@
 {\texttt{luca.heltai@sissa.it}}}
-%
-% \author[6]{Martin Kronbichler}
-% \affil[6]{Institute for Computational Mechanics,
-%   Technical University of Munich,
-%   Boltzmannstr.~15, 85748 Garching, Germany.
-%   {\texttt{kronbichler@lnm.mw.tum.de}}}
-%
+
+\author[6]{Martin Kronbichler}
+\affil[6]{Institute for Computational Mechanics,
+  Technical University of Munich,
+  Boltzmannstr.~15, 85748 Garching, Germany.
+  {\texttt{kronbichler@lnm.mw.tum.de}}}
+
 \author[7]{Matthias Maier}
@@
   iterative inverse), removal of old wrapper classes
+
+\item
+  Significant extension of matrix-free capabilities, including support for face integrals in discontinuous Galerkin schemes, a new Hermite-like polynomial basis adapted to face integrals that involve derivatives of shape functions, and several performance enhancements
 
@@
 under Award No. EAR-0949446 and EAR-1550901 and The University of California -- Davis, and National Science Foundation grant DMS-1522191.
-%
-% M.~Kronbichler was partially supported by the German Research Foundation (DFG)
-% under the project ``High-order discontinuous Galerkin for the exa-scale''
-% (ExaDG) within the priority program ``Software for Exascale Computing''
-% (SPPEXA), grant agreement no.~KR4661/2-1, the Bayerisches Kompetenznetzwerk
-% f\"ur Technisch-Wissenschaftliches Hoch- und H\"ochstleistungsrechnen
-% (KONWIHR), and the Gauss Centre for Supercomputing e.V.~by providing computing
-% time on the GCS Supercomputer SuperMUC at Leibniz Supercomputing Centre (LRZ)
-% through project id pr83te.
-%
+
+M.~Kronbichler was partially supported by the German Research Foundation (DFG)
+under the project ``High-order discontinuous Galerkin for the exa-scale''
+(ExaDG) within the priority program ``Software for Exascale Computing''
+(SPPEXA), grant agreement no.~KR4661/2-1 and the Bayerisches Kompetenznetzwerk
+f\"ur Technisch-Wissenschaftliches Hoch- und H\"ochstleistungsrechnen
+(KONWIHR).
+
 % J-P.~Pelteret was supported by the European Research Council (ERC) through the Advanced Grant 289049 MOCOPOLY.

### 2018-04-05 9.0/paper.tex
Augment description of some of the major changes

@@
 \item
-  C++11 is now required and used in many places, move semantics, ... (...)
+  C++11 is now required and used in many places. For performance reason,
+  move semantics is used when possible and in many places, \texttt{emplace\_back} 
+  replaces \texttt{push\_back()}. C-style raw pointers have been replaced by
+  \texttt{shared\_ptr} and \texttt{unique\_ptr} to avoid the risk of memory leak.
+  Private constructors have been removed and instead were delete. clang-tidy is
+  now being used to ensure the quality of the code.
 
@@
 \item
-  Increase support for dense solvers (LAPACK changes and new ScaLAPACK
+  Increased support for dense solvers (LAPACK changes and new ScaLAPACK
   interface).
@@
 \item
-  Support for Particles
+  A dedicated particles module has been created to provide support for the
+  storage of particles and their properties. This module provides a base class
+  \texttt{Particles} which represents a particle with position, an ID number and
+  a variable number of properties. It also provides a \texttt{ParticleHandler}
+  class that manages the storage and handling of particles. This class
+  distributes the particles in distributed domains (these are the only domains
+  supported by this class so far)
 
 \item
-  Support for automatically differentiable numbers
+  A dedicated differentiation module had been created to provide support for
+  automatic and symbolic differentiation. Automatic differentiation is a set of
+  technique to evaluate the derivates of a function defined by a compute
+  program. Currently the Adol-C and Sacado libraries are supported through a
+  unified interface. In practice, this support means that Adol-c and Sacado
+  data types can be used in the \texttt{FEValues} and \texttt{FEValuesViews}
+  to represent degree-of-freedom values.
 
@@
 \item
-  Support for Nanoflann
+  Support for the nanoflann library. nanoflann is a library for building
+  KD-trees of datasets and for the fast querying of closest neighbors in
+  KD-trees. nanoflann is used to speed up the search of the p nearest neighbors
+  of a given point, or searching the points that fall within a radius of a
+  target point.
 
 \item
-  Support for Assimp
+  Support for the Open Asset Import Library (Assimp). Assimp can be used to read
+  about 40 different 3D graphics formats. A subset of these formats can be used
+  to generate two-dimensional meshes possibly embedded in a three-dimensional
+  space.
 
 \item
-  Initial support for GMHS
+  Initial support for Gmsh
 
@@
 \begin{itemize}
+\item Adol-c \cite{adol-c}
 \item ARPACK \cite{arpack}
@@
 \item muparser \cite{muparser-web-page}
+\item nanoflann \cite{nanoflann}
 \item NetCDF \cite{rew1990netcdf}

### 2018-04-06 9.0/paper.tex
Remove emacs macros from bottom of file.
This is now deprecated and emacs warns about it.
Minor language edits seen at first glance and without actual reading.

@@
 \item
-  A dedicated differentiation module had been created to provide support for
+  A dedicated differentiation module has been created to provide support for
   automatic and symbolic differentiation. Automatic differentiation is a set of
@@
 \item
-  Support for complex-values vectors is now at the same level as real-valued
+  Support for complex-valued vectors is now at the same level as real-valued
   vectors.
@@
 \end{document}
-
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: t
-%%% End:

### 2018-04-06 9.0/paper.tex
Expand major changes section and add references

@@
 \item
-  Support for Sundials and ROL
+  A dedicated sundials module has been created to provide support for the
+  SUNDIALS library. SUNDIALS is a SUite of Nonlinear and
+  DIfferential/ALgebraic equations Solvers. ARKode, IDA, and KINSOL are
+  supported. ARKode is a solver library that provides adaptive-step time
+  integration. IDA is a package for the solution of differential-algebraic
+  equations systems in the form $F(t,y,y')=0$. KINSOL is olver for nonlinear
+  algebraic systems.
+
+\item 
+  A dedicated optimization/rol module has been created to provide support for
+  the Rapid Optimization Library (ROL). ROL is a package for large-scale
+  optimization.
 
@@
 %
-% B.~Turcksin: This material is based upon work supported by the U.S. Department of
-% Energy, Office of Science, under contract number DE-AC05-00OR22725.
+B.~Turcksin: This material is based upon work supported by the U.S. Department of
+Energy, Office of Science, under contract number DE-AC05-00OR22725.
 %

### 2018-04-07 9.0/paper.tex
Add myself to the 9.0 paper.

@@
     Martin Kronbichler
-    Matthias Maier
+    Matthias Maier,
+    David Wells
   },
@@
    {\texttt{turcksinbr@ornl.gov}}}
-%
-% \author[10]{David Wells}
-% \affil[10]{Department of Mathematical Sciences, Rensselaer Polytechnic Institute,
-% Troy, NY 12180, USA.
-%   {\texttt{wellsd2@rpi.edu}}}
+
+\author[10]{David Wells}
+\affil[10]{Department of Mathematical Sciences, Rensselaer Polytechnic Institute,
+Troy, NY 12180, USA.
+  {\texttt{wellsd2@rpi.edu}}}
 
@@
 Energy, Office of Science, under contract number DE-AC05-00OR22725.
-%
-%
-% D.~Wells was supported by the National Science Foundation (NSF) through Grant
-% DMS-1344962.
+
+D.~Wells was supported by the National Science Foundation (NSF) through Grant
+DMS-1344962.
 

### 2018-04-07 9.0/paper.tex
Fill out the LinearOperator paragraph.

--- 2018-04-06 9.0/paper.tex
@@
 \item
-  Improved support for LinearOperator (Trilinos payload, Schur complement,
-  iterative inverse), removal of old wrapper classes
+  \texttt{LinearOperator}, a flexible template class that implements the action of a
+  linear operator (see \cite{MaierBardelloniHeltai-2016-b}), now supports
+  computations with Trilinos, Schur complements, and linear constraints. This
+  class is, as of this release, the official replacement for about half a dozen
+  similar (but less general) classes, such as \texttt{FilteredMatrix},
+  \texttt{IterativeInverse}, and \texttt{PointerMatrix}.
 

### 2018-04-07 9.0/paper.tex
Fix two typos.

--- 2018-04-06 9.0/paper.tex
@@
   integration. IDA is a package for the solution of differential-algebraic
-  equations systems in the form $F(t,y,y')=0$. KINSOL is olver for nonlinear
+  equations systems in the form $F(t,y,y')=0$. KINSOL is solver for nonlinear
   algebraic systems.
@@
   automatic and symbolic differentiation. Automatic differentiation is a set of
-  technique to evaluate the derivates of a function defined by a compute
+  technique to evaluate the derivatives of a function defined by a compute
   program. Currently the Adol-C and Sacado libraries are supported through a

### 2018-04-07 9.0/paper.tex
Fix two typos.

--- 2018-04-07 9.0/paper.tex
@@
 \item
-  C++11 is now required and used in many places. For performance reason,
-  move semantics is used when possible and in many places, \texttt{emplace\_back} 
-  replaces \texttt{push\_back()}. C-style raw pointers have been replaced by
-  \texttt{shared\_ptr} and \texttt{unique\_ptr} to avoid the risk of memory leak.
-  Private constructors have been removed and instead were delete. clang-tidy is
-  now being used to ensure the quality of the code.
+  \dealii{} first offered support for C++11 features in version 6.2, which was
+  published in 2009: This release is the first to \emph{require} a compiler
+  supporting C++11. \dealii{} now uses language improvements such as using
+  \texttt{emplace\_back()} instead of \texttt{push\_back()}, moving objects
+  instead of copying them, using \texttt{nullptr} instead of \texttt{NULL}, and
+  marking unimplemented constructors with \texttt{=delete;} instead of making
+  them \texttt{private}. These changes include some minor incompatibilities: all
+  \texttt{clone} functions (such as \texttt{FiniteElement::clone} and
+  \texttt{Mapping::clone}) now return \texttt{std::unique\_ptr}s instead of
+  C-style raw pointers. Indeed, nearly every interface that returns a pointer
+  now returns either a \texttt{std::shared\_ptr} or \texttt{std::unique\_ptr},
+  which clarifies object ownership responsibilities and avoids memory leaks.
+
+\item \dealii{} has made extensive use of both the clang-tidy and coverity
+      static analysis tools for detecting bugs and other issues in the code.
+      %% TODO is there somewhere we can cite for either of these tools? It would
+      %% be nice to recognize the department of homeland security for offering
+      %% coverity.
+      %% TODO at the time of writing this coverity is down; get an issue count
+      %% when it is back online.
 
 \item
-  Improved support for LinearOperator (Trilinos payload, Schur complement,
-  iterative inverse), removal of old wrapper classes
+  \texttt{LinearOperator}, a flexible template class that implements the action of a
+  linear operator (see \cite{MaierBardelloniHeltai-2016-b}), now supports
+  computations with Trilinos, Schur complements, and linear constraints. This
+  class is, as of this release, the official replacement for about half a dozen
+  similar (but less general) classes, such as \texttt{FilteredMatrix},
+  \texttt{IterativeInverse}, and \texttt{PointerMatrix}.
 

### 2018-04-07 9.0/paper.tex
Use math syntax in two places.

@@
   transformation introduced boundary layers inside cells that prevented
-  convergence rates from exceeding 3.5 in the global L2 errors on typical
+  convergence rates from exceeding \(3.5\) in the global \(L^2\) errors on typical
   settings. This change also considerably improves mesh quality on settings

### 2018-04-07 9.0/paper.tex
Rearrange the C++11 and static analysis text.

--- 2018-04-07 9.0/paper.tex
@@
     Martin Kronbichler
-    Matthias Maier,
-    David Wells
+    Matthias Maier
   },
@@
    {\texttt{turcksinbr@ornl.gov}}}
-
-\author[10]{David Wells}
-\affil[10]{Department of Mathematical Sciences, Rensselaer Polytechnic Institute,
-Troy, NY 12180, USA.
-  {\texttt{wellsd2@rpi.edu}}}
+%
+% \author[10]{David Wells}
+% \affil[10]{Department of Mathematical Sciences, Rensselaer Polytechnic Institute,
+% Troy, NY 12180, USA.
+%   {\texttt{wellsd2@rpi.edu}}}
 
@@
 \item
-  \texttt{LinearOperator}, a flexible template class that implements the action of a
-  linear operator (see \cite{MaierBardelloniHeltai-2016-b}), now supports
-  computations with Trilinos, Schur complements, and linear constraints. This
-  class is, as of this release, the official replacement for about half a dozen
-  similar (but less general) classes, such as \texttt{FilteredMatrix},
-  \texttt{IterativeInverse}, and \texttt{PointerMatrix}.
+  Improved support for LinearOperator (Trilinos payload, Schur complement,
+  iterative inverse), removal of old wrapper classes
 
@@
   integration. IDA is a package for the solution of differential-algebraic
-  equations systems in the form $F(t,y,y')=0$. KINSOL is solver for nonlinear
+  equations systems in the form $F(t,y,y')=0$. KINSOL is olver for nonlinear
   algebraic systems.
@@
   automatic and symbolic differentiation. Automatic differentiation is a set of
-  technique to evaluate the derivatives of a function defined by a compute
+  technique to evaluate the derivates of a function defined by a compute
   program. Currently the Adol-C and Sacado libraries are supported through a
@@
 Energy, Office of Science, under contract number DE-AC05-00OR22725.
-
-D.~Wells was supported by the National Science Foundation (NSF) through Grant
-DMS-1344962.
+%
+%
+% D.~Wells was supported by the National Science Foundation (NSF) through Grant
+% DMS-1344962.
 

### 2018-04-10 9.0/paper.tex
fill-in Lapack/ScaLAPACK

@@
 \item
-  Significant extension of matrix-free capabilities, including support for face integrals in discontinuous Galerkin schemes, a new Hermite-like polynomial basis adapted to face integrals that involve derivatives of shape functions, and several performance enhancements
+  Significant extension of matrix-free capabilities, including support for face
+  integrals in discontinuous Galerkin schemes, a new Hermite-like polynomial
+  basis adapted to face integrals that involve derivatives of shape functions,
+  and several performance enhancements.
 
@@
 \item
-  Increased support for dense solvers (LAPACK changes and new ScaLAPACK
-  interface).
+  In this version we add template wrappers for ScaLAPACK -- parallel dense linear algebra
+  library with block-cyclic matrix distribution over 2D process grid.
+  The functionality and interface is similar to the LAPACK wrappers,
+  matrix-matrix multiplication, Cholesky and LU factorizations,
+  eigensolvers, SVD, pseudoinverse, save/load in HDF5 format and
+  other functionality is implemented.
+  A user can combine the new linear algebra with other MPI-parallel
+  algorithms in \dealii{} without the need to worry about MPI communicator
+  being divisible exactly into 2D process grid. The processes which are not part of
+  the grid are labeled inactive and ScaLAPACK routines will not be called on them.
+  %
+  Additionally, we improved support of LAPACK. We added methods to perform
+  rank-1 update/downdate, calculate Cholesky factorization, trace and determinant,
+  as well as estimate the reciprocal condition number.
+  We also now support configuration with 64-bit BLAS.
 

### 2018-04-11 9.0/paper.tex
Minor edits.

@@
 \item
-  \dealii{} first offered support for C++11 features in version 6.2, which was
-  published in 2009: This release is the first to \emph{require} a compiler
+  \dealii{} first offered support for a subset of C++11 features in
+  version 6.2, released in 2009. The current release is the first to \emph{require} a compiler
   supporting C++11. \dealii{} now uses language improvements such as using
@@
   instead of copying them, using \texttt{nullptr} instead of \texttt{NULL}, and
-  marking unimplemented constructors with \texttt{=delete;} instead of making
-  them \texttt{private}. These changes include some minor incompatibilities: all
-  \texttt{clone} functions (such as \texttt{FiniteElement::clone} and
-  \texttt{Mapping::clone}) now return \texttt{std::unique\_ptr}s instead of
+  marking members with \texttt{=delete;} and \texttt{=default;} where
+  appropriate. These changes include some minor incompatibilities: all
+  \texttt{clone()} functions (such as \texttt{FiniteElement::clone()} and
+  \texttt{Mapping::clone()}) now return \texttt{std::unique\_ptr}s instead of
   C-style raw pointers. Indeed, nearly every interface that returns a pointer

### 2018-04-11 9.0/paper.tex

@@
     Denis Davydov,
+    Rene Gassm\"{o}ller,
     Timo Heister,
@@
 
-\author[4]{Timo Heister}
-\affil[4]{Mathematical Sciences,
+\author[4]{Rene Gassm\"{o}ller}
+\affil[4]{Department of Earth and Planetary Sciences,
+  University of California Davis,
+  One Shields Avenue,
+  CA-95616 Davis, USA.
+  {\texttt{rgassmoeller@ucdavis.edu}}}
+
+\author[5]{Timo Heister}
+\affil[5]{Mathematical Sciences,
   O-110 Martin Hall,
@@
 %
-\author[5]{Luca Heltai}
-\affil[5]{SISSA,
+\author[6]{Luca Heltai}
+\affil[6]{SISSA,
   International School for Advanced Studies,
@@
 
-\author[6]{Martin Kronbichler}
-\affil[6]{Institute for Computational Mechanics,
+\author[7]{Martin Kronbichler}
+\affil[7]{Institute for Computational Mechanics,
   Technical University of Munich,
@@
 
-\author[7]{Matthias Maier}
-\affil[7]{School of Mathematics,
+\author[8]{Matthias Maier}
+\affil[8]{School of Mathematics,
   University of Minnesota,
@@
 
-\author[8]{Jean-Paul Pelteret}
-\affil[8]{Chair of Applied Mechanics,
+\author[9]{Jean-Paul Pelteret}
+\affil[9]{Chair of Applied Mechanics,
   University of Erlangen-Nuremberg,
@@
 
- \author[9]{Bruno Turcksin\footnote{
+ \author[10]{Bruno Turcksin\footnote{
    This manuscript has been authored by UT-Battelle, LLC under Contract No.
@@
    DOE Public Access Plan (http://energy.gov/downloads/doe-public-access-plan).}}
- \affil[9]{Computational Engineering and Energy Sciences Group,
+ \affil[10]{Computational Engineering and Energy Sciences Group,
    Computional Sciences and Engineering Division,
@@
 
-\author[10]{David Wells}
-\affil[10]{Department of Mathematical Sciences, Rensselaer Polytechnic Institute,
+\author[11]{David Wells}
+\affil[11]{Department of Mathematical Sciences, Rensselaer Polytechnic Institute,
 Troy, NY 12180, USA.
@@
   Niklas Fehn,
-  Rene Gassmoeller,
   Joscha Gedicke,

### 2018-04-12 9.0/paper.tex
Add Vishal to the list of authors

@@
     Daniel Arndt,
+    Vishal Boddu,
     Denis Davydov,
@@
 
-\author[3]{Denis Davydov}
-\affil[3]{Chair of Applied Mechanics, University of
+\author[3]{Vishal Boddu}
+\affil[3]{Chair of Applied Mechanics,
+  University of Erlangen-Nuremberg,
+  Egerlandstr.\ 5,
+  91058 Erlangen, Germany.
+  {\texttt{vishal.boddu@fau.de}}}
+
+\author[4]{Denis Davydov}
+\affil[4]{Chair of Applied Mechanics, University of
      Erlangen-Nuremberg, Egerlandstr.\ 5, 91058 Erlangen, Germany.
@@
 
-\author[4]{Rene Gassm\"{o}ller}
-\affil[4]{Department of Earth and Planetary Sciences,
+\author[5]{Rene Gassm\"{o}ller}
+\affil[5]{Department of Earth and Planetary Sciences,
   University of California Davis,
@@
 
-\author[5]{Timo Heister}
-\affil[5]{Mathematical Sciences,
+\author[6]{Timo Heister}
+\affil[6]{Mathematical Sciences,
   O-110 Martin Hall,
@@
 %
-\author[6]{Luca Heltai}
-\affil[6]{SISSA,
+\author[7]{Luca Heltai}
+\affil[7]{SISSA,
   International School for Advanced Studies,
@@
 
-\author[7]{Martin Kronbichler}
-\affil[7]{Institute for Computational Mechanics,
+\author[8]{Martin Kronbichler}
+\affil[8]{Institute for Computational Mechanics,
   Technical University of Munich,
@@
 
-\author[8]{Matthias Maier}
-\affil[8]{School of Mathematics,
+\author[9]{Matthias Maier}
+\affil[9]{School of Mathematics,
   University of Minnesota,
@@
 
-\author[9]{Jean-Paul Pelteret}
-\affil[9]{Chair of Applied Mechanics,
+\author[10]{Jean-Paul Pelteret}
+\affil[10]{Chair of Applied Mechanics,
   University of Erlangen-Nuremberg,
@@
 
- \author[10]{Bruno Turcksin\footnote{
+ \author[11]{Bruno Turcksin\footnote{
    This manuscript has been authored by UT-Battelle, LLC under Contract No.
@@
    DOE Public Access Plan (http://energy.gov/downloads/doe-public-access-plan).}}
- \affil[10]{Computational Engineering and Energy Sciences Group,
+ \affil[11]{Computational Engineering and Energy Sciences Group,
    Computional Sciences and Engineering Division,
@@
 
-\author[11]{David Wells}
-\affil[11]{Department of Mathematical Sciences, Rensselaer Polytechnic Institute,
-Troy, NY 12180, USA.
+\author[12]{David Wells}
+\affil[12]{Department of Mathematical Sciences, Rensselaer Polytechnic
+Institute, Troy, NY 12180, USA.
   {\texttt{wellsd2@rpi.edu}}}
@@
 %
+V.~Boddu was supported by the German Science Foundation (Deutsche 
+Forschungs-Gemeinschaft, DFG) under the research group project FOR 1509.
+
 D.~Davydov was supported by the German Science Foundation

### 2018-04-12 9.0/paper.tex

@@
     Vishal Boddu,
+    Benjamin Brands,
     Denis Davydov,
@@
 \affil[3]{Chair of Applied Mechanics,
-  University of Erlangen-Nuremberg,
+  Friedrich-Alexander-Universit\"{a}t Erlangen-N\"{u}rnberg,
   Egerlandstr.\ 5,
@@
 
-\author[4]{Denis Davydov}
-\affil[4]{Chair of Applied Mechanics, University of
-     Erlangen-Nuremberg, Egerlandstr.\ 5, 91058 Erlangen, Germany.
+\author[4]{Benjamin Brands}
+\affil[4]{Chair of Applied Mechanics,
+          Friedrich-Alexander-Universit\"{a}t Erlangen-N\"{u}rnberg,
+          Egerlandstr.\ 5, 91058 Erlangen, Germany.
+          {\texttt{benjamin.brands@fau.de}}}
+
+\author[5]{Denis Davydov}
+\affil[5]{Chair of Applied Mechanics, Friedrich-Alexander-Universit\"{a}t Erlangen-N\"{u}rnberg,
+            Egerlandstr.\ 5, 91058 Erlangen, Germany.
     {\texttt{denis.davydov@fau.de}}}
 
-\author[5]{Rene Gassm\"{o}ller}
-\affil[5]{Department of Earth and Planetary Sciences,
+\author[6]{Rene Gassm\"{o}ller}
+\affil[6]{Department of Earth and Planetary Sciences,
   University of California Davis,
@@
 
-\author[6]{Timo Heister}
-\affil[6]{Mathematical Sciences,
+\author[7]{Timo Heister}
+\affil[7]{Mathematical Sciences,
   O-110 Martin Hall,
@@
 %
-\author[7]{Luca Heltai}
-\affil[7]{SISSA,
+\author[8]{Luca Heltai}
+\affil[8]{SISSA,
   International School for Advanced Studies,
@@
 
-\author[8]{Martin Kronbichler}
-\affil[8]{Institute for Computational Mechanics,
+\author[9]{Martin Kronbichler}
+\affil[9]{Institute for Computational Mechanics,
   Technical University of Munich,
@@
 
-\author[9]{Matthias Maier}
-\affil[9]{School of Mathematics,
+\author[10]{Matthias Maier}
+\affil[10]{School of Mathematics,
   University of Minnesota,
@@
 
-\author[10]{Jean-Paul Pelteret}
-\affil[10]{Chair of Applied Mechanics,
-  University of Erlangen-Nuremberg,
+\author[11]{Jean-Paul Pelteret}
+\affil[11]{Chair of Applied Mechanics,
+  Friedrich-Alexander-Universit\"{a}t Erlangen-N\"{u}rnberg,
   Egerlandstr.\ 5,
@@
 
- \author[11]{Bruno Turcksin\footnote{
+ \author[12]{Bruno Turcksin\footnote{
    This manuscript has been authored by UT-Battelle, LLC under Contract No.
@@
    DOE Public Access Plan (http://energy.gov/downloads/doe-public-access-plan).}}
- \affil[11]{Computational Engineering and Energy Sciences Group,
+ \affil[12]{Computational Engineering and Energy Sciences Group,
    Computional Sciences and Engineering Division,
@@
 
-\author[12]{David Wells}
-\affil[12]{Department of Mathematical Sciences, Rensselaer Polytechnic
+\author[13]{David Wells}
+\affil[13]{Department of Mathematical Sciences, Rensselaer Polytechnic
 Institute, Troy, NY 12180, USA.
@@
 
+B.~Brands was partially supported by the Indo-German exchange programm ``Multiscale Modeling, Simulation and Optimization for Energy, Advanced Materials and Manufacturing'' (MMSO) funded by DAAD (Germany) and UGC (India).
+
 D.~Davydov was supported by the German Science Foundation

### 2018-04-12 9.0/paper.tex
Added Giovanni Alzetta

@@
   pdfauthor={
+    Giovanni Alzetta,
     Daniel Arndt,
@@
 
-\author[1]{Daniel Arndt}
-\affil[1]{Interdisciplinary Center for Scientific Computing,
+\author[1]{Giovanni Alzetta}
+\affil[1]{SISSA,
+  International School for Advanced Studies,
+  Via Bonomea 265,
+  34136, Trieste, Italy.
+{\texttt{galzetta@sissa.it}}}
+
+\author[2]{Daniel Arndt}
+\affil[2]{Interdisciplinary Center for Scientific Computing,
   Heidelberg University,
@@
 
-\author[2]{Wolfgang Bangerth}
-\affil[2]{Department of Mathematics, Colorado State University, Fort
+\author[3]{Wolfgang Bangerth}
+\affil[3]{Department of Mathematics, Colorado State University, Fort
   Collins, CO 80523-1874, USA.
@@
 
-\author[3]{Vishal Boddu}
-\affil[3]{Chair of Applied Mechanics,
+\author[4]{Vishal Boddu}
+\affil[4]{Chair of Applied Mechanics,
   Friedrich-Alexander-Universit\"{a}t Erlangen-N\"{u}rnberg,
@@
 
-\author[4]{Benjamin Brands}
-\affil[4]{Chair of Applied Mechanics,
+\author[5]{Benjamin Brands}
+\affil[5]{Chair of Applied Mechanics,
           Friedrich-Alexander-Universit\"{a}t Erlangen-N\"{u}rnberg,
@@
 
-\author[5]{Denis Davydov}
-\affil[5]{Chair of Applied Mechanics, Friedrich-Alexander-Universit\"{a}t Erlangen-N\"{u}rnberg,
+\author[6]{Denis Davydov}
+\affil[6]{Chair of Applied Mechanics, Friedrich-Alexander-Universit\"{a}t Erlangen-N\"{u}rnberg,
             Egerlandstr.\ 5, 91058 Erlangen, Germany.
@@
 
-\author[6]{Rene Gassm\"{o}ller}
-\affil[6]{Department of Earth and Planetary Sciences,
+\author[7]{Rene Gassm\"{o}ller}
+\affil[7]{Department of Earth and Planetary Sciences,
   University of California Davis,
@@
 
-\author[7]{Timo Heister}
-\affil[7]{Mathematical Sciences,
+\author[8]{Timo Heister}
+\affil[8]{Mathematical Sciences,
   O-110 Martin Hall,
@@
 %
-\author[8]{Luca Heltai}
-\affil[8]{SISSA,
+\author[9]{Luca Heltai}
+\affil[9]{SISSA,
   International School for Advanced Studies,
@@
 
-\author[9]{Martin Kronbichler}
-\affil[9]{Institute for Computational Mechanics,
+\author[10]{Martin Kronbichler}
+\affil[10]{Institute for Computational Mechanics,
   Technical University of Munich,
@@
 
-\author[10]{Matthias Maier}
-\affil[10]{School of Mathematics,
+\author[11]{Matthias Maier}
+\affil[11]{School of Mathematics,
   University of Minnesota,
@@
 
-\author[11]{Jean-Paul Pelteret}
-\affil[11]{Chair of Applied Mechanics,
+\author[12]{Jean-Paul Pelteret}
+\affil[12]{Chair of Applied Mechanics,
   Friedrich-Alexander-Universit\"{a}t Erlangen-N\"{u}rnberg,
@@
 
- \author[12]{Bruno Turcksin\footnote{
+ \author[13]{Bruno Turcksin\footnote{
    This manuscript has been authored by UT-Battelle, LLC under Contract No.
@@
    DOE Public Access Plan (http://energy.gov/downloads/doe-public-access-plan).}}
- \affil[12]{Computational Engineering and Energy Sciences Group,
+ \affil[13]{Computational Engineering and Energy Sciences Group,
    Computional Sciences and Engineering Division,
@@
 
-\author[13]{David Wells}
-\affil[13]{Department of Mathematical Sciences, Rensselaer Polytechnic
+\author[14]{David Wells}
+\affil[14]{Department of Mathematical Sciences, Rensselaer Polytechnic
 Institute, Troy, NY 12180, USA.

### 2018-04-12 9.0/paper.tex
Add more names to pdfauthor

@@
     Daniel Arndt,
+    Wolfgang Bangerth,
     Vishal Boddu,
@@
     Matthias Maier,
+    Jean-Paul Pelteret,
+    Bruno Turcksin,
     David Wells

### 2018-04-16 9.0/paper.tex
Add funding acknowledgement
Update entries for clang-tidy and coverity

@@
 
-\item \dealii{} has made extensive use of both the clang-tidy and coverity
-      static analysis tools for detecting bugs and other issues in the code.
-      %% TODO is there somewhere we can cite for either of these tools? It would
-      %% be nice to recognize the department of homeland security for offering
-      %% coverity.
-      %% TODO at the time of writing this coverity is down; get an issue count
-      %% when it is back online.
+\item \dealii{} has made extensive use of both the Clang-Tidy \cite{clang-tidy}
+  and Coverity Scan \cite{coverity} static analysis tools for detecting bugs
+  and other issues in the code. In total around, 260 issues were detected and
+  fixed using the latter tool.
+      %% > Is there somewhere we can cite for either of these tools? It would
+      %% > be nice to recognize the department of homeland security for offering
+      %% The contract with homeland security was completed in 2009.
 
@@
 
-D.~Arndt was partially supported by the German Research Foundation (DFG) under the
+D.~Arndt was supported by the German Research Foundation (DFG) under the
 project ``High-order discontinuous Galerkin for the exa-scale'' (ExaDG) within the
@@
 
-% J-P.~Pelteret was supported by the European Research Council (ERC) through the Advanced Grant 289049 MOCOPOLY.
-%
+J-P.~Pelteret was supported by the European Research Council (ERC) through the Advanced Grant 289049 MOCOPOLY.
+
 B.~Turcksin: This material is based upon work supported by the U.S. Department of

### 2018-04-17 9.0/paper.tex
Added Katharina Kormann.

@@
     Luca Heltai,
-    Martin Kronbichler
+    Katharina Kormann,
+    Martin Kronbichler,
     Matthias Maier,
@@
 
-\author[10]{Martin Kronbichler}
-\affil[10]{Institute for Computational Mechanics,
+\author[10]{Katharina Kormann}
+\affil[10]{Max Planck Institute for Plasma Physics,
+Boltzmannstr.~2, 85748 Garching, Germany.
+{\texttt{katharina.kormann@ipp.mpg.de}}}
+
+\author[11]{Martin Kronbichler}
+\affil[11]{Institute for Computational Mechanics,
   Technical University of Munich,
@@
 
-\author[11]{Matthias Maier}
-\affil[11]{School of Mathematics,
+\author[12]{Matthias Maier}
+\affil[12]{School of Mathematics,
   University of Minnesota,
@@
 
-\author[12]{Jean-Paul Pelteret}
-\affil[12]{Chair of Applied Mechanics,
+\author[13]{Jean-Paul Pelteret}
+\affil[13]{Chair of Applied Mechanics,
   Friedrich-Alexander-Universit\"{a}t Erlangen-N\"{u}rnberg,
@@
 
- \author[13]{Bruno Turcksin\footnote{
+ \author[14]{Bruno Turcksin\footnote{
    This manuscript has been authored by UT-Battelle, LLC under Contract No.
@@
    DOE Public Access Plan (http://energy.gov/downloads/doe-public-access-plan).}}
- \affil[13]{Computational Engineering and Energy Sciences Group,
+ \affil[14]{Computational Engineering and Energy Sciences Group,
    Computional Sciences and Engineering Division,
@@
 
-\author[14]{David Wells}
-\affil[14]{Department of Mathematical Sciences, Rensselaer Polytechnic
+\author[15]{David Wells}
+\affil[15]{Department of Mathematical Sciences, Rensselaer Polytechnic
 Institute, Troy, NY 12180, USA.

### 2018-04-19 9.0/paper.tex
Update Acknowledgments
 - break overlong lines

 - The DFG calls itself "German Research Foundation" on their English
   website. Update naming

 - Try to unify naming

 - add my own funding

@@
 % No.~EAR-0949446 and EAR-1550901 and The University of California -- Davis.
-%
-V.~Boddu was supported by the German Science Foundation (Deutsche 
-Forschungs-Gemeinschaft, DFG) under the research group project FOR 1509.
 
-B.~Brands was partially supported by the Indo-German exchange programm ``Multiscale Modeling, Simulation and Optimization for Energy, Advanced Materials and Manufacturing'' (MMSO) funded by DAAD (Germany) and UGC (India).
+V.~Boddu was supported by the German Research Foundation (DFG) under the
+research group project FOR 1509.
 
-D.~Davydov was supported by the German Science Foundation
-(Deutsche Forschungs-Gemeinschaft, DFG), grant DA 1664/2-1.
+B.~Brands was partially supported by the Indo-German exchange programm
+``Multiscale Modeling, Simulation and Optimization for Energy, Advanced
+Materials and Manufacturing'' (MMSO) funded by DAAD (Germany) and UGC
+(India).
+
+D.~Davydov was supported by the German Research Foundation (DFG), grant DA
+1664/2-1.
 
 T.~Heister was partially supported by the Computational Infrastructure in
-Geodynamics initiative (CIG), through the National Science Foundation
-under Award No. EAR-0949446 and EAR-1550901 and The University of California -- Davis, and National Science Foundation grant DMS-1522191.
+Geodynamics initiative (CIG), through the National Science Foundation (NSF)
+under Award No. EAR-0949446 and EAR-1550901 and The University of
+California -- Davis, and NSF Grant DMS-1522191.
 
@@
 
-J-P.~Pelteret was supported by the European Research Council (ERC) through the Advanced Grant 289049 MOCOPOLY.
+M.~Maier was partially supported by ARO MURI Award No. W911NF-14-0247.
 
-B.~Turcksin: This material is based upon work supported by the U.S. Department of
-Energy, Office of Science, under contract number DE-AC05-00OR22725.
+J-P.~Pelteret was supported by the European Research Council (ERC) through
+the Advanced Grant 289049 MOCOPOLY.
+
+B.~Turcksin: This material is based upon work supported by the U.S.
+Department of Energy, Office of Science, under contract number
+DE-AC05-00OR22725.
 
@@
 
-The Interdisciplinary Center for Scientific Computing (IWR) at Heidelberg University has provided
-hosting services for the \dealii{} web page.
+The Interdisciplinary Center for Scientific Computing (IWR) at Heidelberg
+University has provided hosting services for the \dealii{} web page.
 

### 2018-04-19 9.0/paper.tex
minor updates, wrote a paragraph, update contributors

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Overview}
@@
   Support for GPU
+  \marginpar{paragraph for CUDA missing}
 
 \item
-  In this version we add template wrappers for ScaLAPACK -- parallel dense linear algebra
-  library with block-cyclic matrix distribution over 2D process grid.
-  The functionality and interface is similar to the LAPACK wrappers,
-  matrix-matrix multiplication, Cholesky and LU factorizations,
-  eigensolvers, SVD, pseudoinverse, save/load in HDF5 format and
-  other functionality is implemented.
-  A user can combine the new linear algebra with other MPI-parallel
-  algorithms in \dealii{} without the need to worry about MPI communicator
-  being divisible exactly into 2D process grid. The processes which are not part of
-  the grid are labeled inactive and ScaLAPACK routines will not be called on them.
+  In this version we add template wrappers for ScaLAPACK -- parallel dense
+  linear algebra library with block-cyclic matrix distribution over 2D
+  process grid. The functionality and interface is similar to the LAPACK
+  wrappers, matrix-matrix multiplication, Cholesky and LU factorizations,
+  eigensolvers, SVD, pseudoinverse, save/load in HDF5 format and other
+  functionality is implemented. A user can combine the new linear algebra
+  with other MPI-parallel algorithms in \dealii{} without the need to worry
+  about MPI communicator being divisible exactly into 2D process grid. The
+  processes which are not part of the grid are labeled inactive and
+  ScaLAPACK routines will not be called on them.
   %
   Additionally, we improved support of LAPACK. We added methods to perform
-  rank-1 update/downdate, calculate Cholesky factorization, trace and determinant,
-  as well as estimate the reciprocal condition number.
-  We also now support configuration with 64-bit BLAS.
+  rank-1 update/downdate, calculate Cholesky factorization, trace and
+  determinant, as well as estimate the reciprocal condition number. We also
+  now support configuration with 64-bit BLAS.
 
@@
 \item
-  Initial support for Gmsh
+  Initial support for Gmsh.
+  \emph{[--- Do we actually provide any functionality with the GMSH
+  executable yet? ---]}
+  \marginpar{TODO}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Significant changes to the library}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Feature 1}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Feature 2}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{New and updated tutorial programs}
+
 \marginpar{OLD}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{New non-standard quadrature rules}
 
+A number of non-standard, special-purpose quadrature rules have been
+implemented. Among those are quadrature formulas that support
 \begin{itemize}
-\item QSimplex
-\item QDuffy
-\item QTrianglePolar
-\item QSplit
+  \item
+    truncating standard formulas to simplical domains (\texttt{Qsimplex}),
+  \item
+    singular transformations of the unit cell to the unit simplex
+    (\texttt{QDuffy}),
+  \item
+    composition of simplical quadrature rules to a combined rule on the
+    unit cell (\texttt{QSplit}),
+  \item
+    transformation of the unit square to polar coordinates
+    (\texttt{QTrianglePolar}).
 \end{itemize}
+These new quadrature rule greatly help to integrate singular domains. They
+are mainly used in Boundary Element Methods.
 
-Mostly used for BEM problems.
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Incompatible changes}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{How to cite \dealii{}}\label{sec:cite}
@@
 
-There are various ways to reference \dealii{}. To acknowledge the use of the
-current version of the library, \textbf{please reference the present document}. For up
-to date information and bibtex snippets for this document see:
+There are various ways to reference \dealii{}. To acknowledge the use of
+the current version of the library, \textbf{please reference the present
+document}. For up to date information and bibtex snippets for this document
+see:
 \begin{center}
@@
 
-% \begin{minipage}{0.9\textwidth}%no page break in here please
-% \begin{verbatim}
-% @article{dealII82,
-%   title = {The {\tt deal.{I}{I}} Library, Version 8.2},
-%   author = {W. Bangerth and T. Heister and L. Heltai
-%    and G. Kanschat and M. Kronbichler and M. Maier
-%    and B. Turcksin and T. D. Young},
-%   journal = {preprint},
-%   year = {2015},
-% }
-% \end{verbatim} %  journal = {arXiv preprint \url{http://arxiv.org/abs/TODO}},
-% \end{minipage}
-
 The original \texttt{\dealii{}} paper containing an overview of its
-architecture is \cite{BangerthHartmannKanschat2007}. If you rely on specific
-features of the library, please consider citing any of the following:
+architecture is \cite{BangerthHartmannKanschat2007}. If you rely on
+specific features of the library, please consider citing any of the
+following:
 \begin{itemize}
@@
 \dealii{} can interface with many other libraries:
-\begin{multicols}{2}
+\begin{multicols}{3}
 \begin{itemize}
@@
 \item SUNDIALS \cite{sundials}
-\item Threading Building Blocks \cite{Rei07}
+\item TBB \cite{Rei07}
 \item Trilinos \cite{trilinos,trilinos-web-page}
@@
 
-Older releases of \dealii{} can be cited as \cite{dealII80,dealII81,dealII82,dealII83,dealII84,dealII85}.
+Older releases of \dealii{} can be cited as
+\cite{dealII80,dealII81,dealII82,dealII83,dealII84,dealII85}.
 
@@
 %
-\marginpar{update}
+% Updated 2018-04-19
+Julian Andrej,
 Rajat Arora,
-  Mauro Bardelloni,
+Lucas Campos,
+Praveen Chandrashekar,
+Jie Cheng,
+Emma Cinatl,
 Conrad Clevenger,
-  Sam Cox,
-  Toby D. Young,
-  Juliane Dannberg,
-  Nicola Demo,
-  Patrick Esser,
+Ester Comellas,
+Sambit Das,
+Giovanni Di Ilio,
+Nivesh Dommaraju,
+Marc Fehling,
 Niklas Fehn,
-  Joscha Gedicke,
+Menno Fraters,
+Anian Fuchs,
+Daniel Garcia-Sanchez,
 Nicola Giuliani,
-  Sebastian Gonzalez-Pintor,
-  Ryan Grove,
-  Michael Harmon,
+Anne Glerum,
+Christoph Goering,
+Alexander Grayver,
+Samuel Imfeld,
 Daniel Jodlbauer,
 Guido Kanschat,
-  Justin Kauffman,
-  Eldar Khattatov ,
-  Uwe Koecher,
-  Alex Kokomov,
-  Paul Kuberry,
-  Dustin Kumor,
+Vishal Kenchan,
+Andreas Kergassner,
+Eldar Khattatov,
+Ingo Kligge,
+Uwe Köcher,
+Joachim Kopp,
+Ross Kynch,
 Konstantin Ladutenko,
+Tulio Ligneul,
 Karl Ljungkvist,
-  Andrew McBride,
-  Mathias Mentler,
-  Andrea Mola,
-  Dragan Nikolic,
-  Vaibhav Palkar,
-  Spencer Patty,
-  Jonathan Perry-Houts,
-  Giuseppe Pitton,
+Santiago Ospina,
+Alexey Ozeritsky,
+Dirk Peschka,
+Simon Puchert,
+E. G. Puckett,
+Lei Qiao,
 Ce Qin,
 Jonathan Robey,
-  Mayank Sabharwal,
-  Ali Samii,
 Alberto Sartori,
 Daniel Shapero,
-  Martin Steigemann,
-  Jihuan Tian,
-  Jaeryun Yim,
-  Liang Zhao
+Ben Shields,
+Simon Sticko,
+Oliver Sutton,
+Zhuoran Wang,
+Xiaoyu Wei,
+Michał Wichrowski,
+Julius Witte,
+Feimi Yu,
+Weixiong Zheng.
 

### 2018-04-21 9.0/paper.tex
List sources of support for myself and for Rene.

@@
 
-% W.~Bangerth was partially
-% supported by the National Science Foundation under award OCI-1148116
-% as part of the Software Infrastructure for Sustained Innovation (SI2)
-% program; and by the Computational Infrastructure in Geodynamics initiative
-% (CIG), through the National Science Foundation under Awards
-% No.~EAR-0949446 and EAR-1550901 and The University of California -- Davis.
+W.~Bangerth and R.~Gassm\"{o}ller were partially
+supported by the National Science Foundation under award OCI-1148116
+as part of the Software Infrastructure for Sustained Innovation (SI2)
+program; and by the Computational Infrastructure in Geodynamics initiative
+(CIG), through the National Science Foundation under Awards
+No.~EAR-0949446 and EAR-1550901 and The University of California -- Davis.
 

### 2018-04-21 9.0/paper.tex
Rewrite the C++11 section.

@@
 
-\dealii{} version 9.0.0 was released April X, 2018. This paper provides an
+\dealii{} version 9.0.0 was released April X, 2018.
+\marginpar{Update the date}
+This paper provides an
 overview of the new features of this major release and serves as a citable
@@
 
-\item
-  \dealii{} first offered support for a subset of C++11 features in
-  version 6.2, released in 2009. The current release is the first to \emph{require} a compiler
-  supporting C++11. \dealii{} now uses language improvements such as using
-  \texttt{emplace\_back()} instead of \texttt{push\_back()}, moving objects
-  instead of copying them, using \texttt{nullptr} instead of \texttt{NULL}, and
-  marking members with \texttt{=delete;} and \texttt{=default;} where
-  appropriate. These changes include some minor incompatibilities: all
-  \texttt{clone()} functions (such as \texttt{FiniteElement::clone()} and
-  \texttt{Mapping::clone()}) now return \texttt{std::unique\_ptr}s instead of
-  C-style raw pointers. Indeed, nearly every interface that returns a pointer
-  now returns either a \texttt{std::shared\_ptr} or \texttt{std::unique\_ptr},
-  which clarifies object ownership responsibilities and avoids memory leaks.
+\item C++11 is now both required and used.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Feature 2}
+\subsection{Use of C++11}
+\dealii{} first offered support for a subset of C++11 features in
+version 6.2, released in 2009. The current release is the first to
+\emph{require} a C++11 compiler.
+
+Many parts of the code base have been rewritten to both support and
+use the new features of C++11. In particular, \dealii{} now makes
+extensive use of move semantics as well as range-based \texttt{for}
+loops with \texttt{auto} type deduction of iterator variables. We have
+also largely replaced \texttt{push\_back()} by
+\texttt{emplace\_back()} when adding elements to collections more
+efficiently.
+
+Finally, we have changed the entire code base to avoid using raw
+pointers and instead use \texttt{std::unique\_ptr} and
+\texttt{std::shared\_ptr} where possible to make memory management
+more reliable. These changes include some minor incompatibilities: all
+\texttt{clone()} functions (such as \texttt{FiniteElement::clone()} and
+\texttt{Mapping::clone()}) now return \texttt{std::unique\_ptr}s instead of
+C-style raw pointers. Indeed, nearly all interfaces throughout the library that return a pointer
+now return either a \texttt{std::shared\_ptr} or a \texttt{std::unique\_ptr},
+thereby clarifying object ownership responsibilities and avoiding memory leaks.
 

### 2018-04-21 9.0/paper.tex
Reshuffle some of the sections.

@@
 
+\item
+  A dedicated particles module has been created to provide support for the
+  storage of particles and their properties. This module provides a base class
+  \texttt{Particles} which represents a particle with position, an ID number and
+  a variable number of properties. It also provides a \texttt{ParticleHandler}
+  class that manages the storage and handling of particles. This class
+  distributes the particles in distributed domains (these are the only domains
+  supported by this class so far)
+
+\item
+  A dedicated differentiation module has been created to provide support for
+  automatic and symbolic differentiation. Automatic differentiation is a set of
+  technique to evaluate the derivatives of a function defined by a compute
+  program. Currently the Adol-C and Sacado libraries are supported through a
+  unified interface. In practice, this support means that Adol-c and Sacado
+  data types can be used in the \texttt{FEValues} and \texttt{FEValuesViews}
+  to represent degree-of-freedom values.
+
+\item Interfaces to more external libraries and programs.
+
 \item C++11 is now both required and used.
+
+\item
+  Support for GPU computations.
 
@@
   and Coverity Scan \cite{coverity} static analysis tools for detecting bugs
-  and other issues in the code. In total around, 260 issues were detected and
+  and other issues in the code. For example, around 260 issues were detected and
   fixed using the latter tool.
@@
 
-\item
-  Support for GPU
-  \marginpar{paragraph for CUDA missing}
+\item New non-standard quadrature rules:
+A number of non-standard, special-purpose quadrature rules have been
+implemented. Among these are ones for
+(i) truncating standard formulas to simplical domains (\texttt{Qsimplex});
+(ii) singular transformations of the unit cell to the unit simplex
+    (\texttt{QDuffy});
+(iii) composition of simplical quadrature rules to a combined rule on the
+    unit cell (\texttt{QSplit});
+and (iv) transformation of the unit square to polar coordinates
+    (\texttt{QTrianglePolar}).
+These quadrature rules greatly help when integrating singular
+functions or on singular domains. They
+are mainly used in Boundary Element Methods.
 
 \item
-  In this version we add template wrappers for ScaLAPACK -- parallel dense
-  linear algebra library with block-cyclic matrix distribution over 2D
-  process grid. The functionality and interface is similar to the LAPACK
-  wrappers, matrix-matrix multiplication, Cholesky and LU factorizations,
-  eigensolvers, SVD, pseudoinverse, save/load in HDF5 format and other
-  functionality is implemented. A user can combine the new linear algebra
-  with other MPI-parallel algorithms in \dealii{} without the need to worry
-  about MPI communicator being divisible exactly into 2D process grid. The
-  processes which are not part of the grid are labeled inactive and
-  ScaLAPACK routines will not be called on them.
-  %
-  Additionally, we improved support of LAPACK. We added methods to perform
-  rank-1 update/downdate, calculate Cholesky factorization, trace and
-  determinant, as well as estimate the reciprocal condition number. We also
-  now support configuration with 64-bit BLAS.
-
-\item
-  A dedicated sundials module has been created to provide support for the
-  SUNDIALS library. SUNDIALS is a SUite of Nonlinear and
-  DIfferential/ALgebraic equations Solvers. ARKode, IDA, and KINSOL are
-  supported. ARKode is a solver library that provides adaptive-step time
-  integration. IDA is a package for the solution of differential-algebraic
-  equations systems in the form $F(t,y,y')=0$. KINSOL is solver for nonlinear
-  algebraic systems.
-
-\item
-  A dedicated optimization/rol module has been created to provide support for
-  the Rapid Optimization Library (ROL). ROL is a package for large-scale
-  optimization.
-
-\item
-  A dedicated particles module has been created to provide support for the
-  storage of particles and their properties. This module provides a base class
-  \texttt{Particles} which represents a particle with position, an ID number and
-  a variable number of properties. It also provides a \texttt{ParticleHandler}
-  class that manages the storage and handling of particles. This class
-  distributes the particles in distributed domains (these are the only domains
-  supported by this class so far)
-
-\item
-  A dedicated differentiation module has been created to provide support for
-  automatic and symbolic differentiation. Automatic differentiation is a set of
-  technique to evaluate the derivatives of a function defined by a compute
-  program. Currently the Adol-C and Sacado libraries are supported through a
-  unified interface. In practice, this support means that Adol-c and Sacado
-  data types can be used in the \texttt{FEValues} and \texttt{FEValuesViews}
-  to represent degree-of-freedom values.
-
-\item
-  Support for complex-valued vectors is now at the same level as real-valued
+  Support for complex-valued vectors at the same level as real-valued
   vectors.
-
-\item
-  Support for the nanoflann library. nanoflann is a library for building
-  KD-trees of datasets and for the fast querying of closest neighbors in
-  KD-trees. nanoflann is used to speed up the search of the p nearest neighbors
-  of a given point, or searching the points that fall within a radius of a
-  target point.
-
-\item
-  Support for the Open Asset Import Library (Assimp). Assimp can be used to read
-  about 40 different 3D graphics formats. A subset of these formats can be used
-  to generate two-dimensional meshes possibly embedded in a three-dimensional
-  space.
-
-\item
-  Initial support for Gmsh.
-  \emph{[--- Do we actually provide any functionality with the GMSH
-  executable yet? ---]}
-  \marginpar{TODO}
-
-\item
-  Initial support for generalized support points: Internal interpolation
-  rewritten (support for $H(\text{curl})$ and $H(\text{div})$ conforming
-  elements).
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Feature 1}
+\subsection{Manifolds}
+
+...
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Support for particle-in-cell methods}
+
+...
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Dedicated support for automatic and symbolic differentiation}
+
+...
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{New interfaces to external libraries and programs}
+
+\dealii{} has always tried to leverage high-quality implementations of
+algorithms available through other open source software, rather than
+re-implementing their functionality. (A list of interfaces to other
+packages is given in Section~\ref{sec:cite}.) As part of the current
+release, we have written several new interfaces as discussed in the following.
+
+\marginpar{Add references to all of these libraries.}
+
+\paragraph*{ScaLAPACK, a parallel dense linear algebra library.}
+  ScaLAPACK provides block-cyclic matrix distribution over 2D process
+  grid. The functionality and interface of our wrappers is similar to
+  the LAPACK wrappers for serial dense linear algebra, namely
+  matrix-matrix multiplication, Cholesky and LU factorizations,
+  eigensolvers, SVD, pseudoinverses, and save/load operations using
+  the HDF5 format. All of this functionality is available even in
+  cases where the number of MPI processes does not exactly divide the
+  number of tiles stored in the block-cyclic data distribution.
+
+  As part of this effort, we have also improved LAPACK support: there
+  are now methods to perform rank-1 updates/downdates, Cholesky
+  factorizations, to compute the trace and determinant, as well as
+  estimate the reciprocal condition number. We also now support
+  configuration with 64-bit BLAS.
+
+\paragraph*{SUNDIALS, a SUite of Nonlinear and DIfferential/ALgebraic
+  Equation Solvers.}
+  Solving nonlinear algebraic and differential equations is both a
+  common task and one that often requires sophisticated globalization
+  algorithms for efficiency and reliability. SUNDIALS provides these
+  in a widely used format, both sequentially and in parallel.
+
+  \dealii{} now has interfaces to SUNDIALS's ARKode, IDA, and KINSOL sub-packages.
+  ARKode is a solver library that provides adaptive-step time
+  integration. IDA is a package for the solution of differential-algebraic
+  equations systems in the form $F(t,y,y')=0$. KINSOL is solver for nonlinear
+  algebraic systems.
+
+\paragraph*{ROL, a Rapid Optimization Library.}
+ROL is a package for large-scale optimization.
+\marginpar{Can we add a sentence or two about the ROL interface here?}
+
+\paragraph*{nanoflann, a library for building and querying
+  $k$-d trees of datasets.} Operations such as finding the vertex
+or cell closest
+to a given evaluation point occur frequently in many applications that
+use unstructured meshes. While the naive algorithm is linear in the
+number of vertices or cells, many such operations can be made
+significantly faster by building a $k$-d tree data structure that
+recursively subdivides a $k$ dimensional space. The nanoflann library
+provides such a data structure and allows querying it, either for
+closest points (e.g., when finding the closest vertex) or for
+searching the points that fall within a radius of a target point. This
+functionality is now available via \dealii{} interfaces.
+
+\paragraph*{Assimp, the Open Asset Import Library.}
+  Assimp can be used to read about 40 different 3D graphics formats. A
+  subset of these formats can be now be read from within \dealii{} to
+  generate two-dimensional meshes, possibly embedded in a
+  three-dimensional space.
+
+\paragraph*{Gmsh.}
+  \marginpar{Do we actually provide any functionality with the GMSH
+  executable yet? If so, complete this part}
 
@@
 thereby clarifying object ownership responsibilities and avoiding memory leaks.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Support for GPU computations}
+
+\marginpar{Bruno?}
+...
 
@@
   \end{itemize}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{New non-standard quadrature rules}
-
-A number of non-standard, special-purpose quadrature rules have been
-implemented. Among those are quadrature formulas that support
-\begin{itemize}
-  \item
-    truncating standard formulas to simplical domains (\texttt{Qsimplex}),
-  \item
-    singular transformations of the unit cell to the unit simplex
-    (\texttt{QDuffy}),
-  \item
-    composition of simplical quadrature rules to a combined rule on the
-    unit cell (\texttt{QSplit}),
-  \item
-    transformation of the unit square to polar coordinates
-    (\texttt{QTrianglePolar}).
-\end{itemize}
-These new quadrature rule greatly help to integrate singular domains. They
-are mainly used in Boundary Element Methods.
 

### 2018-04-23 9.0/paper.tex
Update the particles section.

@@
 \begin{itemize}
-\item
-  Manifold smoothing: The manifold smoothing algorithms applied in the
-  Triangulation class and MappingQGeneric have been changed from the old
-  Laplace-style smoothing to a transfinite interpolation that linearly
-  blends between the descriptions on the faces around a cell. The old
-  transformation introduced boundary layers inside cells that prevented
-  convergence rates from exceeding \(3.5\) in the global \(L^2\) errors on typical
-  settings. This change also considerably improves mesh quality on settings
-  where curved descriptions are only applied to the boundary rather than
-  the whole volume.
-
-\item
-  A dedicated particles module has been created to provide support for the
-  storage of particles and their properties. This module provides a base class
-  \texttt{Particles} which represents a particle with position, an ID number and
-  a variable number of properties. It also provides a \texttt{ParticleHandler}
-  class that manages the storage and handling of particles. This class
-  distributes the particles in distributed domains (these are the only domains
-  supported by this class so far)
-
-\item
-  A dedicated differentiation module has been created to provide support for
-  automatic and symbolic differentiation. Automatic differentiation is a set of
-  technique to evaluate the derivatives of a function defined by a compute
-  program. Currently the Adol-C and Sacado libraries are supported through a
-  unified interface. In practice, this support means that Adol-c and Sacado
-  data types can be used in the \texttt{FEValues} and \texttt{FEValuesViews}
-  to represent degree-of-freedom values.
-
+\item xxxx something with manifolds xxx
+\item Support for particle-in-cell methods
+\item Dedicated support for automatic and symbolic differentiation.
 \item Interfaces to more external libraries and programs.
-
 \item C++11 is now both required and used.
-
-\item
-  Support for GPU computations.
+\item Support for GPU computations.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Manifolds}
+\subsection{xxx something with Manifolds xxx}
 
-...
+  Manifold smoothing: The manifold smoothing algorithms applied in the
+  Triangulation class and MappingQGeneric have been changed from the old
+  Laplace-style smoothing to a transfinite interpolation that linearly
+  blends between the descriptions on the faces around a cell. The old
+  transformation introduced boundary layers inside cells that prevented
+  convergence rates from exceeding \(3.5\) in the global \(L^2\) errors on typical
+  settings. This change also considerably improves mesh quality on settings
+  where curved descriptions are only applied to the boundary rather than
+  the whole volume.
 
@@
 
-...
+While \dealii{} is a package intended to solve problems with the
+finite element method -- i.e., using continuous or discontinuous
+\textit{fields} --, is often convenient in fluid dynamics problems to
+couple the continuum description of phenomena with particles. These
+particles, advected along with the numerical approximation of the flow
+field, are then either used to visualize properties of the flow, or to
+advect material properties such as the viscosity of inhomogenous
+mixtures of fluids. If each particle is associated with the cells
+of a mesh, these methods are often referred to as particle-in-cell (PIC).
+
+\dealii{} now has a dedicated particles module. The module provides a base class
+\texttt{Particle} that represents a particle with position, an ID number and
+a variable number of properties. They are jointly represented by a \texttt{ParticleHandler}
+class that manages the storage and handling of all particles. In
+parallel simulations, this class also
+distributes the particles among the subdomains of the parallel process
+and supports efficient data transfer during mesh refinement and
+checkpoint/restart phases.
+
+A much more detailed view of the underlying algorithms can be found in
+\cite{GLHPB18}. A longer report is at \cite{GHPB16}.
 
@@
 
-...
+A dedicated differentiation module has been created to provide support for
+automatic and symbolic differentiation. Automatic differentiation is a set of
+technique to evaluate the derivatives of a function defined by a compute
+program. Currently the Adol-C and Sacado libraries are supported through a
+unified interface. In practice, this support means that Adol-c and Sacado
+data types can be used in the \texttt{FEValues} and \texttt{FEValuesViews}
+to represent degree-of-freedom values.
 

### 2018-04-23 9.0/paper.tex
Update the section on tutorial/code gallery programs.

@@
 \item Support for GPU computations.
-
+\end{itemize}
+These will all be discussed in more detail in the
+following section. In addition, this release contains the following changes:
+\begin{itemize}
 \item \dealii{} has made extensive use of both the Clang-Tidy \cite{clang-tidy}
@@
 
-\item New python tutorial program tutorial-1; as well as
+\item A new python tutorial program tutorial-1; as well as
   updates to step-37. In addition, the separate code
   gallery of \dealii{} has gained a number of new entries.
-
-  \item More than 330 other features and bugfixes.
 \end{itemize}
-The more important ones of these changes will be detailed in the
-following section.  Information on how to cite \dealii{} is provided
-in Section \ref{sec:cite}.
+Beyond these changes, the changelog lists more than 330 other features and bugfixes.
 
@@
 
+\marginpar{All: please edit as appropriate}
+
+(i) separation between manifold and boundary ids, (ii) manifold
+smoothing via the transfinite mapping, (iii) maybe: grid generator functions already attach manifolds?
+
+Previous text:
+
 Manifold smoothing: The manifold smoothing algorithms applied in the
@@
 
+\marginpar{Rene: Please edit as appropriate}
 While \dealii{} is a package intended to solve problems with the
@@
 A much more detailed view of the underlying algorithms can be found in
-\cite{GLHPB18}. A longer report is at \cite{GHPB16}.
+\cite{GLHPB18}. A longer report is at \cite{GHPB16}. The
+implementation here originated in the \aspect{} code, see \cite{KHB12,HDGB17}.
 
@@
 
-A dedicated differentiation module has been created to provide support for
-automatic and symbolic differentiation. Automatic differentiation is a set of
-technique to evaluate the derivatives of a function defined by a compute
-program. Currently the Adol-C and Sacado libraries are supported through a
-unified interface. In practice, this support means that Adol-c and Sacado
-data types can be used in the \texttt{FEValues} and \texttt{FEValuesViews}
-to represent degree-of-freedom values.
+\marginpar{J-P: Please edit as appropriate}
+Automatic differentiation is often used to automatically derive
+residuals from the stored energy functional, and to derive Jacobian
+matrices from residual vectors for simulations that use complicated
+material models. Examples can be found widely for nonlinear solid
+mechanics as well as for nonlinear viscosity models in fluid flow.
+
+\dealii{} has had a tutorial program (step-33) since 2007 that
+demonstrates this technique based on the Trilinos Sacado package, but
+the functionality was not available pervasively throughout
+deal.II. This has changed with release 9.0 where support for
+differentiation has been collected in one module that provides both
+automatic and symbolic differentiation. Currently, this module relies
+on either the Adol-C and Sacado libraries, through a unified
+interface. In practice, this support means that Adol-C and Sacado data
+types can be used in the \texttt{FEValues}, \texttt{FEValuesViews},
+and related classes that are generally used to assemble linear systems
+and right hand sides.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{New and updated tutorial programs}
+\subsection{Tutorial and code gallery programs}
 
-\marginpar{OLD}
-In addition to the updated tutorial programs mentioned in the previous
-section, this release of \dealii{} includes three new tutorials TODO: old:
-\begin{itemize}
- \item {\bf step-55} explains how to solve the Stokes
- equations efficiently in parallel. It is a good introduction to solving
- systems of PDEs in parallel, discusses optimal block
- preconditioners, and demonstrates other aspects like error computation.
- Inverses of individual blocks of the linear system are approximated with an algebraic
- multigrid preconditioner.
+This release does not contain any new tutorial programs, though
+several have been updated extensively for the changes to the manifold
+handling as well as to adjust for current functionality
+compared to what that was available when the programs were first written.
 
- \item {\bf step-56} shows how to apply geometric multigrid
- preconditioners on a subset of a system of PDEs. The problem solved here is the
- Stokes equations, like in step-55.
-
- \item {\bf step-57} solves the stationary Navier-Stokes equations.
- The nonlinear system is solved using Newton's method on a sequence of adaptively refined
- grids. The preconditioner is again built on a block factorization of the saddle point
- system like in step-55 and step-56, but the non-symmetric terms stemming from the
- nonlinear convective part requires more sophisticated solvers. The benchmark problem,
- flow in the 2d lid-driven cavity, requires a continuation method for high Reynolds
- numbers.
- \end{itemize}
-
-In addition to tutorials, \dealii{} has a separate ``code gallery'' that
+\dealii{} has a separate ``code gallery'' that
 consists of programs shared by users as examples of what can be
@@
 the following topics:
-\marginpar{OLD}
 \begin{itemize}
-    \item Quasi-static quasi-incompressible visco-elastic material behavior;
-    \item Multiphase Navier-Stokes flow;
-    \item The evolution of global-scale topography on planetary bodies;
-    \item Goal-oriented elastoplasticity.
+\item The multipoint flux mixed finite element method (MFMFE) applied
+  to the Darcy problem of porous media flow;
+\item A linearized active skeletal muscle model with application to
+  the simulation concentric contraction of the human biceps brachii;
+\item A parallel implementation of the Local Discontinuous Galerkin
+  (LDG) method applied to the Poisson equation.
 \end{itemize}
+With these additions, the code gallery now contains 10 different applications.
 

### 2018-04-23 9.0/paper.tex
Work through the incompatibilities section.

@@
 \subsection{xxx something with Manifolds xxx}
+\label{sec:manifolds}
 
@@
   \subsection{Use of C++11}
+  \label{sec:cxx11}
 \dealii{} first offered support for a subset of C++11 features in
@@
 
-\marginpar{OLD}
-The 9.0 release includes around XYZ
-\href{https://www.dealii.org/developer/doxygen/deal.II/changes_between_8_5_and_9_0.html}{
-incompatible changes}; see \cite{changes90}. The majority of these changes
+The 9.0 release includes
+\href{https://www.dealii.org/developer/doxygen/deal.II/changes_between_8_5_and_9_0.html}
+     {around 75 incompatible changes}; see \cite{changes90}. The majority of these changes
 should not be visible to typical user codes; some remove previously
 deprecated classes and functions, and the majority change internal
-interfaces that are not usually used in external applications. However, three
+interfaces that are not usually used in external
+applications. However, some are, such as changes to the interplay between meshes and
+manifolds, as well as the requirement to use a
+C++11 compiler (see Sections~\ref{sec:manifolds} and
+~\ref{sec:cxx11}). In addition, the following
 incompatible changes are worth mentioning:
 \begin{itemize}
-  \item High-order Lagrange elements, both continuous \verb!FE_Q! and
-    discontinuous \verb!FE_DGQ! types, now use the nodal points of the
-    Gauss-Lobatto quadrature formula as support points by default, rather than the
-    previous equidistant ones. For cubic polynomials and higher, the point
-    distribution has thus changed and, consequently, the entries in
-    solution vectors will be different compared to previous
-    versions of \dealii{}. Note, however, that using the Gauss-Lobatto points as nodal
-    points results in a much more stable interpolation, including better
-    iteration counts in most iterative solvers.
-  \item
-    The library no longer instantiates template classes with \texttt{long
-    double}. These were rarely used, but took up a significant
-    fraction of compile and link time, as well as library
-    size. Application programs can, however, still instantiate all
-    template classes with \texttt{long
-    double} as long as they include the corresponding \texttt{.templates.h}
-    header files.
-  \item
-    The \texttt{ParameterGUI} has been moved to a separate repository.
+\item The \texttt{BlockDiagonalMatrix},
+  \texttt{InverseMatrixRichardson},
+  \texttt{IterativeInverse},
+  \texttt{ProductMatrix},
+  \texttt{ProductSparseMatrix},
+  \texttt{TransposeMatrix},
+  \texttt{ScaledMatrix},
+  \texttt{SchurMatrix},
+  \texttt{ShiftedMatrix}, and
+  \texttt{ShiftedMatrixGeneralized} classes have been removed. They are now
+  generalized through the \texttt{LinearOperator} concept. Several
+  other, similar classes have been deprecated.
+
+  \item The default partitioner for the
+    \texttt{parallel::shared::Triangulation} is now the Trilinos
+    package Zoltan. This functionality was previously provided by the
+    METIS partitioner, but the METIS package has not been actively maintained
+    for a long time, and moreover yields subdivisions that depend on
+    system details such as the random number generator and sorting
+    facilities provided by the operating system; consequently, the
+    partition is not consistent across platforms.
+
+  \item  The class \texttt{FE\_DGQHermite}  now uses a more stable,
+    ``Hermite-like'' polynomial basis. The change is highly
+    beneficial because it significantly improves the accuracy (in
+    terms of roundoff) for this basis and also reduces iteration
+    counts for some iterative solvers with simple preconditioners.
+
+  \item Many functions that previously returned a raw, C-style pointer
+    now return a \texttt{std::unique\_ptr} and
+    \texttt{std::shared\_ptr} where possible to make memory management
+    more reliable.
 \end{itemize}

### 2018-04-23 9.0/paper.tex
Fix the spelling of ADOL-C.

@@
 automatic and symbolic differentiation. Currently, this module relies
-on either the Adol-C and Sacado libraries, through a unified
-interface. In practice, this support means that Adol-C and Sacado data
+on either the ADOL-C and Sacado libraries, through a unified
+interface. In practice, this support means that ADOL-C and Sacado data
 types can be used in the \texttt{FEValues}, \texttt{FEValuesViews},
@@
 \begin{itemize}
-\item Adol-c \cite{adol-c}
+\item ADOL-C \cite{adol-c}
 \item ARPACK \cite{arpack}

### 2018-04-24 9.0/paper.tex
More or less finish the main part of the paper.
no need to cite ScaLAPACK, there is no single library, it's API like Blas/LAPACK
cite ROL
Collect multiple authors from same institution on title page.

@@
   34136, Trieste, Italy.
-{\texttt{galzetta@sissa.it}}}
+{\texttt{\{galzetta,luca.heltai\}@sissa.it}}}
 
@@
   91058 Erlangen, Germany.
-  {\texttt{vishal.boddu@fau.de}}}
+  {\texttt{\{vishal.boddu,benjamin.brands,denis.davydov,jean-paul.pelteret\}@fau.de}}}
 
-\author[5]{Benjamin Brands}
-\affil[5]{Chair of Applied Mechanics,
-          Friedrich-Alexander-Universit\"{a}t Erlangen-N\"{u}rnberg,
-          Egerlandstr.\ 5, 91058 Erlangen, Germany.
-          {\texttt{benjamin.brands@fau.de}}}
+\author[4]{Benjamin Brands}
 
-\author[6]{Denis Davydov}
-\affil[6]{Chair of Applied Mechanics, Friedrich-Alexander-Universit\"{a}t Erlangen-N\"{u}rnberg,
-            Egerlandstr.\ 5, 91058 Erlangen, Germany.
-    {\texttt{denis.davydov@fau.de}}}
+\author[4]{Denis Davydov}
 
-\author[7]{Rene Gassm\"{o}ller}
-\affil[7]{Department of Earth and Planetary Sciences,
+\author[5]{Rene Gassm\"{o}ller}
+\affil[5]{Department of Earth and Planetary Sciences,
   University of California Davis,
@@
 
-\author[8]{Timo Heister}
-\affil[8]{Mathematical Sciences,
+\author[6]{Timo Heister}
+\affil[6]{Mathematical Sciences,
   O-110 Martin Hall,
@@
 %
-\author[9]{Luca Heltai}
-\affil[9]{SISSA,
-  International School for Advanced Studies,
-  Via Bonomea 265,
-  34136, Trieste, Italy.
-{\texttt{luca.heltai@sissa.it}}}
+\author[1]{Luca Heltai}
 
-\author[10]{Katharina Kormann}
-\affil[10]{Max Planck Institute for Plasma Physics,
+\author[7]{Katharina Kormann}
+\affil[7]{Max Planck Institute for Plasma Physics,
 Boltzmannstr.~2, 85748 Garching, Germany.
@@
 
-\author[11]{Martin Kronbichler}
-\affil[11]{Institute for Computational Mechanics,
+\author[8]{Martin Kronbichler}
+\affil[8]{Institute for Computational Mechanics,
   Technical University of Munich,
@@
 
-\author[12]{Matthias Maier}
-\affil[12]{School of Mathematics,
+\author[9]{Matthias Maier}
+\affil[9]{School of Mathematics,
   University of Minnesota,
@@
 
-\author[13]{Jean-Paul Pelteret}
-\affil[13]{Chair of Applied Mechanics,
-  Friedrich-Alexander-Universit\"{a}t Erlangen-N\"{u}rnberg,
-  Egerlandstr.\ 5,
-  91058 Erlangen,
-  Germany.
-  {\texttt{jean-paul.pelteret@fau.de}}}
+\author[4]{Jean-Paul Pelteret}
 
- \author[14]{Bruno Turcksin\footnote{
+ \author[10]{Bruno Turcksin\footnote{
    This manuscript has been authored by UT-Battelle, LLC under Contract No.
@@
    DOE Public Access Plan (http://energy.gov/downloads/doe-public-access-plan).}}
- \affil[14]{Computational Engineering and Energy Sciences Group,
+ \affil[10]{Computational Engineering and Energy Sciences Group,
    Computional Sciences and Engineering Division,
@@
 
-\author[15]{David Wells}
-\affil[15]{Department of Mathematical Sciences, Rensselaer Polytechnic
+\author[11]{David Wells}
+\affil[11]{Department of Mathematical Sciences, Rensselaer Polytechnic
 Institute, Troy, NY 12180, USA.
@@
  \item For $hp$ adaptivity: \cite{BangerthKayserHerold2007};
-  \item For $PUM$ and enrichment of the FE space: \cite{Davydov2016};
+  \item For partition-of-unity (PUM) and enrichment methods of the
+    finite element space: \cite{Davydov2016};
  \item For matrix-free and fast assembly techniques:
@@
 
-\marginpar{update}
+\marginpar{All: update gmsh reference}
 \dealii{} can interface with many other libraries:
@@
 \item Assimp \cite{assimp}
-\item BLAS, LAPACK
+\item BLAS and LAPACK
+\item Gmsh \cite{...}
 \item GSL \cite{gsl2016}
@@
 \item PETSc \cite{petsc-user-ref,petsc-web-page}
+\item ROL \cite{ridzal2014rapid}
+\item ScaLAPACK
 \item SLEPc \cite{Hernandez:2005:SSF}
@@
 variety of funding sources:
-\marginpar{update}
+\marginpar{all: update your funding sources}
 
-D.~Arndt was supported by the German Research Foundation (DFG) under the
-project ``High-order discontinuous Galerkin for the exa-scale'' (ExaDG) within the
-priority program ``Software for Exascale Computing'' (SPPEXA).
+D.~Arndt, K.~Kormann and M.~Kronbichler were partially supported by the German
+Research Foundation (DFG) under the project ``High-order discontinuous
+Galerkin for the exa-scale'' (ExaDG) within the priority program ``Software
+for Exascale Computing'' (SPPEXA).
 
@@
 
-M.~Kronbichler was partially supported by the German Research Foundation (DFG)
-under the project ``High-order discontinuous Galerkin for the exa-scale''
-(ExaDG) within the priority program ``Software for Exascale Computing''
-(SPPEXA), grant agreement no.~KR4661/2-1 and the Bayerisches Kompetenznetzwerk
-f\"ur Technisch-Wissenschaftliches Hoch- und H\"ochstleistungsrechnen
-(KONWIHR).
-
 M.~Maier was partially supported by ARO MURI Award No. W911NF-14-0247.

### 2018-04-24 9.0/paper.tex
Small changes here and there.

@@
 \usepackage{graphicx}
+\usepackage{xspace}
 
@@
 \newcommand{\specialword}[1]{\texttt{#1}}
-\newcommand{\dealii}{{\specialword{deal.II}}}
-\newcommand{\pfrst}{{\specialword{p4est}}}
-\newcommand{\trilinos}{{\specialword{Trilinos}}}
-\newcommand{\aspect}{\specialword{Aspect}}
-\newcommand{\petsc}{\specialword{PETSc}}
-\newcommand{\cmake}{{\specialword{CMake}}}
-\newcommand{\autoconf}{{\specialword{autoconf}}}
+\newcommand{\dealii}{{\specialword{deal.II}}\xspace}
+\newcommand{\pfrst}{{\specialword{p4est}}\xspace}
+\newcommand{\trilinos}{{\specialword{Trilinos}}\xspace}
+\newcommand{\aspect}{\specialword{Aspect}\xspace}
+\newcommand{\petsc}{\specialword{PETSc}\xspace}
+\newcommand{\cmake}{{\specialword{CMake}}\xspace}
+\newcommand{\autoconf}{{\specialword{autoconf}}\xspace}
 
@@
 
-\title{The \dealii{} Library, Version 9.0}
+\title{The \dealii Library, Version 9.0}
 
@@
   This paper provides an overview of the new features of the finite element
-  library \dealii{} version 9.0.
+  library \dealii version 9.0.
 \end{abstract}
@@
 
-\dealii{} version 9.0.0 was released April X, 2018.
+\dealii version 9.0.0 was released April X, 2018.
 \marginpar{Update the date}
@@
 overview of the new features of this major release and serves as a citable
-reference for the \dealii{} software library version 9.0. \dealii{} is an
+reference for the \dealii software library version 9.0. \dealii is an
 object-oriented finite element library used around the world in the
 development of finite element solvers. It is available for free under the
-GNU Lesser General Public License (LGPL) from the \dealii{} homepage at
+GNU Lesser General Public License (LGPL) from the \dealii homepage at
 \url{http://www.dealii.org/}.
@@
 \begin{itemize}
-\item xxxx something with manifolds xxx
-\item Support for particle-in-cell methods
-\item Dedicated support for automatic and symbolic differentiation.
-\item Interfaces to more external libraries and programs.
-\item C++11 is now both required and used.
-\item Support for GPU computations.
+\item Improved support for curved geometries;
+\item Support for particle-in-cell methods;
+\item Dedicated support for automatic and symbolic differentiation;
+\item Interfaces to more external libraries and programs;
+\item C++11 is now both required and used;
+\item Support for GPU computations;
 \end{itemize}
@@
 \begin{itemize}
-\item \dealii{} has made extensive use of both the Clang-Tidy \cite{clang-tidy}
+\item \dealii has made extensive use of both the Clang-Tidy \cite{clang-tidy}
   and Coverity Scan \cite{coverity} static analysis tools for detecting bugs
@@
   updates to step-37. In addition, the separate code
-  gallery of \dealii{} has gained a number of new entries.
+  gallery of \dealii has gained a number of new entries.
+
+\item Improved support for user parameters; a new \texttt{ParameterAcceptor}
+  class has been added to the library. The class is intended to be used as a
+  base for any class that wants to handle parameters using the
+  \texttt{ParameterHandler} class. If you derive all your classes from
+  \texttt{ParameterAcceptor}, and declare your parameters either with
+  \texttt{parse/declare\_parameters} methods or via the
+  \texttt{ParameterAcceptor::add\_parameter} method, then the declaring and
+  parsing of your parameter files will be automatically managed by the
+  \texttt{ParameterAcceptor::initialize} function, greatly simplifying
+  the declaration and parsing of parameters in user codes.
+
+\item New caching mechanism for expensive grid computations; We introduced a new
+  class, \texttt{GridTools::Cache}, that caches computationally intensive
+  information about a \texttt{Triangulation}.
+
+This class allows the user to query some of the data structures constructed
+using functions in the GridTools namespace which are computed only once, and
+then cached inside this class for faster access whenever the triangulation has
+not changed. The Cache is marked for update by the \texttt{Triangulation} itself
+using signals.
+
+Some of the methods in \texttt{GridTools} already use the \texttt{Cache} to
+speed up repeated calls to the same expensive methods.
+\item A new \texttt{MeshWorker::mesh\_loop} function has been added that
+  performs the same tasks of the \texttt{MeshWorker::loop} function without
+  forcing the users to adhere to a specific interface.
+
 \end{itemize}
@@
 
-This release of \dealii{} contains a number of large and significant changes
+This release of \dealii contains a number of large and significant changes
 that will be discussed in the following sections. It of course also contains a
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{xxx something with Manifolds xxx}
+\subsection{Improved support for curved geometries}
 \label{sec:manifolds}
@@
 
-(i) separation between manifold and boundary ids, (ii) manifold
-smoothing via the transfinite mapping, (iii) maybe: grid generator functions already attach manifolds?
+Every function in the GridGenerator namespace now attaches a default manifold to
+the curved parts of the domain, and sets reasonable defaults for manifold
+indicators both in the domain and on the boundary, where appropriate. Manifold
+classes are no longer sensible to settings made on boundary indicators, and are
+only affected by manifold indicators, decoupling completely the concept of
+boundary indicators and manifold indicator, leaving the former for boundary
+conditions, and the latter for the geometrical description of the domain.
 
-Previous text:
+All functions that allowed to query for \texttt{Boundary} objects (deprecated
+since version 8.9 of the \dealii library) have been removed, and have been
+replaced by the equivalent methods that query \texttt{Manifold} classes.
 
-Manifold smoothing: The manifold smoothing algorithms applied in the
-Triangulation class and MappingQGeneric have been changed from the old
-Laplace-style smoothing to a transfinite interpolation that linearly
-blends between the descriptions on the faces around a cell. The old
-transformation introduced boundary layers inside cells that prevented
-convergence rates from exceeding \(3.5\) in the global \(L^2\) errors on typical
-settings. This change also considerably improves mesh quality on settings
-where curved descriptions are only applied to the boundary rather than
-the whole volume.
+The manifold smoothing algorithms applied in the Triangulation class and
+MappingQGeneric have been changed from the old Laplace-style smoothing to a
+transfinite interpolation that linearly blends between the descriptions on the
+faces around a cell. The old transformation introduced boundary layers inside
+cells that prevented convergence rates from exceeding \(3.5\) in the global
+\(L^2\) errors on typical settings. This change also considerably improves mesh
+quality on settings where curved descriptions are only applied to the boundary
+rather than the whole volume.
+
+A new manifold class \texttt{TransfiniteInterpolationManifold} implementing an
+interpolation from a curved boundary description to the interior has been added.
+This class enables high-order convergence rates of more than three in the power
+of the mesh size for situations where a curved manifold can only be prescribed
+to the boundary but not in a whole volume.
 
@@
 \marginpar{Rene: Please edit as appropriate}
-While \dealii{} is a package intended to solve problems with the
+While \dealii is a package intended to solve problems with the
 finite element method -- i.e., using continuous or discontinuous
@@
 
-\dealii{} now has a dedicated particles module. The module provides a base class
+\dealii now has a dedicated particles module. The module provides a base class
 \texttt{Particle} that represents a particle with position, an ID number and
@@
 \cite{GLHPB18}. A longer report is at \cite{GHPB16}. The
-implementation here originated in the \aspect{} code, see \cite{KHB12,HDGB17}.
+implementation here originated in the \aspect code, see \cite{KHB12,HDGB17}.
 
@@
 
-\dealii{} has had a tutorial program (step-33) since 2007 that
+\dealii has had a tutorial program (step-33) since 2007 that
 demonstrates this technique based on the Trilinos Sacado package, but
@@
 
-\dealii{} has always tried to leverage high-quality implementations of
+\dealii has always tried to leverage high-quality implementations of
 algorithms available through other open source software, rather than
@@
 
-  \dealii{} now has interfaces to SUNDIALS's ARKode, IDA, and KINSOL sub-packages.
+  \dealii now has interfaces to SUNDIALS's ARKode, IDA, and KINSOL sub-packages.
   ARKode is a solver library that provides adaptive-step time
@@
 searching the points that fall within a radius of a target point. This
-functionality is now available via \dealii{} interfaces.
+functionality is now available via \dealii interfaces.
 
@@
   Assimp can be used to read about 40 different 3D graphics formats. A
-  subset of these formats can be now be read from within \dealii{} to
+  subset of these formats can be now be read from within \dealii to
   generate two-dimensional meshes, possibly embedded in a
@@
   \label{sec:cxx11}
-\dealii{} first offered support for a subset of C++11 features in
+\dealii first offered support for a subset of C++11 features in
 version 6.2, released in 2009. The current release is the first to
@@
 Many parts of the code base have been rewritten to both support and
-use the new features of C++11. In particular, \dealii{} now makes
+use the new features of C++11. In particular, \dealii now makes
 extensive use of move semantics as well as range-based \texttt{for}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{MeshWorker::mesh\_loop}
+
+\marginpar{do we want this? LH}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Tutorial and code gallery programs}
@@
 
-\dealii{} has a separate ``code gallery'' that
+\dealii has a separate ``code gallery'' that
 consists of programs shared by users as examples of what can be
-done with \dealii{}. While not part of the release process, it is nonetheless
+done with \dealii. While not part of the release process, it is nonetheless
 worth mentioning that the set of new programs since the last release covers
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{How to cite \dealii{}}\label{sec:cite}
+\section{How to cite \dealii}\label{sec:cite}
 
-In order to justify the work the developers of \dealii{} put into this
+In order to justify the work the developers of \dealii put into this
 software, we ask that papers using the library reference one of the
-\dealii{} papers. This helps us justify the effort we put into it.
+\dealii papers. This helps us justify the effort we put into it.
 
-There are various ways to reference \dealii{}. To acknowledge the use of
+There are various ways to reference \dealii. To acknowledge the use of
 the current version of the library, \textbf{please reference the present
@@
 
-The original \texttt{\dealii{}} paper containing an overview of its
+The original \texttt{\dealii} paper containing an overview of its
 architecture is \cite{BangerthHartmannKanschat2007}. If you rely on
@@
 \marginpar{All: update gmsh reference}
-\dealii{} can interface with many other libraries:
+\dealii can interface with many other libraries:
 \begin{multicols}{3}
@@
 
-Older releases of \dealii{} can be cited as
+Older releases of \dealii can be cited as
 \cite{dealII80,dealII81,dealII82,dealII83,dealII84,dealII85}.
@@
 
-\dealii{} is a world-wide project with dozens of contributors around the
+\dealii is a world-wide project with dozens of contributors around the
 globe. Other than the authors of this paper, the following people contributed code to
@@
 
-\dealii{} and its developers are financially supported through a
+\dealii and its developers are financially supported through a
 variety of funding sources:
@@
 The Interdisciplinary Center for Scientific Computing (IWR) at Heidelberg
-University has provided hosting services for the \dealii{} web page.
+University has provided hosting services for the \dealii web page.
 

### 2018-04-24 9.0/paper.tex
Update references to newly supported libraries

@@
 \item Support for GPU computations;
+\item Support for face integrals and significant improvements of matrix-free framework;
 \end{itemize}
@@
   fixed using the latter tool.
-      %% > Is there somewhere we can cite for either of these tools? It would
-      %% > be nice to recognize the department of homeland security for offering
-      %% The contract with homeland security was completed in 2009.
 
@@
 
-\item
-  Significant extension of matrix-free capabilities, including support for face
-  integrals in discontinuous Galerkin schemes, a new Hermite-like polynomial
-  basis adapted to face integrals that involve derivatives of shape functions,
-  and several performance enhancements.
-
 \item New non-standard quadrature rules:
@@
 implemented. Among these are ones for
-(i) truncating standard formulas to simplical domains (\texttt{Qsimplex});
+(i) truncating standard formulas to simplical domains (\texttt{QSimplex});
 (ii) singular transformations of the unit cell to the unit simplex
@@
 finite element method -- i.e., using continuous or discontinuous
-\textit{fields} --, is often convenient in fluid dynamics problems to
+\textit{fields} --, it is often convenient in fluid dynamics problems to
 couple the continuum description of phenomena with particles. These
@@
 
-\marginpar{Add references to all of these libraries.}
+\paragraph*{Assimp, the Open Asset Import Library.}
+  Assimp \cite{assimp} can be used to read about 40 different 3D
+  graphics formats. A subset of these formats can be now be read from
+  within \dealii to generate two-dimensional meshes, possibly embedded in
+  a three-dimensional space.
+
+% \paragraph*{Gmsh.}
+%   \marginpar{Do we actually provide any functionality with the GMSH
+%   executable yet? If so, complete this part}
+%   Gmsh \cite{geuzaine2009gmsh}
+% The only thing we currently do is to define DEAL_II_GMSH_EXECUTABLE_PATH
+
+\paragraph*{nanoflann, a library for building and querying
+  $k$-d trees of datasets.} Operations such as finding the vertex or cell
+  closest to a given evaluation point occur frequently in many applications
+  that use unstructured meshes. While the naive algorithm is linear in the
+  number of vertices or cells, many such operations can be made significantly
+  faster by building a $k$-d tree data structure that recursively subdivides
+  a $k$ dimensional space. The nanoflann library \cite{nanoflann} provides
+  such a data structure and allows querying it, either for closest points
+  (e.g., when finding the closest vertex) or for searching the points that
+  fall within a radius of a target point. This functionality is now available
+  via \dealii interfaces.
+
+\paragraph*{ROL, a Rapid Optimization Library.}
+  ROL \cite{ridzal2014rapid} is a package for large-scale optimization.
+  \marginpar{Can we add a sentence or two about the ROL interface here?}
 
 \paragraph*{ScaLAPACK, a parallel dense linear algebra library.}
-  ScaLAPACK provides block-cyclic matrix distribution over 2D process
-  grid. The functionality and interface of our wrappers is similar to
-  the LAPACK wrappers for serial dense linear algebra, namely
+  ScaLAPACK \cite{slug} provides block-cyclic matrix distribution over 2D
+  process grid. The functionality and interface of our wrappers is similar
+  to the LAPACK wrappers for serial dense linear algebra, namely
   matrix-matrix multiplication, Cholesky and LU factorizations,
@@
   common task and one that often requires sophisticated globalization
-  algorithms for efficiency and reliability. SUNDIALS provides these
-  in a widely used format, both sequentially and in parallel.
+  algorithms for efficiency and reliability. SUNDIALS \cite{sundials}
+  provides these in a widely used format, both sequentially and in parallel.
 
@@
   algebraic systems.
-
-\paragraph*{ROL, a Rapid Optimization Library.}
-ROL is a package for large-scale optimization.
-\marginpar{Can we add a sentence or two about the ROL interface here?}
-
-\paragraph*{nanoflann, a library for building and querying
-  $k$-d trees of datasets.} Operations such as finding the vertex
-or cell closest
-to a given evaluation point occur frequently in many applications that
-use unstructured meshes. While the naive algorithm is linear in the
-number of vertices or cells, many such operations can be made
-significantly faster by building a $k$-d tree data structure that
-recursively subdivides a $k$ dimensional space. The nanoflann library
-provides such a data structure and allows querying it, either for
-closest points (e.g., when finding the closest vertex) or for
-searching the points that fall within a radius of a target point. This
-functionality is now available via \dealii interfaces.
-
-\paragraph*{Assimp, the Open Asset Import Library.}
-  Assimp can be used to read about 40 different 3D graphics formats. A
-  subset of these formats can be now be read from within \dealii to
-  generate two-dimensional meshes, possibly embedded in a
-  three-dimensional space.
-
-\paragraph*{Gmsh.}
-  \marginpar{Do we actually provide any functionality with the GMSH
-  executable yet? If so, complete this part}
 
@@
 \marginpar{do we want this? LH}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Extended matrix-free capabilities}
+
+The matrix-free infrastructure in \dealii{} was significantly overhauled for
+the current release. The major new contribution is the support of face
+integrals through a new class \texttt{FEFaceEvaluation}. The new class has a
+similar interface as the previous \texttt{FEEvaluation}, and applies SIMD
+vectorization over several faces in analogy to the intra-cell vectorization in
+FEEvaluation. The data structures have been particularly tuned for typical
+discontinuous Galerkin setups involving operators with first and second
+spatial derivatives. Both data access and computations have been thoroughly
+optimized and compared to the performance boundaries of the hardware.
+
+Furthermore, the internal algorithms for computing the values and gradients on
+all quadrature points for cell integrals have been significantly improved,
+giving around 10--20\% better performance for the case the kernels are compute
+bound. For the example of the reference cell gradient of a solution field
+$\mathbf u$ in three space dimensions, the new release applies the following
+change:
+\begin{equation*}
+  \text{previous: }
+  \begin{bmatrix}
+    D_{1} \otimes S_{2} \otimes S_{3}\\
+    S_{1} \otimes D_{2} \otimes S_{3}\\
+    S_{1} \otimes S_{2} \otimes D_{3}
+  \end{bmatrix}
+  \mathbf u
+  \quad
+  \leadsto
+  \quad
+  \text{new: }
+  \begin{bmatrix}
+    D_{1}^{\mathrm{co}} \otimes I_{2} \otimes I_{3}\\
+    I_{1} \otimes D_{2}^{\mathrm{co}} \otimes I_{3}\\
+    I_{1} \otimes I_{2} \otimes D_{3}^{\mathrm{co}}
+  \end{bmatrix}
+  \begin{bmatrix}
+    S_{1} \otimes S_{2} \otimes S_{3}
+  \end{bmatrix}
+  \mathbf u.
+\end{equation*}
+The matrices $S_i$ contain the values of the one-dimensional shape functions
+in one-dimensional quadrature points and $D_i$ their derivatives. When applied
+with the usual sum factorization implementation described e.g.~in
+\cite{KronbichlerKormann2012}, the old kernels amounted to 9 partial
+summations---or rather 8 in the previous implementation of \dealii{} because
+the application of $S_1$ for the $y$ and $z$ components of the gradient can be
+merged. The new code performs a basis transformation to a related basis with
+derivative matrix $D_i = D_i^{\mathrm{co}} S_i$, which is the basis of
+Lagrange polynomials in the points of the quadrature. This change reduces the
+number of partial sums to only 6 for the gradient, as the action of the unit
+matrices $I_i$ needs not be implemented. In isolation, this spectral
+element-like evaluation was previously available in \dealii{} for collocation
+between nodal points and quadrature, but not used for general bases. A more
+detailed description of this improvement, including an analysis of these
+kernels on modern hardware, is given in the preprint
+\cite{KronbichlerKormann2017}.
 
@@
 \item BLAS and LAPACK
-\item Gmsh \cite{...}
+\item Gmsh \cite{geuzaine2009gmsh}
 \item GSL \cite{gsl2016}
@@
 \item ROL \cite{ridzal2014rapid}
-\item ScaLAPACK
+\item ScaLAPACK \cite{slug}
 \item SLEPc \cite{Hernandez:2005:SSF}
@@
 Research Foundation (DFG) under the project ``High-order discontinuous
-Galerkin for the exa-scale'' (ExaDG) within the priority program ``Software
+Galerkin for the exa-scale'' (\mbox{ExaDG}) within the priority program ``Software
 for Exascale Computing'' (SPPEXA).

### 2018-04-24 9.0/paper.tex
Write more about matrix-free, update references.

@@
   fixed using the latter tool.
+      %% > Is there somewhere we can cite for either of these tools? It would
+      %% > be nice to recognize the department of homeland security for offering
+      %% The contract with homeland security was completed in 2009.
 
@@
 implemented. Among these are ones for
-(i) truncating standard formulas to simplical domains (\texttt{QSimplex});
+(i) truncating standard formulas to simplical domains (\texttt{Qsimplex});
 (ii) singular transformations of the unit cell to the unit simplex
@@
 
-\paragraph*{Assimp, the Open Asset Import Library.}
-  Assimp \cite{assimp} can be used to read about 40 different 3D
-  graphics formats. A subset of these formats can be now be read from
-  within \dealii to generate two-dimensional meshes, possibly embedded in
-  a three-dimensional space.
-
-% \paragraph*{Gmsh.}
-%   \marginpar{Do we actually provide any functionality with the GMSH
-%   executable yet? If so, complete this part}
-%   Gmsh \cite{geuzaine2009gmsh}
-% The only thing we currently do is to define DEAL_II_GMSH_EXECUTABLE_PATH
-
-\paragraph*{nanoflann, a library for building and querying
-  $k$-d trees of datasets.} Operations such as finding the vertex or cell
-  closest to a given evaluation point occur frequently in many applications
-  that use unstructured meshes. While the naive algorithm is linear in the
-  number of vertices or cells, many such operations can be made significantly
-  faster by building a $k$-d tree data structure that recursively subdivides
-  a $k$ dimensional space. The nanoflann library \cite{nanoflann} provides
-  such a data structure and allows querying it, either for closest points
-  (e.g., when finding the closest vertex) or for searching the points that
-  fall within a radius of a target point. This functionality is now available
-  via \dealii interfaces.
-
-\paragraph*{ROL, a Rapid Optimization Library.}
-  ROL \cite{ridzal2014rapid} is a package for large-scale optimization.
-  \marginpar{Can we add a sentence or two about the ROL interface here?}
+\marginpar{Add references to all of these libraries.}
 
 \paragraph*{ScaLAPACK, a parallel dense linear algebra library.}
-  ScaLAPACK \cite{slug} provides block-cyclic matrix distribution over 2D
-  process grid. The functionality and interface of our wrappers is similar
-  to the LAPACK wrappers for serial dense linear algebra, namely
+  ScaLAPACK provides block-cyclic matrix distribution over 2D process
+  grid. The functionality and interface of our wrappers is similar to
+  the LAPACK wrappers for serial dense linear algebra, namely
   matrix-matrix multiplication, Cholesky and LU factorizations,
@@
   common task and one that often requires sophisticated globalization
-  algorithms for efficiency and reliability. SUNDIALS \cite{sundials}
-  provides these in a widely used format, both sequentially and in parallel.
+  algorithms for efficiency and reliability. SUNDIALS provides these
+  in a widely used format, both sequentially and in parallel.
 
@@
   algebraic systems.
+
+\paragraph*{ROL, a Rapid Optimization Library.}
+ROL is a package for large-scale optimization.
+\marginpar{Can we add a sentence or two about the ROL interface here?}
+
+\paragraph*{nanoflann, a library for building and querying
+  $k$-d trees of datasets.} Operations such as finding the vertex
+or cell closest
+to a given evaluation point occur frequently in many applications that
+use unstructured meshes. While the naive algorithm is linear in the
+number of vertices or cells, many such operations can be made
+significantly faster by building a $k$-d tree data structure that
+recursively subdivides a $k$ dimensional space. The nanoflann library
+provides such a data structure and allows querying it, either for
+closest points (e.g., when finding the closest vertex) or for
+searching the points that fall within a radius of a target point. This
+functionality is now available via \dealii interfaces.
+
+\paragraph*{Assimp, the Open Asset Import Library.}
+  Assimp can be used to read about 40 different 3D graphics formats. A
+  subset of these formats can be now be read from within \dealii to
+  generate two-dimensional meshes, possibly embedded in a
+  three-dimensional space.
+
+\paragraph*{Gmsh.}
+  \marginpar{Do we actually provide any functionality with the GMSH
+  executable yet? If so, complete this part}
 
@@
 \item BLAS and LAPACK
-\item Gmsh \cite{geuzaine2009gmsh}
+\item Gmsh \cite{...}
 \item GSL \cite{gsl2016}
@@
 \item ROL \cite{ridzal2014rapid}
-\item ScaLAPACK \cite{slug}
+\item ScaLAPACK
 \item SLEPc \cite{Hernandez:2005:SSF}

### 2018-04-24 9.0/paper.tex
remove sidenote
Some minor markup improvements.

--- 2018-04-24 9.0/paper.tex
@@
 \item Support for GPU computations;
-\item Support for face integrals and significant improvements of matrix-free framework;
+\item Support for face integrals and significant improvements of the matrix-free framework;
 \end{itemize}
@@
 
-\item Improved support for user parameters; a new \texttt{ParameterAcceptor}
+\item Improved support for user parameters: a new \texttt{ParameterAcceptor}
   class has been added to the library. The class is intended to be used as a
@@
   \texttt{parse/declare\_parameters} methods or via the
-  \texttt{ParameterAcceptor::add\_parameter} method, then the declaring and
-  parsing of your parameter files will be automatically managed by the
-  \texttt{ParameterAcceptor::initialize} function, greatly simplifying
-  the declaration and parsing of parameters in user codes.
+  \texttt{ParameterAcceptor::add\_parameter} method, then both the
+  declaration and the
+  parsing of parameter files are automatically managed by
+  \texttt{ParameterAcceptor::initialize}, greatly simplifying
+  dealing with parameters in user codes.
 
-\item New caching mechanism for expensive grid computations; We introduced a new
-  class, \texttt{GridTools::Cache}, that caches computationally intensive
+\item New caching mechanism for expensive grid computations: we introduced a new
+  class \texttt{GridTools::Cache} that caches computationally intensive
   information about a \texttt{Triangulation}.
@@
 This class allows the user to query some of the data structures constructed
-using functions in the GridTools namespace which are computed only once, and
+using functions in the \texttt{GridTools} namespace. This data is then
+computed only once, and
 then cached inside this class for faster access whenever the triangulation has
-not changed. The Cache is marked for update by the \texttt{Triangulation} itself
+not changed. The cache is marked for update by the \texttt{Triangulation} itself
 using signals.
 
-Some of the methods in \texttt{GridTools} already use the \texttt{Cache} to
+Some of the methods in \texttt{GridTools} already use this cache to
 speed up repeated calls to the same expensive methods.
@@
 
-\marginpar{All: update gmsh reference}
 \dealii can interface with many other libraries:

### 2018-04-24 9.0/paper.tex
Revise matrix-free.

@@
 field, are then either used to visualize properties of the flow, or to
-advect material properties such as the viscosity of inhomogenous
+advect material properties such as the viscosity of inhomogeneous
 mixtures of fluids. If each particle is associated with the cells
@@
 vectorization over several faces in analogy to the intra-cell vectorization in
-FEEvaluation. The data structures have been particularly tuned for typical
-discontinuous Galerkin setups involving operators with first and second
-spatial derivatives. Both data access and computations have been thoroughly
-optimized and compared to the performance boundaries of the hardware.
+FEEvaluation. Discontinuous Galerkin operators are implemented defining two
+face functions, one for interior and one for boundary faces, in addition to
+the cell function. These kernels for the matrix-free operator evaluation are
+now collected in the new function \texttt{MatrixFree::loop}. The data
+structures have been particularly tuned for typical discontinuous Galerkin
+setups involving operators with first and second spatial derivatives. Both
+data access and computations have been thoroughly optimized and compared to
+the performance boundaries of the hardware.
 
-Furthermore, the internal algorithms for computing the values and gradients on
-all quadrature points for cell integrals have been significantly improved,
-giving around 10--20\% better performance for the case the kernels are compute
-bound. For the example of the reference cell gradient of a solution field
-$\mathbf u$ in three space dimensions, the new release applies the following
-change:
+To give an example of the algorithmic improvements, the computation of the
+values and gradients on all quadrature points for cell integrals has been
+significantly improved, giving around 10--20\% better performance for the case
+the kernels are compute bound. For the example of the reference cell gradient
+of a solution field $\mathbf u$ in three space dimensions, the new release
+applies the following change:
 \begin{equation*}
@@
 between nodal points and quadrature, but not used for general bases. A more
-detailed description of this improvement, including an analysis of these
-kernels on modern hardware, is given in the preprint
+detailed description of the matrix-free modules is given in the preprint
 \cite{KronbichlerKormann2017}.

### 2018-04-24 9.0/paper.tex
Fix old version number.

--- 2018-04-24 9.0/paper.tex
@@
 All functions that allowed to query for \texttt{Boundary} objects (deprecated
-since version 8.9 of the \dealii library) have been removed, and have been
+since version 8.5 of the \dealii library) have been removed, and have been
 replaced by the equivalent methods that query \texttt{Manifold} classes.

### 2018-04-25 9.0/paper.tex
ScaLAPACK minor changes

@@
 
-\paragraph*{ScaLAPACK, a parallel dense linear algebra library.}
+\paragraph*{ScaLAPACK, a parallel dense linear algebra library for distributed memory machines.}
   ScaLAPACK \cite{slug} provides block-cyclic matrix distribution over 2D
-  process grid. The functionality and interface of our wrappers is similar
-  to the LAPACK wrappers for serial dense linear algebra, namely
+  process grids. The functionality and interface of our wrappers is similar
+  to the LAPACK \cite{lapack} wrappers for serial dense linear algebra, namely
   matrix-matrix multiplication, Cholesky and LU factorizations,
-  eigensolvers, SVD, pseudoinverses, and save/load operations using
-  the HDF5 format. All of this functionality is available even in
-  cases where the number of MPI processes does not exactly divide the
-  number of tiles stored in the block-cyclic data distribution.
+  eigensolvers, SVD, least squares, pseudoinverses and save/load operations using
+  either serial or parallel HDF5 \cite{hdf5}. All of this functionality is available
+  even in cases where the number of MPI processes does not match the numbers of
+  processes in the 2D process grid used to distribute a matrix.
 
@@
 \item Assimp \cite{assimp}
-\item BLAS and LAPACK
+\item BLAS and LAPACK \cite{lapack}
 \item Gmsh \cite{geuzaine2009gmsh}

### 2018-04-25 9.0/paper.tex
added ROL information

@@
   ROL \cite{ridzal2014rapid} is a package for large-scale optimization.
-  \marginpar{Can we add a sentence or two about the ROL interface here?}
+  \dealii can now use the state-of-the-art algorithms in ROL to solve
+  unconstrained and constrained optimization problems as well as optimization
+  problems under uncertainty.
+  \dealii provides an interface to ROL's (abstract) vector class using the
+  adapter software pattern.
+  Through such an interface any vector class in \dealii following certain
+  interface requirements can be used to define a ROL objective function.
+  It is strongly recommended that the user additionally implement the objective
+  function gradient and the action of the Hessian.
 

### 2018-04-25 9.0/paper.tex
Add section on automatic differentiation

--- 2018-04-25 9.0/paper.tex
@@
 \item Support for particle-in-cell methods;
-\item Dedicated support for automatic and symbolic differentiation;
+\item Dedicated support for automatic differentiation;
 \item Interfaces to more external libraries and programs;
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Dedicated support for automatic and symbolic differentiation}
+\subsection{Dedicated support for automatic differentiation}
 
-\marginpar{J-P: Please edit as appropriate}
-Automatic differentiation is often used to automatically derive
-residuals from the stored energy functional, and to derive Jacobian
-matrices from residual vectors for simulations that use complicated
-material models. Examples can be found widely for nonlinear solid
-mechanics as well as for nonlinear viscosity models in fluid flow.
+Automatic differentiation (AD) is often used to automatically derive
+residuals and their linearization from a stored energy functional, and to
+derive Jacobian matrices from residual vectors for simulations that use
+complicated material models. Examples of its application can be found widely
+within nonlinear solid mechanics, coupled multiphysics problems, as well as
+for nonlinear viscosity models in fluid flow.
 
 \dealii has had a tutorial program (step-33) since 2007 that
-demonstrates this technique based on the Trilinos Sacado package, but
+demonstrates this technique based on the Trilinos Sacado \cite{Bartlett2006a} package, but
 the functionality was not available pervasively throughout
-deal.II. This has changed with release 9.0 where support for
-differentiation has been collected in one module that provides both
-automatic and symbolic differentiation. Currently, this module relies
-on either the ADOL-C and Sacado libraries, through a unified
-interface. In practice, this support means that ADOL-C and Sacado data
+deal.II. 
+This has changed with release 9.0 where support is given for differentiation 
+using a selection of ``white-listed'' libraries (namely ADOL-C \citep{Griewank1996a} 
+and Sacado) and a subset of their supported number types. Currently, we offer support for
+\begin{itemize}
+\item ADOL-C taped (n-differentiable),
+\item ADOL-C tapeless (once differentiable),
+\item Sacado dynamic forward (once differentiable),
+\item Sacado reverse (once differentiable),
+\item Sacado nested dynamic forward (twice differentiable), and
+\item Sacado nested reverse and dynamic forward (twice differentiable)
+\end{itemize}
+numbers. In practice, this support means that these ADOL-C and Sacado data
 types can be used in the \texttt{FEValues}, \texttt{FEValuesViews},
+\texttt{Tensor}, \texttt{SymmetricTensor},
 and related classes that are generally used to assemble linear systems
 and right hand sides.
+Given the updated capabilities of the library, there is now a dedicated module 
+that presents the AD compatibility and capabilities of the \dealii libraries. 
+Furthermore, the use of Sacado is now  demonstrated in a much more simplified 
+and transparent manner in an modernized version of an existing ``code gallery'' 
+example \cite{Pelteret2016a}.
+
+To date it remains necessary for the user  to manage the initialization of AD 
+independent variables and the resultant calls  to the AD dependent variables 
+in order to initiate the computation of derivatives.
+In the next release we expect to provide a unified interface to these
+AD libraries, which will hide these library-dependent implementational details 
+and facilitate switching between the supported libraries and AD number types
+based on the user's requirements.
 
@@
 \begin{itemize}
-\item ADOL-C \cite{adol-c}
+\item ADOL-C \cite{Griewank1996a,adol-c}
 \item ARPACK \cite{arpack}

### 2018-04-25 9.0/paper.tex
removing last line

@@
   ROL \cite{ridzal2014rapid} is a package for large-scale optimization.
-  \marginpar{Can we add a sentence or two about the ROL interface here?}
+  \dealii can now use the state-of-the-art algorithms in ROL to solve
+  unconstrained and constrained optimization problems as well as optimization
+  problems under uncertainty.
+  \dealii provides an interface to ROL's (abstract) vector class using the
+  adapter software pattern.
+  Through such an interface any vector class in \dealii following certain
+  interface requirements can be used to define a ROL objective function.
 

### 2018-04-25 9.0/paper.tex
Some edits.

@@
 
-\item Improved support for user parameters: a new \texttt{ParameterAcceptor}
+\item Improved support for user-defined run-time parameters: a new \texttt{ParameterAcceptor}
   class has been added to the library. The class is intended to be used as a
@@
 
-Every function in the GridGenerator namespace now attaches a default manifold to
-the curved parts of the domain, and sets reasonable defaults for manifold
-indicators both in the domain and on the boundary, where appropriate. Manifold
-classes are no longer sensible to settings made on boundary indicators, and are
-only affected by manifold indicators, decoupling completely the concept of
-boundary indicators and manifold indicator, leaving the former for boundary
-conditions, and the latter for the geometrical description of the domain.
+\dealii{} has had the ability to attach \textit{manifold descriptions}
+to all parts of a geometry for several releases already. These
+descriptions are used when considering where to place new vertices
+upon mesh refinement, in determining the mapping from reference cell
+to real cell, and in a number of other contexts. However, for historical
+reasons, manifold descripts have used some of the same code paths also
+used for boundary indicators typically used to identify which parts of
+the boundary correspond to what boundary conditions.
 
-All functions that allowed to query for \texttt{Boundary} objects (deprecated
+This connection has been severed in the current release: Boundary
+indicators and manifold descriptions are now entirely separated. Furthermore,
+all functions that allowed to query for old-style \texttt{Boundary} objects (deprecated
 since version 8.5 of the \dealii library) have been removed, and have been
@@
 
-The manifold smoothing algorithms applied in the Triangulation class and
-MappingQGeneric have been changed from the old Laplace-style smoothing to a
+There are also numerous improvements to the available manifold descriptions.
+First, the manifold smoothing algorithms applied in the \texttt{Triangulation} class and
+\texttt{MappingQGeneric} have been changed from the old Laplace-style smoothing to a
 transfinite interpolation that linearly blends between the descriptions on the
@@
 cells that prevented convergence rates from exceeding \(3.5\) in the global
-\(L^2\) errors on typical settings. This change also considerably improves mesh
-quality on settings where curved descriptions are only applied to the boundary
+\(L^2\) errors for typical settings. This change also considerably improves mesh
+quality in situations where curved descriptions are only applied to the boundary
 rather than the whole volume.
 
-A new manifold class \texttt{TransfiniteInterpolationManifold} implementing an
-interpolation from a curved boundary description to the interior has been added.
-This class enables high-order convergence rates of more than three in the power
-of the mesh size for situations where a curved manifold can only be prescribed
+\marginpar{This paragraph seems to duplicate the previous. Can we
+  merge the information?}
+Secondly, a new manifold class
+\texttt{TransfiniteInterpolationManifold} has been added that implements an
+interpolation from a curved boundary description to the interior.
+This class enables high-order convergence rates of better than ${\cal O}(h^3)$
+for situations where a curved manifold can only be prescribed
 to the boundary but not in a whole volume.
+
+Finally,
+every function in the \texttt{GridGenerator} namespace now attaches a default manifold to
+the curved parts of the domain described by the generated mesh, and sets reasonable defaults for manifold
+indicators both in the domain and on the boundary.
 
@@
 using a selection of ``white-listed'' libraries (namely ADOL-C \citep{Griewank1996a} 
-and Sacado) and a subset of their supported number types. Currently, we offer support for
+and Sacado) and a subset of their supported number types. Currently,
+we offer support for the following cases:
 \begin{itemize}
@@
 \item Sacado nested dynamic forward (twice differentiable), and
-\item Sacado nested reverse and dynamic forward (twice differentiable)
+\item Sacado nested reverse and dynamic forward (twice differentiable).
 \end{itemize}
-numbers. In practice, this support means that these ADOL-C and Sacado data
+In practice, this support means that these ADOL-C and Sacado data
 types can be used in the \texttt{FEValues}, \texttt{FEValuesViews},
@@
 integrals through a new class \texttt{FEFaceEvaluation}. The new class has a
-similar interface as the previous \texttt{FEEvaluation}, and applies SIMD
+similar interface as the existing \texttt{FEEvaluation} class, and applies SIMD
 vectorization over several faces in analogy to the intra-cell vectorization in
@@
 values and gradients on all quadrature points for cell integrals has been
-significantly improved, giving around 10--20\% better performance for the case
+significantly improved, yielding 10--20\% better performance for cases where
 the kernels are compute bound. For the example of the reference cell gradient
@@
 in one-dimensional quadrature points and $D_i$ their derivatives. When applied
-with the usual sum factorization implementation described e.g.~in
+with the usual sum factorization implementation described, for
+example, in
 \cite{KronbichlerKormann2012}, the old kernels amounted to 9 partial
-summations---or rather 8 in the previous implementation of \dealii{} because
+summations -- or rather 8 in the previous implementation of \dealii{} because
 the application of $S_1$ for the $y$ and $z$ components of the gradient can be
@@
 several have been updated extensively for the changes to the manifold
-handling as well as to adjust for current functionality
+handling as well as to adjust for current functionality and coding styles
 compared to what that was available when the programs were first written.

### 2018-04-26 9.0/paper.tex
Added BEM citation.
Add section about GPU support

@@
 
-\marginpar{Bruno?}
-...
+Heterogeneous computing is becoming more prevalent in supercomputing and this is
+a trend that is expected to continue in the future. In particular, the use of
+GPU has been increasing the last few years.
+
+This release of \dealii adds support for GPU both for matrix-based and
+matrix-free applications. For matrix-based applications, we rely on
+cuSPARSE\cite{cusparse} and cuSOLVER\cite{cusolver} for operations on sparse
+matrices such as matrix-vector multiplication and for direct solvers. We have
+introduced a new type of sparse matrix, \texttt{CUDAWrappers::SparseMatrix},
+which moves onto the device a \dealii \texttt{SparseMatrix} and changes the
+format of the underlying data to the appropriate CSR format used by cuSPARSE. We
+also have added wrappers for Cholesky and LU factorizations provided by
+cuSOLVER. In practice, a user would assemble the matrix associated to the system
+on the host and then move the matrix to the device. At this point, the system
+would be solved on the device and the solution would be moved back to the host.
+
+We also have some support for matrix-free computation on GPU. For now, the
+evaluation of the operator is limited to mesh without hanging-nodes.
 
@@
    \cite{HeltaiMola2015};
+ \item For Boundary Elements Computations:
+   \cite{GiulianiMolaHeltai-2018-a};
  \item For \texttt{LinearOperator} and \texttt{PackagedOperation} facilities:
@@
 \item BLAS and LAPACK \cite{lapack}
+\item cuSOLVER \cite{cusolver}
+\item cuSPARSE \cite{cusparse}
 \item Gmsh \cite{geuzaine2009gmsh}

### 2018-04-28 9.0/paper.tex
Write about AVX-512.
Fix part on transfinite interpolation manifold.

@@
 quality in situations where curved descriptions are only applied to the boundary
-rather than the whole volume.
-
-\marginpar{This paragraph seems to duplicate the previous. Can we
-  merge the information?}
-Secondly, a new manifold class
-\texttt{TransfiniteInterpolationManifold} has been added that implements an
-interpolation from a curved boundary description to the interior.
-This class enables high-order convergence rates of better than ${\cal O}(h^3)$
-for situations where a curved manifold can only be prescribed
-to the boundary but not in a whole volume.
+rather than the whole volume. This concept was also introduced as a new manifold class
+\texttt{TransfiniteInterpolationManifold}, which allows to apply this type of smoothing not only in the cells close to the boundary but over a full coarse (level 0) cell.
 
@@
 data access and computations have been thoroughly optimized and compared to
-the performance boundaries of the hardware.
+the performance boundaries of the hardware. Furthermore, the support for
+AVX-512 instructions in the matrix-free framework was extended, adding new
+\texttt{gather} and \texttt{scatter} intrinsics for the indirect access to
+vector entries where appropriate.
 

### 2018-05-09 9.0/paper.tex
Edit the part on manifolds.
We have not yet removed Boundary or things like
Triangulation::get_boundary, but they are now deprecated (as of last
August).

@@
 
-\dealii{} has had the ability to attach \textit{manifold descriptions}
-to all parts of a geometry for several releases already. These
-descriptions are used when considering where to place new vertices
-upon mesh refinement, in determining the mapping from reference cell
-to real cell, and in a number of other contexts. However, for historical
-reasons, manifold descripts have used some of the same code paths also
-used for boundary indicators typically used to identify which parts of
-the boundary correspond to what boundary conditions.
+\dealii{} has had the ability to attach \textit{manifold descriptions} to all
+parts of a geometry since the 8.2 release. These descriptions are used to place
+new vertices during mesh refinement, to determine the mapping between the
+reference and real cells, and in a number of other contexts. These classes,
+inheriting from \texttt{Manifold}, describe coordinate transformations in a
+general way and completely supersede the older classes inheriting from
+\texttt{Boundary}. However, for historical reasons, manifold descriptors have
+used some of the same code paths as boundary indicators, which were only
+intended for marking what parts of the boundary correspond to what boundary
+conditions. Put another way: under certain circumstances a boundary indicator
+was also interpreted as a manifold indicator.
 
-This connection has been severed in the current release: Boundary
-indicators and manifold descriptions are now entirely separated. Furthermore,
-all functions that allowed to query for old-style \texttt{Boundary} objects (deprecated
-since version 8.5 of the \dealii library) have been removed, and have been
-replaced by the equivalent methods that query \texttt{Manifold} classes.
+The current release severs this connection: Boundary indicators and manifold
+descriptions are now entirely separated. This means that \texttt{boundary\_id}s
+are only used to set boundary conditions and \texttt{manifold\_id}s are only
+used to set geometry descriptions. The old compatibility code for using boundary
+indicators as manifold indicators has been removed and all usages of the
+old-style \texttt{Boundary} objects (even with manifold ids) are now deprecated.
 
 There are also numerous improvements to the available manifold descriptions.
-First, the manifold smoothing algorithms applied in the \texttt{Triangulation} class and
-\texttt{MappingQGeneric} have been changed from the old Laplace-style smoothing to a
-transfinite interpolation that linearly blends between the descriptions on the
-faces around a cell. The old transformation introduced boundary layers inside
-cells that prevented convergence rates from exceeding \(3.5\) in the global
-\(L^2\) errors for typical settings. This change also considerably improves mesh
-quality in situations where curved descriptions are only applied to the boundary
-rather than the whole volume. This concept was also introduced as a new manifold class
-\texttt{TransfiniteInterpolationManifold}, which allows to apply this type of smoothing not only in the cells close to the boundary but over a full coarse (level 0) cell.
+First, the manifold smoothing algorithms applied in the \texttt{Triangulation}
+class and \texttt{MappingQGeneric} have been changed from the old Laplace-style
+smoothing to a transfinite interpolation that linearly blends between the
+descriptions on the faces around a cell. The old transformation introduced
+boundary layers inside cells that prevented convergence rates from exceeding
+\(3.5\) in the global \(L^2\) errors for typical settings. This change also
+considerably improves mesh quality in situations where curved descriptions are
+only applied to the boundary rather than the whole volume. This concept was also
+introduced as a new manifold class \texttt{TransfiniteInterpolationManifold},
+which allows to apply this type of smoothing not only in the cells close to the
+boundary but over a full coarse (level 0) cell.
 
-Finally,
-every function in the \texttt{GridGenerator} namespace now attaches a default manifold to
-the curved parts of the domain described by the generated mesh, and sets reasonable defaults for manifold
-indicators both in the domain and on the boundary.
+Finally, every function in the \texttt{GridGenerator} namespace now attaches a
+default manifold to the curved parts of the domain described by the generated
+mesh, and sets reasonable defaults for manifold indicators both in the domain
+and on the boundary.
 

### 2018-05-09 9.0/paper.tex
Add note on adding notes for new tutorial programs.

@@
 
-This release does not contain any new tutorial programs, though
-several have been updated extensively for the changes to the manifold
-handling as well as to adjust for current functionality and coding styles
-compared to what that was available when the programs were first written.
+\marginpar{LH, MK, and others: Please add summaries of the new tutorial programs.}
 

### 2018-05-11 9.0/paper.tex
Write on step-59.

@@
 \item Support for face integrals and significant improvements of the matrix-free framework;
+\item Two new tutorial programs step-59 and step-60.
 \end{itemize}
@@
 
-\marginpar{All: please edit as appropriate}
-
 \dealii{} has had the ability to attach \textit{manifold descriptions} to all
@@
 
-\marginpar{Rene: Please edit as appropriate}
 While \dealii is a package intended to solve problems with the
@@
 a trend that is expected to continue in the future. In particular, the use of
-GPU has been increasing the last few years.
+GPUs has been increasing during the last few years.
 
-This release of \dealii adds support for GPU both for matrix-based and
+This release of \dealii adds support for GPUs both for matrix-based and
 matrix-free applications. For matrix-based applications, we rely on
@@
 
-We also have some support for matrix-free computation on GPU. For now, the
+We also have some support for matrix-free computation on a GPU. For now, the
 evaluation of the operator is limited to mesh without hanging-nodes.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{MeshWorker::mesh\_loop}
-
-\marginpar{do we want this? LH}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Extended matrix-free capabilities}
+\subsection{Extended matrix-free capabilities}\label{sec:matrixfree}
 
 The matrix-free infrastructure in \dealii{} was significantly overhauled for
-the current release. The major new contribution is the support of face
+this release. The major new contribution is the support of face
 integrals through a new class \texttt{FEFaceEvaluation}. The new class has a
@@
 the cell function. These kernels for the matrix-free operator evaluation are
-now collected in the new function \texttt{MatrixFree::loop}. The data
+now invoked by the new function \texttt{MatrixFree::loop}. The data
 structures have been particularly tuned for typical discontinuous Galerkin
@@
 between nodal points and quadrature, but not used for general bases. A more
-detailed description of the matrix-free modules is given in the preprint
-\cite{KronbichlerKormann2017}.
+detailed description of the matrix-free components and their performance
+characteristics is given in the preprint \cite{KronbichlerKormann2017}.
 
@@
 
-\marginpar{LH, MK, and others: Please add summaries of the new tutorial programs.}
+Two new tutorial programs were added to this release of \dealii{}. The
+tutorial program \texttt{step-59} presents a matrix-free solver for the
+Poisson equation discretized with the interior penalty discontinuous Galerkin
+method. The implementation is based on the new matrix-free functions described
+in Subsection~\ref{sec:matrixfree}. The new class
+\texttt{TensorProductMatrixSymmetricSum} is used to construct a block-Jacobi
+smoother in a geometric multigrid preconditioner, based on the definition of
+the inverse of a tensor product matrix through tensor products. As compared to
+a matrix-based block-Jacobi method, this approach considerably reduces the
+memory access and also the complexity at high polynomial degrees, going from
+$\mathcal O(k^{2d})$ arithmetic operations per cell for degree $k$ in $d$
+dimensions to only $\mathcal O(k^{d+1})$. The new tutorial demonstrates the
+outstanding performance of \dealii{}'s matrix-free module, solving the Poisson
+equation with 191 million degrees of freedom at degree $k=8$ on a workstation
+with 12 cores in about a minute.
+
+\marginpar{LH and others: Please add summaries of the new tutorial program.}
+The \texttt{step-60} tutorial program shows \ldots
 

### 2018-05-11 9.0/paper.tex
Added sentence about step-60

@@
 
-\marginpar{LH and others: Please add summaries of the new tutorial program.}
-The \texttt{step-60} tutorial program shows \ldots
+The tutorial program \texttt{step-60} shows how to perform computations on
+non-matching grids, and it presents advanced manipulation of
+\texttt{ParameterHandler} objects, exploiting the new \texttt{ParameterAccessor}
+and \texttt{ParameterAccessorProxy} classes. \texttt{step-60} solves a Poisson
+problem on a domain $\Omega$, subject to equality constraints defined on an
+embedded domain $\Gamma$. The embedded domain can be of co-dimension one or
+co-dimension zero, and its definition is independent with respect to $\Omega$.
+In order to enforce correctly the constraints, a non-matching coupling matrix
+needs to be constructed. This is achieved using the new
+\texttt{NonMatching::create\_coupling\_sparsity\_pattern} and
+\texttt{NonMatching::create\_coupling\_mass\_matrix}, that exploit some new
+functionality in the \texttt{GridTools} namespace.
 

### 2018-05-13 9.0/paper.tex
Fix a typo

@@
 Furthermore, the use of Sacado is now  demonstrated in a much more simplified
-and transparent manner in an modernized version of an existing ``code gallery''
+and transparent manner in a modernized version of an existing ``code gallery''
 example \cite{Pelteret2016a}.

### 2018-05-13 9.0/paper.tex
Add release date.

@@
 
-\dealii version 9.0.0 was released April X, 2018.
-\marginpar{Update the date}
+\dealii version 9.0.0 was released May 11, 2018.
 This paper provides an
@@
 variety of funding sources:
-\marginpar{all: update your funding sources}
 

### 2018-05-14 9.0/paper.tex
Some last edits.

@@
 
-\item New caching mechanism for expensive grid computations: we introduced a new
+\item A new caching mechanism for expensive grid computations: this version introduces a new
   class \texttt{GridTools::Cache} that caches computationally intensive
@@
 computed only once, and
-then cached inside this class for faster access whenever the triangulation has
+cached inside this class for faster access whenever the triangulation has
 not changed. The cache is marked for update by the \texttt{Triangulation} itself
-using signals.
+using signals so that data is properly invalidated upon mesh
+refinement and coarsening.
 
@@
 deal.II.
-This has changed with release 9.0 where support is given for differentiation
+This has changed with release 9.0 where support for differentiation is
+now available
 using a selection of ``white-listed'' libraries (namely ADOL-C \citep{Griewank1996a}
@@
 In practice, this support means that these ADOL-C and Sacado data
-types can be used in the \texttt{FEValues}, \texttt{FEValuesViews},
+types can be used as the underlying ``scalar'' in the \texttt{FEValues}, \texttt{FEValuesViews},
 \texttt{Tensor}, \texttt{SymmetricTensor},
@@
 that presents the AD compatibility and capabilities of the \dealii libraries.
-Furthermore, the use of Sacado is now  demonstrated in a much more simplified
+Furthermore, the use of Sacado is demonstrated in a much more simplified
 and transparent manner in a modernized version of an existing ``code gallery''
@@
   integration. IDA is a package for the solution of differential-algebraic
-  equations systems in the form $F(t,y,y')=0$. KINSOL is solver for nonlinear
+  equations systems in the form $F(t,y,y')=0$. KINSOL is a solver for nonlinear
   algebraic systems.
@@
 now return either a \texttt{std::shared\_ptr} or a \texttt{std::unique\_ptr},
-thereby clarifying object ownership responsibilities and avoiding memory leaks.
+thereby clarifying object ownership and avoiding memory leaks.
 
@@
 matrix-free applications. For matrix-based applications, we rely on
-cuSPARSE\cite{cusparse} and cuSOLVER\cite{cusolver} for operations on sparse
+cuSPARSE \cite{cusparse} and cuSOLVER \cite{cusolver} for operations on sparse
 matrices such as matrix-vector multiplication and for direct solvers. We have
@@
 also have added wrappers for Cholesky and LU factorizations provided by
-cuSOLVER. In practice, a user would assemble the matrix associated to the system
-on the host and then move the matrix to the device. At this point, the system
+cuSOLVER. In practice, a user would assemble the system matrix and
+right hand side vector
+on the host and then move them to the device. At this point, the
+linear system
 would be solved on the device and the solution would be moved back to the host.
@@
 We also have some support for matrix-free computation on a GPU. For now, the
-evaluation of the operator is limited to mesh without hanging-nodes.
+evaluation of the operator is limited to meshes without hanging nodes.
 
@@
 non-matching grids, and it presents advanced manipulation of
-\texttt{ParameterHandler} objects, exploiting the new \texttt{ParameterAccessor}
+\texttt{ParameterHandler} objects using the new \texttt{ParameterAccessor}
 and \texttt{ParameterAccessorProxy} classes. \texttt{step-60} solves a Poisson
@@
 \texttt{NonMatching::create\_coupling\_sparsity\_pattern} and
-\texttt{NonMatching::create\_coupling\_mass\_matrix}, that exploit some new
+\texttt{NonMatching::create\_coupling\_mass\_matrix} functions that exploit new
 functionality in the \texttt{GridTools} namespace.

### 2018-05-14 9.0/paper.tex
Rewrap some text.
The previous version spilled into the margins.

@@
 
-\item Improved support for user-defined run-time parameters: a new \texttt{ParameterAcceptor}
-  class has been added to the library. The class is intended to be used as a
-  base for any class that wants to handle parameters using the
-  \texttt{ParameterHandler} class. If you derive all your classes from
-  \texttt{ParameterAcceptor}, and declare your parameters either with
-  \texttt{parse/declare\_parameters} methods or via the
-  \texttt{ParameterAcceptor::add\_parameter} method, then both the
-  declaration and the
-  parsing of parameter files are automatically managed by
-  \texttt{ParameterAcceptor::initialize}, greatly simplifying
-  dealing with parameters in user codes.
+\item Improved support for user-defined run-time parameters: A new class
+  \texttt{ParameterAcceptor} has been added to the library. Users should write
+  classes inheriting from that class to manage parameters stored by a
+  \texttt{ParameterHandler}. If the managing class is derived from
+  \texttt{ParameterAcceptor} and all parameters are declared by either
+  \texttt{parse\_parameters} and \texttt{declare\_parameters} or
+  \texttt{ParameterAcceptor::add\_parameter}, then both the declaration and
+  parsing of parameter files will be (instead of using ad-hoc calls to
+  \texttt{ParameterHandler} methods) automatically managed by the
+  \texttt{ParameterAcceptor::initialize} method, which greatly simplifies
+  parameter management in user codes.
 

### 2018-05-16 9.0/paper.tex
Modified cache section: added compute point location information

@@
 
-Some of the methods in \texttt{GridTools} already use this cache to
-speed up repeated calls to the same expensive methods.
+Some of the methods in \texttt{GridTools} and \texttt{FEFieldFunction} already
+use this cache to avoid repeated calls to the same expensive methods.
+This results in important speed-ups for computationally heavy methods like
+\texttt{GridTools::compute\_point\_locations}.
+
 \item A new \texttt{MeshWorker::mesh\_loop} function has been added that

### 2018-05-16 9.0/paper.tex
Added item on distributed ctploc and bboxes

--- 2018-05-14 9.0/paper.tex
@@
   forcing the users to adhere to a specific interface.
+
+\item A new method \texttt{GridTools::distributed\_compute\_point\_locations} to
+  perform the same tasks of \texttt{GridTools::compute\_point\_locations} with
+  arbitrary points on a distributed mesh has been added. The current implementation
+  uses vectors of \texttt{BoundingBoxes} to manage the communication to other
+  processes of points which cannot be computed locally.
 

### 2018-05-16 9.0/paper.tex
Fix links.
We use '8_5_0' and '9_0_0' in the URLs.

--- 2018-05-14 9.0/paper.tex
@@
 can be found
-\href{https://www.dealii.org/developer/doxygen/deal.II/changes_between_8_5_and_9_0.html}{
+\href{https://dealii.org/developer/doxygen/deal.II/changes_between_8_5_0_and_9_0_0.html}{
 in the file that lists all changes for this release}, see \cite{changes90}.
@@
 The 9.0 release includes
-\href{https://www.dealii.org/developer/doxygen/deal.II/changes_between_8_5_and_9_0.html}
+\href{https://dealii.org/developer/doxygen/deal.II/changes_between_8_5_0_and_9_0_0.html}
      {around 75 incompatible changes}; see \cite{changes90}. The majority of these changes

### 2018-05-20 9.0/paper.tex
fix my affiliation

@@
 
-T.~Heister was partially supported by the Computational Infrastructure in
-Geodynamics initiative (CIG), through the National Science Foundation (NSF)
-under Award No. EAR-0949446 and EAR-1550901 and The University of
-California -- Davis, and NSF Grant DMS-1522191.
+T.~Heister was partially supported by NSF Award DMS-1522191, by the
+Computational Infrastructure in Geodynamics initiative (CIG), through the NSF
+under Award EAR-0949446 and EAR-1550901 and The University of California --
+Davis, and by Technical Data Analysis, Inc. through US Navy SBIR N16A-T003.
 

### 2019-05-08 9.1/paper.tex
Add first version of release paper

--- 2018-05-20 9.0/paper.tex
@@
   pdfauthor={
-    Giovanni Alzetta,
     Daniel Arndt,
     Wolfgang Bangerth,
-    Vishal Boddu,
-    Benjamin Brands,
     Denis Davydov,
-    Rene Gassm\"{o}ller,
     Timo Heister,
     Luca Heltai,
-    Katharina Kormann,
     Martin Kronbichler,
@@
   },
-  pdftitle={The deal.II Library, Version 9.0, 2018},
+  pdftitle={The deal.II Library, Version 9.1, 2019},
 }
 
-\title{The \dealii Library, Version 9.0}
+\title{The \dealii Library, Version 9.1}
 
-\author[1]{Giovanni Alzetta}
-\affil[1]{SISSA,
-  International School for Advanced Studies,
-  Via Bonomea 265,
-  34136, Trieste, Italy.
-{\texttt{\{galzetta,luca.heltai\}@sissa.it}}}
-
-\author[2]{Daniel Arndt}
-\affil[2]{Interdisciplinary Center for Scientific Computing,
+\author[1]{Daniel Arndt}
+\affil[1]{Interdisciplinary Center for Scientific Computing,
   Heidelberg University,
@@
 
-\author[3]{Wolfgang Bangerth}
-\affil[3]{Department of Mathematics, Colorado State University, Fort
+\author[2]{Wolfgang Bangerth}
+\affil[2]{Department of Mathematics, Colorado State University, Fort
   Collins, CO 80523-1874, USA.
@@
 
-\author[4]{Vishal Boddu}
-\affil[4]{Chair of Applied Mechanics,
+\author[3]{Denis Davydov}
+\affil[3]{Chair of Applied Mechanics,
   Friedrich-Alexander-Universit\"{a}t Erlangen-N\"{u}rnberg,
@@
   91058 Erlangen, Germany.
-  {\texttt{\{vishal.boddu,benjamin.brands,denis.davydov,jean-paul.pelteret\}@fau.de}}}
+  {\texttt{\{denis.davydov,jean-paul.pelteret\}@fau.de}}}
 
-\author[4]{Benjamin Brands}
-
-\author[4]{Denis Davydov}
-
-\author[5]{Rene Gassm\"{o}ller}
-\affil[5]{Department of Earth and Planetary Sciences,
-  University of California Davis,
-  One Shields Avenue,
-  CA-95616 Davis, USA.
-  {\texttt{rgassmoeller@ucdavis.edu}}}
-
-\author[6]{Timo Heister}
-\affil[6]{Mathematical Sciences,
+\author[4]{Timo Heister}
+\affil[4]{Mathematical Sciences,
   O-110 Martin Hall,
@@
 %
-\author[1]{Luca Heltai}
+\author[5]{Luca Heltai}
+\affil[5]{SISSA,
+  International School for Advanced Studies,
+  Via Bonomea 265,
+  34136, Trieste, Italy.
+{\texttt{luca.heltai@sissa.it}}}
 
-\author[7]{Katharina Kormann}
-\affil[7]{Max Planck Institute for Plasma Physics,
-Boltzmannstr.~2, 85748 Garching, Germany.
-{\texttt{katharina.kormann@ipp.mpg.de}}}
-
-\author[8]{Martin Kronbichler}
-\affil[8]{Institute for Computational Mechanics,
+\author[6]{Martin Kronbichler}
+\affil[6]{Institute for Computational Mechanics,
   Technical University of Munich,
@@
 
-\author[9]{Matthias Maier}
-\affil[9]{School of Mathematics,
+\author[7]{Matthias Maier}
+\affil[7]{School of Mathematics,
   University of Minnesota,
@@
 
- \author[10]{Bruno Turcksin\footnote{
+ \author[8]{Bruno Turcksin\footnote{
    This manuscript has been authored by UT-Battelle, LLC under Contract No.
@@
    DOE Public Access Plan (http://energy.gov/downloads/doe-public-access-plan).}}
- \affil[10]{Computational Engineering and Energy Sciences Group,
+ \affil[8]{Computational Engineering and Energy Sciences Group,
    Computional Sciences and Engineering Division,
@@
 
-\author[11]{David Wells}
-\affil[11]{Department of Mathematical Sciences, Rensselaer Polytechnic
+\author[9]{David Wells}
+\affil[9]{Department of Mathematical Sciences, Rensselaer Polytechnic
 Institute, Troy, NY 12180, USA.
@@
   This paper provides an overview of the new features of the finite element
-  library \dealii version 9.0.
+  library \dealii version 9.1.
 \end{abstract}
@@
 
-\dealii version 9.0.0 was released May 11, 2018.
+\dealii version 9.1.0 was released May XX, 2019
 This paper provides an
 overview of the new features of this major release and serves as a citable
-reference for the \dealii software library version 9.0. \dealii is an
+reference for the \dealii software library version 9.1. \dealii is an
 object-oriented finite element library used around the world in the
@@
 \begin{itemize}
-\item Improved support for curved geometries;
-\item Support for particle-in-cell methods;
-\item Dedicated support for automatic differentiation;
-\item Interfaces to more external libraries and programs;
-\item C++11 is now both required and used;
-\item Support for GPU computations;
-\item Support for face integrals and significant improvements of the matrix-free framework;
-\item Two new tutorial programs step-59 and step-60.
+\item Improved support for automatic and symbolic differentiation;
+\item Support for $hp$ adaptivity on MPI-parallel programs with \texttt{parallel::distributed::Triangulation}
+\item Four new tutorial programs step-61, step-62, step-63, step-64.
 \end{itemize}
@@
 \begin{itemize}
-\item \dealii has made extensive use of both the Clang-Tidy \cite{clang-tidy}
-  and Coverity Scan \cite{coverity} static analysis tools for detecting bugs
-  and other issues in the code. For example, around 260 issues were detected and
-  fixed using the latter tool.
-
-\item
-  \texttt{LinearOperator}, a flexible template class that implements the action of a
-  linear operator (see \cite{MaierBardelloniHeltai-2016-b}), now supports
-  computations with Trilinos, Schur complements, and linear constraints. This
-  class is, as of this release, the official replacement for about half a dozen
-  similar (but less general) classes, such as \texttt{FilteredMatrix},
-  \texttt{IterativeInverse}, and \texttt{PointerMatrix}.
-
-\item New non-standard quadrature rules:
-A number of non-standard, special-purpose quadrature rules have been
-implemented. Among these are ones for
-(i) truncating standard formulas to simplical domains (\texttt{QSimplex});
-(ii) singular transformations of the unit cell to the unit simplex
-    (\texttt{QDuffy});
-(iii) composition of simplical quadrature rules to a combined rule on the
-    unit cell (\texttt{QSplit});
-and (iv) transformation of the unit square to polar coordinates
-    (\texttt{QTrianglePolar}).
-These quadrature rules greatly help when integrating singular
-functions or on singular domains. They
-are mainly used in Boundary Element Methods.
-
-\item
-  Support for complex-valued vectors at the same level as real-valued
-  vectors.
-
-\item A new python tutorial program tutorial-1; as well as
-  updates to step-37. In addition, the separate code
-  gallery of \dealii has gained a number of new entries.
-
-\item Improved support for user-defined run-time parameters: A new class
-  \texttt{ParameterAcceptor} has been added to the library. Users should write
-  classes inheriting from that class to manage parameters stored by a
-  \texttt{ParameterHandler}. If the managing class is derived from
-  \texttt{ParameterAcceptor} and all parameters are declared by either
-  \texttt{parse\_parameters} and \texttt{declare\_parameters} or
-  \texttt{ParameterAcceptor::add\_parameter}, then both the declaration and
-  parsing of parameter files will be (instead of using ad-hoc calls to
-  \texttt{ParameterHandler} methods) automatically managed by the
-  \texttt{ParameterAcceptor::initialize} method, which greatly simplifies
-  parameter management in user codes.
-
-\item A new caching mechanism for expensive grid computations: this version introduces a new
-  class \texttt{GridTools::Cache} that caches computationally intensive
-  information about a \texttt{Triangulation}.
-
-This class allows the user to query some of the data structures constructed
-using functions in the \texttt{GridTools} namespace. This data is then
-computed only once, and
-cached inside this class for faster access whenever the triangulation has
-not changed. The cache is marked for update by the \texttt{Triangulation} itself
-using signals so that data is properly invalidated upon mesh
-refinement and coarsening.
-
-Some of the methods in \texttt{GridTools} and \texttt{FEFieldFunction} already
-use this cache to avoid repeated calls to the same expensive methods.
-This results in important speed-ups for computationally heavy methods like
-\texttt{GridTools::compute\_point\_locations}.
-
-\item A new \texttt{MeshWorker::mesh\_loop} function has been added that
-  performs the same tasks of the \texttt{MeshWorker::loop} function without
-  forcing the users to adhere to a specific interface.
-
-\item A new method \texttt{GridTools::distributed\_compute\_point\_locations} to
-  perform the same tasks of \texttt{GridTools::compute\_point\_locations} with
-  arbitrary points on a distributed mesh has been added. The current implementation
-  uses vectors of \texttt{BoundingBoxes} to manage the communication to other
-  processes of points which cannot be computed locally.
-
+\item ParsedConvergenceTable
+\item FE\_BernardiRaugel
+\item ConstraintMatrix class has been renamed to AffineConstraints
+\item FE\_NedelecSZ
+\item A number of bug fixes and performance improvements to the matrix-free framework and related geometric multigrid solvers were added.
+In particular, the implementation of the Chebyshev iteration, an often used smoother in the matrix-free context, has been revised to reduce the vector access.
+The performance improvement speed up matrix-free multigrid solvers by up to 10-15\% on affine geometries and up to 5\% on curved geometries.
 \end{itemize}
-Beyond these changes, the changelog lists more than 330 other features and bugfixes.
+Beyond these changes, the changelog lists more than 200 other features and bugfixes.
 
@@
 can be found
-\href{https://dealii.org/developer/doxygen/deal.II/changes_between_8_5_0_and_9_0_0.html}{
-in the file that lists all changes for this release}, see \cite{changes90}.
+\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_0_1_and_9_1_0.html}{
+in the file that lists all changes for this release}, see \cite{changes91}.
 (The file is also linked to from the web site of each release as well as
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Improved support for curved geometries}
-\label{sec:manifolds}
-
-\dealii{} has had the ability to attach \textit{manifold descriptions} to all
-parts of a geometry since the 8.2 release. These descriptions are used to place
-new vertices during mesh refinement, to determine the mapping between the
-reference and real cells, and in a number of other contexts. These classes,
-inheriting from \texttt{Manifold}, describe coordinate transformations in a
-general way and completely supersede the older classes inheriting from
-\texttt{Boundary}. However, for historical reasons, manifold descriptors have
-used some of the same code paths as boundary indicators, which were only
-intended for marking what parts of the boundary correspond to what boundary
-conditions. Put another way: under certain circumstances a boundary indicator
-was also interpreted as a manifold indicator.
-
-The current release severs this connection: Boundary indicators and manifold
-descriptions are now entirely separated. This means that \texttt{boundary\_id}s
-are only used to set boundary conditions and \texttt{manifold\_id}s are only
-used to set geometry descriptions. The old compatibility code for using boundary
-indicators as manifold indicators has been removed and all usages of the
-old-style \texttt{Boundary} objects (even with manifold ids) are now deprecated.
-
-There are also numerous improvements to the available manifold descriptions.
-First, the manifold smoothing algorithms applied in the \texttt{Triangulation}
-class and \texttt{MappingQGeneric} have been changed from the old Laplace-style
-smoothing to a transfinite interpolation that linearly blends between the
-descriptions on the faces around a cell. The old transformation introduced
-boundary layers inside cells that prevented convergence rates from exceeding
-\(3.5\) in the global \(L^2\) errors for typical settings. This change also
-considerably improves mesh quality in situations where curved descriptions are
-only applied to the boundary rather than the whole volume. This concept was also
-introduced as a new manifold class \texttt{TransfiniteInterpolationManifold},
-which allows to apply this type of smoothing not only in the cells close to the
-boundary but over a full coarse (level 0) cell.
-
-Finally, every function in the \texttt{GridGenerator} namespace now attaches a
-default manifold to the curved parts of the domain described by the generated
-mesh, and sets reasonable defaults for manifold indicators both in the domain
-and on the boundary.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Support for particle-in-cell methods}
-
-While \dealii is a package intended to solve problems with the
-finite element method -- i.e., using continuous or discontinuous
-\textit{fields} --, it is often convenient in fluid dynamics problems to
-couple the continuum description of phenomena with particles. These
-particles, advected along with the numerical approximation of the flow
-field, are then either used to visualize properties of the flow, or to
-advect material properties such as the viscosity of inhomogeneous
-mixtures of fluids. If each particle is associated with the cells
-of a mesh, these methods are often referred to as particle-in-cell (PIC).
-
-\dealii now has a dedicated particles module. The module provides a base class
-\texttt{Particle} that represents a particle with position, an ID number and
-a variable number of properties. They are jointly represented by a \texttt{ParticleHandler}
-class that manages the storage and handling of all particles. In
-parallel simulations, this class also
-distributes the particles among the subdomains of the parallel process
-and supports efficient data transfer during mesh refinement and
-checkpoint/restart phases.
-
-A much more detailed view of the underlying algorithms can be found in
-\cite{GLHPB18}. A longer report is at \cite{GHPB16}. The
-implementation here originated in the \aspect code, see \cite{KHB12,HDGB17}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Dedicated support for automatic differentiation}
-
-Automatic differentiation (AD) is often used to automatically derive
-residuals and their linearization from a stored energy functional, and to
-derive Jacobian matrices from residual vectors for simulations that use
-complicated material models. Examples of its application can be found widely
-within nonlinear solid mechanics, coupled multiphysics problems, as well as
-for nonlinear viscosity models in fluid flow.
-
-\dealii has had a tutorial program (step-33) since 2007 that
-demonstrates this technique based on the Trilinos Sacado \cite{Bartlett2006a} package, but
-the functionality was not available pervasively throughout
-deal.II.
-This has changed with release 9.0 where support for differentiation is
-now available
-using a selection of ``white-listed'' libraries (namely ADOL-C \citep{Griewank1996a}
-and Sacado) and a subset of their supported number types. Currently,
-we offer support for the following cases:
-\begin{itemize}
-\item ADOL-C taped (n-differentiable),
-\item ADOL-C tapeless (once differentiable),
-\item Sacado dynamic forward (once differentiable),
-\item Sacado reverse (once differentiable),
-\item Sacado nested dynamic forward (twice differentiable), and
-\item Sacado nested reverse and dynamic forward (twice differentiable).
-\end{itemize}
-In practice, this support means that these ADOL-C and Sacado data
-types can be used as the underlying ``scalar'' in the \texttt{FEValues}, \texttt{FEValuesViews},
-\texttt{Tensor}, \texttt{SymmetricTensor},
-and related classes that are generally used to assemble linear systems
-and right hand sides.
-Given the updated capabilities of the library, there is now a dedicated module
-that presents the AD compatibility and capabilities of the \dealii libraries.
-Furthermore, the use of Sacado is demonstrated in a much more simplified
-and transparent manner in a modernized version of an existing ``code gallery''
-example \cite{Pelteret2016a}.
-
-To date it remains necessary for the user  to manage the initialization of AD
-independent variables and the resultant calls  to the AD dependent variables
-in order to initiate the computation of derivatives.
-In the next release we expect to provide a unified interface to these
-AD libraries, which will hide these library-dependent implementational details
-and facilitate switching between the supported libraries and AD number types
-based on the user's requirements.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{New interfaces to external libraries and programs}
-
-\dealii has always tried to leverage high-quality implementations of
-algorithms available through other open source software, rather than
-re-implementing their functionality. (A list of interfaces to other
-packages is given in Section~\ref{sec:cite}.) As part of the current
-release, we have written several new interfaces as discussed in the following.
-
-\paragraph*{Assimp, the Open Asset Import Library.}
-  Assimp \cite{assimp} can be used to read about 40 different 3D
-  graphics formats. A subset of these formats can be now be read from
-  within \dealii to generate two-dimensional meshes, possibly embedded in
-  a three-dimensional space.
-
-% \paragraph*{Gmsh.}
-%   \marginpar{Do we actually provide any functionality with the GMSH
-%   executable yet? If so, complete this part}
-%   Gmsh \cite{geuzaine2009gmsh}
-% The only thing we currently do is to define DEAL_II_GMSH_EXECUTABLE_PATH
-
-\paragraph*{nanoflann, a library for building and querying
-  $k$-d trees of datasets.} Operations such as finding the vertex or cell
-  closest to a given evaluation point occur frequently in many applications
-  that use unstructured meshes. While the naive algorithm is linear in the
-  number of vertices or cells, many such operations can be made significantly
-  faster by building a $k$-d tree data structure that recursively subdivides
-  a $k$ dimensional space. The nanoflann library \cite{nanoflann} provides
-  such a data structure and allows querying it, either for closest points
-  (e.g., when finding the closest vertex) or for searching the points that
-  fall within a radius of a target point. This functionality is now available
-  via \dealii interfaces.
-
-\paragraph*{ROL, a Rapid Optimization Library.}
-  ROL \cite{ridzal2014rapid} is a package for large-scale optimization.
-  \dealii can now use the state-of-the-art algorithms in ROL to solve
-  unconstrained and constrained optimization problems as well as optimization
-  problems under uncertainty.
-  \dealii provides an interface to ROL's (abstract) vector class using the
-  adapter software pattern.
-  Through such an interface any vector class in \dealii following certain
-  interface requirements can be used to define a ROL objective function.
-
-\paragraph*{ScaLAPACK, a parallel dense linear algebra library for distributed memory machines.}
-  ScaLAPACK \cite{slug} provides block-cyclic matrix distribution over 2D
-  process grids. The functionality and interface of our wrappers is similar
-  to the LAPACK \cite{lapack} wrappers for serial dense linear algebra, namely
-  matrix-matrix multiplication, Cholesky and LU factorizations,
-  eigensolvers, SVD, least squares, pseudoinverses and save/load operations using
-  either serial or parallel HDF5 \cite{hdf5}. All of this functionality is available
-  even in cases where the number of MPI processes does not match the numbers of
-  processes in the 2D process grid used to distribute a matrix.
-
-  As part of this effort, we have also improved LAPACK support: there
-  are now methods to perform rank-1 updates/downdates, Cholesky
-  factorizations, to compute the trace and determinant, as well as
-  estimate the reciprocal condition number. We also now support
-  configuration with 64-bit BLAS.
-
-\paragraph*{SUNDIALS, a SUite of Nonlinear and DIfferential/ALgebraic
-  Equation Solvers.}
-  Solving nonlinear algebraic and differential equations is both a
-  common task and one that often requires sophisticated globalization
-  algorithms for efficiency and reliability. SUNDIALS \cite{sundials}
-  provides these in a widely used format, both sequentially and in parallel.
-
-  \dealii now has interfaces to SUNDIALS's ARKode, IDA, and KINSOL sub-packages.
-  ARKode is a solver library that provides adaptive-step time
-  integration. IDA is a package for the solution of differential-algebraic
-  equations systems in the form $F(t,y,y')=0$. KINSOL is a solver for nonlinear
-  algebraic systems.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-  \subsection{Use of C++11}
-  \label{sec:cxx11}
-\dealii first offered support for a subset of C++11 features in
-version 6.2, released in 2009. The current release is the first to
-\emph{require} a C++11 compiler.
-
-Many parts of the code base have been rewritten to both support and
-use the new features of C++11. In particular, \dealii now makes
-extensive use of move semantics as well as range-based \texttt{for}
-loops with \texttt{auto} type deduction of iterator variables. We have
-also largely replaced \texttt{push\_back()} by
-\texttt{emplace\_back()} when adding elements to collections more
-efficiently.
-
-Finally, we have changed the entire code base to avoid using raw
-pointers and instead use \texttt{std::unique\_ptr} and
-\texttt{std::shared\_ptr} where possible to make memory management
-more reliable. These changes include some minor incompatibilities: all
-\texttt{clone()} functions (such as \texttt{FiniteElement::clone()} and
-\texttt{Mapping::clone()}) now return \texttt{std::unique\_ptr}s instead of
-C-style raw pointers. Indeed, nearly all interfaces throughout the library that return a pointer
-now return either a \texttt{std::shared\_ptr} or a \texttt{std::unique\_ptr},
-thereby clarifying object ownership and avoiding memory leaks.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Support for GPU computations}
-
-Heterogeneous computing is becoming more prevalent in supercomputing and this is
-a trend that is expected to continue in the future. In particular, the use of
-GPUs has been increasing during the last few years.
-
-This release of \dealii adds support for GPUs both for matrix-based and
-matrix-free applications. For matrix-based applications, we rely on
-cuSPARSE \cite{cusparse} and cuSOLVER \cite{cusolver} for operations on sparse
-matrices such as matrix-vector multiplication and for direct solvers. We have
-introduced a new type of sparse matrix, \texttt{CUDAWrappers::SparseMatrix},
-which moves onto the device a \dealii \texttt{SparseMatrix} and changes the
-format of the underlying data to the appropriate CSR format used by cuSPARSE. We
-also have added wrappers for Cholesky and LU factorizations provided by
-cuSOLVER. In practice, a user would assemble the system matrix and
-right hand side vector
-on the host and then move them to the device. At this point, the
-linear system
-would be solved on the device and the solution would be moved back to the host.
-
-We also have some support for matrix-free computation on a GPU. For now, the
-evaluation of the operator is limited to meshes without hanging nodes.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Extended matrix-free capabilities}\label{sec:matrixfree}
-
-The matrix-free infrastructure in \dealii{} was significantly overhauled for
-this release. The major new contribution is the support of face
-integrals through a new class \texttt{FEFaceEvaluation}. The new class has a
-similar interface as the existing \texttt{FEEvaluation} class, and applies SIMD
-vectorization over several faces in analogy to the intra-cell vectorization in
-FEEvaluation. Discontinuous Galerkin operators are implemented defining two
-face functions, one for interior and one for boundary faces, in addition to
-the cell function. These kernels for the matrix-free operator evaluation are
-now invoked by the new function \texttt{MatrixFree::loop}. The data
-structures have been particularly tuned for typical discontinuous Galerkin
-setups involving operators with first and second spatial derivatives. Both
-data access and computations have been thoroughly optimized and compared to
-the performance boundaries of the hardware. Furthermore, the support for
-AVX-512 instructions in the matrix-free framework was extended, adding new
-\texttt{gather} and \texttt{scatter} intrinsics for the indirect access to
-vector entries where appropriate.
-
-To give an example of the algorithmic improvements, the computation of the
-values and gradients on all quadrature points for cell integrals has been
-significantly improved, yielding 10--20\% better performance for cases where
-the kernels are compute bound. For the example of the reference cell gradient
-of a solution field $\mathbf u$ in three space dimensions, the new release
-applies the following change:
-\begin{equation*}
-  \text{previous: }
-  \begin{bmatrix}
-    D_{1} \otimes S_{2} \otimes S_{3}\\
-    S_{1} \otimes D_{2} \otimes S_{3}\\
-    S_{1} \otimes S_{2} \otimes D_{3}
-  \end{bmatrix}
-  \mathbf u
-  \quad
-  \leadsto
-  \quad
-  \text{new: }
-  \begin{bmatrix}
-    D_{1}^{\mathrm{co}} \otimes I_{2} \otimes I_{3}\\
-    I_{1} \otimes D_{2}^{\mathrm{co}} \otimes I_{3}\\
-    I_{1} \otimes I_{2} \otimes D_{3}^{\mathrm{co}}
-  \end{bmatrix}
-  \begin{bmatrix}
-    S_{1} \otimes S_{2} \otimes S_{3}
-  \end{bmatrix}
-  \mathbf u.
-\end{equation*}
-The matrices $S_i$ contain the values of the one-dimensional shape functions
-in one-dimensional quadrature points and $D_i$ their derivatives. When applied
-with the usual sum factorization implementation described, for
-example, in
-\cite{KronbichlerKormann2012}, the old kernels amounted to 9 partial
-summations -- or rather 8 in the previous implementation of \dealii{} because
-the application of $S_1$ for the $y$ and $z$ components of the gradient can be
-merged. The new code performs a basis transformation to a related basis with
-derivative matrix $D_i = D_i^{\mathrm{co}} S_i$, which is the basis of
-Lagrange polynomials in the points of the quadrature. This change reduces the
-number of partial sums to only 6 for the gradient, as the action of the unit
-matrices $I_i$ needs not be implemented. In isolation, this spectral
-element-like evaluation was previously available in \dealii{} for collocation
-between nodal points and quadrature, but not used for general bases. A more
-detailed description of the matrix-free components and their performance
-characteristics is given in the preprint \cite{KronbichlerKormann2017}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Tutorial and code gallery programs}
-
-Two new tutorial programs were added to this release of \dealii{}. The
-tutorial program \texttt{step-59} presents a matrix-free solver for the
-Poisson equation discretized with the interior penalty discontinuous Galerkin
-method. The implementation is based on the new matrix-free functions described
-in Subsection~\ref{sec:matrixfree}. The new class
-\texttt{TensorProductMatrixSymmetricSum} is used to construct a block-Jacobi
-smoother in a geometric multigrid preconditioner, based on the definition of
-the inverse of a tensor product matrix through tensor products. As compared to
-a matrix-based block-Jacobi method, this approach considerably reduces the
-memory access and also the complexity at high polynomial degrees, going from
-$\mathcal O(k^{2d})$ arithmetic operations per cell for degree $k$ in $d$
-dimensions to only $\mathcal O(k^{d+1})$. The new tutorial demonstrates the
-outstanding performance of \dealii{}'s matrix-free module, solving the Poisson
-equation with 191 million degrees of freedom at degree $k=8$ on a workstation
-with 12 cores in about a minute.
-
-The tutorial program \texttt{step-60} shows how to perform computations on
-non-matching grids, and it presents advanced manipulation of
-\texttt{ParameterHandler} objects using the new \texttt{ParameterAccessor}
-and \texttt{ParameterAccessorProxy} classes. \texttt{step-60} solves a Poisson
-problem on a domain $\Omega$, subject to equality constraints defined on an
-embedded domain $\Gamma$. The embedded domain can be of co-dimension one or
-co-dimension zero, and its definition is independent with respect to $\Omega$.
-In order to enforce correctly the constraints, a non-matching coupling matrix
-needs to be constructed. This is achieved using the new
-\texttt{NonMatching::create\_coupling\_sparsity\_pattern} and
-\texttt{NonMatching::create\_coupling\_mass\_matrix} functions that exploit new
-functionality in the \texttt{GridTools} namespace.
-
-\dealii has a separate ``code gallery'' that
-consists of programs shared by users as examples of what can be
-done with \dealii. While not part of the release process, it is nonetheless
-worth mentioning that the set of new programs since the last release covers
-the following topics:
-\begin{itemize}
-\item The multipoint flux mixed finite element method (MFMFE) applied
-  to the Darcy problem of porous media flow;
-\item A linearized active skeletal muscle model with application to
-  the simulation concentric contraction of the human biceps brachii;
-\item A parallel implementation of the Local Discontinuous Galerkin
-  (LDG) method applied to the Poisson equation.
-\end{itemize}
-With these additions, the code gallery now contains 10 different applications.
 
@@
 
-The 9.0 release includes
+The 9.1 release includes
 \href{https://dealii.org/developer/doxygen/deal.II/changes_between_8_5_0_and_9_0_0.html}
-     {around 75 incompatible changes}; see \cite{changes90}. The majority of these changes
+     {around 15 incompatible changes}; see \cite{changes91}. The majority of these changes
 should not be visible to typical user codes; some remove previously
@@
 interfaces that are not usually used in external
-applications. However, some are, such as changes to the interplay between meshes and
-manifolds, as well as the requirement to use a
-C++11 compiler (see Sections~\ref{sec:manifolds} and
-~\ref{sec:cxx11}). In addition, the following
-incompatible changes are worth mentioning:
+applications. However, some are worth mentioning:
 \begin{itemize}
-\item The \texttt{BlockDiagonalMatrix},
-  \texttt{InverseMatrixRichardson},
-  \texttt{IterativeInverse},
-  \texttt{ProductMatrix},
-  \texttt{ProductSparseMatrix},
-  \texttt{TransposeMatrix},
-  \texttt{ScaledMatrix},
-  \texttt{SchurMatrix},
-  \texttt{ShiftedMatrix}, and
-  \texttt{ShiftedMatrixGeneralized} classes have been removed. They are now
-  generalized through the \texttt{LinearOperator} concept. Several
-  other, similar classes have been deprecated.
-
-  \item The default partitioner for the
-    \texttt{parallel::shared::Triangulation} is now the Trilinos
-    package Zoltan. This functionality was previously provided by the
-    METIS partitioner, but the METIS package has not been actively maintained
-    for a long time, and moreover yields subdivisions that depend on
-    system details such as the random number generator and sorting
-    facilities provided by the operating system; consequently, the
-    partition is not consistent across platforms.
-
-  \item  The class \texttt{FE\_DGQHermite}  now uses a more stable,
-    ``Hermite-like'' polynomial basis. The change is highly
-    beneficial because it significantly improves the accuracy (in
-    terms of roundoff) for this basis and also reduces iteration
-    counts for some iterative solvers with simple preconditioners.
-
-  \item Many functions that previously returned a raw, C-style pointer
-    now return a \texttt{std::unique\_ptr} and
-    \texttt{std::shared\_ptr} where possible to make memory management
-    more reliable.
+\item TODO
 \end{itemize}
@@
 Older releases of \dealii can be cited as
-\cite{dealII80,dealII81,dealII82,dealII83,dealII84,dealII85}.
+\cite{dealII80,dealII81,dealII82,dealII83,dealII84,dealII85,dealII90}.
 
@@
 %
-% Updated 2018-04-19
-Julian Andrej,
-Rajat Arora,
-Lucas Campos,
-Praveen Chandrashekar,
-Jie Cheng,
-Emma Cinatl,
-Conrad Clevenger,
-Ester Comellas,
-Sambit Das,
-Giovanni Di Ilio,
-Nivesh Dommaraju,
-Marc Fehling,
-Niklas Fehn,
-Menno Fraters,
-Anian Fuchs,
-Daniel Garcia-Sanchez,
-Nicola Giuliani,
-Anne Glerum,
-Christoph Goering,
-Alexander Grayver,
-Samuel Imfeld,
-Daniel Jodlbauer,
-Guido Kanschat,
-Vishal Kenchan,
-Andreas Kergassner,
-Eldar Khattatov,
-Ingo Kligge,
-Uwe Köcher,
-Joachim Kopp,
-Ross Kynch,
-Konstantin Ladutenko,
-Tulio Ligneul,
-Karl Ljungkvist,
-Santiago Ospina,
-Alexey Ozeritsky,
-Dirk Peschka,
-Simon Puchert,
-E. G. Puckett,
-Lei Qiao,
-Ce Qin,
-Jonathan Robey,
-Alberto Sartori,
-Daniel Shapero,
-Ben Shields,
-Simon Sticko,
-Oliver Sutton,
-Zhuoran Wang,
-Xiaoyu Wei,
-Michał Wichrowski,
-Julius Witte,
-Feimi Yu,
-Weixiong Zheng.
 
@@
 
-D.~Arndt, K.~Kormann and M.~Kronbichler were partially supported by the German
+D.~Arndt and M.~Kronbichler were partially supported by the German
 Research Foundation (DFG) under the project ``High-order discontinuous
@@
 
-W.~Bangerth and R.~Gassm\"{o}ller were partially
+W.~Bangerth was partially
 supported by the National Science Foundation under award OCI-1148116
@@
 No.~EAR-0949446 and EAR-1550901 and The University of California -- Davis.
-
-V.~Boddu was supported by the German Research Foundation (DFG) under the
-research group project FOR 1509.
-
-B.~Brands was partially supported by the Indo-German exchange programm
-``Multiscale Modeling, Simulation and Optimization for Energy, Advanced
-Materials and Manufacturing'' (MMSO) funded by DAAD (Germany) and UGC
-(India).
 

### 2019-05-08 9.1/paper.tex
Update affiliation

@@
 
-\author[1]{Daniel Arndt}
-\affil[1]{Interdisciplinary Center for Scientific Computing,
-  Heidelberg University,
-  Im Neuenheimer Feld 205,
-  69120 Heidelberg, Germany.
-  {\texttt{daniel.arndt@iwr.uni-heidelberg.de}}}
+\author[1]{Daniel Arndt \footnote{
+  This manuscript has been authored by UT-Battelle, LLC under Contract No.
+  DE-AC05-00OR22725 with the U.S. Department of Energy. The United States
+  Government retains and the publisher, by accepting the article for
+  publication, acknowledges that the United States Government retains a
+  non-exclusive, paid-up, irrevocable, worldwide license to publish or reproduce
+  the published form of this manuscript, or allow others to do so, for United
+  States Government purposes. The Department of Energy will provide public
+  access to these results of federally sponsored research in accordance with the
+  DOE Public Access Plan (http://energy.gov/downloads/doe-public-access-plan).}}
+\affil[1]{Computational Engineering and Energy Sciences Group,
+  Computional Sciences and Engineering Division,
+  Oak Ridge National Laboratory, 1 Bethel Valley Rd.,
+  TN 37831, USA.
+  {\texttt{\{arndtd,turcksinbr\}@ornl.gov}}}
 
@@
 
- \author[8]{Bruno Turcksin\footnote{
-   This manuscript has been authored by UT-Battelle, LLC under Contract No.
-   DE-AC05-00OR22725 with the U.S. Department of Energy. The United States
-   Government retains and the publisher, by accepting the article for
-   publication, acknowledges that the United States Government retains a
-   non-exclusive, paid-up, irrevocable, worldwide license to publish or reproduce
-   the published form of this manuscript, or allow others to do so, for United
-   States Government purposes. The Department of Energy will provide public
-   access to these results of federally sponsored research in accordance with the
-   DOE Public Access Plan (http://energy.gov/downloads/doe-public-access-plan).}}
- \affil[8]{Computational Engineering and Energy Sciences Group,
-   Computional Sciences and Engineering Division,
-   Oak Ridge National Laboratory, 1 Bethel Valley Rd.,
-   TN 37831, USA.
-   {\texttt{turcksinbr@ornl.gov}}}
+\author[1]{Bruno Turcksin}
 

### 2019-05-10 9.1/paper.tex
Update source of funding for myself and Timo.

@@
 
-W.~Bangerth was partially
-supported by the National Science Foundation under award OCI-1148116
-as part of the Software Infrastructure for Sustained Innovation (SI2)
-program; and by the Computational Infrastructure in Geodynamics initiative
-(CIG), through the National Science Foundation under Awards
-No.~EAR-0949446 and EAR-1550901 and The University of California -- Davis.
+W.~Bangerth and T.~Heister were partially
+supported by the National Science Foundation under award OAC-1835673
+as part of the Cyberinfrastructure for Sustained Scientific Innovation (CSSI)
+program; by award DMS-1821210; and by the Computational Infrastructure
+in Geodynamics initiative (CIG), through the National Science
+Foundation under Awards No.~EAR-0949446 and EAR-1550901 and The
+University of California -- Davis.
 
@@
 
-T.~Heister was partially supported by NSF Award DMS-1522191, by the
-Computational Infrastructure in Geodynamics initiative (CIG), through the NSF
-under Award EAR-0949446 and EAR-1550901 and The University of California --
-Davis, and by Technical Data Analysis, Inc. through US Navy SBIR N16A-T003.
+T.~Heister was also partially supported by NSF Award DMS-1522191, and
+by Technical Data Analysis, Inc. through US Navy SBIR N16A-T003.
 

### 2019-05-10 9.1/paper.tex
Address a few of the open issues.

@@
 \item Improved support for automatic and symbolic differentiation;
-\item Support for $hp$ adaptivity on MPI-parallel programs with \texttt{parallel::distributed::Triangulation}
-\item Four new tutorial programs step-61, step-62, step-63, step-64.
+\item Full support for $hp$ adaptivity in parallel computations;
+\item Four new tutorial programs step-61, step-62, step-63, step-64,
+  as well as one new code gallery program.
 \end{itemize}
 These will all be discussed in more detail in the
-following section. In addition, this release contains the following changes:
+following section. There are a number of other noteworthy changes in this release:
 \begin{itemize}
-\item ParsedConvergenceTable
-\item FE\_BernardiRaugel
-\item ConstraintMatrix class has been renamed to AffineConstraints
-\item FE\_NedelecSZ
-\item A number of bug fixes and performance improvements to the matrix-free framework and related geometric multigrid solvers were added.
-In particular, the implementation of the Chebyshev iteration, an often used smoother in the matrix-free context, has been revised to reduce the vector access.
-The performance improvement speed up matrix-free multigrid solvers by up to 10-15\% on affine geometries and up to 5\% on curved geometries.
+\item A number of bug fixes and performance improvements to the
+  matrix-free framework and related geometric multigrid solvers. In
+  particular, the implementation of the Chebyshev iteration, an often
+  used smoother in the matrix-free context, has been revised to reduce
+  vector accesses. This speeds up matrix-free
+  multigrid solvers by up to 10-15\% on geometries with affine
+  (parallelogram and parallelpiped) cells, and up to
+  5\% on geometries with cells bounded by curved edges and faces.
+\marginpar{Luca: Please expand}
+\item ParsedConvergenceTable.  
+\item The \texttt{FE\_BernardiRaugel} class implements the
+  non-standard Bernardi-Raugel (BR) element that can be used as one part of
+  a stable velocity/pressure pair for the Stokes equation
+  \cite{BR85}. The BR element can be seen as either an enriched
+  version of the $Q_1^d$ element with added bubble functions on each
+  edge (in 2d) or face (in 3d), or as a reduced version of the $Q_2^d$
+  element. It addresses the fact that the $Q_1^d\times Q_0$
+  combination is not inf-sup stable (requiring a larger velocity
+  space), and that the $Q_2^d\times Q_1$ combination is stable but
+  sub-optimal since the velocity space is too large relative to the
+  pressure space to provide additional accuracy commensurate with the
+  cost of the large number of velocity unknowns. The BR space is
+  intermediate to the $Q_1^d$ and $Q_2^d$ spaces.
+
+\item The \texttt{FE\_NedelecSZ} class is an alternative
+  implementation of the ideas behind the N{\'e}d{\'e}elec element. It is based
+  on the work of Zaglmayr \cite{Zag06} and overcomes the
+  sign conflict issues present in traditional N{\'e}d{\'e}lec
+  elements that arise from the edge and face parameterizations used
+  in the basis functions. Therefore, this element should provide
+  consistent results for general quadrilateral and hexahedral
+  elements for which the relative orientations of edges and faces as
+  seen from all adjacent cells are often difficult to establish.
+  The way this element addresses the sign conflict problem is to assign
+  local edges and faces a globally defined orientation. The
+  implementation is described in \cite{Kynch2017}.
 \end{itemize}
+
 Beyond these changes, the changelog lists more than 200 other features and bugfixes.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Tutorial and code gallery programs}
+\subsection{Improved support for automatic and symbolic differentiation}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Full support for $hp$ adaptivity in parallel computations}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{New and improved tutorial and code gallery programs}
+
+talk about the new programs; also mention updating for C++11 style
 
@@
 
+\marginpar{Update after the release}
 The 9.1 release includes
-\href{https://dealii.org/developer/doxygen/deal.II/changes_between_8_5_0_and_9_0_0.html}
+\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_0_0_and_9_1_0.html}
      {around 15 incompatible changes}; see \cite{changes91}. The majority of these changes
@@
 
-Older releases of \dealii can be cited as
-\cite{dealII80,dealII81,dealII82,dealII83,dealII84,dealII85,dealII90}.
+The two previous releases of \dealii can be cited as
+\cite{dealII85,dealII90}.
 

### 2019-05-10 9.1/paper.tex
More ideas.
Add another note about the BR element.

@@
 \item Full support for $hp$ adaptivity in parallel computations;
+\item Interfaces to the HDF5 file format and libraries.
 \item Four new tutorial programs step-61, step-62, step-63, step-64,
@@
 
+  The current implementation of the element only works for
+  parallelogram meshes due to difficulties associated with the mapping
+  of shape functions: The shape functions of the $Q_1^d$ part of the
+  element need to be mapped as scalars, as is common for the vector
+  components of the $Q_1^d$ element; on the other hand, the
+  vector-valued edge bubble functions need to be mapped using the
+  Piola transform as is common for the Raviart-Thomas
+  element. \dealii{} does not currently have the ability to use
+  different mappings for individual shape functions, though this
+  functionality is planned for the next release.
+
 \item The \texttt{FE\_NedelecSZ} class is an alternative
@@
   implementation is described in \cite{Kynch2017}.
+
+  \marginpar{Maybe also mention the rtree interfaces?}
 \end{itemize}
@@
 \subsection{Full support for $hp$ adaptivity in parallel computations}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Interfaces to the HDF5 file format and libraries}
 
@@
 
+\marginpar{I'm sure we have new interfaces. HDF5 and the various AD/SD
+  libraries come to mind. Please update.}
 \dealii can interface with many other libraries:

### 2019-05-12 9.1/paper.tex
do not break author names, add new affiliation
add another grant

@@
 
-\author[2]{Wolfgang Bangerth}
+\author[2]{Wolfgang~Bangerth}
 \affil[2]{Department of Mathematics, Colorado State University, Fort
@@
 
-\author[3]{Denis Davydov}
+\author[3]{Denis~Davydov}
 \affil[3]{Chair of Applied Mechanics,
@@
 
-\author[4]{Timo Heister}
+\author[4]{Timo~Heister}
 \affil[4]{Mathematical Sciences,
@@
 %
-\author[5]{Luca Heltai}
+\author[5]{Luca~Heltai}
 \affil[5]{SISSA,
@@
 
-\author[6]{Martin Kronbichler}
+\author[6]{Martin~Kronbichler}
 \affil[6]{Institute for Computational Mechanics,
@@
 
-\author[7]{Matthias Maier}
-\affil[7]{School of Mathematics,
-  University of Minnesota,
-  127 Vincent Hall, 206 Church Street SE,
-  Minneapolis, MN 55455, USA.
-  {\texttt{msmaier@umn.edu}}}
+\author[7]{Matthias~Maier}
+\affil[7]{Department of Mathematics,
+  Texas A\&M University,
+  3368 TAMU,
+  College Station, TX 77845, USA.
+  {\texttt{maier@math.tamu.edu}}}
 
-\author[4]{Jean-Paul Pelteret}
+\author[4]{Jean-Paul~Pelteret}
 
-\author[1]{Bruno Turcksin}
+\author[1]{Bruno~Turcksin}
 
@@
 D.~Davydov was supported by the German Research Foundation (DFG), grant DA
-1664/2-1.
+1664/2-1 and the Bayerisches Kompetenznetzwerk
+f\"ur Technisch-Wissenschaftliches Hoch- und H\"ochstleistungsrechnen
+(KONWIHR).
 

### 2019-05-13 9.1/paper.tex
Remove a pointless citation.
Update the list of contributors.

@@
 
-\nocite{BangerthKanschat1999}
-
 \section{Acknowledgments}
@@
 %
+% updated 5/12/2019
+\marginpar{Remove all other authors of this paper}
+\marginpar{Do we want to spell out who the Gingko devs are?}
+  Giovanni Alzetta,
+  Mathias Anselmann,
+  Daniel Appel,
+  Alexander Blank,
+  Vishal Boddu,
+  Benjamin Brands,
+  Pi-Yueh Chuang,
+  T. Conrad Clevenger,
+  Sambit Das,
+  the Ginkgo developers,
+  Stefano Dominici,
+  Nivesh Dommaraju,
+  Marc Fehling,
+  Niklas Fehn,
+  Isuru Fernando,
+  Andreas Fink,
+  Daniel Garcia-Sanchez,
+  Rene Gassm{\"o}ller,
+  Alexander Grayver,
+  Joshua Hanophy,
+  Logan Harbour,
+  Graham Harper,
+  Daniel Jodlbauer,
+  Stefan Kaessmair,
+  Eldar Khattatov,
+  Alexander Knieps,
+  Uwe K{\"o}cher,
+  Kurt Kremitzki,
+  Dustin Kumor,
+  Ross Kynch,
+  Damien Lebrun-Grandie,
+  Jonathan Matthews,
+  Stefan Meggendorfer,
+  Pratik Nayak,
+  Lei Qiao,
+  Ce Qin,
+  Reza Rastak,
+  Roland Richter,
+  Alberto Sartori,
+  Svenja Schoeder,
+  Sebastian Stark,
+  Antoni Vidal,
+  Jiaxin Wang,
+  Yuxiang Wang,
+  Zhuoran Wang.
 

### 2019-05-13 9.1/paper.tex
Report on the hp situation.

@@
 
+\dealii{} has had support for $hp$ adaptive methods since around 2005
+(documented in \cite{BangerthKayserHerold2007} and for parallel
+computations on distributed meshes since around 2010 (see
+\cite{BangerthBursteddeHeisterKronbichler11}), but not for both at the
+same time. The challenges to combine these are related to a number of
+areas:
+\begin{enumerate}
+\item Data structures: The data structures necessary to store the
+  indices of degrees of freedom are necessarily substantially more
+  complicated for $hp$ algorithms than for the $h$-adaptive schemes
+  that were already implemented. This is because the number of degrees
+  of freedom per cell is now no longer constant. Furthermore,
+  faces and edges may need to store more than one set of indices if
+  the adjacent cells use different polynomial degrees; in the case of
+  edges, the number of sets of indices may also be of variable size.
+
+  All of this poses challenges in the parallel context because some of
+  the information may not be known, or not be known right away, for
+  cells that are not locally owned (i.e., for ghost and ``artificial''
+  cells), and for which the data structures stored on different
+  processors have to reconciled.
+
+\item Algorithms: Already for $h$ adaptive meshes, enumerating all
+  degrees of freedom on the global mesh is difficult as evidenced by
+  the complications of the algorithms shown in Section 3.1 of
+  \cite{BangerthBursteddeHeisterKronbichler11}, which requires more
+  than a page of text and is implemented in many hundreds of lines of
+  code).
+
+  These difficulties are even more pronounced when using $hp$
+  adaptivity. The main obstacle is the desire to unify the indices of
+  matching degrees of freedom on adjacent cells, such as the edge
+  since degree of freedom of a $Q_2$ element with the middle one of
+  the three edge degrees of freedom of a $Q_4$ element on a
+  neighboring cell. Section 4.2 of
+  \cite{BangerthKayserHerold2007} discusses a sequential algorithm
+  that eliminates one of these degrees of freedom in favor of another,
+  but in this introduces a ``master'' and a ``slave'' side of the
+  interface. This is of no major consequence in sequential
+  computations, but is inconvenient in parallel computations if the
+  ``master'' side is a ghost cell whose degree of freedom indices are
+  not (yet) available while enumerating local degrees of freedom, or
+  if the master is an artificial cell whose information will never be
+  available on a processor.
+
+  An earlier implementation of the algorithm enumerating degrees of
+  freedom, already available in \dealii{} 9.0, simply did not unify
+  indices on processor boundaries. However, this makes the total
+  number of degrees of freedom dependent on both the partition of the
+  mesh and the number of processors available. We have therefore
+  re-implemented the algorithm so that the unification does happen
+  also on processor boundaries, and will report on the details
+  elsewhere.
+
+\item Data transfer patterns: An important algorithm in parallel
+  finite element methods is the exchange of information stored on
+  cells during mesh repartitioning. This happens, for example, when
+  interpolating the solution from one mesh to the next,
+  adaptively-refined mesh, or when adapting the polynomial degrees of
+  freedom associated with each cell and repartitioning in order to
+  balance the computational cost of each processor's partition. When
+  using $h$ adaptive methods, the amount of data associated with each
+  cell is fixed and the algorithms that implement the data transfer
+  are consequently relatively simple. On the other hand, in $hp$
+  contexts, each cell may have a different number of unknowns
+  associated with it, and the algorithms that transfer the data are
+  substantially more complicated. Furthermore, the amount of data
+  associated with each cell may be large on cells with higher
+  polynomial degrees, and might profit from compression before
+  sending.
+
+\item Balancing computational cost: For $h$-adaptive algorithms, the
+  amount of work associated with each cell is essentially the same
+  both during the assembly of linear systems as well as during the
+  solver phase. For $hp$-adaptive methods, this is no longer the
+  case. Consequently, balancing the cost of work between different
+  processors' partitions is no longer as easy as ensuring that every
+  processor owns a roughly equal number of cells. Rather, one needs to
+  introduce a weighting factor for each cell that describes its
+  relative cost compared to some reference. To make things work, the
+  relative cost of assembly on a cell might now match the relative
+  cost of the linear solver associated with this cell, leading to
+  difficult trade-offs in defining optimal weighting factors.
+\end{enumerate}
+
+All of these issues have been addressed in the current release and are
+available to users. We will report on the details of the algorithms
+and their performance in a separate publication.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-05-13 9.1/paper.tex
Add three more references to multigrid and matrixfree
fix bib entry for changes
affiliation

@@
   Clemson University,
-  Clemson, SC 29634, USA.
-  {\texttt{heister@clemson.edu}}}
+  Clemson, SC 29634, USA
+  and Scientific Computing and Imaging Institute,
+72 S Central Campus Drive, Room 3750
+Salt Lake City, UT 84112.
+  {\texttt{heister@sci.utah.edu}}}
 %
@@
 the release announcement.)
+\marginpar{check that URL works!}
 
@@
 \begin{itemize}
- \item For geometric multigrid: \cite{Kanschat2004,JanssenKanschat2011};
+ \item For geometric multigrid: \cite{Kanschat2004,JanssenKanschat2011,ClevengerHeisterKanschatKronbichler2019,KronbichlerLjungkvist2019};
  \item For distributed parallel computing: \cite{BangerthBursteddeHeisterKronbichler11};
@@
  \item For matrix-free and fast assembly techniques:
-   \cite{KronbichlerKormann2012};
+   \cite{KronbichlerKormann2012,KronbichlerKormann2019};
  \item For computations on lower-dimensional manifolds:

### 2019-05-14 9.1/paper.tex
List GPU support

@@
 \item Interfaces to the HDF5 file format and libraries.
+\item Significantly extended GPU support.
 \item Four new tutorial programs step-61, step-62, step-63, step-64,
@@
   5\% on geometries with cells bounded by curved edges and faces.
+\item Various variants of geometric multigrid solvers and matrix-free
+  implementations were run on up to 304,128 cores during the acceptance phase of
+  the SuperMUC-NG supercomputer, verifying the scalability of our
+  implementations. Some performance bottlenecks on distributed triangulation
+  with more than 100k MPI ranks have been fixed.
 \marginpar{Luca: Please expand}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{GPU support via CUDA}
+
+The GPU support was significantly extended for the current release. On the one hand, the MPI-parallel vector class
+\texttt{LinearAlgebra::distributed::Vector} has gained a second template
+argument \texttt{MemorySpace} which can either be \texttt{Host} or
+\texttt{CUDA}. In the latter case, the data resides in the GPU memory. Also,
+the ghost exchange can be performed completely via CUDA in case a compatible
+MPI library is found (as is usual on large GPU-based supercomputers).
+
+Furthermore, a new tutorial program \texttt{step-64} has been added that demonstrates the usage of matrix-free methods on Nvidia GPUs.
+GPUs are advantageous for these kind of operations because of their superior hardware characteristics, in particular a higher memory bandwidth than server CPUs within a given power envelope.
+The matrix-free GPU components integrated in \dealii have been analyzed against CPUs in \cite{KronbichlerLjungkvist2019}, where the application to geometric multigrid solvers is discussed.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{New and improved tutorial and code gallery programs}
@@
 \begin{itemize}
- \item For geometric multigrid: \cite{Kanschat2004,JanssenKanschat2011,ClevengerHeisterKanschatKronbichler2019,KronbichlerLjungkvist2019};
+ \item For geometric multigrid: \cite{Kanschat2004,JanssenKanschat2011,ClevengerHeisterKanschatKronbichler2019};
  \item For distributed parallel computing: \cite{BangerthBursteddeHeisterKronbichler11};

### 2019-05-15 9.1/paper.tex
fix changelist URL
add an entry

@@
 \item Significantly extended GPU support.
+\item Parallel GMG:
+  \begin{itemize}
+  \item see \cite{ClevengerHeisterKanschatKronbichler2019}
+  \item performance optimizations
+    \item new features?
+    \item largest runs now at XYZ (Kronbichler?)
+    \end{itemize}
 \item Four new tutorial programs step-61, step-62, step-63, step-64,
@@
 the release announcement.)
-\marginpar{check that URL works!}
 
@@
 The 9.1 release includes
-\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_0_0_and_9_1_0.html}
+\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_0_1_and_9_1_0.html}
      {around 15 incompatible changes}; see \cite{changes91}. The majority of these changes

### 2019-05-16 9.1/paper.tex
deuglify footnote

@@
 
-\author[1]{Daniel Arndt \footnote{
-  This manuscript has been authored by UT-Battelle, LLC under Contract No.
-  DE-AC05-00OR22725 with the U.S. Department of Energy. The United States
-  Government retains and the publisher, by accepting the article for
-  publication, acknowledges that the United States Government retains a
-  non-exclusive, paid-up, irrevocable, worldwide license to publish or reproduce
-  the published form of this manuscript, or allow others to do so, for United
-  States Government purposes. The Department of Energy will provide public
-  access to these results of federally sponsored research in accordance with the
-  DOE Public Access Plan (http://energy.gov/downloads/doe-public-access-plan).}}
+\author[*1]{Daniel Arndt}
 \affil[1]{Computational Engineering and Energy Sciences Group,
@@
 
-\author[1]{Bruno~Turcksin}
+\author[*1]{Bruno~Turcksin}
 
@@
 \maketitle
+
+\footnotetext{%
+  $^\ast$ This manuscript has been authored by UT-Battelle, LLC under Contract No.
+  DE-AC05-00OR22725 with the U.S. Department of Energy. The United States
+  Government retains and the publisher, by accepting the article for
+  publication, acknowledges that the United States Government retains a
+  non-exclusive, paid-up, irrevocable, worldwide license to publish or reproduce
+  the published form of this manuscript, or allow others to do so, for United
+  States Government purposes. The Department of Energy will provide public
+  access to these results of federally sponsored research in accordance with the
+  DOE Public Access Plan (http://energy.gov/downloads/doe-public-access-plan).}
 

### 2019-05-16 9.1/paper.tex
restructure outline

@@
 development of finite element solvers. It is available for free under the
-GNU Lesser General Public License (LGPL) from the \dealii homepage at
-\url{http://www.dealii.org/}.
+GNU Lesser General Public License (LGPL). Downloads are available at
+\url{https://www.dealii.org/} and \url{https://github.com/dealii/dealii}.
 
@@
 \begin{itemize}
-\item Improved support for automatic and symbolic differentiation;
-\item Full support for $hp$ adaptivity in parallel computations;
-\item Interfaces to the HDF5 file format and libraries.
-\item Significantly extended GPU support.
-\item Parallel GMG:
-  \begin{itemize}
-  \item see \cite{ClevengerHeisterKanschatKronbichler2019}
-  \item performance optimizations
-    \item new features?
-    \item largest runs now at XYZ (Kronbichler?)
-    \end{itemize}
-\item Four new tutorial programs step-61, step-62, step-63, step-64,
-  as well as one new code gallery program.
+  \item Improved support for automatic and symbolic differentiation (see
+    Section~\ref{subsec:ad}),
+  \item Full support for $hp$ adaptivity in parallel computations (see
+    Section~\ref{subsec:hp}),
+  \item Interfaces to the HDF5 file format and libraries (see
+    Section~\ref{subsec:hdf5}),
+  \item Significantly extended GPU support (see Section~\ref{subsec:gpu}),
+  \item Parallel geometric multigrid (GMG) improvements (see
+    \cite{ClevengerHeisterKanschatKronbichler2019} and
+    Section~\ref{subsec:gmg}),
+  \item Four new tutorial programs (step-61, step-62, step-63, step-64),
+    as well as one new code gallery program (see
+    Section~\ref{subsec:steps}).
 \end{itemize}
@@
 \subsection{Improved support for automatic and symbolic differentiation}
+\label{subsec:ad}
 
@@
 \subsection{Full support for $hp$ adaptivity in parallel computations}
+\label{subsec:hp}
 
@@
 \subsection{Interfaces to the HDF5 file format and libraries}
+\label{subsec:hdf5}
 
@@
 \subsection{GPU support via CUDA}
+\label{subsec:gpu}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Parallel geometric multigrid improvements}
+\label{subsec:gmg}
+\marginpar{performance optimizations, new features,
+largest runs now at XYZ (Kronbichler?)}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{New and improved tutorial and code gallery programs}
+\label{subsec:steps}
 

### 2019-05-16 9.1/paper.tex
polish Overview, add some unfinished sections

@@
 The major changes of this release are:
+%
 \begin{itemize}
@@
 \end{itemize}
-These will all be discussed in more detail in the
-following section. There are a number of other noteworthy changes in this release:
+%
+The major changes are discussed in detail in Section~\ref{sec:major}. There
+are a number of other noteworthy changes in the current \dealii{} release
+that we briefly outline in the remainder of this section:
+%
 \begin{itemize}
-\item A number of bug fixes and performance improvements to the
+\item
+  The release contains bug fixes and performance improvements of the
   matrix-free framework and related geometric multigrid solvers. In
-  particular, the implementation of the Chebyshev iteration, an often
-  used smoother in the matrix-free context, has been revised to reduce
-  vector accesses. This speeds up matrix-free
-  multigrid solvers by up to 10-15\% on geometries with affine
-  (parallelogram and parallelpiped) cells, and up to
-  5\% on geometries with cells bounded by curved edges and faces.
-\item Various variants of geometric multigrid solvers and matrix-free
-  implementations were run on up to 304,128 cores during the acceptance phase of
-  the SuperMUC-NG supercomputer, verifying the scalability of our
-  implementations. Some performance bottlenecks on distributed triangulation
-  with more than 100k MPI ranks have been fixed.
+  particular, the implementation of the Chebyshev iteration, an often used
+  smoother in the matrix-free context, has been revised to reduce vector
+  accesses. This speeds up matrix-free multigrid solvers by up to 10-15\%
+  on geometries with affine (parallelogram and parallelpiped) cells, and up
+  to 5\% on geometries with cells bounded by curved edges and faces.
+
+\item
+  Various variants of geometric multigrid solvers and matrix-free
+  implementations were run on up to 304,128 cores during the acceptance
+  phase of the SuperMUC-NG supercomputer, verifying the scalability of our
+  implementations. Some performance bottlenecks on distributed
+  triangulation with more than 100k MPI ranks have been fixed.
+
+\item
+  A new class \texttt{ParsedConvergenceTable} has been introduced.
+  \textcolor{red}{TODO}
   \marginpar{Luca: Please expand}
-\item ParsedConvergenceTable.
-\item The \texttt{FE\_BernardiRaugel} class implements the
-  non-standard Bernardi-Raugel (BR) element that can be used as one part of
-  a stable velocity/pressure pair for the Stokes equation
-  \cite{BR85}. The BR element can be seen as either an enriched
-  version of the $Q_1^d$ element with added bubble functions on each
-  edge (in 2d) or face (in 3d), or as a reduced version of the $Q_2^d$
-  element. It addresses the fact that the $Q_1^d\times Q_0$
-  combination is not inf-sup stable (requiring a larger velocity
-  space), and that the $Q_2^d\times Q_1$ combination is stable but
-  sub-optimal since the velocity space is too large relative to the
-  pressure space to provide additional accuracy commensurate with the
+
+\item
+  The \texttt{FE\_BernardiRaugel} class implements the non-standard
+  Bernardi-Raugel (BR) element that can be used to construct a stable
+  velocity-pressure pair for the Stokes equation \cite{BR85}. The BR
+  element is an enriched version of the $Q_1^d$ element with added bubble
+  functions on each edge (in 2d) or face (in 3d). It addresses the fact
+  that the $Q_1^d\times Q_0$ combination is not inf-sup stable (requiring a
+  larger velocity space), and that the $Q_2^d\times Q_1$ combination is
+  stable but sub-optimal since the velocity space is too large relative to
+  the pressure space to provide additional accuracy commensurate with the
   cost of the large number of velocity unknowns. The BR space is
@@
 
-  The current implementation of the element only works for
-  parallelogram meshes due to difficulties associated with the mapping
-  of shape functions: The shape functions of the $Q_1^d$ part of the
-  element need to be mapped as scalars, as is common for the vector
-  components of the $Q_1^d$ element; on the other hand, the
-  vector-valued edge bubble functions need to be mapped using the
-  Piola transform as is common for the Raviart-Thomas
-  element. \dealii{} does not currently have the ability to use
-  different mappings for individual shape functions, though this
-  functionality is planned for the next release.
+  The element is currently only implemented for parallelogram meshes due to
+  difficulties associated with the mapping of shape functions: The shape
+  functions of the $Q_1^d$ part of the element need to be mapped as
+  scalars, as is common for the vector components of the $Q_1^d$ element;
+  on the other hand, the vector-valued edge bubble functions need to be
+  mapped using the Piola transform as is common for the Raviart-Thomas
+  element. \dealii{} does not currently have the ability to use different
+  mappings for individual shape functions, though this functionality is
+  planned for the next release.
 
-\item The \texttt{FE\_NedelecSZ} class is an alternative
-  implementation of the ideas behind the N{\'e}d{\'e}elec element. It is based
-  on the work of Zaglmayr \cite{Zag06} and overcomes the
-  sign conflict issues present in traditional N{\'e}d{\'e}lec
-  elements that arise from the edge and face parameterizations used
-  in the basis functions. Therefore, this element should provide
-  consistent results for general quadrilateral and hexahedral
-  elements for which the relative orientations of edges and faces as
-  seen from all adjacent cells are often difficult to establish.
-  The way this element addresses the sign conflict problem is to assign
-  local edges and faces a globally defined orientation. The
-  implementation is described in \cite{Kynch2017}.
+\item
+  The \texttt{FE\_NedelecSZ} class is a new implementation of the
+  N{\'e}d{\'e}lec element on quadrilaterals and hexahedra. It is based on
+  the work of Zaglmayr \cite{Zag06} and overcomes the sign conflict issues
+  present in traditional N{\'e}d{\'e}lec elements that arise from the edge
+  and face parameterizations used in the basis functions. Therefore, this
+  element should provide consistent results for general quadrilateral and
+  hexahedral elements for which the relative orientations of edges and
+  faces (as seen from all adjacent cells) are often difficult to establish.
+  The \texttt{FE\_NedelecSZ} element addresses the sign conflict problem by
+  assigning a globally defined orientation to local edges and faces. A
+  detailed overview of the implementation of the \texttt{FE\_NedelecSZ}
+  element in \dealii{} can be found in \cite{Kynch2017}.
 
-  \marginpar{Maybe also mention the rtree interfaces?}
+\item
+  A new interface to \texttt{boost::geometry::index::rtree} has been added.
+  \textcolor{red}{TODO}
 \end{itemize}
-
-Beyond these changes, the changelog lists more than 200 other features and bugfixes.
+%
+In addition to these changes, the changelog lists more than 200 other
+features and bugfixes.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Significant changes to the library}
+\section{Major changes to the library}
+\label{sec:major}
 
 This release of \dealii contains a number of large and significant changes
-that will be discussed in the following sections. It of course also contains a
+that will be discussed in this section.
+
+It of course also contains a
 vast number of smaller changes and added functionality; the details of these

### 2019-05-18 9.1/paper.tex
Add text about CUDA support

@@
 
-The GPU support was significantly extended for the current release. On the one hand, the MPI-parallel vector class
-\texttt{LinearAlgebra::distributed::Vector} has gained a second template
-argument \texttt{MemorySpace} which can either be \texttt{Host} or
-\texttt{CUDA}. In the latter case, the data resides in the GPU memory. Also,
-the ghost exchange can be performed completely via CUDA in case a compatible
-MPI library is found (as is usual on large GPU-based supercomputers).
+The GPU support was significantly extended for the current release:
+\begin{itemize}
+  \item \texttt{LinearAlgebra::distributed::Vector}: the MPI-parallel vector
+    class has gained a second template argument \texttt{MemorySpace} which can
+    either be \texttt{Host} or \texttt{CUDA}. In the latter case, the data
+    resides in the GPU memory. By default, the template parameter is
+    \texttt{Host} and the behavior is unchanged compared to previous versions. 
+    When using CUDA, the ghost exchange can be performed either by first copy
+    the relevant data to the host, performing a MPI communication, and finally
+    move the data to the device or, if CUDA-aware MPI is available, by
+    performing the MPI directly from GPU to GPU.
+  \item Constrained degrees of freedom: the matrix-free framework now now
+    supports constraints degrees of freedom. The implementation is based on
+    \cite{ljungkvist2017}. With this addition, not only the user can impose Dirichlet
+    boundary conditions but the matrix-free framework can be used on adapted
+    meshes. The only restriction is that for two-dimensional meshes, the degree
+    of finite elements must be odd.  There is no such restriction in three
+    dimensions.
+  \item MPI matrix-free: using \texttt{LinearAlgebra::distributed::Vector}, the
+    matrix-free framework can scale to multiple multiple GPUS by taking
+    advantage of MPI. Each MPI process can only use GPU and therefore, if
+    multiple GPUs are available in one node, it is necessary to have as many
+    ranks as GPUs. Using Nvidia Multi-Process Device (MPS), it is also possible
+    for multiple processes to use the same GPU. This can be advantageous if the
+    amount of work on one rank is not sufficient to fully utilize GPU.
+\end{itemize}
 

### 2019-05-20 9.1/paper.tex
Add my authorship info.
Summarize incompatible changes.
This list is a bit subjective: I suspect users are not using some of the
changed interfaces that deal with, e.g., implementing hp FEM.

@@
 \author[9]{David Wells}
-\affil[9]{Department of Mathematical Sciences, Rensselaer Polytechnic
-  Institute, Troy, NY 12180, USA.
-  {\texttt{wellsd2@rpi.edu}}}
+\affil[9]{Department of Mathematics, University of North Carolina,
+  Chapel Hill, NC 27516, USA.
+  {\texttt{drwells@email.unc.edu}}}
 
@@
 
-\marginpar{Update after the release}
 The 9.1 release includes
@@
 \begin{itemize}
-\item TODO
+\item The \texttt{VectorView} class was removed. We recommend either copying the
+      vector subset into a \texttt{Vector} or using a \texttt{BlockVector}.
+\item The function \texttt{Subscriptor::subscribe} now requires a pointer to a
+      \texttt{std::atomic<bool>} that tracks whether or not the pointer to the
+      subscribed-to object is still valid.
+\item The \texttt{ConstraintMatrix} class gained a template parameter for scalar
+      type and was been renamed \texttt{AffineConstraints}. Several methods that
+      take vectors or matrices as arguments,
+      such as \texttt{AffineConstraints::distribute\_local\_to\_global()},
+      now require that all matrix and vector arguments have matching number
+      types.
+\item Similarly, the functions \texttt{create\_mass\_matrix} and
+      \texttt{create\_boundary\_mass\_matrix} in the \texttt{MatrixCreator}
+      namespace no longer
+      support matrix and vector objects of different types.
 \end{itemize}

### 2019-05-20 9.1/paper.tex
Improve GPU related sections

@@
 \begin{itemize}
+  \item \texttt{CUDAWrappers::PreconditionILU} and \texttt{CUDAWrappers::PreconditionIC}
+    can be used for preconditioning \texttt{CUDAWrappers::SparseMatrix} objects.
   \item \texttt{LinearAlgebra::distributed::Vector}: the MPI-parallel vector
@@
     When using CUDA, the ghost exchange can be performed either by first copy
-    the relevant data to the host, performing a MPI communication, and finally
+    the relevant data to the host, performing MPI communication, and finally
     move the data to the device or, if CUDA-aware MPI is available, by
-    performing the MPI directly from GPU to GPU.
-  \item Constrained degrees of freedom: the matrix-free framework now now
-    supports constraints degrees of freedom. The implementation is based on
+    performing MPI communication directly between GPUs.
+  \item Constrained degrees of freedom: the matrix-free framework now
+    supports constrained degrees of freedom. The implementation is based on
     \cite{ljungkvist2017}. With this addition, not only the user can impose Dirichlet
-    boundary conditions but the matrix-free framework can be used on adapted
-    meshes. The only restriction is that for two-dimensional meshes, the degree
-    of finite elements must be odd.  There is no such restriction in three
+    boundary conditions but also the matrix-free framework can be used with adaptively
+    refined meshes. The only restriction is that for two-dimensional meshes the
+    finite element degree must be odd.  There is no such restriction in three
     dimensions.
   \item MPI matrix-free: using \texttt{LinearAlgebra::distributed::Vector}, the
-    matrix-free framework can scale to multiple multiple GPUS by taking
-    advantage of MPI. Each MPI process can only use GPU and therefore, if
+    matrix-free framework can scale to multiple GPUs by taking
+    advantage of MPI. Each MPI process can only use one GPU and therefore, if
     multiple GPUs are available in one node, it is necessary to have as many
-    ranks as GPUs. Using Nvidia Multi-Process Device (MPS), it is also possible
+    ranks as there are GPUs. Using Nvidia Multi-Process Service (MPS), it is also possible
     for multiple processes to use the same GPU. This can be advantageous if the
-    amount of work on one rank is not sufficient to fully utilize GPU.
+    amount of work on one rank is not sufficient to fully utilize a GPU.
 \end{itemize}
 
-Furthermore, a new tutorial program \texttt{step-64} has been added that demonstrates the usage of matrix-free methods on Nvidia GPUs.
-GPUs are advantageous for these kind of operations because of their superior hardware characteristics, in particular a higher memory bandwidth than server CPUs within a given power envelope.
-The matrix-free GPU components integrated in \dealii have been analyzed against CPUs in \cite{KronbichlerLjungkvist2019}, where the application to geometric multigrid solvers is discussed.
+The matrix-free GPU components integrated in \dealii have been analyzed against
+CPUs in \cite{KronbichlerLjungkvist2019}, where the application to geometric
+multigrid solvers is discussed.
 
@@
 talk about the new programs; also mention updating for C++11 style
+
+\texttt{step-64} demonstrates the usage of matrix-free methods on Nvidia GPUs.
+GPUs are advantageous for these kind of operations because of their superior
+hardware characteristics, in particular a higher memory bandwidth than server
+CPUs within a given power envelope.
 

### 2019-05-21 9.1/paper.tex
Added two paragraphs on rtree and parsed ct

@@
 \usepackage{multicol}
+\usepackage{pdfsync}
 
@@
 
-\item
-  A new class \texttt{ParsedConvergenceTable} has been introduced.
-  \textcolor{red}{TODO}
-  \marginpar{Luca: Please expand}
+\item A new class \texttt{ParsedConvergenceTable} has been introduced
+  that greatly simplifies the construction of convergence tables,
+  reading the options for the generation of the table from a parameter
+  file, and providing methods that, combined with a parameter file,
+  allow one to generate convergence tables using one-liners in user
+  codes.
 
@@
 
-\item
-  A new interface to \texttt{boost::geometry::index::rtree} has been added.
-  \textcolor{red}{TODO}
+\item All of the elementary geometrical objects of the library (namely
+  \texttt{Point<dim>}, \texttt{Segment<dim>}, and
+  \texttt{BoundingBox<dim>}) have been augmented with the traits
+  needed to comply with \texttt{boost::geometry} concepts. A new
+  interface to \texttt{boost::geometry::index::rtree} has been added
+  that simplifies the construction of spacial indices based on points,
+  bounding boxes, or segments.
 \end{itemize}
@@
  \item For uses of the \texttt{WorkStream} interface:
-   \cite{TKB16}.
+   \cite{TKB16};
+   \item Fur uses of the \texttt{ParameterAcceptor} concept,
+     \texttt{MeshWorker::ScratchData}, and
+     \texttt{ParsedConvergenceTable}: \cite{SartoriGiulianiBardelloni-2018-a}.
 \end{itemize}

### 2019-05-21 9.1/paper.tex
Fix a few wording issues

@@
 \dealii{} has had support for $hp$ adaptive methods since around 2005
-(documented in \cite{BangerthKayserHerold2007} and for parallel
+(documented in \cite{BangerthKayserHerold2007}) and for parallel
 computations on distributed meshes since around 2010 (see
@@
   matching degrees of freedom on adjacent cells, such as the edge
-  since degree of freedom of a $Q_2$ element with the middle one of
+  degree of freedom of a $Q_2$ element with the middle one of
   the three edge degrees of freedom of a $Q_4$ element on a
@@
   interpolating the solution from one mesh to the next,
-  adaptively-refined mesh, or when adapting the polynomial degrees of
-  freedom associated with each cell and repartitioning in order to
+  adaptively-refined mesh, or when adapting the polynomial degrees
+  associated with each cell and repartitioning in order to
   balance the computational cost of each processor's partition. When
@@
   introduce a weighting factor for each cell that describes its
-  relative cost compared to some reference. To make things work, the
-  relative cost of assembly on a cell might now match the relative
+  relative cost compared to some reference. To make things worse, the
+  relative cost of assembly on a cell might not match the relative
   cost of the linear solver associated with this cell, leading to
@@
     \texttt{Host} and the behavior is unchanged compared to previous versions.
-    When using CUDA, the ghost exchange can be performed either by first copy
+    When using CUDA, the ghost exchange can be performed either by first copying
     the relevant data to the host, performing MPI communication, and finally
@@
   \item Constrained degrees of freedom: the matrix-free framework now
-    supports constrained degrees of freedom. The implementation is based on
-    \cite{ljungkvist2017}. With this addition, not only the user can impose Dirichlet
-    boundary conditions but also the matrix-free framework can be used with adaptively
-    refined meshes. The only restriction is that for two-dimensional meshes the
+    supports constrained degrees of freedom. The implementation is based on the algorithms described in
+    \cite{ljungkvist2017}. With this addition, both Dirichlet
+    boundary conditions and the constraints arising for adaptively
+    refined meshes can be imposed with in the matrix-free framework. The only restriction is that for two-dimensional meshes the
     finite element degree must be odd.  There is no such restriction in three
     dimensions.
-  \item MPI matrix-free: using \texttt{LinearAlgebra::distributed::Vector}, the
+  \item MPI matrix-free computations: using \texttt{LinearAlgebra::distributed::Vector}, the
     matrix-free framework can scale to multiple GPUs by taking
@@
 
-The matrix-free GPU components integrated in \dealii have been analyzed against
+The matrix-free GPU components integrated in \dealii have been compared against
 CPUs in \cite{KronbichlerLjungkvist2019}, where the application to geometric

### 2019-05-22 9.1/paper.tex
Replace cores -> MPI ranks.
I will run on 300k cores only next week, up to now it was only 152k cores
and 304k MPI ranks

@@
   Various variants of geometric multigrid solvers and matrix-free
-  implementations were run on up to 304,128 cores during the acceptance
+  implementations were run on up to 304,128 MPI ranks during the acceptance
   phase of the SuperMUC-NG supercomputer, verifying the scalability of our
   implementations. Some performance bottlenecks on distributed
-  triangulation with more than 100k MPI ranks have been fixed.
+  triangulation with more than 100k ranks have been fixed.
 

### 2019-05-23 9.1/paper.tex
update acknowledgement

@@
 
-B.~Turcksin: This material is based upon work supported by the U.S.
-Department of Energy, Office of Science, under contract number
-DE-AC05-00OR22725.
+B.~Turcksin: Research sponsored by the Laboratory Directed Research and
+Development Program of Oak Ridge National Laboratory, managed by UT-Battelle,
+LLC, for the U. S. Department of Energy.
 

### 2019-06-02 9.1/paper.tex
Write description of AD improvements

@@
 \begin{itemize}
-  \item Improved support for automatic and symbolic differentiation (see
+  \item Improved support for automatic differentiation (see
     Section~\ref{subsec:ad}),
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Improved support for automatic and symbolic differentiation}
+\subsection{Improved support for automatic differentiation}
 \label{subsec:ad}
+
+In the previous release, numerous classes that are used to assemble linear systems
+and right hand sides, as well those used to define constitutive laws, were given 
+full support for ``white-listed'' automatically differentiable (AD) number types 
+from the ADOL-C and Sacado libraries.
+In the current release we have provided a unified interface to these
+AD libraries focussing two specific use contexts, namely
+\begin{enumerate}
+\item the construction and linearization of finite element residuals, and
+\item the construction and linearization of constitutive model kinetic variables.
+\end{enumerate}
+
+In the first context, the finite element degrees of freedom are considered the
+independent variables. From these primitives, the \texttt{EnergyFunctional} helper
+class in the namespace \texttt{Differentiation::AD} may be used to compute both the 
+residual and its linearization by directly defining the contribution to the 
+(twice differentiated) scalar total energy functional from each cell. Similarly, 
+the \texttt{ResidualLinearization} class requires the (once-differentiated) finite 
+element residual to be defined on a per cell basis,  and this contribution is 
+automatically linearized.
+
+The second context aims directly at constitutive model formulations, and serves to
+compute the directional derivatives of components of (multi-field) constitutive laws
+with respect to the scalar, vector, tensor and symmetric tensor fields in terms 
+of which they are parameterized. The \texttt{ScalarFunction} class may be used to 
+define a scalar function (such as strain energy function) that may be twice 
+differentiated, while the \texttt{VectorFunction} may be used to define a vector
+function (such as a set of kinematic fields) that may be differentiated once.
+Since the total derivatives of all components are computed at once, these two helper 
+classes provide an interface to retrieve each sub-component of the gradient and 
+Hessian (for a \texttt{ScalarFunction}) or values and Jacobian (for a 
+\texttt{VectorFunction}).
+
+Although these aforementioned helper classes have been documented with a specific
+use in mind, they remain generic and may (with a reinterpretation of the meaning of
+the independent and dependent variables) be used for other purposes as well.
+Furthermore, through the implementation of \texttt{TapedDrivers} and 
+\texttt{TapelessDrivers} classes that interface with the active AD library, the 
+generic helper classes hide library-dependent implementational details and facilitate 
+switching between the supported libraries and AD number types based on the 
+user's requirements.
 

### 2019-06-02 9.1/paper.tex
Write section on symbolic differentiation

@@
     Section~\ref{subsec:ad}),
+  \item Dedicated support for symbolic algebra (see
+    Section~\ref{subsec:sd}),
   \item Full support for $hp$ adaptivity in parallel computations (see
@@
 user's requirements.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Dedicated support for symbolic algebra, including algebra differentiation}
+\label{subsec:sd}
+To complement the automatic differentiation features in \dealii{}, this release sees
+the first step towards integrating and supporting a highly performant computer algebra 
+system (CAS) via the SymEngine library.
+The \texttt{Expression} class in the namespace \texttt{Differentiation::SD} interfaces
+to SymEngine and forms the basis of symbolic computations, offering a full set of 
+overloaded operators and a C++ style interface. This class offers the following 
+basic functionality:
+\begin{itemize}
+\item symbolic variable definition,
+\item symbolic function definition,
+\item expression creation using standard C++ syntax,
+\item expression parsing,
+\item comparison operations,
+\item logical operations,
+\item conditional expression construction,
+\item differentiation,
+\item substitution (partial and complete), and
+\item serialization.
+\end{itemize}
+We also provide an extensive set of math operations, with a syntax mimicing that used 
+in the C++ standard library. Using \dealii{}'s \texttt{Expression} class as a basis, we 
+have developed a set of functions that can be used to create \dealii{} \texttt{Tensor}s
+and \texttt{SymmetricTensor}s of symbolic variables and symbolic functions. This
+gives full symbolic tensor algebra support using the pre-existing \texttt{Tensor}
+and \texttt{SymmetricTensor} class and associated functions. A set of utility functions to
+\begin{itemize}
+\item \texttt{differentiate} scalar expressions, tensor and symmetric tensors of expressions
+with respect to other scalar expressions, tensor and symmetric tensors of expressions,
+\item create symbolic substitution maps,
+\item resolve explicit dependencies between expressions, and
+\item perform scalar and tensor valued substitution (including conversion from symbolic to 
+real-valued scalars and tensors)
+\end{itemize}
+have also been implemented.
+
+In the next release we expect to implement classes to assist in performing assembly operations 
+in the same spirit as that which has been done in the \texttt{Differentiation::AD} namespace,
+although in a fully symbolic manner.
+We will also address increase the performance of the \texttt{Expression} class by leveraging
+the optimization capabilities of SymEngine, including common subexpression elimination (CSE), 
+as well as by generating high performance code-paths to evaluate these expressions through the
+use of a custom-generated \texttt{std::function} (so-called ``lambda'' optimization) or by 
+compiling expressions using the LLVM JIT compiler.
 

### 2019-06-03 9.1/paper.tex
Fix a typo
Smaller edits.

--- 2019-06-02 9.1/paper.tex
@@
   This paper provides an overview of the new features of the finite element
-  library \dealii version 9.1.
+  library \dealii, version 9.1.
 \end{abstract}
@@
 
-\dealii version 9.1.0 was released May XX, 2019
+\dealii version 9.1.0 was released May 21, 2019
 This paper provides an
@@
   cost of the large number of velocity unknowns. The BR space is
-  intermediate to the $Q_1^d$ and $Q_2^d$ spaces.
+  thus intermediate to the $Q_1^d$ and $Q_2^d$ spaces.
 
@@
   interface to \texttt{boost::geometry::index::rtree} has been added
-  that simplifies the construction of spacial indices based on points,
+  that simplifies the construction of spatial indices based on points,
   bounding boxes, or segments.
@@
 full support for ``white-listed'' automatically differentiable (AD) number types 
-from the ADOL-C and Sacado libraries.
+from the ADOL-C and Sacado libraries. This included the classes that
+represent the local contributions of one cell to the global linear
+system (i.e., \texttt{FullMatrix} and \texttt{Vector}) as well as the
+classes commonly used for the computations at individual quadrature
+points (primarily the \texttt{Tensor} and \texttt{SymmetricTensor}
+classes of various ranks).
 In the current release we have provided a unified interface to these
-AD libraries focussing two specific use contexts, namely
+AD libraries, focussing on two specific use contexts:
 \begin{enumerate}
-\item the construction and linearization of finite element residuals, and
-\item the construction and linearization of constitutive model kinetic variables.
+\item The construction and linearization of finite element residuals; and
+\item The construction and linearization of constitutive model kinetic variables.
 \end{enumerate}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Dedicated support for symbolic algebra, including algebra differentiation}
+\label{subsec:sd}
+To complement the automatic differentiation features in \dealii{}, this release sees
+the first step towards integrating and supporting a highly performant computer algebra 
+system (CAS) via the SymEngine library.
+\marginpar{Could use a reference to SymEngine}
+The \texttt{Expression} class in the namespace \texttt{Differentiation::SD} interfaces
+to SymEngine and forms the basis of symbolic computations, offering a full set of 
+overloaded operators and a C++ style interface. This class offers the following 
+basic functionality:
+\begin{itemize}
+\item symbolic variable definition,
+\item symbolic function definition,
+\item expression creation using standard C++ syntax,
+\item expression parsing,
+\item comparison operations,
+\item logical operations,
+\item conditional expression construction,
+\item differentiation,
+\item substitution (partial and complete), and
+\item serialization.
+\end{itemize}
+\dealii{} now also provides an extensive set of math operations, with a syntax mimicking that used 
+in the C++ standard library. Using the \texttt{Expression} class as a basis, we 
+have developed a set of functions that can be used to create \dealii{} \texttt{Tensor}s
+and \texttt{SymmetricTensor}s of symbolic variables and symbolic functions. This
+gives full symbolic tensor algebra support using the pre-existing \texttt{Tensor}
+and \texttt{SymmetricTensor} classes and associated functions. We have
+also implemented a set of utility functions with the following features:
+\begin{itemize}
+\item \texttt{differentiate} scalar expressions, tensor and symmetric tensors of expressions
+with respect to other scalar expressions, tensor and symmetric tensors of expressions;
+\item create symbolic substitution maps;
+\item resolve explicit dependencies between expressions; and
+\item perform scalar and tensor valued substitution (including conversion from symbolic to 
+real-valued scalars and tensors).
+\end{itemize}
+
+In the next release we expect to implement classes to assist in performing assembly operations 
+in the same spirit as that which has been done in the \texttt{Differentiation::AD} namespace,
+although in a fully symbolic manner.
+We will also address performance issues of the \texttt{Expression} class by leveraging
+the optimization capabilities of SymEngine, including common subexpression elimination (CSE), 
+as well as by generating high performance code-paths to evaluate these expressions through the
+use of a custom-generated \texttt{std::function} (so-called ``lambda'' optimization) or by 
+compiling expressions using the LLVM JIT compiler.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Full support for $hp$ adaptivity in parallel computations}
@@
   cells), and for which the data structures stored on different
-  processors have to reconciled.
+  processors have to be reconciled.
 
 \item Algorithms: Already for $h$ adaptive meshes, enumerating all
-  degrees of freedom on the global mesh is difficult as evidenced by
+  degrees of freedom on the global mesh is difficult, as evidenced by
   the complications of the algorithms shown in Section 3.1 of
   \cite{BangerthBursteddeHeisterKronbichler11}, which requires more
-  than a page of text and is implemented in many hundreds of lines of
-  code).
+  than one page of text and is implemented in many hundreds of lines of
+  code.
 
@@
   that eliminates one of these degrees of freedom in favor of another,
-  but in this introduces a ``master'' and a ``slave'' side of the
+  but it introduces a ``master'' and a ``slave'' side of the
   interface. This is of no major consequence in sequential
@@
   interpolating the solution from one mesh to the next,
-  adaptively-refined mesh, or when adapting the polynomial degrees
+  adaptively-refined mesh; or when adapting the polynomial degrees
   associated with each cell and repartitioning in order to
@@
 \item Balancing computational cost: For $h$-adaptive algorithms, the
-  amount of work associated with each cell is essentially the same
+  amount of work associated with each cell is essentially the same,
   both during the assembly of linear systems as well as during the
@@
 
+\marginpar{Missing}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-The GPU support was significantly extended for the current release:
+GPU support was significantly extended for the current release:
 \begin{itemize}
-  \item \texttt{CUDAWrappers::PreconditionILU} and \texttt{CUDAWrappers::PreconditionIC}
-    can be used for preconditioning \texttt{CUDAWrappers::SparseMatrix} objects.
+  \item The \texttt{CUDAWrappers::PreconditionILU} and \texttt{CUDAWrappers::PreconditionIC}
+    classes can now be used for preconditioning \texttt{CUDAWrappers::SparseMatrix} objects.
   \item \texttt{LinearAlgebra::distributed::Vector}: the MPI-parallel vector
@@
     either be \texttt{Host} or \texttt{CUDA}. In the latter case, the data
-    resides in the GPU memory. By default, the template parameter is
+    resides in GPU memory. By default, the template parameter is
     \texttt{Host} and the behavior is unchanged compared to previous versions.
@@
     \cite{ljungkvist2017}. With this addition, both Dirichlet
-    boundary conditions and the constraints arising for adaptively
-    refined meshes can be imposed with in the matrix-free framework. The only restriction is that for two-dimensional meshes the
-    finite element degree must be odd.  There is no such restriction in three
+    boundary conditions and the constraints arising from adaptively
+    refined meshes can be imposed within the matrix-free framework. The only restriction is that for two-dimensional meshes the
+    finite element degree must be odd.  
+    \marginpar{Can we explain this restriction? It seems odd (pun intended)...}
+    There is no such restriction in three
     dimensions.
@@
 \label{subsec:gmg}
-\marginpar{performance optimizations, new features,
+
+\marginpar{Missing: performance optimizations, new features,
 largest runs now at XYZ (Kronbichler?)}
@@
 \label{subsec:steps}
+
+\marginpar{Missing}
 
@@
 should not be visible to typical user codes; some remove previously
-deprecated classes and functions, and the majority change internal
+deprecated classes and functions; and the majority change internal
 interfaces that are not usually used in external
@@
       vector subset into a \texttt{Vector} or using a \texttt{BlockVector}.
-\item The function \texttt{Subscriptor::subscribe} now requires a pointer to a
+\item The function \texttt{Subscriptor::subscribe()}, used through the
+  \texttt{SmartPointer} class, now requires a pointer to a
       \texttt{std::atomic<bool>} that tracks whether or not the pointer to the
@@
 the current version of the library, \textbf{please reference the present
-document}. For up to date information and bibtex snippets for this document
+document}. For up to date information and a bibtex entry for this document
 see:
@@
    \cite{TKB16};
-   \item Fur uses of the \texttt{ParameterAcceptor} concept,
-     \texttt{MeshWorker::ScratchData}, and
-     \texttt{ParsedConvergenceTable}: \cite{SartoriGiulianiBardelloni-2018-a}.
+   \item For uses of the \texttt{ParameterAcceptor} concept, the
+     \texttt{MeshWorker::ScratchData} base class, and the
+     \texttt{ParsedConvergenceTable} class: \cite{SartoriGiulianiBardelloni-2018-a}.
 \end{itemize}

### 2019-06-03 9.1/paper.tex
Document the new tutorial programs.

@@
 
-\marginpar{Missing}
+Many of the \dealii{} tutorial programs were substantially revised as
+part of this release. In particular, we have converted many places
+that now allow for simpler code through the use of C++11 features such
+as range-based for loops and lambda functions.
 
-talk about the new programs; also mention updating for C++11 style
+In addition, there are four new tutorial programs and one new code
+gallery program:
+\begin{itemize}
+\item \texttt{step-61} is a program that implements the ``weak
+  Galerkin'' method to solve the Laplace equation. Weak Galerkin
+  methods are related to the Hybridized Discontinuous Galerkin method
+  in that they introduce degrees of freedom on the interfaces between
+  cells, but they do not require the reformulation of the problem as a
+  first-order system and instead re-define what the gradient of a
+  discontinuous function is.
 
-\texttt{step-64} demonstrates the usage of matrix-free methods on Nvidia GPUs.
+\item \texttt{step-62} demonstrates the solution of problems related
+  to phononic or photonic crystals. Among the techniques shown in this
+  program is the solution of complex-valued linear systems, and the
+  use of absorbing boundary conditions through the Perfectly Matched
+  Layer technique.
+
+\item \texttt{step-63} implements a multigrid preconditioner
+  and solver for the advection-diffusion equation, yielding optimal
+  complexity.
+
+\item \texttt{step-64} demonstrates the usage of matrix-free methods on Nvidia GPUs.
 GPUs are advantageous for these kind of operations because of their superior
@@
 CPUs within a given power envelope.
+
+\item The \texttt{MCMC-Laplace} code gallery program is a code useful
+  for the forward solution that is used as a building block in
+  Bayesian inverse problems, and for sampling the parameter space
+  through a Metropolis-Hastings sampler (a kind of Monte Carlo
+  Markov Chain method).
+\end{itemize}
 

### 2019-06-03 9.1/paper.tex
Update an acknowledgment.

@@
 in Geodynamics initiative (CIG), through the National Science
-Foundation under Awards No.~EAR-0949446 and EAR-1550901 and The
+Foundation under Award No.~EAR-1550901 and The
 University of California -- Davis.

### 2019-06-04 9.1/paper.tex
Remove an old funding acknowledgement
Reply to Wolfgang's comment

@@
     \marginpar{Can we explain this restriction? It seems odd (pun intended)...}
+    \marginpar{BT: No, we cannot. I have spent at least two full weeks trying to
+    understand why it doesn't work, i.e., gives the wrong result, but I don't know. We decided at the time that
+    even with the current limitation this was a useful addition.}
     There is no such restriction in three
@@
 
-J-P.~Pelteret was supported by the European Research Council (ERC) through
-the Advanced Grant 289049 MOCOPOLY.
-
 B.~Turcksin: Research sponsored by the Laboratory Directed Research and

### 2019-06-04 9.1/paper.tex
Added my name and affiliation to author list.

@@
     Martin Kronbichler,
+    Ross Maguire Kynch,
     Matthias Maier,
@@
 
-\author[7]{Matthias~Maier}
-\affil[7]{Department of Mathematics,
+\author[7]{Ross~Maguire~Kynch}
+\affil[7]{Renishaw Plc,
+  Miskin, CF72 8XY, Wales, UK.
+  {\texttt{rkynch@gmail.com}}}
+
+\author[8]{Matthias~Maier}
+\affil[8]{Department of Mathematics,
   Texas A\&M University,

### 2019-06-05 9.1/paper.tex
Add Graham Harper to name and affiliation list

@@
     Denis Davydov,
+    Graham Harper
     Timo Heister,
@@
   Collins, CO 80523-1874, USA.
-  {\texttt{bangerth@colostate.edu}}}
+  {\texttt{bangerth@colostate.edu}, 
+   \texttt{harper@math.colostate.edu}}}
 
@@
   {\texttt{\{denis.davydov,jean-paul.pelteret\}@fau.de}}}
+
+\author[2]{Graham Harper}
 

### 2019-06-05 9.1/paper.tex
Correct a typo in the FE_BernardiRaugel summary

@@
   that the $Q_1^d\times Q_0$ combination is not inf-sup stable (requiring a
-  larger velocity space), and that the $Q_2^d\times Q_1$ combination is
+  larger velocity space), and that the $Q_2^d\times Q_0$ combination is
   stable but sub-optimal since the velocity space is too large relative to

### 2019-06-05 9.1/paper.tex
Fixup comma

@@
     Denis Davydov,
-    Graham Harper
+    Graham Harper,
     Timo Heister,

### 2019-06-05 9.1/paper.tex
Added Marc Fehling as co-author.

--- 2019-06-05 9.1/paper.tex
@@
     Denis Davydov,
-    Graham Harper
+    Marc Fehling,
+    Graham Harper,
     Timo Heister,
@@
 
+\author[4]{Marc~Fehling}
+\affil[4]{Institute for Advanced Simulation,
+  Forschungszentrum J\"{u}lich GmbH,
+  52425 J\"{u}lich, Germany.
+  {\texttt{m.fehling@fz-juelich.de}}}
+
 \author[2]{Graham Harper}
 
-\author[4]{Timo~Heister}
-\affil[4]{Mathematical Sciences,
+\author[5]{Timo~Heister}
+\affil[5]{Mathematical Sciences,
   O-110 Martin Hall,
@@
 %
-\author[5]{Luca~Heltai}
-\affil[5]{SISSA,
+\author[6]{Luca~Heltai}
+\affil[6]{SISSA,
   International School for Advanced Studies,
@@
 
-\author[6]{Martin~Kronbichler}
-\affil[6]{Institute for Computational Mechanics,
+\author[7]{Martin~Kronbichler}
+\affil[7]{Institute for Computational Mechanics,
   Technical University of Munich,
@@
 
-\author[7]{Ross~Maguire~Kynch}
-\affil[7]{Renishaw Plc,
+\author[8]{Ross~Maguire~Kynch}
+\affil[8]{Renishaw Plc,
   Miskin, CF72 8XY, Wales, UK.
@@
 
-\author[8]{Matthias~Maier}
-\affil[8]{Department of Mathematics,
+\author[9]{Matthias~Maier}
+\affil[9]{Department of Mathematics,
   Texas A\&M University,
@@
 
-\author[4]{Jean-Paul~Pelteret}
+\author[3]{Jean-Paul~Pelteret}
 
@@
 
-\author[9]{David Wells}
-\affil[9]{Department of Mathematics, University of North Carolina,
+\author[10]{David Wells}
+\affil[10]{Department of Mathematics, University of North Carolina,
   Chapel Hill, NC 27516, USA.
@@
   Nivesh Dommaraju,
-  Marc Fehling,
   Niklas Fehn,

### 2019-06-05 9.1/paper.tex
Reviewed section on hp adaptivity.

@@
     Section~\ref{subsec:sd}),
-  \item Full support for $hp$ adaptivity in parallel computations (see
+  \item Full support for $hp$~adaptivity in parallel computations (see
     Section~\ref{subsec:hp}),
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Full support for $hp$ adaptivity in parallel computations}
+\subsection{Full support for $hp$~adaptivity in parallel computations}
 \label{subsec:hp}
 
-\dealii{} has had support for $hp$ adaptive methods since around 2005
+\dealii{} has had support for $hp$~adaptive methods since around 2005
 (documented in \cite{BangerthKayserHerold2007}) and for parallel
@@
 \item Data structures: The data structures necessary to store the
-  indices of degrees of freedom are necessarily substantially more
-  complicated for $hp$ algorithms than for the $h$-adaptive schemes
+  indices of degrees of freedom are substantially more
+  complicated for $hp$~algorithms than for the $h$~adaptive schemes
   that were already implemented. This is because the number of degrees
@@
 
-\item Algorithms: Already for $h$ adaptive meshes, enumerating all
-  degrees of freedom on the global mesh is difficult, as evidenced by
+\item Algorithms: Already for $h$~adaptive meshes, enumerating all
+  degrees of freedom uniquely on the global mesh is difficult, as evidenced by
   the complications of the algorithms shown in Section 3.1 of
@@
 
-  These difficulties are even more pronounced when using $hp$
-  adaptivity. The main obstacle is the desire to unify the indices of
-  matching degrees of freedom on adjacent cells, such as the edge
-  degree of freedom of a $Q_2$ element with the middle one of
-  the three edge degrees of freedom of a $Q_4$ element on a
-  neighboring cell. Section 4.2 of
-  \cite{BangerthKayserHerold2007} discusses a sequential algorithm
+  These difficulties are even more pronounced when using $hp$~adaptivity.
+  The main obstacle is the desire to unify the indices of
+  matching degrees of freedom on adjacent cells whenever elements with
+  continuous polynomials are used. For example, the edge degree of freedom
+  of a $Q_2$~element has to be merged with the middle one of the three
+  edge degrees of freedom of a $Q_4$~element on a neighboring cell.
+  Section 4.2 of \cite{BangerthKayserHerold2007} discusses a sequential algorithm
   that eliminates one of these degrees of freedom in favor of another,
@@
   cells during mesh repartitioning. This happens, for example, when
-  interpolating the solution from one mesh to the next,
-  adaptively-refined mesh; or when adapting the polynomial degrees
+  interpolating the solution from one mesh to the next
+  adaptively refined mesh; or when adapting the polynomial degrees
   associated with each cell and repartitioning in order to
   balance the computational cost of each processor's partition. When
-  using $h$ adaptive methods, the amount of data associated with each
+  using $h$~adaptive methods, the amount of data associated with each
   cell is fixed and the algorithms that implement the data transfer
-  are consequently relatively simple. On the other hand, in $hp$
-  contexts, each cell may have a different number of unknowns
+  are consequently relatively simple. On the other hand, in $hp$~contexts,
+  each cell may have a different number of unknowns
   associated with it, and the algorithms that transfer the data are
-  substantially more complicated. Furthermore, the amount of data
-  associated with each cell may be large on cells with higher
-  polynomial degrees, and might profit from compression before
-  sending.
+  substantially more complicated. In order to implement those, we rely
+  on recent enhancements of the \pfrst~library (documented in
+  \cite{Burstedde2018}) to transfer data of variable size across
+  processors. Furthermore, the amount of data associated with each
+  cell may be large on cells with higher polynomial degrees, and
+  might profit from compression before sending.
 
-\item Balancing computational cost: For $h$-adaptive algorithms, the
+\item Balancing computational cost: For $h$~adaptive algorithms, the
   amount of work associated with each cell is essentially the same,
   both during the assembly of linear systems as well as during the
-  solver phase. For $hp$-adaptive methods, this is no longer the
+  solver phase. For $hp$~adaptive methods, this is no longer the
   case. Consequently, balancing the cost of work between different
@@
   cost of the linear solver associated with this cell, leading to
-  difficult trade-offs in defining optimal weighting factors.
+  difficult trade-offs in defining optimal weighting factors. In this
+  release, we supplied the basic functionality to attach any amount
+  of weighting factors to cells, but users still have to find reasonable
+  weights for themselves.
 \end{enumerate}
@@
  \item For distributed parallel computing: \cite{BangerthBursteddeHeisterKronbichler11};
- \item For $hp$ adaptivity: \cite{BangerthKayserHerold2007};
+ \item For $hp$~adaptivity: \cite{BangerthKayserHerold2007};
   \item For partition-of-unity (PUM) and enrichment methods of the
@@
 \item OpenCASCADE \cite{opencascade-web-page}
-\item p4est \cite{p4est}
+\item p4est \cite{Burstedde2018,p4est}
 \item PETSc \cite{petsc-user-ref,petsc-web-page}

### 2019-06-05 9.1/paper.tex
Add Daniel Garcia-Sanchez to the author list

@@
 \usepackage{xspace}
+
+\usepackage[normalem]{ulem}
 
@@
     Marc Fehling,
+    Daniel Garcia-Sanchez,
     Graham Harper,
@@
 
+\author[5]{Daniel Garcia-Sanchez}
+\affil[5]{Sorbonne Universit\'es, UPMC Univ.\ Paris 06, CNRS-UMR 7588,
+  Institut des NanoSciences de Paris, F-75005, Paris, France
+  {\texttt{daniel.garcia-sanchez@insp.upmc.fr}}}
+
 \author[2]{Graham Harper}
 
-\author[5]{Timo~Heister}
-\affil[5]{Mathematical Sciences,
+\author[6]{Timo~Heister}
+\affil[6]{Mathematical Sciences,
   O-110 Martin Hall,
@@
 %
-\author[6]{Luca~Heltai}
-\affil[6]{SISSA,
+\author[7]{Luca~Heltai}
+\affil[7]{SISSA,
   International School for Advanced Studies,
@@
 
-\author[7]{Martin~Kronbichler}
-\affil[7]{Institute for Computational Mechanics,
+\author[8]{Martin~Kronbichler}
+\affil[8]{Institute for Computational Mechanics,
   Technical University of Munich,
@@
 
-\author[8]{Ross~Maguire~Kynch}
-\affil[8]{Renishaw Plc,
+\author[9]{Ross~Maguire~Kynch}
+\affil[9]{Renishaw Plc,
   Miskin, CF72 8XY, Wales, UK.
@@
 
-\author[9]{Matthias~Maier}
-\affil[9]{Department of Mathematics,
+\author[10]{Matthias~Maier}
+\affil[10]{Department of Mathematics,
   Texas A\&M University,
@@
 
-\author[10]{David Wells}
-\affil[10]{Department of Mathematics, University of North Carolina,
+\author[11]{David Wells}
+\affil[11]{Department of Mathematics, University of North Carolina,
   Chapel Hill, NC 27516, USA.
@@
   larger velocity space), and that the $Q_2^d\times Q_0$ combination is
-  stable but sub-optimal since the velocity space is too large relative to
-  the pressure space to provide additional accuracy commensurate with the
-  cost of the large number of velocity unknowns. The BR space is
-  thus intermediate to the $Q_1^d$ and $Q_2^d$ spaces.
+  stable but converges with only first-order at the cost of the large
+  number of velocity unknowns. The BR space is thus intermediate to the
+  $Q_1^d$ and $Q_2^d$ spaces.
 
@@
 the first step towards integrating and supporting a highly performant computer algebra
-system (CAS) via the SymEngine library.
-\marginpar{Could use a reference to SymEngine}
+system (CAS) via the SymEngine library \cite{symengine-web-page}.
 The \texttt{Expression} class in the namespace \texttt{Differentiation::SD} interfaces
@@
   \item Constrained degrees of freedom: the matrix-free framework now
-    supports constrained degrees of freedom. The implementation is based on the algorithms described in
-    \cite{ljungkvist2017}. With this addition, both Dirichlet
-    boundary conditions and the constraints arising from adaptively
-    refined meshes can be imposed within the matrix-free framework. The only restriction is that for two-dimensional meshes the
-    finite element degree must be odd.
-    \marginpar{Can we explain this restriction? It seems odd (pun intended)...}
-    \marginpar{BT: No, we cannot. I have spent at least two full weeks trying to
-    understand why it doesn't work, i.e., gives the wrong result, but I don't know. We decided at the time that
-    even with the current limitation this was a useful addition.}
-    There is no such restriction in three
-    dimensions.
+    supports constrained degrees of freedom. The implementation is based on
+    the algorithms described in \cite{ljungkvist2017}. With this addition,
+    both Dirichlet boundary conditions and the constraints arising from
+    adaptively refined meshes can be imposed within the matrix-free
+    framework.
+    \textcolor{red}{%
+    \sout{The only restriction is that for two-dimensional meshes the
+    finite element degree must be odd. There is no such restriction in
+    three dimensions.}
+    \emph{%
+    WB: Can we explain this restriction? It seems odd (pun intended)...
+    BT: No, we cannot. I have spent at least two full weeks trying to
+    understand why it doesn't work, i.e., gives the wrong result, but I
+    don't know. We decided at the time that even with the current
+    limitation this was a useful addition.
+    MM: I suggest that we then simply not mention the limitation and treat
+    it as a known bug. As it is currently worded it can be easily
+    interpreted as a fundamental limitation, i.\,e. ``FE with even degree
+    are not admissible.''}}
+
   \item MPI matrix-free computations: using \texttt{LinearAlgebra::distributed::Vector}, the
@@
 
-\marginpar{Missing: performance optimizations, new features,
-largest runs now at XYZ (Kronbichler?)}
+\textcolor{red}{Missing: performance optimizations, new features, largest
+runs now at XYZ (Kronbichler?)}
 
@@
 
-\marginpar{I'm sure we have new interfaces. HDF5 and the various AD/SD
-  libraries come to mind. Please update.}
 \dealii can interface with many other libraries:
@@
 \item GSL \cite{gsl2016}
+\item Ginkgo \cite{ginkgo-web-page}
 \item HDF5 \cite{hdf5}
@@
 \item SUNDIALS \cite{sundials}
+\item SymEngine \cite{symengine-web-page}
 \item TBB \cite{Rei07}
@@
 \dealii is a world-wide project with dozens of contributors around the
-globe. Other than the authors of this paper, the following people contributed code to
-this release:\\
-%
-% get this from the changes/*/* files using the command listed in the
-% release-tasks paper and remove the authors of this paper
-%
-% updated 5/12/2019
-\marginpar{Remove all other authors of this paper}
-\marginpar{Do we want to spell out who the Gingko devs are?}
+globe. Other than the authors of this paper, the following people
+contributed code to this release:\\
+% updated 6/05/2019
   Giovanni Alzetta,
@@
   Sambit Das,
-  the Ginkgo developers,
   Stefano Dominici,
@@
   Andreas Fink,
-  Daniel Garcia-Sanchez,
   Rene Gassm{\"o}ller,
@@
   Logan Harbour,
-  Graham Harper,
   Daniel Jodlbauer,
@@
   Dustin Kumor,
-  Ross Kynch,
   Damien Lebrun-Grandie,

### 2019-06-05 9.1/paper.tex
Add punctuation

@@
 
-\dealii version 9.1.0 was released May 21, 2019
+\dealii version 9.1.0 was released May 21, 2019.
 This paper provides an

### 2019-06-05 9.1/paper.tex
some minor changes

--- 2019-06-05 9.1/paper.tex
@@
 \marginpar{I'm sure we have new interfaces. HDF5 and the various AD/SD
-  libraries come to mind. Please update.}
+  libraries come to mind. Please update. MM: The list actually looks pretty
+  complete (I have not verified that we actually cite *all* canonical
+  publications for a given library).}
 \dealii can interface with many other libraries:

### 2019-06-05 9.1/paper.tex
add a reference to the ginkgo project
Same problem here: I haven't found a citable publication
Changed my affiliation to Swansea University where most of the work was undertaken and added support information to the acknowledgements section.

--- 2019-06-05 9.1/paper.tex
@@
 \author[9]{Ross~Maguire~Kynch}
-\affil[9]{Renishaw Plc,
-  Miskin, CF72 8XY, Wales, UK.
+\affil[9]{Zienkiewicz Centre for Computational Engineering,
+  College of Engineering, Swansea University,
+  Bay Campus, Fabian Way, Swansea SA1 8EN, Wales, UK.
   {\texttt{rkynch@gmail.com}}}
@@
 This paper provides an
-overview of the new features of this major release and serves as a citable
+overview of the new features of this release and serves as a citable
 reference for the \dealii software library version 9.1. \dealii is an
@@
     Section~\ref{subsec:hp}),
-  \item Interfaces to the HDF5 file format and libraries (see
+  \item An interface to the HDF5 file format and libraries (see
     Section~\ref{subsec:hdf5}),
@@
 \item
-  The release contains bug fixes and performance improvements of the
-  matrix-free framework and related geometric multigrid solvers. In
-  particular, the implementation of the Chebyshev iteration, an often used
-  smoother in the matrix-free context, has been revised to reduce vector
-  accesses. This speeds up matrix-free multigrid solvers by up to 10-15\%
-  on geometries with affine (parallelogram and parallelpiped) cells, and up
-  to 5\% on geometries with cells bounded by curved edges and faces.
-
-\item
-  Various variants of geometric multigrid solvers and matrix-free
-  implementations were run on up to 304,128 MPI ranks during the acceptance
-  phase of the SuperMUC-NG supercomputer, verifying the scalability of our
-  implementations. Some performance bottlenecks on distributed
-  triangulation with more than 100k ranks have been fixed.
+  The release contains a number of performance improvements and bug fixes for
+  the matrix-free framework. One notable improvement is the support for
+  renumbering of degrees of freedom within the cells for discontinuous
+  elements, avoiding some reshuffling operations across the SIMD lanes
+  with vectorization over several cells and faces, which is especially
+  useful on processors with AVX-512 vectorization (8 doubles), speeding up
+  operations by up to 10\%. Secondly, the strategy for the most efficient
+  tensor product evaluators according to the performance analysis of
+  \cite{KronbichlerKormann2019} in the context of more quadrature points than
+  shape functions has been revised for better performance.
 
@@
 In the current release we have provided a unified interface to these
-AD libraries, focussing on two specific use contexts:
+AD libraries, focusing on two specific use contexts:
 \begin{enumerate}
@@
 the \texttt{ResidualLinearization} class requires the (once-differentiated) finite
-element residual to be defined on a per cell basis,  and this contribution is
+element residual to be defined on a per-cell basis,  and this contribution is
 automatically linearized.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Interfaces to the HDF5 file format and libraries}
+\subsection{Interface to the HDF5 file format and libraries}
 \label{subsec:hdf5}
+\texttt{HDF5} is an open-source library and file format designed to store large amounts of data.
+The \texttt{HDF5} format is specially tailored for high volume parallel I/O operations using MPI.
+\texttt{HDF5} files are self-describing and allow complex data relationships and a large variety
+of datatypes. In addition, the \texttt{HDF5} format is designed for long-term preservation of
+data; a \texttt{HDF5} file created by an HPC system can be easily read by a commodity laptop.
 
-\marginpar{Missing}
+\dealii{}'s new  \texttt{HDF5} interface allows to write \texttt{HDF5} files and manipulate
+datasets, groups, and attributes in serial and in parallel using MPI. The
+\texttt{HDF5} MPI library calls that modify the structure of the file are
+always collective, whereas writing and reading raw data in a dataset can be done
+independently or collectively. In the \dealii{}'s \texttt{HDF5} interface all the calls are
+set to collective in order to maximize performance. This means that all the MPI
+processes have to contribute to every single call, even if they don't have data to write.
+We have added the following classes to the \dealii{}'s \texttt{HDF5} namespace.
+\begin{itemize}
+\item The new \texttt{HDF5::File} class can be used to open and create \texttt{HDF5} files
+in serial or in parallel.
+\item The new \texttt{HDF5::Group} class can be used to open and create groups.
+ \texttt{HDF5} files have a tree structure. The root contains groups and
+the groups can contain other groups.
+\item The new \texttt{HDF5::DataSet} class can be used to open and create datasets which
+can be placed inside groups or at the root of the \texttt{HDF5} file. A dataset can be a
+vector, a matrix, or a tensor. It is possible to read and write in the dataset using
+hyperslabs or unordered data. Hyperslabs are portions of datasets
+which can be a contiguous collection of points in a dataset, or a regular
+pattern of points or blocks in a datataset.
+\item It is possible to define attributes in the \texttt{HDF5::File}, \texttt{HDF5::Group} and
+\texttt{HDF5::DataSet} classes. An attribute is a small metadata such as a number or a string.
+Attributes are commonly used to store simulation parameters.
+\end{itemize}
+As we have shown in \texttt{step-62}, the \dealii{}'s \texttt{HDF5} interface can be easily used
+to exchange data with python and jupyter notebooks.
 
@@
     framework.
-    \textcolor{red}{%
-    \sout{The only restriction is that for two-dimensional meshes the
-    finite element degree must be odd. There is no such restriction in
-    three dimensions.}
-    \emph{%
-    WB: Can we explain this restriction? It seems odd (pun intended)...
-    BT: No, we cannot. I have spent at least two full weeks trying to
-    understand why it doesn't work, i.e., gives the wrong result, but I
-    don't know. We decided at the time that even with the current
-    limitation this was a useful addition.
-    MM: I suggest that we then simply not mention the limitation and treat
-    it as a known bug. As it is currently worded it can be easily
-    interpreted as a fundamental limitation, i.\,e. ``FE with even degree
-    are not admissible.''}}
+%     \textcolor{red}{%
+%     \sout{The only restriction is that for two-dimensional meshes the
+%     finite element degree must be odd. There is no such restriction in
+%     three dimensions.}
+%     \emph{%
+%     WB: Can we explain this restriction? It seems odd (pun intended)...
+%     BT: No, we cannot. I have spent at least two full weeks trying to
+%     understand why it doesn't work, i.e., gives the wrong result, but I
+%     don't know. We decided at the time that even with the current
+%     limitation this was a useful addition.
+%     MM: I suggest that we then simply not mention the limitation and treat
+%     it as a known bug. As it is currently worded it can be easily
+%     interpreted as a fundamental limitation, i.\,e. ``FE with even degree
+%     are not admissible.''}}
 
@@
 
-\textcolor{red}{Missing: performance optimizations, new features, largest
-runs now at XYZ (Kronbichler?)}
+For the 9.1 release, the geometric multigrid facilities have been extended and
+revised for performance. The geometric multigrid algorithms for uniform and
+adaptively refined meshes in \dealii are based on so-called local coarsening,
+i.e., smoothening is applied level-per-level, skipping parts of the domain
+where the mesh is not as refined. The algorithm for the assignment of the
+owner on level cells and the implications on load balancing have been analyzed
+in detail in \cite{ClevengerHeisterKanschatKronbichler2019}. While most of the
+functionality was already available since the 8.5 release of \dealii presented
+in \cite{dealII85}, several components have been finalized, such as the
+support for certain renumbering algorithms that are beneficial for matrix-free
+execution, and interfaces that allow the combination with matrix-free GPU
+computations according to \cite{KronbichlerLjungkvist2019}.
+
+A number of data structures and implementations in \dealii have been adapted
+to ensure scalability of the matrix-free algorithms and geometric multigrid
+infrastructure on more than 100,000 MPI ranks. Geometric multigrid solver for
+the Poisson equation as described in \cite{KronbichlerWall2018} have been used
+as performance tests during the acceptance phase of the SuperMUC-NG
+supercomputer in Garching, Germany. Scaling tests have been performed on up to
+the full machine with 304,128 cores of the Intel Xeon Skylake architecture and
+an arithmetic performance of around 5 PFlop/s for a geometric multigrid solver
+with polynomials of degree 4 has been reached. Compared to the official
+LINPACK performance of the machine of 19.5 PFlop/s (the machine is listed on
+position 8 of the top-500 list of November 2018), this can be considered an
+extremely good value for a PDE solver which have classically only reached a
+few percent of the LINPACK performance, and given the optimization steps
+contained in the matrix-free module of \dealii that aim to minimize the time
+to solution rather than maximize the number of floating point operations. The
+largest Poisson problem that has been solved on 304k cores contained 2.15
+trillion unknowns (or 7.1 million unknowns per MPI rank) and was solved in 3.5
+seconds. Also, CFD production runs with up to $10^{11}$ unknowns and $10^5$
+time steps have been completed in less than seven hours, demonstrating the
+capabilities of \dealii for large-scale parallel computations. The scaling tests
+also revealed several relatively expensive operations in the setup of the
+multigrid unknowns in \dealii's \texttt{DoFHandler} and \texttt{MGTransfer}
+classes. While a few bottlenecks have already been resolved, we plan several
+further improvements in setup times for the next release.
+
+Furthermore, the implementation of the Chebyshev iteration, \dealii's most
+popular smoother in the matrix-free context, has been revised to reduce the
+vector accesses. The vector operations have become an increasing bottleneck
+due to the level of optimization available for the operator evaluation with
+our matrix-free framework \cite{KronbichlerKormann2019}, especially on newer
+CPUs which can perform a large number of arithmetic operations per byte loaded
+from main memory. This speeds up matrix-free multigrid solvers by up to
+10--15\% on geometries with affine (parallelogram and parallelpiped) cells,
+and up to 5\% on deformed cells.
 
@@
 
+M.~Kronbichler was also supported by the Bayerisches Kompetenznetzwerk
+f\"ur Technisch-Wissen\-schaft\-li\-ches Hoch- und H\"ochstleistungsrechnen
+(KONWIHR) in the context of the project
+``Performance tuning of high-order discontinuous Galerkin solvers for
+SuperMUC-NG''.
+
+R.~M.~Kynch was supported by the Engineering and Physical Science Research
+Council (EPSRC) UK through grant EP/K023950/1 while working at Swansea
+University, UK 2013-2015 where the majority of his contribution was
+undertaken.
+
 M.~Maier was partially supported by ARO MURI Award No. W911NF-14-0247.

### 2019-06-06 9.1/paper.tex
Update FE_BernardiRaugel summary

--- 2019-06-05 9.1/paper.tex
@@
 \usepackage{xspace}
-
-\usepackage[normalem]{ulem}
 
@@
     Denis Davydov,
-    Marc Fehling,
-    Graham Harper,
+    Graham Harper
     Timo Heister,
@@
 
-\author[4]{Marc~Fehling}
-\affil[4]{Institute for Advanced Simulation,
-  Forschungszentrum J\"{u}lich GmbH,
-  52425 J\"{u}lich, Germany.
-  {\texttt{m.fehling@fz-juelich.de}}}
-
 \author[2]{Graham Harper}
 
-\author[5]{Timo~Heister}
-\affil[5]{Mathematical Sciences,
+\author[4]{Timo~Heister}
+\affil[4]{Mathematical Sciences,
   O-110 Martin Hall,
@@
 %
-\author[6]{Luca~Heltai}
-\affil[6]{SISSA,
+\author[5]{Luca~Heltai}
+\affil[5]{SISSA,
   International School for Advanced Studies,
@@
 
-\author[7]{Martin~Kronbichler}
-\affil[7]{Institute for Computational Mechanics,
+\author[6]{Martin~Kronbichler}
+\affil[6]{Institute for Computational Mechanics,
   Technical University of Munich,
@@
 
-\author[8]{Ross~Maguire~Kynch}
-\affil[8]{Renishaw Plc,
+\author[7]{Ross~Maguire~Kynch}
+\affil[7]{Renishaw Plc,
   Miskin, CF72 8XY, Wales, UK.
@@
 
-\author[9]{Matthias~Maier}
-\affil[9]{Department of Mathematics,
+\author[8]{Matthias~Maier}
+\affil[8]{Department of Mathematics,
   Texas A\&M University,
@@
 
-\author[3]{Jean-Paul~Pelteret}
+\author[4]{Jean-Paul~Pelteret}
 
@@
 
-\author[10]{David Wells}
-\affil[10]{Department of Mathematics, University of North Carolina,
+\author[9]{David Wells}
+\affil[9]{Department of Mathematics, University of North Carolina,
   Chapel Hill, NC 27516, USA.
@@
     Section~\ref{subsec:sd}),
-  \item Full support for $hp$~adaptivity in parallel computations (see
+  \item Full support for $hp$ adaptivity in parallel computations (see
     Section~\ref{subsec:hp}),
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Full support for $hp$~adaptivity in parallel computations}
+\subsection{Full support for $hp$ adaptivity in parallel computations}
 \label{subsec:hp}
 
-\dealii{} has had support for $hp$~adaptive methods since around 2005
+\dealii{} has had support for $hp$ adaptive methods since around 2005
 (documented in \cite{BangerthKayserHerold2007}) and for parallel
@@
 \item Data structures: The data structures necessary to store the
-  indices of degrees of freedom are substantially more
-  complicated for $hp$~algorithms than for the $h$~adaptive schemes
+  indices of degrees of freedom are necessarily substantially more
+  complicated for $hp$ algorithms than for the $h$-adaptive schemes
   that were already implemented. This is because the number of degrees
@@
 
-\item Algorithms: Already for $h$~adaptive meshes, enumerating all
-  degrees of freedom uniquely on the global mesh is difficult, as evidenced by
+\item Algorithms: Already for $h$ adaptive meshes, enumerating all
+  degrees of freedom on the global mesh is difficult, as evidenced by
   the complications of the algorithms shown in Section 3.1 of
@@
 
-  These difficulties are even more pronounced when using $hp$~adaptivity.
-  The main obstacle is the desire to unify the indices of
-  matching degrees of freedom on adjacent cells whenever elements with
-  continuous polynomials are used. For example, the edge degree of freedom
-  of a $Q_2$~element has to be merged with the middle one of the three
-  edge degrees of freedom of a $Q_4$~element on a neighboring cell.
-  Section 4.2 of \cite{BangerthKayserHerold2007} discusses a sequential algorithm
+  These difficulties are even more pronounced when using $hp$
+  adaptivity. The main obstacle is the desire to unify the indices of
+  matching degrees of freedom on adjacent cells, such as the edge
+  degree of freedom of a $Q_2$ element with the middle one of
+  the three edge degrees of freedom of a $Q_4$ element on a
+  neighboring cell. Section 4.2 of
+  \cite{BangerthKayserHerold2007} discusses a sequential algorithm
   that eliminates one of these degrees of freedom in favor of another,
@@
   cells during mesh repartitioning. This happens, for example, when
-  interpolating the solution from one mesh to the next
-  adaptively refined mesh; or when adapting the polynomial degrees
+  interpolating the solution from one mesh to the next,
+  adaptively-refined mesh; or when adapting the polynomial degrees
   associated with each cell and repartitioning in order to
   balance the computational cost of each processor's partition. When
-  using $h$~adaptive methods, the amount of data associated with each
+  using $h$ adaptive methods, the amount of data associated with each
   cell is fixed and the algorithms that implement the data transfer
-  are consequently relatively simple. On the other hand, in $hp$~contexts,
-  each cell may have a different number of unknowns
+  are consequently relatively simple. On the other hand, in $hp$
+  contexts, each cell may have a different number of unknowns
   associated with it, and the algorithms that transfer the data are
-  substantially more complicated. In order to implement those, we rely
-  on recent enhancements of the \pfrst~library (documented in
-  \cite{Burstedde2018}) to transfer data of variable size across
-  processors. Furthermore, the amount of data associated with each
-  cell may be large on cells with higher polynomial degrees, and
-  might profit from compression before sending.
+  substantially more complicated. Furthermore, the amount of data
+  associated with each cell may be large on cells with higher
+  polynomial degrees, and might profit from compression before
+  sending.
 
-\item Balancing computational cost: For $h$~adaptive algorithms, the
+\item Balancing computational cost: For $h$-adaptive algorithms, the
   amount of work associated with each cell is essentially the same,
   both during the assembly of linear systems as well as during the
-  solver phase. For $hp$~adaptive methods, this is no longer the
+  solver phase. For $hp$-adaptive methods, this is no longer the
   case. Consequently, balancing the cost of work between different
@@
   cost of the linear solver associated with this cell, leading to
-  difficult trade-offs in defining optimal weighting factors. In this
-  release, we supplied the basic functionality to attach any amount
-  of weighting factors to cells, but users still have to find reasonable
-  weights for themselves.
+  difficult trade-offs in defining optimal weighting factors.
 \end{enumerate}
@@
   \item Constrained degrees of freedom: the matrix-free framework now
-    supports constrained degrees of freedom. The implementation is based on
-    the algorithms described in \cite{ljungkvist2017}. With this addition,
-    both Dirichlet boundary conditions and the constraints arising from
-    adaptively refined meshes can be imposed within the matrix-free
-    framework.
-    \textcolor{red}{%
-    \sout{The only restriction is that for two-dimensional meshes the
-    finite element degree must be odd. There is no such restriction in
-    three dimensions.}
-    \emph{%
-    WB: Can we explain this restriction? It seems odd (pun intended)...
-    BT: No, we cannot. I have spent at least two full weeks trying to
-    understand why it doesn't work, i.e., gives the wrong result, but I
-    don't know. We decided at the time that even with the current
-    limitation this was a useful addition.
-    MM: I suggest that we then simply not mention the limitation and treat
-    it as a known bug. As it is currently worded it can be easily
-    interpreted as a fundamental limitation, i.\,e. ``FE with even degree
-    are not admissible.''}}
-
+    supports constrained degrees of freedom. The implementation is based on the algorithms described in
+    \cite{ljungkvist2017}. With this addition, both Dirichlet
+    boundary conditions and the constraints arising from adaptively
+    refined meshes can be imposed within the matrix-free framework. The only restriction is that for two-dimensional meshes the
+    finite element degree must be odd.  
+    \marginpar{Can we explain this restriction? It seems odd (pun intended)...}
+    \marginpar{BT: No, we cannot. I have spent at least two full weeks trying to
+    understand why it doesn't work, i.e., gives the wrong result, but I don't know. We decided at the time that
+    even with the current limitation this was a useful addition.}
+    There is no such restriction in three
+    dimensions.
   \item MPI matrix-free computations: using \texttt{LinearAlgebra::distributed::Vector}, the
@@
 
-\textcolor{red}{Missing: performance optimizations, new features, largest
-runs now at XYZ (Kronbichler?)}
+\marginpar{Missing: performance optimizations, new features,
+largest runs now at XYZ (Kronbichler?)}
 
@@
  \item For distributed parallel computing: \cite{BangerthBursteddeHeisterKronbichler11};
- \item For $hp$~adaptivity: \cite{BangerthKayserHerold2007};
+ \item For $hp$ adaptivity: \cite{BangerthKayserHerold2007};
   \item For partition-of-unity (PUM) and enrichment methods of the
@@
 \marginpar{I'm sure we have new interfaces. HDF5 and the various AD/SD
-  libraries come to mind. Please update. MM: The list actually looks pretty
-  complete (I have not verified that we actually cite *all* canonical
-  publications for a given library).}
+  libraries come to mind. Please update.}
 \dealii can interface with many other libraries:
@@
 \item OpenCASCADE \cite{opencascade-web-page}
-\item p4est \cite{Burstedde2018,p4est}
+\item p4est \cite{p4est}
 \item PETSc \cite{petsc-user-ref,petsc-web-page}
@@
 \dealii is a world-wide project with dozens of contributors around the
-globe. Other than the authors of this paper, the following people
-contributed code to this release:\\
-% updated 6/05/2019
+globe. Other than the authors of this paper, the following people contributed code to
+this release:\\
+%
+% get this from the changes/*/* files using the command listed in the
+% release-tasks paper and remove the authors of this paper
+%
+% updated 5/12/2019
+\marginpar{Remove all other authors of this paper}
+\marginpar{Do we want to spell out who the Gingko devs are?}
   Giovanni Alzetta,
@@
   Sambit Das,
+  the Ginkgo developers,
   Stefano Dominici,
   Nivesh Dommaraju,
+  Marc Fehling,
   Niklas Fehn,
@@
   Logan Harbour,
+  Graham Harper,
   Daniel Jodlbauer,
@@
   Dustin Kumor,
+  Ross Kynch,
   Damien Lebrun-Grandie,

### 2019-06-06 9.1/paper.tex

--- 2019-06-05 9.1/paper.tex
@@
 
-\marginpar{I'm sure we have new interfaces. HDF5 and the various AD/SD
-  libraries come to mind. Please update. MM: The list actually looks pretty
-  complete (I have not verified that we actually cite *all* canonical
-  publications for a given library).}
 \dealii can interface with many other libraries:

### 2019-06-06 9.1/paper.tex
Discuss the HDF5 interface

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Interfaces to the HDF5 file format and libraries}
+\subsection{Interface to the HDF5 file format and libraries}
 \label{subsec:hdf5}
+\texttt{HDF5} is an open-source library and file format designed to store large amounts of data.
+The \texttt{HDF5} format is specially tailored for high volume parallel I/O operations using MPI.
+\texttt{HDF5} files are self-describing and allow complex data relationships and a large variety
+of datatypes. In addition the \texttt{HDF5} format is designed for long-term preservation of
+data; a \texttt{HDF5} file created by a HPC system can be easily read by a commodity laptop.
 
-\marginpar{Missing}
+The new \dealii{}'s \texttt{HDF5} interface allows to write \texttt{HDF5} files and manipulate
+datasets, groups, and attributes in serial and in parallel using MPI. The
+\texttt{HDF5} MPI library calls that modify the structure of the file are
+always collective, whereas writing and reading raw data in a dataset can be done
+independently or collectively. In the \dealii{}'s \texttt{HDF5} interface all the calls are
+set to collective in order to maximize performance. This means that all the MPI
+processes have to contribute to every single call, even if they don't have data to write.
+We have added the following classes to the \dealii{}'s \texttt{HDF5} namespace. 
+\begin{itemize}
+\item The new \texttt{HDF5::File} class can be used to open and create \texttt{HDF5} files
+in serial or in parallel. 
+\item The new \texttt{HDF5::Group} class can be used to open and create groups.
+ \texttt{HDF5} files have a tree structure. The root contains groups and
+the groups can contain other groups.
+\item The new \texttt{HDF5::DataSet} class can be used to open and create datasets which
+can be placed inside groups or at the root of the \texttt{HDF5} file. A dataset can be a
+vector, a matrix, or a tensor. It is possible to read and write in the dataset using
+hyperslabs or unordered data. Hyperslabs are portions of datasets
+which can be a contiguous collection of points in a dataset, or a regular
+pattern of points or blocks in a datataset.
+\item It is possible to define attributes in the \texttt{HDF5::File}, \texttt{HDF5::Group} and
+\texttt{HDF5::DataSet} classes. An attribute is a small metadata such as a number or a string.
+Attributes are commonly used to store simulation parameters.
+\end{itemize}
+As we have shown in \texttt{step-62}, the \dealii{}'s \texttt{HDF5} interface can be easily used
+to exchange data with python and jupyter notebooks.
 

### 2019-06-07 9.1/paper.tex
Fix grammar.

@@
 
-The new \dealii{}'s \texttt{HDF5} interface allows to write \texttt{HDF5} files and manipulate
+\dealii{}'s new  \texttt{HDF5} interface allows to write \texttt{HDF5} files and manipulate
 datasets, groups, and attributes in serial and in parallel using MPI. The

### 2019-06-07 9.1/paper.tex
Add section about geometric multigrid

@@
 \texttt{HDF5} files are self-describing and allow complex data relationships and a large variety
-of datatypes. In addition the \texttt{HDF5} format is designed for long-term preservation of
-data; a \texttt{HDF5} file created by a HPC system can be easily read by a commodity laptop.
+of datatypes. In addition, the \texttt{HDF5} format is designed for long-term preservation of
+data; a \texttt{HDF5} file created by an HPC system can be easily read by a commodity laptop.
 
@@
 
-\textcolor{red}{Missing: performance optimizations, new features, largest
-runs now at XYZ (Kronbichler?)}
+For the 9.1 release, the geometric multigrid facilities have been extended and
+revised for performance. The geometric multigrid algorithms for uniform and
+adaptively refined meshes in \dealii are based on so-called local coarsening,
+i.e., smoothening is applied level-per-level, skipping parts of the domain
+where the mesh is not as refined. The algorithm for the assignment of the
+owner on level cells and the implications on load balancing have been analyzed
+in detail in \cite{ClevengerHeisterKanschatKronbichler2019}. While most of the
+functionality was already available since the 8.5 release of \dealii presented
+in \cite{dealII85}, several components have been finalized, such as the
+support for certain renumbering algorithms that are beneficial for matrix-free
+execution, and interfaces that allow the combination with matrix-free GPU
+computations according to \cite{KronbichlerLjungkvist2019}.
+
+A number of data structures and implementations in \dealii have been adapted
+to ensure scalability of the matrix-free algorithms and geometric multigrid
+infrastructure on more than 100,000 MPI ranks. Geometric multigrid solver for
+the Poisson equation as described in \cite{KronbichlerWall2018} have been used
+as performance tests during the acceptance phase of the SuperMUC-NG
+supercomputer in Garching, Germany. Scaling tests have been performed on up to
+the full machine with 304,128 cores of the Intel Xeon Skylake architecture and
+arithmetic performance of around 5 PFlop/s for a geometric multigrid solver
+with polynomials of degree 4 has been reached. Compared to the official
+LINPACK performance of the machine of 19.5 PFlop (the machine is listed on
+position 8 of the top-500 list of November 2018), this can be considered an
+extremely good value for a PDE solver which have classically only reached a
+few percent of the LINPACK performance, and given the optimization steps
+contained in the matrix-free module of \dealii that aim to minimize the time
+to solution rather than maximize the number of floating point operations. The
+largest Poisson problem that has been solved on 304k cores contained 2.15
+trillion unknowns (or 7.1 million unknowns per MPI rank) and was solved in 3.5
+seconds. Also, CFD production runs with up to $10^{11}$ unknowns and $10^5$
+time steps have been completed in less than seven hours, demonstrating the
+capabilities of \dealii for large-scale parallel computations.
+
+Furthermore, the implementation of the Chebyshev iteration, \dealii's most
+popular smoother in the matrix-free context, has been revised to reduce the
+vector accesses. The vector operations have become an increasing bottleneck
+due to the level of optimization available for the operator evaluation with
+our matrix-free framework \cite{KronbichlerKormann2019}, especially on newer
+CPUs which can perform a large number of arithmetic operations per byte loaded
+from main memory. This speeds up matrix-free multigrid solvers by up to
+10--15\% on geometries with affine (parallelogram and parallelpiped) cells,
+and up to 5\% on deformed cells.
 
@@
 
+M.~Kronbichler was supported by the Bayerisches Kompetenznetzwerk
+f\"ur Technisch-Wissen\-schaft\-li\-ches Hoch- und H\"ochstleistungsrechnen
+(KONWIHR) in the context of the project
+``Performance tuning of high-order discontinuous Galerkin solvers for
+SuperMUC-NG''.
+
 M.~Maier was partially supported by ARO MURI Award No. W911NF-14-0247.

### 2019-06-07 9.1/paper.tex
Fix typo

@@
 the full machine with 304,128 cores of the Intel Xeon Skylake architecture and
-arithmetic performance of around 5 PFlop/s for a geometric multigrid solver
+an arithmetic performance of around 5 PFlop/s for a geometric multigrid solver
 with polynomials of degree 4 has been reached. Compared to the official

### 2019-06-07 9.1/paper.tex
Add a todo
Fix my second funding (due to SuperMUC-NG).

@@
 time steps have been completed in less than seven hours, demonstrating the
-capabilities of \dealii for large-scale parallel computations.
+capabilities of \dealii for large-scale parallel computations. The tests
+also revealed several relatively expensive operations in the setup of the
+multigrid unknowns in \dealii's \texttt{DoFHandler} and \texttt{MGTransfer}
+classes. While a few bottlenecks have already been resolved, we plan several
+further improvements in setup times for the next release.
 
@@
 
-M.~Kronbichler was supported by the Bayerisches Kompetenznetzwerk
+M.~Kronbichler was also supported by the Bayerisches Kompetenznetzwerk
 f\"ur Technisch-Wissen\-schaft\-li\-ches Hoch- und H\"ochstleistungsrechnen

### 2019-06-07 9.1/paper.tex
Fix two more inaccuracies

@@
 with polynomials of degree 4 has been reached. Compared to the official
-LINPACK performance of the machine of 19.5 PFlop (the machine is listed on
+LINPACK performance of the machine of 19.5 PFlop/s (the machine is listed on
 position 8 of the top-500 list of November 2018), this can be considered an
@@
 time steps have been completed in less than seven hours, demonstrating the
-capabilities of \dealii for large-scale parallel computations. The tests
+capabilities of \dealii for large-scale parallel computations. The scaling tests
 also revealed several relatively expensive operations in the setup of the

### 2019-06-07 9.1/paper.tex

@@
   stable but converges with only first-order at the cost of the large
-  number of velocity unknowns. The BR space is thus intermediate to the
+  number of velocity unknowns. The BR space is thus intermediate between the
   $Q_1^d$ and $Q_2^d$ spaces.
@@
 in the file that lists all changes for this release}, see \cite{changes91}.
-(The file is also linked to from the web site of each release as well as
-the release announcement.)
 
@@
 compute the directional derivatives of components of (multi-field) constitutive laws
-with respect to the scalar, vector, tensor and symmetric tensor fields in terms
+with respect to the scalar, vector, tensor, and symmetric tensor fields in terms
 of which they are parameterized. The \texttt{ScalarFunction} class may be used to
@@
 \begin{itemize}
-\item \texttt{differentiate} scalar expressions, tensor and symmetric tensors of expressions
-with respect to other scalar expressions, tensor and symmetric tensors of expressions;
+\item \texttt{differentiate} scalar, tensor, and symmetric tensor expressions
+with respect to other scalar, tensor, or symmetric tensor expressions;
 \item create symbolic substitution maps;
@@
 In the next release we expect to implement classes to assist in performing assembly operations
-in the same spirit as that which has been done in the \texttt{Differentiation::AD} namespace,
+in the same spirit as it is already possible using automatic differentation using the \texttt{Differentiation::AD} namespace,
 although in a fully symbolic manner.
@@
 As we have shown in \texttt{step-62}, the \dealii{}'s \texttt{HDF5} interface can be easily used
-to exchange data with python and jupyter notebooks.
+to exchange data with Python and Jupyter notebooks.
 
@@
     the relevant data to the host, performing MPI communication, and finally
-    move the data to the device or, if CUDA-aware MPI is available, by
+    moving the data to the device or, if CUDA-aware MPI is available, by
     performing MPI communication directly between GPUs.
@@
 extremely good value for a PDE solver which have classically only reached a
-few percent of the LINPACK performance, and given the optimization steps
-contained in the matrix-free module of \dealii that aim to minimize the time
-to solution rather than maximize the number of floating point operations. The
+few percent of the LINPACK performance.
+More importantly, this is achieved within a flexible framework supporting
+arbitrary polynomial order on adaptively refined, unstructured meshes and
+with algorithms in the matrix-free module of \dealii designed to minimize time
+to solution and scalability, rather than maximizing the number of floating point operations. The
 largest Poisson problem that has been solved on 304k cores contained 2.15
@@
 Furthermore, the implementation of the Chebyshev iteration, \dealii's most
-popular smoother in the matrix-free context, has been revised to reduce the
+popular smoother in the matrix-free context, has been revised to reduce the number of
 vector accesses. The vector operations have become an increasing bottleneck
@@
 
-\item \texttt{step-63} implements a multigrid preconditioner
+\item \texttt{step-63} implements a geometric multigrid preconditioner
   and solver for the advection-diffusion equation, yielding optimal
-  complexity.
+  complexity. The tutorial compares
+  point-based smoothers to cell-based smoothers and
+  demonstrates the effect of downstream ordering on smoother performance.
 
 \item \texttt{step-64} demonstrates the usage of matrix-free methods on Nvidia GPUs.
-GPUs are advantageous for these kind of operations because of their superior
+GPUs are shown to be advantageous for these kind of operations because of their superior
 hardware characteristics, in particular a higher memory bandwidth than server
@@
 \item The \texttt{MCMC-Laplace} code gallery program is a code useful
-  for the forward solution that is used as a building block in
+  for the forward solution used as a building block in
   Bayesian inverse problems, and for sampling the parameter space
@@
       subscribed-to object is still valid.
-\item The \texttt{ConstraintMatrix} class gained a template parameter for scalar
+\item The \texttt{ConstraintMatrix} class gained a template parameter for the scalar
       type and was been renamed \texttt{AffineConstraints}. Several methods that
@@
 
-W.~Bangerth and T.~Heister were partially
+W.~Bangerth, T.~C.~Clevenger, and T.~Heister were partially
 supported by the National Science Foundation under award OAC-1835673
 as part of the Cyberinfrastructure for Sustained Scientific Innovation (CSSI)
-program; by award DMS-1821210; and by the Computational Infrastructure
+program  and by the Computational Infrastructure
 in Geodynamics initiative (CIG), through the National Science
@@
 University of California -- Davis.
+
+W.~Bangerth and T.~Heister were also partially supported by award DMS-1821210.
 
@@
 SuperMUC-NG''.
+
+R.~M.~Kynch was supported by the Engineering and Physical Science Research
+Council (EPSRC) UK through grant EP/K023950/1 while working at Swansea
+University, UK 2013-2015 where the majority of his contribution was
+undertaken.
 

### 2019-06-07 9.1/paper.tex
Improve wording of symbolic differentiation ops

@@
 \begin{itemize}
-\item \texttt{differentiate} scalar, tensor, and symmetric tensor expressions
-with respect to other scalar, tensor, or symmetric tensor expressions;
+\item \texttt{differentiate} scalar expressions with respect to other scalar expressions, as well as
+tensors and symmetric tensors of expressions;
+\item \texttt{differentiate} tensors and symmetric tensors of expressions
+with respect to scalar expressions, as well as other tensors and symmetric tensors of expressions;
 \item create symbolic substitution maps;

### 2019-06-08 9.1/paper.tex

--- 2019-06-07 9.1/paper.tex
@@
     Wolfgang Bangerth,
+    Thomas C. Clevenger,
     Denis Davydov,
@@
 
-\author[3]{Denis~Davydov}
-\affil[3]{Chair of Applied Mechanics,
+\author[3]{Thomas~C.~Clevenger}
+\affil[3]{School of Mathematical and Statistical Sciences,
+  Clemson University,
+  Clemson, SC, 29634, USA
+  {\texttt{\{tcleven, heister\}@clemson.edu}}}
+
+\author[4]{Denis~Davydov}
+\affil[4]{Chair of Applied Mechanics,
   Friedrich-Alexander-Universit\"{a}t Erlangen-N\"{u}rnberg,
@@
 
-\author[4]{Marc~Fehling}
-\affil[4]{Institute for Advanced Simulation,
+\author[5]{Marc~Fehling}
+\affil[5]{Institute for Advanced Simulation,
   Forschungszentrum J\"{u}lich GmbH,
@@
 
-\author[5]{Daniel Garcia-Sanchez}
-\affil[5]{Sorbonne Universit\'es, UPMC Univ.\ Paris 06, CNRS-UMR 7588,
+\author[6]{Daniel Garcia-Sanchez}
+\affil[6]{Sorbonne Universit\'es, UPMC Univ.\ Paris 06, CNRS-UMR 7588,
   Institut des NanoSciences de Paris, F-75005, Paris, France
@@
 
-\author[6]{Timo~Heister}
-\affil[6]{Mathematical Sciences,
-  O-110 Martin Hall,
-  Clemson University,
-  Clemson, SC 29634, USA
-  and Scientific Computing and Imaging Institute,
+\author[3,7]{Timo~Heister}
+\affil[7]{Scientific Computing and Imaging Institute,
   72 S Central Campus Drive, Room 3750
@@
 %
-\author[7]{Luca~Heltai}
-\affil[7]{SISSA,
+\author[8]{Luca~Heltai}
+\affil[8]{SISSA,
   International School for Advanced Studies,
@@
 
-\author[8]{Martin~Kronbichler}
-\affil[8]{Institute for Computational Mechanics,
+\author[9]{Martin~Kronbichler}
+\affil[9]{Institute for Computational Mechanics,
   Technical University of Munich,
@@
 
-\author[9]{Ross~Maguire~Kynch}
-\affil[9]{Zienkiewicz Centre for Computational Engineering,
+\author[10]{Ross~Maguire~Kynch}
+\affil[10]{Zienkiewicz Centre for Computational Engineering,
   College of Engineering, Swansea University,
@@
 
-\author[10]{Matthias~Maier}
-\affil[10]{Department of Mathematics,
+\author[11]{Matthias~Maier}
+\affil[11]{Department of Mathematics,
   Texas A\&M University,
@@
 
-\author[11]{David Wells}
-\affil[11]{Department of Mathematics, University of North Carolina,
+\author[12]{David Wells}
+\affil[12]{Department of Mathematics, University of North Carolina,
   Chapel Hill, NC 27516, USA.
@@
   Pi-Yueh Chuang,
-  T. Conrad Clevenger,
   Sambit Das,

### 2019-06-08 9.1/paper.tex
Add context for the Nedelec element

--- 2019-06-07 9.1/paper.tex
@@
   stable but converges with only first-order at the cost of the large
-  number of velocity unknowns. The BR space is thus intermediate between the
+  number of velocity unknowns. The BR space is thus intermediate to the
   $Q_1^d$ and $Q_2^d$ spaces.
@@
   present in traditional N{\'e}d{\'e}lec elements that arise from the edge
-  and face parameterizations used in the basis functions. Therefore, this
+  and face parameterizations used in the basis functions.
+  Because of the sign conflict it is only possible to use the traditional
+  N{\'e}d{\'e}lec element either in a grid configuration or in a complex
+  tetrahedra mesh configuration. It is not possible to use traditional
+  N{\'e}d{\'e}lec in a complex hexahedra mesh configuration. This
   element should provide consistent results for general quadrilateral and
-  hexahedral elements for which the relative orientations of edges and
+  hexahedral elements in a complex mesh configuration
+  for which the relative orientations of edges and
   faces (as seen from all adjacent cells) are often difficult to establish.
@@
 in the file that lists all changes for this release}, see \cite{changes91}.
+(The file is also linked to from the web site of each release as well as
+the release announcement.)
 
@@
 compute the directional derivatives of components of (multi-field) constitutive laws
-with respect to the scalar, vector, tensor, and symmetric tensor fields in terms
+with respect to the scalar, vector, tensor and symmetric tensor fields in terms
 of which they are parameterized. The \texttt{ScalarFunction} class may be used to
@@
 \begin{itemize}
-\item \texttt{differentiate} scalar, tensor, and symmetric tensor expressions
-with respect to other scalar, tensor, or symmetric tensor expressions;
+\item \texttt{differentiate} scalar expressions, tensor and symmetric tensors of expressions
+with respect to other scalar expressions, tensor and symmetric tensors of expressions;
 \item create symbolic substitution maps;
@@
 In the next release we expect to implement classes to assist in performing assembly operations
-in the same spirit as it is already possible using automatic differentation using the \texttt{Differentiation::AD} namespace,
+in the same spirit as that which has been done in the \texttt{Differentiation::AD} namespace,
 although in a fully symbolic manner.
@@
 As we have shown in \texttt{step-62}, the \dealii{}'s \texttt{HDF5} interface can be easily used
-to exchange data with Python and Jupyter notebooks.
+to exchange data with python and jupyter notebooks.
 
@@
     the relevant data to the host, performing MPI communication, and finally
-    moving the data to the device or, if CUDA-aware MPI is available, by
+    move the data to the device or, if CUDA-aware MPI is available, by
     performing MPI communication directly between GPUs.
@@
 extremely good value for a PDE solver which have classically only reached a
-few percent of the LINPACK performance.
-More importantly, this is achieved within a flexible framework supporting
-arbitrary polynomial order on adaptively refined, unstructured meshes and
-with algorithms in the matrix-free module of \dealii designed to minimize time
-to solution and scalability, rather than maximizing the number of floating point operations. The
+few percent of the LINPACK performance, and given the optimization steps
+contained in the matrix-free module of \dealii that aim to minimize the time
+to solution rather than maximize the number of floating point operations. The
 largest Poisson problem that has been solved on 304k cores contained 2.15
@@
 Furthermore, the implementation of the Chebyshev iteration, \dealii's most
-popular smoother in the matrix-free context, has been revised to reduce the number of
+popular smoother in the matrix-free context, has been revised to reduce the
 vector accesses. The vector operations have become an increasing bottleneck
@@
 
-\item \texttt{step-63} implements a geometric multigrid preconditioner
+\item \texttt{step-63} implements a multigrid preconditioner
   and solver for the advection-diffusion equation, yielding optimal
-  complexity. The tutorial compares
-  point-based smoothers to cell-based smoothers and
-  demonstrates the effect of downstream ordering on smoother performance.
+  complexity.
 
 \item \texttt{step-64} demonstrates the usage of matrix-free methods on Nvidia GPUs.
-GPUs are shown to be advantageous for these kind of operations because of their superior
+GPUs are advantageous for these kind of operations because of their superior
 hardware characteristics, in particular a higher memory bandwidth than server
@@
 \item The \texttt{MCMC-Laplace} code gallery program is a code useful
-  for the forward solution used as a building block in
+  for the forward solution that is used as a building block in
   Bayesian inverse problems, and for sampling the parameter space
@@
       subscribed-to object is still valid.
-\item The \texttt{ConstraintMatrix} class gained a template parameter for the scalar
+\item The \texttt{ConstraintMatrix} class gained a template parameter for scalar
       type and was been renamed \texttt{AffineConstraints}. Several methods that
@@
 
-W.~Bangerth, T.~C.~Clevenger, and T.~Heister were partially
+W.~Bangerth and T.~Heister were partially
 supported by the National Science Foundation under award OAC-1835673
 as part of the Cyberinfrastructure for Sustained Scientific Innovation (CSSI)
-program  and by the Computational Infrastructure
+program; by award DMS-1821210; and by the Computational Infrastructure
 in Geodynamics initiative (CIG), through the National Science
@@
 University of California -- Davis.
-
-W.~Bangerth and T.~Heister were also partially supported by award DMS-1821210.
 

### 2019-06-08 9.1/paper.tex
Add comments

@@
   element in \dealii{} can be found in \cite{Kynch2017}.
+
+  \textbf{
+  Matthias. These are suggestions. Feel free to ignore them. I'm planing
+  to do maxwell simulations with \dealii{} and I would like to understand
+  the advantages of NedelecSZ.
+  \begin{itemize}
+  \item I'm not sure what is wrong
+  with the former Nedelec element. Is it correct what I wrote? Feel free
+  to edit it.
+  \item This new NedelecSZ element works with adaptive mesh refinement?
+  \item If it does not work with hanging nodes. Would it be possible to
+  explain in a couple of sentences what is left to make it work with
+  hanging nodes?.
+  \end{itemize}
+  }
 

### 2019-06-08 9.1/paper.tex
Add context algebra_diferentiation

--- 2019-06-08 9.1/paper.tex
@@
   present in traditional N{\'e}d{\'e}lec elements that arise from the edge
-  and face parameterizations used in the basis functions.
-  Because of the sign conflict it is only possible to use the traditional
-  N{\'e}d{\'e}lec element either in a grid configuration or in a complex
-  tetrahedra mesh configuration. It is not possible to use traditional
-  N{\'e}d{\'e}lec in a complex hexahedra mesh configuration. This
+  and face parameterizations used in the basis functions. Therefore, this
   element should provide consistent results for general quadrilateral and
-  hexahedral elements in a complex mesh configuration
-  for which the relative orientations of edges and
+  hexahedral elements for which the relative orientations of edges and
   faces (as seen from all adjacent cells) are often difficult to establish.
@@
 \label{subsec:sd}
+\textbf{Jean-Paul, could you give some context? What is the goal of symbolic algebra?
+The goal is to write an equation in the strong form and automatically get the weak form?}
 To complement the automatic differentiation features in \dealii{}, this release sees

### 2019-06-09 9.1/paper.tex
fix typos

@@
 \label{subsec:sd}
-\textbf{Jean-Paul, could you give some context? What is the goal of symbolic algebra?
-The goal is to write an equation in the strong form and automatically get the weak form?}
 To complement the automatic differentiation features in \dealii{}, this release sees
@@
 In the next release we expect to implement classes to assist in performing assembly operations
-in the same spirit as that which has been done in the \texttt{Differentiation::AD} namespace,
+in the same spirit as it is already possible using automatic differentiation using the \texttt{Differentiation::AD} namespace,
 although in a fully symbolic manner.
@@
 which can be a contiguous collection of points in a dataset, or a regular
-pattern of points or blocks in a datataset.
+pattern of points or blocks in a dataset.
 \item It is possible to define attributes in the \texttt{HDF5::File}, \texttt{HDF5::Group} and

### 2019-06-10 9.1/paper.tex
Fix line break point in a sentance

@@
 In the next release we expect to implement classes to assist in performing assembly operations
-in the same spirit as it is already possible using automatic differentiation using the \texttt{Differentiation::AD} namespace,
-although in a fully symbolic manner.
+in the same spirit as it is already possible using automatic differentiation using the
+\texttt{Differentiation::AD} namespace, although in a fully symbolic manner.
 We will also address performance issues of the \texttt{Expression} class by leveraging

### 2019-06-10 9.1/paper.tex
Add a couple of sentances on what the uses of a CAS are.

@@
 system (CAS) via the SymEngine library \cite{symengine-web-page}.
+This allows the development of exact algebraic expressions using variables that are
+manipulated symbolically and may represent any value (or supported data structure).
+In the context of finite element simulations, typical applications include (but, due to
+the generality of the CAS, are not limited to) the development of constitutive models
+and the implementation of finite element assembly operations through the construction
+and linearization of finite element residuals.
+
 The \texttt{Expression} class in the namespace \texttt{Differentiation::SD} interfaces

### 2019-06-11 9.1/paper.tex
Some additional text improvements

@@
 adaptively refined meshes in \dealii are based on so-called local coarsening,
-i.e., smoothening is applied level-per-level, skipping parts of the domain
+i.e., smoothening is done level-per-level, skipping parts of the domain
 where the mesh is not as refined. The algorithm for the assignment of the
@@
 in detail in \cite{ClevengerHeisterKanschatKronbichler2019}. While most of the
-functionality was already available since the 8.5 release of \dealii presented
+functionality has already been available since the 8.5 release of \dealii presented
 in \cite{dealII85}, several components have been finalized, such as the
@@
 execution, and interfaces that allow the combination with matrix-free GPU
-computations according to \cite{KronbichlerLjungkvist2019}.
+computations as showcased in \cite{KronbichlerLjungkvist2019}.
 
@@
 to ensure scalability of the matrix-free algorithms and geometric multigrid
-infrastructure on more than 100,000 MPI ranks. Geometric multigrid solver for
-the Poisson equation as described in \cite{KronbichlerWall2018} have been used
-as performance tests during the acceptance phase of the SuperMUC-NG
+infrastructure on more than 100,000 MPI ranks. A geometric multigrid solver for
+the Poisson equation as described in \cite{KronbichlerWall2018} has been used
+as a performance test during the acceptance phase of the SuperMUC-NG
 supercomputer in Garching, Germany. Scaling tests have been performed on up to
@@
 position 8 of the top-500 list of November 2018), this can be considered an
-extremely good value for a PDE solver which have classically only reached a
+extremely good value for PDE solvers which have classically only reached a
 few percent of the LINPACK performance.
@@
 trillion unknowns (or 7.1 million unknowns per MPI rank) and was solved in 3.5
-seconds. Also, CFD production runs with up to $10^{11}$ unknowns and $10^5$
+seconds. Also, CFD production runs with $10^{11}$ unknowns and $10^5$
 time steps have been completed in less than seven hours, demonstrating the
@@
 multigrid unknowns in \dealii's \texttt{DoFHandler} and \texttt{MGTransfer}
-classes. While a few bottlenecks have already been resolved, we plan several
-further improvements in setup times for the next release.
+classes. While a few bottlenecks have already been resolved for the present
+release, we plan several further improvements of the setup stage for the next release.
 
@@
   Bayesian inverse problems, and for sampling the parameter space
-  through a Metropolis-Hastings sampler (a kind of Monte Carlo
+  through a Metropolis--Hastings sampler (a kind of Monte Carlo
   Markov Chain method).

### 2019-06-12 9.1/paper.tex
typo

@@
 \texttt{TapelessDrivers} classes that interface with the active AD library, the
-generic helper classes hide library-dependent implementational details and facilitate
+generic helper classes hide library-dependent implementation details and facilitate
 switching between the supported libraries and AD number types based on the

### 2020-04-24 9.2/paper.tex
start 9.2

--- 2019-06-12 9.1/paper.tex
@@
   pdfauthor={
-    Daniel Arndt,
-    Wolfgang Bangerth,
-    Thomas C. Clevenger,
-    Denis Davydov,
-    Marc Fehling,
-    Daniel Garcia-Sanchez,
-    Graham Harper,
+%     Daniel Arndt,
+%     Wolfgang Bangerth,
+%     Thomas C. Clevenger,
+%     Denis Davydov,
+%     Marc Fehling,
+%     Daniel Garcia-Sanchez,
+%     Graham Harper,
     Timo Heister,
-    Luca Heltai,
-    Martin Kronbichler,
-    Ross Maguire Kynch,
-    Matthias Maier,
-    Jean-Paul Pelteret,
-    Bruno Turcksin,
-    David Wells
+%     Luca Heltai,
+%     Martin Kronbichler,
+%     Ross Maguire Kynch,
+%     Matthias Maier,
+%     Jean-Paul Pelteret,
+%     Bruno Turcksin,
+%     David Wells
   },
-  pdftitle={The deal.II Library, Version 9.1, 2019},
+  pdftitle={The deal.II Library, Version 9.2, 2020},
 }
 
-\title{The \dealii Library, Version 9.1}
+\title{The \dealii Library, Version 9.2}
 
-\author[*1]{Daniel Arndt}
-\affil[1]{Computational Engineering and Energy Sciences Group,
-  Computional Sciences and Engineering Division,
-  Oak Ridge National Laboratory, 1 Bethel Valley Rd.,
-  TN 37831, USA.
-  {\texttt{\{arndtd,turcksinbr\}@ornl.gov}}}
-
-\author[2]{Wolfgang~Bangerth}
-\affil[2]{Department of Mathematics, Colorado State University, Fort
-  Collins, CO 80523-1874, USA.
-  {\texttt{bangerth@colostate.edu},
-   \texttt{harper@math.colostate.edu}}}
-
-\author[3]{Thomas~C.~Clevenger}
+% \author[*1]{Daniel Arndt}
+% \affil[1]{Computational Engineering and Energy Sciences Group,
+%   Computional Sciences and Engineering Division,
+%   Oak Ridge National Laboratory, 1 Bethel Valley Rd.,
+%   TN 37831, USA.
+%   {\texttt{\{arndtd,turcksinbr\}@ornl.gov}}}
+% 
+% \author[2]{Wolfgang~Bangerth}
+% \affil[2]{Department of Mathematics, Colorado State University, Fort
+%   Collins, CO 80523-1874, USA.
+%   {\texttt{bangerth@colostate.edu},
+%    \texttt{harper@math.colostate.edu}}}
+% 
+% \author[3]{Thomas~C.~Clevenger}
+% \affil[3]{School of Mathematical and Statistical Sciences,
+%   Clemson University,
+%   Clemson, SC, 29634, USA
+%   {\texttt{\{tcleven, heister\}@clemson.edu}}}
+% 
+% \author[4]{Denis~Davydov}
+% \affil[4]{Chair of Applied Mechanics,
+%   Friedrich-Alexander-Universit\"{a}t Erlangen-N\"{u}rnberg,
+%   Egerlandstr.\ 5,
+%   91058 Erlangen, Germany.
+%   {\texttt{\{denis.davydov,jean-paul.pelteret\}@fau.de}}}
+% 
+% \author[5]{Marc~Fehling}
+% \affil[5]{Institute for Advanced Simulation,
+%   Forschungszentrum J\"{u}lich GmbH,
+%   52425 J\"{u}lich, Germany.
+%   {\texttt{m.fehling@fz-juelich.de}}}
+% 
+% \author[6]{Daniel Garcia-Sanchez}
+% \affil[6]{Sorbonne Universit\'es, UPMC Univ.\ Paris 06, CNRS-UMR 7588,
+%   Institut des NanoSciences de Paris, F-75005, Paris, France
+%   {\texttt{daniel.garcia-sanchez@insp.upmc.fr}}}
+% 
+% \author[2]{Graham Harper}
+% 
+\author[3]{Timo~Heister}
  \affil[3]{School of Mathematical and Statistical Sciences,
@@
    Clemson, SC, 29634, USA
-  {\texttt{\{tcleven, heister\}@clemson.edu}}}
+   {\texttt{heister@clemson.edu}}}
 
-\author[4]{Denis~Davydov}
-\affil[4]{Chair of Applied Mechanics,
-  Friedrich-Alexander-Universit\"{a}t Erlangen-N\"{u}rnberg,
-  Egerlandstr.\ 5,
-  91058 Erlangen, Germany.
-  {\texttt{\{denis.davydov,jean-paul.pelteret\}@fau.de}}}
-
-\author[5]{Marc~Fehling}
-\affil[5]{Institute for Advanced Simulation,
-  Forschungszentrum J\"{u}lich GmbH,
-  52425 J\"{u}lich, Germany.
-  {\texttt{m.fehling@fz-juelich.de}}}
-
-\author[6]{Daniel Garcia-Sanchez}
-\affil[6]{Sorbonne Universit\'es, UPMC Univ.\ Paris 06, CNRS-UMR 7588,
-  Institut des NanoSciences de Paris, F-75005, Paris, France
-  {\texttt{daniel.garcia-sanchez@insp.upmc.fr}}}
-
-\author[2]{Graham Harper}
-
-\author[3,7]{Timo~Heister}
-\affil[7]{Scientific Computing and Imaging Institute,
-  72 S Central Campus Drive, Room 3750
-  Salt Lake City, UT 84112.
-  {\texttt{heister@sci.utah.edu}}}
+% %
+% \author[8]{Luca~Heltai}
+% \affil[8]{SISSA,
+%   International School for Advanced Studies,
+%   Via Bonomea 265,
+%   34136, Trieste, Italy.
+%   {\texttt{luca.heltai@sissa.it}}}
 % 
-\author[8]{Luca~Heltai}
-\affil[8]{SISSA,
-  International School for Advanced Studies,
-  Via Bonomea 265,
-  34136, Trieste, Italy.
-  {\texttt{luca.heltai@sissa.it}}}
-
-\author[9]{Martin~Kronbichler}
-\affil[9]{Institute for Computational Mechanics,
-  Technical University of Munich,
-  Boltzmannstr.~15, 85748 Garching, Germany.
-  {\texttt{kronbichler@lnm.mw.tum.de}}}
-
-\author[10]{Ross~Maguire~Kynch}
-\affil[10]{Zienkiewicz Centre for Computational Engineering,
-  College of Engineering, Swansea University,
-  Bay Campus, Fabian Way, Swansea SA1 8EN, Wales, UK.
-  {\texttt{rkynch@gmail.com}}}
-
-\author[11]{Matthias~Maier}
-\affil[11]{Department of Mathematics,
-  Texas A\&M University,
-  3368 TAMU,
-  College Station, TX 77845, USA.
-  {\texttt{maier@math.tamu.edu}}}
-
-\author[4]{Jean-Paul~Pelteret}
-
-\author[*1]{Bruno~Turcksin}
-
-\author[12]{David Wells}
-\affil[12]{Department of Mathematics, University of North Carolina,
-  Chapel Hill, NC 27516, USA.
-  {\texttt{drwells@email.unc.edu}}}
+% \author[9]{Martin~Kronbichler}
+% \affil[9]{Institute for Computational Mechanics,
+%   Technical University of Munich,
+%   Boltzmannstr.~15, 85748 Garching, Germany.
+%   {\texttt{kronbichler@lnm.mw.tum.de}}}
+% 
+% \author[10]{Ross~Maguire~Kynch}
+% \affil[10]{Zienkiewicz Centre for Computational Engineering,
+%   College of Engineering, Swansea University,
+%   Bay Campus, Fabian Way, Swansea SA1 8EN, Wales, UK.
+%   {\texttt{rkynch@gmail.com}}}
+% 
+% \author[11]{Matthias~Maier}
+% \affil[11]{Department of Mathematics,
+%   Texas A\&M University,
+%   3368 TAMU,
+%   College Station, TX 77845, USA.
+%   {\texttt{maier@math.tamu.edu}}}
+% 
+% \author[4]{Jean-Paul~Pelteret}
+% 
+% \author[*1]{Bruno~Turcksin}
+% 
+% \author[12]{David Wells}
+% \affil[12]{Department of Mathematics, University of North Carolina,
+%   Chapel Hill, NC 27516, USA.
+%   {\texttt{drwells@email.unc.edu}}}
 

### 2020-04-24 9.2/paper.tex
comment out old stuff

@@
   This paper provides an overview of the new features of the finite element
-  library \dealii, version 9.1.
+  library \dealii, version 9.2.
 \end{abstract}
@@
 
-\dealii version 9.1.0 was released May 21, 2019.
+\dealii version 9.2.0 was released May XYZ, 2020.
 This paper provides an
 overview of the new features of this release and serves as a citable
-reference for the \dealii software library version 9.1. \dealii is an
+reference for the \dealii software library version 9.2. \dealii is an
 object-oriented finite element library used around the world in the
@@
 \begin{itemize}
-  \item Improved support for automatic differentiation (see
-    Section~\ref{subsec:ad}),
-  \item Dedicated support for symbolic algebra (see
-    Section~\ref{subsec:sd}),
-  \item Full support for $hp$~adaptivity in parallel computations (see
-    Section~\ref{subsec:hp}),
-  \item An interface to the HDF5 file format and libraries (see
-    Section~\ref{subsec:hdf5}),
-  \item Significantly extended GPU support (see Section~\ref{subsec:gpu}),
-  \item Parallel geometric multigrid (GMG) improvements (see
-    \cite{ClevengerHeisterKanschatKronbichler2019} and
-    Section~\ref{subsec:gmg}),
-  \item Four new tutorial programs (step-61, step-62, step-63, step-64),
-    as well as one new code gallery program (see
-    Section~\ref{subsec:steps}).
+\item
+%   \item Improved support for automatic differentiation (see
+%     Section~\ref{subsec:ad}),
+%   \item Dedicated support for symbolic algebra (see
+%     Section~\ref{subsec:sd}),
+%   \item Full support for $hp$~adaptivity in parallel computations (see
+%     Section~\ref{subsec:hp}),
+%   \item An interface to the HDF5 file format and libraries (see
+%     Section~\ref{subsec:hdf5}),
+%   \item Significantly extended GPU support (see Section~\ref{subsec:gpu}),
+%   \item Parallel geometric multigrid (GMG) improvements (see
+%     \cite{ClevengerHeisterKanschatKronbichler2019} and
+%     Section~\ref{subsec:gmg}),
+%   \item Four new tutorial programs (step-61, step-62, step-63, step-64),
+%     as well as one new code gallery program (see
+%     Section~\ref{subsec:steps}).
 \end{itemize}
@@
 \item
-  The release contains a number of performance improvements and bug fixes for
-  the matrix-free framework. One notable improvement is the support for
-  renumbering of degrees of freedom within the cells for discontinuous
-  elements, avoiding some reshuffling operations across the SIMD lanes
-  with vectorization over several cells and faces, which is especially
-  useful on processors with AVX-512 vectorization (8 doubles), speeding up
-  operations by up to 10\%. Secondly, the strategy for the most efficient
-  tensor product evaluators according to the performance analysis of
-  \cite{KronbichlerKormann2019} in the context of more quadrature points than
-  shape functions has been revised for better performance.
-
-\item A new class \texttt{ParsedConvergenceTable} has been introduced
-  that greatly simplifies the construction of convergence tables,
-  reading the options for the generation of the table from a parameter
-  file, and providing methods that, combined with a parameter file,
-  allow one to generate convergence tables using one-liners in user
-  codes.
-
-\item
-  The \texttt{FE\_BernardiRaugel} class implements the non-standard
-  Bernardi-Raugel (BR) element that can be used to construct a stable
-  velocity-pressure pair for the Stokes equation \cite{BR85}. The BR
-  element is an enriched version of the $Q_1^d$ element with added bubble
-  functions on each edge (in 2d) or face (in 3d). It addresses the fact
-  that the $Q_1^d\times Q_0$ combination is not inf-sup stable (requiring a
-  larger velocity space), and that the $Q_2^d\times Q_0$ combination is
-  stable but converges with only first-order at the cost of the large
-  number of velocity unknowns. The BR space is thus intermediate between the
-  $Q_1^d$ and $Q_2^d$ spaces.
-
-  The element is currently only implemented for parallelogram meshes due to
-  difficulties associated with the mapping of shape functions: The shape
-  functions of the $Q_1^d$ part of the element need to be mapped as
-  scalars, as is common for the vector components of the $Q_1^d$ element;
-  on the other hand, the vector-valued edge bubble functions need to be
-  mapped using the Piola transform as is common for the Raviart-Thomas
-  element. \dealii{} does not currently have the ability to use different
-  mappings for individual shape functions, though this functionality is
-  planned for the next release.
-
-\item
-  The \texttt{FE\_NedelecSZ} class is a new implementation of the
-  N{\'e}d{\'e}lec element on quadrilaterals and hexahedra. It is based on
-  the work of Zaglmayr \cite{Zag06} and overcomes the sign conflict issues
-  present in traditional N{\'e}d{\'e}lec elements that arise from the edge
-  and face parameterizations used in the basis functions. Therefore, this
-  element should provide consistent results for general quadrilateral and
-  hexahedral elements for which the relative orientations of edges and
-  faces (as seen from all adjacent cells) are often difficult to establish.
-  The \texttt{FE\_NedelecSZ} element addresses the sign conflict problem by
-  assigning a globally defined orientation to local edges and faces. A
-  detailed overview of the implementation of the \texttt{FE\_NedelecSZ}
-  element in \dealii{} can be found in \cite{Kynch2017}.
-
-\item All of the elementary geometrical objects of the library (namely
-  \texttt{Point<dim>}, \texttt{Segment<dim>}, and
-  \texttt{BoundingBox<dim>}) have been augmented with the traits
-  needed to comply with \texttt{boost::geometry} concepts. A new
-  interface to \texttt{boost::geometry::index::rtree} has been added
-  that simplifies the construction of spatial indices based on points,
-  bounding boxes, or segments.
+% \item
+%   The release contains a number of performance improvements and bug fixes for
+%   the matrix-free framework. One notable improvement is the support for
+%   renumbering of degrees of freedom within the cells for discontinuous
+%   elements, avoiding some reshuffling operations across the SIMD lanes
+%   with vectorization over several cells and faces, which is especially
+%   useful on processors with AVX-512 vectorization (8 doubles), speeding up
+%   operations by up to 10\%. Secondly, the strategy for the most efficient
+%   tensor product evaluators according to the performance analysis of
+%   \cite{KronbichlerKormann2019} in the context of more quadrature points than
+%   shape functions has been revised for better performance.
+% 
+% \item A new class \texttt{ParsedConvergenceTable} has been introduced
+%   that greatly simplifies the construction of convergence tables,
+%   reading the options for the generation of the table from a parameter
+%   file, and providing methods that, combined with a parameter file,
+%   allow one to generate convergence tables using one-liners in user
+%   codes.
+% 
+% \item
+%   The \texttt{FE\_BernardiRaugel} class implements the non-standard
+%   Bernardi-Raugel (BR) element that can be used to construct a stable
+%   velocity-pressure pair for the Stokes equation \cite{BR85}. The BR
+%   element is an enriched version of the $Q_1^d$ element with added bubble
+%   functions on each edge (in 2d) or face (in 3d). It addresses the fact
+%   that the $Q_1^d\times Q_0$ combination is not inf-sup stable (requiring a
+%   larger velocity space), and that the $Q_2^d\times Q_0$ combination is
+%   stable but converges with only first-order at the cost of the large
+%   number of velocity unknowns. The BR space is thus intermediate between the
+%   $Q_1^d$ and $Q_2^d$ spaces.
+% 
+%   The element is currently only implemented for parallelogram meshes due to
+%   difficulties associated with the mapping of shape functions: The shape
+%   functions of the $Q_1^d$ part of the element need to be mapped as
+%   scalars, as is common for the vector components of the $Q_1^d$ element;
+%   on the other hand, the vector-valued edge bubble functions need to be
+%   mapped using the Piola transform as is common for the Raviart-Thomas
+%   element. \dealii{} does not currently have the ability to use different
+%   mappings for individual shape functions, though this functionality is
+%   planned for the next release.
+% 
+% \item
+%   The \texttt{FE\_NedelecSZ} class is a new implementation of the
+%   N{\'e}d{\'e}lec element on quadrilaterals and hexahedra. It is based on
+%   the work of Zaglmayr \cite{Zag06} and overcomes the sign conflict issues
+%   present in traditional N{\'e}d{\'e}lec elements that arise from the edge
+%   and face parameterizations used in the basis functions. Therefore, this
+%   element should provide consistent results for general quadrilateral and
+%   hexahedral elements for which the relative orientations of edges and
+%   faces (as seen from all adjacent cells) are often difficult to establish.
+%   The \texttt{FE\_NedelecSZ} element addresses the sign conflict problem by
+%   assigning a globally defined orientation to local edges and faces. A
+%   detailed overview of the implementation of the \texttt{FE\_NedelecSZ}
+%   element in \dealii{} can be found in \cite{Kynch2017}.
+% 
+% \item All of the elementary geometrical objects of the library (namely
+%   \texttt{Point<dim>}, \texttt{Segment<dim>}, and
+%   \texttt{BoundingBox<dim>}) have been augmented with the traits
+%   needed to comply with \texttt{boost::geometry} concepts. A new
+%   interface to \texttt{boost::geometry::index::rtree} has been added
+%   that simplifies the construction of spatial indices based on points,
+%   bounding boxes, or segments.
 \end{itemize}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Improved support for automatic differentiation}
-\label{subsec:ad}
-
-In the previous release, numerous classes that are used to assemble linear systems
-and right hand sides, as well those used to define constitutive laws, were given
-full support for ``white-listed'' automatically differentiable (AD) number types
-from the ADOL-C and Sacado libraries. This included the classes that
-represent the local contributions of one cell to the global linear
-system (i.e., \texttt{FullMatrix} and \texttt{Vector}) as well as the
-classes commonly used for the computations at individual quadrature
-points (primarily the \texttt{Tensor} and \texttt{SymmetricTensor}
-classes of various ranks).
-In the current release we have provided a unified interface to these
-AD libraries, focusing on two specific use contexts:
-\begin{enumerate}
-\item The construction and linearization of finite element residuals; and
-\item The construction and linearization of constitutive model kinetic variables.
-\end{enumerate}
-
-In the first context, the finite element degrees of freedom are considered the
-independent variables. From these primitives, the \texttt{EnergyFunctional} helper
-class in the namespace \texttt{Differentiation::AD} may be used to compute both the
-residual and its linearization by directly defining the contribution to the
-(twice differentiated) scalar total energy functional from each cell. Similarly,
-the \texttt{ResidualLinearization} class requires the (once-differentiated) finite
-element residual to be defined on a per-cell basis,  and this contribution is
-automatically linearized.
-
-The second context aims directly at constitutive model formulations, and serves to
-compute the directional derivatives of components of (multi-field) constitutive laws
-with respect to the scalar, vector, tensor, and symmetric tensor fields in terms
-of which they are parameterized. The \texttt{ScalarFunction} class may be used to
-define a scalar function (such as strain energy function) that may be twice
-differentiated, while the \texttt{VectorFunction} may be used to define a vector
-function (such as a set of kinematic fields) that may be differentiated once.
-Since the total derivatives of all components are computed at once, these two helper
-classes provide an interface to retrieve each sub-component of the gradient and
-Hessian (for a \texttt{ScalarFunction}) or values and Jacobian (for a
-\texttt{VectorFunction}).
-
-Although these aforementioned helper classes have been documented with a specific
-use in mind, they remain generic and may (with a reinterpretation of the meaning of
-the independent and dependent variables) be used for other purposes as well.
-Furthermore, through the implementation of \texttt{TapedDrivers} and
-\texttt{TapelessDrivers} classes that interface with the active AD library, the
-generic helper classes hide library-dependent implementation details and facilitate
-switching between the supported libraries and AD number types based on the
-user's requirements.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Dedicated support for symbolic algebra, including algebra differentiation}
-\label{subsec:sd}
-To complement the automatic differentiation features in \dealii{}, this release sees
-the first step towards integrating and supporting a highly performant computer algebra
-system (CAS) via the SymEngine library \cite{symengine-web-page}.
-This allows the development of exact algebraic expressions using variables that are
-manipulated symbolically and may represent any value (or supported data structure).
-In the context of finite element simulations, typical applications include (but, due to
-the generality of the CAS, are not limited to) the development of constitutive models
-and the implementation of finite element assembly operations through the construction
-and linearization of finite element residuals.
-
-The \texttt{Expression} class in the namespace \texttt{Differentiation::SD} interfaces
-to SymEngine and forms the basis of symbolic computations, offering a full set of
-overloaded operators and a C++ style interface. This class offers the following
-basic functionality:
-\begin{itemize}
-\item symbolic variable definition,
-\item symbolic function definition,
-\item expression creation using standard C++ syntax,
-\item expression parsing,
-\item comparison operations,
-\item logical operations,
-\item conditional expression construction,
-\item differentiation,
-\item substitution (partial and complete), and
-\item serialization.
-\end{itemize}
-\dealii{} now also provides an extensive set of math operations, with a syntax mimicking that used
-in the C++ standard library. Using the \texttt{Expression} class as a basis, we
-have developed a set of functions that can be used to create \dealii{} \texttt{Tensor}s
-and \texttt{SymmetricTensor}s of symbolic variables and symbolic functions. This
-gives full symbolic tensor algebra support using the pre-existing \texttt{Tensor}
-and \texttt{SymmetricTensor} classes and associated functions. We have
-also implemented a set of utility functions with the following features:
-\begin{itemize}
-\item \texttt{differentiate} scalar expressions with respect to other scalar expressions, as well as
-tensors and symmetric tensors of expressions;
-\item \texttt{differentiate} tensors and symmetric tensors of expressions
-with respect to scalar expressions, as well as other tensors and symmetric tensors of expressions;
-\item create symbolic substitution maps;
-\item resolve explicit dependencies between expressions; and
-\item perform scalar and tensor valued substitution (including conversion from symbolic to
-real-valued scalars and tensors).
-\end{itemize}
-
-In the next release we expect to implement classes to assist in performing assembly operations
-in the same spirit as it is already possible using automatic differentiation using the
-\texttt{Differentiation::AD} namespace, although in a fully symbolic manner.
-We will also address performance issues of the \texttt{Expression} class by leveraging
-the optimization capabilities of SymEngine, including common subexpression elimination (CSE),
-as well as by generating high performance code-paths to evaluate these expressions through the
-use of a custom-generated \texttt{std::function} (so-called ``lambda'' optimization) or by
-compiling expressions using the LLVM JIT compiler.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Full support for $hp$~adaptivity in parallel computations}
-\label{subsec:hp}
-
-\dealii{} has had support for $hp$~adaptive methods since around 2005
-(documented in \cite{BangerthKayserHerold2007}) and for parallel
-computations on distributed meshes since around 2010 (see
-\cite{BangerthBursteddeHeisterKronbichler11}), but not for both at the
-same time. The challenges to combine these are related to a number of
-areas:
-\begin{enumerate}
-\item Data structures: The data structures necessary to store the
-  indices of degrees of freedom are substantially more
-  complicated for $hp$~algorithms than for the $h$~adaptive schemes
-  that were already implemented. This is because the number of degrees
-  of freedom per cell is now no longer constant. Furthermore,
-  faces and edges may need to store more than one set of indices if
-  the adjacent cells use different polynomial degrees; in the case of
-  edges, the number of sets of indices may also be of variable size.
-
-  All of this poses challenges in the parallel context because some of
-  the information may not be known, or not be known right away, for
-  cells that are not locally owned (i.e., for ghost and ``artificial''
-  cells), and for which the data structures stored on different
-  processors have to be reconciled.
-
-\item Algorithms: Already for $h$~adaptive meshes, enumerating all
-  degrees of freedom uniquely on the global mesh is difficult, as evidenced by
-  the complications of the algorithms shown in Section 3.1 of
-  \cite{BangerthBursteddeHeisterKronbichler11}, which requires more
-  than one page of text and is implemented in many hundreds of lines of
-  code.
-
-  These difficulties are even more pronounced when using $hp$~adaptivity.
-  The main obstacle is the desire to unify the indices of
-  matching degrees of freedom on adjacent cells whenever elements with
-  continuous polynomials are used. For example, the edge degree of freedom
-  of a $Q_2$~element has to be merged with the middle one of the three
-  edge degrees of freedom of a $Q_4$~element on a neighboring cell.
-  Section 4.2 of \cite{BangerthKayserHerold2007} discusses a sequential algorithm
-  that eliminates one of these degrees of freedom in favor of another,
-  but it introduces a ``master'' and a ``slave'' side of the
-  interface. This is of no major consequence in sequential
-  computations, but is inconvenient in parallel computations if the
-  ``master'' side is a ghost cell whose degree of freedom indices are
-  not (yet) available while enumerating local degrees of freedom, or
-  if the master is an artificial cell whose information will never be
-  available on a processor.
-
-  An earlier implementation of the algorithm enumerating degrees of
-  freedom, already available in \dealii{} 9.0, simply did not unify
-  indices on processor boundaries. However, this makes the total
-  number of degrees of freedom dependent on both the partition of the
-  mesh and the number of processors available. We have therefore
-  re-implemented the algorithm so that the unification does happen
-  also on processor boundaries, and will report on the details
-  elsewhere.
-
-\item Data transfer patterns: An important algorithm in parallel
-  finite element methods is the exchange of information stored on
-  cells during mesh repartitioning. This happens, for example, when
-  interpolating the solution from one mesh to the next
-  adaptively refined mesh; or when adapting the polynomial degrees
-  associated with each cell and repartitioning in order to
-  balance the computational cost of each processor's partition. When
-  using $h$~adaptive methods, the amount of data associated with each
-  cell is fixed and the algorithms that implement the data transfer
-  are consequently relatively simple. On the other hand, in $hp$~contexts,
-  each cell may have a different number of unknowns
-  associated with it, and the algorithms that transfer the data are
-  substantially more complicated. In order to implement those, we rely
-  on recent enhancements of the \pfrst~library (documented in
-  \cite{Burstedde2018}) to transfer data of variable size across
-  processors. Furthermore, the amount of data associated with each
-  cell may be large on cells with higher polynomial degrees, and
-  might profit from compression before sending.
-
-\item Balancing computational cost: For $h$~adaptive algorithms, the
-  amount of work associated with each cell is essentially the same,
-  both during the assembly of linear systems as well as during the
-  solver phase. For $hp$~adaptive methods, this is no longer the
-  case. Consequently, balancing the cost of work between different
-  processors' partitions is no longer as easy as ensuring that every
-  processor owns a roughly equal number of cells. Rather, one needs to
-  introduce a weighting factor for each cell that describes its
-  relative cost compared to some reference. To make things worse, the
-  relative cost of assembly on a cell might not match the relative
-  cost of the linear solver associated with this cell, leading to
-  difficult trade-offs in defining optimal weighting factors. In this
-  release, we supplied the basic functionality to attach any amount
-  of weighting factors to cells, but users still have to find reasonable
-  weights for themselves.
-\end{enumerate}
-
-All of these issues have been addressed in the current release and are
-available to users. We will report on the details of the algorithms
-and their performance in a separate publication.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Interface to the HDF5 file format and libraries}
-\label{subsec:hdf5}
-\texttt{HDF5} is an open-source library and file format designed to store large amounts of data.
-The \texttt{HDF5} format is specially tailored for high volume parallel I/O operations using MPI.
-\texttt{HDF5} files are self-describing and allow complex data relationships and a large variety
-of datatypes. In addition, the \texttt{HDF5} format is designed for long-term preservation of
-data; a \texttt{HDF5} file created by an HPC system can be easily read by a commodity laptop.
-
-\dealii{}'s new  \texttt{HDF5} interface allows to write \texttt{HDF5} files and manipulate
-datasets, groups, and attributes in serial and in parallel using MPI. The
-\texttt{HDF5} MPI library calls that modify the structure of the file are
-always collective, whereas writing and reading raw data in a dataset can be done
-independently or collectively. In the \dealii{}'s \texttt{HDF5} interface all the calls are
-set to collective in order to maximize performance. This means that all the MPI
-processes have to contribute to every single call, even if they don't have data to write.
-We have added the following classes to the \dealii{}'s \texttt{HDF5} namespace.
-\begin{itemize}
-\item The new \texttt{HDF5::File} class can be used to open and create \texttt{HDF5} files
-in serial or in parallel.
-\item The new \texttt{HDF5::Group} class can be used to open and create groups.
- \texttt{HDF5} files have a tree structure. The root contains groups and
-the groups can contain other groups.
-\item The new \texttt{HDF5::DataSet} class can be used to open and create datasets which
-can be placed inside groups or at the root of the \texttt{HDF5} file. A dataset can be a
-vector, a matrix, or a tensor. It is possible to read and write in the dataset using
-hyperslabs or unordered data. Hyperslabs are portions of datasets
-which can be a contiguous collection of points in a dataset, or a regular
-pattern of points or blocks in a dataset.
-\item It is possible to define attributes in the \texttt{HDF5::File}, \texttt{HDF5::Group} and
-\texttt{HDF5::DataSet} classes. An attribute is a small metadata such as a number or a string.
-Attributes are commonly used to store simulation parameters.
-\end{itemize}
-As we have shown in \texttt{step-62}, the \dealii{}'s \texttt{HDF5} interface can be easily used
-to exchange data with Python and Jupyter notebooks.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{GPU support via CUDA}
-\label{subsec:gpu}
-
-GPU support was significantly extended for the current release:
-\begin{itemize}
-  \item The \texttt{CUDAWrappers::PreconditionILU} and \texttt{CUDAWrappers::PreconditionIC}
-    classes can now be used for preconditioning \texttt{CUDAWrappers::SparseMatrix} objects.
-  \item \texttt{LinearAlgebra::distributed::Vector}: the MPI-parallel vector
-    class has gained a second template argument \texttt{MemorySpace} which can
-    either be \texttt{Host} or \texttt{CUDA}. In the latter case, the data
-    resides in GPU memory. By default, the template parameter is
-    \texttt{Host} and the behavior is unchanged compared to previous versions.
-    When using CUDA, the ghost exchange can be performed either by first copying
-    the relevant data to the host, performing MPI communication, and finally
-    moving the data to the device or, if CUDA-aware MPI is available, by
-    performing MPI communication directly between GPUs.
-  \item Constrained degrees of freedom: the matrix-free framework now
-    supports constrained degrees of freedom. The implementation is based on
-    the algorithms described in \cite{ljungkvist2017}. With this addition,
-    both Dirichlet boundary conditions and the constraints arising from
-    adaptively refined meshes can be imposed within the matrix-free
-    framework.
-%     \textcolor{red}{%
-%     \sout{The only restriction is that for two-dimensional meshes the
-%     finite element degree must be odd. There is no such restriction in
-%     three dimensions.}
-%     \emph{%
-%     WB: Can we explain this restriction? It seems odd (pun intended)...
-%     BT: No, we cannot. I have spent at least two full weeks trying to
-%     understand why it doesn't work, i.e., gives the wrong result, but I
-%     don't know. We decided at the time that even with the current
-%     limitation this was a useful addition.
-%     MM: I suggest that we then simply not mention the limitation and treat
-%     it as a known bug. As it is currently worded it can be easily
-%     interpreted as a fundamental limitation, i.\,e. ``FE with even degree
-%     are not admissible.''}}
-
-  \item MPI matrix-free computations: using \texttt{LinearAlgebra::distributed::Vector}, the
-    matrix-free framework can scale to multiple GPUs by taking
-    advantage of MPI. Each MPI process can only use one GPU and therefore, if
-    multiple GPUs are available in one node, it is necessary to have as many
-    ranks as there are GPUs. Using Nvidia Multi-Process Service (MPS), it is also possible
-    for multiple processes to use the same GPU. This can be advantageous if the
-    amount of work on one rank is not sufficient to fully utilize a GPU.
-\end{itemize}
-
-The matrix-free GPU components integrated in \dealii have been compared against
-CPUs in \cite{KronbichlerLjungkvist2019}, where the application to geometric
-multigrid solvers is discussed.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Parallel geometric multigrid improvements}
-\label{subsec:gmg}
-
-For the 9.1 release, the geometric multigrid facilities have been extended and
-revised for performance. The geometric multigrid algorithms for uniform and
-adaptively refined meshes in \dealii are based on so-called local coarsening,
-i.e., smoothening is done level-per-level, skipping parts of the domain
-where the mesh is not as refined. The algorithm for the assignment of the
-owner on level cells and the implications on load balancing have been analyzed
-in detail in \cite{ClevengerHeisterKanschatKronbichler2019}. While most of the
-functionality has already been available since the 8.5 release of \dealii presented
-in \cite{dealII85}, several components have been finalized, such as the
-support for certain renumbering algorithms that are beneficial for matrix-free
-execution, and interfaces that allow the combination with matrix-free GPU
-computations as showcased in \cite{KronbichlerLjungkvist2019}.
-
-A number of data structures and implementations in \dealii have been adapted
-to ensure scalability of the matrix-free algorithms and geometric multigrid
-infrastructure on more than 100,000 MPI ranks. A geometric multigrid solver for
-the Poisson equation as described in \cite{KronbichlerWall2018} has been used
-as a performance test during the acceptance phase of the SuperMUC-NG
-supercomputer in Garching, Germany. Scaling tests have been performed on up to
-the full machine with 304,128 cores of the Intel Xeon Skylake architecture and
-an arithmetic performance of around 5 PFlop/s for a geometric multigrid solver
-with polynomials of degree 4 has been reached. Compared to the official
-LINPACK performance of the machine of 19.5 PFlop/s (the machine is listed on
-position 8 of the top-500 list of November 2018), this can be considered an
-extremely good value for PDE solvers which have classically only reached a
-few percent of the LINPACK performance.
-More importantly, this is achieved within a flexible framework supporting
-arbitrary polynomial order on adaptively refined, unstructured meshes and
-with algorithms in the matrix-free module of \dealii designed to minimize time
-to solution and scalability, rather than maximizing the number of floating point operations. The
-largest Poisson problem that has been solved on 304k cores contained 2.15
-trillion unknowns (or 7.1 million unknowns per MPI rank) and was solved in 3.5
-seconds. Also, CFD production runs with $10^{11}$ unknowns and $10^5$
-time steps have been completed in less than seven hours, demonstrating the
-capabilities of \dealii for large-scale parallel computations. The scaling tests
-also revealed several relatively expensive operations in the setup of the
-multigrid unknowns in \dealii's \texttt{DoFHandler} and \texttt{MGTransfer}
-classes. While a few bottlenecks have already been resolved for the present
-release, we plan several further improvements of the setup stage for the next release.
-
-Furthermore, the implementation of the Chebyshev iteration, \dealii's most
-popular smoother in the matrix-free context, has been revised to reduce the number of
-vector accesses. The vector operations have become an increasing bottleneck
-due to the level of optimization available for the operator evaluation with
-our matrix-free framework \cite{KronbichlerKormann2019}, especially on newer
-CPUs which can perform a large number of arithmetic operations per byte loaded
-from main memory. This speeds up matrix-free multigrid solvers by up to
-10--15\% on geometries with affine (parallelogram and parallelpiped) cells,
-and up to 5\% on deformed cells.
+\subsection{bla1}
+\label{subsec:bla1}
 
@@
 \begin{itemize}
-\item \texttt{step-61} is a program that implements the ``weak
-  Galerkin'' method to solve the Laplace equation. Weak Galerkin
-  methods are related to the Hybridized Discontinuous Galerkin method
-  in that they introduce degrees of freedom on the interfaces between
-  cells, but they do not require the reformulation of the problem as a
-  first-order system and instead re-define what the gradient of a
-  discontinuous function is.
-
-\item \texttt{step-62} demonstrates the solution of problems related
-  to phononic or photonic crystals. Among the techniques shown in this
-  program is the solution of complex-valued linear systems, and the
-  use of absorbing boundary conditions through the Perfectly Matched
-  Layer technique.
-
-\item \texttt{step-63} implements a geometric multigrid preconditioner
-  and solver for the advection-diffusion equation, yielding optimal
-  complexity. The tutorial compares
-  point-based smoothers to cell-based smoothers and
-  demonstrates the effect of downstream ordering on smoother performance.
-
-\item \texttt{step-64} demonstrates the usage of matrix-free methods on Nvidia GPUs.
-GPUs are shown to be advantageous for these kind of operations because of their superior
-hardware characteristics, in particular a higher memory bandwidth than server
-CPUs within a given power envelope.
-
-\item The \texttt{MCMC-Laplace} code gallery program is a code useful
-  for the forward solution used as a building block in
-  Bayesian inverse problems, and for sampling the parameter space
-  through a Metropolis--Hastings sampler (a kind of Monte Carlo
-  Markov Chain method).
+\item 
+% \texttt{step-61} is a program that implements the ``weak
+%   Galerkin'' method to solve the Laplace equation. Weak Galerkin
+%   methods are related to the Hybridized Discontinuous Galerkin method
+%   in that they introduce degrees of freedom on the interfaces between
+%   cells, but they do not require the reformulation of the problem as a
+%   first-order system and instead re-define what the gradient of a
+%   discontinuous function is.
+% 
+% \item \texttt{step-62} demonstrates the solution of problems related
+%   to phononic or photonic crystals. Among the techniques shown in this
+%   program is the solution of complex-valued linear systems, and the
+%   use of absorbing boundary conditions through the Perfectly Matched
+%   Layer technique.
+% 
+% \item \texttt{step-63} implements a geometric multigrid preconditioner
+%   and solver for the advection-diffusion equation, yielding optimal
+%   complexity. The tutorial compares
+%   point-based smoothers to cell-based smoothers and
+%   demonstrates the effect of downstream ordering on smoother performance.
+% 
+% \item \texttt{step-64} demonstrates the usage of matrix-free methods on Nvidia GPUs.
+% GPUs are shown to be advantageous for these kind of operations because of their superior
+% hardware characteristics, in particular a higher memory bandwidth than server
+% CPUs within a given power envelope.
+% 
+% \item The \texttt{MCMC-Laplace} code gallery program is a code useful
+%   for the forward solution used as a building block in
+%   Bayesian inverse problems, and for sampling the parameter space
+%   through a Metropolis--Hastings sampler (a kind of Monte Carlo
+%   Markov Chain method).
 \end{itemize}
@@
 
-The 9.1 release includes
+The 9.2 release includes
 \href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_0_1_and_9_1_0.html}
@@
 \begin{itemize}
-\item The \texttt{VectorView} class was removed. We recommend either copying the
-      vector subset into a \texttt{Vector} or using a \texttt{BlockVector}.
-\item The function \texttt{Subscriptor::subscribe()}, used through the
-  \texttt{SmartPointer} class, now requires a pointer to a
-      \texttt{std::atomic<bool>} that tracks whether or not the pointer to the
-      subscribed-to object is still valid.
-\item The \texttt{ConstraintMatrix} class gained a template parameter for the scalar
-      type and was been renamed \texttt{AffineConstraints}. Several methods that
-      take vectors or matrices as arguments,
-      such as \texttt{AffineConstraints::distribute\_local\_to\_global()},
-      now require that all matrix and vector arguments have matching number
-      types.
-\item Similarly, the functions \texttt{create\_mass\_matrix} and
-      \texttt{create\_boundary\_mass\_matrix} in the \texttt{MatrixCreator}
-      namespace no longer
-      support matrix and vector objects of different types.
+\item
+% \item The \texttt{VectorView} class was removed. We recommend either copying the
+%       vector subset into a \texttt{Vector} or using a \texttt{BlockVector}.
+% \item The function \texttt{Subscriptor::subscribe()}, used through the
+%   \texttt{SmartPointer} class, now requires a pointer to a
+%       \texttt{std::atomic<bool>} that tracks whether or not the pointer to the
+%       subscribed-to object is still valid.
+% \item The \texttt{ConstraintMatrix} class gained a template parameter for the scalar
+%       type and was been renamed \texttt{AffineConstraints}. Several methods that
+%       take vectors or matrices as arguments,
+%       such as \texttt{AffineConstraints::distribute\_local\_to\_global()},
+%       now require that all matrix and vector arguments have matching number
+%       types.
+% \item Similarly, the functions \texttt{create\_mass\_matrix} and
+%       \texttt{create\_boundary\_mass\_matrix} in the \texttt{MatrixCreator}
+%       namespace no longer
+%       support matrix and vector objects of different types.
 \end{itemize}

### 2020-04-24 9.2/paper.tex
Make myself co-author

@@
 %     Luca Heltai,
-%     Martin Kronbichler,
+    Martin Kronbichler,
 %     Ross Maguire Kynch,
@@
 %
-% \author[9]{Martin~Kronbichler}
-% \affil[9]{Institute for Computational Mechanics,
-%   Technical University of Munich,
-%   Boltzmannstr.~15, 85748 Garching, Germany.
-%   {\texttt{kronbichler@lnm.mw.tum.de}}}
+ \author[9]{Martin~Kronbichler}
+ \affil[9]{Institute for Computational Mechanics,
+   Technical University of Munich,
+   Boltzmannstr.~15, 85748 Garching, Germany.
+   {\texttt{kronbichler@lnm.mw.tum.de}}}
 %
@@
 
-In addition, there are four new tutorial programs and one new code
-gallery program:
+In addition, there are seven new tutorial programs:
 \begin{itemize}
-\item 
-% \texttt{step-61} is a program that implements the ``weak
-%   Galerkin'' method to solve the Laplace equation. Weak Galerkin
-%   methods are related to the Hybridized Discontinuous Galerkin method
-%   in that they introduce degrees of freedom on the interfaces between
-%   cells, but they do not require the reformulation of the problem as a
-%   first-order system and instead re-define what the gradient of a
-%   discontinuous function is.
-% 
-% \item \texttt{step-62} demonstrates the solution of problems related
-%   to phononic or photonic crystals. Among the techniques shown in this
-%   program is the solution of complex-valued linear systems, and the
-%   use of absorbing boundary conditions through the Perfectly Matched
-%   Layer technique.
-% 
-% \item \texttt{step-63} implements a geometric multigrid preconditioner
-%   and solver for the advection-diffusion equation, yielding optimal
-%   complexity. The tutorial compares
-%   point-based smoothers to cell-based smoothers and
-%   demonstrates the effect of downstream ordering on smoother performance.
-% 
-% \item \texttt{step-64} demonstrates the usage of matrix-free methods on Nvidia GPUs.
-% GPUs are shown to be advantageous for these kind of operations because of their superior
-% hardware characteristics, in particular a higher memory bandwidth than server
-% CPUs within a given power envelope.
-% 
-% \item The \texttt{MCMC-Laplace} code gallery program is a code useful
-%   for the forward solution used as a building block in
-%   Bayesian inverse problems, and for sampling the parameter space
-%   through a Metropolis--Hastings sampler (a kind of Monte Carlo
-%   Markov Chain method).
+\item \texttt{step-47}
+\item \texttt{step-50}
+\item \texttt{step-58}
+\item \texttt{step-65}
+\item \texttt{step-67}
+\item \texttt{step-69}
+\item \texttt{step-70}
 \end{itemize}
@@
 The 9.2 release includes
-\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_0_1_and_9_1_0.html}
-     {around 15 incompatible changes}; see \cite{changes91}. The majority of these changes
+\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_1_1_and_9_2_0.html}
+     {around 15 incompatible changes}; see \cite{changes92}. The majority of these changes
 should not be visible to typical user codes; some remove previously
@@
 The two previous releases of \dealii can be cited as
-\cite{dealII85,dealII90}.
+\cite{dealII90,dealII91}.
 

### 2020-05-08 9.2/paper.tex
Add myself to the author list.
Add myself to the author list
add myself

@@
   pdfauthor={
-%     Daniel Arndt,
-%     Wolfgang Bangerth,
+     Daniel Arndt,
+     Wolfgang Bangerth,
 %     Thomas C. Clevenger,
@@
 
-% \author[*1]{Daniel Arndt}
-% \affil[1]{Computational Engineering and Energy Sciences Group,
-%   Computional Sciences and Engineering Division,
-%   Oak Ridge National Laboratory, 1 Bethel Valley Rd.,
-%   TN 37831, USA.
-%   {\texttt{\{arndtd,turcksinbr\}@ornl.gov}}}
-%
-% \author[2]{Wolfgang~Bangerth}
-% \affil[2]{Department of Mathematics, Colorado State University, Fort
-%   Collins, CO 80523-1874, USA.
-%   {\texttt{bangerth@colostate.edu},
-%    \texttt{harper@math.colostate.edu}}}
-%
+ \author[*1]{Daniel Arndt}
+ \affil[1]{Computational Engineering and Energy Sciences Group,
+   Computional Sciences and Engineering Division,
+   Oak Ridge National Laboratory, 1 Bethel Valley Rd.,
+   TN 37831, USA.
+   \texttt{arndtd@ornl.gov}}
+
+ \author[2]{Wolfgang~Bangerth}
+ \affil[2]{Department of Mathematics, Colorado State University, Fort
+   Collins, CO 80523-1874, USA.
+   \texttt{bangerth@colostate.edu}}
+
 % \author[3]{Thomas~C.~Clevenger}
@@
 %
-% \author[11]{Matthias~Maier}
-% \affil[11]{Department of Mathematics,
-%   Texas A\&M University,
-%   3368 TAMU,
-%   College Station, TX 77845, USA.
-%   {\texttt{maier@math.tamu.edu}}}
+\author[11]{Matthias~Maier}
+\affil[11]{Department of Mathematics,
+  Texas A\&M University,
+  3368 TAMU,
+  College Station, TX 77845, USA.
+  {\texttt{maier@math.tamu.edu}}}
 %

### 2020-05-10 9.2/paper.tex
Add sections: ca, pft, mf

@@
 \usepackage{xspace}
+
+%\renewcommand{\baselinestretch}{2.0}
 
@@
     Martin Kronbichler,
+    Peter Munch,
 %     Ross Maguire Kynch,
@@
    Boltzmannstr.~15, 85748 Garching, Germany.
-   {\texttt{kronbichler@lnm.mw.tum.de}}}
+   {\texttt{kronbichler/munch@lnm.mw.tum.de}}}
 %
@@
   {\texttt{maier@math.tamu.edu}}}
+
+  \author[9,13]{Peter Munch}
+
+ \affil[13]{Institute of Materials Research, Materials Mechanics, 
+ Helmholtz-Zentrum Geesthacht, 
+ Max-Planck-Str. 1, 21502 Geesthacht, Germany.
+   {\texttt{peter.muench@hzg.de}}}
+
 %
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Improved large-scale performance}
+\label{subsec:performance}
+
+Large-scale simulations with 304,128 cores have revealed bottlenecks in release 
+9.1 during initialization due to the usage of expensive collective operations 
+like \texttt{MPI\_Allgather()} and \texttt{MPI\_\allowbreak Alltoall()}. These 
+operations are used to retrieve or even store information about all processes. 
+For example, all processes have stored in  an array the number of degrees of 
+freedom each process owns. This information is in particular needed to set up 
+the  \texttt{Utilities::MPI::Partitioner} class, which contains the 
+point-to-point communication pattern for vector ghost-value updates and 
+compressions. In release 9.2, we have removed such arrays and have replaced 
+the \texttt{MPI\_Allgather}/\allowbreak\texttt{MPI\_\allowbreak Alltoall} 
+function calls by consensus algorithms~\cite{hoefler2010scalable}, which can be 
+found in the namespace \texttt{Utilities::\allowbreak MPI::\allowbreak ConsensusAlgorithms}: now, only the locally relevant information is computed 
+(and recomputed) when needed, using these algorithms.
+
+Consensus algorithms are algorithms dedicated to efficient dynamic-sparse 
+communication patterns. In this context, the term ``dynamic-sparse'' means 
+that by the time this function is called, the other processes do not know 
+yet that they have to answer requests and
+each process only has to communicate with a small subset of processes of the 
+MPI communicator. We provide two flavors of the consensus algorithm: the two-step 
+approach \texttt{ConsensusAlgorithms::PEX} and the \texttt{ConsensusAlgorithms::NBX}, 
+which uses only point-to-point communications and a single \texttt{MPI\_IBarrier()}. 
+The class \texttt{ConsensusAlgorithms::Selector} selects one of the two previous 
+algorithms, depending on the number of processes.
+
+Due to the excellent scalability of the consensus algorithms, users are encouraged 
+to use them for their own dynamic-sparse problems by providing a list of target 
+processes and pack/unpack routines either by implementing the interface 
+\texttt{ConsensusAlgorithms::Process} or by providing \texttt{std::function} 
+objects to \texttt{ConsensusAlgorithms::AnonymousProcess}.
+
+The \texttt{ConsensusAlgorithms} are used by now internally in many places. These 
+places are appropriate starting points for users for their own application of the 
+\texttt{ConsensusAlgorithms} infrastructure.
+For example, to set up the partitioners, the new function 
+\texttt{compute\_index\_owner() } is used: given an index set containing the 
+locally owned indices and an index set containing the ghost indices, it returns 
+the owner of the ghost indices. Consensus algorithms are used now also in the 
+functions \texttt{compute\_point\_to\_point\_communication\_pattern()} and 
+\texttt{compute\_\allowbreak n\_\allowbreak point\_\allowbreak to\_\allowbreak point\_\allowbreak communications()}. 
+Furthermore, it is utilized in the class \texttt{NoncontiguousPartitioner} to 
+efficiently permute distributed solution vectors globally in an arbitrary order, e.g., 
+to interface with external libraries that prescribe a certain partitioning and 
+padding of the data.
+
+By replacing the collective communications during setup and removing the arrays 
+that contain information for each process (enabled by the application of consensus 
+algorithms and other modifications---a full list of modifications leading to this 
+improvement can be found online), we were able to significantly improve the setup 
+time for large-scale simulations and to solve a Poisson problem with multigrid 
+with 12T unknowns. 
+Figure~\ref{} compares the timings of a simulation (incl. setup) with the 
+previous release 9.1 and with the current release 9.2. 
+{\color{red}TODO[Peter/Martin] description of the results}
+
+The new code has been also applied to solve problems with adaptively refined 
+meshes with more than 4B unknowns. {\color{red}TODO[Timo]}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{A new fully distributed triangulation class}
+\label{subsec:pft}
+
+By release 9.1, \texttt{deal.II} had three types of triangulation classes: the 
+serial triangulation class \texttt{Triangulation} as well as the parallel 
+triangulation classes \texttt{parallel::shared::Triangulation} and \texttt{parallel::distributed::Triangulation}. The latter builds around a serial 
+triangulation and uses \texttt{p4est} as an oracle during adaptive mesh refinement. 
+All these triangulation classes have in common that the coarse grid is shared by 
+all processes and the actual mesh used for computations is constructed by repeated 
+local and/or global refinement, which adapts nicely to curved boundaries described 
+by the \texttt{Manifold} class. However, this way to construct a computational 
+mesh has its limitations in industrial applications where, often, the mesh comes 
+from an external CAD program in the form of a file that already contains millions 
+or tens of millions of cells with a similar number of vertices. In such a case, 
+refining a mesh is not practical, since it would increase the computational effort 
+and new vertices would not be placed on curved boundaries. A problem that arises 
+for such large grids in the context how meshes have been treated in \texttt{deal.II} 
+until now is that the coarse grid, i.e., potentially the whole mesh is shared by 
+all processes. It might be a major difficulty in MPI-only parallelized applications 
+on modern multi-core processors, since these applications might already run out of 
+main memory during reading the mesh. Not even increasing the number of processes 
+might help in this situation.
+
+The new class \texttt{parallel::fullydistributed::Triangulation} targets this issue 
+by distributing also the coarse grid, which is the reason for the name of the 
+chosen namespace: it distributes the coarse grid as well as the refinement levels. Such 
+a triangulation can be created by providing a \texttt{TriangulationDescription::Description} struct to each process, containing 
+1) the relevant data to construct the local part of the coarse grid, 2) the 
+translation of the local coarse-cell IDs to globally unique IDs, 3) the hierarchy 
+of mesh refinements, and 4) the owner of the cells on the active mesh level as well 
+as on the multigrid levels. Once the triangulation is set up with this struct, no 
+changes to the mesh are allowed at the moment.
+
+The \texttt{TriangulationDescription::Description} struct can be filled manually or 
+by the utility functions from the \texttt{TriangulationDescription::Utilities} 
+namespace. The function \texttt{create\_\allowbreak description\_\allowbreak 
+from\_ \allowbreak triangulation()} can convert a base triangulation (partitioned 
+serial \texttt{Tri\-angulation} and \texttt{parallel::distributed::Triangulation}) 
+to such a struct. The advantage of this approach is that all known utility 
+functions from the namespaces \texttt{GridIn} and \texttt{GridTools} can be used 
+on the base triangulations before converting them to the structs. Since this 
+function suffers from the same main memory problems as described above, we also 
+provide the function \texttt{create\_description\_from\_triangulation\_in\_groups()}, 
+which creates the structs only on the master process in a process group. These 
+structs are filled one by one and are sent to the relevant processes once they are 
+ready. A sensible process group size might contain all processes of one compute node.
+
+The new (fully) distributed triangulation class works---in contrast to  
+\texttt{parallel::distributed::\allowbreak Tri\-an\-gu\-la\-tion}---not only for 2D- and 3D- but also for 
+1D-problems. It can be used in the context of geometric multigrid methods and 
+supports periodic boundary conditions. Furthermore, hanging nodes are supported.
+
+We intend to extend the usability of the new triangulation class in regard of 
+different aspects, e.g., I/O. In addition, we would like to enable adaptive mesh 
+refinement, a feature of the other triangulation classes, which is very much 
+appreciated by many users. For repartitioning, we plan to use an oracle approach 
+known from \texttt{parallel::distributed::Triangulation}. Here, we would like to 
+rely on a user-provided partitioner, which might also be a graph partitioner.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Advances of the SIMD capabilities and the matrix-free infrastructure}
+\label{subsec:mf}
+
+%\begin{itemize}
+%\item ECL
+%\item VectorizedArrayType
+%\end{itemize}
+
+The class \texttt{VectorizedArray<Number>} is a key ingredient for the high 
+node-level performance of the matrix-free algorithms in deal.II. It is a wrapper 
+class around $n$ vector entries of type \texttt{Number} and delegates relevant 
+function calls to appropriate Intrinsics instructions. Up to release 9.1, the 
+vector length $n$ has been set at compile time of the library to the highest 
+possible value supported by the given processor architecture.
+
+The class \texttt{VectorizedArray} has been made more user-friendly by making 
+it compatible with the STL algorithms found in the header \texttt{<algorithm>}. 
+Now, it has following features:
+\begin{itemize}
+\item \texttt{VectorizedArray::size()} returns the vector length. This function 
+replaces the public static attribute \texttt{VectorizedArray::n\_array\_elements}, 
+which has been deprecated.
+\item \texttt{VectorizedArray::value\_type} contains the underlying number type of 
+the array.
+\item \texttt{VectorizedArray} has an output operator 
+\texttt{std::ostream\& operator<<(\&out, \&p)}.
+\item \texttt{VectorizedArray::begin()} and \texttt{VectorizedArray::end()} allow 
+range-based iteration over all vector entries.
+\end{itemize}
+Furthermore, the \texttt{VectorizedArray} class supports the following (tested) 
+algorithms: \texttt{std::\allowbreak ad\-vance()}, \texttt{std::distance()}, and \texttt{std::max\_element()}.
+
+It has been also extended with the second optional template argument 
+\texttt{VectorizedArray<Number, size>} with \texttt{size} being related to the 
+vector length, i.e., the number of lanes to be used and the instruction set to be 
+used. By default, the number is set to the highest value supported by the given 
+hardware, i.e., \texttt{VectorizedArray<double>} is translated on Skylake-based 
+processors to \texttt{VectorizedArray<double, 8>}. A full list of supported 
+vector lengths are presented in Table~\ref{tab:vectorizedarray}.
+
+All matrix-free related classes (like \texttt{MatrixFree} and \texttt{FEEvaluation}) 
+have been templated with the floating-point number type \texttt{Number} (e.g. \texttt{double} or \texttt{float}); the computations were performed implicitly 
+on \texttt{VectorizedArray<Number>} structs with the highest 
+instruction-set-architecture extension, with each lane responsible for a separate 
+cell (vectoriziation over elements). In release 9.2, all matrix-free classes 
+have been extended with a new optional template argument specifying the 
+\texttt{VectorizedArrayType}. This allows users to select the vector length/ISA and, 
+as a consequence, the number of cells to be processed at once. 
+\begin{table}
+\caption{Supported vector lengths of the class \texttt{VectorizedArray} and 
+the corresponding instruction-set-architecture extensions. }\label{tab:vectorizedarray}
+\centering
+\begin{tabular}{ccc}
+\toprule
+\textbf{double} & \textbf{float} & \textbf{ISA}\\
+\midrule
+VectorizedArray<double, 1> & VectorizedArray<float, 1> & (auto-vectorization) \\
+VectorizedArray<double, 2> & VectorizedArray<float, 4> & SSE2 \\ 
+VectorizedArray<double, 4> & VectorizedArray<float, 8> & AVX/AVX2 \\ 
+VectorizedArray<double, 8> & VectorizedArray<float, 16> & AVX-512 \\ 
+\bottomrule
+\end{tabular}
+
+\caption{Comparison of relevant SIMD-related classes in deal.II and C++20.}\label{tab:simd}
+\centering
+\begin{tabular}{cc}
+\toprule
+\textbf{VectorizedArray (deal.II)} & \textbf{std::simd (C++20)} \\
+\midrule
+VectorizedArray<Number> & std::experimental::native\_simd<Number> \\
+VectorizedArray<Number, size> & std::experimental::fixed\_size\_simd<Number, size> \\ \bottomrule
+\end{tabular}
+\end{table}
+
+In standard (2D/3D) matrix-free applications with moderate polynomial degrees, 
+we found that there is no reason to modify the default vector length of 
+\texttt{VectorizedArray}, since it reaches the highest possible computational 
+throughput despite of increased memory footprint. However, in the deal.II-based 
+library \texttt{hyper.deal}~\cite{munch2020hyperdeal}, where the same matrix-free 
+infrastructure was used for solving the 6D Vlasov--Poisson equation with high-order 
+discontinuous Galerkin methods (with more than 1024 degrees of freedom per cell), the 
+benefit of decreasing the number of cells processed by a single SIMD instruction was 
+shown. That library works with \texttt{MatrixFree} objects of different SIMD-vector 
+length in the same application, which would not have been possible before this 
+release.
+
+In a next step, we intend to support that users could set 
+\texttt{VectorizedArrayType} to \texttt{Number}, which would lead to the usage not 
+of \texttt{VectorizedArray<Number, 1>} but of a specialized code-path exploiting 
+vectorization within an element~\cite{KronbichlerKormann2019}.
+
+A side effect of introducing the new template argument \texttt{VectorizedArrayType} 
+in the \texttt{MatrixFree} classes is that any data structures 
+\texttt{VectorizedArrayType} can be processed if they support required 
+functionalities like \texttt{size()} or \texttt{value\_type}. In this context, we 
+would like to highlight that the new \texttt{C++20} feature \texttt{std::simd} 
+can be processed by the matrix-free infrastructure with minor internal 
+adjustment as an open pull request shows 
+(see \url{https://github.com/dealii/dealii/pull/9994}).  
+Table~\ref{tab:simd} gives a comparison of the deal.II-specific SIMD classes and 
+the equivalent C++20 classes. We welcome the standardization of the SIMD 
+parallelization paradigm in C++ and intend to replace step by step our own 
+wrapper class, which has been continuously developed over the last decade. We 
+would like to emphasize that the work invested in this class was not in vain, 
+since many performance-relevant utility functions implemented with \texttt{VectorizedArray} in mind (e.g., \texttt{vectorized\_load\_and\_transpose} 
+and \texttt{vectorized\_transpose\_and\_store}) will be still used, since they 
+have not become part of the standard.
+
+Further additions to the \texttt{MatrixFree} infrastructure consist of:
+\begin{itemize}
+\item a new variant of \texttt{MatrixFree::cell\_loop()}: It takes two
+\texttt{std::function} objects with ranges on the locally owned degrees of freedom, one
+with work to be scheduled before the cell operation first touches some
+unknowns and another with work to be executed after the cell operation last 
+touches them. The goal of
+these functions is to bring vector operations close to the time when the
+vector entries are accessed by the cell operation, which increases the cache
+hit rate of modern processors by improved temporal locality. 
+\item a new form of loop \texttt{MatrixFree::loop\_cell\_centric()}: This 
+kind of loop can be used in the context of discontinuous Galerkin methods, 
+where both cell and face integrals have to be evaluated. While in the case of 
+the traditional \texttt{loop}, cell and face integrals have been performed 
+independently, the new loop performs all cell and face integrals of a cell in 
+one go. This includes that each face integral has to be evaluated twice, but 
+entries have to be written into the solution vector only once with improved 
+data locality. Previous publications based on \texttt{deal.II} have shown the 
+relevance of the latter aspect for reaching higher performance. 
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Advances in GPU support}
+\label{subsec:gpu}
+
+\begin{itemize}
+\item overlapping of computation and communication in the case of CUDA-aware MPI
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{New and improved tutorial and code gallery programs}
@@
 \item \texttt{step-58}
-\item \texttt{step-65}
-\item \texttt{step-67}
+\item \texttt{step-65} presents \texttt{TransfiniteInterpolationManifold}, a
+manifold class that can propagate curved boundary information into the
+interior of a computational domain, and \texttt{MappingQCache} for fast operations for
+expensive manifolds.
+\item \texttt{step-67} presents an explicit time integrator for the
+compressible Euler equations discretized with a high-order discontinuous
+Galerkin scheme using the matrix-free infrastructure. Besides the use of
+matrix-free evaluators for systems of equations and over-integration, it also
+presents \texttt{MatrixFreeOperators::CellwiseInverseMassMatrix}, a fast implementation
+of the action of the inverse mass matrix in the DG setting using tensor
+products. Furthermore, this tutorial demonstrates i.a. the usage of the new 
+pre and post operations which can be passed to \texttt{cell\_loop()} 
+(see also Subsection~\ref{subsec:mf}) and discusses performance-related aspects.
 \item \texttt{step-69}
@@
 \begin{itemize}
+\item The functions:
+\begin{itemize}
+\item \texttt{DoFHandler::loccaly\_owned\_dofs\_per\_processor()}
+\item \texttt{DoFHandler::loccaly\_owned\_mg\_dofs\_per\_processor()}
+\end{itemize} have been deprecated.
+As discussed in Subsection~\ref{subsec:performance}, deal.II does not store information for all processes on all processes processes, but only the local information or the locally-relevant information.
+Users are asked to construct the global information on their own, e.g.
+by calling \texttt{Utilities::MPI::Allgather(locally\_owned\_info(), comm)}.
 \item
+
 % \item The \texttt{VectorView} class was removed. We recommend either copying the

### 2020-05-11 9.2/paper.tex
Incorporate martins suggestions

--- 2020-05-08 9.2/paper.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Improved large-scale performance}
+\label{subsec:performance}
+
+Large-scale simulations with 304,128 cores have revealed bottlenecks in release 
+9.1 during initialization due to the usage of expensive collective operations 
+like \texttt{MPI\_Allgather()} and \texttt{MPI\_\allowbreak Alltoall()}, e.g., during the 
+pre-computationof the index ranges of all processes, which have been stored in  an array. 
+This information is needed to set up 
+the  \texttt{Utilities::MPI::Par\-ti\-ti\-oner} class. 
+In release 9.2, we have removed such arrays and have replaced 
+the \texttt{MPI\_Allgather}/\allowbreak\texttt{MPI\_\allowbreak Alltoall} 
+function calls by consensus algorithms~\cite{hoefler2010scalable}, which can be 
+found in the namespace \texttt{Utilities::\allowbreak MPI::\allowbreak ConsensusAlgorithms} (short: \texttt{CA}): now, only the locally relevant information about the index ranges is (re)computed when needed, using these algorithms.
+We provide two flavors of the consensus algorithm: the two-step 
+approach \texttt{CA::PEX} and the \texttt{CA::NBX}, 
+which uses only point-to-point communications and a single \texttt{MPI\_IBarrier()}.
+
+%Consensus algorithms are algorithms dedicated to efficient dynamic-sparse 
+%communication patterns. In this context, the term ``dynamic-sparse'' means 
+%that by the time this function is called, the other processes do not know 
+%yet that they have to answer requests and
+%each process only has to communicate with a small subset of processes of the 
+%MPI communicator. We provide two flavors of the consensus algorithm: the two-step 
+%approach \texttt{ConsensusAlgorithms::PEX} and the \texttt{ConsensusAlgorithms::NBX}, 
+%which uses only point-to-point communications and a single \texttt{MPI\_IBarrier()}. 
+%The class \texttt{ConsensusAlgorithms::Selector} selects one of the two previous 
+%algorithms, depending on the number of processes.
+
+Users can use the new algorithms for their own dynamic-sparse problems by 
+providing a list of target 
+processes and pack/unpack routines either by implementing the interface 
+\texttt{CA::\allowbreak Process} or by providing \texttt{std::function} 
+objects to \texttt{CA::AnonymousProcess}.
+
+By replacing the collective communications during setup and removing the arrays 
+that contain information for each process (enabled by the application of consensus 
+algorithms and other modifications---a full list of modifications leading to this 
+improvement can be found online), we were able to significantly improve the setup 
+time for large-scale simulations and to solve a Poisson problem with multigrid 
+with 12T unknowns. 
+Figure~\ref{} compares the timings of a simulation (incl. setup) with the 
+previous release 9.1 and with the current release 9.2. 
+{\color{red}TODO[Peter/Martin] description of the results}
+
+The new code has been also applied to solve problems with adaptively refined 
+meshes with more than 4B unknowns. {\color{red}TODO[Timo]}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{A new fully distributed triangulation class}
+\label{subsec:pft}
+
+By release 9.1, all triangulation classes of \texttt{deal.II} have in common that the coarse grid is shared by 
+all processes and the actual mesh used for computations is constructed by repeated 
+refinement. However, this has its limitations in industrial applications where, often, the mesh comes 
+from an external mesh generator in the form of a file that already contains millions 
+or tens of millions of cells. For such configurations, applications might already 
+run out of  main memory while reading the complete mesh by each MPI process.
+
+The new class \texttt{parallel::fullydistributed::Triangulation} targets this issue 
+by distributing also the coarse grid. Such 
+a triangulation can be created by providing a \texttt{Triangulation\-De\-scrip\-tion::Description} struct to each process, containing 
+1) the relevant data to construct the local part of the coarse grid, 2) the 
+translation of the local coarse-cell IDs to globally unique IDs, 3) the hierarchy 
+of mesh refinements, and 4) the owner of the cells on the active mesh level as well 
+as on the multigrid levels. Once the triangulation is set up with this struct, no adaptive
+changes to the mesh are allowed at the moment.
+
+%The \texttt{TriangulationDescription::Description} struct can be filled manually or 
+%by the utility functions from the \texttt{TriangulationDescription::Utilities} 
+%namespace. The function \texttt{create\_\allowbreak description\_\allowbreak 
+%from\_ \allowbreak triangulation()} can convert a base triangulation (partitioned 
+%serial \texttt{Tri\-angulation} and \texttt{parallel::distributed::Triangulation}) 
+%to such a struct. The advantage of this approach is that all known utility 
+%functions from the namespaces \texttt{GridIn} and \texttt{GridTools} can be used 
+%on the base triangulations before converting them to the structs. Since this 
+%function suffers from the same main memory problems as described above, we also 
+%provide the function \texttt{create\_description\_from\_triangulation\_in\_groups()}, 
+%which creates the structs only on the master process in a process group. These 
+%structs are filled one by one and are sent to the relevant processes once they are 
+%ready. A sensible process group size might contain all processes of one compute node.
+
+The new (fully) distributed triangulation class works---in contrast to  
+\texttt{parallel::distributed::\allowbreak Tri\-an\-gu\-la\-tion}---not only for 2D- and 3D- but also for 
+1D-problems. It can be used in the context of geometric multigrid methods and 
+supports periodic boundary conditions. Furthermore, hanging nodes are supported.
+
+%We intend to extend the usability of the new triangulation class in regard of 
+%different aspects, e.g., I/O. In addition, we would like to enable adaptive mesh 
+%refinement, a feature of the other triangulation classes, which is very much 
+%appreciated by many users. For repartitioning, we plan to use an oracle approach 
+%known from \texttt{parallel::distributed::Triangulation}. Here, we would like to 
+%rely on a user-provided partitioner, which might also be a graph partitioner.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Advances of the SIMD capabilities and the matrix-free infrastructure}
+\label{subsec:mf}
+
+%\begin{itemize}
+%\item ECL
+%\item VectorizedArrayType
+%\end{itemize}
+
+The class \texttt{VectorizedArray<Number>} is a key ingredient for the high 
+node-level performance of the matrix-free algorithms in deal.II~\cite{KronbichlerKormann2012, KronbichlerKormann2019}. It is a wrapper 
+class around $n$ vector entries of type \texttt{Number} and delegates relevant 
+function calls to appropriate Intrinsics instructions. Up to release 9.1, the 
+vector length $n$ has been set at compile time of the library to the highest 
+possible value supported by the given processor architecture.
+
+The class \texttt{VectorizedArray} has been made more user-friendly by making 
+it compatible with the STL algorithms found in the header \texttt{<algorithm>}. 
+The length of the  vector can now be queried by  \texttt{VectorizedArray::size()} and its underlying number type by \texttt{VectorizedArray::value\_type}. 
+Furthermore, the \texttt{VectorizedArray} class supports range-based iterations over its entries and, i.a., the following 
+algorithms: \texttt{std::\allowbreak ad\-vance()}, \texttt{std::distance()}, and \texttt{std::max\_element()}.
+
+It has been also extended with the second optional template argument 
+\texttt{VectorizedArray<Number, size>} with \texttt{size} being related to the 
+vector length, i.e., the number of lanes to be used and the instruction set to be 
+used. By default, the number is set to the highest value supported by the given 
+hardware. A full list of supported 
+vector lengths are presented in Table~\ref{tab:vectorizedarray}.
+
+All matrix-free related classes (like \texttt{MatrixFree} and \texttt{FEEvaluation}) 
+have been templated with the floating-point number type \texttt{Number} (e.g. \texttt{double} or \texttt{float}); the computations were performed implicitly 
+on \texttt{VectorizedArray<Number>} structs with the highest 
+instruction-set-architecture extension, with each lane responsible for a separate 
+cell (vectoriziation over elements). In release 9.2, all matrix-free classes 
+have been extended with a new optional template argument specifying the 
+\texttt{VectorizedArrayType}. This allows users to select the vector length/ISA and, 
+as a consequence, the number of cells to be processed at once directly in their applications: 
+The deal.II-based 
+library \texttt{hyper.deal}~\cite{munch2020hyperdeal}, which solves the 6D Vlasov--Poisson equation with high-order 
+discontinuous Galerkin methods (with more than 1024 degrees of freedom per cell), works with \texttt{MatrixFree} objects of different SIMD-vector 
+length in the same application and benefits---in terms of performance---by the possibility of decreasing the number of cells processed by a single SIMD instruction.
+
+\begin{table}
+\caption{Supported vector lengths of the class \texttt{VectorizedArray} and 
+the corresponding instruction-set-architecture extensions. }\label{tab:vectorizedarray}
+\centering
+\begin{tabular}{ccc}
+\toprule
+\textbf{double} & \textbf{float} & \textbf{ISA}\\
+\midrule
+VectorizedArray<double, 1> & VectorizedArray<float, 1> & (auto-vectorization) \\
+VectorizedArray<double, 2> & VectorizedArray<float, 4> & SSE2 \\ 
+VectorizedArray<double, 4> & VectorizedArray<float, 8> & AVX/AVX2 \\ 
+VectorizedArray<double, 8> & VectorizedArray<float, 16> & AVX-512 \\ 
+\bottomrule
+\end{tabular}
+
+\caption{Comparison of relevant SIMD-related classes in deal.II and C++20.}\label{tab:simd}
+\centering
+\begin{tabular}{cc}
+\toprule
+\textbf{VectorizedArray (deal.II)} & \textbf{std::simd (C++20)} \\
+\midrule
+VectorizedArray<Number> & std::experimental::native\_simd<Number> \\
+VectorizedArray<Number, size> & std::experimental::fixed\_size\_simd<Number, size> \\ \bottomrule
+\end{tabular}
+\end{table}
+
+A side effect of introducing the new template argument \texttt{VectorizedArrayType} 
+in the \texttt{MatrixFree} classes is that any data structures 
+\texttt{VectorizedArrayType} can be processed if they support required 
+functionalities like \texttt{size()} or \texttt{value\_type}. In this context, we 
+would like to highlight that the new \texttt{C++20} feature \texttt{std::simd} 
+can be processed by the matrix-free infrastructure with minor internal 
+adjustment as an open pull request shows 
+(see \url{https://github.com/dealii/dealii/pull/9994}).  
+Table~\ref{tab:simd} gives a comparison of the deal.II-specific SIMD classes and 
+the equivalent C++20 classes. 
+%We welcome the standardization of the SIMD 
+%parallelization paradigm in C++ and intend to replace step by step our own 
+%wrapper class, which has been continuously developed over the last decade. We 
+%would like to emphasize that the work invested in this class was not in vain, 
+%since many performance-relevant utility functions implemented with \texttt{VectorizedArray} in mind (e.g., \texttt{vectorized\_load\_and\_transpose} 
+%and \texttt{vectorized\_transpose\_and\_store}) will be still used, since they 
+%have not become part of the standard.
+
+%Further additions to the \texttt{MatrixFree} infrastructure consist of:
+%\begin{itemize}
+%\item a new variant of \texttt{MatrixFree::cell\_loop()}: It takes two
+%\texttt{std::function} objects with ranges on the locally owned degrees of freedom, one
+%with work to be scheduled before the cell operation first touches some
+%unknowns and another with work to be executed after the cell operation last 
+%touches them. The goal of
+%these functions is to bring vector operations close to the time when the
+%vector entries are accessed by the cell operation, which increases the cache
+%hit rate of modern processors by improved temporal locality. 
+%\item a new form of loop \texttt{MatrixFree::loop\_cell\_centric()}: This 
+%kind of loop can be used in the context of discontinuous Galerkin methods, 
+%where both cell and face integrals have to be evaluated. While in the case of 
+%the traditional \texttt{loop}, cell and face integrals have been performed 
+%independently, the new loop performs all cell and face integrals of a cell in 
+%one go. This includes that each face integral has to be evaluated twice, but 
+%entries have to be written into the solution vector only once with improved 
+%data locality. Previous publications based on \texttt{deal.II} have shown the 
+%relevance of the latter aspect for reaching higher performance. 
+%\end{itemize}
+
+In a next step, we intend to support that users could set 
+\texttt{VectorizedArrayType} to \texttt{Number}, which would lead to the usage not 
+of \texttt{VectorizedArray<Number, 1>} but of a specialized code-path exploiting 
+vectorization within an element~\cite{KronbichlerKormann2019}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Advances in GPU support}
+\label{subsec:gpu}
+
+\begin{itemize}
+\item overlapping of computation and communication in the case of CUDA-aware MPI
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{New and improved tutorial and code gallery programs}
@@
 \item \texttt{step-58}
-\item \texttt{step-65}
-\item \texttt{step-67}
+\item \texttt{step-65} presents \texttt{TransfiniteInterpolationManifold}, a
+manifold class that can propagate curved boundary information into the
+interior of a computational domain, and \texttt{MappingQCache}, which can sample 
+the information of expensive manifolds in the points of a \texttt{MappingQ} and 
+cache it for further use.
+\item \texttt{step-67} presents an explicit time integrator for the
+compressible Euler equations discretized with a high-order discontinuous
+Galerkin scheme using the matrix-free infrastructure. Besides the use of
+matrix-free evaluators for systems of equations and over-integration, it also
+presents \texttt{MatrixFreeOperators::CellwiseInverseMassMatrix}, a fast implementation
+of the action of the inverse mass matrix in the DG setting using tensor
+products. Furthermore, this tutorial demonstrates the usage of new 
+pre and post operations, which can be passed to \texttt{cell\_loop()}, to schedule operations on sections of vectors close 
+to the matrix-vector product to increase data locality 
+and discusses performance-related aspects.
 \item \texttt{step-69}

### 2020-05-11 9.2/paper.tex
Fix some typos

@@
 like \texttt{MPI\_Allgather()} and \texttt{MPI\_\allowbreak Alltoall()}, e.g., during the 
-pre-computationof the index ranges of all processes, which have been stored in  an array. 
+pre-computation of the index ranges of all processes, which have been stored in  an array. 
 This information is needed to set up 
@@
 
-Users can use the new algorithms for their own dynamic-sparse problems by 
+Users can apply the new algorithms for their own dynamic-sparse problems by 
 providing a list of target 
@@
 or tens of millions of cells. For such configurations, applications might already 
-run out of  main memory while reading the complete mesh by each MPI process.
+run out of  main memory, while reading the complete mesh by each MPI process.
 
@@
 by distributing also the coarse grid. Such 
-a triangulation can be created by providing a \texttt{Triangulation\-De\-scrip\-tion::Description} struct to each process, containing 
+a triangulation can be created by providing to each process a \texttt{Triangulation\-De\-scrip\-tion::Description} struct, containing 
 1) the relevant data to construct the local part of the coarse grid, 2) the 
@@
 
-The new (fully) distributed triangulation class works---in contrast to  
+The new fully distributed triangulation class works---in contrast to  
 \texttt{parallel::distributed::\allowbreak Tri\-an\-gu\-la\-tion}---not only for 2D- and 3D- but also for 
@@
 
-It has been also extended with the second optional template argument 
+This class has been also extended with the second optional template argument 
 \texttt{VectorizedArray<Number, size>} with \texttt{size} being related to the 
@@
 hardware. A full list of supported 
-vector lengths are presented in Table~\ref{tab:vectorizedarray}.
+vector lengths is presented in Table~\ref{tab:vectorizedarray}.
 

### 2020-05-11 9.2/paper.tex
update list of contributors

--- 2020-05-11 9.2/paper.tex
@@
 \usepackage{xspace}
-
-%\renewcommand{\baselinestretch}{2.0}
 
@@
     Martin Kronbichler,
-    Peter Munch,
 %     Ross Maguire Kynch,
@@
    Boltzmannstr.~15, 85748 Garching, Germany.
-   {\texttt{kronbichler/munch@lnm.mw.tum.de}}}
+   {\texttt{kronbichler@lnm.mw.tum.de}}}
 %
@@
   {\texttt{maier@math.tamu.edu}}}
-
-  \author[9,13]{Peter Munch}
-
- \affil[13]{Institute of Materials Research, Materials Mechanics, 
- Helmholtz-Zentrum Geesthacht, 
- Max-Planck-Str. 1, 21502 Geesthacht, Germany.
-   {\texttt{peter.muench@hzg.de}}}
-
 %
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Improved large-scale performance}
-\label{subsec:performance}
-
-Large-scale simulations with 304,128 cores have revealed bottlenecks in release 
-9.1 during initialization due to the usage of expensive collective operations 
-like \texttt{MPI\_Allgather()} and \texttt{MPI\_\allowbreak Alltoall()}, e.g., during the 
-pre-computationof the index ranges of all processes, which have been stored in  an array. 
-This information is needed to set up 
-the  \texttt{Utilities::MPI::Par\-ti\-ti\-oner} class. 
-In release 9.2, we have removed such arrays and have replaced 
-the \texttt{MPI\_Allgather}/\allowbreak\texttt{MPI\_\allowbreak Alltoall} 
-function calls by consensus algorithms~\cite{hoefler2010scalable}, which can be 
-found in the namespace \texttt{Utilities::\allowbreak MPI::\allowbreak ConsensusAlgorithms} (short: \texttt{CA}): now, only the locally relevant information about the index ranges is (re)computed when needed, using these algorithms.
-We provide two flavors of the consensus algorithm: the two-step 
-approach \texttt{CA::PEX} and the \texttt{CA::NBX}, 
-which uses only point-to-point communications and a single \texttt{MPI\_IBarrier()}.
-
-%Consensus algorithms are algorithms dedicated to efficient dynamic-sparse 
-%communication patterns. In this context, the term ``dynamic-sparse'' means 
-%that by the time this function is called, the other processes do not know 
-%yet that they have to answer requests and
-%each process only has to communicate with a small subset of processes of the 
-%MPI communicator. We provide two flavors of the consensus algorithm: the two-step 
-%approach \texttt{ConsensusAlgorithms::PEX} and the \texttt{ConsensusAlgorithms::NBX}, 
-%which uses only point-to-point communications and a single \texttt{MPI\_IBarrier()}. 
-%The class \texttt{ConsensusAlgorithms::Selector} selects one of the two previous 
-%algorithms, depending on the number of processes.
-
-Users can use the new algorithms for their own dynamic-sparse problems by 
-providing a list of target 
-processes and pack/unpack routines either by implementing the interface 
-\texttt{CA::\allowbreak Process} or by providing \texttt{std::function} 
-objects to \texttt{CA::AnonymousProcess}.
-
-By replacing the collective communications during setup and removing the arrays 
-that contain information for each process (enabled by the application of consensus 
-algorithms and other modifications---a full list of modifications leading to this 
-improvement can be found online), we were able to significantly improve the setup 
-time for large-scale simulations and to solve a Poisson problem with multigrid 
-with 12T unknowns. 
-Figure~\ref{} compares the timings of a simulation (incl. setup) with the 
-previous release 9.1 and with the current release 9.2. 
-{\color{red}TODO[Peter/Martin] description of the results}
-
-The new code has been also applied to solve problems with adaptively refined 
-meshes with more than 4B unknowns. {\color{red}TODO[Timo]}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{A new fully distributed triangulation class}
-\label{subsec:pft}
-
-By release 9.1, all triangulation classes of \texttt{deal.II} have in common that the coarse grid is shared by 
-all processes and the actual mesh used for computations is constructed by repeated 
-refinement. However, this has its limitations in industrial applications where, often, the mesh comes 
-from an external mesh generator in the form of a file that already contains millions 
-or tens of millions of cells. For such configurations, applications might already 
-run out of  main memory while reading the complete mesh by each MPI process.
-
-The new class \texttt{parallel::fullydistributed::Triangulation} targets this issue 
-by distributing also the coarse grid. Such 
-a triangulation can be created by providing a \texttt{Triangulation\-De\-scrip\-tion::Description} struct to each process, containing 
-1) the relevant data to construct the local part of the coarse grid, 2) the 
-translation of the local coarse-cell IDs to globally unique IDs, 3) the hierarchy 
-of mesh refinements, and 4) the owner of the cells on the active mesh level as well 
-as on the multigrid levels. Once the triangulation is set up with this struct, no adaptive
-changes to the mesh are allowed at the moment.
-
-%The \texttt{TriangulationDescription::Description} struct can be filled manually or 
-%by the utility functions from the \texttt{TriangulationDescription::Utilities} 
-%namespace. The function \texttt{create\_\allowbreak description\_\allowbreak 
-%from\_ \allowbreak triangulation()} can convert a base triangulation (partitioned 
-%serial \texttt{Tri\-angulation} and \texttt{parallel::distributed::Triangulation}) 
-%to such a struct. The advantage of this approach is that all known utility 
-%functions from the namespaces \texttt{GridIn} and \texttt{GridTools} can be used 
-%on the base triangulations before converting them to the structs. Since this 
-%function suffers from the same main memory problems as described above, we also 
-%provide the function \texttt{create\_description\_from\_triangulation\_in\_groups()}, 
-%which creates the structs only on the master process in a process group. These 
-%structs are filled one by one and are sent to the relevant processes once they are 
-%ready. A sensible process group size might contain all processes of one compute node.
-
-The new (fully) distributed triangulation class works---in contrast to  
-\texttt{parallel::distributed::\allowbreak Tri\-an\-gu\-la\-tion}---not only for 2D- and 3D- but also for 
-1D-problems. It can be used in the context of geometric multigrid methods and 
-supports periodic boundary conditions. Furthermore, hanging nodes are supported.
-
-%We intend to extend the usability of the new triangulation class in regard of 
-%different aspects, e.g., I/O. In addition, we would like to enable adaptive mesh 
-%refinement, a feature of the other triangulation classes, which is very much 
-%appreciated by many users. For repartitioning, we plan to use an oracle approach 
-%known from \texttt{parallel::distributed::Triangulation}. Here, we would like to 
-%rely on a user-provided partitioner, which might also be a graph partitioner.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Advances of the SIMD capabilities and the matrix-free infrastructure}
-\label{subsec:mf}
-
-%\begin{itemize}
-%\item ECL
-%\item VectorizedArrayType
-%\end{itemize}
-
-The class \texttt{VectorizedArray<Number>} is a key ingredient for the high 
-node-level performance of the matrix-free algorithms in deal.II~\cite{KronbichlerKormann2012, KronbichlerKormann2019}. It is a wrapper 
-class around $n$ vector entries of type \texttt{Number} and delegates relevant 
-function calls to appropriate Intrinsics instructions. Up to release 9.1, the 
-vector length $n$ has been set at compile time of the library to the highest 
-possible value supported by the given processor architecture.
-
-The class \texttt{VectorizedArray} has been made more user-friendly by making 
-it compatible with the STL algorithms found in the header \texttt{<algorithm>}. 
-The length of the  vector can now be queried by  \texttt{VectorizedArray::size()} and its underlying number type by \texttt{VectorizedArray::value\_type}. 
-Furthermore, the \texttt{VectorizedArray} class supports range-based iterations over its entries and, i.a., the following 
-algorithms: \texttt{std::\allowbreak ad\-vance()}, \texttt{std::distance()}, and \texttt{std::max\_element()}.
-
-It has been also extended with the second optional template argument 
-\texttt{VectorizedArray<Number, size>} with \texttt{size} being related to the 
-vector length, i.e., the number of lanes to be used and the instruction set to be 
-used. By default, the number is set to the highest value supported by the given 
-hardware. A full list of supported 
-vector lengths are presented in Table~\ref{tab:vectorizedarray}.
-
-All matrix-free related classes (like \texttt{MatrixFree} and \texttt{FEEvaluation}) 
-have been templated with the floating-point number type \texttt{Number} (e.g. \texttt{double} or \texttt{float}); the computations were performed implicitly 
-on \texttt{VectorizedArray<Number>} structs with the highest 
-instruction-set-architecture extension, with each lane responsible for a separate 
-cell (vectoriziation over elements). In release 9.2, all matrix-free classes 
-have been extended with a new optional template argument specifying the 
-\texttt{VectorizedArrayType}. This allows users to select the vector length/ISA and, 
-as a consequence, the number of cells to be processed at once directly in their applications: 
-The deal.II-based 
-library \texttt{hyper.deal}~\cite{munch2020hyperdeal}, which solves the 6D Vlasov--Poisson equation with high-order 
-discontinuous Galerkin methods (with more than 1024 degrees of freedom per cell), works with \texttt{MatrixFree} objects of different SIMD-vector 
-length in the same application and benefits---in terms of performance---by the possibility of decreasing the number of cells processed by a single SIMD instruction.
-
-\begin{table}
-\caption{Supported vector lengths of the class \texttt{VectorizedArray} and 
-the corresponding instruction-set-architecture extensions. }\label{tab:vectorizedarray}
-\centering
-\begin{tabular}{ccc}
-\toprule
-\textbf{double} & \textbf{float} & \textbf{ISA}\\
-\midrule
-VectorizedArray<double, 1> & VectorizedArray<float, 1> & (auto-vectorization) \\
-VectorizedArray<double, 2> & VectorizedArray<float, 4> & SSE2 \\ 
-VectorizedArray<double, 4> & VectorizedArray<float, 8> & AVX/AVX2 \\ 
-VectorizedArray<double, 8> & VectorizedArray<float, 16> & AVX-512 \\ 
-\bottomrule
-\end{tabular}
-
-\caption{Comparison of relevant SIMD-related classes in deal.II and C++20.}\label{tab:simd}
-\centering
-\begin{tabular}{cc}
-\toprule
-\textbf{VectorizedArray (deal.II)} & \textbf{std::simd (C++20)} \\
-\midrule
-VectorizedArray<Number> & std::experimental::native\_simd<Number> \\
-VectorizedArray<Number, size> & std::experimental::fixed\_size\_simd<Number, size> \\ \bottomrule
-\end{tabular}
-\end{table}
-
-A side effect of introducing the new template argument \texttt{VectorizedArrayType} 
-in the \texttt{MatrixFree} classes is that any data structures 
-\texttt{VectorizedArrayType} can be processed if they support required 
-functionalities like \texttt{size()} or \texttt{value\_type}. In this context, we 
-would like to highlight that the new \texttt{C++20} feature \texttt{std::simd} 
-can be processed by the matrix-free infrastructure with minor internal 
-adjustment as an open pull request shows 
-(see \url{https://github.com/dealii/dealii/pull/9994}).  
-Table~\ref{tab:simd} gives a comparison of the deal.II-specific SIMD classes and 
-the equivalent C++20 classes. 
-%We welcome the standardization of the SIMD 
-%parallelization paradigm in C++ and intend to replace step by step our own 
-%wrapper class, which has been continuously developed over the last decade. We 
-%would like to emphasize that the work invested in this class was not in vain, 
-%since many performance-relevant utility functions implemented with \texttt{VectorizedArray} in mind (e.g., \texttt{vectorized\_load\_and\_transpose} 
-%and \texttt{vectorized\_transpose\_and\_store}) will be still used, since they 
-%have not become part of the standard.
-
-%Further additions to the \texttt{MatrixFree} infrastructure consist of:
-%\begin{itemize}
-%\item a new variant of \texttt{MatrixFree::cell\_loop()}: It takes two
-%\texttt{std::function} objects with ranges on the locally owned degrees of freedom, one
-%with work to be scheduled before the cell operation first touches some
-%unknowns and another with work to be executed after the cell operation last 
-%touches them. The goal of
-%these functions is to bring vector operations close to the time when the
-%vector entries are accessed by the cell operation, which increases the cache
-%hit rate of modern processors by improved temporal locality. 
-%\item a new form of loop \texttt{MatrixFree::loop\_cell\_centric()}: This 
-%kind of loop can be used in the context of discontinuous Galerkin methods, 
-%where both cell and face integrals have to be evaluated. While in the case of 
-%the traditional \texttt{loop}, cell and face integrals have been performed 
-%independently, the new loop performs all cell and face integrals of a cell in 
-%one go. This includes that each face integral has to be evaluated twice, but 
-%entries have to be written into the solution vector only once with improved 
-%data locality. Previous publications based on \texttt{deal.II} have shown the 
-%relevance of the latter aspect for reaching higher performance. 
-%\end{itemize}
-
-In a next step, we intend to support that users could set 
-\texttt{VectorizedArrayType} to \texttt{Number}, which would lead to the usage not 
-of \texttt{VectorizedArray<Number, 1>} but of a specialized code-path exploiting 
-vectorization within an element~\cite{KronbichlerKormann2019}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Advances in GPU support}
-\label{subsec:gpu}
-
-\begin{itemize}
-\item overlapping of computation and communication in the case of CUDA-aware MPI
-\end{itemize}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{New and improved tutorial and code gallery programs}
@@
 \item \texttt{step-58}
-\item \texttt{step-65} presents \texttt{TransfiniteInterpolationManifold}, a
-manifold class that can propagate curved boundary information into the
-interior of a computational domain, and \texttt{MappingQCache}, which can sample 
-the information of expensive manifolds in the points of a \texttt{MappingQ} and 
-cache it for further use.
-\item \texttt{step-67} presents an explicit time integrator for the
-compressible Euler equations discretized with a high-order discontinuous
-Galerkin scheme using the matrix-free infrastructure. Besides the use of
-matrix-free evaluators for systems of equations and over-integration, it also
-presents \texttt{MatrixFreeOperators::CellwiseInverseMassMatrix}, a fast implementation
-of the action of the inverse mass matrix in the DG setting using tensor
-products. Furthermore, this tutorial demonstrates the usage of new 
-pre and post operations, which can be passed to \texttt{cell\_loop()}, to schedule operations on sections of vectors close 
-to the matrix-vector product to increase data locality 
-and discusses performance-related aspects.
+\item \texttt{step-65}
+\item \texttt{step-67}
 \item \texttt{step-69}
@@
 \begin{itemize}
-\item The functions:
-\begin{itemize}
-\item \texttt{DoFHandler::loccaly\_owned\_dofs\_per\_processor()}
-\item \texttt{DoFHandler::loccaly\_owned\_mg\_dofs\_per\_processor()}
-\end{itemize} have been deprecated.
-As discussed in Subsection~\ref{subsec:performance}, deal.II does not store information for all processes on all processes processes, but only the local information or the locally-relevant information.
-Users are asked to construct the global information on their own, e.g.
-by calling \texttt{Utilities::MPI::Allgather(locally\_owned\_info(), comm)}.
 \item
-
 % \item The \texttt{VectorView} class was removed. We recommend either copying the
@@
 contributed code to this release:\\
-% updated 6/05/2019
+% updated 5/11/2020 MM - TODO: remove authors of paper
+Pasquale Africa,
+Ashna Aggarwal,
 Giovanni Alzetta,
 Mathias Anselmann,
-  Daniel Appel,
-  Alexander Blank,
-  Vishal Boddu,
+Daniel Arndt,
+Wolfgang Bangerth,
+Kirana Bergstrom,
+Manaswinee Bezbaruah,
+Bruno Blais,
 Benjamin Brands,
-  Pi-Yueh Chuang,
-  Sambit Das,
+Thomas C. Clevenger,
+Katherine Cosburn,
+Denis Davydov,
+Elias Dejene,
 Stefano Dominici,
-  Nivesh Dommaraju,
+Brett Dong,
+Luel Emishaw,
+Marc Fehling,
 Niklas Fehn,
-  Isuru Fernando,
-  Andreas Fink,
-  Rene Gassm{\"o}ller,
+Rebecca Fildes,
+Menno Fraters,
+Andres Galindo,
+Daniel Garcia-Sanchez,
+Rene Gassmoeller,
+Melanie Gerault,
+Nicola Giuliani,
+Brandon Gleeson,
+Krishnakumar Gopalakrishnan,
 Alexander Grayver,
-  Joshua Hanophy,
-  Logan Harbour,
+Graham Harper,
+Nicole Hayes,
+Bang He,
+Timo Heister,
+Luca Heltai,
+Jiuhua Hu,
+Lise-Marie Imbert-Gerard,
+Manu Jayadharan,
 Daniel Jodlbauer,
-  Stefan Kaessmair,
-  Eldar Khattatov,
+Marie Kajan,
 Alexander Knieps,
-  Uwe K{\"o}cher,
-  Kurt Kremitzki,
-  Dustin Kumor,
-  Damien Lebrun-Grandie,
-  Jonathan Matthews,
-  Stefan Meggendorfer,
-  Pratik Nayak,
-  Lei Qiao,
-  Ce Qin,
+Martin Kronbichler,
+Konstantin Ladutenko,
+Charu Lata,
+Adam Lee,
+Wenyu Lei,
+Matthias Maier,
+Katrin Mang,
+Mae Markowski,
+Franco Milicchio,
+Adriana Morales,
+Peter Munch,
+Bob Myhill,
+Emily Novak,
+Omotayo Omosebi,
+Alexey Ozeritskiy,
+Jean-Paul Pelteret,
+Rebecca Pereira,
+Jan Philipp Thiele,
+Geneva Porter,
+Laura Prieto Saavedra,
 Reza Rastak,
 Roland Richter,
-  Alberto Sartori,
-  Svenja Schoeder,
+Jonathan Robey,
+Irabiel Romero,
+Matthew Russell,
+Tonatiuh Sanchez-Vizuet,
+Doug Shi-Dong,
+Natasha S. Sharma,
+Konrad Simon,
+Stephanie Sparks,
 Sebastian Stark,
-  Antoni Vidal,
-  Jiaxin Wang,
-  Yuxiang Wang,
-  Zhuoran Wang.
+Simon Sticko,
+Jihuan Tian,
+Ignacio Tomas,
+Sara Tro,
+Bruno Turcksin,
+Ferdinand Vanmaele,
+Zhuoran Wang,
+David Wells,
+Michal Wichrowski,
+Julius Witte,
+Winnifried Wollner,
+Mario Zepeda Aguilar,
+Wenjuan Zhang.
 
@@
 
-M.~Maier was partially supported by ARO MURI Award No. W911NF-14-0247.
+M.~Maier was partially supported by ARO MURI Award No. W911NF-14-0247 and
+NSF Award DMS-1912847.
 

### 2020-05-12 9.2/paper.tex
complete contributors

@@
 Benjamin Brands,
+Yong-Yong Cai,
+Fabian Castelli,
 Thomas C. Clevenger,
@@
 Brandon Gleeson,
+Anne Glerum,
 Krishnakumar Gopalakrishnan,
@@
 Graham Harper,
+Mohammed Hassan,
 Nicole Hayes,
 Bang He,
+Johannes Heinz,
 Timo Heister,
@@
 Marie Kajan,
+Guido Kanschat,
 Alexander Knieps,
 Martin Kronbichler,
+Paras Kumar,
 Konstantin Ladutenko,
@@
 Franco Milicchio,
-Adriana Morales,
+Adriana Morales Miranda,
 Peter Munch,
@@
 Rebecca Pereira,
-Jan Philipp Thiele,
 Geneva Porter,
@@
 Tonatiuh Sanchez-Vizuet,
+Natasha S. Sharma,
 Doug Shi-Dong,
-Natasha S. Sharma,
 Konrad Simon,
@@
 Simon Sticko,
+Jan Philipp Thiele,
 Jihuan Tian,
@@
 Winnifried Wollner,
+Ming Yang,
 Mario Zepeda Aguilar,
-Wenjuan Zhang.
+Wenjuan Zhang,
+Victor Zheng.
 

### 2020-05-13 make_offline_doc.sh
update release tasks

@@ -1,33 +1,0 @@
-#!/bin/bash
-
-if [ "$#" -ne 3 ]; then
-    echo "Offline documentation generator for deal.II"
-    echo "Usage:"
-    echo "  ./make_offline_doc.sh <major> <minor> <patch>"
-    echo "Execute in the base directory of a clean checkout for the release."
-    echo "For an RC use: './make_offline_doc.sh 8 3 0-rc1'"
-    exit 1
-fi
-
-MAJOR=$1
-MINOR=$2
-PATCH=$3
-
-echo release $MAJOR.$MINOR.$PATCH
-
-# Generate documentation
-rm -rf builddoc
-mkdir builddoc
-cd builddoc
-cmake -DDEAL_II_WITH_MPI=OFF -DCMAKE_BUILD_TYPE=Debug -DDEAL_II_COMPONENT_DOCUMENTATION=ON -DCMAKE_INSTALL_PREFIX=`pwd`/../installeddoc/ ../
-make documentation -j 8
-
-cd ../installeddoc
-
-cd doc/doxygen/deal.II
-../../../../contrib/utilities/makeofflinedoc.sh
-
-cd ../../..
-tar czf ../dealii-$MAJOR.$MINOR.$PATCH-offline_documentation.tar.gz doc
-cd ..
-

### 2020-05-13 9.2/paper.tex
update
Make a bit of progress on the release paper.

@@
 \usepackage[normalem]{ulem}
+
+\usepackage{todonotes}
 
@@
 %     Ross Maguire Kynch,
-%     Matthias Maier,
+    Matthias Maier,
 %     Jean-Paul Pelteret,
@@
 \begin{itemize}
-\item
+\item xy \todo[inline]{Update once we have the subsections in Section 2;
+  provide cross-references to each of these subsections}
 %   \item Improved support for automatic differentiation (see
 %     Section~\ref{subsec:ad}), 
-%   \item Dedicated support for symbolic algebra (see
-%     Section~\ref{subsec:sd}),
-%   \item Full support for $hp$~adaptivity in parallel computations (see
-%     Section~\ref{subsec:hp}),
-%   \item An interface to the HDF5 file format and libraries (see
-%     Section~\ref{subsec:hdf5}),
-%   \item Significantly extended GPU support (see Section~\ref{subsec:gpu}),
-%   \item Parallel geometric multigrid (GMG) improvements (see
-%     \cite{ClevengerHeisterKanschatKronbichler2019} and
-%     Section~\ref{subsec:gmg}),
-%   \item Four new tutorial programs (step-61, step-62, step-63, step-64),
-%     as well as one new code gallery program (see
-%     Section~\ref{subsec:steps}).
 \end{itemize}
@@
 \begin{itemize}
-\item
-% \item
-%   The release contains a number of performance improvements and bug fixes for
-%   the matrix-free framework. One notable improvement is the support for
-%   renumbering of degrees of freedom within the cells for discontinuous
-%   elements, avoiding some reshuffling operations across the SIMD lanes
-%   with vectorization over several cells and faces, which is especially
-%   useful on processors with AVX-512 vectorization (8 doubles), speeding up
-%   operations by up to 10\%. Secondly, the strategy for the most efficient
-%   tensor product evaluators according to the performance analysis of
-%   \cite{KronbichlerKormann2019} in the context of more quadrature points than
-%   shape functions has been revised for better performance.
-%
-% \item A new class \texttt{ParsedConvergenceTable} has been introduced
-%   that greatly simplifies the construction of convergence tables,
-%   reading the options for the generation of the table from a parameter
-%   file, and providing methods that, combined with a parameter file,
-%   allow one to generate convergence tables using one-liners in user
-%   codes.
-%
-% \item
-%   The \texttt{FE\_BernardiRaugel} class implements the non-standard
-%   Bernardi-Raugel (BR) element that can be used to construct a stable
-%   velocity-pressure pair for the Stokes equation \cite{BR85}. The BR
-%   element is an enriched version of the $Q_1^d$ element with added bubble
-%   functions on each edge (in 2d) or face (in 3d). It addresses the fact
-%   that the $Q_1^d\times Q_0$ combination is not inf-sup stable (requiring a
-%   larger velocity space), and that the $Q_2^d\times Q_0$ combination is
-%   stable but converges with only first-order at the cost of the large
-%   number of velocity unknowns. The BR space is thus intermediate between the
-%   $Q_1^d$ and $Q_2^d$ spaces.
-%
-%   The element is currently only implemented for parallelogram meshes due to
-%   difficulties associated with the mapping of shape functions: The shape
-%   functions of the $Q_1^d$ part of the element need to be mapped as
-%   scalars, as is common for the vector components of the $Q_1^d$ element;
-%   on the other hand, the vector-valued edge bubble functions need to be
-%   mapped using the Piola transform as is common for the Raviart-Thomas
-%   element. \dealii{} does not currently have the ability to use different
-%   mappings for individual shape functions, though this functionality is
-%   planned for the next release.
-%
-% \item
-%   The \texttt{FE\_NedelecSZ} class is a new implementation of the
-%   N{\'e}d{\'e}lec element on quadrilaterals and hexahedra. It is based on
-%   the work of Zaglmayr \cite{Zag06} and overcomes the sign conflict issues
-%   present in traditional N{\'e}d{\'e}lec elements that arise from the edge
-%   and face parameterizations used in the basis functions. Therefore, this
-%   element should provide consistent results for general quadrilateral and
-%   hexahedral elements for which the relative orientations of edges and
-%   faces (as seen from all adjacent cells) are often difficult to establish.
-%   The \texttt{FE\_NedelecSZ} element addresses the sign conflict problem by
-%   assigning a globally defined orientation to local edges and faces. A
-%   detailed overview of the implementation of the \texttt{FE\_NedelecSZ}
-%   element in \dealii{} can be found in \cite{Kynch2017}.
-%
-% \item All of the elementary geometrical objects of the library (namely
-%   \texttt{Point<dim>}, \texttt{Segment<dim>}, and
-%   \texttt{BoundingBox<dim>}) have been augmented with the traits
-%   needed to comply with \texttt{boost::geometry} concepts. A new
-%   interface to \texttt{boost::geometry::index::rtree} has been added
-%   that simplifies the construction of spatial indices based on points,
-%   bounding boxes, or segments.
+\item x \todo[inline]{Wolfgang to write about complex-valued output}
+\item y \todo[inline]{Timo to write about problems and fixes for
+  computations with more than $2^{32}$ unknowns}
+\item z \todo[inline]{What else?}
 \end{itemize}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{bla1}
-\label{subsec:bla1}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{New and improved tutorial and code gallery programs}
@@
 \item \texttt{step-47}
+\todo[inline]{Zhuoran to write}
 \item \texttt{step-50}
+\todo[inline]{Timo/Conrad/... to write}
 \item \texttt{step-58}
+\todo[inline]{Wolfgang to write}  
 \item \texttt{step-65}
+\todo[inline]{Martin to write}  
 \item \texttt{step-67}
+\todo[inline]{Martin to write}  
 \item \texttt{step-69}
+\todo[inline]{Matthias/Ignacio to write}  
 \item \texttt{step-70}
+  \todo[inline]{Also need to update announce and announce-short if this
+    makes it into the release.}
 \end{itemize}
+
+\todo[inline]{Do we have new code gallery programs}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Support for large, fully distributed meshes}
+\label{subsec:pfT}
+
+\todo[inline]{Peter: Write something about p::f::T}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Better support for parallel $hp$-adaptive algorithms}
+\label{subsec:hp}
+
+\todo[inline]{Marc: Your section}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Support for particle-based methods}
+\label{subsec:particles}
+
+\todo[inline]{Luca: Your section}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Python interfaces}
+\label{subsec:python}
+
+\todo[inline]{What's new here? mention step-49 and step-53 versions written in python.}
 
@@
 \href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_1_1_and_9_2_0.html}
-     {around 15 incompatible changes}; see \cite{changes92}. The majority of these changes
+     {around 60 incompatible changes}; see \cite{changes92}. The majority of these changes
 should not be visible to typical user codes; some remove previously
@@
 \item
-% \item The \texttt{VectorView} class was removed. We recommend either copying the
-%       vector subset into a \texttt{Vector} or using a \texttt{BlockVector}.
-% \item The function \texttt{Subscriptor::subscribe()}, used through the
-%   \texttt{SmartPointer} class, now requires a pointer to a
-%       \texttt{std::atomic<bool>} that tracks whether or not the pointer to the
-%       subscribed-to object is still valid.
-% \item The \texttt{ConstraintMatrix} class gained a template parameter for the scalar
-%       type and was been renamed \texttt{AffineConstraints}. Several methods that
-%       take vectors or matrices as arguments,
-%       such as \texttt{AffineConstraints::distribute\_local\_to\_global()},
-%       now require that all matrix and vector arguments have matching number
-%       types.
-% \item Similarly, the functions \texttt{create\_mass\_matrix} and
-%       \texttt{create\_boundary\_mass\_matrix} in the \texttt{MatrixCreator}
-%       namespace no longer
-%       support matrix and vector objects of different types.
 \end{itemize}
@@
 contributed code to this release:\\
-% updated 5/11/2020 MM - TODO: remove authors of paper
+% updated 5/11/2020 MM
+\todo[inline]{TODO: remove authors of paper}
 Pasquale Africa,

### 2020-05-13 9.2/paper.tex
Add a couple of comments.
Update a comment.

@@
   computations with more than $2^{32}$ unknowns}
-\item z \todo[inline]{What else?}
+\item z \todo[inline]{What else? Maybe mention the updated step-12?}
 \end{itemize}
@@
 
-Many of the \dealii{} tutorial programs were substantially revised as
-part of this release. In particular, we have converted many places
-that now allow for simpler code through the use of C++11 features such
-as range-based for loops and lambda functions.
+Many of the \dealii{} tutorial programs were revised in a variety of
+ways as part of this release. A particular example is that we have
+converted a number of programs to use range-based for loops (a C++11
+feature) for loops over a range of integer indices, given that the
+range-based way of writing loops seems to be the idiomatic approach
+these days.
 
-In addition, there are seven new tutorial programs:
+In addition, there are a number of new tutorial programs:
 \begin{itemize}

### 2020-05-13 9.2/paper.tex
Expand on a comment.

@@
 converted a number of programs to use range-based for loops (a C++11
-feature) for loops over a range of integer indices, given that the
+feature) for loops over a range of integer indices such as loops over
+all quadrature points or all indices of degrees of freedom during
+assembly. This makes sense given that the
 range-based way of writing loops seems to be the idiomatic approach
-these days.
+these days, and that we had previously already converted loops over
+all cells in this way.
 

### 2020-05-14 9.2/paper.tex

--- 2020-05-13 9.2/paper.tex
@@
 
+%\renewcommand{\baselinestretch}{2.0}
+
 \usepackage[normalem]{ulem}
-
-\usepackage{todonotes}
 
@@
 \begin{itemize}
-\item xy \todo[inline]{Update once we have the subsections in Section 2;
-  provide cross-references to each of these subsections}
+\item
 %   \item Improved support for automatic differentiation (see
 %     Section~\ref{subsec:ad}),
+%   \item Dedicated support for symbolic algebra (see
+%     Section~\ref{subsec:sd}),
+%   \item Full support for $hp$~adaptivity in parallel computations (see
+%     Section~\ref{subsec:hp}),
+%   \item An interface to the HDF5 file format and libraries (see
+%     Section~\ref{subsec:hdf5}),
+%   \item Significantly extended GPU support (see Section~\ref{subsec:gpu}),
+%   \item Parallel geometric multigrid (GMG) improvements (see
+%     \cite{ClevengerHeisterKanschatKronbichler2019} and
+%     Section~\ref{subsec:gmg}),
+%   \item Four new tutorial programs (step-61, step-62, step-63, step-64),
+%     as well as one new code gallery program (see
+%     Section~\ref{subsec:steps}).
 \end{itemize}
@@
 \begin{itemize}
-\item x \todo[inline]{Wolfgang to write about complex-valued output}
-\item y \todo[inline]{Timo to write about problems and fixes for
-  computations with more than $2^{32}$ unknowns}
-\item z \todo[inline]{What else?}
+\item
+% \item
+%   The release contains a number of performance improvements and bug fixes for
+%   the matrix-free framework. One notable improvement is the support for
+%   renumbering of degrees of freedom within the cells for discontinuous
+%   elements, avoiding some reshuffling operations across the SIMD lanes
+%   with vectorization over several cells and faces, which is especially
+%   useful on processors with AVX-512 vectorization (8 doubles), speeding up
+%   operations by up to 10\%. Secondly, the strategy for the most efficient
+%   tensor product evaluators according to the performance analysis of
+%   \cite{KronbichlerKormann2019} in the context of more quadrature points than
+%   shape functions has been revised for better performance.
+%
+% \item A new class \texttt{ParsedConvergenceTable} has been introduced
+%   that greatly simplifies the construction of convergence tables,
+%   reading the options for the generation of the table from a parameter
+%   file, and providing methods that, combined with a parameter file,
+%   allow one to generate convergence tables using one-liners in user
+%   codes.
+%
+% \item
+%   The \texttt{FE\_BernardiRaugel} class implements the non-standard
+%   Bernardi-Raugel (BR) element that can be used to construct a stable
+%   velocity-pressure pair for the Stokes equation \cite{BR85}. The BR
+%   element is an enriched version of the $Q_1^d$ element with added bubble
+%   functions on each edge (in 2d) or face (in 3d). It addresses the fact
+%   that the $Q_1^d\times Q_0$ combination is not inf-sup stable (requiring a
+%   larger velocity space), and that the $Q_2^d\times Q_0$ combination is
+%   stable but converges with only first-order at the cost of the large
+%   number of velocity unknowns. The BR space is thus intermediate between the
+%   $Q_1^d$ and $Q_2^d$ spaces.
+%
+%   The element is currently only implemented for parallelogram meshes due to
+%   difficulties associated with the mapping of shape functions: The shape
+%   functions of the $Q_1^d$ part of the element need to be mapped as
+%   scalars, as is common for the vector components of the $Q_1^d$ element;
+%   on the other hand, the vector-valued edge bubble functions need to be
+%   mapped using the Piola transform as is common for the Raviart-Thomas
+%   element. \dealii{} does not currently have the ability to use different
+%   mappings for individual shape functions, though this functionality is
+%   planned for the next release.
+%
+% \item
+%   The \texttt{FE\_NedelecSZ} class is a new implementation of the
+%   N{\'e}d{\'e}lec element on quadrilaterals and hexahedra. It is based on
+%   the work of Zaglmayr \cite{Zag06} and overcomes the sign conflict issues
+%   present in traditional N{\'e}d{\'e}lec elements that arise from the edge
+%   and face parameterizations used in the basis functions. Therefore, this
+%   element should provide consistent results for general quadrilateral and
+%   hexahedral elements for which the relative orientations of edges and
+%   faces (as seen from all adjacent cells) are often difficult to establish.
+%   The \texttt{FE\_NedelecSZ} element addresses the sign conflict problem by
+%   assigning a globally defined orientation to local edges and faces. A
+%   detailed overview of the implementation of the \texttt{FE\_NedelecSZ}
+%   element in \dealii{} can be found in \cite{Kynch2017}.
+%
+% \item All of the elementary geometrical objects of the library (namely
+%   \texttt{Point<dim>}, \texttt{Segment<dim>}, and
+%   \texttt{BoundingBox<dim>}) have been augmented with the traits
+%   needed to comply with \texttt{boost::geometry} concepts. A new
+%   interface to \texttt{boost::geometry::index::rtree} has been added
+%   that simplifies the construction of spatial indices based on points,
+%   bounding boxes, or segments.
 \end{itemize}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{bla1}
+\label{subsec:bla1}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Improved large-scale performance}
+\label{subsec:performance}
+
+Large-scale simulations with 304,128 cores have revealed bottlenecks in release 
+9.1 during initialization due to the usage of expensive collective operations 
+like \texttt{MPI\_Allgather()} and \texttt{MPI\_\allowbreak Alltoall()}, e.g., during the 
+pre-computation of the index ranges of all processes, which have been stored in  an array. 
+This information is needed to set up 
+the  \texttt{Utilities::MPI::Par\-ti\-ti\-oner} class. 
+In release 9.2, we have removed such arrays and have replaced 
+the \texttt{MPI\_Allgather}/\allowbreak\texttt{MPI\_\allowbreak Alltoall} 
+function calls by consensus algorithms~\cite{hoefler2010scalable}, which can be 
+found in the namespace \texttt{Utilities::\allowbreak MPI::\allowbreak ConsensusAlgorithms} (short: \texttt{CA}): now, only the locally relevant information about the index ranges is (re)computed when needed, using these algorithms.
+We provide two flavors of the consensus algorithm: the two-step 
+approach \texttt{CA::PEX} and the \texttt{CA::NBX}, 
+which uses only point-to-point communications and a single \texttt{MPI\_IBarrier()}.
+
+%Consensus algorithms are algorithms dedicated to efficient dynamic-sparse 
+%communication patterns. In this context, the term ``dynamic-sparse'' means 
+%that by the time this function is called, the other processes do not know 
+%yet that they have to answer requests and
+%each process only has to communicate with a small subset of processes of the 
+%MPI communicator. We provide two flavors of the consensus algorithm: the two-step 
+%approach \texttt{ConsensusAlgorithms::PEX} and the \texttt{ConsensusAlgorithms::NBX}, 
+%which uses only point-to-point communications and a single \texttt{MPI\_IBarrier()}. 
+%The class \texttt{ConsensusAlgorithms::Selector} selects one of the two previous 
+%algorithms, depending on the number of processes.
+
+Users can apply the new algorithms for their own dynamic-sparse problems by 
+providing a list of target 
+processes and pack/unpack routines either by implementing the interface 
+\texttt{CA::\allowbreak Process} or by providing \texttt{std::function} 
+objects to \texttt{CA::AnonymousProcess}.
+
+By replacing the collective communications during setup and removing the arrays 
+that contain information for each process (enabled by the application of consensus 
+algorithms and other modifications---a full list of modifications leading to this 
+improvement can be found online), we were able to significantly improve the setup 
+time for large-scale simulations and to solve a Poisson problem with multigrid 
+with 12T unknowns. 
+Figure~\ref{} compares the timings of a simulation (incl. setup) with the 
+previous release 9.1 and with the current release 9.2. 
+{\color{red}TODO[Peter/Martin] description of the results}
+
+The new code has been also applied to solve problems with adaptively refined 
+meshes with more than 4B unknowns. {\color{red}TODO[Timo]}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{A new fully distributed triangulation class}
+\label{subsec:pft}
+
+By release 9.1, all triangulation classes of \texttt{deal.II} have in common that the coarse grid is shared by 
+all processes and the actual mesh used for computations is constructed by repeated 
+refinement. However, this has its limitations in industrial applications where, often, the mesh comes 
+from an external mesh generator in the form of a file that already contains millions 
+or tens of millions of cells. For such configurations, applications might already 
+run out of  main memory, while reading the complete mesh by each MPI process.
+
+The new class \texttt{parallel::fullydistributed::Triangulation} targets this issue 
+by distributing also the coarse grid. Such 
+a triangulation can be created by providing to each process a \texttt{Triangulation\-De\-scrip\-tion::Description} struct, containing 
+1) the relevant data to construct the local part of the coarse grid, 2) the 
+translation of the local coarse-cell IDs to globally unique IDs, 3) the hierarchy 
+of mesh refinements, and 4) the owner of the cells on the active mesh level as well 
+as on the multigrid levels. Once the triangulation is set up with this struct, no adaptive
+changes to the mesh are allowed at the moment.
+
+%The \texttt{TriangulationDescription::Description} struct can be filled manually or 
+%by the utility functions from the \texttt{TriangulationDescription::Utilities} 
+%namespace. The function \texttt{create\_\allowbreak description\_\allowbreak 
+%from\_ \allowbreak triangulation()} can convert a base triangulation (partitioned 
+%serial \texttt{Tri\-angulation} and \texttt{parallel::distributed::Triangulation}) 
+%to such a struct. The advantage of this approach is that all known utility 
+%functions from the namespaces \texttt{GridIn} and \texttt{GridTools} can be used 
+%on the base triangulations before converting them to the structs. Since this 
+%function suffers from the same main memory problems as described above, we also 
+%provide the function \texttt{create\_description\_from\_triangulation\_in\_groups()}, 
+%which creates the structs only on the master process in a process group. These 
+%structs are filled one by one and are sent to the relevant processes once they are 
+%ready. A sensible process group size might contain all processes of one compute node.
+
+The new fully distributed triangulation class works---in contrast to  
+\texttt{parallel::distributed::\allowbreak Tri\-an\-gu\-la\-tion}---not only for 2D- and 3D- but also for 
+1D-problems. It can be used in the context of geometric multigrid methods and 
+supports periodic boundary conditions. Furthermore, hanging nodes are supported.
+
+%We intend to extend the usability of the new triangulation class in regard of 
+%different aspects, e.g., I/O. In addition, we would like to enable adaptive mesh 
+%refinement, a feature of the other triangulation classes, which is very much 
+%appreciated by many users. For repartitioning, we plan to use an oracle approach 
+%known from \texttt{parallel::distributed::Triangulation}. Here, we would like to 
+%rely on a user-provided partitioner, which might also be a graph partitioner.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Advances of the SIMD capabilities and the matrix-free infrastructure}
+\label{subsec:mf}
+
+%\begin{itemize}
+%\item ECL
+%\item VectorizedArrayType
+%\end{itemize}
+
+The class \texttt{VectorizedArray<Number>} is a key ingredient for the high 
+node-level performance of the matrix-free algorithms in deal.II~\cite{KronbichlerKormann2012, KronbichlerKormann2019}. It is a wrapper 
+class around $n$ vector entries of type \texttt{Number} and delegates relevant 
+function calls to appropriate Intrinsics instructions. Up to release 9.1, the 
+vector length $n$ has been set at compile time of the library to the highest 
+possible value supported by the given processor architecture.
+
+The class \texttt{VectorizedArray} has been made more user-friendly by making 
+it compatible with the STL algorithms found in the header \texttt{<algorithm>}. 
+The length of the  vector can now be queried by  \texttt{VectorizedArray::size()} and its underlying number type by \texttt{VectorizedArray::value\_type}. 
+Furthermore, the \texttt{VectorizedArray} class supports range-based iterations over its entries and, i.a., the following 
+algorithms: \texttt{std::\allowbreak ad\-vance()}, \texttt{std::distance()}, and \texttt{std::max\_element()}.
+
+This class has been also extended with the second optional template argument 
+\texttt{VectorizedArray<Number, size>} with \texttt{size} being related to the 
+vector length, i.e., the number of lanes to be used and the instruction set to be 
+used. By default, the number is set to the highest value supported by the given 
+hardware. A full list of supported 
+vector lengths is presented in Table~\ref{tab:vectorizedarray}.
+
+All matrix-free related classes (like \texttt{MatrixFree} and \texttt{FEEvaluation}) 
+have been templated with the floating-point number type \texttt{Number} (e.g. \texttt{double} or \texttt{float}); the computations were performed implicitly 
+on \texttt{VectorizedArray<Number>} structs with the highest 
+instruction-set-architecture extension, with each lane responsible for a separate 
+cell (vectoriziation over elements). In release 9.2, all matrix-free classes 
+have been extended with a new optional template argument specifying the 
+\texttt{VectorizedArrayType}. This allows users to select the vector length/ISA and, 
+as a consequence, the number of cells to be processed at once directly in their applications: 
+The deal.II-based 
+library \texttt{hyper.deal}~\cite{munch2020hyperdeal}, which solves the 6D Vlasov--Poisson equation with high-order 
+discontinuous Galerkin methods (with more than 1024 degrees of freedom per cell), works with \texttt{MatrixFree} objects of different SIMD-vector 
+length in the same application and benefits---in terms of performance---by the possibility of decreasing the number of cells processed by a single SIMD instruction.
+
+\begin{table}
+\caption{Supported vector lengths of the class \texttt{VectorizedArray} and 
+the corresponding instruction-set-architecture extensions. }\label{tab:vectorizedarray}
+\centering
+\begin{tabular}{ccc}
+\toprule
+\textbf{double} & \textbf{float} & \textbf{ISA}\\
+\midrule
+VectorizedArray<double, 1> & VectorizedArray<float, 1> & (auto-vectorization) \\
+VectorizedArray<double, 2> & VectorizedArray<float, 4> & SSE2 \\ 
+VectorizedArray<double, 4> & VectorizedArray<float, 8> & AVX/AVX2 \\ 
+VectorizedArray<double, 8> & VectorizedArray<float, 16> & AVX-512 \\ 
+\bottomrule
+\end{tabular}
+
+\caption{Comparison of relevant SIMD-related classes in deal.II and C++20.}\label{tab:simd}
+\centering
+\begin{tabular}{cc}
+\toprule
+\textbf{VectorizedArray (deal.II)} & \textbf{std::simd (C++20)} \\
+\midrule
+VectorizedArray<Number> & std::experimental::native\_simd<Number> \\
+VectorizedArray<Number, size> & std::experimental::fixed\_size\_simd<Number, size> \\ \bottomrule
+\end{tabular}
+\end{table}
+
+A side effect of introducing the new template argument \texttt{VectorizedArrayType} 
+in the \texttt{MatrixFree} classes is that any data structures 
+\texttt{VectorizedArrayType} can be processed if they support required 
+functionalities like \texttt{size()} or \texttt{value\_type}. In this context, we 
+would like to highlight that the new \texttt{C++20} feature \texttt{std::simd} 
+can be processed by the matrix-free infrastructure with minor internal 
+adjustment as an open pull request shows 
+(see \url{https://github.com/dealii/dealii/pull/9994}).  
+Table~\ref{tab:simd} gives a comparison of the deal.II-specific SIMD classes and 
+the equivalent C++20 classes. 
+%We welcome the standardization of the SIMD 
+%parallelization paradigm in C++ and intend to replace step by step our own 
+%wrapper class, which has been continuously developed over the last decade. We 
+%would like to emphasize that the work invested in this class was not in vain, 
+%since many performance-relevant utility functions implemented with \texttt{VectorizedArray} in mind (e.g., \texttt{vectorized\_load\_and\_transpose} 
+%and \texttt{vectorized\_transpose\_and\_store}) will be still used, since they 
+%have not become part of the standard.
+
+%Further additions to the \texttt{MatrixFree} infrastructure consist of:
+%\begin{itemize}
+%\item a new variant of \texttt{MatrixFree::cell\_loop()}: It takes two
+%\texttt{std::function} objects with ranges on the locally owned degrees of freedom, one
+%with work to be scheduled before the cell operation first touches some
+%unknowns and another with work to be executed after the cell operation last 
+%touches them. The goal of
+%these functions is to bring vector operations close to the time when the
+%vector entries are accessed by the cell operation, which increases the cache
+%hit rate of modern processors by improved temporal locality. 
+%\item a new form of loop \texttt{MatrixFree::loop\_cell\_centric()}: This 
+%kind of loop can be used in the context of discontinuous Galerkin methods, 
+%where both cell and face integrals have to be evaluated. While in the case of 
+%the traditional \texttt{loop}, cell and face integrals have been performed 
+%independently, the new loop performs all cell and face integrals of a cell in 
+%one go. This includes that each face integral has to be evaluated twice, but 
+%entries have to be written into the solution vector only once with improved 
+%data locality. Previous publications based on \texttt{deal.II} have shown the 
+%relevance of the latter aspect for reaching higher performance. 
+%\end{itemize}
+
+In a next step, we intend to support that users could set 
+\texttt{VectorizedArrayType} to \texttt{Number}, which would lead to the usage not 
+of \texttt{VectorizedArray<Number, 1>} but of a specialized code-path exploiting 
+vectorization within an element~\cite{KronbichlerKormann2019}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Advances in GPU support}
+\label{subsec:gpu}
+
+\begin{itemize}
+\item overlapping of computation and communication in the case of CUDA-aware MPI
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{New and improved tutorial and code gallery programs}
@@
 \item \texttt{step-47}
-\todo[inline]{Zhuoran to write}
 \item \texttt{step-50}
-\todo[inline]{Timo/Conrad/... to write}
 \item \texttt{step-58}
-\todo[inline]{Wolfgang to write}  
-\item \texttt{step-65}
-\todo[inline]{Martin to write}  
-\item \texttt{step-67}
-\todo[inline]{Martin to write}  
+\item \texttt{step-65} presents \texttt{TransfiniteInterpolationManifold}, a
+manifold class that can propagate curved boundary information into the
+interior of a computational domain, and \texttt{MappingQCache}, which can sample 
+the information of expensive manifolds in the points of a \texttt{MappingQ} and 
+cache it for further use.
+\item \texttt{step-67} presents an explicit time integrator for the
+compressible Euler equations discretized with a high-order discontinuous
+Galerkin scheme using the matrix-free infrastructure. Besides the use of
+matrix-free evaluators for systems of equations and over-integration, it also
+presents \texttt{MatrixFreeOperators::CellwiseInverseMassMatrix}, a fast implementation
+of the action of the inverse mass matrix in the DG setting using tensor
+products. Furthermore, this tutorial demonstrates the usage of new 
+pre and post operations, which can be passed to \texttt{cell\_loop()}, to schedule operations on sections of vectors close 
+to the matrix-vector product to increase data locality 
+and discusses performance-related aspects.
 \item \texttt{step-69}
-\todo[inline]{Matthias/Ignacio to write}  
 \item \texttt{step-70}
-  \todo[inline]{Also need to update announce and announce-short if this
-    makes it into the release.}
 \end{itemize}
-
-\todo[inline]{Do we have new code gallery programs}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Support for large, fully distributed meshes}
-\label{subsec:pfT}
-
-\todo[inline]{Peter: Write something about p::f::T}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Better support for parallel $hp$-adaptive algorithms}
-\label{subsec:hp}
-
-\todo[inline]{Marc: Your section}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Support for particle-based methods}
-\label{subsec:particles}
-
-\todo[inline]{Luca: Your section}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Python interfaces}
-\label{subsec:python}
-
-\todo[inline]{What's new here? mention step-49 and step-53 versions written in python.}
 
@@
 \href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_1_1_and_9_2_0.html}
-     {around 60 incompatible changes}; see \cite{changes92}. The majority of these changes
+     {around 15 incompatible changes}; see \cite{changes92}. The majority of these changes
 should not be visible to typical user codes; some remove previously
@@
 \begin{itemize}
+\item The functions:
+\begin{itemize}
+\item \texttt{DoFHandler::loccaly\_owned\_dofs\_per\_processor()}
+\item \texttt{DoFHandler::loccaly\_owned\_mg\_dofs\_per\_processor()}
+\end{itemize} have been deprecated.
+As discussed in Subsection~\ref{subsec:performance}, deal.II does not store information for all processes on all processes processes, but only the local information or the locally-relevant information.
+Users are asked to construct the global information on their own, e.g.
+by calling \texttt{Utilities::MPI::Allgather(locally\_owned\_info(), comm)}.
 \item
+
+% \item The \texttt{VectorView} class was removed. We recommend either copying the
+%       vector subset into a \texttt{Vector} or using a \texttt{BlockVector}.
+% \item The function \texttt{Subscriptor::subscribe()}, used through the
+%   \texttt{SmartPointer} class, now requires a pointer to a
+%       \texttt{std::atomic<bool>} that tracks whether or not the pointer to the
+%       subscribed-to object is still valid.
+% \item The \texttt{ConstraintMatrix} class gained a template parameter for the scalar
+%       type and was been renamed \texttt{AffineConstraints}. Several methods that
+%       take vectors or matrices as arguments,
+%       such as \texttt{AffineConstraints::distribute\_local\_to\_global()},
+%       now require that all matrix and vector arguments have matching number
+%       types.
+% \item Similarly, the functions \texttt{create\_mass\_matrix} and
+%       \texttt{create\_boundary\_mass\_matrix} in the \texttt{MatrixCreator}
+%       namespace no longer
+%       support matrix and vector objects of different types.
 \end{itemize}
@@
 contributed code to this release:\\
-% updated 5/11/2020 MM
-\todo[inline]{TODO: remove authors of paper}
+% updated 5/11/2020 MM - TODO: remove authors of paper
 Pasquale Africa,

### 2020-05-14 9.2/paper.tex

@@
 \begin{itemize}
-\item
-% \item
-%   The release contains a number of performance improvements and bug fixes for
-%   the matrix-free framework. One notable improvement is the support for
-%   renumbering of degrees of freedom within the cells for discontinuous
-%   elements, avoiding some reshuffling operations across the SIMD lanes
-%   with vectorization over several cells and faces, which is especially
-%   useful on processors with AVX-512 vectorization (8 doubles), speeding up
-%   operations by up to 10\%. Secondly, the strategy for the most efficient
-%   tensor product evaluators according to the performance analysis of
-%   \cite{KronbichlerKormann2019} in the context of more quadrature points than
-%   shape functions has been revised for better performance.
-%
-% \item A new class \texttt{ParsedConvergenceTable} has been introduced
-%   that greatly simplifies the construction of convergence tables,
-%   reading the options for the generation of the table from a parameter
-%   file, and providing methods that, combined with a parameter file,
-%   allow one to generate convergence tables using one-liners in user
-%   codes.
-%
-% \item
-%   The \texttt{FE\_BernardiRaugel} class implements the non-standard
-%   Bernardi-Raugel (BR) element that can be used to construct a stable
-%   velocity-pressure pair for the Stokes equation \cite{BR85}. The BR
-%   element is an enriched version of the $Q_1^d$ element with added bubble
-%   functions on each edge (in 2d) or face (in 3d). It addresses the fact
-%   that the $Q_1^d\times Q_0$ combination is not inf-sup stable (requiring a
-%   larger velocity space), and that the $Q_2^d\times Q_0$ combination is
-%   stable but converges with only first-order at the cost of the large
-%   number of velocity unknowns. The BR space is thus intermediate between the
-%   $Q_1^d$ and $Q_2^d$ spaces.
-%
-%   The element is currently only implemented for parallelogram meshes due to
-%   difficulties associated with the mapping of shape functions: The shape
-%   functions of the $Q_1^d$ part of the element need to be mapped as
-%   scalars, as is common for the vector components of the $Q_1^d$ element;
-%   on the other hand, the vector-valued edge bubble functions need to be
-%   mapped using the Piola transform as is common for the Raviart-Thomas
-%   element. \dealii{} does not currently have the ability to use different
-%   mappings for individual shape functions, though this functionality is
-%   planned for the next release.
-%
-% \item
-%   The \texttt{FE\_NedelecSZ} class is a new implementation of the
-%   N{\'e}d{\'e}lec element on quadrilaterals and hexahedra. It is based on
-%   the work of Zaglmayr \cite{Zag06} and overcomes the sign conflict issues
-%   present in traditional N{\'e}d{\'e}lec elements that arise from the edge
-%   and face parameterizations used in the basis functions. Therefore, this
-%   element should provide consistent results for general quadrilateral and
-%   hexahedral elements for which the relative orientations of edges and
-%   faces (as seen from all adjacent cells) are often difficult to establish.
-%   The \texttt{FE\_NedelecSZ} element addresses the sign conflict problem by
-%   assigning a globally defined orientation to local edges and faces. A
-%   detailed overview of the implementation of the \texttt{FE\_NedelecSZ}
-%   element in \dealii{} can be found in \cite{Kynch2017}.
-%
-% \item All of the elementary geometrical objects of the library (namely
-%   \texttt{Point<dim>}, \texttt{Segment<dim>}, and
-%   \texttt{BoundingBox<dim>}) have been augmented with the traits
-%   needed to comply with \texttt{boost::geometry} concepts. A new
-%   interface to \texttt{boost::geometry::index::rtree} has been added
-%   that simplifies the construction of spatial indices based on points,
-%   bounding boxes, or segments.
+\item x \todo[inline]{Wolfgang to write about complex-valued output}
+\item y \todo[inline]{Timo to write about problems and fixes for
+  computations with more than $2^{32}$ unknowns}
+\item z \todo[inline]{What else? Maybe mention the updated step-12?}
 \end{itemize}
@@
 in the file that lists all changes for this release}, see \cite{changes91}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{bla1}
-\label{subsec:bla1}
 
@@
 
-Many of the \dealii{} tutorial programs were substantially revised as
-part of this release. In particular, we have converted many places
-that now allow for simpler code through the use of C++11 features such
-as range-based for loops and lambda functions.
+Many of the \dealii{} tutorial programs were revised in a variety of
+ways as part of this release. A particular example is that we have
+converted a number of programs to use range-based for loops (a C++11
+feature) for loops over a range of integer indices such as loops over
+all quadrature points or all indices of degrees of freedom during
+assembly. This makes sense given that the
+range-based way of writing loops seems to be the idiomatic approach
+these days, and that we had previously already converted loops over
+all cells in this way.
 
-In addition, there are seven new tutorial programs:
+In addition, there are a number of new tutorial programs:
 \begin{itemize}
 \item \texttt{step-47}
+\todo[inline]{Zhuoran to write}
 \item \texttt{step-50}
+\todo[inline]{Timo/Conrad/... to write}
 \item \texttt{step-58}
+\todo[inline]{Wolfgang to write}  
 \item \texttt{step-65} presents \texttt{TransfiniteInterpolationManifold}, a
@@
 \item \texttt{step-69}
+\todo[inline]{Matthias/Ignacio to write}  
 \item \texttt{step-70}
+  \todo[inline]{Also need to update announce and announce-short if this
+    makes it into the release.}
 \end{itemize}
+
+\todo[inline]{Do we have new code gallery programs}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Support for large, fully distributed meshes}
+\label{subsec:pfT}
+
+\todo[inline]{Peter: Write something about p::f::T}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Better support for parallel $hp$-adaptive algorithms}
+\label{subsec:hp}
+
+\todo[inline]{Marc: Your section}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Support for particle-based methods}
+\label{subsec:particles}
+
+\todo[inline]{Luca: Your section}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Python interfaces}
+\label{subsec:python}
+
+\todo[inline]{What's new here? mention step-49 and step-53 versions written in python.}
 

### 2020-05-14 9.2/paper.tex
Fix typos.
Remove a section no longer necessary.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Support for large, fully distributed meshes}
-\label{subsec:pfT}
-
-\todo[inline]{Peter: Write something about p::f::T}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Better support for parallel $hp$-adaptive algorithms}
@@
 \begin{itemize}
-\item \texttt{DoFHandler::loccaly\_owned\_dofs\_per\_processor()}
-\item \texttt{DoFHandler::loccaly\_owned\_mg\_dofs\_per\_processor()}
+\item \texttt{DoFHandler::locally\_owned\_dofs\_per\_processor()}
+\item \texttt{DoFHandler::locally\_owned\_mg\_dofs\_per\_processor()}
 \end{itemize} have been deprecated.

### 2020-05-14 9.2/paper.tex
Re-shuffle section.
Briefly introduce step-58.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Improved large-scale performance}
-\label{subsec:performance}
-
-Large-scale simulations with 304,128 cores have revealed bottlenecks in release 
-9.1 during initialization due to the usage of expensive collective operations 
-like \texttt{MPI\_Allgather()} and \texttt{MPI\_\allowbreak Alltoall()}, e.g., during the 
-pre-computation of the index ranges of all processes, which have been stored in  an array. 
-This information is needed to set up 
-the  \texttt{Utilities::MPI::Par\-ti\-ti\-oner} class. 
-In release 9.2, we have removed such arrays and have replaced 
-the \texttt{MPI\_Allgather}/\allowbreak\texttt{MPI\_\allowbreak Alltoall} 
-function calls by consensus algorithms~\cite{hoefler2010scalable}, which can be 
-found in the namespace \texttt{Utilities::\allowbreak MPI::\allowbreak ConsensusAlgorithms} (short: \texttt{CA}): now, only the locally relevant information about the index ranges is (re)computed when needed, using these algorithms.
-We provide two flavors of the consensus algorithm: the two-step 
-approach \texttt{CA::PEX} and the \texttt{CA::NBX}, 
-which uses only point-to-point communications and a single \texttt{MPI\_IBarrier()}.
-
-%Consensus algorithms are algorithms dedicated to efficient dynamic-sparse 
-%communication patterns. In this context, the term ``dynamic-sparse'' means 
-%that by the time this function is called, the other processes do not know 
-%yet that they have to answer requests and
-%each process only has to communicate with a small subset of processes of the 
-%MPI communicator. We provide two flavors of the consensus algorithm: the two-step 
-%approach \texttt{ConsensusAlgorithms::PEX} and the \texttt{ConsensusAlgorithms::NBX}, 
-%which uses only point-to-point communications and a single \texttt{MPI\_IBarrier()}. 
-%The class \texttt{ConsensusAlgorithms::Selector} selects one of the two previous 
-%algorithms, depending on the number of processes.
-
-Users can apply the new algorithms for their own dynamic-sparse problems by 
-providing a list of target 
-processes and pack/unpack routines either by implementing the interface 
-\texttt{CA::\allowbreak Process} or by providing \texttt{std::function} 
-objects to \texttt{CA::AnonymousProcess}.
-
-By replacing the collective communications during setup and removing the arrays 
-that contain information for each process (enabled by the application of consensus 
-algorithms and other modifications---a full list of modifications leading to this 
-improvement can be found online), we were able to significantly improve the setup 
-time for large-scale simulations and to solve a Poisson problem with multigrid 
-with 12T unknowns. 
-Figure~\ref{} compares the timings of a simulation (incl. setup) with the 
-previous release 9.1 and with the current release 9.2. 
-{\color{red}TODO[Peter/Martin] description of the results}
-
-The new code has been also applied to solve problems with adaptively refined 
-meshes with more than 4B unknowns. {\color{red}TODO[Timo]}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{A new fully distributed triangulation class}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Improved large-scale performance}
+\label{subsec:performance}
+
+Large-scale simulations with 304,128 cores have revealed bottlenecks in release 
+9.1 during initialization due to the usage of expensive collective operations 
+like \texttt{MPI\_Allgather()} and \texttt{MPI\_\allowbreak Alltoall()}, e.g., during the 
+pre-computation of the index ranges of all processes, which have been stored in  an array. 
+This information is needed to set up 
+the  \texttt{Utilities::MPI::Par\-ti\-ti\-oner} class. 
+In release 9.2, we have removed such arrays and have replaced 
+the \texttt{MPI\_Allgather}/\allowbreak\texttt{MPI\_\allowbreak Alltoall} 
+function calls by consensus algorithms~\cite{hoefler2010scalable}, which can be 
+found in the namespace \texttt{Utilities::\allowbreak MPI::\allowbreak ConsensusAlgorithms} (short: \texttt{CA}): now, only the locally relevant information about the index ranges is (re)computed when needed, using these algorithms.
+We provide two flavors of the consensus algorithm: the two-step 
+approach \texttt{CA::PEX} and the \texttt{CA::NBX}, 
+which uses only point-to-point communications and a single \texttt{MPI\_IBarrier()}.
+
+%Consensus algorithms are algorithms dedicated to efficient dynamic-sparse 
+%communication patterns. In this context, the term ``dynamic-sparse'' means 
+%that by the time this function is called, the other processes do not know 
+%yet that they have to answer requests and
+%each process only has to communicate with a small subset of processes of the 
+%MPI communicator. We provide two flavors of the consensus algorithm: the two-step 
+%approach \texttt{ConsensusAlgorithms::PEX} and the \texttt{ConsensusAlgorithms::NBX}, 
+%which uses only point-to-point communications and a single \texttt{MPI\_IBarrier()}. 
+%The class \texttt{ConsensusAlgorithms::Selector} selects one of the two previous 
+%algorithms, depending on the number of processes.
+
+Users can apply the new algorithms for their own dynamic-sparse problems by 
+providing a list of target 
+processes and pack/unpack routines either by implementing the interface 
+\texttt{CA::\allowbreak Process} or by providing \texttt{std::function} 
+objects to \texttt{CA::AnonymousProcess}.
+
+By replacing the collective communications during setup and removing the arrays 
+that contain information for each process (enabled by the application of consensus 
+algorithms and other modifications---a full list of modifications leading to this 
+improvement can be found online), we were able to significantly improve the setup 
+time for large-scale simulations and to solve a Poisson problem with multigrid 
+with 12T unknowns. 
+Figure~\ref{} compares the timings of a simulation (incl. setup) with the 
+previous release 9.1 and with the current release 9.2. 
+{\color{red}TODO[Peter/Martin] description of the results}
+
+The new code has been also applied to solve problems with adaptively refined 
+meshes with more than 4B unknowns. {\color{red}TODO[Timo]}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Better support for parallel $hp$-adaptive algorithms}
+\label{subsec:hp}
+
+\todo[inline]{Marc: Your section}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Support for particle-based methods}
+\label{subsec:particles}
+
+\todo[inline]{Luca: Your section}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{New and improved tutorial and code gallery programs}
@@
   \todo[inline]{Timo/Conrad/... to write}
-\item \texttt{step-58}
-\todo[inline]{Wolfgang to write}  
+
+\item \texttt{step-58} is a program that solves the nonlinear
+  Schr{\"o}dinger equation, which in non-dimensional form reads
+  \begin{align*}
+  - i \frac{\partial \psi}{\partial t}
+  - \frac 12 \Delta \psi
+  + V \psi
+  + \kappa |\psi|^2 \psi
+  &= 0,
+  \end{align*}
+  augmented by appropriate initial and boundary conditions and using
+  an appropriate form for the potential $V=V(\mathbf x)$. The
+  tutorial program focused on two specific aspects for which this
+  equation serves as an excellent test case: (i) Solving
+  complex-valued problems without splitting the equation into its
+  real and imaginary parts (as \texttt{step-29} does, for
+  example). (ii) Using operator splitting techniques. The equation is
+  a particularly good test case for this technique because the only
+  nonlinear term, $\kappa |\psi|^2 \psi$, does not contain any
+  derivatives and consequently forms an ODE to be solved at each time
+  step in an operator splitting scheme (for which, furthermore, there
+  exists an analytic solution), whereas the remainder of the
+  equation is linear and easily solved using standard finite element
+  techniques.
+
 \item \texttt{step-65} presents \texttt{TransfiniteInterpolationManifold}, a
@@
 
-\todo[inline]{Do we have new code gallery programs}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Better support for parallel $hp$-adaptive algorithms}
-\label{subsec:hp}
-
-\todo[inline]{Marc: Your section}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Support for particle-based methods}
-\label{subsec:particles}
-
-\todo[inline]{Luca: Your section}
+\todo[inline]{Do we have new code gallery programs?}
 

### 2020-05-14 9.2/paper.tex
Minor edits.

@@
 cache it for further use.
-\item \texttt{step-67} presents an explicit time integrator for the
+
+\item \texttt{step-67} is an explicit time integrator for the
 compressible Euler equations discretized with a high-order discontinuous
@@
 and discusses performance-related aspects.
+
 \item \texttt{step-69}
   \todo[inline]{Matthias/Ignacio to write}
+
 \item \texttt{step-70}
@@
 \begin{itemize}
-\item The functions:
+\item Two functions that provide information about all processes,
+  namely
   \begin{itemize}
@@
  \item \texttt{DoFHandler::locally\_owned\_mg\_dofs\_per\_processor()}
-\end{itemize} have been deprecated.
-As discussed in Subsection~\ref{subsec:performance}, deal.II does not store information for all processes on all processes processes, but only the local information or the locally-relevant information.
-Users are asked to construct the global information on their own, e.g.
-by calling \texttt{Utilities::MPI::Allgather(locally\_owned\_info(), comm)}.
+  \end{itemize}
+have been deprecated. As discussed in
+Subsection~\ref{subsec:performance}, deal.II by default no longer
+stores information for all processes on all processes, but only the
+local information or the locally-relevant information. On the other
+hand, if necessary, global information can still be computed using,
+for example, calling \texttt{Utilities::MPI::Allgather(locally\_owned\_info(), comm)}.
 \item
-
-% \item The \texttt{VectorView} class was removed. We recommend either copying the
-%       vector subset into a \texttt{Vector} or using a \texttt{BlockVector}.
-% \item The function \texttt{Subscriptor::subscribe()}, used through the
-%   \texttt{SmartPointer} class, now requires a pointer to a
-%       \texttt{std::atomic<bool>} that tracks whether or not the pointer to the
-%       subscribed-to object is still valid.
-% \item The \texttt{ConstraintMatrix} class gained a template parameter for the scalar
-%       type and was been renamed \texttt{AffineConstraints}. Several methods that
-%       take vectors or matrices as arguments,
-%       such as \texttt{AffineConstraints::distribute\_local\_to\_global()},
-%       now require that all matrix and vector arguments have matching number
-%       types.
-% \item Similarly, the functions \texttt{create\_mass\_matrix} and
-%       \texttt{create\_boundary\_mass\_matrix} in the \texttt{MatrixCreator}
-%       namespace no longer
-%       support matrix and vector objects of different types.
+  \todo[inline]{What else is worth mentioning? Someone needs to go
+    through the list of incompatible changes...}
 \end{itemize}

### 2020-05-19 9.2/paper.tex
Update.

@@
 \usepackage{xspace}
+
+\usepackage{siunitx}
 
@@
 %     Denis Davydov,
-%     Marc Fehling,
+    Marc Fehling,
 %     Daniel Garcia-Sanchez,
@@
 %     Jean-Paul Pelteret,
-%     Bruno Turcksin,
+    Bruno Turcksin,
 %     David Wells
@@
 %
-% \author[5]{Marc~Fehling}
-% \affil[5]{Institute for Advanced Simulation,
-%   Forschungszentrum J\"{u}lich GmbH,
-%   52425 J\"{u}lich, Germany.
-%   {\texttt{m.fehling@fz-juelich.de}}}
+\author[5]{Marc~Fehling}
+\affil[5]{Institute for Advanced Simulation,
+  Forschungszentrum J\"{u}lich GmbH,
+  52425 J\"{u}lich, Germany.
+  {\texttt{m.fehling@fz-juelich.de}}}
 %
@@
 %
-% \author[*1]{Bruno~Turcksin}
+\author[*1]{Bruno~Turcksin}
 %
@@
 %
-The major changes are discussed in detail in Section~\ref{sec:major}. There
+These major changes are discussed in detail in Section~\ref{sec:major}. There
 are a number of other noteworthy changes in the current \dealii{} release
@@
 \begin{itemize}
-\item x \todo[inline]{Wolfgang to write about complex-valued output}
-\item y \todo[inline]{Timo to write about problems and fixes for
-  computations with more than $2^{32}$ unknowns}
+\item \dealii{} had decent support for solving complex-valued problems
+  (e.g., ones in quantum mechanics -- like the example covered by the
+  \texttt{step-58} tutorial program covered below -- or for
+  time-harmonic problems) for a while already. However, there were two
+  areas in which support was missing. First, the UMFPACK direct solver
+  packaged with \dealii{} did not support solving complex-valued
+  linear problems. This has now been addressed: UMFPACK actually can
+  solve such systems, we just needed to write the appropriate
+  interfaces. Second, the \texttt{DataOut} class that is responsible
+  for converting nodal data into information that can then be written
+  into files for visualization did not know how to deal with vector-
+  and tensor-valued fields whose components are complex numbers. An
+  example for this is to solve the time-harmonic version of the
+  Maxwell equations that has the electric and magnetic fields as
+  solution. This, too, has been addressed in this release.
 \item z \todo[inline]{What else? Maybe mention the updated step-12?}
@@
 from an external mesh generator in the form of a file that already contains millions
-or tens of millions of cells. For such configurations, applications might already 
-run out of  main memory, while reading the complete mesh by each MPI process.
+or tens of millions of cells. For such configurations, applications might exhaust
+available memory already while reading the mesh on each MPI process.
 
@@
 of mesh refinements, and 4) the owner of the cells on the active mesh level as well
-as on the multigrid levels. Once the triangulation is set up with this struct, no adaptive
-changes to the mesh are allowed at the moment.
+as on the multigrid levels. For the current release, triangulations is set up
+this way cannot be adaptively refined after construction, which is planned to be
+improved for the next release.
 
@@
 
-The new fully distributed triangulation class works---in contrast to  
-\texttt{parallel::distributed::\allowbreak Tri\-an\-gu\-la\-tion}---not only for 2D- and 3D- but also for 
-1D-problems. It can be used in the context of geometric multigrid methods and 
-supports periodic boundary conditions. Furthermore, hanging nodes are supported.
+The new fully distributed triangulation class supports 1D, 2D, and 3D meshes
+including geometric multigrid hierarchies, periodic boundary conditions, and
+hanging nodes.
 
@@
 node-level performance of the matrix-free algorithms in deal.II~\cite{KronbichlerKormann2012, KronbichlerKormann2019}. It is a wrapper
-class around $n$ vector entries of type \texttt{Number} and delegates relevant 
-function calls to appropriate Intrinsics instructions. Up to release 9.1, the 
+class around a short vector of $n$ entries of type \texttt{Number} and maps
+arithmetic operations to appropriate single-instruction/multiple-data (SIMD)
+concepts by intrinsic functions.
+The class \texttt{VectorizedArray} has been made more user-friendly in this release by making
+it compatible with the STL algorithms found in the header \texttt{<algorithm>}.
+The length of the  vector can now be queried by  \texttt{VectorizedArray::size()} and its underlying number type by \texttt{VectorizedArray::value\_type}.
+Furthermore, the \texttt{VectorizedArray} class now supports range-based iterations over its entries.
+
+Up to release 9.1, the
 vector length $n$ has been set at compile time of the library to the highest
 possible value supported by the given processor architecture.
-
-The class \texttt{VectorizedArray} has been made more user-friendly by making 
-it compatible with the STL algorithms found in the header \texttt{<algorithm>}. 
-The length of the  vector can now be queried by  \texttt{VectorizedArray::size()} and its underlying number type by \texttt{VectorizedArray::value\_type}. 
-Furthermore, the \texttt{VectorizedArray} class supports range-based iterations over its entries and, i.a., the following 
-algorithms: \texttt{std::\allowbreak ad\-vance()}, \texttt{std::distance()}, and \texttt{std::max\_element()}.
-
-This class has been also extended with the second optional template argument 
-\texttt{VectorizedArray<Number, size>} with \texttt{size} being related to the 
-vector length, i.e., the number of lanes to be used and the instruction set to be 
-used. By default, the number is set to the highest value supported by the given 
-hardware. A full list of supported 
+Now, a second optional template argument
+\texttt{VectorizedArray<Number, size>} can be given with \texttt{size} explicitly controlling
+the vector length within the capabilities of a particular instruction set.
+A full list of supported
 vector lengths is presented in Table~\ref{tab:vectorizedarray}.
 
-All matrix-free related classes (like \texttt{MatrixFree} and \texttt{FEEvaluation}) 
-have been templated with the floating-point number type \texttt{Number} (e.g. \texttt{double} or \texttt{float}); the computations were performed implicitly 
-on \texttt{VectorizedArray<Number>} structs with the highest 
-instruction-set-architecture extension, with each lane responsible for a separate 
-cell (vectoriziation over elements). In release 9.2, all matrix-free classes 
+To account for the variable-size \texttt{VectorizedArray} class, all matrix-free related classes (like \texttt{MatrixFree} and \texttt{FEEvaluation})
 have been extended with a new optional template argument specifying the
@@
 library \texttt{hyper.deal}~\cite{munch2020hyperdeal}, which solves the 6D Vlasov--Poisson equation with high-order
-discontinuous Galerkin methods (with more than 1024 degrees of freedom per cell), works with \texttt{MatrixFree} objects of different SIMD-vector 
+discontinuous Galerkin methods (with more than a thousand degrees of freedom per cell), constructs a tensor product of two \texttt{MatrixFree} objects of different SIMD-vector
 length in the same application and benefits---in terms of performance---by the possibility of decreasing the number of cells processed by a single SIMD instruction.
@@
 VectorizedArray<double, 1> & VectorizedArray<float, 1> & (auto-vectorization) \\
-VectorizedArray<double, 2> & VectorizedArray<float, 4> & SSE2 \\ 
+VectorizedArray<double, 2> & VectorizedArray<float, 4> & SSE2/AltiVec \\
 VectorizedArray<double, 4> & VectorizedArray<float, 8> & AVX/AVX2 \\
@@
 
-\caption{Comparison of relevant SIMD-related classes in deal.II and C++20.}\label{tab:simd}
+\caption{Comparison of relevant SIMD-related classes in deal.II and \texttt{C++23}.}\label{tab:simd}
 \centering
@@
 \toprule
-\textbf{VectorizedArray (deal.II)} & \textbf{std::simd (C++20)} \\
+\textbf{VectorizedArray (deal.II)} & \textbf{std::simd (\texttt{C++23})} \\
 \midrule
@@
 
-A side effect of introducing the new template argument \texttt{VectorizedArrayType} 
-in the \texttt{MatrixFree} classes is that any data structures 
-\texttt{VectorizedArrayType} can be processed if they support required 
-functionalities like \texttt{size()} or \texttt{value\_type}. In this context, we 
-would like to highlight that the new \texttt{C++20} feature \texttt{std::simd} 
-can be processed by the matrix-free infrastructure with minor internal 
-adjustment as an open pull request shows 
-(see \url{https://github.com/dealii/dealii/pull/9994}).  
+Furthermore, the new interfaces enable using any data structure
+\texttt{VectorizedArrayType} as long as it supports required
+functionalities like \texttt{size()} or \texttt{value\_type}. This prepares
+for the \texttt{C++23} feature \texttt{std::simd} that will be enabled in the future.
 Table~\ref{tab:simd} gives a comparison of the deal.II-specific SIMD classes and
-the equivalent C++20 classes. 
+the equivalent \texttt{C++23} classes. Finally, this change also prepares for specialized
+code paths exploiting
+vectorization within an element~\cite{KronbichlerKormann2019} in the future.
+
 %We welcome the standardization of the SIMD
@@
 
-In a next step, we intend to support that users could set 
-\texttt{VectorizedArrayType} to \texttt{Number}, which would lead to the usage not 
-of \texttt{VectorizedArray<Number, 1>} but of a specialized code-path exploiting 
-vectorization within an element~\cite{KronbichlerKormann2019}.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Large-scale simulations with 304,128 cores have revealed bottlenecks in release 
-9.1 during initialization due to the usage of expensive collective operations 
-like \texttt{MPI\_Allgather()} and \texttt{MPI\_\allowbreak Alltoall()}, e.g., during the 
-pre-computation of the index ranges of all processes, which have been stored in  an array. 
+9.1 during initialization of a number of distributed data structures due to the usage of expensive collective operations 
+like \texttt{MPI\_Allgather()} and \texttt{MPI\_\allowbreak
+  Alltoall()}. Typical examples are the
+pre-computation of those indices of the vector entries (or
+other linear index ranges) owned by
+each process, which were previously stored in  an array on every process. 
 This information is needed to set up 
 the  \texttt{Utilities::MPI::Par\-ti\-ti\-oner} class.
-In release 9.2, we have removed such arrays and have replaced 
-the \texttt{MPI\_Allgather}/\allowbreak\texttt{MPI\_\allowbreak Alltoall} 
-function calls by consensus algorithms~\cite{hoefler2010scalable}, which can be 
-found in the namespace \texttt{Utilities::\allowbreak MPI::\allowbreak ConsensusAlgorithms} (short: \texttt{CA}): now, only the locally relevant information about the index ranges is (re)computed when needed, using these algorithms.
-We provide two flavors of the consensus algorithm: the two-step 
-approach \texttt{CA::PEX} and the \texttt{CA::NBX}, 
-which uses only point-to-point communications and a single \texttt{MPI\_IBarrier()}.
+In release 9.2, we have replaced these functions in favor of
+consensus algorithms~\cite{hoefler2010scalable}, which can be
+found in the namespace \texttt{Utilities::\allowbreak MPI::\allowbreak ConsensusAlgorithms} (short: \texttt{CA}).
+Now, only the locally relevant information about the index ranges is
+(re)computed when needed, which, for more than 100 MPI processes, uses
+point-to-point communications and a single \texttt{MPI\_IBarrier()}.
 
@@
 
-By replacing the collective communications during setup and removing the arrays 
+\todo[inline]{Aren't the target processes the result of the operation, and the
+  input global numbers of indices that we request?}
+
+By replacing the collective communications during set up and removing the arrays
 that contain information for each process (enabled by the application of consensus
 algorithms and other modifications---a full list of modifications leading to this
-improvement can be found online), we were able to significantly improve the setup 
+improvement can be found online), we were able to significantly
+improve the set up
 time for large-scale simulations and to solve a Poisson problem with multigrid
-with 12T unknowns. 
-Figure~\ref{} compares the timings of a simulation (incl. setup) with the 
+with \num{12e12} unknowns.
+\todo[inline]{Where did we solve for 12T unknowns? The largest I [Martin] have done is 2.1T unknowns.}
+Figure~\ref{} compares the timings of a simulation (including set up) with the
 previous release 9.1 and with the current release 9.2.
@@
 The new code has been also applied to solve problems with adaptively refined
-meshes with more than 4B unknowns. {\color{red}TODO[Timo]}
+meshes with more than \num{4e9} unknowns. {\color{red}TODO[Timo]}
+
+\todo[inline]{We did solve adaptive problems with more than 4B unknowns before, but some were broken for a while during this release stage due to the consensus algorithms.
+But we did not test systems or block vectors before.
+But we should not make it sound too negative.}
 
@@
 
-\todo[inline]{Marc: Your section}
+With the previous release, \dealii supports $hp$-adaptive finite element methods
+on distributed memory systems \cite{dealII91}. We implemented the bare functionality
+for $hp$-adaptive methods with the objective to offer the greatest flexibility in
+their application. Here, reference finite elements still had to be assigned manually
+to each cell, which may not lead to an optimal mesh and is tedious.
+
+With the current release, $hp$-adaptive finite element methods have been further
+expanded: New features like decision strategies have been added and the user interface
+has been overhauled, effectively making $hp$-methods more attractive to use. We introduced
+many new functions that automatize the general workflow for applying $hp$-decision
+strategies, which run on top of the previous low-level implementation for both serial
+and parallel applications.
+
+The interface is now as simple to use as the one for pure adaptive mesh refinement.
+Consider the following (incomplete) listing as an example: We estimate both error and
+smoothness of the finite element approximation. Further, we flag certain fractions of
+cells with the highest and lowest errors for refinement and coarsening, respectively
+(here: 30\% / 3\%). From those cells listed for adaptation, we specify a subset of them
+for $h$- and $p$-adaptation (here: 10\% / 90\%).
+\begin{c++}
+Vector<float> estimated_error_per_cell (n_active_cells);
+KellyErrorEstimator::estimate(
+  dof_handler, ..., solution, estimated_error_per_cell, ...);
+GridRefinement::refine_and_coarsen_fixed_number(
+  triangulation, estimated_error_per_cell, 0.3, 0.03);
+
+Vector<float> estimated_smoothness_per_cell (n_active_cells);
+SmoothnessEstimator::Legendre::coefficient_decay(
+  ..., dof_handler, solution, estimated_smoothness_per_cell);
+hp::Refinement::p_adaptivity_fixed_number(
+  dof_handler, estimated_smoothness_per_cell, 0.9, 0.9);
+hp::Refinement::choose_p_over_h (dof_handler);
+
+triangulation.execute_coarsening_and_refinement();
+\end{c++}
+
+In particular, we implemented decision strategies based on refinement history and
+smoothness estimation, and made sure that they work for refinement as well as
+coarsening in terms of $h$- and $p$-adaptation in serial and parallel applications.
+
+The former relies on knowing an estimate for the upper error bound \cite[Thm.~3.4]{BabuskaSuri1990}.
+For successive refinements, we can predict how the error will change on basis of
+current error estimates and adaptation flags. In the next refinement cycle, these
+predicted error estimates allow us to decide whether the choice of adaptation in
+the previous cycle was justified, and provide the foundation for it in the next
+one \cite{MelenkWohlmuth2001}.
+
+In general, $p$-refinement is favorable over $h$-refinement in smooth regions of
+the finite element approximation \cite[Thm.~3.4]{BabuskaSuri1990}. Thus, estimating
+its smoothness provides a suitable decision indicator for $hp$-adaptation. For this
+purpose, we express the finite element approximation in an orthogonal
+basis of increasing frequency, and consider the decay of their expansion
+coefficients as the estimation of smoothness. This has been implemented for both
+Fourier coefficients \cite{BangerthKayserHerold2007} and Legendre coefficients
+\cite{Mavriplis1994,HoustonSeniorSueli2003,HoustonSueli2005,EibnerMelenk2007}.
 
@@
 \todo[inline]{Luca: Your section}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Improved performance of symbolic framework}
+\label{subsec:symbdiff}
+
+\todo[inline]{J-P's section}
+
+BatchOptimizer for symbolic expressions
+\begin{itemize}
+\item CSE for dictionary-based expressions
+\item ``lambdify'' (via SymEngine)
+\item Offloading to LLVM JIT compiler (via SymEngine)
+\item Serialization, so complex expressions can be compiled offline
+\end{itemize}
 
@@
 manifold class that can propagate curved boundary information into the
-interior of a computational domain, and \texttt{MappingQCache}, which can sample 
-the information of expensive manifolds in the points of a \texttt{MappingQ} and 
-cache it for further use.
+interior of a computational domain by transfinite interpolation \cite{Gordon82}.
+This manifold is a prototype for many other manifolds in that it is relatively
+expensive to compute the new points, especially for higher order mappings. Since
+typical programs query higher-order geometries in a large variety of contexts,
+the contribution of the mapping to the run time can be significant. As a
+solution, the tutorial also presents the class \texttt{MappingQCache}, which
+samples the information of expensive manifolds in the points of a
+\texttt{MappingQ} and caches it. The tutorial shows that this makes all queries
+to the geometry very cheap.
 
@@
 products. Furthermore, this tutorial demonstrates the usage of new
-pre and post operations, which can be passed to \texttt{cell\_loop()}, to schedule operations on sections of vectors close 
-to the matrix-vector product to increase data locality 
-and discusses performance-related aspects.
+pre and post operations, which can be passed to
+\texttt{MatrixFree::cell\_loop()}, to schedule operations on sections of vectors close
+to the matrix-vector product to increase data locality.
 
@@
 \item \texttt{step-70}
-  \todo[inline]{Also need to update announce and announce-short if this
-    makes it into the release.}
+  \todo[inline]{Luca, please write.}
 \end{itemize}
@@
    \cite{DeSimoneHeltaiManigrasso2009};
+ \item For curved geometry representations and manifolds:
+   \cite{HeltaiBangerthKronbichlerMola2019};
  \item For integration with CAD files and tools:
@@
 \dealii can interface with many other libraries:
+\todo[inline]{We picked up gingko. Anything else?}
 \begin{multicols}{3}
@@
 
-D.~Arndt and M.~Kronbichler were partially supported by the German
-Research Foundation (DFG) under the project ``High-order discontinuous
-Galerkin for the exa-scale'' (\mbox{ExaDG}) within the priority program ``Software
-for Exascale Computing'' (SPPEXA).
-
 W.~Bangerth, T.~C.~Clevenger, and T.~Heister were partially
@@
 
-M.~Kronbichler was also supported by the Bayerisches Kompetenznetzwerk
+M.~Kronbichler was supported by the German
+Research Foundation (DFG) under the project ``High-order discontinuous
+Galerkin for the exa-scale'' (\mbox{ExaDG}) within the priority program ``Software
+for Exascale Computing'' (SPPEXA) and the Bayerisches Kompetenznetzwerk
 f\"ur Technisch-Wissen\-schaft\-li\-ches Hoch- und H\"ochstleistungsrechnen

### 2020-05-23 9.2/paper.tex
Add section for SD::BatchOptimizer
Add myself to the author list
Improve and shorten some text

--- 2020-05-14 9.2/paper.tex
@@
 node-level performance of the matrix-free algorithms in deal.II~\cite{KronbichlerKormann2012, KronbichlerKormann2019}. It is a wrapper
-class around $n$ vector entries of type \texttt{Number} and delegates relevant 
-function calls to appropriate Intrinsics instructions. Up to release 9.1, the 
+class around a short vector of $n$ entries of type \texttt{Number} and maps
+arithmetic operations to appropriate single-instruction/multiple-data (SIMD)
+concepts by intrinsic functions.
+The class \texttt{VectorizedArray} has been made more user-friendly in this release by making
+it compatible with the STL algorithms found in the header \texttt{<algorithm>}.
+The length of the  vector can now be queried by  \texttt{VectorizedArray::size()} and its underlying number type by \texttt{VectorizedArray::value\_type}.
+Furthermore, the \texttt{VectorizedArray} class now supports range-based iterations over its entries.
+
+Up to release 9.1, the
 vector length $n$ has been set at compile time of the library to the highest
 possible value supported by the given processor architecture.
+Now, a second optional template argument
+\texttt{VectorizedArray<Number, size>} can be given with \texttt{size} explicitly controlling
+the vector length within the capabilities of a particular instruction set.
+A full list of supported
+vector lengths is presented in Table~\ref{tab:vectorizedarray}.
+\todo[inline]{This table lacks the AltiVec support}
 
-The class \texttt{VectorizedArray} has been made more user-friendly by making 
-it compatible with the STL algorithms found in the header \texttt{<algorithm>}. 
-The length of the  vector can now be queried by  \texttt{VectorizedArray::size()} and its underlying number type by \texttt{VectorizedArray::value\_type}. 
-Furthermore, the \texttt{VectorizedArray} class supports range-based iterations over its entries and, i.a., the following 
-algorithms: \texttt{std::\allowbreak ad\-vance()}, \texttt{std::distance()}, and \texttt{std::max\_element()}.
-
-This class has been also extended with the second optional template argument 
-\texttt{VectorizedArray<Number, size>} with \texttt{size} being related to the 
-vector length, i.e., the number of lanes to be used and the instruction set to be 
-used. By default, the number is set to the highest value supported by the given 
-hardware. A full list of supported 
-vector lengths is presented in Table~\ref{tab:vectorizedarray}.
-
-All matrix-free related classes (like \texttt{MatrixFree} and \texttt{FEEvaluation}) 
-have been templated with the floating-point number type \texttt{Number} (e.g. \texttt{double} or \texttt{float}); the computations were performed implicitly 
-on \texttt{VectorizedArray<Number>} structs with the highest 
-instruction-set-architecture extension, with each lane responsible for a separate 
-cell (vectoriziation over elements). In release 9.2, all matrix-free classes 
+To account for the variable-size \texttt{VectorizedArray} class, all matrix-free related classes (like \texttt{MatrixFree} and \texttt{FEEvaluation})
 have been extended with a new optional template argument specifying the
@@
 
-A side effect of introducing the new template argument \texttt{VectorizedArrayType} 
-in the \texttt{MatrixFree} classes is that any data structures 
-\texttt{VectorizedArrayType} can be processed if they support required 
-functionalities like \texttt{size()} or \texttt{value\_type}. In this context, we 
-would like to highlight that the new \texttt{C++20} feature \texttt{std::simd} 
-can be processed by the matrix-free infrastructure with minor internal 
-adjustment as an open pull request shows 
-(see \url{https://github.com/dealii/dealii/pull/9994}).  
+Furthermore, the new interfaces enable using any data structure
+\texttt{VectorizedArrayType} as long as it supports required
+functionalities like \texttt{size()} or \texttt{value\_type}. This prepares
+for the \texttt{C++23} feature \texttt{std::simd} that will be enabled in the future.
 Table~\ref{tab:simd} gives a comparison of the deal.II-specific SIMD classes and
-the equivalent C++20 classes. 
+the equivalent C++20 classes. Finally, this change also prepares for specialized
+code paths exploiting
+vectorization within an element~\cite{KronbichlerKormann2019} in the future.
+
 %We welcome the standardization of the SIMD
@@
 
-Large-scale simulations with 304,128 cores have revealed bottlenecks in release 
-9.1 during initialization due to the usage of expensive collective operations 
-like \texttt{MPI\_Allgather()} and \texttt{MPI\_\allowbreak Alltoall()}, e.g., during the 
-pre-computation of the index ranges of all processes, which have been stored in  an array. 
-This information is needed to set up 
-the  \texttt{Utilities::MPI::Par\-ti\-ti\-oner} class. 
-In release 9.2, we have removed such arrays and have replaced 
-the \texttt{MPI\_Allgather}/\allowbreak\texttt{MPI\_\allowbreak Alltoall} 
-function calls by consensus algorithms~\cite{hoefler2010scalable}, which can be 
-found in the namespace \texttt{Utilities::\allowbreak MPI::\allowbreak ConsensusAlgorithms} (short: \texttt{CA}): now, only the locally relevant information about the index ranges is (re)computed when needed, using these algorithms.
-We provide two flavors of the consensus algorithm: the two-step 
-approach \texttt{CA::PEX} and the \texttt{CA::NBX}, 
-which uses only point-to-point communications and a single \texttt{MPI\_IBarrier()}.
+Large-scale simulations with 304,128 cores have revealed bottlenecks in setup
+routines due to the usage of expensive collective operations
+like \texttt{MPI\_Allgather()} and \texttt{MPI\_\allowbreak Alltoall()}, storing
+problem sizes and offsets from all processors in the MPI universe.
+In release 9.2, we have replaced these functions in favor of
+consensus algorithms~\cite{hoefler2010scalable}, which can be
+found in the namespace \texttt{Utilities::\allowbreak MPI::\allowbreak ConsensusAlgorithms} (short: \texttt{CA}).
+Now, only the locally relevant information about the index ranges is
+(re)computed when needed, which, for more than 100 MPI processes, uses
+point-to-point communications and a single \texttt{MPI\_IBarrier()}.
 
@@
 objects to \texttt{CA::AnonymousProcess}.
+\todo[inline]{Aren't the target processes the result of the operation, and the
+  input global numbers of indices that we request?}
 
@@
 with 12T unknowns.
+\todo[inline]{Where did we solve for 12T unknowns? The largest I [Martin] have done is 2.1T unknowns.}
 Figure~\ref{} compares the timings of a simulation (incl. setup) with the

### 2020-05-23 9.2/paper.tex
Add another todo
Improve text on tutorials

@@
 
+\todo[inline]{We did solve adaptive problems with more than 4B unknowns before, but some were broken for a while during this release stage due to the consensus algorithms.
+But we did not test systems or block vectors before.
+But we should not make it sound too negative.}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 manifold class that can propagate curved boundary information into the
-interior of a computational domain, and \texttt{MappingQCache}, which can sample
-the information of expensive manifolds in the points of a \texttt{MappingQ} and
-cache it for further use.
+interior of a computational domain by transfinite interpolation \cite{Gordon82}.
+This manifold is a prototype for many other manifolds in that it is relatively
+expensive to compute the new points, especially for higher order mappings. Since
+typical programs query higher-order geometries in a large variety of contexts,
+the contribution of the mapping to the run time can be significant. As a
+solution, the tutorial also presents the class \texttt{MappingQCache}, which
+samples the information of expensive manifolds in the points of a
+\texttt{MappingQ} and caches it. The tutorial shows that this makes all queries
+to the geometry very cheap.
 
@@
 products. Furthermore, this tutorial demonstrates the usage of new
-pre and post operations, which can be passed to \texttt{cell\_loop()}, to schedule operations on sections of vectors close
-to the matrix-vector product to increase data locality
-and discusses performance-related aspects.
+pre and post operations, which can be passed to
+\texttt{MatrixFree::cell\_loop()}, to schedule operations on sections of vectors close
+to the matrix-vector product to increase data locality.
 

### 2020-05-23 9.2/paper.tex
Add myself to author list.
Extension of hp-Interface.

@@
 
-\todo[inline]{Marc: Your section}
+With the previous release, \dealii supports $hp$-adaptive finite element methods
+on distributed memory systems \cite{dealII91}. We implemented the bare functionality
+for $hp$-adaptive methods with the objective to offer the greatest flexibility in
+their application. Here, reference finite elements still had to be assigned manually
+to each cell, which may not lead to an optimal mesh and is tedious.
+
+With the current release, $hp$-adaptive finite element methods have been further
+expanded in both the range of features in terms of decision strategies and their 
+user interface, effectively making the more attractive to use. We introduced many
+new functions that automatize the general workflow for applying $hp$-decision
+strategies, which run on top of the previous low-level implementation for both serial
+and parallel applications.
+
+The interface is now as simple to use as the one for pure adaptive mesh refinement.
+Consider the following (incomplete) listing as an example: We estimate both error and
+smoothness of the finite element approximation. Further, we flag certain fractions of
+cells with the highest and lowest errors for refinement and coarsening, respectively
+(here: 30\% / 3\%). From those cells listed for adaptation, we specify a subset of them
+for $h$- and $p$-adaptation (here: 10\% / 90\%).
+\begin{c++}
+Vector<float> estimated_error_per_cell (n_active_cells);
+KellyErrorEstimator::estimate(
+  dof_handler, ..., solution, estimated_error_per_cell, ...);
+GridRefinement::refine_and_coarsen_fixed_number(
+  triangulation, estimated_error_per_cell, 0.3, 0.03);
+
+Vector<float> estimated_smoothness_per_cell (n_active_cells);
+SmoothnessEstimator::Legendre::coefficient_decay(
+  ..., dof_handler, solution, estimated_smoothness_per_cell);
+hp::Refinement::p_adaptivity_fixed_number(
+  dof_handler, estimated_smoothness_per_cell, 0.9, 0.9);
+hp::Refinement::choose_p_over_h (dof_handler);
+
+triangulation.execute_coarsening_and_refinement();
+\end{c++}
+
+In particular, we implemented decision strategies based on refinement history and
+smoothness estimation, and made sure that they work for refinement as well as
+coarsening in terms of $h$- and $p$-adaptation in serial and parallel applications.
+
+The former relies on knowing an estimate for the upper error bound \cite[Thm.~3.4]{BabuskaSuri1990}.
+For successive refinements, we can predict how the error will change on basis of
+current error estimates and adaptation flags. In the next refinement cycle, these
+predicted error estimates allow us to decide whether the choice of adaptation in
+the previous cycle was justified, and provide the foundation for it in the next
+one \cite{MelenkWohlmuth2001}.
+
+In general, $p$-refinement is favorable over $h$-refinement in smooth regions of
+the finite element approximation \cite[Thm.~3.4]{BabuskaSuri1990}. Thus, estimating
+its smoothness provides a suitable decision indicator for $hp$-adaptation. For this
+purpose, we expand the finite element approximation into a series of orthogonal
+basis functions of increasing complexity, and consider the decay of their expansion
+coefficients as the estimation of smoothness. This has been implemented for both
+Fourier coefficients \cite{BangerthKayserHerold2007} and Legendre coefficients
+\cite{Mavriplis1994,HoustonSeniorSueli2003,HoustonSueli2005,EibnerMelenk2007}.
 

### 2020-05-23 9.2/paper.tex
Small fixes in simd section

--- 2020-05-23 9.2/paper.tex
@@
 vector lengths is presented in Table~\ref{tab:vectorizedarray}.
-\todo[inline]{This table lacks the AltiVec support}
 
@@
 VectorizedArray<double, 1> & VectorizedArray<float, 1> & (auto-vectorization) \\
-VectorizedArray<double, 2> & VectorizedArray<float, 4> & SSE2 \\
+VectorizedArray<double, 2> & VectorizedArray<float, 4> & SSE2/AltiVec \\
 VectorizedArray<double, 4> & VectorizedArray<float, 8> & AVX/AVX2 \\
@@
 Table~\ref{tab:simd} gives a comparison of the deal.II-specific SIMD classes and
-the equivalent C++20 classes. Finally, this change also prepares for specialized
+the equivalent \texttt{C++23} classes. Finally, this change also prepares for specialized
 code paths exploiting

### 2020-05-24 9.2/paper.tex
Apply suggestions from review
Co-authored-by: Martin Kronbichler <kronbichler@lnm.mw.tum.de>

--- 2020-05-23 9.2/paper.tex
@@
 With the current release, $hp$-adaptive finite element methods have been further
-expanded in both the range of features in terms of decision strategies and their 
-user interface, effectively making the more attractive to use. We introduced many
-new functions that automatize the general workflow for applying $hp$-decision
+expanded: New features like decision strategies have been added and the user interface
+has been overhauled, effectively making $hp$-methods more attractive to use. We introduced
+many new functions that automatize the general workflow for applying $hp$-decision
 strategies, which run on top of the previous low-level implementation for both serial
@@
 its smoothness provides a suitable decision indicator for $hp$-adaptation. For this
-purpose, we expand the finite element approximation into a series of orthogonal
-basis functions of increasing complexity, and consider the decay of their expansion
+purpose, we express the finite element approximation in an orthogonal
+basis of increasing frequency, and consider the decay of their expansion
 coefficients as the estimation of smoothness. This has been implemented for both

### 2020-05-24 9.2/paper.tex
Add a todo.

@@
 
-Large-scale simulations with 304,128 cores have revealed bottlenecks in setup
-routines due to the usage of expensive collective operations
-like \texttt{MPI\_Allgather()} and \texttt{MPI\_\allowbreak Alltoall()}, storing
-problem sizes and offsets from all processors in the MPI universe.
+Large-scale simulations with 304,128 cores have revealed bottlenecks in release 
+9.1 during initialization of a number of distributed data structures due to the usage of expensive collective operations 
+like \texttt{MPI\_Allgather()} and \texttt{MPI\_\allowbreak
+  Alltoall()}. Typical examples are the
+pre-computation of those indices of the vector entries (or
+other linear index ranges) owned by
+each process, which were previously stored in  an array on every process. 
+This information is needed to set up 
+the  \texttt{Utilities::MPI::Par\-ti\-ti\-oner} class.
 In release 9.2, we have replaced these functions in favor of
@@
 objects to \texttt{CA::AnonymousProcess}.
+
 \todo[inline]{Aren't the target processes the result of the operation, and the
@@
 
-By replacing the collective communications during setup and removing the arrays
+By replacing the collective communications during set up and removing the arrays
 that contain information for each process (enabled by the application of consensus
 algorithms and other modifications---a full list of modifications leading to this
-improvement can be found online), we were able to significantly improve the setup
+improvement can be found online), we were able to significantly
+improve the set up
 time for large-scale simulations and to solve a Poisson problem with multigrid
-with 12T unknowns.
+with \num{12e12} unknowns.
 \todo[inline]{Where did we solve for 12T unknowns? The largest I [Martin] have done is 2.1T unknowns.}
-Figure~\ref{} compares the timings of a simulation (incl. setup) with the
+Figure~\ref{} compares the timings of a simulation (including set up) with the
 previous release 9.1 and with the current release 9.2.
@@
 The new code has been also applied to solve problems with adaptively refined
-meshes with more than 4B unknowns. {\color{red}TODO[Timo]}
+meshes with more than \num{4e9} unknowns. {\color{red}TODO[Timo]}
 
@@
 
+\todo[inline]{Need to add the particles paper by Gassmoeller et al.}
+
 \dealii can interface with many other libraries:
+\todo[inline]{We picked up gingko. Anything else?}
 \begin{multicols}{3}

### 2020-05-25 9.2/paper.tex
Added LH to authors.
Leave a note.
Re-sort sections.

@@
     Timo Heister,
-%     Luca Heltai,
+    Luca Heltai,
     Martin Kronbichler,
@@
 
-% %
-% \author[8]{Luca~Heltai}
-% \affil[8]{SISSA,
-%   International School for Advanced Studies,
-%   Via Bonomea 265,
-%   34136, Trieste, Italy.
-%   {\texttt{luca.heltai@sissa.it}}}
-%
+\author[8]{Luca~Heltai}
+\affil[8]{SISSA,
+   International School for Advanced Studies,
+   Via Bonomea 265,
+   34136, Trieste, Italy.
+   {\texttt{luca.heltai@sissa.it}}}
+
  \author[9]{Martin~Kronbichler}
@@
   solution. This, too, has been addressed in this release.
+  \item \todo[inline]{Reza: Please add a short summary of the
+    \texttt{DiscreteTime} class}
 \item z \todo[inline]{What else? Maybe mention the updated step-12?}
@@
 %rely on a user-provided partitioner, which might also be a graph partitioner.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Advances of the SIMD capabilities and the matrix-free infrastructure}
-\label{subsec:mf}
-
-%\begin{itemize}
-%\item ECL
-%\item VectorizedArrayType
-%\end{itemize}
-
-The class \texttt{VectorizedArray<Number>} is a key ingredient for the high
-node-level performance of the matrix-free algorithms in deal.II~\cite{KronbichlerKormann2012, KronbichlerKormann2019}. It is a wrapper
-class around a short vector of $n$ entries of type \texttt{Number} and maps
-arithmetic operations to appropriate single-instruction/multiple-data (SIMD)
-concepts by intrinsic functions.
-The class \texttt{VectorizedArray} has been made more user-friendly in this release by making
-it compatible with the STL algorithms found in the header \texttt{<algorithm>}.
-The length of the  vector can now be queried by  \texttt{VectorizedArray::size()} and its underlying number type by \texttt{VectorizedArray::value\_type}.
-Furthermore, the \texttt{VectorizedArray} class now supports range-based iterations over its entries.
-
-Up to release 9.1, the
-vector length $n$ has been set at compile time of the library to the highest
-possible value supported by the given processor architecture.
-Now, a second optional template argument
-\texttt{VectorizedArray<Number, size>} can be given with \texttt{size} explicitly controlling
-the vector length within the capabilities of a particular instruction set.
-A full list of supported
-vector lengths is presented in Table~\ref{tab:vectorizedarray}.
-
-To account for the variable-size \texttt{VectorizedArray} class, all matrix-free related classes (like \texttt{MatrixFree} and \texttt{FEEvaluation})
-have been extended with a new optional template argument specifying the
-\texttt{VectorizedArrayType}. This allows users to select the vector length/ISA and,
-as a consequence, the number of cells to be processed at once directly in their applications:
-The deal.II-based
-library \texttt{hyper.deal}~\cite{munch2020hyperdeal}, which solves the 6D Vlasov--Poisson equation with high-order
-discontinuous Galerkin methods (with more than a thousand degrees of freedom per cell), constructs a tensor product of two \texttt{MatrixFree} objects of different SIMD-vector
-length in the same application and benefits---in terms of performance---by the possibility of decreasing the number of cells processed by a single SIMD instruction.
-
-\begin{table}
-\caption{Supported vector lengths of the class \texttt{VectorizedArray} and
-the corresponding instruction-set-architecture extensions. }\label{tab:vectorizedarray}
-\centering
-\begin{tabular}{ccc}
-\toprule
-\textbf{double} & \textbf{float} & \textbf{ISA}\\
-\midrule
-VectorizedArray<double, 1> & VectorizedArray<float, 1> & (auto-vectorization) \\
-VectorizedArray<double, 2> & VectorizedArray<float, 4> & SSE2/AltiVec \\
-VectorizedArray<double, 4> & VectorizedArray<float, 8> & AVX/AVX2 \\
-VectorizedArray<double, 8> & VectorizedArray<float, 16> & AVX-512 \\
-\bottomrule
-\end{tabular}
-
-\caption{Comparison of relevant SIMD-related classes in deal.II and \texttt{C++23}.}\label{tab:simd}
-\centering
-\begin{tabular}{cc}
-\toprule
-\textbf{VectorizedArray (deal.II)} & \textbf{std::simd (\texttt{C++23})} \\
-\midrule
-VectorizedArray<Number> & std::experimental::native\_simd<Number> \\
-VectorizedArray<Number, size> & std::experimental::fixed\_size\_simd<Number, size> \\ \bottomrule
-\end{tabular}
-\end{table}
-
-Furthermore, the new interfaces enable using any data structure
-\texttt{VectorizedArrayType} as long as it supports required
-functionalities like \texttt{size()} or \texttt{value\_type}. This prepares
-for the \texttt{C++23} feature \texttt{std::simd} that will be enabled in the future.
-Table~\ref{tab:simd} gives a comparison of the deal.II-specific SIMD classes and
-the equivalent \texttt{C++23} classes. Finally, this change also prepares for specialized
-code paths exploiting
-vectorization within an element~\cite{KronbichlerKormann2019} in the future.
-
-%We welcome the standardization of the SIMD
-%parallelization paradigm in C++ and intend to replace step by step our own
-%wrapper class, which has been continuously developed over the last decade. We
-%would like to emphasize that the work invested in this class was not in vain,
-%since many performance-relevant utility functions implemented with \texttt{VectorizedArray} in mind (e.g., \texttt{vectorized\_load\_and\_transpose}
-%and \texttt{vectorized\_transpose\_and\_store}) will be still used, since they
-%have not become part of the standard.
-
-%Further additions to the \texttt{MatrixFree} infrastructure consist of:
-%\begin{itemize}
-%\item a new variant of \texttt{MatrixFree::cell\_loop()}: It takes two
-%\texttt{std::function} objects with ranges on the locally owned degrees of freedom, one
-%with work to be scheduled before the cell operation first touches some
-%unknowns and another with work to be executed after the cell operation last
-%touches them. The goal of
-%these functions is to bring vector operations close to the time when the
-%vector entries are accessed by the cell operation, which increases the cache
-%hit rate of modern processors by improved temporal locality.
-%\item a new form of loop \texttt{MatrixFree::loop\_cell\_centric()}: This
-%kind of loop can be used in the context of discontinuous Galerkin methods,
-%where both cell and face integrals have to be evaluated. While in the case of
-%the traditional \texttt{loop}, cell and face integrals have been performed
-%independently, the new loop performs all cell and face integrals of a cell in
-%one go. This includes that each face integral has to be evaluated twice, but
-%entries have to be written into the solution vector only once with improved
-%data locality. Previous publications based on \texttt{deal.II} have shown the
-%relevance of the latter aspect for reaching higher performance.
-%\end{itemize}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Advances in GPU support}
-\label{subsec:gpu}
-
-\begin{itemize}
-\item overlapping of computation and communication in the case of CUDA-aware MPI
-\end{itemize}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Advances of the SIMD capabilities and the matrix-free infrastructure}
+\label{subsec:mf}
+
+%\begin{itemize}
+%\item ECL
+%\item VectorizedArrayType
+%\end{itemize}
+
+The class \texttt{VectorizedArray<Number>} is a key ingredient for the high
+node-level performance of the matrix-free algorithms in deal.II~\cite{KronbichlerKormann2012, KronbichlerKormann2019}. It is a wrapper
+class around a short vector of $n$ entries of type \texttt{Number} and maps
+arithmetic operations to appropriate single-instruction/multiple-data (SIMD)
+concepts by intrinsic functions.
+The class \texttt{VectorizedArray} has been made more user-friendly in this release by making
+it compatible with the STL algorithms found in the header \texttt{<algorithm>}.
+The length of the  vector can now be queried by  \texttt{VectorizedArray::size()} and its underlying number type by \texttt{VectorizedArray::value\_type}.
+Furthermore, the \texttt{VectorizedArray} class now supports range-based iterations over its entries.
+
+Up to release 9.1, the
+vector length $n$ has been set at compile time of the library to the highest
+possible value supported by the given processor architecture.
+Now, a second optional template argument
+\texttt{VectorizedArray<Number, size>} can be given with \texttt{size} explicitly controlling
+the vector length within the capabilities of a particular instruction set.
+A full list of supported
+vector lengths is presented in Table~\ref{tab:vectorizedarray}.
+
+To account for the variable-size \texttt{VectorizedArray} class, all matrix-free related classes (like \texttt{MatrixFree} and \texttt{FEEvaluation})
+have been extended with a new optional template argument specifying the
+\texttt{VectorizedArrayType}. This allows users to select the vector length/ISA and,
+as a consequence, the number of cells to be processed at once directly in their applications:
+The deal.II-based
+library \texttt{hyper.deal}~\cite{munch2020hyperdeal}, which solves the 6D Vlasov--Poisson equation with high-order
+discontinuous Galerkin methods (with more than a thousand degrees of freedom per cell), constructs a tensor product of two \texttt{MatrixFree} objects of different SIMD-vector
+length in the same application and benefits---in terms of performance---by the possibility of decreasing the number of cells processed by a single SIMD instruction.
+
+\begin{table}
+\caption{Supported vector lengths of the class \texttt{VectorizedArray} and
+the corresponding instruction-set-architecture extensions. }\label{tab:vectorizedarray}
+\centering
+\begin{tabular}{ccc}
+\toprule
+\textbf{double} & \textbf{float} & \textbf{ISA}\\
+\midrule
+VectorizedArray<double, 1> & VectorizedArray<float, 1> & (auto-vectorization) \\
+VectorizedArray<double, 2> & VectorizedArray<float, 4> & SSE2/AltiVec \\
+VectorizedArray<double, 4> & VectorizedArray<float, 8> & AVX/AVX2 \\
+VectorizedArray<double, 8> & VectorizedArray<float, 16> & AVX-512 \\
+\bottomrule
+\end{tabular}
+
+\caption{Comparison of relevant SIMD-related classes in deal.II and \texttt{C++23}.}\label{tab:simd}
+\centering
+\begin{tabular}{cc}
+\toprule
+\textbf{VectorizedArray (deal.II)} & \textbf{std::simd (\texttt{C++23})} \\
+\midrule
+VectorizedArray<Number> & std::experimental::native\_simd<Number> \\
+VectorizedArray<Number, size> & std::experimental::fixed\_size\_simd<Number, size> \\ \bottomrule
+\end{tabular}
+\end{table}
+
+Furthermore, the new interfaces enable using any data structure
+\texttt{VectorizedArrayType} as long as it supports required
+functionalities like \texttt{size()} or \texttt{value\_type}. This prepares
+for the \texttt{C++23} feature \texttt{std::simd} that will be enabled in the future.
+Table~\ref{tab:simd} gives a comparison of the deal.II-specific SIMD classes and
+the equivalent \texttt{C++23} classes. Finally, this change also prepares for specialized
+code paths exploiting
+vectorization within an element~\cite{KronbichlerKormann2019} in the future.
+
+%We welcome the standardization of the SIMD
+%parallelization paradigm in C++ and intend to replace step by step our own
+%wrapper class, which has been continuously developed over the last decade. We
+%would like to emphasize that the work invested in this class was not in vain,
+%since many performance-relevant utility functions implemented with \texttt{VectorizedArray} in mind (e.g., \texttt{vectorized\_load\_and\_transpose}
+%and \texttt{vectorized\_transpose\_and\_store}) will be still used, since they
+%have not become part of the standard.
+
+%Further additions to the \texttt{MatrixFree} infrastructure consist of:
+%\begin{itemize}
+%\item a new variant of \texttt{MatrixFree::cell\_loop()}: It takes two
+%\texttt{std::function} objects with ranges on the locally owned degrees of freedom, one
+%with work to be scheduled before the cell operation first touches some
+%unknowns and another with work to be executed after the cell operation last
+%touches them. The goal of
+%these functions is to bring vector operations close to the time when the
+%vector entries are accessed by the cell operation, which increases the cache
+%hit rate of modern processors by improved temporal locality.
+%\item a new form of loop \texttt{MatrixFree::loop\_cell\_centric()}: This
+%kind of loop can be used in the context of discontinuous Galerkin methods,
+%where both cell and face integrals have to be evaluated. While in the case of
+%the traditional \texttt{loop}, cell and face integrals have been performed
+%independently, the new loop performs all cell and face integrals of a cell in
+%one go. This includes that each face integral has to be evaluated twice, but
+%entries have to be written into the solution vector only once with improved
+%data locality. Previous publications based on \texttt{deal.II} have shown the
+%relevance of the latter aspect for reaching higher performance.
+%\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Advances in GPU support}
+\label{subsec:gpu}
+
+\begin{itemize}
+\item overlapping of computation and communication in the case of CUDA-aware MPI
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Expanded use of C++11 facilities}
+\label{subsec:cxx}
+
+\todo[inline]{Reza: Short paragraph about constexpr}
+
+\todo[inline]{Mention as last sentence that next release will use C++14.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{New and improved tutorial and code gallery programs}
@@
 
-\todo[inline]{What's new here? mention step-49 and step-53 versions written in python.}
+\todo[inline]{Alexander: What's new here? mention step-49 and step-53 versions written in python.}
 

### 2020-05-25 9.2/paper.tex
Update one bib entry.
add to author list

@@
      Wolfgang Bangerth,
-%     Thomas C. Clevenger,
+     Thomas C. Clevenger,
 %     Denis Davydov,
@@
 
-% \author[3]{Thomas~C.~Clevenger}
-% \affil[3]{School of Mathematical and Statistical Sciences,
-%   Clemson University,
-%   Clemson, SC, 29634, USA
-%   {\texttt{\{tcleven, heister\}@clemson.edu}}}
+ \author[3]{Thomas~C.~Clevenger}
+ \affil[3]{School of Mathematical and Statistical Sciences,
+   Clemson University,
+   Clemson, SC, 29634, USA
+   {\texttt{tcleven/heister@clemson.edu}}}
 %
@@
 \author[3]{Timo~Heister}
- \affil[3]{School of Mathematical and Statistical Sciences,
-   Clemson University,
-   Clemson, SC, 29634, USA
-   {\texttt{heister@clemson.edu}}}
 
@@
    \cite{HeltaiMola2015};
- \item For Boundary Elements Computations:
+ \item For boundary element computations:
    \cite{GiulianiMolaHeltai-2018-a};
  \item For \texttt{LinearOperator} and \texttt{PackagedOperation} facilities:
-   \cite{MaierBardelloniHeltai-2016-a,MaierBardelloniHeltai-2016-b}.
+   \cite{MaierBardelloniHeltai-2016-a,MaierBardelloniHeltai-2016-b};
  \item For uses of the \texttt{WorkStream} interface:
@@
      \texttt{MeshWorker::ScratchData} base class, and the
-     \texttt{ParsedConvergenceTable} class: \cite{SartoriGiulianiBardelloni-2018-a}.
+     \texttt{ParsedConvergenceTable} class:
+     \cite{SartoriGiulianiBardelloni-2018-a};
+   \item For uses of the particle functionality in \dealii{}: \cite{GLHPB18}.
 \end{itemize}
-
-\todo[inline]{Need to add the particles paper by Gassmoeller et al.}
 

### 2020-05-25 9.2/paper.tex
Add author

@@
 %     Daniel Garcia-Sanchez,
+    Alexander V. Grayver,
 %     Graham Harper,
@@
 %
+
+\author[6]{Alexander~V.~Grayver}
+\affil[6]{Institute of Geophysics,
+  ETH Zurich,
+  8092 Z\"{u}rich, Switzerland.
+  {\texttt{agrayver@ethz.ch}}}
+
 \author[3]{Timo~Heister}
@@
 Krishnakumar Gopalakrishnan,
-Alexander Grayver,
 Graham Harper,

### 2020-05-25 9.2/paper.tex
Leave a note.

@@
 \todo[inline]{Where did we solve for 12T unknowns? The largest I [Martin] have done is 2.1T unknowns.}
-Figure~\ref{} compares the timings of a simulation (including set up) with the
+Figure~\ref{...} {\color{red}Update reference} compares the timings of a simulation (including set up) with the
 previous release 9.1 and with the current release 9.2.

### 2020-05-25 9.2/paper.tex
Update my dual affiliation.
Add a note.
Added to author list

@@
      Wolfgang Bangerth,
+     Bruno Blais,
      Thomas C. Clevenger,
@@
  \author[2]{Wolfgang~Bangerth}
- \affil[2]{Department of Mathematics, Colorado State University, Fort
+ \affil[2]{Department of Mathematics and Department of Geosciences, Colorado State University, Fort
    Collins, CO 80523-1874, USA.
@@
 
- \author[3]{Thomas~C.~Clevenger}
- \affil[3]{School of Mathematical and Statistical Sciences,
+\author[3]{Bruno Blais}
+\affil[3]{Research Unit for Industrial Flows Processes (URPEI),
+          Polytechnique Montréal,
+          PO Box 6079, Stn Centre-Ville, Montréal, Québec, Canada, H3C 3A7.
+          {\texttt{bruno.blais@polymtl.ca}}}
+
+ \author[4]{Thomas~C.~Clevenger}
+ \affil[4]{School of Mathematical and Statistical Sciences,
    Clemson University,

### 2020-05-26 9.2/paper.tex
Language fix.
Add one missing entry.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Improved performance of symbolic framework}
+\subsection{Improved performance of the symbolic differentiation framework}
 \label{subsec:symbdiff}
@@
 contributed code to this release:\\
-% updated 5/11/2020 MM
+%
+% Uwe Koecher doesn't usually show up in the changelog, but
+% we should make sure he's listed.
+%
+% 9.2: updated 5/11/2020 MM
 \todo[inline]{TODO: remove authors of paper}
@@
 Alexander Knieps,
+Uwe K{\"o}cher,
 Martin Kronbichler,
@@
 W.~Bangerth and T.~Heister were also partially supported by award DMS-1821210.
+\todo[inline]{Need to add FRES project}
 

### 2020-05-26 9.2/paper.tex
Add section on python

@@
   ETH Zurich,
-  8092 Z\"{u}rich, Switzerland.
+  Sonneggstrasse 5, 8092 Z\"{u}rich, Switzerland.
   {\texttt{agrayver@ethz.ch}}}
@@
 
-\todo[inline]{Alexander: What's new here? mention step-49 and step-53 versions written in python.}
+\begin{figure}
+\renewcommand\figurename{Listing}
+
+\centering
+
+\begin{python}
+import PyDealII.Release as dealii
+
+triangulation = dealii.Triangulation('2D')
+triangulation.generate_hyper_shell(center = dealii.Point([0, 0]), 
+                    inner_radius = 0.5, outer_radius = 1.,
+                    n_cells = 0, colorize = True)
+
+triangulation.refine_global(2)
+
+for cell in triangulation.active_cells():
+    cell.material_id = 1 if cell.center().x > 0 else 2
+
+    for face in cell.faces():
+        if face.at_boundary() and\
+           face.boundary_id == 1:
+            cell.refine_flag ='isotropic'
+
+triangulation.execute_coarsening_and_refinement()
+\end{python}
+
+\includegraphics{python_mesh.png}
+
+\caption{Python code that uses \texttt{deal.II}'s Python interface to generate a mesh shown at the bottom (coloured by the cell's material id).}
+\label{python_wrapper}
+\end{figure}
+
+The initial support for Python has come in \texttt{deal.II 9.0}.
+The present release significantly extends the Python interface.
+Specficically, a large number of methods from the C++ classes such as \texttt{Triangulation, CellAccessor, TriaAccessor, Mapping, Manifold, GridTools} can now be invoked from within the Python.
+The accent was made on methods and functions that are widely used at a mesh preparation stage when a mesh is created and parameters related to the boundary, manifold and material identifiers are assigned.
+
+A triangulation object can be read in or created using a number of simplistic geometries offered by the \texttt{GridGenerator}'s functions.
+Further information such as boundary, material and manifold identifiers can be assigned to cells and respective faces or edges (see Listing \ref{python_wrapper}).
+The introspective nature of the Python language makes it easy to infer the list of supported methods from the Python objects, for example by typing \texttt{dir(PyDealII.Release.Triangulation)}.
+
+All triangulations created within the Python are serial.
+Once the mesh is designed, triangulation can be serialized along with the auxiliary information about possible refinement, boundaries, materials and manifolds.
+This object can be easily deserialized within a C++ program for subsequent production runs.
+Although saved triangulations are serial, it is possible to use them for construction of \texttt{parallel::shared, parallel::disrtibuted, and parallel::fullydistributed} (see also Section \ref{subsec:pft}) types of triangulations respecting the refinement and all the auxiliary information.
+
+To facilitate the illustration of the new Python bindings, tutorial programs \href{https://github.com/dealii/dealii/blob/dealii-9.2/examples/step-49/step-49.ipynb}{step-49} and \href{https://github.com/dealii/dealii/blob/dealii-9.2/examples/step-53/step-53.ipynb}{step-53} were replicated in a form of the Jupyter notebooks.
+
+Note that the current Python interface does not yet provide access to the actual \texttt{deal.II}'s finite element machinery, that is classes such as \texttt{DoFHandler, FE\_*, FEValues}, etc.
+It is envisaged that a progress towards this will be made for the next release.
 

### 2020-05-26 9.2/paper.tex
Add experiment for setup times

--- 2020-05-26 9.2/paper.tex
@@
 \pgfplotsset{compat=1.9}
+\definecolor{gnuplot@lightblue}{RGB}{87,181,232}
+\definecolor{gnuplot@green}{RGB}{0,158,115}
+\definecolor{gnuplot@purple}{RGB}{148,0,212}
 
@@
 time for large-scale simulations and to solve a Poisson problem with multigrid
-with \num{12e12} unknowns.
-\todo[inline]{Where did we solve for 12T unknowns? The largest I [Martin] have done is 2.1T unknowns.}
-Figure~\ref{...} {\color{red}Update reference} compares the timings of a simulation (including set up) with the
-previous release 9.1 and with the current release 9.2.
-{\color{red}TODO[Peter/Martin] description of the results}
+with \num{2.1e12} unknowns.
+Figure~\ref{fig:init_costs} compares the timings of simulations of various problem
+sizes (including set up) on 49,152 MPI ranks using a matrix-free solver according
+to~\cite{KronbichlerKormann2019,KronbichlerWall2018} discontinuous elements of
+degree $5$ in a geometric multigrid (GMG) scheme with the previous release 9.1
+and with the current release 9.2. While the times for the V-cycle have shown
+excellent scaling before, many initialization routines have been considerably
+improved, especially the enumeration of unknowns on the multigrid levels and
+the setup of the multigrid transfer.
+
+\pgfplotstableread{
+dofs  grid  distdofs  mfsetup  mgtransfer  smoother  vcycle
+14155776  0.347457  0.02424084  0.03346427  0.0105426  0.0108817  0.0035033
+28311552  0.35134  0.0215024  0.0340963  0.0095201  0.010616  0.0035887
+56623104  0.41577  0.0243593  0.0432297  0.011557  0.011667  0.0041244
+113246208  0.42172  0.0323919  0.0427806  0.012983  0.011834  0.0043853
+226492416  0.43565  0.0404079  0.0421606  0.012867  0.017305  0.0056340
+452984832  0.5055  0.0500101  0.0562977  0.017739  0.022763  0.0061992
+905969664  0.54552  0.065865  0.0641994  0.019445  0.037431  0.0087932
+1811939328  0.5672  0.104909  0.080807  0.025395  0.063548  0.0133252
+3623878656  0.65541  0.165071  0.129717  0.042294  0.12644  0.0218619
+7247757312  0.74521  0.251417  0.221072  0.073314  0.21807  0.0339264
+14495514624  0.84789  0.403463  0.36894  0.11425  0.4091  0.0669798
+28991029248  0.91983  0.52435  0.5363  0.14189  0.79178  0.1322448
+57982058496  1.0339  0.87448  0.93134  0.23863  1.5896  0.2539445
+115964116992  1.2667  1.46016  1.70806  0.44101  3.1409  0.4956625
+231928233984  1.7597  2.58121  3.3741  0.77624  6.2628  0.9857200
+}\tableinit
+
+\pgfplotstableread{
+dofs  grid  distdofs  mfsetup  mgtransfer  smoother  vcycle
+14155776  2.0148  0.9086  0.433886  1.3099  0.010685  0.0035033
+28311552  1.8746  0.91078  0.576777  1.3093  0.010752  0.0035887
+56623104  2.0924  0.98058  0.574183  1.4459  0.011831  0.0041244
+113246208  2.3788  1.22025  0.547492  1.7919  0.011979  0.0043853
+226492416  2.1254  1.50329  0.839219  1.7832  0.017819  0.0056340
+452984832  2.4325  2.10171  0.711708  2.0263  0.023393  0.0061992
+905969664  3.8911  3.58355  1.08478  2.5844  0.045697  0.0087932
+1811939328  4.1803  4.379  0.784804  1.8947  0.061552  0.0133252
+3623878656  2.6654  5.6073  0.833896  1.8912  0.1188  0.0218619
+7247757312  2.7422  7.8463  0.741733  1.8477  0.21291  0.0339264
+14495514624  2.765  9.9848  0.99218  1.8248  0.40665  0.0669798
+28991029248  3.2901  15.0526  1.5124  2.6723  0.81013  0.1322448
+57982058496  3.1927  21.0562  1.72831  2.1998  1.5905  0.2539445
+115964116992  4.1485  31.0087  2.67998  2.9091  3.2009  0.4956625
+231928233984  4.1787  47.195  4.2821  3.9558  6.3362  0.9857200
+}\tableinitold
+
+\begin{figure}
+  \begin{tikzpicture}
+    \begin{loglogaxis}[
+      width=0.51\textwidth,
+      height=0.38\textwidth,
+          xlabel={\# DoFs},
+          ylabel={time [s]},
+          title={9.1 release},
+          xtick={1.4e7,5.7e7,2.3e8,9.1e8,3.6e9,1.4e10,5.8e10,2.3e11},
+          xticklabels={1.4$\times 10^7$,5.7$\times 10^7$,2.3$\times 10^8$,9.1$\times 10^8$,3.6$\times 10^9$,1.4$\times 10^{10}$,5.8$\times 10^{10}$,2.3$\times 10^{11}$},
+          tick label style={font=\scriptsize},
+          xticklabel style={rotate=25,anchor=north east},
+          label style={font=\scriptsize},
+          legend style={font=\scriptsize},
+          %legend to name=legend:scalingMG,
+          %legend cell align=left,
+          %legend columns = 3,
+          ymin=2e-3,ymax=8e1,
+          grid, thick]
+          \addplot table[x index = 0, y index={1}] {\tableinitold};
+          \addplot[red,mark=square*] table[x index = 0, y index={2}] {\tableinitold};
+          \addplot[gnuplot@green,mark=otimes] table[x index = 0, y index={3}] {\tableinitold};
+          \addplot[gnuplot@lightblue,mark=triangle] table[x index = 0, y index={4}] {\tableinitold};
+          \addplot[gnuplot@purple,mark=diamond] table[x index = 0, y index={5}] {\tableinitold};
+          \addplot[black,mark=x] table[x index = 0, y index={6}] {\tableinitold};
+    \end{loglogaxis}
+  \end{tikzpicture}
+  \hfill
+  \begin{tikzpicture}
+    \begin{loglogaxis}[
+      width=0.51\textwidth,
+      height=0.38\textwidth,
+          xlabel={\# DoFs},
+          ylabel={time [s]},
+          title={9.2 release},
+          xtick={1.4e7,5.7e7,2.3e8,9.1e8,3.6e9,1.4e10,5.8e10,2.3e11},
+          xticklabels={1.4$\times 10^7$,5.7$\times 10^7$,2.3$\times 10^8$,9.1$\times 10^8$,3.6$\times 10^9$,1.4$\times 10^{10}$,5.8$\times 10^{10}$,2.3$\times 10^{11}$},
+          tick label style={font=\scriptsize},
+          xticklabel style={rotate=25,anchor=north east}, thick,
+          label style={font=\scriptsize},
+          legend style={font=\scriptsize},
+          legend to name=legend:init,
+          legend cell align=left,
+          legend columns = 6,
+          ymin=2e-3,ymax=8e1,
+          grid]
+          \addplot table[x index = 0, y index={1}] {\tableinit};
+          \addlegendentry{create mesh};
+          \addplot[red,mark=square*] table[x index = 0, y index={2}] {\tableinit};
+          \addlegendentry{enumerate (mg) dofs};
+          \addplot[gnuplot@green,mark=otimes] table[x index = 0, y index={3}] {\tableinit};
+          \addlegendentry{init matrix-free};
+          \addplot[gnuplot@lightblue,mark=triangle] table[x index = 0, y index={4}] {\tableinit};
+          \addlegendentry{init GMG transfer};
+          \addplot[gnuplot@purple,mark=diamond] table[x index = 0, y index={5}] {\tableinit};
+          \addlegendentry{init smoother};
+          \addplot[black,mark=x] table[x index = 0, y index={6}] {\tableinit};
+          \addlegendentry{GMG V-cycle};
+    \end{loglogaxis}
+  \end{tikzpicture}
+  \\
+  \strut\hfill\pgfplotslegendfromname{legend:init}\hfill\strut
+  \caption{Comparison of initialization costs of various data structures in the 9.1 release (left) and the new 9.2 release (right).}
+  \label{fig:init_costs}
+\end{figure}
 

### 2020-05-27 9.2/paper.tex
Leave a note for 9.3
Leave a second note for 9.3
Correctly sort names.
Acknowledge Zhuoran's support.

@@
 %     David Wells
+    Zhuoran Wang,
   },
@@
 
+\author[14]{Zhuoran Wang}
+\affil[14]{Department of Mathematics, Colorado State University, Fort Collins,
+          CO 80523-1874, USA.
+{\texttt{zhrwang@colostate.edu}}}         
 \renewcommand{\labelitemi}{--}
@@
 \item nanoflann \cite{nanoflann}
+  %
+  % nanoflann has been deprecated and will be removed for 9.3. remove
+  % this line in the 9.3 release paper
+  %
 \item NetCDF \cite{rew1990netcdf}
+  %
+  % netcdf has been deprecated and will be removed for 9.3. remove
+  % this line in the 9.3 release paper
+  %
 \item OpenCASCADE \cite{opencascade-web-page}
@@
 
+Z.~Wang was partially
+supported by the National Science Foundation under award OAC-1835673.
+
 The Interdisciplinary Center for Scientific Computing (IWR) at Heidelberg

### 2020-05-28 9.2/paper.tex
add myself to 9.2
brief discreption of the DiscreteTime class is added

@@
 %     Jean-Paul Pelteret,
+    Reza Rastak,
     Bruno Turcksin,
-%     David Wells
+    David Wells
     Zhuoran Wang,
@@
 %
+ \author[14]{Reza Rastak}
+ \affil[14]{Department of Civil and Environmental Engineering,
+   Stanford University,
+   Stanford, CA 94305, USA.
+   {\texttt{rastak@stanford.edu}}}
+
 \author[*1]{Bruno~Turcksin}
 %
-% \author[12]{David Wells}
-% \affil[12]{Department of Mathematics, University of North Carolina,
-%   Chapel Hill, NC 27516, USA.
-%   {\texttt{drwells@email.unc.edu}}}
+\author[12]{David Wells}
+\affil[12]{Department of Mathematics, University of North Carolina,
+  Chapel Hill, NC 27516, USA.
+  {\texttt{drwells@email.unc.edu}}}
 
@@
   solution. This, too, has been addressed in this release.
-  \item \todo[inline]{Reza: Please add a short summary of the
-    \texttt{DiscreteTime} class}
+\item The class \texttt{DiscreteTime} is introduced to provide a more
+  consistent, more readable, and less error-prone approach to control
+  time-stepping algorithms within time-dependent simulations.
+  The mutating interface of this class is designed to be minimal
+  to enforce a number of important programming invariants, reducing
+  the possibility of mistakes in the user code.
+  When time-incrementation is requested within the user code, the class makes
+  sure that time increases by a non-zero step size and the current step
+  number is incremented accordingly.
+  Furthermore, \texttt{DiscreteTime} ensures that the final time step ends
+  precisely on a predefined end time. Therefore, the final step size is
+  automatically lengthened or shortened to accommodate this feature.
+  In addition, the class provides useful access methods which return the step
+  number $n$, the values of the simulation time corresponding to $t_{n-1}$,
+  $t_n$, and $t_{n+1}$, and the step-size values $t_n - t_{n-1}$ and
+  $t_{n+1} - t_n$.
 \item z \todo[inline]{What else? Maybe mention the updated step-12?}
@@
 degree $5$ in a geometric multigrid (GMG) scheme with the previous release 9.1
-and with the current release 9.2. While the times for the V-cycle have shown
-excellent scaling before, many initialization routines have been considerably
+and with the current release 9.2. While the scaling for the V-cycle had been
+very good before, many initialization routines have been considerably
 improved, especially the enumeration of unknowns on the multigrid levels and
@@
   \strut\hfill\pgfplotslegendfromname{legend:init}\hfill\strut
-  \caption{Comparison of initialization costs of various data structures in the 9.1 release (left) and the new 9.2 release (right).}
+  \caption{Comparison of initialization costs of various data structures in the 9.1 release (left) and the new 9.2 release (right) when run on 49,152 MPI ranks.}
   \label{fig:init_costs}
@@
 
-\todo[inline]{Reza: Short paragraph about constexpr}
-
-\todo[inline]{Mention as last sentence that next release will use C++14.}
+Certain types of quantities in a simulation are constants fully known
+at compile time. They can be pre-calculated and fully stored inside the
+compiled program binary in order to avoid unnecessary initialization during
+runtime. This optimization is now enabled for the class templates
+\texttt{Tensor} and \texttt{SymmetricTensor} by qualifying their constructor,
+member functions, and overloaded operators as \texttt{constexpr}.
+For instance, the linear mechanical constitutive model for elastic solids
+uses a constant fourth-order elasticity tensor
+$\mathbb{C} = \lambda \boldsymbol{I} \otimes \boldsymbol{I} + 2 \mu \mathbb{I}$
+which does not depend on the current state of strain.
+This tensor can be statically initialized by defining it as
+\texttt{constexpr SymmetricTensor<4, dim>}.
+As another example, the lattice vectors in a crystal plasticity model are generally
+constant and known during compilation time, enabling their efficient definition as
+\texttt{constexpr Tensor<1, dim>}.
+The capability of defining \texttt{constexpr} variables, functions, and methods
+was introduced by the C++11 standard and was later expanded by the C++14 standard.
+Therefore, the extent of \texttt{constexpr} support in \dealii{} depends on the C++
+standard which is used to compile the library. The next release of \dealii{}
+will fully adopt the features of the C++14 standard.
 
@@
 
-R.~M.~Kynch was supported by the Engineering and Physical Science Research
-Council (EPSRC) UK through grant EP/K023950/1 while working at Swansea
-University, UK 2013-2015 where the majority of his contribution was
-undertaken.
-
 M.~Maier was partially supported by ARO MURI Award No. W911NF-14-0247 and

### 2020-05-28 9.2/paper.tex
Reza added to the list of authors
Information about constexpr capabilities of dealii added

--- 2020-05-26 9.2/paper.tex
@@
     \texttt{DiscreteTime} class}
+\item Certain types of quantities in a simulation are constants fully known
+  at compile time. They can be pre-calculated and fully stored inside the
+  compiled program binary in order to avoid unnecessary initialization during
+  runtime. This optimization is now enabled for the class templates
+  \texttt{Tensor} and \texttt{SymmetricTensor} by qualifying their constructor,
+  member functions, and overloaded operators as \texttt{constexpr}.
+  For instance, the linear mechanical constitutive model for elastic solids
+  uses a constant fourth-order elasticity tensor
+  $\mathbb{C} = \lambda \boldsymbol{I} \otimes \boldsymbol{I} + 2 \mu \mathbb{I}$
+  which does not depend on the current state of strain.
+  This tensor can be statically initialized by defining it as
+  \texttt{constexpr SymmetricTensor<4, dim>}.
+  As another example, the lattice vectors in a crystal plasticity model are generally
+  constant and known during compilation time, enabling their efficient definition as
+  \texttt{constexpr Tensor<1, dim>}.
 \item z \todo[inline]{What else? Maybe mention the updated step-12?}

### 2020-05-28 9.2/paper.tex

--- 2020-05-28 9.2/paper.tex
@@
 %     Jean-Paul Pelteret,
+    Reza Rastak,
     Bruno Turcksin,
-%     David Wells
+    David Wells
+    Zhuoran Wang,
   },
@@
 %
+ \author[14]{Reza Rastak}
+ \affil[14]{Department of Civil and Environmental Engineering,
+   Stanford University,
+   Stanford, CA 94305, USA.
+   {\texttt{rastak@stanford.edu}}}
+
 \author[*1]{Bruno~Turcksin}
 %
-% \author[12]{David Wells}
-% \affil[12]{Department of Mathematics, University of North Carolina,
-%   Chapel Hill, NC 27516, USA.
-%   {\texttt{drwells@email.unc.edu}}}
-
+\author[12]{David Wells}
+\affil[12]{Department of Mathematics, University of North Carolina,
+  Chapel Hill, NC 27516, USA.
+  {\texttt{drwells@email.unc.edu}}}
+
+\author[14]{Zhuoran Wang}
+\affil[14]{Department of Mathematics, Colorado State University, Fort Collins,
+          CO 80523-1874, USA.
+{\texttt{zhrwang@colostate.edu}}}         
 \renewcommand{\labelitemi}{--}

### 2020-05-28 9.2/paper.tex

@@
     \texttt{DiscreteTime} class}
-\item Certain types of quantities in a simulation are constants fully known
-  at compile time. They can be pre-calculated and fully stored inside the
-  compiled program binary in order to avoid unnecessary initialization during
-  runtime. This optimization is now enabled for the class templates
-  \texttt{Tensor} and \texttt{SymmetricTensor} by qualifying their constructor,
-  member functions, and overloaded operators as \texttt{constexpr}.
-  For instance, the linear mechanical constitutive model for elastic solids
-  uses a constant fourth-order elasticity tensor
-  $\mathbb{C} = \lambda \boldsymbol{I} \otimes \boldsymbol{I} + 2 \mu \mathbb{I}$
-  which does not depend on the current state of strain.
-  This tensor can be statically initialized by defining it as
-  \texttt{constexpr SymmetricTensor<4, dim>}.
-  As another example, the lattice vectors in a crystal plasticity model are generally
-  constant and known during compilation time, enabling their efficient definition as
-  \texttt{constexpr Tensor<1, dim>}.
 \item z \todo[inline]{What else? Maybe mention the updated step-12?}

### 2020-05-28 9.2/paper.tex
Fix affiliations.

@@
 
-\author[3]{Timo~Heister}
-
-\author[8]{Luca~Heltai}
-\affil[8]{SISSA,
+\author[4]{Timo~Heister}
+
+\author[7]{Luca~Heltai}
+\affil[7]{SISSA,
    International School for Advanced Studies,
@@
 
- \author[9]{Martin~Kronbichler}
- \affil[9]{Institute for Computational Mechanics,
+ \author[8]{Martin~Kronbichler}
+ \affil[8]{Institute for Computational Mechanics,
    Technical University of Munich,
@@
    {\texttt{kronbichler/munch@lnm.mw.tum.de}}}
-%
-% \author[10]{Ross~Maguire~Kynch}
-% \affil[10]{Zienkiewicz Centre for Computational Engineering,
-%   College of Engineering, Swansea University,
-%   Bay Campus, Fabian Way, Swansea SA1 8EN, Wales, UK.
-%   {\texttt{rkynch@gmail.com}}}
-%
-\author[11]{Matthias~Maier}
-\affil[11]{Department of Mathematics,
+
+\author[9]{Matthias~Maier}
+\affil[9]{Department of Mathematics,
   Texas A\&M University,
@@
 
-  \author[9,13]{Peter Munch}
-
- \affil[13]{Institute of Materials Research, Materials Mechanics,
+\author[8,10]{Peter Munch}
+ \affil[10]{Institute of Materials Research, Materials Mechanics,
  Helmholtz-Zentrum Geesthacht,
@@
 %
- \author[14]{Reza Rastak}
- \affil[14]{Department of Civil and Environmental Engineering,
+ \author[11]{Reza Rastak}
+ \affil[11]{Department of Civil and Environmental Engineering,
    Stanford University,
@@
 
-\author[14]{Zhuoran Wang}
-\affil[14]{Department of Mathematics, Colorado State University, Fort Collins,
+\author[13]{Zhuoran Wang}
+\affil[13]{Department of Mathematics, Colorado State University, Fort Collins,
           CO 80523-1874, USA.

### 2020-05-28 9.2/paper.tex
add description for step-47.

@@
           CO 80523-1874, USA.
-{\texttt{zhrwang@colostate.edu}}}         
+{\texttt{zhrwang@math.colostate.edu}}}
 \renewcommand{\labelitemi}{--}
@@
 \begin{itemize}
-\item \texttt{step-47}
-\todo[inline]{Zhuoran to write}
+\item \texttt{step-47} solves the biharmonic equation with the
+  clamped boundary condition. This program is based on the $C^0$ interior
+  penalty method for fourth order problems. In order to overcome
+  shortcomings of classical approaches, this method uses $C^0$ Lagrange finite
+  elements and introduces ``jump`` and ``average`` on interfaces of elements.
+  The resulting system is obtained through integration by parts, symmetrization,
+  and stabilization. Appropriate choices for the penalty parameter are
+  discussed based on numerical results.
+
 \item \texttt{step-50}

### 2020-05-28 9.2/paper.tex
Work through the first half of the paper.

--- 2020-05-28 9.2/paper.tex
@@
 \dealii version 9.2.0 was released May XYZ, 2020.
+\todo[inline]{Need to update date.}
 This paper provides an
@@
 \item \dealii{} had decent support for solving complex-valued problems
-  (e.g., ones in quantum mechanics -- like the example covered by the
+  for a while already
+  (e.g., ones in quantum mechanics -- like the equation used in the
   \texttt{step-58} tutorial program covered below -- or for
-  time-harmonic problems) for a while already. However, there were two
+  time-harmonic problems). However, there were two
   areas in which support was missing. First, the UMFPACK direct solver
@@
   solution. This, too, has been addressed in this release.
-\item The class \texttt{DiscreteTime} is introduced to provide a more
+\item The new \texttt{DiscreteTime} class provides a more
   consistent, more readable, and less error-prone approach to control
   time-stepping algorithms within time-dependent simulations.
-  The mutating interface of this class is designed to be minimal
+  The interface of this class is designed to be minimal
   to enforce a number of important programming invariants, reducing
   the possibility of mistakes in the user code.
-  When time-incrementation is requested within the user code, the class makes
-  sure that time increases by a non-zero step size and the current step
-  number is incremented accordingly.
-  Furthermore, \texttt{DiscreteTime} ensures that the final time step ends
-  precisely on a predefined end time. Therefore, the final step size is
-  automatically lengthened or shortened to accommodate this feature.
-  In addition, the class provides useful access methods which return the step
-  number $n$, the values of the simulation time corresponding to $t_{n-1}$,
-  $t_n$, and $t_{n+1}$, and the step-size values $t_n - t_{n-1}$ and
-  $t_{n+1} - t_n$.
+  An example is that \texttt{DiscreteTime} ensures that the final time step ends
+  precisely on a predefined end time, automatically
+  lengthening or shortening to final time step.
 \item z \todo[inline]{What else? Maybe mention the updated step-12?}
@@
 %
-In addition to these changes, the changelog lists more than 200 other
+The changelog lists more than 240 other
 features and bugfixes.
@@
 that will be discussed in this section.
-
 It of course also contains a
@@
 
-By release 9.1, all triangulation classes of \texttt{deal.II} have in common that the coarse grid is shared by
-all processes and the actual mesh used for computations is constructed by repeated
-refinement. However, this has its limitations in industrial applications where, often, the mesh comes
+Previously, all triangulation classes of \texttt{deal.II} had in common that the coarse grid is replicated by
+all processes in a parallel environment, and the actual mesh used for computations is constructed by repeated
+refinement. However, this has its limitations in industrial
+applications, where frequently the mesh comes
 from an external mesh generator in the form of a file that already contains millions
@@
 
-The new class \texttt{parallel::fullydistributed::Triangulation} targets this issue
+The new \texttt{parallel::fullydistributed::Triangulation} class targets this issue
 by distributing also the coarse grid. Such
 a triangulation can be created by providing to each process a \texttt{Triangulation\-De\-scrip\-tion::Description} struct, containing
-1) the relevant data to construct the local part of the coarse grid, 2) the
-translation of the local coarse-cell IDs to globally unique IDs, 3) the hierarchy
-of mesh refinements, and 4) the owner of the cells on the active mesh level as well
-as on the multigrid levels. For the current release, triangulations is set up
-this way cannot be adaptively refined after construction, which is planned to be
-improved for the next release.
+(i) the relevant data to construct the local part of the coarse grid, (ii) the
+translation of the local coarse-cell IDs to globally unique IDs, (iii) the hierarchy
+of mesh refinement steps, and (iv) the owner of the cells on the active mesh level as well
+as on the multigrid levels. For the current release, triangulations set up
+this way cannot be adaptively refined after construction, though we plann to
+improve this for the next release.
 
@@
 
-Large-scale simulations with 304,128 cores have revealed bottlenecks in release
-9.1 during initialization of a number of distributed data structures due to the usage of expensive collective operations
+Large-scale simulations with up to 304,128 cores have revealed bottlenecks in release
+9.1 during initialization of a number of distributed data structures, due to the usage of expensive collective operations
 like \texttt{MPI\_Allgather()} and \texttt{MPI\_\allowbreak
   Alltoall()}. Typical examples are the
-pre-computation of those indices of the vector entries (or
+pre-computation of the indices of those vector entries (or
 other linear index ranges) owned by
@@
 with \num{2.1e12} unknowns.
-Figure~\ref{fig:init_costs} compares the timings of simulations of various problem
-sizes (including set up) on 49,152 MPI ranks using a matrix-free solver according
-to~\cite{KronbichlerKormann2019,KronbichlerWall2018} discontinuous elements of
-degree $5$ in a geometric multigrid (GMG) scheme with the previous release 9.1
-and with the current release 9.2. While the scaling for the V-cycle had been
+Figure~\ref{fig:init_costs} compares timings of simulations of various problem
+sizes (including set up) on 49,152 MPI ranks using a matrix-free
+solver~\cite{KronbichlerKormann2019,KronbichlerWall2018}; this solver uses discontinuous elements of
+degree $5$ in a geometric multigrid (GMG) scheme. The comparison between the previous release 9.1
+and the current release 9.2 shows that while the scaling for the V-cycle had been
 very good before, many initialization routines have been considerably
@@
   \strut\hfill\pgfplotslegendfromname{legend:init}\hfill\strut
-  \caption{Comparison of initialization costs of various data structures in the 9.1 release (left) and the new 9.2 release (right) when run on 49,152 MPI ranks.}
+  \caption{\it Comparison of initialization costs of various data structures in the 9.1 release (left) and the new 9.2 release (right) when run on 49,152 MPI ranks.}
   \label{fig:init_costs}
@@
 
-The new code has been also applied to solve problems with adaptively refined
-meshes with more than \num{4e9} unknowns. {\color{red}TODO[Timo]}
-
-\todo[inline]{We did solve adaptive problems with more than 4B unknowns before, but some were broken for a while during this release stage due to the consensus algorithms.
-But we did not test systems or block vectors before.
-But we should not make it sound too negative.}
+\todo[inline]{Timo: Do you want to say something about that the new code has been also applied to solve problems with adaptively refined
+meshes with more than \num{4e9} unknowns?}
+
+\todo[inline]{Timo says: We did solve adaptive problems with more than
+  4B unknowns before, but some were broken for a while during this
+  release stage due to the consensus algorithms. But we did not test
+  systems or block vectors before. But we should not make it sound too
+  negative.}
+
+\todo[inline]{Wolfgang says: Should we just not mention any of this at
+  all? The section stands well on its own.}
 
@@
 
-With the previous release, \dealii supports $hp$-adaptive finite element methods
+Since the previous release, \dealii has had support for $hp$-adaptive finite element methods
 on distributed memory systems \cite{dealII91}. We implemented the bare functionality
@@
 
-The interface is now as simple to use as the one for pure adaptive mesh refinement.
+The interface is now as simple to use as the one for $h$-adaptive mesh refinement.
 Consider the following (incomplete) listing as an example: We estimate both error and
@@
 cells with the highest and lowest errors for refinement and coarsening, respectively
-(here: 30\% / 3\%). From those cells listed for adaptation, we specify a subset of them
-for $h$- and $p$-adaptation (here: 10\% / 90\%).
+(here: 30\%/3\%). From those cells listed for adaptation, we designate a subset
+for $h$- and $p$-adaptation (here: 10\%/90\%).
 \begin{c++}
@@
 The former relies on knowing an estimate for the upper error bound \cite[Thm.~3.4]{BabuskaSuri1990}.
-For successive refinements, we can predict how the error will change on basis of
+For successive refinements, we can predict how the error will change
+based on
 current error estimates and adaptation flags. In the next refinement cycle, these
 predicted error estimates allow us to decide whether the choice of adaptation in
-the previous cycle was justified, and provide the foundation for it in the next
-one \cite{MelenkWohlmuth2001}.
+the previous cycle was justified, and provide a criterion for the choice in the next
+cycle \cite{MelenkWohlmuth2001}.
 

### 2020-05-28 9.2/paper.tex
Add GPU support section

@@
    TN 37831, USA.
-   \texttt{arndtd@ornl.gov}}
+   \texttt{arndtd/turcksinbr@ornl.gov}}
 
@@
 as on the multigrid levels. For the current release, triangulations set up
-this way cannot be adaptively refined after construction, though we plann to
+this way cannot be adaptively refined after construction, though we plan to
 improve this for the next release.
@@
 
-\begin{itemize}
-\item overlapping of computation and communication in the case of CUDA-aware MPI
-\end{itemize}
+For this release, the most noteworthy improvements for the GPU support are the
+simplification of the kernel written by user, improvement of the error messages,
+and overlapping of computation and communication when using CUDA-aware MPI with
+the matrix-free framework. In order to simplify user code, we know recompute the
+local degree of freedom and quadrature point indices instead of having the user
+keeping track of them. A new option to overlap computation and communication is
+now available when using CUDA-aware MPI and matrix-free. The underlying idea is
+that only the degrees of freedom (dofs) on the boundary of the local domain
+require communication. Before evaluating the matrix-free operator at these
+degrees of freedom, we need to communicate the ghosted dofs from other
+processors.  Similarly once the operator has been evaluated, we need to update
+the resulting global vector with values from other processors. The strategy that
+we are using consist in splitting the dofs in three groups: one group of dofs
+that are on the local boundary and two groups each owning half of the interior
+dofs. When evaluating the matrix-free operator, we start the MPI communication
+to get the ghosted dofs and evaluate the operator on one of the two interior
+dofs group. When the evaluation is done, we wait for the MPI to be over and
+then, evaluate the operator on the boundary dofs group. When this evaluation is
+over, we start the communication to update the global vector, we evaluate the
+operator on the last of group of dofs, and finally wait for the MPI
+communication to finish.
 
@@
 The present release significantly extends the Python interface.
-Specficically, a large number of methods from the C++ classes such as \texttt{Triangulation, CellAccessor, TriaAccessor, Mapping, Manifold, GridTools} can now be invoked from within the Python.
+Specifically, a large number of methods from the C++ classes such as \texttt{Triangulation, CellAccessor, TriaAccessor, Mapping, Manifold, GridTools} can now be invoked from Python.
 The accent was made on methods and functions that are widely used at a mesh preparation stage when a mesh is created and parameters related to the boundary, manifold and material identifiers are assigned.

### 2020-05-28 9.2/paper.tex

@@
 (here: 30\%/3\%). From those cells listed for adaptation, we designate a subset
-for $h$- and $p$-adaptation (here: 10\%/90\%).
+for $h$- and $p$-adaptation (here: 10\%/90\%). The parameters of the corresponding
+\texttt{hp::Refinement} function specify the fraction of cells to be $p$-adapted from
+those subsets flagged for refinement and coarsening, respectively.
 \begin{c++}
-Vector<float> estimated_error_per_cell (n_active_cells);
+Vector<float> estimated_error_per_cell(n_active_cells);
 KellyErrorEstimator::estimate(
@@
 
-Vector<float> estimated_smoothness_per_cell (n_active_cells);
+Vector<float> estimated_smoothness_per_cell(n_active_cells);
 SmoothnessEstimator::Legendre::coefficient_decay(
@@
   dof_handler, estimated_smoothness_per_cell, 0.9, 0.9);
-hp::Refinement::choose_p_over_h (dof_handler);
+hp::Refinement::choose_p_over_h(dof_handler);
 
@@
 The former relies on knowing an estimate for the upper error bound \cite[Thm.~3.4]{BabuskaSuri1990}.
-For successive refinements, we can predict how the error will change
-based on
+For successive refinements, we can predict how the error will change based on
 current error estimates and adaptation flags. In the next refinement cycle, these

### 2020-05-28 9.2/paper.tex
add Ignacio

@@
 
+ \author[**12]{Ignacio Tomas}
+ \affil[12]{Sandia National Laboratories, Org 1442,
+ P.O. Box 5800, MS 1320, Albuquerque, NM, 87185-1320,
+   \texttt{itomas@sandia.gov}}
+
 \author[*1]{Bruno~Turcksin}
 %
-\author[12]{David Wells}
-\affil[12]{Department of Mathematics, University of North Carolina,
+\author[13]{David Wells}
+\affil[13]{Department of Mathematics, University of North Carolina,
   Chapel Hill, NC 27516, USA.
@@
 
-\author[13]{Zhuoran Wang}
-\affil[13]{Department of Mathematics, Colorado State University, Fort Collins,
+\author[14]{Zhuoran Wang}
+\affil[14]{Department of Mathematics, Colorado State University, Fort Collins,
           CO 80523-1874, USA.
-{\texttt{zhrwang@colostate.edu}}}         
+{\texttt{zhrwang@math.colostate.edu}}}
 \renewcommand{\labelitemi}{--}
@@
   DOE Public Access Plan (http://energy.gov/downloads/doe-public-access-plan).}
+
+\footnotetext{%
+  $^{**}$ Sandia National Laboratories is a multimission laboratory managed and
+  operated by National Technology \& Engineering Solutions of Sandia, LLC, a
+  wholly owned subsidiary of Honeywell International Inc., for the U.S.
+  Department of Energy's National Nuclear Security Administration under contract
+  DE-NA0003525. This document describes objective technical results and analysis.
+  Any subjective views or opinions that might be expressed in the paper do not
+  necessarily represent the views of the U.S. Department of Energy or the United
+  States Government.}
 

### 2020-05-28 9.2/paper.tex
Edit whatever text we have.

--- 2020-05-28 9.2/paper.tex
@@
 
-The class \texttt{VectorizedArray<Number>} is a key ingredient for the high
+The class \texttt{VectorizedArray<Number>} is a key component to achieve the high
 node-level performance of the matrix-free algorithms in deal.II~\cite{KronbichlerKormann2012, KronbichlerKormann2019}. It is a wrapper
@@
 The length of the  vector can now be queried by  \texttt{VectorizedArray::size()} and its underlying number type by \texttt{VectorizedArray::value\_type}.
-Furthermore, the \texttt{VectorizedArray} class now supports range-based iterations over its entries.
-
-Up to release 9.1, the
-vector length $n$ has been set at compile time of the library to the highest
-possible value supported by the given processor architecture.
+Furthermore, the \texttt{VectorizedArray} class now supports range-based iteration over its entries.
+
+In previous \dealii{} releases, the
+vector length was set at compile time of the library to match the highest
+value supported by the given processor architecture.
 Now, a second optional template argument
-\texttt{VectorizedArray<Number, size>} can be given with \texttt{size} explicitly controlling
+can be specified as \texttt{VectorizedArray<Number, size>}, where \texttt{size} explicitly controls
 the vector length within the capabilities of a particular instruction set.
-A full list of supported
-vector lengths is presented in Table~\ref{tab:vectorizedarray}.
-
-To account for the variable-size \texttt{VectorizedArray} class, all matrix-free related classes (like \texttt{MatrixFree} and \texttt{FEEvaluation})
-have been extended with a new optional template argument specifying the
-\texttt{VectorizedArrayType}. This allows users to select the vector length/ISA and,
-as a consequence, the number of cells to be processed at once directly in their applications:
-The deal.II-based
+(A full list of supported
+vector lengths is presented in Table~\ref{tab:vectorizedarray}.)
+This allows users to select the vector length/ISA and,
+as a consequence, the number of cells to be processed at once in matrix-free
+operator evaluations. For example, the deal.II-based
 library \texttt{hyper.deal}~\cite{munch2020hyperdeal}, which solves the 6D Vlasov--Poisson equation with high-order
@@
 \begin{table}
-\caption{Supported vector lengths of the class \texttt{VectorizedArray} and
+\caption{\it Supported vector lengths of the class \texttt{VectorizedArray} and
 the corresponding instruction-set-architecture extensions. }\label{tab:vectorizedarray}
@@
 
-\caption{Comparison of relevant SIMD-related classes in deal.II and \texttt{C++23}.}\label{tab:simd}
+\caption{\it Comparison of relevant SIMD-related classes in deal.II and \texttt{C++23}.}\label{tab:simd}
 \centering
@@
 
-Furthermore, the new interfaces enable using any data structure
-\texttt{VectorizedArrayType} as long as it supports required
-functionalities like \texttt{size()} or \texttt{value\_type}. This prepares
-for the \texttt{C++23} feature \texttt{std::simd} that will be enabled in the future.
-Table~\ref{tab:simd} gives a comparison of the deal.II-specific SIMD classes and
-the equivalent \texttt{C++23} classes. Finally, this change also prepares for specialized
+The new interface of \texttt{VectorizedArray} also enables replacement
+by any type with a matching interface. Specifically, this prepares
+\dealii{} for the \texttt{std::simd} class that is slated to become
+part of the \texttt{C++23} standard.
+Table~\ref{tab:simd} compares the deal.II-specific SIMD classes and
+the equivalent \texttt{C++23} classes. These changes also prepare for specialized
 code paths exploiting
-vectorization within an element~\cite{KronbichlerKormann2019} in the future.
+vectorization within an element, see~\cite{KronbichlerKormann2019}.
 
@@
 
-For this release, the most noteworthy improvements for the GPU support are the
-simplification of the kernel written by user, improvement of the error messages,
+For this release, the most noteworthy improvements in GPU support are the
+simplification of the kernel written by user, improvement of error messaging,
 and overlapping of computation and communication when using CUDA-aware MPI with
-the matrix-free framework. In order to simplify user code, we know recompute the
+the matrix-free framework.
+
+In order to simplify user code, we now recompute
 local degree of freedom and quadrature point indices instead of having the user
@@
 now available when using CUDA-aware MPI and matrix-free. The underlying idea is
-that only the degrees of freedom (dofs) on the boundary of the local domain
+that only the degrees of freedom (DoFs) on the boundary of the local domain
 require communication. Before evaluating the matrix-free operator at these
-degrees of freedom, we need to communicate the ghosted dofs from other
+degrees of freedom, we need to communicate ghost DoFs from other
 processors.  Similarly once the operator has been evaluated, we need to update
 the resulting global vector with values from other processors. The strategy that
-we are using consist in splitting the dofs in three groups: one group of dofs
+we are now using consists in splitting the DoFs into three groups: one group of DoFs
 that are on the local boundary and two groups each owning half of the interior
-dofs. When evaluating the matrix-free operator, we start the MPI communication
-to get the ghosted dofs and evaluate the operator on one of the two interior
-dofs group. When the evaluation is done, we wait for the MPI to be over and
-then, evaluate the operator on the boundary dofs group. When this evaluation is
-over, we start the communication to update the global vector, we evaluate the
-operator on the last of group of dofs, and finally wait for the MPI
+DoFs. When evaluating the matrix-free operator, we start the MPI communication
+to get the ghosted DoFs and evaluate the operator on one of the two interior
+DoFs group. When the evaluation is done, we wait for the MPI
+communication to be over, and
+then evaluate the operator on the boundary DoFs group. When this evaluation is
+completed, we start the communication to update the global vector, we evaluate the
+operator on the last group of DoFs, and finally wait for the MPI
 communication to finish.
@@
 Certain types of quantities in a simulation are constants fully known
-at compile time. They can be pre-calculated and fully stored inside the
-compiled program binary in order to avoid unnecessary initialization during
-runtime. This optimization is now enabled for the class templates
-\texttt{Tensor} and \texttt{SymmetricTensor} by qualifying their constructor,
-member functions, and overloaded operators as \texttt{constexpr}.
-For instance, the linear mechanical constitutive model for elastic solids
+at compile time. They can be pre-calculated and stored in the
+compiled executable in order to avoid unnecessary initialization during
+runtime. C++11 and later standards enables such computations by
+marking variables and functions with the \texttt{constexpr} keyword.
+
+This optimization is now enabled for the class templates
+\texttt{Tensor} and \texttt{SymmetricTensor}.
+For instance, the linear mechanical constitutive model for isotropic elastic solids
 uses a constant fourth-order elasticity tensor
@@
 which does not depend on the current state of strain.
-This tensor can be statically initialized by defining it as
+This tensor can now be statically initialized by defining it as
 \texttt{constexpr SymmetricTensor<4, dim>}.
@@
 \texttt{constexpr Tensor<1, dim>}.
-The capability of defining \texttt{constexpr} variables, functions, and methods
-was introduced by the C++11 standard and was later expanded by the C++14 standard.
-Therefore, the extent of \texttt{constexpr} support in \dealii{} depends on the C++
-standard which is used to compile the library. The next release of \dealii{}
-will fully adopt the features of the C++14 standard.
+
+Declaring variables, functions, and methods as \texttt{constexpr}
+is a C++11 feature that was later expanded by the C++14 standard.
+Thus, parts of the \texttt{constexpr} support in \dealii{} depend on the C++
+standard supported by the compiler used to install the library.
+
+The next release of \dealii{} will require compiler support for the C++14 standard.
 
@@
 \begin{itemize}
-\item \texttt{step-47} solves the biharmonic equation with the
-  clamped boundary condition. This program is based on the $C^0$ interior
-  penalty method for fourth order problems. In order to overcome
+\item \texttt{step-47} is a new program that solves the biharmonic
+  equation $\Delta^2 u = f$ with
+  ``clamped'' boundary condition given by $u=g, \partial u/\partial
+  n=h$. This program is based on the $C^0$ interior
+  penalty ($C^0$IP) method for fourth order problems
+  \cite{Brenner2005}. In order to overcome
   shortcomings of classical approaches, this method uses $C^0$ Lagrange finite
-  elements and introduces ``jump`` and ``average`` on interfaces of elements.
-  The resulting system is obtained through integration by parts, symmetrization,
-  and stabilization. Appropriate choices for the penalty parameter are
-  discussed based on numerical results.
+  elements and introduces ``jump'' and ``average'' operators on
+  interfaces of elements that penalize the jump of the gradient of the
+  solution in order to obtain convergence to the $H^2$-regular
+  solution of the equation.
+
+  The $C^0$IP approach is a modern alternative to classical methods that
+  use $C^1$-conforming elements such as the Argyris
+  element, the Clough-Tocher element and others, all developed in the
+  late 1960s. From a twenty-first century perspective, they can only be
+  described as bizarre in their construction. They are also exceedingly
+  cumbersome to implement if one wants to use general meshes. As a
+  consequence, they have largely fallen out of favor and deal.II currently
+  does not contain implementations of these shape functions.
 
@@
   an appropriate form for the potential $V=V(\mathbf x)$. The
-  tutorial program focused on two specific aspects for which this
+  tutorial program focuses on two specific aspects for which this
   equation serves as an excellent test case: (i) Solving
@@
   real and imaginary parts (as \texttt{step-29} does, for
-  example). (ii) Using operator splitting techniques. The equation is
+  example). (ii)~Using operator splitting techniques. The equation is
   a particularly good test case for this technique because the only
   nonlinear term, $\kappa |\psi|^2 \psi$, does not contain any
-  derivatives and consequently forms an ODE to be solved at each time
+  derivatives and consequently forms an ODE at each node point, to be
+  solved in each time 
   step in an operator splitting scheme (for which, furthermore, there
@@
 \begin{figure}
-\renewcommand\figurename{Listing}
-
 \centering
-
+\includegraphics[scale=0.5]{python_mesh.png}
+\caption{\it The mesh generated by the Python code shown in the main
+  text. Cells are colored by material id.}
+\label{fig:pymesh}
+\end{figure}
+
+Initial support for Python has existed in \dealii{} since version
+9.0. The present release significantly extends the Python
+interface. Specifically, a large number of methods from classes such
+as \texttt{Triangulation, CellAccessor, TriaAccessor, Mapping,
+  Manifold, GridTools} can now be invoked from Python. We have focused
+on methods and functions that are widely used when a mesh is created
+and parameters related to the boundary, manifold, and material
+identifiers are assigned. The following listing gives an idea of how
+such code looks:
 \begin{python}
@@
 \end{python}
-
-\includegraphics{python_mesh.png}
-
-\caption{Python code that uses \texttt{deal.II}'s Python interface to generate a mesh shown at the bottom (coloured by the cell's material id).}
-\label{python_wrapper}
-\end{figure}
-
-The initial support for Python has come in \texttt{deal.II 9.0}.
-The present release significantly extends the Python interface.
-Specifically, a large number of methods from the C++ classes such as \texttt{Triangulation, CellAccessor, TriaAccessor, Mapping, Manifold, GridTools} can now be invoked from Python.
-The accent was made on methods and functions that are widely used at a mesh preparation stage when a mesh is created and parameters related to the boundary, manifold and material identifiers are assigned.
-
-A triangulation object can be read in or created using a number of simplistic geometries offered by the \texttt{GridGenerator}'s functions.
-Further information such as boundary, material and manifold identifiers can be assigned to cells and respective faces or edges (see Listing \ref{python_wrapper}).
-The introspective nature of the Python language makes it easy to infer the list of supported methods from the Python objects, for example by typing \texttt{dir(PyDealII.Release.Triangulation)}.
-
-All triangulations created within the Python are serial.
-Once the mesh is designed, triangulation can be serialized along with the auxiliary information about possible refinement, boundaries, materials and manifolds.
-This object can be easily deserialized within a C++ program for subsequent production runs.
-Although saved triangulations are serial, it is possible to use them for construction of \texttt{parallel::shared, parallel::disrtibuted, and parallel::fullydistributed} (see also Section \ref{subsec:pft}) types of triangulations respecting the refinement and all the auxiliary information.
-
-To facilitate the illustration of the new Python bindings, tutorial programs \href{https://github.com/dealii/dealii/blob/dealii-9.2/examples/step-49/step-49.ipynb}{step-49} and \href{https://github.com/dealii/dealii/blob/dealii-9.2/examples/step-53/step-53.ipynb}{step-53} were replicated in a form of the Jupyter notebooks.
-
-Note that the current Python interface does not yet provide access to the actual \texttt{deal.II}'s finite element machinery, that is classes such as \texttt{DoFHandler, FE\_*, FEValues}, etc.
-It is envisaged that a progress towards this will be made for the next release.
+The mesh that results from this code is shown in Fig.~\ref{fig:pymesh}.
+
+All triangulations created from within Python are serial. However,
+once the mesh is designed, the triangulation can be serialized along
+with the auxiliary information about possible refinement, boundaries,
+materials and manifolds. This object can be easily deserialized within
+a C++ program for subsequent production runs. Furthermore, such a
+serialized triangulation can also be used in the construction of
+\texttt{parallel::shared}, \texttt{parallel::disrtibuted}, and
+\texttt{parallel::fullydistributed} triangulations (see also Section~\ref{subsec:pft}).
+
+To facilitate the illustration of the new Python bindings, tutorial programs \href{https://github.com/dealii/dealii/blob/dealii-9.2/examples/step-49/step-49.ipynb}{step-49} and \href{https://github.com/dealii/dealii/blob/dealii-9.2/examples/step-53/step-53.ipynb}{step-53} were replicated as Jupyter notebooks.
+
+The introspective nature of the Python language makes it easy to infer
+the list of supported methods from the Python objects, for example by
+typing \texttt{dir(PyDealII.Release.Triangulation)}. 
+The current Python interface does not yet provide access to \dealii{}'s finite element machinery, i.e., classes such as \texttt{DoFHandler, FE\_*, FEValues}, etc.
 
@@
 Subsection~\ref{subsec:performance}, deal.II by default no longer
-stores information for all processes on all processes, but only the
-local information or the locally-relevant information. On the other
+stores information for all processes on all processes, but only
+local or locally-relevant information. On the other
 hand, if necessary, global information can still be computed using,
@@
 the current version of the library, \textbf{please reference the present
-document}. For up to date information and a bibtex entry for this document
-see:
+document}. For up to date information and a bibtex entry
+see
 \begin{center}
@@
 following:
+\begin{multicols}{2}
+  \vspace*{-36pt}
   \begin{itemize}
@@
    \cite{HeltaiBangerthKronbichlerMola2019};
+   \vfill\null  \columnbreak
  \item For integration with CAD files and tools:
@@
    \cite{SartoriGiulianiBardelloni-2018-a};
-   \item For uses of the particle functionality in \dealii{}: \cite{GLHPB18}.
+ \item For uses of the particle functionality in \dealii{}:
+   \cite{GLHPB18}.
+   \vfill\null  
 \end{itemize}
+\end{multicols}
 

### 2020-05-28 9.2/paper.tex
Inline a figure.
Typos and support

--- 2020-05-28 9.2/paper.tex
@@
 all processes in a parallel environment, and the actual mesh used for computations is constructed by repeated
-refinement. However, this has its limitations in industrial
-applications, where frequently the mesh comes
+refinement. However, this has its limitations in
+applications where the mesh comes
 from an external mesh generator in the form of a file that already contains millions
@@
 and overlapping of computation and communication when using CUDA-aware MPI with
-the matrix-free framework. In order to simplify user code, we know recompute the
+the matrix-free framework. In order to simplify user code, we now recompute the
 local degree of freedom and quadrature point indices instead of having the user
@@
 
+A.~V.~Grayver was partially supported by the European Space Agency Swarm DISC program.
+
 M.~Maier was partially supported by ARO MURI Award No. W911NF-14-0247 and

### 2020-05-28 9.2/paper.tex
Resolve a todo.

--- 2020-05-28 9.2/paper.tex
@@
 \usepackage{siunitx}
+
+\usepackage{floatflt}
 
@@
 
-\begin{figure}
-\centering
-\includegraphics[scale=0.5]{python_mesh.png}
-\caption{\it The mesh generated by the Python code shown in the main
-  text. Cells are colored by material id.}
-\label{fig:pymesh}
-\end{figure}
-
 Initial support for Python has existed in \dealii{} since version
@@
 
+\begin{floatingfigure}[r]{0.3\textwidth}
+  \centering
+  \vspace*{-1.4cm}
+  \includegraphics[width=0.3\textwidth]{python_mesh.png}
+  \vspace*{-5mm}
+  \caption{\it The mesh generated by the Python code shown in the main
+    text. Cells are colored by material id.}
+  \vspace*{0.2cm}
+\label{fig:pymesh}
+\end{floatingfigure}
 All triangulations created from within Python are serial. However,
@@
 interfaces that are not usually used in external
-applications. However, some are worth mentioning:
+applications. That said, the following are worth mentioning since they
+may have been more widely used:
 \begin{itemize}
@@
 hand, if necessary, global information can still be computed using,
-for example, calling \texttt{Utilities::MPI::Allgather(locally\_owned\_info(), comm)}.
-\item
-  \todo[inline]{What else is worth mentioning? Someone needs to go
-    through the list of incompatible changes...}
+for example, calling
+\texttt{Utilities::MPI::Allgather(locally\_owned\_info(), comm)}.
+
+\item A corresponding change has been made in parallel triangulation
+  classes:
+  \texttt{parallel::} \texttt{TriangulationBase::compute\_n\_locally\_owned\_active\_cells\_per\_processor()}
+  can be used to obtain information about how many cells each
+  process owns.
 \end{itemize}

### 2020-05-28 9.2/paper.tex
Resolve another todo.

@@
   lengthening or shortening to final time step.
-\item z \todo[inline]{What else? Maybe mention the updated step-12?}
+\item A key component of \dealii{} are the \texttt{FEValues} and
+  \texttt{FEFaceValues} classes that evaluate finite element functions
+  at quadrature points located on cells and faces of a cell,
+  respectively~\cite{BangerthHartmannKanschat2007}. This release now
+  contains a class \texttt{FEInterfaceValues} that considers the
+  restriction of the shape functions from \textit{both} sides of a face and
+  allows evaluating jumps and averages of shape functions along this
+  face. These are common components of the bilinear forms of
+  discontinuous Galerkin schemes (as well as schemes for fourth-order
+  equations, see the discussion of \texttt{step-47} below) and greatly
+  simplify the implementation of these methods.
 \end{itemize}

### 2020-05-28 9.2/paper.tex
Take over some of the edits from #50.

@@
 all processes in a parallel environment, and the actual mesh used for computations is constructed by repeated
-refinement. However, this has its limitations in industrial
-applications, where frequently the mesh comes
-from an external mesh generator in the form of a file that already contains millions
+refinement. However, this has its limitations in many
+applications where the mesh comes
+from an external mesh generator in the form of a file that frequently
+already contains millions
 or tens of millions of cells. For such configurations, applications might exhaust
@@
 
-D.~Davydov was supported by the German Research Foundation (DFG), grant DA
-1664/2-1 and the Bayerisches Kompetenznetzwerk
-f\"ur Technisch-Wissenschaftliches Hoch- und H\"ochstleistungsrechnen
-(KONWIHR).
+%D.~Davydov was supported by the German Research Foundation (DFG), grant DA
+%1664/2-1 and the Bayerisches Kompetenznetzwerk
+%f\"ur Technisch-Wissenschaftliches Hoch- und H\"ochstleistungsrechnen
+%(KONWIHR).
+
+A.~V.~Grayver was partially supported by the European Space Agency
+Swarm DISC program.
 

### 2020-05-28 9.2/paper.tex
Unify typesetting of deal.II

@@
 
-\title{The \dealii Library, Version 9.2}
+\title{The \dealii{} Library, Version 9.2}
 
@@
 
-\dealii version 9.2.0 was released May XYZ, 2020.
+\dealii{} version 9.2.0 was released May XYZ, 2020.
 \todo[inline]{Need to update date.}
@@
 overview of the new features of this release and serves as a citable
-reference for the \dealii software library version 9.2. \dealii is an
+reference for the \dealii{} software library version 9.2. \dealii{} is an
 object-oriented finite element library used around the world in the
@@
 
-This release of \dealii contains a number of large and significant changes
+This release of \dealii{} contains a number of large and significant changes
 that will be discussed in this section.
@@
 
-Previously, all triangulation classes of \texttt{deal.II} had in common that the coarse grid is replicated by
+Previously, all triangulation classes of \dealii{} had in common that the coarse grid is replicated by
 all processes in a parallel environment, and the actual mesh used for computations is constructed by repeated
@@
 
-Since the previous release, \dealii has had support for $hp$-adaptive finite element methods
+Since the previous release, \dealii{} has had support for $hp$-adaptive finite element methods
 on distributed memory systems \cite{dealII91}. We implemented the bare functionality
@@
 The class \texttt{VectorizedArray<Number>} is a key component to achieve the high
-node-level performance of the matrix-free algorithms in deal.II~\cite{KronbichlerKormann2012, KronbichlerKormann2019}. It is a wrapper
+node-level performance of the matrix-free algorithms in \dealii{}~\cite{KronbichlerKormann2012, KronbichlerKormann2019}. It is a wrapper
 class around a short vector of $n$ entries of type \texttt{Number} and maps
@@
 as a consequence, the number of cells to be processed at once in matrix-free
-operator evaluations. For example, the deal.II-based
+operator evaluations. For example, the \dealii{}-based
 library \texttt{hyper.deal}~\cite{munch2020hyperdeal}, which solves the 6D Vlasov--Poisson equation with high-order
@@
 
-\caption{\it Comparison of relevant SIMD-related classes in deal.II and \texttt{C++23}.}\label{tab:simd}
+\caption{\it Comparison of relevant SIMD-related classes in \dealii{} and \texttt{C++23}.}\label{tab:simd}
 \centering
@@
 \toprule
-\textbf{VectorizedArray (deal.II)} & \textbf{std::simd (\texttt{C++23})} \\
+\textbf{VectorizedArray (\dealii{})} & \textbf{std::simd (\texttt{C++23})} \\
 \midrule
@@
 part of the \texttt{C++23} standard.
-Table~\ref{tab:simd} compares the deal.II-specific SIMD classes and
+Table~\ref{tab:simd} compares the \dealii{}-specific SIMD classes and
 the equivalent \texttt{C++23} classes. These changes also prepare for specialized
@@
 %entries have to be written into the solution vector only once with improved
-%data locality. Previous publications based on \texttt{deal.II} have shown the
+%data locality. Previous publications based on \dealii{} have shown the
 %relevance of the latter aspect for reaching higher performance.
@@
   cumbersome to implement if one wants to use general meshes. As a
-  consequence, they have largely fallen out of favor and deal.II currently
+  consequence, they have largely fallen out of favor and \dealii{} currently
   does not contain implementations of these shape functions.
@@
 have been deprecated. As discussed in
-Subsection~\ref{subsec:performance}, deal.II by default no longer
+Subection~\ref{subsec:performance}, \dealii{} by default no longer
 stores information for all processes on all processes, but only
@@
 
-In order to justify the work the developers of \dealii put into this
+In order to justify the work the developers of \dealii{} put into this
 software, we ask that papers using the library reference one of the
-\dealii papers. This helps us justify the effort we put into it.
-
-There are various ways to reference \dealii. To acknowledge the use of
+\dealii{} papers. This helps us justify the effort we put into it.
+
+There are various ways to reference \dealii{}. To acknowledge the use of
 the current version of the library, \textbf{please reference the present
@@
 
-The original \texttt{\dealii} paper containing an overview of its
+The original \dealii{} paper containing an overview of its
 architecture is \cite{BangerthHartmannKanschat2007}. If you rely on
@@
 
-\dealii can interface with many other libraries:
+\dealii{} can interface with many other libraries:
 \todo[inline]{We picked up gingko. Anything else?}
@@
 
-The two previous releases of \dealii can be cited as
+The two previous releases of \dealii{} can be cited as
 \cite{dealII90,dealII91}.
@@
 
-\dealii is a world-wide project with dozens of contributors around the
+\dealii{} is a world-wide project with dozens of contributors around the
 globe. Other than the authors of this paper, the following people
@@
 
-\dealii and its developers are financially supported through a
+\dealii{} and its developers are financially supported through a
 variety of funding sources:
@@
 The Interdisciplinary Center for Scientific Computing (IWR) at Heidelberg
-University has provided hosting services for the \dealii web page.
+University has provided hosting services for the \dealii{} web page.
 

### 2020-05-28 9.2/paper.tex
Fix link changes

@@
 can be found
-\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_0_1_and_9_1_0.html}{
-in the file that lists all changes for this release}, see \cite{changes91}.
+\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_1_1_and_9_2_0.html}{
+in the file that lists all changes for this release}, see \cite{changes92}.
 

### 2020-05-28 9.2/paper.tex
Introduce abbreviation for DG

@@
 compressible Euler equations discretized with a high-order discontinuous
-Galerkin scheme using the matrix-free infrastructure. Besides the use of
+Galerkin (DG) scheme using the matrix-free infrastructure. Besides the use of
 matrix-free evaluators for systems of equations and over-integration, it also

### 2020-05-28 9.2/paper.tex
Update funding

@@
 
+D.~Arndt and B.~Turcksin: Research sponsored by the Laboratory Directed Research and
+Development Program of Oak Ridge National Laboratory, managed by UT-Battelle,
+LLC, for the U. S. Department of Energy.
+
 W.~Bangerth, T.~C.~Clevenger, and T.~Heister were partially
@@
 
-B.~Turcksin: Research sponsored by the Laboratory Directed Research and
-Development Program of Oak Ridge National Laboratory, managed by UT-Battelle,
-LLC, for the U. S. Department of Energy.
-
 D.~Wells was supported by the National Science Foundation (NSF) through Grant

### 2020-05-28 9.2/paper.tex
add some text about step-69

@@
 
-\item \texttt{step-69}
-  \todo[inline]{Matthias/Ignacio to write}
+\item \texttt{step-69} presents a first-order scheme solving compressible
+Euler's equations of gas dynamics using a graph-based scheme. The time-update
+at each node requires the evaluation of a right-hand side that depends
+(nonlinearly) on information from the previous time-step that spans more than
+one cell. Therefore, assembly loops operate directly on the sparsity graph in
+order to retrieve information from the entire stencil associated to each
+node. From a programming perspective, \texttt{step-69} features a number of
+techniques that are of interest for a wider audience: It discusses a hybrid
+thread and MPI parallelized scheme with efficient (MPI node) local
+numbering of degrees of freedom. It showcases how to perform asynchronous
+write-out of results using a background thread with \texttt{std::async},
+and discusses a simple but effective checkpointing and restart technique.
 

### 2020-05-29 9.2/paper.tex
add a small section on our ternary operator support for VectorizedArray

@@
 
-The class \texttt{VectorizedArray<Number>} is a key component to achieve the high
-node-level performance of the matrix-free algorithms in \dealii{}~\cite{KronbichlerKormann2012, KronbichlerKormann2019}. It is a wrapper
-class around a short vector of $n$ entries of type \texttt{Number} and maps
-arithmetic operations to appropriate single-instruction/multiple-data (SIMD)
-concepts by intrinsic functions.
-The class \texttt{VectorizedArray} has been made more user-friendly in this release by making
-it compatible with the STL algorithms found in the header \texttt{<algorithm>}.
-The length of the  vector can now be queried by  \texttt{VectorizedArray::size()} and its underlying number type by \texttt{VectorizedArray::value\_type}.
-Furthermore, the \texttt{VectorizedArray} class now supports range-based iteration over its entries.
-
-In previous \dealii{} releases, the
-vector length was set at compile time of the library to match the highest
-value supported by the given processor architecture.
-Now, a second optional template argument
-can be specified as \texttt{VectorizedArray<Number, size>}, where \texttt{size} explicitly controls
-the vector length within the capabilities of a particular instruction set.
-(A full list of supported
-vector lengths is presented in Table~\ref{tab:vectorizedarray}.)
-This allows users to select the vector length/ISA and,
-as a consequence, the number of cells to be processed at once in matrix-free
-operator evaluations. For example, the \dealii{}-based
-library \texttt{hyper.deal}~\cite{munch2020hyperdeal}, which solves the 6D Vlasov--Poisson equation with high-order
-discontinuous Galerkin methods (with more than a thousand degrees of freedom per cell), constructs a tensor product of two \texttt{MatrixFree} objects of different SIMD-vector
-length in the same application and benefits---in terms of performance---by the possibility of decreasing the number of cells processed by a single SIMD instruction.
+The class \texttt{VectorizedArray<Number>} is a key component to achieve
+the high node-level performance of the matrix-free algorithms in
+\dealii{}~\cite{KronbichlerKormann2012, KronbichlerKormann2019}. It is a
+wrapper class around a short vector of $n$ entries of type \texttt{Number}
+and maps arithmetic operations to appropriate
+single-instruction/multiple-data (SIMD) concepts by intrinsic functions.
+The class \texttt{VectorizedArray} has been made more user-friendly in this
+release by making it compatible with the STL algorithms found in the header
+\texttt{<algorithm>}. The length of the  vector can now be queried by
+\texttt{VectorizedArray::size()} and its underlying number type by
+\texttt{VectorizedArray::value\_type}. The \texttt{VectorizedArray} class
+now supports range-based iteration over its entries. In addition \dealii{}
+now supports ternary operations on vectorized data, where for a given
+binary comparison operation a true or false value is selected. For example,
+the vectorized equivalent of \texttt{(left < right) ? true\_value :
+false\_value} can be expressed as
+\begin{c++}
+auto result = compare_and_apply_mask<SIMDComparison::less_than>(
+  left, right, true_value, false_value);
+\end{c++}
+which compares every element of the vectorized array individually and
+selects the corresponding value from the \texttt{true\_value}, or
+\texttt{false\_value} array.
+
+In previous \dealii{} releases, the vector length was set at compile time
+of the library to match the highest value supported by the given processor
+architecture. Now, a second optional template argument can be specified as
+\texttt{VectorizedArray<Number, size>}, where \texttt{size} explicitly
+controls the vector length within the capabilities of a particular
+instruction set. (A full list of supported vector lengths is presented in
+Table~\ref{tab:vectorizedarray}.) This allows users to select the vector
+length/ISA and, as a consequence, the number of cells to be processed at
+once in matrix-free operator evaluations. For example, the \dealii{}-based
+library \texttt{hyper.deal}~\cite{munch2020hyperdeal}, which solves the 6D
+Vlasov--Poisson equation with high-order discontinuous Galerkin methods
+(with more than a thousand degrees of freedom per cell), constructs a
+tensor product of two \texttt{MatrixFree} objects of different SIMD-vector
+length in the same application and benefits---in terms of performance---by
+the possibility of decreasing the number of cells processed by a single
+SIMD instruction.
 

### 2020-05-29 9.2/paper.tex
reword step-69 slightly

@@
 \item \texttt{step-69} presents a first-order scheme solving compressible
-Euler's equations of gas dynamics using a graph-based scheme. The time-update
-at each node requires the evaluation of a right-hand side that depends
-(nonlinearly) on information from the previous time-step that spans more than
-one cell. Therefore, assembly loops operate directly on the sparsity graph in
-order to retrieve information from the entire stencil associated to each
-node. From a programming perspective, \texttt{step-69} features a number of
-techniques that are of interest for a wider audience: It discusses a hybrid
-thread and MPI parallelized scheme with efficient (MPI node) local
-numbering of degrees of freedom. It showcases how to perform asynchronous
-write-out of results using a background thread with \texttt{std::async},
-and discusses a simple but effective checkpointing and restart technique.
+Euler's equations of gas dynamics with a graph-viscosity stabilization
+technique. Beside the usual conservation properties of mass, momentum and
+total energy, the method also guarantees that the constructed solution
+obeys pointwise stability constraints (positivity of density, internal
+energy and a local minimum principle on the specific entropy). As such
+\text{step-69} is strictly speaking more a collocation-type discretization
+than a variational formulation, even though it is implemented with finite
+elements.
+
+The time-update at each node requires the evaluation of a right-hand side
+that depends (nonlinearly) on information from the previous time-step that
+spans more than one cell. Therefore, assembly loops operate directly on the
+sparsity graph in order to retrieve information from the entire stencil
+associated to each node. From a programming perspective, \texttt{step-69}
+features a number of techniques that are of interest for a wider audience:
+It discusses a hybrid thread and MPI parallelized scheme with efficient
+(MPI node) local numbering of degrees of freedom. It showcases how to
+perform asynchronous write-out of results using a background thread with
+\texttt{std::async}, and discusses a simple but effective checkpointing and
+restart technique.
 

### 2020-05-30 9.2/paper.tex
Add author: JPP

@@
     Matthias Maier,
-%     Jean-Paul Pelteret,
+    Jean-Paul Pelteret,
     Reza Rastak,
@@
 
-%
-% \author[4]{Jean-Paul~Pelteret}
-%
- \author[11]{Reza Rastak}
- \affil[11]{Department of Civil and Environmental Engineering,
+\author[11]{Jean-Paul~Pelteret}
+\affil[11]{Independent researcher.
+{\texttt{jppelteret@gmail.com}}}
+
+ \author[12]{Reza Rastak}
+ \affil[12]{Department of Civil and Environmental Engineering,
    Stanford University,
@@
 
- \author[**12]{Ignacio Tomas}
- \affil[12]{Sandia National Laboratories, Org 1442,
+ \author[**13]{Ignacio Tomas}
+ \affil[13]{Sandia National Laboratories, Org 1442,
  P.O. Box 5800, MS 1320, Albuquerque, NM, 87185-1320,
@@
 %
-\author[13]{David Wells}
-\affil[13]{Department of Mathematics, University of North Carolina,
+\author[14]{David Wells}
+\affil[14]{Department of Mathematics, University of North Carolina,
   Chapel Hill, NC 27516, USA.
@@
 
-\author[14]{Zhuoran Wang}
-\affil[14]{Department of Mathematics, Colorado State University, Fort Collins,
+\author[15]{Zhuoran Wang}
+\affil[15]{Department of Mathematics, Colorado State University, Fort Collins,
           CO 80523-1874, USA.

### 2020-05-30 9.2/paper.tex
Notes for (future) new additions to code-gallery

@@
 \todo[inline]{Do we have new code gallery programs?}
+\begin{itemize}
+  \item Parallel in time (\url{https://github.com/dealii/code-gallery/pull/49})
+  \item Linear elastoplasticity (\url{https://github.com/dealii/code-gallery/pull/62})
+  \item Thermo-electro-elastic pump (\url{https://github.com/dealii/code-gallery/pull/63})
+  \item Nonlinear poro-viscoelasticity (\url{https://github.com/dealii/code-gallery/pull/72})
+\end{itemize}
 

### 2020-05-31 9.2/paper.tex
Added a first version of section 2.4 related t the support for particle-based methods

@@
 \author[3]{Bruno Blais}
-\affil[3]{Research Unit for Industrial Flows Processes (URPEI),
+\affil[3]{Research Unit for Industrial Flows Processes (URPEI), Department of Chemical Engineering,
           Polytechnique Montréal,
@@
 \label{subsec:particles}
+
+Support for particles was originally introduced in \dealii{} version 9.0. In \dealii,
+particles can be used as passive tracers, or as part of more complex
+models such as those based on Particle-In-Cells (PIC) approaches \cite{gassmoller2018flexible}.
+
+With the current release, the support for particles has been further expanded: New parallel
+ insertion mechanisms and a basic interface to post-process particles have
+been added, effectively making their usage more flexible and enabling a larger
+range of use cases (e.g. immersed boundaries in step-70).
+
+Through the addition of the \texttt{Particles::ParticleHandler::insert\_global\_particles} member function,
+ particles can now be inserted in parallel from a vector of points even if these points do not lie
+ on the subdomain from which the insertion is called. This operation requires extensive 
+ communication between the processes to locate   the MPI process the cell in which the particle
+  is located. However, it is made significantly faster  through the usage of bounding boxes
+  that surround the cells of each subdomain. 
+  This function also takes care of transferring the properties attached to the particles to their new owner.
+ This new capacity enables particle generators that insert particles at the location
+ of the support points (\texttt{Particles::Generator::dof\_support\_points}) and at the quadrature points (\texttt{Particles::Generator::quadrature\_points}) of a
+ possibly non-matching triangulation. Consequently, complex particle patterns can be inserted
+ using unstructured hexahedral grids generated outside of \dealii{}.
+
+To visualize the motion of particles, the \texttt{Particles::DataOut} class was added to the library.
+In its current version, this class supports writing the particles positions and id in all format supported
+by \dealii, but does not allow writing the properties attached to the particles.
 
@@
 
+B.~Blais was partially supported by the National Science and Engineering Research Council of Canada(NSERC)  through the RGPIN-2020-04510 Discovery Grant
+
 A.~V.~Grayver was partially supported by the European Space Agency

### 2020-05-31 9.2/paper.tex
Introdusing transfer of quadrature data of different lengths

--- 2020-05-30 9.2/paper.tex
@@
   simplify the implementation of these methods.
+\item Previously, the \texttt{parallel::distributed::ContinuousQuadratureDataTransfer}
+  class, which transfers local quadrature point data onto the
+  children of newly refined cells, did not allow different regions of the
+  triangulation store quadrature data of different lengths. This release lifts
+  this restriction, which enables efficient quadrature data storage and data
+  transfer within a triangulation containing multiple material models, each with
+  their own number of local state variables and history variables.
+  As an example, in a solid mechanics simulation, a hyper-elastic material
+  model can be assigned to a region of the mesh requiring no history variables
+  while another part of the triangulation uses an elasto-plastic constitutive
+  model, requiring the storage of local plasticity data at quadrature points.
+  During each refinement of the mesh, \texttt{ContinuousQuadratureDataTransfer}
+  can be used to transfer and interpolate plasticity data from parent cells to
+  their children, ignoring the cells with hyper-elastic constitutive model.
 \end{itemize}

### 2020-05-31 9.2/paper.tex
sentences changed to an active voice

--- 2020-05-30 9.2/paper.tex
@@
   simplify the implementation of these methods.
+\item Previously, the \texttt{parallel::distributed::ContinuousQuadratureDataTransfer}
+  class, which transfers local quadrature point data onto the
+  children of newly refined cells, did not allow different regions of the
+  triangulation store quadrature data of different lengths. This release lifts
+  this restriction, which enables efficient quadrature data storage and data
+  transfer within a triangulation containing multiple material models, each with
+  their own number of local state variables and history variables.
+  As an example, in a solid mechanics simulation, we can assign a hyper-elastic
+  material model can be assigned to a region of the mesh with no associated
+  history variables while in another part of the triangulation we incorporate an
+  elasto-plastic constitutive model, requiring the storage of local plasticity
+  data at quadrature points.
+  During each refinement of the mesh, we can use
+  \texttt{ContinuousQuadratureDataTransfer} to transfer and interpolate
+  plasticity data from parent cells to their children, ignoring the cells
+  with the hyper-elastic constitutive model.
 \end{itemize}

### 2020-06-01 9.2/paper.tex
Avoid a duplicate bibtex entry.

@@
 particles can be used as passive tracers, or as part of more complex
-models such as those based on Particle-In-Cells (PIC) approaches \cite{gassmoller2018flexible}.
+models such as those based on Particle-In-Cells (PIC) approaches \cite{GLHPB18}.
 

### 2020-06-01 9.2/paper.tex
Minor edits.

--- 2020-05-31 9.2/paper.tex
@@
   class, which transfers local quadrature point data onto the
-  children of newly refined cells, did not allow different regions of the
-  triangulation store quadrature data of different lengths. This release lifts
+  children of newly refined cells, did not allow different cells
+  to store quadrature data of different lengths. This release lifts
   this restriction, which enables efficient quadrature data storage and data
@@
   As an example, in a solid mechanics simulation, we can assign a hyper-elastic
-  material model can be assigned to a region of the mesh with no associated
-  history variables while in another part of the triangulation we incorporate an
+  material model to a region of the mesh with no associated
+  history variables, while in another part of the triangulation we incorporate an
   elasto-plastic constitutive model, requiring the storage of local plasticity
   data at quadrature points.
-  During each refinement of the mesh, we can use
+  During each refinement of the mesh, we can then use
   \texttt{ContinuousQuadratureDataTransfer} to transfer and interpolate
@@
 
+Support for particles was originally introduced in \dealii{} version 9.0. These
+particles can be used as passive tracers, or as part of more complex
+models such as those based on Particle-In-Cells (PIC) approaches \cite{GLHPB18}.
+
+With the current release, support for particles has been further expanded: New parallel
+insertion mechanisms and a basic interface to post-process particles have
+been added, effectively making their usage more flexible and enabling a larger
+range of use cases (such as the immersed boundaries in step-70, see below).
+
+Through the addition of the \texttt{Particles::ParticleHandler::insert\_global\_particles()} member function,
+particles can now be inserted in parallel from a vector of points even if these points do not lie
+on the subdomain from which the insertion is called. This operation requires extensive 
+communication between the processes to locate   the MPI process that
+owns the cell in which the particle
+is located. However, it is made significantly faster  through the usage of bounding boxes
+that surround the cells of each subdomain. 
+This function also takes care of transferring the properties attached to the particles to their new owner.
+This new capability enables particle generators that insert particles at the location
+of the support points (\texttt{Particles::Generator::dof\_support\_points()}) and at the quadrature points (\texttt{Particles::Generator::quadrature\_points()}) of a
+possibly non-matching triangulation. Consequently, complex particle patterns can be inserted
+using unstructured grids generated outside of \dealii{}.
+
+To visualize the motion of particles, the \texttt{Particles::DataOut} class was added to the library.
+
 \todo[inline]{Luca: Your section}
@@
 
-\item \texttt{step-69} presents a first-order scheme solving compressible
-Euler's equations of gas dynamics with a graph-viscosity stabilization
-technique. Beside the usual conservation properties of mass, momentum and
+\item \texttt{step-69} presents a first-order scheme solving the compressible
+Euler equations of gas dynamics with a graph-viscosity stabilization
+technique. Beside the usual conservation properties of mass, momentum, and
 total energy, the method also guarantees that the constructed solution
-obeys pointwise stability constraints (positivity of density, internal
-energy and a local minimum principle on the specific entropy). As such
+obeys pointwise stability constraints (in particular positivity of
+density and internal
+energy, and a local minimum principle on the specific entropy). As such
 \text{step-69} is strictly speaking more a collocation-type discretization
@@
 sparsity graph in order to retrieve information from the entire stencil
-associated to each node. From a programming perspective, \texttt{step-69}
+associated with each node. From a programming perspective, \texttt{step-69}
 features a number of techniques that are of interest for a wider audience:
 It discusses a hybrid thread and MPI parallelized scheme with efficient
-(MPI node) local numbering of degrees of freedom. It showcases how to
+MPI node-local numbering of degrees of freedom. It showcases how to
 perform asynchronous write-out of results using a background thread with

### 2020-06-01 9.2/paper.tex
Add step-70 description.

@@
     Bruno Turcksin,
+    Zhuoran Wang,
     David Wells
-    Zhuoran Wang,
   },
@@
 %
-\author[14]{David Wells}
-\affil[14]{Department of Mathematics, University of North Carolina,
+
+\author[14]{Zhuoran Wang}
+\affil[14]{Department of Mathematics, Colorado State University, Fort Collins,
+          CO 80523-1874, USA.
+{\texttt{zhrwang@math.colostate.edu}}}
+
+\author[15]{David Wells}
+\affil[15]{Department of Mathematics, University of North Carolina,
   Chapel Hill, NC 27516, USA.
@@
 
-\author[15]{Zhuoran Wang}
-\affil[15]{Department of Mathematics, Colorado State University, Fort Collins,
-          CO 80523-1874, USA.
-{\texttt{zhrwang@math.colostate.edu}}}
 \renewcommand{\labelitemi}{--}
@@
 is located. However, it is made significantly faster  through the usage of bounding boxes
-that surround the cells of each subdomain. 
+that provide a coarse description of the geometrical shape of each subdomain.
 This function also takes care of transferring the properties attached to the particles to their new owner.
@@
     \toprule
-\textbf{double} & \textbf{float} & \textbf{ISA}\\
+    \textbf{double}            & \textbf{float}             & \textbf{ISA}         \\
     \midrule
@@
           + \kappa |\psi|^2 \psi
-  &= 0,
+           & = 0,
         \end{align*}
@@
 
-\item \texttt{step-70}
-  \todo[inline]{Luca, please write.}
+  \item \texttt{step-70} solves a fluid structure interaction problem on
+        non-matching parallel distributed triangulations, showing the usage of
+        the \texttt{particles::ParticleHandler} class for two different tasks:
+        i) to track the position of quadrature points of a non-matching
+        solid grid immersed in a fluid grid, and ii) to track the position of a
+        collection of massless tracers.
+
+        In \texttt{step-70} we consider a mixing problem in the laminar flow
+        regime. Such problems occur in a wide range of applications ranging
+        from chemical engineering to power generation (e.g. turbomachinery).
+        Mixing problems are particularly hard to solve numerically, because
+        they often involve a container (with fixed boundaries, and possibly
+        complex geometries such as baffles), and one (or more) immersed and
+        rotating impellers, making it difficult to use Arbitrary Lagrangian
+        Eulerian formulations, where the fluid domain – along with the mesh! –
+        is smoothly deformed to follow the deformations of the immersed solid.
+        We show how to solve such problems using a penalization technique
+        where the flow problem is solved in the union of the fluid and solid
+        domain, and the movement of the solid is imposed weakly using a
+        Nitsche-like penalization.
+
+        The \texttt{particles::ParticleHandler} class is used in this context
+        to allow integration of the fluid basis functions on the solid domain
+        (which is immersed and non-matching w.r.t. the fluid grid). This is
+        achieved by attaching the information required to perform the
+        integration on the solid grid, to the property field of the particles
+        associated with the quadrature points of the solid, and advecting those
+        particles according to the solid velocity. In this program, we assign
+        the solid velocity externally, and we do not solve for it, even though
+        the program can be easily extended to allow for the solution of
+        coupled elasticity equations on the solid domain.
 \end{itemize}
@@
 % 9.2: updated 5/11/2020 MM
-\todo[inline]{TODO: remove authors of paper}
 Pasquale Africa,
@@
 Mathias Anselmann,
-Daniel Arndt,
-Wolfgang Bangerth,
 Kirana Bergstrom,
 Manaswinee Bezbaruah,
-Bruno Blais,
 Benjamin Brands,
@@
 Fabian Castelli,
-Thomas C. Clevenger,
 Katherine Cosburn,
@@
 Luel Emishaw,
-Marc Fehling,
 Niklas Fehn,
@@
 Johannes Heinz,
-Timo Heister,
-Luca Heltai,
 Jiuhua Hu,
@@
 Uwe K{\"o}cher,
-Martin Kronbichler,
 Paras Kumar,
@@
 Wenyu Lei,
-Matthias Maier,
 Katrin Mang,
@@
 Adriana Morales Miranda,
-Peter Munch,
 Bob Myhill,
@@
 Alexey Ozeritskiy,
-Jean-Paul Pelteret,
 Rebecca Pereira,
@@
 Laura Prieto Saavedra,
-Reza Rastak,
 Roland Richter,
@@
 Jihuan Tian,
-Ignacio Tomas,
 Sara Tro,
-Bruno Turcksin,
 Ferdinand Vanmaele,
-Zhuoran Wang,
-David Wells,
 Michal Wichrowski,
@@
 by Technical Data Analysis, Inc. through US Navy SBIR N16A-T003.
+
+L.~Heltai was partially supported by the Italian Ministry of Instruction,
+University and Research (MIUR), under the 2017 PRIN project NA-FROM-PDEs MIUR
+PE1, ``Numerical Analysis for Full and Reduced Order Methods for the efficient
+and accurate solution of complex systems governed by Partial Differential
+Equations''.
 

### 2020-06-01 9.2/paper.tex
Add section on SD::BatchOptimizer

@@
 
-\todo[inline]{J-P's section}
-
-BatchOptimizer for symbolic expressions
-\begin{itemize}
-  \item CSE for dictionary-based expressions
-  \item ``lambdify'' (via SymEngine)
-  \item Offloading to LLVM JIT compiler (via SymEngine)
-  \item Serialization, so complex expressions can be compiled offline
-\end{itemize}
+In the previous release we added support for symbolic expressions, leveraging the
+SymEngine library \cite{symengine-web-page}.
+Although effective, evaluating lengthy expressions could be a bottle-neck as this
+was performed using dictionary-based substitution.
+We have improved on this by implementing a \texttt{BatchOptimizer} class in the 
+namespace \texttt{Differentiation::SD} that collects several \texttt{Expression}s
+and transforms them in such a way that the equivalent result is returned through
+a quicker code path.
+This may be done by simply using common subexpression elimination (CSE) for the
+dictionary-based expressions, by transformation to a set of nested 
+\texttt{std::function}s (the equivalent to \texttt{SymPy}'s "lambdify", with or
+without using CSE), or by offloading to these expressions to the \texttt{LLVM}
+just-in-time (JIT) compiler.
+Although, each of these features is implemented and tested in the SymEngine
+library itself, the \texttt{BatchOptimizer} class provides both uniform
+interface to their classes and a convenient interface for scalar expressions,
+as well as tensorial expressions formed using the \texttt{deal.II} tensor and 
+symmetric tensor classes.
+It, like the \texttt{Expression} class, is also serializable.
+
+The way that the batch optimizer may be employed within a user's code is shown
+in the pseudo-code below.
+As per usual, one would first define some independent variables, and 
+subsequently compute some symbolic expressions that are dependent on these 
+independent variables. 
+These expression could be, for example, scalar  expressions or tensors of 
+expressions.
+Instead of evaluating these expressions directly, the user would now create an
+optimizer to evaluate the dependent functions.
+In this example, the selected arithmetic type numerical result will be of type
+\texttt{double}, and the \texttt{LLVM} JIT optimizer will be invoked.
+It will employ common subexpression elimination and aggressive optimizations
+during compilation.
+The user then informs the optimizer of all of the independent variables and the
+dependent expressions, and invokes the optimization process.
+This is an expensive call, as it determines an equivalent code path to evaluate
+all of the dependent functions at once.
+However, in many cases each evaluation has significantly less computational cost
+than evaluating the symbolic expressions directly.
+Evaluation is performed when the user constructs a substitution map, giving each
+independent variable a numerical representation, and passes those to the 
+optimizer.
+After this step, the numerical equivalent of the individual dependent expressions
+may finally be retreived from the optimizer.
+
+\begin{c++}
+using namespace Differentiation::SD;
+
+const Expression x("x");
+const Expression y("y");
+...
+const auto f = calculate_f(x, y, ...); // User function
+const auto g = calculate_g(x, y, ...); // User function
+...
+
+BatchOptimizer<double> optimizer (OptimizerType::llvm,
+                                  OptimizationFlags::optimize_all);
+optimizer.register_symbols(x, y, ...);
+optimizer.register_functions(f, g, ...);
+optimizer.optimize();
+
+const auto substitution_map
+  = make_substitution_map({x, ...}, {y, ...}, ...);
+optimizer.substitute(substitution_map);
+
+const auto result_f = optimizer.evaluate(f);
+const auto result_g = optimizer.evaluate(g);
+\end{c++}
+
+This expense of invoking the optimizer may be offset not only by
+the number of evaluations performed, but also by the amount of reuse each
+instance of a \texttt{BatchOptimizer} has.
+In certain circumstances, this can be maximized by generalizing the way in which
+the dependent expressions are formulated.
+For example, in the context of constitutive modelling the material coefficients
+may be made symbolic rather than encoding these into the dependent expressions
+as numerical values.
+The optimizer may then be used to evaluate an entire family of constitutive laws,
+and not a specific one that describes the response of a single material.
+Thereafter, serializing the optimizer instance and reloading the contents during
+subsequent simulations permits the user to skip the optimization process entirely.
+Serialization also enables these complex expressions to be compiled offline.
 

### 2020-06-01 9.2/paper.tex
add date, leave a question

@@
 
-\dealii{} version 9.2.0 was released May XYZ, 2020.
-\todo[inline]{Need to update date.}
+\dealii{} version 9.2.0 was released May 20, 2020.
 This paper provides an
@@
 (here: 30\%/3\%). From those cells listed for adaptation, we designate a subset
-for $h$- and $p$-adaptation (here: 10\%/90\%). The parameters of the corresponding
+for $h$- and $p$-adaptation (here: 10\%/90\%).
+\todo[inline]{Timo asks: is this a typo in the code where you say 0.9, 0.9?}
+The parameters of the corresponding
 \texttt{hp::Refinement} function specify the fraction of cells to be $p$-adapted from

### 2020-06-01 9.2/paper.tex
add a section about large scale Stokes

@@
 with \num{2.1e12} unknowns.
+\todo[inline]{the figure only shows 2e11 unknowns. Is this a typo? If not, do we have a reference?}
 Figure~\ref{fig:init_costs} compares timings of simulations of various problem
@@
 
-\todo[inline]{Timo: Do you want to say something about that the new code has been also applied to solve problems with adaptively refined
-  meshes with more than \num{4e9} unknowns?}
-
-\todo[inline]{Timo says: We did solve adaptive problems with more than
-  4B unknowns before, but some were broken for a while during this
-  release stage due to the consensus algorithms. But we did not test
-  systems or block vectors before. But we should not make it sound too
-  negative.}
-
-\todo[inline]{Wolfgang says: Should we just not mention any of this at
-  all? The section stands well on its own.}
+As part of this effort, we ran benchmarks on the TACC Frontera system, where we were able to apply the matrix-free geometric multigrid framework to a variable viscosity Stokes system and achieved weak and strong scaling up to 114k MPI ranks with up to
+\num{2.1e11} unknowns. This is likely the currently largest block system solved with \dealii{} and required various optimizations
+and fixes on top of the ones mentioned above:
+1) Bug fixes to concurrent point to point communications.
+2) Fixes multigrid transfer with adaptive refinement and more than \num{4e9} unknowns.
+3) Fixes to index sets in block indices with more than \num{4e9} unknowns.
+4) Fixes to computations with more than \num{4e9} active cells.
+5) Implementation of IDR(s) solvers to reduce memory overhead.
+For more details, see \cite{clevenger_stokes19}.
+
+\todo[inline]{Timo added the paragraph above. Thoughts on how much we should mention (none of this worked with 9.1)?}
 

### 2020-06-01 9.2/paper.tex
add draft for step-50

@@
 
-  \item \texttt{step-50}
+  \item \texttt{step-50} is a program that demonstrates the parallel geometric multigrid features
+  in \dealii{} as described in \cite{ClevengerHeisterKanschatKronbichler2019}.
+  The problem considered is a variable viscosity Laplace equation and it is solved
+  with three different approaches:
+  1. With a matrix-based geometric multigrid based on Trilinos or PETSc. 2. With a matrix-free
+  geometric multigrid. 3. Using algebraic multigrid (Trilinos ML).
+
   \todo[inline]{Timo/Conrad/... to write}

### 2020-06-01 9.2/paper.tex
Be more verbose on hp interface.

@@
 (here: 30\%/3\%). From those cells listed for adaptation, we designate a subset
-for $h$- and $p$-adaptation (here: 10\%/90\%).
-\todo[inline]{Timo asks: is this a typo in the code where you say 0.9, 0.9?}
-The parameters of the corresponding
+for $h$- and $p$-adaptation. The parameters of the corresponding
 \texttt{hp::Refinement} function specify the fraction of cells to be $p$-adapted from
-those subsets flagged for refinement and coarsening, respectively.
+those subsets flagged for refinement and coarsening, respectively (here: 90\%/80\%),
+while the remaining fraction will be $h$-adapted (here: 10\%/20\%).
 \begin{c++}
@@
 hp::Refinement::p_adaptivity_fixed_number(
-dof_handler, estimated_smoothness_per_cell, 0.9, 0.9);
+dof_handler, estimated_smoothness_per_cell, 0.9, 0.8);
 hp::Refinement::choose_p_over_h(dof_handler);

### 2020-06-01 9.2/paper.tex
Minor edits.

@@
 and fixes on top of the ones mentioned above:
-1) Bug fixes to concurrent point to point communications.
-2) Fixes multigrid transfer with adaptive refinement and more than \num{4e9} unknowns.
-3) Fixes to index sets in block indices with more than \num{4e9} unknowns.
-4) Fixes to computations with more than \num{4e9} active cells.
-5) Implementation of IDR(s) solvers to reduce memory overhead.
+(i) Bug fixes to concurrent point to point communications.
+(ii) Fixes multigrid transfer with adaptive refinement and more than \num{4e9} unknowns.
+(iii) Fixes to index sets in block indices with more than \num{4e9} unknowns.
+(iv) Fixes to computations with more than \num{4e9} active cells.
+(v) Implementation of IDR(s) solvers to reduce memory overhead.
 For more details, see \cite{clevenger_stokes19}.
@@
 SymEngine library \cite{symengine-web-page}.
-Although effective, evaluating lengthy expressions could be a bottle-neck as this
+Although effective, evaluating lengthy expressions could be a bottleneck as this
 was performed using dictionary-based substitution.
@@
 dictionary-based expressions, by transformation to a set of nested 
-\texttt{std::function}s (the equivalent to \texttt{SymPy}'s "lambdify", with or
-without using CSE), or by offloading to these expressions to the \texttt{LLVM}
+\texttt{std::function} objects (the equivalent to \texttt{SymPy}'s ``lambdify'', with or
+without using CSE), or by offloading these expressions to the \texttt{LLVM}
 just-in-time (JIT) compiler.
-Although, each of these features is implemented and tested in the SymEngine
-library itself, the \texttt{BatchOptimizer} class provides both uniform
+Although each of these features is implemented and tested in the SymEngine
+library itself, the \texttt{BatchOptimizer} class provides both a uniform
 interface to their classes and a convenient interface for scalar expressions,
@@
 
-The way that the batch optimizer may be employed within a user's code is shown
+The way the batch optimizer may be employed within a user's code is shown
 in the pseudo-code below.
@@
   with three different approaches:
-  1. With a matrix-based geometric multigrid based on Trilinos or PETSc. 2. With a matrix-free
-  geometric multigrid. 3. Using algebraic multigrid (Trilinos ML).
+  (i) With a matrix-based geometric multigrid based on Trilinos or
+  PETSc; (ii) with a matrix-free
+  geometric multigrid; (iii) using algebraic multigrid (Trilinos ML).
 

### 2020-06-01 9.2/paper.tex
More edits.

@@
         the \texttt{particles::ParticleHandler} class for two different tasks:
-        i) to track the position of quadrature points of a non-matching
-        solid grid immersed in a fluid grid, and ii) to track the position of a
+        (i) to track the position of quadrature points of a non-matching
+        solid grid immersed in a fluid grid, and (ii) to track the position of a
         collection of massless tracers.
 
-        In \texttt{step-70} we consider a mixing problem in the laminar flow
+        The program considers a mixing problem in the laminar flow
         regime. Such problems occur in a wide range of applications ranging
@@
         to allow integration of the fluid basis functions on the solid domain
-        (which is immersed and non-matching w.r.t. the fluid grid). This is
+        (which is immersed and non-matching with regard to the fluid grid). This is
         achieved by attaching the information required to perform the
@@
         associated with the quadrature points of the solid, and advecting those
-        particles according to the solid velocity. In this program, we assign
-        the solid velocity externally, and we do not solve for it, even though
-        the program can be easily extended to allow for the solution of
+        particles according to the solid velocity. In this program, we
+        externally prescribe rather than solve for the
+        the solid velocity; however,
+        the program can easily be extended to allow for the solution of
         coupled elasticity equations on the solid domain.

### 2020-06-02 9.2/paper.tex
Add one acknowledgment.

@@
 
-As part of this effort, we ran benchmarks on the TACC Frontera system, where we were able to apply the matrix-free geometric multigrid framework to a variable viscosity Stokes system and achieved weak and strong scaling up to 114k MPI ranks with up to
-\num{2.1e11} unknowns. This is likely the currently largest block system solved with \dealii{} and required various optimizations
+As part of this effort, we ran benchmarks on the TACC Frontera system, where we were
+able to apply the matrix-free geometric multigrid framework to a variable viscosity
+Stokes system and achieved weak and strong scaling up to 114K MPI ranks with up to
+\num{2.1e11} unknowns. This is likely the largest block system currently solved with
+\dealii{} and required various optimizations
 and fixes on top of the ones mentioned above:
@@
   with three different approaches:
-  (i) With a matrix-based geometric multigrid based on Trilinos or
-  PETSc; (ii) with a matrix-free
-  geometric multigrid; (iii) using algebraic multigrid (Trilinos ML).
-
-  \todo[inline]{Timo/Conrad/... to write}
+  (i) using a matrix-based geometric multigrid based on Trilinos or
+  PETSc; (ii) using a matrix-free
+  geometric multigrid; (iii) using algebraic multigrid (Trilinos ML). The tutorial
+  demonstrates the superiority of the matrix-free method for the problem under consideration,
+  and shows that, for matrix-based formulations, the performance of algebraic and
+  geometric multigrid methods are roughly comparable.
 
@@
 
-W.~Bangerth and T.~Heister were also partially supported by award DMS-1821210.
-\todo[inline]{Need to add FRES project}
+W.~Bangerth and T.~Heister were also partially supported by award DMS-1821210
+and EAR-1925595.
 

### 2020-06-02 9.2/paper.tex
Add reference

@@
 
-\todo[inline]{Aren't the target processes the result of the operation, and the
-  input global numbers of indices that we request?}
-
 By replacing the collective communications during set up and removing the arrays
@@
 time for large-scale simulations and to solve a Poisson problem with multigrid
-with \num{2.1e12} unknowns.
-\todo[inline]{the figure only shows 2e11 unknowns. Is this a typo? If not, do we have a reference?}
+with \num{2.1e12} unknowns on the SuperMUC-NG supercomputer with 304,152
+cores~\cite{dealII2020design,Arndt2020sppexa}.
 Figure~\ref{fig:init_costs} compares timings of simulations of various problem
 sizes (including set up) on 49,152 MPI ranks using a matrix-free
-solver~\cite{KronbichlerKormann2019,KronbichlerWall2018}; this solver uses discontinuous elements of
+solver~\cite{Arndt2020sppexa,KronbichlerKormann2019,KronbichlerWall2018}; this solver uses discontinuous elements of
 degree $5$ in a geometric multigrid (GMG) scheme. The comparison between the previous release 9.1
@@
 \todo[inline]{Timo added the paragraph above. Thoughts on how much we should mention (none of this worked with 9.1)?}
+
+\todo[inline]{Martin: I think (ii) worked in 9.1/9.0/8.5, and (iv) definitely worked in the uniform case for 8.5, see Fig 1 in that paper, data line `8B cells'.
+We just happened to break things in between.
+But ``fixes'' is a good wording here because it doesn't exclude that some problems with those configurations have been solved before.
+So I like the text.}
 

### 2020-06-02 9.2/paper.tex
Fix python code indentation broken 9cd9120
fix large scale Stokes section, remove todo

@@
 (i) Bug fixes to concurrent point to point communications.
-(ii) Fixes multigrid transfer with adaptive refinement and more than \num{4e9} unknowns.
+(ii) Fixes to multigrid transfer with adaptive refinement and more than \num{4e9} unknowns.
 (iii) Fixes to index sets in block indices with more than \num{4e9} unknowns.
@@
 For more details, see \cite{clevenger_stokes19}.
-
-\todo[inline]{Timo added the paragraph above. Thoughts on how much we should mention (none of this worked with 9.1)?}
-
-\todo[inline]{Martin: I think (ii) worked in 9.1/9.0/8.5, and (iv) definitely worked in the uniform case for 8.5, see Fig 1 in that paper, data line `8B cells'.
-We just happened to break things in between.
-But ``fixes'' is a good wording here because it doesn't exclude that some problems with those configurations have been solved before.
-So I like the text.}
 
@@
     for face in cell.faces():
-  if face.at_boundary() and\
-  face.boundary_id == 1:
-  cell.refine_flag ='isotropic'
+        if face.at_boundary() and face.boundary_id == 1:
+            cell.refine_flag = 'isotropic'
 

### 2020-06-03 9.2/paper.tex
leave a todo
Fix some typos

@@
  \affil[1]{Computational Engineering and Energy Sciences Group,
-   Computional Sciences and Engineering Division,
+   Computational Sciences and Engineering Division,
    Oak Ridge National Laboratory, 1 Bethel Valley Rd.,
@@
 After this step, the numerical equivalent of the individual dependent expressions
-may finally be retreived from the optimizer.
+may finally be retrieved from the optimizer.
 
@@
 serialized triangulation can also be used in the construction of
-\texttt{parallel::shared}, \texttt{parallel::disrtibuted}, and
+\texttt{parallel::shared}, \texttt{parallel::distributed}, and
 \texttt{parallel::fullydistributed} triangulations (see also Section~\ref{subsec:pft}).
@@
         have been deprecated. As discussed in
-        Subection~\ref{subsec:performance}, \dealii{} by default no longer
+        Subsection~\ref{subsec:performance}, \dealii{} by default no longer
         stores information for all processes on all processes, but only
@@
 
+\todo[inline]{Timo needs to fix his support}
+
 W.~Bangerth and T.~Heister were also partially supported by award DMS-1821210

### 2020-06-03 9.2/paper.tex
add todo
fix support for Timo, Conrad, and Wolfgang

@@
 \todo[inline]{We picked up gingko. Anything else?}
+\todo[inline]{Timo says: should we remove deprecated modules already? (nanoflann, netcdf)}
 \begin{multicols}{3}
@@
 W.~Bangerth, T.~C.~Clevenger, and T.~Heister were partially
-supported by the National Science Foundation under award OAC-1835673
-as part of the Cyberinfrastructure for Sustained Scientific Innovation (CSSI)
-program  and by the Computational Infrastructure
+supported by the Computational Infrastructure
 in Geodynamics initiative (CIG), through the National Science
@@
 
-\todo[inline]{Timo needs to fix his support}
-
-W.~Bangerth and T.~Heister were also partially supported by award DMS-1821210
+W.~Bangerth was also partially supported by award OAC-1835673 as part of the Cyberinfrastructure for Sustained Scientific Innovation (CSSI)
+program, DMS-1821210,
 and EAR-1925595.
@@
 
+T.~C.~Clevenger was also partially supported EAR-1925575 and OAC-2015848.
+
 A.~V.~Grayver was partially supported by the European Space Agency
@@
 
-T.~Heister was also partially supported by NSF Award DMS-1522191, and
-by Technical Data Analysis, Inc. through US Navy SBIR N16A-T003.
+Timo Heister was also partially supported by the National Science Foundation (NSF)
+Award DMS-2028346, OAC-2015848, EAR-1925575, and by
+Technical Data Analysis, Inc. through US Navy STTR Contract N68335-18-C-0011.
 
@@
 
+Clemson University is acknowledged for generous allotment of compute time on Palmetto cluster.
+
 \bibliography{paper}{}

### 2020-06-04 9.2/paper.tex
Resolve one TODO.

@@
 \begin{itemize}
-  \item xy \todo[inline]{Update once we have the subsections in Section 2;
-          provide cross-references to each of these subsections}
-        %   \item Improved support for automatic differentiation (see
-        %     Section~\ref{subsec:ad}),
+  \item A new, parallel, fully distributed triangulation class (see
+        Section~\ref{subsec:pft});
+  \item Substantially improved performance for very large computations
+    on tens or hundreds of thousands of processor cores and up to
+    trillions of unknowns (see
+        Section~\ref{subsec:performance});
+  \item Better support for parallel $hp$-adaptive algorithms (see
+        Section~\ref{subsec:hp});
+  \item Support for particle-based methods (see
+        Section~\ref{subsec:particles});
+  \item Improved performance of the symbolic differentiation framework (see
+        Section~\ref{subsec:symbdiff});
+  \item Advances in SIMD capabilities and the matrix-free infrastructure (see
+        Section~\ref{subsec:mf});
+  \item Advances in GPU support (see
+        Section~\ref{subsec:gpu});
+  \item Better use of modern C++ language features (see
+        Section~\ref{subsec:cxx});
+  \item Seven new tutorial programs (see
+        Section~\ref{subsec:steps}).
 \end{itemize}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Advances of the SIMD capabilities and the matrix-free infrastructure}
+\subsection{Advances in SIMD capabilities and the matrix-free infrastructure}
 \label{subsec:mf}

### 2020-06-07 9.2/paper.tex
Resolve another TODO.
Minor typos fixed

@@
         the possibility of mistakes in the user code.
-        An example is that \texttt{DiscreteTime} ensures that the final time step ends
+        For instance, \texttt{DiscreteTime} ensures that the final time step ends
         precisely on a predefined end time, automatically
-        lengthening or shortening to final time step.
+        lengthening or shortening the final time step.
   \item A key component of \dealii{} are the \texttt{FEValues} and
@@
 
-\todo[inline]{Luca: Your section}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 compiled executable in order to avoid unnecessary initialization during
-runtime. C++11 and later standards enables such computations by
+runtime. C++11 and later standards enable such computations by
 marking variables and functions with the \texttt{constexpr} keyword.

### 2020-06-08 9.2/paper.tex
Distinguish const vs non-const interface
also acknowledge TACC for frontera access

@@
         time-stepping algorithms within time-dependent simulations.
-        The interface of this class is designed to be minimal
+        While providing a rich read-only interface, the non-const interface
+        of this class is designed to be minimal
         to enforce a number of important programming invariants, reducing
@@
 
-Clemson University is acknowledged for generous allotment of compute time on Palmetto cluster.
+The authors acknowledge the Texas Advanced Computing Center (TACC) at The
+University of Texas at Austin for providing access to HPC resources that have
+contributed to the research results reported within this paper.
+
+Clemson University is acknowledged for generous allotment of compute time on
+Palmetto cluster.
 

### 2020-06-10 9.2/paper.tex
Discuss the new code gallery examples.

@@
 
-\todo[inline]{Do we have new code gallery programs?}
+There are also new programs in the code gallery (a collection of
+user-contributed programs that often solve more complicated problems
+than tutorial programs, and intended as starting points for further
+research rather than as teaching tools):
 \begin{itemize}
-  \item Parallel in time (\url{https://github.com/dealii/code-gallery/pull/49})
-  \item Linear elastoplasticity (\url{https://github.com/dealii/code-gallery/pull/62})
-  \item Thermo-electro-elastic pump (\url{https://github.com/dealii/code-gallery/pull/63})
-  \item Nonlinear poro-viscoelasticity (\url{https://github.com/dealii/code-gallery/pull/72})
+  \item A program based on the XBraids library to solve time-dependent
+    problems in a parallel-in-time fashion. This program was
+    contributed by Joshua Christopher.
+  \item A program that solves the biphasic nonlinear
+    poro-viscoelasticity equations based on Ogden hyperelasticity, to
+    explore the porous and viscous contributions in brain mechanics
+    applications.  This program was contributed by Ester Comellas and
+    Jean-Paul Pelteret. The program also serves as a demonstration of
+    the automatic differentiation capabilities of \dealii{}.
 \end{itemize}
@@
 Fabian Castelli,
+Joshua Christopher,
+Ester Comellas,
 Katherine Cosburn,
@@
 Clemson University is acknowledged for generous allotment of compute time on
-Palmetto cluster.
+the Palmetto cluster.
 

### 2020-06-11 9.2/paper.tex
Deal with deprecated interfaces.
Fix the author list.

@@
      Thomas C. Clevenger,
-%     Denis Davydov,
     Marc Fehling,
-%     Daniel Garcia-Sanchez,
     Alexander V. Grayver,
-%     Graham Harper,
     Timo Heister,
@@
     Martin Kronbichler,
+    Matthias Maier,
     Peter Munch,
-%     Ross Maguire Kynch,
-    Matthias Maier,
     Jean-Paul Pelteret,
     Reza Rastak,
+    Ignacio Thomas,
     Bruno Turcksin,
@@
 \dealii{} can interface with many other libraries:
-\todo[inline]{We picked up gingko. Anything else?}
-\todo[inline]{Timo says: should we remove deprecated modules already? (nanoflann, netcdf)}
 \begin{multicols}{3}
@@
     \item muparser \cite{muparser-web-page}
-    \item nanoflann \cite{nanoflann}
+    \item nanoflann \cite{nanoflann}%
           %
@@
 \end{multicols}
-Please consider citing the appropriate references if you use interfaces to these
-libraries.
+Please consider citing the appropriate references if you use
+interfaces to these libraries. We note that the nanoflann and NetCDF
+interfaces are now deprecated and will be removed in \dealii{} version
+9.3.
 
@@
 Niklas Fehn,
+Isuru Fernando,
 Rebecca Fildes,

### 2020-06-15 9.2/paper.tex
Add another name to the Acknowledgments section

@@
      Thomas C. Clevenger,
+%     Denis Davydov,
     Marc Fehling,
+%     Daniel Garcia-Sanchez,
     Alexander V. Grayver,
+%     Graham Harper,
     Timo Heister,
@@
     Martin Kronbichler,
+    Peter Munch,
+%     Ross Maguire Kynch,
     Matthias Maier,
-    Peter Munch,
     Jean-Paul Pelteret,
     Reza Rastak,
-    Ignacio Thomas,
     Bruno Turcksin,

### 2020-06-16 9.3/taskflow-vs-tbb/2d/step-9.cc
Add source files for the experiment.

@@ -1,0 +1,2142 @@
+/* ---------------------------------------------------------------------
+ *
+ * Copyright (C) 2000 - 2020 by the deal.II authors
+ *
+ * This file is part of the deal.II library.
+ *
+ * The deal.II library is free software; you can use it, redistribute
+ * it, and/or modify it under the terms of the GNU Lesser General
+ * Public License as published by the Free Software Foundation; either
+ * version 2.1 of the License, or (at your option) any later version.
+ * The full text of the license can be found in the file LICENSE.md at
+ * the top level directory of deal.II.
+ *
+ * ---------------------------------------------------------------------
+
+ *
+ * Author: Wolfgang Bangerth, University of Heidelberg, 2000
+ */
+
+
+// Just as in previous examples, we have to include several files of which the
+// meaning has already been discussed:
+#include <deal.II/base/quadrature_lib.h>
+#include <deal.II/base/function.h>
+#include <deal.II/base/logstream.h>
+#include <deal.II/lac/vector.h>
+#include <deal.II/lac/full_matrix.h>
+#include <deal.II/lac/sparse_matrix.h>
+#include <deal.II/lac/dynamic_sparsity_pattern.h>
+#include <deal.II/lac/solver_gmres.h>
+#include <deal.II/lac/precondition.h>
+#include <deal.II/lac/affine_constraints.h>
+#include <deal.II/grid/tria.h>
+#include <deal.II/grid/grid_generator.h>
+#include <deal.II/grid/grid_refinement.h>
+#include <deal.II/grid/tria_accessor.h>
+#include <deal.II/grid/tria_iterator.h>
+#include <deal.II/dofs/dof_handler.h>
+#include <deal.II/dofs/dof_accessor.h>
+#include <deal.II/dofs/dof_tools.h>
+#include <deal.II/fe/fe_values.h>
+#include <deal.II/numerics/vector_tools.h>
+#include <deal.II/numerics/matrix_tools.h>
+#include <deal.II/numerics/data_out.h>
+#include <deal.II/fe/fe_q.h>
+#include <deal.II/grid/grid_out.h>
+
+// The following two files provide classes and information for multithreaded
+// programs. In the first one, the classes and functions are declared which we
+// need to do assembly in parallel (i.e. the
+// <code>WorkStream</code> namespace). The
+// second file has a class MultithreadInfo which can be used to query the
+// number of processors in your system, which is often useful when deciding
+// how many threads to start in parallel.
[…]

### 2020-06-16 9.3/taskflow-vs-tbb/3d/step-9.cc

@@ -1,0 +1,2142 @@
+/* ---------------------------------------------------------------------
+ *
+ * Copyright (C) 2000 - 2020 by the deal.II authors
+ *
+ * This file is part of the deal.II library.
+ *
+ * The deal.II library is free software; you can use it, redistribute
+ * it, and/or modify it under the terms of the GNU Lesser General
+ * Public License as published by the Free Software Foundation; either
+ * version 2.1 of the License, or (at your option) any later version.
+ * The full text of the license can be found in the file LICENSE.md at
+ * the top level directory of deal.II.
+ *
+ * ---------------------------------------------------------------------
+
+ *
+ * Author: Wolfgang Bangerth, University of Heidelberg, 2000
+ */
+
+
+// Just as in previous examples, we have to include several files of which the
+// meaning has already been discussed:
+#include <deal.II/base/quadrature_lib.h>
+#include <deal.II/base/function.h>
+#include <deal.II/base/logstream.h>
+#include <deal.II/lac/vector.h>
+#include <deal.II/lac/full_matrix.h>
+#include <deal.II/lac/sparse_matrix.h>
+#include <deal.II/lac/dynamic_sparsity_pattern.h>
+#include <deal.II/lac/solver_gmres.h>
+#include <deal.II/lac/precondition.h>
+#include <deal.II/lac/affine_constraints.h>
+#include <deal.II/grid/tria.h>
+#include <deal.II/grid/grid_generator.h>
+#include <deal.II/grid/grid_refinement.h>
+#include <deal.II/grid/tria_accessor.h>
+#include <deal.II/grid/tria_iterator.h>
+#include <deal.II/dofs/dof_handler.h>
+#include <deal.II/dofs/dof_accessor.h>
+#include <deal.II/dofs/dof_tools.h>
+#include <deal.II/fe/fe_values.h>
+#include <deal.II/numerics/vector_tools.h>
+#include <deal.II/numerics/matrix_tools.h>
+#include <deal.II/numerics/data_out.h>
+#include <deal.II/fe/fe_q.h>
+#include <deal.II/grid/grid_out.h>
+
+// The following two files provide classes and information for multithreaded
+// programs. In the first one, the classes and functions are declared which we
+// need to do assembly in parallel (i.e. the
+// <code>WorkStream</code> namespace). The
+// second file has a class MultithreadInfo which can be used to query the
+// number of processors in your system, which is often useful when deciding
+// how many threads to start in parallel.
[…]

### 2020-07-02 9.2/paper.tex
9.2 revision without DoE statement

--- 2020-06-11 9.2/paper.tex
@@
   $^\ast$ This manuscript has been authored by UT-Battelle, LLC under Contract No.
-  DE-AC05-00OR22725 with the U.S. Department of Energy. The United States
-  Government retains and the publisher, by accepting the article for
-  publication, acknowledges that the United States Government retains a
-  non-exclusive, paid-up, irrevocable, worldwide license to publish or reproduce
-  the published form of this manuscript, or allow others to do so, for United
-  States Government purposes. The Department of Energy will provide public
-  access to these results of federally sponsored research in accordance with the
-  DOE Public Access Plan (http://energy.gov/downloads/doe-public-access-plan).}
+  DE-AC05-00OR22725 with the U.S. Department of Energy.
+  %The United States
+  %Government retains and the publisher, by accepting the article for
+  %publication, acknowledges that the United States Government retains a
+  %non-exclusive, paid-up, irrevocable, worldwide license to publish or reproduce
+  %the published form of this manuscript, or allow others to do so, for United
+  %States Government purposes. The Department of Energy will provide public
+  %access to these results of federally sponsored research in accordance with the
+  %DOE Public Access Plan (http://energy.gov/downloads/doe-public-access-plan).
+}
 

### 2021-02-12 9.3/paper.tex
Add discussion related simplex, multigrid, MatrixFree, FEPointValues

--- 2020-07-02 9.2/paper.tex
@@
 \newcommand{\autoconf}{{\specialword{autoconf}}\xspace}
+
+\usetikzlibrary{shapes.misc}
+\tikzset{cross/.style={cross out, draw=black, minimum size=2*(#1-\pgflinewidth), inner sep=0pt, outer sep=0pt},
+%default radius will be 1pt. 
+cross/.default={2pt}}
 
@@
   },
-  pdftitle={The deal.II Library, Version 9.2, 2020},
+  pdftitle={The deal.II Library, Version 9.3, 2021},
 }
 
-\title{The \dealii{} Library, Version 9.2}
+\title{The \dealii{} Library, Version 9.3}
 
@@
 \begin{itemize}
-  \item \dealii{} had decent support for solving complex-valued problems
-        for a while already
-        (e.g., ones in quantum mechanics -- like the equation used in the
-        \texttt{step-58} tutorial program covered below -- or for
-        time-harmonic problems). However, there were two
-        areas in which support was missing. First, the UMFPACK direct solver
-        packaged with \dealii{} did not support solving complex-valued
-        linear problems. This has now been addressed: UMFPACK actually can
-        solve such systems, we just needed to write the appropriate
-        interfaces. Second, the \texttt{DataOut} class that is responsible
-        for converting nodal data into information that can then be written
-        into files for visualization did not know how to deal with vector-
-        and tensor-valued fields whose components are complex numbers. An
-        example for this is to solve the time-harmonic version of the
-        Maxwell equations that has the electric and magnetic fields as
-        solution. This, too, has been addressed in this release.
-  \item The new \texttt{DiscreteTime} class provides a more
-        consistent, more readable, and less error-prone approach to control
-        time-stepping algorithms within time-dependent simulations.
-        While providing a rich read-only interface, the non-const interface
-        of this class is designed to be minimal
-        to enforce a number of important programming invariants, reducing
-        the possibility of mistakes in the user code.
-        For instance, \texttt{DiscreteTime} ensures that the final time step ends
-        precisely on a predefined end time, automatically
-        lengthening or shortening the final time step.
-  \item A key component of \dealii{} are the \texttt{FEValues} and
-        \texttt{FEFaceValues} classes that evaluate finite element functions
-        at quadrature points located on cells and faces of a cell,
-        respectively~\cite{BangerthHartmannKanschat2007}. This release now
-        contains a class \texttt{FEInterfaceValues} that considers the
-        restriction of the shape functions from \textit{both} sides of a face and
-        allows evaluating jumps and averages of shape functions along this
-        face. These are common components of the bilinear forms of
-        discontinuous Galerkin schemes (as well as schemes for fourth-order
-        equations, see the discussion of \texttt{step-47} below) and greatly
-        simplify the implementation of these methods.
-  \item Previously, the \texttt{parallel::distributed::ContinuousQuadratureDataTransfer}
-        class, which transfers local quadrature point data onto the
-        children of newly refined cells, did not allow different cells
-        to store quadrature data of different lengths. This release lifts
-        this restriction, which enables efficient quadrature data storage and data
-        transfer within a triangulation containing multiple material models, each with
-        their own number of local state variables and history variables.
-        As an example, in a solid mechanics simulation, we can assign a hyper-elastic
-        material model to a region of the mesh with no associated
-        history variables, while in another part of the triangulation we incorporate an
-        elasto-plastic constitutive model, requiring the storage of local plasticity
-        data at quadrature points.
-        During each refinement of the mesh, we can then use
-        \texttt{ContinuousQuadratureDataTransfer} to transfer and interpolate
-        plasticity data from parent cells to their children, ignoring the cells
-        with the hyper-elastic constitutive model.
+  \item globally unique cell IDs
 \end{itemize}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{A new fully distributed triangulation class}
-\label{subsec:pft}
-
-Previously, all triangulation classes of \dealii{} had in common that the coarse grid is replicated by
-all processes in a parallel environment, and the actual mesh used for computations is constructed by repeated
-refinement. However, this has its limitations in many
-applications where the mesh comes
-from an external mesh generator in the form of a file that frequently
-already contains millions
-or tens of millions of cells. For such configurations, applications might exhaust
-available memory already while reading the mesh on each MPI process.
-
-The new \texttt{parallel::fullydistributed::Triangulation} class targets this issue
-by distributing also the coarse grid. Such
-a triangulation can be created by providing to each process a \texttt{Triangulation\-De\-scrip\-tion::Description} struct, containing
-(i) the relevant data to construct the local part of the coarse grid, (ii) the
-translation of the local coarse-cell IDs to globally unique IDs, (iii) the hierarchy
-of mesh refinement steps, and (iv) the owner of the cells on the active mesh level as well
-as on the multigrid levels. For the current release, triangulations set up
-this way cannot be adaptively refined after construction, though we plan to
-improve this for the next release.
-
-%The \texttt{TriangulationDescription::Description} struct can be filled manually or
-%by the utility functions from the \texttt{TriangulationDescription::Utilities}
-%namespace. The function \texttt{create\_\allowbreak description\_\allowbreak
-%from\_ \allowbreak triangulation()} can convert a base triangulation (partitioned
-%serial \texttt{Tri\-angulation} and \texttt{parallel::distributed::Triangulation})
-%to such a struct. The advantage of this approach is that all known utility
-%functions from the namespaces \texttt{GridIn} and \texttt{GridTools} can be used
-%on the base triangulations before converting them to the structs. Since this
-%function suffers from the same main memory problems as described above, we also
-%provide the function \texttt{create\_description\_from\_triangulation\_in\_groups()},
-%which creates the structs only on the master process in a process group. These
-%structs are filled one by one and are sent to the relevant processes once they are
-%ready. A sensible process group size might contain all processes of one compute node.
-
-The new fully distributed triangulation class supports 1D, 2D, and 3D meshes
-including geometric multigrid hierarchies, periodic boundary conditions, and
-hanging nodes.
-
-%We intend to extend the usability of the new triangulation class in regard of
-%different aspects, e.g., I/O. In addition, we would like to enable adaptive mesh
-%refinement, a feature of the other triangulation classes, which is very much
-%appreciated by many users. For repartitioning, we plan to use an oracle approach
-%known from \texttt{parallel::distributed::Triangulation}. Here, we would like to
-%rely on a user-provided partitioner, which might also be a graph partitioner.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Improved large-scale performance}
-\label{subsec:performance}
-
-Large-scale simulations with up to 304,128 cores have revealed bottlenecks in release
-9.1 during initialization of a number of distributed data structures, due to the usage of expensive collective operations
-like \texttt{MPI\_Allgather()} and \texttt{MPI\_\allowbreak
-  Alltoall()}. Typical examples are the
-pre-computation of the indices of those vector entries (or
-other linear index ranges) owned by
-each process, which were previously stored in  an array on every process.
-This information is needed to set up
-the  \texttt{Utilities::MPI::Par\-ti\-ti\-oner} class.
-In release 9.2, we have replaced these functions in favor of
-consensus algorithms~\cite{hoefler2010scalable}, which can be
-found in the namespace \texttt{Utilities::\allowbreak MPI::\allowbreak ConsensusAlgorithms} (short: \texttt{CA}).
-Now, only the locally relevant information about the index ranges is
-(re)computed when needed, which, for more than 100 MPI processes, uses
-point-to-point communications and a single \texttt{MPI\_IBarrier()}.
-
-%Consensus algorithms are algorithms dedicated to efficient dynamic-sparse
-%communication patterns. In this context, the term ``dynamic-sparse'' means
-%that by the time this function is called, the other processes do not know
-%yet that they have to answer requests and
-%each process only has to communicate with a small subset of processes of the
-%MPI communicator. We provide two flavors of the consensus algorithm: the two-step
-%approach \texttt{ConsensusAlgorithms::PEX} and the \texttt{ConsensusAlgorithms::NBX},
-%which uses only point-to-point communications and a single \texttt{MPI\_IBarrier()}.
-%The class \texttt{ConsensusAlgorithms::Selector} selects one of the two previous
-%algorithms, depending on the number of processes.
-
-Users can apply the new algorithms for their own dynamic-sparse problems by
-providing a list of target
-processes and pack/unpack routines either by implementing the interface
-\texttt{CA::\allowbreak Process} or by providing \texttt{std::function}
-objects to \texttt{CA::AnonymousProcess}.
-
-By replacing the collective communications during set up and removing the arrays
-that contain information for each process (enabled by the application of consensus
-algorithms and other modifications---a full list of modifications leading to this
-improvement can be found online), we were able to significantly
-improve the set up
-time for large-scale simulations and to solve a Poisson problem with multigrid
-with \num{2.1e12} unknowns on the SuperMUC-NG supercomputer with 304,152
-cores~\cite{dealII2020design,Arndt2020sppexa}.
-Figure~\ref{fig:init_costs} compares timings of simulations of various problem
-sizes (including set up) on 49,152 MPI ranks using a matrix-free
-solver~\cite{Arndt2020sppexa,KronbichlerKormann2019,KronbichlerWall2018}; this solver uses discontinuous elements of
-degree $5$ in a geometric multigrid (GMG) scheme. The comparison between the previous release 9.1
-and the current release 9.2 shows that while the scaling for the V-cycle had been
-very good before, many initialization routines have been considerably
-improved, especially the enumeration of unknowns on the multigrid levels and
-the setup of the multigrid transfer.
-
-\pgfplotstableread{
-  dofs  grid  distdofs  mfsetup  mgtransfer  smoother  vcycle
-  14155776  0.347457  0.02424084  0.03346427  0.0105426  0.0108817  0.0035033
-  28311552  0.35134  0.0215024  0.0340963  0.0095201  0.010616  0.0035887
-  56623104  0.41577  0.0243593  0.0432297  0.011557  0.011667  0.0041244
-  113246208  0.42172  0.0323919  0.0427806  0.012983  0.011834  0.0043853
-  226492416  0.43565  0.0404079  0.0421606  0.012867  0.017305  0.0056340
-  452984832  0.5055  0.0500101  0.0562977  0.017739  0.022763  0.0061992
-  905969664  0.54552  0.065865  0.0641994  0.019445  0.037431  0.0087932
-  1811939328  0.5672  0.104909  0.080807  0.025395  0.063548  0.0133252
-  3623878656  0.65541  0.165071  0.129717  0.042294  0.12644  0.0218619
-  7247757312  0.74521  0.251417  0.221072  0.073314  0.21807  0.0339264
-  14495514624  0.84789  0.403463  0.36894  0.11425  0.4091  0.0669798
-  28991029248  0.91983  0.52435  0.5363  0.14189  0.79178  0.1322448
-  57982058496  1.0339  0.87448  0.93134  0.23863  1.5896  0.2539445
-  115964116992  1.2667  1.46016  1.70806  0.44101  3.1409  0.4956625
-  231928233984  1.7597  2.58121  3.3741  0.77624  6.2628  0.9857200
-}\tableinit
-
-\pgfplotstableread{
-  dofs  grid  distdofs  mfsetup  mgtransfer  smoother  vcycle
-  14155776  2.0148  0.9086  0.433886  1.3099  0.010685  0.0035033
-  28311552  1.8746  0.91078  0.576777  1.3093  0.010752  0.0035887
-  56623104  2.0924  0.98058  0.574183  1.4459  0.011831  0.0041244
-  113246208  2.3788  1.22025  0.547492  1.7919  0.011979  0.0043853
-  226492416  2.1254  1.50329  0.839219  1.7832  0.017819  0.0056340
-  452984832  2.4325  2.10171  0.711708  2.0263  0.023393  0.0061992
-  905969664  3.8911  3.58355  1.08478  2.5844  0.045697  0.0087932
-  1811939328  4.1803  4.379  0.784804  1.8947  0.061552  0.0133252
-  3623878656  2.6654  5.6073  0.833896  1.8912  0.1188  0.0218619
-  7247757312  2.7422  7.8463  0.741733  1.8477  0.21291  0.0339264
-  14495514624  2.765  9.9848  0.99218  1.8248  0.40665  0.0669798
-  28991029248  3.2901  15.0526  1.5124  2.6723  0.81013  0.1322448
-  57982058496  3.1927  21.0562  1.72831  2.1998  1.5905  0.2539445
-  115964116992  4.1485  31.0087  2.67998  2.9091  3.2009  0.4956625
-  231928233984  4.1787  47.195  4.2821  3.9558  6.3362  0.9857200
-}\tableinitold
-
-\begin{figure}
-  \begin{tikzpicture}
-    \begin{loglogaxis}[
-        width=0.51\textwidth,
-        height=0.38\textwidth,
-        xlabel={\# DoFs},
-        ylabel={time [s]},
-        title={9.1 release},
-        xtick={1.4e7,5.7e7,2.3e8,9.1e8,3.6e9,1.4e10,5.8e10,2.3e11},
-        xticklabels={1.4$\times 10^7$,5.7$\times 10^7$,2.3$\times 10^8$,9.1$\times 10^8$,3.6$\times 10^9$,1.4$\times 10^{10}$,5.8$\times 10^{10}$,2.3$\times 10^{11}$},
-        tick label style={font=\scriptsize},
-        xticklabel style={rotate=25,anchor=north east},
-        label style={font=\scriptsize},
-        legend style={font=\scriptsize},
-        %legend to name=legend:scalingMG,
-        %legend cell align=left,
-        %legend columns = 3,
-        ymin=2e-3,ymax=8e1,
-        grid, thick]
-      \addplot table[x index = 0, y index={1}] {\tableinitold};
-      \addplot[red,mark=square*] table[x index = 0, y index={2}] {\tableinitold};
-      \addplot[gnuplot@green,mark=otimes] table[x index = 0, y index={3}] {\tableinitold};
-      \addplot[gnuplot@lightblue,mark=triangle] table[x index = 0, y index={4}] {\tableinitold};
-      \addplot[gnuplot@purple,mark=diamond] table[x index = 0, y index={5}] {\tableinitold};
-      \addplot[black,mark=x] table[x index = 0, y index={6}] {\tableinitold};
-    \end{loglogaxis}
+\newpage
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Simplex and mixed mesh support}
+\label{subsec:simplex}
+
+\begin{figure}[!h]
+
+\centering
+
+\begin{tikzpicture}[scale=1.7]
+
+\coordinate (0) at (0,0);
+\coordinate (1) at (1,0);
+\coordinate (2) at (0,1);
+
+\coordinate (3) at (0,0.5);
+\coordinate (4) at (0.5,0.5);
+\coordinate (5) at (0.5,0);
+
+\foreach \i in {0,1, 2}
+    \fill (\i) circle (1.1pt) node [below] {};
+
+\foreach \i in {3,4,5}
+    \draw (\i) node[cross, draw=red, rotate=-20] {};
+
+\draw (0) --(1) -- (2) -- (0);
+
 \end{tikzpicture}
-  \hfill
-  \begin{tikzpicture}
-    \begin{loglogaxis}[
-        width=0.51\textwidth,
-        height=0.38\textwidth,
-        xlabel={\# DoFs},
-        ylabel={time [s]},
-        title={9.2 release},
-        xtick={1.4e7,5.7e7,2.3e8,9.1e8,3.6e9,1.4e10,5.8e10,2.3e11},
-        xticklabels={1.4$\times 10^7$,5.7$\times 10^7$,2.3$\times 10^8$,9.1$\times 10^8$,3.6$\times 10^9$,1.4$\times 10^{10}$,5.8$\times 10^{10}$,2.3$\times 10^{11}$},
-        tick label style={font=\scriptsize},
-        xticklabel style={rotate=25,anchor=north east}, thick,
-        label style={font=\scriptsize},
-        legend style={font=\scriptsize},
-        legend to name=legend:init,
-        legend cell align=left,
-        legend columns = 6,
-        ymin=2e-3,ymax=8e1,
-        grid]
-      \addplot table[x index = 0, y index={1}] {\tableinit};
-      \addlegendentry{create mesh};
-      \addplot[red,mark=square*] table[x index = 0, y index={2}] {\tableinit};
-      \addlegendentry{enumerate (mg) dofs};
-      \addplot[gnuplot@green,mark=otimes] table[x index = 0, y index={3}] {\tableinit};
-      \addlegendentry{init matrix-free};
-      \addplot[gnuplot@lightblue,mark=triangle] table[x index = 0, y index={4}] {\tableinit};
-      \addlegendentry{init GMG transfer};
-      \addplot[gnuplot@purple,mark=diamond] table[x index = 0, y index={5}] {\tableinit};
-      \addlegendentry{init smoother};
-      \addplot[black,mark=x] table[x index = 0, y index={6}] {\tableinit};
-      \addlegendentry{GMG V-cycle};
-    \end{loglogaxis}
+\begin{tikzpicture}[scale=1.7]
+
+\coordinate (0) at (0,0);
+\coordinate (1) at (1,0);
+\coordinate (2) at (1,1);
+\coordinate (3) at (0,1);
+
+\coordinate (4) at (0.5,0,0);
+\coordinate (5) at (1.0,0.5);
+\coordinate (6) at (0.0,0.5);
+\coordinate (7) at (0.5,1.0);
+\coordinate (8) at (0.5,0.5);
+
+\foreach \i in {0,1, 2, 3}
+    \fill (\i) circle (1.1pt) node [below] {};
+
+\foreach \i in {4,5,6,7,8}
+    \draw (\i) node[cross, draw=red, rotate=-20] {};
+
+\draw (0) --(1) -- (2) -- (3) -- (0);
+
 \end{tikzpicture}
-  \\
-  \strut\hfill\pgfplotslegendfromname{legend:init}\hfill\strut
-  \caption{\it Comparison of initialization costs of various data structures in the 9.1 release (left) and the new 9.2 release (right) when run on 49,152 MPI ranks.}
-  \label{fig:init_costs}
+\qquad\qquad
+\begin{tikzpicture}[scale=1.3]
+
+\coordinate (0) at (0,0,0);
+\coordinate (1) at (1,0,0);
+\coordinate (2) at (0,1,0);
+\coordinate (3) at (0,0,1);
+
+\coordinate (4) at (0.5,0,0);
+\coordinate (5) at (0,0.5,0);
+\coordinate (6) at (0,0,0.5);
+\coordinate (7) at (0.5,0.5,0.0);
+\coordinate (8) at (0.5,0.0,0.5);
+\coordinate (9) at (0.0,0.5,0.5);
+
+\foreach \i in {0,1,2,3}
+    \fill (\i) circle (1.5pt) node [below] {};
+
+\foreach \i in {4,5,6,7,8,9}
+    \draw (\i) node[cross, draw=red, rotate=-20] {};
+
+\draw (0) -- (1) -- (2) -- (0);
+\draw (0) -- (1) -- (3) -- (0);
+\draw (1) -- (2) -- (3) -- (1);
+
+\end{tikzpicture}
+\begin{tikzpicture}[scale=1.5]
+
+\coordinate (0) at (0.0, 0.0, 0.0);
+\coordinate (1) at (0.0, 0.0, 1.0);
+\coordinate (2) at (1.0, 0.0, 0.0);
+\coordinate (3) at (1.0, 0.0, 1.0);
+\coordinate (4) at (0.5, 1.0, 0.5);
+
+\coordinate  (5) at (0.5, 0.0, 0.0);
+\coordinate  (6) at (0.0, 0.0, 0.5);
+\coordinate  (7) at (0.5, 0.0, 1.0);
+\coordinate  (8) at (1.0, 0.0, 0.5);
+
+\coordinate  (9) at (0.25,0.5,0.25);
+\coordinate (10) at (0.25,0.5,0.75);
+\coordinate (11) at (0.75,0.5,0.25);
+\coordinate (12) at (0.75,0.5,0.75);
+
+\coordinate (13) at (0.5, 0.0, 0.5);
+
+\foreach \i in {0,1,2,3,4}
+    \fill (\i) circle (1.3pt) node [below] {};
+
+%\foreach \i in {5, 6, 7, 8, 9, 10, 11, 12, 13}
+%    \draw (\i) node[cross] {};
+
+\draw (0) -- (1) -- (3) -- (2) -- (0);
+\draw (1) -- (4) -- (3);
+\draw (0) -- (4) -- (2);
+
+\end{tikzpicture}
+\begin{tikzpicture}[scale=1.3]
+
+\coordinate (0) at (0,0,1.5);
+\coordinate (1) at (1,0,1.5);
+\coordinate (2) at (0.0,1.0,1.5);
+
+\coordinate (3) at (0,0,0);
+\coordinate (4) at (1,0,0);
+\coordinate (5) at (0.0,1.0,0);
+
+\coordinate  (6) at (0.5, 0.0, 0.0);
+\coordinate  (7) at (0.0, 0.5, 0.0);
+\coordinate  (8) at (0.5, 0.5, 0.0);
+
+\coordinate   (9) at (0.5, 0.0, 0.75);
+\coordinate  (10) at (0.0, 0.5, 0.75);
+\coordinate  (11) at (0.5, 0.5, 0.75);
+
+\coordinate  (12) at (0.5, 0.0, 1.5);
+\coordinate  (13) at (0.0, 0.5, 1.5);
+\coordinate  (14) at (0.5, 0.5, 1.5);
+
+\foreach \i in {0,1,2,3,4,5}
+    \fill (\i) circle (1.5pt) node [below] {};
+
+\foreach \i in {6,7,8,9,10,11,12,13,14}
+    \draw (\i) node[cross, draw=red, rotate=-20] {};
+
+\draw (0) -- (1) -- (2) -- (0);
+\draw (3) -- (4) -- (5) -- (3);
+\draw (0) -- (3);
+\draw (1) -- (4);
+\draw (2) -- (5);
+
+\end{tikzpicture}
+\begin{tikzpicture}[scale=1.5]
+
+\coordinate (0) at (0,0,1);
+\coordinate (1) at (1,0,1);
+\coordinate (2) at (0,0,0);
+\coordinate (3) at (1,0,0);
+
+\coordinate (4) at (0,1,1);
+\coordinate (5) at (1,1,1);
+\coordinate (6) at (0,1,0);
+\coordinate (7) at (1,1,0);
+
+\coordinate  (8) at (0,0,0.5);
+\coordinate  (9) at (1,0,0.5);
+\coordinate (10) at (0.5,0,1);
+\coordinate (11) at (0.5,0,0);
+\coordinate (12) at (0.5,0,0.5);
+
+\coordinate (13) at (0,1,0.5);
+\coordinate (14) at (1,1,0.5);
+\coordinate (15) at (0.5,1,1);
+\coordinate (16) at (0.5,1,0);
+\coordinate (17) at (0.5,1,0.5);
+
+\coordinate (18) at (0,0.5,1);
+\coordinate (19) at (1,0.5,1);
+\coordinate (20) at (0,0.5,0);
+\coordinate (21) at (1,0.5,0);
+\coordinate (22) at (0,0.5,0.5);
+\coordinate (23) at (1,0.5,0.5);
+\coordinate (24) at (0.5,0.5,1);
+\coordinate (25) at (0.5,0.5,0);
+\coordinate (26) at (0.5,0.5,0.5);
+
+\foreach \i in {0,1,2,3,4,5,6,7}
+    \fill (\i) circle (1.3pt) node [below] {};
+
+\foreach \i in {8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26}
+    \draw (\i) node[cross, draw=red, rotate=-20] {};
+
+\draw (0) -- (1) -- (3) -- (2) -- (0);
+\draw (4) -- (5) -- (7) -- (6) -- (4);
+\draw (0) -- (4);
+\draw (1) -- (5);
+\draw (2) -- (6);
+\draw (3) -- (7);
+
+\end{tikzpicture}
+
+\caption{All reference cells in 2D (triangle, quadrilateral) and 3D (tetrahedron, pyramid, wedge, hexahedron) with the support points indicated both for linear ($\bullet$) and quadratic ($\times$) shape functions.}
+
 \end{figure}
 
-As part of this effort, we ran benchmarks on the TACC Frontera system, where we were
-able to apply the matrix-free geometric multigrid framework to a variable viscosity
-Stokes system and achieved weak and strong scaling up to 114K MPI ranks with up to
-\num{2.1e11} unknowns. This is likely the largest block system currently solved with
-\dealii{} and required various optimizations
-and fixes on top of the ones mentioned above:
-(i) Bug fixes to concurrent point to point communications.
-(ii) Fixes to multigrid transfer with adaptive refinement and more than \num{4e9} unknowns.
-(iii) Fixes to index sets in block indices with more than \num{4e9} unknowns.
-(iv) Fixes to computations with more than \num{4e9} active cells.
-(v) Implementation of IDR(s) solvers to reduce memory overhead.
-For more details, see \cite{clevenger_stokes19}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Better support for parallel $hp$-adaptive algorithms}
-\label{subsec:hp}
-
-Since the previous release, \dealii{} has had support for $hp$-adaptive finite element methods
-on distributed memory systems \cite{dealII91}. We implemented the bare functionality
-for $hp$-adaptive methods with the objective to offer the greatest flexibility in
-their application. Here, reference finite elements still had to be assigned manually
-to each cell, which may not lead to an optimal mesh and is tedious.
-
-With the current release, $hp$-adaptive finite element methods have been further
-expanded: New features like decision strategies have been added and the user interface
-has been overhauled, effectively making $hp$-methods more attractive to use. We introduced
-many new functions that automatize the general workflow for applying $hp$-decision
-strategies, which run on top of the previous low-level implementation for both serial
-and parallel applications.
-
-The interface is now as simple to use as the one for $h$-adaptive mesh refinement.
-Consider the following (incomplete) listing as an example: We estimate both error and
-smoothness of the finite element approximation. Further, we flag certain fractions of
-cells with the highest and lowest errors for refinement and coarsening, respectively
-(here: 30\%/3\%). From those cells listed for adaptation, we designate a subset
-for $h$- and $p$-adaptation. The parameters of the corresponding
-\texttt{hp::Refinement} function specify the fraction of cells to be $p$-adapted from
-those subsets flagged for refinement and coarsening, respectively (here: 90\%/80\%),
-while the remaining fraction will be $h$-adapted (here: 10\%/20\%).
-\begin{c++}
-Vector<float> estimated_error_per_cell(n_active_cells);
-KellyErrorEstimator::estimate(
-  dof_handler, ..., solution, estimated_error_per_cell, ...);
-GridRefinement::refine_and_coarsen_fixed_number(
-  triangulation, estimated_error_per_cell, 0.3, 0.03);
-
-Vector<float> estimated_smoothness_per_cell(n_active_cells);
-SmoothnessEstimator::Legendre::coefficient_decay(
-  ..., dof_handler, solution, estimated_smoothness_per_cell);
-hp::Refinement::p_adaptivity_fixed_number(
-  dof_handler, estimated_smoothness_per_cell, 0.9, 0.8);
-hp::Refinement::choose_p_over_h(dof_handler);
-
-triangulation.execute_coarsening_and_refinement();
-\end{c++}
-
-In particular, we implemented decision strategies based on refinement history and
-smoothness estimation, and made sure that they work for refinement as well as
-coarsening in terms of $h$- and $p$-adaptation in serial and parallel applications.
-
-The former relies on knowing an estimate for the upper error bound \cite[Thm.~3.4]{BabuskaSuri1990}.
-For successive refinements, we can predict how the error will change based on
-current error estimates and adaptation flags. In the next refinement cycle, these
-predicted error estimates allow us to decide whether the choice of adaptation in
-the previous cycle was justified, and provide a criterion for the choice in the next
-cycle \cite{MelenkWohlmuth2001}.
-
-In general, $p$-refinement is favorable over $h$-refinement in smooth regions of
-the finite element approximation \cite[Thm.~3.4]{BabuskaSuri1990}. Thus, estimating
-its smoothness provides a suitable decision indicator for $hp$-adaptation. For this
-purpose, we express the finite element approximation in an orthogonal
-basis of increasing frequency, and consider the decay of their expansion
-coefficients as the estimation of smoothness. This has been implemented for both
-Fourier coefficients \cite{BangerthKayserHerold2007} and Legendre coefficients
-\cite{Mavriplis1994,HoustonSeniorSueli2003,HoustonSueli2005,EibnerMelenk2007}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Support for particle-based methods}
-\label{subsec:particles}
-
-Support for particles was originally introduced in \dealii{} version 9.0. These
-particles can be used as passive tracers, or as part of more complex
-models such as those based on Particle-In-Cells (PIC) approaches \cite{GLHPB18}.
-
-With the current release, support for particles has been further expanded: New parallel
-insertion mechanisms and a basic interface to post-process particles have
-been added, effectively making their usage more flexible and enabling a larger
-range of use cases (such as the immersed boundaries in step-70, see below).
-
-Through the addition of the \texttt{Particles::ParticleHandler::insert\_global\_particles()} member function,
-particles can now be inserted in parallel from a vector of points even if these points do not lie
-on the subdomain from which the insertion is called. This operation requires extensive
-communication between the processes to locate   the MPI process that
-owns the cell in which the particle
-is located. However, it is made significantly faster  through the usage of bounding boxes
-that provide a coarse description of the geometrical shape of each subdomain.
-This function also takes care of transferring the properties attached to the particles to their new owner.
-This new capability enables particle generators that insert particles at the location
-of the support points (\texttt{Particles::Generator::dof\_support\_points()}) and at the quadrature points (\texttt{Particles::Generator::quadrature\_points()}) of a
-possibly non-matching triangulation. Consequently, complex particle patterns can be inserted
-using unstructured grids generated outside of \dealii{}.
-
-To visualize the motion of particles, the \texttt{Particles::DataOut} class was added to the library.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Improved performance of the symbolic differentiation framework}
-\label{subsec:symbdiff}
-
-In the previous release we added support for symbolic expressions, leveraging the
-SymEngine library \cite{symengine-web-page}.
-Although effective, evaluating lengthy expressions could be a bottleneck as this
-was performed using dictionary-based substitution.
-We have improved on this by implementing a \texttt{BatchOptimizer} class in the
-namespace \texttt{Differentiation::SD} that collects several \texttt{Expression}s
-and transforms them in such a way that the equivalent result is returned through
-a quicker code path.
-This may be done by simply using common subexpression elimination (CSE) for the
-dictionary-based expressions, by transformation to a set of nested
-\texttt{std::function} objects (the equivalent to \texttt{SymPy}'s ``lambdify'', with or
-without using CSE), or by offloading these expressions to the \texttt{LLVM}
-just-in-time (JIT) compiler.
-Although each of these features is implemented and tested in the SymEngine
-library itself, the \texttt{BatchOptimizer} class provides both a uniform
-interface to their classes and a convenient interface for scalar expressions,
-as well as tensorial expressions formed using the \texttt{deal.II} tensor and
-symmetric tensor classes.
-It, like the \texttt{Expression} class, is also serializable.
-
-The way the batch optimizer may be employed within a user's code is shown
-in the pseudo-code below.
-As per usual, one would first define some independent variables, and
-subsequently compute some symbolic expressions that are dependent on these
-independent variables.
-These expression could be, for example, scalar  expressions or tensors of
-expressions.
-Instead of evaluating these expressions directly, the user would now create an
-optimizer to evaluate the dependent functions.
-In this example, the selected arithmetic type numerical result will be of type
-\texttt{double}, and the \texttt{LLVM} JIT optimizer will be invoked.
-It will employ common subexpression elimination and aggressive optimizations
-during compilation.
-The user then informs the optimizer of all of the independent variables and the
-dependent expressions, and invokes the optimization process.
-This is an expensive call, as it determines an equivalent code path to evaluate
-all of the dependent functions at once.
-However, in many cases each evaluation has significantly less computational cost
-than evaluating the symbolic expressions directly.
-Evaluation is performed when the user constructs a substitution map, giving each
-independent variable a numerical representation, and passes those to the
-optimizer.
-After this step, the numerical equivalent of the individual dependent expressions
-may finally be retrieved from the optimizer.
-
-\begin{c++}
-using namespace Differentiation::SD;
-
-const Expression x("x");
-const Expression y("y");
-...
-const auto f = calculate_f(x, y, ...); // User function
-const auto g = calculate_g(x, y, ...); // User function
-...
-
-BatchOptimizer<double> optimizer (OptimizerType::llvm,
-                                  OptimizationFlags::optimize_all);
-optimizer.register_symbols(x, y, ...);
-optimizer.register_functions(f, g, ...);
-optimizer.optimize();
-
-const auto substitution_map
-  = make_substitution_map({x, ...}, {y, ...}, ...);
-optimizer.substitute(substitution_map);
-
-const auto result_f = optimizer.evaluate(f);
-const auto result_g = optimizer.evaluate(g);
-\end{c++}
-
-This expense of invoking the optimizer may be offset not only by
-the number of evaluations performed, but also by the amount of reuse each
-instance of a \texttt{BatchOptimizer} has.
-In certain circumstances, this can be maximized by generalizing the way in which
-the dependent expressions are formulated.
-For example, in the context of constitutive modelling the material coefficients
-may be made symbolic rather than encoding these into the dependent expressions
-as numerical values.
-The optimizer may then be used to evaluate an entire family of constitutive laws,
-and not a specific one that describes the response of a single material.
-Thereafter, serializing the optimizer instance and reloading the contents during
-subsequent simulations permits the user to skip the optimization process entirely.
-Serialization also enables these complex expressions to be compiled offline.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Advances in SIMD capabilities and the matrix-free infrastructure}
-\label{subsec:mf}
-
-%\begin{itemize}
-%\item ECL
-%\item VectorizedArrayType
-%\end{itemize}
-
-The class \texttt{VectorizedArray<Number>} is a key component to achieve
-the high node-level performance of the matrix-free algorithms in
-\dealii{}~\cite{KronbichlerKormann2012, KronbichlerKormann2019}. It is a
-wrapper class around a short vector of $n$ entries of type \texttt{Number}
-and maps arithmetic operations to appropriate
-single-instruction/multiple-data (SIMD) concepts by intrinsic functions.
-The class \texttt{VectorizedArray} has been made more user-friendly in this
-release by making it compatible with the STL algorithms found in the header
-\texttt{<algorithm>}. The length of the  vector can now be queried by
-\texttt{VectorizedArray::size()} and its underlying number type by
-\texttt{VectorizedArray::value\_type}. The \texttt{VectorizedArray} class
-now supports range-based iteration over its entries. In addition \dealii{}
-now supports ternary operations on vectorized data, where for a given
-binary comparison operation a true or false value is selected. For example,
-the vectorized equivalent of \texttt{(left < right) ? true\_value :
-  false\_value} can be expressed as
-\begin{c++}
-auto result = compare_and_apply_mask<SIMDComparison::less_than>(
-left, right, true_value, false_value);
-\end{c++}
-which compares every element of the vectorized array individually and
-selects the corresponding value from the \texttt{true\_value}, or
-\texttt{false\_value} array.
-
-In previous \dealii{} releases, the vector length was set at compile time
-of the library to match the highest value supported by the given processor
-architecture. Now, a second optional template argument can be specified as
-\texttt{VectorizedArray<Number, size>}, where \texttt{size} explicitly
-controls the vector length within the capabilities of a particular
-instruction set. (A full list of supported vector lengths is presented in
-Table~\ref{tab:vectorizedarray}.) This allows users to select the vector
-length/ISA and, as a consequence, the number of cells to be processed at
-once in matrix-free operator evaluations. For example, the \dealii{}-based
-library \texttt{hyper.deal}~\cite{munch2020hyperdeal}, which solves the 6D
-Vlasov--Poisson equation with high-order discontinuous Galerkin methods
-(with more than a thousand degrees of freedom per cell), constructs a
-tensor product of two \texttt{MatrixFree} objects of different SIMD-vector
-length in the same application and benefits---in terms of performance---by
-the possibility of decreasing the number of cells processed by a single
-SIMD instruction.
-
-\begin{table}
-  \caption{\it Supported vector lengths of the class \texttt{VectorizedArray} and
-    the corresponding instruction-set-architecture extensions. }\label{tab:vectorizedarray}
+\begin{table}[!h]
+  \caption{\it List of new scalar \texttt{FiniteElement} classes for the new
+  reference-cell types. Vectorial elements can be constructed based on these classes
+  via \texttt{FE\_Systems}.}\label{tab:simplex:fe}
   \centering
-  \begin{tabular}{ccc}
+  \begin{tabular}{cccc}
     \toprule
-    \textbf{double}            & \textbf{float}             & \textbf{ISA}         \\
+    \textbf{reference cell} & \textbf{finite element} & \textbf{dimension} & \textbf{degree}\\
     \midrule
-    VectorizedArray<double, 1> & VectorizedArray<float, 1>  & (auto-vectorization) \\
-    VectorizedArray<double, 2> & VectorizedArray<float, 4>  & SSE2/AltiVec         \\
-    VectorizedArray<double, 4> & VectorizedArray<float, 8>  & AVX/AVX2             \\
-    VectorizedArray<double, 8> & VectorizedArray<float, 16> & AVX-512              \\
+    simplex (line, triangle, tetrahedron) & FE\_P, FE\_DGP  & 1--3 & 1--2 \\
+    pyramid & FE\_PyramidP, FE\_PyramidDGP  & 3 & 1 \\
+    wedge & FE\_WedgeP, FE\_WedgeDGP  & 3 & 1--2 \\
     \bottomrule
   \end{tabular}
-
-  \caption{\it Comparison of relevant SIMD-related classes in \dealii{} and \texttt{C++23}.}\label{tab:simd}
-  \centering
-  \begin{tabular}{cc}
-    \toprule
-    \textbf{VectorizedArray (\dealii{})} & \textbf{std::simd (\texttt{C++23})}                \\
-    \midrule
-    VectorizedArray<Number>              & std::experimental::native\_simd<Number>            \\
-    VectorizedArray<Number, size>        & std::experimental::fixed\_size\_simd<Number, size> \\ \bottomrule
-  \end{tabular}
 \end{table}
 
-The new interface of \texttt{VectorizedArray} also enables replacement
-by any type with a matching interface. Specifically, this prepares
-\dealii{} for the \texttt{std::simd} class that is slated to become
-part of the \texttt{C++23} standard.
-Table~\ref{tab:simd} compares the \dealii{}-specific SIMD classes and
-the equivalent \texttt{C++23} classes. These changes also prepare for specialized
-code paths exploiting
-vectorization within an element, see~\cite{KronbichlerKormann2019}.
-
-%We welcome the standardization of the SIMD
-%parallelization paradigm in C++ and intend to replace step by step our own
-%wrapper class, which has been continuously developed over the last decade. We
-%would like to emphasize that the work invested in this class was not in vain,
-%since many performance-relevant utility functions implemented with \texttt{VectorizedArray} in mind (e.g., \texttt{vectorized\_load\_and\_transpose}
-%and \texttt{vectorized\_transpose\_and\_store}) will be still used, since they
-%have not become part of the standard.
-
-%Further additions to the \texttt{MatrixFree} infrastructure consist of:
-%\begin{itemize}
-%\item a new variant of \texttt{MatrixFree::cell\_loop()}: It takes two
-%\texttt{std::function} objects with ranges on the locally owned degrees of freedom, one
-%with work to be scheduled before the cell operation first touches some
-%unknowns and another with work to be executed after the cell operation last
-%touches them. The goal of
-%these functions is to bring vector operations close to the time when the
-%vector entries are accessed by the cell operation, which increases the cache
-%hit rate of modern processors by improved temporal locality.
-%\item a new form of loop \texttt{MatrixFree::loop\_cell\_centric()}: This
-%kind of loop can be used in the context of discontinuous Galerkin methods,
-%where both cell and face integrals have to be evaluated. While in the case of
-%the traditional \texttt{loop}, cell and face integrals have been performed
-%independently, the new loop performs all cell and face integrals of a cell in
-%one go. This includes that each face integral has to be evaluated twice, but
-%entries have to be written into the solution vector only once with improved
-%data locality. Previous publications based on \dealii{} have shown the
-%relevance of the latter aspect for reaching higher performance.
-%\end{itemize}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Advances in GPU support}
-\label{subsec:gpu}
-
-For this release, the most noteworthy improvements in GPU support are the
-simplification of the kernel written by user, improvement of error messaging,
-and overlapping of computation and communication when using CUDA-aware MPI with
-the matrix-free framework.
-
-In order to simplify user code, we now recompute
-local degree of freedom and quadrature point indices instead of having the user
-keeping track of them. A new option to overlap computation and communication is
-now available when using CUDA-aware MPI and matrix-free. The underlying idea is
-that only the degrees of freedom (DoFs) on the boundary of the local domain
-require communication. Before evaluating the matrix-free operator at these
-degrees of freedom, we need to communicate ghost DoFs from other
-processors.  Similarly once the operator has been evaluated, we need to update
-the resulting global vector with values from other processors. The strategy that
-we are now using consists in splitting the DoFs into three groups: one group of DoFs
-that are on the local boundary and two groups each owning half of the interior
-DoFs. When evaluating the matrix-free operator, we start the MPI communication
-to get the ghosted DoFs and evaluate the operator on one of the two interior
-DoFs group. When the evaluation is done, we wait for the MPI
-communication to be over, and
-then evaluate the operator on the boundary DoFs group. When this evaluation is
-completed, we start the communication to update the global vector, we evaluate the
-operator on the last group of DoFs, and finally wait for the MPI
-communication to finish.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Expanded use of C++11 facilities}
-\label{subsec:cxx}
-
-Certain types of quantities in a simulation are constants fully known
-at compile time. They can be pre-calculated and stored in the
-compiled executable in order to avoid unnecessary initialization during
-runtime. C++11 and later standards enable such computations by
-marking variables and functions with the \texttt{constexpr} keyword.
-
-This optimization is now enabled for the class templates
-\texttt{Tensor} and \texttt{SymmetricTensor}.
-For instance, the linear mechanical constitutive model for isotropic elastic solids
-uses a constant fourth-order elasticity tensor
-$\mathbb{C} = \lambda \boldsymbol{I} \otimes \boldsymbol{I} + 2 \mu \mathbb{I}$
-which does not depend on the current state of strain.
-This tensor can now be statically initialized by defining it as
-\texttt{constexpr SymmetricTensor<4, dim>}.
-As another example, the lattice vectors in a crystal plasticity model are generally
-constant and known during compilation time, enabling their efficient definition as
-\texttt{constexpr Tensor<1, dim>}.
-
-Declaring variables, functions, and methods as \texttt{constexpr}
-is a C++11 feature that was later expanded by the C++14 standard.
-Thus, parts of the \texttt{constexpr} support in \dealii{} depend on the C++
-standard supported by the compiler used to install the library.
-
-The next release of \dealii{} will require compiler support for the C++14 standard.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{New and improved tutorial and code gallery programs}
+The current release of \dealii adds support for simplex meshes (consisting of triangles in 2D; tetrahedra in 3D) and mixed meshes (consisting of triangles and/or 
+quadrilaterals in 2D; tetrahedra, pyramids, wedges, and/or hexahedra in 3D). 
+Many freely available mesh-generation tools produce such kind of meshes; while users of \dealii used to have to pre-process such meshes and convert them to pure hex meshes, they
+can now directly work with them.
+
+\subsubsection{Refactoring of internal data structures}
+
+To enable simplex and mixed mesh support, we performed a major refactoring of the internal data structures of \dealii. In particular, the \texttt{Triangulation} class and 
+the \texttt{DoFHandler} class have undergone large changes.
+
+The template parameters of the internal data structures of \texttt{Triangulation}
+have been removed and the type of each cell and of each face (only in 3D) is stored. The function
+\texttt{Triangulation::create\_\allowbreak triangulation()}, which converts a given list of
+cells and vertices to the internal data structures, has been rewritten 
+inspired by \citep{logg09} and a speed-up of up to 5 has been reached. Minor adjustments 
+have been made to \texttt{parallel::shared::Triangulation} and \texttt{parallel::\allowbreak fullydistributed::\allowbreak Triangulation} such that 
+the new mesh types can be processed also in parallel.
+
+The internal data structures of \texttt{DoFHandler} used to be hard-coded for pure
+hypercube meshes, while the \texttt{hp::DoFHandler} used to be built around
+CRS-like data structures. Due to the need of CRS data structures in
+\texttt{DoFHandler} in the context of more general meshes, we have merged
+\texttt{hp::DoFHandler} into the \texttt{DoFHandler}. The class \texttt{hp::DoFHandler}, which is now a dummy derivation of \texttt{DoFHandler}, currently only exists for compatibility reasons. For
+more details see Subsection~\ref{TODO}.
+
+\subsubsection{Generating meshes }
+
+The most obvious way to generate a simplex or a mixed mesh is to read the mesh from a file
+generated by an external mesh generator. Currently, we support the 
+file formats \texttt{VTK}, \texttt{MSH}, \texttt{EXODUS II}.
+
+Alternatively, one can create a pure hypercube mesh with the known functions
+in the \texttt{GridGenerator} namespace and convert the obtained mesh to a
+pure simplex mesh with the function 
+\texttt{convert\_\allowbreak hypercube\_\allowbreak to\_\allowbreak simplex\_\allowbreak mesh()} from the \texttt{GridGenerator} namespace.
+
+\subsubsection{Simplex mesh}
+
+After having created a triangulation, one can proceed as usual in the case of
+a pure simplex mesh by selecting an appropriate finite element, mapping, and
+quadrature class:
+
+\begin{c++}
+Simplex::FE_P<dim, spacedim> fe(degree);
+MappingFE<dim, spacedim> mapping(Simplex::FE_P<dim, spacedim>(1));
+Simplex::QGauss<dim> quad(degree + 1);
+
+DoFHandler<dim> dof_handler(tria);
+dof_handler.distribute_dofs(fe);
+
+FEValues<dim, spacedim> fe_values(mapping, fe, quad, flags);
+\end{c++}
+The  list of currently supported finite-element classes is provided in Table~\ref{tab:simplex:fe}. Currently,
+only linear iso-parametric mapping (\texttt{MappingFE}) is available, which takes one of the 
+finite elements mentioned before. For quadrature, the classes \texttt{QGauss}, 
+\texttt{QGaussPyamid}, and \texttt{QGaussWedge} from the \texttt{simplex} namespace are
+available.
+
+\subsubsection{Mixed mesh}
+For mixed meshes, concepts known from the hp context have been applied: 
+ different finite-element classes are assigned to different cell types. In the 
+case of a 2D mixed mesh, which can only consists of triangles and
+quadrilaterals, the finite element defined on a triangle (e.g., \texttt{Simplex::FE\_P})
+and on a quadrilateral (e.g., \texttt{FE\_Q}) can be collected in a \texttt{hp::FECollection}:
+
+\begin{c++}  
+hp::FECollection<dim, spacedim> fe
+  {Simplex::FE_P<dim, spacedim>(degree), FE_Q<dim, spacedim>(degree)};
+\end{c++}
+
+Similarly, \texttt{hp::QCollection} and \texttt{hp::MappingCollection} can be used
+to construct appropriate collections. Furthermore, the right active finite-element index, which points to the right 
+finite element of that cell, has to be assigned to each cell:
+
+\begin{c++}  
+DoFHandler<dim> dof_handler(tria);
+
+for (const auto &cell : dof_handler.active_cell_iterators())
+  switch(cell->reference_cell_type())
+    {
+      case ReferenceCell::Type::Tri:  cell->set_active_fe_index(0); break;
+      case ReferenceCell::Type::Quad: cell->set_active_fe_index(1); break;
+      // 3D (Tet, Pyramid, Wedge, Hex) not shown
+      default: Assert(false, ExcNotImplemented());
+    }
+
+dof_handler.distribute_dofs(fe);
+\end{c++}
+
+\subsubsection{Practical implications}
+
+The introduction of simplex and mixed meshes leads to some implications
+for the user if these features should be used. For instance, each cell might have a different type with 
+different number of vertices, lines, and faces so that these quantities cannot be 
+compile-time constants anymore. This information used to be queried from
+the \texttt{GeometryInfo} class. To avoid using this class, we have extended
+relevant classes, e.g., \texttt{TriaAccessor} or \texttt{TriaCellAccessor},
+with useful new functions like \texttt{n\_vertices()}, \texttt{n\_lines()}, or
+\texttt{n\_faces()} so that users can simply write:
+\begin{c++}
+for(const auto & cell : tria.active_cell_iterators())
+  for(unsigned int f = 0; f < cell->n_faces(); ++f)
+    (void) cell->face(f);
+\end{c++}
+Alternatively, one can use an iterator-based approach to loop over all faces
+of a cell, introduced in the last release. The relevant functions have been adjusted to be able to deal with
+various number of faces. 
+\begin{c++}
+for(const auto & cell : tria.active_cell_iterators())
+  for(const auto & face : cell->face_iterators())
+    (void) face;
+\end{c++}
+Furthermore, the number of degrees of freedom might differ between cells so that
+local data storage units needed for assembly might need to be resized for each 
+cell as it is already good practice in the hp context:
+\begin{c++}
+std::vector<double> local_rhs;
+for(const auto & cell : tria.active_cell_iterators())
+ {
+   hp_fe_values.reinit(cell);
+   local_rhs.resize(hp_fe_values.get_present_fe_values().n_dofs_per_cell());
+ }
+\end{c++}
+
+What is true for cells is also true for faces in 3D: faces can be either triangles
+or quadrilaterals.
+This is the case even if no mixed mesh is used.
+As a consequence, some functions, e.g.
+\texttt{FiniteElementData::n\_dofs\_per\_face()}, have been extended with a new optional argument for the face number.
+
+We tried---where possible---to provide utility functions so that users do not need to access a \texttt{GeometryInfo}-like data structure.
+However, these functions internally rely on the \texttt{ReferenceCell::\allowbreak internal::\allowbreak Info} classes, which are similarly structured as \texttt{GeometryInfo} and even use it for hypercube cells.
+Users can query the reference-cell type of geometric entity via:
+\begin{c++}
+const auto reference_cell_type = cell->reference_cell_type();
+\end{c++}
+and can access the relevant struct describing the current reference cell or its i-th face via:
+\begin{c++}
+ReferenceCell::internal::Info::get_cell(reference_cell_type);
+ReferenceCell::internal::Info::get_face(reference_cell_type, i);
+\end{c++}
+
+Furthermore, it is crucial to use the right FE--mapping--quadrature family 
+for the right cell type. In particular, this means that users cannot 
+rely on default parameters regarding mapping and/or quadrature in the 
+context of simplex meshes since these are defined for hypercube cells.
+We recommend to choose the needed mapping,
+FE, and quadrature rules at a single central place and pass these throughout the program
+in the form of ``scratch data''.
+
+\subsubsection{Matrix-free support}
+
+We provide matrix-free support for simplex and mixed meshes as well as for
+continuous and discontinuous elements. From the user perspective, the setup has
+hardly changed:
+%\begin{c++}
+%Simplex::QGauss<dim> quad(degree + 1)
+%
+%MatrixFree matrix_free;
+%matrix_free.reinit(mapping, dof_handler, constraints, quad);
+%\end{c++}
+Instead of providing a 1D quadrature rule, which is extended to higher dimensional
+spaces via a tensor product, one provides d-dimensional quadrature rules.
+The templated versions of \texttt{FEEvaluation} and \texttt{FEFaceEvaluation} cannot be used: the number of degrees of freedom and the number of quadrature points are determined at runtime.
+More details can be found in Subsection~\ref{subsec:mf}.
+
+By the time of writing, we do not use internally any advance techniques for evaluating values and 
+gradients at the quadrature points as we do for tensor-product elements and 
+instead rely on full interpolation matrices. This is for now acceptable since only low-order elements are 
+supported.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Advances in the multigrid infrastructure}
+\label{subsec:mg}
+
+Until now, \dealii has only supported local smoothing multigrid algorithms~\citep{ClevengerHeisterKanschatKronbichler2019}.
+With release 9.3, the support for global coarsening of continuous (\texttt{FE\_Q}) and discontinuous (\texttt{FE\_DGQ}) elements  in a geometric and a polynomial sense has been added.
+These multigrid variants promise less iteration numbers and better parallel scalability with the disadvantage of having to deal with hanging nodes during smoothing and of more work per iteration overall.
+
+Global-coarsening algorithms~\citep{becker00} either coarsen each cell of a triangulation or reduce the polynomial degree of the finite element of each cell and hereby obtain a new system with less unknowns (also known as "coarse level").
+Repeating such kind of coarsening results in a sequence of multigrid levels.
+
+The transfer between two levels has been encoded in the new class \texttt{MGTwoLevelTransfer}, which can be set up via the functions \texttt{MGTwoLevelTransfer::\allowbreak reinit\_\allowbreak geometric\_\allowbreak transfer()} or \texttt{MGTwo\allowbreak LevelTransfer::\allowbreak reinit\_\allowbreak polynomial\_\allowbreak transfer()} for given
+\texttt{DoFHandler} and \texttt{AffineConstraint} classes of two levels. The resulting transfer operators
+between two levels can be collected in a single object 
+(\texttt{MGTransferGlobalCoarsening}) that can be used just as \texttt{MGTransferMatrixFree} and can be passed to the actual \texttt{Multigrid}
+algorithm. The option that users can set up the transfer operators between two levels on their own enables
+the mix of polynomial  and geometric coarsening, as needed. It should also be noted that a transfer between continuous
+and discontinuous elements is possible by providing \texttt{DoFHandler}s set up with this kind of elements.
+
+Internally, the transfer operators categorize fine/coarse cell pairs according to the 
+occurrence of refinement 
+(refinement or no-refinement) and to the polynomial-degree pair ($k_f$-$k_c$) into groups. 
+The projection and restriction are performed with efficient matrix-free kernels applying
+1D projection and restriction matrices.
+
+The user is responsible for setting up the levels (including the operators and the
+smoothers). Please note that global coarsening can currently only be performed between active levels. 
+The user can apply utility
+functions from the \texttt{MGTransferGlobalCoarseningTools} namespace to set up the levels. For 
+constructing the diagonal or the system matrix for a matrix-free operator,
+needed for the smoother or the coarse-grid solver,
+the functions \texttt{create\_diagonal()} and \texttt{create\_matrix()} from
+the \texttt{MatrixFreeTools} namespace can be used (see also Subsection~\ref{subsec:mf}).
+
+The usage of the new transfer operators (and of some of the utility functions) in the context of a hybrid multigrid algorithm (hp-MG with AMG as coarse-grid solver) for hp problems is demonstrated in the new tutorial \texttt{step-75} (see also Subsection~\ref{subsec:steps}).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Advances in the matrix-free infrastructure}
+\label{subsec:mf}
+
+\subsubsection{Precompilation of evaluation kernels}
+
+The classes \texttt{FEEvaluation} and \texttt{FEFaceEvaluation} are highly
+templated to reach high performance. In particular, the template parameters include
+the polynomial degree of the finite element $k$ and the number of the 1D quadrature points $q$.
+For applications codes that rely on operators of many different degrees (e.g., because
+they use p-multigrid or hp algorithms), the instantiation
+means a long compilation time.
+
+In the current release, we have improved the \texttt{FEEvaluation} and the \texttt{FEFaceEvaluation} class implementations, which do
+not rely on the template parameters $k$ and $q$ (expressed in the code
+with "-1" and "0"):
+\begin{c++}
+FEEvaluation<dim, -1, 0, n_components, Number, VectorizedArrayType>
+  phi(range, dofhandler_index, quadrature_index, first_selected_component);
+\end{c++}
+The meaning of the new first argument is 
+discussed in Subsubsection~\ref{subsubsection:mf:hp}.
+
+These classes select at runtime---for common low and medium polynomial degree/quadrature combinations ($k\le 6$ and $q\in\{ k+1, k+2, \lfloor (3k)/2 \rfloor \}$)---efficient precompiled implementations and default to non-templated
+evaluation kernels else.
+
+In the case that even higher polynomial degrees are needed, one can precompile the
+relevant internal classes
+(\texttt{FEEvaluationFactory}, \texttt{FEFaceEvaluationFactory}, \texttt{CellwiseInverseMassFactory}) in the user code for needed \texttt{degree}s
+and \texttt{VectorizedArrayType}s in the following way:
+\begin{c++}
+#define FE_EVAL_FACTORY_DEGREE_MAX 12
+
+#include <deal.II/matrix_free/evaluation_template_factory.templates.h>
+
+DEAL_II_NAMESPACE_OPEN
+template struct dealii::internal::FEEvaluationFactory
+  <dim, VectorizedArrayType::value_type, VectorizedArrayType>;  
+DEAL_II_NAMESPACE_CLOSE
+\end{c++}
+
+\subsubsection{Parallel matrix-free hp implementation}\label{subsubsection:mf:hp}
+
+With release 9.1, large parts of the hp algorithms in \dealii were parallelized so that
+parallel matrix-based simulations can be performed with the hp infrastructure. In the present
+release, we have extended the setup routines of \texttt{MatrixFree} so that it now also
+provides parallel hp support.
+
+Until now, the \texttt{FEEvaluation} classes used the template parameters $k$ and $q$ to select the correct active hp FE and quadrature index and users had 
+to use the function \texttt{MatrixFree::create\_cell\_\allowbreak subrange\_\allowbreak hp()} or
+\texttt{::create\_cell\_\allowbreak subrange\_\allowbreak hp\_\allowbreak by\_index()} to create subranges of cells with 
+the same polynomial degree. This led to user codes that were hard to read due to complicated jump tables.
+
+The creation of subranges is now performed internally, and the non-templated versions
+of the \texttt{FEEvaluation} classes are extended for the hp case. To nevertheless determine
+the active hp FE and quadrature index, the current cell/face range has to be provided
+the constructors of the \texttt{FEEvaluation} classes, from which the relevant information
+can be deduced (in the simplex case also the face type). These changes enable
+users to write matrix-free code independently of whether hp is used or not.
+
+The new tutorial \texttt{step-75} presents how to use the new hp-related features in \texttt{MatrixFree} 
+in the context of a hybrid-multigrid solver.
+
+\subsubsection{MPI-3.0 shared-memory support} 
+The classes \texttt{MatrixFree} and \texttt{LinearAlgebra::\allowbreak distributed::\allowbreak Vector} have been extended with 
+MPI-3.0 shared-memory features. If \texttt{MatrixFree} has been configured by setting
+\texttt{MatrixFree::AdditionalData::communicator\_sm} by an appropriate user-provided subcommunicator,
+vectors are created by \texttt{MatrixFree::create\_dof\_vector()} in such a way that the
+memory of processes in this subcommunicator can be accessed. In particular, this means 
+that the \texttt{FEEvaluation} classes can access the degrees of freedom owned by 
+other processes and in certain cases local 
+communication can be skipped. To prevent race conditions, \texttt{MatrixFree} uses local
+barriers at the beginning and the end of loops (\texttt{loop()}, \texttt{cell\_loop()}, \texttt{loop\_cell\_centric()}).
+
+The novel tutorial \texttt{step-67b} shows the usage of the new MPI-3.0 feature
+ in the context of the solution
+of the Euler equations. Here, a speed-up of 27\% could be reached compared to the
+regular version \texttt{step-67}, by using the new feature.
+For more details and a successful application in the library \texttt{hyper.deal}, see \citep{munch2020hyperdeal}.
+
+\subsubsection{Miscellaneous}
+
+The namespace \texttt{MatrixFreeTools} has been introduced. It collects various utility
+functions useful in the context of application of \texttt{MatrixFree}. In particular,
+it contains different functions to extract quantities related to a matrix, e.g., the 
+computation of the diagonal or a matrix with a matrix-free approach.
+
+With version 9.2, we have introduced cell-centric loops. They have been extended to unstructured
+meshes. The tutorial \texttt{step-67b} shows their usage in the context of the solution
+of the Euler equations.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Evaluation and integration at arbitrary points}
+\label{subsec:fepointvalues}
+
+Example: testing of surface tension in the context of sharp-interface methods:
+\begin{align*}
+\left(\vec{v}, \kappa \, \vec{n}\right)_\Gamma
+\approx
+\sum\left(\vec{v}, \kappa(\vec{p}_q) \, \vec{n}(\vec{p}_q) (JxW)_q\right)
+\end{align*}
+in code:
+\begin{c++}
+phi_normal.evaluate(cell, points, normal_values, EvaluationFlags::values);
+for (unsigned int q = 0; q < n_points; ++q)
+   phi_force.submit_value(phi_normal.get_value(q) * 
+                          phi_curvature.get_value(q) * JxW[q], q);
+phi_force.integrate(cell, points, force_values, EvaluationFlags::values);
+\end{c++}
+The quadrature points and the connected \texttt{JxW} can, e.g., come from a 
+codim-1 mesh. Determining to which \texttt{cell} a quadrature point belongs
+to on the background mesh, incl. the reference-cell coordinates \texttt{points},
+can be determined with know functions like \texttt{GridTools::find\_all\_active\_cells\_around\_point()}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{New and improved tutorials and code gallery programs}
 \label{subsec:steps}
@@
 \begin{itemize}
-  \item \texttt{step-47} is a new program that solves the biharmonic
-        equation $\Delta^2 u = f$ with
-        ``clamped'' boundary condition given by $u=g, \partial u/\partial
-          n=h$. This program is based on the $C^0$ interior
-        penalty ($C^0$IP) method for fourth order problems
-        \cite{Brenner2005}. In order to overcome
-        shortcomings of classical approaches, this method uses $C^0$ Lagrange finite
-        elements and introduces ``jump'' and ``average'' operators on
-        interfaces of elements that penalize the jump of the gradient of the
-        solution in order to obtain convergence to the $H^2$-regular
-        solution of the equation.
-
-        The $C^0$IP approach is a modern alternative to classical methods that
-        use $C^1$-conforming elements such as the Argyris
-        element, the Clough-Tocher element and others, all developed in the
-        late 1960s. From a twenty-first century perspective, they can only be
-        described as bizarre in their construction. They are also exceedingly
-        cumbersome to implement if one wants to use general meshes. As a
-        consequence, they have largely fallen out of favor and \dealii{} currently
-        does not contain implementations of these shape functions.
-
-  \item \texttt{step-50} is a program that demonstrates the parallel geometric multigrid features
-  in \dealii{} as described in \cite{ClevengerHeisterKanschatKronbichler2019}.
-  The problem considered is a variable viscosity Laplace equation and it is solved
-  with three different approaches:
-  (i) using a matrix-based geometric multigrid based on Trilinos or
-  PETSc; (ii) using a matrix-free
-  geometric multigrid; (iii) using algebraic multigrid (Trilinos ML). The tutorial
-  demonstrates the superiority of the matrix-free method for the problem under consideration,
-  and shows that, for matrix-based formulations, the performance of algebraic and
-  geometric multigrid methods are roughly comparable.
-
-  \item \texttt{step-58} is a program that solves the nonlinear
-        Schr{\"o}dinger equation, which in non-dimensional form reads
-        \begin{align*}
-          - i \frac{\partial \psi}{\partial t}
-          - \frac 12 \Delta \psi
-          + V \psi
-          + \kappa |\psi|^2 \psi
-           & = 0,
-        \end{align*}
-        augmented by appropriate initial and boundary conditions and using
-        an appropriate form for the potential $V=V(\mathbf x)$. The
-        tutorial program focuses on two specific aspects for which this
-        equation serves as an excellent test case: (i) Solving
-        complex-valued problems without splitting the equation into its
-        real and imaginary parts (as \texttt{step-29} does, for
-        example). (ii)~Using operator splitting techniques. The equation is
-        a particularly good test case for this technique because the only
-        nonlinear term, $\kappa |\psi|^2 \psi$, does not contain any
-        derivatives and consequently forms an ODE at each node point, to be
-        solved in each time
-        step in an operator splitting scheme (for which, furthermore, there
-        exists an analytic solution), whereas the remainder of the
-        equation is linear and easily solved using standard finite element
-        techniques.
-
-  \item \texttt{step-65} presents \texttt{TransfiniteInterpolationManifold}, a
-        manifold class that can propagate curved boundary information into the
-        interior of a computational domain by transfinite interpolation \cite{Gordon82}.
-        This manifold is a prototype for many other manifolds in that it is relatively
-        expensive to compute the new points, especially for higher order mappings. Since
-        typical programs query higher-order geometries in a large variety of contexts,
-        the contribution of the mapping to the run time can be significant. As a
-        solution, the tutorial also presents the class \texttt{MappingQCache}, which
-        samples the information of expensive manifolds in the points of a
-        \texttt{MappingQ} and caches it. The tutorial shows that this makes all queries
-        to the geometry very cheap.
-
-  \item \texttt{step-67} is an explicit time integrator for the
+\item \texttt{step-19} TODO
+\item \texttt{step-67b} is an explicit time integrator for the
         compressible Euler equations discretized with a high-order discontinuous
-        Galerkin (DG) scheme using the matrix-free infrastructure. Besides the use of
-        matrix-free evaluators for systems of equations and over-integration, it also
-        presents \texttt{MatrixFreeOperators::CellwiseInverseMassMatrix}, a fast implementation
-        of the action of the inverse mass matrix in the DG setting using tensor
-        products. Furthermore, this tutorial demonstrates the usage of new
-        pre and post operations, which can be passed to
-        \texttt{MatrixFree::cell\_loop()}, to schedule operations on sections of vectors close
-        to the matrix-vector product to increase data locality.
-
-  \item \texttt{step-69} presents a first-order scheme solving the compressible
-        Euler equations of gas dynamics with a graph-viscosity stabilization
-        technique. Beside the usual conservation properties of mass, momentum, and
-        total energy, the method also guarantees that the constructed solution
-        obeys pointwise stability constraints (in particular positivity of
-        density and internal
-        energy, and a local minimum principle on the specific entropy). As such
-        \text{step-69} is strictly speaking more a collocation-type discretization
-        than a variational formulation, even though it is implemented with finite
-        elements.
-
-        The time-update at each node requires the evaluation of a right-hand side
-        that depends (nonlinearly) on information from the previous time-step that
-        spans more than one cell. Therefore, assembly loops operate directly on the
-        sparsity graph in order to retrieve information from the entire stencil
-        associated with each node. From a programming perspective, \texttt{step-69}
-        features a number of techniques that are of interest for a wider audience:
-        It discusses a hybrid thread and MPI parallelized scheme with efficient
-        MPI node-local numbering of degrees of freedom. It showcases how to
-        perform asynchronous write-out of results using a background thread with
-        \texttt{std::async}, and discusses a simple but effective checkpointing and
-        restart technique.
-
-  \item \texttt{step-70} solves a fluid structure interaction problem on
-        non-matching parallel distributed triangulations, showing the usage of
-        the \texttt{particles::ParticleHandler} class for two different tasks:
-        (i) to track the position of quadrature points of a non-matching
-        solid grid immersed in a fluid grid, and (ii) to track the position of a
-        collection of massless tracers.
-
-        The program considers a mixing problem in the laminar flow
-        regime. Such problems occur in a wide range of applications ranging
-        from chemical engineering to power generation (e.g. turbomachinery).
-        Mixing problems are particularly hard to solve numerically, because
-        they often involve a container (with fixed boundaries, and possibly
-        complex geometries such as baffles), and one (or more) immersed and
-        rotating impellers, making it difficult to use Arbitrary Lagrangian
-        Eulerian formulations, where the fluid domain – along with the mesh! –
-        is smoothly deformed to follow the deformations of the immersed solid.
-        We show how to solve such problems using a penalization technique
-        where the flow problem is solved in the union of the fluid and solid
-        domain, and the movement of the solid is imposed weakly using a
-        Nitsche-like penalization.
-
-        The \texttt{particles::ParticleHandler} class is used in this context
-        to allow integration of the fluid basis functions on the solid domain
-        (which is immersed and non-matching with regard to the fluid grid). This is
-        achieved by attaching the information required to perform the
-        integration on the solid grid, to the property field of the particles
-        associated with the quadrature points of the solid, and advecting those
-        particles according to the solid velocity. In this program, we
-        externally prescribe rather than solve for the
-        the solid velocity; however,
-        the program can easily be extended to allow for the solution of
-        coupled elasticity equations on the solid domain.
+        Galerkin (DG) scheme, using the matrix-free infrastructure just as \texttt{step67} does.
+        The tutorial presents advanced topics, like the usage of cell-centric loops and
+        the new MPI-3.0 shared-memory capabilities of \texttt{MatrixFree} tor each high
+        throughput. Furthermore, the utilization of the template parameter
+        \texttt{VectorizedArrayType} and the application of lambdas to capture cell and face
+        integrals are discussed.
+\item \texttt{step-68} TODO
+\item \texttt{step-75} TODO
 \end{itemize}
@@
 \begin{itemize}
-  \item A program based on the XBraids library to solve time-dependent
-    problems in a parallel-in-time fashion. This program was
-    contributed by Joshua Christopher.
-  \item A program that solves the biphasic nonlinear
-    poro-viscoelasticity equations based on Ogden hyperelasticity, to
-    explore the porous and viscous contributions in brain mechanics
-    applications.  This program was contributed by Ester Comellas and
-    Jean-Paul Pelteret. The program also serves as a demonstration of
-    the automatic differentiation capabilities of \dealii{}.
+  \item 
 \end{itemize}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Python interfaces}
-\label{subsec:python}
-
-Initial support for Python has existed in \dealii{} since version
-9.0. The present release significantly extends the Python
-interface. Specifically, a large number of methods from classes such
-as \texttt{Triangulation, CellAccessor, TriaAccessor, Mapping,
-  Manifold, GridTools} can now be invoked from Python. We have focused
-on methods and functions that are widely used when a mesh is created
-and parameters related to the boundary, manifold, and material
-identifiers are assigned. The following listing gives an idea of how
-such code looks:
-\begin{python}
-import PyDealII.Release as dealii
-
-triangulation = dealii.Triangulation('2D')
-triangulation.generate_hyper_shell(center = dealii.Point([0, 0]),
-                                   inner_radius = 0.5, outer_radius = 1.,
-                                   n_cells = 0, colorize = True)
-
-triangulation.refine_global(2)
-
-for cell in triangulation.active_cells():
-    cell.material_id = 1 if cell.center().x > 0 else 2
-
-    for face in cell.faces():
-        if face.at_boundary() and face.boundary_id == 1:
-            cell.refine_flag = 'isotropic'
-
-triangulation.execute_coarsening_and_refinement()
-\end{python}
-The mesh that results from this code is shown in Fig.~\ref{fig:pymesh}.
-
-\begin{floatingfigure}[r]{0.3\textwidth}
-  \centering
-  \vspace*{-1.4cm}
-  \includegraphics[width=0.3\textwidth]{python_mesh.png}
-  \vspace*{-5mm}
-  \caption{\it The mesh generated by the Python code shown in the main
-    text. Cells are colored by material id.}
-  \vspace*{0.2cm}
-  \label{fig:pymesh}
-\end{floatingfigure}
-All triangulations created from within Python are serial. However,
-once the mesh is designed, the triangulation can be serialized along
-with the auxiliary information about possible refinement, boundaries,
-materials and manifolds. This object can be easily deserialized within
-a C++ program for subsequent production runs. Furthermore, such a
-serialized triangulation can also be used in the construction of
-\texttt{parallel::shared}, \texttt{parallel::distributed}, and
-\texttt{parallel::fullydistributed} triangulations (see also Section~\ref{subsec:pft}).
-
-To facilitate the illustration of the new Python bindings, tutorial programs \href{https://github.com/dealii/dealii/blob/dealii-9.2/examples/step-49/step-49.ipynb}{step-49} and \href{https://github.com/dealii/dealii/blob/dealii-9.2/examples/step-53/step-53.ipynb}{step-53} were replicated as Jupyter notebooks.
-
-The introspective nature of the Python language makes it easy to infer
-the list of supported methods from the Python objects, for example by
-typing \texttt{dir(PyDealII.Release.Triangulation)}.
-The current Python interface does not yet provide access to \dealii{}'s finite element machinery, i.e., classes such as \texttt{DoFHandler, FE\_*, FEValues}, etc.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Incompatible changes}
 
-The 9.2 release includes
+The 9.3 release includes
 \href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_1_1_and_9_2_0.html}
@@
 \begin{itemize}
-  \item Two functions that provide information about all processes,
-        namely
-        \begin{itemize}
-          \item \texttt{DoFHandler::locally\_owned\_dofs\_per\_processor()}
-          \item \texttt{DoFHandler::locally\_owned\_mg\_dofs\_per\_processor()}
-        \end{itemize}
-        have been deprecated. As discussed in
-        Subsection~\ref{subsec:performance}, \dealii{} by default no longer
-        stores information for all processes on all processes, but only
-        local or locally-relevant information. On the other
-        hand, if necessary, global information can still be computed using,
-        for example, calling
-        \texttt{Utilities::MPI::Allgather(locally\_owned\_info(), comm)}.
-
-  \item A corresponding change has been made in parallel triangulation
-        classes:
-        \texttt{parallel::} \texttt{TriangulationBase::compute\_n\_locally\_owned\_active\_cells\_per\_processor()}
-        can be used to obtain information about how many cells each
-        process owns.
+  \item 
 \end{itemize}

### 2021-02-18 9.3/paper.tex

@@
      Wolfgang Bangerth,
-     Bruno Blais,
-     Thomas C. Clevenger,
+     %Bruno Blais,
+     %Thomas C. Clevenger,
     Marc Fehling,
-    Alexander V. Grayver,
+    %Alexander V. Grayver,
     Timo Heister,
@@
     Jean-Paul Pelteret,
-    Reza Rastak,
-    Ignacio Thomas,
+    %Reza Rastak,
+    %Ignacio Thomas,
     Bruno Turcksin,
-    Zhuoran Wang,
+    %Zhuoran Wang,
     David Wells
@@
 
-\author[3]{Bruno Blais}
-\affil[3]{Research Unit for Industrial Flows Processes (URPEI), Department of Chemical Engineering,
-          Polytechnique Montréal,
-          PO Box 6079, Stn Centre-Ville, Montréal, Québec, Canada, H3C 3A7.
-          {\texttt{bruno.blais@polymtl.ca}}}
-
- \author[4]{Thomas~C.~Clevenger}
- \affil[4]{School of Mathematical and Statistical Sciences,
-   Clemson University,
-   Clemson, SC, 29634, USA
-   {\texttt{tcleven/heister@clemson.edu}}}
+%\author[3]{Bruno Blais}
+%\affil[3]{Research Unit for Industrial Flows Processes (URPEI), Department of Chemical Engineering,
+%          Polytechnique Montréal,
+%          PO Box 6079, Stn Centre-Ville, Montréal, Québec, Canada, H3C 3A7.
+%          {\texttt{bruno.blais@polymtl.ca}}}
+
+% \author[4]{Thomas~C.~Clevenger}
+% \affil[4]{School of Mathematical and Statistical Sciences,
+%   Clemson University,
+%   Clemson, SC, 29634, USA
+%   {\texttt{tcleven/heister@clemson.edu}}}
 %
@@
 
-\author[6]{Alexander~V.~Grayver}
-\affil[6]{Institute of Geophysics,
-  ETH Zurich,
-  Sonneggstrasse 5, 8092 Z\"{u}rich, Switzerland.
-  {\texttt{agrayver@ethz.ch}}}
+%\author[6]{Alexander~V.~Grayver}
+%\affil[6]{Institute of Geophysics,
+%  ETH Zurich,
+%  Sonneggstrasse 5, 8092 Z\"{u}rich, Switzerland.
+%  {\texttt{agrayver@ethz.ch}}}
 
@@
 
- \author[12]{Reza Rastak}
- \affil[12]{Department of Civil and Environmental Engineering,
-   Stanford University,
-   Stanford, CA 94305, USA.
-   {\texttt{rastak@stanford.edu}}}
-
- \author[**13]{Ignacio Tomas}
- \affil[13]{Sandia National Laboratories, Org 1442,
- P.O. Box 5800, MS 1320, Albuquerque, NM, 87185-1320,
-   \texttt{itomas@sandia.gov}}
+% \author[12]{Reza Rastak}
+% \affil[12]{Department of Civil and Environmental Engineering,
+%   Stanford University,
+%   Stanford, CA 94305, USA.
+%   {\texttt{rastak@stanford.edu}}}
+
+% \author[**13]{Ignacio Tomas}
+% \affil[13]{Sandia National Laboratories, Org 1442,
+% P.O. Box 5800, MS 1320, Albuquerque, NM, 87185-1320,
+%   \texttt{itomas@sandia.gov}}
 
@@
 
-\author[14]{Zhuoran Wang}
-\affil[14]{Department of Mathematics, Colorado State University, Fort Collins,
-          CO 80523-1874, USA.
-{\texttt{zhrwang@math.colostate.edu}}}
+%\author[14]{Zhuoran Wang}
+%\affil[14]{Department of Mathematics, Colorado State University, Fort Collins,
+%          CO 80523-1874, USA.
+%{\texttt{zhrwang@math.colostate.edu}}}
 
@@
   This paper provides an overview of the new features of the finite element
-  library \dealii, version 9.2.
+  library \dealii, version 9.3.
 \end{abstract}
@@
 
-\dealii{} version 9.2.0 was released May 20, 2020.
+\dealii{} version 9.3.0 was released May ??, 2021.
 This paper provides an
 overview of the new features of this release and serves as a citable
-reference for the \dealii{} software library version 9.2. \dealii{} is an
+reference for the \dealii{} software library version 9.3. \dealii{} is an
 object-oriented finite element library used around the world in the
@@
 \begin{itemize}
-  \item A new, parallel, fully distributed triangulation class (see
-        Section~\ref{subsec:pft});
-  \item Substantially improved performance for very large computations
-    on tens or hundreds of thousands of processor cores and up to
-    trillions of unknowns (see
-        Section~\ref{subsec:performance});
-  \item Better support for parallel $hp$-adaptive algorithms (see
-        Section~\ref{subsec:hp});
-  \item Support for particle-based methods (see
-        Section~\ref{subsec:particles});
-  \item Improved performance of the symbolic differentiation framework (see
-        Section~\ref{subsec:symbdiff});
-  \item Advances in SIMD capabilities and the matrix-free infrastructure (see
-        Section~\ref{subsec:mf});
-  \item Advances in GPU support (see
-        Section~\ref{subsec:gpu});
-  \item Better use of modern C++ language features (see
-        Section~\ref{subsec:cxx});
-  \item Seven new tutorial programs (see
-        Section~\ref{subsec:steps}).
+  \item TODO.
 \end{itemize}
@@
 \begin{itemize}
-  \item globally unique cell IDs
+  \item globally unique cell IDs: TODO
 \end{itemize}
@@
 
-\newpage
+%\newpage
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Advances in the particle infrastructure}
+\label{subsec:particles}
+
+TODO
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Advances in the multigrid infrastructure}
@@
 \item \texttt{step-19} TODO
-\item \texttt{step-67b} is an explicit time integrator for the
+\item \texttt{step-68} TODO
+\item \texttt{step-75} TODO
+\item \texttt{step-76} is an explicit time integrator for the
         compressible Euler equations discretized with a high-order discontinuous
@@
         integrals are discussed.
-\item \texttt{step-68} TODO
-\item \texttt{step-75} TODO
 \end{itemize}
@@
 \begin{itemize}
-  \item 
+  \item preCICE: TODO
 \end{itemize}
@@
 % 9.2: updated 5/11/2020 MM
-Pasquale Africa,
-Ashna Aggarwal,
-Giovanni Alzetta,
-Mathias Anselmann,
-Kirana Bergstrom,
-Manaswinee Bezbaruah,
-Benjamin Brands,
-Yong-Yong Cai,
-Fabian Castelli,
-Joshua Christopher,
-Ester Comellas,
-Katherine Cosburn,
-Denis Davydov,
-Elias Dejene,
-Stefano Dominici,
-Brett Dong,
-Luel Emishaw,
-Niklas Fehn,
-Isuru Fernando,
-Rebecca Fildes,
-Menno Fraters,
-Andres Galindo,
-Daniel Garcia-Sanchez,
-Rene Gassmoeller,
-Melanie Gerault,
-Nicola Giuliani,
-Brandon Gleeson,
-Anne Glerum,
-Krishnakumar Gopalakrishnan,
-Graham Harper,
-Mohammed Hassan,
-Nicole Hayes,
-Bang He,
-Johannes Heinz,
-Jiuhua Hu,
-Lise-Marie Imbert-Gerard,
-Manu Jayadharan,
-Daniel Jodlbauer,
-Marie Kajan,
-Guido Kanschat,
-Alexander Knieps,
-Uwe K{\"o}cher,
-Paras Kumar,
-Konstantin Ladutenko,
-Charu Lata,
-Adam Lee,
-Wenyu Lei,
-Katrin Mang,
-Mae Markowski,
-Franco Milicchio,
-Adriana Morales Miranda,
-Bob Myhill,
-Emily Novak,
-Omotayo Omosebi,
-Alexey Ozeritskiy,
-Rebecca Pereira,
-Geneva Porter,
-Laura Prieto Saavedra,
-Roland Richter,
-Jonathan Robey,
-Irabiel Romero,
-Matthew Russell,
-Tonatiuh Sanchez-Vizuet,
-Natasha S. Sharma,
-Doug Shi-Dong,
-Konrad Simon,
-Stephanie Sparks,
-Sebastian Stark,
-Simon Sticko,
-Jan Philipp Thiele,
-Jihuan Tian,
-Sara Tro,
-Ferdinand Vanmaele,
-Michal Wichrowski,
-Julius Witte,
-Winnifried Wollner,
-Ming Yang,
-Mario Zepeda Aguilar,
-Wenjuan Zhang,
-Victor Zheng.
+TODO
 

### 2021-03-07 9.3/paper.tex
Update simplex section

@@
 \usepackage{floatflt}
+
+\usepackage{multirow}
 
@@
     \midrule
-    simplex (line, triangle, tetrahedron) & FE\_P, FE\_DGP  & 1--3 & 1--2 \\
+    \multirow{2}{*}{simplex (line, triangle, tetrahedron)} & FE\_SimplexP, FE\_SimplexDGP  & 1--3 & 1--2 \\
+                                          & FE\_SimplexP\_Bubbles  & 1--3 & 1--2 \\
     pyramid & FE\_PyramidP, FE\_PyramidDGP  & 3 & 1 \\
@@
 hypercube meshes, while the \texttt{hp::DoFHandler} used to be built around
-CRS-like data structures. Due to the need of CRS data structures in
+CRS-like data structures. Due to the need of CRS data structures in the
 \texttt{DoFHandler} in the context of more general meshes, we have merged
@@
 \begin{c++}
-Simplex::FE_P<dim, spacedim> fe(degree);
-MappingFE<dim, spacedim> mapping(Simplex::FE_P<dim, spacedim>(1));
-Simplex::QGauss<dim> quad(degree + 1);
+FE_SimplexP<dim, spacedim> fe(degree);
+MappingFE<dim, spacedim> mapping(FE_SimplexP<dim, spacedim>(1));
+QGaussSimplex<dim> quad(degree + 1);
 
@@
 The  list of currently supported finite-element classes is provided in Table~\ref{tab:simplex:fe}. Currently,
-only linear iso-parametric mapping (\texttt{MappingFE}) is available, which takes one of the 
-finite elements mentioned before. For quadrature, the classes \texttt{QGauss}, 
-\texttt{QGaussPyamid}, and \texttt{QGaussWedge} from the \texttt{simplex} namespace are
+only linear iso-parametric mapping (\texttt{MappingFE} and \texttt{MappingFEFields}) is available. For quadrature, the classes \texttt{QGaussSimplex}, \texttt{QWitherdenVincentSimplex},  
+\texttt{QGaussPyamid}, and \texttt{QGaussWedge} are
 available.
@@
 case of a 2D mixed mesh, which can only consists of triangles and
-quadrilaterals, the finite element defined on a triangle (e.g., \texttt{Simplex::FE\_P})
+quadrilaterals, the finite element defined on a triangle (e.g., \texttt{FE\_SimplexP})
 and on a quadrilateral (e.g., \texttt{FE\_Q}) can be collected in a \texttt{hp::FECollection}:
@@
 hp::FECollection<dim, spacedim> fe
-  {Simplex::FE_P<dim, spacedim>(degree), FE_Q<dim, spacedim>(degree)};
+  {FE_SimplexP<dim, spacedim>(degree), FE_Q<dim, spacedim>(degree)};
 \end{c++}
@@
 for the user if these features should be used. For instance, each cell might have a different type with 
-different number of vertices, lines, and faces so that these quantities cannot be 
+different number of vertices, lines, and faces so that these quantities can not be 
 compile-time constants anymore. This information used to be queried from
@@
 We tried---where possible---to provide utility functions so that users do not need to access a \texttt{GeometryInfo}-like data structure.
-However, these functions internally rely on the \texttt{ReferenceCell::\allowbreak internal::\allowbreak Info} classes, which are similarly structured as \texttt{GeometryInfo} and even use it for hypercube cells.
-Users can query the reference-cell type of geometric entity via:
+However, these functions internally rely on the \texttt{ReferenceCell} class, which is similarly structured as \texttt{GeometryInfo} and even use it for hypercube cells.
+Users can query the right \texttt{ReferenceCell} of a geometric entity via:
 \begin{c++}
-const auto reference_cell_type = cell->reference_cell_type();
+const auto reference_cell = cell->reference_cell();
 \end{c++}
-and can access the relevant struct describing the current reference cell or its i-th face via:
+and of its i-th face via:
 \begin{c++}
-ReferenceCell::internal::Info::get_cell(reference_cell_type);
-ReferenceCell::internal::Info::get_face(reference_cell_type, i);
+const auto face_reference_cell = reference_cell.face_reference_cell(i);
 \end{c++}
@@
 
+\subsubsection{Miscellaneous}
+
+We have created two new tutorials presenting the new simplex (\texttt{step-3b}) and mixed mesh features
+(\texttt{step-3c}). Similarly as \texttt{step-3}, they solve a Poisson problem, however, focus
+on the changed workflow.
+
+By the time of writing, there have been 74 tests (in the folder \texttt{tests/simplex}) 
+targeting the new simplex and mixed mesh support. In particular, the folder also
+contains ported variants of the following tutorials: 1, 2, 3, 4, 6, 7, 8, 12, 17, 18, 20, 23, 31, 38,
+40, 55, 67, 68, and 74. These might be also good starting points.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Until now, \dealii has only supported local smoothing multigrid algorithms~\citep{ClevengerHeisterKanschatKronbichler2019}.
-With release 9.3, the support for global coarsening of continuous (\texttt{FE\_Q}) and discontinuous (\texttt{FE\_DGQ}) elements  in a geometric and a polynomial sense has been added.
+With release 9.3, the support for global coarsening of continuous (\texttt{FE\_Q}, \texttt{FE\_SimplexP}) and discontinuous (\texttt{FE\_DGQ}, \texttt{FE\_SimplexDGP}) elements  in a geometric and a polynomial sense has been added.
 These multigrid variants promise less iteration numbers and better parallel scalability with the disadvantage of having to deal with hanging nodes during smoothing and of more work per iteration overall.
@@
 The projection and restriction are performed with efficient matrix-free kernels applying
-1D projection and restriction matrices.
+1D projection and restriction matrices in the case of hypercube cells and via full projection and
+restriction matrices in the case of simplex cells.
 
@@
 the polynomial degree of the finite element $k$ and the number of the 1D quadrature points $q$.
-For applications codes that rely on operators of many different degrees (e.g., because
+For application codes that rely on operators of many different degrees (e.g., because
 they use p-multigrid or hp algorithms), the instantiation
@@
 
-In the current release, we have improved the \texttt{FEEvaluation} and the \texttt{FEFaceEvaluation} class implementations, which do
+In the current release, we have improved the \texttt{FEEvaluation} and the \texttt{FEFaceEvaluation} class implementations that do
 not rely on the template parameters $k$ and $q$ (expressed in the code
@@
 
-The novel tutorial \texttt{step-67b} shows the usage of the new MPI-3.0 feature
+The novel tutorial \texttt{step-76} shows the usage of the new MPI-3.0 feature
  in the context of the solution
 of the Euler equations. Here, a speed-up of 27\% could be reached compared to the
-regular version \texttt{step-67}, by using the new feature.
+original version, \texttt{step-67}, by using the new feature.
 For more details and a successful application in the library \texttt{hyper.deal}, see \citep{munch2020hyperdeal}.

### 2021-03-10 9.3/paper.tex
Adjust funding

@@
 
-M.~Kronbichler was supported by the German
-Research Foundation (DFG) under the project ``High-order discontinuous
-Galerkin for the exa-scale'' (\mbox{ExaDG}) within the priority program ``Software
-for Exascale Computing'' (SPPEXA) and the Bayerisches Kompetenznetzwerk
+M.~Kronbichler was partially supported by the
+Bayerisches Kompetenznetzwerk
 f\"ur Technisch-Wissen\-schaft\-li\-ches Hoch- und H\"ochstleistungsrechnen
-(KONWIHR) in the context of the project
+(KONWIHR) in the context of the projects
 ``Performance tuning of high-order discontinuous Galerkin solvers for
-SuperMUC-NG''.
+SuperMUC-NG'' and ``High-order matrix-free finite element implementations with
+hybrid parallelization and improved data locality''.
 

### 2021-03-15 9.3/paper.tex
Adjust funding

@@
 
-M.~Kronbichler was partially supported by the
+M.~Kronbichler and P.~Munch wer partially supported by the
 Bayerisches Kompetenznetzwerk

### 2021-03-15 9.3/paper.tex
Fix typo

--- 2021-03-10 9.3/paper.tex
@@
 
-M.~Kronbichler was partially supported by the
+M.~Kronbichler and P.~Munch were partially supported by the
 Bayerisches Kompetenznetzwerk

### 2021-04-23 9.3/paper.tex
Added deal.II 9.2 publication and changes between 9.2 and 9.3.

@@
 can be found
-\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_1_1_and_9_2_0.html}{
-  in the file that lists all changes for this release}, see \cite{changes92}.
+\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_2_0_and_9_3_0.html}
+{in the file that lists all changes for this release}; see \cite{changes93}.
 
@@
 The 9.3 release includes
-\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_1_1_and_9_2_0.html}
-{around 60 incompatible changes}; see \cite{changes92}. The majority of these changes
+\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_2_0_and_9_3_0.html}
+{around 45 incompatible changes}; see \cite{changes93}. The majority of these changes
 should not be visible to typical user codes; some remove previously
@@
 The two previous releases of \dealii{} can be cited as
-\cite{dealII90,dealII91}.
+\cite{dealII91,dealII92}.
 

### 2021-04-23 9.3/paper.tex
Change affiliation for marcfehling

@@
    Collins, CO 80523-1874, USA.
-   \texttt{bangerth@colostate.edu}}
+   \texttt{bangerth/marc.fehling@colostate.edu}}
 
@@
 %
-\author[5]{Marc~Fehling}
-\affil[5]{Institute for Advanced Simulation,
-  Forschungszentrum J\"{u}lich GmbH,
-  52425 J\"{u}lich, Germany.
-  {\texttt{m.fehling@fz-juelich.de}}}
+\author[2]{Marc~Fehling}
+
 %
@@
 
-\subsubsection{Generating meshes }
+\subsubsection{Generating meshes}
 
@@
 \subsubsection{Mixed mesh}
-For mixed meshes, concepts known from the hp context have been applied: 
+For mixed meshes, concepts known from the $hp$-context have been applied:
  different finite-element classes are assigned to different cell types. In the
@@
 local data storage units needed for assembly might need to be resized for each
-cell as it is already good practice in the hp context:
+cell as it is already good practice in the $hp$-context:
 \begin{c++}
@@
 
-The usage of the new transfer operators (and of some of the utility functions) in the context of a hybrid multigrid algorithm (hp-MG with AMG as coarse-grid solver) for hp problems is demonstrated in the new tutorial \texttt{step-75} (see also Subsection~\ref{subsec:steps}).
+The usage of the new transfer operators (and of some of the utility functions) in the context of a hybrid multigrid algorithm ($hp$-MG with AMG as coarse-grid solver) for $hp$-problems is demonstrated in the new tutorial \texttt{step-75} (see also Subsection~\ref{subsec:steps}).
 
@@
 For application codes that rely on operators of many different degrees (e.g., because
-they use p-multigrid or hp algorithms), the instantiation
+they use $p$-multigrid or $hp$-algorithms), the instantiation
 means a long compilation time.
@@
 
-\subsubsection{Parallel matrix-free hp implementation}\label{subsubsection:mf:hp}
-
-With release 9.1, large parts of the hp algorithms in \dealii were parallelized so that
-parallel matrix-based simulations can be performed with the hp infrastructure. In the present
+\subsubsection{Parallel matrix-free $hp$-implementation}\label{subsubsection:mf:hp}
+
+With release 9.1, large parts of the $hp$-algorithms in \dealii were parallelized so that
+parallel matrix-based simulations can be performed with the $hp$-infrastructure. In the present
 release, we have extended the setup routines of \texttt{MatrixFree} so that it now also
-provides parallel hp support.
-
-Until now, the \texttt{FEEvaluation} classes used the template parameters $k$ and $q$ to select the correct active hp FE and quadrature index and users had 
+provides parallel $hp$-support.
+
+Until now, the \texttt{FEEvaluation} classes used the template parameters $k$ and $q$ to select the correct active FE and quadrature index and users had
 to use the function \texttt{MatrixFree::create\_cell\_\allowbreak subrange\_\allowbreak hp()} or
@@
 The creation of subranges is now performed internally, and the non-templated versions
-of the \texttt{FEEvaluation} classes are extended for the hp case. To nevertheless determine
-the active hp FE and quadrature index, the current cell/face range has to be provided
+of the \texttt{FEEvaluation} classes are extended for the $hp$-case. To nevertheless determine
+the active FE and quadrature index, the current cell/face range has to be provided
 the constructors of the \texttt{FEEvaluation} classes, from which the relevant information
 can be deduced (in the simplex case also the face type). These changes enable
-users to write matrix-free code independently of whether hp is used or not.
-
-The new tutorial \texttt{step-75} presents how to use the new hp-related features in \texttt{MatrixFree} 
+users to write matrix-free code independently of whether $hp$-capabilities are used or not.
+
+The new tutorial \texttt{step-75} presents how to use the new $hp$-related features in \texttt{MatrixFree}
 in the context of a hybrid-multigrid solver.
@@
 \item \texttt{step-19} TODO
+
 \item \texttt{step-68} TODO
-\item \texttt{step-75} TODO
+
+\item \texttt{step-75} demonstrates a state-of-the-art way of solving a simple
+      Laplace problem using $hp$-adaptation and hybrid multigrid methods on machines
+      with distributed memory. This tutorial points out particularities in porting
+      serial $hp$-adaptive code for parallelization. Furthermore, it guides through
+      the process of writing an efficient preconditioner for $hp$-adaptive applications.
+
 \item \texttt{step-76} is an explicit time integrator for the
       compressible Euler equations discretized with a high-order discontinuous
-        Galerkin (DG) scheme, using the matrix-free infrastructure just as \texttt{step67} does.
+      Galerkin (DG) scheme, using the matrix-free infrastructure just as \texttt{step-67} does.
       The tutorial presents advanced topics, like the usage of cell-centric loops and
@@
       integrals are discussed.
+
+\item \texttt{step-77} TODO
+
+\item \texttt{step-78} TODO
+
+\item \texttt{step-79} TODO
 \end{itemize}
@@
 \begin{itemize}
-  \item 
+  \item The \texttt{hp::DoFHandler} class has been deprecated. The standard
+        \texttt{DoFHandler} class is now capable of all $hp$-functionalities.
+  \item Consequently, all template arguments \texttt{DoFHandlerType} are obsolete,
+        and classes like \texttt{DataOut} or \texttt{SolutionTransfer} for example
+        no longer require them. If you rely on these template arguments, an interim
+        namespace \texttt{Legacy} has been introduced that provides all affected
+        classes with the old interface for the transition phase.
 \end{itemize}
@@
     \item For distributed parallel computing: \cite{BangerthBursteddeHeisterKronbichler11};
-    \item For $hp$~adaptivity: \cite{BangerthKayserHerold2007};
+    \item For $hp$-adaptivity: \cite{BangerthKayserHerold2007};
     \item For partition-of-unity (PUM) and enrichment methods of the

### 2021-04-23 9.3/paper.tex
Correctly format `hp`.

@@
    Collins, CO 80523-1874, USA.
-   \texttt{bangerth/marc.fehling@colostate.edu}}
+   \texttt{bangerth@colostate.edu}}
 
@@
 %
-\author[2]{Marc~Fehling}
-
+\author[5]{Marc~Fehling}
+\affil[5]{Institute for Advanced Simulation,
+  Forschungszentrum J\"{u}lich GmbH,
+  52425 J\"{u}lich, Germany.
+  {\texttt{m.fehling@fz-juelich.de}}}
 %
@@
 
-\subsubsection{Generating meshes}
+\subsubsection{Generating meshes }
 
@@
 \item \texttt{step-19} TODO
-
 \item \texttt{step-68} TODO
-
-\item \texttt{step-75} demonstrates a state-of-the-art way of solving a simple
-      Laplace problem using $hp$-adaptation and hybrid multigrid methods on machines
-      with distributed memory. This tutorial points out particularities in porting
-      serial $hp$-adaptive code for parallelization. Furthermore, it guides through
-      the process of writing an efficient preconditioner for $hp$-adaptive applications.
-
+\item \texttt{step-75} TODO
 \item \texttt{step-76} is an explicit time integrator for the
         compressible Euler equations discretized with a high-order discontinuous
-      Galerkin (DG) scheme, using the matrix-free infrastructure just as \texttt{step-67} does.
+        Galerkin (DG) scheme, using the matrix-free infrastructure just as \texttt{step67} does.
         The tutorial presents advanced topics, like the usage of cell-centric loops and
@@
         integrals are discussed.
-
-\item \texttt{step-77} TODO
-
-\item \texttt{step-78} TODO
-
-\item \texttt{step-79} TODO
 \end{itemize}
@@
 \begin{itemize}
-  \item The \texttt{hp::DoFHandler} class has been deprecated. The standard
-        \texttt{DoFHandler} class is now capable of all $hp$-functionalities.
-  \item Consequently, all template arguments \texttt{DoFHandlerType} are obsolete,
-        and classes like \texttt{DataOut} or \texttt{SolutionTransfer} for example
-        no longer require them. If you rely on these template arguments, an interim
-        namespace \texttt{Legacy} has been introduced that provides all affected
-        classes with the old interface for the transition phase.
+  \item 
 \end{itemize}

### 2021-04-23 9.3/paper.tex
Describe step-75
Write a bit more material.

--- 2021-04-23 9.3/paper.tex
@@
 
-\subsubsection{Miscellaneous}
+\subsubsection{Miscellanea}
 
@@
 
-Until now, \dealii has only supported local smoothing multigrid algorithms~\citep{ClevengerHeisterKanschatKronbichler2019}.
-With release 9.3, the support for global coarsening of continuous (\texttt{FE\_Q}, \texttt{FE\_SimplexP}) and discontinuous (\texttt{FE\_DGQ}, \texttt{FE\_SimplexDGP}) elements  in a geometric and a polynomial sense has been added.
-These multigrid variants promise less iteration numbers and better parallel scalability with the disadvantage of having to deal with hanging nodes during smoothing and of more work per iteration overall.
-
-Global-coarsening algorithms~\citep{becker00} either coarsen each cell of a triangulation or reduce the polynomial degree of the finite element of each cell and hereby obtain a new system with less unknowns (also known as "coarse level").
-Repeating such kind of coarsening results in a sequence of multigrid levels.
-
-The transfer between two levels has been encoded in the new class \texttt{MGTwoLevelTransfer}, which can be set up via the functions \texttt{MGTwoLevelTransfer::\allowbreak reinit\_\allowbreak geometric\_\allowbreak transfer()} or \texttt{MGTwo\allowbreak LevelTransfer::\allowbreak reinit\_\allowbreak polynomial\_\allowbreak transfer()} for given
+Until now, \dealii has only supported local smoothing multigrid
+algorithms~\citep{ClevengerHeisterKanschatKronbichler2019}; in local
+smoothing algorithms, one only applies the smoother to that part of an
+adaptively refined mesh that has cells on a given refinement level,
+but not to parts of the mesh that are coarser than that level.
+The current release now also has support for global coarsening~\citep{becker00} when
+using continuous (\texttt{FE\_Q}, \texttt{FE\_SimplexP}) and
+discontinuous (\texttt{FE\_DGQ}, \texttt{FE\_SimplexDGP})
+elements. Global coarsening builds multigrid levels for the entire
+domain, coarsening \textit{all} cells of a triangulation regardless of
+how many times they have been refined. In addition, the framework now available in
+\dealii{} in not only applicable to geometric coarsening, but can also
+perform coarsening by reducing the polynomial degree used, for example
+to support $hp$-adaptive meshes where it has been shown that one can
+efficiently precondition high-order spaces with lower-order finite
+element operators.
+Finally, the implementation also supports transfer between continuous
+and discontinuous elements by providing \texttt{DoFHandler}s set up
+with these kinds of elements; the scheme therefore allows
+preconditioning of discontinuous schemes with a continuous
+discretization with fewer unknowns, for example.
+
+These new multigrid variants promise fewer solver
+iterations and better parallel scalability than the existing local
+smoothing algorithms, but have to deal with
+hanging nodes within each level and general require more computational
+work per iteration overall.
+
+The transfer operators between two levels has been implemented in the new class \texttt{MGTwoLevelTransfer}, which can be set up via the functions \texttt{MGTwoLevelTransfer::\allowbreak reinit\_\allowbreak geometric\_\allowbreak transfer()} or \texttt{MGTwo\allowbreak LevelTransfer::\allowbreak reinit\_\allowbreak polynomial\_\allowbreak transfer()} for given
 \texttt{DoFHandler} and \texttt{AffineConstraint} classes of two levels. The resulting transfer operators
-between two levels can be collected in a single object
-(\texttt{MGTransferGlobalCoarsening}) that can be used just as \texttt{MGTransferMatrixFree} and can be passed to the actual \texttt{Multigrid}
-algorithm. The option that users can set up the transfer operators between two levels on their own enables
-the mix of polynomial  and geometric coarsening, as needed. It should also be noted that a transfer between continuous
-and discontinuous elements is possible by providing \texttt{DoFHandler}s set up with this kind of elements.
-
-Internally, the transfer operators categorize fine/coarse cell pairs according to the
-occurrence of refinement
-(refinement or no-refinement) and to the polynomial-degree pair ($k_f$-$k_c$) into groups.
-The projection and restriction are performed with efficient matrix-free kernels applying
-1D projection and restriction matrices in the case of hypercube cells and via full projection and
-restriction matrices in the case of simplex cells.
-
-The user is responsible for setting up the levels (including the operators and the
-smoothers). Please note that global coarsening can currently only be performed between active levels.
-The user can apply utility
-functions from the \texttt{MGTransferGlobalCoarseningTools} namespace to set up the levels. For
-constructing the diagonal or the system matrix for a matrix-free operator,
-needed for the smoother or the coarse-grid solver,
+can then be collected in a single
+\texttt{MGTransferGlobalCoarsening} object that can be used just as \texttt{MGTransferMatrixFree} and can be passed to the actual \texttt{Multigrid}
+algorithm.
+Several common operations are encoded in utility
+functions in the \texttt{MGTransferGlobalCoarseningTools}
+namespace. In practice, matrix-free methods typically require
+constructing the diagonal of the system matrix for smoothing
+operations, and the full matrix on the coarse level for the coarse-grid solver;
 the functions \texttt{create\_diagonal()} and \texttt{create\_matrix()} from
-the \texttt{MatrixFreeTools} namespace can be used (see also Subsection~\ref{subsec:mf}).
-
-The usage of the new transfer operators (and of some of the utility functions) in the context of a hybrid multigrid algorithm ($hp$-MG with AMG as coarse-grid solver) for $hp$-problems is demonstrated in the new tutorial \texttt{step-75} (see also Subsection~\ref{subsec:steps}).
+the \texttt{MatrixFreeTools} namespace can be used to this end (see also Subsection~\ref{subsec:mf}).
+
+The usage of the new transfer operators (and of some of the utility
+functions) in the context of a hybrid multigrid algorithm
+($hp$-multigrid with algebraic multigrid as coarse-grid solver) for $hp$-problems is demonstrated in the new tutorial \texttt{step-75}, see also Section~\ref{subsec:steps}.
 
@@
 For application codes that rely on operators of many different degrees (e.g., because
-they use $p$-multigrid or $hp$-algorithms), the instantiation
-means a long compilation time.
-
-In the current release, we have improved the \texttt{FEEvaluation} and the \texttt{FEFaceEvaluation} class implementations that do
+they use $p$-multigrid or $hp$-algorithms), the many instantiations
+necessary and their complexity
+imply long compile times.
+
+In the current release, we have improved specializations of these classes that do
 not rely on the template parameters $k$ and $q$ (expressed in the code
-with "-1" and "0"):
+with ``-1'' and ``0'') -- for example:
 \begin{c++}
@@
 \end{c++}
-The meaning of the new first argument is
-discussed in Subsubsection~\ref{subsubsection:mf:hp}.
-
-These classes select at runtime---for common low and medium polynomial degree/quadrature combinations ($k\le 6$ and $q\in\{ k+1, k+2, \lfloor (3k)/2 \rfloor \}$)---efficient precompiled implementations and default to non-templated
+These classes select at runtime -- for common low and medium
+polynomial degree/quadrature combinations ($k\le 6$ and $q\in\{ k+1,
+k+2, \left\lfloor (3k)/2 \right\rfloor \}$) -- efficient precompiled implementations and default to non-templated
 evaluation kernels else.
@@
 
-With release 9.1, large parts of the $hp$-algorithms in \dealii were parallelized so that
+With release 9.1 \cite{dealII91}, large parts of the $hp$-algorithms in \dealii were parallelized so that
 parallel matrix-based simulations can be performed with the $hp$-infrastructure. In the present
-release, we have extended the setup routines of \texttt{MatrixFree} so that it now also
-provides parallel $hp$-support.
+release, we have extended the setup routines of \texttt{MatrixFree} so that they now also
+provide parallel $hp$-support.
 
@@
 the active FE and quadrature index, the current cell/face range has to be provided
-the constructors of the \texttt{FEEvaluation} classes, from which the relevant information
+to the constructors of the \texttt{FEEvaluation} classes, from which the relevant information
 can be deduced (in the simplex case also the face type). These changes enable
@@
 
-\subsubsection{MPI-3.0 shared-memory support}
-The classes \texttt{MatrixFree} and \texttt{LinearAlgebra::\allowbreak distributed::\allowbreak Vector} have been extended with
-MPI-3.0 shared-memory features. If \texttt{MatrixFree} has been configured by setting
-\texttt{MatrixFree::AdditionalData::communicator\_sm} by an appropriate user-provided subcommunicator,
-vectors are created by \texttt{MatrixFree::create\_dof\_vector()} in such a way that the
-memory of processes in this subcommunicator can be accessed. In particular, this means
-that the \texttt{FEEvaluation} classes can access the degrees of freedom owned by
+\subsection{MPI-3.0 shared-memory support}
+In many large computations, certain pieces of data are computed once
+and then threated as read-only. If this information is needed by more
+than one MPI process, it is more efficient to store this information only
+once in shared memory among all processes located on a multicore
+node. MPI supports the creation of such shared memory windows since
+version 3.0, and deal.II can now use this in the
+\texttt{AlignedVector} and \texttt{Table} classes that are often used
+for large lookup tables.
+
+The feature is also used in the \texttt{MatrixFree} and
+\texttt{LinearAlgebra::\allowbreak distributed::\allowbreak Vector} classes. If \texttt{MatrixFree} has been configured by setting
+\texttt{MatrixFree::AdditionalData::communicator\_sm} appropriately, then
+\texttt{MatrixFree::create\_dof\_vector()} creates vectors that share
+information among all processes on one node. As a consequence, the
+\texttt{FEEvaluation} classes can access vector elements owned by
 other processes and in certain cases local
@@
 
-The novel tutorial \texttt{step-76} shows the usage of the new MPI-3.0 feature
- in the context of the solution
-of the Euler equations. Here, a speed-up of 27\% could be reached compared to the
+The new \texttt{step-76} tutorial program illustrates this latter case
+in the context of the solution of the Euler equations. \texttt{step-76} reaches a
+speed-up of 27\% compared to the 
 original version, \texttt{step-67}, by using the new feature.
-For more details and a successful application in the library \texttt{hyper.deal}, see \citep{munch2020hyperdeal}.
-
-\subsubsection{Miscellaneous}
-
-The namespace \texttt{MatrixFreeTools} has been introduced. It collects various utility
-functions useful in the context of application of \texttt{MatrixFree}. In particular,
-it contains different functions to extract quantities related to a matrix, e.g., the
-computation of the diagonal or a matrix with a matrix-free approach.
-
-With version 9.2, we have introduced cell-centric loops. They have been extended to unstructured
-meshes. The tutorial \texttt{step-67b} shows their usage in the context of the solution
-of the Euler equations.
+For more details and use of the feature in the library \texttt{hyper.deal}, see \citep{munch2020hyperdeal}.
 
@@
 \begin{itemize}
-\item \texttt{step-19} TODO
+\item \texttt{step-19} is an introductory demonstration of \dealii{}'s
+  particle functionality. It solves the coupled problem of
+  charged particles and an electric field, using a cathode tube as an
+  example.
 
 \item \texttt{step-68} TODO
+
+\item \texttt{step-71} focused on automatic and symbolic
+  differentiation (AD and SD, in short) as a tool to make solvers for complex, nonlinear
+  problems possible. To this end, \dealii{} can interface to a number
+  of AD and SD libraries, including Trilinos' Sacado package
+  \cite{...}, ...\todo{which others??}, and SymEngine \cite{symengine-web-page}. The tutorial
+  illustrates how these techniques can be used to compute derivatives
+  of first a rather simple function, and then of the much more complex
+  energy functions of two magneto-elastic and magneto-visco-elastic
+  material formulations in which just the scalar energy functional
+  takes up the better part of a page, and even first derivatives can
+  only be computed with heroic effort.
+
+\item \texttt{step-72} illustrates the use of automatic
+  differentiation to simplify the computation of derivatives in the
+  context of nonlinear partial differential equations where one needs
+  to compute the Jacobian from the residual operator for efficient
+  Newton iterations. \texttt{step-72} builds on the minimal surface
+  solver \texttt{step-15} and replaces the hand-construction of the
+  Jacobian by either computing it as the derivative of the residual,
+  or alternatively as the second derivative of an energy functional
+  that then also yields the residual itself.
 
@@
 
-\item \texttt{step-77} TODO
+\item \texttt{step-77} is a program that illustrate \texttt{dealii}'s
+  interfaces to the SUNDIALS library \cite{sundials}, and specifically
+  the KINSOL nonlinear solver. Like the \texttt{step-72} program
+  mentioned above, it is a variation of the minimal surface solver
+  \texttt{step-15}: Instead of implementing the nonlinear Newton and
+  line search loop ourselves, \texttt{step-77} relies on KINSOL for
+  decisions such as when to rebuild the Jacobian matrix, when to
+  actually solve linear systems with it, and how to form updates that
+  drive the residual to convergence. The program illustrates the
+  substantial savings that can be obtained by not re-inventing the
+  wheel but instead building on an existing and well-tuned software
+  such as KINSOL.
+
+  \dealii{}'s interfaces were also updated to the latest SUNDIALS release, 5.7.
 

### 2021-04-23 9.3/paper.tex
Fix a grammar error.

@@
 
-\item \texttt{step-71} focused on automatic and symbolic
+\item \texttt{step-71} focuses on automatic and symbolic
   differentiation (AD and SD, in short) as a tool to make solvers for complex, nonlinear

### 2021-05-08 9.3/paper.tex
Fix a typo.
Update contributors

@@
 In many large computations, certain pieces of data are computed once
-and then threated as read-only. If this information is needed by more
+and then treated as read-only. If this information is needed by more
 than one MPI process, it is more efficient to store this information only
@@
 %
-% 9.2: updated 5/11/2020 MM
-TODO
+% 9.3: updated 5/08/2021 PM
+Alexander Grayver,
+Andrew Davis,
+Arpit Babbar,
+Bruno Blais,
+Bruno Turcksin,
+Ce Qin,
+Christoph Kammer,
+Conrad Clevenger,
+Dominic Soldner,
+Daniel Arndt,
+Daniel Garcia-Sanchez,
+Daniel Paukner,
+David F. Castellanos,
+David Wells,
+Doug Shi-Dong,
+David Schneider,
+Elias Dejene,
+Graham Harper,
+Ignacio Tomas,
+Ivan Fumagalli,
+Jake Harmon,
+Jean-Paul Pelteret,
+Jiaqi Zhang,
+Julian Roth,
+Katharina Kormann,
+Konrad Simon,
+Krishnakumar Gopalakrishnan,
+Laura Prieto Saavedra,
+Lei Qiao,
+Luca Heltai,
+Magdalena Schreter,
+Malhar Tidke,
+Marc Fehling,
+Marcus Calhoun-Lopez,
+Martin Kronbichler ,
+Mathias Anselmann,
+Matthias Maier,
+Maximilian Bergbauer,
+Menno Fraters,
+Michele Bucelli,
+Nicola Giuliani,
+Nicolas Barnafi,
+Niklas Fehn,
+Nils Much,
+Olivier Guevremont,
+Pasquale Africa,
+Peter Munch,
+Peter Westerbaan,
+Phillip Mobley,
+Pratik Nayak,
+Praveen Chandrashekar,
+Rene Gassmoeller,
+Reza Rastak,
+Richard Schussnig,
+Sebastian Proell,
+Simon Sticko,
+Timo Heister,
+Toni El Geitani Nehme,
+Wenyu Lei,
+Wolfgang Bangerth,
+Zhou Lei
 

### 2021-05-08 9.3/paper.tex
Lightly edit section on step-71

@@
 \item \texttt{step-71} focuses on automatic and symbolic
-  differentiation (AD and SD, in short) as a tool to make solvers for complex, nonlinear
-  problems possible. To this end, \dealii{} can interface to a number
-  of AD and SD libraries, including Trilinos' Sacado package
-  \cite{...}, ...\todo{which others??}, and SymEngine \cite{symengine-web-page}. The tutorial
+  differentiation (AD and SD, in short) as a tool to make solvers for complex,
+  nonlinear problems possible. To this end, \dealii{} can interface to a number
+  of AD and SD libraries, specifically Trilinos' Sacado package
+  \cite{Bartlett2006a}, ADOL-C \cite{Griewank1996a}, and 
+  SymEngine \cite{symengine-web-page}. The tutorial
   illustrates how these techniques can be used to compute derivatives
   of first a rather simple function, and then of the much more complex
-  energy functions of two magneto-elastic and magneto-visco-elastic
+  energy functions of two magnetoelastic and magneto-viscoelastic
   material formulations in which just the scalar energy functional

### 2021-05-22 9.3/paper.tex
add my affil
Improve some text. Write about FEPointEvaluation. Add references.

@@
 \author[4]{Timo~Heister}
-
-\author[7]{Luca~Heltai}
-\affil[7]{SISSA,
+ \affil[4]{School of Mathematical and Statistical Sciences,
+   Clemson University,
+   Clemson, SC, 29634, USA
+   {\texttt{tcleven/heister@clemson.edu}}}
+
+\author[5]{Luca~Heltai}
+\affil[5]{SISSA,
    International School for Advanced Studies,
@@
 
- \author[8]{Martin~Kronbichler}
- \affil[8]{Institute for Computational Mechanics,
+ \author[6,7]{Martin~Kronbichler}
+ \affil[6]{Institute for Computational Mechanics,
    Technical University of Munich,
@@
    {\texttt{kronbichler/munch@lnm.mw.tum.de}}}
-
-\author[9]{Matthias~Maier}
-\affil[9]{Department of Mathematics,
+ \affil[7]{Department of Information Technology,
+   Uppsala University,
+   Box 337, 751\,05 Uppsala, Sweden.
+   {\texttt{martin.kronbichler@it.uu.se}}}
+
+\author[8]{Matthias~Maier}
+\affil[8]{Department of Mathematics,
   Texas A\&M University,
@@
 
-\author[8,10]{Peter Munch}
+\author[6,9]{Peter Munch}
  \affil[10]{Institute of Materials Research, Materials Mechanics,
@@
 
-\author[11]{Jean-Paul~Pelteret}
-\affil[11]{Independent researcher.
+\author[10]{Jean-Paul~Pelteret}
+\affil[10]{Independent researcher.
 {\texttt{jppelteret@gmail.com}}}
@@
 We provide matrix-free support for simplex and mixed meshes as well as for
-continuous and discontinuous elements. From the user perspective, the setup has
-hardly changed:
-%\begin{c++}
-%Simplex::QGauss<dim> quad(degree + 1)
-%
-%MatrixFree matrix_free;
-%matrix_free.reinit(mapping, dof_handler, constraints, quad);
-%\end{c++}
-Instead of providing a 1D quadrature rule, which is extended to higher dimensional
-spaces via a tensor product, one provides d-dimensional quadrature rules.
-The templated versions of \texttt{FEEvaluation} and \texttt{FEFaceEvaluation} cannot be used: the number of degrees of freedom and the number of quadrature points are determined at runtime.
-More details can be found in Subsection~\ref{subsec:mf}.
-
-By the time of writing, we do not use internally any advance techniques for evaluating values and
-gradients at the quadrature points as we do for tensor-product elements and
-instead rely on full interpolation matrices. This is for now acceptable since only low-order elements are
+continuous and discontinuous elements. From a user perspective, the main
+changes are to pass a d-dimensional quadrature rules rather than a 1D formula,
+and the fact that \texttt{FEEvaluation} and \texttt{FEFaceEvaluation} must not
+specify the polynomial degree via template arguments, determining all
+information at runtime. More details can be found in
+Subsection~\ref{subsec:mf}.
+
+By the time of writing, no advanced algorithms for evaluating values and
+gradients at the quadrature points such as sum factorization are used.
+The use of full interpolation matrices is for now acceptable since only low-order elements are
 supported.
@@
 algorithms~\citep{ClevengerHeisterKanschatKronbichler2019}; in local
-smoothing algorithms, one only applies the smoother to that part of an
-adaptively refined mesh that has cells on a given refinement level,
-but not to parts of the mesh that are coarser than that level.
+smoothing algorithms, smoothers only act on the cells of a given refinement level,
+skipping those parts of the mesh that are not adaptively refined to that level.
 The current release now also has support for global coarsening~\citep{becker00} when
@@
 \dealii{} in not only applicable to geometric coarsening, but can also
-perform coarsening by reducing the polynomial degree used, for example
-to support $hp$-adaptive meshes where it has been shown that one can
-efficiently precondition high-order spaces with lower-order finite
-element operators.
+perform coarsening by reducing the polynomial degree ($p$-multigrid,
+see~\cite{Fehn2020}), for example
+to support $hp$-adaptive meshes.
+%where it has been shown that one can
+%efficiently precondition high-order spaces with lower-order finite
+%element operators.
+% MK: I do not think "precondition" is a precise enough description of p-multigrid
+% because one still needs smoothing on the fine levels in addition to the
+% "solve" at the coarser degree, a fact not transported by the mere
+% 'precondition' term; rather than trying to discuss this in a more precise way,
+% let us simply skip it as that is not central to the presentation here.
 Finally, the implementation also supports transfer between continuous
-and discontinuous elements by providing \texttt{DoFHandler}s set up
-with these kinds of elements; the scheme therefore allows
-preconditioning of discontinuous schemes with a continuous
-discretization with fewer unknowns, for example.
+and discontinuous elements as a further way to create multigrid levels.
 
@@
 can then be collected in a single
-\texttt{MGTransferGlobalCoarsening} object that can be used just as \texttt{MGTransferMatrixFree} and can be passed to the actual \texttt{Multigrid}
+\texttt{MGTransferGlobalCoarsening} object that can be used just as the previous workhorse \texttt{MGTransferMatrixFree} within the \texttt{Multigrid}
 algorithm.
-Several common operations are encoded in utility
-functions in the \texttt{MGTransferGlobalCoarseningTools}
-namespace. In practice, matrix-free methods typically require
-constructing the diagonal of the system matrix for smoothing
-operations, and the full matrix on the coarse level for the coarse-grid solver;
-the functions \texttt{create\_diagonal()} and \texttt{create\_matrix()} from
-the \texttt{MatrixFreeTools} namespace can be used to this end (see also Subsection~\ref{subsec:mf}).
+%Several common operations are encoded in utility
+%functions in the \texttt{MGTransferGlobalCoarseningTools}
+%namespace.
+To facilitate the construction of matrix diagonals with matrix-free methods as well as a matrix representation of the coarse level matrix, new utility functions \texttt{create\_diagonal()} and \texttt{create\_matrix()} from
+the \texttt{MatrixFreeTools} namespace have been added. (see also Subsection~\ref{subsec:mf}).
 
@@
 
-The classes \texttt{FEEvaluation} and \texttt{FEFaceEvaluation} are highly
-templated to reach high performance. In particular, the template parameters include
-the polynomial degree of the finite element $k$ and the number of the 1D quadrature points $q$.
+The classes \texttt{FEEvaluation} and \texttt{FEFaceEvaluation} use template parameters for
+the polynomial degree of the finite element $k$ and the number of the 1D quadrature points $q$
+to generate near-optimal code for these operations.
 For application codes that rely on operators of many different degrees (e.g., because
-they use $p$-multigrid or $hp$-algorithms), the many instantiations
-necessary and their complexity
-imply long compile times.
-
-In the current release, we have improved specializations of these classes that do
+they use $p$-multigrid or $hp$-algorithms), creating all instantiations
+can be overly complex and incur long compile times.
+
+In the current release, specializations of these classes that do
 not rely on the template parameters $k$ and $q$ (expressed in the code
-with ``-1'' and ``0'') -- for example:
+with ``-1'' and ``0'') have been added. For example:
 \begin{c++}
@@
 
-With release 9.1 \cite{dealII91}, large parts of the $hp$-algorithms in \dealii were parallelized so that
-parallel matrix-based simulations can be performed with the $hp$-infrastructure. In the present
-release, we have extended the setup routines of \texttt{MatrixFree} so that they now also
-provide parallel $hp$-support.
-
-Until now, the \texttt{FEEvaluation} classes used the template parameters $k$ and $q$ to select the correct active FE and quadrature index and users had
-to use the function \texttt{MatrixFree::create\_cell\_\allowbreak subrange\_\allowbreak hp()} or
-\texttt{::create\_cell\_\allowbreak subrange\_\allowbreak hp\_\allowbreak by\_index()} to create subranges of cells with
-the same polynomial degree. This led to user codes that were hard to read due to complicated jump tables.
-
+With release 9.1 \cite{dealII91}, large parts of the $hp$-algorithms in \dealii were parallelized, enabling
+parallel matrix-based simulations with the $hp$-infrastructure. In the present
+release, the parallel $hp$ support was extended to \texttt{MatrixFree}.
+
+Until now, the \texttt{FEEvaluation} classes used the template parameters $k$ and $q$ to select the correct active FE and quadrature index and cumbersome detection of subranges of the same degree.
 The creation of subranges is now performed internally, and the non-templated versions
-of the \texttt{FEEvaluation} classes are extended for the $hp$-case. To nevertheless determine
-the active FE and quadrature index, the current cell/face range has to be provided
-to the constructors of the \texttt{FEEvaluation} classes, from which the relevant information
-can be deduced (in the simplex case also the face type). These changes enable
+of the \texttt{FEEvaluation} classes have been extended for the $hp$-case. To determine
+the desired FE and quadrature index of a subrange, the current cell/face range has to be provided
+to the constructors of the \texttt{FEEvaluation} classes. These changes enable
 users to write matrix-free code independently of whether $hp$-capabilities are used or not.
@@
 
-Example: testing of surface tension in the context of sharp-interface methods:
+In a number of circumstances, finite element solutions need to be evaluated on
+arbitrary reference points that change from one element to next. Two important
+examples are particle simulations coupled to a finite element solution, or
+algorithms on non-matching grids. The existing \texttt{FEValues} class a poor
+fit for this task, as it relies on tabulating all information before use,
+necessitating a separate setup for every cell. The new class
+\texttt{FEPointEvaluation} provides a more convenient interface for this
+task. For tensor product finite elements (\texttt{FE\_Q}, \texttt{FE\_DGQ})
+and tensor product mappings (\texttt{MappingQGeneric} and derived classes),
+the new approach is also very fast, as it can use some of the matrix-free
+infrastructure and vectorization facilities.
+
+To give an example of the new features, let us consider the evaluation of a
+surface tension force in the context of sharp-interface methods, whose
+contribution is added to a fluid solver by multiplication with test function
+and addition over quadrature points:
 \begin{align*}
@@
 \approx
-\sum\left(\vec{v}, \kappa(\vec{p}_q) \, \vec{n}(\vec{p}_q) (JxW)_q\right)
+\sum_q\left(\vec{v}, \kappa(\vec{p}_q) \, \vec{n}(\vec{p}_q) (JxW)_q\right).
 \end{align*}
-in code:
+In \dealii{}, this can now be conveniently written as
 \begin{c++}
-phi_normal.evaluate(cell, points, normal_values, EvaluationFlags::values);
+phi_normal.reinit(cell, reference_points);
+phi_force.reinit(cell, reference_points);
+phi_normal.evaluate(normal_values, EvaluationFlags::values);
 for (unsigned int q = 0; q < n_points; ++q)
@@
                           phi_curvature.get_value(q) * JxW[q], q);
-phi_force.integrate(cell, points, force_values, EvaluationFlags::values);
+phi_force.integrate(force_values, EvaluationFlags::values);
 \end{c++}
-The quadrature points and the connected \texttt{JxW} can, e.g., come from a
-codim-1 mesh. Determining to which \texttt{cell} a quadrature point belongs
-to on the background mesh, incl. the reference-cell coordinates \texttt{points},
-can be determined with know functions like \texttt{GridTools::find\_all\_active\_cells\_around\_point()}.
+The quadrature points and the related \texttt{JxW} value can, e.g., come from
+a mesh of codimension one. Determining to which \texttt{cell} a quadrature
+point belongs to on the background mesh, including the reference-cell
+coordinates \texttt{reference\_points}, can be determined with functions like
+\texttt{GridTools::find\_all\_active\_cells\_around\_point()}. While these
+functions have been available in \dealii{} previously, their performance has
+been considerably enhanced with the aforementioned more optimized code paths
+for selected mappings.
 

### 2021-05-24 9.3/paper.tex
Add some relevant minor-change topics

@@
   \item globally unique cell IDs: TODO
+  \item In addition to the macro \texttt{DEAL\_II\_DEPRECATED}, we have introduced
+  the \texttt{DEAL\_II\_DEPRECATED\_EARLY} macro, which indicates that a feature will be
+  deprecated in the next release. In contrast to the first macro, it will only give
+  warnings if \dealii{} has been configured with \texttt{-D DEAL\_II\_EARLY\_DEPRECATIONS=ON}.
+  \item After each update of the master branch of \dealii{}, we build a new Docker image
+  with all features enabled. It is
+  accessible on \texttt{Docker Hub} via \texttt{dealii/dealii:master-focal}. This image is
+  particular useful when used in the contiguous-integration processes of user codes.
 \end{itemize}

### 2021-05-24 9.3/paper.tex
Write text

@@
 \begin{itemize}
-  \item TODO.
+  \item An experimental support for simplex and mixed meshes (see Section~\ref{subsec:simplex});
+  \item Improved flexibility of the particle infrastructure (see Section~\ref{subsec:particles});
+  \item Support for global-coarsening multigrid algorithms (see Section~\ref{subsec:mg});
+  \item Advances in the matrix-free infrastructure  (see Section~\ref{subsec:mf});
+  \item Usage of MPI-3.0 shared-memory features to reduce memory footprint  (see Section~\ref{subsec:sm});
+  \item Improved support for evaluation and integration at arbitrary points (see Section~\ref{subsec:fepointvalues});
+  \item Nine new tutorial programs and a new code gallery program (see Section~\ref{subsec:steps}).
 \end{itemize}
@@
 \begin{itemize}
-  \item globally unique cell IDs: TODO
+  \item Each non-artificial cell posses now an globally unique index, which can be queried
+  for active cells via \texttt{CellAccessor::global\_active\_cell\_index()} and for level cells 
+  via \texttt{::global\_\allowbreak level\_\allowbreak cell\_\allowbreak index()}. The information
+  can be used to efficiently access global cell vectors. Users have been able to construct the 
+  same information by distributing \texttt{FE\_DGQ(0)} in a \texttt{DoFHandler}.
   \item In addition to the macro \texttt{DEAL\_II\_DEPRECATED}, we have introduced
-  the \texttt{DEAL\_II\_DEPRECATED\_EARLY} macro, which indicates that a feature will be
+  the \texttt{DEAL\_II\_DEPRECATED\_\allowbreak EARLY} macro, which indicates that a feature will be
   deprecated in the next release. In contrast to the first macro, it will only give
@@
 %
-The changelog lists more than 240 other
+The changelog lists more than ?? other
 features and bugfixes.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Simplex and mixed mesh support}
+\subsection{Experimental simplex and mixed mesh support}
 \label{subsec:simplex}
 
-\begin{figure}[!h]
+\begin{figure}[!t]
 
@@
 
-The current release of \dealii adds support for simplex meshes (consisting of triangles in 2D; tetrahedra in 3D) and mixed meshes (consisting of triangles and/or
+The current release of \dealii adds an experimental support for simplex meshes (consisting of triangles in 2D; tetrahedra in 3D) and mixed meshes (consisting of triangles and/or
 quadrilaterals in 2D; tetrahedra, pyramids, wedges, and/or hexahedra in 3D).
@@
 cells and vertices to the internal data structures, has been rewritten
-inspired by \citep{logg09} and a speed-up of up to 5 has been reached. Minor adjustments
+inspired by \citep{logg2012} and a speed-up of up to 5 has been reached. Minor adjustments
 have been made to \texttt{parallel::shared::Triangulation} and \texttt{parallel::\allowbreak fullydistributed::\allowbreak Triangulation} such that
@@
 \texttt{DoFHandler} in the context of more general meshes, we have merged
-\texttt{hp::DoFHandler} into the \texttt{DoFHandler}. The class \texttt{hp::DoFHandler}, which is now a dummy derivation of \texttt{DoFHandler}, currently only exists for compatibility reasons. For
-more details see Subsection~\ref{TODO}.
+\texttt{hp::DoFHandler} into the \texttt{DoFHandler}. The class \texttt{hp::DoFHandler}, which is now a dummy derivation of \texttt{DoFHandler}, currently only exists for compatibility reasons and has
+been deprecated, see Section~\ref{subsec:deprecated}.
 
@@
 generated by an external mesh generator. Currently, we support the
-file formats \texttt{VTK}, \texttt{MSH}, \texttt{EXODUS II}.
+file formats \texttt{VTK}, \texttt{MSH}, and \texttt{EXODUS II}.
 
@@
 The  list of currently supported finite-element classes is provided in Table~\ref{tab:simplex:fe}. Currently,
-only linear iso-parametric mapping (\texttt{MappingFE} and \texttt{MappingFEFields}) is available. For quadrature, the classes \texttt{QGaussSimplex}, \texttt{QWitherdenVincentSimplex},
+only linear iso-parametric mapping (\texttt{MappingFE} and \texttt{MappingFEFields}) is available.
+For quadrature, the classes \texttt{QDuffy}, \texttt{QGaussSimplex}, \texttt{QWitherdenVincentSimplex},
 \texttt{QGaussPyamid}, and \texttt{QGaussWedge} are
@@
 
-We have created two new tutorials presenting the new simplex (\texttt{step-3b}) and mixed mesh features
-(\texttt{step-3c}). Similarly as \texttt{step-3}, they solve a Poisson problem, however, focus
-on the changed workflow.
-
-By the time of writing, there have been 74 tests (in the folder \texttt{tests/simplex})
+Further information can be found on the new module page ``Simplex support (experimental)'', see
+\begin{center}
+  \url{https://www.dealii.org/current/doxygen/deal.II/group__simplex.html}
+\end{center}
+As an example, it shows how to solve a simple Poisson problem like in \texttt{step-3}
+on simplex and mixed meshes, with the focus on on the changed workflow.
+
+By the time of writing, there have been 92 tests (in the folder \texttt{tests/simplex})
 targeting the new simplex and mixed mesh support. In particular, the folder also
@@
 skipping those parts of the mesh that are not adaptively refined to that level.
-The current release now also has support for global coarsening~\citep{becker00} when
+The current release now also has support for global coarsening~\citep{becker2000multigrid, sundar2012parallel} when
 using continuous (\texttt{FE\_Q}, \texttt{FE\_SimplexP}) and
@@
 
-The transfer operators between two levels has been implemented in the new class \texttt{MGTwoLevelTransfer}, which can be set up via the functions \texttt{MGTwoLevelTransfer::\allowbreak reinit\_\allowbreak geometric\_\allowbreak transfer()} or \texttt{MGTwo\allowbreak LevelTransfer::\allowbreak reinit\_\allowbreak polynomial\_\allowbreak transfer()} for given
+The transfer operators between two levels have been implemented in the new class \texttt{MGTwoLevel\allowbreak Transfer}, which can be set up via the functions \texttt{MGTwoLevel\allowbreak Transfer::\allowbreak reinit\_\allowbreak geometric\_\allowbreak transfer()} or \texttt{MGTwo\allowbreak LevelTransfer::\allowbreak reinit\_\allowbreak polynomial\_\allowbreak transfer()} for given
 \texttt{DoFHandler} and \texttt{AffineConstraint} classes of two levels. The resulting transfer operators
 can then be collected in a single
-\texttt{MGTransferGlobalCoarsening} object that can be used just as the previous workhorse \texttt{MGTransferMatrixFree} within the \texttt{Multigrid}
+\texttt{MGTransfer\allowbreak GlobalCoarsening} object that can be used just as the previous workhorse \texttt{MGTransferMatrixFree} within the \texttt{Multigrid}
 algorithm.
@@
 To facilitate the construction of matrix diagonals with matrix-free methods as well as a matrix representation of the coarse level matrix, new utility functions \texttt{create\_diagonal()} and \texttt{create\_matrix()} from
-the \texttt{MatrixFreeTools} namespace have been added. (see also Subsection~\ref{subsec:mf}).
+the \texttt{MatrixFreeTools} namespace have been added (see also Subsection~\ref{subsec:mf}).
 
@@
 
-\subsection{MPI-3.0 shared-memory support}
+\subsection{MPI-3.0 shared-memory support}\label{subsec:sm}
 In many large computations, certain pieces of data are computed once
@@
 \texttt{FEEvaluation} classes can access vector elements owned by
-other processes and in certain cases local
+other processes and in certain cases node-local
 communication can be skipped. To prevent race conditions, \texttt{MatrixFree} uses local
@@
 original version, \texttt{step-67}, by using the new feature.
-For more details and use of the feature in the library \texttt{hyper.deal}, see \citep{munch2020hyperdeal}.
+For more details and the usage of the feature in the library \texttt{hyper.deal}, see \citep{munch2020hyperdeal}.
 
@@
 examples are particle simulations coupled to a finite element solution, or
-algorithms on non-matching grids. The existing \texttt{FEValues} class a poor
-fit for this task, as it relies on tabulating all information before use,
+algorithms on non-matching grids. The existing \texttt{FEValues} class is a poor
+fit for this task, as it relies on tabulating all information before its usage,
 necessitating a separate setup for every cell. The new class
@@
 for (unsigned int q = 0; q < n_points; ++q)
-   phi_force.submit_value(phi_normal.get_value(q) *
-                          phi_curvature.get_value(q) * JxW[q], q);
+   phi_force.submit_value(phi_curvature.get_value(q) *
+                          phi_normal.get_value(q) * JxW[q], q);
 phi_force.integrate(force_values, EvaluationFlags::values);
 \end{c++}
-The quadrature points and the related \texttt{JxW} value can, e.g., come from
+The quadrature points (at reference positions - \texttt{reference\_points}) and the related \texttt{JxW} value can, e.g., come from
 a mesh of codimension one. Determining to which \texttt{cell} a quadrature
@@
 
+While \texttt{FEPointEvaluation} assumes that evaluation points are already sorted according to 
+the owning cells and such can concentrate on cell-local operations, the new class
+\texttt{RemovePointEvaluation} is responsible to determine the owning cells in a distributed
+context and for providing efficient communication patterns for the data exchange. In \texttt{deal.II},
+the class has been successfully applied together with \texttt{FEPointEvaluation} to evaluate a distributed
+solution vector at arbitrary points (see \texttt{VectorTools::evaluate\_at\_points()}).
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Many of the \dealii{} tutorial programs were revised in a variety of
-ways as part of this release. A particular example is that we have
-converted a number of programs to use range-based for loops (a C++11
-feature) for loops over a range of integer indices such as loops over
-all quadrature points or all indices of degrees of freedom during
-assembly. This makes sense given that the
-range-based way of writing loops seems to be the idiomatic approach
-these days, and that we had previously already converted loops over
-all cells in this way.
+ways as part of this release. TODO
 
@@
 
-There are also new programs in the code gallery (a collection of
+There is also a new program in the code gallery (a collection of
 user-contributed programs that often solve more complicated problems
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Incompatible changes}
+\subsection{Incompatible changes}\label{subsec:deprecated}
 

### 2021-05-28 9.3/paper.tex
Go through the simplex section.

@@
      Wolfgang Bangerth,
-     %Bruno Blais,
-     %Thomas C. Clevenger,
     Marc Fehling,
-    %Alexander V. Grayver,
     Timo Heister,
@@
     Jean-Paul Pelteret,
-    %Reza Rastak,
-    %Ignacio Thomas,
     Bruno Turcksin,
-    %Zhuoran Wang,
     David Wells
@@
 
- \author[*1]{Daniel Arndt}
+ \author[1*]{Daniel Arndt}
  \affil[1]{Computational Engineering and Energy Sciences Group,
@@
 
- \author[2]{Wolfgang~Bangerth}
- \affil[2]{Department of Mathematics and Department of Geosciences, Colorado State University, Fort
+ \author[2,3]{Wolfgang~Bangerth}
+ \affil[2]{Department of Mathematics, Colorado State University, Fort
    Collins, CO 80523-1874, USA.
    \texttt{bangerth/marc.fehling@colostate.edu}}
-
-%\author[3]{Bruno Blais}
-%\affil[3]{Research Unit for Industrial Flows Processes (URPEI), Department of Chemical Engineering,
-%          Polytechnique Montréal,
-%          PO Box 6079, Stn Centre-Ville, Montréal, Québec, Canada, H3C 3A7.
-%          {\texttt{bruno.blais@polymtl.ca}}}
-
-% \author[4]{Thomas~C.~Clevenger}
-% \affil[4]{School of Mathematical and Statistical Sciences,
-%   Clemson University,
-%   Clemson, SC, 29634, USA
-%   {\texttt{tcleven/heister@clemson.edu}}}
-%
-% \author[4]{Denis~Davydov}
-% \affil[4]{Chair of Applied Mechanics,
-%   Friedrich-Alexander-Universit\"{a}t Erlangen-N\"{u}rnberg,
-%   Egerlandstr.\ 5,
-%   91058 Erlangen, Germany.
-%   {\texttt{\{denis.davydov,jean-paul.pelteret\}@fau.de}}}
-%
+ \affil[3]{Department of Geosciences, Colorado State University, Fort
+   Collins, CO 80523, USA.}
+
 \author[2]{Marc~Fehling}
-
-%
-% \author[6]{Daniel Garcia-Sanchez}
-% \affil[6]{Sorbonne Universit\'es, UPMC Univ.\ Paris 06, CNRS-UMR 7588,
-%   Institut des NanoSciences de Paris, F-75005, Paris, France
-%   {\texttt{daniel.garcia-sanchez@insp.upmc.fr}}}
-%
-% \author[2]{Graham Harper}
-%
-
-%\author[6]{Alexander~V.~Grayver}
-%\affil[6]{Institute of Geophysics,
-%  ETH Zurich,
-%  Sonneggstrasse 5, 8092 Z\"{u}rich, Switzerland.
-%  {\texttt{agrayver@ethz.ch}}}
 
@@
 
-\author[*1]{Bruno~Turcksin}
+\author[1*]{Bruno~Turcksin}
 %
@@
 
-\footnotetext{%
-  $^{**}$ Sandia National Laboratories is a multimission laboratory managed and
-  operated by National Technology \& Engineering Solutions of Sandia, LLC, a
-  wholly owned subsidiary of Honeywell International Inc., for the U.S.
-  Department of Energy's National Nuclear Security Administration under contract
-  DE-NA0003525. This document describes objective technical results and analysis.
-  Any subjective views or opinions that might be expressed in the paper do not
-  necessarily represent the views of the U.S. Department of Energy or the United
-  States Government.}
-
 \begin{abstract}
@@
 \dealii{} version 9.3.0 was released May ??, 2021.
+\todo{Update date}
 This paper provides an
@@
 \begin{itemize}
-  \item An experimental support for simplex and mixed meshes (see Section~\ref{subsec:simplex});
+  \item Experimental support for simplex and mixed meshes (see Section~\ref{subsec:simplex});
   \item Improved flexibility of the particle infrastructure (see Section~\ref{subsec:particles});
@@
 \begin{itemize}
-  \item Each non-artificial cell posses now an globally unique index, which can be queried
+  \item Each non-artificial cell now has a globally unique index that can be queried
   for active cells via \texttt{CellAccessor::global\_active\_cell\_index()} and for level cells 
   via \texttt{::global\_\allowbreak level\_\allowbreak cell\_\allowbreak index()}. The information
-  can be used to efficiently access global cell vectors. Users have been able to construct the 
-  same information by distributing \texttt{FE\_DGQ(0)} in a \texttt{DoFHandler}.
-  \item In addition to the macro \texttt{DEAL\_II\_DEPRECATED}, we have introduced
-  the \texttt{DEAL\_II\_DEPRECATED\_\allowbreak EARLY} macro, which indicates that a feature will be
+  can be used to efficiently index into global vectors storing data
+  for each cell, rather than for each degree of freedom corresponding to a finite
+  element field.
+
+  \item Previously, functions and classes were marked using the 
+  macro \texttt{DEAL\_II\_DEPRECATED} and then typically removed in
+  the release after the one in which these deprecation notices were
+  available to users. We have now extended this policy by introducing
+  the \texttt{DEAL\_II\_\allowbreak DEPRECATED\_EARLY} macro, which indicates that a feature will be
   deprecated in the next release. In contrast to the first macro, it will only give
@@
 %
-The changelog lists more than ?? other
-features and bugfixes.
+The changelog lists more than 200 other features and bugfixes.
 
@@
 
-\begin{figure}[!t]
+The current release of \dealii adds experimental support for simplex
+meshes (consisting of triangles in 2D; tetrahedra in 3D) and mixed meshes (consisting of triangles and/or
+quadrilaterals in 2D; tetrahedra, pyramids, wedges, and/or hexahedra in 3D).
+Many freely available mesh-generation tools produce such kind of
+meshes and they are widely used in industry and applications, but were
+previously unsupported by \dealii. As a consequence, users of \dealii
+had to pre-process such meshes and convert them to pure quadrilateral
+or hexahedral meshes.
+
+Support for simplex and mixed meshes is not universal in \dealii{} at
+this point. While \dealii{} can read such meshes, write output for
+them, and solve partial differential equations with certain finite
+elements, there are also many areas that have not been fully converted
+to the new functionality. In particular, \dealii{} currently only
+offers low-order finite elements on such meshes, and many tool
+functions might throw exceptions when used with such meshes.
+
+A user-focused summary of information around simplex and mixed mesh
+support is also available on the new module page at
+  \url{https://www.dealii.org/current/doxygen/deal.II/group__simplex.html}
+In particular, it shows how to solve a simple Poisson problem like in
+the \texttt{step-3} tutorial program
+on simplex and mixed meshes, with a focus on the necessary changes to
+the workflow.
+At the time of this release, there are also 92 tests (in the folder \texttt{tests/simplex})
+targeting the new simplex and mixed mesh support. In particular, the folder also
+contains ported variants of a number of existing tutorials: 1, 2, 3, 4, 6, 7, 8, 12, 17, 18, 20, 23, 31, 38,
+40, 55, 67, 68, and 74.
+
+\begin{figure}
 
@@
     \foreach \i in {0,1, 2}
-    \fill (\i) circle (1.1pt) node [below] {};
+    \draw[blue,fill=blue] (\i) circle (1.1pt) node [below] {};
 
@@
     \foreach \i in {0,1, 2, 3}
-    \fill (\i) circle (1.1pt) node [below] {};
+    \draw[blue,fill=blue] (\i) circle (1.1pt) node [below] {};
 
@@
     \foreach \i in {0,1,2,3}
-    \fill (\i) circle (1.5pt) node [below] {};
+    \draw[blue,fill=blue] (\i) circle (1.5pt) node [below] {};
 
@@
     \foreach \i in {0,1,2,3,4}
-    \fill (\i) circle (1.3pt) node [below] {};
+    \draw[blue,fill=blue] (\i) circle (1.3pt) node [below] {};
 
@@
     \foreach \i in {0,1,2,3,4,5}
-    \fill (\i) circle (1.5pt) node [below] {};
+    \draw[blue,fill=blue] (\i) circle (1.5pt) node [below] {};
 
@@
     \foreach \i in {0,1,2,3,4,5,6,7}
-    \fill (\i) circle (1.3pt) node [below] {};
+    \draw[blue,fill=blue] (\i) circle (1.3pt) node [below] {};
 
@@
 
-\caption{All reference cells in 2D (triangle, quadrilateral) and 3D (tetrahedron, pyramid, wedge, hexahedron) with the support points indicated both for linear ($\bullet$) and quadratic ($\times$) shape functions.}
-
+  \caption{\it Reference cells in 2D (triangle, quadrilateral) and 3D
+    (tetrahedron, pyramid, wedge, hexahedron) with the support points
+    indicated both for linear ({\color{blue}$\bullet$}) and quadratic
+    ({\color{red}$\times$}) shape functions.
+  \todo[inline]{The pyramid is missing red crosses. The wedge is
+    missing them on three edges.}}
+  \label{fig:simplex}
 \end{figure}
 
-\begin{table}[!h]
+\begin{table}
   \caption{\it List of new scalar \texttt{FiniteElement} classes for the new
@@
 
-The current release of \dealii adds an experimental support for simplex meshes (consisting of triangles in 2D; tetrahedra in 3D) and mixed meshes (consisting of triangles and/or
-quadrilaterals in 2D; tetrahedra, pyramids, wedges, and/or hexahedra in 3D).
-Many freely available mesh-generation tools produce such kind of meshes; while users of \dealii used to have to pre-process such meshes and convert them to pure hex meshes, they
-can now directly work with them.
-
 \subsubsection{Refactoring of internal data structures}
@@
 To enable simplex and mixed mesh support, we performed a major refactoring of the internal data structures of \dealii. In particular, the \texttt{Triangulation} class and
-the \texttt{DoFHandler} class have undergone large changes.
+the \texttt{DoFHandler} class have undergone large changes and now
+support meshes composed of all of the cells shown in Fig.~\ref{fig:simplex}.
 
@@
 cells and vertices to the internal data structures, has been rewritten
-inspired by \citep{logg2012} and a speed-up of up to 5 has been reached. Minor adjustments
-have been made to \texttt{parallel::shared::Triangulation} and \texttt{parallel::\allowbreak fullydistributed::\allowbreak Triangulation} such that
-the new mesh types can be processed also in parallel.
+inspired by \citep{logg2012} and this has had the side effect of a speed-up of up to 5. Minor adjustments
+have also been made to the \texttt{parallel::shared::Triangulation}
+and \texttt{parallel::\allowbreak fullydistributed::\allowbreak
+  Triangulation} classes such that
+the new mesh types can also be used in parallel.
 
@@
 \texttt{hp::DoFHandler} into the \texttt{DoFHandler}. The class \texttt{hp::DoFHandler}, which is now a dummy derivation of \texttt{DoFHandler}, currently only exists for compatibility reasons and has
-been deprecated, see Section~\ref{subsec:deprecated}.
+been deprecated, see Section~\ref{subsec:deprecated}. It will be
+removed in the next release.
 
@@
 generated by an external mesh generator. Currently, we support the
-file formats \texttt{VTK}, \texttt{MSH}, and \texttt{EXODUS II}.
-
-Alternatively, one can create a pure hypercube mesh with the known functions
-in the \texttt{GridGenerator} namespace and convert the obtained mesh to a
+\texttt{VTK}, \texttt{MSH} (generated by the \textsc{gmsh} program \cite{geuzaine2009gmsh}), and \texttt{EXODUS II} file formats.
+
+Alternatively, one can create a pure hypercube mesh with the existing functions
+in the \texttt{GridGenerator} namespace and convert it to a
 pure simplex mesh with the function
-\texttt{convert\_\allowbreak hypercube\_\allowbreak to\_\allowbreak simplex\_\allowbreak mesh()} from the \texttt{GridGenerator} namespace.
-
-\subsubsection{Simplex mesh}
-
-After having created a triangulation, one can proceed as usual in the case of
-a pure simplex mesh by selecting an appropriate finite element, mapping, and
-quadrature class:
+\texttt{GridGenerator::convert\_\allowbreak hypercube\_\allowbreak to\_\allowbreak simplex\_\allowbreak mesh()}.
+
+\subsubsection{Using simplex meshes}
+
+After having created a triangulation, one can proceed in the same way
+as for hypercube meshes. In particular, one selects an appropriate finite element, mapping, and
+quadrature class as follows:
 
@@
 The  list of currently supported finite-element classes is provided in Table~\ref{tab:simplex:fe}. Currently,
-only linear iso-parametric mapping (\texttt{MappingFE} and \texttt{MappingFEFields}) is available.
+only linear iso-parametric mappings (via the \texttt{MappingFE} and
+\texttt{MappingFEField} classes) are available.
+\todo[inline]{I don't understand the previous sentence. "linear" means
+  p=1, but isoparametric means that it has the same p as the finite
+  element. It can't be both.}
 For quadrature, the classes \texttt{QDuffy}, \texttt{QGaussSimplex}, \texttt{QWitherdenVincentSimplex},
@@
 
-\subsubsection{Mixed mesh}
+\subsubsection{Using mixed meshes}
 For mixed meshes, concepts known from the $hp$-context have been applied:
- different finite-element classes are assigned to different cell types. In the
-case of a 2D mixed mesh, which can only consists of triangles and
+ different finite-element classes are assigned to different cells
+ based on their respective kind of reference cell. In the
+case of a 2D mixed mesh, which can only consist of triangles and
 quadrilaterals, the finite element defined on a triangle (e.g., \texttt{FE\_SimplexP})
@@
 Similarly, \texttt{hp::QCollection} and \texttt{hp::MappingCollection} can be used
-to construct appropriate collections. Furthermore, the right active finite-element index, which points to the right
-finite element of that cell, has to be assigned to each cell:
+to construct appropriate collections. Furthermore, the correct active
+finite element index, which points to the correct
+finite element of that cell, has to be assigned to each cell. The
+following piece of code will then correctly enumerate all degrees of
+freedom on the mesh:
 
@@
 The introduction of simplex and mixed meshes leads to some implications
-for the user if these features should be used. For instance, each cell might have a different type with
+for the user if these features are to be used. For instance, each cell might have a different type with
 different number of vertices, lines, and faces so that these quantities can not be
@@
   for(unsigned int f = 0; f < cell->n_faces(); ++f)
-    (void) cell->face(f);
+    // do something with cell->face(f);
 \end{c++}
 Alternatively, one can use an iterator-based approach to loop over all faces
-of a cell, introduced in the last release. The relevant functions have been adjusted to be able to deal with
-various number of faces.
+of a cell, introduced in the previous release. The relevant functions have been adjusted to be able to deal with
+the now variable number of faces per cell:
 \begin{c++}
@@
   for(const auto & face : cell->face_iterators())
-    (void) face;
+    // do something with face
 \end{c++}
-Furthermore, the number of degrees of freedom might differ between cells so that
-local data storage units needed for assembly might need to be resized for each
-cell as it is already good practice in the $hp$-context:
+Furthermore, for mixed meshes, the number of degrees of freedom will differ between cells so that
+cell-local arrays need to be resized for each
+cell (as has previously already been the case in the $hp$-context):
 \begin{c++}
@@
    hp_fe_values.reinit(cell);
-   local_rhs.resize(hp_fe_values.get_present_fe_values().n_dofs_per_cell());
+   local_rhs.resize(cell->get_fe().dofs_per_cell);
  }
@@
 What is true for cells is also true for faces in 3D: faces can be either triangles
-or quadrilaterals.
-This is the case even if no mixed mesh is used.
+or quadrilaterals. This is the case even if no mixed mesh is used if
+the mesh consists exclusively of pyramids or wedges.
 As a consequence, some functions, e.g.
@@
 
-We tried---where possible---to provide utility functions so that users do not need to access a \texttt{GeometryInfo}-like data structure.
-However, these functions internally rely on the \texttt{ReferenceCell} class, which is similarly structured as \texttt{GeometryInfo} and even use it for hypercube cells.
-Users can query the right \texttt{ReferenceCell} of a geometric entity via:
-\begin{c++}
-const auto reference_cell = cell->reference_cell();
-\end{c++}
-and of its i-th face via:
-\begin{c++}
-const auto face_reference_cell = reference_cell.face_reference_cell(i);
-\end{c++}
-
-Furthermore, it is crucial to use the right FE--mapping--quadrature family
-for the right cell type. In particular, this means that users cannot
-rely on default parameters regarding mapping and/or quadrature in the
-context of simplex meshes since these are defined for hypercube cells.
-We recommend to choose the needed mapping,
-FE, and quadrature rules at a single central place and pass these throughout the program
-in the form of ``scratch data''.
+Geometric information about cells and faces -- previously provided by
+the \texttt{GeometryInfo} class for hypercube-type cells -- is now
+available via the  \texttt{ReferenceCell} class that can be
+instantiated for each of the seven possible reference cell. The
+correct reference cell object for a cell or face can be obtained, respectively, using
+\texttt{cell->reference\_cell()}, and either the call
+\texttt{cell->reference\_cell().face\_reference\_cell(f)}
+or
+\texttt{cell->face(f)->reference\_cell()}.
+
+Furthermore, many functions in \dealii{} used mappings that, when not
+given explicitly, defaulted to (bi-/tri-)linear ones. These no longer
+work for simplex or mixed meshes, and in these cases users will need
+to explicitly provide the correct mapping for the mesh to be used.
 
@@
 
-We provide matrix-free support for simplex and mixed meshes as well as for
+\dealii{}'s matrix-free support has also been extended to simplex and
+mixed meshes, for both
 continuous and discontinuous elements. From a user perspective, the main
-changes are to pass a d-dimensional quadrature rules rather than a 1D formula,
+changes are to pass a $d$-dimensional quadrature object (rather than
+one for 1D),
 and the fact that \texttt{FEEvaluation} and \texttt{FEFaceEvaluation} must not
@@
 information at runtime. More details can be found in
-Subsection~\ref{subsec:mf}.
-
-By the time of writing, no advanced algorithms for evaluating values and
-gradients at the quadrature points such as sum factorization are used.
+Section~\ref{subsec:mf}.
+
+For the current release, no advanced algorithms for evaluating values and
+gradients at quadrature points, such as sum factorization, are used.
 The use of full interpolation matrices is for now acceptable since only low-order elements are
 supported.
-
-\subsubsection{Miscellanea}
-
-Further information can be found on the new module page ``Simplex support (experimental)'', see
-\begin{center}
-  \url{https://www.dealii.org/current/doxygen/deal.II/group__simplex.html}
-\end{center}
-As an example, it shows how to solve a simple Poisson problem like in \texttt{step-3}
-on simplex and mixed meshes, with the focus on on the changed workflow.
-
-By the time of writing, there have been 92 tests (in the folder \texttt{tests/simplex})
-targeting the new simplex and mixed mesh support. In particular, the folder also
-contains ported variants of the following tutorials: 1, 2, 3, 4, 6, 7, 8, 12, 17, 18, 20, 23, 31, 38,
-40, 55, 67, 68, and 74. These might be also good starting points.
 

### 2021-05-28 9.3/paper.tex
Go through the rest of the paper.

@@
 
-TODO
+\todo[inline]{TODO}
 
@@
 
-Until now, \dealii has only supported local smoothing multigrid
-algorithms~\citep{ClevengerHeisterKanschatKronbichler2019}; in local
-smoothing algorithms, smoothers only act on the cells of a given refinement level,
-skipping those parts of the mesh that are not adaptively refined to that level.
-The current release now also has support for global coarsening~\citep{becker2000multigrid, sundar2012parallel} when
+Until now, \dealii has only supported ``local smoothing'' multigrid
+algorithms~\citep{ClevengerHeisterKanschatKronbichler2019} wherein
+smoothers only act on the cells of a given refinement level,
+skipping those parts of the mesh that have not been adaptively refined
+to that level. This approach guarantees that the work done summed up
+over all levels is proportional to the number of unknowns, and is
+consequently necessary so that the overall multigrid preconditioner
+can have a complexity of ${\cal O}(N)$.
+
+In contrast, the current release now also supports ``global
+coarsening'' algorithms \citep{becker2000multigrid, sundar2012parallel} when
 using continuous (\texttt{FE\_Q}, \texttt{FE\_SimplexP}) and
@@
 The transfer operators between two levels have been implemented in the new class \texttt{MGTwoLevel\allowbreak Transfer}, which can be set up via the functions \texttt{MGTwoLevel\allowbreak Transfer::\allowbreak reinit\_\allowbreak geometric\_\allowbreak transfer()} or \texttt{MGTwo\allowbreak LevelTransfer::\allowbreak reinit\_\allowbreak polynomial\_\allowbreak transfer()} for given
-\texttt{DoFHandler} and \texttt{AffineConstraint} classes of two levels. The resulting transfer operators
+\texttt{DoFHandler} and \texttt{AffineConstraint} objects
+corresponding to the two levels. The resulting transfer operators
 can then be collected in a single
@@
 
+\dealii{}'s matrix-free framework enables high-throughput operations
+for applications in which only the availability of the \textit{action}
+of a matrix, but not the entries of the matrix, are necessary. This
+framework has been substantially extended in the current release.
+
 \subsubsection{Precompilation of evaluation kernels}
 
-The classes \texttt{FEEvaluation} and \texttt{FEFaceEvaluation} use template parameters for
+The \texttt{FEEvaluation} and \texttt{FEFaceEvaluation} classes use template parameters for
 the polynomial degree of the finite element $k$ and the number of the 1D quadrature points $q$
@@
 not rely on the template parameters $k$ and $q$ (expressed in the code
-with ``-1'' and ``0'') have been added. For example:
+using special values ``-1'' and ``0'') have been added. For example:
 \begin{c++}
@@
 
-With release 9.1 \cite{dealII91}, large parts of the $hp$-algorithms in \dealii were parallelized, enabling
-parallel matrix-based simulations with the $hp$-infrastructure. In the present
+With release 9.1, large parts of the $hp$-algorithms
+in \dealii were ported to a model that allows for
+parallel matrix-based simulations \cite{dealII91}. In the present
 release, the parallel $hp$ support was extended to \texttt{MatrixFree}.
 
-Until now, the \texttt{FEEvaluation} classes used the template parameters $k$ and $q$ to select the correct active FE and quadrature index and cumbersome detection of subranges of the same degree.
+Until now, the \texttt{FEEvaluation} classes used the template
+parameters $k$ and $q$ to select the correct active FE and quadrature
+index and cumbersome detection of subranges of the same degree.
+\todo{I don't understand what ``detection of subranges of the same
+  degree'' refers to. Please clarify. In particular, I don't know what
+``subranges'' mean in this paragraph.}
 The creation of subranges is now performed internally, and the non-templated versions
@@
 \subsection{MPI-3.0 shared-memory support}\label{subsec:sm}
-In many large computations, certain pieces of data are computed once
-and then treated as read-only. If this information is needed by more
+In many large computations, certain pieces of data are computed
+(or read from disk) once and then treated as read-only. If this information is needed by more
 than one MPI process, it is more efficient to store this information only
@@
 \texttt{AlignedVector} and \texttt{Table} classes that are often used
-for large lookup tables.
-
-The feature is also used in the \texttt{MatrixFree} and
+for large lookup tables for classes such as \texttt{InterpolatedTensorProductGridData}.
+
+Shared memory storage is also used in the \texttt{MatrixFree} and
 \texttt{LinearAlgebra::\allowbreak distributed::\allowbreak Vector} classes. If \texttt{MatrixFree} has been configured by setting
-\texttt{MatrixFree::AdditionalData::communicator\_sm} appropriately, then
+\texttt{MatrixFree::\allowbreak AdditionalData::\allowbreak communicator\_sm} appropriately, then
 \texttt{MatrixFree::create\_dof\_vector()} creates vectors that share
@@
 
-In a number of circumstances, finite element solutions need to be evaluated on
+In a number of circumstances, finite element solutions need to be evaluated at
 arbitrary reference points that change from one element to next. Two important
@@
 algorithms on non-matching grids. The existing \texttt{FEValues} class is a poor
-fit for this task, as it relies on tabulating all information before its usage,
-necessitating a separate setup for every cell. The new class
-\texttt{FEPointEvaluation} provides a more convenient interface for this
-task. For tensor product finite elements (\texttt{FE\_Q}, \texttt{FE\_DGQ})
+fit for this task, as it is based on the assumption that evaluation of
+shape functions and their derivatives happens at the same quadrature
+points on every cell, and that consequently expensive computations can
+be done once and results re-used for many subsequent cells. The new class
+\texttt{FEPointEvaluation} provides a more convenient interface for
+cases where the evaluation on different cells does not happen at the
+same points mapped from the reference cell.
+For tensor product finite elements (\texttt{FE\_Q}, \texttt{FE\_DGQ})
 and tensor product mappings (\texttt{MappingQGeneric} and derived classes),
@@
 contribution is added to a fluid solver by multiplication with test function
-and addition over quadrature points:
+and addition over quadrature points located at the interface section
+$\Gamma_K=\Gamma \cup K$ located on the current cell $K$ and that is,
+in general, located differently within $K$ than for any other cell:
 \begin{align*}
-\left(\vec{v}, \kappa \, \vec{n}\right)_\Gamma
+\left(\mathbf{v}, \kappa \, \mathbf{n}\right)_\Gamma
 \approx
-\sum_q\left(\vec{v}, \kappa(\vec{p}_q) \, \vec{n}(\vec{p}_q) (JxW)_q\right).
+\sum_q \mathbf{v}(\mathbf{x}_q) \cdot \left(\kappa(\mathbf{x}_q) \, \mathbf{n}(\mathbf{x}_q)\right) (JxW)_q.
 \end{align*}
@@
 \end{c++}
-The quadrature points (at reference positions - \texttt{reference\_points}) and the related \texttt{JxW} value can, e.g., come from
-a mesh of codimension one. Determining to which \texttt{cell} a quadrature
-point belongs to on the background mesh, including the reference-cell
-coordinates \texttt{reference\_points}, can be determined with functions like
+The quadrature points (at reference positions
+\texttt{reference\_points}) and the related \texttt{JxW} value can,
+for example, come from
+a mesh of lower dimension. Determining to which \texttt{cell} a quadrature
+point belongs to on the background mesh, including the reference cell
+coordinates \texttt{reference\_points}, is aided by functions like
 \texttt{GridTools::find\_all\_active\_cells\_around\_point()}. While these
@@
 While \texttt{FEPointEvaluation} assumes that evaluation points are already sorted according to 
-the owning cells and such can concentrate on cell-local operations, the new class
-\texttt{RemovePointEvaluation} is responsible to determine the owning cells in a distributed
+the owning cells and thus can focus on cell-local operations, the new class
+\texttt{RemotePointEvaluation} is responsible for determine the owning cells in a distributed
 context and for providing efficient communication patterns for the data exchange. In \texttt{deal.II},
 the class has been successfully applied together with \texttt{FEPointEvaluation} to evaluate a distributed
-solution vector at arbitrary points (see \texttt{VectorTools::evaluate\_at\_points()}).
+solution vector at arbitrary points (see also the
+\texttt{VectorTools::evaluate\_at\_points()} function).
 
@@
 Many of the \dealii{} tutorial programs were revised in a variety of
-ways as part of this release. TODO
-
-In addition, there are a number of new tutorial programs:
+ways as part of this release. In addition, there are a number of new tutorial programs:
 \begin{itemize}
@@
 
-\item \texttt{step-68} TODO
+\item \texttt{step-68} \todo[inline]{TODO}
 
@@
       The tutorial presents advanced topics, like the usage of cell-centric loops and
-      the new MPI-3.0 shared-memory capabilities of \texttt{MatrixFree} tor each high
+      the new MPI-3.0 shared-memory capabilities of \texttt{MatrixFree} to reach high
       throughput. Furthermore, the utilization of the template parameter
-      \texttt{VectorizedArrayType} and the application of lambdas to capture cell and face
+      \texttt{VectorizedArrayType} and the application of lambda functions to describe cell and face
       integrals are discussed.
 
-\item \texttt{step-77} is a program that illustrate \texttt{dealii}'s
+\item \texttt{step-77} is a program that illustrate \dealii{}'s
   interfaces to the SUNDIALS library \cite{sundials}, and specifically
@@
 
-  \dealii{}'s interfaces were also updated to the latest SUNDIALS release, 5.7.
-
-\item \texttt{step-78} TODO
-
-\item \texttt{step-79} TODO
+  \dealii{}'s interfaces to the various SUNDIALS sub-packages were also updated to the latest SUNDIALS release, 5.7.
+
+\item \texttt{step-78} \todo[inline]{TODO}
+
+\item \texttt{step-79} \todo[inline]{TODO}
 \end{itemize}
@@
 \begin{itemize}
-  \item preCICE: TODO
+  \item preCICE: \todo[inline]{TODO}
 \end{itemize}
@@
         namespace \texttt{Legacy} has been introduced that provides all affected
-        classes with the old interface for the transition phase.
+        classes with the old interface for a transition period.
 \end{itemize}
+
+\todo[inline]{Check whether there are other things worth listing.}
 
@@
 software, we ask that papers using the library reference one of the
-\dealii{} papers. This helps us justify the effort we put into it.
+\dealii{} papers. This helps us justify the effort we put into this library.
 
@@
 The original \dealii{} paper containing an overview of its
-architecture is \cite{BangerthHartmannKanschat2007}. If you rely on
+architecture is \cite{BangerthHartmannKanschat2007}, and a more recent
+publication documenting \dealii{}'s design decisions is available as \cite{dealII2020design}. If you rely on
 specific features of the library, please consider citing any of the
@@
     \item For $hp$-adaptivity: \cite{BangerthKayserHerold2007};
-    \item For partition-of-unity (PUM) and enrichment methods of the
-          finite element space: \cite{Davydov2016};
+    \item For partition-of-unity (PUM) and finite element enrichment methods:
+           \cite{Davydov2016};
     \item For matrix-free and fast assembly techniques:
@@
           \cite{HeltaiBangerthKronbichlerMola2019};
-          \vfill\null  \columnbreak
     \item For integration with CAD files and tools:
@@
           \cite{GiulianiMolaHeltai-2018-a};
-    \item For \texttt{LinearOperator} and \texttt{PackagedOperation} facilities:
+    \item For the \texttt{LinearOperator} and
+      \texttt{Packaged\-Operation} facilities:
           \cite{MaierBardelloniHeltai-2016-a,MaierBardelloniHeltai-2016-b};
@@
 % 9.3: updated 5/08/2021 PM
+%      sorted by WB 5/28
+Pasquale Africa,
+Tyler Anderson,
+Mathias Anselmann,
+Arpit Babbar,
+Maximilian Bergbauer,
+Nicolas Barnafi,
+Bruno Blais,
+Michele Bucelli,
+Marcus Calhoun-Lopez,
+Praveen Chandrashekar,
+Conrad Clevenger,
+David F. Castellanos,
+Andrew Davis,
+Elias Dejene,
+Niklas Fehn,
+Menno Fraters,
+Ivan Fumagalli,
+Daniel Garcia-Sanchez,
+Rene Gassmoeller,
+Nicola Giuliani,
+Krishnakumar Gopalakrishnan,
 Alexander Grayver,
-Andrew Davis,
-Arpit Babbar,
-Bruno Blais,
-Bruno Turcksin,
+Olivier Guevremont,
+Jake Harmon,
+Graham Harper,
+Uwe K\"ocher,
+Katharina Kormann,
+Christoph Kammer,
+Wenyu Lei,
+Zhou Lei,
+Phillip Mobley,
+Nils Much,
+Pratik Nayak,
+Toni El Geitani Nehme,
+Justin O'Connor,
+Daniel Paukner,
+Sebastian Proell,
+Lei Qiao,
 Ce Qin,
-Christoph Kammer,
-Conrad Clevenger,
-Dominic Soldner,
-Daniel Arndt,
-Daniel Garcia-Sanchez,
-Daniel Paukner,
-David F. Castellanos,
-David Wells,
+Reza Rastak,
+Julian Roth,
+Laura Prieto Saavedra,
 Doug Shi-Dong,
 David Schneider,
-Elias Dejene,
-Graham Harper,
+Magdalena Schreter,
+Richard Schussnig,
+Konrad Simon,
+Dominic Soldner,
+Simon Sticko,
+Malhar Tidke,
 Ignacio Tomas,
-Ivan Fumagalli,
-Jake Harmon,
-Jean-Paul Pelteret,
-Jiaqi Zhang,
-Julian Roth,
-Katharina Kormann,
-Konrad Simon,
-Krishnakumar Gopalakrishnan,
-Laura Prieto Saavedra,
-Lei Qiao,
-Luca Heltai,
-Magdalena Schreter,
-Malhar Tidke,
-Marc Fehling,
-Marcus Calhoun-Lopez,
-Martin Kronbichler ,
-Mathias Anselmann,
-Matthias Maier,
-Maximilian Bergbauer,
-Menno Fraters,
-Michele Bucelli,
-Nicola Giuliani,
-Nicolas Barnafi,
-Niklas Fehn,
-Nils Much,
-Olivier Guevremont,
-Pasquale Africa,
-Peter Munch,
 Peter Westerbaan,
-Phillip Mobley,
-Pratik Nayak,
-Praveen Chandrashekar,
-Rene Gassmoeller,
-Reza Rastak,
-Richard Schussnig,
-Sebastian Proell,
-Simon Sticko,
-Timo Heister,
-Toni El Geitani Nehme,
-Wenyu Lei,
-Wolfgang Bangerth,
-Zhou Lei
+Jiaqi Zhang.
+
+\todo[inline]{Remove everyone who may end up being a co-author.}
 
@@
 
-W.~Bangerth, T.~C.~Clevenger, and T.~Heister were partially
+W.~Bangerth and T.~Heister were partially
 supported by the Computational Infrastructure
@@
 and EAR-1925595.
-
-%D.~Davydov was supported by the German Research Foundation (DFG), grant DA
-%1664/2-1 and the Bayerisches Kompetenznetzwerk
-%f\"ur Technisch-Wissenschaftliches Hoch- und H\"ochstleistungsrechnen
-%(KONWIHR).
-
-B.~Blais was partially supported by the National Science and Engineering Research Council of Canada(NSERC)  through the RGPIN-2020-04510 Discovery Grant
-
-T.~C.~Clevenger was also partially supported EAR-1925575 and OAC-2015848.
-
-A.~V.~Grayver was partially supported by the European Space Agency
-Swarm DISC program.
 
@@
 
-Z.~Wang was partially
-supported by the National Science Foundation under award OAC-1835673.
-
 The Interdisciplinary Center for Scientific Computing (IWR) at Heidelberg
@@
 
-The authors acknowledge the Texas Advanced Computing Center (TACC) at The
-University of Texas at Austin for providing access to HPC resources that have
-contributed to the research results reported within this paper.
-
-Clemson University is acknowledged for generous allotment of compute time on
-the Palmetto cluster.
+\todo[inline]{Everyone update. Add support for any additional authors.}
 

### 2021-05-28 9.3/paper.tex
Some more small updates.
I checked.
Add an entry for step-74.

@@
 
-\item \texttt{step-68} \todo[inline]{TODO}
+\item \texttt{step-66} nonlinear matrix-free solver \todo[inline]{TODO}
+
+\item \texttt{step-68} parallel particles, load balancing \todo[inline]{TODO}
 
@@
   that then also yields the residual itself.
+
+\item \texttt{step-74} SIPG for the Poisson equation \todo[inline]{TODO}
 
@@
 \end{itemize}
-
-\todo[inline]{Check whether there are other things worth listing.}
 
@@
 \dealii{} can interface with many other libraries:
+ \todo{Bruno: Please add some kind of reference for ArboxX}
 \begin{multicols}{3}
@@
     \item ADOL-C \cite{Griewank1996a,adol-c}
+    \item ArborX
     \item ARPACK \cite{arpack}
@@
     \item muparser \cite{muparser-web-page}
-    \item nanoflann \cite{nanoflann}%
-          %
-          % nanoflann has been deprecated and will be removed for 9.3. remove
-          % this line in the 9.3 release paper
-          %
-    \item NetCDF \cite{rew1990netcdf}
-          %
-          % netcdf has been deprecated and will be removed for 9.3. remove
-          % this line in the 9.3 release paper
-          %
     \item OpenCASCADE \cite{opencascade-web-page}

### 2021-05-31 9.3/paper.tex
Further changes

@@
 support is also available on the new module page at
-  \url{https://www.dealii.org/current/doxygen/deal.II/group__simplex.html}
+  \url{https://www.dealii.org/current/doxygen/deal.II/group__simplex.html}.
 In particular, it shows how to solve a simple Poisson problem like in
@@
 
+    \coordinate (15) at (0,0,0.75);
+    \coordinate (16) at (1,0,0.75);
+    \coordinate (17) at (0.0,1.0,0.75);
+
     \foreach \i in {0,1,2,3,4,5}
@@
 
-    \foreach \i in {6,7,8,9,10,11,12,13,14}
+    \foreach \i in {6,7,8,9,10,11,12,13,14,15,16,17}
     \draw (\i) node[cross, draw=red, rotate=-20] {};
@@
     ({\color{red}$\times$}) shape functions.
-  \todo[inline]{The pyramid is missing red crosses. The wedge is
-    missing them on three edges.}}
+  %\todo[inline]{The pyramid is missing red crosses. The wedge is
+  %  missing them on three edges.}
+  % PM: we only have linear pyramids for now.
+  }
   \label{fig:simplex}
@@
 The  list of currently supported finite-element classes is provided in Table~\ref{tab:simplex:fe}. Currently,
-only linear iso-parametric mappings (via the \texttt{MappingFE} and
-\texttt{MappingFEField} classes) are available.
-\todo[inline]{I don't understand the previous sentence. "linear" means
-  p=1, but isoparametric means that it has the same p as the finite
-  element. It can't be both.}
-For quadrature, the classes \texttt{QDuffy}, \texttt{QGaussSimplex}, \texttt{QWitherdenVincentSimplex},
+only linear and quadratic mappings via the \texttt{MappingFE} and
+\texttt{MappingFEField} classes built around the listed nodal elements are available.
+%\todo[inline]{I don't understand the previous sentence. "linear" means
+%  p=1, but isoparametric means that it has the same p as the finite
+%  element. It can't be both.}
+For quadrature, the classes \texttt{QDuffy}, \texttt{QGaussSimplex}, \texttt{QWitherdenVincentSimplex} \cite{schloemer21, witherden2015identification},
 \texttt{QGaussPyamid}, and \texttt{QGaussWedge} are
@@
    hp_fe_values.reinit(cell);
-   local_rhs.resize(cell->get_fe().dofs_per_cell);
+   local_rhs.resize(cell->get_fe().n_dofs_per_cell());
  }
@@
 how many times they have been refined. In addition, the framework now available in
-\dealii{} in not only applicable to geometric coarsening, but can also
+\dealii{} is not only applicable to geometric coarsening, but can also
 perform coarsening by reducing the polynomial degree ($p$-multigrid,
@@
 k+2, \left\lfloor (3k)/2 \right\rfloor \}$) -- efficient precompiled implementations and default to non-templated
-evaluation kernels else.
+evaluation kernels else (see also \texttt{FEEvaluation::fast\_evaluation\_\allowbreak supported()}).
 
@@
   <dim, VectorizedArrayType::value_type, VectorizedArrayType>;
+// same for FEFaceEvaluationFactory and CellwiseInverseMassFactory (skipped)
 DEAL_II_NAMESPACE_CLOSE
@@
 parameters $k$ and $q$ to select the correct active FE and quadrature
-index and cumbersome detection of subranges of the same degree.
-\todo{I don't understand what ``detection of subranges of the same
-  degree'' refers to. Please clarify. In particular, I don't know what
-``subranges'' mean in this paragraph.}
+index and users were responsible for the cumbersome detection of 
+subranges of the same $k$ and $q$ within the cell ranges returned by the 
+matrix-free loops.
+%\todo{I don't understand what ``detection of subranges of the same
+%  degree'' refers to. Please clarify. In particular, I don't know what
+%``subranges'' mean in this paragraph.}
 The creation of subranges is now performed internally, and the non-templated versions
@@
 point belongs to on the background mesh, including the reference cell
-coordinates \texttt{reference\_points}, is aided by functions like
-\texttt{GridTools::find\_all\_active\_cells\_around\_point()}. While these
+coordinates \texttt{reference\_\allowbreak points}, is aided by functions like
+\texttt{find\_active\_cell\_around\_point()}
+\texttt{find\_all\_active\_cells\_\allowbreak around\_\allowbreak point()} from the \texttt{GridTools} namespace. While these
 functions have been available in \dealii{} previously, their performance has
@@
 solution vector at arbitrary points (see also the
-\texttt{VectorTools::evaluate\_at\_points()} function).
+\texttt{VectorTools::point\_values()} function).
 
@@
   illustrates how these techniques can be used to compute derivatives
-  of first a rather simple function, and then of the much more complex
+  first of a rather simple function, and then of the much more complex
   energy functions of two magnetoelastic and magneto-viscoelastic

### 2021-06-09 9.3/paper.tex
Minor text edit.

@@
 k+2, \left\lfloor (3k)/2 \right\rfloor \}$) -- efficient precompiled implementations and default to non-templated
-evaluation kernels else (see also \texttt{FEEvaluation::fast\_evaluation\_\allowbreak supported()}).
+evaluation kernels otherwise (see also \texttt{FEEvaluation::fast\_evaluation\_\allowbreak supported()}).
 

### 2021-06-17 9.3/paper.tex
Remove old names.
Updates list of coauthors; introduces new Sec. 2.7 for candi

@@
 \newcommand{\cmake}{{\specialword{CMake}}\xspace}
-\newcommand{\autoconf}{{\specialword{autoconf}}\xspace}
+\newcommand{\candi}{{\specialword{candi}}\xspace}
 
@@
 
- \author[6,7]{Martin~Kronbichler}
- \affil[6]{Institute for Computational Mechanics,
+\author[6]{Uwe~K{\"o}cher}
+\affil[6]{Chair of Numerical Mathematics,
+  Helmut-Schmidt-University,
+  University of the Federal Armed Forces Hamburg,
+  Holstenhofweg~85, 22043 Hamburg, Germany.
+  {\texttt{uwe.koecher@hsu-hh.de}}}
+
+ \author[7,8]{Martin~Kronbichler}
+ \affil[7]{Institute for Computational Mechanics,
    Technical University of Munich,
@@
    {\texttt{kronbichler/munch@lnm.mw.tum.de}}}
- \affil[7]{Department of Information Technology,
+ \affil[8]{Department of Information Technology,
    Uppsala University,
@@
 
-\author[6,9]{Peter Munch}
+\author[7,10]{Peter Munch}
  \affil[10]{Institute of Materials Research, Materials Mechanics,
@@
 
-\author[10]{Jean-Paul~Pelteret}
-\affil[10]{Independent researcher.
+\author[11]{Jean-Paul~Pelteret}
+\affil[11]{Independent researcher.
 {\texttt{jppelteret@gmail.com}}}
 
-% \author[12]{Reza Rastak}
-% \affil[12]{Department of Civil and Environmental Engineering,
-%   Stanford University,
-%   Stanford, CA 94305, USA.
-%   {\texttt{rastak@stanford.edu}}}
-
-% \author[**13]{Ignacio Tomas}
-% \affil[13]{Sandia National Laboratories, Org 1442,
-% P.O. Box 5800, MS 1320, Albuquerque, NM, 87185-1320,
-%   \texttt{itomas@sandia.gov}}
-
 \author[1*]{Bruno~Turcksin}
-%
-
-%\author[14]{Zhuoran Wang}
-%\affil[14]{Department of Mathematics, Colorado State University, Fort Collins,
-%          CO 80523-1874, USA.
-%{\texttt{zhrwang@math.colostate.edu}}}
-
-\author[15]{David Wells}
-\affil[15]{Department of Mathematics, University of North Carolina,
+
+\author[16]{David Wells}
+\affil[16]{Department of Mathematics, University of North Carolina,
   Chapel Hill, NC 27516, USA.
@@
 \texttt{VectorTools::point\_values()} function).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Terminal based toolchain installer}
+\label{subsec:candi}
+
+Half a page up to a page on candi.
 

### 2021-06-17 9.3/paper.tex
9.3 release

--- 2021-06-09 9.3/paper.tex
@@
     Bruno Turcksin,
-    David Wells
+    David Wells,
+    Jiaqi Zhang
   },
@@
    Clemson, SC, 29634, USA
-   {\texttt{tcleven/heister@clemson.edu}}}
+   {\texttt{heister/jiaqi2@clemson.edu}}}
 
@@
 
-% \author[12]{Reza Rastak}
-% \affil[12]{Department of Civil and Environmental Engineering,
-%   Stanford University,
-%   Stanford, CA 94305, USA.
-%   {\texttt{rastak@stanford.edu}}}
-
-% \author[**13]{Ignacio Tomas}
-% \affil[13]{Sandia National Laboratories, Org 1442,
-% P.O. Box 5800, MS 1320, Albuquerque, NM, 87185-1320,
-%   \texttt{itomas@sandia.gov}}
-
 \author[1*]{Bruno~Turcksin}
-%
-
-%\author[14]{Zhuoran Wang}
-%\affil[14]{Department of Mathematics, Colorado State University, Fort Collins,
-%          CO 80523-1874, USA.
-%{\texttt{zhrwang@math.colostate.edu}}}
 
@@
   {\texttt{drwells@email.unc.edu}}}
+
+\author[4]{Jiaqi~Zhang}
 
@@
 Ignacio Tomas,
-Peter Westerbaan,
-Jiaqi Zhang.
+Peter Westerbaan.
 

### 2021-06-17 9.3/paper.tex

@@
 
-\author[10]{Jean-Paul~Pelteret}
-\affil[10]{Independent researcher.
+\author[11]{Jean-Paul~Pelteret}
+\affil[11]{Independent researcher.
 {\texttt{jppelteret@gmail.com}}}
@@
 
-\author[15]{David Wells}
-\affil[15]{Department of Mathematics, University of North Carolina,
+\author[16]{David Wells}
+\affil[16]{Department of Mathematics, University of North Carolina,
   Chapel Hill, NC 27516, USA.

### 2021-06-17 9.3/paper.tex
Add name + fix typo

@@
     Jean-Paul Pelteret,
+    Sebastian Proell,
     Bruno Turcksin,
@@
    Boltzmannstr.~15, 85748 Garching, Germany.
-   {\texttt{kronbichler/munch@lnm.mw.tum.de}}}
+   {\texttt{kronbichler/munch/proell@lnm.mw.tum.de}}}
  \affil[8]{Department of Information Technology,
@@
 {\texttt{jppelteret@gmail.com}}}
+
+\author[7]{Sebastian~Proell}
 
@@
   accessible on \texttt{Docker Hub} via \texttt{dealii/dealii:master-focal}. This image is
-  particular useful when used in the contiguous-integration processes of user codes.
+  particular useful when used in the continuous-integration processes of user codes.
 \end{itemize}
@@
 Daniel Paukner,
-Sebastian Proell,
 Lei Qiao,

### 2021-06-17 9.3/paper.tex
updates list of acknowledgments

--- 2021-06-17 9.3/paper.tex
@@
     Bruno Turcksin,
-    David Wells,
-    Jiaqi Zhang
+    David Wells
   },
@@
    Clemson, SC, 29634, USA
-   {\texttt{heister/jiaqi2@clemson.edu}}}
+   {\texttt{tcleven/heister@clemson.edu}}}
 
@@
   {\texttt{drwells@email.unc.edu}}}
-
-\author[4]{Jiaqi~Zhang}
 
@@
 
-Half a page up to a page on candi.
+Half a page up to a page on \candi.
 
@@
 Marcus Calhoun-Lopez,
+Fabian Castelli, %He provided major contributions to candi for v9.3
 Praveen Chandrashekar,
@@
 Graham Harper,
-Uwe K\"ocher,
 Katharina Kormann,
@@
 Ignacio Tomas,
-Peter Westerbaan.
+Peter Westerbaan,
+Jiaqi Zhang.
 

### 2021-06-17 9.3/paper.tex
write text (Sec 2.7 source based toolchain installer)
follow up

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Terminal based toolchain installer}
+\subsection{Source based toolchain installer}
 \label{subsec:candi}
 
-Half a page up to a page on \candi.
+The source based installation of \dealii{} or
+complex toolchains of multiple libraries working together with \dealii{}
+can be done with the \candi{} script tool. \candi is a bash-script based tool,
+it is the abbreviation of ``compile and install'' and it is released under the
+GNU LGPL v3.0.
+%
+The origin of \candi{} is a fork made in 2013 from \texttt{dorsal},
+a retired source based installer for the \texttt{FEniCS} library.
+%
+The \candi{} tool can be found via the download page of \dealii{} since version 8.5
+and is under continuous and active maintenance. To install older releases of
+\dealii{} toolchains one can checkout the branch corresponding to the
+\dealii{}-version with \texttt{git}.
+
+The installer supports the user to download, unpack, compile and install a certain
+library, a list of libraries or a complete toolchain.
+A library is called a package within \candi{}.
+The toolchain installation is the default behaviour and the default configuration
+ensures that most of the \dealii{} step tutorials can be used directly.
+The package installation mode can be handy to generate docker containers
+for instance.
+Especially in the toolchain mode we can check for dependencies from one library to
+another and handle them accordingly.
+
+Each package for a library is defined by variables for its name and version,
+a (remote or local) download location,
+its packaging format (e.g. tar, gz or zip),
+a checksum to prevent from incomplete downloads,
+its general build chain (e.g. using \texttt{autotools}, \texttt{cmake} and others)
+as well as configuration options.
+%
+Moreover, one can give specific instructions for each of the steps being
+necessary for the unpacking, configuration or building of the package.
+
+An internet connection is not necessary for the run, if the source files
+are downloaded and repositories are checked out to its unpacking location.
+The download of a package can be supported via a local or remote mirror server
+similar as we are doing it already for several of the packages.
+%
+If the temporary files of \candi{} are not removed, one can use a developer
+mode to prepare patches or developments for any of the packages.
+%
+A new feature allows \candi{} to skip user prompts which can be used for
+a batch mode run on a compute cluster job for instance.
+
+This source based installer is implemented in a modular fashion. The main
+script is implemented without any knowledge on a package or toolchain for
+a specific operating system (platform).
+
+The installer is organised in the following way:
+\begin{itemize}
+\item \texttt{candi.sh}: the bash-script which controls the overall process.
+
+\item \texttt{candi.cfg} (or \texttt{local.cfg}): the default toolchain
+configuration file. Here one can easily switch on/off important features,
+give additional configuration options to \dealii{} and
+define the set of packages for the toolchain.
+
+\item \texttt{deal.II-toolchain} folder: the central project folder for
+\dealii{}-based toolchains.
+It contains subfolders for individual packages, platforms (operating systems)
+and for patches to be applied after unpacking a package.
+
+\end{itemize}
+
+The \texttt{dealii/candi} \texttt{github} repository includes continuous
+integration (CI) to ensure that new features of the script or new packages
+are working as expected.
 

### 2021-06-17 9.3/paper.tex
Add Rene as author. Add particle text.

--- 2021-06-17 9.3/paper.tex
@@
     Marc Fehling,
+    Rene Gassm{\"o}ller
     Timo Heister,
@@
 
-\author[4]{Timo~Heister}
- \affil[4]{School of Mathematical and Statistical Sciences,
+\author[4]{Rene~Gassm{\"o}ller}
+\affil[4]{Department of Geological Sciences,
+   University of Florida,
+   1843 Stadium Road,
+   Gainesville, FL, 32611, USA.
+  {\texttt{rene.gassmoeller@ufl.edu}}}
+
+\author[5]{Timo~Heister}
+ \affil[5]{School of Mathematical and Statistical Sciences,
    Clemson University,
@@
 
-\author[5]{Luca~Heltai}
-\affil[5]{SISSA,
+\author[6]{Luca~Heltai}
+\affil[6]{SISSA,
    International School for Advanced Studies,
@@
 
-\author[6]{Uwe~K{\"o}cher}
-\affil[6]{Chair of Numerical Mathematics,
+\author[7]{Uwe~K{\"o}cher}
+\affil[7]{Chair of Numerical Mathematics,
   Helmut-Schmidt-University,
@@
 
- \author[7,8]{Martin~Kronbichler}
- \affil[7]{Institute for Computational Mechanics,
+ \author[8,9]{Martin~Kronbichler}
+ \affil[8]{Institute for Computational Mechanics,
    Technical University of Munich,
@@
    {\texttt{kronbichler/munch/proell@lnm.mw.tum.de}}}
- \affil[8]{Department of Information Technology,
+ \affil[9]{Department of Information Technology,
    Uppsala University,
@@
 
-\author[7,10]{Peter Munch}
- \affil[10]{Institute of Materials Research, Materials Mechanics,
+\author[8,11]{Peter Munch}
+ \affil[11]{Institute of Materials Research, Materials Mechanics,
  Helmholtz-Zentrum Geesthacht,
@@
 
-\author[11]{Jean-Paul~Pelteret}
-\affil[11]{Independent researcher.
+\author[12]{Jean-Paul~Pelteret}
+\affil[12]{Independent researcher.
 {\texttt{jppelteret@gmail.com}}}
@@
 
-\todo[inline]{TODO}
+\dealii 's particle infrastructure has been modernized to store nearly all
+particle data as a set of continuous data arrays instead of a container of
+objects. This reorganization improves cache efficiency when iterating over
+particles and reduces the amount of data that needs to be moved when a particle
+moves to a different cell. Keeping track of unused data slots allows to reuse
+them for new particles, significantly reducing memory allocations when
+particles are created after other particles have left the local domain. The
+data arrays are rebuild and sorted during every mesh refinement cycle.
+
+Additionally the particle infrastructure now supports a faster search algorithm
+for particles that moved farther than one cell width between particle sorting
+operations, and support for updating ghost particles (particles that live in
+ghost cells around the local domain) by updating their properties instead of
+destroying and rebuilding their container was added, which improves the
+efficiency of ghost particle exchange significantly.
+
+Finally the \texttt{Particles::DataOut} class now supports writing particle
+properties as vectors or tensors instead of scalars if the properties are
+marked as such.
 
@@
 
-\item \texttt{step-68} parallel particles, load balancing \todo[inline]{TODO}
+\item \texttt{step-68} is a demonstration of how to embed and distribute
+  particles in a parallel triangulation. It tracks the movement of
+  a set of particles in a fluid flow and illustrates how to distribute
+  a parallel domain according to the number of locally owned
+  particles instead of the number of locally owned cells.
 
@@
 Daniel Garcia-Sanchez,
-Rene Gassmoeller,
 Nicola Giuliani,
@@
 
-W.~Bangerth and T.~Heister were partially
+W.~Bangerth, T.~Heister, and R.~Gassm{\"o}ller were partially
 supported by the Computational Infrastructure
-in Geodynamics initiative (CIG), through the National Science
+for Geodynamics initiative (CIG), through the National Science
 Foundation under Award No.~EAR-1550901 and The
@@
 Technical Data Analysis, Inc. through US Navy STTR Contract N68335-18-C-0011.
+
+R.~Gassm{\"o}ller was also partially supported by the National Science
+Foundation (NSF) award EAR-1925595.
 

### 2021-06-17 9.3/paper.tex
updates introduction of sec 2.7

--- 2021-06-17 9.3/paper.tex
@@
 
-The source based installation of \dealii{} or
-complex toolchains of multiple libraries working together with \dealii{}
-can be done with the \candi{} script tool. \candi is a bash-script based tool,
-it is the abbreviation of ``compile and install'' and it is released under the
-GNU LGPL v3.0.
+The requirement to download, compile and install \dealii{} and its dependencies from
+source is of mayor interest for many \dealii{}-users.
+%
+This can be of importance on arbitrary Linux-based operating systems for which
+binary-install package aren't available or on compute clusters and other machines
+without root privileges for the user to install system dependencies.
+
+The source based installation of \dealii{} or complex toolchains of multiple libraries
+working together with \dealii{} can be done with the \candi{} script tool for various
+Linux operating systems, within the Windows Subsystem Linux (WSL) and on OSX (experimental).
+%
+The general assumption is that a C, C++ and Fortran compiler and suitable MPI-compilers
+for the base compilers as well as the corresponding development system packages are
+available.
+
+\candi is a bash-script based tool, it is the abbreviation of ``compile and install''
+and it is released under the GNU LGPL v3.0.
 %
@@
 The \candi{} tool can be found via the download page of \dealii{} since version 8.5
-and is under continuous and active maintenance. To install older releases of
-\dealii{} toolchains one can checkout the branch corresponding to the
-\dealii{}-version with \texttt{git}.
+and is under active maintenance.
+To install older releases of \dealii{} toolchains one can checkout the branch
+corresponding to the \dealii{}-version with \texttt{git}.
 

### 2021-06-17 9.3/paper.tex
Fix some affiliations
Fix typo

--- 2021-06-17 9.3/paper.tex
@@
 
-\author[7]{Sebastian~Proell}
+\author[8]{Sebastian~Proell}
 
@@
 
-\author[4]{Jiaqi~Zhang}
+\author[5]{Jiaqi~Zhang}
 
@@
 the owning cells and thus can focus on cell-local operations, the new class
-\texttt{RemotePointEvaluation} is responsible for determine the owning cells in a distributed
+\texttt{RemotePointEvaluation} is responsible for determining the owning cells in a distributed
 context and for providing efficient communication patterns for the data exchange. In \texttt{deal.II},

### 2021-06-17 9.3/paper.tex
updates to sec. 2.7

--- 2021-06-17 9.3/paper.tex
@@
 
-An internet connection is not necessary for the run, if the source files
-are downloaded and repositories are checked out to its unpacking location.
-The download of a package can be supported via a local or remote mirror server
+An internet connection isn't necessary
+if the source files are downloaded and
+repositories are checked out to its unpacking location,
+e.g. from a previous run on a machine having an internet connection.
+%
+The download of a package can be supported via local or remote mirror servers
 similar as we are doing it already for several of the packages.
 %
-If the temporary files of \candi{} are not removed, one can use a developer
+If the temporary files of \candi{} aren't removed, one can use the developer
 mode to prepare patches or developments for any of the packages.
@@
 
-This source based installer is implemented in a modular fashion. The main
-script is implemented without any knowledge on a package or toolchain for
-a specific operating system (platform).
-
-The installer is organised in the following way:
+This source based installer is implemented in a modular fashion.
+The \texttt{candi.sh} script is implemented without any knowledge
+on a package or toolchain for a specific operating system.
+Therefore, a couple of checks and setting internal variables
+for known platforms are done.
+
+The \candi{} source based installer is organised in the following way:
 \begin{itemize}
 \item \texttt{candi.sh}: the bash-script which controls the overall process.
+The command line options are given by the output of ``\texttt{candi.sh -h}''
+and explained in the \texttt{README.md} file.
 
@@
 \dealii{}-based toolchains.
-It contains subfolders for individual packages, platforms (operating systems)
-and for patches to be applied after unpacking a package.
+It contains subfolders
+for packages (libraries, dependencies, general tools in a specific version),
+platforms (operating systems) and
+for patches to be applied after unpacking a package.
 

### 2021-06-18 9.3/paper.tex
Minor updates.

--- 2021-06-17 9.3/paper.tex
@@
     Luca Heltai,
+    Uwe K{\"o}cher,
     Martin Kronbichler,
@@
 
-\dealii{} version 9.3.0 was released May ??, 2021.
+\dealii{} version 9.3.0 was released June ??, 2021.
 \todo{Update date}

### 2021-06-18 9.3/paper.tex
added author name

@@
     Sebastian Proell,
+    Konrad Simon,
     Bruno Turcksin,
@@
 \author[8]{Sebastian~Proell}
+
+\author[17]{Konrad Simon}
+  \affil[17]{Department of Mathematics /
+  Center for Earth System Research and Sustainability (CEN),
+  University of Hamburg, Grindelberg 5, 20144 Hamburg, Germany}
 
@@
 Richard Schussnig,
-Konrad Simon,
 Dominic Soldner,

### 2021-06-18 9.3/paper.tex
Minor edits to the particle section.

--- 2021-06-18 9.3/paper.tex
@@
 %
-These major changes are discussed in detail in Section~\ref{sec:major}. There
+In addition, we discuss the \texttt{candi} installation program in Section~\ref{subsec:candi}.
+
+While all of these major changes are discussed in detail in
+Section~\ref{sec:major}, there
 are a number of other noteworthy changes in the current \dealii{} release
@@
 \dealii 's particle infrastructure has been modernized to store nearly all
-particle data as a set of continuous data arrays instead of a container of
+particle data as a collection of continuous data arrays instead of a container of
 objects. This reorganization improves cache efficiency when iterating over
@@
 particles are created after other particles have left the local domain. The
-data arrays are rebuild and sorted during every mesh refinement cycle.
-
-Additionally the particle infrastructure now supports a faster search algorithm
+data arrays are rebuilt and sorted during every mesh refinement cycle.
+
+Additionally, the particle infrastructure now supports a faster search algorithm
 for particles that moved farther than one cell width between particle sorting
@@
 ghost cells around the local domain) by updating their properties instead of
-destroying and rebuilding their container was added, which improves the
+destroying and rebuilding their container. The latter step improves the
 efficiency of ghost particle exchange significantly.
 
-Finally the \texttt{Particles::DataOut} class now supports writing particle
-properties as vectors or tensors instead of scalars if the properties are
+Finally, the \texttt{Particles::DataOut} class now supports writing particle
+properties as vectors or tensors instead of a collection of scalars if the properties are
 marked as such.

### 2021-06-18 9.3/paper.tex
Update 9.3/paper.tex
Co-authored-by: Timo Heister <timo.heister@gmail.com>

--- 2021-06-18 9.3/paper.tex
@@
     Marc Fehling,
-    Rene Gassm{\"o}ller
     Timo Heister,
     Luca Heltai,
-    Uwe K{\"o}cher,
     Martin Kronbichler,
@@
     Jean-Paul Pelteret,
-    Sebastian Proell,
-    Konrad Simon,
     Bruno Turcksin,
-    David Wells,
-    Jiaqi Zhang
+    David Wells
   },
@@
 
-\author[4]{Rene~Gassm{\"o}ller}
-\affil[4]{Department of Geological Sciences,
-   University of Florida,
-   1843 Stadium Road,
-   Gainesville, FL, 32611, USA.
-  {\texttt{rene.gassmoeller@ufl.edu}}}
-
-\author[5]{Timo~Heister}
- \affil[5]{School of Mathematical and Statistical Sciences,
+\author[4]{Timo~Heister}
+ \affil[4]{School of Mathematical and Statistical Sciences,
    Clemson University,
    Clemson, SC, 29634, USA
-   {\texttt{heister/jiaqi2@clemson.edu}}}
-
-\author[6]{Luca~Heltai}
-\affil[6]{SISSA,
+   {\texttt{tcleven/heister@clemson.edu}}}
+
+\author[5]{Luca~Heltai}
+\affil[5]{SISSA,
    International School for Advanced Studies,
@@
 
-\author[7]{Uwe~K{\"o}cher}
-\affil[7]{Chair of Numerical Mathematics,
+\author[6]{Uwe~K{\"o}cher}
+\affil[6]{Chair of Numerical Mathematics,
   Helmut-Schmidt-University,
@@
 
- \author[8,9]{Martin~Kronbichler}
- \affil[8]{Institute for Computational Mechanics,
+ \author[7,8]{Martin~Kronbichler}
+ \affil[7]{Institute for Computational Mechanics,
    Technical University of Munich,
    Boltzmannstr.~15, 85748 Garching, Germany.
-   {\texttt{kronbichler/munch/proell@lnm.mw.tum.de}}}
- \affil[9]{Department of Information Technology,
+   {\texttt{kronbichler/munch@lnm.mw.tum.de}}}
+ \affil[8]{Department of Information Technology,
    Uppsala University,
@@
 
-\author[10]{Matthias~Maier}
-\affil[10]{Department of Mathematics,
+\author[9]{Matthias~Maier}
+\affil[9]{Department of Mathematics,
   Texas A\&M University,
@@
 
-\author[8,11]{Peter Munch}
- \affil[11]{Institute of Materials Research, Materials Mechanics,
+\author[7,10]{Peter Munch}
+ \affil[10]{Institute of Materials Research, Materials Mechanics,
  Helmholtz-Zentrum Geesthacht,
@@
 
-\author[12]{Jean-Paul~Pelteret}
-\affil[12]{Independent researcher.
+\author[11]{Jean-Paul~Pelteret}
+\affil[11]{Independent researcher.
 {\texttt{jppelteret@gmail.com}}}
-
-\author[8]{Sebastian~Proell}
-
-\author[17]{Konrad Simon}
-  \affil[17]{Department of Mathematics /
-  Center for Earth System Research and Sustainability (CEN),
-  University of Hamburg, Grindelberg 5, 20144 Hamburg, Germany}
 
@@
   {\texttt{drwells@email.unc.edu}}}
-
-\author[5]{Jiaqi~Zhang}
 
@@
 
-\dealii{} version 9.3.0 was released June ??, 2021.
+\dealii{} version 9.3.0 was released May ??, 2021.
 \todo{Update date}
@@
   accessible on \texttt{Docker Hub} via \texttt{dealii/dealii:master-focal}. This image is
-  particular useful when used in the continuous-integration processes of user codes.
+  particular useful when used in the contiguous-integration processes of user codes.
 \end{itemize}
@@
 
-\dealii 's particle infrastructure has been modernized to store nearly all
-particle data as a set of continuous data arrays instead of a container of
-objects. This reorganization improves cache efficiency when iterating over
-particles and reduces the amount of data that needs to be moved when a particle
-moves to a different cell. Keeping track of unused data slots allows to reuse
-them for new particles, significantly reducing memory allocations when
-particles are created after other particles have left the local domain. The
-data arrays are rebuild and sorted during every mesh refinement cycle.
-
-Additionally the particle infrastructure now supports a faster search algorithm
-for particles that moved farther than one cell width between particle sorting
-operations, and support for updating ghost particles (particles that live in
-ghost cells around the local domain) by updating their properties instead of
-destroying and rebuilding their container was added, which improves the
-efficiency of ghost particle exchange significantly.
-
-Finally the \texttt{Particles::DataOut} class now supports writing particle
-properties as vectors or tensors instead of scalars if the properties are
-marked as such.
+\todo[inline]{TODO}
 
@@
 the owning cells and thus can focus on cell-local operations, the new class
-\texttt{RemotePointEvaluation} is responsible for determining the owning cells in a distributed
+\texttt{RemotePointEvaluation} is responsible for determine the owning cells in a distributed
 context and for providing efficient communication patterns for the data exchange. In \texttt{deal.II},
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Terminal based toolchain installer}
+\subsection{Source based toolchain installer}
 \label{subsec:candi}
 
-Half a page up to a page on candi.
+The requirement to download, compile and install \dealii{} and its dependencies from
+source is of major interest for many \dealii{}-users.
+%
+This can be of importance on arbitrary Linux-based operating systems for which
+binary-install package aren't available or on compute clusters and other machines
+without root privileges for the user to install system dependencies.
+
+The source based installation of \dealii{} or complex toolchains of multiple libraries
+working together with \dealii{} can be done with the \candi{} script tool for various
+Linux operating systems, within the Windows Subsystem Linux (WSL) and on OSX (experimental).
+%
+The general assumption is that a C, C++ and Fortran compiler and suitable MPI-compilers
+for the base compilers as well as the corresponding development system packages are
+available.
+
+\candi is a bash-script based tool, it is the abbreviation of ``compile and install''
+and it is released under the GNU LGPL v3.0.
+%
+The origin of \candi{} is a fork made in 2013 from \texttt{dorsal},
+a retired source based installer for the \texttt{FEniCS} library.
+%
+The \candi{} tool can be found via the download page of \dealii{} since version 8.5
+and is under active maintenance.
+To install older releases of \dealii{} toolchains one can checkout the branch
+corresponding to the \dealii{}-version with \texttt{git}.
+
+The installer supports the user to download, unpack, compile and install a certain
+library, a list of libraries or a complete toolchain.
+A library is called a package within \candi{}.
+The toolchain installation is the default behaviour and the default configuration
+ensures that most of the \dealii{} step tutorials can be used directly.
+The package installation mode can be handy to generate docker containers
+for instance.
+Especially in the toolchain mode we can check for dependencies from one library to
+another and handle them accordingly.
+
+Each package for a library is defined by variables for its name and version,
+a (remote or local) download location,
+its packaging format (e.g. tar, gz or zip),
+a checksum to prevent from incomplete downloads,
+its general build chain (e.g. using \texttt{autotools}, \texttt{cmake} and others)
+as well as configuration options.
+%
+Moreover, one can give specific instructions for each of the steps being
+necessary for the unpacking, configuration or building of the package.
+
+An internet connection isn't necessary
+if the source files are downloaded and
+repositories are checked out to its unpacking location,
+e.g. from a previous run on a machine having an internet connection.
+%
+The download of a package can be supported via local or remote mirror servers
+similar as we are doing it already for several of the packages.
+%
+If the temporary files of \candi{} aren't removed, one can use the developer
+mode to prepare patches or developments for any of the packages.
+%
+A new feature allows \candi{} to skip user prompts which can be used for
+a batch mode run on a compute cluster job for instance.
+
+This source based installer is implemented in a modular fashion.
+The \texttt{candi.sh} script is implemented without any knowledge
+on a package or toolchain for a specific operating system.
+Therefore, a couple of checks and setting internal variables
+for known platforms are done.
+
+The \candi{} source based installer is organised in the following way:
+\begin{itemize}
+\item \texttt{candi.sh}: the bash-script which controls the overall process.
+The command line options are given by the output of ``\texttt{candi.sh -h}''
+and explained in the \texttt{README.md} file.
+
+\item \texttt{candi.cfg} (or \texttt{local.cfg}): the default toolchain
+configuration file. Here one can easily switch on/off important features,
+give additional configuration options to \dealii{} and
+define the set of packages for the toolchain.
+
+\item \texttt{deal.II-toolchain} folder: the central project folder for
+\dealii{}-based toolchains.
+It contains subfolders
+for packages (libraries, dependencies, general tools in a specific version),
+platforms (operating systems) and
+for patches to be applied after unpacking a package.
+
+\end{itemize}
+
+The \texttt{dealii/candi} \texttt{github} repository includes continuous
+integration (CI) to ensure that new features of the script or new packages
+are working as expected.
 
@@
 
-\item \texttt{step-68} is a demonstration of how to embed and distribute
-  particles in a parallel triangulation. It tracks the movement of
-  a set of particles in a fluid flow and illustrates how to distribute
-  a parallel domain according to the number of locally owned
-  particles instead of the number of locally owned cells.
+\item \texttt{step-68} parallel particles, load balancing \todo[inline]{TODO}
 
@@
 Daniel Garcia-Sanchez,
+Rene Gassmoeller,
 Nicola Giuliani,
@@
 Daniel Paukner,
+Sebastian Proell,
 Lei Qiao,
@@
 Richard Schussnig,
+Konrad Simon,
 Dominic Soldner,
@@
 
-W.~Bangerth, T.~Heister, and R.~Gassm{\"o}ller were partially
+W.~Bangerth and T.~Heister were partially
 supported by the Computational Infrastructure
-for Geodynamics initiative (CIG), through the National Science
+in Geodynamics initiative (CIG), through the National Science
 Foundation under Award No.~EAR-1550901 and The
@@
 Technical Data Analysis, Inc. through US Navy STTR Contract N68335-18-C-0011.
-
-R.~Gassm{\"o}ller was also partially supported by the National Science
-Foundation (NSF) award EAR-1925595.
 

### 2021-06-18 9.3/paper.tex
Update 9.3/paper.tex
Co-authored-by: Timo Heister <timo.heister@gmail.com>

@@
 This can be of importance on arbitrary Linux-based operating systems for which
-binary-install package aren't available or on compute clusters and other machines
+binary packages aren't available or on compute clusters and other machines
 without root privileges for the user to install system dependencies.

### 2021-06-18 9.3/paper.tex
Update 9.3/paper.tex
Co-authored-by: Timo Heister <timo.heister@gmail.com>

@@
 The source based installation of \dealii{} or complex toolchains of multiple libraries
-working together with \dealii{} can be done with the \candi{} script tool for various
+can be done with the \candi{} script tool for various
 Linux operating systems, within the Windows Subsystem Linux (WSL) and on OSX (experimental).

### 2021-06-18 9.3/paper.tex
Update 9.3/paper.tex
Co-authored-by: Timo Heister <timo.heister@gmail.com>

@@
 
-The source based installation of \dealii{} or complex toolchains of multiple libraries
+The source based installation of \dealii{} and it's many dependent libraries
 can be done with the \candi{} script tool for various

### 2021-06-18 9.3/paper.tex
Update 9.3/paper.tex
Co-authored-by: Timo Heister <timo.heister@gmail.com>

@@
 %
-This can be of importance on arbitrary Linux-based operating systems for which
+Compiling dependencies from source can be of importance on arbitrary Linux-based operating systems for which
 binary packages aren't available or on compute clusters and other machines
@@
 The \candi{} tool can be found via the download page of \dealii{} since version 8.5
-and is under active maintenance.
+or from github.com/dealii/candi and is under active development.
 To install older releases of \dealii{} toolchains one can checkout the branch

### 2021-06-18 9.3/paper.tex

@@
     Marc Fehling,
+    Rene Gassm{\"o}ller
     Timo Heister,
     Luca Heltai,
+    Uwe K{\"o}cher,
     Martin Kronbichler,
@@
     Jean-Paul Pelteret,
+    Sebastian Proell,
+    Konrad Simon,
     Bruno Turcksin,
-    David Wells
+    David Wells,
+    Jiaqi Zhang
   },
@@
 
-\author[4]{Timo~Heister}
- \affil[4]{School of Mathematical and Statistical Sciences,
+\author[4]{Rene~Gassm{\"o}ller}
+\affil[4]{Department of Geological Sciences,
+   University of Florida,
+   1843 Stadium Road,
+   Gainesville, FL, 32611, USA.
+  {\texttt{rene.gassmoeller@ufl.edu}}}
+
+\author[5]{Timo~Heister}
+ \affil[5]{School of Mathematical and Statistical Sciences,
    Clemson University,
    Clemson, SC, 29634, USA
-   {\texttt{tcleven/heister@clemson.edu}}}
-
-\author[5]{Luca~Heltai}
-\affil[5]{SISSA,
+   {\texttt{heister/jiaqi2@clemson.edu}}}
+
+\author[6]{Luca~Heltai}
+\affil[6]{SISSA,
    International School for Advanced Studies,
@@
 
- \author[7,8]{Martin~Kronbichler}
- \affil[7]{Institute for Computational Mechanics,
+ \author[8,9]{Martin~Kronbichler}
+ \affil[8]{Institute for Computational Mechanics,
    Technical University of Munich,
    Boltzmannstr.~15, 85748 Garching, Germany.
-   {\texttt{kronbichler/munch@lnm.mw.tum.de}}}
- \affil[8]{Department of Information Technology,
+   {\texttt{kronbichler/munch/proell@lnm.mw.tum.de}}}
+ \affil[9]{Department of Information Technology,
    Uppsala University,
@@
 
-\author[11]{Jean-Paul~Pelteret}
-\affil[11]{Independent researcher.
+\author[12]{Jean-Paul~Pelteret}
+\affil[12]{Independent researcher.
 {\texttt{jppelteret@gmail.com}}}
+
+\author[8]{Sebastian~Proell}
+
+\author[17]{Konrad Simon}
+  \affil[17]{Department of Mathematics /
+  Center for Earth System Research and Sustainability (CEN),
+  University of Hamburg, Grindelberg 5, 20144 Hamburg, Germany}
 
@@
   {\texttt{drwells@email.unc.edu}}}
+
+\author[5]{Jiaqi~Zhang}
 
@@
 
-\todo[inline]{TODO}
+\dealii 's particle infrastructure has been modernized to store nearly all
+particle data as a collection of continuous data arrays instead of a container of
+objects. This reorganization improves cache efficiency when iterating over
+particles and reduces the amount of data that needs to be moved when a particle
+moves to a different cell. Keeping track of unused data slots allows to reuse
+them for new particles, significantly reducing memory allocations when
+particles are created after other particles have left the local domain. The
+data arrays are rebuilt and sorted during every mesh refinement cycle.
+
+Additionally, the particle infrastructure now supports a faster search algorithm
+for particles that moved farther than one cell width between particle sorting
+operations, and support for updating ghost particles (particles that live in
+ghost cells around the local domain) by updating their properties instead of
+destroying and rebuilding their container. The latter step improves the
+efficiency of ghost particle exchange significantly.
+
+Finally, the \texttt{Particles::DataOut} class now supports writing particle
+properties as vectors or tensors instead of a collection of scalars if the properties are
+marked as such.
 

### 2021-06-18 9.3/paper.tex
Fix minor issues.

@@
 
-\author[17]{Konrad Simon}
-  \affil[17]{Department of Mathematics /
-  Center for Earth System Research and Sustainability (CEN),
+\author[13]{Konrad Simon}
+  \affil[13]{Department of Mathematics/Center for Earth System Research and Sustainability (CEN),
   University of Hamburg, Grindelberg 5, 20144 Hamburg, Germany}
@@
 
-\author[16]{David Wells}
-\affil[16]{Department of Mathematics, University of North Carolina,
+\author[14]{David Wells}
+\affil[14]{Department of Mathematics, University of North Carolina,
   Chapel Hill, NC 27516, USA.
@@
 
-The source based installation of \dealii{} and it's many dependent libraries
+The source based installation of \dealii{} and its many dependent libraries
 can be done with the \candi{} script tool for various

### 2021-06-18 9.3/paper.tex
Edit the candi section.

@@
 %
-In addition, we discuss the \texttt{candi} installation program in Section~\ref{subsec:candi}.
+In addition, we discuss the \candi{} installation program in Section~\ref{subsec:candi}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Source based toolchain installer}
+\subsection{The ource based toolchain installer \candi{}}
 \label{subsec:candi}
 
-The requirement to download, compile and install \dealii{} and its dependencies from
-source is of major interest for many \dealii{}-users.
-%
-Compiling dependencies from source can be of importance on arbitrary Linux-based operating systems for which
+The requirement to download, compile, and install \dealii{} and its dependencies from
+source is a major obstacle to many \dealii{} users.
+Compiling all dependencies from source can be difficult on operating systems for which
 binary packages aren't available or on compute clusters and other machines
@@
 can be done with the \candi{} script tool for various
-Linux operating systems, within the Windows Subsystem Linux (WSL) and on OSX (experimental).
+Linux operating systems, within the Windows Subsystem Linux (WSL) and
+on OS X (experimental).
 %
@@
 
-\candi is a bash-script based tool, it is the abbreviation of ``compile and install''
-and it is released under the GNU LGPL v3.0.
+\candi is a bash-script based tool, is an abbreviation of ``compile and install''
+and is released under the GNU LGPL v3.0.
 %
 The origin of \candi{} is a fork made in 2013 from \texttt{dorsal},
-a retired source based installer for the \texttt{FEniCS} library.
+a now-retired source based installer for the \texttt{FEniCS} library.
 %
 The \candi{} tool can be found via the download page of \dealii{} since version 8.5
-or from github.com/dealii/candi and is under active development.
-To install older releases of \dealii{} toolchains one can checkout the branch
-corresponding to the \dealii{}-version with \texttt{git}.
-
-The installer supports the user to download, unpack, compile and install a certain
-library, a list of libraries or a complete toolchain.
-A library is called a package within \candi{}.
-The toolchain installation is the default behaviour and the default configuration
+or from \url{github.com/dealii/candi}, and is under active development.
+To install older releases of \dealii{} toolchains, one can check out
+the \candi{} branch of the \texttt{git} repository that
+corresponds to the desired \dealii{} version.
+
+\candi{} downloads, unpacks, compiles, and installs an individual
+library (called a ``package'' in \candi{}), a list of libraries, or a complete toolchain.
+The toolchain installation is the default behavior and the default configuration
 ensures that most of the \dealii{} step tutorials can be used directly.
-The package installation mode can be handy to generate docker containers
-for instance.
-Especially in the toolchain mode we can check for dependencies from one library to
-another and handle them accordingly.
+The package installation mode is also useful to generate docker containers.
+In toolchain mode, \candi{} checks for and deals appropriately with
+dependencies between libraries.
 
@@
 a (remote or local) download location,
-its packaging format (e.g. tar, gz or zip),
-a checksum to prevent from incomplete downloads,
-its general build chain (e.g. using \texttt{autotools}, \texttt{cmake} and others)
+its packaging format (e.g. tar, gz, or zip),
+a checksum,
+its general build chain (e.g., using \texttt{autotools}, \texttt{cmake}, or others),
 as well as configuration options.
-%
 Moreover, one can give specific instructions for each of the steps being
-necessary for the unpacking, configuration or building of the package.
-
-An internet connection isn't necessary
-if the source files are downloaded and
-repositories are checked out to its unpacking location,
-e.g. from a previous run on a machine having an internet connection.
-%
-The download of a package can be supported via local or remote mirror servers
-similar as we are doing it already for several of the packages.
-%
-If the temporary files of \candi{} aren't removed, one can use the developer
+necessary for the unpacking, configuration, or building of a package.
+A new feature allows \candi{} to skip user prompts, allowing its use in
+a batch mode.
+
+\candi{} does all this by either downloading packages from the
+internet, including through mirrors, or re-using previously downloaded
+or checked out from a repository.
+If temporary files of \candi{} are not removed, one can use the developer
 mode to prepare patches or developments for any of the packages.
-%
-A new feature allows \candi{} to skip user prompts which can be used for
-a batch mode run on a compute cluster job for instance.
-
-This source based installer is implemented in a modular fashion.
-The \texttt{candi.sh} script is implemented without any knowledge
-on a package or toolchain for a specific operating system.
-Therefore, a couple of checks and setting internal variables
-for known platforms are done.
-
-The \candi{} source based installer is organised in the following way:
+
+\candi{} is organized in the following way:
 \begin{itemize}
-\item \texttt{candi.sh}: the bash-script which controls the overall process.
-The command line options are given by the output of ``\texttt{candi.sh -h}''
+\item \texttt{candi.sh}: the bash script which controls the overall process.
+Available command line options are listed by calling ``\texttt{candi.sh -h}''
 and explained in the \texttt{README.md} file.
@@
 \item \texttt{candi.cfg} (or \texttt{local.cfg}): the default toolchain
-configuration file. Here one can easily switch on/off important features,
-give additional configuration options to \dealii{} and
-define the set of packages for the toolchain.
+configuration file. Here one can easily switch on/off features,
+give additional configuration, and
+list the packages for a toolchain.
 
@@
 for patches to be applied after unpacking a package.
-
 \end{itemize}
 
-The \texttt{dealii/candi} \texttt{github} repository includes continuous
-integration (CI) to ensure that new features of the script or new packages
+The installer is developed in the \texttt{dealii/candi}
+\texttt{github} repository, and uses continuous
+integration (CI) to ensure that old and new features
 are working as expected.

### 2021-06-18 9.3/paper.tex
Fix typo.
Write text for steps 78 and 79.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The ource based toolchain installer \candi{}}
+\subsection{The source-based toolchain installer \candi{}}
 \label{subsec:candi}
@@
 
-\item \texttt{step-78} \todo[inline]{TODO}
-
-\item \texttt{step-79} \todo[inline]{TODO}
+\item \texttt{step-78}, a program written by Tyler Anderson at
+  Colorado State University, solves the
+  one-dimensional Black-Scholes equations to model the price of stock
+  options.
+
+\item \texttt{step-79} is a program for topology optimization. Written
+  by Justin O'Connor at Colorado State University, the program asks
+  what the optimal distribution of a finite amount of material in a
+  domain is to maximize its strength (or optimize some other objective
+  functional). It uses advanced optimization techniques.
 \end{itemize}

### 2021-06-18 9.3/paper.tex
Add a sentence.
Add Bruno's name & address.

@@
     Wolfgang Bangerth,
+    Bruno Blais,
     Marc Fehling,
@@
 
+\author[4]{Bruno Blais}
+\affil[4]{Research Unit for Industrial Flows Processes (URPEI), Department of Chemical Engineering,
+          Polytechnique Montréal,
+          PO Box 6079, Stn Centre-Ville, Montréal, Québec, Canada, H3C 3A7.
+          {\texttt{bruno.blais@polymtl.ca}}}
+
 \author[2]{Marc~Fehling}
 
-\author[4]{Rene~Gassm{\"o}ller}
-\affil[4]{Department of Geological Sciences,
+\author[5]{Rene~Gassm{\"o}ller}
+\affil[5]{Department of Geological Sciences,
    University of Florida,
@@
 
-\author[5]{Timo~Heister}
- \affil[5]{School of Mathematical and Statistical Sciences,
+\author[6]{Timo~Heister}
+ \affil[6]{School of Mathematical and Statistical Sciences,
    Clemson University,
@@
 
-\author[6]{Luca~Heltai}
-\affil[6]{SISSA,
+\author[7]{Luca~Heltai}
+\affil[7]{SISSA,
    International School for Advanced Studies,
@@
 
-\author[7]{Uwe~K{\"o}cher}
-\affil[7]{Chair of Numerical Mathematics,
+\author[8]{Uwe~K{\"o}cher}
+\affil[8]{Chair of Numerical Mathematics,
   Helmut-Schmidt-University,
@@
 
- \author[8,9]{Martin~Kronbichler}
- \affil[8]{Institute for Computational Mechanics,
+ \author[9,10]{Martin~Kronbichler}
+ \affil[9]{Institute for Computational Mechanics,
    Technical University of Munich,
@@
    {\texttt{kronbichler/munch/proell@lnm.mw.tum.de}}}
- \affil[9]{Department of Information Technology,
+ \affil[10]{Department of Information Technology,
    Uppsala University,
@@
 
-\author[8,11]{Peter Munch}
- \affil[11]{Institute of Materials Research, Materials Mechanics,
+\author[9,12]{Peter Munch}
+ \affil[12]{Institute of Materials Research, Materials Mechanics,
  Helmholtz-Zentrum Geesthacht,
@@
 
-\author[12]{Jean-Paul~Pelteret}
-\affil[12]{Independent researcher.
+\author[13]{Jean-Paul~Pelteret}
+\affil[13]{Independent researcher.
 {\texttt{jppelteret@gmail.com}}}
 
-\author[8]{Sebastian~Proell}
-
-\author[13]{Konrad Simon}
-  \affil[13]{Department of Mathematics/Center for Earth System Research and Sustainability (CEN),
+\author[9]{Sebastian~Proell}
+
+\author[14]{Konrad Simon}
+  \affil[14]{Department of Mathematics/Center for Earth System Research and Sustainability (CEN),
   University of Hamburg, Grindelberg 5, 20144 Hamburg, Germany}
@@
 
-\author[14]{David Wells}
-\affil[14]{Department of Mathematics, University of North Carolina,
+\author[15]{David Wells}
+\affil[15]{Department of Mathematics, University of North Carolina,
   Chapel Hill, NC 27516, USA.
@@
 
-\author[5]{Jiaqi~Zhang}
+\author[6]{Jiaqi~Zhang}
 
@@
   domain is to maximize its strength (or optimize some other objective
-  functional). It uses advanced optimization techniques.
+  functional). It uses advanced optimization techniques to deal with
+  the nonlinearity of the problem as well as with the equality and
+  inequality constraints that characterize the application.
 \end{itemize}

### 2021-06-18 9.3/paper.tex
Remove Bruno Blais from the "other contributors".

@@
 Nicolas Barnafi,
-Bruno Blais,
 Michele Bucelli,
@@
 
-\todo[inline]{Remove everyone who may end up being a co-author.}
-
 Their contributions are much appreciated!

### 2021-06-18 9.3/paper.tex
Use 6/17 as the release date, just like on the homepage.

@@
 
-\dealii{} version 9.3.0 was released June ??, 2021.
-\todo{Update date}
+\dealii{} version 9.3.0 was released June 17, 2021.
 This paper provides an
@@
 \dealii{} can interface with many other libraries:
- \todo{Bruno: Please add some kind of reference for ArboxX}
 \begin{multicols}{3}
@@
     \item ADOL-C \cite{Griewank1996a,adol-c}
-    \item ArborX
+    \item ArborX \cite{lebrun2020arborx}
     \item ARPACK \cite{arpack}

### 2021-06-18 9.3/paper.tex
Add a reference to ArborX
funding jiaqi

@@
 
-W.~Bangerth, T.~Heister, and R.~Gassm{\"o}ller were partially
+W.~Bangerth, T.~Heister, R.~Gassm{\"o}ller, and J.~Zhang were partially
 supported by the Computational Infrastructure
@@
 
-Timo Heister was also partially supported by the National Science Foundation (NSF)
-Award DMS-2028346, OAC-2015848, EAR-1925575, and by
+T.~Heister and J.~Zhang were also partially supported by NSF OAC-2015848.
+
+T.~Heister was also partially supported by the NSF awards
+DMS-2028346, EAR-1925575, and by
 Technical Data Analysis, Inc. through US Navy STTR Contract N68335-18-C-0011.

### 2021-06-18 9.3/paper.tex
step-66
remove sentence

@@
 
-\item \texttt{step-66} nonlinear matrix-free solver \todo[inline]{TODO}
+\item \texttt{step-66} shows how to solve a nonlinear problem using Newton's method in the matrix-free
+framework in parallel. The PDE considered is the Gelfand problem $-\triangle u = \exp(u)$.
 
@@
 Please consider citing the appropriate references if you use
-interfaces to these libraries. We note that the nanoflann and NetCDF
-interfaces are now deprecated and will be removed in \dealii{} version
-9.3.
+interfaces to these libraries.
 

### 2021-06-19 9.3/paper.tex
improve candi wording

@@
 source is a major obstacle to many \dealii{} users.
-Compiling all dependencies from source can be difficult on operating systems for which
+Compiling all dependencies from source can be difficult but is a necessity on operating systems for which
 binary packages aren't available or on compute clusters and other machines
@@
 can be done with the \candi{} script tool for various
-Linux operating systems, within the Windows Subsystem Linux (WSL) and
+Linux operating systems, within the Windows Subsystem Linux (WSL), and
 on OS X (experimental).
@@
 The package installation mode is also useful to generate docker containers.
-In toolchain mode, \candi{} checks for and deals appropriately with
-dependencies between libraries.
+In toolchain mode, \candi{} checks for and deals with
+dependencies between libraries appropriately.
 

### 2021-06-19 9.3/paper.tex
step-74 description

@@
 
-\item \texttt{step-74} SIPG for the Poisson equation \todo[inline]{TODO}
+\item \texttt{step-74} implements the symmetric interior penalty Galerkin 
+(SIPG) method for Poisson's equation using the \texttt{FEInterfaceValues} class
+in combination with \texttt{MeshWorker::mesh\_loop()}. With \texttt{FEInterfaceValues},
+we can directly obtain the jump and the average of shape functions and of the solution 
+across cell faces, which simplifies the assembling of face terms in DG methods especially on 
+adaptive meshes.
 

### 2021-06-19 9.3/paper.tex
Write text about preCICE

--- 2021-06-19 9.3/paper.tex
@@
 \begin{itemize}
-  \item preCICE: \todo[inline]{TODO}
+\item ``\texttt{Laplace equation coupled to an external simulation program}'' was
+  contributed by David Schneider and Benjamin Uekermann at Technical University
+  of Munich. It shows by extending \texttt{step-4}---a time-dependent Poisson
+  problem~---how to surface-couple a \dealii application with a simple external 
+  application, using the \texttt{preCICE} 
+  library~\cite{bungartz2016precice, lindner2020exafsa}. \texttt{preCICE},
+  however, also allows to couple \dealii to actual external simulation softwares, 
+  such as OpenFOAM, SU2, CalculiX, or FEniCS, using provided adapter classes. 
+  For more details visit their webpage
+  \url{https://precice.org/} and check out the numerous \dealii tutorials there,
+  showing, i.a., interesting fluid-structure interaction application cases.
 \end{itemize}

### 2021-06-19 9.3/paper.tex
Fix fpe code

--- 2021-06-19 9.3/paper.tex
@@
 \begin{c++}
+phi_curvature.reinit(cell, reference_points);
 phi_normal.reinit(cell, reference_points);
 phi_force.reinit(cell, reference_points);
+
+phi_curvature.evaluate(curvature_values, EvaluationFlags::values);
 phi_normal.evaluate(normal_values, EvaluationFlags::values);
+
 for (unsigned int q = 0; q < n_points; ++q)
@@
                          phi_normal.get_value(q) * JxW[q], q);
+
 phi_force.integrate(force_values, EvaluationFlags::values);

### 2021-06-19 9.3/paper.tex
HZG -> Hereon
elaborate on changes, add email

@@
 \author[9,12]{Peter Munch}
- \affil[12]{Institute of Materials Research, Materials Mechanics,
- Helmholtz-Zentrum Geesthacht,
+ \affil[12]{Institute of Material Systems Modeling,
+ Helmholtz-Zentrum Hereon,
  Max-Planck-Str. 1, 21502 Geesthacht, Germany.
-   {\texttt{peter.muench@hzg.de}}}
+   {\texttt{peter.muench@hereon.de}}}
 
@@
 \author[14]{Konrad Simon}
-  \affil[14]{Department of Mathematics/Center for Earth System Research and Sustainability (CEN),
-  University of Hamburg, Grindelberg 5, 20144 Hamburg, Germany}
+  \affil[14]{Department of Mathematics/Center
+  for Earth System Research and Sustainability (CEN), University of
+  Hamburg, Grindelberg 5, 20144 Hamburg, Germany.
+  \texttt{konrad.simon@uni-hamburg.de}}}
 
@@
   warnings if \dealii{} has been configured with \texttt{-D DEAL\_II\_EARLY\_DEPRECATIONS=ON}.
+
   \item After each update of the master branch of \dealii{}, we build a new Docker image
@@
   particular useful when used in the continuous-integration processes of user codes.
+
+  \item A long standing orientation issue for the vector valued finite
+  element \texttt{FE\_RaviartThomas<3>} was resolved. This issue
+  prevented the user from using this class on meshes with cells that
+  do not have standard orientation, i.e., cells that have faces that
+  are flipped and rotated relative to their neighbors. In all eight
+  possible neighboring configurations we can now guarantee that the
+  necessary $H(\mathrm{div})$-conformity condition is met for all
+  polynomaial orders. This is verified through a new conformity test
+  on non-standard meshes.
 \end{itemize}

### 2021-06-20 9.3/paper.tex
Small addition to particles + funding

--- 2021-06-19 9.3/paper.tex
@@
 destroying and rebuilding their container. The latter step improves the
-efficiency of ghost particle exchange significantly.
+efficiency of ghost particle exchange significantly. With this feature,
+\dealii can be used to generate scalable parallel code for Lagrangian
+models such as the Discrete Element Method (DEM) \cite{golshan2021lethedem}
+or Molecular Dynamics.
 
@@
 
+B.~Blais was partially supported by the National Science and Engineering Research Council of Canada (NSERC)  through the RGPIN-2020-04510 Discovery Grant
+
 T.~Heister and J.~Zhang were also partially supported by NSF OAC-2015848.

### 2021-06-20 9.3/paper.tex
Fix compilation
Change fliped -> reflected

@@
   Hamburg, Grindelberg 5, 20144 Hamburg, Germany.
-  \texttt{konrad.simon@uni-hamburg.de}}}
+  \texttt{konrad.simon@uni-hamburg.de}}
 
@@
   do not have standard orientation, i.e., cells that have faces that
-  are flipped and rotated relative to their neighbors. In all eight
-  possible neighboring configurations we can now guarantee that the
+  are reflected and/or rotated relative to their neighbors. In all eight
+  possible neighboring configurations, we can now guarantee that the
   necessary $H(\mathrm{div})$-conformity condition is met for all

### 2021-06-20 9.3/paper.tex
description of FEInterfaceValues

--- 2021-06-20 9.3/paper.tex
@@
 \author[9,12]{Peter Munch}
- \affil[12]{Institute of Material Systems Modeling,
- Helmholtz-Zentrum Hereon,
+ \affil[12]{Institute of Materials Research, Materials Mechanics,
+ Helmholtz-Zentrum Geesthacht,
  Max-Planck-Str. 1, 21502 Geesthacht, Germany.
-   {\texttt{peter.muench@hereon.de}}}
+   {\texttt{peter.muench@hzg.de}}}
 
@@
   \item Improved support for evaluation and integration at arbitrary points (see Section~\ref{subsec:fepointvalues});
+  \item Simplified implementation for cell face integrals (see Section~\ref{subsec:feinterfacevalues});
   \item Nine new tutorial programs and a new code gallery program (see Section~\ref{subsec:steps}).
@@
 destroying and rebuilding their container. The latter step improves the
-efficiency of ghost particle exchange significantly. With this feature,
-\dealii can be used to generate scalable parallel code for Lagrangian
-models such as the Discrete Element Method (DEM) \cite{golshan2021lethedem}
-or Molecular Dynamics.
+efficiency of ghost particle exchange significantly.
 
@@
 \begin{c++}
-phi_curvature.reinit(cell, reference_points);
 phi_normal.reinit(cell, reference_points);
 phi_force.reinit(cell, reference_points);
-
-phi_curvature.evaluate(curvature_values, EvaluationFlags::values);
 phi_normal.evaluate(normal_values, EvaluationFlags::values);
-
 for (unsigned int q = 0; q < n_points; ++q)
@@
                           phi_normal.get_value(q) * JxW[q], q);
-
 phi_force.integrate(force_values, EvaluationFlags::values);
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Simplied implementation for cell face integrals}
+\label{subsec:feinterfacevalues}
+Discontinuous Galerkin (DG) methods require evaluations of averages and jumps accross cell faces, therefore we need to compute values of the shape functions and solutions from two adjacent cells.
+The \texttt{FEInterfaceValues} class is designed to provide an interface to directly obtain jumps and averages.
+When used with \texttt{MeshWorker::mesh\_loop()}, we can assemble cell face integrals in a uniform way independent of mesh types, which significantly simplifies the implementation.
+
+Internally, this class provides an abstraction for two \texttt{FEFaceValues} objects (or \texttt{FESubfaceValues} when using adaptive refinement).
+The class introduces new interface degrees of freedom indices that are the union of the degrees of freedom indices of the two \texttt{FEFaceValues} objects.
+The interface degrees of freedom indices can be converted to the corresponding local degrees of freedom indices of the two cells using a helper function.
+Furthermore, this class supports vector-valued problems in a similar way as \texttt{FEFaceValues} does.
+We can easily obtain the corresponding scalar or vector values by providing an \texttt{FEValuesExtractors} object.
+Overall, this class is intended to be a low level replacement for \texttt{MeshWorker} and \texttt{LocalIntegrators}, and a higher level abstraction compared to assembling face terms manually.
+
+Examples of \texttt{FEInterfaceValues} for scalar problems can be found in \texttt{step-12}, \texttt{step-47}, \texttt{step-50}, and the new tutorial \texttt{step-74}.
+While examples of vector-valued problems will be added in the future.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{The source-based toolchain installer \candi{}}
@@
 
-\item \texttt{step-74} SIPG for the Poisson equation \todo[inline]{TODO}
+\item \texttt{step-74} implements the symmetric interior penalty Galerkin 
+(SIPG) method for Poisson's equation using the \texttt{FEInterfaceValues} class
+in combination with \texttt{MeshWorker::mesh\_loop()}. This tutorial demonstrates a
+simple and mesh-independent way to assemble cell face integrals.
 
@@
 
-B.~Blais was partially supported by the National Science and Engineering Research Council of Canada (NSERC)  through the RGPIN-2020-04510 Discovery Grant
-
 T.~Heister and J.~Zhang were also partially supported by NSF OAC-2015848.

### 2021-06-20 9.3/paper.tex
Update 9.3/paper.tex
Co-authored-by: Martin Kronbichler <kronbichler@lnm.mw.tum.de>

@@
   \item Improved support for evaluation and integration at arbitrary points (see Section~\ref{subsec:fepointvalues});
-  \item Simplified implementation for cell face integrals (see Section~\ref{subsec:feinterfacevalues});
+  \item Simplified implementation for face integrals (see Section~\ref{subsec:feinterfacevalues});
   \item Nine new tutorial programs and a new code gallery program (see Section~\ref{subsec:steps}).
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Simplied implementation for cell face integrals}
+\subsection{Simplied implementation for face integrals}
 \label{subsec:feinterfacevalues}
@@
 The \texttt{FEInterfaceValues} class is designed to provide an interface to directly obtain jumps and averages.
-When used with \texttt{MeshWorker::mesh\_loop()}, we can assemble cell face integrals in a uniform way independent of mesh types, which significantly simplifies the implementation.
+When used with \texttt{MeshWorker::mesh\_loop()}, we can assemble face integrals in a uniform way independent of mesh types, which significantly simplifies the implementation.
 
@@
 Examples of \texttt{FEInterfaceValues} for scalar problems can be found in \texttt{step-12}, \texttt{step-47}, \texttt{step-50}, and the new tutorial \texttt{step-74}.
-While examples of vector-valued problems will be added in the future.
+Examples of vector-valued problems will be added in the future.
 
@@
 in combination with \texttt{MeshWorker::mesh\_loop()}. This tutorial demonstrates a
-simple and mesh-independent way to assemble cell face integrals.
+simple and mesh-independent way to assemble face integrals.
 

### 2021-06-21 9.3/paper.tex
Update name of Daniel and Bruno's group

@@
  \author[1*]{Daniel Arndt}
- \affil[1]{Computational Engineering and Energy Sciences Group,
+ \affil[1]{Scalable Algorithms and Coupled Physics Group,
    Computational Sciences and Engineering Division,

### 2021-06-21 9.3/paper.tex

@@
   \item Improved support for evaluation and integration at arbitrary points (see Section~\ref{subsec:fepointvalues});
+  \item Simplified implementation for face integrals (see Section~\ref{subsec:feinterfacevalues});
   \item Nine new tutorial programs and a new code gallery program (see Section~\ref{subsec:steps}).
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Simplied implementation for face integrals}
+\label{subsec:feinterfacevalues}
+Discontinuous Galerkin (DG) methods require evaluations of averages and jumps accross cell faces, therefore we need to compute values of the shape functions and solutions from two adjacent cells.
+The \texttt{FEInterfaceValues} class is designed to provide an interface to directly obtain jumps and averages.
+When used with \texttt{MeshWorker::mesh\_loop()}, we can assemble face integrals in a uniform way independent of mesh types, which significantly simplifies the implementation.
+
+Internally, this class provides an abstraction for two \texttt{FEFaceValues} objects (or \texttt{FESubfaceValues} when using adaptive refinement).
+The class introduces new interface degrees of freedom indices that are the union of the degrees of freedom indices of the two \texttt{FEFaceValues} objects.
+The interface degrees of freedom indices can be converted to the corresponding local degrees of freedom indices of the two cells using a helper function.
+Furthermore, this class supports vector-valued problems in a similar way as \texttt{FEFaceValues} does.
+We can easily obtain the corresponding scalar or vector values by providing an \texttt{FEValuesExtractors} object.
+Overall, this class is intended to be a low level replacement for \texttt{MeshWorker} and \texttt{LocalIntegrators}, and a higher level abstraction compared to assembling face terms manually.
+
+Examples of \texttt{FEInterfaceValues} for scalar problems can be found in \texttt{step-12}, \texttt{step-47}, \texttt{step-50}, and the new tutorial \texttt{step-74}.
+Examples of vector-valued problems will be added in the future.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{The source-based toolchain installer \candi{}}
@@
 
-\item \texttt{step-74} SIPG for the Poisson equation \todo[inline]{TODO}
+\item \texttt{step-74} implements the symmetric interior penalty Galerkin 
+(SIPG) method for Poisson's equation using the \texttt{FEInterfaceValues} class
+in combination with \texttt{MeshWorker::mesh\_loop()}. This tutorial demonstrates a
+simple and mesh-independent way to assemble face integrals.
 

### 2021-06-21 9.3/paper.tex
Shorten some text to focus on what is new.

--- 2021-06-21 9.3/paper.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Simplified implementation for face integrals}
+\label{subsec:feinterfacevalues}
+Discontinuous Galerkin (DG) methods -- and other methods with penalty
+terms defined on faces -- require the evaluation of averages and jumps
+accross cell faces, involving values and derivatives of the shape
+functions and solutions from two adjacent cells. The
+\texttt{FEInterfaceValues} class, first introduced in \dealii{} 9.2,
+is designed to provide the necessary interface.
+
+Internally, this class provides an abstraction for two
+\texttt{FEFaceValues} objects (or \texttt{FESubfaceValues} when using
+adaptive refinement). The class introduces new interface degrees of
+freedom indices that are the union of the degrees of freedom indices
+of the two \texttt{FEFaceValues} objects. The interface degrees of
+freedom indices can be converted to the corresponding local degrees of
+freedom indices of the two cells using a helper function. New in the
+current release is better support for vector-valued problems: scalar
+or vector components of shape functions can now be extracted by providing an \texttt{FEValuesExtractors} object.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{The source-based toolchain installer \candi{}}
@@
 
-\item \texttt{step-74} SIPG for the Poisson equation \todo[inline]{TODO}
+\item \texttt{step-74} implements the symmetric interior penalty Galerkin 
+(SIPG) method for Poisson's equation using the \texttt{FEInterfaceValues} class
+within the \texttt{MeshWorker::mesh\_loop()} framework. This tutorial demonstrates a
+simple to assemble face integrals.
 

### 2021-06-21 9.3/paper.tex
More minor edits.

@@
 
-The new \texttt{step-76} tutorial program illustrates this latter case
-in the context of the solution of the Euler equations. \texttt{step-76} reaches a
+The new \texttt{step-76} tutorial program illustrates this case
+in the context of the solution of the Euler equations. It reaches a
 speed-up of 27\% compared to the
@@
 
-To give an example of the new features, let us consider the evaluation of a
+As an example, let us consider the evaluation of a
 surface tension force in the context of sharp-interface methods, whose
@@
 and addition over quadrature points located at the interface section
-$\Gamma_K=\Gamma \cup K$ located on the current cell $K$ and that is,
-in general, located differently within $K$ than for any other cell:
+$\Gamma_K=\Gamma \cup K$ of the current cell $K$ and that is,
+in general, positioned differently within $K$ than for any other cell:
 \begin{align*}
@@
   contributed by David Schneider and Benjamin Uekermann at Technical University
-  of Munich. It shows by extending \texttt{step-4}---a time-dependent Poisson
-  problem~---how to surface-couple a \dealii application with a simple external 
+  of Munich. It extends \texttt{step-4} to a time-dependent Poisson
+  problem and couples it with a simple external surface-based
   application, using the \texttt{preCICE} 
   library~\cite{bungartz2016precice, lindner2020exafsa}. \texttt{preCICE},
-  however, also allows to couple \dealii to actual external simulation softwares, 
-  such as OpenFOAM, SU2, CalculiX, or FEniCS, using provided adapter classes. 
-  For more details visit their webpage
-  \url{https://precice.org/} and check out the numerous \dealii tutorials there,
-  showing, i.a., interesting fluid-structure interaction application cases.
+  also allows to couple \dealii to external simulation packages, 
+  such as OpenFOAM, SU2, CalculiX, or FEniCS, using adapter classes. 
+  The webpage
+  \url{https://precice.org/} provides several more \dealii{}-based tutorials,
+  including interesting fluid-structure interaction applications.
 \end{itemize}

### 2021-06-21 9.3/paper.tex
Another minor edit.

@@
 efficiency of ghost particle exchange significantly. With this feature,
-\dealii can be used to generate scalable parallel code for Lagrangian
+\dealii can be used for scalable parallel Lagrangian
 models such as the Discrete Element Method (DEM) \cite{golshan2021lethedem}

### 2021-06-21 9.3/paper.tex
Remove a TODO.
Fix typo

@@
   necessary $H(\mathrm{div})$-conformity condition is met for all
-  polynomaial orders. This is verified through a new conformity test
+  polynomial orders. This is verified through a new conformity test
   on non-standard meshes.
@@
 
-\todo[inline]{Everyone update. Add support for any additional authors.}
-
 \bibliography{paper}{}

### 2021-06-21 9.3/paper.tex
update funding, unify NSF acknowledgments

@@
 W.~Bangerth, T.~Heister, R.~Gassm{\"o}ller, and J.~Zhang were partially
-supported by the Computational Infrastructure
-for Geodynamics initiative (CIG), through the National Science
-Foundation under Award No.~EAR-1550901 and The
-University of California -- Davis.
-
-W.~Bangerth was also partially supported by award OAC-1835673 as part of the Cyberinfrastructure for Sustained Scientific Innovation (CSSI)
-program, DMS-1821210,
-and EAR-1925595.
-
-B.~Blais was partially supported by the National Science and Engineering Research Council of Canada (NSERC)  through the RGPIN-2020-04510 Discovery Grant
-
-T.~Heister and J.~Zhang were also partially supported by NSF OAC-2015848.
-
-T.~Heister was also partially supported by the NSF awards
-DMS-2028346, EAR-1925575, and by
-Technical Data Analysis, Inc. through US Navy STTR Contract N68335-18-C-0011.
-
-R.~Gassm{\"o}ller was also partially supported by the National Science
-Foundation (NSF) award EAR-1925595.
+supported by the Computational Infrastructure for Geodynamics initiative
+(CIG), through the National Science Foundation (NSF) under Award
+No.~EAR-1550901 and The University of California -- Davis.
+
+W.~Bangerth was also partially supported by Award OAC-1835673 as part of
+the Cyberinfrastructure for Sustained Scientific Innovation (CSSI) program,
+DMS-1821210, and EAR-1925595.
+
+B.~Blais was partially supported by the National Science and Engineering
+Research Council of Canada (NSERC)  through the RGPIN-2020-04510 Discovery
+Grant
+
+T.~Heister and J.~Zhang were also partially supported by NSF
+Award OAC-2015848.
+
+T.~Heister was also partially supported by the NSF Awards DMS-2028346,
+EAR-1925575, and by Technical Data Analysis, Inc. through US Navy STTR
+Contract N68335-18-C-0011.
+
+R.~Gassm{\"o}ller was also partially supported by the NSF Award
+EAR-1925595.
 
@@
 
-M.~Maier was partially supported by ARO MURI Award No. W911NF-14-0247 and
-NSF Award DMS-1912847.
-
-D.~Wells was supported by the National Science Foundation (NSF) through Grant
-DMS-1344962.
+M.~Maier was partially supported by NSF Awards DMS-1912847 and DMS-2045636.
+
+D.~Wells was supported by NSF through Grant DMS-1344962.
 

### 2021-06-21 9.3/paper.tex
Added funding for Marc Fehling.

@@
 
-W.~Bangerth was also partially supported by Award OAC-1835673 as part of
-the Cyberinfrastructure for Sustained Scientific Innovation (CSSI) program,
-DMS-1821210, and EAR-1925595.
+W.~Bangerth and M.~Fehling were partially supported by Award OAC-1835673
+as part of the Cyberinfrastructure for Sustained Scientific Innovation (CSSI)
+program.
+
+W.~Bangerth was also partially supported by Awards DMS-1821210 and EAR-1925595.
 

### 2021-06-21 9.3/paper.tex
Fix two minor typos

--- 2021-06-21 9.3/paper.tex
@@
 The transfer operators between two levels have been implemented in the new class \texttt{MGTwoLevel\allowbreak Transfer}, which can be set up via the functions \texttt{MGTwoLevel\allowbreak Transfer::\allowbreak reinit\_\allowbreak geometric\_\allowbreak transfer()} or \texttt{MGTwo\allowbreak LevelTransfer::\allowbreak reinit\_\allowbreak polynomial\_\allowbreak transfer()} for given
-\texttt{DoFHandler} and \texttt{AffineConstraint} objects
+\texttt{DoFHandler} and \texttt{AffineConstraints} objects
 corresponding to the two levels. The resulting transfer operators
@@
 terms defined on faces -- require the evaluation of averages and jumps
-accross cell faces, involving values and derivatives of the shape
+across cell faces, involving values and derivatives of the shape
 functions and solutions from two adjacent cells. The

### 2021-06-21 9.3/paper.tex
Make some code spacing more consistent.

@@
 for (const auto &cell : dof_handler.active_cell_iterators())
-  switch(cell->reference_cell_type())
+  switch (cell->reference_cell_type())
     {
@@
 \begin{c++}
-for(const auto & cell : tria.active_cell_iterators())
-  for(unsigned int f = 0; f < cell->n_faces(); ++f)
+for (const auto &cell : tria.active_cell_iterators())
+  for (unsigned int f = 0; f < cell->n_faces(); ++f)
     // do something with cell->face(f);
@@
 \begin{c++}
-for(const auto & cell : tria.active_cell_iterators())
-  for(const auto & face : cell->face_iterators())
+for (const auto &cell : tria.active_cell_iterators())
+  for (const auto &face : cell->face_iterators())
     // do something with face
@@
 std::vector<double> local_rhs;
-for(const auto & cell : tria.active_cell_iterators())
+for (const auto &cell : tria.active_cell_iterators())
  {

### 2021-06-21 9.3/paper.tex
Fix another minor mistake.
Only DoFHandler iterators have get_fe().

@@
 std::vector<double> local_rhs;
-for (const auto &cell : tria.active_cell_iterators())
+for (const auto &cell : dof_handler.active_cell_iterators())
  {

### 2021-06-21 9.3/paper.tex
Reword a sentence to avoid spilling into the margins.

@@
 
-The \texttt{FEEvaluation} and \texttt{FEFaceEvaluation} classes use template parameters for
+Both the \texttt{FEEvaluation} and \texttt{FEFaceEvaluation} classes use template parameters for
 the polynomial degree of the finite element $k$ and the number of the 1D quadrature points $q$

### 2021-06-21 9.3/paper.tex
Reword the ReferenceCell section.
'instantiated' isn't quite right here because this is a glorified enum (not a
template class). I had to rearrange a sentence to avoid things spilling into
margins.

@@
 the \texttt{GeometryInfo} class for hypercube-type cells -- is now
-available via the  \texttt{ReferenceCell} class that can be
-instantiated for each of the seven possible reference cell. The
-correct reference cell object for a cell or face can be obtained, respectively, using
-\texttt{cell->reference\_cell()}, and either the call
-\texttt{cell->reference\_cell().face\_reference\_cell(f)}
+available via the \texttt{ReferenceCell} class that is defined
+for each of the seven possible reference cells. The correct
+\texttt{ReferenceCell} for a cell or face can be obtained,
+respectively, using
+\texttt{cell->reference\_cell()}, and either
+\texttt{cell->face(f)->reference\_cell()}
 or
-\texttt{cell->face(f)->reference\_cell()}.
+\texttt{cell->reference\_cell().face\_reference\_cell(f)}.
 

### 2021-06-21 9.3/paper.tex
Fix hbox overflows with better wordings.
Also move the caption below the table.

@@
 \begin{table}
-  \caption{\it List of new scalar \texttt{FiniteElement} classes for the new
-  reference-cell types. Vectorial elements can be constructed based on these classes
-  via \texttt{FE\_Systems}.}\label{tab:simplex:fe}
   \centering
-  \begin{tabular}{cccc}
+  \begin{tabular}{@{} cccc @{}}
     \toprule
@@
   \end{tabular}
+  \caption{\it List of new scalar \texttt{FiniteElement} classes for the new
+  reference-cell types. Vectorial elements can be constructed based on these classes
+  via \texttt{FE\_Systems}.}\label{tab:simplex:fe}
 \end{table}
@@
 
-Alternatively, one can create a pure hypercube mesh with the existing functions
+Alternatively, one can create a pure hypercube mesh with the functions
 in the \texttt{GridGenerator} namespace and convert it to a
@@
 
-With release 9.1, large parts of the $hp$-algorithms
+In release 9.1, large parts of the $hp$-algorithms
 in \dealii were ported to a model that allows for
@@
         \texttt{DoFHandler} class is now capable of all $hp$-functionalities.
-  \item Consequently, all template arguments \texttt{DoFHandlerType} are obsolete,
+  \item Consequently, all template arguments \texttt{DoFHandlerType} are now obsolete,
         and classes like \texttt{DataOut} or \texttt{SolutionTransfer} for example

### 2021-06-22 9.3/paper.tex
Numerous minor changes

@@
 ghost cells around the local domain) by updating their properties instead of
-destroying and rebuilding their container. The latter step improves the
+destroying and rebuilding their container has been added. The latter step improves the
 efficiency of ghost particle exchange significantly. With this feature,
@@
 smoothing algorithms, but have to deal with
-hanging nodes within each level and general require more computational
+hanging nodes within each level and generally require more computational
 work per iteration overall.
@@
 %namespace.
-To facilitate the construction of matrix diagonals with matrix-free methods as well as a matrix representation of the coarse level matrix, new utility functions \texttt{create\_diagonal()} and \texttt{create\_matrix()} from
-the \texttt{MatrixFreeTools} namespace have been added (see also Subsection~\ref{subsec:mf}).
+To facilitate the construction of matrix diagonals with matrix-free methods as well as a matrix representation of the coarse level matrix, new utility functions \texttt{create\_diagonal()} and \texttt{create\_matrix()} have been added to
+the \texttt{MatrixFreeTools} namespace (see also Subsection~\ref{subsec:mf}).
 
@@
 
-In the case that even higher polynomial degrees are needed, one can precompile the
+In the case that even higher polynomial degrees are needed (e.g., $k\le 12$), one can precompile the
 relevant internal classes
-(\texttt{FEEvaluationFactory}, \texttt{FEFaceEvaluationFactory}, \texttt{CellwiseInverseMassFactory}) in the user code for needed \texttt{degree}s
+(\texttt{FEEvaluationFactory}, \texttt{FEFaceEvaluationFactory}, \texttt{CellwiseInverse\allowbreak MassFactory}) in the user code for needed \texttt{degree}s
 and \texttt{VectorizedArrayType}s in the following way:
@@
 
-\item \texttt{step-66} shows how to solve a nonlinear problem using Newton's method in the matrix-free
+\item \texttt{step-66}, a program written by Fabian Castelli at Karlsruhe Institute of Technology, shows how to solve a nonlinear problem using Newton's method in the matrix-free
 framework in parallel. The PDE considered is the Gelfand problem $-\triangle u = \exp(u)$.
@@
 within the \texttt{MeshWorker::mesh\_loop()} framework. This tutorial demonstrates a
-simple to assemble face integrals.
+simple way to assemble face integrals.
 
@@
       serial $hp$-adaptive code for parallelization. Furthermore, it guides through
-      the process of writing an efficient preconditioner for $hp$-adaptive applications.
+      the process of writing an efficient matrix-free preconditioner for $hp$-adaptive applications.
 
@@
 Conrad Clevenger,
+Fabian Castelli,
 David F. Castellanos,
@@
 Ignacio Tomas,
+Benjamin Uekermann,
 Peter Westerbaan.

### 2021-06-22 9.3/paper.tex
Fix order of names.

@@
 
-After having created a triangulation, one can proceed in the same way
-as for hypercube meshes. In particular, one selects an appropriate finite element, mapping, and
+After creating a triangulation, one can proceed in the same way as for hypercube
+meshes. In particular, one selects an appropriate finite element, mapping, and
 quadrature class as follows:
@@
 \end{c++}
-The  list of currently supported finite-element classes is provided in Table~\ref{tab:simplex:fe}. Currently,
+The list of currently supported finite element classes is provided in Table~\ref{tab:simplex:fe}. Currently,
 only linear and quadratic mappings via the \texttt{MappingFE} and
@@
 For mixed meshes, concepts known from the $hp$-context have been applied:
- different finite-element classes are assigned to different cells
+ different finite element classes are assigned to different cells
  based on their respective kind of reference cell. In the
@@
 coordinates \texttt{reference\_\allowbreak points}, is aided by functions like
-\texttt{find\_active\_cell\_around\_point()}
-\texttt{find\_all\_active\_cells\_\allowbreak around\_\allowbreak point()} from the \texttt{GridTools} namespace. While these
+\texttt{find\_active\_cell\_around\_point()} and
+\texttt{find\_all\_active\_\allowbreak cells\_\allowbreak around\_\allowbreak point()} from the \texttt{GridTools} namespace. While these
 functions have been available in \dealii{} previously, their performance has
@@
 Marcus Calhoun-Lopez,
+David F. Castellanos,
 Fabian Castelli, %He provided major contributions to candi for v9.3
@@
 Conrad Clevenger,
-Fabian Castelli,
-David F. Castellanos,
 Andrew Davis,

### 2021-06-23 9.3/paper.tex
Mention that we now require C++14.

--- 2021-06-22 9.3/paper.tex
@@
 
-  \item A long standing orientation issue for the vector valued finite
+  \item A long-standing orientation issue for the vector valued finite
   element \texttt{FE\_RaviartThomas<3>} was resolved. This issue
@@
   on non-standard meshes.
+
+  \item \dealii{} now requires compilers to support the C++14 standard
+    \cite{cpp14}.
 \end{itemize}

### 2021-06-23 9.3/paper.tex
simplex code suggestions

@@
 MappingFE<dim, spacedim> mapping(FE_SimplexP<dim, spacedim>(1));
-QGaussSimplex<dim> quad(degree + 1);
-
-DoFHandler<dim> dof_handler(tria);
+QGaussSimplex<dim> quadrature(degree + 1);
+
+DoFHandler<dim, spacedim> dof_handler(tria);
 dof_handler.distribute_dofs(fe);
 
-FEValues<dim, spacedim> fe_values(mapping, fe, quad, flags);
+FEValues<dim, spacedim> fe_values(mapping, fe, quadrature, flags);
 \end{c++}
@@
 \begin{c++}
-std::vector<double> local_rhs;
+Vector<double> local_rhs;
 for (const auto &cell : dof_handler.active_cell_iterators())
@@
    local_rhs.resize(cell->get_fe().n_dofs_per_cell());
+   // ...
  }

### 2021-06-24 9.3/paper.tex
fix code

@@
    hp_fe_values.reinit(cell);
-   local_rhs.resize(cell->get_fe().n_dofs_per_cell());
+   local_rhs.reinit(cell->get_fe().n_dofs_per_cell());
    // ...

### 2021-06-24 9.3/paper.tex
read through, many smaller changes

@@
 them, and solve partial differential equations with certain finite
-elements, there are also many areas that have not been fully converted
+elements, there are also many areas that have not been fully adapted
 to the new functionality. In particular, \dealii{} currently only
-offers low-order finite elements on such meshes, and many tool
+offers low-order finite elements on such meshes, and many utility
 functions might throw exceptions when used with such meshes.
@@
 The template parameters of the internal data structures of \texttt{Triangulation}
-have been removed and the type of each cell and of each face (only in 3D) is stored. The function
+have been removed and the type of each cell and of each face (in 3D) is stored. The function
 \texttt{Triangulation::create\_\allowbreak triangulation()}, which converts a given list of
 cells and vertices to the internal data structures, has been rewritten
-inspired by \citep{logg2012} and this has had the side effect of a speed-up of up to 5. Minor adjustments
+inspired by \citep{logg2012} and this has had the side effect of a speed-up of up to a factor of 5. Minor adjustments
 have also been made to the \texttt{parallel::shared::Triangulation}
@@
 quadrilaterals, the finite element defined on a triangle (e.g., \texttt{FE\_SimplexP})
-and on a quadrilateral (e.g., \texttt{FE\_Q}) can be collected in a \texttt{hp::FECollection}:
+and on a quadrilateral (e.g., \texttt{FE\_Q}) can be collected in a \texttt{hp::FECollection} as follows:
 
@@
 for the user if these features are to be used. For instance, each cell might have a different type with
-different number of vertices, lines, and faces so that these quantities can not be
-compile-time constants anymore. This information used to be queried from
-the \texttt{GeometryInfo} class. To avoid using this class, we have extended
+different number of vertices, lines, and faces so that these quantities can no longer be 
+compile-time constants. This information used to be queried from
+the \texttt{GeometryInfo} class. Instead, we have extended
 relevant classes, e.g., \texttt{TriaAccessor} or \texttt{TriaCellAccessor},
@@
 given explicitly, defaulted to (bi-/tri-)linear ones. These no longer
-work for simplex or mixed meshes, and in these cases users will need
+work for simplex or mixed meshes, so users will need
 to explicitly provide the correct mapping for the mesh to be used.
@@
 \texttt{FEEvaluation} classes can access vector elements owned by
-other processes and in certain cases node-local
-communication can be skipped. To prevent race conditions, \texttt{MatrixFree} uses local
+other processes and node-local
+communication can be skipped in certain cases. To prevent race conditions, \texttt{MatrixFree} uses local
 barriers at the beginning and the end of loops (\texttt{loop()}, \texttt{cell\_loop()}, \texttt{loop\_cell\_centric()}).
@@
 In a number of circumstances, finite element solutions need to be evaluated at
-arbitrary reference points that change from one element to next. Two important
+arbitrary reference points that change from one element to the next. Two important
 examples are particle simulations coupled to a finite element solution, or
@@
 or vector components of shape functions can now be extracted by providing an \texttt{FEValuesExtractors} object.
+
+See \texttt{step-12}, \texttt{step-47}, \texttt{step-50}, and \texttt{step-74} for more details.
 
@@
         classes with the old interface for a transition period.
+  \item \texttt{GridTools::find\_active\_cell\_around\_point()} no longer throws an exception when no cell if found, but returns an invalid iterator.
+  User codes previously catching an exception will need to be changed.
+
 \end{itemize}

### 2021-06-24 9.3/paper.tex
FEIV example

@@
 is designed to provide the necessary interface.
+
+\newcommand{\jump}[1]{\left[\!\left[ #1 \right]\!\right]}
+\newcommand{\average}[1]{\left\{\!\left\{ #1 \right\}\!\right\}}
+For example, the interface terms of the SIPG formulation for a Laplace problem 
+\[
+\sum_{F} -
+    \left< \jump{v_h}, \average{ \nabla u_h} \cdot  \mathbf n \right>_F
+   -\left<\average{ \nabla v_h }\cdot \mathbf n,\jump{u_h}\right>_F
+   +\left<\jump{v_h}, \sigma \jump{u_h} \right>_F
+\]
+with face jump $\jump{\cdot}$ and average $\average{\cdot}$
+can be implemented as (also see \texttt{step-74}):
+\begin{c++}
+  cell_matrix(i, j) +=
+  (- fe_iv.jump(i, q) * (fe_iv.average_gradient(j, q) * n)
+   - (fe_iv.average_gradient(i, q) * n) * fe_iv.jump(j, q)
+   + penalty * fe_iv.jump(i, q) * fe_iv.jump(j, q)  
+  ) * JxW[q];
+\end{c++}
 

### 2021-06-25 9.3/paper.tex
fix code alignment

@@
   cell_matrix(i, j) +=
-  (- fe_iv.jump(i, q) * (fe_iv.average_gradient(j, q) * n)
+    ( - fe_iv.jump(i, q) * (fe_iv.average_gradient(j, q) * n)
       - (fe_iv.average_gradient(i, q) * n) * fe_iv.jump(j, q)

### 2022-04-10 9.4/paper.tex
Add empty document for 9.4 release

--- 2021-06-24 9.3/paper.tex
@@
     Wolfgang Bangerth,
-    Bruno Blais,
     Marc Fehling,
-    Rene Gassm{\"o}ller
     Timo Heister,
     Luca Heltai,
-    Uwe K{\"o}cher,
     Martin Kronbichler,
@@
     Jean-Paul Pelteret,
-    Sebastian Proell,
-    Konrad Simon,
     Bruno Turcksin,
-    David Wells,
-    Jiaqi Zhang
+    David Wells
   },
-  pdftitle={The deal.II Library, Version 9.3, 2021},
+  pdftitle={The deal.II Library, Version 9.4, 2022},
 }
 
-\title{The \dealii{} Library, Version 9.3}
+\title{The \dealii{} Library, Version 9.4}
 
@@
 
-\author[4]{Bruno Blais}
-\affil[4]{Research Unit for Industrial Flows Processes (URPEI), Department of Chemical Engineering,
-          Polytechnique Montréal,
-          PO Box 6079, Stn Centre-Ville, Montréal, Québec, Canada, H3C 3A7.
-          {\texttt{bruno.blais@polymtl.ca}}}
-
 \author[2]{Marc~Fehling}
-
-\author[5]{Rene~Gassm{\"o}ller}
-\affil[5]{Department of Geological Sciences,
-   University of Florida,
-   1843 Stadium Road,
-   Gainesville, FL, 32611, USA.
-  {\texttt{rene.gassmoeller@ufl.edu}}}
 
@@
    {\texttt{luca.heltai@sissa.it}}}
-
-\author[8]{Uwe~K{\"o}cher}
-\affil[8]{Chair of Numerical Mathematics,
-  Helmut-Schmidt-University,
-  University of the Federal Armed Forces Hamburg,
-  Holstenhofweg~85, 22043 Hamburg, Germany.
-  {\texttt{uwe.koecher@hsu-hh.de}}}
 
@@
 
-\author[9]{Sebastian~Proell}
-
-\author[14]{Konrad Simon}
-  \affil[14]{Department of Mathematics/Center
-  for Earth System Research and Sustainability (CEN), University of
-  Hamburg, Grindelberg 5, 20144 Hamburg, Germany.
-  \texttt{konrad.simon@uni-hamburg.de}}
-
 \author[1*]{Bruno~Turcksin}
@@
   {\texttt{drwells@email.unc.edu}}}
-
-\author[6]{Jiaqi~Zhang}
 
@@
 
-\dealii{} version 9.3.0 was released June 17, 2021.
+\dealii{} version 9.4.0 was released June TODO, 2022.
 This paper provides an
 overview of the new features of this release and serves as a citable
-reference for the \dealii{} software library version 9.3. \dealii{} is an
+reference for the \dealii{} software library version 9.4. \dealii{} is an
 object-oriented finite element library used around the world in the
@@
 \begin{itemize}
-  \item Experimental support for simplex and mixed meshes (see Section~\ref{subsec:simplex});
-  \item Improved flexibility of the particle infrastructure (see Section~\ref{subsec:particles});
-  \item Support for global-coarsening multigrid algorithms (see Section~\ref{subsec:mg});
-  \item Advances in the matrix-free infrastructure  (see Section~\ref{subsec:mf});
-  \item Usage of MPI-3.0 shared-memory features to reduce memory footprint  (see Section~\ref{subsec:sm});
-  \item Improved support for evaluation and integration at arbitrary points (see Section~\ref{subsec:fepointvalues});
-  \item Simplified implementation for face integrals (see Section~\ref{subsec:feinterfacevalues});
-  \item Nine new tutorial programs and a new code gallery program (see Section~\ref{subsec:steps}).
+  \item TODO;
+  \item TODO new tutorial programs and a new code gallery program (see Section~\ref{subsec:steps}).
 \end{itemize}
@@
 \begin{itemize}
-  \item Each non-artificial cell now has a globally unique index that can be queried
-  for active cells via \texttt{CellAccessor::global\_active\_cell\_index()} and for level cells 
-  via \texttt{::global\_\allowbreak level\_\allowbreak cell\_\allowbreak index()}. The information
-  can be used to efficiently index into global vectors storing data
-  for each cell, rather than for each degree of freedom corresponding to a finite
-  element field.
-
-  \item Previously, functions and classes were marked using the 
-  macro \texttt{DEAL\_II\_DEPRECATED} and then typically removed in
-  the release after the one in which these deprecation notices were
-  available to users. We have now extended this policy by introducing
-  the \texttt{DEAL\_II\_\allowbreak DEPRECATED\_EARLY} macro, which indicates that a feature will be
-  deprecated in the next release. In contrast to the first macro, it will only give
-  warnings if \dealii{} has been configured with \texttt{-D DEAL\_II\_EARLY\_DEPRECATIONS=ON}.
-
-  \item After each update of the master branch of \dealii{}, we build a new Docker image
-  with all features enabled. It is
-  accessible on \texttt{Docker Hub} via \texttt{dealii/dealii:master-focal}. This image is
-  particular useful when used in the continuous-integration processes of user codes.
-
-  \item A long-standing orientation issue for the vector valued finite
-  element \texttt{FE\_RaviartThomas<3>} was resolved. This issue
-  prevented the user from using this class on meshes with cells that
-  do not have standard orientation, i.e., cells that have faces that
-  are reflected and/or rotated relative to their neighbors. In all eight
-  possible neighboring configurations, we can now guarantee that the
-  necessary $H(\mathrm{div})$-conformity condition is met for all
-  polynomial orders. This is verified through a new conformity test
-  on non-standard meshes.
-
-  \item \dealii{} now requires compilers to support the C++14 standard
-    \cite{cpp14}.
+  \item TODO.
 \end{itemize}
@@
 can be found
-\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_2_0_and_9_3_0.html}
-{in the file that lists all changes for this release}; see \cite{changes93}.
+\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_3_0_and_9_4_0.html}
+{in the file that lists all changes for this release}; see \cite{changes94}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Experimental simplex and mixed mesh support}
-\label{subsec:simplex}
-
-The current release of \dealii adds experimental support for simplex
-meshes (consisting of triangles in 2D; tetrahedra in 3D) and mixed meshes (consisting of triangles and/or
-quadrilaterals in 2D; tetrahedra, pyramids, wedges, and/or hexahedra in 3D).
-Many freely available mesh-generation tools produce such kind of
-meshes and they are widely used in industry and applications, but were
-previously unsupported by \dealii. As a consequence, users of \dealii
-had to pre-process such meshes and convert them to pure quadrilateral
-or hexahedral meshes.
-
-Support for simplex and mixed meshes is not universal in \dealii{} at
-this point. While \dealii{} can read such meshes, write output for
-them, and solve partial differential equations with certain finite
-elements, there are also many areas that have not been fully adapted
-to the new functionality. In particular, \dealii{} currently only
-offers low-order finite elements on such meshes, and many utility
-functions might throw exceptions when used with such meshes.
-
-A user-focused summary of information around simplex and mixed mesh
-support is also available on the new module page at
-  \url{https://www.dealii.org/current/doxygen/deal.II/group__simplex.html}.
-In particular, it shows how to solve a simple Poisson problem like in
-the \texttt{step-3} tutorial program
-on simplex and mixed meshes, with a focus on the necessary changes to
-the workflow.
-At the time of this release, there are also 92 tests (in the folder \texttt{tests/simplex})
-targeting the new simplex and mixed mesh support. In particular, the folder also
-contains ported variants of a number of existing tutorials: 1, 2, 3, 4, 6, 7, 8, 12, 17, 18, 20, 23, 31, 38,
-40, 55, 67, 68, and 74.
-
-\begin{figure}
-
-  \centering
-
-  \begin{tikzpicture}[scale=1.7]
-
-    \coordinate (0) at (0,0);
-    \coordinate (1) at (1,0);
-    \coordinate (2) at (0,1);
-
-    \coordinate (3) at (0,0.5);
-    \coordinate (4) at (0.5,0.5);
-    \coordinate (5) at (0.5,0);
-
-    \foreach \i in {0,1, 2}
-    \draw[blue,fill=blue] (\i) circle (1.1pt) node [below] {};
-
-    \foreach \i in {3,4,5}
-    \draw (\i) node[cross, draw=red, rotate=-20] {};
-
-    \draw (0) --(1) -- (2) -- (0);
-
-  \end{tikzpicture}
-  \begin{tikzpicture}[scale=1.7]
-
-    \coordinate (0) at (0,0);
-    \coordinate (1) at (1,0);
-    \coordinate (2) at (1,1);
-    \coordinate (3) at (0,1);
-
-    \coordinate (4) at (0.5,0,0);
-    \coordinate (5) at (1.0,0.5);
-    \coordinate (6) at (0.0,0.5);
-    \coordinate (7) at (0.5,1.0);
-    \coordinate (8) at (0.5,0.5);
-
-    \foreach \i in {0,1, 2, 3}
-    \draw[blue,fill=blue] (\i) circle (1.1pt) node [below] {};
-
-    \foreach \i in {4,5,6,7,8}
-    \draw (\i) node[cross, draw=red, rotate=-20] {};
-
-    \draw (0) --(1) -- (2) -- (3) -- (0);
-
-  \end{tikzpicture}
-  \qquad\qquad
-  \begin{tikzpicture}[scale=1.3]
-
-    \coordinate (0) at (0,0,0);
-    \coordinate (1) at (1,0,0);
-    \coordinate (2) at (0,1,0);
-    \coordinate (3) at (0,0,1);
-
-    \coordinate (4) at (0.5,0,0);
-    \coordinate (5) at (0,0.5,0);
-    \coordinate (6) at (0,0,0.5);
-    \coordinate (7) at (0.5,0.5,0.0);
-    \coordinate (8) at (0.5,0.0,0.5);
-    \coordinate (9) at (0.0,0.5,0.5);
-
-    \foreach \i in {0,1,2,3}
-    \draw[blue,fill=blue] (\i) circle (1.5pt) node [below] {};
-
-    \foreach \i in {4,5,6,7,8,9}
-    \draw (\i) node[cross, draw=red, rotate=-20] {};
-
-    \draw (0) -- (1) -- (2) -- (0);
-    \draw (0) -- (1) -- (3) -- (0);
-    \draw (1) -- (2) -- (3) -- (1);
-
-  \end{tikzpicture}
-  \begin{tikzpicture}[scale=1.5]
-
-    \coordinate (0) at (0.0, 0.0, 0.0);
-    \coordinate (1) at (0.0, 0.0, 1.0);
-    \coordinate (2) at (1.0, 0.0, 0.0);
-    \coordinate (3) at (1.0, 0.0, 1.0);
-    \coordinate (4) at (0.5, 1.0, 0.5);
-
-    \coordinate  (5) at (0.5, 0.0, 0.0);
-    \coordinate  (6) at (0.0, 0.0, 0.5);
-    \coordinate  (7) at (0.5, 0.0, 1.0);
-    \coordinate  (8) at (1.0, 0.0, 0.5);
-
-    \coordinate  (9) at (0.25,0.5,0.25);
-    \coordinate (10) at (0.25,0.5,0.75);
-    \coordinate (11) at (0.75,0.5,0.25);
-    \coordinate (12) at (0.75,0.5,0.75);
-
-    \coordinate (13) at (0.5, 0.0, 0.5);
-
-    \foreach \i in {0,1,2,3,4}
-    \draw[blue,fill=blue] (\i) circle (1.3pt) node [below] {};
-
-    %\foreach \i in {5, 6, 7, 8, 9, 10, 11, 12, 13}
-    %    \draw (\i) node[cross] {};
-
-    \draw (0) -- (1) -- (3) -- (2) -- (0);
-    \draw (1) -- (4) -- (3);
-    \draw (0) -- (4) -- (2);
-
-  \end{tikzpicture}
-  \begin{tikzpicture}[scale=1.3]
-
-    \coordinate (0) at (0,0,1.5);
-    \coordinate (1) at (1,0,1.5);
-    \coordinate (2) at (0.0,1.0,1.5);
-
-    \coordinate (3) at (0,0,0);
-    \coordinate (4) at (1,0,0);
-    \coordinate (5) at (0.0,1.0,0);
-
-    \coordinate  (6) at (0.5, 0.0, 0.0);
-    \coordinate  (7) at (0.0, 0.5, 0.0);
-    \coordinate  (8) at (0.5, 0.5, 0.0);
-
-    \coordinate   (9) at (0.5, 0.0, 0.75);
-    \coordinate  (10) at (0.0, 0.5, 0.75);
-    \coordinate  (11) at (0.5, 0.5, 0.75);
-
-    \coordinate  (12) at (0.5, 0.0, 1.5);
-    \coordinate  (13) at (0.0, 0.5, 1.5);
-    \coordinate  (14) at (0.5, 0.5, 1.5);
-
-    \coordinate (15) at (0,0,0.75);
-    \coordinate (16) at (1,0,0.75);
-    \coordinate (17) at (0.0,1.0,0.75);
-
-    \foreach \i in {0,1,2,3,4,5}
-    \draw[blue,fill=blue] (\i) circle (1.5pt) node [below] {};
-
-    \foreach \i in {6,7,8,9,10,11,12,13,14,15,16,17}
-    \draw (\i) node[cross, draw=red, rotate=-20] {};
-
-    \draw (0) -- (1) -- (2) -- (0);
-    \draw (3) -- (4) -- (5) -- (3);
-    \draw (0) -- (3);
-    \draw (1) -- (4);
-    \draw (2) -- (5);
-
-  \end{tikzpicture}
-  \begin{tikzpicture}[scale=1.5]
-
-    \coordinate (0) at (0,0,1);
-    \coordinate (1) at (1,0,1);
-    \coordinate (2) at (0,0,0);
-    \coordinate (3) at (1,0,0);
-
-    \coordinate (4) at (0,1,1);
-    \coordinate (5) at (1,1,1);
-    \coordinate (6) at (0,1,0);
-    \coordinate (7) at (1,1,0);
-
-    \coordinate  (8) at (0,0,0.5);
-    \coordinate  (9) at (1,0,0.5);
-    \coordinate (10) at (0.5,0,1);
-    \coordinate (11) at (0.5,0,0);
-    \coordinate (12) at (0.5,0,0.5);
-
-    \coordinate (13) at (0,1,0.5);
-    \coordinate (14) at (1,1,0.5);
-    \coordinate (15) at (0.5,1,1);
-    \coordinate (16) at (0.5,1,0);
-    \coordinate (17) at (0.5,1,0.5);
-
-    \coordinate (18) at (0,0.5,1);
-    \coordinate (19) at (1,0.5,1);
-    \coordinate (20) at (0,0.5,0);
-    \coordinate (21) at (1,0.5,0);
-    \coordinate (22) at (0,0.5,0.5);
-    \coordinate (23) at (1,0.5,0.5);
-    \coordinate (24) at (0.5,0.5,1);
-    \coordinate (25) at (0.5,0.5,0);
-    \coordinate (26) at (0.5,0.5,0.5);
-
-    \foreach \i in {0,1,2,3,4,5,6,7}
-    \draw[blue,fill=blue] (\i) circle (1.3pt) node [below] {};
-
-    \foreach \i in {8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26}
-    \draw (\i) node[cross, draw=red, rotate=-20] {};
-
-    \draw (0) -- (1) -- (3) -- (2) -- (0);
-    \draw (4) -- (5) -- (7) -- (6) -- (4);
-    \draw (0) -- (4);
-    \draw (1) -- (5);
-    \draw (2) -- (6);
-    \draw (3) -- (7);
-
-  \end{tikzpicture}
-
-  \caption{\it Reference cells in 2D (triangle, quadrilateral) and 3D
-    (tetrahedron, pyramid, wedge, hexahedron) with the support points
-    indicated both for linear ({\color{blue}$\bullet$}) and quadratic
-    ({\color{red}$\times$}) shape functions.
-  %\todo[inline]{The pyramid is missing red crosses. The wedge is
-  %  missing them on three edges.}
-  % PM: we only have linear pyramids for now.
-  }
-  \label{fig:simplex}
-\end{figure}
-
-\begin{table}
-  \centering
-  \begin{tabular}{@{} cccc @{}}
-    \toprule
-    \textbf{reference cell} & \textbf{finite element} & \textbf{dimension} & \textbf{degree}\\
-    \midrule
-    \multirow{2}{*}{simplex (line, triangle, tetrahedron)} & FE\_SimplexP, FE\_SimplexDGP  & 1--3 & 1--2 \\
-                                          & FE\_SimplexP\_Bubbles  & 1--3 & 1--2 \\
-    pyramid & FE\_PyramidP, FE\_PyramidDGP  & 3 & 1 \\
-    wedge & FE\_WedgeP, FE\_WedgeDGP  & 3 & 1--2 \\
-    \bottomrule
-  \end{tabular}
-  \caption{\it List of new scalar \texttt{FiniteElement} classes for the new
-  reference-cell types. Vectorial elements can be constructed based on these classes
-  via \texttt{FE\_Systems}.}\label{tab:simplex:fe}
-\end{table}
-
-\subsubsection{Refactoring of internal data structures}
-
-To enable simplex and mixed mesh support, we performed a major refactoring of the internal data structures of \dealii. In particular, the \texttt{Triangulation} class and
-the \texttt{DoFHandler} class have undergone large changes and now
-support meshes composed of all of the cells shown in Fig.~\ref{fig:simplex}.
-
-The template parameters of the internal data structures of \texttt{Triangulation}
-have been removed and the type of each cell and of each face (in 3D) is stored. The function
-\texttt{Triangulation::create\_\allowbreak triangulation()}, which converts a given list of
-cells and vertices to the internal data structures, has been rewritten
-inspired by \citep{logg2012} and this has had the side effect of a speed-up of up to a factor of 5. Minor adjustments
-have also been made to the \texttt{parallel::shared::Triangulation}
-and \texttt{parallel::\allowbreak fullydistributed::\allowbreak
-  Triangulation} classes such that
-the new mesh types can also be used in parallel.
-
-The internal data structures of \texttt{DoFHandler} used to be hard-coded for pure
-hypercube meshes, while the \texttt{hp::DoFHandler} used to be built around
-CRS-like data structures. Due to the need of CRS data structures in the
-\texttt{DoFHandler} in the context of more general meshes, we have merged
-\texttt{hp::DoFHandler} into the \texttt{DoFHandler}. The class \texttt{hp::DoFHandler}, which is now a dummy derivation of \texttt{DoFHandler}, currently only exists for compatibility reasons and has
-been deprecated, see Section~\ref{subsec:deprecated}. It will be
-removed in the next release.
-
-\subsubsection{Generating meshes}
-
-The most obvious way to generate a simplex or a mixed mesh is to read the mesh from a file
-generated by an external mesh generator. Currently, we support the
-\texttt{VTK}, \texttt{MSH} (generated by the \textsc{gmsh} program \cite{geuzaine2009gmsh}), and \texttt{EXODUS II} file formats.
-
-Alternatively, one can create a pure hypercube mesh with the functions
-in the \texttt{GridGenerator} namespace and convert it to a
-pure simplex mesh with the function
-\texttt{GridGenerator::convert\_\allowbreak hypercube\_\allowbreak to\_\allowbreak simplex\_\allowbreak mesh()}.
-
-\subsubsection{Using simplex meshes}
-
-After creating a triangulation, one can proceed in the same way as for hypercube
-meshes. In particular, one selects an appropriate finite element, mapping, and
-quadrature class as follows:
-
-\begin{c++}
-FE_SimplexP<dim, spacedim> fe(degree);
-MappingFE<dim, spacedim> mapping(FE_SimplexP<dim, spacedim>(1));
-QGaussSimplex<dim> quadrature(degree + 1);
-
-DoFHandler<dim, spacedim> dof_handler(tria);
-dof_handler.distribute_dofs(fe);
-
-FEValues<dim, spacedim> fe_values(mapping, fe, quadrature, flags);
-\end{c++}
-The list of currently supported finite element classes is provided in Table~\ref{tab:simplex:fe}. Currently,
-only linear and quadratic mappings via the \texttt{MappingFE} and
-\texttt{MappingFEField} classes built around the listed nodal elements are available.
-%\todo[inline]{I don't understand the previous sentence. "linear" means
-%  p=1, but isoparametric means that it has the same p as the finite
-%  element. It can't be both.}
-For quadrature, the classes \texttt{QDuffy}, \texttt{QGaussSimplex}, \texttt{QWitherdenVincentSimplex} \cite{schloemer21, witherden2015identification},
-\texttt{QGaussPyamid}, and \texttt{QGaussWedge} are
-available.
-
-\subsubsection{Using mixed meshes}
-For mixed meshes, concepts known from the $hp$-context have been applied:
- different finite element classes are assigned to different cells
- based on their respective kind of reference cell. In the
-case of a 2D mixed mesh, which can only consist of triangles and
-quadrilaterals, the finite element defined on a triangle (e.g., \texttt{FE\_SimplexP})
-and on a quadrilateral (e.g., \texttt{FE\_Q}) can be collected in a \texttt{hp::FECollection} as follows:
-
-\begin{c++}
-hp::FECollection<dim, spacedim> fe
-  {FE_SimplexP<dim, spacedim>(degree), FE_Q<dim, spacedim>(degree)};
-\end{c++}
-
-Similarly, \texttt{hp::QCollection} and \texttt{hp::MappingCollection} can be used
-to construct appropriate collections. Furthermore, the correct active
-finite element index, which points to the correct
-finite element of that cell, has to be assigned to each cell. The
-following piece of code will then correctly enumerate all degrees of
-freedom on the mesh:
-
-\begin{c++}
-DoFHandler<dim> dof_handler(tria);
-
-for (const auto &cell : dof_handler.active_cell_iterators())
-  switch (cell->reference_cell_type())
-    {
-      case ReferenceCell::Type::Tri:  cell->set_active_fe_index(0); break;
-      case ReferenceCell::Type::Quad: cell->set_active_fe_index(1); break;
-      // 3D (Tet, Pyramid, Wedge, Hex) not shown
-      default: Assert(false, ExcNotImplemented());
-    }
-
-dof_handler.distribute_dofs(fe);
-\end{c++}
-
-\subsubsection{Practical implications}
-
-The introduction of simplex and mixed meshes leads to some implications
-for the user if these features are to be used. For instance, each cell might have a different type with
-different number of vertices, lines, and faces so that these quantities can no longer be 
-compile-time constants. This information used to be queried from
-the \texttt{GeometryInfo} class. Instead, we have extended
-relevant classes, e.g., \texttt{TriaAccessor} or \texttt{TriaCellAccessor},
-with useful new functions like \texttt{n\_vertices()}, \texttt{n\_lines()}, or
-\texttt{n\_faces()} so that users can simply write:
-\begin{c++}
-for (const auto &cell : tria.active_cell_iterators())
-  for (unsigned int f = 0; f < cell->n_faces(); ++f)
-    // do something with cell->face(f);
-\end{c++}
-Alternatively, one can use an iterator-based approach to loop over all faces
-of a cell, introduced in the previous release. The relevant functions have been adjusted to be able to deal with
-the now variable number of faces per cell:
-\begin{c++}
-for (const auto &cell : tria.active_cell_iterators())
-  for (const auto &face : cell->face_iterators())
-    // do something with face
-\end{c++}
-Furthermore, for mixed meshes, the number of degrees of freedom will differ between cells so that
-cell-local arrays need to be resized for each
-cell (as has previously already been the case in the $hp$-context):
-\begin{c++}
-Vector<double> local_rhs;
-for (const auto &cell : dof_handler.active_cell_iterators())
- {
-   hp_fe_values.reinit(cell);
-   local_rhs.reinit(cell->get_fe().n_dofs_per_cell());
-   // ...
- }
-\end{c++}
-
-What is true for cells is also true for faces in 3D: faces can be either triangles
-or quadrilaterals. This is the case even if no mixed mesh is used if
-the mesh consists exclusively of pyramids or wedges.
-As a consequence, some functions, e.g.
-\texttt{FiniteElementData::n\_dofs\_per\_face()}, have been extended with a new optional argument for the face number.
-
-Geometric information about cells and faces -- previously provided by
-the \texttt{GeometryInfo} class for hypercube-type cells -- is now
-available via the \texttt{ReferenceCell} class that is defined
-for each of the seven possible reference cells. The correct
-\texttt{ReferenceCell} for a cell or face can be obtained,
-respectively, using
-\texttt{cell->reference\_cell()}, and either
-\texttt{cell->face(f)->reference\_cell()}
-or
-\texttt{cell->reference\_cell().face\_reference\_cell(f)}.
-
-Furthermore, many functions in \dealii{} used mappings that, when not
-given explicitly, defaulted to (bi-/tri-)linear ones. These no longer
-work for simplex or mixed meshes, so users will need
-to explicitly provide the correct mapping for the mesh to be used.
-
-\subsubsection{Matrix-free support}
-
-\dealii{}'s matrix-free support has also been extended to simplex and
-mixed meshes, for both
-continuous and discontinuous elements. From a user perspective, the main
-changes are to pass a $d$-dimensional quadrature object (rather than
-one for 1D),
-and the fact that \texttt{FEEvaluation} and \texttt{FEFaceEvaluation} must not
-specify the polynomial degree via template arguments, determining all
-information at runtime. More details can be found in
-Section~\ref{subsec:mf}.
-
-For the current release, no advanced algorithms for evaluating values and
-gradients at quadrature points, such as sum factorization, are used.
-The use of full interpolation matrices is for now acceptable since only low-order elements are
-supported.
+\subsection{Topic 1}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Advances in the particle infrastructure}
-\label{subsec:particles}
-
-\dealii 's particle infrastructure has been modernized to store nearly all
-particle data as a collection of continuous data arrays instead of a container of
-objects. This reorganization improves cache efficiency when iterating over
-particles and reduces the amount of data that needs to be moved when a particle
-moves to a different cell. Keeping track of unused data slots allows to reuse
-them for new particles, significantly reducing memory allocations when
-particles are created after other particles have left the local domain. The
-data arrays are rebuilt and sorted during every mesh refinement cycle.
-
-Additionally, the particle infrastructure now supports a faster search algorithm
-for particles that moved farther than one cell width between particle sorting
-operations, and support for updating ghost particles (particles that live in
-ghost cells around the local domain) by updating their properties instead of
-destroying and rebuilding their container has been added. The latter step improves the
-efficiency of ghost particle exchange significantly. With this feature,
-\dealii can be used for scalable parallel Lagrangian
-models such as the Discrete Element Method (DEM) \cite{golshan2021lethedem}
-or Molecular Dynamics.
-
-Finally, the \texttt{Particles::DataOut} class now supports writing particle
-properties as vectors or tensors instead of a collection of scalars if the properties are
-marked as such.
+\subsection{Topic 2}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Advances in the multigrid infrastructure}
-\label{subsec:mg}
-
-Until now, \dealii has only supported ``local smoothing'' multigrid
-algorithms~\citep{ClevengerHeisterKanschatKronbichler2019} wherein
-smoothers only act on the cells of a given refinement level,
-skipping those parts of the mesh that have not been adaptively refined
-to that level. This approach guarantees that the work done summed up
-over all levels is proportional to the number of unknowns, and is
-consequently necessary so that the overall multigrid preconditioner
-can have a complexity of ${\cal O}(N)$.
-
-In contrast, the current release now also supports ``global
-coarsening'' algorithms \citep{becker2000multigrid, sundar2012parallel} when
-using continuous (\texttt{FE\_Q}, \texttt{FE\_SimplexP}) and
-discontinuous (\texttt{FE\_DGQ}, \texttt{FE\_SimplexDGP})
-elements. Global coarsening builds multigrid levels for the entire
-domain, coarsening \textit{all} cells of a triangulation regardless of
-how many times they have been refined. In addition, the framework now available in
-\dealii{} is not only applicable to geometric coarsening, but can also
-perform coarsening by reducing the polynomial degree ($p$-multigrid,
-see~\cite{Fehn2020}), for example
-to support $hp$-adaptive meshes.
-%where it has been shown that one can
-%efficiently precondition high-order spaces with lower-order finite
-%element operators.
-% MK: I do not think "precondition" is a precise enough description of p-multigrid
-% because one still needs smoothing on the fine levels in addition to the
-% "solve" at the coarser degree, a fact not transported by the mere
-% 'precondition' term; rather than trying to discuss this in a more precise way,
-% let us simply skip it as that is not central to the presentation here.
-Finally, the implementation also supports transfer between continuous
-and discontinuous elements as a further way to create multigrid levels.
-
-These new multigrid variants promise fewer solver
-iterations and better parallel scalability than the existing local
-smoothing algorithms, but have to deal with
-hanging nodes within each level and generally require more computational
-work per iteration overall.
-
-The transfer operators between two levels have been implemented in the new class \texttt{MGTwoLevel\allowbreak Transfer}, which can be set up via the functions \texttt{MGTwoLevel\allowbreak Transfer::\allowbreak reinit\_\allowbreak geometric\_\allowbreak transfer()} or \texttt{MGTwo\allowbreak LevelTransfer::\allowbreak reinit\_\allowbreak polynomial\_\allowbreak transfer()} for given
-\texttt{DoFHandler} and \texttt{AffineConstraints} objects
-corresponding to the two levels. The resulting transfer operators
-can then be collected in a single
-\texttt{MGTransfer\allowbreak GlobalCoarsening} object that can be used just as the previous workhorse \texttt{MGTransferMatrixFree} within the \texttt{Multigrid}
-algorithm.
-%Several common operations are encoded in utility
-%functions in the \texttt{MGTransferGlobalCoarseningTools}
-%namespace.
-To facilitate the construction of matrix diagonals with matrix-free methods as well as a matrix representation of the coarse level matrix, new utility functions \texttt{create\_diagonal()} and \texttt{create\_matrix()} have been added to
-the \texttt{MatrixFreeTools} namespace (see also Subsection~\ref{subsec:mf}).
-
-The usage of the new transfer operators (and of some of the utility
-functions) in the context of a hybrid multigrid algorithm
-($hp$-multigrid with algebraic multigrid as coarse-grid solver) for $hp$-problems is demonstrated in the new tutorial \texttt{step-75}, see also Section~\ref{subsec:steps}.
+\subsection{Topic 3}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Advances in the matrix-free infrastructure}
-\label{subsec:mf}
-
-\dealii{}'s matrix-free framework enables high-throughput operations
-for applications in which only the availability of the \textit{action}
-of a matrix, but not the entries of the matrix, are necessary. This
-framework has been substantially extended in the current release.
-
-\subsubsection{Precompilation of evaluation kernels}
-
-Both the \texttt{FEEvaluation} and \texttt{FEFaceEvaluation} classes use template parameters for
-the polynomial degree of the finite element $k$ and the number of the 1D quadrature points $q$
-to generate near-optimal code for these operations.
-For application codes that rely on operators of many different degrees (e.g., because
-they use $p$-multigrid or $hp$-algorithms), creating all instantiations
-can be overly complex and incur long compile times.
-
-In the current release, specializations of these classes that do
-not rely on the template parameters $k$ and $q$ (expressed in the code
-using special values ``-1'' and ``0'') have been added. For example:
-\begin{c++}
-FEEvaluation<dim, -1, 0, n_components, Number, VectorizedArrayType>
-  phi(range, dofhandler_index, quadrature_index, first_selected_component);
-\end{c++}
-These classes select at runtime -- for common low and medium
-polynomial degree/quadrature combinations ($k\le 6$ and $q\in\{ k+1,
-k+2, \left\lfloor (3k)/2 \right\rfloor \}$) -- efficient precompiled implementations and default to non-templated
-evaluation kernels otherwise (see also \texttt{FEEvaluation::fast\_evaluation\_\allowbreak supported()}).
-
-In the case that even higher polynomial degrees are needed (e.g., $k\le 12$), one can precompile the
-relevant internal classes
-(\texttt{FEEvaluationFactory}, \texttt{FEFaceEvaluationFactory}, \texttt{CellwiseInverse\allowbreak MassFactory}) in the user code for needed \texttt{degree}s
-and \texttt{VectorizedArrayType}s in the following way:
-\begin{c++}
-#define FE_EVAL_FACTORY_DEGREE_MAX 12
-
-#include <deal.II/matrix_free/evaluation_template_factory.templates.h>
-
-DEAL_II_NAMESPACE_OPEN
-template struct dealii::internal::FEEvaluationFactory
-  <dim, VectorizedArrayType::value_type, VectorizedArrayType>;
-// same for FEFaceEvaluationFactory and CellwiseInverseMassFactory (skipped)
-DEAL_II_NAMESPACE_CLOSE
-\end{c++}
-
-\subsubsection{Parallel matrix-free $hp$-implementation}\label{subsubsection:mf:hp}
-
-In release 9.1, large parts of the $hp$-algorithms
-in \dealii were ported to a model that allows for
-parallel matrix-based simulations \cite{dealII91}. In the present
-release, the parallel $hp$ support was extended to \texttt{MatrixFree}.
-
-Until now, the \texttt{FEEvaluation} classes used the template
-parameters $k$ and $q$ to select the correct active FE and quadrature
-index and users were responsible for the cumbersome detection of 
-subranges of the same $k$ and $q$ within the cell ranges returned by the 
-matrix-free loops.
-%\todo{I don't understand what ``detection of subranges of the same
-%  degree'' refers to. Please clarify. In particular, I don't know what
-%``subranges'' mean in this paragraph.}
-The creation of subranges is now performed internally, and the non-templated versions
-of the \texttt{FEEvaluation} classes have been extended for the $hp$-case. To determine
-the desired FE and quadrature index of a subrange, the current cell/face range has to be provided
-to the constructors of the \texttt{FEEvaluation} classes. These changes enable
-users to write matrix-free code independently of whether $hp$-capabilities are used or not.
-
-The new tutorial \texttt{step-75} presents how to use the new $hp$-related features in \texttt{MatrixFree}
-in the context of a hybrid-multigrid solver.
-
-\subsection{MPI-3.0 shared-memory support}\label{subsec:sm}
-In many large computations, certain pieces of data are computed
-(or read from disk) once and then treated as read-only. If this information is needed by more
-than one MPI process, it is more efficient to store this information only
-once in shared memory among all processes located on a multicore
-node. MPI supports the creation of such shared memory windows since
-version 3.0, and deal.II can now use this in the
-\texttt{AlignedVector} and \texttt{Table} classes that are often used
-for large lookup tables for classes such as \texttt{InterpolatedTensorProductGridData}.
-
-Shared memory storage is also used in the \texttt{MatrixFree} and
-\texttt{LinearAlgebra::\allowbreak distributed::\allowbreak Vector} classes. If \texttt{MatrixFree} has been configured by setting
-\texttt{MatrixFree::\allowbreak AdditionalData::\allowbreak communicator\_sm} appropriately, then
-\texttt{MatrixFree::create\_dof\_vector()} creates vectors that share
-information among all processes on one node. As a consequence, the
-\texttt{FEEvaluation} classes can access vector elements owned by
-other processes and node-local
-communication can be skipped in certain cases. To prevent race conditions, \texttt{MatrixFree} uses local
-barriers at the beginning and the end of loops (\texttt{loop()}, \texttt{cell\_loop()}, \texttt{loop\_cell\_centric()}).
-
-The new \texttt{step-76} tutorial program illustrates this case
-in the context of the solution of the Euler equations. It reaches a
-speed-up of 27\% compared to the
-original version, \texttt{step-67}, by using the new feature.
-For more details and the usage of the feature in the library \texttt{hyper.deal}, see \citep{munch2020hyperdeal}.
+\subsection{Topic 4}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Evaluation and integration at arbitrary points}
-\label{subsec:fepointvalues}
-
-In a number of circumstances, finite element solutions need to be evaluated at
-arbitrary reference points that change from one element to the next. Two important
-examples are particle simulations coupled to a finite element solution, or
-algorithms on non-matching grids. The existing \texttt{FEValues} class is a poor
-fit for this task, as it is based on the assumption that evaluation of
-shape functions and their derivatives happens at the same quadrature
-points on every cell, and that consequently expensive computations can
-be done once and results re-used for many subsequent cells. The new class
-\texttt{FEPointEvaluation} provides a more convenient interface for
-cases where the evaluation on different cells does not happen at the
-same points mapped from the reference cell.
-For tensor product finite elements (\texttt{FE\_Q}, \texttt{FE\_DGQ})
-and tensor product mappings (\texttt{MappingQGeneric} and derived classes),
-the new approach is also very fast, as it can use some of the matrix-free
-infrastructure and vectorization facilities.
-
-As an example, let us consider the evaluation of a
-surface tension force in the context of sharp-interface methods, whose
-contribution is added to a fluid solver by multiplication with test function
-and addition over quadrature points located at the interface section
-$\Gamma_K=\Gamma \cup K$ of the current cell $K$ and that is,
-in general, positioned differently within $K$ than for any other cell:
-\begin{align*}
-\left(\mathbf{v}, \kappa \, \mathbf{n}\right)_\Gamma
-\approx
-\sum_q \mathbf{v}(\mathbf{x}_q) \cdot \left(\kappa(\mathbf{x}_q) \, \mathbf{n}(\mathbf{x}_q)\right) (JxW)_q.
-\end{align*}
-In \dealii{}, this can now be conveniently written as
-\begin{c++}
-phi_curvature.reinit(cell, reference_points);
-phi_normal.reinit(cell, reference_points);
-phi_force.reinit(cell, reference_points);
-
-phi_curvature.evaluate(curvature_values, EvaluationFlags::values);
-phi_normal.evaluate(normal_values, EvaluationFlags::values);
-
-for (unsigned int q = 0; q < n_points; ++q)
-  phi_force.submit_value(phi_curvature.get_value(q) *
-                         phi_normal.get_value(q) * JxW[q], q);
-
-phi_force.integrate(force_values, EvaluationFlags::values);
-\end{c++}
-The quadrature points (at reference positions
-\texttt{reference\_points}) and the related \texttt{JxW} value can,
-for example, come from
-a mesh of lower dimension. Determining to which \texttt{cell} a quadrature
-point belongs to on the background mesh, including the reference cell
-coordinates \texttt{reference\_\allowbreak points}, is aided by functions like
-\texttt{find\_active\_cell\_around\_point()} and
-\texttt{find\_all\_active\_\allowbreak cells\_\allowbreak around\_\allowbreak point()} from the \texttt{GridTools} namespace. While these
-functions have been available in \dealii{} previously, their performance has
-been considerably enhanced with the aforementioned more optimized code paths
-for selected mappings.
-
-While \texttt{FEPointEvaluation} assumes that evaluation points are already sorted according to 
-the owning cells and thus can focus on cell-local operations, the new class
-\texttt{RemotePointEvaluation} is responsible for determining the owning cells in a distributed
-context and for providing efficient communication patterns for the data exchange. In \texttt{deal.II},
-the class has been successfully applied together with \texttt{FEPointEvaluation} to evaluate a distributed
-solution vector at arbitrary points (see also the
-\texttt{VectorTools::point\_values()} function).
+\subsection{Topic 5}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Simplified implementation for face integrals}
-\label{subsec:feinterfacevalues}
-Discontinuous Galerkin (DG) methods -- and other methods with penalty
-terms defined on faces -- require the evaluation of averages and jumps
-across cell faces, involving values and derivatives of the shape
-functions and solutions from two adjacent cells. The
-\texttt{FEInterfaceValues} class, first introduced in \dealii{} 9.2,
-is designed to provide the necessary interface.
-
-Internally, this class provides an abstraction for two
-\texttt{FEFaceValues} objects (or \texttt{FESubfaceValues} when using
-adaptive refinement). The class introduces new interface degrees of
-freedom indices that are the union of the degrees of freedom indices
-of the two \texttt{FEFaceValues} objects. The interface degrees of
-freedom indices can be converted to the corresponding local degrees of
-freedom indices of the two cells using a helper function. New in the
-current release is better support for vector-valued problems: scalar
-or vector components of shape functions can now be extracted by providing an \texttt{FEValuesExtractors} object.
-
-See \texttt{step-12}, \texttt{step-47}, \texttt{step-50}, and \texttt{step-74} for more details.
+\subsection{Topic 6}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The source-based toolchain installer \candi{}}
-\label{subsec:candi}
-
-The requirement to download, compile, and install \dealii{} and its dependencies from
-source is a major obstacle to many \dealii{} users.
-Compiling all dependencies from source can be difficult but is a necessity on operating systems for which
-binary packages aren't available or on compute clusters and other machines
-without root privileges for the user to install system dependencies.
-
-The source based installation of \dealii{} and its many dependent libraries
-can be done with the \candi{} script tool for various
-Linux operating systems, within the Windows Subsystem Linux (WSL), and
-on OS X (experimental).
-%
-The general assumption is that a C, C++ and Fortran compiler and suitable MPI-compilers
-for the base compilers as well as the corresponding development system packages are
-available.
-
-\candi is a bash-script based tool, is an abbreviation of ``compile and install''
-and is released under the GNU LGPL v3.0.
-%
-The origin of \candi{} is a fork made in 2013 from \texttt{dorsal},
-a now-retired source based installer for the \texttt{FEniCS} library.
-%
-The \candi{} tool can be found via the download page of \dealii{} since version 8.5
-or from \url{github.com/dealii/candi}, and is under active development.
-To install older releases of \dealii{} toolchains, one can check out
-the \candi{} branch of the \texttt{git} repository that
-corresponds to the desired \dealii{} version.
-
-\candi{} downloads, unpacks, compiles, and installs an individual
-library (called a ``package'' in \candi{}), a list of libraries, or a complete toolchain.
-The toolchain installation is the default behavior and the default configuration
-ensures that most of the \dealii{} step tutorials can be used directly.
-The package installation mode is also useful to generate docker containers.
-In toolchain mode, \candi{} checks for and deals with
-dependencies between libraries appropriately.
-
-Each package for a library is defined by variables for its name and version,
-a (remote or local) download location,
-its packaging format (e.g. tar, gz, or zip),
-a checksum,
-its general build chain (e.g., using \texttt{autotools}, \texttt{cmake}, or others),
-as well as configuration options.
-Moreover, one can give specific instructions for each of the steps being
-necessary for the unpacking, configuration, or building of a package.
-A new feature allows \candi{} to skip user prompts, allowing its use in
-a batch mode.
-
-\candi{} does all this by either downloading packages from the
-internet, including through mirrors, or re-using previously downloaded
-or checked out from a repository.
-If temporary files of \candi{} are not removed, one can use the developer
-mode to prepare patches or developments for any of the packages.
-
-\candi{} is organized in the following way:
-\begin{itemize}
-\item \texttt{candi.sh}: the bash script which controls the overall process.
-Available command line options are listed by calling ``\texttt{candi.sh -h}''
-and explained in the \texttt{README.md} file.
-
-\item \texttt{candi.cfg} (or \texttt{local.cfg}): the default toolchain
-configuration file. Here one can easily switch on/off features,
-give additional configuration, and
-list the packages for a toolchain.
-
-\item \texttt{deal.II-toolchain} folder: the central project folder for
-\dealii{}-based toolchains.
-It contains subfolders
-for packages (libraries, dependencies, general tools in a specific version),
-platforms (operating systems) and
-for patches to be applied after unpacking a package.
-\end{itemize}
-
-The installer is developed in the \texttt{dealii/candi}
-\texttt{github} repository, and uses continuous
-integration (CI) to ensure that old and new features
-are working as expected.
+\subsection{Topic 7}
 
@@
 \begin{itemize}
-\item \texttt{step-19} is an introductory demonstration of \dealii{}'s
-  particle functionality. It solves the coupled problem of
-  charged particles and an electric field, using a cathode tube as an
-  example.
+\item \texttt{TODO} TODO.
 
-\item \texttt{step-66}, a program written by Fabian Castelli at Karlsruhe Institute of Technology, shows how to solve a nonlinear problem using Newton's method in the matrix-free
-framework in parallel. The PDE considered is the Gelfand problem $-\triangle u = \exp(u)$.
-
-\item \texttt{step-68} is a demonstration of how to embed and distribute
-  particles in a parallel triangulation. It tracks the movement of
-  a set of particles in a fluid flow and illustrates how to distribute
-  a parallel domain according to the number of locally owned
-  particles instead of the number of locally owned cells.
-
-\item \texttt{step-71} focuses on automatic and symbolic
-  differentiation (AD and SD, in short) as a tool to make solvers for complex,
-  nonlinear problems possible. To this end, \dealii{} can interface to a number
-  of AD and SD libraries, specifically Trilinos' Sacado package
-  \cite{Bartlett2006a}, ADOL-C \cite{Griewank1996a}, and
-  SymEngine \cite{symengine-web-page}. The tutorial
-  illustrates how these techniques can be used to compute derivatives
-  first of a rather simple function, and then of the much more complex
-  energy functions of two magnetoelastic and magneto-viscoelastic
-  material formulations in which just the scalar energy functional
-  takes up the better part of a page, and even first derivatives can
-  only be computed with heroic effort.
-
-\item \texttt{step-72} illustrates the use of automatic
-  differentiation to simplify the computation of derivatives in the
-  context of nonlinear partial differential equations where one needs
-  to compute the Jacobian from the residual operator for efficient
-  Newton iterations. \texttt{step-72} builds on the minimal surface
-  solver \texttt{step-15} and replaces the hand-construction of the
-  Jacobian by either computing it as the derivative of the residual,
-  or alternatively as the second derivative of an energy functional
-  that then also yields the residual itself.
-
-\item \texttt{step-74} implements the symmetric interior penalty Galerkin 
-(SIPG) method for Poisson's equation using the \texttt{FEInterfaceValues} class
-within the \texttt{MeshWorker::mesh\_loop()} framework. This tutorial demonstrates a
-simple way to assemble face integrals.
-
-\item \texttt{step-75} demonstrates a state-of-the-art way of solving a simple
-      Laplace problem using $hp$-adaptation and hybrid multigrid methods on machines
-      with distributed memory. This tutorial points out particularities in porting
-      serial $hp$-adaptive code for parallelization. Furthermore, it guides through
-      the process of writing an efficient matrix-free preconditioner for $hp$-adaptive applications.
-
-\item \texttt{step-76} is an explicit time integrator for the
-      compressible Euler equations discretized with a high-order discontinuous
-      Galerkin (DG) scheme, using the matrix-free infrastructure just as \texttt{step-67} does.
-      The tutorial presents advanced topics, like the usage of cell-centric loops and
-      the new MPI-3.0 shared-memory capabilities of \texttt{MatrixFree} to reach high
-      throughput. Furthermore, the utilization of the template parameter
-      \texttt{VectorizedArrayType} and the application of lambda functions to describe cell and face
-      integrals are discussed.
-
-\item \texttt{step-77} is a program that illustrate \dealii{}'s
-  interfaces to the SUNDIALS library \cite{sundials}, and specifically
-  the KINSOL nonlinear solver. Like the \texttt{step-72} program
-  mentioned above, it is a variation of the minimal surface solver
-  \texttt{step-15}: Instead of implementing the nonlinear Newton and
-  line search loop ourselves, \texttt{step-77} relies on KINSOL for
-  decisions such as when to rebuild the Jacobian matrix, when to
-  actually solve linear systems with it, and how to form updates that
-  drive the residual to convergence. The program illustrates the
-  substantial savings that can be obtained by not re-inventing the
-  wheel but instead building on an existing and well-tuned software
-  such as KINSOL.
-
-  \dealii{}'s interfaces to the various SUNDIALS sub-packages were also updated to the latest SUNDIALS release, 5.7.
-
-\item \texttt{step-78}, a program written by Tyler Anderson at
-  Colorado State University, solves the
-  one-dimensional Black-Scholes equations to model the price of stock
-  options.
-
-\item \texttt{step-79} is a program for topology optimization. Written
-  by Justin O'Connor at Colorado State University, the program asks
-  what the optimal distribution of a finite amount of material in a
-  domain is to maximize its strength (or optimize some other objective
-  functional). It uses advanced optimization techniques to deal with
-  the nonlinearity of the problem as well as with the equality and
-  inequality constraints that characterize the application.
 \end{itemize}
@@
 \begin{itemize}
-\item ``\texttt{Laplace equation coupled to an external simulation program}'' was
-  contributed by David Schneider and Benjamin Uekermann at Technical University
-  of Munich. It extends \texttt{step-4} to a time-dependent Poisson
-  problem and couples it with a simple external surface-based
-  application, using the \texttt{preCICE} 
-  library~\cite{bungartz2016precice, lindner2020exafsa}. \texttt{preCICE},
-  also allows to couple \dealii to external simulation packages, 
-  such as OpenFOAM, SU2, CalculiX, or FEniCS, using adapter classes. 
-  The webpage
-  \url{https://precice.org/} provides several more \dealii{}-based tutorials,
-  including interesting fluid-structure interaction applications.
+\item ``\texttt{TODO}'' TODO.
 \end{itemize}
@@
 
-The 9.3 release includes
-\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_2_0_and_9_3_0.html}
-{around 45 incompatible changes}; see \cite{changes93}. The majority of these changes
+The 9.4 release includes
+\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_3_0_and_9_4_0.html}
+{around TODO incompatible changes}; see \cite{changes94}. The majority of these changes
 should not be visible to typical user codes; some remove previously
@@
 \begin{itemize}
-  \item The \texttt{hp::DoFHandler} class has been deprecated. The standard
-        \texttt{DoFHandler} class is now capable of all $hp$-functionalities.
-  \item Consequently, all template arguments \texttt{DoFHandlerType} are now obsolete,
-        and classes like \texttt{DataOut} or \texttt{SolutionTransfer} for example
-        no longer require them. If you rely on these template arguments, an interim
-        namespace \texttt{Legacy} has been introduced that provides all affected
-        classes with the old interface for a transition period.
-  \item \texttt{GridTools::find\_active\_cell\_around\_point()} no longer throws an exception when no cell if found, but returns an invalid iterator.
-  User codes previously catching an exception will need to be changed.
+  \item TODO
+  \item TODO
 
@@
 %      sorted by WB 5/28
-Pasquale Africa,
-Tyler Anderson,
-Mathias Anselmann,
-Arpit Babbar,
-Maximilian Bergbauer,
-Nicolas Barnafi,
-Michele Bucelli,
-Marcus Calhoun-Lopez,
-David F. Castellanos,
-Fabian Castelli, %He provided major contributions to candi for v9.3
-Praveen Chandrashekar,
-Conrad Clevenger,
-Andrew Davis,
-Elias Dejene,
-Niklas Fehn,
-Menno Fraters,
-Ivan Fumagalli,
-Daniel Garcia-Sanchez,
-Nicola Giuliani,
-Krishnakumar Gopalakrishnan,
-Alexander Grayver,
-Olivier Guevremont,
-Jake Harmon,
-Graham Harper,
-Katharina Kormann,
-Christoph Kammer,
-Wenyu Lei,
-Zhou Lei,
-Phillip Mobley,
-Nils Much,
-Pratik Nayak,
-Toni El Geitani Nehme,
-Justin O'Connor,
-Daniel Paukner,
-Lei Qiao,
-Ce Qin,
-Reza Rastak,
-Julian Roth,
-Laura Prieto Saavedra,
-Doug Shi-Dong,
-David Schneider,
-Magdalena Schreter,
-Richard Schussnig,
-Dominic Soldner,
-Simon Sticko,
-Malhar Tidke,
-Ignacio Tomas,
-Benjamin Uekermann,
-Peter Westerbaan.
+TODO.
 

### 2022-04-23 9.4/paper.tex
Draft

@@
 %\renewcommand{\baselinestretch}{2.0}
+
+\graphicspath{{svg/}}
 
@@
    Clemson, SC, 29634, USA
-   {\texttt{heister/jiaqi2@clemson.edu}}}
+   {\texttt{heister@clemson.edu}}}
 
@@
    Boltzmannstr.~15, 85748 Garching, Germany.
-   {\texttt{kronbichler/munch/proell@lnm.mw.tum.de}}}
+   {\texttt{kronbichler/munch@lnm.mw.tum.de}}}
  \affil[10]{Department of Information Technology,
@@
   This paper provides an overview of the new features of the finite element
-  library \dealii, version 9.3.
+  library \dealii, version 9.4.
 \end{abstract}
@@
 \begin{itemize}
-  \item TODO;
-  \item TODO new tutorial programs and a new code gallery program (see Section~\ref{subsec:steps}).
+  \item Advances in simplex- and mixed-mesh support (see Section~\ref{sec:simplex});
+  \item Repartitioning of distributed meshes (see Section~\ref{sec:repartitioning});
+  \item Advances in matrix-free infrastructure (see Section~\ref{sec:mf});
+  \item Advances in multigrid infrastructure (see Section~\ref{sec:multigrid});
+  \item CutFEM support (see Section~\ref{sec:cut});
+  \item Performance improvement of particle infrastructure (see Section~\ref{sec:particles});
+  \item Two new tutorial programs and a new code gallery program (see Section~\ref{subsec:steps}).
 \end{itemize}
 %
-In addition, we discuss the \candi{} installation program in Section~\ref{subsec:candi}.
 
@@
 \begin{itemize}
-  \item TODO.
+%  \item \texttt{AffineConstraints::make\_consistent\_in\_parallel()} allows
+%  to make constraints consistent in parallel.
+  \item \texttt{DataOutResample} does not output a numerical solution 
+  on the cells of the original triangulation but interpolates the result 
+  onto a second triangulation (that can be completely unrelated). 
+  By using this class, one can output the result obtained on an 
+  unstructured mesh on a structured one, which might be a more
+  memory-efficient storage format, or one can create a slice in 3D.
+  \item The new function \texttt{find\_point\_owner\_rank()} of \texttt{parallel::distributed::Triangulation} allows to find the MPI
+  rank of the subdomain of a distributed mesh that contains a given point.
+  It is communication-free, leverages the functionality of p4est (>v.2.2),
+  and its algorithm is described in \cite{burstedde2020parallel}. Based on the obtained information,
+  communication pattern in \texttt{Utilties::MPI::RemotePointEvaluation} can be set up efficiently. Furthermore, this feature could be used in the future to allow
+  particle simulations, where particle movement is not
+  limited by CFL conditions, as done in \cite{mirzadeh2016parallel}.
 \end{itemize}
 %
-The changelog lists more than 200 other features and bugfixes.
+The changelog lists more than 150 other features and bugfixes.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Topic 1}
+\subsection{Advances in simplex- and mixed-mesh support}\label{sec:simplex}
+
+We continued to work on the simplex- and mixed-mesh support. We fixed
+many bugs and generalized existing functions that only worked for
+hypercube shaped cells. Most notable new functions are:
+
+\begin{figure}[!t]
+
+  \centering
+
+  \begin{tikzpicture}[scale=1.7]
+
+    \coordinate (0) at (0,0);
+    \coordinate (1) at (1,0);
+    \coordinate (2) at (0,1);
+
+    \coordinate (3) at (0,0.5);
+    \coordinate (4) at (0.5,0.5);
+    \coordinate (5) at (0.5,0);
+
+    \foreach \i in {0,1, 2, 3, 4, 5}
+    \draw[blue,fill=blue] (\i) circle (1.1pt) node [below] {};
+
+    \draw (0) --(1) -- (2) -- (0);
+    \draw (3) --(4) -- (5) -- (3);
+
+  \end{tikzpicture}
+  \qquad\qquad
+  \begin{tikzpicture}[scale=1.3]
+
+    \coordinate (0) at (0,0,0);
+    \coordinate (1) at (0,0,1);
+    \coordinate (2) at (1,0,0);
+    \coordinate (3) at (0,1,0);
+    \coordinate (4) at (0,0,0.5);
+    \coordinate (6) at (0.5,0,0);
+    \coordinate (7) at (0,0.5,0);
+    \coordinate (5) at (0.5,0.0,0.5);
+    \coordinate (8) at (0.0,0.5,0.5);
+    \coordinate (9) at (0.5,0.5,0.0);
+
+    \draw (0) -- (1) -- (2) -- (0);
+    \draw (0) -- (1) -- (3) -- (0);
+    \draw (1) -- (2) -- (3) -- (1);
+
+    \draw[-, fill=red!30, opacity=.5] (4)--(6)--(7)--cycle;
+    \draw[-, fill=red!30, opacity=.5] (0)--(6)--(7)--cycle;
+    \draw[-, fill=red!30, opacity=.5] (0)--(4)--(7)--cycle;
+    \draw[-, fill=red!30, opacity=.5] (0)--(4)--(6)--cycle;
+
+    \draw[-, fill=blue!30, opacity=.5] (1)--(4)--(5)--cycle;
+    \draw[-, fill=blue!30, opacity=.5] (1)--(4)--(8)--cycle;
+    \draw[-, fill=blue!30, opacity=.5] (4)--(5)--(8)--cycle;
+    \draw[-, fill=blue!30, opacity=.5] (1)--(5)--(8)--cycle;
+
+    \draw[-, fill=green!30, opacity=.5] (2)--(5)--(6)--cycle;
+    \draw[-, fill=green!30, opacity=.5] (2)--(6)--(9)--cycle;
+    \draw[-, fill=green!30, opacity=.5] (2)--(5)--(9)--cycle;
+    \draw[-, fill=green!30, opacity=.5] (5)--(6)--(9)--cycle;
+
+    \draw[-, fill=orange!30, opacity=.5] (7)--(8)--(9)--cycle;
+    \draw[-, fill=orange!30, opacity=.5] (3)--(7)--(9)--cycle;
+    \draw[-, fill=orange!30, opacity=.5] (3)--(7)--(8)--cycle;
+    \draw[-, fill=orange!30, opacity=.5] (3)--(8)--(9)--cycle;
+
+    \foreach \i in {0,1,2,3,4,5,6,7,8,9}
+    \draw[blue,fill=blue] (\i) circle (1.0pt) node [below] {};
+
+  \end{tikzpicture}
+  \begin{tikzpicture}[scale=1.3]
+
+    \coordinate (0) at (0,0,0);
+    \coordinate (1) at (0,0,1);
+    \coordinate (2) at (1,0,0);
+    \coordinate (3) at (0,1,0);
+    \coordinate (4) at (0,0,0.5);
+    \coordinate (6) at (0.5,0,0);
+    \coordinate (7) at (0,0.5,0);
+    \coordinate (5) at (0.5,0.0,0.5);
+    \coordinate (8) at (0.0,0.5,0.5);
+    \coordinate (9) at (0.5,0.5,0.0);
+
+    \draw (0) -- (1) -- (2) -- (0);
+    \draw (0) -- (1) -- (3) -- (0);
+    \draw (1) -- (2) -- (3) -- (1);
+
+    \draw[-, fill=red!30, opacity=.5] (4)--(5)--(6)--cycle;
+    \draw[-, fill=red!30, opacity=.5] (4)--(5)--(8)--cycle;
+    \draw[-, fill=red!30, opacity=.5] (4)--(6)--(8)--cycle;
+    \draw[-, fill=red!30, opacity=.5] (5)--(6)--(8)--cycle;
+
+%    \draw[-, fill=blue!30, opacity=.5] (4)--(6)--(7)--cycle;
+%    \draw[-, fill=blue!30, opacity=.5] (4)--(6)--(8)--cycle;
+%    \draw[-, fill=blue!30, opacity=.5] (4)--(7)--(8)--cycle;
+%    \draw[-, fill=blue!30, opacity=.5] (6)--(7)--(8)--cycle;
+
+%    \draw[-, fill=green!30, opacity=.5] (6)--(7)--(8)--cycle;
+%    \draw[-, fill=green!30, opacity=.5] (6)--(7)--(9)--cycle;
+%    \draw[-, fill=green!30, opacity=.5] (6)--(8)--(9)--cycle;
+%    \draw[-, fill=green!30, opacity=.5] (7)--(8)--(9)--cycle; 
+%        
+    \draw[-, fill=blue!30, opacity=.5] (5)--(6)--(8)--cycle;
+    \draw[-, fill=blue!30, opacity=.5] (5)--(6)--(9)--cycle;
+    \draw[-, fill=blue!30, opacity=.5] (5)--(8)--(9)--cycle;
+    \draw[-, fill=blue!30, opacity=.5] (6)--(8)--(9)--cycle;
+
+    \foreach \i in {0,1,2,3,4,5,6,7,8,9}
+    \draw[blue,fill=blue] (\i) circle (1.0pt) node [below] {};
+
+  \end{tikzpicture}
+  \begin{tikzpicture}[scale=1.3]
+
+    \coordinate (0) at (0,0,0);
+    \coordinate (1) at (0,0,1);
+    \coordinate (2) at (1,0,0);
+    \coordinate (3) at (0,1,0);
+    \coordinate (4) at (0,0,0.5);
+    \coordinate (6) at (0.5,0,0);
+    \coordinate (7) at (0,0.5,0);
+    \coordinate (5) at (0.5,0.0,0.5);
+    \coordinate (8) at (0.0,0.5,0.5);
+    \coordinate (9) at (0.5,0.5,0.0);
+
+    \draw (0) -- (1) -- (2) -- (0);
+    \draw (0) -- (1) -- (3) -- (0);
+    \draw (1) -- (2) -- (3) -- (1);
+
+%    \draw[-, fill=red!30, opacity=.5] (4)--(5)--(6)--cycle;
+%    \draw[-, fill=red!30, opacity=.5] (4)--(5)--(8)--cycle;
+%    \draw[-, fill=red!30, opacity=.5] (4)--(6)--(8)--cycle;
+%    \draw[-, fill=red!30, opacity=.5] (5)--(6)--(8)--cycle;
+
+    \draw[-, fill=red!30, opacity=.5] (4)--(6)--(7)--cycle;
+    \draw[-, fill=red!30, opacity=.5] (4)--(6)--(8)--cycle;
+    \draw[-, fill=red!30, opacity=.5] (4)--(7)--(8)--cycle;
+    \draw[-, fill=red!30, opacity=.5] (6)--(7)--(8)--cycle;
+
+    \draw[-, fill=blue!30, opacity=.5] (6)--(7)--(8)--cycle;
+    \draw[-, fill=blue!30, opacity=.5] (6)--(7)--(9)--cycle;
+    \draw[-, fill=blue!30, opacity=.5] (6)--(8)--(9)--cycle;
+    \draw[-, fill=blue!30, opacity=.5] (7)--(8)--(9)--cycle;
+
+%    \draw[-, fill=orange!30, opacity=.5] (5)--(6)--(8)--cycle;
+%    \draw[-, fill=orange!30, opacity=.5] (5)--(6)--(9)--cycle;
+%    \draw[-, fill=orange!30, opacity=.5] (5)--(8)--(9)--cycle;
+%    \draw[-, fill=orange!30, opacity=.5] (6)--(8)--(9)--cycle;
+
+    \foreach \i in {0,1,2,3,4,5,6,7,8,9}
+    \draw[blue,fill=blue] (\i) circle (1.0pt) node [below] {};
+
+  \end{tikzpicture}
+
+  \caption{New refinement strategies for triangles and tetrahedrons: triangles
+  are subdivided in 4 children and tetrahedrons 8 ones.}\label{fig:refinement}
+
+\end{figure}
+
+\begin{itemize}
+\item Experimental support of locally refined meshes. FEM on locally refined
+meshes requires 1) the possibility to locally refine the mesh (see Figure~\ref{fig:refinement})
+and 2) appropriate hanging-node constraints. For 3D, we are in progress to
+implement the missing constraint definitions.
+\item TODO
+\end{itemize}
+
+Furthermore, we continued to remove the usage of  \texttt{GeometryInfo} from
+the library and to replace the instances by more general equivalent functions, which
+in many cases rely of \texttt{ReferenceCell}. Once all instance of 
+\texttt{GeometryInfo} are remove, we will deprecate the class.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Topic 2}
+\subsection{Repartitioning of distributed meshes}\label{sec:repartitioning}
+
+Until now, distributed meshes were either partitioned statically (\texttt{parallel::\allowbreak fullydistributed::\allowbreak Triangulation}; short \texttt{p::f::T}) or used a
+fixed policy to partition the cells among processes
+(\texttt{parallel::\allowbreak distributed::\allowbreak Triangulation}; short \texttt{p::d::T}). For latter, we
+use a Morton-order as the backend of that class---\texttt{p4est}---does. Fixed Morton-order
+might provide very good performance in terms in communication and
+setup costs, but might be non-optimal when interfacing with other
+libraries that have fixed partitioning, most notably a Cartesian
+partitioning, themselves.
+
+New utility functions from the \texttt{RepartitioningPolicyTools} can be
+used now to create a new \texttt{p::f::T} instances,
+given a distributed triangulation (\texttt{p::f::T} or \texttt{p::d::T}) and a vector with the new owners of locally
+owned cells. The workflow is shown in the following listing:
+\begin{c++}
+// original triangulation
+parallel::distributed::Triangulation<dim> tria(comm);
+
+// select a partitioning policy and use it
+const RepartitioningPolicyTools::Cartesian<dim> policy(tria);
+const auto construction_data = TriangulationDescription::Utilities::
+  create_description_from_triangulation(tria, policy.partition(tria));
+
+// create the new triangulation
+parallel::fullydistributed::Triangulation<dim> tria_pft(comm);
+tria_pft.create_triangulation(construction_data);
+\end{c++}
+Instead of using one of the predefined partitioning policies, users
+can write their own by implementing the \texttt{RepartitioningPolicyTools::Base} interface. Similarly to the
+active level, also the multigrid levels can be repartitioned arbitrarily.
+
+\begin{figure}
+    \centering
+    \def\svgwidth{0.8\columnwidth}
+    \input{svg/repartitioning.pdf_tex}
+    \caption{Visualization of the repartitioning process: after 
+    cells are assigned new ranks, each process collect sends the cells incl. 
+    ghost cells to the new owner. There incoming cells are processed, duplicates
+    are removed and the local part of the triangulation is built.}\label{fig:repartitioning}
+\end{figure}
+
+The setup process is visualized in Figure~\ref{fig:repartitioning}. First cells, incl. their
+surrounding (ghost) cells and their parent cells, are collected and 
+sent to the new owner. On the
+receiving side, the set of all cells are combined and possible duplicates
+are removed. This information is enough to construct a new triangulation.
+For sending/receiving, we apply consensus-based algorithms~\cite{hoefler2010scalable}, which
+we introduced in the library in release 9.2~\cite{dealII92}. Consensus-based 
+algorithms are used also in \cite{ibanez2016pumi} for repartitioning.
+
+In future releases, we plan to add support for repartitioning based on distributed graph
+partitioning libraries, e.g., \texttt{ParMETIS} of \texttt{Zotan}. Furthermore, we intend to extend the 
+\texttt{p::f::T} with adaptivity support.
+With this and and the new repartitionging features, \texttt{p::f::T}
+could be selfconsistent like \texttt{p::d::T}, which is an important
+step towards the support of AMR also for distributed simplex and mixed meshes
+(see also Subsection~\ref{sec:simplex}).
+
+%We would like to note that we provide a similar class also for repartitioning
+%of vectors: \texttt{Noncontiguous\allowbreak Partitioner} from the
+%\texttt{Utilities::MPI} namespace.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Topic 3}
+\subsection{Advances in matrix-free infrastructure}\label{sec:mf}
+
+In the matrix-free infrastructure, we have added many new features. Most
+notably:
+\begin{itemize}
+\item Improved support of Hessians: just like values and gradients can be
+evaluated and integrated during matrix-free loops, one can do this now
+also for Hessians both for cells and faces. This enables, e.g., to write
+a matrix-free version of step-47, which solves the biharmonic equation with DG.
+\item Cell-centric loops now also allow to access to gradients and Hessians
+of neighboring cells on faces. The major obstacle here was to reorientation
+of theses in the case of arbitrary orientated cells on unstructured meshes.
+\item Users can now create their own cell batches, by providing \texttt{FEEvaluation::reinit()} a list of cell IDs. \texttt{FEEvaluation}
+accesses the appropriate data and reshuffles mapping data accordingly on
+the fly to enable vectorization over cells. The new feature is useful in many
+contexts. Examples are application with sharp interfaces (e.g., two phase flow
+or shock capturing) where one needs to treat cells that are ``cut'' by
+the interface in a special way. A challenge is that cell batchs
+might contain cut or not-cut cell, making vectorization potentially more complication. One way to deal with such cell batches it to apply masks if
+the code paths do not diverge too much
+(many functions of \texttt{FEEvaluation} allows this). Another way is to
+categorize cells during \texttt{MatrixFree::reinit()} in such a way that mixed cell batches do not occur. However, calling  \texttt{MatrixFree::reinit()}
+might be too expensive if a very dynamic system is given, which requires
+recatorization in each time step. Doing this on-the-fly might be a cheep alternative, even if not computationally optimal during matrix-free loops.
+\end{itemize}
+Besides these new features, we improved the performance of 
+hanging-node-constraint evaluation on the CPU. Instead of performing
+quasi-dense matrix-vector multiplications~\cite{KronbichlerKormann2012}, we now use an
+approach based on inplace interpolation and sumfactorization, similarly
+as we already did in the GPU code~\cite{ljungkvist2017matrix}. The algorithm is described
+in \cite{munch2022hn} and performance numbers are shown, indicating a reduction
+of overhead of cells with hanging nodes by a factor of 10$\times$.
+
+Furthermore, we have performed a major restructuring of internals
+of the \texttt{FEEvaluation} classes. This will enable us to add more
+simply the support of new element types in the future, e.g., of 
+Raviart--Thomas and Nedelec elements.
+
+%\begin{c++}
+%additional_data.mapping_update_flags = ... | update_hessians;
+%\end{c++}
+%
+%\begin{c++}
+%FEEvaluation<dim> phi(matrix_free);
+%
+%phi.reinit(cell);
+%phi.gather_evaluate(src, ... | EvaluationFlags::hessians);
+%
+%for(const auto q : phi.quadrature_point_indices())
+%  phi.submit_hessian(phi.get_hessian(q), q);
+%  
+%phi.integrate_scatter(... | EvaluationFlags::hessians, dst);
+%\end{c++}
+%
+%... similar for faces; can be used to implement biharmonic equation (like in
+%step-X)
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Topic 4}
+\subsection{Advances in multigrid infrastructure}\label{sec:multigrid}
+
+In release 9.3~\cite{dealII93}, we added support for global coarsening multigrid in 
+addition to the established local smoothing infrastructure. Global 
+coarsening smoothens on the whole computational domain on each
+multigrid level, which are obtained by coarsening the finest mesh globally.
+For this purpose, we use a sequence of triangulations, on which we perform
+the smoothing only on the active levels. To create the sequence of 
+triangulations, one can use the functions \texttt{create\_geometric\_coarsening\_sequence()} from the
+\texttt{MGTransferGlobalCoarseningTools} namespace. A new version takes an 
+instance of \texttt{RepartitioningPolicyTools::Base} (see Subsection~\ref{sec:repartitioning}, which allow to specify the parallel 
+distribution of each multigrid level (in contrast to the fixed---first-child---policy in the case of local smoothing). Furthermore, we added support of block vectors,
+fixed a number of limitations, and performed performance optimizations of 
+the transfer operator; particularly, the redundant copy from/to temporary vectors
+has been eliminated and the application of hanging-node constraints are applied
+efficiently the same way as in the matrix-free loops (see Subsection~\ref{sec:mf}).
+
+In \cite{munch2022gc}, the performance of \texttt{deal.II}'s local-smoothing and global-coarsening
+infrastructure was compared for locally refined meshes. The results indicate that
+the local definition of multirgrid levels might introduce load imbalances
+in the case of local smoothing so that global coarsening is favorable despite
+of potentially more expensive inter-grid transfers. To judge the benefit
+of one approach against the other, \texttt{deal.II} provides new functions
+\texttt{workload\_imbalance()} and \texttt{vertical\_communication\_efficiency()} 
+in the \texttt{MGTools} namespace to estimate the workload imbalance and the
+vertical communication efficiency purely based on the given mesh. The publication also points out that not
+all types of smoothers are applicable for global coarsening, due to the
+presence of hanging-nodes, which is a motivation to add new smoother types
+to \dealii.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Topic 5}
+\subsection{CutFEM support}\label{sec:cut}
+
+\begin{itemize}
+\item new classes
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Performance improvement of particle infrastructure}\label{sec:particles}
+
+\begin{itemize}
+\item describe the new data structure
+\item do we have performance numbers?
+\end{itemize}
 
@@
 \begin{itemize}
-\item \texttt{TODO} TODO.
-
+\item \texttt{step-82} was contributed by Andrea Bonito (Texas A\&M University) and Diane Guignard (University of Ottawa). It shows how
+\dealii{} can be used to implement the
+local discontinuous Galerkin (LDG) method for approximating the solution to the bi-Laplacian problem.
+\item \texttt{step-85} was contributed by Simon Sticko (Uppsala University). It shows how to use the new CutFEM infrastructure to solve a Poisson 
+problem on a circular domain that is embedded in a Cartesian background mesh.
 \end{itemize}
 
-There is also a new program in the code gallery (a collection of
-user-contributed programs that often solve more complicated problems
-than tutorial programs, and intended as starting points for further
-research rather than as teaching tools):
-\begin{itemize}
-\item ``\texttt{TODO}'' TODO.
-\end{itemize}
+%There is also a new program in the code gallery (a collection of
+%user-contributed programs that often solve more complicated problems
+%than tutorial programs, and intended as starting points for further
+%research rather than as teaching tools):
+%\begin{itemize}
+%\item ``\texttt{TODO}'' TODO.
+%\end{itemize}
 
@@
 \href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_3_0_and_9_4_0.html}
-{around TODO incompatible changes}; see \cite{changes94}. The majority of these changes
+{around 40 incompatible changes}; see \cite{changes94}. The majority of these changes
 should not be visible to typical user codes; some remove previously
@@
 \begin{itemize}
-  \item TODO
-  \item TODO
+  \item In continuation of our attempt to merge the classes \texttt{DoFHandler} and \texttt{hp::DoFHandler}, we have removed the
+  template parameter \texttt{DoFHandlerType} form many classes and
+  functions and replaced it by \texttt{dim}/\texttt{spacedim}. Affected
+  classes are, e.g., \texttt{SolutionTransfer} and \texttt{DataOut}. 
+  \item \texttt{FE\_RaviartThomasNodal} now uses a different polynomial space to allow
+for a simpler use for faces in non-standard orientation. The new polynomials
+are anisotropic tensor products of Lagrange polynomials on the points of a
+Gauss--Lobatto quadrature formula. This change leads to different entries in
+the matrices and constraints, for example, but as the resulting polynomial
+space spans the same polynomials, no change in accuracy should be expected.
+\item We made different changes to the repartitoning of \texttt{parallel::distributed::Triangulation}. Most notably, we have have
+removed the default weight (10,000) for consistency reasons with the rest
+of the library.
 
@@
   \begin{itemize}
-    \item For geometric multigrid: \cite{Kanschat2004,JanssenKanschat2011,ClevengerHeisterKanschatKronbichler2019};
+    \item For geometric multigrid: \cite{Kanschat2004,JanssenKanschat2011,ClevengerHeisterKanschatKronbichler2019, munch2022gc};
     \item For distributed parallel computing: \cite{BangerthBursteddeHeisterKronbichler11};
@@
 %
-% 9.3: updated 5/08/2021 PM
-%      sorted by WB 5/28
-TODO.
+% 9.4: updated 23/04/2022 PM
+Pasquale	Africa,
+Tyler	Anderson,
+Francesco	Andreuzzi,
+Mathias	Anselmann,
+Maximilian	Bergbauer,
+Bruno	Blais,
+Till	Budde,
+Fabian	Castelli,
+Praveen	Chandrashekar,
+Cu	Cui,
+Ivo	Dravins,
+Marco	Feder,
+Niklas	Fehn,
+Johannes	Friedlein,
+Sebastian	Fuchs,
+Rene	Gassmoeller,
+Nicola	Giuliani,
+Alexander	Grayver,
+Diane	Guignard,
+Jake	Harmon,
+Sean	Ingimarson,
+Pengfei	Jia,
+Sebastian	Kinnewig,
+Simon	Konrad,
+Katharina	Kormann,
+Paras	Kumar,
+Wenyu	Lei,
+Alberto F.	Martin,
+Nils	Much,
+Lucas	Myers,
+Justin	O'Connor,
+Judith	Pauen,
+Vachan	Potluri,
+Raghunandan	Pratoori,
+Sebastian	Proell,
+Ce	Qin,
+Reza	Rastak,
+Jose E.	Roman,
+Raphael	Schoof,
+Magdalena	Schreter,
+Konrad	Simon,
+Simon	Sticko,
+Bruno	Turcksin,
+Kuljit S.	Virk,
+Michał	Wichrowski,
+Niklas Wik,
+Jiaqi	Zhang.
 

### 2022-04-24 9.4/paper.tex
Fix typos

@@
 
-\dealii{} version 9.4.0 was released June TODO, 2022.
+\dealii{} version 9.4.0 was released June 10, 2022.
 This paper provides an
@@
   \item Performance improvement of particle infrastructure (see Section~\ref{sec:particles});
-  \item Two new tutorial programs and a new code gallery program (see Section~\ref{subsec:steps}).
+  \item Two new tutorial programs (see Section~\ref{subsec:steps}).
 \end{itemize}
@@
 Section~\ref{sec:major}, there
-are a number of other noteworthy changes in the current \dealii{} release
-that we briefly outline in the remainder of this section:
+are a number of other noteworthy changes in the current \dealii{} release,
+which we briefly outline in the remainder of this section:
 %
@@
 %  to make constraints consistent in parallel.
-  \item \texttt{DataOutResample} does not output a numerical solution 
-  on the cells of the original triangulation but interpolates the result 
-  onto a second triangulation (that can be completely unrelated). 
+  \item The \texttt{DataOutResample} class does not output a numerical solution 
+  on the cells of the original triangulation, but interpolates the result 
+  onto a second triangulation (which can be completely unrelated). 
   By using this class, one can output the result obtained on an 
@@
   rank of the subdomain of a distributed mesh that contains a given point.
-  It is communication-free, leverages the functionality of p4est (>v.2.2),
-  and its algorithm is described in \cite{burstedde2020parallel}. Based on the obtained information,
-  communication pattern in \texttt{Utilties::MPI::RemotePointEvaluation} can be set up efficiently. Furthermore, this feature could be used in the future to allow
-  particle simulations, where particle movement is not
+  It is communication-free and leverages the functionality of p4est (>v.2.2).
+  Its algorithm is described in \cite{burstedde2020parallel}. Based on the information obtained,
+  communication pattern in \texttt{Utilties::MPI::RemotePointEvaluation} can be set up efficiently. Furthermore, this function could be used in the future to allow
+  particle simulations where particle movement is not
   limited by CFL conditions, as done in \cite{mirzadeh2016parallel}.
@@
 
-This release of \dealii{} contains a number of large and significant changes
-that will be discussed in this section.
-It of course also contains a
+This release of \dealii{} contains a number of large and significant changes,
+which will be discussed in this section.
+It of course also includes a
 vast number of smaller changes and added functionality; the details of these
@@
 many bugs and generalized existing functions that only worked for
-hypercube shaped cells. Most notable new functions are:
+hypercube-shaped cells. The most notable new functions are:
 
@@
   \caption{New refinement strategies for triangles and tetrahedrons: triangles
-  are subdivided in 4 children and tetrahedrons 8 ones.}\label{fig:refinement}
+  are subdivided in 4 children and tetrahedrons in 8 ones.}\label{fig:refinement}
 
@@
 meshes requires 1) the possibility to locally refine the mesh (see Figure~\ref{fig:refinement})
-and 2) appropriate hanging-node constraints. For 3D, we are in progress to
-implement the missing constraint definitions.
+and 2) appropriate hanging-node constraints. For 3D, the implementation of the missing constraint definitions is in progress.
 \item TODO
@@
 
-Furthermore, we continued to remove the usage of  \texttt{GeometryInfo} from
-the library and to replace the instances by more general equivalent functions, which
-in many cases rely of \texttt{ReferenceCell}. Once all instance of 
-\texttt{GeometryInfo} are remove, we will deprecate the class.
+Furthermore, we continued to remove the instance of usage of the \texttt{GeometryInfo} class from
+the library and to replace them by more general equivalent functions, which
+rely on \texttt{ReferenceCell} in many cases. Once all instances of 
+\texttt{GeometryInfo} are removed, we will deprecate the class.
 
@@
 
-Until now, distributed meshes were either partitioned statically (\texttt{parallel::\allowbreak fullydistributed::\allowbreak Triangulation}; short \texttt{p::f::T}) or used a
+Until now, distributed meshes were either partitioned statically (\texttt{parallel::\allowbreak fullydistributed::\allowbreak Triangulation}; short \texttt{p::f::T}) or they used a
 fixed policy to partition the cells among processes
-(\texttt{parallel::\allowbreak distributed::\allowbreak Triangulation}; short \texttt{p::d::T}). For latter, we
-use a Morton-order as the backend of that class---\texttt{p4est}---does. Fixed Morton-order
+(\texttt{parallel::\allowbreak distributed::\allowbreak Triangulation}; short \texttt{p::d::T}). For the latter, we
+use the Morton-order as the backend of \texttt{p::d::T}---\texttt{p4est}---does. A fixed Morton-order
 might provide very good performance in terms in communication and
 setup costs, but might be non-optimal when interfacing with other
-libraries that have fixed partitioning, most notably a Cartesian
+libraries that have fixed partitioning, e.g., a Cartesian
 partitioning, themselves.
 
-New utility functions from the \texttt{RepartitioningPolicyTools} can be
-used now to create a new \texttt{p::f::T} instances,
+New utility functions from the \texttt{RepartitioningPolicyTools} namepace can be
+used now to create a new \texttt{p::f::T} instance,
 given a distributed triangulation (\texttt{p::f::T} or \texttt{p::d::T}) and a vector with the new owners of locally
@@
 Instead of using one of the predefined partitioning policies, users
-can write their own by implementing the \texttt{RepartitioningPolicyTools::Base} interface. Similarly to the
+can write their own ones by implementing the \texttt{RepartitioningPolicyTools::Base} interface. Similarly to the
 active level, also the multigrid levels can be repartitioned arbitrarily.
@@
     \input{svg/repartitioning.pdf_tex}
-    \caption{Visualization of the repartitioning process: after 
-    cells are assigned new ranks, each process collect sends the cells incl. 
-    ghost cells to the new owner. There incoming cells are processed, duplicates
-    are removed and the local part of the triangulation is built.}\label{fig:repartitioning}
+    \caption{Visualization of the repartitioning process: after new ranks 
+    are assigned to cells, each process collects and sends the cells, incl. 
+    ghost cells, to the new owner. There, incoming cells are processed, duplicates
+    are removed, and the local part of the triangulation is built.}\label{fig:repartitioning}
 \end{figure}
 
-The setup process is visualized in Figure~\ref{fig:repartitioning}. First cells, incl. their
-surrounding (ghost) cells and their parent cells, are collected and 
+The setup process is visualized in Figure~\ref{fig:repartitioning}. At first, cells, incl. their
+surrounding (ghost) cells and parent cells, are collected and 
 sent to the new owner. On the
-receiving side, the set of all cells are combined and possible duplicates
+receiving side, the sets of all cells are combined and possible duplicates
 are removed. This information is enough to construct a new triangulation.
 For sending/receiving, we apply consensus-based algorithms~\cite{hoefler2010scalable}, which
-we introduced in the library in release 9.2~\cite{dealII92}. Consensus-based 
+we introduced into the library in release 9.2~\cite{dealII92}. Consensus-based 
 algorithms are used also in \cite{ibanez2016pumi} for repartitioning.
@@
 In future releases, we plan to add support for repartitioning based on distributed graph
-partitioning libraries, e.g., \texttt{ParMETIS} of \texttt{Zotan}. Furthermore, we intend to extend the 
+partitioning libraries, e.g., \texttt{ParMETIS} or \texttt{Zoltan}. Furthermore, we intend to extend 
 \texttt{p::f::T} with adaptivity support.
-With this and and the new repartitionging features, \texttt{p::f::T}
-could be selfconsistent like \texttt{p::d::T}, which is an important
+With this support and and the new repartitioning features, \texttt{p::f::T}
+could be self-consistent like \texttt{p::d::T}, which is an important
 step towards the support of AMR also for distributed simplex and mixed meshes
@@
 
-In the matrix-free infrastructure, we have added many new features. Most
-notably:
+In the matrix-free infrastructure, we have added a number of new features. The most
+notable ones are:
 \begin{itemize}
-\item Improved support of Hessians: just like values and gradients can be
-evaluated and integrated during matrix-free loops, one can do this now
-also for Hessians both for cells and faces. This enables, e.g., to write
+\item Improved support of Hessians: just like values and gradients, Hessians can be
+now evaluated and integrated during matrix-free loops
+both for cells and faces. This enables, e.g., to write
 a matrix-free version of step-47, which solves the biharmonic equation with DG.
-\item Cell-centric loops now also allow to access to gradients and Hessians
-of neighboring cells on faces. The major obstacle here was to reorientation
-of theses in the case of arbitrary orientated cells on unstructured meshes.
+\item Cell-centric loops now also allow to access gradients and Hessians
+of neighboring cells on faces. The major difficulty here was their reorientation
+in the case of arbitrary orientated cells on unstructured meshes.
 \item Users can now create their own cell batches, by providing \texttt{FEEvaluation::reinit()} a list of cell IDs. \texttt{FEEvaluation}
 accesses the appropriate data and reshuffles mapping data accordingly on
-the fly to enable vectorization over cells. The new feature is useful in many
-contexts. Examples are application with sharp interfaces (e.g., two phase flow
-or shock capturing) where one needs to treat cells that are ``cut'' by
-the interface in a special way. A challenge is that cell batchs
-might contain cut or not-cut cell, making vectorization potentially more complication. One way to deal with such cell batches it to apply masks if
+the fly in order to enable vectorization over cells. The new feature is useful in many
+contexts. Examples are application with sharp interfaces (e.g., two-phase flow
+or shock capturing), where one needs to treat cells that are ``cut'' by
+the interface in a special way. A challenge is that cell batches
+might contain cut or not-cut cells, making vectorization potentially more complicated. One way to deal with such cell batches is to apply masks if
 the code paths do not diverge too much
-(many functions of \texttt{FEEvaluation} allows this). Another way is to
+(a lot of functions of \texttt{FEEvaluation} allows this). Another way is to
 categorize cells during \texttt{MatrixFree::reinit()} in such a way that mixed cell batches do not occur. However, calling  \texttt{MatrixFree::reinit()}
 might be too expensive if a very dynamic system is given, which requires
-recatorization in each time step. Doing this on-the-fly might be a cheep alternative, even if not computationally optimal during matrix-free loops.
+recategorization in each time step. Doing this on the fly might be a cheep alternative, even if not computationally optimal during matrix-free loops.
 \end{itemize}
@@
 approach based on inplace interpolation and sumfactorization, similarly
-as we already did in the GPU code~\cite{ljungkvist2017matrix}. The algorithm is described
-in \cite{munch2022hn} and performance numbers are shown, indicating a reduction
+as we already did in the GPU code~\cite{ljungkvist2017matrix}. In \cite{munch2022hn}, the algorithm is described 
+and performance numbers are shown, indicating a reduction
 of overhead of cells with hanging nodes by a factor of 10$\times$.
@@
 Furthermore, we have performed a major restructuring of internals
-of the \texttt{FEEvaluation} classes. This will enable us to add more
-simply the support of new element types in the future, e.g., of 
-Raviart--Thomas and Nedelec elements.
+of the \texttt{FEEvaluation} classes. This will enable us to add the support of new element types, e.g., of 
+Raviart--Thomas and Nedelec elements,  more
+simply in the future.
 
@@
 
-In release 9.3~\cite{dealII93}, we added support for global coarsening multigrid in 
-addition to the established local smoothing infrastructure. Global 
+In release 9.3~\cite{dealII93}, we added support for global-coarsening multigrid in 
+addition to the established local-smoothing infrastructure. Global 
 coarsening smoothens on the whole computational domain on each
-multigrid level, which are obtained by coarsening the finest mesh globally.
-For this purpose, we use a sequence of triangulations, on which we perform
-the smoothing only on the active levels. To create the sequence of 
+multigrid level, which is obtained by coarsening the finest mesh globally.
+For this purpose, we use a sequence of triangulations, and we perform
+the smoothing only on their active levels. To create the sequence of 
 triangulations, one can use the functions \texttt{create\_geometric\_coarsening\_sequence()} from the
 \texttt{MGTransferGlobalCoarseningTools} namespace. A new version takes an 
-instance of \texttt{RepartitioningPolicyTools::Base} (see Subsection~\ref{sec:repartitioning}, which allow to specify the parallel 
-distribution of each multigrid level (in contrast to the fixed---first-child---policy in the case of local smoothing). Furthermore, we added support of block vectors,
+instance of \texttt{RepartitioningPolicyTools::Base} (see Subsection~\ref{sec:repartitioning}) as argument, which allows to specify the parallel 
+distribution of each multigrid level (in contrast to the fixed first-child policy in the case of local smoothing). Furthermore, we added support for block vectors,
 fixed a number of limitations, and performed performance optimizations of 
 the transfer operator; particularly, the redundant copy from/to temporary vectors
-has been eliminated and the application of hanging-node constraints are applied
-efficiently the same way as in the matrix-free loops (see Subsection~\ref{sec:mf}).
+has been eliminated. Now, hanging-node constraints are applied
+efficiently in the same way as in the matrix-free loops (see Subsection~\ref{sec:mf}).
 
-In \cite{munch2022gc}, the performance of \texttt{deal.II}'s local-smoothing and global-coarsening
-infrastructure was compared for locally refined meshes. The results indicate that
-the local definition of multirgrid levels might introduce load imbalances
+In \cite{munch2022gc}, the performance of the local-smoothing and global-coarsening
+infrastructure of \texttt{deal.II} was compared for locally refined meshes. The results indicate that
+the local definition of multigrid levels might introduce load imbalances
 in the case of local smoothing so that global coarsening is favorable despite
-of potentially more expensive inter-grid transfers. To judge the benefit
+potentially more expensive intergrid transfers. In order, to judge the benefits
 of one approach against the other, \texttt{deal.II} provides new functions
 \texttt{workload\_imbalance()} and \texttt{vertical\_communication\_efficiency()} 
-in the \texttt{MGTools} namespace to estimate the workload imbalance and the
-vertical communication efficiency purely based on the given mesh. The publication also points out that not
-all types of smoothers are applicable for global coarsening, due to the
-presence of hanging-nodes, which is a motivation to add new smoother types
-to \dealii.
+in the \texttt{MGTools} namespace for the  estimation of the workload imbalance and the
+vertical communication efficiency purely based on the given mesh. The publication \cite{munch2022gc} also points out that not
+all types of smoothers are applicable for global coarsening due to the
+presence of hanging nodes, which is a motivation to add new smoother types
+to \dealii in the future.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{New and improved tutorials and code gallery programs}
+\subsection{New and improved tutorials}
 \label{subsec:steps}
@@
 should not be visible to typical user codes; some remove previously
-deprecated classes and functions; and the majority change internal
+deprecated classes and functions; and the majority changes internal
 interfaces that are not usually used in external
@@
   \item In continuation of our attempt to merge the classes \texttt{DoFHandler} and \texttt{hp::DoFHandler}, we have removed the
-  template parameter \texttt{DoFHandlerType} form many classes and
+  template parameter \texttt{DoFHandlerType} from number of classes and
   functions and replaced it by \texttt{dim}/\texttt{spacedim}. Affected
@@
 for a simpler use for faces in non-standard orientation. The new polynomials
-are anisotropic tensor products of Lagrange polynomials on the points of a
-Gauss--Lobatto quadrature formula. This change leads to different entries in
-the matrices and constraints, for example, but as the resulting polynomial
-space spans the same polynomials, no change in accuracy should be expected.
-\item We made different changes to the repartitoning of \texttt{parallel::distributed::Triangulation}. Most notably, we have have
-removed the default weight (10,000) for consistency reasons with the rest
+are anisotropic tensor products of Lagrange polynomials on the points of the
+Gauss--Lobatto quadrature formula. This change leads to different entries, for example, in
+the matrices and constraints, but no change in accuracy should be expected as the resulting polynomial
+space spans the same polynomials.
+\item We made different changes in the repartitioning of \texttt{parallel::distributed::Triangulation}. Most notably, we have
+removed the default weight (10,000) in order to ensure consistency with the rest
 of the library.

### 2022-04-26 9.4/paper.tex
Add code gallery section back

@@
   \item Performance improvement of particle infrastructure (see Section~\ref{sec:particles});
-  \item Two new tutorial programs (see Section~\ref{subsec:steps}).
+  \item Two new tutorial programs and a new code gallery program (see Section~\ref{subsec:steps}).
 \end{itemize}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{New and improved tutorials}
+\subsection{New and improved tutorials and code gallery programs}
 \label{subsec:steps}
@@
 
-%There is also a new program in the code gallery (a collection of
-%user-contributed programs that often solve more complicated problems
-%than tutorial programs, and intended as starting points for further
-%research rather than as teaching tools):
-%\begin{itemize}
-%\item ``\texttt{TODO}'' TODO.
-%\end{itemize}
+There is also a new program in the code gallery (a collection of
+user-contributed programs that often solve more complicated problems
+than tutorial programs, and intended as starting points for further
+research rather than as teaching tools):
+\begin{itemize}
+\item ``\texttt{MCMC for the Laplace equation}'' TODO.
+\end{itemize}
 

### 2022-05-03 9.4/paper.tex
Removed value of weight. Not important in this context. (Default value is 1000.)
Adjust indent.
Added bullet-point for pipe_junction.

@@
   limited by CFL conditions, as done in \cite{mirzadeh2016parallel}.
+  \item The new function
+  \texttt{GridGenerator::pipe\_junction}
+  generates a triangulation of three cone-shaped pipes that cross at a bifurcation point in any possible configuration.
+  A manifold description will be applied to the boundary, which can be extended into the volume via transfinite interpolation \cite{Gordon82} using the \texttt{TransfiniteInterpolationManifold} class \cite{dealII90}.
+  \todo[inline]{[MF] If desired, I can extend this part into its own section of the main chapter. With figures, this would probably occupy about 3/4 of a page.}
 \end{itemize}
@@
   space spans the same polynomials.
-\item We made different changes in the repartitioning of \texttt{parallel::distributed::Triangulation}. Most notably, we have
-removed the default weight (10,000) in order to ensure consistency with the rest
-of the library.
-
+  \item We made different changes in the repartitioning of \texttt{parallel::distributed::Triangulation}.
+  Most notably, we have removed the default weight in order to ensure consistency with the rest of the
+  library.
 \end{itemize}

### 2022-05-25 9.4/paper.tex
Leave a placeholder for another topic.
Add name

@@
     Wolfgang Bangerth,
+    Marco Feder,
     Marc Fehling,
@@
 
+   \author[4]{Marco~Feder}
+\affil[4]{SISSA,
+   International School for Advanced Studies,
+   Via Bonomea 265,
+   34136, Trieste, Italy.
+   {\texttt{ \{marco.feder,luca.heltai\}@sissa.it}}}
+
 \author[2]{Marc~Fehling}
 
-\author[6]{Timo~Heister}
- \affil[6]{School of Mathematical and Statistical Sciences,
+\author[5]{Timo~Heister}
+ \affil[5]{School of Mathematical and Statistical Sciences,
    Clemson University,
@@
 
-\author[7]{Luca~Heltai}
-\affil[7]{SISSA,
-   International School for Advanced Studies,
-   Via Bonomea 265,
-   34136, Trieste, Italy.
-   {\texttt{luca.heltai@sissa.it}}}
+\author[4]{Luca~Heltai}
 
- \author[9,10]{Martin~Kronbichler}
- \affil[9]{Institute for Computational Mechanics,
+ \author[6,7]{Martin~Kronbichler}
+ \affil[6]{Institute for Computational Mechanics,
    Technical University of Munich,
@@
    {\texttt{kronbichler/munch@lnm.mw.tum.de}}}
- \affil[10]{Department of Information Technology,
+ \affil[7]{Department of Information Technology,
    Uppsala University,
@@
 
-\author[11]{Matthias~Maier}
-\affil[11]{Department of Mathematics,
+\author[8]{Matthias~Maier}
+\affil[8]{Department of Mathematics,
   Texas A\&M University,
@@
 
-\author[9,12]{Peter Munch}
- \affil[12]{Institute of Material Systems Modeling,
+\author[6,9]{Peter Munch}
+ \affil[9]{Institute of Material Systems Modeling,
  Helmholtz-Zentrum Hereon,
@@
 
-\author[13]{Jean-Paul~Pelteret}
-\affil[13]{Independent researcher.
+\author[10]{Jean-Paul~Pelteret}
+\affil[10]{Independent researcher.
 {\texttt{jppelteret@gmail.com}}}
@@
 
-\author[15]{David Wells}
-\affil[15]{Department of Mathematics, University of North Carolina,
+\author[11]{David Wells}
+\affil[11]{Department of Mathematics, University of North Carolina,
   Chapel Hill, NC 27516, USA.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Topic 6}
+\subsection{Improvements to unstructured communication}
+
+\todo[inline]{Wolfgang to write something about the CA algorithms (and to give
+credit to the previous implementation by Peter).}
 
@@
 Ivo	Dravins,
-Marco	Feder,
 Niklas	Fehn,

### 2022-05-25 9.4/paper.tex
Leave a note.

--- 2022-05-03 9.4/paper.tex
@@
 For sending/receiving, we apply consensus-based algorithms~\cite{hoefler2010scalable}, which
-we introduced into the library in release 9.2~\cite{dealII92}. Consensus-based 
+we introduced into the library in release 9.2~\cite{dealII92} -- see
+also Section~\ref{sec:CA}. Consensus-based
 algorithms are used also in \cite{ibanez2016pumi} for repartitioning.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Topic 6}
+\subsection{Improvements to unstructured communication}
+\label{sec:CA}
+
+\todo[inline]{Wolfgang to write something about the CA algorithms (and to give
+credit to the previous implementation by Peter).}
+
+Also reference Section~\ref{sec:repartitioning}.
 

### 2022-05-26 9.4/paper.tex

--- 2022-05-25 9.4/paper.tex
@@
 \subsection{Improvements to unstructured communication}
+\label{sec:CA}
 
@@
 credit to the previous implementation by Peter).}
+
+Also reference Section~\ref{sec:repartitioning}.
 

### 2022-05-26 9.4/paper.tex
Leave a few comments.

@@
 %
+\todo[inline]{Update}
 \begin{itemize}
@@
   A manifold description will be applied to the boundary, which can be extended into the volume via transfinite interpolation \cite{Gordon82} using the \texttt{TransfiniteInterpolationManifold} class \cite{dealII90}.
-  \todo[inline]{[MF] If desired, I can extend this part into its own section of the main chapter. With figures, this would probably occupy about 3/4 of a page.}
+  \todo[inline]{[MF] If desired, I can extend this part into its own
+    section of the main chapter. With figures, this would probably
+    occupy about 3/4 of a page.}
+  \item \todo[inline]{Add something about the FEInterfaceValues improvements.}
 \end{itemize}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Integration with the Computational Geometry Algorithms
+  Library (CGAL)}\label{sec:cgal}
+
+\begin{itemize}
+\item new classes
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Performance improvement of particle infrastructure}\label{sec:particles}
@@
 \todo[inline]{Wolfgang to write something about the CA algorithms (and to give
-credit to the previous implementation by Peter).}
+credit to the previous implementation by Peter). Mention improvements
+to pack(), unpack(), broadcast() too.}
 

### 2022-05-26 9.4/paper.tex
More small improvements.

@@
 %
-The changelog lists more than 150 other features and bugfixes.
+The changelog lists more than 175 other features and bugfixes.
+\todo{This number is up to date.}
 
@@
 \dealii{} can interface with many other libraries:
+\todo[inline]{Need to update, in particular add CGAL.}
 \begin{multicols}{3}
@@
 The two previous releases of \dealii{} can be cited as
-\cite{dealII91,dealII92}.
+\cite{dealII92,dealII93}.
 
@@
 %
-% 9.4: updated 23/04/2022 PM
+\todo[inline]{updated 23/04/2022 PM. We should probably make sure that
+  nobody else came after that and before the release.}
 Pasquale	Africa,
@@
 variety of funding sources:
+
+\todo[inline]{Update. In particular make sure that we remove those who
+  are not authors on this paper.}
 

### 2022-05-30 9.4/paper.tex

@@
 \usepackage[normalem]{ulem}
+
+\usepackage{caption}
+\usepackage{subcaption}
 
@@
     Jean-Paul Pelteret,
+    Simon Sticko,
     Bruno Turcksin,
@@
    Box 337, 751\,05 Uppsala, Sweden.
-   {\texttt{martin.kronbichler@it.uu.se}}}
+   {\texttt{martin.kronbichler/simon.sticko@it.uu.se}}}
 
@@
 {\texttt{jppelteret@gmail.com}}}
+
+\author[7]{Simon~Sticko}
 
@@
 
+Several classes have been added to the \texttt{NonMatching} namespace to enable the use of cut finite element methods, where the domain is immersed in the background mesh, as illustrated in Figure~\ref{fig:immersed-domain}.
+Many of the new classes assume that the domain, $\Omega$, is described by a level set function,
+$\psi : \mathbb{R}^d \to \mathbb{R}$, such that
+\begin{align}
+  \Omega = \{x \in \mathbb{R}^d : \psi(x)<0\},
+  \qquad
+  \Gamma = \partial \Omega = \{x \in \mathbb{R}^d : \psi(x) = 0\}.
+\end{align}
 \begin{itemize}
-\item new classes
+  \item The \texttt{MeshClassifier} class identifies how the active cells and faces are located relative to the zero contour of the level set function, as illustrated in Figure~\ref{fig:location-to-level-set}.
+Its \texttt{location\_to\_level\_set} function takes a cell/face and
+        returns an enum, \texttt{ LocationToLevelSet}, which takes values \{\texttt{inside}, \texttt{outside}, \texttt{intersected}\}.
+        These represent that $\psi$ is negative, positive, or varies in sign over the cell/face.
+        That is, the cell/face is located inside/outside $\Omega$, or is intersected by $\Gamma$.
+        This information is typically needed when choosing what element a cell of the \texttt{DoFHandler} should use.
+
+  \item The \texttt{QuadratureGenerator} class, which implements the algorithm in \cite{saye2015}, generates high-order quadrature rules for the 3 different regions of a \texttt{BoundingBox}, $B$, defined by the sign of the level set function:
+        \begin{align}
+          B \cap \Omega = \{ x\in B:  \psi(x) < 0 \},  \quad
+          B \cap \Gamma = \{ x\in B:  \psi(x) = 0 \},  \quad
+          \{ x\in B:  \psi(x) > 0 \}.
+        \end{align}
+        An example of these quadratures is shown in Figure~\ref{fig:immersed_quadratures}.
+        Correspondingly, the \newline \texttt{FaceQuadratureGenerator} class generates the same type of quadratures but over a face of $B$.
+        Furthermore, the new classes \texttt{DiscreteQuadratureGenerator} and \newline \texttt{DiscreteFaceQuadratureGenerator} can be used to generate these quadrature rules over a cell/face when the level set function lies in a finite element space: $\psi_h \in V_h$, and when the reference cell of the cell/face is a hypercube.
+
+  \item \texttt{ImmersedSurfaceQuadrature} is a class representing a quadrature rule over a $(d-1)$-dimensional surface embedded in $\mathbb{R}^d$ ($\psi = 0$ in Figure~\ref{fig:immersed_quadratures}).
+In addition to the weight, it stores the unit normal to the surface, for each quadrature point.
+This is needed to be able to transform the quadrature rule from reference space to real space.
+
+  \item \texttt{FEImmersedSurfaceValues} is a \texttt{FEValues}-like class for evaluating real space values based on an \texttt{ImmersedSurfaceQuadrature}.
+
+  \item \texttt{NonMatching::FEValues} combines the functionality of several of the above classes to simplify assembling. It works similarly to \texttt{hp::FEValues}.
+  When calling the \texttt{reinit} function, the immersed quadrature rules are generated in the background and 
+  \texttt{dealii::FEValues} objects for the inside/outside region and a \texttt{FEImmersedSurfaceValues} object for the surface regions are set up internally.
+These can then be obtained using getter-functions and used for the assembly.
+  Since the generation of immersed quadrature rules is not for free,
+  \texttt{NonMatching::FEValues} calls  \texttt{QuadratureGenerator} only if it is needed, i.e.
+if the cell is intersected.
+If the cell passed to \texttt{reinit} is not intersected, already cached \texttt{dealii::FEValues} objects will be returned by the getter functions.
+  Correspondingly, the class \texttt{NonMatching::FEInterfaceValues} generates
+  \texttt{dealii::FEInterfaceValues} objects for assembling face terms of the type $F \cap \{x : \psi(x) < 0 \}$ or $F \cap \{x : \psi(x) > 0 \}$.
 \end{itemize}
+
+\begin{figure}[h]
+  \centering
+  \begin{subfigure}[b]{0.3\textwidth}
+    \centering
+    \includegraphics[height=.12\paperheight]{svg/immersed-domain.pdf}
+    \caption{\label{fig:immersed-domain}}
+  \end{subfigure}
+  \qquad
+  \begin{subfigure}[b]{0.3\textwidth}
+    \centering
+    \includegraphics[height=.12\paperheight]{svg/location-to-level-set.pdf}
+    \caption{ \label{fig:location-to-level-set}}
+  \end{subfigure}
+  \caption{a) Domain immersed in a background mesh. b) Value of \texttt{LocationToLevelSet} for each cell}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.4\paperwidth]{svg/immersed_quadratures.pdf}
+  \caption{Quadrature points for integrating over the three different regions of a cell cut by the zero contour of the level set function, $\psi$. \label{fig:immersed_quadratures}}
+\end{figure}
 
@@
 
+S.~Sticko was partially supported by eSSENCE of e-Science.
+
 D.~Wells was supported by NSF through Grant DMS-1344962.

### 2022-06-01 9.4/paper.tex

@@
 
-Several classes have been added to the \texttt{NonMatching} namespace to enable the use of cut finite element methods, where the domain is immersed in the background mesh, as illustrated in Figure~\ref{fig:immersed-domain}.
-Many of the new classes assume that the domain, $\Omega$, is described by a level set function,
+Several classes have been added to the \texttt{NonMatching} namespace to enable the use of cut finite element methods\cite{burman_cutfem_2015}.
+In the literature, these types of methods are also referred to as immersed, extended, or fictitious finite element methods.
+Here, the domain, $\Omega$, is immersed in the background mesh, as illustrated in Figure~\ref{fig:immersed-domain}.
+Often, one solves for the degrees of freedom of the smallest submesh which completely covers the domain.
+The bilinear form in the weak form can, for example, look like
+\begin{equation}
+  a(u,v) = (\nabla u, \nabla v)_\Omega - (\partial_n u, v)_\Gamma + \ldots
+\end{equation}
+Thus, when assembling, we are on each cell
+required to integrate over the part of the domain and the part of the boundary, $\Gamma = \partial \Omega$, that falls inside the cell:
+$K\cap \Omega$ and $K \cap \Gamma$.
+Many of the new classes assume that the domain is described by a level set function,
 $\psi : \mathbb{R}^d \to \mathbb{R}$, such that
@@
   \qquad
-  \Gamma = \partial \Omega = \{x \in \mathbb{R}^d : \psi(x) = 0\}.
+  \Gamma = \{x \in \mathbb{R}^d : \psi(x) = 0\}.
 \end{align}
@@
 \end{itemize}
+The new \texttt{step-85} tutorial shows how many of these classes go together.
 

### 2022-06-01 9.4/paper.tex
Update my affiliations + some text

@@
  \author[6,7]{Martin~Kronbichler}
- \affil[6]{Institute for Computational Mechanics,
-   Technical University of Munich,
-   Boltzmannstr.~15, 85748 Garching, Germany.
-   {\texttt{kronbichler/munch@lnm.mw.tum.de}}}
- \affil[7]{Department of Information Technology,
+ \affil[6]{Department of Information Technology,
    Uppsala University,
@@
    {\texttt{martin.kronbichler/simon.sticko@it.uu.se}}}
+ \affil[7]{Institute of Mathematics,
+   University of Augsburg,
+   Universit\"atsstr.~12a, 86159 Augsburg, Germany.
+   {\texttt{martin.kronbichler@uni-a.de}}}
 
@@
 
-\author[6,9]{Peter Munch}
+\author[7,9]{Peter Munch}
  \affil[9]{Institute of Material Systems Modeling,
@@
 
-\author[7]{Simon~Sticko}
+\author[6]{Simon~Sticko}
 
@@
 (\texttt{parallel::\allowbreak distributed::\allowbreak Triangulation}; short \texttt{p::d::T}). For the latter, we
-use the Morton-order as the backend of \texttt{p::d::T}---\texttt{p4est}---does. A fixed Morton-order
+use the Morton-order as the backend \texttt{p4est} does. A fixed Morton-order
 might provide very good performance in terms in communication and
@@
 
-The setup process is visualized in Figure~\ref{fig:repartitioning}. At first, cells, incl. their
+The setup process is visualized in Figure~\ref{fig:repartitioning}. At first, cells, including their
 surrounding (ghost) cells and parent cells, are collected and
 sent to the new owner. On the
-receiving side, the sets of all cells are combined and possible duplicates
+receiving site, the sets of all cells are combined and possible duplicates
 are removed. This information is enough to construct a new triangulation.
@@
 \begin{itemize}
-\item Improved support of Hessians: just like values and gradients, Hessians can be
+\item Improved support of computations with Hessians of shape functions: just like values and gradients, Hessians can be
 now evaluated and integrated during matrix-free loops
@@
 recategorization in each time step. Doing this on the fly might be a cheep alternative, even if not computationally optimal during matrix-free loops.
+\item Initial support for H(div)-conforming elements with Piola transform, based on Raviart--Thomas finite element spaces with the updated class \texttt{FE\_RaviartThomasNodal}, has been added.
+The features are currently limited to meshes in standard orientations and affine geometries.
+Full support will be provided in the next release.
+\item Selected matrix-free algorithms can now exploit additional data locality
+  between the matrix-vector product and vector operations happening nearby in
+  an algorithms. These have been added to both the
+  \texttt{PreconditionChebyshev} class often used as a smoother in multigrid
+  methods and the conjugate gradient implementation in \texttt{SolverCG}. To
+  use these features, an operator needs to define a \texttt{vmult} operation,
+  taking two additional \texttt{std::function} arguments. The first function
+  defines what should be scheduled on the vector entries before the
+  matrix-vector product touches them, and the second what happens
+  afterwards. The new features also include a renumbering to maximize data
+  locality. The theory is described in the
+  contribution~\cite{kronbichler2022cg}.
+\todo[inline]{We could also make this last a separate section or in the introduction, but it feels it fits well here.}
 \end{itemize}
@@
 in the case of local smoothing so that global coarsening is favorable despite
-potentially more expensive intergrid transfers. In order, to judge the benefits
+potentially more expensive intergrid transfers. In order to judge the benefits
 of one approach against the other, \texttt{deal.II} provides new functions
@@
 
-Several classes have been added to the \texttt{NonMatching} namespace to enable the use of cut finite element methods\cite{burman_cutfem_2015}.
+Several classes have been added to the \texttt{NonMatching} namespace to enable the use of cut finite element methods~\cite{burman_cutfem_2015}.
 In the literature, these types of methods are also referred to as immersed, extended, or fictitious finite element methods.
@@
 Its \texttt{location\_to\_level\_set} function takes a cell/face and
-        returns an enum, \texttt{ LocationToLevelSet}, which takes values \{\texttt{inside}, \texttt{outside}, \texttt{intersected}\}.
-        These represent that $\psi$ is negative, positive, or varies in sign over the cell/face.
+        returns an enum, \texttt{LocationToLevelSet}, which takes values \{\texttt{inside}, \texttt{outside}, \texttt{intersected}\}.
+        These indicate whether $\psi$ is negative, positive, or varies in sign over the cell/face.
         That is, the cell/face is located inside/outside $\Omega$, or is intersected by $\Gamma$.
@@
   space spans the same polynomials.
+  \item The class \texttt{MappingQ} now applies a high-order mapping to all cells, not just the cells near the boundary, functionality that was previously provided by the \texttt{MappingQGeneric} class.
+The latter has been marked as deprecated.
+In order to use different degrees for the mapping, users can use \texttt{hp::MappingCollection} or \texttt{MappingFEField} with appropriate definitions of function spaces.
   \item We made different changes in the repartitioning of \texttt{parallel::distributed::Triangulation}.
@@
 (KONWIHR) in the context of the projects
-``Performance tuning of high-order discontinuous Galerkin solvers for
-SuperMUC-NG'' and ``High-order matrix-free finite element implementations with
-hybrid parallelization and improved data locality''.
+``High-order matrix-free finite element implementations with
+hybrid parallelization and improved data locality'' and ``Fast and scalable finite element algorithms for coupled multiphysics problems and non-matching grids''.
 

### 2022-06-14 9.4/paper.tex
Add a second affiliation.

@@
 
-\author[6]{Simon~Sticko}
+\author[6,11]{Simon~Sticko}
+\affil[11]{Department of Mathematics and Mathematical Statistics,
+   Umeå University,
+   SE-90187 Umeå, Sweden}
 
@@
   \item CutFEM support (see Section~\ref{sec:cut});
+  \item Integration with the Computational Geometry Algorithms Library (CGAL) (see Section~\ref{sec:cgalwrappers});
   \item Performance improvement of particle infrastructure (see Section~\ref{sec:particles});
@@
 \subsection{Integration with the Computational Geometry Algorithms
-  Library (CGAL)}\label{sec:cgal}
-
-\begin{itemize}
-\item new classes
-\end{itemize}
+  Library (CGAL)}\label{sec:cgalwrappers}
+  If you have CGAL (\href{https://www.cgal.org/}{{https://www.cgal.org/}}) installed \cite{cgal}, you can use wrappers around CGAL classes and functions contained in the new namespace \texttt{CGALWrappers}, allowing capabilities spanning from mesh generation to boolean operations between \dealii{} triangulations and cells.
+CGAL wrappers are enabled only if \dealii{} is compiled with \texttt{C++17}.
+We have introduced the following utilities to the \dealii's \texttt{CGALWrappers} namespace.
+\begin{itemize}
+\item Refinement criteria for tetrahedra or surface facets can be stored and specified in a struct named \texttt{AdditionalData}. 
+Depending on the template parameter \texttt{dim}, different criteria can be chosen. See the documentation of this struct for a list and an explanation of the possible parameters.
+\item \texttt{GridGenerator::implicit\_function()} creates a \texttt{Triangulation<dim,3>} out of the zero level set of an implicit function $f$. 
+Apart from the final \texttt{Triangulation<dim,3>} and the implicit funcion $f$ itself, also an interior point for which $f<0$ must be supplied by the user, along with the radius of a ball that will contain the generated triangulation.
+Mesh-related entities must be specified with \texttt{AdditionalData<dim>}.
+When \texttt{dim} is \texttt{3}, the mesh is filled with tetrahedra.
+A prototypical use case is the following, where the surface is the zero level set of Taubin's heart function $f=\bigl ( x^2 + \frac{9y^2}{4} +z^2 -1 \bigr ) -x^2 z^3 - \frac{9y^2z^3}{80}$.
+The resulting \texttt{Triangulation<3>} can be appreciated in Figure~\ref{fig:heart_tria}.
+\begin{c++}
+// Empty triangulation.
+Triangulation<3, 3>  tria;
+//Taubin's heart surface
+ImplicitFunction                implicit_function;
+CGALWrappers::AdditionalData<3> data;
+data.cell_size = .05;
+GridGenerator::implicit_function(
+  tria, implicit_function, data, Point<3>(0, 0, 0), 10.0);
+\end{c++}
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=.6\paperwidth]{png/heart_implicit.png}
+  \caption{\texttt{Triangulation<3>} of a heart surface starting from an implicit function $f$. \label{fig:heart_tria}}
+\end{figure}
+
+\item \texttt{GridGenerator::surface\_mesh\_to\_volumetric\_mesh()} fills a \texttt{Triangulation<3>} starting from a \texttt{Triangulation<2,3>}.  
+The surface triangulation bounds the three dimensional volume, which is meshed inside with tetrahedra.
+\item Corefinement and boolean operations between \texttt{CGAL::Surface\_mesh} types are exploited to perform boolean operations between \dealii{}
+triangulations, and are available in the utility \texttt{CGALWrappers::compute\_boolean\_operation()}.
+The available operations are \texttt{corefinement}, \texttt{intersection}, \texttt{union} and \texttt{difference}, and they are
+described in the enum class \texttt{BooleanOperation}.
+Also, several wrappers allow to go back and forth from a \dealii{} Triangulation type to a CGAL Triangulation type.
+In general, after a \texttt{BooleanOperation} the corefinement around the intersection gives most of the times
+badly shaped elements. To overcome this issue, one can use \texttt{CGALWrappers::remesh\_surface()}. See Figure~\ref{fig:corefinement_remeshed} for a graphical example.
+A possible workflow is the following:
+\begin{c++}
+  Triangulation<spacedim> tria0, tria1;
+  GridGenerator::hyper_cube(tria0,-1.,1.);
+  GridGenerator::hyper_ball(tria1,{1.,1.,1.},0.6);
+  tria0.refine_global(3);
+  tria1.refine_global(3);
+
+  // Move to CGAL surfaces, assume Kernel is already defined. 
+  CGAL::Surface_mesh<Kernel> surface_mesh0, surface_mesh1;
+  CGALWrappers::dealii_tria_to_cgal_surface_mesh(tria0, surface_mesh0);
+  CGALWrappers::dealii_tria_to_cgal_surface_mesh(tria1, surface_mesh1);
+
+  // Compute the union of the two meshes
+  CGALWrappers::compute_boolean_operation(surface_mesh0,
+                            surface_mesh1,
+                            BooleanOperation::compute_union,
+                            out_mesh);
+  // Now back to deal.II
+  Triangulation<2, 3>     tria_out;
+  CGALWrappers::cgal_surface_mesh_to_dealii_triangulation(out_mesh, tria_out);
+  // tria_out is now a valid deal.II triangulation
+  \end{c++}
+  The output of the boolean operation can be appreciated in Figure~\ref{fig:corefinement}, while in Figure~\ref{fig:corefinement_remeshed}
+  the same mesh has been remeshed.
+  \begin{figure}[h]
+    \centering
+    \begin{subfigure}[b]{0.4\textwidth}
+      \centering
+      \includegraphics[height=.2\paperheight]{png/intersection_cube_sphere_mesh.png}
+      \caption{\label{fig:corefinement}}
+    \end{subfigure}
+    \qquad
+    \begin{subfigure}[b]{0.5\textwidth}
+      \centering
+      \includegraphics[height=.2\paperheight]{png/cube_sphere_remeshed.png}
+      \caption{ \label{fig:corefinement_remeshed}}
+    \end{subfigure}
+    \caption{a) Union of the unit cube with an hyper ball. Notice how the corefinement around the intersection gives badly shaped elements. b) The remeshed \dealii{} Triangulation.}
+  \end{figure}
+
+\item \texttt{CGALWrappers::compute\_quadrature\_on\_boolean\_operation()} returns a \texttt{Quadrature<3>} that allows to integrate exactly on polyhedral elements coming out of a \texttt{BooleanOperation} between \dealii cells.
+The quadrature rule is built by meshing the polyhedral region with tetrahedra, computing on each tetrahedron a \texttt{QGaussSimplex<3>} quadrature rule by using \texttt{QSimplex<3>::compute\_affine\_transformation()}, and finally
+collecting all of the rules together, giving a \texttt{Quadrature<3>} formula on the \emph{physical} element.
+\end{itemize}
+
+These utilities will be the building blocks for adding functions to the \texttt{NonMatching} namespace that will assemble coupling terms like $(u,v)_{B}$, with $B$ a domain immersed in a fixed background mesh $\Omega$ and $u,v$ finite element functions on $V_h(\Omega)$, as usually happens using Nitsche's method to weakly impose boundary conditions at an interface.
+The same applies to coupling terms of the form $(u,q)_B$ in formulations using Lagrange multipliers, where now $q \in Q_h(B)$, with $Q_h(B)$ the space of the multiplier variable.
 
@@
 \dealii{} can interface with many other libraries:
-\todo[inline]{Need to update, in particular add CGAL.}
 \begin{multicols}{3}
@@
     \item BLAS and LAPACK \cite{lapack}
+    \item CGAL \cite{cgal}
     \item cuSOLVER \cite{cusolver}

### 2022-06-20 9.4/paper.tex
Start the editing process.
Also write the communications section.

@@
 \dealii{} version 9.4.0 was released June 10, 2022.
+\todo{Update release date.}
 This paper provides an
@@
 %
-\todo[inline]{Update}
 \begin{itemize}
@@
   \item Integration with the Computational Geometry Algorithms Library (CGAL) (see Section~\ref{sec:cgalwrappers});
-  \item Performance improvement of particle infrastructure (see Section~\ref{sec:particles});
-  \item Two new tutorial programs and a new code gallery program (see Section~\ref{subsec:steps}).
+  \item Performance improvements in the particle infrastructure (see Section~\ref{sec:particles});
+  \item Two new tutorial programs and a new code gallery program (see
+    Section~\ref{subsec:steps}).
+    \todo[inline]{Update the number of code gallery programs when
+      we're done.}
 \end{itemize}
@@
   By using this class, one can output the result obtained on an
-  unstructured mesh on a structured one, which might be a more
-  memory-efficient storage format, or one can create a slice in 3D.
+  unstructured mesh on a structured one (which might facilitate a more
+  memory-efficient storage format, for example if this second
+  triangulation is a uniformly refined rectangle or box), or one can create a slice in 3D.
   \item The new function \texttt{find\_point\_owner\_rank()} of \texttt{parallel::distributed::Triangulation} allows to find the MPI
@@
   Its algorithm is described in \cite{burstedde2020parallel}. Based on the information obtained,
-  communication pattern in \texttt{Utilties::MPI::RemotePointEvaluation} can be set up efficiently. Furthermore, this function could be used in the future to allow
+  the communication pattern used in \texttt{Utilties::MPI::RemotePointEvaluation} can be set up efficiently. Furthermore, this function could be used in the future to allow
   particle simulations where particle movement is not
@@
   \item The new function
-  \texttt{GridGenerator::pipe\_junction}
+  \texttt{GridGenerator::pipe\_junction()}
   generates a triangulation of three cone-shaped pipes that cross at a bifurcation point in any possible configuration.
-  A manifold description will be applied to the boundary, which can be extended into the volume via transfinite interpolation \cite{Gordon82} using the \texttt{TransfiniteInterpolationManifold} class \cite{dealII90}.
-  \todo[inline]{[MF] If desired, I can extend this part into its own
-    section of the main chapter. With figures, this would probably
-    occupy about 3/4 of a page.}
-  \item \todo[inline]{Add something about the FEInterfaceValues improvements.}
+  A manifold description is applied to the boundary, which can be extended into the volume via transfinite interpolation \cite{Gordon82} using the \texttt{TransfiniteInterpolationManifold} class \cite{dealII90}.
+  \item \todo[inline]{Add something about the FEInterfaceValues improvements?}
 \end{itemize}
@@
 The changelog lists more than 175 other features and bugfixes.
-\todo{This number is up to date.}
 
@@
 
-We continued to work on the simplex- and mixed-mesh support. We fixed
-many bugs and generalized existing functions that only worked for
-hypercube-shaped cells. The most notable new functions are:
-
-\begin{figure}[!t]
+\dealii{} has supported simplex and mixed meshes since the previous
+release, 9.3, on an experimental basis.
+We have continued to work on this support; the current state is
+sufficient to support several larger-scale programs, although not all
+functionality in \dealii{} works with such meshes yet. Specifically,
+we have fixed many bugs and generalized existing functions that
+previously only worked for hypercube-shaped cells. The most notable
+new functionality is:
+
+\begin{figure}
 
@@
 
+  \phantom{.}
+  \hfill
   \begin{tikzpicture}[scale=1.7]
@@
   \end{tikzpicture}
-  \qquad\qquad
+  \hfill\hfill
   \begin{tikzpicture}[scale=1.3]
@@
   \end{tikzpicture}
-
-  \caption{New refinement strategies for triangles and tetrahedrons: triangles
-  are subdivided in 4 children and tetrahedrons in 8 ones.}\label{fig:refinement}
-
+  \hfill
+  \phantom{.}
+
+  \caption{\it New refinement strategies for triangles and tetrahedra: Triangles
+    are subdivided into four children and tetrahedra into eight.
+  \todo[inline]{Are these ``new strategies'', or is it ``New:
+    Refinement strategies'' because we did not have refinement at all before?}}
+  \label{fig:refinement}
 \end{figure}
@@
 \begin{itemize}
-\item Experimental support of locally refined meshes. FEM on locally refined
-meshes requires 1) the possibility to locally refine the mesh (see Figure~\ref{fig:refinement})
-and 2) appropriate hanging-node constraints. For 3D, the implementation of the missing constraint definitions is in progress.
-\item TODO
-\end{itemize}
-
-Furthermore, we continued to remove the instance of usage of the \texttt{GeometryInfo} class from
-the library and to replace them by more general equivalent functions, which
-rely on \texttt{ReferenceCell} in many cases. Once all instances of
+\item Experimental support for locally refined meshes: For finite
+  element computations on locally refined
+meshes, one needs (i) the possibility to locally refine the mesh (see
+Figure~\ref{fig:refinement}), and (ii) appropriate hanging-node
+constraints. Both are now available in 2d; for 3D, the implementation
+of the constraint definitions is still in progress.
+\item \todo[inline]{TODO}
+\end{itemize}
+
+Furthermore, we have continued to remove uses of the
+\texttt{GeometryInfo} class (which is specific to hypercube cells) from
+the library, and to replace them by more general equivalent
+functionality based on the
+\texttt{ReferenceCell} class. Once all instances of
 \texttt{GeometryInfo} are removed, we will deprecate the class.
@@
     \input{svg/repartitioning.pdf_tex}
-    \caption{Visualization of the repartitioning process: after new ranks
+    \caption{\it Visualization of the repartitioning process: after new ranks
     are assigned to cells, each process collects and sends the cells, incl.
@@
   \end{subfigure}
-  \caption{a) Domain immersed in a background mesh. b) Value of \texttt{LocationToLevelSet} for each cell}
+  \caption{\it (a) Domain immersed in a background mesh. (b) Value of \texttt{LocationToLevelSet} for each cell}
 \end{figure}
@@
   \includegraphics[width=.4\paperwidth]{svg/immersed_quadratures.pdf}
-  \caption{Quadrature points for integrating over the three different regions of a cell cut by the zero contour of the level set function, $\psi$. \label{fig:immersed_quadratures}}
+  \caption{\it Quadrature points for integrating over the three different regions of a cell cut by the zero contour of the level set function, $\psi$. \label{fig:immersed_quadratures}}
 \end{figure}
@@
   \includegraphics[width=.6\paperwidth]{png/heart_implicit.png}
-  \caption{\texttt{Triangulation<3>} of a heart surface starting from an implicit function $f$. \label{fig:heart_tria}}
+  \caption{\it \texttt{Triangulation<3>} of a heart surface starting from an implicit function $f$. \label{fig:heart_tria}}
 \end{figure}
@@
     \end{subfigure}
-    \caption{a) Union of the unit cube with an hyper ball. Notice how the corefinement around the intersection gives badly shaped elements. b) The remeshed \dealii{} Triangulation.}
+    \caption{\it (a) Union of the unit cube with an hyper ball. Notice how the corefinement around the intersection gives badly shaped elements. (b) The remeshed \dealii{} Triangulation.}
   \end{figure}
@@
 
-\todo[inline]{Wolfgang to write something about the CA algorithms (and to give
-credit to the previous implementation by Peter). Mention improvements
-to pack(), unpack(), broadcast() too.}
-
-Also reference Section~\ref{sec:repartitioning}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Topic 7}
+Many problems in parallel computations can be stated in the following
+way: Each process in a parallel universe has a number of queries to
+send to other processes who do not know that they will be asked, and
+who will then have to respond with replies. This problem is solved by
+``consensus algorithms''. An example of where this problem appears is
+given in Section~\ref{sec:repartitioning}.
+
+\dealii{} has had an implementation of these algorithms for some time,
+but the current version substantially expands it. Specifically, the
+updated interfaces -- now based on function objects such as lambda
+functions to formulate and process queries and replies -- can now deal
+with arbitrary data types for queries and replies, rather than only
+arrays of data types natively supported by MPI. To make this possible,
+the implementation packs and unpacks these objects into character
+arrays via the \texttt{Utilities::pack()} and
+\texttt{Utilities::unpack()} functions. We have also worked on making
+these functions more efficient: if the object to be packed is an array
+(or array of arrays) of elements that satisfy the
+\texttt{std::is\_trivially\_copyable} type trait, then the data is
+copied into the character array via \texttt{std::memcpy}; only for
+other objects do the packing and unpacking functions rely on BOOST's
+serialization library.
+
+A special case of consensus algorithms is where the sender does not
+actually require and answer. This occasionally happens during ghost
+exchange where each process sends information about cells it owns to
+some of its neighboring processes where these cells are ghost
+cells. Previously, this case was implemented by sending an empty
+message. The rewritten interfaces now support this case: Code using
+these interfaces no longer have to provide functions that formulate
+and read the (empty) replies, though internally these functions still
+send around an empty reply; this case will be implemented in the
+future, using the interfaces now already in place.
+
+Finally, in the spirit of optimizing communication, the
+\texttt{Utilities::MPI::broadcast()} function has been optimized for
+objects that are arrays of data types natively supported by MPI.
 

### 2022-06-20 9.4/paper.tex
Some more editing.

@@
 
-Until now, distributed meshes were either partitioned statically (\texttt{parallel::\allowbreak fullydistributed::\allowbreak Triangulation}; short \texttt{p::f::T}) or they used a
-fixed policy to partition the cells among processes
-(\texttt{parallel::\allowbreak distributed::\allowbreak Triangulation}; short \texttt{p::d::T}). For the latter, we
-use the Morton-order as the backend \texttt{p4est} does. A fixed Morton-order
-might provide very good performance in terms in communication and
-setup costs, but might be non-optimal when interfacing with other
-libraries that have fixed partitioning, e.g., a Cartesian
-partitioning, themselves.
-
-New utility functions from the \texttt{RepartitioningPolicyTools} namepace can be
-used now to create a new \texttt{p::f::T} instance,
-given a distributed triangulation (\texttt{p::f::T} or \texttt{p::d::T}) and a vector with the new owners of locally
-owned cells. The workflow is shown in the following listing:
+\dealii{} has two classes that support distributed storage of meshes: 
+\texttt{parallel::\allowbreak distributed::\allowbreak Triangulation}
+(which in the following we will abbreviate as \texttt{p::d::T}) and
+\texttt{parallel::\allowbreak fullydistributed::\allowbreak
+  Triangulation} (or, in short, \texttt{p::f::T}). The former is
+partitioned based on the space-filling Morton (or Z-order) curve
+implemented by the \texttt{p4est} backend. The latter, more recent
+class is currently statically partitioned at creation time.
+
+There are now new utility functions in the
+\texttt{RepartitioningPolicyTools} namepace that can be
+used to create a new \texttt{p::f::T} instance,
+given a distributed triangulation (\texttt{p::f::T} or
+\texttt{p::d::T}) and a vector that indicates the designated owner processes of locally
+owned cells. This approach allows for partitioning \texttt{p::f::T}
+based on arbitrary criteria. The workflow is shown in the following listing:
 \begin{c++}
-// original triangulation
-parallel::distributed::Triangulation<dim> tria(comm);
-
-// select a partitioning policy and use it
+// original triangulation:
+parallel::distributed::Triangulation<dim> tria(communicator);
+
+// select a partitioning policy and use it:
 const RepartitioningPolicyTools::Cartesian<dim> policy(tria);
@@
 
-// create the new triangulation
+// create the new triangulation:
 parallel::fullydistributed::Triangulation<dim> tria_pft(comm);
@@
 \end{c++}
-Instead of using one of the predefined partitioning policies, users
-can write their own ones by implementing the \texttt{RepartitioningPolicyTools::Base} interface. Similarly to the
-active level, also the multigrid levels can be repartitioned arbitrarily.
 
@@
     \input{svg/repartitioning.pdf_tex}
-    \caption{\it Visualization of the repartitioning process: after new ranks
-    are assigned to cells, each process collects and sends the cells, incl.
-    ghost cells, to the new owner. There, incoming cells are processed, duplicates
-    are removed, and the local part of the triangulation is built.}\label{fig:repartitioning}
+    \caption{\it Visualization of the repartitioning process. Top
+      left: Existing ownership of the cells of the mesh, distributed
+      on four processes. Bottom left:
+      Requested new ownership. Second column: Processes 0 and 1
+      collect information about the cells they own or that are ghost
+      cells. Processes 2 and 3 would do the same. Third column: What
+      processes 0 and 1 would send to process 0. Fourth column: The
+      combined knowledge on process 0.}\label{fig:repartitioning}
 \end{figure}
 
-The setup process is visualized in Figure~\ref{fig:repartitioning}. At first, cells, including their
-surrounding (ghost) cells and parent cells, are collected and
+The setup process is visualized in Figure~\ref{fig:repartitioning}. At
+first, locally owned cells and their
+surrounding (ghost) cells are collected on each process and
 sent to the new owner. On the
@@
 we introduced into the library in release 9.2~\cite{dealII92} -- see
-also Section~\ref{sec:CA}. Consensus-based
-algorithms are used also in \cite{ibanez2016pumi} for repartitioning.
+also Section~\ref{sec:CA}. This kind of consensus-based
+algorithm is used also in \cite{ibanez2016pumi} for repartitioning.
+
+In addition to the predefined partitioning policies, users
+can write their own by implementing the
+\texttt{RepartitioningPolicyTools::Base} interface. Like the
+``active'' level of finest cells, the scheme also allows for
+arbitrarily repartitioned multigrid levels.
 
@@
 partitioning libraries, e.g., \texttt{ParMETIS} or \texttt{Zoltan}. Furthermore, we intend to extend
-\texttt{p::f::T} with adaptivity support.
-With this support and and the new repartitioning features, \texttt{p::f::T}
-could be self-consistent like \texttt{p::d::T}, which is an important
-step towards the support of AMR also for distributed simplex and mixed meshes
-(see also Subsection~\ref{sec:simplex}).
-
-%We would like to note that we provide a similar class also for repartitioning
-%of vectors: \texttt{Noncontiguous\allowbreak Partitioner} from the
-%\texttt{Utilities::MPI} namespace.
+\texttt{p::f::T} to support adaptive mesh refinement, completing \texttt{p::f::T}
+as a complement to \texttt{p::d::T}.
 

### 2022-06-21 9.4/paper.tex
More updates.

@@
 
-In the matrix-free infrastructure, we have added a number of new features. The most
-notable ones are:
+The matrix-free infrastructure of \dealii{} is used to solve problems
+without assembling matrices, providing only the action of the matrix
+instead. It has seen a number of new features in this release, the most
+notable ones being:
 \begin{itemize}
-\item Improved support of computations with Hessians of shape functions: just like values and gradients, Hessians can be
+\item Improved support for computations with Hessians of shape functions: just like values and gradients, Hessians can be
 now evaluated and integrated during matrix-free loops
-both for cells and faces. This enables, e.g., to write
+both for cells and faces. This could enable, for example, writing
 a matrix-free version of step-47, which solves the biharmonic equation with DG.
-\item Cell-centric loops now also allow to access gradients and Hessians
-of neighboring cells on faces. The major difficulty here was their reorientation
-in the case of arbitrary orientated cells on unstructured meshes.
+\item Cell-centric loops now also allow access to gradients and Hessians
+of neighboring cells on faces. The major difficulty here was the relative orientation
+of cells on unstructured meshes.
 \item Users can now create their own cell batches, by providing \texttt{FEEvaluation::reinit()} a list of cell IDs. \texttt{FEEvaluation}
@@
 might be too expensive if a very dynamic system is given, which requires
-recategorization in each time step. Doing this on the fly might be a cheep alternative, even if not computationally optimal during matrix-free loops.
-\item Initial support for H(div)-conforming elements with Piola transform, based on Raviart--Thomas finite element spaces with the updated class \texttt{FE\_RaviartThomasNodal}, has been added.
-The features are currently limited to meshes in standard orientations and affine geometries.
-Full support will be provided in the next release.
+recategorization in each time step. Doing this on the fly might be a cheap alternative, even if not computationally optimal during matrix-free loops.
+\item Initial support for H(div)-conforming elements with Piola transform, based on Raviart--Thomas finite element spaces with the updated class \texttt{FE\_RaviartThomasNodal}.
+This feature is currently limited to meshes in standard orientation and affine geometries.
+Full support will be provided in a future release.
 \item Selected matrix-free algorithms can now exploit additional data locality
@@
   contribution~\cite{kronbichler2022cg}.
-\todo[inline]{We could also make this last a separate section or in the introduction, but it feels it fits well here.}
 \end{itemize}
@@
 quasi-dense matrix-vector multiplications~\cite{KronbichlerKormann2012}, we now use an
-approach based on inplace interpolation and sumfactorization, similarly
-as we already did in the GPU code~\cite{ljungkvist2017matrix}. In \cite{munch2022hn}, the algorithm is described
+approach based on in-place interpolation and sum factorization, similar
+to what was already done in the GPU code~\cite{ljungkvist2017matrix}. In \cite{munch2022hn}, the algorithm is described
 and performance numbers are shown, indicating a reduction
-of overhead of cells with hanging nodes by a factor of 10$\times$.
-
-Furthermore, we have performed a major restructuring of internals
-of the \texttt{FEEvaluation} classes. This will enable us to add the support of new element types, e.g., of
-Raviart--Thomas and Nedelec elements,  more
-simply in the future.
+of overhead of cells with hanging nodes by a factor of ten.
+
+Finally, we have performed a major restructuring of internals
+of the \texttt{FEEvaluation} classes. This will enable us to add support for new element types, e.g., of
+Raviart-Thomas and Nedelec elements.
 
@@
 coarsening smoothens on the whole computational domain on each
-multigrid level, which is obtained by coarsening the finest mesh globally.
+multigrid level, which is obtained by coarsening the finest cells of
+the next finer multigrid level.
 For this purpose, we use a sequence of triangulations, and we perform
 the smoothing only on their active levels. To create the sequence of
-triangulations, one can use the functions \texttt{create\_geometric\_coarsening\_sequence()} from the
-\texttt{MGTransferGlobalCoarseningTools} namespace. A new version takes an
+triangulations, one can use the functions \texttt{MGTransferGlobalCoarseningTools::create\_geometric\_coarsening\_sequence()}. A new version takes an
 instance of \texttt{RepartitioningPolicyTools::Base} (see Subsection~\ref{sec:repartitioning}) as argument, which allows to specify the parallel
@@
 
-Several classes have been added to the \texttt{NonMatching} namespace to enable the use of cut finite element methods~\cite{burman_cutfem_2015}.
-In the literature, these types of methods are also referred to as immersed, extended, or fictitious finite element methods.
-Here, the domain, $\Omega$, is immersed in the background mesh, as illustrated in Figure~\ref{fig:immersed-domain}.
-Often, one solves for the degrees of freedom of the smallest submesh which completely covers the domain.
-The bilinear form in the weak form can, for example, look like
-\begin{equation}
-  a(u,v) = (\nabla u, \nabla v)_\Omega - (\partial_n u, v)_\Gamma + \ldots
-\end{equation}
-Thus, when assembling, we are on each cell
-required to integrate over the part of the domain and the part of the boundary, $\Gamma = \partial \Omega$, that falls inside the cell:
-$K\cap \Omega$ and $K \cap \Gamma$.
-Many of the new classes assume that the domain is described by a level set function,
-$\psi : \mathbb{R}^d \to \mathbb{R}$, such that
-\begin{align}
-  \Omega = \{x \in \mathbb{R}^d : \psi(x)<0\},
-  \qquad
-  \Gamma = \{x \in \mathbb{R}^d : \psi(x) = 0\}.
-\end{align}
-\begin{itemize}
-  \item The \texttt{MeshClassifier} class identifies how the active cells and faces are located relative to the zero contour of the level set function, as illustrated in Figure~\ref{fig:location-to-level-set}.
-Its \texttt{location\_to\_level\_set} function takes a cell/face and
-        returns an enum, \texttt{LocationToLevelSet}, which takes values \{\texttt{inside}, \texttt{outside}, \texttt{intersected}\}.
-        These indicate whether $\psi$ is negative, positive, or varies in sign over the cell/face.
-        That is, the cell/face is located inside/outside $\Omega$, or is intersected by $\Gamma$.
-        This information is typically needed when choosing what element a cell of the \texttt{DoFHandler} should use.
-
-  \item The \texttt{QuadratureGenerator} class, which implements the algorithm in \cite{saye2015}, generates high-order quadrature rules for the 3 different regions of a \texttt{BoundingBox}, $B$, defined by the sign of the level set function:
-        \begin{align}
-          B \cap \Omega = \{ x\in B:  \psi(x) < 0 \},  \quad
-          B \cap \Gamma = \{ x\in B:  \psi(x) = 0 \},  \quad
-          \{ x\in B:  \psi(x) > 0 \}.
-        \end{align}
-        An example of these quadratures is shown in Figure~\ref{fig:immersed_quadratures}.
-        Correspondingly, the \newline \texttt{FaceQuadratureGenerator} class generates the same type of quadratures but over a face of $B$.
-        Furthermore, the new classes \texttt{DiscreteQuadratureGenerator} and \newline \texttt{DiscreteFaceQuadratureGenerator} can be used to generate these quadrature rules over a cell/face when the level set function lies in a finite element space: $\psi_h \in V_h$, and when the reference cell of the cell/face is a hypercube.
-
-  \item \texttt{ImmersedSurfaceQuadrature} is a class representing a quadrature rule over a $(d-1)$-dimensional surface embedded in $\mathbb{R}^d$ ($\psi = 0$ in Figure~\ref{fig:immersed_quadratures}).
-In addition to the weight, it stores the unit normal to the surface, for each quadrature point.
-This is needed to be able to transform the quadrature rule from reference space to real space.
-
-  \item \texttt{FEImmersedSurfaceValues} is a \texttt{FEValues}-like class for evaluating real space values based on an \texttt{ImmersedSurfaceQuadrature}.
-
-  \item \texttt{NonMatching::FEValues} combines the functionality of several of the above classes to simplify assembling. It works similarly to \texttt{hp::FEValues}.
-  When calling the \texttt{reinit} function, the immersed quadrature rules are generated in the background and
-  \texttt{dealii::FEValues} objects for the inside/outside region and a \texttt{FEImmersedSurfaceValues} object for the surface regions are set up internally.
-These can then be obtained using getter-functions and used for the assembly.
-  Since the generation of immersed quadrature rules is not for free,
-  \texttt{NonMatching::FEValues} calls  \texttt{QuadratureGenerator} only if it is needed, i.e.
-if the cell is intersected.
-If the cell passed to \texttt{reinit} is not intersected, already cached \texttt{dealii::FEValues} objects will be returned by the getter functions.
-  Correspondingly, the class \texttt{NonMatching::FEInterfaceValues} generates
-  \texttt{dealii::FEInterfaceValues} objects for assembling face terms of the type $F \cap \{x : \psi(x) < 0 \}$ or $F \cap \{x : \psi(x) > 0 \}$.
-\end{itemize}
-The new \texttt{step-85} tutorial shows how many of these classes go together.
-
-\begin{figure}[h]
+\begin{figure}
   \centering
@@
   \end{subfigure}
-  \caption{\it (a) Domain immersed in a background mesh. (b) Value of \texttt{LocationToLevelSet} for each cell}
+  \caption{\it (a) Domain immersed in a background mesh. (b) Value of \texttt{LocationToLevelSet} for each cell.}
 \end{figure}
 
-\begin{figure}[h]
+Several classes have been added to the \texttt{NonMatching} namespace to enable the use of cut finite element methods~\cite{burman_cutfem_2015}.
+In the literature, these types of methods are also referred to as immersed, extended, or fictitious finite element methods.
+Here, the domain, $\Omega$, is immersed in the background mesh, as illustrated in Fig.~\ref{fig:immersed-domain}.
+Often, one solves for the degrees of freedom of the smallest submesh
+which completely covers the domain, i.e., the blue and yellow cells of
+Fig.~\ref{fig:location-to-level-set}.
+The bilinear form in the weak form would then, for example, look like
+\begin{equation}
+  a(u,v) = (\nabla u, \nabla v)_\Omega - (\partial_n u, v)_\Gamma + \ldots
+\end{equation}
+Thus, when assembling on a cut call, we are
+required to integrate over the part of the domain and the part of the boundary, $\Gamma = \partial \Omega$, that falls inside the cell:
+$K\cap \Omega$ and $K \cap \Gamma$.
+Many of the new classes that supports these operations assume that the domain is described by a level set function,
+$\psi : \mathbb{R}^d \to \mathbb{R}$, such that
+\begin{align}
+  \Omega = \{x \in \mathbb{R}^d : \psi(x)<0\},
+  \qquad
+  \Gamma = \{x \in \mathbb{R}^d : \psi(x) = 0\}.
+\end{align}
+Specifically, the following are the key new classes and functions:
+\begin{itemize}
+  \item The \texttt{MeshClassifier} class identifies how the active cells and faces are located relative to the zero contour of the level set function, as illustrated in Figure~\ref{fig:location-to-level-set}.
+Its \texttt{location\_to\_level\_set()} function takes a cell/face and
+        returns an enum, \texttt{LocationToLevelSet}, which takes values \{\texttt{inside}, \texttt{outside}, \texttt{intersected}\}.
+        This information is typically needed when choosing what element a cell of the \texttt{DoFHandler} should use.
+
+  \item The \texttt{QuadratureGenerator} class, which implements the algorithm in \cite{saye2015}, generates high-order quadrature rules for the three different regions of a \texttt{BoundingBox}, $B$, defined by the sign of the level set function:
+        \begin{align}
+          B \cap \Omega = \{ x\in B:  \psi(x) < 0 \},  \quad
+          B \cap \Gamma = \{ x\in B:  \psi(x) = 0 \},  \quad
+          \{ x\in B:  \psi(x) > 0 \}.
+        \end{align}
+        An example of these quadratures is shown in Figure~\ref{fig:immersed_quadratures}.
+        The \texttt{FaceQuadratureGenerator} class does the same for faces.
+        Furthermore, the new classes \texttt{DiscreteQuadratureGenerator} and \newline \texttt{DiscreteFaceQuadratureGenerator} can be used to generate these quadrature rules over a cell/face when the level set function lies in a finite element space: $\psi_h \in V_h$, and when the reference cell of the cell/face is a hypercube.
+
+  \item \texttt{ImmersedSurfaceQuadrature} is a class representing a quadrature rule over a $(d-1)$-dimensional surface embedded in $\mathbb{R}^d$ ($\psi = 0$ in Figure~\ref{fig:immersed_quadratures}).
+In addition to the weight, it stores the unit normal to the surface, for each quadrature point.
+This is needed to transform the quadrature rule from reference space to real space.
+
+  \item \texttt{FEImmersedSurfaceValues} is an \texttt{FEValues}-like class for evaluating real space values based on an \texttt{ImmersedSurfaceQuadrature}.
+
+  \item \texttt{NonMatching::FEValues} combines the functionality of
+    several of the above classes to simplify assembly of linear systems. It works similarly to \texttt{hp::FEValues}:
+  When calling the \texttt{reinit()} function, immersed quadrature rules are generated in the background and
+  \texttt{FEValues} objects for the inside/outside region and a \texttt{FEImmersedSurfaceValues} object for the surface regions are set up internally.
+These can then be obtained using getter-functions and used for the assembly.
+  Since the generation of immersed quadrature rules is not cheap,
+  \texttt{NonMatching::FEValues} calls  \texttt{QuadratureGenerator} only if needed, i.e.
+if the cell is intersected.
+If not, already cached \texttt{FEValues} objects will be returned by the getter functions.
+  Correspondingly, the class \texttt{NonMatching::FEInterfaceValues} generates
+  \texttt{FEInterfaceValues} objects for assembling face terms over $F \cap \{x : \psi(x) < 0 \}$ or $F \cap \{x : \psi(x) > 0 \}$.
+\end{itemize}
+The new \texttt{step-85} tutorial illustrates how many of these classes work together.
+
+\begin{figure}
   \centering
@@
   Library (CGAL)}\label{sec:cgalwrappers}
-  If you have CGAL (\href{https://www.cgal.org/}{{https://www.cgal.org/}}) installed \cite{cgal}, you can use wrappers around CGAL classes and functions contained in the new namespace \texttt{CGALWrappers}, allowing capabilities spanning from mesh generation to boolean operations between \dealii{} triangulations and cells.
-CGAL wrappers are enabled only if \dealii{} is compiled with \texttt{C++17}.
-We have introduced the following utilities to the \dealii's \texttt{CGALWrappers} namespace.
-\begin{itemize}
-\item Refinement criteria for tetrahedra or surface facets can be stored and specified in a struct named \texttt{AdditionalData}. 
-Depending on the template parameter \texttt{dim}, different criteria can be chosen. See the documentation of this struct for a list and an explanation of the possible parameters.
-\item \texttt{GridGenerator::implicit\_function()} creates a \texttt{Triangulation<dim,3>} out of the zero level set of an implicit function $f$. 
-Apart from the final \texttt{Triangulation<dim,3>} and the implicit funcion $f$ itself, also an interior point for which $f<0$ must be supplied by the user, along with the radius of a ball that will contain the generated triangulation.
-Mesh-related entities must be specified with \texttt{AdditionalData<dim>}.
-When \texttt{dim} is \texttt{3}, the mesh is filled with tetrahedra.
+
+The Computational Geometry Algorithms Library (CGAL, \url{https://www.cgal.org/}) is a widely used
+library to describe geometries and meshes \cite{cgal}. \dealii{} now
+has wrappers for CGAL classes and functions, provided in the new
+namespace \texttt{CGALWrappers}, and implementing functionality
+spanning from mesh generation to boolean operations between \dealii{}
+triangulations and cells. These wrappers are enabled only if \dealii{}
+is compiled with \texttt{C++17} and, for the first time in \dealii{}'s
+25-year history provides a built-in ability to generate meshes for
+arbitrary geometries.
+
+The main mesh generation function is
+\texttt{GridGenerator::implicit\_function()}, which creates a \texttt{Triangulation<dim,3>} out of the zero level set of an implicit function $f$. 
+For \texttt{dim==3}, the mesh consists of tetrahedra.
 A prototypical use case is the following, where the surface is the zero level set of Taubin's heart function $f=\bigl ( x^2 + \frac{9y^2}{4} +z^2 -1 \bigr ) -x^2 z^3 - \frac{9y^2z^3}{80}$.
@@
 data.cell_size = .05;
-GridGenerator::implicit_function(
-  tria, implicit_function, data, Point<3>(0, 0, 0), 10.0);
+GridGenerator::implicit_function(tria, implicit_function, data,
+                                 {0, 0, 0}, 10.0);
 \end{c++}
 
-\begin{figure}[h!]
+\begin{figure}
   \centering
@@
 
-\item \texttt{GridGenerator::surface\_mesh\_to\_volumetric\_mesh()} fills a \texttt{Triangulation<3>} starting from a \texttt{Triangulation<2,3>}.  
-The surface triangulation bounds the three dimensional volume, which is meshed inside with tetrahedra.
-\item Corefinement and boolean operations between \texttt{CGAL::Surface\_mesh} types are exploited to perform boolean operations between \dealii{}
-triangulations, and are available in the utility \texttt{CGALWrappers::compute\_boolean\_operation()}.
-The available operations are \texttt{corefinement}, \texttt{intersection}, \texttt{union} and \texttt{difference}, and they are
-described in the enum class \texttt{BooleanOperation}.
-Also, several wrappers allow to go back and forth from a \dealii{} Triangulation type to a CGAL Triangulation type.
-In general, after a \texttt{BooleanOperation} the corefinement around the intersection gives most of the times
-badly shaped elements. To overcome this issue, one can use \texttt{CGALWrappers::remesh\_surface()}. See Figure~\ref{fig:corefinement_remeshed} for a graphical example.
+A related function is
+\texttt{GridGenerator::surface\_mesh\_to\_volumetric\_mesh()}, which
+computes a tetrahedral volume triangulation \texttt{Triangulation<3>}, based on a
+given surface triangulation \texttt{Triangulation<2,3>} that bounds
+the three dimensional volume.
+
+CGAL also provides us with boolean operations on meshes, as available in the
+utility function
+\texttt{CGALWrappers::\allowbreak{}compute\_boolean\_operation()}. The
+available operations are \textit{corefinement}, \textit{intersection},
+\textit{union}, and \textit{difference}. Oftentimes, boolean
+operations and corefinement around the intersection produces
+badly shaped elements. To overcome this issue, one can use \texttt{CGALWrappers::\allowbreak{}remesh\_surface()}. See Figure~\ref{fig:corefinement_remeshed} for a graphical example.
 A possible workflow is the following:
@@
   \end{c++}
-  The output of the boolean operation can be appreciated in Figure~\ref{fig:corefinement}, while in Figure~\ref{fig:corefinement_remeshed}
+  The output of the boolean operation can be seen in Fig.~\ref{fig:corefinement}, while in Fig.~\ref{fig:corefinement_remeshed}
   the same mesh has been remeshed.
-  \begin{figure}[h]
+  \begin{figure}
     \centering
@@
     \end{subfigure}
-    \caption{\it (a) Union of the unit cube with an hyper ball. Notice how the corefinement around the intersection gives badly shaped elements. (b) The remeshed \dealii{} Triangulation.}
+    \caption{\it (a) Union of the unit cube with an hyper ball. Notice
+      how the corefinement around the intersection results in badly shaped elements. (b) The remeshed \dealii{} Triangulation.}
   \end{figure}
 
-\item \texttt{CGALWrappers::compute\_quadrature\_on\_boolean\_operation()} returns a \texttt{Quadrature<3>} that allows to integrate exactly on polyhedral elements coming out of a \texttt{BooleanOperation} between \dealii cells.
-The quadrature rule is built by meshing the polyhedral region with tetrahedra, computing on each tetrahedron a \texttt{QGaussSimplex<3>} quadrature rule by using \texttt{QSimplex<3>::compute\_affine\_transformation()}, and finally
+\texttt{CGALWrappers::compute\_quadrature\_on\_boolean\_operation()} returns a \texttt{Quadrature<3>} that allows exact integration on polyhedral elements coming out of a \texttt{BooleanOperation} between \dealii cells.
+The quadrature rule is built by meshing the polyhedral region with tetrahedra, computing on each tetrahedron a \texttt{QGaussSimplex<3>} quadrature rule by using \texttt{QSimplex<3>::\allowbreak{}compute\_affine\_transformation()}, and finally
 collecting all of the rules together, giving a \texttt{Quadrature<3>} formula on the \emph{physical} element.
-\end{itemize}
 
@@
 
-\begin{itemize}
-\item describe the new data structure
-\item do we have performance numbers?
-\end{itemize}
+\todo[inline]{
+* describe the new data structure
+
+* do we have performance numbers?
+}
 
@@
 \dealii{} has had an implementation of these algorithms for some time,
-but the current version substantially expands it. Specifically, the
+but the current release substantially expands on it. Specifically, the
 updated interfaces -- now based on function objects such as lambda
@@
 A special case of consensus algorithms is where the sender does not
-actually require and answer. This occasionally happens during ghost
+actually require an answer. This happens, for example, during ghost
 exchange where each process sends information about cells it owns to
 some of its neighboring processes where these cells are ghost
-cells. Previously, this case was implemented by sending an empty
-message. The rewritten interfaces now support this case: Code using
+cells. Previously, this case was implemented through a consensus
+algorithm where the reply message is simply empty.
+The rewritten interfaces now support this case more explicitly: Code using
 these interfaces no longer have to provide functions that formulate
@@
 \texttt{Utilities::MPI::broadcast()} function has been optimized for
-objects that are arrays of data types natively supported by MPI.
+objects that are arrays of data types natively supported by MPI and
+which consequently can be sent without packing and unpacking.
 
@@
 \dealii{} can be used to implement the
-local discontinuous Galerkin (LDG) method for approximating the solution to the bi-Laplacian problem.
+local discontinuous Galerkin (LDG) method for approximating the
+solution to the bi-Laplacian problem. The method is an alternative to
+the C0IP method used in \texttt{step-47}.
+
 \item \texttt{step-85} was contributed by Simon Sticko (Uppsala University). It shows how to use the new CutFEM infrastructure to solve a Poisson
@@
 \begin{itemize}
-\item ``\texttt{MCMC for the Laplace equation}'' TODO.
+  \item \todo[inline]{Is there actually a new program?}
 \end{itemize}
+Finally, the ``\texttt{MCMC for the Laplace equation}'' code gallery
+program has been updated by providing MATLAB and Python versions of
+the benchmark that is implemented in this code.
 
@@
   \item In continuation of our attempt to merge the classes \texttt{DoFHandler} and \texttt{hp::DoFHandler}, we have removed the
-  template parameter \texttt{DoFHandlerType} from number of classes and
-  functions and replaced it by \texttt{dim}/\texttt{spacedim}. Affected
-  classes are, e.g., \texttt{SolutionTransfer} and \texttt{DataOut}.
+  template parameter \texttt{DoFHandlerType} from a number of classes and
+  functions, instead using \texttt{dim}/\texttt{spacedim} template arguments. Affected
+  classes include \texttt{SolutionTransfer} and \texttt{DataOut}.
   \item \texttt{FE\_RaviartThomasNodal} now uses a different polynomial space to allow
-  for a simpler use for faces in non-standard orientation. The new polynomials
+  for a simpler use on faces in non-standard orientation. The new polynomials
   are anisotropic tensor products of Lagrange polynomials on the points of the
@@
   space spans the same polynomials.
-  \item The class \texttt{MappingQ} now applies a high-order mapping to all cells, not just the cells near the boundary, functionality that was previously provided by the \texttt{MappingQGeneric} class.
-The latter has been marked as deprecated.
-In order to use different degrees for the mapping, users can use \texttt{hp::MappingCollection} or \texttt{MappingFEField} with appropriate definitions of function spaces.
-  \item We made different changes in the repartitioning of \texttt{parallel::distributed::Triangulation}.
-  Most notably, we have removed the default weight in order to ensure consistency with the rest of the
-  library.
+  \item The class \texttt{MappingQ} now applies a high-order mapping
+    to all cells, not just the cells near the boundary, functionality
+    that was previously provided by the \texttt{MappingQGeneric}
+    class. The latter has been marked as deprecated.
+  \item Changes to weighted repartitioning of
+    \texttt{parallel::distributed::Triangulation} objects include.
+  the removal the default weight of each cell in order to ensure
+  consistency with the rest of the library.
 \end{itemize}

### 2022-06-21 9.4/paper.tex
Finish editing.

@@
     \item For distributed parallel computing: \cite{BangerthBursteddeHeisterKronbichler11};
-    \item For $hp$-adaptivity: \cite{BangerthKayserHerold2007};
+    \item For $hp$-adaptivity: \cite{BangerthKayserHerold2007,fehling2022};
     \item For partition-of-unity (PUM) and finite element enrichment methods:
@@
 
-\todo[inline]{Update. In particular make sure that we remove those who
-  are not authors on this paper.}
+\todo[inline]{Update. In particular make sure that we list how Marco
+  Feder is supported.}
 
@@
 
-W.~Bangerth, T.~Heister, R.~Gassm{\"o}ller, and J.~Zhang were partially
+W.~Bangerth, T.~Heister, and R.~Gassm{\"o}ller were partially
 supported by the Computational Infrastructure for Geodynamics initiative
@@
 
-B.~Blais was partially supported by the National Science and Engineering
-Research Council of Canada (NSERC)  through the RGPIN-2020-04510 Discovery
-Grant
-
-T.~Heister and J.~Zhang were also partially supported by NSF
-Award OAC-2015848.
-
-T.~Heister was also partially supported by the NSF Awards DMS-2028346,
+T.~Heister was also partially supported by NSF
+Awards OAC-2015848, DMS-2028346, and
 EAR-1925575, and by Technical Data Analysis, Inc. through US Navy STTR

### 2022-06-21 9.4/paper.tex
Improve CGAL section

@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=.6\paperwidth]{png/heart_implicit.png}
-  \caption{\it \texttt{Triangulation<3>} of a heart surface starting from an implicit function $f$. \label{fig:heart_tria}}
-\end{figure}
-
 A related function is
@@
     \centering
-    \begin{subfigure}[b]{0.4\textwidth}
+    \begin{subfigure}[b]{0.3\textwidth}
       \centering
-      \includegraphics[height=.2\paperheight]{png/intersection_cube_sphere_mesh.png}
+      \includegraphics[width=.2\paperwidth]{png/heart_implicit.png}
+      \caption{\label{fig:heart_tria}}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.3\textwidth}
+      \centering
+      \includegraphics[height=.14\paperheight]{png/intersection_cube_sphere_mesh.png}
       \caption{\label{fig:corefinement}}
@@
     \qquad
-    \begin{subfigure}[b]{0.5\textwidth}
+    \begin{subfigure}[b]{0.3\textwidth}
       \centering
-      \includegraphics[height=.2\paperheight]{png/cube_sphere_remeshed.png}
+      \includegraphics[height=.14\paperheight]{png/cube_sphere_remeshed.png}
       \caption{ \label{fig:corefinement_remeshed}}
     \end{subfigure}
-    \caption{\it (a) Union of the unit cube with an hyper ball. Notice
-      how the corefinement around the intersection results in badly shaped elements. (b) The remeshed \dealii{} Triangulation.}
+    \caption{\it (a) \texttt{Triangulation<3>} of a heart surface starting from an implicit function $f$.
+(b) Union of the unit cube with an hyper ball.
+Notice how the corefinement around the intersection gives badly shaped elements.
+(c) The remeshed \dealii{} Triangulation.}
   \end{figure}
@@
 The same applies to coupling terms of the form $(u,q)_B$ in formulations using Lagrange multipliers, where now $q \in Q_h(B)$, with $Q_h(B)$ the space of the multiplier variable.
+Notice that the most relevant difference with the quadrature rules provided by cutFEM support in \texttt{QuadratureGenerator} is that \texttt{Quadrature} objects are not created using a level-set approach, as described in the previous section, but directly with the two overlapped grids.
 
@@
 
-\todo[inline]{Update. In particular make sure that we list how Marco
-  Feder is supported.}
-
 D.~Arndt and B.~Turcksin: Research sponsored by the Laboratory Directed Research and

### 2022-06-21 9.4/paper.tex
Update my funding source.
Add particle section

--- 2022-06-21 9.4/paper.tex
@@
     Marc Fehling,
+    Rene Gassm\"{o}ller,
     Timo Heister,
@@
 \author[2]{Marc~Fehling}
+
+\author[13]{Rene~Gassm{\"o}ller}
+\affil[13]{Department of Geological Sciences,
+   University of Florida,
+   1843 Stadium Road,
+   Gainesville, FL, 32611, USA.
+  {\texttt{rene.gassmoeller@ufl.edu}}}
 
@@
 \subsection{Performance improvement of particle infrastructure}\label{sec:particles}
+
+We have improved upon the particle structure discussed in our previous release \citep{dealII93} by restructuring the storage of the particle identifiers.
+In our previous optimizations we stored all particle data as a separate continuous array for each particle property, but particle identifiers were still stored in a multimap tree-like structure.
+
+Our new particle storage is organized as follows: Particle IDs are stored in a list of dynamic arrays, each array contains the particle IDs of all particles in a unique cell.
+Each ID is a handle that determines the location of the data of this unique particle in the property arrays.
+The list of ID arrays only contains entries for cells that contain particles.
+We keep a separate cache structure that contains pointers to the particular list entries for each cell that contains particles, and invalid entries for all cells that do not contain particles.
+
+This new structure allows for the following significant performance improvements:
+All particle data (both their identifiers and their actual data) are now stored as separate and contiguous arrays in memory, which improves prefetching of data and makes iterating over particles extremely cache efficient.
+The choice of a list container that only includes entries for cells that contain particles means iteration is efficient, even if many cells in the domain do not contain any particles (as can be the case for discrete-element models).
+Creating separate arrays for each cell allows us to easily move particle IDs from one cell to another as a local operation, affecting only the two cell containers in question.
+We take care to reuse allocated memory to minimize the number of memory reallocations.
+The separate cache structure that contains entries for each cell allows quick random-access to the particles of a particular cell, and also allows to quickly determine if a particular cell has particles at all.
+
+In addition to the new storage structure we have made the following algorithmic improvements:
+Determining if a particle is inside a cell after changing its position involves inverting the mapping of this cell.
+We have reorganized our algorithms to perform these inversions on a batch of particles of a cell instead of particle-by-particle, which allows us to make use of vectorized instructions during the inversion.
+In addition, after sorting all particles into their new cells, the arrays that store particle properties are now sorted in the same order as the particle IDs in the list of arrays, which allows for cache efficient iteration over particle properties.
+To avoid a costly sorting operation this operation is executed as a copy of the existing data into a new data container while iterating over particle IDs.
 

### 2022-06-21 9.4/paper.tex
More edits

--- 2022-06-21 9.4/paper.tex
@@
 %  to make constraints consistent in parallel.
-  \item The \texttt{DataOutResample} class does not output a numerical solution
-  on the cells of the original triangulation, but interpolates the result
-  onto a second triangulation (which can be completely unrelated).
+  \item The \texttt{DataOutResample} class interpolates values defined on one
+  triangulation onto a second potentially unrelated triangulation.
   By using this class, one can output the result obtained on an
@@
   triangulation is a uniformly refined rectangle or box), or one can create a slice in 3D.
-  \item The new function \texttt{find\_point\_owner\_rank()} of \texttt{parallel::distributed::Triangulation} allows to find the MPI
-  rank of the subdomain of a distributed mesh that contains a given point.
+  \item The new member function \texttt{find\_point\_owner\_rank()} of \texttt{parallel::distributed::Triangulation} allows one to find the MPI
+  ranks of cells containing specified points.
   It is communication-free and leverages the functionality of p4est (>v.2.2).
-  Its algorithm is described in \cite{burstedde2020parallel}. Based on the information obtained,
-  the communication pattern used in \texttt{Utilties::MPI::RemotePointEvaluation} can be set up efficiently. Furthermore, this function could be used in the future to allow
-  particle simulations where particle movement is not
+  Its algorithm is described in \cite{burstedde2020parallel}. This information
+  enables efficient construction of the
+  communication pattern used in the class \texttt{Utilties::\allowbreak MPI::\allowbreak RemotePointEvaluation}. Furthermore, this function could be used in the future to allow
+  particle simulations in which particle movement is not
   limited by CFL conditions, as done in \cite{mirzadeh2016parallel}.
@@
   A manifold description is applied to the boundary, which can be extended into the volume via transfinite interpolation \cite{Gordon82} using the \texttt{TransfiniteInterpolationManifold} class \cite{dealII90}.
+  \item A new DoF renumbering function \texttt{DoFRenumbering::support\_point\_wise()} which groups together
+  shape functions by their support point. This functionality is useful in both developing nodal schemes since, e.g., the $x$, $y$, and $z$
+  velocities at a point will be consecutive in the solution vector. It also improves interoperability with external libraries which expect
+  data in this format.
   \item \todo[inline]{Add something about the FEInterfaceValues improvements?}
@@
 
-  \caption{\it New refinement strategies for triangles and tetrahedra: Triangles
-    are subdivided into four children and tetrahedra into eight.
-  \todo[inline]{Are these ``new strategies'', or is it ``New:
-    Refinement strategies'' because we did not have refinement at all before?}}
+  % We definitely supported tri refinement in 9.3 - see 6ceb8b70559dc2274761a28dc7836f1e9602da8c (Dec 4, 2020)
+  % and also 74b8171dd030a4b2de41955b8d913cde06974b91 (Feb 4, 2021)
+  \caption{\it New: The previous release \cite{dealII93} added support for adaptive mesh
+  refinement with triangles. This release adds support for global tetrahedral refinement.}
   \label{fig:refinement}
@@
 There are now new utility functions in the
-\texttt{RepartitioningPolicyTools} namepace that can be
+\texttt{RepartitioningPolicyTools} namespace that can be
 used to create a new \texttt{p::f::T} instance,
@@
 addition to the established local-smoothing infrastructure. Global
-coarsening smoothens on the whole computational domain on each
+coarsening smooths on the whole computational domain on each
 multigrid level, which is obtained by coarsening the finest cells of
@@
 given surface triangulation \texttt{Triangulation<2,3>} that bounds
-the three dimensional volume.
+the three dimensional shape.
 
@@
 
-D.~Wells was supported by NSF through Grant DMS-1344962.
+D.~Wells was supported by the NSF Award OAC-1931516.
 

### 2022-06-21 9.4/paper.tex
Clean up the list of contributors.
Rene is a coauthor so he and Bruno (Turcksin) should not be on this list.

@@
 %
-\todo[inline]{updated 23/04/2022 PM. We should probably make sure that
-  nobody else came after that and before the release.}
+
+% This is up-to-date as of 9.4 RC1 - should be the final list.
 Pasquale      Africa,
@@
 Maximilian    Bergbauer,
+Manaswinee    Bezbaruah,
 Bruno         Blais,
@@
 Praveen       Chandrashekar,
+Huimin        Chen,
 Cu            Cui,
@@
 Niklas        Fehn,
+Corbin        Foucart,
 Johannes      Friedlein,
 Sebastian     Fuchs,
-Rene	Gassmoeller,
+Daniel        Garcia-Sanchez,
 Nicola        Giuliani,
@@
 Sebastian     Kinnewig,
-Simon	Konrad,
+Uwe           K{\"o}cher,
 Katharina     Kormann,
@@
 Simon         Sticko,
-Bruno	Turcksin,
+Daniel        Sun,
 Kuljit S.     Virk,

### 2022-06-21 9.4/paper.tex
Improve particle section. Update funding

--- 2022-06-21 9.4/paper.tex
@@
 
-We have improved upon the particle structure discussed in our previous release \citep{dealII93} by restructuring the storage of the particle identifiers.
-In our previous optimizations we stored all particle data as a separate continuous array for each particle property, but particle identifiers were still stored in a multimap tree-like structure.
-
-Our new particle storage is organized as follows: Particle IDs are stored in a list of dynamic arrays, each array contains the particle IDs of all particles in a unique cell.
+We have enhanced the capabilities of \dealii to include particle methods by reorganizing storage structures and optimizing algorithms.
+In our previously reported improvements~\citep{dealII93} we stored all particle data as a separate continuous array for each particle property, but particle identifiers (IDs) were still stored in a multimap tree-like structure.
+
+Our new particle containers are organized as follows: Particle IDs are stored in a list of dynamic arrays, each array contains the particle IDs of all particles in a unique cell.
 Each ID is a handle that determines the location of the data of this unique particle in the property arrays.
 The list of ID arrays only contains entries for cells that contain particles.
-We keep a separate cache structure that contains pointers to the particular list entries for each cell that contains particles, and invalid entries for all cells that do not contain particles.
-
-This new structure allows for the following significant performance improvements:
-All particle data (both their identifiers and their actual data) are now stored as separate and contiguous arrays in memory, which improves prefetching of data and makes iterating over particles extremely cache efficient.
-The choice of a list container that only includes entries for cells that contain particles means iteration is efficient, even if many cells in the domain do not contain any particles (as can be the case for discrete-element models).
-Creating separate arrays for each cell allows us to easily move particle IDs from one cell to another as a local operation, affecting only the two cell containers in question.
+We keep a separate cache structure that contains pointers to the particular list entries for each cell.
+If a cell has no particles this pointer is invalid.
+
+This structure allows for the following significant performance improvements:
+
+\begin{itemize}
+\item All particle data (both their identifiers and their actual data) are now stored as separate and contiguous arrays in memory, which improves prefetching of data and makes iterating over particles extremely cache efficient.
+\item The choice of a list container that only includes entries for cells that contain particles means iteration is efficient, even if many cells in the domain do not contain any particles (as can be the case for discrete-element methods).
+\item Creating separate arrays for each cell allows us to easily move particle IDs from one cell to another as a local operation, affecting only the two cell containers in question.
 We take care to reuse allocated memory to minimize the number of memory reallocations.
-The separate cache structure that contains entries for each cell allows quick random-access to the particles of a particular cell, and also allows to quickly determine if a particular cell has particles at all.
+\item The separate cache structure that contains entries for each cell allows quick random-access to the particles of a particular cell, and also allows to quickly determine if a particular cell has particles at all.
+\end{itemize}
 
 In addition to the new storage structure we have made the following algorithmic improvements:
-Determining if a particle is inside a cell after changing its position involves inverting the mapping of this cell.
-We have reorganized our algorithms to perform these inversions on a batch of particles of a cell instead of particle-by-particle, which allows us to make use of vectorized instructions during the inversion.
+Determining if a particle is inside a cell after changing its position involves inverting the mapping for this cell.
+We have reorganized our algorithms to perform these inversions on a batch of particles in the same cell instead of particle-by-particle, which allows us to make use of vectorized instructions during the inversion.
 In addition, after sorting all particles into their new cells, the arrays that store particle properties are now sorted in the same order as the particle IDs in the list of arrays, which allows for cache efficient iteration over particle properties.
-To avoid a costly sorting operation this operation is executed as a copy of the existing data into a new data container while iterating over particle IDs.
-
-\todo[inline]{
-* describe the new data structure
-
-* do we have performance numbers?
-}
+To avoid a costly sorting operation this operation is executed as a copy of the existing data into a new data container that replaces the existing container.
+
+\begin{table}
+  \caption{Timing of various particle operations for tutorial program \texttt{step-68} (particle advection in a 2D, cartesian box) using 400,000 particles on a single process.}
+  \label{tab:particle_timing}
+
+  \begin{tabular}{|c|c|c|c|}
+    \hline
+   Particle Operation & \dealii 9.3 & \dealii 9.4 & Speedup \\
+   \hline
+   Generation & 444 ms & 235 ms & 1.9x \\
+   Iteration & 4.18 ms & 0.638 ms & 6.6x \\
+   Advection & 37.8 ms & 33.9 ms & 1.15x \\
+   Sorting & 21.9 ms & 9.27 ms & 2.4x \\
+   \hline
+  \end{tabular}
+  \end{table}
+
+We illustrate the combined effect of these performance improvements in Table~\ref{tab:particle_timing}.
+We have measured the averaged compute time for four particle operations in a slightly modified version of the \dealii tutorial program \texttt{step-68} when advecting 400,000 particles on a single process (we have not observed any influence of the described changes on the parallel scalability of the algorithms).
+The four operations we have measured are:
+\begin{itemize}
+\item Generation of a set of 400,000 particles at positions that are not aligned with the background mesh, i.e. the containing cell of each particle has to be found.
+\item Iteration without significant computation over the whole set of created particles (without accessing particle data).
+\item Advection of all particles, which involves iteration over all particles, evaluation of the finite element solution at the location of the particles, read and write access to the position of all particles to modify their location, and write access to the particle properties (to store their velocity for visualization purposes).
+\item Sorting, meaning the inversion of the mapping of each cell to find the new particle locations relative to this cell, and moving all particles that have left their original cell into new cells (both \dealii 9.3 and \dealii 9.4).
+In \dealii 9.4 this operation also includes reordering the particle properties for optimal iteration.
+\end{itemize}
+
+Table~\ref{tab:particle_timing} shows that all particle operations are much faster faster in \dealii 9.4 than \dealii 9.3.
+In particular operations that depend strongly on particle storage structure and require few fixed computations (like iteration and sorting) benefit massively from the above mentioned optimizations.
+Please note that the exact gains will depend strongly on the exact combination of geometry, mapping, dimensionality, and number of particles per cell in any specific model, and can be smaller or larger than the measurements provided here.
 
@@
 
-R.~Gassm{\"o}ller was also partially supported by the NSF Award
-EAR-1925595.
+R.~Gassm{\"o}ller was also partially supported by the NSF Awards
+EAR-1925677, and EAR-2054605.
 

### 2022-06-21 9.4/paper.tex
More edits

--- 2022-06-21 9.4/paper.tex
@@
   triangulation is a uniformly refined rectangle or box), or one can create a slice in 3D.
-  \item The new member function \texttt{find\_point\_owner\_rank()} of \texttt{parallel::distributed::Triangulation} allows one to find the MPI
+  \item The new member function \texttt{find\_point\_owner\_rank()} of
+        \texttt{parallel\allowbreak ::distributed::\allowbreak Triangulation} allows one to find the MPI
   ranks of cells containing specified points.
@@
   data in this format.
-  \item \todo[inline]{Add something about the FEInterfaceValues improvements?}
+  \item A new module \texttt{Utilities::MPI::LargeCount} which enables sending and receiving MPI messages containing more than $2^31$ objects.
+  This library either uses the new MPI-4 functions, such as \texttt{MPI\_Send\_c()}, or an internal implementation of large objects for
+  compatibility with MPI-3.
+  \item The \texttt{FEInterfaceValues} class, which computes common quantities at the interface of two cells, has been overhauled to make it
+  more consistent with the rest of the library and use more intuitive names for functions. For example,
+  \texttt{FEInterfaceValues::jump\_gradient()} is now \texttt{FEInterfaceValues::jump\_in\_shape\_gradients()}. Several new
+  functions, such as \texttt{FEInterfaceValues::get\_jump\_in\_function\_values()}, have also been added.
 \end{itemize}
 %
-The changelog lists more than 175 other features and bugfixes.
+The changelog lists more than 100 other features and bugfixes.
 
@@
 both for cells and faces. This could enable, for example, writing
-a matrix-free version of step-47, which solves the biharmonic equation with DG.
+a matrix-free version of step-47, which solves the biharmonic equation with the discontinuous Galerkin method.
 \item Cell-centric loops now also allow access to gradients and Hessians

### 2022-06-22 9.4/paper.tex

@@
 
-\todo[inline]{
-* describe the new data structure
-
-* do we have performance numbers?
-}
+We have enhanced the capabilities of \dealii to include particle methods by reorganizing storage structures and optimizing algorithms.
+In our previously reported improvements~\citep{dealII93} we stored all particle data as a separate continuous array for each particle property, but particle identifiers (IDs) were still stored in a multimap tree-like structure.
+
+Our new particle containers are organized as follows: Particle IDs are stored in a list of dynamic arrays, each array contains the particle IDs of all particles in a unique cell.
+Each ID is a handle that determines the location of the data of this unique particle in the property arrays.
+The list of ID arrays only contains entries for cells that contain particles.
+We keep a separate cache structure that contains pointers to the particular list entries for each cell.
+If a cell has no particles this pointer is invalid.
+
+This structure allows for the following significant performance improvements:
+
+\begin{itemize}
+\item All particle data (both their identifiers and their actual data) are now stored as separate and contiguous arrays in memory, which improves prefetching of data and makes iterating over particles extremely cache efficient.
+\item The choice of a list container that only includes entries for cells that contain particles means iteration is efficient, even if many cells in the domain do not contain any particles (as can be the case for discrete-element methods).
+\item Creating separate arrays for each cell allows us to easily move particle IDs from one cell to another as a local operation, affecting only the two cell containers in question.
+We take care to reuse allocated memory to minimize the number of memory reallocations.
+\item The separate cache structure that contains entries for each cell allows quick random-access to the particles of a particular cell, and also allows to quickly determine if a particular cell has particles at all.
+\end{itemize}
+
+In addition to the new storage structure we have made the following algorithmic improvements:
+Determining if a particle is inside a cell after changing its position involves inverting the mapping for this cell.
+We have reorganized our algorithms to perform these inversions on a batch of particles in the same cell instead of particle-by-particle, which allows us to make use of vectorized instructions during the inversion.
+In addition, after sorting all particles into their new cells, the arrays that store particle properties are now sorted in the same order as the particle IDs in the list of arrays, which allows for cache efficient iteration over particle properties.
+To avoid a costly sorting operation this operation is executed as a copy of the existing data into a new data container that replaces the existing container.
+
+\begin{table}
+  \caption{Timing of various particle operations for tutorial program \texttt{step-68} (particle advection in a 2D, cartesian box) using 400,000 particles on a single process.}
+  \label{tab:particle_timing}
+
+  \begin{tabular}{|c|c|c|c|}
+    \hline
+   Particle Operation & \dealii 9.3 & \dealii 9.4 & Speedup \\
+   \hline
+   Generation & 444 ms & 235 ms & 1.9x \\
+   Iteration & 4.18 ms & 0.638 ms & 6.6x \\
+   Advection & 37.8 ms & 33.9 ms & 1.15x \\
+   Sorting & 21.9 ms & 9.27 ms & 2.4x \\
+   \hline
+  \end{tabular}
+  \end{table}
+
+We illustrate the combined effect of these performance improvements in Table~\ref{tab:particle_timing}.
+We have measured the averaged compute time for four particle operations in a slightly modified version of the \dealii tutorial program \texttt{step-68} when advecting 400,000 particles on a single process (we have not observed any influence of the described changes on the parallel scalability of the algorithms).
+The four operations we have measured are:
+\begin{itemize}
+\item Generation of a set of 400,000 particles at positions that are not aligned with the background mesh, i.e. the containing cell of each particle has to be found.
+\item Iteration without significant computation over the whole set of created particles (without accessing particle data).
+\item Advection of all particles, which involves iteration over all particles, evaluation of the finite element solution at the location of the particles, read and write access to the position of all particles to modify their location, and write access to the particle properties (to store their velocity for visualization purposes).
+\item Sorting, meaning the inversion of the mapping of each cell to find the new particle locations relative to this cell, and moving all particles that have left their original cell into new cells (both \dealii 9.3 and \dealii 9.4).
+In \dealii 9.4 this operation also includes reordering the particle properties for optimal iteration.
+\end{itemize}
+
+Table~\ref{tab:particle_timing} shows that all particle operations are much faster faster in \dealii 9.4 than \dealii 9.3.
+In particular operations that depend strongly on particle storage structure and require few fixed computations (like iteration and sorting) benefit massively from the above mentioned optimizations.
+Please note that the exact gains will depend strongly on the exact combination of geometry, mapping, dimensionality, and number of particles per cell in any specific model, and can be smaller or larger than the measurements provided here.
 

### 2022-06-22 9.4/paper.tex
Edit the particle section.

@@
   % and also 74b8171dd030a4b2de41955b8d913cde06974b91 (Feb 4, 2021)
-  \caption{\it New: The previous release \cite{dealII93} added support for adaptive mesh
-  refinement with triangles. This release adds support for global tetrahedral refinement.}
+  \caption{\it The previous release, 9.3 \cite{dealII93}, had
+    added support for adaptive mesh 
+  refinement with triangles. This release adds support for global
+  tetrahedral refinement by subdividing each tetrahedron into eight
+  children, as shown on the right.}
   \label{fig:refinement}
@@
 
-We have enhanced the capabilities of \dealii to include particle methods by reorganizing storage structures and optimizing algorithms.
-In our previously reported improvements~\citep{dealII93} we stored all particle data as a separate continuous array for each particle property, but particle identifiers (IDs) were still stored in a multimap tree-like structure.
-
-Our new particle containers are organized as follows: Particle IDs are stored in a list of dynamic arrays, each array contains the particle IDs of all particles in a unique cell.
+In the release described herein, we have reorganized the data
+structures and optimized the algorithms that support using particles
+in \dealii{}.
+In our previously reported improvements~\citep{dealII93} we stored all particle data as a separate contiguous array for each particle property, but particle identifiers (IDs) were still stored in a multimap tree-like structure.
+
+Our new particle containers are organized as follows: Particle IDs are stored in a list of dynamic arrays, each array containing the particle IDs of all particles in a unique cell.
 Each ID is a handle that determines the location of the data of this unique particle in the property arrays.
@@
 \begin{table}
-  \caption{Timing of various particle operations for tutorial program \texttt{step-68} (particle advection in a 2D, cartesian box) using 400,000 particles on a single process.}
+  \caption{\it Timing of various particle operations for tutorial program \texttt{step-68} (particle advection in a 2D, Cartesian box) using 400,000 particles on a single process.}
   \label{tab:particle_timing}
 
+  \centering
   \begin{tabular}{|c|c|c|c|}
@@
    \hline
-   Generation & 444 ms & 235 ms & 1.9x \\
-   Iteration & 4.18 ms & 0.638 ms & 6.6x \\
-   Advection & 37.8 ms & 33.9 ms & 1.15x \\
-   Sorting & 21.9 ms & 9.27 ms & 2.4x \\
+   Generation & 444 ms & 235 ms & 1.9$\times$ \\
+   Iteration & 4.18 ms & 0.638 ms & 6.6$\times$ \\
+   Advection & 37.8 ms & 33.9 ms & 1.15$\times$ \\
+   Sorting & 21.9 ms & 9.27 ms & 2.4$\times$ \\
    \hline
@@
 We illustrate the combined effect of these performance improvements in Table~\ref{tab:particle_timing}.
-We have measured the averaged compute time for four particle operations in a slightly modified version of the \dealii tutorial program \texttt{step-68} when advecting 400,000 particles on a single process (we have not observed any influence of the described changes on the parallel scalability of the algorithms).
+We measure the averaged compute time for four particle operations in a slightly modified version of the \dealii tutorial program \texttt{step-68} when advecting 400,000 particles on a single process (we have not observed any influence of the described changes on the parallel scalability of the algorithms).
 The four operations we have measured are:
@@
 \item Generation of a set of 400,000 particles at positions that are not aligned with the background mesh, i.e. the containing cell of each particle has to be found.
-\item Iteration without significant computation over the whole set of created particles (without accessing particle data).
+\item Iteration over the whole set of created particles, without
+  significant computation and in particular without accessing particle data.
 \item Advection of all particles, which involves iteration over all particles, evaluation of the finite element solution at the location of the particles, read and write access to the position of all particles to modify their location, and write access to the particle properties (to store their velocity for visualization purposes).
-\item Sorting, meaning the inversion of the mapping of each cell to find the new particle locations relative to this cell, and moving all particles that have left their original cell into new cells (both \dealii 9.3 and \dealii 9.4).
+\item Sorting, i.e., the inversion of the mapping of each cell to find the new particle locations relative to this cell, and moving all particles that have left their original cell into new cells (both \dealii 9.3 and \dealii 9.4).
 In \dealii 9.4 this operation also includes reordering the particle properties for optimal iteration.
@@
 
-Table~\ref{tab:particle_timing} shows that all particle operations are much faster faster in \dealii 9.4 than \dealii 9.3.
+Table~\ref{tab:particle_timing} shows that all particle operations are
+much faster faster in \dealii 9.4 than in version 9.3.
 In particular operations that depend strongly on particle storage structure and require few fixed computations (like iteration and sorting) benefit massively from the above mentioned optimizations.
-Please note that the exact gains will depend strongly on the exact combination of geometry, mapping, dimensionality, and number of particles per cell in any specific model, and can be smaller or larger than the measurements provided here.
+We note that the exact gains will depend strongly on the exact combination of geometry, mapping, dimensionality, and number of particles per cell in any specific model, and can be smaller or larger than the measurements provided here.
 

### 2022-06-22 9.4/paper.tex
Adjust figure spacing.
Go through document and make some updates

--- 2022-06-22 9.4/paper.tex
@@
   data in this format.
-  \item A new module \texttt{Utilities::MPI::LargeCount} which enables sending and receiving MPI messages containing more than $2^31$ objects.
+  \item A new module \texttt{Utilities::MPI::LargeCount} which enables sending and receiving MPI messages containing more than $2^{31}$ objects.
   This library either uses the new MPI-4 functions, such as \texttt{MPI\_Send\_c()}, or an internal implementation of large objects for
@@
 accesses the appropriate data and reshuffles mapping data accordingly on
-the fly in order to enable vectorization over cells. The new feature is useful in many
-contexts. Examples are application with sharp interfaces (e.g., two-phase flow
+the fly in order to enable vectorization over cells. The new feature is useful in several
+contexts. Examples are simulations with sharp interfaces (e.g., two-phase flow
 or shock capturing), where one needs to treat cells that are ``cut'' by
 the interface in a special way. A challenge is that cell batches
-might contain cut or not-cut cells, making vectorization potentially more complicated. One way to deal with such cell batches is to apply masks if
-the code paths do not diverge too much
-(a lot of functions of \texttt{FEEvaluation} allows this). Another way is to
-categorize cells during \texttt{MatrixFree::reinit()} in such a way that mixed cell batches do not occur. However, calling  \texttt{MatrixFree::reinit()}
-might be too expensive if a very dynamic system is given, which requires
-recategorization in each time step. Doing this on the fly might be a cheap alternative, even if not computationally optimal during matrix-free loops.
-\item Initial support for H(div)-conforming elements with Piola transform, based on Raviart--Thomas finite element spaces with the updated class \texttt{FE\_RaviartThomasNodal}.
+might contain cut or not-cut cells, making vectorization of these operations potentially more complicated.
+Previous functionality has provided the option of masking certain cells in cell batches, which works well if
+the code paths do not diverge too much. Another way is to
+categorize cells during \texttt{MatrixFree::reinit()} in such a way that mixed cell batches do not occur. However, \texttt{MatrixFree::reinit()}
+might be too expensive if recategorization needs to happen very frequently to follow the dynamics of a system, e.g., in each time step.
+Despite some overhead compared to static matrix-free loops, the new feature can be the best options in these scenarios.
+\item Initial support for H(div)-conforming elements with Piola transform, based on Raviart--Thomas finite element spaces with the updated class \texttt{FE\_RaviartThomasNodal}, has been added.
 This feature is currently limited to meshes in standard orientation and affine geometries.
-Full support will be provided in a future release.
+Full support and performance optimizations will be provided in a future release.
 \item Selected matrix-free algorithms can now exploit additional data locality
   between the matrix-vector product and vector operations happening nearby in
-  an algorithms. These have been added to both the
-  \texttt{PreconditionChebyshev} class often used as a smoother in multigrid
-  methods and the conjugate gradient implementation in \texttt{SolverCG}. To
-  use these features, an operator needs to define a \texttt{vmult} operation,
+  an algorithm. Interfaces have been added to both the
+  \texttt{PreconditionChebyshev} class (a frequently used smoother in multigrid
+  methods) and the conjugate gradient implementation in \texttt{SolverCG}.
+  From the user's perspective, an operator needs to define a \texttt{vmult} operation,
   taking two additional \texttt{std::function} arguments. The first function
-  defines what should be scheduled on the vector entries before the
+  defines the operation to be scheduled on the vector entries before the
   matrix-vector product touches them, and the second what happens
@@
 Finally, we have performed a major restructuring of internals
-of the \texttt{FEEvaluation} classes. This will enable us to add support for new element types, e.g., of
-Raviart-Thomas and Nedelec elements.
+of the \texttt{FEEvaluation} classes. This reduces some overheads for low polynomial degrees and will enable us to add support for new element types in the future.
 
@@
 addition to the established local-smoothing infrastructure. Global
-coarsening smooths on the whole computational domain on each
+coarsening algorithms smoothen on the whole computational domain on each
 multigrid level, which is obtained by coarsening the finest cells of
@@
 instance of \texttt{RepartitioningPolicyTools::Base} (see Subsection~\ref{sec:repartitioning}) as argument, which allows to specify the parallel
-distribution of each multigrid level (in contrast to the fixed first-child policy in the case of local smoothing). Furthermore, we added support for block vectors,
+distribution of each multigrid level (in contrast to the fixed first-child policy in the case of local smoothing). These features have been developed and tuned for runs on supercomputer scale
+with complicated coarse meshes as presented in \cite{kronbichler2021next}.
+Furthermore, we added support for block vectors,
 fixed a number of limitations, and performed performance optimizations of
 the transfer operator; particularly, the redundant copy from/to temporary vectors
-has been eliminated. Now, hanging-node constraints are applied
+has been eliminated. Furthermore, hanging-node constraints are applied
 efficiently in the same way as in the matrix-free loops (see Subsection~\ref{sec:mf}).
@@
 in the \texttt{MGTools} namespace for the  estimation of the workload imbalance and the
-vertical communication efficiency purely based on the given mesh. The publication \cite{munch2022gc} also points out that not
-all types of smoothers are applicable for global coarsening due to the
-presence of hanging nodes, which is a motivation to add new smoother types
-to \dealii in the future.
+vertical communication efficiency purely based on the given mesh.
+% MK: I would not add this part, it does not really fit into this paper as
+% there is nothing to report at this point (and we should then add references
+% to the actual literature).
+%The publication \cite{munch2022gc} also points out that not
+%all types of smoothers are applicable for global coarsening due to the
+%presence of hanging nodes, which is a motivation to add new smoother types
+%to \dealii in the future.
 
@@
 \end{equation}
-Thus, when assembling on a cut call, we are
+Thus, when assembling on a cut cell, we are
 required to integrate over the part of the domain and the part of the boundary, $\Gamma = \partial \Omega$, that falls inside the cell:
 $K\cap \Omega$ and $K \cap \Gamma$.
-Many of the new classes that supports these operations assume that the domain is described by a level set function,
+Many of the new classes that support these operations assume that the domain is described by a level set function,
 $\psi : \mathbb{R}^d \to \mathbb{R}$, such that
-\begin{align}
+\begin{align}\label{eq:levelset}
   \Omega = \{x \in \mathbb{R}^d : \psi(x)<0\},
@@
   \item The \texttt{MeshClassifier} class identifies how the active cells and faces are located relative to the zero contour of the level set function, as illustrated in Figure~\ref{fig:location-to-level-set}.
-Its \texttt{location\_to\_level\_set()} function takes a cell/face and
-        returns an enum, \texttt{LocationToLevelSet}, which takes values \{\texttt{inside}, \texttt{outside}, \texttt{intersected}\}.
+Its member function \texttt{location\_to\_level\_set()} takes a cell or face and
+        returns an enum, \texttt{LocationToLevelSet}, with values \{\texttt{inside}, \texttt{outside}, \texttt{intersected}\}.
         This information is typically needed when choosing what element a cell of the \texttt{DoFHandler} should use.
@@
   \item The \texttt{QuadratureGenerator} class, which implements the algorithm in \cite{saye2015}, generates high-order quadrature rules for the three different regions of a \texttt{BoundingBox}, $B$, defined by the sign of the level set function:
-        \begin{align}
+        \begin{align}\label{eq:boundingbox}
           B \cap \Omega = \{ x\in B:  \psi(x) < 0 \},  \quad
@@
         The \texttt{FaceQuadratureGenerator} class does the same for faces.
-        Furthermore, the new classes \texttt{DiscreteQuadratureGenerator} and \newline \texttt{DiscreteFaceQuadratureGenerator} can be used to generate these quadrature rules over a cell/face when the level set function lies in a finite element space: $\psi_h \in V_h$, and when the reference cell of the cell/face is a hypercube.
+        Furthermore, the new classes \texttt{DiscreteQuadratureGenerator} and \texttt{DiscreteFaceQuadratureGenerator} can be used to generate these quadrature rules over a cell or face when the level set function lies in a finite element space: $\psi_h \in V_h$, and when the reference cell of the cell or face is a hypercube.
 
@@
   Since the generation of immersed quadrature rules is not cheap,
-  \texttt{NonMatching::FEValues} calls  \texttt{QuadratureGenerator} only if needed, i.e.
-if the cell is intersected.
+  \texttt{NonMatching::FEValues} calls  \texttt{QuadratureGenerator} only if needed, i.e., if the cell is intersected.
 If not, already cached \texttt{FEValues} objects will be returned by the getter functions.
@@
 is compiled with \texttt{C++17} and, for the first time in \dealii{}'s
-25-year history provides a built-in ability to generate meshes for
+25-year history, provides a built-in ability to generate meshes for
 arbitrary geometries.
@@
 The main mesh generation function is
-\texttt{GridGenerator::implicit\_function()}, which creates a \texttt{Triangulation<dim,3>} out of the zero level set of an implicit function $f$. 
+\texttt{GridGenerator::implicit\_function()}, which creates a \texttt{Triangulation<dim,3>} out of the zero level set of an implicit function $\psi$ similar to \eqref{eq:levelset}.
 For \texttt{dim==3}, the mesh consists of tetrahedra.
@@
 operations and corefinement around the intersection produces
-badly shaped elements. To overcome this issue, one can use \texttt{CGALWrappers::\allowbreak{}remesh\_surface()}. See Figure~\ref{fig:corefinement_remeshed} for a graphical example.
+badly shaped mesh cells. To overcome this issue, one can use \texttt{CGALWrappers::\allowbreak{}remesh\_surface()}. Figure~\ref{fig:corefinement_remeshed} shows a graphical example.
 A possible workflow is the following:
@@
       \centering
-      \includegraphics[width=.2\paperwidth]{png/heart_implicit.png}
+      \includegraphics[width=\textwidth]{png/heart_implicit.png}
       \caption{\label{fig:heart_tria}}
     \end{subfigure}
+    \hfill
     \begin{subfigure}[b]{0.3\textwidth}
       \centering
-      \includegraphics[height=.14\paperheight]{png/intersection_cube_sphere_mesh.png}
+      \includegraphics[width=\textwidth]{png/intersection_cube_sphere_mesh.png}
       \caption{\label{fig:corefinement}}
     \end{subfigure}
-    \qquad
-    \begin{subfigure}[b]{0.3\textwidth}
+    \hfill
+    \begin{subfigure}[b]{0.37\textwidth}
       \centering
-      \includegraphics[height=.14\paperheight]{png/cube_sphere_remeshed.png}
+      \includegraphics[width=\textwidth]{png/cube_sphere_remeshed.png}
       \caption{ \label{fig:corefinement_remeshed}}
@@
 
-These utilities will be the building blocks for adding functions to the \texttt{NonMatching} namespace that will assemble coupling terms like $(u,v)_{B}$, with $B$ a domain immersed in a fixed background mesh $\Omega$ and $u,v$ finite element functions on $V_h(\Omega)$, as usually happens using Nitsche's method to weakly impose boundary conditions at an interface.
+These utilities will be the building blocks for adding functions to the \texttt{NonMatching} namespace that will assemble coupling terms like $(u,v)_{B}$,\marginpar{The $B$ and $\Omega$ here need to be streamlined with \eqref{eq:boundingbox}, right now they are the opposite} with $B$ a domain immersed in a fixed background mesh $\Omega$ and $u,v$ finite element functions on $V_h(\Omega)$, as usually happens using Nitsche's method to weakly impose boundary conditions at an interface.
 The same applies to coupling terms of the form $(u,q)_B$ in formulations using Lagrange multipliers, where now $q \in Q_h(B)$, with $Q_h(B)$ the space of the multiplier variable.
@@
 \begin{itemize}
-\item All particle data (both their identifiers and their actual data) are now stored as separate and contiguous arrays in memory, which improves prefetching of data and makes iterating over particles extremely cache efficient.
+\item All particle data (both their identifiers and their actual data) are now stored as separate and contiguous arrays in memory, which improves spatial locality for better prefetching of data and makes iterating over particles extremely efficient.
 \item The choice of a list container that only includes entries for cells that contain particles means iteration is efficient, even if many cells in the domain do not contain any particles (as can be the case for discrete-element methods).
@@
 
-In addition to the new storage structure we have made the following algorithmic improvements:
+In addition to the new storage structure, we have made the following algorithmic improvements:
 Determining if a particle is inside a cell after changing its position involves inverting the mapping for this cell.
-We have reorganized our algorithms to perform these inversions on a batch of particles in the same cell instead of particle-by-particle, which allows us to make use of vectorized instructions during the inversion.
+We have reorganized our algorithms to perform these inversions on a batch of particles in the same cell instead of particle-by-particle, which allows us to make use of vectorized instructions during the inversion using the generic scheme of~\cite{KronbichlerKormann2012}.
 In addition, after sorting all particles into their new cells, the arrays that store particle properties are now sorted in the same order as the particle IDs in the list of arrays, which allows for cache efficient iteration over particle properties.
@@
 \begin{itemize}
-\item Generation of a set of 400,000 particles at positions that are not aligned with the background mesh, i.e. the containing cell of each particle has to be found.
+\item Generation of a set of 400,000 particles at positions that are not aligned with the background mesh, i.e., the containing cell of each particle has to be identified with a search algorithm.
 \item Iteration over the whole set of created particles, without
@@
 updated interfaces -- now based on function objects such as lambda
-functions to formulate and process queries and replies -- can now deal
+functions to formulate and process queries and replies -- can deal
 with arbitrary data types for queries and replies, rather than only
@@
 \texttt{Utilities::unpack()} functions. We have also worked on making
-these functions more efficient: if the object to be packed is an array
+these functions efficient: if the object to be packed is an array
 (or array of arrays) of elements that satisfy the
@@
   Gauss--Lobatto quadrature formula. This change leads to different entries, for example, in
-  the matrices and constraints, but no change in accuracy should be expected as the resulting polynomial
-  space spans the same polynomials.
+  the matrices and constraints, but no change in accuracy should be expected as
+  the resulting basis spans the same polynomial space.
   \item The class \texttt{MappingQ} now applies a high-order mapping
@@
   \item Changes to weighted repartitioning of
-    \texttt{parallel::distributed::Triangulation} objects include.
+    \texttt{parallel::distributed::Triangulation} objects include
   the removal the default weight of each cell in order to ensure

### 2022-06-22 9.4/paper.tex
corefinement->co-refinement

@@
 \texttt{CGALWrappers::\allowbreak{}compute\_boolean\_operation()}. The
-available operations are \textit{corefinement}, \textit{intersection},
+available operations are \textit{co-refinement}, \textit{intersection},
 \textit{union}, and \textit{difference}. Oftentimes, boolean
-operations and corefinement around the intersection produces
+operations and co-refinement around the intersection produces
 badly shaped mesh cells. To overcome this issue, one can use \texttt{CGALWrappers::\allowbreak{}remesh\_surface()}. Figure~\ref{fig:corefinement_remeshed} shows a graphical example.
@@
 (b) Union of the unit cube with an hyper ball.
-Notice how the corefinement around the intersection gives badly shaped elements.
+Notice how the co-refinement around the intersection gives badly shaped elements.
 (c) The remeshed \dealii{} Triangulation.}

### 2022-06-23 9.4/paper.tex

@@
 The same applies to coupling terms of the form $(u,q)_B$ in formulations using Lagrange multipliers, where now $q \in Q_h(B)$, with $Q_h(B)$ the space of the multiplier variable.
-Notice that the most relevant difference with the quadrature rules provided by cutFEM support in \texttt{QuadratureGenerator} is that \texttt{Quadrature} objects are not created using a level-set approach, as described in the previous section, but directly with the two overlapped grids.
+Note that the most relevant difference between this and the \texttt{QuadratureGenerator} in Section~\ref{sec:cut} is that the \texttt{Quadrature} objects are created directly from the two overlapping grids and not from a level set function.
 

### 2022-06-23 9.4/paper.tex
Numerous changes

@@
   \item Performance improvements in the particle infrastructure (see Section~\ref{sec:particles});
-  \item Two new tutorial programs and a new code gallery program (see
+  \item Two new tutorial programs and one new code gallery program (see
     Section~\ref{subsec:steps}).
-    \todo[inline]{Update the number of code gallery programs when
-      we're done.}
 \end{itemize}
@@
   \item The new member function \texttt{find\_point\_owner\_rank()} of
-        \texttt{parallel\allowbreak ::distributed::\allowbreak Triangulation} allows one to find the MPI
+        \texttt{parallel\allowbreak ::distributed::\allowbreak Tri\allowbreak angulation} allows one to find the MPI
   ranks of cells containing specified points.
   It is communication-free and leverages the functionality of p4est (>v.2.2).
-  Its algorithm is described in \cite{burstedde2020parallel}. This information
-  enables efficient construction of the
+  Its algorithm is described in \cite{burstedde2020parallel}. This information will
+  enable efficient construction of the
   communication pattern used in the class \texttt{Utilties::\allowbreak MPI::\allowbreak RemotePointEvaluation}. Furthermore, this function could be used in the future to allow
@@
   functions, such as \texttt{FEInterfaceValues::get\_jump\_in\_function\_values()}, have also been added.
+  \item Vectors attached to \texttt{DataOut} do not need to be in ghosted state anymore. Internally, we create a copy of the vector with appropriate ghosting.
 \end{itemize}
@@
 previously only worked for hypercube-shaped cells. The most notable
-new functionality is:
+new functionalities are:
 
@@
 Figure~\ref{fig:refinement}), and (ii) appropriate hanging-node
-constraints. Both are now available in 2d; for 3D, the implementation
+constraints. Both are now available in 2D; for 3D, the implementation
 of the constraint definitions is still in progress.
-\item \todo[inline]{TODO}
+\item \texttt{QIteratedSimplex} allows to build composite simplex quadrature rules.
+\item The new wrappers to the \texttt{CGAL} library allow to simply create simplex
+meshes. For more details, see Section~\ref{sec:cgalwrappers}.
 \end{itemize}
@@
 (which in the following we will abbreviate as \texttt{p::d::T}) and
-\texttt{parallel::\allowbreak fullydistributed::\allowbreak
+\texttt{parallel::\allowbreak fully\allowbreak distributed::\allowbreak
   Triangulation} (or, in short, \texttt{p::f::T}). The former is
@@
 // select a partitioning policy and use it:
-const RepartitioningPolicyTools::Cartesian<dim> policy(tria);
+const RepartitioningPolicyTools::CellWeightPolicy<dim> policy(tria, fu);
 const auto construction_data = TriangulationDescription::Utilities::
@@
     \input{svg/repartitioning.pdf_tex}
-    \caption{\it Visualization of the repartitioning process. Top
+    \caption{\it Visualization of the repartitioning process and the
+      construction of the new mesh of process 0. Top
       left: Existing ownership of the cells of the mesh, distributed
@@
       collect information about the cells they own or that are ghost
-      cells. Processes 2 and 3 would do the same. Third column: What
+      cells. Processes 2 and 3 do not contribute to process 0. Third column: What
       processes 0 and 1 would send to process 0. Fourth column: The
@@
 
-The setup process is visualized in Figure~\ref{fig:repartitioning}. At
+The setup process (of \texttt{construction\_data}) is visualized in Figure~\ref{fig:repartitioning}. At
 first, locally owned cells and their
@@
 
+We would like to remind users that we transitioned from the use of Booleans
+to flags to configure the evaluation and integration process of \texttt{FEEvaluation}
+and \texttt{FEFaceEvaluation}:
+
+\begin{c++}
+fe\_eval.evaluate(false, true, true)          // old (deprecated)
+fe\_eval.evaluate(EvaluationFlags::gradients) // new
+\end{c++}
+
 %\begin{c++}
@@
 \texttt{workload\_imbalance()} and \texttt{vertical\_communication\_efficiency()}
-in the \texttt{MGTools} namespace for the  estimation of the workload imbalance and the
-vertical communication efficiency purely based on the given mesh.
+in the \texttt{MGTools} namespace for the  estimation of the imbalance during, e.g.,
+smoothing and the
+communication efficiency during intergrid transfer purely based on the given mesh.
 % MK: I would not add this part, it does not really fit into this paper as
@@
 \subsection{CutFEM support}\label{sec:cut}
-
-\begin{figure}
-  \centering
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \includegraphics[height=.12\paperheight]{svg/immersed-domain.pdf}
-    \caption{\label{fig:immersed-domain}}
-  \end{subfigure}
-  \qquad
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \includegraphics[height=.12\paperheight]{svg/location-to-level-set.pdf}
-    \caption{ \label{fig:location-to-level-set}}
-  \end{subfigure}
-  \caption{\it (a) Domain immersed in a background mesh. (b) Value of \texttt{LocationToLevelSet} for each cell.}
-\end{figure}
 
@@
 \end{align}
+
+\begin{figure}
+  \centering
+  \begin{subfigure}[b]{0.3\textwidth}
+    \centering
+    \includegraphics[height=.12\paperheight]{svg/immersed-domain.pdf}
+    \caption{\label{fig:immersed-domain}}
+  \end{subfigure}
+  \qquad
+  \begin{subfigure}[b]{0.3\textwidth}
+    \centering
+    \includegraphics[height=.12\paperheight]{svg/location-to-level-set.pdf}
+    \caption{ \label{fig:location-to-level-set}}
+  \end{subfigure}
+  \caption{\it (a) Domain immersed in a background mesh. (b) Value of \texttt{LocationToLevelSet} for each cell.}
+%\end{figure}
+%
+%\begin{figure}
+  \centering
+  \includegraphics[width=.4\paperwidth]{svg/immersed_quadratures.pdf}
+  \caption{\it Quadrature points for integrating over the three different regions of a cell cut by the zero contour of the level set function, $\psi$. \label{fig:immersed_quadratures}}
+\end{figure}
+
 Specifically, the following are the key new classes and functions:
@@
 
-  \item \texttt{FEImmersedSurfaceValues} is an \texttt{FEValues}-like class for evaluating real space values based on an \texttt{ImmersedSurfaceQuadrature}.
+  \item \texttt{FEImmersedSurfaceValues} is an \texttt{FEFaceValues}-like class for evaluating real space values based on an \texttt{ImmersedSurfaceQuadrature}.
 
@@
   \texttt{FEValues} objects for the inside/outside region and a \texttt{FEImmersedSurfaceValues} object for the surface regions are set up internally.
-These can then be obtained using getter-functions and used for the assembly.
+These can then be obtained using getter-functions (\texttt{get\_inside/outside/surface\_fe\_values()}) and used for the assembly.
   Since the generation of immersed quadrature rules is not cheap,
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=.4\paperwidth]{svg/immersed_quadratures.pdf}
-  \caption{\it Quadrature points for integrating over the three different regions of a cell cut by the zero contour of the level set function, $\psi$. \label{fig:immersed_quadratures}}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Integration with the Computational Geometry Algorithms
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Integration of the Computational Geometry Algorithms
   Library (CGAL)}\label{sec:cgalwrappers}
@@
 A prototypical use case is the following, where the surface is the zero level set of Taubin's heart function $f=\bigl ( x^2 + \frac{9y^2}{4} +z^2 -1 \bigr ) -x^2 z^3 - \frac{9y^2z^3}{80}$.
-The resulting \texttt{Triangulation<3>} can be appreciated in Figure~\ref{fig:heart_tria}.
+The resulting \texttt{Triangulation<3>} is shown in Figure~\ref{fig:heart_tria}
+and the required steps are:
 \begin{c++}
-// Empty triangulation.
-Triangulation<3, 3>  tria;
-//Taubin's heart surface
-ImplicitFunction                implicit_function;
-CGALWrappers::AdditionalData<3> data;
-data.cell_size = .05;
-GridGenerator::implicit_function(tria, implicit_function, data,
-                                 {0, 0, 0}, 10.0);
+// 1) An implicit function, e.g., Taubin's heart surface (not shown)
+ImplicitFunction implicit_fu;
+
+// 2) configure output mesh (optional)
+CGALWrappers::AdditionalData<3> data; data.cell_size = .05;
+
+// 3) create mesh
+Triangulation<3>  tria;
+GridGenerator::implicit_function(tria, implicit_fu, data, {0, 0, 0}, 10.0);
 \end{c++}
@@
 
-CGAL also provides us with boolean operations on meshes, as available in the
+CGAL also provides Boolean operations on meshes, as available in the
 utility function
@@
 \begin{c++}
+// 1) create deal.II triangulations, e.g., cube and sphere (not shown)  
 Triangulation<spacedim> tria0, tria1;
-  GridGenerator::hyper_cube(tria0,-1.,1.);
-  GridGenerator::hyper_ball(tria1,{1.,1.,1.},0.6);
-  tria0.refine_global(3);
-  tria1.refine_global(3);
-
-  // Move to CGAL surfaces, assume Kernel is already defined. 
+
+// 2) convert to CGAL surface meshes (assuming Kernel is already defined) 
 CGAL::Surface_mesh<Kernel> surface_mesh0, surface_mesh1;
@@
 
-  // Compute the union of the two meshes
-  CGALWrappers::compute_boolean_operation(surface_mesh0,
-                            surface_mesh1,
-                            BooleanOperation::compute_union,
-                            out_mesh);
-  // Now back to deal.II
+// 3) compute the union of the two meshes
+CGALWrappers::compute_boolean_operation(surface_mesh0, surface_mesh1,
+  BooleanOperation::compute_union, out_mesh);
+
+// 4) convert CGAL surface mesh to deal.II surface mesh
 Triangulation<2, 3>     tria_out;
 CGALWrappers::cgal_surface_mesh_to_dealii_triangulation(out_mesh, tria_out);
-  // tria_out is now a valid deal.II triangulation
+
+// 5) convert surface to volume mesh via surface_mesh_to_volumetric_mesh()
   \end{c++}
@@
     \centering
-    \begin{subfigure}[b]{0.3\textwidth}
+    \begin{subfigure}[b]{0.28\textwidth}
       \centering
@@
       \caption{\label{fig:heart_tria}}
-    \end{subfigure}
+    \end{subfigure}\qquad
     \hfill
-    \begin{subfigure}[b]{0.3\textwidth}
+    \begin{subfigure}[b]{0.28\textwidth}
       \centering
@@
     \hfill
-    \begin{subfigure}[b]{0.37\textwidth}
+    \begin{subfigure}[b]{0.35\textwidth}
       \centering
@@
 
-These utilities will be the building blocks for adding functions to the \texttt{NonMatching} namespace that will assemble coupling terms like $(u,v)_{B}$,\marginpar{The $B$ and $\Omega$ here need to be streamlined with \eqref{eq:boundingbox}, right now they are the opposite} with $B$ a domain immersed in a fixed background mesh $\Omega$ and $u,v$ finite element functions on $V_h(\Omega)$, as usually happens using Nitsche's method to weakly impose boundary conditions at an interface.
+These utilities will be the building blocks for adding functions to the \todo{NonMatching} namespace that will assemble coupling terms like $(u,v)_{B}$,\marginpar{The $B$ and $\Omega$ here need to be streamlined with \eqref{eq:boundingbox}, right now they are the opposite} with $B$ a domain immersed in a fixed background mesh $\Omega$ and $u,v$ finite element functions on $V_h(\Omega)$, as needed, e.g., in
+the context of CutFEM (see Section~\ref{sec:cut}) of Nitsche's method to weakly impose boundary conditions at an interface.
 The same applies to coupling terms of the form $(u,q)_B$ in formulations using Lagrange multipliers, where now $q \in Q_h(B)$, with $Q_h(B)$ the space of the multiplier variable.
-Note that the most relevant difference between this and the \texttt{QuadratureGenerator} in Section~\ref{sec:cut} is that the \texttt{Quadrature} objects are created directly from the two overlapping grids and not from a level set function.
+Note that the most relevant difference between this and the \texttt{QuadratureGenerator} in Section~\ref{sec:cut} is that the \texttt{Quadrature} objects are created directly from two overlapping grids, one
+spanning over $B$ and the other one over $\Omega$, and not from a level set function.
 
@@
 We keep a separate cache structure that contains pointers to the particular list entries for each cell.
-If a cell has no particles this pointer is invalid.
+If a cell has no particles, this pointer is invalid.
 
@@
 \item All particle data (both their identifiers and their actual data) are now stored as separate and contiguous arrays in memory, which improves spatial locality for better prefetching of data and makes iterating over particles extremely efficient.
-\item The choice of a list container that only includes entries for cells that contain particles means iteration is efficient, even if many cells in the domain do not contain any particles (as can be the case for discrete-element methods).
+\item The choice of a list container that only includes entries for cells that contain particles means iteration is efficient, even if many cells in the domain do not contain any particles (as can be the case for discrete element methods~\cite{golshan2022lethe}).
 \item Creating separate arrays for each cell allows us to easily move particle IDs from one cell to another as a local operation, affecting only the two cell containers in question.
@@
 In addition, after sorting all particles into their new cells, the arrays that store particle properties are now sorted in the same order as the particle IDs in the list of arrays, which allows for cache efficient iteration over particle properties.
-To avoid a costly sorting operation this operation is executed as a copy of the existing data into a new data container that replaces the existing container.
+To avoid a costly sorting operation this operation is executed as a copy of the existing data into a new data container that replaces the existing container\todo{@RG: I don't get this sentence.}.
 
@@
 much faster faster in \dealii 9.4 than in version 9.3.
-In particular operations that depend strongly on particle storage structure and require few fixed computations (like iteration and sorting) benefit massively from the above mentioned optimizations.
+In particular, operations that depend strongly on particle storage structure and require few fixed computations (like iteration and sorting) benefit massively from the above mentioned optimizations.
 We note that the exact gains will depend strongly on the exact combination of geometry, mapping, dimensionality, and number of particles per cell in any specific model, and can be smaller or larger than the measurements provided here.
@@
 who will then have to respond with replies. This problem is solved by
-``consensus algorithms''. An example of where this problem appears is
+``consensus algorithms''~\cite{hoefler2010scalable}. An example of where this problem appears is
 given in Section~\ref{sec:repartitioning}.
@@
 A special case of consensus algorithms is where the sender does not
-actually require an answer. This happens, for example, during ghost
-exchange where each process sends information about cells it owns to
-some of its neighboring processes where these cells are ghost
-cells. Previously, this case was implemented through a consensus
+actually require an answer. This happens, for example, during repartitioning
+of meshes (Section~\ref{sec:repartitioning}), where a process sends parts of
+the new mesh to the new owners: the new owner does not know how many processes
+will send it mesh parts and the sender only needs an acknowledgment that the
+data has been received. Previously, such a case was implemented through a consensus
 algorithm where the reply message is simply empty.
@@
 future, using the interfaces now already in place.
+
+By the time of writing, \dealii uses consensus-based algorithms to determine
+the owners of distributed index sets (\texttt{Utilities::MPI::Partitioner},
+\texttt{Utilities::MPI::Noncontiguous\allowbreak Partitioner}, \texttt{internal::MatrixFreeFunctions::VectorDataExchange}; see~\cite{dealII91}),
+to setup the global-coarsening transfer operators (see~\cite{dealII92} and
+Section~\ref{sec:multigrid}), to repartition distributed meshes (see Section~\ref{sec:repartitioning}), and basis coupling algorithms between non-matching
+meshes, based on the communication patters in \texttt{RemotePointeEvaluation} (see~\cite{dealII92}).
 
@@
 \begin{itemize}
-  \item \todo[inline]{Is there actually a new program?}
+  \item ``\texttt{TRBDF2-DG projection solver for the incompressible Navier--Stokes equations}'' was contributed by Giuseppe Orlando Politecnico di Milano. It shows
+  how to solve the incompressible Navier--Stokes equations efficiently 
+  with \texttt{deal.II}'s matrix-free DG infrastructure, multigrid, and adaptive-mesh
+  refinement. Interested readers are referred to~\cite{orlando2021efficient}.
 \end{itemize}

### 2022-06-23 9.4/paper.tex
Remove duplicate p4est publication. Adjust p4est references.

@@
 \usepackage{pdfsync}
+\usepackage{enumitem}
 
@@
 
-\title{The \dealii{} Library, Version 9.4}
+\title{The \dealii Library, Version 9.4}
 
@@
    34136, Trieste, Italy.
-   {\texttt{ \{marco.feder,luca.heltai\}@sissa.it}}}
+   {\texttt{marco.feder/luca.heltai@sissa.it}}}
 
@@
 
-\author[13]{Rene~Gassm{\"o}ller}
-\affil[13]{Department of Geological Sciences,
+\author[5]{Rene~Gassm{\"o}ller}
+\affil[5]{Department of Geological Sciences,
    University of Florida,
@@
 
- \author[6,7]{Martin~Kronbichler}
- \affil[6]{Department of Information Technology,
+ \author[7,8]{Martin~Kronbichler}
+ \affil[7]{Department of Information Technology,
    Uppsala University,
@@
    {\texttt{martin.kronbichler/simon.sticko@it.uu.se}}}
- \affil[7]{Institute of Mathematics,
+ \affil[8]{Institute of Mathematics,
    University of Augsburg,
@@
 
-\author[7,9]{Peter Munch}
- \affil[9]{Institute of Material Systems Modeling,
+\author[8,10]{Peter Munch}
+ \affil[10]{Institute of Material Systems Modeling,
  Helmholtz-Zentrum Hereon,
@@
 
-\author[10]{Jean-Paul~Pelteret}
-\affil[10]{Independent researcher.
+\author[11]{Jean-Paul~Pelteret}
+\affil[11]{Independent researcher.
 {\texttt{jppelteret@gmail.com}}}
 
-\author[6,11]{Simon~Sticko}
-\affil[11]{Department of Mathematics and Mathematical Statistics,
+\author[7,12]{Simon~Sticko}
+\affil[12]{Department of Mathematics and Mathematical Statistics,
    Umeå University,
@@
 
-\author[12]{David Wells}
-\affil[12]{Department of Mathematics, University of North Carolina,
+\author[13]{David Wells}
+\affil[13]{Department of Mathematics, University of North Carolina,
   Chapel Hill, NC 27516, USA.
@@
 
-\dealii{} version 9.4.0 was released June 10, 2022.
+\dealii version 9.4.0 was released June 10, 2022.
 \todo{Update release date.}
@@
 overview of the new features of this release and serves as a citable
-reference for the \dealii{} software library version 9.4. \dealii{} is an
+reference for the \dealii software library version 9.4. \dealii is an
 object-oriented finite element library used around the world in the
@@
   \item CutFEM support (see Section~\ref{sec:cut});
-  \item Integration with the Computational Geometry Algorithms Library (CGAL) (see Section~\ref{sec:cgalwrappers});
+  \item Experimental integration of the Computational Geometry Algorithms Library (CGAL) (see Section~\ref{sec:cgalwrappers});
   \item Performance improvements in the particle infrastructure (see Section~\ref{sec:particles});
-  \item Two new tutorial programs and one new code gallery program (see
-    Section~\ref{subsec:steps}).
+  \item Improvements to unstructured communication (see Section~\ref{sec:CA});
+  \item Three new tutorial programs and one new code gallery program (see Section~\ref{subsec:steps}).
 \end{itemize}
@@
 Section~\ref{sec:major}, there
-are a number of other noteworthy changes in the current \dealii{} release,
+are a number of other noteworthy changes in the current \dealii release,
 which we briefly outline in the remainder of this section:
@@
   ranks of cells containing specified points.
-  It is communication-free and leverages the functionality of p4est (>v.2.2).
+  It is communication-free and leverages the functionality of \pfrst (>v.2.2).
   Its algorithm is described in \cite{burstedde2020parallel}. This information will
@@
 
-This release of \dealii{} contains a number of large and significant changes,
+This release of \dealii contains a number of large and significant changes,
 which will be discussed in this section.
@@
 
-\dealii{} has supported simplex and mixed meshes since the previous
+\dealii has supported simplex and mixed meshes since the previous
 release, 9.3, on an experimental basis.
@@
 sufficient to support several larger-scale programs, although not all
-functionality in \dealii{} works with such meshes yet. Specifically,
+functionality in \dealii works with such meshes yet. Specifically,
 we have fixed many bugs and generalized existing functions that
@@
 
-\dealii{} has two classes that support distributed storage of meshes: 
+\dealii has two classes that support distributed storage of meshes:
 \texttt{parallel::\allowbreak distributed::\allowbreak Triangulation}
@@
 partitioned based on the space-filling Morton (or Z-order) curve
-implemented by the \texttt{p4est} backend. The latter, more recent
+implemented by the \pfrst backend. The latter, more recent
 class is currently statically partitioned at creation time.
@@
 
-The matrix-free infrastructure of \dealii{} is used to solve problems
+The matrix-free infrastructure of \dealii is used to solve problems
 without assembling matrices, providing only the action of the matrix
@@
 the interface in a special way. A challenge is that cell batches
-might contain cut or not-cut cells, making vectorization of these operations potentially more complicated.
+might contain cut or non-cut cells, making the vectorization of these operations potentially more complicated.
 Previous functionality has provided the option of masking certain cells in cell batches, which works well if
@@
 might be too expensive if recategorization needs to happen very frequently to follow the dynamics of a system, e.g., in each time step.
-Despite some overhead compared to static matrix-free loops, the new feature can be the best options in these scenarios.
+Despite some overhead compared to static matrix-free loops, the new feature can be the best option in these scenarios.
 \item Initial support for H(div)-conforming elements with Piola transform, based on Raviart--Thomas finite element spaces with the updated class \texttt{FE\_RaviartThomasNodal}, has been added.
@@
   matrix-vector product touches them, and the second what happens
-  afterwards. The new features also include a renumbering to maximize data
+  afterward. The new features also include a renumbering to maximize data
   locality. The theory is described in the
@@
 
-Finally, we have performed a major restructuring of internals
+Finally, we have performed a major restructuring of the internals
 of the \texttt{FEEvaluation} classes. This reduces some overheads for low polynomial degrees and will enable us to add support for new element types in the future.
@@
 \begin{c++}
-fe\_eval.evaluate(false, true, true)          // old (deprecated)
-fe\_eval.evaluate(EvaluationFlags::gradients) // new
+fe_eval.evaluate(false, true, false)         // old (deprecated)
+fe_eval.evaluate(EvaluationFlags::gradients) // new
 \end{c++}
@@
 addition to the established local-smoothing infrastructure. Global
-coarsening algorithms smoothen on the whole computational domain on each
+coarsening algorithms smoothen over the whole computational domain on each
 multigrid level, which is obtained by coarsening the finest cells of
@@
 triangulations, one can use the functions \texttt{MGTransferGlobalCoarseningTools::create\_geometric\_coarsening\_sequence()}. A new version takes an
-instance of \texttt{RepartitioningPolicyTools::Base} (see Subsection~\ref{sec:repartitioning}) as argument, which allows to specify the parallel
-distribution of each multigrid level (in contrast to the fixed first-child policy in the case of local smoothing). These features have been developed and tuned for runs on supercomputer scale
+instance of \texttt{RepartitioningPolicyTools::Base} (see Subsection~\ref{sec:repartitioning}) as argument, which allows specifying the parallel
+distribution of each multigrid level (in contrast to the fixed first-child policy in the case of local smoothing). These features have been developed and tuned for running on a supercomputer scale
 with complicated coarse meshes as presented in \cite{kronbichler2021next}.
@@
 In \cite{munch2022gc}, the performance of the local-smoothing and global-coarsening
-infrastructure of \texttt{deal.II} was compared for locally refined meshes. The results indicate that
+infrastructure of \dealii was compared for locally refined meshes. The results indicate that
 the local definition of multigrid levels might introduce load imbalances
@@
 potentially more expensive intergrid transfers. In order to judge the benefits
-of one approach against the other, \texttt{deal.II} provides new functions
+of one approach against the other, \dealii provides new functions
 \texttt{workload\_imbalance()} and \texttt{vertical\_communication\_efficiency()}
 in the \texttt{MGTools} namespace for the  estimation of the imbalance during, e.g.,
-smoothing and the
-communication efficiency during intergrid transfer purely based on the given mesh.
+smoothing or the
+communication efficiency during intergrid transfer, purely based on the given mesh.
 % MK: I would not add this part, it does not really fit into this paper as
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Integration of the Computational Geometry Algorithms
+\subsection{Experimental integration of the Computational Geometry Algorithms
   Library (CGAL)}\label{sec:cgalwrappers}
@@
 The Computational Geometry Algorithms Library (CGAL, \url{https://www.cgal.org/}) is a widely used
-library to describe geometries and meshes \cite{cgal}. \dealii{} now
+library to describe geometries and meshes \cite{cgal}. \dealii now
 has wrappers for CGAL classes and functions, provided in the new
 namespace \texttt{CGALWrappers}, and implementing functionality
-spanning from mesh generation to boolean operations between \dealii{}
-triangulations and cells. These wrappers are enabled only if \dealii{}
-is compiled with \texttt{C++17} and, for the first time in \dealii{}'s
-25-year history, provides a built-in ability to generate meshes for
-arbitrary geometries.
+spanning from mesh generation to boolean operations between \dealii
+triangulations and cells. These wrappers are enabled only if \dealii
+is compiled with \texttt{C++17}. \textit{Note: This feature is still experimental and
+interfaces might change during the next release cycle.}
 
@@
   \end{subfigure}
-    \caption{\it (a) \texttt{Triangulation<3>} of a heart surface starting from an implicit function $f$.
-(b) Union of the unit cube with an hyper ball.
-Notice how the co-refinement around the intersection gives badly shaped elements.
-(c) The remeshed \dealii{} Triangulation.}
+  \caption{\it (a) Triangulation created by filling a heart-shaped surface implicitly described by a function $f$.
+(b) Union of a cube with a sphere with badly shaped cells at the intersection.
+(c) Remeshed version of the same triangulation.}
 \end{figure}
 
-\texttt{CGALWrappers::compute\_quadrature\_on\_boolean\_operation()} returns a \texttt{Quadrature<3>} that allows exact integration on polyhedral elements coming out of a \texttt{BooleanOperation} between \dealii cells.
+\texttt{CGALWrappers::compute\_quadrature\_on\_boolean\_operation()} returns a \texttt{Quadrature<3>} that allows exact integration of polynomials on polyhedral elements coming out of a \texttt{BooleanOperation} between \dealii cells.
 The quadrature rule is built by meshing the polyhedral region with tetrahedra, computing on each tetrahedron a \texttt{QGaussSimplex<3>} quadrature rule by using \texttt{QSimplex<3>::\allowbreak{}compute\_affine\_transformation()}, and finally
@@
 
-These utilities will be the building blocks for adding functions to the \todo{NonMatching} namespace that will assemble coupling terms like $(u,v)_{B}$,\marginpar{The $B$ and $\Omega$ here need to be streamlined with \eqref{eq:boundingbox}, right now they are the opposite} with $B$ a domain immersed in a fixed background mesh $\Omega$ and $u,v$ finite element functions on $V_h(\Omega)$, as needed, e.g., in
-the context of CutFEM (see Section~\ref{sec:cut}) of Nitsche's method to weakly impose boundary conditions at an interface.
-The same applies to coupling terms of the form $(u,q)_B$ in formulations using Lagrange multipliers, where now $q \in Q_h(B)$, with $Q_h(B)$ the space of the multiplier variable.
+These utility functions will be the building blocks for functions in the \texttt{NonMatching} namespace that will, e.g.,  assemble coupling terms like $(u,v)_{\Omega}$, with $\Omega$ a domain immersed in a fixed background mesh $B$ and $u,v$ finite element functions on $V_h(B)$, as needed, e.g., in
+the context of CutFEM (see Section~\ref{sec:cut}) or Nitsche's method to weakly impose boundary conditions at an interface.
+The same applies to coupling terms of the form $(u,q)_{\Omega}$ in formulations using Lagrange multipliers, where now $q \in Q_h(\Omega)$, with $Q_h(\Omega)$ the space of the multiplier variable.
 Note that the most relevant difference between this and the \texttt{QuadratureGenerator} in Section~\ref{sec:cut} is that the \texttt{Quadrature} objects are created directly from two overlapping grids, one
@@
 structures and optimized the algorithms that support using particles
-in \dealii{}.
+in \dealii.
 In our previously reported improvements~\citep{dealII93} we stored all particle data as a separate contiguous array for each particle property, but particle identifiers (IDs) were still stored in a multimap tree-like structure.
@@
 In addition, after sorting all particles into their new cells, the arrays that store particle properties are now sorted in the same order as the particle IDs in the list of arrays, which allows for cache efficient iteration over particle properties.
-To avoid a costly sorting operation this operation is executed as a copy of the existing data into a new data container that replaces the existing container\todo{@RG: I don't get this sentence.}.
+Because the particle IDs are already sorted in the intended order at this time, we can reorder the particle properties as part of a copy operation into a new data container that replaces the existing container.
+This approach avoids a costly in-place sort of the particle properties.
 
@@
 Table~\ref{tab:particle_timing} shows that all particle operations are
-much faster faster in \dealii 9.4 than in version 9.3.
-In particular, operations that depend strongly on particle storage structure and require few fixed computations (like iteration and sorting) benefit massively from the above mentioned optimizations.
-We note that the exact gains will depend strongly on the exact combination of geometry, mapping, dimensionality, and number of particles per cell in any specific model, and can be smaller or larger than the measurements provided here.
+much faster in \dealii 9.4 than in version 9.3.
+In particular, operations that depend strongly on particle storage structure and require few fixed computations (like iteration and sorting) benefit massively from the above-mentioned optimizations.
+We note that the exact gains will depend strongly on the exact combination of geometry, mapping, dimensionality, and the number of particles per cell in any specific model, and can be smaller or larger than the measurements provided here.
 
@@
 way: Each process in a parallel universe has a number of queries to
-send to other processes who do not know that they will be asked, and
+send to other processes that do not know that they will be asked, and
 who will then have to respond with replies. This problem is solved by
@@
 
-\dealii{} has had an implementation of these algorithms for some time,
+\dealii has an implementation of these algorithms for some time,
 but the current release substantially expands on it. Specifically, the
@@
 The rewritten interfaces now support this case more explicitly: Code using
-these interfaces no longer have to provide functions that formulate
+these interfaces no longer has to provide functions that formulate
 and read the (empty) replies, though internally these functions still
@@
 
-Many of the \dealii{} tutorial programs were revised in a variety of
-ways as part of this release. In addition, there are a number of new tutorial programs:
+Many of the \dealii tutorial programs were revised in a variety of ways
+as part of this release. In addition, there are a number of new tutorial
+programs:
 \begin{itemize}
-\item \texttt{step-82} was contributed by Andrea Bonito (Texas A\&M University) and Diane Guignard (University of Ottawa). It shows how
-\dealii{} can be used to implement the
-local discontinuous Galerkin (LDG) method for approximating the
-solution to the bi-Laplacian problem. The method is an alternative to
-the C0IP method used in \texttt{step-47}.
-
-\item \texttt{step-85} was contributed by Simon Sticko (Uppsala University). It shows how to use the new CutFEM infrastructure to solve a Poisson
-problem on a circular domain that is embedded in a Cartesian background mesh.
+  \item
+    \texttt{step-81} was contributed by Manaswinee Bezbaruah (Texas A\&M
+    University) and Matthias Maier (Texas A\&M University). It explains how
+    to solve the complex-valued time-harmonic Maxwell's equations for an
+    optical scattering problem.
+  \item
+    \texttt{step-82} was contributed by Andrea Bonito (Texas A\&M
+    University) and Diane Guignard (University of Ottawa). It shows how
+    \dealii can be used to implement the local discontinuous Galerkin
+    (LDG) method for approximating the solution to the bi-Laplacian
+    problem. The method is an alternative to the C0IP method used in
+    \texttt{step-47}.
+  \item
+    \texttt{step-85} was contributed by Simon Sticko (Uppsala University).
+    It shows how to use the new CutFEM infrastructure to solve a Poisson
+    problem on a circular domain that is embedded in a Cartesian background
+    mesh.
 \end{itemize}
@@
 \begin{itemize}
-  \item ``\texttt{TRBDF2-DG projection solver for the incompressible Navier--Stokes equations}'' was contributed by Giuseppe Orlando Politecnico di Milano. It shows
+  \item ``TRBDF2-DG projection solver for the incompressible Navier--Stokes equations'' was contributed by Giuseppe Orlando (Politecnico di Milano). It shows
   how to solve the incompressible Navier--Stokes equations efficiently 
-  with \texttt{deal.II}'s matrix-free DG infrastructure, multigrid, and adaptive-mesh
+  with \dealii's matrix-free DG infrastructure, multigrid, and adaptive-mesh
   refinement. Interested readers are referred to~\cite{orlando2021efficient}.
 \end{itemize}
-Finally, the ``\texttt{MCMC for the Laplace equation}'' code gallery
+Finally, the ``MCMC for the Laplace equation'' code gallery
 program has been updated by providing MATLAB and Python versions of
@@
     \texttt{parallel::distributed::Triangulation} objects include
-  the removal the default weight of each cell in order to ensure
-  consistency with the rest of the library.
+  the removal of the default weight of each cell in order to ensure
+  consistency with the rest of the library and to improve flexibility.
 \end{itemize}
@@
 
-In order to justify the work the developers of \dealii{} put into this
+In order to justify the work the developers of \dealii put into this
 software, we ask that papers using the library reference one of the
-\dealii{} papers. This helps us justify the effort we put into this library.
-
-There are various ways to reference \dealii{}. To acknowledge the use of
+\dealii papers. This helps us justify the effort we put into this library.
+
+There are various ways to reference \dealii. To acknowledge the use of
 the current version of the library, \textbf{please reference the present
-  document}. For up to date information and a bibtex entry
+  document}. For up-to-date information and a bibtex entry
 see
@@
 
-The original \dealii{} paper containing an overview of its
+The original \dealii paper containing an overview of its
 architecture is \cite{BangerthHartmannKanschat2007}, and a more recent
-publication documenting \dealii{}'s design decisions is available as \cite{dealII2020design}. If you rely on
+publication documenting \dealii's design decisions is available as \cite{dealII2020design}. If you rely on
 specific features of the library, please consider citing any of the
@@
   \vspace*{-36pt}
-  \begin{itemize}
+  \begin{itemize}[leftmargin=4mm]
     \item For geometric multigrid: \cite{Kanschat2004,JanssenKanschat2011,ClevengerHeisterKanschatKronbichler2019, munch2022gc};
@@
           \cite{SartoriGiulianiBardelloni-2018-a};
-    \item For uses of the particle functionality in \dealii{}:
+    \item For uses of the particle functionality in \dealii:
           \cite{GLHPB18}.
@@
 
-\dealii{} can interface with many other libraries:
+\dealii can interface with many other libraries:
 \begin{multicols}{3}
-  \begin{itemize}
+  \begin{itemize}[leftmargin=4mm]
     \item ADOL-C \cite{Griewank1996a,adol-c}
@@
     \item OpenCASCADE \cite{opencascade-web-page}
-    \item p4est \cite{p4est}
+    \item p4est \cite{p4est,burstedde2020parallel}
     \item PETSc \cite{petsc-user-ref,petsc-web-page}
@@
 
-The two previous releases of \dealii{} can be cited as
+The two previous releases of \dealii can be cited as
 \cite{dealII92,dealII93}.
@@
 
-\dealii{} is a world-wide project with dozens of contributors around the
+\dealii is a world-wide project with dozens of contributors around the
 globe. Other than the authors of this paper, the following people
@@
 Konrad        Simon,
-Simon         Sticko,
 Daniel        Sun,

### 2022-06-23 9.4/paper.tex
Update references for Trilinos, PETSc, Mumps, Ginkgo, GSL, CGAL, Gmesh, hdf5.
Use deal.II command for abbreviation.

--- 2022-06-23 9.4/paper.tex
@@
 In \cite{munch2022gc}, the performance of the local-smoothing and global-coarsening
-infrastructure of \texttt{deal.II} was compared for locally refined meshes. The results indicate that
+infrastructure of \dealii{} was compared for locally refined meshes. The results indicate that
 the local definition of multigrid levels might introduce load imbalances
@@
 potentially more expensive intergrid transfers. In order to judge the benefits
-of one approach against the other, \texttt{deal.II} provides new functions
+of one approach against the other, \dealii{} provides new functions
 \texttt{workload\_imbalance()} and \texttt{vertical\_communication\_efficiency()}
@@
   how to solve the incompressible Navier--Stokes equations efficiently 
-  with \texttt{deal.II}'s matrix-free DG infrastructure, multigrid, and adaptive-mesh
+  with \dealii{}'s matrix-free DG infrastructure, multigrid, and adaptive-mesh
   refinement. Interested readers are referred to~\cite{orlando2021efficient}.

### 2022-06-23 9.4/paper.tex
Fix code
More minor modifications

@@
 \begin{c++}
-fe\_eval.evaluate(false, true, true)          // old (deprecated)
+fe\_eval.evaluate(false, true, false)         // old (deprecated)
 fe\_eval.evaluate(EvaluationFlags::gradients) // new
@@
     \end{subfigure}
-    \caption{\it (a) \texttt{Triangulation<3>} of a heart surface starting from an implicit function $f$.
-(b) Union of the unit cube with an hyper ball.
-Notice how the co-refinement around the intersection gives badly shaped elements.
-(c) The remeshed \dealii{} Triangulation.}
+    \caption{\it (a) Triangulation created by filling a heart-shaped surface implicitly described by a function $f$.
+(b) Union of a cube with a sphere with badly shaped cells at the intersection.
+(c) Remeshed version of the same triangulation.}
   \end{figure}

### 2022-06-23 9.4/paper.tex
Fix some typos

@@
 
-These utilities will be the building blocks for adding functions to the \todo{NonMatching} namespace that will assemble coupling terms like $(u,v)_{B}$,\marginpar{The $B$ and $\Omega$ here need to be streamlined with \eqref{eq:boundingbox}, right now they are the opposite} with $B$ a domain immersed in a fixed background mesh $\Omega$ and $u,v$ finite element functions on $V_h(\Omega)$, as needed, e.g., in
-the context of CutFEM (see Section~\ref{sec:cut}) of Nitsche's method to weakly impose boundary conditions at an interface.
+These utility functions will be the building blocks for functions in the \texttt{NonMatching} namespace that will, e.g.,  assemble coupling terms like $(u,v)_{B}$,\todo{The $B$ and $\Omega$ here need to be streamlined with \eqref{eq:boundingbox}, right now they are the opposite} with $B$ a domain immersed in a fixed background mesh $\Omega$ and $u,v$ finite element functions on $V_h(\Omega)$, as needed, e.g., in
+the context of CutFEM (see Section~\ref{sec:cut}) or Nitsche's method to weakly impose boundary conditions at an interface.
 The same applies to coupling terms of the form $(u,q)_B$ in formulations using Lagrange multipliers, where now $q \in Q_h(B)$, with $Q_h(B)$ the space of the multiplier variable.

### 2022-06-23 9.4/paper.tex
One more typo

@@
   \item CutFEM support (see Section~\ref{sec:cut});
-  \item Integration with the Computational Geometry Algorithms Library (CGAL) (see Section~\ref{sec:cgalwrappers});
+  \item Integration of the Computational Geometry Algorithms Library (CGAL) (see Section~\ref{sec:cgalwrappers});
   \item Performance improvements in the particle infrastructure (see Section~\ref{sec:particles});
+  \item Improvements to unstructured communication (see Section~\ref{sec:CA});
   \item Two new tutorial programs and one new code gallery program (see
@@
 \begin{c++}
-fe\_eval.evaluate(false, true, false)         // old (deprecated)
-fe\_eval.evaluate(EvaluationFlags::gradients) // new
+fe_eval.evaluate(false, true, false)         // old (deprecated)
+fe_eval.evaluate(EvaluationFlags::gradients) // new
 \end{c++}

### 2022-06-23 9.4/paper.tex
Make author mail addresses consistent.
Fix some typos.

@@
 partitioned based on the space-filling Morton (or Z-order) curve
-implemented by the \texttt{p4est} backend. The latter, more recent
+implemented by the \pfrst{} backend. The latter, more recent
 class is currently statically partitioned at creation time.
@@
 the interface in a special way. A challenge is that cell batches
-might contain cut or not-cut cells, making vectorization of these operations potentially more complicated.
+might contain cut or non-cut cells, making vectorization of these operations potentially more complicated.
 Previous functionality has provided the option of masking certain cells in cell batches, which works well if
@@
 instance of \texttt{RepartitioningPolicyTools::Base} (see Subsection~\ref{sec:repartitioning}) as argument, which allows to specify the parallel
-distribution of each multigrid level (in contrast to the fixed first-child policy in the case of local smoothing). These features have been developed and tuned for runs on supercomputer scale
+distribution of each multigrid level (in contrast to the fixed first-child policy in the case of local smoothing). These features have been developed and tuned for running on supercomputer scale
 with complicated coarse meshes as presented in \cite{kronbichler2021next}.
@@
 Table~\ref{tab:particle_timing} shows that all particle operations are
-much faster faster in \dealii 9.4 than in version 9.3.
+much faster in \dealii 9.4 than in version 9.3.
 In particular, operations that depend strongly on particle storage structure and require few fixed computations (like iteration and sorting) benefit massively from the above mentioned optimizations.
@@
 
-\dealii{} has had an implementation of these algorithms for some time,
+\dealii{} has an implementation of these algorithms for some time,
 but the current release substantially expands on it. Specifically, the

### 2022-06-23 9.4/paper.tex
Minor language fixes.

--- 2022-06-23 9.4/paper.tex
@@
 the interface in a special way. A challenge is that cell batches
-might contain cut or non-cut cells, making vectorization of these operations potentially more complicated.
+might contain cut or non-cut cells, making the vectorization of these operations potentially more complicated.
 Previous functionality has provided the option of masking certain cells in cell batches, which works well if
@@
 triangulations, one can use the functions \texttt{MGTransferGlobalCoarseningTools::create\_geometric\_coarsening\_sequence()}. A new version takes an
-instance of \texttt{RepartitioningPolicyTools::Base} (see Subsection~\ref{sec:repartitioning}) as argument, which allows to specify the parallel
-distribution of each multigrid level (in contrast to the fixed first-child policy in the case of local smoothing). These features have been developed and tuned for running on supercomputer scale
+instance of \texttt{RepartitioningPolicyTools::Base} (see Subsection~\ref{sec:repartitioning}) as argument, which allows specifying the parallel
+distribution of each multigrid level (in contrast to the fixed first-child policy in the case of local smoothing). These features have been developed and tuned for running on a supercomputer scale
 with complicated coarse meshes as presented in \cite{kronbichler2021next}.
@@
 in the \texttt{MGTools} namespace for the  estimation of the imbalance during, e.g.,
-smoothing and the
-communication efficiency during intergrid transfer purely based on the given mesh.
+smoothing or the
+communication efficiency during intergrid transfer, purely based on the given mesh.
 % MK: I would not add this part, it does not really fit into this paper as
@@
 much faster in \dealii 9.4 than in version 9.3.
-In particular, operations that depend strongly on particle storage structure and require few fixed computations (like iteration and sorting) benefit massively from the above mentioned optimizations.
-We note that the exact gains will depend strongly on the exact combination of geometry, mapping, dimensionality, and number of particles per cell in any specific model, and can be smaller or larger than the measurements provided here.
+In particular, operations that depend strongly on particle storage structure and require few fixed computations (like iteration and sorting) benefit massively from the above-mentioned optimizations.
+We note that the exact gains will depend strongly on the exact combination of geometry, mapping, dimensionality, and the number of particles per cell in any specific model, and can be smaller or larger than the measurements provided here.
 
@@
 the current version of the library, \textbf{please reference the present
-  document}. For up to date information and a bibtex entry
+  document}. For up-to-date information and a bibtex entry
 see

### 2022-06-23 9.4/paper.tex

--- 2022-06-23 9.4/paper.tex
@@
 
-\texttt{CGALWrappers::compute\_quadrature\_on\_boolean\_operation()} returns a \texttt{Quadrature<3>} that allows exact integration on polyhedral elements coming out of a \texttt{BooleanOperation} between \dealii cells.
+\texttt{CGALWrappers::compute\_quadrature\_on\_boolean\_operation()} returns a \texttt{Quadrature<3>} that allows exact integration of polynomials on polyhedral elements coming out of a \texttt{BooleanOperation} between \dealii cells.
 The quadrature rule is built by meshing the polyhedral region with tetrahedra, computing on each tetrahedron a \texttt{QGaussSimplex<3>} quadrature rule by using \texttt{QSimplex<3>::\allowbreak{}compute\_affine\_transformation()}, and finally

### 2022-06-23 9.4/paper.tex
Remove myself from the contributor list.
Since I'm already on the author list.
Minor changes.

@@
 \begin{itemize}
-  \item ``\texttt{TRBDF2-DG projection solver for the incompressible Navier--Stokes equations}'' was contributed by Giuseppe Orlando Politecnico di Milano. It shows
+  \item ``TRBDF2-DG projection solver for the incompressible Navier--Stokes equations'' was contributed by Giuseppe Orlando (Politecnico di Milano). It shows
   how to solve the incompressible Navier--Stokes equations efficiently 
@@
 \end{itemize}
-Finally, the ``\texttt{MCMC for the Laplace equation}'' code gallery
+Finally, the ``MCMC for the Laplace equation'' code gallery
 program has been updated by providing MATLAB and Python versions of
@@
 \begin{multicols}{3}
-  \begin{itemize}
+  \begin{itemize}[leftmargin=4mm]
     \item ADOL-C \cite{Griewank1996a,adol-c}

### 2022-06-23 9.4/paper.tex
Add comment regarding experimental

@@
   \item CutFEM support (see Section~\ref{sec:cut});
-  \item Integration of the Computational Geometry Algorithms Library (CGAL) (see Section~\ref{sec:cgalwrappers});
+  \item Experimental integration of the Computational Geometry Algorithms Library (CGAL) (see Section~\ref{sec:cgalwrappers});
   \item Performance improvements in the particle infrastructure (see Section~\ref{sec:particles});
@@
 triangulations and cells. These wrappers are enabled only if \dealii{}
-is compiled with \texttt{C++17} and, for the first time in \dealii{}'s
-25-year history, provides a built-in ability to generate meshes for
-arbitrary geometries.
+is compiled with \texttt{C++17}. \textit{Note: This feature is still experimental and
+interfaces might change during the next release cycle.}
 

### 2022-06-23 9.4/paper.tex
Swap immersed and background mesh

@@
 
-These utility functions will be the building blocks for functions in the \texttt{NonMatching} namespace that will, e.g.,  assemble coupling terms like $(u,v)_{B}$,\todo{The $B$ and $\Omega$ here need to be streamlined with \eqref{eq:boundingbox}, right now they are the opposite} with $B$ a domain immersed in a fixed background mesh $\Omega$ and $u,v$ finite element functions on $V_h(\Omega)$, as needed, e.g., in
+These utility functions will be the building blocks for functions in the \texttt{NonMatching} namespace that will, e.g.,  assemble coupling terms like $(u,v)_{\Omega}$, with $\Omega$ a domain immersed in a fixed background mesh $B$ and $u,v$ finite element functions on $V_h(B)$, as needed, e.g., in
 the context of CutFEM (see Section~\ref{sec:cut}) or Nitsche's method to weakly impose boundary conditions at an interface.
-The same applies to coupling terms of the form $(u,q)_B$ in formulations using Lagrange multipliers, where now $q \in Q_h(B)$, with $Q_h(B)$ the space of the multiplier variable.
+The same applies to coupling terms of the form $(u,q)_{\Omega}$ in formulations using Lagrange multipliers, where now $q \in Q_h(\Omega)$, with $Q_h(\Omega)$ the space of the multiplier variable.
 Note that the most relevant difference between this and the \texttt{QuadratureGenerator} in Section~\ref{sec:cut} is that the \texttt{Quadrature} objects are created directly from two overlapping grids, one

### 2022-06-24 9.4/paper.tex
add short announcement and mention step-81 in paper

@@
   \item Improvements to unstructured communication (see Section~\ref{sec:CA});
-  \item Two new tutorial programs and one new code gallery program (see
-    Section~\ref{subsec:steps}).
+  \item Three new tutorial programs and one new code gallery program (see Section~\ref{subsec:steps}).
 \end{itemize}
@@
 
-Many of the \dealii{} tutorial programs were revised in a variety of
-ways as part of this release. In addition, there are a number of new tutorial programs:
+Many of the \dealii{} tutorial programs were revised in a variety of ways
+as part of this release. In addition, there are a number of new tutorial
+programs:
 \begin{itemize}
-\item \texttt{step-82} was contributed by Andrea Bonito (Texas A\&M University) and Diane Guignard (University of Ottawa). It shows how
-\dealii{} can be used to implement the
-local discontinuous Galerkin (LDG) method for approximating the
-solution to the bi-Laplacian problem. The method is an alternative to
-the C0IP method used in \texttt{step-47}.
-
-\item \texttt{step-85} was contributed by Simon Sticko (Uppsala University). It shows how to use the new CutFEM infrastructure to solve a Poisson
-problem on a circular domain that is embedded in a Cartesian background mesh.
+  \item
+    \texttt{step-81} was contributed by Manaswinee Bezbaruah (Texas A\&M
+    University) and Matthias Maier (Texas A\&M University). It explains how
+    to solve the complex-valued time-harmonic Maxwell's equations for an
+    optical scattering problem.
+  \item
+    \texttt{step-82} was contributed by Andrea Bonito (Texas A\&M
+    University) and Diane Guignard (University of Ottawa). It shows how
+    \dealii{} can be used to implement the local discontinuous Galerkin
+    (LDG) method for approximating the solution to the bi-Laplacian
+    problem. The method is an alternative to the C0IP method used in
+    \texttt{step-47}.
+  \item
+    \texttt{step-85} was contributed by Simon Sticko (Uppsala University).
+    It shows how to use the new CutFEM infrastructure to solve a Poisson
+    problem on a circular domain that is embedded in a Cartesian background
+    mesh.
 \end{itemize}

### 2022-06-24 9.4/paper.tex
Tidy specialwords.

@@
 partitioned based on the space-filling Morton (or Z-order) curve
-implemented by the \pfrst{} backend. The latter, more recent
+implemented by the \pfrst backend. The latter, more recent
 class is currently statically partitioned at creation time.
@@
 In \cite{munch2022gc}, the performance of the local-smoothing and global-coarsening
-infrastructure of \dealii{} was compared for locally refined meshes. The results indicate that
+infrastructure of \dealii was compared for locally refined meshes. The results indicate that
 the local definition of multigrid levels might introduce load imbalances
@@
 potentially more expensive intergrid transfers. In order to judge the benefits
-of one approach against the other, \dealii{} provides new functions
+of one approach against the other, \dealii provides new functions
 \texttt{workload\_imbalance()} and \texttt{vertical\_communication\_efficiency()}
@@
 The Computational Geometry Algorithms Library (CGAL, \url{https://www.cgal.org/}) is a widely used
-library to describe geometries and meshes \cite{cgal}. \dealii{} now
+library to describe geometries and meshes \cite{cgal}. \dealii now
 has wrappers for CGAL classes and functions, provided in the new
 namespace \texttt{CGALWrappers}, and implementing functionality
-spanning from mesh generation to boolean operations between \dealii{}
-triangulations and cells. These wrappers are enabled only if \dealii{}
+spanning from mesh generation to boolean operations between \dealii
+triangulations and cells. These wrappers are enabled only if \dealii
 is compiled with \texttt{C++17}. \textit{Note: This feature is still experimental and
@@
 
-\dealii{} has an implementation of these algorithms for some time,
+\dealii has an implementation of these algorithms for some time,
 but the current release substantially expands on it. Specifically, the
@@
 
-Many of the \dealii{} tutorial programs were revised in a variety of ways
+Many of the \dealii tutorial programs were revised in a variety of ways
 as part of this release. In addition, there are a number of new tutorial
@@
     University) and Diane Guignard (University of Ottawa). It shows how
-    \dealii{} can be used to implement the local discontinuous Galerkin
+    \dealii can be used to implement the local discontinuous Galerkin
     (LDG) method for approximating the solution to the bi-Laplacian
@@
   how to solve the incompressible Navier--Stokes equations efficiently 
-  with \dealii{}'s matrix-free DG infrastructure, multigrid, and adaptive-mesh
+  with \dealii's matrix-free DG infrastructure, multigrid, and adaptive-mesh
   refinement. Interested readers are referred to~\cite{orlando2021efficient}.
@@
 
-In order to justify the work the developers of \dealii{} put into this
+In order to justify the work the developers of \dealii put into this
 software, we ask that papers using the library reference one of the
-\dealii{} papers. This helps us justify the effort we put into this library.
-
-There are various ways to reference \dealii{}. To acknowledge the use of
+\dealii papers. This helps us justify the effort we put into this library.
+
+There are various ways to reference \dealii. To acknowledge the use of
 the current version of the library, \textbf{please reference the present
@@
 
-\dealii{} can interface with many other libraries:
+\dealii can interface with many other libraries:
 \begin{multicols}{3}

### 2022-06-24 9.4/paper.tex
Added boost.

@@
 \usepackage{multirow}
+
+\usepackage{booktabs}
 
@@
 
-\dealii version 9.4.0 was released June 10, 2022.
-\todo{Update release date.}
+\dealii version 9.4.0 was released June 24, 2022.
 This paper provides an
@@
   \item Performance improvements in the particle infrastructure (see Section~\ref{sec:particles});
+  \item Support for large MPI buffers and parallel I/O (see Section~\ref{sec:large-mpi});
   \item Improvements to unstructured communication (see Section~\ref{sec:CA});
@@
   enable efficient construction of the
-  communication pattern used in the class \texttt{Utilties::\allowbreak MPI::\allowbreak RemotePointEvaluation}. Furthermore, this function could be used in the future to allow
+  communication pattern used in the class \texttt{Utilities::\allowbreak MPI::\allowbreak RemotePointEvaluation}. Furthermore, this function could be used in the future to allow
   particle simulations in which particle movement is not
@@
   data in this format.
-  \item A new module \texttt{Utilities::MPI::LargeCount} which enables sending and receiving MPI messages containing more than $2^{31}$ objects.
-  This library either uses the new MPI-4 functions, such as \texttt{MPI\_Send\_c()}, or an internal implementation of large objects for
-  compatibility with MPI-3.
   \item The \texttt{FEInterfaceValues} class, which computes common quantities at the interface of two cells, has been overhauled to make it
@@
   functions, such as \texttt{FEInterfaceValues::get\_jump\_in\_function\_values()}, have also been added.
+  \item The\texttt{MeshWorker::ScratchData} and \texttt{MeshWorker::CopyData} have been made $hp$-compatible,
+  and supports face integration where the integration rule and mapping differs on either side of an interface.
+  The \texttt{MeshWorker::CopyData} class has also been made compatible with complex numbers.
   \item Vectors attached to \texttt{DataOut} do not need to be in ghosted state anymore. Internally, we create a copy of the vector with appropriate ghosting.
@@
         returns an enum, \texttt{LocationToLevelSet}, with values \{\texttt{inside}, \texttt{outside}, \texttt{intersected}\}.
-        This information is typically needed when choosing what element a cell of the \texttt{DoFHandler} should use.
+        This information is typically needed when choosing what element (e.g., \texttt{FE\_Q} or \texttt{FE\_Nothing}) and/or what quadrature (defined over the 
+        complete cell or over a part---see below) a cell
+        should use.
 
@@
 has wrappers for CGAL classes and functions, provided in the new
-namespace \texttt{CGALWrappers}, and implementing functionality
-spanning from mesh generation to boolean operations between \dealii
+namespace \texttt{CGALWrappers}: they implement functionality
+spanning from mesh generation to boolean operations between
 triangulations and cells. These wrappers are enabled only if \dealii
@@
 \texttt{GridGenerator::surface\_mesh\_to\_volumetric\_mesh()}, which
-computes a tetrahedral volume triangulation \texttt{Triangulation<3>}, based on a
-given surface triangulation \texttt{Triangulation<2,3>} that bounds
+computes a tetrahedral volume mesh \texttt{Triangulation<3>}, based on a
+given surface mesh \texttt{Triangulation<2,3>} that bounds
 the three dimensional shape.
 
-CGAL also provides Boolean operations on meshes, as available in the
-utility function
+CGAL also provides Boolean operations on meshes: they can be accessed via 
+the utility function
 \texttt{CGALWrappers::\allowbreak{}compute\_boolean\_operation()}. The
 available operations are \textit{co-refinement}, \textit{intersection},
-\textit{union}, and \textit{difference}. Oftentimes, boolean
+\textit{union}, and \textit{difference}. Oftentimes, Boolean
 operations and co-refinement around the intersection produces
-badly shaped mesh cells. To overcome this issue, one can use \texttt{CGALWrappers::\allowbreak{}remesh\_surface()}. Figure~\ref{fig:corefinement_remeshed} shows a graphical example.
-A possible workflow is the following:
+badly shaped mesh cells. To overcome this issue, one can use \texttt{CGALWrappers::\allowbreak{}remesh\_surface()}.
+A possible workflow to create a good-quality mesh of the union of a cube and
+a sphere mesh is given in the following listing:
 \begin{c++}
@@
 \end{c++}
-The output of the boolean operation can be seen in Fig.~\ref{fig:corefinement}, while in Fig.~\ref{fig:corefinement_remeshed}
-the same mesh has been remeshed.
+The output of the Boolean operation is shown in Fig.~\ref{fig:corefinement};
+Fig.~\ref{fig:corefinement_remeshed} shows the mesh after remeshing.
 \begin{figure}
@@
 
-\texttt{CGALWrappers::compute\_quadrature\_on\_boolean\_operation()} returns a \texttt{Quadrature<3>} that allows exact integration of polynomials on polyhedral elements coming out of a \texttt{BooleanOperation} between \dealii cells.
-The quadrature rule is built by meshing the polyhedral region with tetrahedra, computing on each tetrahedron a \texttt{QGaussSimplex<3>} quadrature rule by using \texttt{QSimplex<3>::\allowbreak{}compute\_affine\_transformation()}, and finally
-collecting all of the rules together, giving a \texttt{Quadrature<3>} formula on the \emph{physical} element.
-
-These utility functions will be the building blocks for functions in the \texttt{NonMatching} namespace that will, e.g.,  assemble coupling terms like $(u,v)_{\Omega}$, with $\Omega$ a domain immersed in a fixed background mesh $B$ and $u,v$ finite element functions on $V_h(B)$, as needed, e.g., in
-the context of CutFEM (see Section~\ref{sec:cut}) or Nitsche's method to weakly impose boundary conditions at an interface.
+The function \texttt{CGALWrappers::compute\_quadrature\_on\_boolean\_operation()} returns a quadrature that allows exact integration of polynomials on polyhedral elements created by a \texttt{Boolean\allowbreak Operation} between \dealii cells.
+The quadrature rule is built by 1) subdividing the polyhedral region in tetrahedra, 2) computing on each a \texttt{QGaussSimplex<3>} quadrature rule via \texttt{QSimplex<3>::\allowbreak{}compute\_affine\_transformation()}, and 3)
+collecting all in a single \texttt{Quadrature<3>} object on the \emph{physical} cell.
+
+The new utility functions will be the building blocks for functions in 
+the \texttt{NonMatching} namespace in future releases. While our current 
+CutFEM implementation (see Section~\ref{sec:cut}) relies on a level set description of the domain to determine
+the boundary and the quadrature rules, one could do that also based on non-matching overlapping
+grids~\cite{massing2013efficient}, for which the newly introduced, CGAL-based Boolean operations and 
+quadrature-generation functionality might be helpful. We would like to point out
+that the functionalities are not limited to CutFEM but are also applicable
+for weakly imposing boundary conditions on the interface of an immersed boundary
+using Nitsche's method and coupling terms using Lagrange multipliers.
+
+%These utility functions will be the building blocks for functions in the \texttt{NonMatching} namespace that will, e.g.,  assemble coupling terms like $(u,v)_{\Omega}$, with $\Omega$ a domain immersed in a fixed background mesh $B$ and $u,v$ finite element functions on $V_h(B)$, as needed, e.g., in
+%the context of CutFEM (see Section~\ref{sec:cut}) or Nitsche's method to weakly impose boundary conditions at an interface.
 The same applies to coupling terms of the form $(u,q)_{\Omega}$ in formulations using Lagrange multipliers, where now $q \in Q_h(\Omega)$, with $Q_h(\Omega)$ the space of the multiplier variable.
-Note that the most relevant difference between this and the \texttt{QuadratureGenerator} in Section~\ref{sec:cut} is that the \texttt{Quadrature} objects are created directly from two overlapping grids, one
-spanning over $B$ and the other one over $\Omega$, and not from a level set function.
+%Note that the most relevant difference between this and the \texttt{QuadratureGenerator} in Section~\ref{sec:cut} is that the \texttt{Quadrature} objects are created directly from two overlapping grids, one
+%spanning over $B$ and the other one over $\Omega$, and not from a level set function.
 
@@
   \label{tab:particle_timing}
-
   \centering
-  \begin{tabular}{|c|c|c|c|}
-    \hline
-   Particle Operation & \dealii 9.3 & \dealii 9.4 & Speedup \\
-   \hline
+  \begin{tabular}{lccc}
+    \toprule
+    Particle Operation & \dealii 9.3 & \dealii 9.4 & Speedup      \\[0.3em]
     Generation         & 444 ms      & 235 ms      & 1.9$\times$  \\
@@
     Sorting            & 21.9 ms     & 9.27 ms     & 2.4$\times$  \\
-   \hline
+    \bottomrule
   \end{tabular}
@@
 We note that the exact gains will depend strongly on the exact combination of geometry, mapping, dimensionality, and the number of particles per cell in any specific model, and can be smaller or larger than the measurements provided here.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Support for large MPI buffers and parallel I/O}
+\label{sec:large-mpi}
+
+   Sending large messages over MPI or reading and writing large buffers using MPI I/O,
+   especially when dealing with small datatypes like \texttt{MPI\_CHAR},
+   can often require processing more than $2^{31}$ objects at once. Prior
+   to the recent MPI-4 standard, which introduced a new set of \texttt{MPI\_*\_c()} functions, all MPI functions expected the ``count''
+   to be a signed integer. This limits the number of objects to $2^{31}$,
+   i.e., writing of up to 2 GB at once if the datatype is \texttt{MPI\_CHAR}.
+
+   Several places in the library could run into the issue mentioned above,
+   for example when producing large graphical output (per rank), large
+   checkpointing (per rank), or when broadcasting large datasets for lookup tables.
+
+   With this release we introduce a new module \texttt{Utilities::MPI::LargeCount} which enables sending and receiving MPI messages and I/O containing more than $2^{31}$ objects by implementing
+   the before-mentioned \texttt{MPI\_*\_c()} functions, i.e., \texttt{MPI\_Bcast\_c()}, using MPI-3 features if necessary. The
+   solution is based on custom datatypes as described in~\ref{hammond2014int_max}.
+
+   This functionality is now used in all places in the library, where
+   large counts might be necessary (\texttt{DataOut::write\_vtu\_in\_parallel()} for graphical output, checkpointing of \texttt{Triangulation} objects, etc.).
+   %
+   While implementing these changes we also worked on the following:
+   \begin{itemize}
+    \item Testing of large I/O with large chunks per MPI rank (>2 GB) and
+    large total sizes (4 GB +). Several instances of 32 bit datatypes for
+    offsets were changed to 64 bit to correctly support files larger
+    than 4 GB (HDF5 output, VTU output, checkpointing).
+    \item Performance testing of MPI I/O routines used for parallel VTU output with large performance improvements by switching from a shared
+    file pointer to individual file pointers, see Table~\ref{tab:mpi-io}.
+   \end{itemize}
+
+\begin{table}
+  \caption{\it Performance comparison of writing large VTU graphical output using MPI I/O on
+  TACC Frontera /scratch1/ with 16 file servers (``OSTs'') and a theoretical peak performance of 50 GB/s.
+The ROMIO version used enforces sequential writes when using \texttt{MPI\_File\_write\_ordered()}.}
+  \label{tab:mpi-io}
+
+  \centering
+
+  \begin{tabular}{llcc}
+    \toprule
+    Version       & MPI routine used                     & 1 file, striping 16 & 16 files, no striping \\[0.3em]
+    \dealii{} 9.3 & \texttt{MPI\_File\_write\_ordered()} & 3 GB/s              & 15 GB/s               \\
+    \dealii{} 9.4 & \texttt{MPI\_File\_write\_at\_all()} & 17 GB/s             & 21 GB/s               \\
+    \bottomrule
+\end{tabular}
+
+  \end{table}
 
@@
     University) and Matthias Maier (Texas A\&M University). It explains how
-    to solve the complex-valued time-harmonic Maxwell's equations for an
+    to solve the complex-valued time-harmonic Maxwell equations for an
     optical scattering problem.
@@
     \item cuSPARSE \cite{cusparse}
+    \item Boost \cite{boost-web-page}
     \item Gmsh \cite{geuzaine2009gmsh}
-    \item GSL \cite{gsl2016}
-    \item Ginkgo \cite{ginkgo-web-page}
+    \item GSL \cite{Galassi2009,gsl-web-page}
+    \item Ginkgo \cite{Anzt2022,ginkgo-web-page}
     \item HDF5 \cite{hdf5}
     \item METIS \cite{karypis1998fast}
-    \item MUMPS \cite{ADE00,MUMPS:1,MUMPS:2,mumps-web-page}
+    \item MUMPS \cite{ADE00,MUMPS:1,MUMPS:2,MUMPS:3,mumps-web-page}
     \item muparser \cite{muparser-web-page}
     \item OpenCASCADE \cite{opencascade-web-page}
-    \item p4est \cite{p4est}
+    \item p4est \cite{p4est,burstedde2020parallel}
     \item PETSc \cite{petsc-user-ref,petsc-web-page}

### 2022-06-24 9.4/paper.tex
Added links to libraries as notes.
Reduced references in interface section.

Removed duplicate entries for webpages.

@@
 The Computational Geometry Algorithms Library (CGAL, \url{https://www.cgal.org/}) is a widely used
-library to describe geometries and meshes \cite{cgal}. \dealii now
+library to describe geometries and meshes \cite{cgal-user-ref}. \dealii now
 has wrappers for CGAL classes and functions, provided in the new
@@
   \begin{itemize}[leftmargin=4mm]
-    \item ADOL-C \cite{Griewank1996a,adol-c}
+    \item ADOL-C \cite{griewank1996adolc}
     \item ArborX \cite{lebrun2020arborx}
-    \item ARPACK \cite{arpack}
-    \item Assimp \cite{assimp}
-    \item BLAS and LAPACK \cite{lapack}
-    \item CGAL \cite{cgal}
-    \item cuSOLVER \cite{cusolver}
-    \item cuSPARSE \cite{cusparse}
+    \item ARPACK \cite{lehoucq1998arpack}
+    \item Assimp \cite{schulze2021assimp}
+    \item BLAS and LAPACK \cite{anderson1999lapack}
     \item Boost \cite{boost-web-page}
+    \item CGAL \cite{cgal-user-ref}
+    \item cuSOLVER \cite{cusolver-web-page}
+    \item cuSPARSE \cite{cusparse-web-page}
     \item Gmsh \cite{geuzaine2009gmsh}
-    \item GSL \cite{Galassi2009,gsl-web-page}
-    \item Ginkgo \cite{Anzt2022,ginkgo-web-page}
-    \item HDF5 \cite{hdf5}
-    \item METIS \cite{karypis1998fast}
-    \item MUMPS \cite{ADE00,MUMPS:1,MUMPS:2,MUMPS:3,mumps-web-page}
+    \item GSL \cite{galassi2009gsl}
+    \item Ginkgo \cite{anzt2022ginkgo}
+    \item HDF5 \cite{hdf5-web-page}
+    \item METIS \cite{karypis1998metis}
+    \item MUMPS \cite{amestoy2001mumps}
     \item muparser \cite{muparser-web-page}
     \item OpenCASCADE \cite{opencascade-web-page}
-    \item p4est \cite{p4est,burstedde2020parallel}
-    \item PETSc \cite{petsc-user-ref,petsc-web-page}
-    \item ROL \cite{ridzal2014rapid}
-    \item ScaLAPACK \cite{slug}
-    \item SLEPc \cite{Hernandez:2005:SSF}
-    \item SUNDIALS \cite{sundials}
+    \item p4est \cite{burstedde2011p4est,burstedde2020parallel}
+    \item PETSc \cite{petsc-user-ref}
+    \item ROL \cite{ridzal2014rol}
+    \item ScaLAPACK \cite{blackford1997scalapack}
+    \item SLEPc \cite{hernandez2005slepc}
+    \item SUNDIALS \cite{hindmarsh2005sundials}
     \item SymEngine \cite{symengine-web-page}
-    \item TBB \cite{Rei07}
-    \item Trilinos \cite{trilinos,trilinos-web-page}
-    \item UMFPACK \cite{umfpack}
+    \item TBB \cite{reinders2007tbb}
+    \item Trilinos \cite{heroux2005trilinos}
+    \item UMFPACK \cite{davis2004umfpack}
   \end{itemize}

### 2022-06-24 9.4/paper.tex
promote large-MPI to section

@@
 We note that the exact gains will depend strongly on the exact combination of geometry, mapping, dimensionality, and the number of particles per cell in any specific model, and can be smaller or larger than the measurements provided here.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Support for large MPI buffers and parallel I/O}
+\label{sec:large-mpi}
+
+   A new module \texttt{Utilities::MPI::LargeCount} which enables sending and receiving MPI messages containing more than $2^{31}$ objects.
+  This library either uses the new MPI-4 functions, such as \texttt{MPI\_Send\_c()}, or an internal implementation of large objects for
+  compatibility with MPI-3.
 

### 2022-06-24 9.4/paper.tex
draft for sec 2.7 large MPI

@@
 
-   A new module \texttt{Utilities::MPI::LargeCount} which enables sending and receiving MPI messages containing more than $2^{31}$ objects.
-  This library either uses the new MPI-4 functions, such as \texttt{MPI\_Send\_c()}, or an internal implementation of large objects for
-  compatibility with MPI-3.
+   Sending large messages over MPI or reading and writing large buffers using MPI I/O,
+   especially when dealing with small datatypes like \texttt{MPI\_CHAR},
+   can often require processing more than $2^{31}$ objects at once. Prior
+   to the recent MPI-4 standard, which introduced a new set of \texttt{MPI\_*\_c()} functions, all MPI functions expected the ``count''
+   to be a signed integer. This limits the number of objects to $2^{31}$,
+   i.e., writing of up to 2 GB at once if the datatype is \texttt{MPI\_CHAR}.
+
+   Several places in the library could run into the issue mentioned above,
+   for example when producing large graphical output (per rank), large
+   checkpointing (per rank), or when broadcasting large datasets for lookup tables.
+
+   With this release we introduce a new module \texttt{Utilities::MPI::LargeCount} which enables sending and receiving MPI messages and I/O containing more than $2^{31}$ objects by implementing
+   the before-mentioned \texttt{MPI\_*\_c()} functions, i.e., \texttt{MPI\_Bcast\_c()}, using MPI-3 features if necessary. The
+   solution is based on custom datatypes as described in~\ref{hammond2014int_max}.
+
+   This functionality is now used in all places in the library, where
+   large counts might be necessary (\texttt{DataOut::write\_vtu\_in\_parallel()} for graphical output, checkpointing of \texttt{Triangulation} objects, etc.).
+   %
+   While implementing these changes we also worked on the following:
+   \begin{itemize}
+    \item Testing of large I/O with large chunks per MPI rank (>2 GB) and
+    large total sizes (4 GB +). Several instances of 32 bit datatypes for
+    offsets were changed to 64 bit to correctly support files larger
+    than 4 GB (HDF5 output, VTU output, checkpointing).
+    \item Performance testing of MPI I/O routines used for parallel VTU output with large performance improvements by switching from a shared
+    file pointer to individual file pointers, see Table~\ref{tab:mpi-io}.
+   \end{itemize}
+
+\begin{table}
+  \caption{\it Performance comparison of writing large VTU graphical output using MPI I/O on
+  Frontera /scratch1/ with 16 file servers (``OSTs'') and a theoretical peak performance of 50 GB/s. ROMIO seems to enforce sequential writes when using \texttt{MPI\_File\_write\_ordered()}.}
+  \label{tab:mpi-io}
+
+  \centering
+
+  \begin{tabular}{|c|c|c|c|}
+  \hline
+Version       & MPI routine used            & 1 file, striping 16 & 16 files, no striping \\
+\hline
+old VTU code  & \texttt{MPI\_File\_write\_ordered()} & 3 GB/s              & 15 GB/s              \\
+new in 9.4 & \texttt{MPI\_File\_write\_at\_all()} & 17 GB/s             & 21 GB/s
+\\
+\hline
+\end{tabular}
+
+  \end{table}
 

### 2022-06-24 9.4/paper.tex
update table

@@
   \caption{\it Performance comparison of writing large VTU graphical output using MPI I/O on
-  Frontera /scratch1/ with 16 file servers (``OSTs'') and a theoretical peak performance of 50 GB/s. ROMIO seems to enforce sequential writes when using \texttt{MPI\_File\_write\_ordered()}.}
+  TACC Frontera /scratch1/ with 16 file servers (``OSTs'') and a theoretical peak performance of 50 GB/s.
+The ROMIO version used enforces sequential writes when using \texttt{MPI\_File\_write\_ordered()}.}
   \label{tab:mpi-io}
@@
 \hline
-old VTU code  & \texttt{MPI\_File\_write\_ordered()} & 3 GB/s              & 15 GB/s              \\
-new in 9.4 & \texttt{MPI\_File\_write\_at\_all()} & 17 GB/s             & 21 GB/s
+\dealii{} 9.3 & \texttt{MPI\_File\_write\_ordered()} & 3 GB/s              & 15 GB/s              \\
+\dealii{} 9.4 & \texttt{MPI\_File\_write\_at\_all()} & 17 GB/s             & 21 GB/s
 \\

### 2022-06-24 9.4/paper.tex
fix typo
improve readability of tables

@@
 
-  \begin{tabular}{|c|c|c|c|}
-  \hline
-Version       & MPI routine used            & 1 file, striping 16 & 16 files, no striping \\
-\hline
+  \begin{tabular}{llcc}
+    \toprule
+    Version       & MPI routine used                     & 1 file, striping 16 & 16 files, no striping \\[0.3em]
     \dealii{} 9.3 & \texttt{MPI\_File\_write\_ordered()} & 3 GB/s              & 15 GB/s               \\
-\dealii{} 9.4 & \texttt{MPI\_File\_write\_at\_all()} & 17 GB/s             & 21 GB/s
-\\
-\hline
+    \dealii{} 9.4 & \texttt{MPI\_File\_write\_at\_all()} & 17 GB/s             & 21 GB/s               \\
+    \bottomrule
 \end{tabular}

### 2022-06-26 9.4/paper.tex
Small improvements to text

--- 2022-06-25 9.4/paper.tex
@@
   functions, such as \texttt{FEInterfaceValues::get\_jump\_in\_function\_values()}, have also been added.
+  \item The\texttt{MeshWorker::ScratchData} and \texttt{MeshWorker::CopyData} have been made $hp$-compatible,
+  and support face integration where the integration rule and mapping differs on either side of an interface.
+  The \texttt{MeshWorker::CopyData} class has also been made compatible with complex numbers.
   \item Vectors attached to \texttt{DataOut} do not need to be in ghosted state anymore. Internally, we create a copy of the vector with appropriate ghosting.
@@
 the fly in order to enable vectorization over cells. The new feature is useful in several
-contexts. Examples are simulations with sharp interfaces (e.g., two-phase flow
+contexts, for example for sharp interfaces (e.g., two-phase flow
 or shock capturing), where one needs to treat cells that are ``cut'' by
 the interface in a special way. A challenge is that cell batches
-might contain cut or non-cut cells, making the vectorization of these operations potentially more complicated.
+might contain cut or non-cut cells, effectively counteracting the cross-cell vectorization.
 Previous functionality has provided the option of masking certain cells in cell batches, which works well if
-the code paths do not diverge too much. Another way is to
+the code paths do not diverge too much, or to
 categorize cells during \texttt{MatrixFree::reinit()} in such a way that mixed cell batches do not occur. However, \texttt{MatrixFree::reinit()}
 might be too expensive if recategorization needs to happen very frequently to follow the dynamics of a system, e.g., in each time step.
-Despite some overhead compared to static matrix-free loops, the new feature can be the best option in these scenarios.
+Despite some overhead compared to static matrix-free loops, the new feature can be the best option in such dynamic scenarios.
 \item Initial support for H(div)-conforming elements with Piola transform, based on Raviart--Thomas finite element spaces with the updated class \texttt{FE\_RaviartThomasNodal}, has been added.
@@
 grids~\cite{massing2013efficient}, for which the newly introduced, CGAL-based Boolean operations and 
-quadrature-generation functionality might be helpful. We would like to point out
-that the functionalities are not limited to CutFEM but are also applicable
+quadrature-generation functionality might be helpful.
+The functionalities are applicable in a broad context of immersed problems, such as
 for weakly imposing boundary conditions on the interface of an immersed boundary
-using Nitsche's method and coupling terms using Lagrange multipliers.
+using Nitsche's method and for coupling terms using Lagrange multipliers.
 
@@
 \texttt{Utilities::MPI::Noncontiguous\allowbreak Partitioner}, \texttt{internal::MatrixFreeFunctions::VectorDataExchange}; see~\cite{dealII91}),
-to setup the global-coarsening transfer operators (see~\cite{dealII92} and
+to set up the global-coarsening transfer operators (see~\cite{dealII92} and
 Section~\ref{sec:multigrid}), to repartition distributed meshes (see Section~\ref{sec:repartitioning}), and basis coupling algorithms between non-matching
@@
     \item Gmsh \cite{geuzaine2009gmsh}
-    \item GSL \cite{gsl2016}
-    \item Ginkgo \cite{ginkgo-web-page}
+    \item GSL \cite{Galassi2009,gsl-web-page}
+    \item Ginkgo \cite{Anzt2022,ginkgo-web-page}
     \item HDF5 \cite{hdf5}
     \item METIS \cite{karypis1998fast}
-    \item MUMPS \cite{ADE00,MUMPS:1,MUMPS:2,mumps-web-page}
+    \item MUMPS \cite{ADE00,MUMPS:1,MUMPS:2,MUMPS:3,mumps-web-page}
     \item muparser \cite{muparser-web-page}
     \item OpenCASCADE \cite{opencascade-web-page}
-    \item p4est \cite{p4est}
+    \item p4est \cite{p4est,burstedde2020parallel}
     \item PETSc \cite{petsc-user-ref,petsc-web-page}

### 2022-06-27 9.4/paper.tex
Add word

@@
 quadrature-generation functionality might be helpful.
-The functionalities are applicable in a broad context of immersed problems, such as
+The functionalities are also applicable in a broad context of immersed problems, such as
 for weakly imposing boundary conditions on the interface of an immersed boundary

### 2022-06-27 9.4/paper.tex
Use midrule

@@
     \toprule
-    Particle Operation & \dealii 9.3 & \dealii 9.4 & Speedup      \\[0.3em]
+    Particle Operation & \dealii 9.3 & \dealii 9.4 & Speedup      \\ \midrule
     Generation         & 444 ms      & 235 ms      & 1.9$\times$  \\
@@
     \toprule
-    Version       & MPI routine used                     & 1 file, striping 16 & 16 files, no striping \\[0.3em]
+    Version       & MPI routine used                     & 1 file, striping 16 & 16 files, no striping \\\midrule
     \dealii{} 9.3 & \texttt{MPI\_File\_write\_ordered()} & 3 GB/s              & 15 GB/s               \\

### 2022-06-27 9.4/paper.tex
Fix some minor language mistakes

@@
   unstructured mesh on a structured one (which might facilitate a more
-  memory-efficient storage format, for example if this second
+  memory-efficient storage format, for example, if this second
   triangulation is a uniformly refined rectangle or box), or one can create a slice in 3D.
@@
   shape functions by their support point. This functionality is useful in both developing nodal schemes since, e.g., the $x$, $y$, and $z$
-  velocities at a point will be consecutive in the solution vector. It also improves interoperability with external libraries which expect
+  components at a point will be consecutive in the solution vector. It also improves interoperability with external libraries which expect
   data in this format.
@@
   functions, such as \texttt{FEInterfaceValues::get\_jump\_in\_function\_values()}, have also been added.
-  \item The\texttt{MeshWorker::ScratchData} and \texttt{MeshWorker::CopyData} have been made $hp$-compatible,
+  \item The \texttt{MeshWorker::ScratchData} and \texttt{MeshWorker::CopyData} have been made $hp$-compatible,
   and support face integration where the integration rule and mapping differs on either side of an interface.
@@
 \texttt{GeometryInfo} class (which is specific to hypercube cells) from
-the library, and to replace them by more general equivalent
+the library, and to replace them with more general equivalent
 functionality based on the
@@
 sent to the new owner. On the
-receiving site, the sets of all cells are combined and possible duplicates
+receiving side, the sets of all cells are combined and possible duplicates
 are removed. This information is enough to construct a new triangulation.
@@
 We take care to reuse allocated memory to minimize the number of memory reallocations.
-\item The separate cache structure that contains entries for each cell allows quick random-access to the particles of a particular cell, and also allows to quickly determine if a particular cell has particles at all.
+\item The separate cache structure that contains entries for each cell allows quick random access to the particles of a particular cell, and also allows to quickly determine if a particular cell has particles at all.
 \end{itemize}
@@
 \item Sorting, i.e., the inversion of the mapping of each cell to find the new particle locations relative to this cell, and moving all particles that have left their original cell into new cells (both \dealii 9.3 and \dealii 9.4).
-In \dealii 9.4 this operation also includes reordering the particle properties for optimal iteration.
+In \dealii 9.4, this operation also includes reordering the particle properties for optimal iteration.
 \end{itemize}
@@
    With this release we introduce a new module \texttt{Utilities::MPI::LargeCount} which enables sending and receiving MPI messages and I/O containing more than $2^{31}$ objects by implementing
-   the before-mentioned \texttt{MPI\_*\_c()} functions, i.e., \texttt{MPI\_Bcast\_c()}, using MPI-3 features if necessary. The
+   the before-mentioned \texttt{MPI\_*\_c()} functions, e.g., \texttt{MPI\_Bcast\_c()}, using MPI-3 features if necessary. The
    solution is based on custom datatypes as described in~\ref{hammond2014int_max}.
@@
     \item Testing of large I/O with large chunks per MPI rank (>2 GB) and
-    large total sizes (4 GB +). Several instances of 32 bit datatypes for
-    offsets were changed to 64 bit to correctly support files larger
+    large total sizes (4 GB +). Several instances of 32-bit data types for
+    offsets were changed to 64-bit to correctly support files larger
     than 4 GB (HDF5 output, VTU output, checkpointing).
@@
 
-\dealii has an implementation of these algorithms for some time,
+\dealii has had an implementation of these algorithms for some time,
 but the current release substantially expands on it. Specifically, the
@@
 
-\dealii is a world-wide project with dozens of contributors around the
+\dealii is a worldwide project with dozens of contributors around the
 globe. Other than the authors of this paper, the following people
@@
 R.~Gassm{\"o}ller was also partially supported by the NSF Awards
-EAR-1925677, and EAR-2054605.
+EAR-1925677 and EAR-2054605.
 

### 2022-06-27 9.4/paper.tex
Added references as requested by third party libraries. Removed notes with url.

--- 2022-06-27 9.4/paper.tex
@@
   \item The\texttt{MeshWorker::ScratchData} and \texttt{MeshWorker::CopyData} have been made $hp$-compatible,
-  and support face integration where the integration rule and mapping differs on either side of an interface.
+  and supports face integration where the integration rule and mapping differs on either side of an interface.
   The \texttt{MeshWorker::CopyData} class has also been made compatible with complex numbers.
@@
 the fly in order to enable vectorization over cells. The new feature is useful in several
-contexts, for example for sharp interfaces (e.g., two-phase flow
+contexts. Examples are simulations with sharp interfaces (e.g., two-phase flow
 or shock capturing), where one needs to treat cells that are ``cut'' by
 the interface in a special way. A challenge is that cell batches
-might contain cut or non-cut cells, effectively counteracting the cross-cell vectorization.
+might contain cut or non-cut cells, making the vectorization of these operations potentially more complicated.
 Previous functionality has provided the option of masking certain cells in cell batches, which works well if
-the code paths do not diverge too much, or to
+the code paths do not diverge too much. Another way is to
 categorize cells during \texttt{MatrixFree::reinit()} in such a way that mixed cell batches do not occur. However, \texttt{MatrixFree::reinit()}
 might be too expensive if recategorization needs to happen very frequently to follow the dynamics of a system, e.g., in each time step.
-Despite some overhead compared to static matrix-free loops, the new feature can be the best option in such dynamic scenarios.
+Despite some overhead compared to static matrix-free loops, the new feature can be the best option in these scenarios.
 \item Initial support for H(div)-conforming elements with Piola transform, based on Raviart--Thomas finite element spaces with the updated class \texttt{FE\_RaviartThomasNodal}, has been added.
@@
 grids~\cite{massing2013efficient}, for which the newly introduced, CGAL-based Boolean operations and 
-quadrature-generation functionality might be helpful.
-The functionalities are also applicable in a broad context of immersed problems, such as
+quadrature-generation functionality might be helpful. We would like to point out
+that the functionalities are not limited to CutFEM but are also applicable
 for weakly imposing boundary conditions on the interface of an immersed boundary
-using Nitsche's method and for coupling terms using Lagrange multipliers.
+using Nitsche's method and coupling terms using Lagrange multipliers.
 
@@
 \texttt{Utilities::MPI::Noncontiguous\allowbreak Partitioner}, \texttt{internal::MatrixFreeFunctions::VectorDataExchange}; see~\cite{dealII91}),
-to set up the global-coarsening transfer operators (see~\cite{dealII92} and
+to setup the global-coarsening transfer operators (see~\cite{dealII92} and
 Section~\ref{sec:multigrid}), to repartition distributed meshes (see Section~\ref{sec:repartitioning}), and basis coupling algorithms between non-matching
@@
   \begin{itemize}[leftmargin=4mm]
-    \item ADOL-C \cite{Griewank1996a,adol-c}
+    \item ADOL-C \cite{griewank1996adolc}
     \item ArborX \cite{lebrun2020arborx}
-    \item ARPACK \cite{arpack}
-    \item Assimp \cite{assimp}
-    \item BLAS and LAPACK \cite{lapack}
-    \item CGAL \cite{cgal}
-    \item cuSOLVER \cite{cusolver}
-    \item cuSPARSE \cite{cusparse}
+    \item ARPACK \cite{lehoucq1998arpack}
+    \item Assimp \cite{schulze2021assimp}
+    \item BLAS and LAPACK \cite{anderson1999lapack}
+    \item Boost \cite{boost-web-page}
+    \item CGAL \cite{cgal-user-ref}
+    \item cuSOLVER \cite{cusolver-web-page}
+    \item cuSPARSE \cite{cusparse-web-page}
     \item Gmsh \cite{geuzaine2009gmsh}
-    \item GSL \cite{Galassi2009,gsl-web-page}
-    \item Ginkgo \cite{Anzt2022,ginkgo-web-page}
-    \item HDF5 \cite{hdf5}
-    \item METIS \cite{karypis1998fast}
-    \item MUMPS \cite{ADE00,MUMPS:1,MUMPS:2,MUMPS:3,mumps-web-page}
+    \item GSL \cite{galassi2009gsl,gsl-web-page}
+    \item Ginkgo \cite{anzt2020ginkgo,anzt2022ginkgo}
+    \item HDF5 \cite{hdf5-web-page}
+    \item METIS \cite{karypis1998metis}
+    \item MUMPS \cite{amestoy2001mumps,amestoy2019mumps}
     \item muparser \cite{muparser-web-page}
     \item OpenCASCADE \cite{opencascade-web-page}
-    \item p4est \cite{p4est,burstedde2020parallel}
+    \item p4est \cite{burstedde2011p4est,burstedde2020parallel}
     \item PETSc \cite{petsc-user-ref,petsc-web-page}
-    \item ROL \cite{ridzal2014rapid}
-    \item ScaLAPACK \cite{slug}
-    \item SLEPc \cite{Hernandez:2005:SSF}
-    \item SUNDIALS \cite{sundials}
+    \item ROL \cite{ridzal2014rol}
+    \item ScaLAPACK \cite{blackford1997scalapack}
+    \item SLEPc \cite{hernandez2005slepc}
+    \item SUNDIALS \cite{hindmarsh2005sundials}
     \item SymEngine \cite{symengine-web-page}
-    \item TBB \cite{Rei07}
-    \item Trilinos \cite{trilinos,trilinos-web-page}
-    \item UMFPACK \cite{umfpack}
+    \item TBB \cite{reinders2007tbb}
+    \item Trilinos \cite{heroux2005trilinos,trilinos-web-page}
+    \item UMFPACK \cite{davis2004umfpack}
   \end{itemize}

### 2022-06-27 9.4/paper.tex
A paper reference should be managed with \cite
Last read.

@@
   \item The \texttt{DataOutResample} class interpolates values defined on one
-  triangulation onto a second potentially unrelated triangulation.
+  triangulation onto a second, potentially unrelated triangulation.
   By using this class, one can output the result obtained on an
@@
   triangulation is a uniformly refined rectangle or box), or one can create a slice in 3D.
+
 \item The new member function \texttt{find\_point\_owner\_rank()} of
-        \texttt{parallel\allowbreak ::distributed::\allowbreak Tri\allowbreak angulation} allows one to find the MPI
+        \texttt{parallel\allowbreak ::distributed::\allowbreak Tri-\allowbreak angulation} allows one to find the MPI
   ranks of cells containing specified points.
@@
   limited by CFL conditions, as done in \cite{mirzadeh2016parallel}.
+
 \item The new function
@@
   shape functions by their support point. This functionality is useful in both developing nodal schemes since, e.g., the $x$, $y$, and $z$
-  components at a point will be consecutive in the solution vector. It also improves interoperability with external libraries which expect
+  components at a point will be consecutive in the solution vector. It
+  also improves interoperability with those external libraries that expect
   data in this format.
@@
   The \texttt{MeshWorker::CopyData} class has also been made compatible with complex numbers.
-  \item Vectors attached to \texttt{DataOut} do not need to be in ghosted state anymore. Internally, we create a copy of the vector with appropriate ghosting.
+  \item Vectors attached to \texttt{DataOut} do not need to be in
+    ghosted state anymore. Internally, we create a copy of the vector
+    with appropriate ghost elements.
 \end{itemize}
@@
 \item \texttt{QIteratedSimplex} allows to build composite simplex quadrature rules.
-\item The new wrappers to the \texttt{CGAL} library allow to simply create simplex
+\item The new wrappers to the \texttt{CGAL} library allow for the easy
+  creation of simplex
 meshes. For more details, see Section~\ref{sec:cgalwrappers}.
@@
 \begin{c++}
-// original triangulation:
+// Original triangulation:
 parallel::distributed::Triangulation<dim> tria(communicator);
 
-// select a partitioning policy and use it:
+// Select a partitioning policy and use it:
 const RepartitioningPolicyTools::CellWeightPolicy<dim> policy(tria, fu);
@@
 
-// create the new triangulation:
+// Create the new triangulation:
 parallel::fullydistributed::Triangulation<dim> tria_pft(comm);
@@
 The matrix-free infrastructure of \dealii is used to solve problems
-without assembling matrices, providing only the action of the matrix
+without assembling matrices, providing only the \textit{action} of the matrix
 instead. It has seen a number of new features in this release, the most
@@
 \begin{itemize}
-\item Improved support for computations with Hessians of shape functions: just like values and gradients, Hessians can be
+\item Improved support for computations with Hessians of shape functions: Just like values and gradients, Hessians can be
 now evaluated and integrated during matrix-free loops
@@
 Finally, we have performed a major restructuring of the internals
-of the \texttt{FEEvaluation} classes. This reduces some overheads for low polynomial degrees and will enable us to add support for new element types in the future.
+of the \texttt{FEEvaluation} classes. This reduces some overhead for low polynomial degrees and will enable us to add support for new element types in the future.
 
@@
 in the case of local smoothing so that global coarsening is favorable despite
-potentially more expensive intergrid transfers. In order to judge the benefits
+potentially more expensive intergrid transfers. In order to assess the benefits
 of one approach against the other, \dealii provides new functions
@@
 \end{equation}
-Thus, when assembling on a cut cell, we are
+Thus, when assembling on a cut cell $K$, we are
 required to integrate over the part of the domain and the part of the boundary, $\Gamma = \partial \Omega$, that falls inside the cell:
@@
 triangulations and cells. These wrappers are enabled only if \dealii
-is compiled with \texttt{C++17}. \textit{Note: This feature is still experimental and
+is compiled with \texttt{C++17}. \textit{This feature is still experimental and
 interfaces might change during the next release cycle.}
@@
 // 1) An implicit function, e.g., Taubin's heart surface (not shown)
-ImplicitFunction implicit_fu;
-
-// 2) configure output mesh (optional)
+ImplicitFunction f;
+
+// 2) Configure output mesh (optional)
 CGALWrappers::AdditionalData<3> data; data.cell_size = .05;
 
-// 3) create mesh
+// 3) Create mesh
 Triangulation<3> tria;
-GridGenerator::implicit_function(tria, implicit_fu, data, {0, 0, 0}, 10.0);
+GridGenerator::implicit_function(tria, f, data, /* more arguments */...);
 \end{c++}
@@
 badly shaped mesh cells. To overcome this issue, one can use \texttt{CGALWrappers::\allowbreak{}remesh\_surface()}.
-A possible workflow to create a good-quality mesh of the union of a cube and
+A possible workflow to create a high-quality mesh of the union of a cube and
 a sphere mesh is given in the following listing:
 \begin{c++}
-// 1) create deal.II triangulations, e.g., cube and sphere (not shown)  
+// 1) Create deal.II triangulations, e.g., cube and sphere (not shown)  
 Triangulation<spacedim> tria0, tria1;
 
-// 2) convert to CGAL surface meshes (assuming Kernel is already defined) 
+// 2) Convert to CGAL surface meshes (assuming Kernel is already defined) 
 CGAL::Surface_mesh<Kernel> surface_mesh0, surface_mesh1;
@@
 
-// 3) compute the union of the two meshes
+// 3) Compute the union of the two meshes
 CGALWrappers::compute_boolean_operation(surface_mesh0, surface_mesh1,
@@
 
-// 4) convert CGAL surface mesh to deal.II surface mesh
+// 4) Convert CGAL surface mesh to deal.II surface mesh
 Triangulation<2, 3> tria_out;
@@
 
-// 5) convert surface to volume mesh via surface_mesh_to_volumetric_mesh()
+// 5) Convert surface to volume mesh via surface_mesh_to_volumetric_mesh()
 \end{c++}
@@
 
-The function \texttt{CGALWrappers::compute\_quadrature\_on\_boolean\_operation()} returns a quadrature that allows exact integration of polynomials on polyhedral elements created by a \texttt{Boolean\allowbreak Operation} between \dealii cells.
-The quadrature rule is built by 1) subdividing the polyhedral region in tetrahedra, 2) computing on each a \texttt{QGaussSimplex<3>} quadrature rule via \texttt{QSimplex<3>::\allowbreak{}compute\_affine\_transformation()}, and 3)
-collecting all in a single \texttt{Quadrature<3>} object on the \emph{physical} cell.
+The function
+\texttt{CGALWrappers::compute\_quadrature\_on\_boolean\_operation()}
+returns a quadrature that allows exact integration of polynomials on
+polyhedral elements created by a Boolean operation between \dealii cells.
+The quadrature rule is built by (i) subdividing the polyhedral region in tetrahedra, (ii) computing on each a \texttt{QGaussSimplex<3>} quadrature rule via \texttt{QSimplex<3>::\allowbreak{}compute\_affine\_transformation()}, and (iii)
+collecting all in a single quadrature object on the \emph{physical} cell.
 
@@
 in \dealii.
-In our previously reported improvements~\citep{dealII93} we stored all particle data as a separate contiguous array for each particle property, but particle identifiers (IDs) were still stored in a multimap tree-like structure.
+In our previously reported implementation~\citep{dealII93} we stored all particle data as a separate contiguous array for each particle property, but particle identifiers (IDs) were still stored in a multimap tree-like structure.
 
@@
 We take care to reuse allocated memory to minimize the number of memory reallocations.
-\item The separate cache structure that contains entries for each cell allows quick random access to the particles of a particular cell, and also allows to quickly determine if a particular cell has particles at all.
+\item The separate cache structure that contains entries for each cell allows quick random access to the particles of a particular cell, and also allows quickly determining if a particular cell has particles at all.
 \end{itemize}
@@
    i.e., writing of up to 2 GB at once if the datatype is \texttt{MPI\_CHAR}.
-
-   Several places in the library could run into the issue mentioned above,
-   for example when producing large graphical output (per rank), large
-   checkpointing (per rank), or when broadcasting large datasets for lookup tables.
+   Several places in the library could exceed this limit,
+   for example when producing large graphical output, large
+   checkpoint files, or when broadcasting large datasets such as lookup tables.
 
@@
    the before-mentioned \texttt{MPI\_*\_c()} functions, e.g., \texttt{MPI\_Bcast\_c()}, using MPI-3 features if necessary. The
-   solution is based on custom datatypes as described in~\ref{hammond2014int_max}.
-
-   This functionality is now used in all places in the library, where
+   solution is based on custom datatypes as described in~\cite{hammond2014int_max}.
+This functionality is now used in all places in the library where
    large counts might be necessary (\texttt{DataOut::write\_vtu\_in\_parallel()} for graphical output, checkpointing of \texttt{Triangulation} objects, etc.).
-   %
+
    While implementing these changes we also worked on the following:
@@
     \item Testing of large I/O with large chunks per MPI rank (>2 GB) and
-    large total sizes (4 GB +). Several instances of 32-bit data types for
+    large total sizes (>4 GB). Several instances of 32-bit data types for
     offsets were changed to 64-bit to correctly support files larger
@@
   \caption{\it Performance comparison of writing large VTU graphical output using MPI I/O on
-  TACC Frontera /scratch1/ with 16 file servers (``OSTs'') and a theoretical peak performance of 50 GB/s.
+  TACC Frontera \texttt{/scratch1} file system with 16 file servers (``OSTs'') and a theoretical peak performance of 50 GB/s.
 The ROMIO version used enforces sequential writes when using \texttt{MPI\_File\_write\_ordered()}.}
@@
 user-contributed programs that often solve more complicated problems
-than tutorial programs, and intended as starting points for further
+than tutorial programs, and that are intended as starting points for further
 research rather than as teaching tools):

### 2022-06-28 9.4/paper.tex
Improve a sentence.

@@
   \item A new DoF renumbering function \texttt{DoFRenumbering::support\_point\_wise()} which groups together
-  shape functions by their support point. This functionality is useful in both developing nodal schemes since, e.g., the $x$, $y$, and $z$
-  components at a point will be consecutive in the solution vector. It
-  also improves interoperability with those external libraries that expect
+  shape functions by their support point. This functionality is useful in developing nodal schemes since, e.g., the $x$, $y$, and $z$
+  components at a point will be consecutive in the solution vector. It also improves interoperability with external libraries which expect
   data in this format.

### 2022-06-28 9.4/paper.tex
Fix several dead links

@@
 can be found
-\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_3_0_and_9_4_0.html}
+\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_3_3_and_9_4_0.html}
 {in the file that lists all changes for this release}; see \cite{changes94}.
@@
 The 9.4 release includes
-\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_3_0_and_9_4_0.html}
+\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_3_3_and_9_4_0.html}
 {around 40 incompatible changes}; see \cite{changes94}. The majority of these changes

### 2022-06-28 9.4/paper.tex
Use past perfect to describe previous releases
remove one funding source

@@
   % and also 74b8171dd030a4b2de41955b8d913cde06974b91 (Feb 4, 2021)
-  \caption{\it The previous release, 9.3 \cite{dealII93}, had
+  \caption{\it The previous release, 9.3 \cite{dealII93},
     added support for adaptive mesh 
@@
 Awards OAC-2015848, DMS-2028346, and
-EAR-1925575, and by Technical Data Analysis, Inc. through US Navy STTR
-Contract N68335-18-C-0011.
+EAR-1925575.
 

### 2023-03-19 9.5/paper.tex
Copy files -> 9.5
Add sections

--- 2022-06-28 9.4/paper.tex
@@
     Wolfgang Bangerth,
-    Marco Feder,
     Marc Fehling,
-    Rene Gassm\"{o}ller,
     Timo Heister,
@@
     Jean-Paul Pelteret,
-    Simon Sticko,
     Bruno Turcksin,
@@
   },
-  pdftitle={The deal.II Library, Version 9.4, 2022},
+  pdftitle={The deal.II Library, Version 9.5, 2023},
 }
 
-\title{The \dealii Library, Version 9.4}
+\title{The \dealii Library, Version 9.5}
 
@@
 
-   \author[4]{Marco~Feder}
-\affil[4]{SISSA,
+\author[2]{Marc~Fehling}
+
+\author[4]{Timo~Heister}
+ \affil[4]{School of Mathematical and Statistical Sciences,
+   Clemson University,
+   Clemson, SC, 29634, USA
+   {\texttt{heister@clemson.edu}}}
+
+\author[5]{Luca~Heltai}
+\affil[5]{SISSA,
    International School for Advanced Studies,
@@
 
-\author[2]{Marc~Fehling}
-
-\author[5]{Rene~Gassm{\"o}ller}
-\affil[5]{Department of Geological Sciences,
-   University of Florida,
-   1843 Stadium Road,
-   Gainesville, FL, 32611, USA.
-  {\texttt{rene.gassmoeller@ufl.edu}}}
-
-\author[6]{Timo~Heister}
- \affil[6]{School of Mathematical and Statistical Sciences,
-   Clemson University,
-   Clemson, SC, 29634, USA
-   {\texttt{heister@clemson.edu}}}
-
-\author[4]{Luca~Heltai}
-
- \author[7,8]{Martin~Kronbichler}
- \affil[7]{Department of Information Technology,
-   Uppsala University,
-   Box 337, 751\,05 Uppsala, Sweden.
-   {\texttt{martin.kronbichler/simon.sticko@it.uu.se}}}
- \affil[8]{Institute of Mathematics,
+ \author[6]{Martin~Kronbichler}
+ \affil[6]{Institute of Mathematics,
    University of Augsburg,
@@
 
-\author[9]{Matthias~Maier}
-\affil[9]{Department of Mathematics,
+\author[7]{Matthias~Maier}
+\affil[7]{Department of Mathematics,
   Texas A\&M University,
@@
 
-\author[8,10]{Peter Munch}
- \affil[10]{Institute of Material Systems Modeling,
+\author[6,8]{Peter Munch}
+ \affil[8]{Institute of Material Systems Modeling,
  Helmholtz-Zentrum Hereon,
@@
 
-\author[11]{Jean-Paul~Pelteret}
-\affil[11]{Independent researcher.
+\author[9]{Jean-Paul~Pelteret}
+\affil[9]{Independent researcher.
 {\texttt{jppelteret@gmail.com}}}
-
-\author[7,12]{Simon~Sticko}
-\affil[12]{Department of Mathematics and Mathematical Statistics,
-   Umeå University,
-   SE-90187 Umeå, Sweden}
 
@@
 
-\author[13]{David Wells}
-\affil[13]{Department of Mathematics, University of North Carolina,
+\author[10]{David Wells}
+\affil[10]{Department of Mathematics, University of North Carolina,
   Chapel Hill, NC 27516, USA.
@@
   This paper provides an overview of the new features of the finite element
-  library \dealii, version 9.4.
+  library \dealii, version 9.5.
 \end{abstract}
@@
 
-\dealii version 9.4.0 was released June 24, 2022.
+\dealii version 9.5.0 was released XX XX, 2023.
 This paper provides an
 overview of the new features of this release and serves as a citable
-reference for the \dealii software library version 9.4. \dealii is an
+reference for the \dealii software library version 9.5. \dealii is an
 object-oriented finite element library used around the world in the
@@
 \begin{itemize}
-  \item Advances in simplex- and mixed-mesh support (see Section~\ref{sec:simplex});
-  \item Repartitioning of distributed meshes (see Section~\ref{sec:repartitioning});
+  \item Integration of \texttt{Kokkos} (see Section~\ref{sec:kokkos});
+  \item Update of the \texttt{PETSc} infrastructure (see Section~\ref{sec:petsc});
+  \item Addition of new \texttt{Trilinos} wrappers (see Section~\ref{sec:trilinos});
   \item Advances in matrix-free infrastructure (see Section~\ref{sec:mf});
-  \item Advances in multigrid infrastructure (see Section~\ref{sec:multigrid});
-  \item CutFEM support (see Section~\ref{sec:cut});
-  \item Experimental integration of the Computational Geometry Algorithms Library (CGAL) (see Section~\ref{sec:cgalwrappers});
-  \item Performance improvements in the particle infrastructure (see Section~\ref{sec:particles});
-  \item Support for large MPI buffers and parallel I/O (see Section~\ref{sec:large-mpi});
-  \item Improvements to unstructured communication (see Section~\ref{sec:CA});
-  \item Three new tutorial programs and one new code gallery program (see Section~\ref{subsec:steps}).
+  \item Advance in non-matching support (see Section~\ref{sec:nonmatching});
+  \item New liner algebra features (see Section~\ref{sec:lac}).
 \end{itemize}
@@
 \begin{itemize}
-%  \item \texttt{AffineConstraints::make\_consistent\_in\_parallel()} allows
-%  to make constraints consistent in parallel.
-  \item The \texttt{DataOutResample} class interpolates values defined on one
-  triangulation onto a second, potentially unrelated triangulation.
-  By using this class, one can output the result obtained on an
-  unstructured mesh on a structured one (which might facilitate a more
-  memory-efficient storage format, for example, if this second
-  triangulation is a uniformly refined rectangle or box), or one can create a slice in 3D.
-
-\item The new member function \texttt{find\_point\_owner\_rank()} of
-        \texttt{parallel\allowbreak ::distributed::\allowbreak Tri-\allowbreak angulation} allows one to find the MPI
-  ranks of cells containing specified points.
-  It is communication-free and leverages the functionality of \pfrst (>v.2.2).
-  Its algorithm is described in \cite{burstedde2020parallel}. This information will
-  enable efficient construction of the
-  communication pattern used in the class \texttt{Utilities::\allowbreak MPI::\allowbreak RemotePointEvaluation}. Furthermore, this function could be used in the future to allow
-  particle simulations in which particle movement is not
-  limited by CFL conditions, as done in \cite{mirzadeh2016parallel}.
-
-\item The new function
-  \texttt{GridGenerator::pipe\_junction()}
-  generates a triangulation of three cone-shaped pipes that cross at a bifurcation point in any possible configuration.
-  A manifold description is applied to the boundary, which can be extended into the volume via transfinite interpolation \cite{Gordon82} using the \texttt{TransfiniteInterpolationManifold} class \cite{dealII90}.
-  \item A new DoF renumbering function \texttt{DoFRenumbering::support\_point\_wise()} which groups together
-  shape functions by their support point. This functionality is useful in developing nodal schemes since, e.g., the $x$, $y$, and $z$
-  components at a point will be consecutive in the solution vector. It also improves interoperability with external libraries which expect
-  data in this format.
-  \item The \texttt{FEInterfaceValues} class, which computes common quantities at the interface of two cells, has been overhauled to make it
-  more consistent with the rest of the library and use more intuitive names for functions. For example,
-  \texttt{FEInterfaceValues::jump\_gradient()} is now \texttt{FEInterfaceValues::jump\_in\_shape\_gradients()}. Several new
-  functions, such as \texttt{FEInterfaceValues::get\_jump\_in\_function\_values()}, have also been added.
-  \item The \texttt{MeshWorker::ScratchData} and \texttt{MeshWorker::CopyData} have been made $hp$-compatible,
-  and support face integration where the integration rule and mapping differs on either side of an interface.
-  The \texttt{MeshWorker::CopyData} class has also been made compatible with complex numbers.
-  \item Vectors attached to \texttt{DataOut} do not need to be in
-    ghosted state anymore. Internally, we create a copy of the vector
-    with appropriate ghost elements.
+  \item The \texttt{CellAccessor::as\_dof\_handler\_iterator()}
 \end{itemize}
 %
-The changelog lists more than 100 other features and bugfixes.
+The changelog lists more than X other features and bugfixes.
 
@@
 can be found
-\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_3_3_and_9_4_0.html}
-{in the file that lists all changes for this release}; see \cite{changes94}.
+\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_4_2_and_9_5_0.html}
+{in the file that lists all changes for this release}; see \cite{changes95}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Advances in simplex- and mixed-mesh support}\label{sec:simplex}
+\subsection{Integration of Kokkos}\label{sec:kokkos}
 
-\dealii has supported simplex and mixed meshes since the previous
-release, 9.3, on an experimental basis.
-We have continued to work on this support; the current state is
-sufficient to support several larger-scale programs, although not all
-functionality in \dealii works with such meshes yet. Specifically,
-we have fixed many bugs and generalized existing functions that
-previously only worked for hypercube-shaped cells. The most notable
-new functionalities are:
-
-\begin{figure}
-
-  \centering
-
-  \phantom{.}
-  \hfill
-  \begin{tikzpicture}[scale=1.7]
-
-    \coordinate (0) at (0,0);
-    \coordinate (1) at (1,0);
-    \coordinate (2) at (0,1);
-
-    \coordinate (3) at (0,0.5);
-    \coordinate (4) at (0.5,0.5);
-    \coordinate (5) at (0.5,0);
-
-    \foreach \i in {0,1, 2, 3, 4, 5}
-    \draw[blue,fill=blue] (\i) circle (1.1pt) node [below] {};
-
-    \draw (0) --(1) -- (2) -- (0);
-    \draw (3) --(4) -- (5) -- (3);
-
-  \end{tikzpicture}
-  \hfill\hfill
-  \begin{tikzpicture}[scale=1.3]
-
-    \coordinate (0) at (0,0,0);
-    \coordinate (1) at (0,0,1);
-    \coordinate (2) at (1,0,0);
-    \coordinate (3) at (0,1,0);
-    \coordinate (4) at (0,0,0.5);
-    \coordinate (6) at (0.5,0,0);
-    \coordinate (7) at (0,0.5,0);
-    \coordinate (5) at (0.5,0.0,0.5);
-    \coordinate (8) at (0.0,0.5,0.5);
-    \coordinate (9) at (0.5,0.5,0.0);
-
-    \draw (0) -- (1) -- (2) -- (0);
-    \draw (0) -- (1) -- (3) -- (0);
-    \draw (1) -- (2) -- (3) -- (1);
-
-    \draw[-, fill=red!30, opacity=.5] (4)--(6)--(7)--cycle;
-    \draw[-, fill=red!30, opacity=.5] (0)--(6)--(7)--cycle;
-    \draw[-, fill=red!30, opacity=.5] (0)--(4)--(7)--cycle;
-    \draw[-, fill=red!30, opacity=.5] (0)--(4)--(6)--cycle;
-
-    \draw[-, fill=blue!30, opacity=.5] (1)--(4)--(5)--cycle;
-    \draw[-, fill=blue!30, opacity=.5] (1)--(4)--(8)--cycle;
-    \draw[-, fill=blue!30, opacity=.5] (4)--(5)--(8)--cycle;
-    \draw[-, fill=blue!30, opacity=.5] (1)--(5)--(8)--cycle;
-
-    \draw[-, fill=green!30, opacity=.5] (2)--(5)--(6)--cycle;
-    \draw[-, fill=green!30, opacity=.5] (2)--(6)--(9)--cycle;
-    \draw[-, fill=green!30, opacity=.5] (2)--(5)--(9)--cycle;
-    \draw[-, fill=green!30, opacity=.5] (5)--(6)--(9)--cycle;
-
-    \draw[-, fill=orange!30, opacity=.5] (7)--(8)--(9)--cycle;
-    \draw[-, fill=orange!30, opacity=.5] (3)--(7)--(9)--cycle;
-    \draw[-, fill=orange!30, opacity=.5] (3)--(7)--(8)--cycle;
-    \draw[-, fill=orange!30, opacity=.5] (3)--(8)--(9)--cycle;
-
-    \foreach \i in {0,1,2,3,4,5,6,7,8,9}
-    \draw[blue,fill=blue] (\i) circle (1.0pt) node [below] {};
-
-  \end{tikzpicture}
-  \begin{tikzpicture}[scale=1.3]
-
-    \coordinate (0) at (0,0,0);
-    \coordinate (1) at (0,0,1);
-    \coordinate (2) at (1,0,0);
-    \coordinate (3) at (0,1,0);
-    \coordinate (4) at (0,0,0.5);
-    \coordinate (6) at (0.5,0,0);
-    \coordinate (7) at (0,0.5,0);
-    \coordinate (5) at (0.5,0.0,0.5);
-    \coordinate (8) at (0.0,0.5,0.5);
-    \coordinate (9) at (0.5,0.5,0.0);
-
-    \draw (0) -- (1) -- (2) -- (0);
-    \draw (0) -- (1) -- (3) -- (0);
-    \draw (1) -- (2) -- (3) -- (1);
-
-    \draw[-, fill=red!30, opacity=.5] (4)--(5)--(6)--cycle;
-    \draw[-, fill=red!30, opacity=.5] (4)--(5)--(8)--cycle;
-    \draw[-, fill=red!30, opacity=.5] (4)--(6)--(8)--cycle;
-    \draw[-, fill=red!30, opacity=.5] (5)--(6)--(8)--cycle;
-
-%    \draw[-, fill=blue!30, opacity=.5] (4)--(6)--(7)--cycle;
-%    \draw[-, fill=blue!30, opacity=.5] (4)--(6)--(8)--cycle;
-%    \draw[-, fill=blue!30, opacity=.5] (4)--(7)--(8)--cycle;
-%    \draw[-, fill=blue!30, opacity=.5] (6)--(7)--(8)--cycle;
-
-%    \draw[-, fill=green!30, opacity=.5] (6)--(7)--(8)--cycle;
-%    \draw[-, fill=green!30, opacity=.5] (6)--(7)--(9)--cycle;
-%    \draw[-, fill=green!30, opacity=.5] (6)--(8)--(9)--cycle;
-%    \draw[-, fill=green!30, opacity=.5] (7)--(8)--(9)--cycle;
-%
-    \draw[-, fill=blue!30, opacity=.5] (5)--(6)--(8)--cycle;
-    \draw[-, fill=blue!30, opacity=.5] (5)--(6)--(9)--cycle;
-    \draw[-, fill=blue!30, opacity=.5] (5)--(8)--(9)--cycle;
-    \draw[-, fill=blue!30, opacity=.5] (6)--(8)--(9)--cycle;
-
-    \foreach \i in {0,1,2,3,4,5,6,7,8,9}
-    \draw[blue,fill=blue] (\i) circle (1.0pt) node [below] {};
-
-  \end{tikzpicture}
-  \begin{tikzpicture}[scale=1.3]
-
-    \coordinate (0) at (0,0,0);
-    \coordinate (1) at (0,0,1);
-    \coordinate (2) at (1,0,0);
-    \coordinate (3) at (0,1,0);
-    \coordinate (4) at (0,0,0.5);
-    \coordinate (6) at (0.5,0,0);
-    \coordinate (7) at (0,0.5,0);
-    \coordinate (5) at (0.5,0.0,0.5);
-    \coordinate (8) at (0.0,0.5,0.5);
-    \coordinate (9) at (0.5,0.5,0.0);
-
-    \draw (0) -- (1) -- (2) -- (0);
-    \draw (0) -- (1) -- (3) -- (0);
-    \draw (1) -- (2) -- (3) -- (1);
-
-%    \draw[-, fill=red!30, opacity=.5] (4)--(5)--(6)--cycle;
-%    \draw[-, fill=red!30, opacity=.5] (4)--(5)--(8)--cycle;
-%    \draw[-, fill=red!30, opacity=.5] (4)--(6)--(8)--cycle;
-%    \draw[-, fill=red!30, opacity=.5] (5)--(6)--(8)--cycle;
-
-    \draw[-, fill=red!30, opacity=.5] (4)--(6)--(7)--cycle;
-    \draw[-, fill=red!30, opacity=.5] (4)--(6)--(8)--cycle;
-    \draw[-, fill=red!30, opacity=.5] (4)--(7)--(8)--cycle;
-    \draw[-, fill=red!30, opacity=.5] (6)--(7)--(8)--cycle;
-
-    \draw[-, fill=blue!30, opacity=.5] (6)--(7)--(8)--cycle;
-    \draw[-, fill=blue!30, opacity=.5] (6)--(7)--(9)--cycle;
-    \draw[-, fill=blue!30, opacity=.5] (6)--(8)--(9)--cycle;
-    \draw[-, fill=blue!30, opacity=.5] (7)--(8)--(9)--cycle;
-
-%    \draw[-, fill=orange!30, opacity=.5] (5)--(6)--(8)--cycle;
-%    \draw[-, fill=orange!30, opacity=.5] (5)--(6)--(9)--cycle;
-%    \draw[-, fill=orange!30, opacity=.5] (5)--(8)--(9)--cycle;
-%    \draw[-, fill=orange!30, opacity=.5] (6)--(8)--(9)--cycle;
-
-    \foreach \i in {0,1,2,3,4,5,6,7,8,9}
-    \draw[blue,fill=blue] (\i) circle (1.0pt) node [below] {};
-
-  \end{tikzpicture}
-  \hfill
-  \phantom{.}
-
-  % We definitely supported tri refinement in 9.3 - see 6ceb8b70559dc2274761a28dc7836f1e9602da8c (Dec 4, 2020)
-  % and also 74b8171dd030a4b2de41955b8d913cde06974b91 (Feb 4, 2021)
-  \caption{\it The previous release, 9.3 \cite{dealII93},
-    added support for adaptive mesh 
-  refinement with triangles. This release adds support for global
-  tetrahedral refinement by subdividing each tetrahedron into eight
-  children, as shown on the right.}
-  \label{fig:refinement}
-\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Update of the PETSc infrastructure}\label{sec:petsc}
 
 \begin{itemize}
-\item Experimental support for locally refined meshes: For finite
-  element computations on locally refined
-meshes, one needs (i) the possibility to locally refine the mesh (see
-Figure~\ref{fig:refinement}), and (ii) appropriate hanging-node
-constraints. Both are now available in 2D; for 3D, the implementation
-of the constraint definitions is still in progress.
-\item \texttt{QIteratedSimplex} allows to build composite simplex quadrature rules.
-\item The new wrappers to the \texttt{CGAL} library allow for the easy
-  creation of simplex
-meshes. For more details, see Section~\ref{sec:cgalwrappers}.
+\item PreconditionBDDC
 \end{itemize}
 
-Furthermore, we have continued to remove uses of the
-\texttt{GeometryInfo} class (which is specific to hypercube cells) from
-the library, and to replace them with more general equivalent
-functionality based on the
-\texttt{ReferenceCell} class. Once all instances of
-\texttt{GeometryInfo} are removed, we will deprecate the class.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{New Trilinos wrappers}\label{sec:trilinos}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Repartitioning of distributed meshes}\label{sec:repartitioning}
-
-\dealii has two classes that support distributed storage of meshes:
-\texttt{parallel::\allowbreak distributed::\allowbreak Triangulation}
-(which in the following we will abbreviate as \texttt{p::d::T}) and
-\texttt{parallel::\allowbreak fully\allowbreak distributed::\allowbreak
-  Triangulation} (or, in short, \texttt{p::f::T}). The former is
-partitioned based on the space-filling Morton (or Z-order) curve
-implemented by the \pfrst backend. The latter, more recent
-class is currently statically partitioned at creation time.
-
-There are now new utility functions in the
-\texttt{RepartitioningPolicyTools} namespace that can be
-used to create a new \texttt{p::f::T} instance,
-given a distributed triangulation (\texttt{p::f::T} or
-\texttt{p::d::T}) and a vector that indicates the designated owner processes of locally
-owned cells. This approach allows for partitioning \texttt{p::f::T}
-based on arbitrary criteria. The workflow is shown in the following listing:
-\begin{c++}
-// Original triangulation:
-parallel::distributed::Triangulation<dim> tria(communicator);
-
-// Select a partitioning policy and use it:
-const RepartitioningPolicyTools::CellWeightPolicy<dim> policy(tria, fu);
-const auto construction_data = TriangulationDescription::Utilities::
-  create_description_from_triangulation(tria, policy.partition(tria));
-
-// Create the new triangulation:
-parallel::fullydistributed::Triangulation<dim> tria_pft(comm);
-tria_pft.create_triangulation(construction_data);
-\end{c++}
-
-\begin{figure}
-    \centering
-    \def\svgwidth{0.8\columnwidth}
-    \input{svg/repartitioning.pdf_tex}
-    \caption{\it Visualization of the repartitioning process and the
-      construction of the new mesh of process 0. Top
-      left: Existing ownership of the cells of the mesh, distributed
-      on four processes. Bottom left:
-      Requested new ownership. Second column: Processes 0 and 1
-      collect information about the cells they own or that are ghost
-      cells. Processes 2 and 3 do not contribute to process 0. Third column: What
-      processes 0 and 1 would send to process 0. Fourth column: The
-      combined knowledge on process 0.}\label{fig:repartitioning}
-\end{figure}
-
-The setup process (of \texttt{construction\_data}) is visualized in Figure~\ref{fig:repartitioning}. At
-first, locally owned cells and their
-surrounding (ghost) cells are collected on each process and
-sent to the new owner. On the
-receiving side, the sets of all cells are combined and possible duplicates
-are removed. This information is enough to construct a new triangulation.
-For sending/receiving, we apply consensus-based algorithms~\cite{hoefler2010scalable}, which
-we introduced into the library in release 9.2~\cite{dealII92} -- see
-also Section~\ref{sec:CA}. This kind of consensus-based
-algorithm is used also in \cite{ibanez2016pumi} for repartitioning.
-
-In addition to the predefined partitioning policies, users
-can write their own by implementing the
-\texttt{RepartitioningPolicyTools::Base} interface. Like the
-``active'' level of finest cells, the scheme also allows for
-arbitrarily repartitioned multigrid levels.
-
-In future releases, we plan to add support for repartitioning based on distributed graph
-partitioning libraries, e.g., \texttt{ParMETIS} or \texttt{Zoltan}. Furthermore, we intend to extend
-\texttt{p::f::T} to support adaptive mesh refinement, completing \texttt{p::f::T}
-as a complement to \texttt{p::d::T}.
+\begin{itemize}
+\item Belos
+\item NOX
+\end{itemize}
 
@@
 
-The matrix-free infrastructure of \dealii is used to solve problems
-without assembling matrices, providing only the \textit{action} of the matrix
-instead. It has seen a number of new features in this release, the most
-notable ones being:
 \begin{itemize}
-\item Improved support for computations with Hessians of shape functions: Just like values and gradients, Hessians can be
-now evaluated and integrated during matrix-free loops
-both for cells and faces. This could enable, for example, writing
-a matrix-free version of step-47, which solves the biharmonic equation with the discontinuous Galerkin method.
-\item Cell-centric loops now also allow access to gradients and Hessians
-of neighboring cells on faces. The major difficulty here was the relative orientation
-of cells on unstructured meshes.
-\item Users can now create their own cell batches, by providing \texttt{FEEvaluation::reinit()} a list of cell IDs. \texttt{FEEvaluation}
-accesses the appropriate data and reshuffles mapping data accordingly on
-the fly in order to enable vectorization over cells. The new feature is useful in several
-contexts, for example for sharp interfaces (e.g., two-phase flow
-or shock capturing), where one needs to treat cells that are ``cut'' by
-the interface in a special way. A challenge is that cell batches
-might contain cut or non-cut cells, effectively counteracting the cross-cell vectorization.
-Previous functionality has provided the option of masking certain cells in cell batches, which works well if
-the code paths do not diverge too much, or to
-categorize cells during \texttt{MatrixFree::reinit()} in such a way that mixed cell batches do not occur. However, \texttt{MatrixFree::reinit()}
-might be too expensive if recategorization needs to happen very frequently to follow the dynamics of a system, e.g., in each time step.
-Despite some overhead compared to static matrix-free loops, the new feature can be the best option in such dynamic scenarios.
-\item Initial support for H(div)-conforming elements with Piola transform, based on Raviart--Thomas finite element spaces with the updated class \texttt{FE\_RaviartThomasNodal}, has been added.
-This feature is currently limited to meshes in standard orientation and affine geometries.
-Full support and performance optimizations will be provided in a future release.
-\item Selected matrix-free algorithms can now exploit additional data locality
-  between the matrix-vector product and vector operations happening nearby in
-  an algorithm. Interfaces have been added to both the
-  \texttt{PreconditionChebyshev} class (a frequently used smoother in multigrid
-  methods) and the conjugate gradient implementation in \texttt{SolverCG}.
-  From the user's perspective, an operator needs to define a \texttt{vmult} operation,
-  taking two additional \texttt{std::function} arguments. The first function
-  defines the operation to be scheduled on the vector entries before the
-  matrix-vector product touches them, and the second what happens
-  afterward. The new features also include a renumbering to maximize data
-  locality. The theory is described in the
-  contribution~\cite{kronbichler2022cg}.
+\item element activation and deactivation, \texttt{FE\_Nothing}, categorization, DG, face integrals 
+\item pre/post for DG
+\item \texttt{CellwiseInverseMassMatrix} operator for coupling (dyadic) coefficients
 \end{itemize}
-Besides these new features, we improved the performance of
-hanging-node-constraint evaluation on the CPU. Instead of performing
-quasi-dense matrix-vector multiplications~\cite{KronbichlerKormann2012}, we now use an
-approach based on in-place interpolation and sum factorization, similar
-to what was already done in the GPU code~\cite{ljungkvist2017matrix}. In \cite{munch2022hn}, the algorithm is described
-and performance numbers are shown, indicating a reduction
-of overhead of cells with hanging nodes by a factor of ten.
-
-Finally, we have performed a major restructuring of the internals
-of the \texttt{FEEvaluation} classes. This reduces some overhead for low polynomial degrees and will enable us to add support for new element types in the future.
-
-We would like to remind users that we transitioned from the use of Booleans
-to flags to configure the evaluation and integration process of \texttt{FEEvaluation}
-and \texttt{FEFaceEvaluation}:
-
-\begin{c++}
-fe_eval.evaluate(false, true, false)         // old (deprecated)
-fe_eval.evaluate(EvaluationFlags::gradients) // new
-\end{c++}
-
-%\begin{c++}
-%additional_data.mapping_update_flags = ... | update_hessians;
-%\end{c++}
-%
-%\begin{c++}
-%FEEvaluation<dim> phi(matrix_free);
-%
-%phi.reinit(cell);
-%phi.gather_evaluate(src, ... | EvaluationFlags::hessians);
-%
-%for(const auto q : phi.quadrature_point_indices())
-%  phi.submit_hessian(phi.get_hessian(q), q);
-%
-%phi.integrate_scatter(... | EvaluationFlags::hessians, dst);
-%\end{c++}
-%
-%... similar for faces; can be used to implement biharmonic equation (like in
-%step-X)
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Advances in multigrid infrastructure}\label{sec:multigrid}
-
-In release 9.3~\cite{dealII93}, we added support for global-coarsening multigrid in
-addition to the established local-smoothing infrastructure. Global
-coarsening algorithms smoothen over the whole computational domain on each
-multigrid level, which is obtained by coarsening the finest cells of
-the next finer multigrid level.
-For this purpose, we use a sequence of triangulations, and we perform
-the smoothing only on their active levels. To create the sequence of
-triangulations, one can use the functions \texttt{MGTransferGlobalCoarseningTools::create\_geometric\_coarsening\_sequence()}. A new version takes an
-instance of \texttt{RepartitioningPolicyTools::Base} (see Subsection~\ref{sec:repartitioning}) as argument, which allows specifying the parallel
-distribution of each multigrid level (in contrast to the fixed first-child policy in the case of local smoothing). These features have been developed and tuned for running on a supercomputer scale
-with complicated coarse meshes as presented in \cite{kronbichler2021next}.
-Furthermore, we added support for block vectors,
-fixed a number of limitations, and performed performance optimizations of
-the transfer operator; particularly, the redundant copy from/to temporary vectors
-has been eliminated. Furthermore, hanging-node constraints are applied
-efficiently in the same way as in the matrix-free loops (see Subsection~\ref{sec:mf}).
-
-In \cite{munch2022gc}, the performance of the local-smoothing and global-coarsening
-infrastructure of \dealii was compared for locally refined meshes. The results indicate that
-the local definition of multigrid levels might introduce load imbalances
-in the case of local smoothing so that global coarsening is favorable despite
-potentially more expensive intergrid transfers. In order to assess the benefits
-of one approach against the other, \dealii provides new functions
-\texttt{workload\_imbalance()} and \texttt{vertical\_communication\_efficiency()}
-in the \texttt{MGTools} namespace for the  estimation of the imbalance during, e.g.,
-smoothing or the
-communication efficiency during intergrid transfer, purely based on the given mesh.
-% MK: I would not add this part, it does not really fit into this paper as
-% there is nothing to report at this point (and we should then add references
-% to the actual literature).
-%The publication \cite{munch2022gc} also points out that not
-%all types of smoothers are applicable for global coarsening due to the
-%presence of hanging nodes, which is a motivation to add new smoother types
-%to \dealii in the future.
+\subsection{Advance in non-matching support}\label{sec:nonmatching}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{CutFEM support}\label{sec:cut}
-
-Several classes have been added to the \texttt{NonMatching} namespace to enable the use of cut finite element methods~\cite{burman_cutfem_2015}.
-In the literature, these types of methods are also referred to as immersed, extended, or fictitious finite element methods.
-Here, the domain, $\Omega$, is immersed in the background mesh, as illustrated in Fig.~\ref{fig:immersed-domain}.
-Often, one solves for the degrees of freedom of the smallest submesh
-which completely covers the domain, i.e., the blue and yellow cells of
-Fig.~\ref{fig:location-to-level-set}.
-The bilinear form in the weak form would then, for example, look like
-\begin{equation}
-  a(u,v) = (\nabla u, \nabla v)_\Omega - (\partial_n u, v)_\Gamma + \ldots
-\end{equation}
-Thus, when assembling on a cut cell $K$, we are
-required to integrate over the part of the domain and the part of the boundary, $\Gamma = \partial \Omega$, that falls inside the cell:
-$K\cap \Omega$ and $K \cap \Gamma$.
-Many of the new classes that support these operations assume that the domain is described by a level set function,
-$\psi : \mathbb{R}^d \to \mathbb{R}$, such that
-\begin{align}\label{eq:levelset}
-  \Omega = \{x \in \mathbb{R}^d : \psi(x)<0\},
-  \qquad
-  \Gamma = \{x \in \mathbb{R}^d : \psi(x) = 0\}.
-\end{align}
-
-\begin{figure}
-  \centering
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \includegraphics[height=.12\paperheight]{svg/immersed-domain.pdf}
-    \caption{\label{fig:immersed-domain}}
-  \end{subfigure}
-  \qquad
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \includegraphics[height=.12\paperheight]{svg/location-to-level-set.pdf}
-    \caption{ \label{fig:location-to-level-set}}
-  \end{subfigure}
-  \caption{\it (a) Domain immersed in a background mesh. (b) Value of \texttt{LocationToLevelSet} for each cell.}
-%\end{figure}
-%
-%\begin{figure}
-  \centering
-  \includegraphics[width=.4\paperwidth]{svg/immersed_quadratures.pdf}
-  \caption{\it Quadrature points for integrating over the three different regions of a cell cut by the zero contour of the level set function, $\psi$. \label{fig:immersed_quadratures}}
-\end{figure}
-
-Specifically, the following are the key new classes and functions:
-\begin{itemize}
-  \item The \texttt{MeshClassifier} class identifies how the active cells and faces are located relative to the zero contour of the level set function, as illustrated in Figure~\ref{fig:location-to-level-set}.
-Its member function \texttt{location\_to\_level\_set()} takes a cell or face and
-        returns an enum, \texttt{LocationToLevelSet}, with values \{\texttt{inside}, \texttt{outside}, \texttt{intersected}\}.
-        This information is typically needed when choosing what element (e.g., \texttt{FE\_Q} or \texttt{FE\_Nothing}) and/or what quadrature (defined over the 
-        complete cell or over a part---see below) a cell
-        should use.
-
-  \item The \texttt{QuadratureGenerator} class, which implements the algorithm in \cite{saye2015}, generates high-order quadrature rules for the three different regions of a \texttt{BoundingBox}, $B$, defined by the sign of the level set function:
-        \begin{align}\label{eq:boundingbox}
-          B \cap \Omega = \{ x\in B:  \psi(x) < 0 \},  \quad
-          B \cap \Gamma = \{ x\in B:  \psi(x) = 0 \},  \quad
-          \{ x\in B:  \psi(x) > 0 \}.
-        \end{align}
-        An example of these quadratures is shown in Figure~\ref{fig:immersed_quadratures}.
-        The \texttt{FaceQuadratureGenerator} class does the same for faces.
-        Furthermore, the new classes \texttt{DiscreteQuadratureGenerator} and \texttt{DiscreteFaceQuadratureGenerator} can be used to generate these quadrature rules over a cell or face when the level set function lies in a finite element space: $\psi_h \in V_h$, and when the reference cell of the cell or face is a hypercube.
-
-  \item \texttt{ImmersedSurfaceQuadrature} is a class representing a quadrature rule over a $(d-1)$-dimensional surface embedded in $\mathbb{R}^d$ ($\psi = 0$ in Figure~\ref{fig:immersed_quadratures}).
-In addition to the weight, it stores the unit normal to the surface, for each quadrature point.
-This is needed to transform the quadrature rule from reference space to real space.
-
-  \item \texttt{FEImmersedSurfaceValues} is an \texttt{FEFaceValues}-like class for evaluating real space values based on an \texttt{ImmersedSurfaceQuadrature}.
-
-  \item \texttt{NonMatching::FEValues} combines the functionality of
-    several of the above classes to simplify assembly of linear systems. It works similarly to \texttt{hp::FEValues}:
-  When calling the \texttt{reinit()} function, immersed quadrature rules are generated in the background and
-  \texttt{FEValues} objects for the inside/outside region and a \texttt{FEImmersedSurfaceValues} object for the surface regions are set up internally.
-These can then be obtained using getter-functions (\texttt{get\_inside/outside/surface\_fe\_values()}) and used for the assembly.
-  Since the generation of immersed quadrature rules is not cheap,
-  \texttt{NonMatching::FEValues} calls  \texttt{QuadratureGenerator} only if needed, i.e., if the cell is intersected.
-If not, already cached \texttt{FEValues} objects will be returned by the getter functions.
-  Correspondingly, the class \texttt{NonMatching::FEInterfaceValues} generates
-  \texttt{FEInterfaceValues} objects for assembling face terms over $F \cap \{x : \psi(x) < 0 \}$ or $F \cap \{x : \psi(x) > 0 \}$.
-\end{itemize}
-The new \texttt{step-85} tutorial illustrates how many of these classes work together.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Experimental integration of the Computational Geometry Algorithms
-  Library (CGAL)}\label{sec:cgalwrappers}
-
-The Computational Geometry Algorithms Library (CGAL, \url{https://www.cgal.org/}) is a widely used
-library to describe geometries and meshes \cite{cgal-user-ref}. \dealii now
-has wrappers for CGAL classes and functions, provided in the new
-namespace \texttt{CGALWrappers}: they implement functionality
-spanning from mesh generation to boolean operations between
-triangulations and cells. These wrappers are enabled only if \dealii
-is compiled with \texttt{C++17}. \textit{This feature is still experimental and
-interfaces might change during the next release cycle.}
-
-The main mesh generation function is
-\texttt{GridGenerator::implicit\_function()}, which creates a \texttt{Triangulation<dim,3>} out of the zero level set of an implicit function $\psi$ similar to \eqref{eq:levelset}.
-For \texttt{dim==3}, the mesh consists of tetrahedra.
-A prototypical use case is the following, where the surface is the zero level set of Taubin's heart function $f=\bigl ( x^2 + \frac{9y^2}{4} +z^2 -1 \bigr ) -x^2 z^3 - \frac{9y^2z^3}{80}$.
-The resulting \texttt{Triangulation<3>} is shown in Figure~\ref{fig:heart_tria}
-and the required steps are:
-\begin{c++}
-// 1) An implicit function, e.g., Taubin's heart surface (not shown)
-ImplicitFunction f;
-
-// 2) Configure output mesh (optional)
-CGALWrappers::AdditionalData<3> data; data.cell_size = .05;
-
-// 3) Create mesh
-Triangulation<3> tria;
-GridGenerator::implicit_function(tria, f, data, /* more arguments */...);
-\end{c++}
-
-A related function is
-\texttt{GridGenerator::surface\_mesh\_to\_volumetric\_mesh()}, which
-computes a tetrahedral volume mesh \texttt{Triangulation<3>}, based on a
-given surface mesh \texttt{Triangulation<2,3>} that bounds
-the three dimensional shape.
-
-CGAL also provides Boolean operations on meshes: they can be accessed via 
-the utility function
-\texttt{CGALWrappers::\allowbreak{}compute\_boolean\_operation()}. The
-available operations are \textit{co-refinement}, \textit{intersection},
-\textit{union}, and \textit{difference}. Oftentimes, Boolean
-operations and co-refinement around the intersection produces
-badly shaped mesh cells. To overcome this issue, one can use \texttt{CGALWrappers::\allowbreak{}remesh\_surface()}.
-A possible workflow to create a high-quality mesh of the union of a cube and
-a sphere mesh is given in the following listing:
-\begin{c++}
-// 1) Create deal.II triangulations, e.g., cube and sphere (not shown)  
-Triangulation<spacedim> tria0, tria1;
-
-// 2) Convert to CGAL surface meshes (assuming Kernel is already defined) 
-CGAL::Surface_mesh<Kernel> surface_mesh0, surface_mesh1;
-CGALWrappers::dealii_tria_to_cgal_surface_mesh(tria0, surface_mesh0);
-CGALWrappers::dealii_tria_to_cgal_surface_mesh(tria1, surface_mesh1);
-
-// 3) Compute the union of the two meshes
-CGALWrappers::compute_boolean_operation(surface_mesh0, surface_mesh1,
-  BooleanOperation::compute_union, out_mesh);
-
-// 4) Convert CGAL surface mesh to deal.II surface mesh
-Triangulation<2, 3> tria_out;
-CGALWrappers::cgal_surface_mesh_to_dealii_triangulation(out_mesh, tria_out);
-
-// 5) Convert surface to volume mesh via surface_mesh_to_volumetric_mesh()
-\end{c++}
-The output of the Boolean operation is shown in Fig.~\ref{fig:corefinement};
-Fig.~\ref{fig:corefinement_remeshed} shows the mesh after remeshing.
-\begin{figure}
-  \centering
-  \begin{subfigure}[b]{0.28\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{png/heart_implicit.png}
-    \caption{\label{fig:heart_tria}}
-  \end{subfigure}\qquad
-  \hfill
-  \begin{subfigure}[b]{0.28\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{png/intersection_cube_sphere_mesh.png}
-    \caption{\label{fig:corefinement}}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.35\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{png/cube_sphere_remeshed.png}
-    \caption{ \label{fig:corefinement_remeshed}}
-  \end{subfigure}
-  \caption{\it (a) Triangulation created by filling a heart-shaped surface implicitly described by a function $f$.
-(b) Union of a cube with a sphere with badly shaped cells at the intersection.
-(c) Remeshed version of the same triangulation.}
-\end{figure}
-
-The function
-\texttt{CGALWrappers::compute\_quadrature\_on\_boolean\_operation()}
-returns a quadrature that allows exact integration of polynomials on
-polyhedral elements created by a Boolean operation between \dealii cells.
-The quadrature rule is built by (i) subdividing the polyhedral region in tetrahedra, (ii) computing on each a \texttt{QGaussSimplex<3>} quadrature rule via \texttt{QSimplex<3>::\allowbreak{}compute\_affine\_transformation()}, and (iii)
-collecting all in a single quadrature object on the \emph{physical} cell.
-
-The new utility functions will be the building blocks for functions in 
-the \texttt{NonMatching} namespace in future releases. While our current 
-CutFEM implementation (see Section~\ref{sec:cut}) relies on a level set description of the domain to determine
-the boundary and the quadrature rules, one could do that also based on non-matching overlapping
-grids~\cite{massing2013efficient}, for which the newly introduced, CGAL-based Boolean operations and 
-quadrature-generation functionality might be helpful.
-The functionalities are also applicable in a broad context of immersed problems, such as
-for weakly imposing boundary conditions on the interface of an immersed boundary
-using Nitsche's method and for coupling terms using Lagrange multipliers.
-
-%These utility functions will be the building blocks for functions in the \texttt{NonMatching} namespace that will, e.g.,  assemble coupling terms like $(u,v)_{\Omega}$, with $\Omega$ a domain immersed in a fixed background mesh $B$ and $u,v$ finite element functions on $V_h(B)$, as needed, e.g., in
-%the context of CutFEM (see Section~\ref{sec:cut}) or Nitsche's method to weakly impose boundary conditions at an interface.
-The same applies to coupling terms of the form $(u,q)_{\Omega}$ in formulations using Lagrange multipliers, where now $q \in Q_h(\Omega)$, with $Q_h(\Omega)$ the space of the multiplier variable.
-%Note that the most relevant difference between this and the \texttt{QuadratureGenerator} in Section~\ref{sec:cut} is that the \texttt{Quadrature} objects are created directly from two overlapping grids, one
-%spanning over $B$ and the other one over $\Omega$, and not from a level set function.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Performance improvement of particle infrastructure}\label{sec:particles}
-
-In the release described herein, we have reorganized the data
-structures and optimized the algorithms that support using particles
-in \dealii.
-In our previously reported implementation~\citep{dealII93} we stored all particle data as a separate contiguous array for each particle property, but particle identifiers (IDs) were still stored in a multimap tree-like structure.
-
-Our new particle containers are organized as follows: Particle IDs are stored in a list of dynamic arrays, each array containing the particle IDs of all particles in a unique cell.
-Each ID is a handle that determines the location of the data of this unique particle in the property arrays.
-The list of ID arrays only contains entries for cells that contain particles.
-We keep a separate cache structure that contains pointers to the particular list entries for each cell.
-If a cell has no particles, this pointer is invalid.
-
-This structure allows for the following significant performance improvements:
+\subsection{New linear algebra features}\label{sec:lac}
 
 \begin{itemize}
-\item All particle data (both their identifiers and their actual data) are now stored as separate and contiguous arrays in memory, which improves spatial locality for better prefetching of data and makes iterating over particles extremely efficient.
-\item The choice of a list container that only includes entries for cells that contain particles means iteration is efficient, even if many cells in the domain do not contain any particles (as can be the case for discrete element methods~\cite{golshan2022lethe}).
-\item Creating separate arrays for each cell allows us to easily move particle IDs from one cell to another as a local operation, affecting only the two cell containers in question.
-We take care to reuse allocated memory to minimize the number of memory reallocations.
-\item The separate cache structure that contains entries for each cell allows quick random access to the particles of a particular cell, and also allows quickly determining if a particular cell has particles at all.
+  \item \texttt{SolverGMRES}: classical/modified Gram Schmidt process
+  \item James Lottes’s fourth-kind Chebyshev
+  \item \texttt{PreconditionChebyshev}, \texttt{PreconditionRelaxation}
+  \item restrict matrices
+  \item fast diagonalization method
 \end{itemize}
-
-In addition to the new storage structure, we have made the following algorithmic improvements:
-Determining if a particle is inside a cell after changing its position involves inverting the mapping for this cell.
-We have reorganized our algorithms to perform these inversions on a batch of particles in the same cell instead of particle-by-particle, which allows us to make use of vectorized instructions during the inversion using the generic scheme of~\cite{KronbichlerKormann2012}.
-In addition, after sorting all particles into their new cells, the arrays that store particle properties are now sorted in the same order as the particle IDs in the list of arrays, which allows for cache efficient iteration over particle properties.
-Because the particle IDs are already sorted in the intended order at this time, we can reorder the particle properties as part of a copy operation into a new data container that replaces the existing container.
-This approach avoids a costly in-place sort of the particle properties.
-
-\begin{table}
-  \caption{\it Timing of various particle operations for tutorial program \texttt{step-68} (particle advection in a 2D, Cartesian box) using 400,000 particles on a single process.}
-  \label{tab:particle_timing}
-  \centering
-  \begin{tabular}{lccc}
-    \toprule
-    Particle Operation & \dealii 9.3 & \dealii 9.4 & Speedup      \\ \midrule
-    Generation         & 444 ms      & 235 ms      & 1.9$\times$  \\
-    Iteration          & 4.18 ms     & 0.638 ms    & 6.6$\times$  \\
-    Advection          & 37.8 ms     & 33.9 ms     & 1.15$\times$ \\
-    Sorting            & 21.9 ms     & 9.27 ms     & 2.4$\times$  \\
-    \bottomrule
-  \end{tabular}
-  \end{table}
-
-We illustrate the combined effect of these performance improvements in Table~\ref{tab:particle_timing}.
-We measure the averaged compute time for four particle operations in a slightly modified version of the \dealii tutorial program \texttt{step-68} when advecting 400,000 particles on a single process (we have not observed any influence of the described changes on the parallel scalability of the algorithms).
-The four operations we have measured are:
-\begin{itemize}
-\item Generation of a set of 400,000 particles at positions that are not aligned with the background mesh, i.e., the containing cell of each particle has to be identified with a search algorithm.
-\item Iteration over the whole set of created particles, without
-  significant computation and in particular without accessing particle data.
-\item Advection of all particles, which involves iteration over all particles, evaluation of the finite element solution at the location of the particles, read and write access to the position of all particles to modify their location, and write access to the particle properties (to store their velocity for visualization purposes).
-\item Sorting, i.e., the inversion of the mapping of each cell to find the new particle locations relative to this cell, and moving all particles that have left their original cell into new cells (both \dealii 9.3 and \dealii 9.4).
-In \dealii 9.4, this operation also includes reordering the particle properties for optimal iteration.
-\end{itemize}
-
-Table~\ref{tab:particle_timing} shows that all particle operations are
-much faster in \dealii 9.4 than in version 9.3.
-In particular, operations that depend strongly on particle storage structure and require few fixed computations (like iteration and sorting) benefit massively from the above-mentioned optimizations.
-We note that the exact gains will depend strongly on the exact combination of geometry, mapping, dimensionality, and the number of particles per cell in any specific model, and can be smaller or larger than the measurements provided here.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Support for large MPI buffers and parallel I/O}
-\label{sec:large-mpi}
-
-   Sending large messages over MPI or reading and writing large buffers using MPI I/O,
-   especially when dealing with small datatypes like \texttt{MPI\_CHAR},
-   can often require processing more than $2^{31}$ objects at once. Prior
-   to the recent MPI-4 standard, which introduced a new set of \texttt{MPI\_*\_c()} functions, all MPI functions expected the ``count''
-   to be a signed integer. This limits the number of objects to $2^{31}$,
-   i.e., writing of up to 2 GB at once if the datatype is \texttt{MPI\_CHAR}.
-   Several places in the library could exceed this limit,
-   for example when producing large graphical output, large
-   checkpoint files, or when broadcasting large datasets such as lookup tables.
-
-   With this release we introduce a new module \texttt{Utilities::MPI::LargeCount} which enables sending and receiving MPI messages and I/O containing more than $2^{31}$ objects by implementing
-   the before-mentioned \texttt{MPI\_*\_c()} functions, e.g., \texttt{MPI\_Bcast\_c()}, using MPI-3 features if necessary. The
-   solution is based on custom datatypes as described in~\cite{hammond2014int_max}.
-This functionality is now used in all places in the library where
-   large counts might be necessary (\texttt{DataOut::write\_vtu\_in\_parallel()} for graphical output, checkpointing of \texttt{Triangulation} objects, etc.).
-
-   While implementing these changes we also worked on the following:
-   \begin{itemize}
-    \item Testing of large I/O with large chunks per MPI rank (>2 GB) and
-    large total sizes (>4 GB). Several instances of 32-bit data types for
-    offsets were changed to 64-bit to correctly support files larger
-    than 4 GB (HDF5 output, VTU output, checkpointing).
-    \item Performance testing of MPI I/O routines used for parallel VTU output with large performance improvements by switching from a shared
-    file pointer to individual file pointers, see Table~\ref{tab:mpi-io}.
-   \end{itemize}
-
-\begin{table}
-  \caption{\it Performance comparison of writing large VTU graphical output using MPI I/O on
-  TACC Frontera \texttt{/scratch1} file system with 16 file servers (``OSTs'') and a theoretical peak performance of 50 GB/s.
-The ROMIO version used enforces sequential writes when using \texttt{MPI\_File\_write\_ordered()}.}
-  \label{tab:mpi-io}
-
-  \centering
-
-  \begin{tabular}{llcc}
-    \toprule
-    Version       & MPI routine used                     & 1 file, striping 16 & 16 files, no striping \\\midrule
-    \dealii{} 9.3 & \texttt{MPI\_File\_write\_ordered()} & 3 GB/s              & 15 GB/s               \\
-    \dealii{} 9.4 & \texttt{MPI\_File\_write\_at\_all()} & 17 GB/s             & 21 GB/s               \\
-    \bottomrule
-\end{tabular}
-
-  \end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Improvements to unstructured communication}
-\label{sec:CA}
-
-Many problems in parallel computations can be stated in the following
-way: Each process in a parallel universe has a number of queries to
-send to other processes that do not know that they will be asked, and
-who will then have to respond with replies. This problem is solved by
-``consensus algorithms''~\cite{hoefler2010scalable}. An example of where this problem appears is
-given in Section~\ref{sec:repartitioning}.
-
-\dealii has had an implementation of these algorithms for some time,
-but the current release substantially expands on it. Specifically, the
-updated interfaces -- now based on function objects such as lambda
-functions to formulate and process queries and replies -- can deal
-with arbitrary data types for queries and replies, rather than only
-arrays of data types natively supported by MPI. To make this possible,
-the implementation packs and unpacks these objects into character
-arrays via the \texttt{Utilities::pack()} and
-\texttt{Utilities::unpack()} functions. We have also worked on making
-these functions efficient: if the object to be packed is an array
-(or array of arrays) of elements that satisfy the
-\texttt{std::is\_trivially\_copyable} type trait, then the data is
-copied into the character array via \texttt{std::memcpy}; only for
-other objects do the packing and unpacking functions rely on BOOST's
-serialization library.
-
-A special case of consensus algorithms is where the sender does not
-actually require an answer. This happens, for example, during repartitioning
-of meshes (Section~\ref{sec:repartitioning}), where a process sends parts of
-the new mesh to the new owners: the new owner does not know how many processes
-will send it mesh parts and the sender only needs an acknowledgment that the
-data has been received. Previously, such a case was implemented through a consensus
-algorithm where the reply message is simply empty.
-The rewritten interfaces now support this case more explicitly: Code using
-these interfaces no longer has to provide functions that formulate
-and read the (empty) replies, though internally these functions still
-send around an empty reply; this case will be implemented in the
-future, using the interfaces now already in place.
-
-By the time of writing, \dealii uses consensus-based algorithms to determine
-the owners of distributed index sets (\texttt{Utilities::MPI::Partitioner},
-\texttt{Utilities::MPI::Noncontiguous\allowbreak Partitioner}, \texttt{internal::MatrixFreeFunctions::VectorDataExchange}; see~\cite{dealII91}),
-to set up the global-coarsening transfer operators (see~\cite{dealII92} and
-Section~\ref{sec:multigrid}), to repartition distributed meshes (see Section~\ref{sec:repartitioning}), and basis coupling algorithms between non-matching
-meshes, based on the communication patters in \texttt{RemotePointeEvaluation} (see~\cite{dealII92}).
-
-Finally, in the spirit of optimizing communication, the
-\texttt{Utilities::MPI::broadcast()} function has been optimized for
-objects that are arrays of data types natively supported by MPI and
-which consequently can be sent without packing and unpacking.
 
@@
   \item
-    \texttt{step-81} was contributed by Manaswinee Bezbaruah (Texas A\&M
-    University) and Matthias Maier (Texas A\&M University). It explains how
-    to solve the complex-valued time-harmonic Maxwell equations for an
-    optical scattering problem.
-  \item
-    \texttt{step-82} was contributed by Andrea Bonito (Texas A\&M
-    University) and Diane Guignard (University of Ottawa). It shows how
-    \dealii can be used to implement the local discontinuous Galerkin
-    (LDG) method for approximating the solution to the bi-Laplacian
-    problem. The method is an alternative to the C0IP method used in
-    \texttt{step-47}.
-  \item
-    \texttt{step-85} was contributed by Simon Sticko (Uppsala University).
-    It shows how to use the new CutFEM infrastructure to solve a Poisson
-    problem on a circular domain that is embedded in a Cartesian background
-    mesh.
 \end{itemize}
@@
 \begin{itemize}
-  \item ``TRBDF2-DG projection solver for the incompressible Navier--Stokes equations'' was contributed by Giuseppe Orlando (Politecnico di Milano). It shows
-  how to solve the incompressible Navier--Stokes equations efficiently 
-  with \dealii's matrix-free DG infrastructure, multigrid, and adaptive-mesh
-  refinement. Interested readers are referred to~\cite{orlando2021efficient}.
+  \item 
 \end{itemize}
-Finally, the ``MCMC for the Laplace equation'' code gallery
-program has been updated by providing MATLAB and Python versions of
-the benchmark that is implemented in this code.
+Finally,
 
@@
 
-The 9.4 release includes
-\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_3_3_and_9_4_0.html}
-{around 40 incompatible changes}; see \cite{changes94}. The majority of these changes
+The 9.5 release includes
+\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_4_2_and_9_5_0.html}
+{around X incompatible changes}; see \cite{changes95}. The majority of these changes
 should not be visible to typical user codes; some remove previously
@@
 \begin{itemize}
-  \item In continuation of our attempt to merge the classes \texttt{DoFHandler} and \texttt{hp::DoFHandler}, we have removed the
-  template parameter \texttt{DoFHandlerType} from a number of classes and
-  functions, instead using \texttt{dim}/\texttt{spacedim} template arguments. Affected
-  classes include \texttt{SolutionTransfer} and \texttt{DataOut}.
-  \item \texttt{FE\_RaviartThomasNodal} now uses a different polynomial space to allow
-  for a simpler use on faces in non-standard orientation. The new polynomials
-  are anisotropic tensor products of Lagrange polynomials on the points of the
-  Gauss--Lobatto quadrature formula. This change leads to different entries, for example, in
-  the matrices and constraints, but no change in accuracy should be expected as
-  the resulting basis spans the same polynomial space.
-  \item The class \texttt{MappingQ} now applies a high-order mapping
-    to all cells, not just the cells near the boundary, functionality
-    that was previously provided by the \texttt{MappingQGeneric}
-    class. The latter has been marked as deprecated.
-  \item Changes to weighted repartitioning of
-    \texttt{parallel::distributed::Triangulation} objects include
-  the removal of the default weight of each cell in order to ensure
-  consistency with the rest of the library and to improve flexibility.
+  \item 
 \end{itemize}
@@
 % This is up-to-date as of 9.4 RC1 - should be the final list.
-Pasquale      Africa,
-Tyler         Anderson,
-Francesco     Andreuzzi,
-Mathias       Anselmann,
-Maximilian    Bergbauer,
-Manaswinee    Bezbaruah,
-Bruno         Blais,
-Till          Budde,
-Fabian        Castelli,
-Praveen       Chandrashekar,
-Huimin        Chen,
-Cu            Cui,
-Ivo           Dravins,
-Niklas        Fehn,
-Corbin        Foucart,
-Johannes      Friedlein,
-Sebastian     Fuchs,
-Daniel        Garcia-Sanchez,
-Nicola        Giuliani,
-Alexander     Grayver,
-Diane         Guignard,
-Jake          Harmon,
-Sean          Ingimarson,
-Pengfei       Jia,
-Sebastian     Kinnewig,
-Uwe           K{\"o}cher,
-Katharina     Kormann,
-Paras         Kumar,
-Wenyu         Lei,
-Alberto F.    Martin,
-Nils          Much,
-Lucas         Myers,
-Justin        O'Connor,
-Judith        Pauen,
-Vachan        Potluri,
-Raghunandan   Pratoori,
-Sebastian     Proell,
-Ce            Qin,
-Reza          Rastak,
-Jose E.       Roman,
-Raphael       Schoof,
-Magdalena     Schreter,
-Konrad        Simon,
-Daniel        Sun,
-Kuljit S.     Virk,
-Michał        Wichrowski,
-Niklas        Wik,
-Jiaqi         Zhang.
+.
 

### 2023-05-04 9.5/paper.tex
Stub for a language modernization section.
Some updates.

@@
   \item Integration of \texttt{Kokkos} (see Section~\ref{sec:kokkos});
-  \item Update of the \texttt{PETSc} infrastructure (see Section~\ref{sec:petsc});
+  \item Update to and extension of the \texttt{PETSc} wrappers (see Section~\ref{sec:petsc});
   \item Addition of new \texttt{Trilinos} wrappers (see Section~\ref{sec:trilinos});
@@
   \item Advance in non-matching support (see Section~\ref{sec:nonmatching});
-  \item New liner algebra features (see Section~\ref{sec:lac}).
+  \item New liner algebra features (see Section~\ref{sec:lac})l
+  \item C++ language modernization (see Section~\ref{sec:language}).
 \end{itemize}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Update of the PETSc infrastructure}\label{sec:petsc}
+\subsection{Update of and additions to the PETSc wrappers}\label{sec:petsc}
 
@@
 \item PreconditionBDDC
+\item SNES
+\item TS
+\item Wrapping PETSc matrices and vectors
 \end{itemize}
@@
 \end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{C++ language modernization}\label{sec:language}
+
+Mention the C++20 work. Starting point: \dealii{} relies heavily on
+templates that are explicitly instantiated.
 

### 2023-05-22 9.5/paper.tex
Write text (Augsburg/Munich)

@@
 %\renewcommand{\baselinestretch}{2.0}	
+%\usepackage{lineno}
+%\renewcommand\linenumberfont{\normalfont\tiny}
+%\linenumbers
 
@@
    Universit\"atsstr.~12a, 86159 Augsburg, Germany.
-   {\texttt{martin.kronbichler@uni-a.de}}}
+   {\texttt{martin.kronbichler/peter.muench@uni-a.de}}}
 
@@
  Helmholtz-Zentrum Hereon,
- Max-Planck-Str. 1, 21502 Geesthacht, Germany.
-   {\texttt{peter.muench@hereon.de}}}
+ Max-Planck-Str. 1, 21502 Geesthacht, Germany}
 
@@
   \item Advances in matrix-free infrastructure (see Section~\ref{sec:mf});
-  \item Advance in non-matching support (see Section~\ref{sec:nonmatching});
-  \item New liner algebra features (see Section~\ref{sec:lac})l
+  \item Advances in non-matching support (see Section~\ref{sec:nonmatching});
+  \item New features related to liner algebra (see Section~\ref{sec:lac})
   \item C++ language modernization (see Section~\ref{sec:language}).
@@
 \begin{itemize}
-  \item The \texttt{CellAccessor::as\_dof\_handler\_iterator()}
+  \item The new function \texttt{CellAccessor::as\_dof\_handler\_iterator()}
+  simplifies the conversion from a \texttt{Cell\-Accessor} to a \texttt{DoF\-Cell\-Accessor}.
+  The old way 
+\begin{c++}
+const auto cell_dof = typename DoFHandler<dim, spacedim>::
+  active_cell_iterator(&dof_handler.get_triangulation(),
+    cell->level(), cell->index(), &dof_handler);
+\end{c++}
+was lengthy and error-prone. In contrast, the new way:
+\begin{c++}
+const auto cell_dof = cell->as_dof_handler_iterator(dof_handler);
+\end{c++}
 \end{itemize}
@@
 
-\begin{itemize}
-\item Belos
-\item NOX
-\end{itemize}
+We have added a new wrapper to \texttt{Belos}. The package
+\texttt{Belos}  is the
+successor of \texttt{AztecOO} and provides basic and advanced iterative solvers
+that heavily rely on multivector operations. The interface of the
+wrapper is similar to the one of the native iterative solvers
+provided by \dealii:
+
+\begin{c++}
+TrilinosWrappers::SolverBelos<VectorType> solver(/*...*/);
+solver.solve(matrix, x, r, preconditioner);
+\end{c++}
+
+Via additional data, the user can select the actual iterative
+solver to be used and set its configurations.
+
+Furthermore, we have added a wrapper to \texttt{NOX}. This is a
+nonlinear-solver library similar to \texttt{KINSOL} from \texttt{SUNDIALS}
+and to \texttt{SNES} from \texttt{PETSc}. The basic interface
+of the \texttt{NOX} wrapper is similar to the one of the other wrappers
+mentioned before:
+
+\begin{c++}
+TrilinosWrappers::NOXSolver<VectorType> solver(/*...*/);
+
+solver.residual            = [ ](const auto &src, auto &dst) {/*...*/};
+solver.setup_jacobian      = [&](const auto &src) {/*...*/};
+solver.apply_jacobian      = [&](const auto &src, auto &dst) {/*...*/};
+solver.solve_with_jacobian = 
+  [&](const auto &src, auto &dst, const auto tol) {/*...*/};
+
+solver.solve(solution);
+\end{c++}
+
+We refer interested readers to our documentation for other more advanced
+functions of this wrapper which are related to the differences in features of these libraries.
+In particular, we have added the possibility to reuse the preconditioner between
+nonlinear steps (also known as \textit{preconditioner lagging}), which is
+unfortunately natively only  supported in the official \texttt{EPetra} implementations.
 
@@
 
+In the matrix-free infrastructure, numerous advances have been made. Some of these
+are:
 \begin{itemize}
-\item element activation and deactivation, \texttt{FE\_Nothing}, categorization, DG, face integrals 
-\item pre/post for DG
-\item \texttt{CellwiseInverseMassMatrix} operator for coupling (dyadic) coefficients
+\item In release 9.3, we enabled parallel $hp$-operations in the matrix-free infrastructure.
+The infrastructure did not work properly in the case that certain cells
+did not get any degrees of freedom due to the usage of \texttt{FE\_Nothing}. This has been
+fixed now. Furthermore, \texttt{FE\_Nothing} now also works together with DG (\texttt{FE\_DGQ}). Due to the popularity of \texttt{FE\_Nothing} as a mean to enable
+or disable cells, we have introduced the new class
+\texttt{ElementActivationAndDeactivationMatrixFree}, which wraps a \texttt{MatrixFree} object, only loops over all
+active cells and optionally interprets faces between active and deactivated cells
+as boundary face. A use case is shown in~\cite{proell2023highly} in the context of powder-bed-fusion additive
+manufacturing.
+\item The matrix-free infrastructure allows to interleave cell loops with vector updates
+by providing \texttt{pre}/\texttt{post} functions that are run on index ranges. This
+feature is used, e.g., in \dealii to improve the performance of (preconditioned)
+conjugate gradient solvers~\cite{kronbichler2022cg} as well as of relaxation and Chebyshev iterations (see Subsection~\ref{sec:lac}).
+Up to release~9.3, the \texttt{pre}/\texttt{post} infrastructure was only supported for
+continuous elements (cell loop); now, it also works for discontinuous elements which require
+face loops additionally.
+\item The operator \texttt{CellwiseInverseMassMatrix} now also efficiently
+evaluates the inverse for coupling (dyadic) coefficients in the case of  multiple
+components:
+\begin{align*}
+\left(v_j, D_{ji} u_i  \right)_{\Omega^{(K)}}
+\quad  1\le i,j \le c,
+\end{align*}
+with $c$ being the number of components and $D\in \mathbb{R}^{c\times c}$ the tensorial
+coefficient.
+This is possible due to the tensor-product structure of the resulting element matrix:
+\begin{align*}
+M =  ( I_1 \otimes N^T) (D \otimes I_2 ) ( I_1 \otimes N),
+\end{align*}
+with $N$ being the tabulated shape functions and $I_1$/$I_2$ the appropriate identity matrices.
+For square and invertible $N$, the inverse is explicitly given as:
+\begin{align*}
+M^{-1} =  ( I_1 \otimes N^{-1}) (D^{-1} \otimes I_2 ) ( I_1 \otimes N^{-T}).
+\end{align*}
+Please note that $N^{-1} = N_{1D}^{-1} \otimes N_{1D}^{-1} \otimes N_{1D}^{-1}$ is given
+for hypercube-shaped cells, allowing to use sum factorization.
 \end{itemize}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Advance in non-matching support}\label{sec:nonmatching}
+\subsection{Advances in non-matching support}\label{sec:nonmatching}
+
+The (matrix-free) non-matching support of \texttt{deal.II} heavily relies
+on the classes \texttt{FE\-Point\-Eval\-u\-ation} and \texttt{RemotePointEvaluation},
+which have been introduced in release 9.3. While \texttt{FE\-Point\-Eval\-u\-a\-tion}
+is responsible for efficient evaluation/integration at arbitrary (reference)
+points within a cell, \texttt{RemotePointEvaluation} is responsible for
+sorting the (real) points within cells and for the successive communication.
+
+In the current release, we considerably optimized \texttt{FEPointEvaluation}, e.g,
+by caching the evaluated shape functions, templating loop bounds, and
+exploiting the tensor-product structure of the shape functions if all points are
+positioned on a face, a common use case in the context of fluid-structure
+interaction. Furthermore, the extended class \texttt{NonMatching::MappingInfo}
+allows to precompute and store metric terms, like the Jacobian, its determinant,
+or the normals. This is useful in cases in which these metric terms do not change
+and can be reused, e.g., in the context of iterative solvers. This development
+is part of the efforts to make the interfaces of the 
+(matrix-free) non-matching support more similar to the ones of the 
+established matrix-free infrastructure of \dealii for structured (quadrature) points.
+
+In addition to these node-level performance optimizations, we added experimental
+support 1) for generating intersections of distributed
+non-matching grids and working on them
+and 2) for multigrid with non-nested levels. In the following, we describe these
+features in detail.
+
+\subsubsection*{Intersected meshes}
+
+\begin{itemize}
+\item uses \texttt{CGAL}~\cite{cgal-user-ref}
+\item data can be used to fill \texttt{RemotePointEvaluation} (communication-free)
+\item used successfully in \cite{heinz2023high} to perform Nitsche-type mortaring in the context
+of the conservative formulation of acoustic equations discretized with
+DG, in order to suppress artificial modes
+\end{itemize}
+
+\subsubsection*{Non-nested multigrid}
+
+The library \dealii provides traditionally strong support for multigrid methods.
+We support, in the context of geometric multigrid, both local-smoothing~\cite{ClevengerHeisterKanschatKronbichler2019}
+and global-coarsening algorithms~\cite{munch2022gc} for locally refined meshes. The global-coarsening
+infrastructure, furthermore, allows to globally coarsen the polynomial degree ($p$-multigrid),
+whose applicability to $hp$-adaptive problems has been demonstrated. Algebraic
+multigrid is supported by interfacing to the external libraries \texttt{PETSc}
+and \texttt{Trilinos}. In the current release, we have added support for the case
+that multigrid levels are given by non-nested meshes~\cite{adams2002evaluation, bittencourt2001nonnested, bramble1991analysis}. An example for such meshes is
+presented in Figure~\ref{fig:nonnested}.
+
+\begin{figure}
+
+\centering
+
+\fbox{\begin{minipage}{0.31\textwidth}\centering 0\vspace{5cm}\end{minipage}}
+\fbox{\begin{minipage}{0.31\textwidth}\centering 1\vspace{5cm}\end{minipage}}
+\fbox{\begin{minipage}{0.31\textwidth}\centering 2\vspace{5cm}\end{minipage}}
+
+\caption{Example of non-nested multigrid levels.}\label{fig:nonnested}
+
+\end{figure}
+
+The current implementation extends the existing global-coarsening infrastructure by
+introducing, in addition to the (conformal) \texttt{MGTwoLevelTransfer},
+ a new (non-conformal) two-level transfer operator
+
+\begin{c++}
+MGTwoLevelTransferNonNested two_level_transfer(/*...*/);
+
+two_level_transfer.reinit(dof_handler_fine, dof_handler_coarse,
+                          // the following parameters are optional
+                          mapping_fine, mapping_coarse, 
+                          constraints_fine, constraints_coarse)
+\end{c++}
+
+which can be passed to \texttt{MGTransferGlobalCoarsening} (and
+\texttt{MGTransferBlockGlobalCoarsening}) in the usual way. Please note
+the similarity between the interfaces of \texttt{MGTwoLevelTransfer} and
+of \texttt{MGTwoLevel\-Transfer\-Non\-Nested}, with the difference that the latter also
+takes \texttt{Mapping} instances. Furthermore, it is not limited to the case
+that the \texttt{DoFHandler} instances need to share the same coarse triangulation
+and the coarser mesh has to be able to be generated by a coarsening step from the fine
+mesh.
+
+At the time of writing, the \texttt{MGTwoLevelTransferNonNested} operator performs
+a pointwise interpolation (injection) from the coarse mesh to the support points
+of the fine mesh. Adopting the notation of~\cite{munch2022gc}, we perform
+for prolongation
+\begin{align*}
+x^{(f)} = \mathcal{W}^{(f)} \circ \sum_{e \in \{\text{coarse cells}\}} \mathcal{S}_e^{(f)} \circ \mathcal{P}_e^{(f, c)}
+\circ \mathcal{C}_e^{(c)} \circ \mathcal{G}_e^{(c)} x^{(c)},
+\end{align*}
+i.e., loop over all coarse cells and interpolate to all (fine support) points that fall
+into the cell with \texttt{FEPointEvaluation}. The local result is scattered into a global
+vector, which is finalized by a communication and a weighting step.
+
+The current implementation works for scalar and vectorial continuous (\texttt{FE\_Q})
+and discontinuous elements (\texttt{FE\_DGQ}) for hypercube-shaped cells. We plan to support
+simplices and to add, in addition to the pointwise interpolation, also support
+for projection. Furthermore, we envision to provide utility tools to generate coarser
+meshes, based on a fine mesh. Currently, the users themselves 
+have to provide such meshes,
+e.g., by generating meshes with different cell sizes in external mesh-generation tools.
+
+Note that the new class \texttt{MGTwoLevelTransferNonNested} is not limited to
+multigrid but can be used to prolongate results and restrict residuals between any
+two meshes. Indeed, the infrastructure has been successfully applied to perform
+conservative interpolation between a fine (near) mesh and a coarse (far) mesh in the
+context of aeroacoustic problems.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{New linear algebra features}\label{sec:lac}
+\subsection{New features regarding linear algebra}\label{sec:lac}
+
+In addition to the new wrappers for linear-algebra functionalities of \texttt{PETSc}
+and \texttt{Trilinos} (see Sections~\ref{sec:petsc} and \ref{sec:trilinos}) as well
+as the new non-nested multigrid infrastructure, 
+we made multiple additions to our own linear-algebra infrastructure.
 
 \begin{itemize}
-  \item \texttt{SolverGMRES}: classical/modified Gram Schmidt process
-  \item James Lottes’s fourth-kind Chebyshev
-  \item \texttt{PreconditionChebyshev}, \texttt{PreconditionRelaxation}
-  \item restrict matrices
-  \item fast diagonalization method
+  \item Both \texttt{SolverGMRES} and \texttt{SolverFGMRES} now support the classical
+  Gram--Schmidt orthonormalization in addition to the existing modified one. This
+  allows to reduce the cost of vector operations in terms of
+  communication latency and memory transfer significantly.
+  \item Our Chebyshev preconditioner (\texttt{PreconditionChebyshev}) now also
+  supports James Lottes’s novel fourth-kind Chebyshev
+  polynomial~\cite{lottes2022optimal, phillips2022optimal}.
+  \item Our relaxation preconditioner (\texttt{PreconditionRelaxation}) now also
+  allows to interleave cell loops and vector updates related to relaxation. The
+  relaxation iteration reads as
+  \begin{align*}
+  x^{(i+1)} \gets x^{(i)} + \omega P^{-1}(b-Ax^{(i)}).
+  \end{align*}
+  E.g, in the case that the preconditioner $P$ is a diagonal matrix, the zeroing of
+  the destination vector $x_{i+1}$ can be performed during a \texttt{pre}-operation  of $A$
+  and the vector update $x^{(i+1)}_j \gets x^{(i)}_j + \omega P^{-1}_{j,j}(b_j-(Ax^{(i)})_j)$
+  during a \texttt{post} operation, allowing to reduce the number of read and write
+  accesses from 8 and 4 to 1 and 3, respectively. The existing \texttt{pre}/\texttt{post}
+  support in our Chebyshev-preconditioner implementation (\texttt{PreconditionChebyshev})
+  has been improved. In addition, both \texttt{PreconditionRelaxation} and 
+  \texttt{PreconditionChebyshev} support \texttt{pre}/\texttt{post} optimizations
+  now not only for diagonal preconditioners but also for preconditioners that are
+  built around cell loops and, as a consequence, support interleaving. An example of
+  such a preconditioner are patch-based additive Schwarz preconditioners.
+  \item In the context of additive Schwarz methods, the preconditioner application
+  is defined as
+  \begin{align*}
+  v = P^{-1} u = \sum R_i^\top A_i^{-1} R_i u 
+  \end{align*}
+  with $A_i = R_i A R_i^T$ being a block of the assembled system matrix restricted
+  to an index set (inverse of the assembly step). During the restriction step,
+  rows of the system matrix that are potentially owned by other
+  processes are needed. In \texttt{deal.II}, it is not possible to access remote entries
+  of sparse matrices. Two novel functions allow to query this information. The
+  function \texttt{restrict\_to\_serial\_sparse\_matrix()} creates, based
+  on a given index set, a serial
+  sparse matrix from a distributed one on each process:
+
+\begin{c++}
+SparseMatrixTools::restrict_to_serial_sparse_matrix (
+  sparse_matrix_in, sparsity_pattern, requested_is, 
+  system_matrix_out, sparsity_pattern_out)
+\end{c++}
+
+This function can be used, e.g., if the granularity of the additive Schwarz preconditioner
+is a complete subdomain, potentially, with a fixed overlap.
+
+In contrast, \texttt{restrict\_to\_full\_matrices()} performs the restriction
+for arbitrary number of patches/blocks:
+
+\begin{c++}
+SparseMatrixTools::restrict_to_full_matrices (
+  sparse_matrix_in, sparsity_pattern, indices_of_blocks, blocks)
+\end{c++}
+
+The data is stored in full matrices, since the typical granularity is a (rather small) cell-centric or
+vertex-star patch.
+
+  \item For certain types of configurations, there are computationally more efficient
+  approaches than extracting submatrices from an assembled matrix. For example, the Laplace
+  operator on 2D Cartesian meshes, the (element/patch) matrix is given as
+  \begin{align*}
+  A_i^{\text{cart}} = K_1 \otimes M_0 + M_1 \otimes K_0,
+  \end{align*}
+  i.e., as the tensor product of 1D mass and stiffness matrices. The inverse is
+  explicitly given according to the fast diagnilization method~\cite{lynch1964direct}, as
+  \begin{align*}
+  \left(A_i^{\text{cart}}\right)^{-1} = (T_1 \otimes T_0) (\Lambda_1 \otimes I + I \otimes \Lambda_0)^{-1} (T_1^\top \otimes T_0^\top),
+  \end{align*}
+  with $T_i$ and $\Lambda_i$, being the (orthonormal) eigenvectors and the diagonal
+  matrix of eigenvalues, obtained from a generalized eigendecomposition
+  $K_iT_i = \Lambda_i M_i T_i$. Since $A_i \approx A_i^{\text{cart}}$
+  might be a good approximation also in the case of non-Cartesian meshes and 
+  $A_i^{\text{cart}}$ has an explicit inverse, it is considered in the
+  literature as patch preconditioners in the context of additive
+  Schwarz~\cite{witte2021fast, phillips2021auto, couzy1995spectral} and block-Jacobi methods~\cite{kronbichler2019hermite}.
+  In \dealii, the new function
+  \texttt{Tensor\-Product\-Matrix\-Creator::create\_\allowbreak laplace\_\allowbreak tensor\_\allowbreak product\_\allowbreak matrix()} computes $T_i$ and $\Lambda_i$
+  for cell-centric patches with a specified overlap and given boundary conditions.
+  A set of $T_i$ and $\Lambda_i$ is applied to a cell via
+  \texttt{Tensor\-Product\-Matrix\-Symmetric\-Sum} or to a collection of cells
+  via the new class \texttt{Tensor\-Product\-Matrix\-Symmetric\-Sum\-Collection}, which
+  tries to reuse the eigenvalues and eigenvectors between cells.
 \end{itemize}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Build-system modernization}\label{sec:buildsystem}
+
+?
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{New and improved tutorials and code gallery programs}
@@
 Many of the \dealii tutorial programs were revised in a variety of ways
-as part of this release. In addition, there are a number of new tutorial
-programs:
-\begin{itemize}
-  \item
-\end{itemize}
+as part of this release. 
+%In addition, there are a number of new tutorial
+%programs:
+%\begin{itemize}
+%  \item
+%\end{itemize}
 
-There is also a new program in the code gallery (a collection of
+There is also two new programs in the code gallery (a collection of
 user-contributed programs that often solve more complicated problems
@@
 \begin{itemize}
-  \item 
+  \item The program \texttt{A posteriori error estimator for first order hyperbolic problems}
+  was contributed by Marco Feder.
+  \item The program \texttt{Distributed moving laser heating} was contributed by
+  Hongfeng Ma and Tatiana E. Itina.
 \end{itemize}
-Finally,
 
@@
 
-R.~Gassm{\"o}ller was also partially supported by the NSF Awards
-EAR-1925677 and EAR-2054605.
-
 L.~Heltai was partially supported by the Italian Ministry of Instruction,
@@
 
-S.~Sticko was partially supported by eSSENCE of e-Science.
-
 D.~Wells was supported by the NSF Award OAC-1931516.

### 2023-05-22 9.5/paper.tex
Minor changes to the document structure.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Integration of Kokkos}\label{sec:kokkos}
+\subsection{Updates to interfaces to other packages}\label{sec:kokkos}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Update of and additions to the PETSc wrappers}\label{sec:petsc}
+\subsubsection{Integration of Kokkos}\label{sec:kokkos}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Updates and additions to the PETSc wrappers}\label{sec:petsc}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{New Trilinos wrappers}\label{sec:trilinos}
+\subsubsection{Interfaces to Trilinos' Belos and NOX packages}\label{sec:trilinos}
 
@@
 unfortunately natively only  supported in the official \texttt{EPetra} implementations.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Uniform structure of callbacks}\label{sec:callbacks}
 

### 2023-05-22 9.5/paper.tex
Update the Trilinos section.

@@
 
-We have added a new wrapper to \texttt{Belos}. The package
+Trilinos is a large collection of individual sub-packages. \dealii{}
+has long had interfaces to the Trilinos packages that provide parallel
+vector and matrix classes (both \texttt{Epetra} and \texttt{Tpetra}),
+as well as a small number of linear algebra packages for iterative
+solvers and preconditions. In the current release, there are now also
+interfaces to two additional packages: \texttt{Belos} and \texttt{NOX}.
+
 \texttt{Belos} is the
-successor of \texttt{AztecOO} and provides basic and advanced iterative solvers
+successor of the Trilinos package \texttt{AztecOO} and provides basic and advanced iterative solvers
 that heavily rely on multivector operations. The interface of the
-wrapper is similar to the one of the native iterative solvers
-provided by \dealii:
-
+wrapper is similar to \dealii{}'s own iterative solvers:
 \begin{c++}
@@
 \end{c++}
+The omitted constructor arguments allow selecting the actual iterative
+solver to be used, along with other configuration options.
 
-Via additional data, the user can select the actual iterative
-solver to be used and set its configurations.
-
-Furthermore, we have added a wrapper to \texttt{NOX}. This is a
-nonlinear-solver library similar to \texttt{KINSOL} from \texttt{SUNDIALS}
-and to \texttt{SNES} from \texttt{PETSc}. The basic interface
-of the \texttt{NOX} wrapper is similar to the one of the other wrappers
-mentioned before:
-
+Secondly, we have added a wrapper to \texttt{NOX}, a
+nonlinear solver library similar to the \texttt{KINSOL} solver from
+the \texttt{SUNDIALS} package to which we already had interfaces, as
+well as to \texttt{SNES} from \texttt{PETSc} (see also
+Section~\ref{sec:petsc}). The interfaces to these three solvers are
+essentially the same and require setting function objects for
+computing the residual of the nonlinear problem, along with setting
+up, applying, and solving with the Jacobian matrix:
 \begin{c++}
@@
 
-solver.residual            = [ ](const auto &src, auto &dst) {/*...*/};
-solver.setup_jacobian      = [&](const auto &src) {/*...*/};
-solver.apply_jacobian      = [&](const auto &src, auto &dst) {/*...*/};
-solver.solve_with_jacobian = 
-  [&](const auto &src, auto &dst, const auto tol) {/*...*/};
+solver.residual            = [ ](const VectorType &src,
+                                 VectorType       &dst) {/*...*/};
+solver.setup_jacobian      = [&](const VectorType &src) {/*...*/};
+solver.apply_jacobian      = [&](const VectorType &src,
+                                 VectorType       &dst) {/*...*/};
+solver.solve_with_jacobian = [&](const VectorType &src,
+                                 VectorType       &dst,
+                                 const double      tol) {/*...*/};
 
@@
 unfortunately natively only  supported in the official \texttt{EPetra} implementations.
+
+\todo[inline]{I don't understand this last sentence. Can you
+  elaborate? At least in the KINSOL interfaces, you only update the
+  preconditioner whenever setup\_jacobian() is called.}
 

### 2023-05-22 9.5/paper.tex
Talk about the callback thing.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Uniform structure of callbacks}\label{sec:callbacks}
+\subsubsection{Uniform error reporting in callbacks}\label{sec:callbacks}
+
+With the current release, \dealii{} has gained interfaces to several
+external packages that are largely driven via \textit{callbacks} --
+see for example the code example in Section~\ref{sec:trilinos} on how
+NOX gains information about the residual vector, how it indicates to
+user code that the Jacobian matrix needs to be rebuilt,
+and how to solve linearized systems of equations as the sub-steps of
+solving a nonlinear problem. Indeed, the interfaces to the nonlinear
+solver KINSOL (as part of SUNDIALS) and SNES (as part of PETSc, see
+Section~\ref{sec:petsc} function in exactly the same way; the
+ODE solver interfaces to ARKode and IDA (also part of SUNDIALS) and TS
+(as part of PETSc) also use this style.
+
+This substantially enlarged use of callbacks used by different backend
+libraries raises the issue that each underlying package has its own
+convention on how success or error codes of these callbacks should be
+encoded. In the case of PETSc's SNES and TS, and for Trilinos's NOX,
+success is indicated by a zero integer return value, whereas failure is
+indicated by a nonzero return value. On the other hand, the SUNDIALS
+packages indicate success by a zero integer return value, a
+recoverable failure with a positive value, and an irrecoverable failure
+with a negative value. Neither of these conventions mesh well with C++
+where error codes are generally indicated via exceptions.
+It is conceivable that libraries we want to
+interface with in the future use yet other conventions.
+
+In order to shield those who write these callbacks from having to
+learn the intricacies of the underlying libraries, we have adopted a
+convention whereby user-provided callbacks are just regular functions
+that return errors via exception as is common in C++. Internally, the
+interfaces to different underlying libraries then translate these
+exceptions into the appropriate error codes, saving the thrown exception for possible
+later use; if an underlying library supports
+recoverable errors, then a callback should indicate such an error by
+throwing an object of the special type
+\texttt{RecoverableUserCallbackError}. If one of the underlying
+packages returns with an error caused by a user-provided callback
+throwing an exception, then the wrapper code re-throws the previously
+saved exception, allowing the calling site to catch it to obtain information
+about what might have gone wrong.
+
+This convention for user callbacks is documented in the glossary.
 

### 2023-05-23 9.5/paper.tex
Write the language standard section.

@@
 
-Mention the C++20 work. Starting point: \dealii{} relies heavily on
-templates that are explicitly instantiated.
+\dealii{} currently uses C++14 as the language standard to which its
+code base is written. It can also use the classes
+\texttt{std::optional} and \texttt{std::variant} if the compiler
+supports C++17, but falls back to implementations obtained via the
+BOOST library otherwise, and this is true also for a number of
+individual functions that were introduced in C++17 or C++20.
+
+As part of the current release, \dealii{} now also uses some C++20
+features to annotate classes and functions with regard to properties
+template arguments need to satisfy. This aids in situations such as
+with the
+following function:
+\begin{c++}
+  template <class MeshType>
+  std::vector<typename MeshType::active_cell_iterator>
+  find_cells_adjacent_to_vertex(const MeshType &   mesh,
+                                const unsigned int vertex);
+\end{c++}
+This function is intended to be called with either a
+\texttt{Triangulation} or \texttt{DoFHandler} object as first
+argument (and documents this requirement), but the compiler can not check this requirement at the call
+site and will gladly call it with any other kind of object as
+well. Because the implementation of the function is in a \texttt{.cc}
+file, and the template is only instantiated for the two classes
+mentioned above, calling the function (erroneously) with anything else
+as first argument is not detected at compile time, but only later when
+the linker reports an undefined symbol.
+
+We have started to address this by annotating functions using
+C++20-style ``\texttt{requires}'' clauses:
+\begin{c++}
+  template <class MeshType>
+    requires (concepts::is_triangulation_or_dof_handler<MeshType>)
+  std::vector<typename MeshType::active_cell_iterator>
+  find_cells_adjacent_to_vertex(const MeshType &   mesh,
+                                const unsigned int vertex)
+\end{c++}
+If the compiler supports C++20, then the added clause will cause the
+compiler to reject any call to the function for which the first
+argument \texttt{MeshType} does not satisfy the named concept, which
+is internally defined as the type being either a \texttt{Triangulation} or \texttt{DoFHandler}
+object, as intended. (The \texttt{requires} annotation is suppressed if the compiler
+does not support C++20.)
+
+Given the heavy dependence of \dealii{} on templates, there are likely
+hundreds or thousands of locations that should be annotated with
+requirements on template arguments over time; for the moment, the
+library contains some 300 of these \texttt{requires} clauses.
 

### 2023-05-23 9.5/paper.tex
Update the overview.

@@
 \begin{itemize}
-  \item Integration of \texttt{Kokkos} (see Section~\ref{sec:kokkos});
-  \item Update to and extension of the \texttt{PETSc} wrappers (see Section~\ref{sec:petsc});
-  \item Addition of new \texttt{Trilinos} wrappers (see Section~\ref{sec:trilinos});
+  \item Substantial updates and extensions to \dealii{}'s interfaces
+    to other libraries (see Section~\ref{sec:external}). This
+    includes, in particular, the integration of \texttt{Kokkos} (Section~\ref{sec:kokkos});
+    additions and updates to the \texttt{PETSc} and Trilinos
+    interfaces
+    (Sections~\ref{sec:petsc} and \ref{sec:trilinos}); and a uniform
+    way of defining callbacks used by external libraries
+    (Section~\ref{sec:callbacks}).
   \item Advances in matrix-free infrastructure (see Section~\ref{sec:mf});
@@
 \end{c++}
-was lengthy and error-prone. In contrast, the new way:
+was lengthy and error-prone. In contrast, the following function is
+substantially clearer and more concise:
 \begin{c++}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Updates to interfaces to other packages}\label{sec:kokkos}
+\subsection{Updates to interfaces to other packages}\label{sec:external}
 

### 2023-05-24 9.5/paper.tex
Minor text updates.

@@
 
+\todo[inline]{We are not consistent in whether or not we print PETSc,
+  SUNDIALS, Trilinos, SNES, ... with a TT font (texttt) or
+  not. Someone should go through the paper and make it consistent.}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \subsubsection{Updates and additions to the PETSc wrappers}\label{sec:petsc}
+
+\todo[inline]{Stefano Z.: Please write this section.}
 
@@
 Secondly, we have added a wrapper to \texttt{NOX}, a
-nonlinear solver library similar to the \texttt{KINSOL} solver from
-the \texttt{SUNDIALS} package to which we already had interfaces, as
-well as to \texttt{SNES} from \texttt{PETSc} (see also
+nonlinear solver library that is similar to both the \texttt{KINSOL} solver from
+the \texttt{SUNDIALS} package to which we already had interfaces, and also
+to the \texttt{SNES} collections of functions from \texttt{PETSc} (see also
 Section~\ref{sec:petsc}). The interfaces to these three solvers are
@@
 solving a nonlinear problem. Indeed, the interfaces to the nonlinear
-solver KINSOL (as part of SUNDIALS) and SNES (as part of PETSc, see
+solver KINSOL (as part of SUNDIALS) and SNES (as part of \texttt{PETSc}, see
 Section~\ref{sec:petsc} function in exactly the same way; the
@@
 convention on how success or error codes of these callbacks should be
-encoded. In the case of PETSc's SNES and TS, and for Trilinos's NOX,
+encoded. In the case of PETSc's SNES and TS, and for Trilinos's \texttt{NOX},
 success is indicated by a zero integer return value, whereas failure is
@@
 
-This convention for user callbacks is documented in the glossary.
+This convention for user callbacks is documented in a glossary entry
+that is also linked to from the documentation of all variables storing
+callbacks.
 

### 2023-05-25 9.5/paper.tex
Read through most of the text.

@@
 
+For many operations, \dealii{} relies on external libraries -- some of
+these are optional, others are mandatory (such as BOOST and, now,
+Kokkos); a complete list of external dependencies is provided in
+Section~\ref{sec:cite}. A substantial amount of work has gone into
+overhauling and extending these interfaces for the current release, as
+detailed in the following sub-sections.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Integration of Kokkos}\label{sec:kokkos}
+
+\todo[inline]{Those of you involved in the Kokkos effort, please see
+  if the following text is useful as a starting point, and complete
+  the section. You might also want to add references as appropriate.}
+
+Kokkos -- originally a part of Trilinos \cite{heroux2005trilinos,trilinos-web-page} but now an
+independent project -- has in recent years emerged as the most
+versatile and platform-independent library for interfacing CPU-based
+codes with the platform-dependent programming environments available
+for GPUs from a variety of vendors. As such, it makes it unnecessary
+to program to either NVidia's CUDA, AMD's HiP, or Intel's SYCL
+standards, and instead makes their functionality available using a
+common interface that provides abstractions for memory operations as
+well as kernels that act on this memory. It can also utilize CPU-based
+resources by dispatching operations via OpenMP directives.
+
+\dealii{} has, for several releases already, used CUDA to offload some
+operations onto GPUs. It has also had interfaces to CUDA-based linear
+algebra libraries. Yet, the diversification of GPU platforms away from
+a single vendor (NVidia) has made it clear that we need a different
+strategy to support what users want. As a consequence, Kokkos has
+become a mandatory dependency of \dealii{} as part of the current
+release; if it is not found on a given system during configuration
+time, then the library will fall back on a copy of Kokkos stored in
+the \texttt{bundled/} directory in the same way as we already
+interface with BOOST.
+
+\todo[inline]{Say what we use Kokkos for at this point, and perhaps
+  point out longer-term plans.}
 
@@
 
-Trilinos is a large collection of individual sub-packages. \dealii{}
+Trilinos is a large collection of individual sub-packages \cite{heroux2005trilinos,trilinos-web-page}. \dealii{}
 has long had interfaces to the Trilinos packages that provide parallel
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Advances in matrix-free infrastructure}\label{sec:mf}
+\subsection{Updates to matrix-free operations}\label{sec:mf}
 
-In the matrix-free infrastructure, numerous advances have been made. Some of these
-are:
+The current release includes numerous updates to the matrix-free
+infrastructure, including:
 \begin{itemize}
@@
 did not get any degrees of freedom due to the usage of \texttt{FE\_Nothing}. This has been
-fixed now. Furthermore, \texttt{FE\_Nothing} now also works together with DG (\texttt{FE\_DGQ}). Due to the popularity of \texttt{FE\_Nothing} as a mean to enable
+fixed now. Furthermore, \texttt{FE\_Nothing} now also works together
+with discontinuous Lagrange elements (i.e., with the \texttt{FE\_DGQ} class). Due to the popularity of \texttt{FE\_Nothing} as a mean to enable
 or disable cells, we have introduced the new class
@@
 active cells and optionally interprets faces between active and deactivated cells
-as boundary face. A use case is shown in~\cite{proell2023highly} in the context of powder-bed-fusion additive
+as boundary faces. A use case is shown in~\cite{proell2023highly} in the context of powder-bed-fusion additive
 manufacturing.
@@
 Up to release~9.3, the \texttt{pre}/\texttt{post} infrastructure was only supported for
-continuous elements (cell loop); now, it also works for discontinuous elements which require
-face loops additionally.
+continuous elements (cell loop); now, it also works for discontinuous
+elements which also require face loops to assemble jump and penalty terms.
 \item The operator \texttt{CellwiseInverseMassMatrix} now also efficiently
@@
 \begin{align*}
-\left(v_j, D_{ji} u_i  \right)_{\Omega^{(K)}}
+\left(v_i, D_{ij} u_j  \right)_{\Omega^{(K)}}
 \quad  1\le i,j \le c,
 \end{align*}
-with $c$ being the number of components and $D\in \mathbb{R}^{c\times c}$ the tensorial
+with $c$ being the number of components and $D\in \mathbb{R}^{c\times c}$ a tensorial
 coefficient.
@@
 \end{align*}
-with $N$ being the tabulated shape functions and $I_1$/$I_2$ the appropriate identity matrices.
+with $N$ being the tabulated shape functions and $I_1,I_2$ the appropriate identity matrices.
 For square and invertible $N$, the inverse is explicitly given as:
@@
 \end{align*}
-Please note that $N^{-1} = N_{1D}^{-1} \otimes N_{1D}^{-1} \otimes N_{1D}^{-1}$ is given
-for hypercube-shaped cells, allowing to use sum factorization.
+For hypercube-shaped cells, $N^{-1}$ has an explicit representation
+again in terms of tensor products; for example, in 3d it can be
+expressed as $N^{-1} = N_{1D}^{-1} \otimes N_{1D}^{-1} \otimes
+N_{1D}^{-1}$, allowing the use sum factorization.
 \end{itemize}
@@
 points within a cell, \texttt{RemotePointEvaluation} is responsible for
-sorting the (real) points within cells and for the successive communication.
+sorting points with regards to the cells they reside in, and for the
+communication necessary to evaluate solutions on cells owned by other
+MPI processes.
 
@@
 interaction. Furthermore, the extended class \texttt{NonMatching::MappingInfo}
-allows to precompute and store metric terms, like the Jacobian, its determinant,
+allows for precomputing and storing metric terms, like the Jacobian, its determinant,
 or the normals. This is useful in cases in which these metric terms do not change
 and can be reused, e.g., in the context of iterative solvers. This development
-is part of the efforts to make the interfaces of the 
+is part of an effort to make the interfaces of the 
 (matrix-free) non-matching support more similar to the ones of the 
@@
 In addition to these node-level performance optimizations, we added experimental
-support 1) for generating intersections of distributed
-non-matching grids and working on them
-and 2) for multigrid with non-nested levels. In the following, we describe these
-features in detail.
+support for (i) generating intersections of distributed
+non-matching grids and working on them,
+and (ii) multigrid with non-nested levels. In the following, we describe these
+in more detail.
 
-\subsubsection*{Intersected meshes}
+\subsubsection{Intersected meshes}
+
+\todo[inline]{Complete.}
 
@@
 
-\subsubsection*{Non-nested multigrid}
+\subsubsection{Non-nested multigrid}
 
-The library \dealii provides traditionally strong support for multigrid methods.
-We support, in the context of geometric multigrid, both local-smoothing~\cite{ClevengerHeisterKanschatKronbichler2019}
-and global-coarsening algorithms~\cite{munch2022gc} for locally refined meshes. The global-coarsening
+\dealii has provided support for geometric multigrid methods%
+\footnote{\dealii{} also supports algebraic multigrid methods
+by interfacing to the external libraries \texttt{PETSc}
+and \texttt{Trilinos}.}
+for locally refined meshes nearly since its inception.
+Traditionally, these were based on local-smoothing
+methods (described many years later in \cite{JanssenKanschat2011}, see also
+~\cite{ClevengerHeisterKanschatKronbichler2019}), and more
+recently also global-coarsening algorithms~\cite{munch2022gc}. The global-coarsening
 infrastructure, furthermore, allows to globally coarsen the polynomial degree ($p$-multigrid),
-whose applicability to $hp$-adaptive problems has been demonstrated. Algebraic
-multigrid is supported by interfacing to the external libraries \texttt{PETSc}
-and \texttt{Trilinos}. In the current release, we have added support for the case
+with specific applicability to $hp$-adaptive methods.
+In the current release, we have added support for the case
 that multigrid levels are given by non-nested meshes~\cite{adams2002evaluation, bittencourt2001nonnested, bramble1991analysis}. An example for such meshes is
@@
 
+\todo[inline]{Complete these figures.}
+
 \fbox{\begin{minipage}{0.31\textwidth}\centering 0\vspace{5cm}\end{minipage}}
@@
 
-\caption{Example of non-nested multigrid levels.}\label{fig:nonnested}
+\caption{\it Example of non-nested multigrid levels.}\label{fig:nonnested}
 
@@
 introducing, in addition to the (conformal) \texttt{MGTwoLevelTransfer},
- a new (non-conformal) two-level transfer operator
-
+ a new (non-conformal) two-level transfer operator:
 \begin{c++}
@@
 \end{c++}
-
-which can be passed to \texttt{MGTransferGlobalCoarsening} (and
-\texttt{MGTransferBlockGlobalCoarsening}) in the usual way. Please note
-the similarity between the interfaces of \texttt{MGTwoLevelTransfer} and
-of \texttt{MGTwoLevel\-Transfer\-Non\-Nested}, with the difference that the latter also
-takes \texttt{Mapping} instances. Furthermore, it is not limited to the case
-that the \texttt{DoFHandler} instances need to share the same coarse triangulation
-and the coarser mesh has to be able to be generated by a coarsening step from the fine
+The resulting object can be passed to \texttt{MGTransferGlobalCoarsening} (and
+\texttt{MGTransferBlockGlobalCoarsening}) in the usual way. The
+interface of the new class is purposefully similar to
+\texttt{MGTwoLevelTransfer},
+with the difference that the former (i) also
+takes \texttt{Mapping} instances, and (ii) is not limited to the case
+that the coarser of the two \texttt{DoFHandler} instances can be generated by a coarsening step from the fine
 mesh.
@@
 for prolongation
+\todo{I think you need to at least explain in a few words what each of
+  the components of this formula is.}
 \begin{align*}
@@
 
-The current implementation works for scalar and vectorial continuous (\texttt{FE\_Q})
-and discontinuous elements (\texttt{FE\_DGQ}) for hypercube-shaped cells. We plan to support
-simplices and to add, in addition to the pointwise interpolation, also support
-for projection. Furthermore, we envision to provide utility tools to generate coarser
+The current implementation works for scalar and vectorial, continuous (\texttt{FE\_Q})
+and discontinuous (\texttt{FE\_DGQ}) elements on hypercube-shaped cells. We plan to support
+simplex elements and to add support
+for projection, in addition to the pointwise interpolation outlined
+above. Furthermore, we envision providing utility tools to generate
+coarser 
 meshes, based on a fine mesh. Currently, the users themselves 
@@
 context of aeroacoustic problems.
+\todo{Can we add a reference for this?}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{New features regarding linear algebra}\label{sec:lac}
+\subsection{New linear algebra features}\label{sec:lac}
 
-In addition to the new wrappers for linear-algebra functionalities of \texttt{PETSc}
-and \texttt{Trilinos} (see Sections~\ref{sec:petsc} and \ref{sec:trilinos}) as well
-as the new non-nested multigrid infrastructure, 
-we made multiple additions to our own linear-algebra infrastructure.
-
+There are a number of linear-algebra related new features in this release:
 \begin{itemize}
@@
   Gram--Schmidt orthonormalization in addition to the existing modified one. This
-  allows to reduce the cost of vector operations in terms of
+  reduces the cost of vector operations in terms of
   communication latency and memory transfer significantly.
@@
   \item Our relaxation preconditioner (\texttt{PreconditionRelaxation}) now also
-  allows to interleave cell loops and vector updates related to relaxation. The
+  allows interleaving cell loops and vector updates related to relaxation. The
   relaxation iteration reads as
@@
   \end{align*}
-  E.g, in the case that the preconditioner $P$ is a diagonal matrix, the zeroing of
-  the destination vector $x_{i+1}$ can be performed during a \texttt{pre}-operation  of $A$
+  In the case that the preconditioner $P$ is a diagonal matrix, the zeroing of
+  the destination vector $x_{i+1}$ can be performed during a
+  \texttt{pre}-operation  of the application of $A$
   and the vector update $x^{(i+1)}_j \gets x^{(i)}_j + \omega P^{-1}_{j,j}(b_j-(Ax^{(i)})_j)$
-  during a \texttt{post} operation, allowing to reduce the number of read and write
-  accesses from 8 and 4 to 1 and 3, respectively. The existing \texttt{pre}/\texttt{post}
+  during a \texttt{post} operation, allowing for a reduction in the number of read and write
+  accesses from 8 and 4, to 1 and 3, respectively. The existing \texttt{pre}/\texttt{post}
   support in our Chebyshev-preconditioner implementation (\texttt{PreconditionChebyshev})
@@
   now not only for diagonal preconditioners but also for preconditioners that are
-  built around cell loops and, as a consequence, support interleaving. An example of
-  such a preconditioner are patch-based additive Schwarz preconditioners.
+  built around cell loops and, as a consequence, support interleaving. Examples of
+  such preconditioners are patch-based additive Schwarz preconditioners.
   \item In the context of additive Schwarz methods, the preconditioner application
@@
   \end{align*}
-  with $A_i = R_i A R_i^T$ being a block of the assembled system matrix restricted
-  to an index set (inverse of the assembly step). During the restriction step,
+  with $A_i = R_i A R_i^T$ being a block of the assembled system
+  matrix $A$ restricted
+  to an index set (inverse of the assembly step).
+\todo{I don't think this is true. The assembly step adds information
+  together. The restriction of the matrix cannot untangle these
+  different contributions. So the operation is similar, but not the
+  exact inverse.}
+  During the restriction step,
   rows of the system matrix that are potentially owned by other
   processes are needed. In \texttt{deal.II}, it is not possible to access remote entries
-  of sparse matrices. Two novel functions allow to query this information. The
-  function \texttt{restrict\_to\_serial\_sparse\_matrix()} creates, based
+  of sparse matrices. Two new functions query this information. First,
+  \texttt{restrict\_to\_serial\_sparse\_matrix()} creates, based
   on a given index set, a serial
   sparse matrix from a distributed one on each process:
-
 \begin{c++}
 SparseMatrixTools::restrict_to_serial_sparse_matrix (
-  sparse_matrix_in, sparsity_pattern, requested_is, 
+  sparse_matrix_in, sparsity_pattern, requested_index_set, 
   system_matrix_out, sparsity_pattern_out)
 \end{c++}
-
 This function can be used, e.g., if the granularity of the additive Schwarz preconditioner
@@
   \item For certain types of configurations, there are computationally more efficient
-  approaches than extracting submatrices from an assembled matrix. For example, the Laplace
+  approaches than extracting submatrices from an assembled matrix. For
+  example, for the Laplace
   operator on 2D Cartesian meshes, the (element/patch) matrix is given as
@@
   i.e., as the tensor product of 1D mass and stiffness matrices. The inverse is
-  explicitly given according to the fast diagnilization method~\cite{lynch1964direct}, as
+  explicitly available via the fast diagonalization method~\cite{lynch1964direct} as
   \begin{align*}
@@
   \end{align*}
-  with $T_i$ and $\Lambda_i$, being the (orthonormal) eigenvectors and the diagonal
+  with $T_i$ and $\Lambda_i$ being the (orthonormal) eigenvectors and the diagonal
   matrix of eigenvalues, obtained from a generalized eigendecomposition
@@
   $A_i^{\text{cart}}$ has an explicit inverse, it is considered in the
-  literature as patch preconditioners in the context of additive
+  literature as a patch preconditioner in the context of additive
   Schwarz~\cite{witte2021fast, phillips2021auto, couzy1995spectral} and block-Jacobi methods~\cite{kronbichler2019hermite}.
@@
 
-?
+\todo[inline]{Write}
 
@@
 
-Many of the \dealii tutorial programs were revised in a variety of ways
-as part of this release. 
-%In addition, there are a number of new tutorial
-%programs:
-%\begin{itemize}
-%  \item
-%\end{itemize}
+While there are no new \dealii tutorial programs in this release, many
+were extensively revised.
+\todo[inline]{See if we can say something more concrete.}
 
-There is also two new programs in the code gallery (a collection of
+There are also two new programs in the code gallery (a collection of
 user-contributed programs that often solve more complicated problems
@@
 \begin{itemize}
-  \item The program \texttt{A posteriori error estimator for first order hyperbolic problems}
+  \item The program \textit{``A posteriori error estimator for first order hyperbolic problems''}
   was contributed by Marco Feder.
-  \item The program \texttt{Distributed moving laser heating} was contributed by
-  Hongfeng Ma and Tatiana E. Itina.
+  \item The program \textit{``Distributed moving laser heating''} was contributed by
+  Hongfeng Ma and Tatiana E.~Itina.
 \end{itemize}
+
+\todo[inline]{There may be another one by Sam Scheuerman. Let's see
+  whether it still arrives in time.}
 
@@
   \item 
+\todo[inline]{Write}
 \end{itemize}
@@
     \item ROL \cite{ridzal2014rol}
+      \todo{ROL is part of Trilinos. Do we also need to cite something
+        for NOX?}
     \item ScaLAPACK \cite{blackford1997scalapack}
@@
 % This is up-to-date as of 9.4 RC1 - should be the final list.
-.
+
+\todo[inline]{Add.}
 
@@
 
+\todo[inline]{Need to update. Also remove those not on the current paper.}
+
 D.~Arndt and B.~Turcksin: Research sponsored by the Laboratory Directed Research and
@@
 
-W.~Bangerth, T.~Heister, and R.~Gassm{\"o}ller were partially
+W.~Bangerth and T.~Heister were partially
 supported by the Computational Infrastructure for Geodynamics initiative
@@
 No.~EAR-1550901 and The University of California -- Davis.
+\todo{This is the CIG-II grant. We need to update this to the CIG-III
+  (and CIG-IV?) grants.}
 
@@
 University has provided hosting services for the \dealii web page.
+\todo{Is this still true?}
 

### 2023-05-25 9.5/paper.tex
Fix grammar.

@@
 manufacturing.
-\item The matrix-free infrastructure allows to interleave cell loops with vector updates
+\item The matrix-free infrastructure allows interleaving cell loops with vector updates
 by providing \texttt{pre}/\texttt{post} functions that are run on index ranges. This
@@
 recently also global-coarsening algorithms~\cite{munch2022gc}. The global-coarsening
-infrastructure, furthermore, allows to globally coarsen the polynomial degree ($p$-multigrid),
+infrastructure, furthermore, allows globally coarsening the polynomial degree ($p$-multigrid),
 with specific applicability to $hp$-adaptive methods.

### 2023-05-25 9.5/paper.tex
Work on the Kokkos section

@@
  \author[1*]{Daniel Arndt}
- \affil[1]{Scalable Algorithms and Coupled Physics Group,
+ \affil[1]{Computational Coupled Physics Group,
    Computational Sciences and Engineering Division,
@@
   DE-AC05-00OR22725 with the U.S. Department of Energy.
-  %The United States
-  %Government retains and the publisher, by accepting the article for
-  %publication, acknowledges that the United States Government retains a
-  %non-exclusive, paid-up, irrevocable, worldwide license to publish or reproduce
-  %the published form of this manuscript, or allow others to do so, for United
-  %States Government purposes. The Department of Energy will provide public
-  %access to these results of federally sponsored research in accordance with the
-  %DOE Public Access Plan (http://energy.gov/downloads/doe-public-access-plan).
+  % We need to submit the manuscript with the text below. If the editor
+  % complains we can remove it. 
+  The United States
+  Government retains and the publisher, by accepting the article for
+  publication, acknowledges that the United States Government retains a
+  non-exclusive, paid-up, irrevocable, worldwide license to publish or reproduce
+  the published form of this manuscript, or allow others to do so, for United
+  States Government purposes. The Department of Energy will provide public
+  access to these results of federally sponsored research in accordance with the
+  DOE Public Access Plan (http://energy.gov/downloads/doe-public-access-plan).
 }
@@
 
-\todo[inline]{Those of you involved in the Kokkos effort, please see
-  if the following text is useful as a starting point, and complete
-  the section. You might also want to add references as appropriate.}
-
-Kokkos -- originally a part of Trilinos \cite{heroux2005trilinos,trilinos-web-page} but now an
-independent project -- has in recent years emerged as the most
-versatile and platform-independent library for interfacing CPU-based
-codes with the platform-dependent programming environments available
-for GPUs from a variety of vendors. As such, it makes it unnecessary
-to program to either NVidia's CUDA, AMD's HiP, or Intel's SYCL
-standards, and instead makes their functionality available using a
-common interface that provides abstractions for memory operations as
-well as kernels that act on this memory. It can also utilize CPU-based
-resources by dispatching operations via OpenMP directives.
+Kokkos \cite{trott2022} is a C++ library that enables the creation of
+performance portable applications for all major high-performance computing (HPC)
+platforms. It implements a programming model that allows developers to write
+code that can efficiently run on diverse architectures. Kokkos provides
+abstractions for both parallel execution of code and data management. It
+supports a wide range of backend programming models, including CUDA, HIP, SYCL,
+HPX, OpenMP, and C++ threads, and continues to evolve with the development of
+new backend options.
 
@@
 algebra libraries. Yet, the diversification of GPU platforms away from
-a single vendor (NVidia) has made it clear that we need a different
+a single vendor (Nvidia) has made it clear that we need a different
 strategy to support what users want. As a consequence, Kokkos has
@@
 
-\todo[inline]{Say what we use Kokkos for at this point, and perhaps
-  point out longer-term plans.}
+In the current release, LinearAlgebra::distributed::Vector and the
+CUDAWrappers::MatrixFree framework are using Kokkos. This allows them to work on
+all the architectures supported by Kokkos. The Kokkos backend used in deal.II is
+`Kokkos::DefaultExecutionSpace` which is the highest available in the hierarchy
+device, host-parallel, and host-serial.
 

### 2023-05-26 9.5/paper.tex
Unify typesetting for deal.II

@@
 CUDAWrappers::MatrixFree framework are using Kokkos. This allows them to work on
-all the architectures supported by Kokkos. The Kokkos backend used in deal.II is
+all the architectures supported by Kokkos. The Kokkos backend used \dealii is
 `Kokkos::DefaultExecutionSpace` which is the highest available in the hierarchy
@@
 
-The (matrix-free) non-matching support of \texttt{deal.II} heavily relies
+The (matrix-free) non-matching support of \dealii heavily relies
 on the classes \texttt{FE\-Point\-Eval\-u\-ation} and \texttt{RemotePointEvaluation},
@@
   rows of the system matrix that are potentially owned by other
-  processes are needed. In \texttt{deal.II}, it is not possible to access remote entries
+  processes are needed. In \dealii, it is not possible to access remote entries
   of sparse matrices. Two new functions query this information. First,

### 2023-05-27 9.5/paper.tex
More minor changes.

@@
 CUDAWrappers::MatrixFree framework are using Kokkos. This allows them to work on
-all the architectures supported by Kokkos. The Kokkos backend used \dealii is
-`Kokkos::DefaultExecutionSpace` which is the highest available in the hierarchy
+all the architectures supported by Kokkos. The Kokkos backend used by \dealii is
+\texttt{Kokkos::DefaultExecutionSpace} which is the highest available in the hierarchy
 device, host-parallel, and host-serial.
+
+\todo[inline]{Are you saying that all three of these are used? If so,
+  maybe replace ``backend'' by ``backends''?}
+
+\todo[inline]{Would it be useful to add a paragraph about future plans?}
 
@@
 
-There are also two new programs in the code gallery (a collection of
+There are also three new programs in the code gallery (a collection of
 user-contributed programs that often solve more complicated problems
@@
 \begin{itemize}
-  \item The program \textit{``A posteriori error estimator for first order hyperbolic problems''}
-  was contributed by Marco Feder.
-  \item The program \textit{``Distributed moving laser heating''} was contributed by
-  Hongfeng Ma and Tatiana E.~Itina.
+  \item \textit{``A posteriori error estimator for first order hyperbolic problems''},
+  contributed by Marco Feder;
+  \item \textit{``Distributed moving laser heating''}, contributed by
+  Hongfeng Ma and Tatiana E.~Itina;
+  \item \textit{``Generalized Swift-Hohenberg Equation Solver''}, contributed by
+  Sam Scheuerman.
 \end{itemize}
-
-\todo[inline]{There may be another one by Sam Scheuerman. Let's see
-  whether it still arrives in time.}
 

### 2023-05-27 9.5/paper.tex

@@
   \item
+  The class \texttt{hp::DoFHandler} has been removed. The standard \texttt{DoFHandler} now implements all hp-related functionality.
+
+  \item
+  We improved the type safety for active and future FE indices in the \texttt{DoFHandler} implementation by introducing a new datatype \texttt{types::fe\textunderscore{}index}.
+Corresponding functions like \texttt{DoFCellAccessor::active\textunderscore{}fe\textunderscore{}index()} and \texttt{DoFHandler::get\textunderscore{}active\textunderscore{}fe\textunderscore{}indices()} have changed their interface.
+
+  \item
+  This rework also affects the serialization process of active FE indices.
+You will need to recreate your serialized data \textit{if and only if} you work in hp-mode.
+You can continue to use previously generated data if you do not use hp-mode.
+A special case forms \texttt{parallel::distributed::Triangulation} for non-hp data, in which it is sufficient to increase the version in the metadata file by one from ``4'' to ``5''.
+
+  \item
+  ...
+\end{itemize}
 \todo[inline]{Write}
-\end{itemize}
 

### 2023-05-27 9.5/paper.tex
Fix typo.

@@
   \item
-  We improved the type safety for active and future FE indices in the \texttt{DoFHandler} implementation by introducing a new datatype \texttt{types::fe\textunderscore{}index}.
+  We improved the type safety for active and future FE indices in the \texttt{DoFHandler} implementation by introducing a new data type \texttt{types::fe\textunderscore{}index}.
 Corresponding functions like \texttt{DoFCellAccessor::active\textunderscore{}fe\textunderscore{}index()} and \texttt{DoFHandler::get\textunderscore{}active\textunderscore{}fe\textunderscore{}indices()} have changed their interface.

### 2023-06-05 9.5/paper.tex
Minor edits.

@@
 HPX, OpenMP, and C++ threads, and continues to evolve with the development of
-new backend options.
+new hardware and corresponding backend options.
 
@@
   \item
-  The class \texttt{hp::DoFHandler} has been removed. The standard \texttt{DoFHandler} now implements all hp-related functionality.
+  The class \texttt{hp::DoFHandler} has been removed. The
+  \texttt{DoFHandler} class now implements all $hp$-related
+  functionality, as it has for the past few releases.
 
@@
   This rework also affects the serialization process of active FE indices.
-You will need to recreate your serialized data \textit{if and only if} you work in hp-mode.
-You can continue to use previously generated data if you do not use hp-mode.
-A special case forms \texttt{parallel::distributed::Triangulation} for non-hp data, in which it is sufficient to increase the version in the metadata file by one from ``4'' to ``5''.
+You will need to recreate your serialized data \textit{if and only if} you work in $hp$-mode.
+You can continue to use previously generated data if you do not use $hp$-mode.
+A special case forms \texttt{parallel::distributed::Triangulation} for non-$hp$ data, in which it is sufficient to increase the version in the metadata file by one from ``4'' to ``5''.
 

### 2023-06-05 9.5/paper.tex
update funding

@@
 f\"ur Technisch-Wissen\-schaft\-li\-ches Hoch- und H\"ochstleistungsrechnen
-(KONWIHR) in the context of the projects
-``High-order matrix-free finite element implementations with
-hybrid parallelization and improved data locality'' and ``Fast and scalable finite element algorithms for coupled multiphysics problems and non-matching grids''.
+(KONWIHR) in the context of the projects ``High-order matrix-free finite
+element implementations with hybrid parallelization and improved data
+locality'' and ``Fast and scalable finite element algorithms for coupled
+multiphysics problems and non-matching grids''.
 
-M.~Maier was partially supported by NSF Awards DMS-1912847 and DMS-2045636.
+M.~Maier was partially supported by NSF Award DMS-2045636 and and by the
+Air Force Office of Scientific Research under grant/contract number
+FA9550-23-1-0007.
 
 D.~Wells was supported by the NSF Award OAC-1931516.
-
-The Interdisciplinary Center for Scientific Computing (IWR) at Heidelberg
-University has provided hosting services for the \dealii web page.
-\todo{Is this still true?}
 

### 2023-06-15 9.5/paper.tex
Add a couple of numbers.

@@
 While there are no new \dealii tutorial programs in this release, many
-were extensively revised.
-\todo[inline]{See if we can say something more concrete.}
+were extensively revised: Around 130 of the more than 2000 (non-merge)
+commits that went into this release touched the tutorial, in some
+cases adding substantial amounts of text.
+
+% data generated using these commands:
+% - tutorial: git log --since 2022/06/01 --no-merges examples | grep commit | wc -l
+% - total:    git log --since 2022/06/01 --no-merges | grep commit | wc -l
 

### 2023-06-15 9.5/paper.tex
Add bergbauer as author

@@
     Wolfgang Bangerth,
+    Maximilian Bergbauer,
     Marc Fehling,
@@
 
+\author[4]{Maximilian~Bergbauer}
+\affil[4]{Institute for Computational Mechanics,
+Technical University of Munich,
+Boltzmannstr.~15, 85748 Garching, Germany.
+{\texttt{maximilian.bergbauer@tum.de}}}
+
 \author[2]{Marc~Fehling}
 
-\author[4]{Timo~Heister}
- \affil[4]{School of Mathematical and Statistical Sciences,
+\author[5]{Timo~Heister}
+ \affil[5]{School of Mathematical and Statistical Sciences,
    Clemson University,
@@
 
- \author[6]{Martin~Kronbichler}
- \affil[6]{Institute of Mathematics,
+ \author[7]{Martin~Kronbichler}
+ \affil[7]{Institute of Mathematics,
    University of Augsburg,
@@
 
-\author[7]{Matthias~Maier}
-\affil[7]{Department of Mathematics,
+\author[8]{Matthias~Maier}
+\affil[8]{Department of Mathematics,
   Texas A\&M University,
@@
 
-\author[6,8]{Peter Munch}
- \affil[8]{Institute of Material Systems Modeling,
+\author[7,9]{Peter Munch}
+ \affil[9]{Institute of Material Systems Modeling,
  Helmholtz-Zentrum Hereon,
@@
 
-\author[9]{Jean-Paul~Pelteret}
-\affil[9]{Independent researcher.
+\author[10]{Jean-Paul~Pelteret}
+\affil[10]{Independent researcher.
 {\texttt{jppelteret@gmail.com}}}

### 2023-06-15 9.5/paper.tex
add heinz as author

@@
     Marc Fehling,
+    Johannes Heinz,
     Timo Heister,
@@
 
-\author[5]{Timo~Heister}
- \affil[5]{School of Mathematical and Statistical Sciences,
+\author[5]{Johannes~Heinz}
+ \affil[5]{Institute of Mechanics and Mechatronics,
+   TU Wien,
+   Getreidemarkt 9, 1060 Vienna, Austria
+   {\texttt{johannes.heinz@tuwien.ac.at}}}
+
+\author[6]{Timo~Heister}
+ \affil[6]{School of Mathematical and Statistical Sciences,
    Clemson University,
@@
 
- \author[7]{Martin~Kronbichler}
- \affil[7]{Institute of Mathematics,
+ \author[8]{Martin~Kronbichler}
+ \affil[8]{Institute of Mathematics,
    University of Augsburg,
@@
 
+M.~Bergbauer was supported by the German Research Foundation (DFG) 
+under the project ``High-Performance Cut Discontinuous Galerkin Methods
+for Flow Problems and Surface-Coupled Multiphysics Problems''.
+
+J.~Heinz was supported by the European Union’s Framework Programme for Research
+and Innovation Horizon 2020 (2014-2020) under the Marie Sk\l{}odowska--Curie Grant
+Agreement No. [812719].
+
 T.~Heister was also partially supported by NSF

### 2023-06-15 9.5/paper.tex
Add Feder as author

@@
     Marc Fehling,
+    Marco Feder,
     Johannes Heinz,
@@
 
-\author[5]{Johannes~Heinz}
- \affil[5]{Institute of Mechanics and Mechatronics,
+\author[5]{Marco~Feder}
+\affil[5]{SISSA,
+International School for Advanced Studies,
+Via Bonomea 265,
+34136, Trieste, Italy.
+{\texttt{marco.feder/luca.heltai@sissa.it}}}
+
+\author[6]{Johannes~Heinz}
+ \affil[6]{Institute of Mechanics and Mechatronics,
    TU Wien,
@@
 
-\author[6]{Timo~Heister}
- \affil[6]{School of Mathematical and Statistical Sciences,
+\author[7]{Timo~Heister}
+ \affil[7]{School of Mathematical and Statistical Sciences,
    Clemson University,
@@
 
-\author[7]{Luca~Heltai}
-\affil[7]{SISSA,
-   International School for Advanced Studies,
-   Via Bonomea 265,
-   34136, Trieste, Italy.
-   {\texttt{marco.feder/luca.heltai@sissa.it}}}
+\author[5]{Luca~Heltai}
 

### 2023-06-16 9.5/paper.tex
add text for Intersected meshes section

--- 2023-06-15 9.5/paper.tex
@@
 \subsubsection{Intersected meshes}
+In the current release, we added experimental support to compute intersections on parallel::distributed::Triangulation objects using \texttt{CGAL}~\cite{cgal-user-ref}.
+For this purpose we introduced a free function which computes intersections and relevant information for communication from \texttt{intersection\_requests}.
+\texttt{intersection\_requests} is a vector that contains entities for which intersections are computed on a given triangulation (in the form of \texttt{GridTools::Cache}).
+Each entity (faces or cells) is described by a vector of vertices.
 
-\todo[inline]{Complete.}
+\begin{c++}
+// compute intersections on distributed triangulation  
+auto intersection_data =
+       GridTools::internal::distributed_compute_intersection_locations(
+         cache, intersection_requests, global_bboxes, marked_vertices,
+         tolerance);
+\end{c++}
 
-\begin{itemize}
-\item uses \texttt{CGAL}~\cite{cgal-user-ref}
-\item data can be used to fill \texttt{RemotePointEvaluation} (communication-free)
-\item used successfully in \cite{heinz2023high} to perform Nitsche-type mortaring in the context
-of the conservative formulation of acoustic equations discretized with
-DG, in order to suppress artificial modes
-\end{itemize}
+For the common case of Nitsche-type mortaring, quadrature points must be distributed on the intersections where relevant quantities are evaluated.
+The \texttt{intersection\_data} returned by the function above can convert itself to data which can be used to fill \texttt{RemotePointEvaluation}.
+The whole procedure is done communication-free.
 
+\begin{c++}
+// distribute n_points_1D quadrature points to the intersections
+auto point_data = intersection_data.template 
+       convert_to_distributed_compute_point_locations_internal<dim>(
+         n_points_1D, tria, mapping,
+         // following parameter is optional would need additional
+         // communication 
+         consistent_numbering_of_sender_and_receiver);
+
+// use point_data to reinit RemotePointEvaluation
+Utilities::MPI::RemotePointEvaluation<dim> rpe;
+rpe.reinit(point_data,tria, mapping);
+\end{c++}
+
+This functionality is handy since \texttt{RemotePointEvaluation} can now be used to access quantities in quadrature points on intersections without further ado.
+To reduce the user's effort, we plan to add a wrapper that takes care of the described procedure and provides interfaces like \texttt{FEEvaluation} to access quantities easily, e.g., in a matrix-free loop.
+
+Described implementation and an early version of the wrapper have been used successfully in~\cite{heinz2023high} to perform Nitsche-type mortaring in the context of the conservative formulation of acoustic equations discretized with DG to suppress artificial modes.
 \subsubsection{Non-nested multigrid}
@@
 
+M.~Bergbauer was supported by the German Research Foundation (DFG) 
+under the project ``High-Performance Cut Discontinuous Galerkin Methods
+for Flow Problems and Surface-Coupled Multiphysics Problems''.
+
 T.~Heister was also partially supported by NSF

### 2023-06-16 9.5/paper.tex

@@
     Marc Fehling,
+    Marco Feder,
+    Johannes Heinz,
     Timo Heister,
@@
 
-\author[5]{Timo~Heister}
- \affil[5]{School of Mathematical and Statistical Sciences,
-   Clemson University,
-   Clemson, SC, 29634, USA
-   {\texttt{heister@clemson.edu}}}
-
-\author[6]{Luca~Heltai}
-\affil[6]{SISSA,
+\author[5]{Marco~Feder}
+\affil[5]{SISSA,
 International School for Advanced Studies,
@@
 
- \author[7]{Martin~Kronbichler}
- \affil[7]{Institute of Mathematics,
+\author[6]{Johannes~Heinz}
+ \affil[6]{Institute of Mechanics and Mechatronics,
+   TU Wien,
+   Getreidemarkt 9, 1060 Vienna, Austria
+   {\texttt{johannes.heinz@tuwien.ac.at}}}
+
+\author[7]{Timo~Heister}
+ \affil[7]{School of Mathematical and Statistical Sciences,
+   Clemson University,
+   Clemson, SC, 29634, USA
+   {\texttt{heister@clemson.edu}}}
+
+\author[5]{Luca~Heltai}
+
+ \author[8]{Martin~Kronbichler}
+ \affil[8]{Institute of Mathematics,
    University of Augsburg,
@@
 
+J.~Heinz was supported by the European Union’s Framework Programme for Research
+and Innovation Horizon 2020 (2014-2020) under the Marie Sk\l{}odowska--Curie Grant
+Agreement No. [812719].
+
 T.~Heister was also partially supported by NSF

### 2023-06-16 9.5/paper.tex
Update text, references, and acknowledgements

@@
 \end{c++}
+\item Several functions used intensively during initialization of
+  \dealii-based programs, such as the refinement of triangulations, the
+  enumeration of degrees of freedom, the setup of global-coarsening multigrid
+  algorithms, and several evaluation functions of the \texttt{MappingQ} class
+  representing a polynomial mapping of quadrilateral and hexahedral, have been
+  overhauled to run more quickly and sometimes also consume less memory. These
+  and related improvements are guided by several performance tests that are
+  used to monitor the performance of the library over time.
 \end{itemize}
@@
 convention whereby user-provided callbacks are just regular functions
-that return errors via exception as is common in C++. Internally, the
+that return errors via exceptions as is common in C++. Internally, the
 interfaces to different underlying libraries then translate these
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Updates to matrix-free operations}\label{sec:mf}
+\subsection{Updates to matrix-free algorithms}\label{sec:mf}
 
@@
 active cells and optionally interprets faces between active and deactivated cells
-as boundary faces. A use case is shown in~\cite{proell2023highly} in the context of powder-bed-fusion additive
-manufacturing.
+as boundary faces. This functionality has enabled simulations in powder-bed-fusion additive
+manufacturing in~\cite{proell2023highly}.
 \item The matrix-free infrastructure allows interleaving cell loops with vector updates
-by providing \texttt{pre}/\texttt{post} functions that are run on index ranges. This
-feature is used, e.g., in \dealii to improve the performance of (preconditioned)
+  by providing \texttt{pre}/\texttt{post} functions that are run on index ranges.  The \dealii
+  library uses this feature, e.g., to improve the performance of (preconditioned)
 conjugate gradient solvers~\cite{kronbichler2022cg} as well as of relaxation and Chebyshev iterations (see Subsection~\ref{sec:lac}).
@@
 with $c$ being the number of components and $D\in \mathbb{R}^{c\times c}$ a tensorial
-coefficient.
-This is possible due to the tensor-product structure of the resulting element matrix:
+coefficient. The algorithm relies on the construction of the element mass matrix,
 \begin{align*}
@@
 \end{align*}
-with $N$ being the tabulated shape functions and $I_1,I_2$ the appropriate identity matrices.
-For square and invertible $N$, the inverse is explicitly given as:
+with $N$ being the tabulated values of shape functions at quadrature and
+$I_1,I_2$ identity matrices associated to $c$ vector components and the
+quadrature points, respectively.
+The algorithms assumes a square $N$:
 \begin{align*}
@@
 expressed as $N^{-1} = N_{1D}^{-1} \otimes N_{1D}^{-1} \otimes
-N_{1D}^{-1}$, allowing the use sum factorization.
+N_{1D}^{-1}$, allowing the use sum factorization \cite{kronbichler2016comparison}.
 \end{itemize}
@@
 
-In the current release, we considerably optimized \texttt{FEPointEvaluation}, e.g,
+In the current release, we have considerably optimized \texttt{FEPointEvaluation}, e.g,
 by caching the evaluated shape functions, templating loop bounds, and
@@
 allows for precomputing and storing metric terms, like the Jacobian, its determinant,
-or the normals. This is useful in cases in which these metric terms do not change
+or the unit outer normal vectors. This is useful in cases in which these metric terms do not change
 and can be reused, e.g., in the context of iterative solvers. This development
@@
 (matrix-free) non-matching support more similar to the ones of the 
-established matrix-free infrastructure of \dealii for structured (quadrature) points.
+established matrix-free infrastructure of \dealii for fixed quadrature formulas.
 
@@
 In the current release, we added experimental support to compute intersections on parallel::distributed::Triangulation objects using \texttt{CGAL}~\cite{cgal-user-ref}.
-For this purpose we introduced a free function which computes intersections and relevant information for communication from \texttt{intersection\_requests}.
-\texttt{intersection\_requests} is a vector that contains entities for which intersections are computed on a given triangulation (in the form of \texttt{GridTools::Cache}).
-Each entity (faces or cells) is described by a vector of vertices.
+For this purpose we introduced a free function, which computes intersections and relevant information for communication from \texttt{intersection\_requests}.
+\texttt{intersection\_requests} is a vector indicating entities of a given triangulation (in the form of \texttt{GridTools::Cache}) that intersections are computed upon.
+Each entity (face or cell) is described by a vector of vertices.
 
@@
 
-For the common case of Nitsche-type mortaring, quadrature points must be distributed on the intersections where relevant quantities are evaluated.
+For the common case of Nitsche-type mortaring, quadrature points must be distributed on the intersections to evaluate the underlying physical coupling terms.
 The \texttt{intersection\_data} returned by the function above can convert itself to data which can be used to fill \texttt{RemotePointEvaluation}.
@@
 
-This functionality is handy since \texttt{RemotePointEvaluation} can now be used to access quantities in quadrature points on intersections without further ado.
+This functionality is handy since \texttt{RemotePointEvaluation} can now be used to access quantities at quadrature points on intersections without further ado.
 To reduce the user's effort, we plan to add a wrapper that takes care of the described procedure and provides interfaces like \texttt{FEEvaluation} to access quantities easily, e.g., in a matrix-free loop.
@@
 Traditionally, these were based on local-smoothing
-methods (described many years later in \cite{JanssenKanschat2011}, see also
-~\cite{ClevengerHeisterKanschatKronbichler2019}), and more
+methods (described many years later in \cite{Kanschat2004,JanssenKanschat2011}, see
+also~\cite{ClevengerHeisterKanschatKronbichler2019}), and more
 recently also global-coarsening algorithms~\cite{munch2022gc}. The global-coarsening
@@
   the components of this formula is.}
+\todo{Alternatively (MK's preference), skip the parts that are just unnecessary
+  abstractions that do not help the reader's understanding.}
 \begin{align*}
@@
   \item Both \texttt{SolverGMRES} and \texttt{SolverFGMRES} now support the classical
-  Gram--Schmidt orthonormalization in addition to the existing modified one. This
+  Gram--Schmidt orthonormalization in addition to the modified Gram--Schmidt algorithm. This
   reduces the cost of vector operations in terms of
@@
   matrix $A$ restricted
-  to an index set (inverse of the assembly step).
-\todo{I don't think this is true. The assembly step adds information
-  together. The restriction of the matrix cannot untangle these
-  different contributions. So the operation is similar, but not the
-  exact inverse.}
+  to an index set described by $R_i$. In the special case when $R_i$ denotes the unknowns
+  of cells, the expression $R_i A R_i^T$ resembles the reverse of matrix
+  assembly.
   During the restriction step,
@@
   on a given index set, a serial
-  sparse matrix from a distributed one on each process:
+  sparse matrix from a distributed matrix:
 \begin{c++}
@@
 
-The data is stored in full matrices, since the typical granularity is a (rather small) cell-centric or
+The data is stored in dense matrices, since the typical granularity is a (rather small) cell-centric or
 vertex-star patch.
@@
   matrix of eigenvalues, obtained from a generalized eigendecomposition
-  $K_iT_i = \Lambda_i M_i T_i$. Since $A_i \approx A_i^{\text{cart}}$
+  $K_iT_i =  M_i T_i \Lambda_i$, as also used by \dealii's \texttt{step-59} tutorial program.
+  Since $A_i \approx A_i^{\text{cart}}$
   might be a good approximation also in the case of non-Cartesian meshes and 
@@
   In \dealii, the new function
-  \texttt{Tensor\-Product\-Matrix\-Creator::create\_\allowbreak laplace\_\allowbreak tensor\_\allowbreak product\_\allowbreak matrix()} computes $T_i$ and $\Lambda_i$
+  \texttt{Tensor\-Product\-Matrix\-Creator::create\_\allowbreak laplace\_\allowbreak tensor\_\allowbreak product\_\allowbreak matrix()} computes $K_i$ and $M_i$
   for cell-centric patches with a specified overlap and given boundary conditions.
@@
 \href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_4_2_and_9_5_0.html}
-{around X incompatible changes}; see \cite{changes95}. The majority of these changes
-should not be visible to typical user codes; some remove previously
-deprecated classes and functions; and the majority changes internal
+{around X incompatible changes}; see \cite{changes95}. Many of the
+incompatible change internal
 interfaces that are not usually used in external
@@
 A special case forms \texttt{parallel::distributed::Triangulation} for non-$hp$ data, in which it is sufficient to increase the version in the metadata file by one from ``4'' to ``5''.
+
+\item Several old interfaces to \texttt{MatrixFree} have been removed, e.g.,
+  initialization functions without \texttt{Mapping} argument, the query
+  to the number of cell batches, \texttt{DoFHandler} objects. In
+  each case, new interfaces are available.
 
@@
 
-M.~Bergbauer was supported by the German Research Foundation (DFG) 
-under the project ``High-Performance Cut Discontinuous Galerkin Methods
-for Flow Problems and Surface-Coupled Multiphysics Problems''.
+M.~Bergbauer was supported by the German Research Foundation (DFG) under the
+project ``High-Performance Cut Discontinuous Galerkin Methods for Flow
+Problems and Surface-Coupled Multiphysics Problems'' Grant Agreement
+No.~456365667.
 
@@
 M.~Kronbichler and P.~Munch were partially supported by the
-Bayerisches Kompetenznetzwerk
+German Ministry of Education and Research, project
+``PDExa: Optimized software methods for solving partial differential
+equations on exascale supercomputers'' and the Bayerisches Kompetenznetzwerk
 f\"ur Technisch-Wissen\-schaft\-li\-ches Hoch- und H\"ochstleistungsrechnen
-(KONWIHR) in the context of the projects ``High-order matrix-free finite
+(KONWIHR), projects ``High-order matrix-free finite
 element implementations with hybrid parallelization and improved data

### 2023-06-20 9.5/paper.tex

@@
 \texttt{intersection\_requests} is a vector indicating entities of a given triangulation (in the form of \texttt{GridTools::Cache}) that intersections are computed upon.
-Each entity (face or cell) is described by a vector of vertices.
+Each entity (face or cell) is described by a vector of vertices. The actual computation of the intersection of two geometric
+entities is performed by the new function \texttt{CGALWrappers::compute\_intersection\_of\_cells()}.
 
@@
 \begin{figure}
-
   \centering
-
-\todo[inline]{Complete these figures.}
-
-\fbox{\begin{minipage}{0.31\textwidth}\centering 0\vspace{5cm}\end{minipage}}
-\fbox{\begin{minipage}{0.31\textwidth}\centering 1\vspace{5cm}\end{minipage}}
-\fbox{\begin{minipage}{0.31\textwidth}\centering 2\vspace{5cm}\end{minipage}}
-
+  \begin{subfigure}[b]{0.28\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{png/mesh_0.png}
+  \end{subfigure}\qquad
+  \hfill
+  \begin{subfigure}[b]{0.28\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{png/mesh_1.png}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.28\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{png/mesh_2.png}
+  \end{subfigure}
   \caption{\it Example of non-nested multigrid levels.}\label{fig:nonnested}
-
 \end{figure}
@@
 
-At the time of writing, the \texttt{MGTwoLevelTransferNonNested} operator performs
-a pointwise interpolation (injection) from the coarse mesh to the support points
-of the fine mesh. Adopting the notation of~\cite{munch2022gc}, we perform
-for prolongation
-\todo{I think you need to at least explain in a few words what each of
-  the components of this formula is.}
-\todo{Alternatively (MK's preference), skip the parts that are just unnecessary
-  abstractions that do not help the reader's understanding.}
+At the time of writing, the \texttt{MGTwoLevelTransferNonNested} operator provides a prolongation operator $\mathcal{P}^{(f,c)}$
+that prolongates a vector $x^{(c)}$ from the coarse space to the fine space:
 \begin{align*}
-x^{(f)} = \mathcal{W}^{(f)} \circ \sum_{e \in \{\text{coarse cells}\}} \mathcal{S}_e^{(f)} \circ \mathcal{P}_e^{(f, c)}
-\circ \mathcal{C}_e^{(c)} \circ \mathcal{G}_e^{(c)} x^{(c)},
+x^{(f)} = \mathcal{P}^{(f, c)}  x^{(c)},
 \end{align*}
-i.e., loop over all coarse cells and interpolate to all (fine support) points that fall
-into the cell with \texttt{FEPointEvaluation}. The local result is scattered into a global
-vector, which is finalized by a communication and a weighting step.
+by a pointwise interpolation (injection) from the coarse mesh to the support points of the fine mesh. This is done
+by looping over all coarse cells and interpolating to all (fine support) points that fall into the cell with
+\texttt{FEPointEvaluation}. The local result is scattered into $x^{(f)}$, which is finalized by a
+communication and a weighting step that is necessary as multiple elements could add to the same global entry.
 

### 2023-06-21 9.5/paper.tex
Add PETSc stuff

--- 2023-06-16 9.5/paper.tex
@@
 \newcommand{\petsc}{\specialword{PETSc}\xspace}
+\newcommand{\snes}{{\specialword{SNES}}\xspace}
+\newcommand{\ts}{{\specialword{TS}}\xspace}
+\newcommand{\petscsf}{{\specialword{SF}}\xspace}
 \newcommand{\cmake}{{\specialword{CMake}}\xspace}
 \newcommand{\candi}{{\specialword{candi}}\xspace}
+\newcommand{\sundials}{{\specialword{SUNDIALS}}\xspace}
+\newcommand{\kinsol}{{\specialword{KINSOL}}\xspace}
+\newcommand{\ida}{{\specialword{IDA}}\xspace}
+\newcommand{\arkode}{{\specialword{ARKODE}}\xspace}
 
@@
 
-\todo[inline]{Stefano Z.: Please write this section.}
-
-\begin{itemize}
-\item PreconditionBDDC
-\item SNES
-\item TS
-\item Wrapping PETSc matrices and vectors
-\end{itemize}
+The \dealii classes wrapping \petsc objects have been heavily rewritten
+to support the nonlinear solver \snes and the Ordinary Differential Equations (ODE)
+solver \ts \cite{abhyankar2018petsc}.
+First, we briefly describe the most important improvements to the existing
+classes and then outline the newly designed interfaces to the nonlinear solvers, together
+with a new interface to the communication module \petscsf in \petsc \cite{zhang2021petscsf}.
+
+Vector and matrix classes of the \petsc wrappers have been extended with an additional constructor
+that can wrap an already existing \petsc vector or matrix, respectively.
+The \texttt{BlockVector} and
+\texttt{BlockSparseMatrix} classes now internally use
+\petsc nested objects, i.e. \texttt{VECNEST} and  \texttt{MATNEST} respectively.
+We have added a new class \texttt{PETScWrappers::PreconditionShell} to support user-defined
+preconditioning that can be simply customized as:
+\begin{c++}
+PETScWrappers::PreconditionShell preconditioner(/*...*/);
+
+preconditioner.vmult = [&](const VectorType &src,
+                           VectorType       &dst) {/*...*/};
+
+\end{c++}
+The resulting object can be passed to \petsc and used within the nonlinear solver hierarchy.
+See \ref{sec:callbacks} for additional informations on such kind of callbacks.
+
+The \petsc~\snes class solves system of nonlinear equations of the form $F(x) = 0$.
+The interface to \snes has been modeled on the already existing interface
+to the \kinsol solver from the \sundials package, and the nonlinear problem
+is specified with a set of callbacks that are used by the solver:
+\begin{c++}
+PETScWrappers::NonlinearSolver<VectorType> solver(/*...*/);
+
+solver.residual            = [&](const VectorType &x,
+                                 VectorType       &F) {/*...*/};
+solver.setup_jacobian      = [&](const VectorType &x) {/*...*/};
+solver.solve_with_jacobian = [&](const VectorType &rhs,
+                                 VectorType       &sol) {/*...*/};
+solver.solve(x);
+\end{c++}
+The default configuration is set up to use a Jacobian-free Newton-Krylov (JFNK)
+approach~\cite{knoll2004jacobian}, using \texttt{solve\_with\_jacobian} as linear solver.
+Numerous other solver configurations are possible and can be selected
+programmatically, or via the powerful command line customization of \petsc.
+This includes for example Quasi-Newton methods, Anderson's acceleration, and
+nonlinear preconditioning~\cite{brune2015composing}.
+
+The \petsc~\ts class solves ODEs in explicit or implicit form \cite{abhyankar2018petsc}, i.e.:
+\begin{eqnarray*}
+%\frac{\partial u}{\partial t} = G(t,u), &\text{explicit}\\
+%F(t,u,\frac{\partial u}{\partial t}) = 0, &\text{implicit}\\
+\dot{u} = G(t,u), &\text{(explicit)}\\
+F(t,u,\dot{u}) = 0, &\text{(implicit)}.\\
+\end{eqnarray*}
+The interface to \ts has been modeled on the already existing interfaces
+to the \ida and \arkode solvers from the \sundials package. Specifically:
+\begin{c++}
+PETScWrappers::TimeDependentSolver<VectorType> solver(/*...*/);
+
+// If solving udot = G(t,u)
+solver.explicit_function = [&](const double     t,
+                               const VectorType &u,
+                               VectorType       &G) {/*...*/};
+// If solving F(t,u,udot) = 0
+solver.implicit_function = [&](const double     t,
+                               const VectorType &u,
+                               const VectorType &udot,
+                               VectorType       &F) {/*...*/};
+\end{c++}
+In addition to specifying the function callbacks, users can further customize
+the solution of the linearized equations $\alpha * \frac{\partial{F}}{\partial{\dot{u}}} + \frac{\partial{F}}{\partial{u}}$
+via the callbacks:
+\begin{c++}
+solver.setup_jacobian = [&](const double     t,
+                            const VectorType &u,
+                            const VectorType &udot,
+                            const double     alpha) {/*...*/};
+solver.solve_with_jacobian = [&](const VectorType &rhs,
+                                 VectorType       &sol) {/*...*/};
+\end{c++}
+As for \snes, the default configuration of an implicit solver is set up to
+a JFNK approach, and the entire suite of solvers offered by PETSc is available,
+including adaptive time-stepping and Implicit-Explicit schemes.
+
+We refer interested readers to our documentation for other more advanced
+functions of the \ts and \snes wrappers, and to the {\tt tests/petsc/} folder
+for examples on how to use them.
+
+We close this section by introducing the interface to the \petscsf class,
+the abstract communication model of \petsc. The \dealii interface to
+\petscsf has been modelled on those of \texttt{Utilities::MPI::Partitioner} and
+\texttt{Utilities::MPI::NoncontiguousPartitioner}. Future developments
+will add support for GPU buffers.
 

### 2023-06-22 9.5/paper.tex
Add myself to the list of authors
minor rewriting of the PETSc section

@@
     Bruno Turcksin,
-    David Wells
+    David Wells,
+    Stefano Zampini
   },
@@
   {\texttt{drwells@email.unc.edu}}}
+
+\author[13]{Stefano Zampini}
+\affil[13]{Extreme Computing Research Center, King Abdullah University of Science and Technology,
+  23955-6900 Thuwal, Saudi Arabia.
+  {\texttt{stefano.zampini@gmail.com}}}
 
@@
 The \dealii classes wrapping \petsc objects have been heavily rewritten
-to support the nonlinear solver \snes and the Ordinary Differential Equations (ODE)
+to support the System of Nonlinear Equations Solver \snes and the Ordinary Differential Equations (ODE)
 solver \ts \cite{abhyankar2018petsc}.
@@
 
-We refer interested readers to our documentation for other more advanced
-functions of the \ts and \snes wrappers, and to the {\tt tests/petsc/} folder
-for examples on how to use them.
-
 We close this section by introducing the interface to the \petscsf class,
 the abstract communication model of \petsc. The \dealii interface to
-\petscsf has been modelled on those of \texttt{Utilities::MPI::Partitioner} and
-\texttt{Utilities::MPI::NoncontiguousPartitioner}. Future developments
-will add support for GPU buffers.
+\petscsf has been modelled on \texttt{Utilities::MPI::Partitioner} and
+\texttt{Utilities::MPI::NoncontiguousPartitioner}, with minimal changes for the
+communication routines API; the equivalent classes based
+on \petscsf are \texttt{PETScWrappers::Partitioner} and
+\texttt{PETScWrappers::CommunicationPattern}.
+Future developments will add support for GPU buffers.
+
+We refer interested readers to our documentation for other more advanced
+functions of the \snes, \ts and \petscsf wrappers, and to the {\tt tests/petsc/} folder
+for examples on how to use them.
 

### 2023-06-22 9.5/paper.tex
minor

@@
 Numerous other solver configurations are possible and can be selected
-programmatically, or via the powerful command line customization of \petsc.
+programmatically using the constructor arguments,
+or via the powerful command line customization of \petsc.
 This includes for example Quasi-Newton methods, Anderson's acceleration, and
@@
 As for \snes, the default configuration of an implicit solver is set up to
-a JFNK approach, and the entire suite of solvers offered by PETSc is available,
+use a JFNK approach, and the entire suite of solvers offered by PETSc is available
+programmatically or via the command line interface,
 including adaptive time-stepping and Implicit-Explicit schemes.
@@
 
-We refer interested readers to our documentation for other more advanced
+We refer interested readers to our documentation for more advanced
 functions of the \snes, \ts and \petscsf wrappers, and to the {\tt tests/petsc/} folder

### 2023-06-23 9.5/paper.tex
One more code gallery program.

@@
 
-There are also three new programs in the code gallery (a collection of
+There are four new programs in the code gallery (a collection of
 user-contributed programs that often solve more complicated problems
@@
     Sam Scheuerman.
+  \item \textit{``Information density-based mesh refinement''},
+    contributed by Wolfgang Bangerth.
 \end{itemize}

### 2023-06-23 9.5/paper.tex
Correctly sort author list.

@@
     Maximilian Bergbauer,
+    Marco Feder,
     Marc Fehling,
-    Marco Feder,
     Johannes Heinz,
@@
 
-\author[2]{Marc~Fehling}
-
 \author[5]{Marco~Feder}
@@
 {\texttt{marco.feder/luca.heltai@sissa.it}}}
+
+\author[2]{Marc~Fehling}
 

### 2023-06-23 9.5/paper.tex
Work through it one more time.

@@
 
-\dealii version 9.5.0 was released XX XX, 2023.
+\dealii version 9.5.0 was released June XX, 2023.
+\todo{Add exact date.}
 This paper provides an
@@
   simplifies the conversion from a \texttt{Cell\-Accessor} to a \texttt{DoF\-Cell\-Accessor}.
-  The old way 
+  The old way,
 \begin{c++}
@@
 
-The \dealii classes wrapping \petsc objects have been heavily rewritten
-to support the System of Nonlinear Equations Solver \snes and the Ordinary Differential Equations (ODE)
+The \dealii classes wrapping \petsc objects have been rewritten in
+substantial ways
+to support \petsc's System of Nonlinear Equations Solver \snes and the Ordinary Differential Equations (ODE)
 solver \ts \cite{abhyankar2018petsc}.
@@
 We have added a new class \texttt{PETScWrappers::PreconditionShell} to support user-defined
-preconditioning that can be simply customized as:
+preconditioning that can be as simply customized as the following code
+snippet:
 \begin{c++}
@@
 
-The \petsc~\snes class solves system of nonlinear equations of the form $F(x) = 0$.
+The \petsc~\snes sub-package solves systems of nonlinear equations of the form $F(x) = 0$.
 The interface to \snes has been modeled on the already existing interface
@@
 or via the powerful command line customization of \petsc.
-This includes for example Quasi-Newton methods, Anderson's acceleration, and
-nonlinear preconditioning~\cite{brune2015composing}.
-
-The \petsc~\ts class solves ODEs in explicit or implicit form \cite{abhyankar2018petsc}, i.e.:
+This includes, for example, Quasi-Newton methods, Anderson's acceleration, and
+nonlinear preconditioning~\cite{brune2015composing}. The step-77
+tutorial program has gained a section that discusses the use of these
+new \snes interfaces.
+
+The \petsc~\ts sub-package solves ODEs in explicit or implicit form \cite{abhyankar2018petsc}, i.e.:
 \begin{eqnarray*}
-%\frac{\partial u}{\partial t} = G(t,u), &\text{explicit}\\
-%F(t,u,\frac{\partial u}{\partial t}) = 0, &\text{implicit}\\
-\dot{u} = G(t,u), &\text{(explicit)}\\
-F(t,u,\dot{u}) = 0, &\text{(implicit)}.\\
+%\frac{\partial u}{\partial t} = G(t,u) &\text{explicit},\\
+%F(t,u,\frac{\partial u}{\partial t}) = 0 &\text{implicit}\\
+\dot{u} = G(t,u) &\text{(explicit)},\\
+F(t,u,\dot{u}) = 0 &\text{(implicit)}.\\
 \end{eqnarray*}
@@
 In addition to specifying the function callbacks, users can further customize
-the solution of the linearized equations $\alpha * \frac{\partial{F}}{\partial{\dot{u}}} + \frac{\partial{F}}{\partial{u}}$
-via the callbacks:
+the solution of the linearized equations $\alpha \frac{\partial{F}}{\partial{\dot{u}}} + \frac{\partial{F}}{\partial{u}}$
+via the following callbacks:
 \begin{c++}
@@
 \end{c++}
-As for \snes, the default configuration of an implicit solver is set up to
+As with \snes, the default configuration of an implicit solver is set up to
 use a JFNK approach, and the entire suite of solvers offered by PETSc is available
@@
 
-We close this section by introducing the interface to the \petscsf class,
+We close this section by introducing the interface to the \petscsf sub-package,
 the abstract communication model of \petsc. The \dealii interface to
-\petscsf has been modelled on \texttt{Utilities::MPI::Partitioner} and
-\texttt{Utilities::MPI::NoncontiguousPartitioner}, with minimal changes for the
+\petscsf has been modeled on the existing \texttt{Utilities::MPI::Partitioner} and
+\texttt{Utilities::MPI::NoncontiguousPartitioner} classes, with minimal changes for the
 communication routines API; the equivalent classes based
@@
 on the classes \texttt{FE\-Point\-Eval\-u\-ation} and \texttt{RemotePointEvaluation},
-which have been introduced in release 9.3. While \texttt{FE\-Point\-Eval\-u\-a\-tion}
+which were introduced in release 9.3. While \texttt{FE\-Point\-Eval\-u\-a\-tion}
 is responsible for efficient evaluation/integration at arbitrary (reference)
@@
 \subsubsection{Intersected meshes}
-In the current release, we added experimental support to compute intersections on parallel::distributed::Triangulation objects using \texttt{CGAL}~\cite{cgal-user-ref}.
-For this purpose we introduced a free function, which computes intersections and relevant information for communication from \texttt{intersection\_requests}.
+In the current release, we added experimental support to compute intersections on \texttt{parallel::distributed::Triangulation} objects using \texttt{CGAL}~\cite{cgal-user-ref}.
+For this purpose we introduce a free function that computes intersections and relevant information for communication from \texttt{intersection\_requests}.
 \texttt{intersection\_requests} is a vector indicating entities of a given triangulation (in the form of \texttt{GridTools::Cache}) that intersections are computed upon.
@@
 entities is performed by the new function \texttt{CGALWrappers::compute\_intersection\_of\_cells()}.
+
+\todo[inline]{The following code snippet uses a different function
+  that the text above. The connection between the two is not clear to me.}
 
@@
 
+\todo[inline]{Marco: It is unfortunate that we advertise a function in
+  namespace `internal' here. We should think about moving it, see \url{https://github.com/dealii/dealii/issues/15463}.}
+
 For the common case of Nitsche-type mortaring, quadrature points must be distributed on the intersections to evaluate the underlying physical coupling terms.
@@
 The whole procedure is done communication-free.
-
 \begin{c++}
@@
 
-This functionality is handy since \texttt{RemotePointEvaluation} can now be used to access quantities at quadrature points on intersections without further ado.
+This functionality is useful since \texttt{RemotePointEvaluation} can now be used to access quantities at quadrature points on intersections without further ado.
 To reduce the user's effort, we plan to add a wrapper that takes care of the described procedure and provides interfaces like \texttt{FEEvaluation} to access quantities easily, e.g., in a matrix-free loop.
 
-Described implementation and an early version of the wrapper have been used successfully in~\cite{heinz2023high} to perform Nitsche-type mortaring in the context of the conservative formulation of acoustic equations discretized with DG to suppress artificial modes.
+The interface described above and an early version of the wrapper have
+been used successfully in~\cite{heinz2023high} to perform Nitsche-type
+mortaring in the context of the conservative formulation of acoustic
+equations discretized with a discontinuous Galerkin method to suppress
+artificial modes.
+
 \subsubsection{Non-nested multigrid}
 
-\dealii has provided support for geometric multigrid methods%
-\footnote{\dealii{} also supports algebraic multigrid methods
-by interfacing to the external libraries \texttt{PETSc}
-and \texttt{Trilinos}.}
-for locally refined meshes nearly since its inception.
-Traditionally, these were based on local-smoothing
+\dealii has provided support for geometric multigrid methods (GMG)
+for locally refined meshes nearly since its inception. (\dealii{} also
+supports algebraic multigrid methods by interfacing to the external
+libraries \texttt{PETSc} and \texttt{Trilinos}.)
+Traditionally, these GMGs were based on local-smoothing
 methods (described many years later in \cite{Kanschat2004,JanssenKanschat2011}, see
@@
 argument \texttt{MeshType} does not satisfy the named concept, which
-is internally defined as the type being either a \texttt{Triangulation} or \texttt{DoFHandler}
+is defined as the type being either a \texttt{Triangulation} or \texttt{DoFHandler}
 object, as intended. (The \texttt{requires} annotation is suppressed if the compiler
@@
 
-\todo[inline]{Write}
+\todo[inline]{Matthias: Write}
 
@@
 While there are no new \dealii tutorial programs in this release, many
-were extensively revised: Around 130 of the more than 2000 (non-merge)
+were extensively revised: Around 145 of the more than 2100 (non-merge)
 commits that went into this release touched the tutorial, in some
@@
 \href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_4_2_and_9_5_0.html}
-{around X incompatible changes}; see \cite{changes95}. Many of the
-incompatible change internal
+     {around XXX incompatible changes};
+\todo{Fix number before release}
+see \cite{changes95}. Many of these
+incompatibilities change internal
 interfaces that are not usually used in external
@@
   \texttt{DoFHandler} class now implements all $hp$-related
-  functionality, as it has for the past few releases.
+  functionality, as it has for the past several releases.
 
@@
   \item
-  This rework also affects the serialization process of active FE indices.
-You will need to recreate your serialized data \textit{if and only if} you work in $hp$-mode.
-You can continue to use previously generated data if you do not use $hp$-mode.
-A special case forms \texttt{parallel::distributed::Triangulation} for non-$hp$ data, in which it is sufficient to increase the version in the metadata file by one from ``4'' to ``5''.
+  This rework also affects the serialization process of active FE
+  indices. You will need to recreate your serialized data \textit{if
+    and only if} you work in $hp$-mode, but can continue to use
+  previously generated data otherwise. For
+  \texttt{parallel::distributed::Triangulation} when used with non-$hp$ data, it is sufficient to increase the version in the metadata file from ``4'' to ``5''.
 
@@
   each case, new interfaces are available.
-
-\item
-  ...
+\todo[inline]{The sentence part ``e.g., initialization
+  functions...objects'' does not sound grammatically correct. Should
+  it have been ``and the functions querying \texttt{DoFHandler}
+  objects''? Please fix and remove this box.}
 \end{itemize}
-\todo[inline]{Write}
 
@@
 
-\todo[inline]{Need to update. Also remove those not on the current paper.}
-
 D.~Arndt and B.~Turcksin: Research sponsored by the Laboratory Directed Research and
@@
 No.~EAR-1550901 and The University of California -- Davis.
-\todo{This is the CIG-II grant. We need to update this to the CIG-III
-  (and CIG-IV?) grants.}
 
@@
 No.~456365667.
+
+\todo[inline]{Marco: Anything to add here?.}
 
@@
 
+\todo[inline]{Stefano: Anything to add here?.}
+
 \bibliography{paper}{}

### 2023-06-24 9.5/paper.tex
fix comment in code snippet

@@
          n_points_1D, tria, mapping,
-         // following parameter is optional would need additional
-         // communication 
+         // following parameter is optional
+         // setting it true requires additional communication
          consistent_numbering_of_sender_and_receiver);

### 2023-06-24 9.5/paper.tex
shift text below cpp code to avoid confusion

--- 2023-06-23 9.5/paper.tex
@@
 \texttt{intersection\_requests} is a vector indicating entities of a given triangulation (in the form of \texttt{GridTools::Cache}) that intersections are computed upon.
-Each entity (face or cell) is described by a vector of vertices. The actual computation of the intersection of two geometric
-entities is performed by the new function \texttt{CGALWrappers::compute\_intersection\_of\_cells()}.
-
-\todo[inline]{The following code snippet uses a different function
-  that the text above. The connection between the two is not clear to me.}
+Each entity (face or cell) is described by a vector of vertices.
 
@@
   namespace `internal' here. We should think about moving it, see \url{https://github.com/dealii/dealii/issues/15463}.}
+
+Internally above function uses the new function \texttt{CGALWrappers::compute\_intersection\_of\_cells()}, to actually compute intersections between two geometric entities.
 

### 2023-06-24 9.5/paper.tex
nonlinear solver

@@
     interfaces
-    (Sections~\ref{sec:petsc} and \ref{sec:trilinos}); and a uniform
-    way of defining callbacks used by external libraries
-    (Section~\ref{sec:callbacks}).
-  \item Advances in matrix-free infrastructure (see Section~\ref{sec:mf});
-  \item Advances in non-matching support (see Section~\ref{sec:nonmatching});
-  \item New features related to liner algebra (see Section~\ref{sec:lac})
+    (Sections~\ref{sec:petsc} and \ref{sec:trilinos}).
+  \item Uniform handling of nonlinear solver
+  packages (Section~\ref{sec:nonlinear}) and a uniform
+    way of defining callbacks used by external libraries (Section~\ref{sec:callbacks}).
+  \item Advances in matrix-free infrastructure (see Section~\ref{sec:mf}).
+  \item Advances in non-matching support (see Section~\ref{sec:nonmatching}).
+  \item New features related to linear algebra (see Section~\ref{sec:lac}).
   \item C++ language modernization (see Section~\ref{sec:language}).
@@
   preconditioner whenever setup\_jacobian() is called.}
+
+\subsubsection{Uniform interface for nonlinear solvers}\label{sec:nonlinear}
+
+With the current release, \dealii{} now supports solvers for nonlinear problems
+provided by several different external packages:
+1. KINSOL (part of SUNDIALS), 2. SNES (part of PETSc, see Section~\ref{sec:petsc}), 3. NOX (part of Trilinos, see Section~\ref{sec:trilinos}).
+The wrappers are provided by the classes \texttt{PETScWrappers::NonlinearSolver<VectorType>}, \texttt{SUNDIALS::KINSOL<VectorType>}, and \texttt{TrilinosWrappers::NOXSolver<VectorType>}, respectively.
+
+All classes have a very similar interface, except that they vary in the kind of
+algorithms and parameters offered. The new class \texttt{NonlinearSolverSelector<VectorType>} provides a wrapper on top of
+the three external solvers with a unified interface. The user can either let
+\dealii{} decide which of the packages to use (depending on the current availability of
+the external packages) or specify it manually.
+
+Here is a complete example of an MPI parallel Newton solver that automatically chooses one of the available packages:
+\begin{c++}
+    using NLSolver = NonlinearSolverSelector<LA::MPI::Vector>;
+
+    NLSolver::AdditionalData additional_data(NLSolver::automatic,
+                                             NLSolver::newton);
+    NLSolver nonlinear_solver(additional_data, mpi_communicator);
+
+    nonlinear_solver.reinit_vector = [&](LA::MPI::Vector &x)
+                                     {/*...*/};
+    nonlinear_solver.residual = [&](const LA::MPI::Vector &evaluation_point,
+                                    LA::MPI::Vector &      residual)
+                                {/*...*/};
+    nonlinear_solver.setup_jacobian = [&](const LA::MPI::Vector &current_u)
+                                      {/*...*/};
+    nonlinear_solver.solve_with_jacobian = [&](const LA::MPI::Vector &rhs,
+                                               LA::MPI::Vector &      dst,
+                                               const double tolerance)
+                                           {/*...*/};
+
+    nonlinear_solver.solve(current_solution);
+\end{c++}
 
@@
 
-Internally above function uses the new function \texttt{CGALWrappers::compute\_intersection\_of\_cells()}, to actually compute intersections between two geometric entities.
+The above function uses the new function \texttt{CGALWrappers::compute\_intersection\_of\_cells()} internally to compute intersections between two geometric entities.
 
@@
          n_points_1D, tria, mapping,
-         // following parameter is optional would need additional
-         // communication 
+         // The following parameter is optional and setting it to 
+         // true requires additional communication
          consistent_numbering_of_sender_and_receiver);
@@
 context of aeroacoustic problems.
-\todo{Can we add a reference for this?}
 

### 2023-06-24 9.5/paper.tex
Update 9.5/paper.tex
Co-authored-by: Timo Heister <timo.heister@gmail.com>

@@
 \texttt{intersection\_requests} is a vector indicating entities of a given triangulation (in the form of \texttt{GridTools::Cache}) that intersections are computed upon.
-Each entity (face or cell) is described by a vector of vertices.
+Each entity (face or cell) is described by a vector of vertices. The actual computation of the intersection of two geometric
+entities is performed by the new function \texttt{CGALWrappers::compute\_intersection\_of\_cells()}.
+
+\todo[inline]{The following code snippet uses a different function
+  that the text above. The connection between the two is not clear to me.}
 
@@
 
-The above function uses the new function \texttt{CGALWrappers::compute\_intersection\_of\_cells()} internally to compute intersections between two geometric entities.
-
 For the common case of Nitsche-type mortaring, quadrature points must be distributed on the intersections to evaluate the underlying physical coupling terms.

### 2023-06-24 9.5/paper.tex

@@
   namespace `internal' here. We should think about moving it, see \url{https://github.com/dealii/dealii/issues/15463}.}
+
+The above function uses the new function \texttt{CGALWrappers::compute\_intersection\_of\_cells()} internally to compute intersections between two geometric entities.
 

### 2023-06-26 9.5/paper.tex

@@
   preconditioner whenever setup\_jacobian() is called.}
+
+\subsubsection{Uniform interface for nonlinear solvers}\label{sec:nonlinear}
+
+With the current release, \dealii{} now supports solvers for nonlinear problems
+provided by several different external packages:
+(i) KINSOL (part of SUNDIALS); (ii) SNES (part of PETSc, see Section~\ref{sec:petsc}); and (iii) NOX (part of Trilinos, see Section~\ref{sec:trilinos}).
+The wrappers are provided by the classes \texttt{PETScWrappers::NonlinearSolver<VectorType>}, \texttt{SUNDIALS::KINSOL<VectorType>}, and \texttt{TrilinosWrappers::NOXSolver<VectorType>}, respectively.
+
+All three of these classes have a very similar interface, except that they vary in the kind of
+algorithms and parameters offered. The new class \texttt{NonlinearSolverSelector<VectorType>} provides a wrapper on top of
+the three external solvers with a unified interface. The user can either let
+\dealii{} decide which of the packages to use (depending on the current availability of
+the external packages) or specify it manually.
+
+Here is a complete example of an MPI parallel Newton solver that automatically chooses one of the available packages:
+\begin{c++}
+    using NLSolver = NonlinearSolverSelector<LA::MPI::Vector>;
+
+    NLSolver::AdditionalData additional_data(NLSolver::automatic,
+                                             NLSolver::newton);
+    NLSolver nonlinear_solver(additional_data, mpi_communicator);
+
+    nonlinear_solver.reinit_vector = [&](LA::MPI::Vector &x)
+                                     {/*...*/};
+    nonlinear_solver.residual = [&](const LA::MPI::Vector &evaluation_point,
+                                    LA::MPI::Vector &      residual)
+                                {/*...*/};
+    nonlinear_solver.setup_jacobian = [&](const LA::MPI::Vector &current_u)
+                                      {/*...*/};
+    nonlinear_solver.solve_with_jacobian = [&](const LA::MPI::Vector &rhs,
+                                               LA::MPI::Vector &      dst,
+                                               const double tolerance)
+                                           {/*...*/};
+
+    nonlinear_solver.solve(current_solution);
+\end{c++}
 

### 2023-06-26 9.5/paper.tex
dont show internal interfaces

@@
 In the current release, we added experimental support to compute intersections on \texttt{parallel::distributed::Triangulation} objects using \texttt{CGAL}~\cite{cgal-user-ref}.
-For this purpose we introduce a free function that computes intersections and relevant information for communication from \texttt{intersection\_requests}.
-\texttt{intersection\_requests} is a vector indicating entities of a given triangulation (in the form of \texttt{GridTools::Cache}) that intersections are computed upon.
+For this purpose, we introduced a free function \texttt{distributed\_compute\_intersection\_locations()} that computes intersections and relevant information for communication from \texttt{intersection\_requests}.
+\texttt{intersection\_requests} is a vector indicating entities of a given triangulation that intersections are computed upon.
 Each entity (face or cell) is described by a vector of vertices.
-
-\begin{c++}
-// compute intersections on distributed triangulation
-auto intersection_data =
-       GridTools::internal::distributed_compute_intersection_locations(
-         cache, intersection_requests, global_bboxes, marked_vertices,
-         tolerance);
-\end{c++}
-
-\todo[inline]{Marco: It is unfortunate that we advertise a function in
-  namespace `internal' here. We should think about moving it, see \url{https://github.com/dealii/dealii/issues/15463}.}
-
-The above function uses the new function \texttt{CGALWrappers::compute\_intersection\_of\_cells()} internally to compute intersections between two geometric entities.
+Currently, the function is placed in \texttt{GridTools::internal}, but the location and arguments of the function might still change in the future.
+To compute intersections between two geometric entities, the function internally uses the new function \texttt{CGALWrappers::compute\_intersection\_of\_cells()}.
 
 For the common case of Nitsche-type mortaring, quadrature points must be distributed on the intersections to evaluate the underlying physical coupling terms.
-The \texttt{intersection\_data} returned by the function above can convert itself to data which can be used to fill \texttt{RemotePointEvaluation}.
+The data structure returned by \texttt{distributed\_compute\_intersection\_locations()} can convert itself to a data structure that can be used to fill \texttt{RemotePointEvaluation}.
+This conversion is triggered by the member function~\texttt{convert\_to\_distributed\_compute\_point\_locations\_internal()}, given the number of quadrature points per intersection.
 The whole procedure is done communication-free.
-\begin{c++}
-// distribute n_points_1D quadrature points to the intersections
-auto point_data = intersection_data.template
-       convert_to_distributed_compute_point_locations_internal<dim>(
-         n_points_1D, tria, mapping,
-         // The following parameter is optional and setting it to 
-         // true requires additional communication
-         consistent_numbering_of_sender_and_receiver);
-
-// use point_data to reinit RemotePointEvaluation
-Utilities::MPI::RemotePointEvaluation<dim> rpe;
-rpe.reinit(point_data,tria, mapping);
-\end{c++}
 
@@
 
-The interface described above and an early version of the wrapper have
+The procedure described above and an early version of the wrapper have
 been used successfully in~\cite{heinz2023high} to perform Nitsche-type

### 2023-06-26 9.5/paper.tex
Use a TT font for library names

--- 2023-06-26 9.5/paper.tex
@@
 \newcommand{\arkode}{{\specialword{ARKODE}}\xspace}
+\newcommand{\boost}{{\specialword{Boost}}\xspace}
+\newcommand{\kokkos}{{\specialword{Kokkos}}\xspace}
 
@@
     to other libraries (see Section~\ref{sec:external}). This
-    includes, in particular, the integration of \texttt{Kokkos} (Section~\ref{sec:kokkos});
-    additions and updates to the \texttt{PETSc} and Trilinos
+    includes, in particular, the integration of \kokkos (Section~\ref{sec:kokkos});
+    additions and updates to the \petsc and \trilinos
     interfaces
@@
 
-\todo[inline]{We are not consistent in whether or not we print PETSc,
-  SUNDIALS, Trilinos, SNES, ... with a TT font (texttt) or
-  not. Someone should go through the paper and make it consistent.}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 For many operations, \dealii{} relies on external libraries -- some of
-these are optional, others are mandatory (such as BOOST and, now,
-Kokkos); a complete list of external dependencies is provided in
+these are optional, others are mandatory (such as \boost and, now,
+\kokkos); a complete list of external dependencies is provided in
 Section~\ref{sec:cite}. A substantial amount of work has gone into
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Integration of Kokkos}\label{sec:kokkos}
-
-Kokkos \cite{trott2022} is a C++ library that enables the creation of
+\subsubsection{Integration of \kokkos}\label{sec:kokkos}
+
+\kokkos \cite{trott2022} is a C++ library that enables the creation of
 performance portable applications for all major high-performance computing (HPC)
 platforms. It implements a programming model that allows developers to write
-code that can efficiently run on diverse architectures. Kokkos provides
+code that can efficiently run on diverse architectures. \kokkos provides
 abstractions for both parallel execution of code and data management. It
@@
 a single vendor (Nvidia) has made it clear that we need a different
-strategy to support what users want. As a consequence, Kokkos has
+strategy to support what users want. As a consequence, \kokkos has
 become a mandatory dependency of \dealii{} as part of the current
 release; if it is not found on a given system during configuration
-time, then the library will fall back on a copy of Kokkos stored in
+time, then the library will fall back on a copy of \kokkos stored in
 the \texttt{bundled/} directory in the same way as we already
-interface with BOOST.
-
-In the current release, LinearAlgebra::distributed::Vector and the
-CUDAWrappers::MatrixFree framework are using Kokkos. This allows them to work on
-all the architectures supported by Kokkos. The Kokkos backend used by \dealii is
+interface with \boost.
+
+In the current release, \texttt{LinearAlgebra::distributed::Vector} and the \texttt{CUDAWrappers::MatrixFree} framework are using \kokkos. This allows them to work on
+all the architectures supported by \kokkos. The \kokkos backend used by \dealii is
 \texttt{Kokkos::DefaultExecutionSpace} which is the highest available in the hierarchy
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Updates and additions to the PETSc wrappers}\label{sec:petsc}
+\subsubsection{Updates and additions to the \petsc wrappers}\label{sec:petsc}
 
@@
 As with \snes, the default configuration of an implicit solver is set up to
-use a JFNK approach, and the entire suite of solvers offered by PETSc is available
+use a JFNK approach, and the entire suite of solvers offered by \petsc is available
 programmatically or via the command line interface,
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Interfaces to Trilinos' Belos and NOX packages}\label{sec:trilinos}
-
-Trilinos is a large collection of individual sub-packages \cite{heroux2005trilinos,trilinos-web-page}. \dealii{}
-has long had interfaces to the Trilinos packages that provide parallel
+\subsubsection{Interfaces to \trilinos{}' \texttt{Belos} and \texttt{NOX} packages}\label{sec:trilinos}
+
+\trilinos is a large collection of individual sub-packages \cite{heroux2005trilinos,trilinos-web-page}. \dealii{}
+has long had interfaces to the \trilinos packages that provide parallel
 vector and matrix classes (both \texttt{Epetra} and \texttt{Tpetra}),
@@
 \texttt{Belos} is the
-successor of the Trilinos package \texttt{AztecOO} and provides basic and advanced iterative solvers
+successor of the \trilinos package \texttt{AztecOO} and provides basic and advanced iterative solvers
 that heavily rely on multivector operations. The interface of the
@@
 Secondly, we have added a wrapper to \texttt{NOX}, a
-nonlinear solver library that is similar to both the \texttt{KINSOL} solver from
-the \texttt{SUNDIALS} package to which we already had interfaces, and also
-to the \texttt{SNES} collections of functions from \texttt{PETSc} (see also
+nonlinear solver library that is similar to both the \kinsol{} solver from
+the \sundials{} package to which we already had interfaces, and also
+to the \snes{} collections of functions from \petsc (see also
 Section~\ref{sec:petsc}). The interfaces to these three solvers are
@@
 \todo[inline]{I don't understand this last sentence. Can you
-  elaborate? At least in the KINSOL interfaces, you only update the
+  elaborate? At least in the \kinsol{} interfaces, you only update the
   preconditioner whenever setup\_jacobian() is called.}
@@
 provided by several different external packages:
-(i) KINSOL (part of SUNDIALS); (ii) SNES (part of PETSc, see Section~\ref{sec:petsc}); and (iii) NOX (part of Trilinos, see Section~\ref{sec:trilinos}).
+(i) \kinsol{} (part of \sundials{}); (ii) \snes{} (part of \petsc{}, see Section~\ref{sec:petsc}); and (iii) \texttt{NOX} (part of \trilinos{}, see Section~\ref{sec:trilinos}).
 The wrappers are provided by the classes \texttt{PETScWrappers::NonlinearSolver<VectorType>}, \texttt{SUNDIALS::KINSOL<VectorType>}, and \texttt{TrilinosWrappers::NOXSolver<VectorType>}, respectively.
@@
 see for example the code example in Section~\ref{sec:trilinos} on how
-NOX gains information about the residual vector, how it indicates to
+\texttt{NOX} gains information about the residual vector, how it indicates to
 user code that the Jacobian matrix needs to be rebuilt,
@@
 solving a nonlinear problem. Indeed, the interfaces to the nonlinear
-solver KINSOL (as part of SUNDIALS) and SNES (as part of \texttt{PETSc}, see
+solver \kinsol{} (as part of \sundials{}) and \snes{} (as part of \petsc, see
 Section~\ref{sec:petsc} function in exactly the same way; the
-ODE solver interfaces to ARKode and IDA (also part of SUNDIALS) and TS
-(as part of PETSc) also use this style.
+ODE solver interfaces to \arkode{} and \ida{} (also part of \sundials{}) and \ts{}
+(as part of \petsc{}) also use this style.
 
@@
 convention on how success or error codes of these callbacks should be
-encoded. In the case of PETSc's SNES and TS, and for Trilinos's \texttt{NOX},
+encoded. In the case of \petsc{}'s \snes{} and \ts{}, and for \trilinos{}'s \texttt{NOX},
 success is indicated by a zero integer return value, whereas failure is
-indicated by a nonzero return value. On the other hand, the SUNDIALS
+indicated by a nonzero return value. On the other hand, the \sundials{}
 packages indicate success by a zero integer return value, a
@@
 supports algebraic multigrid methods by interfacing to the external
-libraries \texttt{PETSc} and \texttt{Trilinos}.)
+libraries \petsc and \trilinos.)
 Traditionally, these GMGs were based on local-smoothing
@@
 supports C++17, but falls back to implementations obtained via the
-BOOST library otherwise, and this is true also for a number of
+\boost library otherwise, and this is true also for a number of
 individual functions that were introduced in C++17 or C++20.

### 2023-06-26 9.5/paper.tex
Fix typos

@@
 The resulting object can be passed to \petsc and used within the nonlinear solver hierarchy.
-See \ref{sec:callbacks} for additional informations on such kind of callbacks.
+See \ref{sec:callbacks} for additional information on such kind of callbacks.
 
@@
 fixed now. Furthermore, \texttt{FE\_Nothing} now also works together
-with discontinuous Lagrange elements (i.e., with the \texttt{FE\_DGQ} class). Due to the popularity of \texttt{FE\_Nothing} as a mean to enable
+with discontinuous Lagrange elements (i.e., with the \texttt{FE\_DGQ} class). Due to the popularity of \texttt{FE\_Nothing} as a means to enable
 or disable cells, we have introduced the new class

### 2023-06-27 9.5/paper.tex
add list of contributors

@@
 
-% This is up-to-date as of 9.4 RC1 - should be the final list.
-
-\todo[inline]{Add.}
+% This is up-to-date as of 2023-06-26 7:00pm central
+
+Pasquale    Africa,
+Nicolas     Barnafi,
+Nistha      Bhawsinka,
+Bruno       Blais,
+Fabian      Castelli,
+Terry       Cojean,
+Niklas      Fehn,
+Emmanuel    Ferdman,
+Vadim       Gallyamov,
+Daniel      Garcia-Sanchez,
+Rene        Gassmoeller,
+Robin       Goermer,
+Graham      Harper,
+Quang       Hoang,
+Sean        Ingimarson,
+Vladimir    Ivannikov,
+Pengfei     Jia,
+Tao         Jin,
+Nils        Margenberg,
+Luz         Paz,
+Laura       Prieto,
+Sebastian   Proell,
+Ce          Qin,
+Oleg        Rogozin,
+Andrew      Salmon,
+Michael     Schlottke-Lakemper,
+Christoph   Schmidt,
+Magdalena   Schreter,
+Richard     Schussnig,
+Nils        Schween,
+Ahmad       Shahba,
+Simon       Sticko,
+Buğrahan    Temür,
+Ivy         Weber,
+Niklas      Wik,
+Vladimir    Yushutin,
+Jiaqi       Zhang.
 

### 2023-06-27 9.5/paper.tex
Add two annotations.

@@
 
-Here is a complete example of an MPI parallel Newton solver that automatically chooses one of the available packages:
+Here is a complete example of an MPI parallel Newton solver that
+automatically chooses one of the available packages:
+\todo[inline]{The NOX example was nicely indented. Can we do the same
+  here? Perhaps if we call the variable \texttt{nl\_solver}?}
+\todo[inline]{But separately: Does the code snippet really provide any
+  additional value? It is for all practical purposes identical to the
+  NOX one, and the NOX one is already identical to the SNES one. Do we
+need to repeat things a third time, or could one just use one or two
+sentences to express the same?}
 \begin{c++}

### 2023-06-27 9.5/paper.tex
Update Kokkos section

@@
 
-In the current release, \texttt{LinearAlgebra::distributed::Vector} and the \texttt{CUDAWrappers::MatrixFree} framework are using \kokkos. This allows them to work on
-all the architectures supported by \kokkos. The \kokkos backend used by \dealii is
-\texttt{Kokkos::DefaultExecutionSpace} which is the highest available in the hierarchy
-device, host-parallel, and host-serial.
-
-\todo[inline]{Are you saying that all three of these are used? If so,
-  maybe replace ``backend'' by ``backends''?}
-
-\todo[inline]{Would it be useful to add a paragraph about future plans?}
+In the current release, \texttt{LinearAlgebra::distributed::Vector} and the 
+\texttt{CUDAWrappers::MatrixFree} framework are using \kokkos. This allows them 
+to work on all the architectures supported by \kokkos. The \kokkos backend used 
+by \dealii is \texttt{Kokkos::DefaultExecutionSpace} which corresponds to the 
+highest available backend in the hierarchy device, host-parallel, and host-serial 
+at the time Kokkos was configured.
 

### 2023-06-29 9.5/paper.tex
Diverse

@@
     interfaces
-    (Sections~\ref{sec:petsc} and \ref{sec:trilinos}).
+    (Sections~\ref{sec:petsc} and \ref{sec:trilinos}, respectively).
   \item Uniform handling of nonlinear solver
@@
 In the current release, \texttt{LinearAlgebra::distributed::Vector} and the 
-\texttt{CUDAWrappers::MatrixFree} framework are using \kokkos. This allows them 
+\texttt{CUDAWrappers::Matrix\allowbreak{}Free} framework are using \kokkos. This allows them 
 to work on all the architectures supported by \kokkos. The \kokkos backend used 
@@
 We have added a new class \texttt{PETScWrappers::PreconditionShell} to support user-defined
-preconditioning that can be as simply customized as the following code
-snippet:
+preconditioning that can be simply customized as the following code
+snippet shows:
 \begin{c++}
@@
 to the \kinsol solver from the \sundials package, and the nonlinear problem
-is specified with a set of callbacks that are used by the solver:
-\begin{c++}
-PETScWrappers::NonlinearSolver<VectorType> solver(/*...*/);
-
-solver.residual            = [&](const VectorType &x,
-                                 VectorType       &F) {/*...*/};
-solver.setup_jacobian      = [&](const VectorType &x) {/*...*/};
-solver.solve_with_jacobian = [&](const VectorType &rhs,
-                                 VectorType       &sol) {/*...*/};
-solver.solve(x);
-\end{c++}
+is specified with a set of callbacks that are used by the solver.
+%\begin{c++}
+%PETScWrappers::NonlinearSolver<VectorType> solver(/*...*/);
+%
+%solver.residual            = [&](const VectorType &x,
+%                                 VectorType       &F) {/*...*/};
+%solver.setup_jacobian      = [&](const VectorType &x) {/*...*/};
+%solver.solve_with_jacobian = [&](const VectorType &rhs,
+%                                 VectorType       &sol) {/*...*/};
+%solver.solve(x);
+%\end{c++}
 The default configuration is set up to use a Jacobian-free Newton-Krylov (JFNK)
@@
 to the \snes{} collections of functions from \petsc (see also
-Section~\ref{sec:petsc}). The interfaces to these three solvers are
-essentially the same and require setting function objects for
-computing the residual of the nonlinear problem, along with setting
-up, applying, and solving with the Jacobian matrix:
+Section~\ref{sec:petsc}). 
+%The interfaces to these three solvers are
+%essentially the same and require setting function objects for
+%computing the residual of the nonlinear problem, along with setting
+%up, applying, and solving with the Jacobian matrix:
+%\begin{c++}
+%TrilinosWrappers::NOXSolver<VectorType> solver(/*...*/);
+%
+%solver.residual            = [ ](const VectorType &src,
+%                                 VectorType       &dst) {/*...*/};
+%solver.setup_jacobian      = [&](const VectorType &src) {/*...*/};
+%solver.apply_jacobian      = [&](const VectorType &src,
+%                                 VectorType       &dst) {/*...*/};
+%solver.solve_with_jacobian = [&](const VectorType &src,
+%                                 VectorType       &dst,
+%                                 const double      tol) {/*...*/};
+%
+%solver.solve(solution);
+%\end{c++}
+%
+We refer interested readers to our documentation for other more advanced
+functions of this wrapper which are related to the differences in features of these libraries.
+In particular, we have added the possibility to reuse the preconditioner between
+nonlinear steps (also known as \textit{preconditioner lagging}), which is
+unfortunately natively only  supported in the official \texttt{EPetra} implementations.
+
+\todo[inline]{I don't understand this last sentence. Can you
+  elaborate? At least in the \kinsol{} interfaces, you only update the
+  preconditioner whenever setup\_jacobian() is called.}
+
+\subsubsection{Uniform interface for nonlinear solvers}\label{sec:nonlinear}
+
+With the current release, \dealii{} now supports solvers for nonlinear problems
+provided by several different external packages:
+(i) \kinsol{} (part of \sundials{}); (ii) \snes{} (part of \petsc{}, see Section~\ref{sec:petsc}); and (iii) \texttt{NOX} (part of \trilinos{}, see Section~\ref{sec:trilinos}).
+The wrappers are provided by the classes \texttt{PETScWrappers::NonlinearSolver\allowbreak{}<VectorType>}, \texttt{SUNDIALS::KINSOL<VectorType>}, and \texttt{TrilinosWrappers::NOXSolver<Vector\allowbreak{}Type>}, respectively.
+
+All three of these classes have a very similar interface, except that they vary in the kind of
+algorithms and parameters offered. The new class \texttt{NonlinearSolverSelector<VectorType>} provides a wrapper on top of
+the three external solvers with a unified interface. The user can either let
+\dealii{} decide which of the packages to use (depending on the current availability of
+the external packages) or specify it manually.
+
+The following code snippet shows a complete example using the new class for applying a Newton solver
+by automatically choosing one of the available packages:
 \begin{c++}
-TrilinosWrappers::NOXSolver<VectorType> solver(/*...*/);
-
+using NLS = NonlinearSolverSelector<VectorType>;
+
+NLS::AdditionalData additional_data(
+  NLS::automatic, // other options: kinsol, nox, petsc_snes 
+  NLS::newton);
+
+NLS nonlinear_solver(additional_data, mpi_communicator);
+
+solver.reinit_vector       = [ ](VectorType &x) {/*...*/};    
 solver.residual            = [ ](const VectorType &src,
@@
 
-solver.solve(solution);
+solver.solve(current_solution);    
 \end{c++}
 
-We refer interested readers to our documentation for other more advanced
-functions of this wrapper which are related to the differences in features of these libraries.
-In particular, we have added the possibility to reuse the preconditioner between
-nonlinear steps (also known as \textit{preconditioner lagging}), which is
-unfortunately natively only  supported in the official \texttt{EPetra} implementations.
-
-\todo[inline]{I don't understand this last sentence. Can you
-  elaborate? At least in the \kinsol{} interfaces, you only update the
-  preconditioner whenever setup\_jacobian() is called.}
-
-\subsubsection{Uniform interface for nonlinear solvers}\label{sec:nonlinear}
-
-With the current release, \dealii{} now supports solvers for nonlinear problems
-provided by several different external packages:
-(i) \kinsol{} (part of \sundials{}); (ii) \snes{} (part of \petsc{}, see Section~\ref{sec:petsc}); and (iii) \texttt{NOX} (part of \trilinos{}, see Section~\ref{sec:trilinos}).
-The wrappers are provided by the classes \texttt{PETScWrappers::NonlinearSolver<VectorType>}, \texttt{SUNDIALS::KINSOL<VectorType>}, and \texttt{TrilinosWrappers::NOXSolver<VectorType>}, respectively.
-
-All three of these classes have a very similar interface, except that they vary in the kind of
-algorithms and parameters offered. The new class \texttt{NonlinearSolverSelector<VectorType>} provides a wrapper on top of
-the three external solvers with a unified interface. The user can either let
-\dealii{} decide which of the packages to use (depending on the current availability of
-the external packages) or specify it manually.
-
-Here is a complete example of an MPI parallel Newton solver that
-automatically chooses one of the available packages:
-\todo[inline]{The NOX example was nicely indented. Can we do the same
-  here? Perhaps if we call the variable \texttt{nl\_solver}?}
-\todo[inline]{But separately: Does the code snippet really provide any
-  additional value? It is for all practical purposes identical to the
-  NOX one, and the NOX one is already identical to the SNES one. Do we
-need to repeat things a third time, or could one just use one or two
-sentences to express the same?}
-\begin{c++}
-    using NLSolver = NonlinearSolverSelector<LA::MPI::Vector>;
-
-    NLSolver::AdditionalData additional_data(NLSolver::automatic,
-                                             NLSolver::newton);
-    NLSolver nonlinear_solver(additional_data, mpi_communicator);
-
-    nonlinear_solver.reinit_vector = [&](LA::MPI::Vector &x)
-                                     {/*...*/};
-    nonlinear_solver.residual = [&](const LA::MPI::Vector &evaluation_point,
-                                    LA::MPI::Vector &      residual)
-                                {/*...*/};
-    nonlinear_solver.setup_jacobian = [&](const LA::MPI::Vector &current_u)
-                                      {/*...*/};
-    nonlinear_solver.solve_with_jacobian = [&](const LA::MPI::Vector &rhs,
-                                               LA::MPI::Vector &      dst,
-                                               const double tolerance)
-                                           {/*...*/};
-
-    nonlinear_solver.solve(current_solution);
-\end{c++}
+Not that while the selector class provides a unified interface and therefore allows to
+easily switch between backends, there are cases when users need to use the underlying
+wrapper classes, if, e.g., the functionality is only provided by one implementation.
 
@@
   \end{subfigure}
-  \caption{\it Example of non-nested multigrid levels.}\label{fig:nonnested}
+  \caption{\it Example of non-nested multigrid levels.\todo[inline]{Make gap uniform!}}\label{fig:nonnested}
 \end{figure}
@@
 two_level_transfer.reinit(dof_handler_fine, dof_handler_coarse,
-                          // the following parameters are optional
                           mapping_fine, mapping_coarse,
@@
 The resulting object can be passed to \texttt{MGTransferGlobalCoarsening} (and
-\texttt{MGTransferBlockGlobalCoarsening}) in the usual way. The
+\texttt{MGTransferBlockGlobal\allowbreak{}Coarsening}) in the usual way. The
 interface of the new class is purposefully similar to
-\texttt{MGTwoLevelTransfer},
+\texttt{MGTwoLevel\allowbreak{}Transfer},
 with the difference that the former (i) also
@@
 
-At the time of writing, the \texttt{MGTwoLevelTransferNonNested} operator provides a prolongation operator $\mathcal{P}^{(f,c)}$
-that prolongates a vector $x^{(c)}$ from the coarse space to the fine space:
-\begin{align*}
-x^{(f)} = \mathcal{P}^{(f, c)}  x^{(c)},
-\end{align*}
+At the time of writing, the \texttt{MGTwoLevelTransferNonNested} operator provides a prolongation operator
+that prolongates a vector from the coarse space to the fine space,
 by a pointwise interpolation (injection) from the coarse mesh to the support points of the fine mesh. This is done
 by looping over all coarse cells and interpolating to all (fine support) points that fall into the cell with
-\texttt{FEPointEvaluation}. The local result is scattered into $x^{(f)}$, which is finalized by a
+\texttt{FEPointEvaluation}. The local result is scattered into the fine vector, which is finalized by a
 communication and a weighting step that is necessary as multiple elements could add to the same global entry.
 
-The current implementation works for scalar and vectorial, continuous (\texttt{FE\_Q})
+The current implementation works for scalar, continuous (\texttt{FE\_Q})
 and discontinuous (\texttt{FE\_DGQ}) elements on hypercube-shaped cells. We plan to support
-simplex elements and to add support
+vectorial and simplex elements and to add support
 for projection, in addition to the pointwise interpolation outlined
@@
 
-\todo[inline]{Matthias: Write}
+\todo[inline]{Matthias: Write?}
 
@@
   \texttt{DoFHandler} class now implements all $hp$-related
-  functionality, as it has for the past several releases.
+  functionality, as it has for the past two releases.
 
@@
 
-  \item
+%  \item
   This rework also affects the serialization process of active FE
@@
 \item Several old interfaces to \texttt{MatrixFree} have been removed, e.g.,
-  initialization functions without \texttt{Mapping} argument, the query
-  to the number of cell batches, \texttt{DoFHandler} objects. In
+  initialization functions without \texttt{Mapping} argument and the query
+  to the number of cell batches. In
   each case, new interfaces are available.
-\todo[inline]{The sentence part ``e.g., initialization
-  functions...objects'' does not sound grammatically correct. Should
-  it have been ``and the functions querying \texttt{DoFHandler}
-  objects''? Please fix and remove this box.}
 \end{itemize}

### 2023-06-29 9.5/paper.tex
Extend info

@@
 \item Several old interfaces to \texttt{MatrixFree} have been removed, e.g.,
-  initialization functions without \texttt{Mapping} argument and the query
-  to the number of cell batches. In
+  initialization functions without \texttt{Mapping} argument and some queries
+  to the number of cell batches as well as \texttt{DoFHandler} objects. In
   each case, new interfaces are available.

### 2023-06-29 9.5/paper.tex
Adjust spaces
Update funding for LH and MF.

@@
     \includegraphics[width=\textwidth]{png/mesh_0.png}
-  \end{subfigure}\qquad
+  \end{subfigure}
   \hfill
@@
   \end{subfigure}
-  \caption{\it Example of non-nested multigrid levels.\todo[inline]{Make gap uniform!}}\label{fig:nonnested}
+  \caption{\it Example of non-nested multigrid levels.}\label{fig:nonnested}
 \end{figure}
@@
 
-\todo[inline]{Marco: Anything to add here?.}
-
 J.~Heinz was supported by the European Union’s Framework Programme for Research
@@
 
-L.~Heltai was partially supported by the Italian Ministry of Instruction,
-University and Research (MIUR), under the 2017 PRIN project NA-FROM-PDEs MIUR
-PE1, ``Numerical Analysis for Full and Reduced Order Methods for the efficient
-and accurate solution of complex systems governed by Partial Differential
-Equations''.
+L.~Heltai and M~.Feder were partially supported by the Italian Ministry of
+University and Research (MUR), under the grant MUR PRIN 2022 No. 2022WKWZA8
+``Immersed methods for multiscale and multiphysics problems (IMMEDIATE)''.
 

### 2023-07-06 9.5/paper.tex
Continue

@@
   \item C++ language modernization (see Section~\ref{sec:language}).
+  \item Build-system modernization (see Section~\ref{sec:buildsystem}).
 \end{itemize}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Uniform interface for nonlinear solvers}\label{sec:nonlinear}
+
+With the current release, \dealii{} now supports solvers for nonlinear problems
+provided by several different external packages:
+(i) \kinsol{} (part of \sundials{}); (ii) \snes{} (part of \petsc{}, see Section~\ref{sec:petsc}); and (iii) \texttt{NOX} (part of \trilinos{}, see Section~\ref{sec:trilinos}).
+The wrappers are provided by the classes \texttt{PETScWrappers::NonlinearSolver\allowbreak{}<VectorType>}, \texttt{SUNDIALS::KINSOL<VectorType>}, and \texttt{TrilinosWrappers::NOXSolver<Vector\allowbreak{}Type>}, respectively.
+
+All three of these classes have a very similar interface, except that they vary in the kind of
+algorithms and parameters offered. The new class \texttt{NonlinearSolverSelector<VectorType>} provides a wrapper on top of
+the three external solvers with a unified interface. The user can either let
+\dealii{} decide which of the packages to use (depending on the current availability of
+the external packages) or specify it manually.
+
+The following code snippet shows a complete example using the new class for applying a Newton solver
+by automatically choosing one of the available packages:
+\begin{c++}
+using NLS = NonlinearSolverSelector<VectorType>;
+
+NLS::AdditionalData additional_data(
+  NLS::automatic, // other options: kinsol, nox, petsc_snes 
+  NLS::newton);
+
+NLS nonlinear_solver(additional_data, mpi_communicator);
+
+solver.reinit_vector       = [ ](VectorType &x) {/*...*/};    
+solver.residual            = [ ](const VectorType &src,
+                                 VectorType       &dst) {/*...*/};
+solver.setup_jacobian      = [&](const VectorType &src) {/*...*/};
+solver.apply_jacobian      = [&](const VectorType &src,
+                                 VectorType       &dst) {/*...*/};
+solver.solve_with_jacobian = [&](const VectorType &src,
+                                 VectorType       &dst,
+                                 const double      tol) {/*...*/};
+
+solver.solve(current_solution);    
+\end{c++}
+
+Note that while the selector class provides a unified interface and therefore allows to
+easily switch between backends, there are cases when users need to use the underlying
+wrapper classes, if, e.g., the functionality is only provided by one implementation.
+Details of each of the implementations are discussed in subsequent sections.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Integration of \kokkos}\label{sec:kokkos}
@@
 \kokkos \cite{trott2022} is a C++ library that enables the creation of
-performance portable applications for all major high-performance computing (HPC)
+performance portable applications for all major high-performance computing
 platforms. It implements a programming model that allows developers to write
@@
 time, then the library will fall back on a copy of \kokkos stored in
-the \texttt{bundled/} directory in the same way as we already
+the \texttt{bundled} directory in the same way as we already
 interface with \boost.
@@
 The resulting object can be passed to \petsc and used within the nonlinear solver hierarchy.
-See \ref{sec:callbacks} for additional information on such kind of callbacks.
+See Section~\ref{sec:callbacks} for additional information on such kind of callbacks.
 
@@
   preconditioner whenever setup\_jacobian() is called.}
-
-\subsubsection{Uniform interface for nonlinear solvers}\label{sec:nonlinear}
-
-With the current release, \dealii{} now supports solvers for nonlinear problems
-provided by several different external packages:
-(i) \kinsol{} (part of \sundials{}); (ii) \snes{} (part of \petsc{}, see Section~\ref{sec:petsc}); and (iii) \texttt{NOX} (part of \trilinos{}, see Section~\ref{sec:trilinos}).
-The wrappers are provided by the classes \texttt{PETScWrappers::NonlinearSolver\allowbreak{}<VectorType>}, \texttt{SUNDIALS::KINSOL<VectorType>}, and \texttt{TrilinosWrappers::NOXSolver<Vector\allowbreak{}Type>}, respectively.
-
-All three of these classes have a very similar interface, except that they vary in the kind of
-algorithms and parameters offered. The new class \texttt{NonlinearSolverSelector<VectorType>} provides a wrapper on top of
-the three external solvers with a unified interface. The user can either let
-\dealii{} decide which of the packages to use (depending on the current availability of
-the external packages) or specify it manually.
-
-The following code snippet shows a complete example using the new class for applying a Newton solver
-by automatically choosing one of the available packages:
-\begin{c++}
-using NLS = NonlinearSolverSelector<VectorType>;
-
-NLS::AdditionalData additional_data(
-  NLS::automatic, // other options: kinsol, nox, petsc_snes 
-  NLS::newton);
-
-NLS nonlinear_solver(additional_data, mpi_communicator);
-
-solver.reinit_vector       = [ ](VectorType &x) {/*...*/};    
-solver.residual            = [ ](const VectorType &src,
-                                 VectorType       &dst) {/*...*/};
-solver.setup_jacobian      = [&](const VectorType &src) {/*...*/};
-solver.apply_jacobian      = [&](const VectorType &src,
-                                 VectorType       &dst) {/*...*/};
-solver.solve_with_jacobian = [&](const VectorType &src,
-                                 VectorType       &dst,
-                                 const double      tol) {/*...*/};
-
-solver.solve(current_solution);    
-\end{c++}
-
-Not that while the selector class provides a unified interface and therefore allows to
-easily switch between backends, there are cases when users need to use the underlying
-wrapper classes, if, e.g., the functionality is only provided by one implementation.
 
@@
 In the current release, we added experimental support to compute intersections on \texttt{parallel::distributed::Triangulation} objects using \texttt{CGAL}~\cite{cgal-user-ref}.
-For this purpose, we introduced a free function \texttt{distributed\_compute\_intersection\_locations()} that computes intersections and relevant information for communication from \texttt{intersection\_requests}.
-\texttt{intersection\_requests} is a vector indicating entities of a given triangulation that intersections are computed upon.
+For this purpose, we introduced a free function \texttt{distributed\_compute\_intersection\_locations()} that computes intersections and relevant information for communication from \texttt{intersection\_re\-quests}.
+The data structure \texttt{intersection\_requests} is a vector indicating entities of a given triangulation that intersections are computed upon.
 Each entity (face or cell) is described by a vector of vertices.
-Currently, the function is placed in \texttt{GridTools::internal}, but the location and arguments of the function might still change in the future.
-To compute intersections between two geometric entities, the function internally uses the new function \texttt{CGALWrappers::compute\_intersection\_of\_cells()}.
+Currently, the function is placed in \texttt{GridTools::internal}, since the location and arguments of the function might still change in the future.
+To compute intersections between two geometric entities, the function internally uses the new function \texttt{CGALWrappers::compute\_intersec\-tion\_of\_cells()}.
 
@@
 This conversion is triggered by the member function~\texttt{convert\_to\_distributed\_compute\_point\_locations\_internal()}, given the number of quadrature points per intersection.
-The whole procedure is done communication-free.
-
+The whole procedure is done without communication.
+%
 This functionality is useful since \texttt{RemotePointEvaluation} can now be used to access quantities at quadrature points on intersections without further ado.

### 2023-07-06 9.5/paper.tex
Last edits.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Uniform interface for nonlinear solvers}\label{sec:nonlinear}
-
-With the current release, \dealii{} now supports solvers for nonlinear problems
-provided by several different external packages:
-(i) \kinsol{} (part of \sundials{}); (ii) \snes{} (part of \petsc{}, see Section~\ref{sec:petsc}); and (iii) \texttt{NOX} (part of \trilinos{}, see Section~\ref{sec:trilinos}).
-The wrappers are provided by the classes \texttt{PETScWrappers::NonlinearSolver\allowbreak{}<VectorType>}, \texttt{SUNDIALS::KINSOL<VectorType>}, and \texttt{TrilinosWrappers::NOXSolver<Vector\allowbreak{}Type>}, respectively.
-
-All three of these classes have a very similar interface, except that they vary in the kind of
-algorithms and parameters offered. The new class \texttt{NonlinearSolverSelector<VectorType>} provides a wrapper on top of
-the three external solvers with a unified interface. The user can either let
-\dealii{} decide which of the packages to use (depending on the current availability of
-the external packages) or specify it manually.
-
-The following code snippet shows a complete example using the new class for applying a Newton solver
-by automatically choosing one of the available packages:
-\begin{c++}
-using NLS = NonlinearSolverSelector<VectorType>;
-
-NLS::AdditionalData additional_data(
-  NLS::automatic, // other options: kinsol, nox, petsc_snes 
-  NLS::newton);
-
-NLS nonlinear_solver(additional_data, mpi_communicator);
-
-solver.reinit_vector       = [ ](VectorType &x) {/*...*/};    
-solver.residual            = [ ](const VectorType &src,
-                                 VectorType       &dst) {/*...*/};
-solver.setup_jacobian      = [&](const VectorType &src) {/*...*/};
-solver.apply_jacobian      = [&](const VectorType &src,
-                                 VectorType       &dst) {/*...*/};
-solver.solve_with_jacobian = [&](const VectorType &src,
-                                 VectorType       &dst,
-                                 const double      tol) {/*...*/};
-
-solver.solve(current_solution);    
-\end{c++}
-
-Note that while the selector class provides a unified interface and therefore allows to
-easily switch between backends, there are cases when users need to use the underlying
-wrapper classes, if, e.g., the functionality is only provided by one implementation.
-Details of each of the implementations are discussed in subsequent sections.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Integration of \kokkos}\label{sec:kokkos}
@@
 time, then the library will fall back on a copy of \kokkos stored in
-the \texttt{bundled} directory in the same way as we already
+the \texttt{bundled/} directory in the same way as we already
 interface with \boost.
@@
 at the time Kokkos was configured.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Uniform interface for nonlinear solvers}\label{sec:nonlinear}
+
+With the current release, \dealii{} now supports solvers for nonlinear problems
+provided by several different external packages:
+(i) \kinsol{} (part of \sundials{}); (ii) \snes{} (part of \petsc{}, see Section~\ref{sec:petsc}); and (iii) \texttt{NOX} (part of \trilinos{}, see Section~\ref{sec:trilinos}).
+The wrappers are provided by the classes \texttt{PETScWrappers::NonlinearSolver\allowbreak{}<VectorType>}, \texttt{SUNDIALS::KINSOL<VectorType>}, and \texttt{TrilinosWrappers::NOXSolver<Vector\allowbreak{}Type>}, respectively.
+
+All three of these classes have a very similar interface, except that they vary in the kind of
+algorithms and parameters offered. The new class \texttt{NonlinearSolverSelector<VectorType>} provides a wrapper on top of
+the three external solvers with a unified interface. The user can either let
+\dealii{} decide which of the packages to use (depending on the current availability of
+the external packages) or specify it manually.
+
+The following code snippet shows a complete example using the new class for applying a Newton solver
+by automatically choosing one of the available packages:
+\begin{c++}
+using NLS = NonlinearSolverSelector<VectorType>;
+
+NLS::AdditionalData additional_data(
+  NLS::automatic, // other options: kinsol, nox, petsc_snes 
+  NLS::newton);
+
+NLS nonlinear_solver(additional_data, mpi_communicator);
+
+solver.reinit_vector       = [&](VectorType &x) {/*...*/};    
+solver.residual            = [&](const VectorType &src,
+                                 VectorType       &dst) {/*...*/};
+solver.setup_jacobian      = [&](const VectorType &src) {/*...*/};
+solver.apply_jacobian      = [&](const VectorType &src,
+                                 VectorType       &dst) {/*...*/};
+solver.solve_with_jacobian = [&](const VectorType &src,
+                                 VectorType       &dst,
+                                 const double      tol) {/*...*/};
+
+solver.solve(current_solution);    
+\end{c++}
+
+Note that while the selector class provides a unified interface and therefore allows to
+easily switch between backends, there are cases when users need to use the underlying
+wrapper classes, if, e.g., the functionality is only provided by one implementation.
+Details of each of the implementations are discussed in subsequent sections.
 
@@
 PETScWrappers::PreconditionShell preconditioner(/*...*/);
-
 preconditioner.vmult = [&](const VectorType &src,
@@
 The \petsc~\snes sub-package solves systems of nonlinear equations of the form $F(x) = 0$.
-The interface to \snes has been modeled on the already existing interface
-to the \kinsol solver from the \sundials package, and the nonlinear problem
-is specified with a set of callbacks that are used by the solver.
+The interface to \snes in the class \texttt{PETScWrappers::NonlinearSolver} has been modeled on the already existing interface
+to the \kinsol solver from the \sundials package: the nonlinear problem
+is specified via a set of callbacks as shown above in Section~\ref{sec:nonlinear}.
 %\begin{c++}
@@
 \dot{u} = G(t,u) &\text{(explicit)},\\
-F(t,u,\dot{u}) = 0 &\text{(implicit)}.\\
+F(t,u,\dot{u}) = 0 &\text{(implicit)}.
 \end{eqnarray*}
@@
 %
-We refer interested readers to our documentation for other more advanced
-functions of this wrapper which are related to the differences in features of these libraries.
-In particular, we have added the possibility to reuse the preconditioner between
-nonlinear steps (also known as \textit{preconditioner lagging}), which is
-unfortunately natively only  supported in the official \texttt{EPetra} implementations.
-
-\todo[inline]{I don't understand this last sentence. Can you
-  elaborate? At least in the \kinsol{} interfaces, you only update the
-  preconditioner whenever setup\_jacobian() is called.}
+As for the \snes{} interface above, the \texttt{NOX} interface is
+drive by the callbacks shown in Section~\ref{sec:nonlinear}, and
+implemented in the \texttt{TrilinosWrappers::NOXSolver} class.
 
@@
 external packages that are largely driven via \textit{callbacks} --
-see for example the code example in Section~\ref{sec:trilinos} on how
-\texttt{NOX} gains information about the residual vector, how it indicates to
-user code that the Jacobian matrix needs to be rebuilt,
-and how to solve linearized systems of equations as the sub-steps of
-solving a nonlinear problem. Indeed, the interfaces to the nonlinear
-solver \kinsol{} (as part of \sundials{}) and \snes{} (as part of \petsc, see
-Section~\ref{sec:petsc} function in exactly the same way; the
-ODE solver interfaces to \arkode{} and \ida{} (also part of \sundials{}) and \ts{}
-(as part of \petsc{}) also use this style.
+see for example the code examples for nonlinear or ODE solvers in the previous sections.
 
@@
 recoverable failure with a positive value, and an irrecoverable failure
-with a negative value. Neither of these conventions mesh well with C++
+with a negative value. Newer PETSc versions can deal with recoverable
+failures as well, in some cases by calling back into PETSc from a
+callback to set an error flag, in others by setting the elements of a
+returned vector to NaN.
+None of these conventions mesh well with C++
 where error codes are generally indicated via exceptions.
@@
 \item In release 9.3, we enabled parallel $hp$-operations in the matrix-free infrastructure.
-The infrastructure did not work properly in the case that certain cells
-did not get any degrees of freedom due to the usage of \texttt{FE\_Nothing}. This has been
+The infrastructure did not work properly for cells
+that do not have any degrees of freedom because they use \texttt{FE\_Nothing}. This has been
 fixed now. Furthermore, \texttt{FE\_Nothing} now also works together
@@
 \texttt{ElementActivationAndDeactivationMatrixFree}, which wraps a \texttt{MatrixFree} object, only loops over all
-active cells and optionally interprets faces between active and deactivated cells
+active cells, and optionally interprets faces between active and deactivated cells
 as boundary faces. This functionality has enabled simulations in powder-bed-fusion additive
@@
 
-\dealii{} currently uses C++14 as the language standard to which its
+This version of \dealii{} uses C++14 as the language standard to which its
 code base is written. It can also use the classes
@@
 
+The next release of \dealii{} will build upon C++17.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 While there are no new \dealii tutorial programs in this release, many
-were extensively revised: Around 145 of the more than 2100 (non-merge)
+were extensively revised: Around 145 of the more than 2200 (non-merge)
 commits that went into this release touched the tutorial, in some
@@
 \href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_4_2_and_9_5_0.html}
-     {around XXX incompatible changes};
-\todo{Fix number before release}
+     {around 35 incompatible changes};
 see \cite{changes95}. Many of these

### 2023-07-06 9.5/paper.tex
write second CMake paragraph

@@
 
-\todo[inline]{Matthias: Write?}
+\dealii's CMake project configuration itself now exports three targets,
+\texttt{dealii::dealii\_debug}, \texttt{dealii::dealii\_release}, and
+\texttt{dealii::dealii}. Linking against one of the first two variants via
+\texttt{target\_link\_libraries()} populates the target with all necessary
+include directories and the full link interface necessary for \dealii. The
+third variant, \texttt{dealii::dealii}, automatically switches between
+debug and release interface depending on the build type set via
+\texttt{CMAKE\_BUILD\_TYPE}. The \texttt{dealii::dealii} third variant also
+populates compiler and linker options, whereas for the first two variants a
+client project has to ensure that the compiler and linker are properly set
+up. The new import targets make it now possible to configure a dependent
+project without the use of any \dealii specific macros:
+%
+\begin{anycode}
+cmake_minimum_required(VERSION 3.13.4)
+set(CMAKE_BUILD_TYPE Debug CACHE STRING "")
+project(step CXX)
+
+find_package(deal.II 9.5.0 REQUIRED)
+
+add_executable(step step.cc)
+target_link_libraries(step dealii::dealii)
+\end{anycode}
 

### 2023-07-06 9.5/paper.tex
write first CMake paragraph

@@
 
+\dealii's original CMake build system that was merged for the 8.0.0 release
+in 2013 was written for CMake 2.8.8, which, naturally, lacks many of the
+modern features introduced into CMake over the last couple of years. This
+created a particular issue with an increasing number of dependencies
+switching their CMake configuration to \emph{import targets}, where all
+necessary information for using an external resource, such as include
+directories and library link interfaces, are associated with an imported
+CMake targets rather than being provided with individual (and
+inconsistently) named CMake variables. With the \dealii 9.5 release the
+handling of external dependencies has been rewritten entirely to support
+imported targets. The \trilinos and \kokkos interfaces have been
+modernized. We plan to migrate all remaining dependencies to import targets
+wherever possible for the next release.
+
 \dealii's CMake project configuration itself now exports three targets,

### 2023-07-07 9.5/paper.tex
List Stefano's acknowledgments. Plus a couple of edits.

--- 2023-07-06 9.5/paper.tex
@@
 
+\dealii's original CMake build system, merged for the 8.0.0 release
+in 2013, was written for CMake 2.8.8 and thus lacks many of the
+modern features introduced into CMake over the last several years. This
+created a particular issue with an increasing number of dependencies
+switching their CMake configuration to \emph{import targets}, where all
+necessary information for using an external resource, such as include
+directories and library link interfaces, are associated with an imported
+CMake target rather than being provided with individual (and
+inconsistently) named CMake variables. With the \dealii 9.5 release the
+handling of external dependencies has been rewritten entirely to support
+imported targets. The \trilinos and \kokkos interfaces have been
+modernized. We plan to migrate all remaining dependencies to import targets
+wherever possible for the next release.
+
 \dealii's CMake project configuration itself now exports three targets,
@@
 client project has to ensure that the compiler and linker are properly set
-up. The new import targets make it now possible to configure a dependent
+up. The new import targets now make it possible to configure a dependent
 project without the use of any \dealii specific macros:
@@
 
-\todo[inline]{Stefano: Anything to add here?.}
+S. Zampini was supported by the KAUST Extreme Computing Research Center.
 

### 2023-07-07 9.5/paper.tex
Remove a TODO. I cannot find a reference for NOX on their website.
Specify the number of changelog items
Fix the release date.

@@
 
-\dealii version 9.5.0 was released June XX, 2023.
-\todo{Add exact date.}
+\dealii version 9.5.0 was released July 7, 2023.
 This paper provides an
@@
 %
-The changelog lists more than X other features and bugfixes.
+The changelog lists more than 120 other features and bugfixes.
 
@@
     \item ROL \cite{ridzal2014rol}
-      \todo{ROL is part of Trilinos. Do we also need to cite something
-        for NOX?}
     \item ScaLAPACK \cite{blackford1997scalapack}

### 2023-07-08 9.5/paper.tex
Update names of contributors

@@
 Luz         Paz,
-Laura       Prieto,
+Laura       Prieto Saavedra,
 Sebastian   Proell,
@@
 Christoph   Schmidt,
-Magdalena   Schreter,
+Magdalena   Schreter-Fleischhacker,
 Richard     Schussnig,

### 2023-07-19 9.5/paper.tex
Mention that step-64 can be used on the CPU

@@
 \texttt{CUDAWrappers::Matrix\allowbreak{}Free} framework are using \kokkos. This allows them 
-to work on all the architectures supported by \kokkos. The \kokkos backend used 
-by \dealii is \texttt{Kokkos::DefaultExecutionSpace} which corresponds to the 
-highest available backend in the hierarchy device, host-parallel, and host-serial 
+to work on all the architectures supported by \kokkos. In particular, the step-64 example
+can now also be run on the CPU and doesn't require a GPU anymore. \\
+The \kokkos backend used by \dealii is \texttt{Kokkos::DefaultExecutionSpace} which corresponds
+to the highest available backend in the hierarchy device, host-parallel, and host-serial 
 at the time Kokkos was configured.

### 2023-07-27 9.5/paper.tex
computing acks

@@
 (CIG), through the National Science Foundation (NSF) under Award
-No.~EAR-1550901 and The University of California -- Davis.
+No.~EAR-1550901 and EAR-2149126 via The University of California -- Davis.
 
@@
 
+Clemson University is acknowledged for generous allotment of compute time on Palmetto cluster.
+
+The authors acknowledge the Texas Advanced Computing Center (TACC) at The University of Texas at Austin for providing HPC resources that have contributed to the research results reported within this paper.
+\url{http://www.tacc.utexas.edu}
+
+This work used the Extreme Science and Engineering Discovery Environment (XSEDE), which is supported by National Science Foundation grant number ACI-1053575 access through the CIG Science Gateway and Community Codes for the Geodynamics Community MCA08X011 allocation.
+
 \bibliography{paper}{}

### 2023-12-28 9.5/paper.tex
Make the list of contributors easier to read in the flow of
the paper.

@@
 % This is up-to-date as of 2023-06-26 7:00pm central
-
+\begin{quote}
 Pasquale    Africa,
@@
 Jiaqi       Zhang.
-
+\end{quote}
 Their contributions are much appreciated!

### 2024-02-28 2024-licensing_acknowledgment/relicensing.tex
add licensing acknowledgment for relicensing

\RequirePackage{fix-cm}
\documentclass[oneside, 12pt, letterpaper]{article}
\usepackage{fixltx2e}[2006/03/24]

\usepackage{geometry}
\geometry{left=1.5in, right=1.5in, top=1.25in, bottom=1.0in}
\usepackage{parskip}
\usepackage{graphicx}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[american]{babel}
\usepackage{enumitem}
\usepackage{amssymb}

% \usepackage[colorlinks=true,linkcolor=black,citecolor=black,urlcolor=black]{hyperref}
\usepackage{url}

\usepackage{microtype}
\microtypesetup{
  tracking=true,
  kerning=true
}
\microtypecontext{spacing=nonfrench}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\thispagestyle{empty}
\enlargethispage{3em}

\begin{center}
  {\bfseries deal.II Licensing Acknowledgment}
\end{center}

By signing below, I acknowledge and agree to the following:

{\bfseries (a)} "{\bfseries Contributions to the deal.II Project}" shall
mean any copyrightable material (including, but not limited to, software
source code, documentation source, configuration files, data, and images)
that I have made available to the deal.II project whether provided before,
on, or after the date hereof. This specifically includes prior
contributions to
\vspace{-0.5em}
\begin{itemize}[noitemsep]
  \item[--]
    the {\bfseries deal.II source code repository} currently hosted at
    \url{https://github.com/dealii/dealii} as part of the development
    history leading up to commit
    \texttt{65d590d6e8fb94ce1fd704d69cc1ed5b874857fb}, and the
  \item[--]
    the {\bfseries website repository} currently hosted at
    \url{https://github.com/dealii/old-website} as part of the development
    history leading up to commit
    \texttt{6e54856bae6ed17385989d7b6f08a659a5fa2d43},
\end{itemize}
\vspace{-0.5em}
but also any pending or future pull requests to these repositories.

{\bfseries (b)} The "{\bfseries Apache License 2.0 with LLVM-exception}"
and "{\bfseries GNU Lesser General Public License version 2.1 or later}"
are open source licenses whose full license texts are publicly available at
\vspace{-0.5em}
\begin{itemize}[noitemsep]
  \item[--] \url{https://spdx.org/licenses/Apache-2.0.html}
  \item[--] \url{https://spdx.org/licenses/LLVM-exception.html}
  \item[--] \url{https://spdx.org/licenses/LGPL-2.1-or-later.html}
\end{itemize}
\vspace{-0.5em}
The full license texts are also available in the file \texttt{LICENSE.md}
in the source code repository with commit
\texttt{65d590d6e8fb94ce1fd704d69cc1ed5b874857fb}.

{\bfseries (c)} I am legally entitled to grant the licenses granted herein.

{\bfseries (d)} I hereby license all {\bfseries Contributions to the
deal.II Project} under the terms of the {\bfseries Apache License 2.0 with
LLVM-exception}. Furthermore, I affirm that my {\bfseries Contributions to
the deal.II Project} which have been previously licensed under the terms of
the {\bfseries GNU Lesser General Public License v2.1 or later} will
continue to also be available under these terms.

{\bfseries (e)} I choose to {\bfseries $\square$\, include /
$\square$\, exclude} my contributions to the {\bfseries Code Gallery
repository} currently hosted at
\url{https://github.com/dealii/code-gallery} into/from this license grant.
If no selection is made, then ``exclude'' applies.
\vspace{0.5em}

Full name:
\vspace{0.5em}

E-Mail address(es):
\vspace{2.75em}

Date:
\vspace{0.5em}

Signature:

\end{document}

### 2024-06-01 9.6/paper.tex
First draft of 9.6 release paper

--- 2023-12-28 9.5/paper.tex
@@
     Wolfgang Bangerth,
-    Maximilian Bergbauer,
-    Marco Feder,
     Marc Fehling,
@@
     Peter Munch,
-    Jean-Paul Pelteret,
     Bruno Turcksin,
-    David Wells,
-    Stefano Zampini
+    David Wells
   },
-  pdftitle={The deal.II Library, Version 9.5, 2023},
+  pdftitle={The deal.II Library, Version 9.6, 2024},
 }
 
-\title{The \dealii Library, Version 9.5}
+\title{The \dealii Library, Version 9.6}
 
@@
 
-\author[4]{Maximilian~Bergbauer}
-\affil[4]{Institute for Computational Mechanics,
-Technical University of Munich,
-Boltzmannstr.~15, 85748 Garching, Germany.
-{\texttt{maximilian.bergbauer@tum.de}}}
+\author[4]{Marc~Fehling}
+ \affil[4]{Charles University, Czech Republic}
 
-\author[5]{Marco~Feder}
-\affil[5]{SISSA,
-International School for Advanced Studies,
-Via Bonomea 265,
-34136, Trieste, Italy.
-{\texttt{marco.feder/luca.heltai@sissa.it}}}
-
-\author[2]{Marc~Fehling}
-
-\author[6]{Johannes~Heinz}
- \affil[6]{Institute of Mechanics and Mechatronics,
-   TU Wien,
-   Getreidemarkt 9, 1060 Vienna, Austria
-   {\texttt{johannes.heinz@tuwien.ac.at}}}
-
-\author[7]{Timo~Heister}
- \affil[7]{School of Mathematical and Statistical Sciences,
+\author[5]{Timo~Heister}
+ \affil[5]{School of Mathematical and Statistical Sciences,
    Clemson University,
@@
 
-\author[5]{Luca~Heltai}
+\author[6]{Luca~Heltai}
+\affil[6]{University of Pisa, Italy}
 
- \author[8]{Martin~Kronbichler}
- \affil[8]{Institute of Mathematics,
-   University of Augsburg,
-   Universit\"atsstr.~12a, 86159 Augsburg, Germany.
-   {\texttt{martin.kronbichler/peter.muench@uni-a.de}}}
+ \author[7]{Martin~Kronbichler}
+ \affil[7]{Ruhr University Bochum, Germany}
 
-\author[9]{Matthias~Maier}
-\affil[9]{Department of Mathematics,
+\author[8]{Matthias~Maier}
+\affil[8]{Department of Mathematics,
   Texas A\&M University,
@@
 
-\author[8,10]{Peter Munch}
- \affil[10]{Institute of Material Systems Modeling,
- Helmholtz-Zentrum Hereon,
- Max-Planck-Str. 1, 21502 Geesthacht, Germany}
-
-\author[11]{Jean-Paul~Pelteret}
-\affil[11]{Independent researcher.
-{\texttt{jppelteret@gmail.com}}}
+\author[9]{Peter Munch}
+ \affil[9]{Uppsala University, Sweden.
+  {\texttt{peter.munch@it.uu.se}}}
 
@@
 
-\author[12]{David Wells}
-\affil[12]{Department of Mathematics, University of North Carolina,
+\author[10]{David Wells}
+\affil[10]{Department of Mathematics, University of North Carolina,
   Chapel Hill, NC 27516, USA.
   {\texttt{drwells@email.unc.edu}}}
-
-\author[13]{Stefano Zampini}
-\affil[13]{Extreme Computing Research Center, King Abdullah University of Science and Technology,
-  23955-6900 Thuwal, Saudi Arabia.
-  {\texttt{stefano.zampini@gmail.com}}}
 
@@
   This paper provides an overview of the new features of the finite element
-  library \dealii, version 9.5.
+  library \dealii, version 9.6.
 \end{abstract}
@@
 
-\dealii version 9.5.0 was released July 7, 2023.
+\dealii version 9.6.0 was released X X, 2024.
 This paper provides an
 overview of the new features of this release and serves as a citable
-reference for the \dealii software library version 9.5. \dealii is an
+reference for the \dealii software library version 9.6. \dealii is an
 object-oriented finite element library used around the world in the
@@
 \begin{itemize}
-  \item Substantial updates and extensions to \dealii{}'s interfaces
-    to other libraries (see Section~\ref{sec:external}). This
-    includes, in particular, the integration of \kokkos (Section~\ref{sec:kokkos});
-    additions and updates to the \petsc and \trilinos
-    interfaces
-    (Sections~\ref{sec:petsc} and \ref{sec:trilinos}, respectively).
-  \item Uniform handling of nonlinear solver
-  packages (Section~\ref{sec:nonlinear}) and a uniform
-    way of defining callbacks used by external libraries (Section~\ref{sec:callbacks}).
-  \item Advances in matrix-free infrastructure (see Section~\ref{sec:mf}).
-  \item Advances in non-matching support (see Section~\ref{sec:nonmatching}).
-  \item New features related to linear algebra (see Section~\ref{sec:lac}).
-  \item C++ language modernization (see Section~\ref{sec:language}).
-  \item Build-system modernization (see Section~\ref{sec:buildsystem}).
+  \item 
 \end{itemize}
@@
 \begin{itemize}
-  \item The new function \texttt{CellAccessor::as\_dof\_handler\_iterator()}
-  simplifies the conversion from a \texttt{Cell\-Accessor} to a \texttt{DoF\-Cell\-Accessor}.
-  The old way,
-\begin{c++}
-const auto cell_dof = typename DoFHandler<dim, spacedim>::
-  active_cell_iterator(&dof_handler.get_triangulation(),
-    cell->level(), cell->index(), &dof_handler);
-\end{c++}
-was lengthy and error-prone. In contrast, the following function is
-substantially clearer and more concise:
-\begin{c++}
-const auto cell_dof = cell->as_dof_handler_iterator(dof_handler);
-\end{c++}
-\item Several functions used intensively during initialization of
-  \dealii-based programs, such as the refinement of triangulations, the
-  enumeration of degrees of freedom, the setup of global-coarsening multigrid
-  algorithms, and several evaluation functions of the \texttt{MappingQ} class
-  representing a polynomial mapping of quadrilateral and hexahedral, have been
-  overhauled to run more quickly and sometimes also consume less memory. These
-  and related improvements are guided by several performance tests that are
-  used to monitor the performance of the library over time.
+  \item changes of interface of AffineConstraints
+  \item clean up: orientation
 \end{itemize}
 %
-The changelog lists more than 120 other features and bugfixes.
+The changelog lists more than X other features and bugfixes.
 
@@
 
-%\newpage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Updates to interfaces to other packages}\label{sec:external}
-
-For many operations, \dealii{} relies on external libraries -- some of
-these are optional, others are mandatory (such as \boost and, now,
-\kokkos); a complete list of external dependencies is provided in
-Section~\ref{sec:cite}. A substantial amount of work has gone into
-overhauling and extending these interfaces for the current release, as
-detailed in the following sub-sections.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Integration of \kokkos}\label{sec:kokkos}
-
-\kokkos \cite{trott2022} is a C++ library that enables the creation of
-performance portable applications for all major high-performance computing
-platforms. It implements a programming model that allows developers to write
-code that can efficiently run on diverse architectures. \kokkos provides
-abstractions for both parallel execution of code and data management. It
-supports a wide range of backend programming models, including CUDA, HIP, SYCL,
-HPX, OpenMP, and C++ threads, and continues to evolve with the development of
-new hardware and corresponding backend options.
-
-\dealii{} has, for several releases already, used CUDA to offload some
-operations onto GPUs. It has also had interfaces to CUDA-based linear
-algebra libraries. Yet, the diversification of GPU platforms away from
-a single vendor (Nvidia) has made it clear that we need a different
-strategy to support what users want. As a consequence, \kokkos has
-become a mandatory dependency of \dealii{} as part of the current
-release; if it is not found on a given system during configuration
-time, then the library will fall back on a copy of \kokkos stored in
-the \texttt{bundled/} directory in the same way as we already
-interface with \boost.
-
-In the current release, \texttt{LinearAlgebra::distributed::Vector} and the 
-\texttt{CUDAWrappers::Matrix\allowbreak{}Free} framework are using \kokkos. This allows them 
-to work on all the architectures supported by \kokkos. In particular, the step-64 example
-can now also be run on the CPU and doesn't require a GPU anymore. \\
-The \kokkos backend used by \dealii is \texttt{Kokkos::DefaultExecutionSpace} which corresponds
-to the highest available backend in the hierarchy device, host-parallel, and host-serial 
-at the time Kokkos was configured.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Uniform interface for nonlinear solvers}\label{sec:nonlinear}
-
-With the current release, \dealii{} now supports solvers for nonlinear problems
-provided by several different external packages:
-(i) \kinsol{} (part of \sundials{}); (ii) \snes{} (part of \petsc{}, see Section~\ref{sec:petsc}); and (iii) \texttt{NOX} (part of \trilinos{}, see Section~\ref{sec:trilinos}).
-The wrappers are provided by the classes \texttt{PETScWrappers::NonlinearSolver\allowbreak{}<VectorType>}, \texttt{SUNDIALS::KINSOL<VectorType>}, and \texttt{TrilinosWrappers::NOXSolver<Vector\allowbreak{}Type>}, respectively.
-
-All three of these classes have a very similar interface, except that they vary in the kind of
-algorithms and parameters offered. The new class \texttt{NonlinearSolverSelector<VectorType>} provides a wrapper on top of
-the three external solvers with a unified interface. The user can either let
-\dealii{} decide which of the packages to use (depending on the current availability of
-the external packages) or specify it manually.
-
-The following code snippet shows a complete example using the new class for applying a Newton solver
-by automatically choosing one of the available packages:
-\begin{c++}
-using NLS = NonlinearSolverSelector<VectorType>;
-
-NLS::AdditionalData additional_data(
-  NLS::automatic, // other options: kinsol, nox, petsc_snes 
-  NLS::newton);
-
-NLS nonlinear_solver(additional_data, mpi_communicator);
-
-solver.reinit_vector       = [&](VectorType &x) {/*...*/};    
-solver.residual            = [&](const VectorType &src,
-                                 VectorType       &dst) {/*...*/};
-solver.setup_jacobian      = [&](const VectorType &src) {/*...*/};
-solver.apply_jacobian      = [&](const VectorType &src,
-                                 VectorType       &dst) {/*...*/};
-solver.solve_with_jacobian = [&](const VectorType &src,
-                                 VectorType       &dst,
-                                 const double      tol) {/*...*/};
-
-solver.solve(current_solution);    
-\end{c++}
-
-Note that while the selector class provides a unified interface and therefore allows to
-easily switch between backends, there are cases when users need to use the underlying
-wrapper classes, if, e.g., the functionality is only provided by one implementation.
-Details of each of the implementations are discussed in subsequent sections.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Updates and additions to the \petsc wrappers}\label{sec:petsc}
-
-The \dealii classes wrapping \petsc objects have been rewritten in
-substantial ways
-to support \petsc's System of Nonlinear Equations Solver \snes and the Ordinary Differential Equations (ODE)
-solver \ts \cite{abhyankar2018petsc}.
-First, we briefly describe the most important improvements to the existing
-classes and then outline the newly designed interfaces to the nonlinear solvers, together
-with a new interface to the communication module \petscsf in \petsc \cite{zhang2021petscsf}.
-
-Vector and matrix classes of the \petsc wrappers have been extended with an additional constructor
-that can wrap an already existing \petsc vector or matrix, respectively.
-The \texttt{BlockVector} and
-\texttt{BlockSparseMatrix} classes now internally use
-\petsc nested objects, i.e. \texttt{VECNEST} and  \texttt{MATNEST} respectively.
-We have added a new class \texttt{PETScWrappers::PreconditionShell} to support user-defined
-preconditioning that can be simply customized as the following code
-snippet shows:
-\begin{c++}
-PETScWrappers::PreconditionShell preconditioner(/*...*/);
-preconditioner.vmult = [&](const VectorType &src,
-                           VectorType       &dst) {/*...*/};
-
-\end{c++}
-The resulting object can be passed to \petsc and used within the nonlinear solver hierarchy.
-See Section~\ref{sec:callbacks} for additional information on such kind of callbacks.
-
-The \petsc~\snes sub-package solves systems of nonlinear equations of the form $F(x) = 0$.
-The interface to \snes in the class \texttt{PETScWrappers::NonlinearSolver} has been modeled on the already existing interface
-to the \kinsol solver from the \sundials package: the nonlinear problem
-is specified via a set of callbacks as shown above in Section~\ref{sec:nonlinear}.
-%\begin{c++}
-%PETScWrappers::NonlinearSolver<VectorType> solver(/*...*/);
-%
-%solver.residual            = [&](const VectorType &x,
-%                                 VectorType       &F) {/*...*/};
-%solver.setup_jacobian      = [&](const VectorType &x) {/*...*/};
-%solver.solve_with_jacobian = [&](const VectorType &rhs,
-%                                 VectorType       &sol) {/*...*/};
-%solver.solve(x);
-%\end{c++}
-The default configuration is set up to use a Jacobian-free Newton-Krylov (JFNK)
-approach~\cite{knoll2004jacobian}, using \texttt{solve\_with\_jacobian} as linear solver.
-Numerous other solver configurations are possible and can be selected
-programmatically using the constructor arguments,
-or via the powerful command line customization of \petsc.
-This includes, for example, Quasi-Newton methods, Anderson's acceleration, and
-nonlinear preconditioning~\cite{brune2015composing}. The step-77
-tutorial program has gained a section that discusses the use of these
-new \snes interfaces.
-
-The \petsc~\ts sub-package solves ODEs in explicit or implicit form \cite{abhyankar2018petsc}, i.e.:
-\begin{eqnarray*}
-%\frac{\partial u}{\partial t} = G(t,u) &\text{explicit},\\
-%F(t,u,\frac{\partial u}{\partial t}) = 0 &\text{implicit}\\
-\dot{u} = G(t,u) &\text{(explicit)},\\
-F(t,u,\dot{u}) = 0 &\text{(implicit)}.
-\end{eqnarray*}
-The interface to \ts has been modeled on the already existing interfaces
-to the \ida and \arkode solvers from the \sundials package. Specifically:
-\begin{c++}
-PETScWrappers::TimeDependentSolver<VectorType> solver(/*...*/);
-
-// If solving udot = G(t,u)
-solver.explicit_function = [&](const double     t,
-                               const VectorType &u,
-                               VectorType       &G) {/*...*/};
-// If solving F(t,u,udot) = 0
-solver.implicit_function = [&](const double     t,
-                               const VectorType &u,
-                               const VectorType &udot,
-                               VectorType       &F) {/*...*/};
-\end{c++}
-In addition to specifying the function callbacks, users can further customize
-the solution of the linearized equations $\alpha \frac{\partial{F}}{\partial{\dot{u}}} + \frac{\partial{F}}{\partial{u}}$
-via the following callbacks:
-\begin{c++}
-solver.setup_jacobian = [&](const double     t,
-                            const VectorType &u,
-                            const VectorType &udot,
-                            const double     alpha) {/*...*/};
-solver.solve_with_jacobian = [&](const VectorType &rhs,
-                                 VectorType       &sol) {/*...*/};
-\end{c++}
-As with \snes, the default configuration of an implicit solver is set up to
-use a JFNK approach, and the entire suite of solvers offered by \petsc is available
-programmatically or via the command line interface,
-including adaptive time-stepping and Implicit-Explicit schemes.
-
-We close this section by introducing the interface to the \petscsf sub-package,
-the abstract communication model of \petsc. The \dealii interface to
-\petscsf has been modeled on the existing \texttt{Utilities::MPI::Partitioner} and
-\texttt{Utilities::MPI::NoncontiguousPartitioner} classes, with minimal changes for the
-communication routines API; the equivalent classes based
-on \petscsf are \texttt{PETScWrappers::Partitioner} and
-\texttt{PETScWrappers::CommunicationPattern}.
-Future developments will add support for GPU buffers.
-
-We refer interested readers to our documentation for more advanced
-functions of the \snes, \ts and \petscsf wrappers, and to the {\tt tests/petsc/} folder
-for examples on how to use them.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Interfaces to \trilinos{}' \texttt{Belos} and \texttt{NOX} packages}\label{sec:trilinos}
-
-\trilinos is a large collection of individual sub-packages \cite{heroux2005trilinos,trilinos-web-page}. \dealii{}
-has long had interfaces to the \trilinos packages that provide parallel
-vector and matrix classes (both \texttt{Epetra} and \texttt{Tpetra}),
-as well as a small number of linear algebra packages for iterative
-solvers and preconditions. In the current release, there are now also
-interfaces to two additional packages: \texttt{Belos} and \texttt{NOX}.
-
-\texttt{Belos} is the
-successor of the \trilinos package \texttt{AztecOO} and provides basic and advanced iterative solvers
-that heavily rely on multivector operations. The interface of the
-wrapper is similar to \dealii{}'s own iterative solvers:
-\begin{c++}
-TrilinosWrappers::SolverBelos<VectorType> solver(/*...*/);
-solver.solve(matrix, x, r, preconditioner);
-\end{c++}
-The omitted constructor arguments allow selecting the actual iterative
-solver to be used, along with other configuration options.
-
-Secondly, we have added a wrapper to \texttt{NOX}, a
-nonlinear solver library that is similar to both the \kinsol{} solver from
-the \sundials{} package to which we already had interfaces, and also
-to the \snes{} collections of functions from \petsc (see also
-Section~\ref{sec:petsc}). 
-%The interfaces to these three solvers are
-%essentially the same and require setting function objects for
-%computing the residual of the nonlinear problem, along with setting
-%up, applying, and solving with the Jacobian matrix:
-%\begin{c++}
-%TrilinosWrappers::NOXSolver<VectorType> solver(/*...*/);
-%
-%solver.residual            = [ ](const VectorType &src,
-%                                 VectorType       &dst) {/*...*/};
-%solver.setup_jacobian      = [&](const VectorType &src) {/*...*/};
-%solver.apply_jacobian      = [&](const VectorType &src,
-%                                 VectorType       &dst) {/*...*/};
-%solver.solve_with_jacobian = [&](const VectorType &src,
-%                                 VectorType       &dst,
-%                                 const double      tol) {/*...*/};
-%
-%solver.solve(solution);
-%\end{c++}
-%
-As for the \snes{} interface above, the \texttt{NOX} interface is
-drive by the callbacks shown in Section~\ref{sec:nonlinear}, and
-implemented in the \texttt{TrilinosWrappers::NOXSolver} class.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Uniform error reporting in callbacks}\label{sec:callbacks}
-
-With the current release, \dealii{} has gained interfaces to several
-external packages that are largely driven via \textit{callbacks} --
-see for example the code examples for nonlinear or ODE solvers in the previous sections.
-
-This substantially enlarged use of callbacks used by different backend
-libraries raises the issue that each underlying package has its own
-convention on how success or error codes of these callbacks should be
-encoded. In the case of \petsc{}'s \snes{} and \ts{}, and for \trilinos{}'s \texttt{NOX},
-success is indicated by a zero integer return value, whereas failure is
-indicated by a nonzero return value. On the other hand, the \sundials{}
-packages indicate success by a zero integer return value, a
-recoverable failure with a positive value, and an irrecoverable failure
-with a negative value. Newer PETSc versions can deal with recoverable
-failures as well, in some cases by calling back into PETSc from a
-callback to set an error flag, in others by setting the elements of a
-returned vector to NaN.
-None of these conventions mesh well with C++
-where error codes are generally indicated via exceptions.
-It is conceivable that libraries we want to
-interface with in the future use yet other conventions.
-
-In order to shield those who write these callbacks from having to
-learn the intricacies of the underlying libraries, we have adopted a
-convention whereby user-provided callbacks are just regular functions
-that return errors via exceptions as is common in C++. Internally, the
-interfaces to different underlying libraries then translate these
-exceptions into the appropriate error codes, saving the thrown exception for possible
-later use; if an underlying library supports
-recoverable errors, then a callback should indicate such an error by
-throwing an object of the special type
-\texttt{RecoverableUserCallbackError}. If one of the underlying
-packages returns with an error caused by a user-provided callback
-throwing an exception, then the wrapper code re-throws the previously
-saved exception, allowing the calling site to catch it to obtain information
-about what might have gone wrong.
-
-This convention for user callbacks is documented in a glossary entry
-that is also linked to from the documentation of all variables storing
-callbacks.
+\begin{itemize}
+\item Hermite element
+\item simplices: cubic simplex elements, improved refinement of tetrahedron cells
+\item Tpetra advances
+\item Nedelec for meshes with hanging nodes
+\end{itemize}
 
@@
 
-The current release includes numerous updates to the matrix-free
-infrastructure, including:
 \begin{itemize}
-\item In release 9.3, we enabled parallel $hp$-operations in the matrix-free infrastructure.
-The infrastructure did not work properly for cells
-that do not have any degrees of freedom because they use \texttt{FE\_Nothing}. This has been
-fixed now. Furthermore, \texttt{FE\_Nothing} now also works together
-with discontinuous Lagrange elements (i.e., with the \texttt{FE\_DGQ} class). Due to the popularity of \texttt{FE\_Nothing} as a means to enable
-or disable cells, we have introduced the new class
-\texttt{ElementActivationAndDeactivationMatrixFree}, which wraps a \texttt{MatrixFree} object, only loops over all
-active cells, and optionally interprets faces between active and deactivated cells
-as boundary faces. This functionality has enabled simulations in powder-bed-fusion additive
-manufacturing in~\cite{proell2023highly}.
-\item The matrix-free infrastructure allows interleaving cell loops with vector updates
-  by providing \texttt{pre}/\texttt{post} functions that are run on index ranges.  The \dealii
-  library uses this feature, e.g., to improve the performance of (preconditioned)
-conjugate gradient solvers~\cite{kronbichler2022cg} as well as of relaxation and Chebyshev iterations (see Subsection~\ref{sec:lac}).
-Up to release~9.3, the \texttt{pre}/\texttt{post} infrastructure was only supported for
-continuous elements (cell loop); now, it also works for discontinuous
-elements which also require face loops to assemble jump and penalty terms.
-\item The operator \texttt{CellwiseInverseMassMatrix} now also efficiently
-evaluates the inverse for coupling (dyadic) coefficients in the case of multiple
-components:
-\begin{align*}
-\left(v_i, D_{ij} u_j  \right)_{\Omega^{(K)}}
-\quad  1\le i,j \le c,
-\end{align*}
-with $c$ being the number of components and $D\in \mathbb{R}^{c\times c}$ a tensorial
-coefficient. The algorithm relies on the construction of the element mass matrix,
-\begin{align*}
-M =  ( I_1 \otimes N^T) (D \otimes I_2 ) ( I_1 \otimes N),
-\end{align*}
-with $N$ being the tabulated values of shape functions at quadrature and
-$I_1,I_2$ identity matrices associated to $c$ vector components and the
-quadrature points, respectively.
-The algorithms assumes a square $N$:
-\begin{align*}
-M^{-1} =  ( I_1 \otimes N^{-1}) (D^{-1} \otimes I_2 ) ( I_1 \otimes N^{-T}).
-\end{align*}
-For hypercube-shaped cells, $N^{-1}$ has an explicit representation
-again in terms of tensor products; for example, in 3d it can be
-expressed as $N^{-1} = N_{1D}^{-1} \otimes N_{1D}^{-1} \otimes
-N_{1D}^{-1}$, allowing the use sum factorization \cite{kronbichler2016comparison}.
+\item VectorizedArray for ARM Neon
 \end{itemize}
@@
 
-The (matrix-free) non-matching support of \dealii heavily relies
-on the classes \texttt{FE\-Point\-Eval\-u\-ation} and \texttt{RemotePointEvaluation},
-which were introduced in release 9.3. While \texttt{FE\-Point\-Eval\-u\-a\-tion}
-is responsible for efficient evaluation/integration at arbitrary (reference)
-points within a cell, \texttt{RemotePointEvaluation} is responsible for
-sorting points with regards to the cells they reside in, and for the
-communication necessary to evaluate solutions on cells owned by other
-MPI processes.
-
-In the current release, we have considerably optimized \texttt{FEPointEvaluation}, e.g,
-by caching the evaluated shape functions, templating loop bounds, and
-exploiting the tensor-product structure of the shape functions if all points are
-positioned on a face, a common use case in the context of fluid-structure
-interaction. Furthermore, the extended class \texttt{NonMatching::MappingInfo}
-allows for precomputing and storing metric terms, like the Jacobian, its determinant,
-or the unit outer normal vectors. This is useful in cases in which these metric terms do not change
-and can be reused, e.g., in the context of iterative solvers. This development
-is part of an effort to make the interfaces of the
-(matrix-free) non-matching support more similar to the ones of the
-established matrix-free infrastructure of \dealii for fixed quadrature formulas.
-
-In addition to these node-level performance optimizations, we added experimental
-support for (i) generating intersections of distributed
-non-matching grids and working on them,
-and (ii) multigrid with non-nested levels. In the following, we describe these
-in more detail.
-
-\subsubsection{Intersected meshes}
-In the current release, we added experimental support to compute intersections on \texttt{parallel::distributed::Triangulation} objects using \texttt{CGAL}~\cite{cgal-user-ref}.
-For this purpose, we introduced a free function \texttt{distributed\_compute\_intersection\_locations()} that computes intersections and relevant information for communication from \texttt{intersection\_re\-quests}.
-The data structure \texttt{intersection\_requests} is a vector indicating entities of a given triangulation that intersections are computed upon.
-Each entity (face or cell) is described by a vector of vertices.
-Currently, the function is placed in \texttt{GridTools::internal}, since the location and arguments of the function might still change in the future.
-To compute intersections between two geometric entities, the function internally uses the new function \texttt{CGALWrappers::compute\_intersec\-tion\_of\_cells()}.
-
-For the common case of Nitsche-type mortaring, quadrature points must be distributed on the intersections to evaluate the underlying physical coupling terms.
-The data structure returned by \texttt{distributed\_compute\_intersection\_locations()} can convert itself to a data structure that can be used to fill \texttt{RemotePointEvaluation}.
-This conversion is triggered by the member function~\texttt{convert\_to\_distributed\_compute\_point\_locations\_internal()}, given the number of quadrature points per intersection.
-The whole procedure is done without communication.
-%
-This functionality is useful since \texttt{RemotePointEvaluation} can now be used to access quantities at quadrature points on intersections without further ado.
-To reduce the user's effort, we plan to add a wrapper that takes care of the described procedure and provides interfaces like \texttt{FEEvaluation} to access quantities easily, e.g., in a matrix-free loop.
-
-The procedure described above and an early version of the wrapper have
-been used successfully in~\cite{heinz2023high} to perform Nitsche-type
-mortaring in the context of the conservative formulation of acoustic
-equations discretized with a discontinuous Galerkin method to suppress
-artificial modes.
-
-\subsubsection{Non-nested multigrid}
-
-\dealii has provided support for geometric multigrid methods (GMG)
-for locally refined meshes nearly since its inception. (\dealii{} also
-supports algebraic multigrid methods by interfacing to the external
-libraries \petsc and \trilinos.)
-Traditionally, these GMGs were based on local-smoothing
-methods (described many years later in \cite{Kanschat2004,JanssenKanschat2011}, see
-also~\cite{ClevengerHeisterKanschatKronbichler2019}), and more
-recently also global-coarsening algorithms~\cite{munch2022gc}. The global-coarsening
-infrastructure, furthermore, allows globally coarsening the polynomial degree ($p$-multigrid),
-with specific applicability to $hp$-adaptive methods.
-In the current release, we have added support for the case
-that multigrid levels are given by non-nested meshes~\cite{adams2002evaluation, bittencourt2001nonnested, bramble1991analysis}. An example for such meshes is
-presented in Figure~\ref{fig:nonnested}.
-
-\begin{figure}
-  \centering
-  \begin{subfigure}[b]{0.28\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{png/mesh_0.png}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.28\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{png/mesh_1.png}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{0.28\textwidth}
-    \centering
-    \includegraphics[width=\textwidth]{png/mesh_2.png}
-  \end{subfigure}
-  \caption{\it Example of non-nested multigrid levels.}\label{fig:nonnested}
-\end{figure}
-
-The current implementation extends the existing global-coarsening infrastructure by
-introducing, in addition to the (conformal) \texttt{MGTwoLevelTransfer},
- a new (non-conformal) two-level transfer operator:
-\begin{c++}
-MGTwoLevelTransferNonNested two_level_transfer(/*...*/);
-
-two_level_transfer.reinit(dof_handler_fine, dof_handler_coarse,
-                          mapping_fine, mapping_coarse,
-                          constraints_fine, constraints_coarse)
-\end{c++}
-The resulting object can be passed to \texttt{MGTransferGlobalCoarsening} (and
-\texttt{MGTransferBlockGlobal\allowbreak{}Coarsening}) in the usual way. The
-interface of the new class is purposefully similar to
-\texttt{MGTwoLevel\allowbreak{}Transfer},
-with the difference that the former (i) also
-takes \texttt{Mapping} instances, and (ii) is not limited to the case
-that the coarser of the two \texttt{DoFHandler} instances can be generated by a coarsening step from the fine
-mesh.
-
-At the time of writing, the \texttt{MGTwoLevelTransferNonNested} operator provides a prolongation operator
-that prolongates a vector from the coarse space to the fine space,
-by a pointwise interpolation (injection) from the coarse mesh to the support points of the fine mesh. This is done
-by looping over all coarse cells and interpolating to all (fine support) points that fall into the cell with
-\texttt{FEPointEvaluation}. The local result is scattered into the fine vector, which is finalized by a
-communication and a weighting step that is necessary as multiple elements could add to the same global entry.
-
-The current implementation works for scalar, continuous (\texttt{FE\_Q})
-and discontinuous (\texttt{FE\_DGQ}) elements on hypercube-shaped cells. We plan to support
-vectorial and simplex elements and to add support
-for projection, in addition to the pointwise interpolation outlined
-above. Furthermore, we envision providing utility tools to generate
-coarser
-meshes, based on a fine mesh. Currently, the users themselves
-have to provide such meshes,
-e.g., by generating meshes with different cell sizes in external mesh-generation tools.
-
-Note that the new class \texttt{MGTwoLevelTransferNonNested} is not limited to
-multigrid but can be used to prolongate results and restrict residuals between any
-two meshes. Indeed, the infrastructure has been successfully applied to perform
-conservative interpolation between a fine (near) mesh and a coarse (far) mesh in the
-context of aeroacoustic problems.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{New linear algebra features}\label{sec:lac}
-
-There are a number of linear-algebra related new features in this release:
 \begin{itemize}
-  \item Both \texttt{SolverGMRES} and \texttt{SolverFGMRES} now support the classical
-  Gram--Schmidt orthonormalization in addition to the modified Gram--Schmidt algorithm. This
-  reduces the cost of vector operations in terms of
-  communication latency and memory transfer significantly.
-  \item Our Chebyshev preconditioner (\texttt{PreconditionChebyshev}) now also
-  supports James Lottes’s novel fourth-kind Chebyshev
-  polynomial~\cite{lottes2022optimal, phillips2022optimal}.
-  \item Our relaxation preconditioner (\texttt{PreconditionRelaxation}) now also
-  allows interleaving cell loops and vector updates related to relaxation. The
-  relaxation iteration reads as
-  \begin{align*}
-  x^{(i+1)} \gets x^{(i)} + \omega P^{-1}(b-Ax^{(i)}).
-  \end{align*}
-  In the case that the preconditioner $P$ is a diagonal matrix, the zeroing of
-  the destination vector $x_{i+1}$ can be performed during a
-  \texttt{pre}-operation  of the application of $A$
-  and the vector update $x^{(i+1)}_j \gets x^{(i)}_j + \omega P^{-1}_{j,j}(b_j-(Ax^{(i)})_j)$
-  during a \texttt{post} operation, allowing for a reduction in the number of read and write
-  accesses from 8 and 4, to 1 and 3, respectively. The existing \texttt{pre}/\texttt{post}
-  support in our Chebyshev-preconditioner implementation (\texttt{PreconditionChebyshev})
-  has been improved. In addition, both \texttt{PreconditionRelaxation} and
-  \texttt{PreconditionChebyshev} support \texttt{pre}/\texttt{post} optimizations
-  now not only for diagonal preconditioners but also for preconditioners that are
-  built around cell loops and, as a consequence, support interleaving. Examples of
-  such preconditioners are patch-based additive Schwarz preconditioners.
-  \item In the context of additive Schwarz methods, the preconditioner application
-  is defined as
-  \begin{align*}
-  v = P^{-1} u = \sum R_i^\top A_i^{-1} R_i u
-  \end{align*}
-  with $A_i = R_i A R_i^T$ being a block of the assembled system
-  matrix $A$ restricted
-  to an index set described by $R_i$. In the special case when $R_i$ denotes the unknowns
-  of cells, the expression $R_i A R_i^T$ resembles the reverse of matrix
-  assembly.
-  During the restriction step,
-  rows of the system matrix that are potentially owned by other
-  processes are needed.  In \dealii, it is not possible to access remote entries
-  of sparse matrices. Two new functions query this information. First,
-  \texttt{restrict\_to\_serial\_sparse\_matrix()} creates, based
-  on a given index set, a serial
-  sparse matrix from a distributed matrix:
-\begin{c++}
-SparseMatrixTools::restrict_to_serial_sparse_matrix (
-  sparse_matrix_in, sparsity_pattern, requested_index_set,
-  system_matrix_out, sparsity_pattern_out)
-\end{c++}
-This function can be used, e.g., if the granularity of the additive Schwarz preconditioner
-is a complete subdomain, potentially, with a fixed overlap.
-
-In contrast, \texttt{restrict\_to\_full\_matrices()} performs the restriction
-for arbitrary number of patches/blocks:
-
-\begin{c++}
-SparseMatrixTools::restrict_to_full_matrices (
-  sparse_matrix_in, sparsity_pattern, indices_of_blocks, blocks)
-\end{c++}
-
-The data is stored in dense matrices, since the typical granularity is a (rather small) cell-centric or
-vertex-star patch.
-
-  \item For certain types of configurations, there are computationally more efficient
-  approaches than extracting submatrices from an assembled matrix. For
-  example, for the Laplace
-  operator on 2D Cartesian meshes, the (element/patch) matrix is given as
-  \begin{align*}
-  A_i^{\text{cart}} = K_1 \otimes M_0 + M_1 \otimes K_0,
-  \end{align*}
-  i.e., as the tensor product of 1D mass and stiffness matrices. The inverse is
-  explicitly available via the fast diagonalization method~\cite{lynch1964direct} as
-  \begin{align*}
-  \left(A_i^{\text{cart}}\right)^{-1} = (T_1 \otimes T_0) (\Lambda_1 \otimes I + I \otimes \Lambda_0)^{-1} (T_1^\top \otimes T_0^\top),
-  \end{align*}
-  with $T_i$ and $\Lambda_i$ being the (orthonormal) eigenvectors and the diagonal
-  matrix of eigenvalues, obtained from a generalized eigendecomposition
-  $K_iT_i =  M_i T_i \Lambda_i$, as also used by \dealii's \texttt{step-59} tutorial program.
-  Since $A_i \approx A_i^{\text{cart}}$
-  might be a good approximation also in the case of non-Cartesian meshes and
-  $A_i^{\text{cart}}$ has an explicit inverse, it is considered in the
-  literature as a patch preconditioner in the context of additive
-  Schwarz~\cite{witte2021fast, phillips2021auto, couzy1995spectral} and block-Jacobi methods~\cite{kronbichler2019hermite}.
-  In \dealii, the new function
-  \texttt{Tensor\-Product\-Matrix\-Creator::create\_\allowbreak laplace\_\allowbreak tensor\_\allowbreak product\_\allowbreak matrix()} computes $K_i$ and $M_i$
-  for cell-centric patches with a specified overlap and given boundary conditions.
-  A set of $T_i$ and $\Lambda_i$ is applied to a cell via
-  \texttt{Tensor\-Product\-Matrix\-Symmetric\-Sum} or to a collection of cells
-  via the new class \texttt{Tensor\-Product\-Matrix\-Symmetric\-Sum\-Collection}, which
-  tries to reuse the eigenvalues and eigenvectors between cells.
+\item non-nested multigrid: new features and performance improvement
+\item FERemoteEvaluation
+\item FEColungValues
 \end{itemize}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{C++ language modernization}\label{sec:language}
-
-This version of \dealii{} uses C++14 as the language standard to which its
-code base is written. It can also use the classes
-\texttt{std::optional} and \texttt{std::variant} if the compiler
-supports C++17, but falls back to implementations obtained via the
-\boost library otherwise, and this is true also for a number of
-individual functions that were introduced in C++17 or C++20.
-
-As part of the current release, \dealii{} now also uses some C++20
-features to annotate classes and functions with regard to properties
-template arguments need to satisfy. This aids in situations such as
-with the
-following function:
-\begin{c++}
-template <class MeshType>
-std::vector<typename MeshType::active_cell_iterator>
-find_cells_adjacent_to_vertex(const MeshType &   mesh,
-                              const unsigned int vertex);
-\end{c++}
-This function is intended to be called with either a
-\texttt{Triangulation} or \texttt{DoFHandler} object as first
-argument (and documents this requirement), but the compiler can not check this requirement at the call
-site and will gladly call it with any other kind of object as
-well. Because the implementation of the function is in a \texttt{.cc}
-file, and the template is only instantiated for the two classes
-mentioned above, calling the function (erroneously) with anything else
-as first argument is not detected at compile time, but only later when
-the linker reports an undefined symbol.
-
-We have started to address this by annotating functions using
-C++20-style ``\texttt{requires}'' clauses:
-\begin{c++}
-template <class MeshType>
-  requires (concepts::is_triangulation_or_dof_handler<MeshType>)
-std::vector<typename MeshType::active_cell_iterator>
-find_cells_adjacent_to_vertex(const MeshType &   mesh,
-                              const unsigned int vertex)
-\end{c++}
-If the compiler supports C++20, then the added clause will cause the
-compiler to reject any call to the function for which the first
-argument \texttt{MeshType} does not satisfy the named concept, which
-is defined as the type being either a \texttt{Triangulation} or \texttt{DoFHandler}
-object, as intended. (The \texttt{requires} annotation is suppressed if the compiler
-does not support C++20.)
-
-Given the heavy dependence of \dealii{} on templates, there are likely
-hundreds or thousands of locations that should be annotated with
-requirements on template arguments over time; for the moment, the
-library contains some 300 of these \texttt{requires} clauses.
-
-The next release of \dealii{} will build upon C++17.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Build-system modernization}\label{sec:buildsystem}
-
-\dealii's original CMake build system, merged for the 8.0.0 release
-in 2013, was written for CMake 2.8.8 and thus lacks many of the
-modern features introduced into CMake over the last several years. This
-created a particular issue with an increasing number of dependencies
-switching their CMake configuration to \emph{import targets}, where all
-necessary information for using an external resource, such as include
-directories and library link interfaces, are associated with an imported
-CMake target rather than being provided with individual (and
-inconsistently) named CMake variables. With the \dealii 9.5 release the
-handling of external dependencies has been rewritten entirely to support
-imported targets. The \trilinos and \kokkos interfaces have been
-modernized. We plan to migrate all remaining dependencies to import targets
-wherever possible for the next release.
-
-\dealii's CMake project configuration itself now exports three targets,
-\texttt{dealii::dealii\_debug}, \texttt{dealii::dealii\_release}, and
-\texttt{dealii::dealii}. Linking against one of the first two variants via
-\texttt{target\_link\_libraries()} populates the target with all necessary
-include directories and the full link interface necessary for \dealii. The
-third variant, \texttt{dealii::dealii}, automatically switches between
-debug and release interface depending on the build type set via
-\texttt{CMAKE\_BUILD\_TYPE}. The \texttt{dealii::dealii} third variant also
-populates compiler and linker options, whereas for the first two variants a
-client project has to ensure that the compiler and linker are properly set
-up. The new import targets now make it possible to configure a dependent
-project without the use of any \dealii specific macros:
-%
-\begin{anycode}
-cmake_minimum_required(VERSION 3.13.4)
-set(CMAKE_BUILD_TYPE Debug CACHE STRING "")
-project(step CXX)
-
-find_package(deal.II 9.5.0 REQUIRED)
-
-add_executable(step step.cc)
-target_link_libraries(step dealii::dealii)
-\end{anycode}
 
@@
 
-While there are no new \dealii tutorial programs in this release, many
-were extensively revised: Around 145 of the more than 2200 (non-merge)
-commits that went into this release touched the tutorial, in some
-cases adding substantial amounts of text.
-
-% data generated using these commands:
-% - tutorial: git log --since 2022/06/01 --no-merges examples | grep commit | wc -l
-% - total:    git log --since 2022/06/01 --no-merges | grep commit | wc -l
-
-There are four new programs in the code gallery (a collection of
-user-contributed programs that often solve more complicated problems
-than tutorial programs, and that are intended as starting points for further
-research rather than as teaching tools):
+Many of the \dealii tutorial programs were revised in a variety of ways
+as part of this release. In addition, there are a number of new tutorial
+programs:
 \begin{itemize}
-  \item \textit{``A posteriori error estimator for first order hyperbolic problems''},
-  contributed by Marco Feder;
-  \item \textit{``Distributed moving laser heating''}, contributed by
-  Hongfeng Ma and Tatiana E.~Itina;
-  \item \textit{``Generalized Swift-Hohenberg Equation Solver''}, contributed by
-    Sam Scheuerman.
-  \item \textit{``Information density-based mesh refinement''},
-    contributed by Wolfgang Bangerth.
+  \item
+    \texttt{step-87} was contributed by Magdalena Schreter-Fleischhacker
+    (Technical University of Munich) and Peter Munch
+    (University of Augsburg/Uppsala University). It
+    presents the advanced point-evaluation functionalities of \dealii,
+    specifically useful for evaluating finite element solutions at
+    arbitrary points on finite element meshes that can be distributed among processes.
+  \item
+    \texttt{step-89} was contributed by Johannes Heinz (TU Wien),
+    Maximilian Bergbauer (Technical University of Munich),
+    Marco Feder (SISSA), and Peter Munch (University of Augsburg/Uppsala University).
+    It shows one way how to apply non-matching and/or Chimera methods
+    within matrix-free loops in \dealii.
+  \item
+    \texttt{step-90} was contributed by Vladimir Yushutin and Timo Heister (Clemson University).
+    It implements the trace finite element method (TraceFEM). TraceFEM solves PDEs
+    posed on a possibly evolving $(dim-1)$-dimensional surface $\Gamma$ employing
+    a fixed uniform background mesh of a $dim$-dimensional domain in which
+    the surface is embedded. Such surface PDEs arise in problems involving
+    material films with complex properties and in other situations in which
+    a non-trivial condition is imposed on either a stationary or a moving interface.
+    Here we consider a steady, complex, non-trivial surface and the prototypical
+    Laplace-Beltrami equation which is a counterpart of
+    the Poisson problem on flat domains.
 \end{itemize}
@@
 
-The 9.5 release includes
-\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_4_2_and_9_5_0.html}
-     {around 35 incompatible changes};
-see \cite{changes95}. Many of these
+The 9.6 release includes
+\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_5_2_and_9_6_0.html}
+     {around X incompatible changes};
+see \cite{changes96}. Many of these
 incompatibilities change internal
@@
 \begin{itemize}
-  \item
-  The class \texttt{hp::DoFHandler} has been removed. The
-  \texttt{DoFHandler} class now implements all $hp$-related
-  functionality, as it has for the past two releases.
-
-  \item
-  We improved the type safety for active and future FE indices in the \texttt{DoFHandler} implementation by introducing a new data type \texttt{types::fe\textunderscore{}index}.
-Corresponding functions like \texttt{DoFCellAccessor::active\textunderscore{}fe\textunderscore{}index()} and \texttt{DoFHandler::get\textunderscore{}active\textunderscore{}fe\textunderscore{}indices()} have changed their interface.
-
-%  \item
-  This rework also affects the serialization process of active FE
-  indices. You will need to recreate your serialized data \textit{if
-    and only if} you work in $hp$-mode, but can continue to use
-  previously generated data otherwise. For
-  \texttt{parallel::distributed::Triangulation} when used with non-$hp$ data, it is sufficient to increase the version in the metadata file from ``4'' to ``5''.
-
-\item Several old interfaces to \texttt{MatrixFree} have been removed, e.g.,
-  initialization functions without \texttt{Mapping} argument and some queries
-  to the number of cell batches as well as \texttt{DoFHandler} objects. In
-  each case, new interfaces are available.
+  \item CUDAWrappers
 \end{itemize}
@@
 \begin{quote}
-Pasquale    Africa,
-Nicolas     Barnafi,
-Nistha      Bhawsinka,
-Bruno       Blais,
-Fabian      Castelli,
-Terry       Cojean,
-Niklas      Fehn,
-Emmanuel    Ferdman,
-Vadim       Gallyamov,
-Daniel      Garcia-Sanchez,
-Rene        Gassmoeller,
-Robin       Goermer,
-Graham      Harper,
-Quang       Hoang,
-Sean        Ingimarson,
-Vladimir    Ivannikov,
-Pengfei     Jia,
-Tao         Jin,
-Nils        Margenberg,
-Luz         Paz,
-Laura       Prieto Saavedra,
-Sebastian   Proell,
-Ce          Qin,
-Oleg        Rogozin,
-Andrew      Salmon,
-Michael     Schlottke-Lakemper,
-Christoph   Schmidt,
-Magdalena   Schreter-Fleischhacker,
-Richard     Schussnig,
-Nils        Schween,
-Ahmad       Shahba,
-Simon       Sticko,
-Buğrahan    Temür,
-Ivy         Weber,
-Niklas      Wik,
-Vladimir    Yushutin,
-Jiaqi       Zhang.
+TODO
 \end{quote}

### 2024-07-02 9.6/paper.tex
Update p4est references.
Update affiliation of Marc.
Update parallel hp paper.
Add todo about access.

@@
    Collins, CO 80523-1874, USA.
-   \texttt{bangerth/marc.fehling@colostate.edu}}
+   \texttt{bangerth@colostate.edu}}
  \affil[3]{Department of Geosciences, Colorado State University, Fort
@@
 \author[4]{Marc~Fehling}
- \affil[4]{Charles University, Czech Republic}
+ \affil[4]{Department of Mathematical Analysis,
+    Faculty of Mathematics and Physics, Charles University,
+    Sokolovsk{\'a} 49/83, 186\,75 Prague 8, Czech Republic.
+    {\texttt{marc.fehling@matfyz.cuni.cz}}}
 
@@
     \item For distributed parallel computing: \cite{BangerthBursteddeHeisterKronbichler11};
-    \item For $hp$-adaptivity: \cite{BangerthKayserHerold2007,fehling2022};
+    \item For $hp$-adaptivity: \cite{BangerthKayserHerold2007,fehlingbangerth2023};
     \item For partition-of-unity (PUM) and finite element enrichment methods:
@@
     \item OpenCASCADE \cite{opencascade-web-page}
-    \item p4est \cite{burstedde2011p4est,burstedde2020parallel}
+    \item p4est \cite{burstedde2011p4est}
     \item PETSc \cite{petsc-user-ref,petsc-web-page}
@@
 
+M.~Fehling was also partially supported by the ERC-CZ grant LL2105
+CONTACT funded by the Czech Ministry of Education, Youth and Sports.
+
 M.~Bergbauer was supported by the German Research Foundation (DFG) under the
@@
 
+\todo[inline]{Update the following to ACCESS, see https://access-ci.org/about/acknowledging-access/}
+
 This work used the Extreme Science and Engineering Discovery Environment (XSEDE), which is supported by National Science Foundation grant number ACI-1053575 access through the CIG Science Gateway and Community Codes for the Geodynamics Community MCA08X011 allocation.

### 2024-07-10 9.6/paper.tex
Minor work on the tutorial/code gallery section.

@@
   \item
+    \texttt{step-83} if we can get it ready TODO  
+  \item
+    \texttt{step-86} if we can get it merged TODO  
+  \item
     \texttt{step-87} was contributed by Magdalena Schreter-Fleischhacker
@@
     (University of Augsburg/Uppsala University). It
-    presents the advanced point-evaluation functionalities of \dealii,
-    specifically useful for evaluating finite element solutions at
+    presents the advanced point-evaluation functionalities of \dealii
+    that are useful for evaluating finite element solutions at
     arbitrary points on finite element meshes that can be distributed among processes.
@@
     It implements the trace finite element method (TraceFEM). TraceFEM solves PDEs
-    posed on a possibly evolving $(dim-1)$-dimensional surface $\Gamma$ employing
+    posed on a, possibly evolving, $(dim-1)$-dimensional surface $\Gamma$ employing
     a fixed uniform background mesh of a $dim$-dimensional domain in which
@@
     a non-trivial condition is imposed on either a stationary or a moving interface.
-    Here we consider a steady, complex, non-trivial surface and the prototypical
+    The program considers a steady, complex, non-trivial surface and the prototypical
     Laplace-Beltrami equation which is a counterpart of
     the Poisson problem on flat domains.
+\end{itemize}
+
+In addition, there are three new programs in the code gallery (a collection of
+user-contributed programs that often solve more complicated problems
+than tutorial programs, and that are intended as starting points for further
+research rather than as teaching tools):
+UPDATE
+\begin{itemize}
+  \item \textit{``Crystal growth phase field model'''},
+    contributed by Umair Hussain;
+  \item \textit{``Nonlinear heat transfer problem''}, contributed by
+    Narasimhan Swaminathan;
+  \item \textit{``Traveling-wave solutions of a qualitative model for combustion waves''}, contributed by
+    Shamil Magomedov.
 \end{itemize}

### 2024-07-11 9.6/paper.tex
Describe steps 83 and 86.

@@
   \item
-    \texttt{step-83} if we can get it ready TODO  
+    \texttt{step-83} \todo{if we can get it merged}
+    demonstrates how one can implement
+    checkpoint/restart functionality in \dealii{}-based programs,
+    using the BOOST serialization functionality as a
+    foundation. step-83 was written by Pasquale Africa, Wolfgang
+    Bangerth, and Bruno Blais and uses step-19 as its basis.
   \item
-    \texttt{step-86} if we can get it merged TODO  
+    \texttt{step-86} \todo{if we can get it merged}
+    is a program that solves the heat equation using PETSc's TS (time
+    stepping) framework for the solution of ordinary differential
+    equations. Written by Wolfgang Bangerth (Colorado State
+    University), Luca Heltai (University of Pisa), and Stefano Zampini
+    (King Abdullah University of Science and Technology), it
+    illustrates how PDE solvers for time-dependent problems can be
+    integrated with existing ODE solver packages to use advanced ODE
+    solver concepts (such as higher-order time integration methods and
+    adaptive time step control), all without sacrificing the things
+    that have traditionally led code authors toward writing their own
+    time stepping routines (such as wanting to change the mesh every
+    once in a while, or having to deal with boundary conditions).
   \item

### 2024-07-19 9.6/paper.tex
Leave a note.
step-86 was merged.

@@
 \item FEColungValues
+  \todo[inline]{This can not be the right class name, it does not
+    exist in deal.II.}
 \end{itemize}
@@
   \item
-    \texttt{step-86} \todo{if we can get it merged}
+    \texttt{step-86}
     is a program that solves the heat equation using PETSc's TS (time

### 2024-07-23 9.6/paper.tex
add contributors

@@
 
-% This is up-to-date as of 2023-06-26 7:00pm central
+% This is up-to-date as of 2024-07-23 11:52PM CEST
 \begin{quote}
-TODO
+Laryssa      Abdala,
+Pasquale     Africa,
+Mathias      Anselmann,
+Abbas        Ballout,
+Maximilian   Bergbauer,
+Julian       Brotz,
+Marco        Feder,
+Niklas       Fehn,
+Menno        Fraters,
+Quang        Hoang,
+Vladimir     Ivannikov,
+Tao          Jin,
+Yimin        Jin,
+Sebastian    Kinnewig,
+Paras        Kumar,
+Sébastien    Loriot,
+Nils         Much,
+Abdullah     Mujahid,
+Bob          Myhill,
+Paul A.      Patience,
+Luz          Paz,
+Laura        Prieto Saavedra,
+Sebastian    Proell,
+Hendrik      Ranocha,
+Johannes     Resch,
+Andreas      Ritthaler,
+Malik        Scheifinger,
+David        Schneider,
+Magdalena    Schreter,
+Richard      Schussnig,
+Nils         Schween,
+Kyle         Schwiebert,
+Simranjeet   Singh,
+Simon        Sticko,
+Dominik      Still,
+Jan Philipp  Thiele,
+Thierry      Thomas,
+Vinayak      Vijay,
+Ivy          Weber,
+Simon        Wiesheier,
+Chengjiang   Yin,
+Vladimir     Yushutin,
+Stefano      Zampini.
 \end{quote}

### 2024-08-11 9.6/paper.tex
Fix name.
Add B. Blais to release 9.6 paper

@@
   {\texttt{drwells@email.unc.edu}}}
+
+\author[11]{Bruno Blais}
+\affil[11]{Chemical Engineering High-performance Analysis, Optimization and Simulation (CHAOS) laboratory, Department of Chemical Engineering,
+          Polytechnique Montréal,
+          PO Box 6079, Stn Centre-Ville, Montréal, Québec, Canada, H3C 3A7.
+          {\texttt{bruno.blais@polymtl.ca}}}
 
@@
 \item FERemoteEvaluation
-\item FEColungValues
-  \todo[inline]{This can not be the right class name, it does not
-    exist in deal.II.}
+\item FECouplingValues
 \end{itemize}

### 2024-08-11 9.6/paper.tex
Removed non main developer + added Rene

@@
     Marc Fehling,
-    Johannes Heinz,
+    Rene Gassmöller,
     Timo Heister,
@@
   \item
-    \texttt{step-83} \todo{if we can get it merged}
+    \texttt{step-83} 
     demonstrates how one can implement

### 2024-08-11 9.6/paper.tex
Add Rene + his affiliation

@@
     Marc Fehling,
-    Rene Gassmöller,
+    Rene Gassm"{o}ller,
     Timo Heister,
@@
 
-\author[4]{Marc~Fehling}
- \affil[4]{Department of Mathematical Analysis,
+\author[4]{Bruno Blais}
+\affil[4]{Chemical Engineering High-performance Analysis, Optimization and Simulation (CHAOS) laboratory, Department of Chemical Engineering,
+             Polytechnique Montréal,
+             PO Box 6079, Stn Centre-Ville, Montréal, Québec, Canada, H3C 3A7.
+             {\texttt{bruno.blais@polymtl.ca}}}
+
+\author[5]{Marc~Fehling}
+ \affil[5]{Department of Mathematical Analysis,
     Faculty of Mathematics and Physics, Charles University,
@@
 
-\author[5]{Timo~Heister}
- \affil[5]{School of Mathematical and Statistical Sciences,
+\author[6]{Rene~Gassm"{o}ller}
+\affil[6]{GEOMAR Helmholtz Centre for Ocean Research Kiel, 24148 Kiel, Germany}
+
+\author[7]{Timo~Heister}
+ \affil[7]{School of Mathematical and Statistical Sciences,
    Clemson University,
@@
 
-\author[6]{Luca~Heltai}
-\affil[6]{University of Pisa, Italy}
+\author[8]{Luca~Heltai}
+\affil[8]{University of Pisa, Italy}
 
- \author[7]{Martin~Kronbichler}
- \affil[7]{Ruhr University Bochum, Germany}
+ \author[9]{Martin~Kronbichler}
+ \affil[9]{Ruhr University Bochum, Germany}
 
-\author[8]{Matthias~Maier}
-\affil[8]{Department of Mathematics,
+\author[10]{Matthias~Maier}
+\affil[10]{Department of Mathematics,
   Texas A\&M University,
@@
 
-\author[9]{Peter Munch}
- \affil[9]{Uppsala University, Sweden.
+\author[11]{Peter Munch}
+ \affil[11]{Uppsala University, Sweden.
   {\texttt{peter.munch@it.uu.se}}}
 
-\author[1*]{Bruno~Turcksin}
+\author[12]{Bruno~Turcksin}
 
-\author[10]{David Wells}
-\affil[10]{Department of Mathematics, University of North Carolina,
+\author[13]{David Wells}
+\affil[13]{Department of Mathematics, University of North Carolina,
   Chapel Hill, NC 27516, USA.
   {\texttt{drwells@email.unc.edu}}}
-
-\author[11]{Bruno Blais}
-\affil[11]{Chemical Engineering High-performance Analysis, Optimization and Simulation (CHAOS) laboratory, Department of Chemical Engineering,
-          Polytechnique Montréal,
-          PO Box 6079, Stn Centre-Ville, Montréal, Québec, Canada, H3C 3A7.
-          {\texttt{bruno.blais@polymtl.ca}}}
 
@@
 \end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Relicensing to Apache 2.0}\label{sec:license}
 
@@
 
+B.~Blais was supported by the National Science and Engineering Research Council of Canada (NSERC)  through the RGPIN-2020-04510 Discovery Grant and the MMIAOW Canada Research Level 2 in Computer-Assisted Design and Scale-up of Alternative Energy Vectors for Sustainable Chemical Processes
+
 Clemson University is acknowledged for generous allotment of compute time on Palmetto cluster.

### 2024-08-12 9.6/paper.tex
Add affiliation of Jan Philipp Thiele

@@
     Peter Munch,
+    Jan Philipp Thiele,
     Bruno Turcksin,
@@
 
-\author[12]{Bruno~Turcksin}
+\author[12]{Jan Philipp Thiele}
+ \affil[12]{Weierstrass Institute for Applied Analysis and Stochastics,\newline
+ Leibniz Institute in Forschungsverbund Berlin e.V.
+  {\texttt{thiele@wias-berlin.de}}}
 
-\author[13]{David Wells}
-\affil[13]{Department of Mathematics, University of North Carolina,
+\author[13]{Bruno~Turcksin}
+
+\author[14]{David Wells}
+\affil[14]{Department of Mathematics, University of North Carolina,
   Chapel Hill, NC 27516, USA.

### 2024-08-12 9.6/paper.tex
Add authorship and funding

--- 2024-08-11 9.6/paper.tex
@@
   pdfauthor={
+    Pasquale Claudio Africa,
     Daniel Arndt,
@@
     Peter Munch,
+    Magdalena Schreter-Fleischhacker,
+    Jan Philipp Thiele,
     Bruno Turcksin,
@@
 
- \author[1*]{Daniel Arndt}
- \affil[1]{Computational Coupled Physics Group,
+ \author[1*]{Pasquale~Claudio~Africa}
+ \affil[1]{SISSA International School for Advanced Studies,
+   mathLab,
+   Via Bonomea, 265,
+   34136O, Trieste, Italy.
+   \texttt{pafrica@sissa.it}}
+
+ \author[2*]{Daniel Arndt}
+ \affil[2]{Computational Coupled Physics Group,
    Computational Sciences and Engineering Division,
@@
 
- \author[2,3]{Wolfgang~Bangerth}
- \affil[2]{Department of Mathematics, Colorado State University, Fort
+ \author[3,4]{Wolfgang~Bangerth}
+ \affil[3]{Department of Mathematics, Colorado State University, Fort
    Collins, CO 80523-1874, USA.
    \texttt{bangerth@colostate.edu}}
- \affil[3]{Department of Geosciences, Colorado State University, Fort
+ \affil[4]{Department of Geosciences, Colorado State University, Fort
    Collins, CO 80523, USA.}
 
-\author[4]{Bruno Blais}
-\affil[4]{Chemical Engineering High-performance Analysis, Optimization and Simulation (CHAOS) laboratory, Department of Chemical Engineering,
+\author[5]{Bruno Blais}
+\affil[5]{Chemical Engineering High-performance Analysis, Optimization and Simulation (CHAOS) laboratory, Department of Chemical Engineering,
              Polytechnique Montréal,
@@
 
-\author[5]{Marc~Fehling}
- \affil[5]{Department of Mathematical Analysis,
+\author[6]{Marc~Fehling}
+ \affil[6]{Department of Mathematical Analysis,
     Faculty of Mathematics and Physics, Charles University,
@@
 
-\author[6]{Rene~Gassm"{o}ller}
-\affil[6]{GEOMAR Helmholtz Centre for Ocean Research Kiel, 24148 Kiel, Germany}
+\author[7]{Rene~Gassm"{o}ller}
+\affil[7]{GEOMAR Helmholtz Centre for Ocean Research Kiel, 24148 Kiel, Germany}
 
-\author[7]{Timo~Heister}
- \affil[7]{School of Mathematical and Statistical Sciences,
+\author[8]{Timo~Heister}
+ \affil[8]{School of Mathematical and Statistical Sciences,
    Clemson University,
@@
 
-\author[8]{Luca~Heltai}
-\affil[8]{University of Pisa, Italy}
+\author[9]{Luca~Heltai}
+\affil[9]{University of Pisa, Italy}
 
- \author[9]{Martin~Kronbichler}
- \affil[9]{Ruhr University Bochum, Germany}
+ \author[10]{Martin~Kronbichler}
+ \affil[10]{Ruhr University Bochum, Germany}
 
-\author[10]{Matthias~Maier}
-\affil[10]{Department of Mathematics,
+\author[11]{Matthias~Maier}
+\affil[11]{Department of Mathematics,
   Texas A\&M University,
@@
 
-\author[11]{Peter Munch}
- \affil[11]{Uppsala University, Sweden.
+\author[12]{Peter Munch}
+ \affil[12]{Uppsala University, Sweden.
   {\texttt{peter.munch@it.uu.se}}}
 
-\author[12]{Bruno~Turcksin}
+\author[13]{Magdalena Schreter-Fleischhacker}
+ \affil[13]{Institute for Computational Mechanics, Technical University of Munich, Boltzmannstraße 15, 85748 Garching, Germany.
+  {\texttt{magdalena.schreter@tum.de}}}
 
-\author[13]{David Wells}
-\affil[13]{Department of Mathematics, University of North Carolina,
+\author[14]{Jan Philipp Thiele}
+ \affil[14]{Weierstrass Institute for Applied Analysis and Stochastics,\newline
+ Leibniz Institute in Forschungsverbund Berlin e.V.
+  {\texttt{thiele@wias-berlin.de}}}
+
+\author[15]{Bruno~Turcksin}
+
+\author[16]{David Wells}
+\affil[16]{Department of Mathematics, University of North Carolina,
   Chapel Hill, NC 27516, USA.
@@
 Laryssa      Abdala,
-Pasquale     Africa,
 Mathias      Anselmann,
@@
 David        Schneider,
-Magdalena    Schreter,
+Magdalena    Schreter-Fleischhacker,
 Richard      Schussnig,
@@
 variety of funding sources:
+
+P.C.~Africa was partially supported by the consortium iNEST (Interconnected North-East Innovation Ecosystem),
+Piano Nazionale di Ripresa e Resilienza (PNRR) - Missione 4 Componente 2, Investimento 1.5 - D.D. 1058 23/06/2022,
+ECS00000043, supported by the European Union's NextGenerationEU program.
 

### 2024-08-12 9.6/paper.tex

--- 2024-08-11 9.6/paper.tex
@@
 
-\author[12]{Bruno~Turcksin}
+\author[12]{Magdalena Schreter-Fleischhacker}
+ \affil[12]{Institute for Computational Mechanics, Technical University of Munich, Boltzmannstraße 15, 85748 Garching, Germany.
+  {\texttt{magdalena.schreter@tum.de}}}
 
-\author[13]{David Wells}
-\affil[13]{Department of Mathematics, University of North Carolina,
+\author[13]{Jan Philipp Thiele}
+ \affil[13]{Weierstrass Institute for Applied Analysis and Stochastics,\newline
+ Leibniz Institute in Forschungsverbund Berlin e.V.
+  {\texttt{thiele@wias-berlin.de}}}
+
+\author[14]{Bruno~Turcksin}
+
+\author[15]{David Wells}
+\affil[15]{Department of Mathematics, University of North Carolina,
   Chapel Hill, NC 27516, USA.

### 2024-08-14 9.6/paper.tex
add funding of Magdalena Schreter-Fleischhacker
... and affiliation

--- 2024-08-12 9.6/paper.tex
@@
     Bruno Turcksin,
-    David Wells
+    David Wells,
+    Vladimir Yushutin
   },
@@
   {\texttt{drwells@email.unc.edu}}}
+
+\author[17]{Vladimir Yushutin}
+\affil[17]{Department of Mathematics, University of Tennessee at Knoxville,
+	 Knoxville TN 37996-1320, USA
+{\texttt{vyushuti@utk.edu}}}
 
@@
 
+M.~Schreter-Fleischhacker was supported by the Austrian Science Fund (FWF) Schrödinger Fellowship (project number: J4577) and by the European Research Council
+through the ERC Starting Grant ExcelAM (project number: 101117579).
+
 The authors acknowledge the Texas Advanced Computing Center (TACC) at The University of Texas at Austin for providing HPC resources that have contributed to the research results reported within this paper.

### 2024-08-16 9.6/paper.tex
Added subsection about the tpetra stack.

--- 2024-08-12 9.6/paper.tex
@@
 \newcommand{\trilinos}{{\specialword{Trilinos}}\xspace}
+\newcommand{\epetra}{{\specialword{Epetra}}\xspace}
+\newcommand{\tpetra}{{\specialword{Tpetra}}\xspace}
 \newcommand{\aspect}{\specialword{Aspect}\xspace}
@@
 \item FECouplingValues
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Interface to the \trilinos \tpetra stack}\label{sec:tpetra}
+
+\begin{itemize}
+  \item What is a \textit{stack}? [Thiele]
+  \item The \epetra stack is deprecated [Thiele]
+  \item Short overview capabilities of \epetra (MPI parallel, fixed types) [Thiele]
+  \item Short overview capabilities of \tpetra (MPI+X via Kokkos, auto diff types, Templated petra) [Thiele]
+  \item Current state of the interface (Tpertra based linear algebra, direct solver from Amesos2, preconditioners from Ifpack2) [Thiele \& Kinnewig]
+  \item User relevant changes from \epetra: incompatibilities, removed options, changes in the constructor. [Thiele \& Kinnewig]
 \end{itemize}

### 2024-08-16 9.6/paper.tex
Update affiliation

@@
 \newcommand{\trilinos}{{\specialword{Trilinos}}\xspace}
-\newcommand{\epetra}{{\specialword{Epetra}}\xspace}
-\newcommand{\tpetra}{{\specialword{Tpetra}}\xspace}
 \newcommand{\aspect}{\specialword{Aspect}\xspace}
@@
    Clemson University,
-   Clemson, SC, 29634, USA
+   Clemson, SC, 29634, USA.
    {\texttt{heister@clemson.edu}}}
@@
 \author[9]{Luca~Heltai}
-\affil[9]{University of Pisa, Italy}
+\affil[9]{University of Pisa, Italy.}
 
- \author[10]{Martin~Kronbichler}
- \affil[10]{Ruhr University Bochum, Germany}
+ \author[10,11]{Martin~Kronbichler}
+ \affil[10]{Faculty of Mathematics, Ruhr University Bochum,
+   Universit\"atsstr.~150, 44780 Bochum, Germany.
+ {\texttt{martin.kronbichler@rub.de}}}
+ \affil[11]{Institute of Mathematics,
+   University of Augsburg,
+   Universit\"atsstr.~12a, 86159 Augsburg, Germany.
+   }
 
-\author[11]{Matthias~Maier}
-\affil[11]{Department of Mathematics,
+\author[12]{Matthias~Maier}
+\affil[12]{Department of Mathematics,
   Texas A\&M University,
@@
 
-\author[12]{Peter Munch}
- \affil[12]{Uppsala University, Sweden.
+\author[11,13]{Peter Munch}
+ \affil[13]{Uppsala University, Sweden.
   {\texttt{peter.munch@it.uu.se}}}
 
-\author[13]{Magdalena Schreter-Fleischhacker}
- \affil[13]{Institute for Computational Mechanics, Technical University of Munich, Boltzmannstraße 15, 85748 Garching, Germany.
+\author[14]{Magdalena Schreter-Fleischhacker}
+ \affil[14]{Institute for Computational Mechanics, Technical University of Munich, Boltzmannstraße 15, 85748 Garching, Germany.
   {\texttt{magdalena.schreter@tum.de}}}
 
-\author[14]{Jan Philipp Thiele}
- \affil[14]{Weierstrass Institute for Applied Analysis and Stochastics,\newline
+\author[15]{Jan Philipp Thiele}
+ \affil[15]{Weierstrass Institute for Applied Analysis and Stochastics,\newline
  Leibniz Institute in Forschungsverbund Berlin e.V.
@@
 
-\author[15]{Bruno~Turcksin}
+\author[16]{Bruno~Turcksin}
 
-\author[16]{David Wells}
-\affil[16]{Department of Mathematics, University of North Carolina,
+\author[17]{David Wells}
+\affil[17]{Department of Mathematics, University of North Carolina,
   Chapel Hill, NC 27516, USA.
@@
 \item clean up: orientation
+\item various performance improvements throughout the library
+\item the \texttt{SolverGMRES} class now offers a third orthogonalization
+  method, the classical Gram--Schmidt method with delayed orthogonalization
+  \cite{Bielich2022}. Furthermore, the solver now specifies the maximal basis
+  size of the Arnoldi basis, rather than the number of auxiliary vectors.
 \end{itemize}
@@
 \item VectorizedArray for ARM Neon
+\item Improvement for evaluation of values and gradients for
+  $H$(div)-conforming Raviart--Thomas elements on non-Cartesian elements
+\item Improved internal data structures of the tensor-product evaluators as
+  well as the evaluators for simplex elements, which speed up the evaluation
+  in several scenarios, especially multi-component systems.
 \end{itemize}
@@
 \item FECouplingValues
-\end{itemize}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Interface to the \trilinos \tpetra stack}\label{sec:tpetra}
-
-\begin{itemize}
-  \item What is a \textit{stack}? [Thiele]
-  \item The \epetra stack is deprecated [Thiele]
-  \item Short overview capabilities of \epetra (MPI parallel, fixed types) [Thiele]
-  \item Short overview capabilities of \tpetra (MPI+X via Kokkos, auto diff types, Templated petra) [Thiele]
-  \item Current state of the interface (Tpertra based linear algebra, direct solver from Amesos2, preconditioners from Ifpack2) [Thiele \& Kinnewig]
-  \item User relevant changes from \epetra: incompatibilities, removed options, changes in the constructor. [Thiele \& Kinnewig]
 \end{itemize}

### 2024-08-16 9.6/paper.tex
Fix author affiliations

@@
 
-\author[16]{Bruno~Turcksin}
+\author[2]{Bruno~Turcksin}
 
-\author[17]{David Wells}
-\affil[17]{Department of Mathematics, University of North Carolina,
+\author[16]{David Wells}
+\affil[16]{Department of Mathematics, University of North Carolina,
   Chapel Hill, NC 27516, USA.
@@
 
-\author[17]{Vladimir Yushutin}
+\author[17,8]{Vladimir Yushutin}
 \affil[17]{Department of Mathematics, University of Tennessee at Knoxville,
-	 Knoxville TN 37996-1320, USA
+	 Knoxville TN 37996-1320, USA.
 {\texttt{vyushuti@utk.edu}}}

### 2024-08-19 9.6/paper.tex
vladimir was at clemson
Fix typo in post code

@@
    Via Bonomea, 265,
-   34136O, Trieste, Italy.
+   34136, Trieste, Italy.
    \texttt{pafrica@sissa.it}}
@@
 
-\dealii version 9.6.0 was released X X, 2024.
+\dealii version 9.6.0 was released August 11, 2024.
 This paper provides an
@@
 \begin{itemize}
-  \item CUDAWrappers
+  \item CUDAWrappers, that enable the usage of cuSPARSE algorithms, have been deprecated and will be removed in the next release. Kokkos is now used for device-specific optimizations.
 \end{itemize}
@@
     \item SymEngine \cite{symengine-web-page}
+    \item Taskflow \cite{huang2021taskflow}
     \item TBB \cite{reinders2007tbb}
@@
 contributed code to this release:\\
-%
-% Uwe Koecher doesn't usually show up in the changelog, but
-% we should make sure he's listed.
-%
 
@@
 David        Schneider,
-Magdalena    Schreter-Fleischhacker,
+%Magdalena    Schreter-Fleischhacker,
 Richard      Schussnig,
@@
 Dominik      Still,
-Jan Philipp  Thiele,
+%Jan Philipp  Thiele,
 Thierry      Thomas,
@@
 Chengjiang   Yin,
-Vladimir     Yushutin,
+%Vladimir     Yushutin,
 Stefano      Zampini.
@@
 
-M.~Bergbauer was supported by the German Research Foundation (DFG) under the
-project ``High-Performance Cut Discontinuous Galerkin Methods for Flow
-Problems and Surface-Coupled Multiphysics Problems'' Grant Agreement
-No.~456365667.
+T.~Heister and V.~Yushutin were also partiallty supported by NSF Awards OAC-2015848 and EAR-1925575.
 
-J.~Heinz was supported by the European Union’s Framework Programme for Research
-and Innovation Horizon 2020 (2014-2020) under the Marie Sk\l{}odowska--Curie Grant
-Agreement No. [812719].
+T.~Heister was also partially supported by NSF OAC-2410848.
 
-T.~Heister was also partially supported by NSF
-Awards OAC-2015848, DMS-2028346, and
-EAR-1925575.
-
-L.~Heltai and M~.Feder were partially supported by the Italian Ministry of
+L.~Heltai was partially supported by the Italian Ministry of
 University and Research (MUR), under the grant MUR PRIN 2022 No. 2022WKWZA8
@@
 
-S. Zampini was supported by the KAUST Extreme Computing Research Center.
-
 B.~Blais was supported by the National Science and Engineering Research Council of Canada (NSERC)  through the RGPIN-2020-04510 Discovery Grant and the MMIAOW Canada Research Level 2 in Computer-Assisted Design and Scale-up of Alternative Energy Vectors for Sustainable Chemical Processes
-
-Clemson University is acknowledged for generous allotment of compute time on Palmetto cluster.
 
@@
 through the ERC Starting Grant ExcelAM (project number: 101117579).
+
+Clemson University is acknowledged for generous allotment of compute time on Palmetto cluster.
 

### 2024-08-19 9.6/paper.tex
remove authors from contributors

--- 2024-08-19 9.6/paper.tex
@@
    Via Bonomea, 265,
-   34136, Trieste, Italy.
+   34136O, Trieste, Italy.
    \texttt{pafrica@sissa.it}}
@@
 
-\dealii version 9.6.0 was released August 11, 2024.
+\dealii version 9.6.0 was released X X, 2024.
 This paper provides an
@@
 \begin{itemize}
-  \item CUDAWrappers, that enable the usage of cuSPARSE algorithms, have been deprecated and will be removed in the next release. Kokkos is now used for device-specific optimizations.
+  \item CUDAWrappers
 \end{itemize}

### 2024-08-19 9.6/paper.tex
some text
fix order of acknowledgements

--- 2024-08-19 9.6/paper.tex
@@
 
+W.~Bangerth and M.~Fehling were partially supported by Award OAC-1835673
+as part of the Cyberinfrastructure for Sustained Scientific Innovation (CSSI)
+program.
+
 W.~Bangerth and T.~Heister were partially
@@
 
-W.~Bangerth and M.~Fehling were partially supported by Award OAC-1835673
-as part of the Cyberinfrastructure for Sustained Scientific Innovation (CSSI)
-program.
+W.~Bangerth was also partially supported by Awards DMS-1821210 and EAR-1925595.
 
-W.~Bangerth was also partially supported by Awards DMS-1821210 and EAR-1925595.
+B.~Blais was supported by the National Science and Engineering Research Council of Canada (NSERC)  through the RGPIN-2020-04510 Discovery Grant and the MMIAOW Canada Research Level 2 in Computer-Assisted Design and Scale-up of Alternative Energy Vectors for Sustainable Chemical Processes.
 
@@
 
-D.~Wells was supported by the NSF Award OAC-1931516.
-
-B.~Blais was supported by the National Science and Engineering Research Council of Canada (NSERC)  through the RGPIN-2020-04510 Discovery Grant and the MMIAOW Canada Research Level 2 in Computer-Assisted Design and Scale-up of Alternative Energy Vectors for Sustainable Chemical Processes
-
 M.~Schreter-Fleischhacker was supported by the Austrian Science Fund (FWF) Schrödinger Fellowship (project number: J4577) and by the European Research Council
 through the ERC Starting Grant ExcelAM (project number: 101117579).
+
+D.~Wells was supported by the NSF Award OAC-1931516.
 

### 2024-08-19 9.6/paper.tex
Write on libCEED

@@
 
+Furthermore, we added an example to the \texttt{libCEED}
+library\footnote{\url{https://github.com/CEED/libCEED}}. \texttt{libCEED} is a library
+that provides matrix-free evaluation routines that work on different hardware. The
+examples show how to interface the \dealii data structures to the \texttt{libCEED} ones.
+The BP1-BP6 benchmarks (scalar/vector Laplace/mass matrix are solved with
+regular integration and over-integration) are solved.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2024-08-23 9.6/paper.tex
Add affiliation of Sebastian Kinnewig.

@@
 
- \author[10,11]{Martin~Kronbichler}
- \affil[10]{Faculty of Mathematics, Ruhr University Bochum,
+\author[10]{Sebastian~Kinnewig}
+\affil[10]{Institute for Applied Mathematics, Scientific Computing, 
+           Leibniz University Hannover, 
+           Welfengarten 1, 30167 Hannover, Germany
+           {\texttt{kinnewig@ifam.uni-hannover.de}}}
+
+ \author[11,12]{Martin~Kronbichler}
+ \affil[11]{Faculty of Mathematics, Ruhr University Bochum,
    Universit\"atsstr.~150, 44780 Bochum, Germany.
  {\texttt{martin.kronbichler@rub.de}}}
- \affil[11]{Institute of Mathematics,
+ \affil[12]{Institute of Mathematics,
    University of Augsburg,
@@
 
-\author[12]{Matthias~Maier}
-\affil[12]{Department of Mathematics,
+\author[13]{Matthias~Maier}
+\affil[13]{Department of Mathematics,
   Texas A\&M University,
@@
 
-\author[11,13]{Peter Munch}
- \affil[13]{Uppsala University, Sweden.
+\author[12,14]{Peter Munch}
+ \affil[14]{Uppsala University, Sweden.
   {\texttt{peter.munch@it.uu.se}}}
 
-\author[14]{Magdalena Schreter-Fleischhacker}
- \affil[14]{Institute for Computational Mechanics, Technical University of Munich, Boltzmannstraße 15, 85748 Garching, Germany.
+\author[15]{Magdalena Schreter-Fleischhacker}
+ \affil[15]{Institute for Computational Mechanics, Technical University of Munich, Boltzmannstraße 15, 85748 Garching, Germany.
   {\texttt{magdalena.schreter@tum.de}}}
 
-\author[15]{Jan Philipp Thiele}
- \affil[15]{Weierstrass Institute for Applied Analysis and Stochastics,\newline
+\author[16]{Jan Philipp Thiele}
+ \affil[16]{Weierstrass Institute for Applied Analysis and Stochastics,\newline
  Leibniz Institute in Forschungsverbund Berlin e.V.
@@
 
-\author[16]{David Wells}
-\affil[16]{Department of Mathematics, University of North Carolina,
+\author[17]{David Wells}
+\affil[17]{Department of Mathematics, University of North Carolina,
   Chapel Hill, NC 27516, USA.
@@
 
-\author[17,8]{Vladimir Yushutin}
-\affil[17]{Department of Mathematics, University of Tennessee at Knoxville,
+\author[18,8]{Vladimir Yushutin}
+\affil[18]{Department of Mathematics, University of Tennessee at Knoxville,
 	 Knoxville TN 37996-1320, USA.

### 2024-08-29 9.6/paper.tex
Add information about Trilinos stacks

@@
 \newcommand{\candi}{{\specialword{candi}}\xspace}
+\newcommand{\MPI}{{\specialword{MPI}}\xspace}
+\newcommand{\MPIx}{{\specialword{MPI+X}}\xspace}
 \newcommand{\sundials}{{\specialword{SUNDIALS}}\xspace}
@@
 \newcommand{\kokkos}{{\specialword{Kokkos}}\xspace}
+
+%Trilinos Packages
+\newcommand{\epetra}{{\specialword{Epetra}}\xspace}
+\newcommand{\petra}{{\specialword{Petra}}\xspace}
+\newcommand{\teuchos}{{\specialword{Teuchos}}\xspace}
+\newcommand{\tpetra}{{\specialword{Tpetra}}\xspace}
 
@@
 
+As a software project, \trilinos{} is designed as a stack of semi-independent
+packages building upon each other and working together. 
+As a foundation, \teuchos{} offers a set of basic tools,
+including parameter handling through XML files or command line interfaces 
+and the reference counted pointer \texttt{RCP} which predates 
+the smart pointer types introduced in \texttt{C++11}.
+The second layer in the stack is a package providing distributed linear algebra 
+classes for vectors and (sparse) matrices based on the \petra{}
+object model. The third layer is a set of interconnected packages
+providing preconditioners, various types of solvers 
+as well as more advanced techniques like automatic differentiation, 
+time integration and discretization.
+
+For now, there are two choices for the second layer 
+with \epetra{} providing \MPI{} only parallelism 
+and the newer \tpetra{} (templated Petra) providing additional shared memory 
+parallelism and GPU capabilities through \kokkos{}, referred to as \MPIx.
+Since the interface of \tpetra{} was overhauled and modernized 
+many of the packages in the third layer now have two generations as well.
+Therefore, the \MPI-only set of packages is also referred to as the 
+\epetra{} stack and the \MPIx{} packages form the \tpetra{} stack.
+As a consequence of the different interfaces, 
+a decision was made in the \trilinos{} project to deprecate the \epetra{} stack 
+and only provide basic maintenance until a full 
+removal from their repository in the course of 2025.
+
 \begin{itemize}
-  \item What is a \textit{stack}? [Thiele]
-  \item The \epetra stack is deprecated [Thiele]
   \item Short overview capabilities of \epetra (MPI parallel, fixed types) [Thiele]

### 2024-09-12 9.6/paper.tex
Work on the paper.

@@
 \begin{itemize}
-  \item 
+  \item Substantial improvements to the multigrid two-level transfer
+    operators, including matrix-free based versions and versions for
+    p-refinement (Section~\ref{sec:mf}).
+  \item A class FERemoteEvaluation that provides support for evaluating
+    finite element shape functions and solutions on parts of the
+    domain stored by other MPI processes (Section~\ref{sec:remote-eval}).
+  \item Much work has gone into writing wrappers for Trilinos' Tpetra
+    stack of linear algebra classes. Tpetra is Trilinos' Kokkos-based
+    replacement for the now-deprecated Epetra stack. See
+    Section~\ref{sec:tpetra} for more on this.
+  \item Tool classes \texttt{TaskResult} and \texttt{Lazy} that provide ways to compute
+    values on a separate thread or only when first accessed (Section~\ref{sec:tools}).
+  \item There are six new tutorial programs, on checkpointing simulations
+    (step-83), integrating time-dependent solvers with external time
+    stepping libraries (step-86, using PETSc's TS library), advanced
+    point evaluation techniques (step-87), non-matching grids
+    (step-89), and trace-based methods for PDEs on embedded surfaces
+    (step-90). See Section~\ref{subsec:steps} for more details.
 \end{itemize}
 %
+
+\todo[inline]{If we keep the section on relicensing
+  (Section~\ref{sec:license}), mention it in the list above.}
 
@@
 \begin{itemize}
+  \item deal.II now requires and uses of C++17.
+  \item We have continued to make progress in supporting simplex and
+    mixed meshes -- mesh types that \dealii{} has traditionally not
+    supported at all. Specifically, the current release uses
+    better strategies for refinement of tetrahedra that result in
+    better-shaped child cells. It also contains
+    support for cubic finite elements on simplices.
+  \item The \texttt{FE\_NedelecSZ} class that contains our
+    implementation of the N\'ed\'elec element using the orientation
+    scheme of \cite{Zag06} now supports the computation of hanging
+    node constraints for locally refined meshes.
+    \todo[inline]{Sebastian: Is there anything more to say here? If
+      you want, we can also list this point in the list above, and add
+      a whole section 2.x for it.}
   \item changes of interface of AffineConstraints
+    \todo[inline]{Wolfgang to write}
 \item clean up: orientation
-\item various performance improvements throughout the library
-\item the \texttt{SolverGMRES} class now offers a third orthogonalization
+  \todo[inline]{DavidW: Please describe some details here.}
+\item Various performance improvements throughout the library.
+  \todo[inline]{Everyone: This sentence would be more useful if we can
+    provide a few examples.}
+\item The \texttt{SolverGMRES} class now offers a third orthogonalization
   method, the classical Gram--Schmidt method with delayed orthogonalization
   \cite{Bielich2022}. Furthermore, the solver now specifies the maximal basis
-  size of the Arnoldi basis, rather than the number of auxiliary vectors.
+  size of the Arnoldi basis, rather than the number of auxiliary
+  vectors. Some changes have also been made to the GMRES and F-GMRES
+  implementations to make them more similar.
+  \todo[inline]{Martin: Does this capture the relevant changes? Please
+    fix as appropriate, or if you'd like put them into a section of
+    its own.}
 \end{itemize}
 %
-The changelog lists more than X other features and bugfixes.
+The changelog lists more than 180 other features and bugfixes.
 
@@
 This release of \dealii contains a number of large and significant changes,
-which will be discussed in this section.
+which we will discuss in this section.
 It of course also includes a
@@
 can be found
-\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_4_2_and_9_5_0.html}
-{in the file that lists all changes for this release}; see \cite{changes95}.
-
-\begin{itemize}
-\item Hermite element
-\item simplices: cubic simplex elements, improved refinement of tetrahedron cells
-\item Tpetra advances
-\item Nedelec for meshes with hanging nodes
-\end{itemize}
+\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_5_2_and_9_6_0.html}
+{in the file that lists all changes for this release}; see \cite{changes96}.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Relicensing to Apache 2.0}\label{sec:license}
+\subsection{Updates to multigrid and matrix-free algorithms}\label{sec:mf}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Updates to matrix-free algorithms}\label{sec:mf}
+\todo[inline]{Martin/Peter: The list in the introduction also mentions the two-level
+  operators. Is this the right place to also discuss these?}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{The \texttt{FERemoteEvaluation} class}\label{sec:remote-eval}
+
+\todo[inline]{Peter/Martin/...: Would you be willing to describe this
+  class and its uses? Or, if you don't think it's important enough,
+  remove this section and move the part in the top list of major
+  changes to the bottom list?}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Advances in non-matching support}\label{sec:nonmatching}
+
+\todo[inline]{All: If you contributed to this functionality, please
+  help flesh out this section.}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Interface to the \trilinos \tpetra stack}\label{sec:tpetra}
+\subsection{Interface to the \trilinos{} \tpetra stack}\label{sec:tpetra}
 
+\dealii{}'s parallel linear algebra facilities are largely built on
+wrappers around functionality provided by the \petsc{} and \trilinos{} libraries.
 As a software project, \trilinos{} is designed as a stack of semi-independent
@@
 
-For now, there are two choices for the second layer 
-with \epetra{} providing \MPI{} only parallelism 
-and the newer \tpetra{} (templated Petra) providing additional shared memory 
-parallelism and GPU capabilities through \kokkos{}, referred to as \MPIx.
+Historically, \trilinos{} has implemented \petra{} objects -- such as
+parallel vector, sparsity pattern, and sparse matrix classes -- in its
+\epetra{} package that uses \MPI{} as its only source of
+parallelism. The \dealii{} interfaces to \trilinos{} are therefore all
+built using \epetra{}.
+However, several years ago, \trilinos also introduced the
+newer \tpetra{} (templated Petra) package that provides additional shared memory 
+parallelism and GPU capabilities by building on \kokkos{}. \tpetra{}
+is slated to eventually replace \epetra{}, and the latter is indeed
+now deprecated. As a consequence, we will eventually have to switch
+all of our \trilinos{} interfaces to \tpetra{} and \tpetra{}-based sub-packages.
+
+In the current release, we have put substantial work into this switch.
 Since the interface of \tpetra{} was overhauled and modernized 
@@
   \item Short overview capabilities of \tpetra (MPI+X via Kokkos, auto diff types, Templated petra) [Thiele]
-  \item Current state of the interface (Tpertra based linear algebra, direct solver from Amesos2, preconditioners from Ifpack2) [Thiele \& Kinnewig]
+  \item Current state of the interface (Tpetra based linear algebra, direct solver from Amesos2, preconditioners from Ifpack2) [Thiele \& Kinnewig]
   \item User relevant changes from \epetra: incompatibilities, removed options, changes in the constructor. [Thiele \& Kinnewig]
 \end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{More support for advanced programming idioms}\label{sec:tools}
+
+\todo[inline]{Wolfgang to write. Mention Lazy and TaskResult.}
 
@@
 Furthermore, we added an example to the \texttt{libCEED}
-library\footnote{\url{https://github.com/CEED/libCEED}}. \texttt{libCEED} is a library
+library.\footnote{\url{https://github.com/CEED/libCEED}} \texttt{libCEED} is a library
 that provides matrix-free evaluation routines that work on different hardware. The
 examples show how to interface the \dealii data structures to the \texttt{libCEED} ones.
-The BP1-BP6 benchmarks (scalar/vector Laplace/mass matrix are solved with
-regular integration and over-integration) are solved.
+The example solves the BP1-BP6 benchmarks (scalar/vector Laplace/mass matrix with
+regular integration and over-integration).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Relicensing to Apache 2.0}\label{sec:license}
+
+\todo[inline]{Matthias: Where are we with this, and what of it is
+  worth describing here?}
 
@@
 \href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_5_2_and_9_6_0.html}
-     {around X incompatible changes};
+     {around 40 incompatible changes};
 see \cite{changes96}. Many of these
@@
 \begin{itemize}
-  \item CUDAWrappers, that enable the usage of cuSPARSE algorithms, have been deprecated and will be removed in the next release. Kokkos is now used for device-specific optimizations.
+  \item \dealii{} now requires compilers to support C++17, and has
+    started to extensively use C++17 features.
+  \item The \texttt{CUDAWrappers} namespace and its contents --
+    notably things that enable the usage of cuSPARSE algorithms --
+    have been deprecated and will be removed in the next
+    release. Kokkos is now used for device-specific optimizations.
 \end{itemize}

### 2024-09-13 9.6/paper.tex
Add many TODOs.

@@
     Marc Fehling,
-    Rene Gassm"{o}ller,
+    Rene Gassm\"{o}ller,
     Timo Heister,
@@
 
-\author[7]{Rene~Gassm"{o}ller}
+\author[7]{Rene~Gassm\"{o}ller}
 \affil[7]{GEOMAR Helmholtz Centre for Ocean Research Kiel, 24148 Kiel, Germany}
@@
 However, several years ago, \trilinos also introduced the
-newer \tpetra{} (templated Petra) package that provides additional shared memory 
+newer \tpetra{} (``templated'' Petra) package that provides additional shared memory 
 parallelism and GPU capabilities by building on \kokkos{}. \tpetra{}
 is slated to eventually replace \epetra{}, and the latter is indeed
-now deprecated. As a consequence, we will eventually have to switch
-all of our \trilinos{} interfaces to \tpetra{} and \tpetra{}-based sub-packages.
+now deprecated with removal slated for 2025. As a consequence, we will
+eventually have to switch all of our \trilinos{} interfaces to
+\tpetra{}; the same is true for the need to switch to \tpetra{}-based
+sub-packages replacing existing \trilinos sub-packages (for example,
+Ifpack2 instead of Ifpack).
 
 In the current release, we have put substantial work into this switch.
-Since the interface of \tpetra{} was overhauled and modernized 
-many of the packages in the third layer now have two generations as well.
-Therefore, the \MPI-only set of packages is also referred to as the 
-\epetra{} stack and the \MPIx{} packages form the \tpetra{} stack.
-As a consequence of the different interfaces, 
-a decision was made in the \trilinos{} project to deprecate the \epetra{} stack 
-and only provide basic maintenance until a full 
-removal from their repository in the course of 2025.
+
+\todo[inline]{Sebastian \& Jan Philipp: Please complete}
 
@@
   \item Short overview capabilities of \tpetra (MPI+X via Kokkos, auto diff types, Templated petra) [Thiele]
+    \todo[inline]{Keep discussion of the two items above as short as
+      you can -- we're describing the changes in deal.II, not some
+      other package. Perhaps roll what you want to say into the
+      paragraphs above?}
   \item Current state of the interface (Tpetra based linear algebra, direct solver from Amesos2, preconditioners from Ifpack2) [Thiele \& Kinnewig]
-  \item User relevant changes from \epetra: incompatibilities, removed options, changes in the constructor. [Thiele \& Kinnewig]
+  \item User relevant changes from \epetra: incompatibilities, removed
+    options, changes in the constructor. [Thiele \& Kinnewig]
+    \todo[inline]{These last two items are the important ones.}
 \end{itemize}
@@
 
-P.C.~Africa was partially supported by the consortium iNEST (Interconnected North-East Innovation Ecosystem),
+P.~C.~Africa was partially supported by the consortium iNEST (Interconnected North-East Innovation Ecosystem),
 Piano Nazionale di Ripresa e Resilienza (PNRR) - Missione 4 Componente 2, Investimento 1.5 - D.D. 1058 23/06/2022,
@@
 (CIG), through the National Science Foundation (NSF) under Award
-No.~EAR-1550901 and EAR-2149126 via The University of California -- Davis.
+No.~EAR-2149126 via The University of California -- Davis.
 
-W.~Bangerth was also partially supported by Awards DMS-1821210 and EAR-1925595.
+W.~Bangerth was also partially supported by Award EAR-1925595.
 
@@
 M.~Fehling was also partially supported by the ERC-CZ grant LL2105
-CONTACT funded by the Czech Ministry of Education, Youth and Sports.
+CONTACT, funded by the Czech Ministry of Education, Youth and Sports.
+
+\todo[inline]{Rene: anything to list here?}
 
@@
 ``Immersed methods for multiscale and multiphysics problems (IMMEDIATE)''.
+
+\todo[inline]{Sebastian: Anything to list here?}
 
@@
 
-D.~Wells was supported by the NSF Award OAC-1931516.
+\todo[inline]{Jan Philipp: Anything to list here?}
 
-Clemson University is acknowledged for generous allotment of compute time on Palmetto cluster.
+D.~Wells was supported by NSF Award OAC-1931516.
+
+Clemson University is acknowledged for generous allotment of compute
+time on the Palmetto cluster.
+
+\todo[inline]{Did anyone actually use TACC resources? Timo or Marc? If
+not, we need to delete the following sentence.}
 
@@
 
-\todo[inline]{Update the following to ACCESS, see https://access-ci.org/about/acknowledging-access/}
+\todo[inline]{Marc: Update the following to ACCESS, see https://access-ci.org/about/acknowledging-access/}
 

### 2024-09-14 9.6/paper.tex
ACCESS

@@
 
-\todo[inline]{Did anyone actually use TACC resources? Timo or Marc? If
-not, we need to delete the following sentence.}
-
 The authors acknowledge the Texas Advanced Computing Center (TACC) at The University of Texas at Austin for providing HPC resources that have contributed to the research results reported within this paper.
@@
 
-\todo[inline]{Marc: Update the following to ACCESS, see https://access-ci.org/about/acknowledging-access/}
+\todo[inline]{Marc: please insert allocation}
+
+This work used [resource-name] at [resource provider] through allocation [allocation number] from the Advanced Cyberinfrastructure Coordination Ecosystem: Services \& Support (ACCESS) program, which is supported by National Science Foundation grants \#2138259, \#2138286, \#2138307, \#2137603, and \#2138296.
+See \cite{Boerner2023}.
 

### 2024-09-15 9.6/paper.tex
Fix one of the todos

@@
     node constraints for locally refined meshes.
-    \todo[inline]{Sebastian: Is there anything more to say here? If
+    \todo[inline]{Sebastian: Is there anything more to say here, maybe some reference? If
       you want, we can also list this point in the list above, and add
@@
   vectors. Some changes have also been made to the GMRES and F-GMRES
-  implementations to make them more similar.
-  \todo[inline]{Martin: Does this capture the relevant changes? Please
-    fix as appropriate, or if you'd like put them into a section of
-    its own.}
+  implementations, making them use the same underlying kernels as much as possible.
 \end{itemize}

### 2024-09-17 9.6/paper.tex
Remove section

@@
     p-refinement (Section~\ref{sec:mf}).
-  \item A class FERemoteEvaluation that provides support for evaluating
+  \item Different additions to the non-matching infrastructure (see
+    Section~\ref{sec:nonmatching}) In particular, the new class
+    FERemoteEvaluation provides support for evaluating
     finite element shape functions and solutions on parts of the
-    domain stored by other MPI processes (Section~\ref{sec:remote-eval}).
+    domain stored by other MPI processes.
   \item Much work has gone into writing wrappers for Trilinos' Tpetra
@@
 \end{itemize}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The \texttt{FERemoteEvaluation} class}\label{sec:remote-eval}
-
-\todo[inline]{Peter/Martin/...: Would you be willing to describe this
-  class and its uses? Or, if you don't think it's important enough,
-  remove this section and move the part in the top list of major
-  changes to the bottom list?}
 

### 2024-09-17 9.6/paper.tex
Write text

@@
   \item Different additions to the non-matching infrastructure (see
-    Section~\ref{sec:nonmatching}) In particular, the new class
+    Section~\ref{sec:nonmatching}). In particular, the new class
     FERemoteEvaluation provides support for evaluating
@@
     \todo[inline]{Wolfgang to write}
-\item clean up: orientation
+\item We have clean up the orientation definition of faces in (2D and) 3D. Three
+    Booleans used to be used to define it for hexedras: \textit{orientation},
+    \textit{flip}, and \textit{rotation}. While introducing simplex and mixed
+     mesh support in \dealii, we merged these three into a single variable of type
+     char internally. Now, we also expose these to the outside, leading to
+     multiple incompatibilities, since the template type of different data
+     structures has changed and the number of input/output arguments have decreased.
   \todo[inline]{DavidW: Please describe some details here.}
-\item Various performance improvements throughout the library.
-  \todo[inline]{Everyone: This sentence would be more useful if we can
-    provide a few examples.}
+\item We have performed various performance improvements throughout the library,
+     which we detail in the following sections.
 \item The \texttt{SolverGMRES} class now offers a third orthogonalization
@@
 
-\todo[inline]{Martin/Peter: The list in the introduction also mentions the two-level
-  operators. Is this the right place to also discuss these?}
+%\todo[inline]{Martin/Peter: The list in the introduction also mentions the two-level
+%  operators. Is this the right place to also discuss these?}
 
+We made different updates to the matrix-free infrastructure in \dealii. These include:
 \begin{itemize}
-\item VectorizedArray for ARM Neon
-\item Improvement for evaluation of values and gradients for
-  $H$(div)-conforming Raviart--Thomas elements on non-Cartesian elements
-\item Improved internal data structures of the tensor-product evaluators as
-  well as the evaluators for simplex elements, which speed up the evaluation
-  in several scenarios, especially multi-component systems.
+\item Our own implementation of \texttt{std::simd} (\texttt{VectorizedArray}) now
+  also Arm Neon. Arm Neon is an architecture extension for
+  the Arm Cortex-A and Arm Cortex-R series of processors, commonly used in Apple
+  products. With these instructions 2 doubles or 4 floats can be processed in one
+  go. Since matrix-free infrastructure works directly with \texttt{VectorizedArray}
+  data structures, it automatically benefits from this extension.
+
+\item We made some improvement for evaluation of values and gradients for
+  $H$(div)-conforming Raviart--Thomas elements on non-Cartesian elements.
+
+\item We furthermore improved the internal data structures of the
+  tensor-product evaluators as well as the evaluators for simplex elements,
+  which speed up the evaluation in several scenarios, especially multi-component systems.
 \end{itemize}
+
+Furthermore, we performed substantial improvements to the global-coarsening
+multigrid infrastructure: \texttt{MGTransferMF} (previously:
+\texttt{MGTransferGlobalCoarsening}) and \texttt{MGTwoLevelTransfer}. They
+now allow to perform local smoothing, which is a key step towards unifying
+all transfer operators in \dealii. \texttt{MGTwoLevelTransfer} now also
+can be set up with an existing \texttt{MatrixFree} object in the case
+of $p$-multigrid, allowing to reduce the setup costs significantly.
 
@@
 
-\todo[inline]{All: If you contributed to this functionality, please
-  help flesh out this section.}
+%\todo[inline]{All: If you contributed to this functionality, please
+%  help flesh out this section.}
+
+We made different advances to the non-matching infrastructure in \dealii. These include:
 
 \begin{itemize}
-\item non-nested multigrid: new features and performance improvement
-\item FERemoteEvaluation
+\item The performance of the non-nested multigrid infrastructure
+(\texttt{MGTwoLevelTransferNon\-Nested}) has been improved significantly by avoiding
+redundant copy operations. Furthermore, the support for simplex-shaped cells and
+multiple-component elements have been added.
+\item The new class \texttt{FERemoteEvaluation} has been added. 
+This is a class to access data
+in a distributed matrix-free loop for non-matching discretizations.
+Interfaces are named with \texttt{FEEvaluation} in mind. Key component is the
+underlying MPI communication, which is done via \texttt{RemotePointEvaluation}.
+Tutorial \texttt{step-89} has been added to present its usage
+in the context of an application to acoustic conservation equations~\cite{heinz2023high}.
 \item FECouplingValues
+
+\todo[inline]{Luca: Please complete}
 \end{itemize}
@@
 
+\todo[inline]{Peter: Anything to list here?}
+
 M.~Schreter-Fleischhacker was supported by the Austrian Science Fund (FWF) Schrödinger Fellowship (project number: J4577) and by the European Research Council

### 2024-09-17 9.6/paper.tex
Fix some typos

@@
 \begin{itemize}
-  \item Substantial improvements to the multigrid two-level transfer
-    operators, including matrix-free based versions and versions for
-    p-refinement (Section~\ref{sec:mf}).
+  \item Substantial performance improvements to the matrix-free and multigrid
+   infrastructure (Section~\ref{sec:mf}).
   \item Different additions to the non-matching infrastructure (see
@@
 \item Our own implementation of \texttt{std::simd} (\texttt{VectorizedArray}) now
-  also Arm Neon. Arm Neon is an architecture extension for
+  also supports Arm Neon. Arm Neon is an architecture extension for
   the Arm Cortex-A and Arm Cortex-R series of processors, commonly used in Apple
@@
   go. Since matrix-free infrastructure works directly with \texttt{VectorizedArray}
-  data structures, it automatically benefits from this extension.
+  data structures, it automatically benefits from this implementation extension.
 

### 2024-09-18 9.6/paper.tex
Add funding support

@@
 
-W.~Bangerth and T.~Heister were partially
+W.~Bangerth, T.~Heister, and R.~Gassm\"{o}ller were partially
 supported by the Computational Infrastructure for Geodynamics initiative
@@
 
-\todo[inline]{Rene: anything to list here?}
+R.~Gassm\"{o}ller was also partially supported by NSF Awards EAR-1925677
+and EAR-2054605.
 

### 2024-09-18 9.6/paper.tex
Edit the section about orientations.

@@
     \todo[inline]{Wolfgang to write}
-\item We have clean up the orientation definition of faces in (2D and) 3D. Three
-    Booleans used to be used to define it for hexedras: \textit{orientation},
-    \textit{flip}, and \textit{rotation}. While introducing simplex and mixed
-     mesh support in \dealii, we merged these three into a single variable of type
-     char internally. Now, we also expose these to the outside, leading to
-     multiple incompatibilities, since the template type of different data
-     structures has changed and the number of input/output arguments have decreased.
-  \todo[inline]{DavidW: Please describe some details here.}
-\item We have performed various performance improvements throughout the library,
-     which we detail in the following sections.
+  \item \dealii{} internally needs to keep track of the relative
+    orientations of the coordinate systems associated with neighboring
+    cells, as well as of the coordinate systems of faces and edges
+    within the coordinate system of the cells they are part
+    of. Historically, different places within the library that had to
+    do that had grown their own, independent ways of representating
+    this information, much of which was difficult to understand.
+    We have now cleaned up this area. Specifically, instead of using three
+    Booleans to define relative orientations for hexedra (\textit{orientation},
+    \textit{flip}, and \textit{rotation}), we have internally merged these three values into a single variable of type
+     \texttt{char} -- i.e., the smallest integer data type that has at
+     least three bits to represent the information. Because the way to represent
+     orientations is part of the public interface, the change causes
+     some incompatibilities in types and function signatures.
+  \todo[inline]{DavidW: Please describe some details here. -- When
+    done, please remove this comment.}
 \item The \texttt{SolverGMRES} class now offers a third orthogonalization

### 2024-09-18 9.6/paper.tex
Resolve one TODO.

@@
       a whole section 2.x for it.}
-  \item changes of interface of AffineConstraints
-    \todo[inline]{Wolfgang to write}
+  \item The \texttt{AffineConstraints} class stores and processes
+    constraints on degrees of freedom in \dealii{}. Such constraints
+    can be of the form $x_3 = \frac 12 x_{14} + \frac 12 x_{15}$ as is
+    common when using hanging node constraints (here, $x_3$ could
+    represent the value of the solution at a hanging node, and $x_{14}$
+    and $x_{15}$ are the values of the two adjacent degrees of freedom on
+    the parent edge that contains the hanging node); of they can be of
+    the form $x_{12}=42$ as is common when using Dirichlet boundary
+    conditions.
+
+    Since very early in the history of the library, the
+    \texttt{AffineConstraints} class interface required building such
+    constraints in multiple steps: First, one declared a degree of
+    freedom as constrained; then one added the dependencies one after
+    the other (e.g., by adding pairs $(\frac 12, 13)$ and $(\frac 12,
+    14)$ in the hanging node example above); then one added
+    inhomogeneities (by setting it to $42$ in the Dirichlet example
+    above). This piecemeal approach is cumbersome and prevents the
+    library from performing certain error checking steps because a
+    constraint is not known to be completely built at any given point. The new
+    \textt{add\_constraint()} function now allows defining a
+    constraint in one step.
   \item \dealii{} internally needs to keep track of the relative

### 2024-09-18 9.6/paper.tex
Fix a latex problem I just introduced :-(
Write a missing section.

@@
     constraint is not known to be completely built at any given point. The new
-    \textt{add\_constraint()} function now allows defining a
+    \texttt{add\_constraint()} function now allows defining a
     constraint in one step.
@@
 
-\todo[inline]{Wolfgang to write. Mention Lazy and TaskResult.}
+Over the years, \dealii{} has accumulated many classes and functions
+that support modern programming idioms and make it easier to write
+code at higher levels of abstraction.
+
+In the current release, we have added two classes to the list of tools
+of this kind:
+\begin{itemize}
+  \item
+  \texttt{Lazy<T>} is a class that supports the lazy computation and
+  initialization of variables. Its intended use is for member
+  variables of classes that are \textit{sometimes} needed, but perhaps
+  not for all uses of an object. For example, all finite element classes
+  provide interpolation and restriction matrices to support multigrid
+  and other algorithms. One could (i) always compute and store these
+  matrices in the constructor of the class; or one could (ii) re-compute these
+  matrices every time they are requested. The first of these
+  approaches costs memory and compute time even though most places
+  where one creates a finite element object will not actually query
+  these matrices; the second of these approaches is costly in places
+  that \textit{do} query these matrices repeatedly because they are re-computed
+  every time. \texttt{Lazy<T>} provides a middle ground: It provides an
+  abstraction for an object that is initialized upon first use (that
+  is, the first time the value is requested), and then stores the
+  computed value for cheap use later on.
+
+  (C++ provides functionality via \texttt{std::async} with
+  launch policy \texttt{std::launch::deferred} that can achieve
+  similar outcomes. But this functionality is more difficult to use than
+  \texttt{Lazy<T>} because, among other reasons, the code generating the
+  object has to be specified at the place of construction of the
+  object holding the result, rather than at the place of use; and
+  because the holder object -- \texttt{std::future} -- can only be
+  asked once for its computed value.)
+
+  \item \texttt{TaskResult<T>} is a class that represents the outcome
+    of a task possibly evaluated on a separate thread. It can be
+    thought of as a ``deferred'' result of a computation in that one
+    wants to state ``This job needs to be done, do it when convenient,
+    and then put the result of the operation into this
+    variable''. Accessing the variable then waits for the operation to
+    complete, if it has not already. \texttt{TaskResult<T>} allows
+    classes to efficiently compute member variables in the background,
+    assuming that they may not be needed right away but only later on.
+
+    (Similar to above, the same effect as \texttt{TaskResult<T>} can
+    be achieved using \texttt{std::async}, this time using the
+    launch policy \texttt{std::launch::async}. This approach suffers
+    from the same issue that one can only query the resulting object
+    once. Moreover, \texttt{std::async} does not integrate with the
+    thread pool that underlies \dealii{}'s approach to parallel
+    processing on modern multi-core machines, whereas
+    \texttt{TaskResult<T>} does.)
+\end{itemize}
 

### 2024-09-19 9.6/paper.tex
A few improvements to the text

@@
 \begin{itemize}
-\item Our own implementation of \texttt{std::simd} (\texttt{VectorizedArray}) now
-  also supports Arm Neon. Arm Neon is an architecture extension for
-  the Arm Cortex-A and Arm Cortex-R series of processors, commonly used in Apple
-  products. With these instructions 2 doubles or 4 floats can be processed in one
-  go. Since matrix-free infrastructure works directly with \texttt{VectorizedArray}
-  data structures, it automatically benefits from this implementation extension.
+\item Our own implementation of \texttt{std::simd}, called
+  \texttt{VectorizedArray}, now also supports Arm Neon. Arm Neon is an
+  architecture extension of the ARMv7, ARMv8 and ARMv9 architecture families,
+  used e.g.~for the Arm Cortex-A and Arm Cortex-R series of processors or in
+  the Apple 64-bit silicon series (A7--A18, M1--M4 as of 2024).
+  With these instructions, 2 doubles or 4 floats can be processed in one
+  go. Since the matrix-free infrastructure works directly with the \texttt{VectorizedArray}
+  abstraction as data structure, it automatically benefits from this new implementation.
 
-\item We made some improvement for evaluation of values and gradients for
-  $H$(div)-conforming Raviart--Thomas elements on non-Cartesian elements.
+\item The application of the Piola transformation for values and gradients of
+  $H$(div)-conforming Raviart--Thomas elements on non-Cartesian elements has
+  been rewritten for better performance. Together with additional changes in
+  the sum-factorization algorithms that are now more similar to the kernels
+  described in~\cite{KronbichlerKormann2019}, the matrix-free operator
+  evaluation is now three to five times faster in these cases.
 
-\item We furthermore improved the internal data structures of the
-  tensor-product evaluators as well as the evaluators for simplex elements,
-  which speed up the evaluation in several scenarios, especially multi-component systems.
+\item Furthermore, improved the internal data structures of the tensor-product
+  evaluators as well as the evaluators for simplex elements.  This speeds up
+  the operator-evaluation in several scenarios, especially for simplices
+  (around two times higher throughput for operator evaluation) and
+  multi-component systems.
 \end{itemize}
@@
 now allow to perform local smoothing, which is a key step towards unifying
-all transfer operators in \dealii. \texttt{MGTwoLevelTransfer} now also
-can be set up with an existing \texttt{MatrixFree} object in the case
-of $p$-multigrid, allowing to reduce the setup costs significantly.
+all transfer operators in \dealii. Finally, \texttt{MGTwoLevelTransfer} can now
+be set up with an existing \texttt{MatrixFree} object in the case
+of $p$-multigrid, allowing to reduce the setup costs and memory consumption significantly.
 
@@
 
-We made different advances to the non-matching infrastructure in \dealii. These include:
+For the non-matching infrastructure in \dealii, the following improvements were made:
 
@@
 (\texttt{MGTwoLevelTransferNon\-Nested}) has been improved significantly by avoiding
-redundant copy operations. Furthermore, the support for simplex-shaped cells and
-multiple-component elements have been added.
+redundant copy operations. Furthermore, support for simplex-shaped cells and
+multiple-component elements has been added.
+\item Several minor performance improvements for \dealii's evaluator class on
+  sets of unstructured points, \texttt{FEPointEvaluation}, have been
+  made. This is particularly useful for evaluating several quantities on the
+  same set of points, which is enabled by \texttt{NonMatching::MappingInfo}.
 \item The new class \texttt{FERemoteEvaluation} has been added. 
@@
 in a distributed matrix-free loop for non-matching discretizations.
-Interfaces are named with \texttt{FEEvaluation} in mind. Key component is the
-underlying MPI communication, which is done via \texttt{RemotePointEvaluation}.
+Interfaces are named in analogy to \texttt{FEEvaluation}, in order to seamlessly switch
+local evaluator functionality at quadrature points of a computation. The key component is the
+underlying MPI communication infrastructure, which is done via \texttt{RemotePointEvaluation}.
 Tutorial \texttt{step-89} has been added to present its usage
@@
 \epetra{} package that uses \MPI{} as its only source of
-parallelism. The \dealii{} interfaces to \trilinos{} are therefore all
-built using \epetra{}.
+parallelism. The \dealii{} interfaces to \trilinos{} have therefore traditionally all
+been built using \epetra{}.
 However, several years ago, \trilinos also introduced the
@@
     foundation. step-83 was written by Pasquale Africa, Wolfgang
-    Bangerth, and Bruno Blais and uses step-19 as its basis.
+    Bangerth, and Bruno Blais using step-19 as its basis.
   \item
@@
 f\"ur Technisch-Wissen\-schaft\-li\-ches Hoch- und H\"ochstleistungsrechnen
-(KONWIHR), projects ``High-order matrix-free finite
-element implementations with hybrid parallelization and improved data
-locality'' and ``Fast and scalable finite element algorithms for coupled
+(KONWIHR), project ``Fast and scalable finite element algorithms for coupled
 multiphysics problems and non-matching grids''.

### 2024-09-19 9.6/paper.tex
Adjust reference to two previous versions
Improve Tpetra section

@@
 
-\dealii{}'s parallel linear algebra facilities are largely built on
+\dealii{}'s parallel linear algebra facilities (apart from matrix-free capabilities) are largely built on
 wrappers around functionality provided by the \petsc{} and \trilinos{} libraries.
-As a software project, \trilinos{} is designed as a stack of semi-independent
-packages building upon each other and working together. 
-As a foundation, \teuchos{} offers a set of basic tools,
-including parameter handling through XML files or command line interfaces 
-and the reference counted pointer \texttt{RCP} which predates 
-the smart pointer types introduced in \texttt{C++11}.
-The second layer in the stack is a package providing distributed linear algebra 
-classes for vectors and (sparse) matrices based on the \petra{}
-object model. The third layer is a set of interconnected packages
-providing preconditioners, various types of solvers 
-as well as more advanced techniques like automatic differentiation, 
-time integration and discretization.
 
-Historically, \trilinos{} has implemented \petra{} objects -- such as
-parallel vector, sparsity pattern, and sparse matrix classes -- in its
+Historically, \trilinos{} has implemented distributed linear algebra 
+-classes for vectors and (sparse) matrices in its
 \epetra{} package that uses \MPI{} as its only source of
@@
 
-In the current release, we have put substantial work into this switch.
+In the current release, we have put substantial work into this switch. The following omits the \texttt{LinearAlgebra} prefix for readability.
+\begin{itemize}
+  \item The \texttt{TpetraWrappers::Vector} class has been overhauled to also allow specifying
+    a memory space, and the \texttt{TpetraWrappers::BlockVector} class has been added.
+  \item \texttt{TpetraWrappers::SparseMatrix} and \texttt{TpetraWrappers::SparsityPattern} has been implemented mirroring
+    the functionality of the \texttt{TrilinosWrappers::SparseMatrix} \\ and \texttt{TrilinosWrappers::SparsityPattern} classes.
+  \item Various \texttt{TpetraWrappers::Precondition*} have been added largely
+    mirroring the exising preconditioners in the \texttt{LinearAlgbera::TrilinosWrappers} namespace.
+  \item \texttt{TpetraWrappers::SolverDirect} and \texttt{TpetraWrappers::SolverDirectKLU2}
+    have been implemented as direct solvers.
+\end{itemize}
+Most of the interfaces are usable at this point even if some functionality existing in the \texttt{TrilinosWrappers} classes
+is still missing, most noticably a wrapper for an algebraic-multigrid preconditioner.
 
-\todo[inline]{Sebastian \& Jan Philipp: Please complete}
-
-\begin{itemize}
-  \item Short overview capabilities of \epetra (MPI parallel, fixed types) [Thiele]
-  \item Short overview capabilities of \tpetra (MPI+X via Kokkos, auto diff types, Templated petra) [Thiele]
-    \todo[inline]{Keep discussion of the two items above as short as
-      you can -- we're describing the changes in deal.II, not some
-      other package. Perhaps roll what you want to say into the
-      paragraphs above?}
-  \item Current state of the interface (Tpetra based linear algebra, direct solver from Amesos2, preconditioners from Ifpack2) [Thiele \& Kinnewig]
-  \item User relevant changes from \epetra: incompatibilities, removed
-    options, changes in the constructor. [Thiele \& Kinnewig]
-    \todo[inline]{These last two items are the important ones.}
-\end{itemize}
+All the \texttt{TpetraWrappers} classes take \texttt{Number} and \texttt{MemorySpace} template arguments
+allowing to control the underlying number type and the memory space used for storing the data. Note that the memory space type
+also determines whether \tpetra{} takes advantage of GPUs if \texttt{Kokkos} has been configure with a GPU backend.
 
@@
 The two previous releases of \dealii can be cited as
-\cite{dealII92,dealII93}.
+\cite{dealII94,dealII95}.
 

### 2024-09-19 9.6/paper.tex
Remove trailing whitespaces
Typo

@@
     a memory space, and the \texttt{TpetraWrappers::BlockVector} class has been added.
-  \item \texttt{TpetraWrappers::SparseMatrix} and \texttt{TpetraWrappers::SparsityPattern} has been implemented mirroring
+  \item \texttt{TpetraWrappers::SparseMatrix} and \texttt{TpetraWrappers::SparsityPattern} have been implemented mirroring
     the functionality of the \texttt{TrilinosWrappers::SparseMatrix} \\ and \texttt{TrilinosWrappers::SparsityPattern} classes.

### 2024-09-19 9.6/paper.tex
Add missing word

@@
     the functionality of the \texttt{TrilinosWrappers::SparseMatrix} \\ and \texttt{TrilinosWrappers::SparsityPattern} classes.
-  \item Various \texttt{TpetraWrappers::Precondition*} have been added largely
+  \item Various \texttt{TpetraWrappers::Precondition*} classes have been added largely
     mirroring the exising preconditioners in the \texttt{LinearAlgbera::TrilinosWrappers} namespace.

### 2024-09-19 9.6/paper.tex
Improve the orientation paragraphs.

@@
     constraint in one step.
-  \item \dealii{} internally needs to keep track of the relative
-    orientations of the coordinate systems associated with neighboring
-    cells, as well as of the coordinate systems of faces and edges
-    within the coordinate system of the cells they are part
-    of. Historically, different places within the library that had to
-    do that had grown their own, independent ways of representating
-    this information, much of which was difficult to understand.
-    We have now cleaned up this area. Specifically, instead of using three
-    Booleans to define relative orientations for hexedra (\textit{orientation},
-    \textit{flip}, and \textit{rotation}), we have internally merged these three values into a single variable of type
-     \texttt{char} -- i.e., the smallest integer data type that has at
-     least three bits to represent the information. Because the way to represent
-     orientations is part of the public interface, the change causes
-     some incompatibilities in types and function signatures.
-  \todo[inline]{DavidW: Please describe some details here. -- When
-    done, please remove this comment.}
+  \item We refactored the systems for managing relative line and face
+    orientations to significantly improve consistency across several library
+    modules, including finite element classes, periodicity, and the
+    \texttt{p4est} interface. \dealii{} has supported using unstructured
+    three-dimensional meshes for a long time. Features like discontinuous
+    Galerkin methods and higher-order elements require that lines and faces of
+    adjacent elements are consistently oriented. For example, consider a line
+    containing two degrees of freedom, which, in reference coordinates, are
+    located at $x_1 = \frac 1 3$ and $x_2 = \frac 2 3$. If the line is read from
+    left-to-right then we get points $(x_1, x_2)$ whereas if it is read from
+    right-to-left we get $(x_2, x_1)$. \dealii{} guarantees that this order will
+    be consistent across all cells sharing that line by storing a flag
+    indicating whether (relative to the vertex enumeration on the present cell)
+    that particular line is in the standard or reversed orientation. Faces in 3d
+    are treated in a similar way. The only significant difference between lines
+    and 3d faces is that, instead of two possible relative orientations,
+    triangular faces have six and quadrilateral faces have eight.
+
+    Historically, different places within the library implemented their own
+    ad-hoc orientation logic, data encodings, and default values. We cleaned up
+    this subsystem and instead of using three different booleans
+    (\emph{orientation}, \emph{rotation}, and \emph{flip}), bitsets, or several
+    other encodings, we use a single \texttt{unsigned char} which we named the
+    \emph{combined orientation}. Since the representation in some library
+    interfaces has changed, most of the old interfaces have been deprecated and
+    a few (such as the low-level interface to periodic boundary conditions) had
+    to be incompatibly altered or removed.
 \item The \texttt{SolverGMRES} class now offers a third orthogonalization

### 2024-09-20 9.6/paper.tex
Minor edits.
Extend the description of the TpetraWrappers classes.

@@
 
-In the current release, we have put substantial work into this switch. The following omits the \texttt{LinearAlgebra} prefix for readability.
+In the current release, we have put substantial work into this switch,
+as outlined below. In the following list, we omit the common namespace
+\texttt{LinearAlgebra::} prefix on all mentioned
+\texttt{TpetraWrappers} symbols for readability:
 \begin{itemize}
@@
 \end{itemize}
-Most of the interfaces are usable at this point even if some functionality existing in the \texttt{TrilinosWrappers} classes
-is still missing, most noticably a wrapper for an algebraic-multigrid preconditioner.
+The preconditioner available in the \texttt{TpetraWrappers::Precondition*} classes can be used along with the existing 
+iterative solvers, e.g., CG, GMRES, etc.; therefore, the interface is already usable. 
+Even if some functionality exists, the \texttt{TrilinosWrappers} classes are missing, most noticably a wrapper for an 
+algebraic-multigrid preconditioner.
 
-All the \texttt{TpetraWrappers} classes take \texttt{Number} and \texttt{MemorySpace} template arguments
-allowing to control the underlying number type and the memory space used for storing the data. Note that the memory space type
-also determines whether \tpetra{} takes advantage of GPUs if \texttt{Kokkos} has been configure with a GPU backend.
+The design goal was to introduce as few changes as possible for the user to allow for an easy transition to the 
+\texttt{TpetraWrappers} classes. 
+However, based on the difference between \epetra{} and \tpetra{}, the following changes where necessary:
+\begin{itemize}
+  \item The constructor for the class \texttt{TrilinosWrappers::MPI::Vector} only required the \texttt{IndexSet locally\_owned}, 
+    which describes the set of indices locally owned by the current rank. 
+    However, if one wants to create a vector that has read or write access to non-locally owned indices,  the constructor of the 
+    \texttt{TpetraWrappers::Vector} requires \texttt{IndexSet locally\_owned}, the \texttt{IndexSet locally\_relevat}, and 
+    a boolean flag, whether to initialize the vector in the read-only or the write-only state.
+  \item A vector created without providing a \texttt{IndexSet locally\_relevat} is purely local and cannot access non-local indices. 
+    Such a vector can not be copied to a vector that can access non-local indices, as the \texttt{IndexSet locally\_relevat} must 
+    be provided at creating the vector object.
+\end{itemize}
+
+One other important change is that
+all the \texttt{TpetraWrappers} classes take \texttt{Number} and \texttt{MemorySpace} template arguments,
+which control the underlying number type and memory space. 
+By default, the number type is set to  \texttt{double}, and memory is allocated on the CPU, mirroring the 
+behavior of the \texttt{TrilinosWrappers}. 
+The support for different number types also enables the usage of automatic differentiation-based tools in 
+\trilinos{}, such as \texttt{NOX}. 
+Additionally, the memory space type determines whether \tpetra{} can utilize GPUs if \kokkos{} is configured 
+with a GPU backend. 
+To make the automatic differentiation tools of \trilinos{} accessible through the \tpetra{} interface without 
+requiring users to have an in-depth understanding of \trilinos{}, additional wrappers for the corresponding 
+\trilinos{} packages will be added in future releases. 
+Furthermore, further work is needed to fully leverage the GPU backend and unlock its full potential for users.
+
+Moreover, the \tpetra{} interface allows access to more functionality and preconditioners, e.g., \texttt{FROSch}, 
+available in \trilinos{}, which will be made accessible through the \texttt{TpetraWrappers} in the future.
 

### 2024-09-20 9.6/paper.tex
Add founding support.

--- 2024-09-19 9.6/paper.tex
@@
 
-\todo[inline]{Sebastian: Anything to list here?}
+Sebastian Kinnewig is supported by the Deutsche Forschungsgemeinschaft (DFG) under Germany’s Excellence
+Strategy within the Cluster of Excellence PhoenixD (EXC 2122, Project ID 390833453)
 

### 2024-09-21 9.6/paper.tex
Small changes

@@
 
- \author[1*]{Pasquale~Claudio~Africa}
+ \author[1]{Pasquale~Claudio~Africa}
  \affil[1]{SISSA International School for Advanced Studies,
@@
 
-\author[2]{Bruno~Turcksin}
+\author[2*]{Bruno~Turcksin}
 
@@
 
-\author[18,8]{Vladimir Yushutin}
+\author[8, 18]{Vladimir Yushutin}
 \affil[18]{Department of Mathematics, University of Tennessee at Knoxville,
@@
     Section~\ref{sec:nonmatching}). In particular, the new class
-    FERemoteEvaluation provides support for evaluating
+    \texttt{FERemoteEvaluation} provides support for evaluating
     finite element shape functions and solutions on parts of the
@@
   implementations, making them use the same underlying kernels as much as possible.
+\item The new class \texttt{FE\_Hermite} implements a Hermite interpolation basis of
+maximum regularity. These bases are always of odd polynomial degree $p$ and have
+regularity $r=(p-1)/2$.  
 \end{itemize}
@@
 
-Furthermore, we performed substantial improvements to the global-coarsening
+In addition, we performed substantial improvements to the global-coarsening
 multigrid infrastructure: \texttt{MGTransferMF} (previously:
@@
 Furthermore, we added an example to the \texttt{libCEED}
-library.\footnote{\url{https://github.com/CEED/libCEED}} \texttt{libCEED} is a library
+library~\cite{brown2021libceed}: \url{https://github.com/CEED/libCEED/tree/main/examples/deal.II}. \texttt{libCEED} is a library
 that provides matrix-free evaluation routines that work on different hardware. The
-examples show how to interface the \dealii data structures to the \texttt{libCEED} ones.
+example show how to interface the \dealii data structures to the \texttt{libCEED} ones.
 The example solves the BP1-BP6 benchmarks (scalar/vector Laplace/mass matrix with

### 2024-09-21 9.6/paper.tex
Small fixes

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Updates to multigrid and matrix-free algorithms}\label{sec:mf}
+\subsection{Updates of multigrid and matrix-free algorithms}\label{sec:mf}
 
-%\todo[inline]{Martin/Peter: The list in the introduction also mentions the two-level
-%  operators. Is this the right place to also discuss these?}
-
-We made different updates to the matrix-free infrastructure in \dealii. These include:
+We made different updates of the matrix-free infrastructure in \dealii. These include:
 \begin{itemize}
@@
   architecture extension of the ARMv7, ARMv8 and ARMv9 architecture families,
-  used e.g.~for the Arm Cortex-A and Arm Cortex-R series of processors or in
+  used, e.g., for the Arm Cortex-A and Arm Cortex-R series of processors or in
   the Apple 64-bit silicon series (A7--A18, M1--M4 as of 2024).
@@
 \item The application of the Piola transformation for values and gradients of
-  $H$(div)-conforming Raviart--Thomas elements on non-Cartesian elements has
+  $H$(div)-conforming Raviart--Thomas elements on non-Cartesian cells has
   been rewritten for better performance. Together with additional changes in
-  the sum-factorization algorithms that are now more similar to the kernels
+  the sum-factorization algorithms, which became more similar to the kernels
   described in~\cite{KronbichlerKormann2019}, the matrix-free operator
@@
 
-\item Furthermore, improved the internal data structures of the tensor-product
+\item Furthermore, we improved the internal data structures of the tensor-product
   evaluators as well as the evaluators for simplex elements.  This speeds up
-  the operator-evaluation in several scenarios, especially for simplices
+  the operator evaluation in several scenarios, especially for simplices
   (around two times higher throughput for operator evaluation) and
@@
 
-For the non-matching infrastructure in \dealii, the following improvements were made:
+In the non-matching infrastructure of \dealii, the following improvements were made:
 
@@
 multiple-component elements has been added.
-\item Several minor performance improvements for \dealii's evaluator class on
+\item Several minor performance improvements in \dealii's evaluator class on
   sets of unstructured points, \texttt{FEPointEvaluation}, have been
@@
 in a distributed matrix-free loop for non-matching discretizations.
-Interfaces are named in analogy to \texttt{FEEvaluation}, in order to seamlessly switch
+Interfaces are named in analogy to \texttt{FEEvaluation}, in order to seamlessly switch the
 local evaluator functionality at quadrature points of a computation. The key component is the
-underlying MPI communication infrastructure, which is done via \texttt{RemotePointEvaluation}.
+underlying MPI communication infrastructure, which is performed via \texttt{RemotePointEvaluation}.
 Tutorial \texttt{step-89} has been added to present its usage
-in the context of an application to acoustic conservation equations~\cite{heinz2023high}.
+in the context of the application to acoustic conservation equations~\cite{heinz2023high}.
 \item FECouplingValues
@@
     (University of Augsburg/Uppsala University). It
-    presents the advanced point-evaluation functionalities of \dealii
-    that are useful for evaluating finite element solutions at
-    arbitrary points on finite element meshes that can be distributed among processes.
+    presents the advanced point-evaluation functionalities of \dealii,
+    which are useful for evaluating finite element solutions at
+    arbitrary points on meshes that can be distributed among processes.
   \item
@@
     Marco Feder (SISSA), and Peter Munch (University of Augsburg/Uppsala University).
-    It shows one way how to apply non-matching and/or Chimera methods
+    It shows a way how to apply non-matching and/or Chimera methods
     within matrix-free loops in \dealii.
@@
 library~\cite{brown2021libceed}: \url{https://github.com/CEED/libCEED/tree/main/examples/deal.II}. \texttt{libCEED} is a library
-that provides matrix-free evaluation routines that work on different hardware. The
-example show how to interface the \dealii data structures to the \texttt{libCEED} ones.
-The example solves the BP1-BP6 benchmarks (scalar/vector Laplace/mass matrix with
+that provides matrix-free evaluation routines for different hardware. The
+example shows how to interface the \dealii data structures with the \texttt{libCEED} ones
+and solves the BP1-BP6 benchmarks (scalar/vector Laplace/mass matrix with
 regular integration and over-integration).

### 2024-09-22 9.6/paper.tex
Minor corrections for the description of the TpetraWrapper classes.

--- 2024-09-20 9.6/paper.tex
@@
 \end{itemize}
-Most of the interfaces are usable at this point even if some functionality existing in the \texttt{TrilinosWrappers} classes
-is still missing, most noticably a wrapper for an algebraic-multigrid preconditioner.
+The preconditioner available in the \texttt{TpetraWrappers::Precondition*} classes can be used along with the existing 
+iterative solvers, e.g., CG, GMRES, etc.; therefore, the interface is already usable. 
+Even if some functionality exists, the \texttt{TrilinosWrappers} classes are missing, most noticably a wrapper for an 
+algebraic-multigrid preconditioner.
 
-All the \texttt{TpetraWrappers} classes take \texttt{Number} and \texttt{MemorySpace} template arguments
-allowing to control the underlying number type and the memory space used for storing the data. Note that the memory space type
-also determines whether \tpetra{} takes advantage of GPUs if \texttt{Kokkos} has been configure with a GPU backend.
+The design goal was to introduce as few changes as possible for the user to allow for an easy transition to the 
+\texttt{TpetraWrappers} classes. 
+However, based on the difference between \epetra{} and \tpetra{}, the following changes were necessary:
+\begin{itemize}
+  \item The constructor for the class \texttt{TrilinosWrappers::MPI::Vector} only required the \texttt{IndexSet locally\_owned}, 
+    which describes the set of indices locally owned by the current rank. 
+    However, if one wants to create a vector that has read or write access to non-locally owned indices,  the constructor of the 
+    \texttt{TpetraWrappers::Vector} requires \texttt{IndexSet locally\_owned}, the \texttt{IndexSet locally\_relevat}, and 
+    a boolean flag, whether to initialize the vector in the read-only or the write-only state.
+  \item A vector created without providing a \texttt{IndexSet locally\_relevat} is purely local and cannot access non-local indices. 
+    Such a vector can not be copied to a vector that can access non-local indices, as the \texttt{IndexSet locally\_relevat} must 
+    be provided at creating the vector object.
+\end{itemize}
+
+One other important change is that
+all the \texttt{TpetraWrappers} classes take \texttt{Number} and \texttt{MemorySpace} template arguments,
+which control the underlying number type and memory space. 
+By default, the memory is allocated on the CPU, mirroring the behavior of the \texttt{TrilinosWrappers}. 
+The support for different number types also enables the usage of automatic differentiation-based tools in 
+\trilinos{}, such as \texttt{NOX}. 
+Additionally, the memory space type determines whether \tpetra{} can utilize GPUs if \kokkos{} is configured 
+with a GPU backend. 
+To make the automatic differentiation tools of \trilinos{} accessible through the \tpetra{} interface without 
+requiring users to have an in-depth understanding of \trilinos{}, additional wrappers for the corresponding 
+\trilinos{} packages will be added in future releases. 
+Furthermore, to fully leverage the GPU backend and unlock its full potential for users will require further work.
 
@@
 
-Sebastian Kinnewig is supported by the Deutsche Forschungsgemeinschaft (DFG) under Germany’s Excellence
-Strategy within the Cluster of Excellence PhoenixD (EXC 2122, Project ID 390833453)
+\todo[inline]{Sebastian: Anything to list here?}
 

### 2024-09-22 9.6/paper.tex
Various minor changes.

@@
 development of finite element solvers. It is available for free under the
-GNU Lesser General Public License (LGPL). Downloads are available at
+GNU Lesser General Public License (LGPL). \todo{Update license} Downloads are available at
 \url{https://www.dealii.org/} and \url{https://github.com/dealii/dealii}.
@@
     domain stored by other MPI processes.
-  \item Much work has gone into writing wrappers for Trilinos' Tpetra
-    stack of linear algebra classes. Tpetra is Trilinos' Kokkos-based
-    replacement for the now-deprecated Epetra stack. See
+  \item Much work has gone into writing wrappers for \trilinos' \tpetra{}
+    stack of linear algebra classes. \tpetra{} is \trilinos' \kokkos-based
+    replacement for the now-deprecated \epetra{} stack. See
     Section~\ref{sec:tpetra} for more on this.
@@
     (step-83), integrating time-dependent solvers with external time
-    stepping libraries (step-86, using PETSc's TS library), advanced
+    stepping libraries (step-86, using \petsc's \ts{} library), advanced
     point evaluation techniques (step-87), non-matching grids
@@
 \begin{itemize}
-  \item deal.II now requires and uses of C++17.
+  \item \dealii{} now requires and makes use of the C++17 language standard.
   \item We have continued to make progress in supporting simplex and
@@
     modules, including finite element classes, periodicity, and the
-    \texttt{p4est} interface. \dealii{} has supported using unstructured
+    \pfrst{} interface. \dealii{} has supported using unstructured
     three-dimensional meshes for a long time. Features like discontinuous
@@
 %
-The changelog lists more than 180 other features and bugfixes.
 
@@
 {in the file that lists all changes for this release}; see \cite{changes96}.
+The changelog lists more than 180 features and bugfixes.
 
@@
   multi-component systems.
-\end{itemize}
 
-In addition, we performed substantial improvements to the global-coarsening
+\item In addition, we performed substantial improvements to the global-coarsening
   multigrid infrastructure: \texttt{MGTransferMF} (previously:
-\texttt{MGTransferGlobalCoarsening}) and \texttt{MGTwoLevelTransfer}. They
+  \texttt{MGTransferGlobalCoarsening}) and \texttt{MGTwoLevel\-Transfer}. They
   now allow to perform local smoothing, which is a key step towards unifying
@@
   of $p$-multigrid, allowing to reduce the setup costs and memory consumption significantly.
+\end{itemize}
 
@@
 underlying MPI communication infrastructure, which is performed via \texttt{RemotePointEvaluation}.
-Tutorial \texttt{step-89} has been added to present its usage
+Tutorial step-89 has been added to present its usage
 in the context of the application to acoustic conservation equations~\cite{heinz2023high}.
@@
 iterative solvers, e.g., CG, GMRES, etc.; therefore, the interface is already usable.
-Even if some functionality exists, the \texttt{TrilinosWrappers} classes are missing, most noticably a wrapper for an 
+However, some functionality in the \texttt{TrilinosWrappers} classes is still missing, most noticeably a wrapper for an
 algebraic-multigrid preconditioner.
@@
     However, if one wants to create a vector that has read or write access to non-locally owned indices,  the constructor of the
-    \texttt{TpetraWrappers::Vector} requires \texttt{IndexSet locally\_owned}, the \texttt{IndexSet locally\_relevat}, and 
+    \texttt{TpetraWrappers::Vector} requires \texttt{IndexSet locally\_owned}, the \texttt{IndexSet locally\_relevant}, and
     a boolean flag, whether to initialize the vector in the read-only or the write-only state.
-  \item A vector created without providing a \texttt{IndexSet locally\_relevat} is purely local and cannot access non-local indices. 
-    Such a vector can not be copied to a vector that can access non-local indices, as the \texttt{IndexSet locally\_relevat} must 
+  \item A vector created without providing a \texttt{IndexSet locally\_relevant} is purely local and cannot access non-local indices.
+    Such a vector can not be copied to a vector that can access non-local indices, as the \texttt{IndexSet locally\_relevant} must
     be provided at creating the vector object.
@@
     demonstrates how one can implement
-    checkpoint/restart functionality in \dealii{}-based programs,
-    using the BOOST serialization functionality as a
+    checkpoint/restart functionality in \dealii-based programs,
+    using the \boost{} serialization functionality as a
     foundation. step-83 was written by Pasquale Africa, Wolfgang
     Bangerth, and Bruno Blais using step-19 as its basis.
+    \todo{Also add affiliations like for the other tutorials?}
   \item
     \texttt{step-86}
-    is a program that solves the heat equation using PETSc's TS (time
+    is a program that solves the heat equation using \petsc's \ts{} (time
     stepping) framework for the solution of ordinary differential
@@
 research rather than as teaching tools):
-UPDATE
+\todo{UPDATE}
 \begin{itemize}
@@
 that provides matrix-free evaluation routines for different hardware. The
-example shows how to interface the \dealii data structures with the \texttt{libCEED} ones
+example shows how to interface the \dealii{} data structures with the \texttt{libCEED} ones
 and solves the BP1-BP6 benchmarks (scalar/vector Laplace/mass matrix with
@@
     have been deprecated and will be removed in the next
-    release. Kokkos is now used for device-specific optimizations.
+    release. \kokkos{} is now used for device-specific optimizations.
 \end{itemize}

### 2024-09-23 9.6/paper.tex
Move changelog mention to old place.
Add Tpetra solver and preconditioner details

@@
 %
+The changelog lists more than 180 features and bugfixes.
 
@@
 {in the file that lists all changes for this release}; see \cite{changes96}.
-The changelog lists more than 180 features and bugfixes.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Interface to the \trilinos{} \tpetra stack}\label{sec:tpetra}
+\subsection{Interface to the \trilinos{} \tpetra{} stack}\label{sec:tpetra}
 
@@
 
-Historically, \trilinos{} has implemented distributed linear algebra
--classes for vectors and (sparse) matrices in its
+Historically, \trilinos{} has implemented distributed linear algebra-classes for vectors and (sparse) matrices in its
 \epetra{} package that uses \MPI{} as its only source of
-parallelism. The \dealii{} interfaces to \trilinos{} have therefore traditionally all
-been built using \epetra{}.
-However, several years ago, \trilinos also introduced the
+parallelism. Building on that, there are multiple interconnected \trilinos{} packages, e.g.\ for (non)-linear solvers 
+and preconditioners, commonly referred to as the 
+\epetra{} stack. For sake of brevity we will also refer to the whole stack as \epetra{} in the following.
+As \epetra{} was the first stack, \dealii{} interfaces to \trilinos{} have traditionally 
+been implemented to use it.
+
+However, several years ago, \trilinos{} also introduced the
 newer \tpetra{} (``templated'' Petra) package that provides additional shared memory
-parallelism and GPU capabilities by building on \kokkos{}. \tpetra{}
-is slated to eventually replace \epetra{}, and the latter is indeed
-now deprecated with removal slated for 2025. As a consequence, we will
-eventually have to switch all of our \trilinos{} interfaces to
-\tpetra{}; the same is true for the need to switch to \tpetra{}-based
-sub-packages replacing existing \trilinos sub-packages (for example,
-Ifpack2 instead of Ifpack).
+parallelism and GPU capabilities, referred to as \MPIx, by building on \kokkos{}. 
+Since a few years, new features have only been implemented in \tpetra{} as 
+it is slated to replace \epetra{}, and the latter is indeed
+now deprecated with removal slated for 2025. 
+As a consequence, 
+we will eventually have to switch all of our \trilinos{} interfaces to \tpetra{}; 
+the same is true for the need to switch to \tpetra{}-based
+sub-packages replacing existing \trilinos{} sub-packages 
+(for example, Ifpack2 instead of Ifpack).
 
 In the current release, we have put substantial work into this switch,
-as outlined below. In the following list, we omit the common namespace
-\texttt{LinearAlgebra::} prefix on all mentioned
+as outlined below. As this is not yet finished, it is important to note that 
+all these features are optional and their individual availability
+depends on the installed \trilinos{} packages. 
+As an example, a \dealii{} build will pass with a \trilinos{} installation 
+without \texttt{Ifpack2}, but its preconditioners will not be included.
+In the following list, we omit the common namespace prefix
+\texttt{LinearAlgebra::} on all mentioned
 \texttt{TpetraWrappers} symbols for readability:
@@
     the functionality of the \texttt{TrilinosWrappers::SparseMatrix} \\ and \texttt{TrilinosWrappers::SparsityPattern} classes.
-  \item Various \texttt{TpetraWrappers::Precondition*} classes have been added largely
-    mirroring the exising preconditioners in the \texttt{LinearAlgbera::TrilinosWrappers} namespace.
-  \item \texttt{TpetraWrappers::SolverDirect} and \texttt{TpetraWrappers::SolverDirectKLU2}
-    have been implemented as direct solvers.
+  \item Various incomplete factorization and relaxation preconditioners (from \texttt{Ifpack2}) have been wrapped,
+    mirroring the existing classes in \texttt{LinearAlgbera::TrilinosWrappers} as closely as possible.
+  \item Two new preconditioner variants \texttt{TpetraWrappers::PreconditionL1Jacobi} and \\
+    \texttt{TpetraWrappers::PreconditionL1GaussSeidel}, based on~\cite{BFKY2011}, have been added.
+  \item \texttt{TpetraWrappers::SolverDirectKLU2} has been wrapped as a direct solver.
 \end{itemize}
-The preconditioner available in the \texttt{TpetraWrappers::Precondition*} classes can be used along with the existing
-iterative solvers, e.g., CG, GMRES, etc.; therefore, the interface is already usable.
-However, some functionality in the \texttt{TrilinosWrappers} classes is still missing, most noticeably a wrapper for an
-algebraic-multigrid preconditioner.
+Some functionality in the \texttt{TrilinosWrappers} classes is still missing, 
+most noticeably wrappers for the algebraic-multigrid preconditioner \texttt{MueLu}
+and the iterative solvers from \texttt{Belos}.
+However, the wrapped \texttt{Ifpack2} preconditioners can already be used with the iterative solvers of \dealii{}.
+
+The solver and preconditioner classes mentioned above provide reasonable parameter subsets 
+through \texttt{AdditionalData} objects, as in \texttt{TrilinosWrappers}. 
+Additionally, there are new generic classes \texttt{TpetraWrappers::SolverDirect} 
+and \texttt{TpetraWrappers::PreconditionIfpack} which expose the internal interface through
+a \texttt{Teuchos::ParameterList} and thereby offering the full set of parameters for 
+more experienced \trilinos{} users. This also allows the use of preconditioners or 
+solvers not (yet) wrapped, e.g.\ \texttt{SuperLU\_dist} or \texttt{MUMPS}.
 
@@
     be provided at creating the vector object. 
+  \item Some of solver and preconditioner parameters are not available in \tpetra{},
+    such that the \texttt{AdditionalData} objects are not identical. 
+    Since there are many preconditioners we will not list each individual change but instead refer 
+    to our Doxygen documentation. 
+  \item The incomplete Cholesky (IC) factorization preconditioner is not available 
+   in \tpetra{}.
 \end{itemize}
+
+Another goal was increasing the interoperability of \dealii{} and \trilinos{},
+allowing users familiar with both to write `pure' \trilinos{} code within their 
+applications, e.g.\ to test or develop a new feature.
+Therefore, the internal data is stored as it would be in a \trilinos{} 
+application code (as \texttt{Teuchos::RCP}) and can be accessed through member functions. 
+In fact, one of the authors has already used this to apply the domain decomposition preconditioner
+\texttt{FROSch} to various problems discretized using \dealii{}.
+\todo[inline]{Kinnewig: Add citation to preprint?}
 

### 2024-09-23 9.6/paper.tex
Remove funding todo for JPT
update intro

@@
 
-\dealii version 9.6.0 was released August 11, 2024.
-This paper provides an
+\dealii version 9.6.0 was released August 11, 2024. This paper provides an
 overview of the new features of this release and serves as a citable
@@
 development of finite element solvers. It is available for free under the
-GNU Lesser General Public License (LGPL). \todo{Update license} Downloads are available at
-\url{https://www.dealii.org/} and \url{https://github.com/dealii/dealii}.
+terms of the \emph{GNU Lesser General Public License} (LGPL). The \dealii
+project is in the process of relicsensing the library under the terms of
+the \emph{Apache License} 2.0 with \emph{LLVM Exception}. Downloads are
+available at \url{https://www.dealii.org/} and
+\url{https://github.com/dealii/dealii}.
 
@@
 \begin{itemize}
-  \item Substantial performance improvements to the matrix-free and multigrid
+  \item
+    Substantial performance improvements to the matrix-free and multigrid
     infrastructure (Section~\ref{sec:mf}).
-  \item Different additions to the non-matching infrastructure (see
+  \item
+    Different additions to the non-matching infrastructure (see
     Section~\ref{sec:nonmatching}). In particular, the new class
-    \texttt{FERemoteEvaluation} provides support for evaluating
-    finite element shape functions and solutions on parts of the
-    domain stored by other MPI processes.
-  \item Much work has gone into writing wrappers for \trilinos' \tpetra{}
-    stack of linear algebra classes. \tpetra{} is \trilinos' \kokkos-based
+    \texttt{FERemoteEvaluation} provides support for evaluating finite
+    element shape functions and solutions on parts of the domain stored by
+    other MPI processes.
+  \item
+    Much work has gone into writing wrappers for \trilinos' \tpetra{} stack
+    of linear algebra classes. \tpetra{} is \trilinos' \kokkos-based
     replacement for the now-deprecated \epetra{} stack. See
     Section~\ref{sec:tpetra} for more on this.
-  \item Tool classes \texttt{TaskResult} and \texttt{Lazy} that provide ways to compute
-    values on a separate thread or only when first accessed (Section~\ref{sec:tools}).
-  \item There are six new tutorial programs, on checkpointing simulations
+  \item
+    Tool classes \texttt{TaskResult} and \texttt{Lazy} that provide ways to
+    compute values on a separate thread or only when first accessed
+    (Section~\ref{sec:tools}).
+  \item
+    There are six new tutorial programs, on checkpointing simulations
     (step-83), integrating time-dependent solvers with external time
     stepping libraries (step-86, using \petsc's \ts{} library), advanced
-    point evaluation techniques (step-87), non-matching grids
-    (step-89), and trace-based methods for PDEs on embedded surfaces
-    (step-90). See Section~\ref{subsec:steps} for more details.
+    point evaluation techniques (step-87), non-matching grids (step-89),
+    and trace-based methods for PDEs on embedded surfaces (step-90). See
+    Section~\ref{subsec:steps} for more details.
+  \item
+    In Section~\ref{sec:license} we summarize the motivation and approach
+    behind our relicensing effort to make \dealii available under the terms
+    of the the Apache License 2.0 with LLVM Exception.
 \end{itemize}
 %
-
-\todo[inline]{If we keep the section on relicensing
-  (Section~\ref{sec:license}), mention it in the list above.}
 
@@
 
-\todo[inline]{Jan Philipp: Anything to list here?}
-
 D.~Wells was supported by NSF Award OAC-1931516.

### 2024-09-23 9.6/paper.tex
add my todo list

@@
 Furthermore, we added an example to the \texttt{libCEED}
-library~\cite{brown2021libceed}: \url{https://github.com/CEED/libCEED/tree/main/examples/deal.II}. \texttt{libCEED} is a library
-that provides matrix-free evaluation routines for different hardware. The
-example shows how to interface the \dealii{} data structures with the \texttt{libCEED} ones
-and solves the BP1-BP6 benchmarks (scalar/vector Laplace/mass matrix with
-regular integration and over-integration).
+library~\cite{brown2021libceed}:
+\url{https://github.com/CEED/libCEED/tree/main/examples/deal.II}.
+\texttt{libCEED} is a library that provides matrix-free evaluation routines
+for different hardware. The example shows how to interface the \dealii{}
+data structures with the \texttt{libCEED} ones and solves the BP1-BP6
+benchmarks (scalar/vector Laplace/mass matrix with regular integration and
+over-integration).
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Relicensing to Apache 2.0}\label{sec:license}
+\subsection{Relicensing to Apache License 2.0 with LLVM exception}
+\label{sec:license}
 
-\todo[inline]{Matthias: Where are we with this, and what of it is
-  worth describing here?}
+\color{red}
+\begin{itemize}
+  \item
+    Motivation for license change, goals: make contributions from industry
+    possible / improve deal.II adoption. Mention reasons against LGPL 2.1,
+    Explaing why GPL-3/LGPL-3 is not a good fit. Explain why Apache 2.0
+    with LLVM exception.
+  \item
+    Successful examples LLVM and Kokkos (with same license)
+  \item
+    Status of external collaborators
+  \item
+    Approach: Require all new contributions to be dual licensed
+\end{itemize}
+\color{black}
 

### 2024-09-23 9.6/paper.tex
Add PM ackn
some text explaining the relicensing effort

@@
 \newcommand{\kokkos}{{\specialword{Kokkos}}\xspace}
+\newcommand{\llvm}{{\specialword{LLVM}}\xspace}
 
@@
 
-\color{red}
-\begin{itemize}
-  \item
-    Motivation for license change, goals: make contributions from industry
-    possible / improve deal.II adoption. Mention reasons against LGPL 2.1,
-    Explaing why GPL-3/LGPL-3 is not a good fit. Explain why Apache 2.0
-    with LLVM exception.
-  \item
-    Successful examples LLVM and Kokkos (with same license)
-  \item
-    Status of external collaborators
-  \item
-    Approach: Require all new contributions to be dual licensed
-\end{itemize}
-\color{black}
+The \dealii project is in the process of relicensing the library under the
+terms of the \emph{Apache License 2.0 with LLVM
+Exception} \cite{Apache-2.0,LLVM-Exception}. To this end we require all new
+code contributions to be dual licensed under the current license (GNU
+Lesser General Public License v2.1 \cite{LGPL-2.1}) and the new license
+(Apache-2.0 with LLVM-exception).
+
+We have chosen to pursue a relicensing effort for deal.II because of some
+long standing problems with the LGPL v2.1 license. Most notably, despite
+our clear intent that the strong copyleft principle should only apply to
+deal.II source code and not user projects (which should be free to choose
+their own license freely) this is not necessarily the case for LGPL v2.1
+due to the language used in the library. This makes potential industry
+partners that evaluate the use of deal.II for their own projects nervous.
+Secondly, our previous choice to use an open source license with strong
+copyleft was motivated by our desire to receive back code contributions
+from third parties, but it is our observation over the last 10 years that
+it hardly helps achieving this goal: Copyleft only applies when a derived
+software is sold or published. But almost all of our code contributions are
+voluntary and come from individual contributors with an academic
+background, where these considerations do not apply.
+
+We have thus decided to switch away from the LGPL v2.1 license towards a
+more permissible license. We settled on the Apache License 2.0 with
+LLVM-exception in large parts because it is (a) a permissible open source
+license with patent clause, (b) considered to be a verbose, well-written
+license, and (c) has been recently chosen by other large open source
+projects, notably \llvm{} and \kokkos{}.
+
+As a first step we now require all new code contributions to be dual
+licensed under the old and new licenses. As a second step we are now in the
+process of contacting more than 200 contributors with copyrighted
+contributions to the library. As of September 2024 roughly 80\% of the
+commits and 80\% of the 1.88 million source lines of code have been
+relicensed.
 
@@
 
-\todo[inline]{Peter: Anything to list here?}
+P.~Munch acknowledges the funding by the Swedish Research Council (VR) under grant 2021-04620 and the strategic funding from the IT department of Uppsala University.
 

### 2024-09-23 9.6/paper.tex

@@
 \newcommand{\llvm}{{\specialword{LLVM}}\xspace}
+\newcommand{\step}[1]{{\specialword{step-#1}}\xspace}
 
@@
 terms of the \emph{GNU Lesser General Public License} (LGPL). The \dealii
-project is in the process of relicsensing the library under the terms of
+project is in the process of relicensing the library under the terms of
 the \emph{Apache License} 2.0 with \emph{LLVM Exception}. Downloads are
@@
     There are six new tutorial programs, on checkpointing simulations
-    (step-83), integrating time-dependent solvers with external time
-    stepping libraries (step-86, using \petsc's \ts{} library), advanced
-    point evaluation techniques (step-87), non-matching grids (step-89),
-    and trace-based methods for PDEs on embedded surfaces (step-90). See
+    (\step{83}), integrating time-dependent solvers with external time
+    stepping libraries (\step{86}, using \petsc's \ts{} library), advanced
+    point evaluation techniques (\step{87}), non-matching grids (\step{89}),
+    and trace-based methods for PDEs on embedded surfaces (\step{90}). See
     Section~\ref{subsec:steps} for more details.
@@
     and $x_{15}$ are the values of the two adjacent degrees of freedom on
-    the parent edge that contains the hanging node); of they can be of
+    the parent edge that contains the hanging node); or they can be of
     the form $x_{12}=42$ as is common when using Dirichlet boundary
@@
     freedom as constrained; then one added the dependencies one after
-    the other (e.g., by adding pairs $(\frac 12, 13)$ and $(\frac 12,
-    14)$ in the hanging node example above); then one added
+    the other (e.g., by adding pairs $(\frac 12, 14)$ and $(\frac 12,
+    15)$ in the hanging node example above); then one added
     inhomogeneities (by setting it to $42$ in the Dirichlet example
@@
 
-We made different updates of the matrix-free infrastructure in \dealii. These include:
+We made different updates of the multigrid and matrix-free infrastructure in \dealii. These include:
 \begin{itemize}
@@
 underlying MPI communication infrastructure, which is performed via \texttt{RemotePointEvaluation}.
-Tutorial step-89 has been added to present its usage
+Tutorial \step{89} has been added to present its usage
 in the context of the application to acoustic conservation equations~\cite{heinz2023high}.
@@
 
-Historically, \trilinos{} has implemented distributed linear algebra-classes for vectors and (sparse) matrices in its
+Historically, \trilinos{} has implemented distributed linear algebra classes for vectors and (sparse) matrices in its
 \epetra{} package that uses \MPI{} as its only source of
@@
 sub-packages replacing existing \trilinos{} sub-packages 
-(for example, Ifpack2 instead of Ifpack).
+(for example, \texttt{Ifpack2} instead of \texttt{Ifpack}).
 
@@
     However, if one wants to create a vector that has read or write access to non-locally owned indices,  the constructor of the
-    \texttt{TpetraWrappers::Vector} requires \texttt{IndexSet locally\_owned}, the \texttt{IndexSet locally\_relevant}, and
+    \texttt{TpetraWrappers::Vector} requires the \texttt{IndexSet locally\_owned}, the \texttt{IndexSet locally\_relevant}, and
     a boolean flag, whether to initialize the vector in the read-only or the write-only state.
-  \item A vector created without providing a \texttt{IndexSet locally\_relevant} is purely local and cannot access non-local indices.
-    Such a vector can not be copied to a vector that can access non-local indices, as the \texttt{IndexSet locally\_relevant} must
+  \item A vector created without providing the \texttt{IndexSet locally\_relevant} is purely local and cannot access non-local indices.
+    Such a vector cannot be copied to a vector that can access non-local indices, as the \texttt{IndexSet locally\_relevant} must
     be provided at creating the vector object. 
@@
   \item
-    \texttt{step-83}
+    \step{83}
     demonstrates how one can implement
@@
     using the \boost{} serialization functionality as a
-    foundation. step-83 was written by Pasquale Africa, Wolfgang
-    Bangerth, and Bruno Blais using step-19 as its basis.
+    foundation. \step{83} was written by Pasquale Africa, Wolfgang
+    Bangerth, and Bruno Blais using \step{19} as its basis.
     \todo{Also add affiliations like for the other tutorials?}
   \item
-    \texttt{step-86}
+    \step{86}
     is a program that solves the heat equation using \petsc's \ts{} (time
@@
   \item
-    \texttt{step-87} was contributed by Magdalena Schreter-Fleischhacker
+    \step{87}~\cite{schreter2023step87} was contributed by Magdalena Schreter-Fleischhacker
     (Technical University of Munich) and Peter Munch
@@
     arbitrary points on meshes that can be distributed among processes.
+The presented mini-examples are motivated by the application to two-phase-flow simulations and demonstrate, for example, the evaluation of solution quantities at a surface mesh embedded in a background mesh, as is the case in front-tracking.
   \item
-    \texttt{step-89} was contributed by Johannes Heinz (TU Wien),
+    \step{89} was contributed by Johannes Heinz (TU Wien),
     Maximilian Bergbauer (Technical University of Munich),
@@
   \item
-    \texttt{step-90} was contributed by Vladimir Yushutin and Timo Heister (Clemson University).
+    \step{90} was contributed by Vladimir Yushutin and Timo Heister (Clemson University).
     It implements the trace finite element method (TraceFEM). TraceFEM solves PDEs
@@
   \item The \texttt{CUDAWrappers} namespace and its contents --
-    notably things that enable the usage of cuSPARSE algorithms --
+    notably things that enable the usage of \texttt{cuSPARSE} algorithms --
     have been deprecated and will be removed in the next
@@
 
-Sebastian Kinnewig is supported by the Deutsche Forschungsgemeinschaft (DFG) under Germany’s Excellence
+S.~Kinnewig was supported by the Deutsche Forschungsgemeinschaft (DFG) under Germany’s Excellence
 Strategy within the Cluster of Excellence PhoenixD (EXC 2122, Project ID 390833453)

### 2024-09-23 9.6/paper.tex
Edit some more.

@@
   \item
-    Different additions to the non-matching infrastructure (see
+    Additions to the non-matching infrastructure (see
     Section~\ref{sec:nonmatching}). In particular, the new class
@@
     constraints on degrees of freedom in \dealii{}. Such constraints
-    can be of the form $x_3 = \frac 12 x_{14} + \frac 12 x_{15}$ as is
+    can be of the (homogeneous) form $x_3 = \frac 12 x_{14} + \frac 12 x_{15}$ as is
     common when using hanging node constraints (here, $x_3$ could
@@
     the form $x_{12}=42$ as is common when using Dirichlet boundary
+    the (inhomogeneous) form $x_{12}=42$ as is common when using Dirichlet boundary
     conditions.
@@
     and 3d faces is that, instead of two possible relative orientations,
-    triangular faces have six and quadrilateral faces have eight.
+    triangular faces have six and quadrilateral faces have eight
+    possible orientation.
 
@@
     (\emph{orientation}, \emph{rotation}, and \emph{flip}), bitsets, or several
-    other encodings, we use a single \texttt{unsigned char} which we named the
+    other encodings, we now use a single \texttt{unsigned char} which we named the
     \emph{combined orientation}. Since the representation in some library
@@
 %
-The changelog lists more than 180 features and bugfixes.
+The
+\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_5_2_and_9_6_0.html}{changelog}
+-- listing more than 180 features and bugfixes --
+contains a complete record of all changes; see \cite{changes96}.
 
@@
 which we will discuss in this section.
-It of course also includes a
-vast number of smaller changes and added functionality; the details of these
-can be found
-\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_5_2_and_9_6_0.html}
-{in the file that lists all changes for this release}; see \cite{changes96}.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Updates of multigrid and matrix-free algorithms}\label{sec:mf}
+\subsection{Updates to the multigrid and matrix-free algorithms}\label{sec:mf}
 
-We made different updates of the multigrid and matrix-free infrastructure in \dealii. These include:
+We updated a number of significant parts of the multigrid and matrix-free
+infrastructure in \dealii. These changes include:
 \begin{itemize}
-\item Our own implementation of \texttt{std::simd}, called
+\item Our own implementation of the \texttt{std::experimental::simd}
+  class proposed for inclusion into the C++ standard, called
   \texttt{VectorizedArray}, now also supports Arm Neon. Arm Neon is an
@@
   described in~\cite{KronbichlerKormann2019}, the matrix-free operator
-  evaluation is now three to five times faster in these cases.
+  evaluation is now 3--5 times faster in these cases.
 
@@
   \texttt{MGTransferGlobalCoarsening}) and \texttt{MGTwoLevel\-Transfer}. They
-  now allow to perform local smoothing, which is a key step towards unifying
+  now allow performing local smoothing, which is a key step towards unifying
   all transfer operators in \dealii. Finally, \texttt{MGTwoLevelTransfer} can now
   be set up with an existing \texttt{MatrixFree} object in the case
-  of $p$-multigrid, allowing to reduce the setup costs and memory consumption significantly.
+  of $p$-multigrid, reducing the setup costs and memory consumption significantly.
 \end{itemize}
@@
 
-In the non-matching infrastructure of \dealii, the following improvements were made:
+In the non-matching infrastructure of \dealii, we made the following improvements:
 
@@
   same set of points, which is enabled by \texttt{NonMatching::MappingInfo}.
-\item The new class \texttt{FERemoteEvaluation} has been added.
+\item We added the new class \texttt{FERemoteEvaluation}.
 This is a class to access data
@@
 \epetra{} package that uses \MPI{} as its only source of
-parallelism. Building on that, there are multiple interconnected \trilinos{} packages, e.g.\ for (non)-linear solvers 
+parallelism. Building on that, there are multiple interconnected \trilinos{} packages, e.g.\ for (non-)linear solvers 
 and preconditioners, commonly referred to as the 
-\epetra{} stack. For sake of brevity we will also refer to the whole stack as \epetra{} in the following.
-As \epetra{} was the first stack, \dealii{} interfaces to \trilinos{} have traditionally 
+``\epetra{} stack''. For sake of brevity we will also refer to the whole stack as \epetra{} in the following.
+\dealii{}'s interfaces to \trilinos{} have traditionally 
 been implemented to use it.
@@
 parallelism and GPU capabilities, referred to as \MPIx, by building on \kokkos{}. 
-Since a few years, new features have only been implemented in \tpetra{} as 
+In recent years, new features have only been implemented in \tpetra{} as 
 it is slated to replace \epetra{}, and the latter is indeed
@@
 the same is true for the need to switch to \tpetra{}-based
-sub-packages replacing existing \trilinos{} sub-packages 
+sub-packages (the ``Tpetra stack'') that are intended to replace existing \trilinos{} sub-packages 
 (for example, \texttt{Ifpack2} instead of \texttt{Ifpack}).
@@
     \texttt{TpetraWrappers::PreconditionL1GaussSeidel}, based on~\cite{BFKY2011}, have been added.
-  \item \texttt{TpetraWrappers::SolverDirectKLU2} has been wrapped as a direct solver.
+  \item \texttt{TpetraWrappers::SolverDirectKLU2} is now available as a direct solver.
 \end{itemize}
-Some functionality in the \texttt{TrilinosWrappers} classes is still missing, 
+Some functionality in these wrapper classes is still missing, 
 most noticeably wrappers for the algebraic-multigrid preconditioner \texttt{MueLu}
@@
 The solver and preconditioner classes mentioned above provide reasonable parameter subsets 
-through \texttt{AdditionalData} objects, as in \texttt{TrilinosWrappers}. 
+through \texttt{AdditionalData} objects, just as the existing
+\texttt{Epetra} wrappers. 
 Additionally, there are new generic classes \texttt{TpetraWrappers::SolverDirect} 
@@
     be provided at creating the vector object. 
-  \item Some of solver and preconditioner parameters are not available in \tpetra{},
-    such that the \texttt{AdditionalData} objects are not identical. 
+  \item Some of parameters of existing solver and preconditioner are
+    not available in the \tpetra{} wrappers,
+    such that the corresponding \texttt{AdditionalData} objects are not identical. 
     Since there are many preconditioners we will not list each individual change but instead refer 
@@
 
-One other important change is that
-all the \texttt{TpetraWrappers} classes take \texttt{Number} and \texttt{MemorySpace} template arguments,
-which control the underlying number type and memory space.
-By default, the memory is allocated on the CPU, mirroring the behavior of the \texttt{TrilinosWrappers}.
-The support for different number types also enables the usage of automatic differentiation-based tools in
+Finally,
+all \texttt{TpetraWrappers} classes take \texttt{Number} and \texttt{MemorySpace} template arguments,
+in order to control the underlying scalar type and memory space.
+By default, memory is allocated on the CPU, mirroring the behavior of
+the \texttt{TrilinosWrappers}, but choosing a different memory space
+also allows \tpetra{} to utilize GPUs if \kokkos{} is configured
+with a GPU backend.
+Furthermore, to fully leverage the GPU backend and unlock its full potential for users will require further work.
+The support for different number types in vector and matrix classes also enables the usage of automatic differentiation-based tools in
 \trilinos{}, such as \texttt{NOX}.
-Additionally, the memory space type determines whether \tpetra{} can utilize GPUs if \kokkos{} is configured
-with a GPU backend.
 To make the automatic differentiation tools of \trilinos{} accessible through the \tpetra{} interface without
@@
 \trilinos{} packages will be added in future releases.
-Furthermore, to fully leverage the GPU backend and unlock its full potential for users will require further work.
 
@@
   because the holder object -- \texttt{std::future} -- can only be
-  asked once for its computed value.)
+  asked \textit{once} for its computed value.)
 
@@
 Many of the \dealii tutorial programs were revised in a variety of ways
-as part of this release. In addition, there are a number of new tutorial
+as part of this release: Around 190 of the more than 1900 (non-merge)
+commits that went into this release touched the tutorial.
+% data generated using these commands:
+% - tutorial: git log --since 2023/07/07 --until 2024/08/11 --no-merges examples | grep commit | wc -l
+% - total:    git log --since 2023/07/07 --until 2024/08/11 --no-merges | grep commit | wc -l
+In addition, there are a number of new tutorial
 programs:
@@
     Bangerth, and Bruno Blais using \step{19} as its basis.
-    \todo{Also add affiliations like for the other tutorials?}
   \item
@@
     stepping) framework for the solution of ordinary differential
-    equations. Written by Wolfgang Bangerth (Colorado State
-    University), Luca Heltai (University of Pisa), and Stefano Zampini
+    equations. Written by Wolfgang Bangerth, Luca Heltai, and Stefano Zampini
     (King Abdullah University of Science and Technology), it
@@
     that have traditionally led code authors toward writing their own
-    time stepping routines (such as wanting to change the mesh every
+    time stepping routines (like wanting to change the mesh every
     once in a while, or having to deal with boundary conditions).
@@
     \step{87}~\cite{schreter2023step87} was contributed by Magdalena Schreter-Fleischhacker
-    (Technical University of Munich) and Peter Munch
-    (University of Augsburg/Uppsala University). It
+    and Peter Munch. It
     presents the advanced point-evaluation functionalities of \dealii,
     which are useful for evaluating finite element solutions at
-    arbitrary points on meshes that can be distributed among processes.
-The presented mini-examples are motivated by the application to two-phase-flow simulations and demonstrate, for example, the evaluation of solution quantities at a surface mesh embedded in a background mesh, as is the case in front-tracking.
+    arbitrary points on meshes that can be distributed among
+    processes. The presented mini-examples are motivated by the
+    application to two-phase flow simulations and demonstrate, for
+    example, the evaluation of solution quantities at a surface mesh
+    embedded in a background mesh, as needed in the case in front-tracking.
   \item
@@
     Maximilian Bergbauer (Technical University of Munich),
-    Marco Feder (SISSA), and Peter Munch (University of Augsburg/Uppsala University).
-    It shows a way how to apply non-matching and/or Chimera methods
+    Marco Feder (SISSA), and Peter Munch.
+    It shows how to apply non-matching and/or Chimera methods
     within matrix-free loops in \dealii.
   \item
-    \step{90} was contributed by Vladimir Yushutin and Timo Heister (Clemson University).
+    \step{90} was contributed by Vladimir Yushutin (Clemson University) and Timo Heister.
     It implements the trace finite element method (TraceFEM). TraceFEM solves PDEs
@@
 research rather than as teaching tools):
-\todo{UPDATE}
+\todo[inline]{Marc Fehling: UPDATE ([W] Marc, you added this comment
+  -- what did you want to update here? Perhaps add the affiliations of
+the authors?)}
 \begin{itemize}
-  \item \textit{``Crystal growth phase field model'''},
+  \item \textit{``Crystal growth phase field model''},
     contributed by Umair Hussain;
@@
 Secondly, our previous choice to use an open source license with strong
-copyleft was motivated by our desire to receive back code contributions
-from third parties, but it is our observation over the last 10 years that
-it hardly helps achieving this goal: Copyleft only applies when a derived
+copyleft was motivated by our hope that we receive back code contributions
+from third parties. In practice, however, our observation over the
+last 10 years is that
+our choice of license has not helped achieve this goal: Copyleft only applies when a derived
 software is sold or published. But almost all of our code contributions are
@@
 more permissible license. We settled on the Apache License 2.0 with
-LLVM-exception in large parts because it is (a) a permissible open source
-license with patent clause, (b) considered to be a verbose, well-written
-license, and (c) has been recently chosen by other large open source
-projects, notably \llvm{} and \kokkos{}.
+LLVM-exception in large parts because it is (a)~a permissible open source
+license with patent clause, (b)~considered to be a verbose, well-written
+license, and (c)~has been recently chosen by other large open source
+projects, notably \llvm{} and \kokkos{}. Our hope going forward is
+that the new license makes it easier for industry partners to use the library.
 
 As a first step we now require all new code contributions to be dual
-licensed under the old and new licenses. As a second step we are now in the
-process of contacting more than 200 contributors with copyrighted
+licensed under the old and new licenses. As a second step we are in the
+process of contacting the more than 200 past contributors with copyrighted
 contributions to the library. As of September 2024 roughly 80\% of the
@@
 applications. That said, the following are worth mentioning since they
-may have been more widely used:
+are more broadly visible:
 \begin{itemize}

### 2024-09-24 9.6/paper.tex
Augment a note.
Two minor additions

@@
       you want, we can also list this point in the list above, and add
-      a whole section 2.x for it.}
+      a whole section 2.x for it. If you think there is nothing more
+      that needs to be said, just delete this note.}
   \item The \texttt{AffineConstraints} class stores and processes
@@
   (around two times higher throughput for operator evaluation) and
-  multi-component systems.
+  multi-component systems. The restructuring made to enable these optimizations
+  also reduces the compile times and the size of the generated code slightly.
 
@@
   sets of unstructured points, \texttt{FEPointEvaluation}, have been
-  made. This is particularly useful for evaluating several quantities on the
+  made. The changes are particularly useful for evaluating several quantities on the
   same set of points, which is enabled by \texttt{NonMatching::MappingInfo}.
@@
     a boolean flag, whether to initialize the vector in the read-only or the write-only state.
+    This interface is in fact close to the one of \dealii{}'s own parallel vector class.
   \item A vector created without providing the \texttt{IndexSet locally\_relevant} is purely local and cannot access non-local indices.

### 2024-09-24 9.6/paper.tex
Typo

@@
     triangular faces have six and quadrilateral faces have eight
-    possible orientation.
+    possible orientations.
 

### 2024-09-24 9.6/paper.tex
Remove reference to deal.II-FROSch interface.
FECouplingValues.

@@
   regularity $r=(p-1)/2$.
+
 \end{itemize}
@@
 in the context of the application to acoustic conservation equations~\cite{heinz2023high}.
-\item FECouplingValues
 
-\todo[inline]{Luca: Please complete}
+\item The \texttt{FECouplingValues} class provides a powerful tool for
+ integrating finite element data between two different objects, even if they
+ live on separate grids or have different topological dimensions (e.g., cells,
+ faces, edges). This is particularly useful in the following scenarios: 
+ \begin{itemize}
+  \item Non-local Differential Operators: Evaluating fractional laplacian or
+  boundary element methods.  
+  \item Non-matching Discretizations: Integrating data between two independent
+  discretization schemes that overlap on some areas. 
+  \item Bulk-surface Coupling: Combining data from a bulk discretization and
+  surface discretization.
+ \end{itemize}
+
+This class enables the combination of degrees of freedom indices, shape
+functions, and quadrature points from two existing \texttt{FEValuesBase}
+objects. The way this combination is performed is controlled by user-provided
+\texttt{DoFCouplingType} and \texttt{QuadratureCouplingType} objects, which
+define how the degrees of freedom and quadrature points are combined.
+
+The flexibility offered by these coupling types allows users to:
+
+\begin{itemize}
+ \item Rearrange quadrature points in a tensor product structure for double
+integration.
+\item Reorder quadrature points to easily access shape function values and
+gradients from two different finite element spaces located on the same point.
+\item Identify and integrate over a reordered subset of the quadrature points.
+\end{itemize}
 \end{itemize}

### 2024-09-24 9.6/paper.tex
Briefly describe the hanging node support for Nedelec_SZ.

--- 2024-09-24 9.6/paper.tex
@@
     node constraints for locally refined meshes. 
-    \todo[inline]{Sebastian: Is there anything more to say here, maybe some reference? If
-      you want, we can also list this point in the list above, and add
-      a whole section 2.x for it. If you think there is nothing more
-      that needs to be said, just delete this note.}
+    Therefore, the sign-conflict that arises for hp quadrilateral and 
+    hexahedral N\'ed\'elec elements in the presence of hanging edges 
+    and hanging faces was addressed for the implementation details; 
+    we refer to \cite{Kinnewig2024}. 
+    The \texttt{FE\_NedelecSZ} class and the hanging node constraints 
+    automatically account for the presence of hanging edges and hanging 
+    faces; therefore, there is no difference for the user when using 
+    \texttt{FE\_NedelecSZ} as finite elements compared to other finite 
+    element classes when used on triangulations with hanging nodes. 
+    Note that the special case where, in 3D, more than four cells with 
+    different refinement levels share a common edge is not covered yet.
   \item The \texttt{AffineConstraints} class stores and processes

### 2024-09-24 9.6/paper.tex

--- 2024-09-24 9.6/paper.tex
@@
 application code (as \texttt{Teuchos::RCP}) and can be accessed through member functions.
-In fact, one of the authors has already used this to apply the domain decomposition preconditioner
-\texttt{FROSch} to various problems discretized using \dealii{}.
-\todo[inline]{Kinnewig: Add citation to preprint?}
 

### 2024-09-24 9.6/paper.tex
Minor typos and styling consistency

@@
 project is in the process of relicensing the library under the terms of
-the \emph{Apache License} 2.0 with \emph{LLVM Exception}. Downloads are
+the \emph{Apache License 2.0 with LLVM Exception}. Downloads are
 available at \url{https://www.dealii.org/} and
@@
     behind our relicensing effort to make \dealii available under the terms
-    of the the Apache License 2.0 with LLVM Exception.
+    of the the \emph{Apache License 2.0 with LLVM Exception}.
 \end{itemize}
@@
     scheme of \cite{Zag06} now supports the computation of hanging 
-    node constraints for locally refined meshes. 
+    node constraints for locally refined \todo[inline]{hexahedral?} meshes. 
     Therefore, the sign-conflict that arises for hp quadrilateral and 
@@
     the parent edge that contains the hanging node); or they can be of
-    the form $x_{12}=42$ as is common when using Dirichlet boundary
     the (inhomogeneous) form $x_{12}=42$ as is common when using Dirichlet boundary
@@
   initialization of variables. Its intended use is for member
-  variables of classes that are \textit{sometimes} needed, but perhaps
+  variables of classes that are \emph{sometimes} needed, but perhaps
   not for all uses of an object. For example, all finite element classes
@@
   these matrices; the second of these approaches is costly in places
-  that \textit{do} query these matrices repeatedly because they are re-computed
+  that \emph{do} query these matrices repeatedly because they are re-computed
   every time. \texttt{Lazy<T>} provides a middle ground: It provides an
@@
   because the holder object -- \texttt{std::future} -- can only be
-  asked \textit{once} for its computed value.)
+  asked \emph{once} for its computed value.)
 
@@
 \begin{itemize}
-  \item \textit{``Crystal growth phase field model''},
+  \item \emph{``Crystal growth phase field model''},
     contributed by Umair Hussain;
-  \item \textit{``Nonlinear heat transfer problem''}, contributed by
+  \item \emph{``Nonlinear heat transfer problem''}, contributed by
     Narasimhan Swaminathan;
-  \item \textit{``Traveling-wave solutions of a qualitative model for combustion waves''}, contributed by
+  \item \emph{``Traveling-wave solutions of a qualitative model for combustion waves''}, contributed by
     Shamil Magomedov.
@@
 from third parties. In practice, however, our observation over the
-last 10 years is that
-our choice of license has not helped achieve this goal: Copyleft only applies when a derived
-software is sold or published. But almost all of our code contributions are
-voluntary and come from individual contributors with an academic
-background, where these considerations do not apply.
+last 10 years is that our choice of license has not helped achieve this goal:
+Copyleft only applies when a derived software is sold or published. But almost
+all of our code contributions are voluntary and come from individual contributors
+with an academic background, where these considerations do not apply.
 

### 2024-09-24 9.6/paper.tex
Minor edits.

@@
     scheme of \cite{Zag06} now supports the computation of hanging 
-    node constraints for locally refined \todo[inline]{hexahedral?} meshes. 
-    Therefore, the sign-conflict that arises for hp quadrilateral and 
-    hexahedral N\'ed\'elec elements in the presence of hanging edges 
-    and hanging faces was addressed for the implementation details; 
-    we refer to \cite{Kinnewig2024}. 
-    The \texttt{FE\_NedelecSZ} class and the hanging node constraints 
-    automatically account for the presence of hanging edges and hanging 
-    faces; therefore, there is no difference for the user when using 
-    \texttt{FE\_NedelecSZ} as finite elements compared to other finite 
-    element classes when used on triangulations with hanging nodes. 
-    Note that the special case where, in 3D, more than four cells with 
+    node constraints for locally refined, hexahedral meshes. 
+    That is, the implementation of hanging node constraints now
+    correctly addresses the sign-conflict that arises for $hp$ quadrilateral and 
+    hexahedral N\'ed\'elec elements; 
+    see \cite{Kinnewig2024} for details. 
+    The other implementation of N\'ed\'elec elements, in the \texttt{FE\_NedelecSZ} class, already
+    implements hanging node constraints; therefore, there is no longer
+    a difference for the user between the two classes as far as
+    constraints are concerned. 
+    However, the special case where, in 3D, more than four cells with 
     different refinement levels share a common edge is not covered yet.
@@
   (around two times higher throughput for operator evaluation) and
-  multi-component systems. The restructuring made to enable these optimizations
-  also reduces the compile times and the size of the generated code slightly.
+  multi-component systems. The restructuring that led to these optimizations
+  also reduced the compile times and the size of the generated code slightly.
 
@@
 \subsection{Advances in non-matching support}\label{sec:nonmatching}
-
-%\todo[inline]{All: If you contributed to this functionality, please
-%  help flesh out this section.}
 
@@
 \item The \texttt{FECouplingValues} class provides a powerful tool for
- integrating finite element data between two different objects, even if they
+ computing
+ integrals of functions that contain finite element fields defined on
+ different objects, for example if these fields
  live on separate grids or have different topological dimensions (e.g., cells,
@@
  \begin{itemize}
-  \item Non-local Differential Operators: Evaluating fractional laplacian or
+  \item Non-local differential operators: Evaluating fractional Laplacian or
   boundary element methods.  
-  \item Non-matching Discretizations: Integrating data between two independent
-  discretization schemes that overlap on some areas. 
-  \item Bulk-surface Coupling: Combining data from a bulk discretization and
-  surface discretization.
- \end{itemize}
-
-This class enables the combination of degrees of freedom indices, shape
+  \item Non-matching discretizations: Integrating data between two independent
+  discretization schemes that overlap in some areas. 
+  \item Bulk-surface coupling: Combining data from a bulk discretization and
+  a surface discretization.
+ \end{itemize}
+
+The new class enables the combination of degrees of freedom indices, shape
 functions, and quadrature points from two existing \texttt{FEValuesBase}

### 2024-09-25 9.6/paper.tex

@@
     The other implementation of N\'ed\'elec elements, in the \texttt{FE\_NedelecSZ} class, already
+    \todo[inline]{JPT: I think this should be \texttt{FE\_Nedelec}, without SZ?}
     implements hanging node constraints; therefore, there is no longer
@@
 However, the wrapped \texttt{Ifpack2} preconditioners can already be used with the iterative solvers of \dealii{}.
+To do this you will need to explicitly specify the vector type parameter,
+e.g.\\ \texttt{SolverCG<LinearAlgebra::TpetraWrappers::Vector<double, MemorySpace::Default> >}
 
@@
     be provided at creating the vector object. 
-  \item Some of parameters of existing solver and preconditioner are
+  \item Some parameters of existing solver and preconditioner are
     not available in the \tpetra{} wrappers,
@@
 Yimin        Jin,
-Sebastian    Kinnewig,
+%Sebastian    Kinnewig,
 Paras        Kumar,
@@
 See \cite{Boerner2023}.
+\todo[inline]{Missing ACCESS information}
 

### 2024-09-25 9.6/paper.tex
Fix up a mistake.

@@
     see \cite{Kinnewig2024} for details. 
-    The other implementation of N\'ed\'elec elements, in the \texttt{FE\_NedelecSZ} class, already
-    \todo[inline]{JPT: I think this should be \texttt{FE\_Nedelec}, without SZ?}
+    The other implementation of N\'ed\'elec elements, in the \texttt{FE\_Nedelec} class, already
     implements hanging node constraints; therefore, there is no longer

### 2024-09-26 9.6/paper.tex
Address remaining TODOs.

@@
 research rather than as teaching tools):
-\todo[inline]{Marc Fehling: UPDATE ([W] Marc, you added this comment
-  -- what did you want to update here? Perhaps add the affiliations of
-the authors?)}
 \begin{itemize}
@@
 
-\todo[inline]{Marc: please insert allocation}
-
-This work used [resource-name] at [resource provider] through allocation [allocation number] from the Advanced Cyberinfrastructure Coordination Ecosystem: Services \& Support (ACCESS) program, which is supported by National Science Foundation grants \#2138259, \#2138286, \#2138307, \#2137603, and \#2138296.
+This work used the \textit{Expanse} HPC system at the San Diego Supercomputer Center (SDSC) at UC San Diego through the CIG Science Gateway and Community Codes for the Geodynamics Community MCA08X011 allocation from the Advanced Cyberinfrastructure Coordination Ecosystem: Services \& Support (ACCESS) program, which is supported by National Science Foundation grants \#2138259, \#2138286, \#2138307, \#2137603, and \#2138296.
 See \cite{Boerner2023}.
-\todo[inline]{Missing ACCESS information}
-
-This work used the Extreme Science and Engineering Discovery Environment (XSEDE), which is supported by National Science Foundation grant number ACI-1053575 access through the CIG Science Gateway and Community Codes for the Geodynamics Community MCA08X011 allocation.
 

### 2024-09-27 9.6/paper.tex
Fix a number.

@@
   \item
-    There are six new tutorial programs, on checkpointing simulations
+    There are five new tutorial programs, on checkpointing simulations
     (\step{83}), integrating time-dependent solvers with external time

### 2025-07-03 9.7/paper.tex
Add an outline for the 9.7 release paper.

--- 2024-09-26 9.6/paper.tex
@@
   },
-  pdftitle={The deal.II Library, Version 9.6, 2024},
+  pdftitle={The deal.II Library, Version 9.7, 2025},
 }
 
-\title{The \dealii Library, Version 9.6}
+\title{The \dealii Library, Version 9.7}
 
- \author[1]{Pasquale~Claudio~Africa}
- \affil[1]{SISSA International School for Advanced Studies,
-   mathLab,
-   Via Bonomea, 265,
-   34136, Trieste, Italy.
-   \texttt{pafrica@sissa.it}}
-
- \author[2*]{Daniel Arndt}
- \affil[2]{Computational Coupled Physics Group,
+ \author[1*]{Daniel Arndt}
+ \affil[1]{Computational Coupled Physics Group,
    Computational Sciences and Engineering Division,
@@
 
- \author[3,4]{Wolfgang~Bangerth}
- \affil[3]{Department of Mathematics, Colorado State University, Fort
+ \author[2,3]{Wolfgang~Bangerth}
+ \affil[2]{Department of Mathematics, Colorado State University, Fort
    Collins, CO 80523-1874, USA.
    \texttt{bangerth@colostate.edu}}
- \affil[4]{Department of Geosciences, Colorado State University, Fort
+ \affil[3]{Department of Geosciences, Colorado State University, Fort
    Collins, CO 80523, USA.}
 
-\author[5]{Bruno Blais}
-\affil[5]{Chemical Engineering High-performance Analysis, Optimization and Simulation (CHAOS) laboratory, Department of Chemical Engineering,
+\author[4]{Bruno Blais}
+\affil[4]{Chemical Engineering High-performance Analysis, Optimization and Simulation (CHAOS) laboratory, Department of Chemical Engineering,
              Polytechnique Montréal,
@@
 
-\author[6]{Marc~Fehling}
- \affil[6]{Department of Mathematical Analysis,
+\author[5]{Marc~Fehling}
+ \affil[5]{Department of Mathematical Analysis,
     Faculty of Mathematics and Physics, Charles University,
@@
 
-\author[7]{Rene~Gassm\"{o}ller}
-\affil[7]{GEOMAR Helmholtz Centre for Ocean Research Kiel, 24148 Kiel, Germany}
+\author[6]{Rene~Gassm\"{o}ller}
+\affil[6]{GEOMAR Helmholtz Centre for Ocean Research Kiel, 24148 Kiel, Germany}
 
-\author[8]{Timo~Heister}
- \affil[8]{School of Mathematical and Statistical Sciences,
+\author[7]{Timo~Heister}
+ \affil[7]{School of Mathematical and Statistical Sciences,
    Clemson University,
@@
 
-\author[9]{Luca~Heltai}
-\affil[9]{University of Pisa, Italy.}
+\author[8]{Luca~Heltai}
+\affil[8]{University of Pisa, Italy.}
 
-\author[10]{Sebastian~Kinnewig}
-\affil[10]{Institute for Applied Mathematics, Scientific Computing,
-           Leibniz University Hannover,
-           Welfengarten 1, 30167 Hannover, Germany
-           {\texttt{kinnewig@ifam.uni-hannover.de}}}
-
- \author[11,12]{Martin~Kronbichler}
- \affil[11]{Faculty of Mathematics, Ruhr University Bochum,
+ \author[9,10]{Martin~Kronbichler}
+ \affil[9]{Faculty of Mathematics, Ruhr University Bochum,
    Universit\"atsstr.~150, 44780 Bochum, Germany.
  {\texttt{martin.kronbichler@rub.de}}}
- \affil[12]{Institute of Mathematics,
+ \affil[10]{Institute of Mathematics,
    University of Augsburg,
@@
 
-\author[13]{Matthias~Maier}
-\affil[13]{Department of Mathematics,
+\author[11]{Matthias~Maier}
+\affil[11]{Department of Mathematics,
   Texas A\&M University,
@@
 
-\author[12,14]{Peter Munch}
- \affil[14]{Uppsala University, Sweden.
+\author[10,12]{Peter Munch}
+ \affil[12]{Uppsala University, Sweden.
   {\texttt{peter.munch@it.uu.se}}}
 
-\author[15]{Magdalena Schreter-Fleischhacker}
- \affil[15]{Institute for Computational Mechanics, Technical University of Munich, Boltzmannstraße 15, 85748 Garching, Germany.
-  {\texttt{magdalena.schreter@tum.de}}}
+\author[1*]{Bruno~Turcksin}
 
-\author[16]{Jan Philipp Thiele}
- \affil[16]{Weierstrass Institute for Applied Analysis and Stochastics,\newline
- Leibniz Institute in Forschungsverbund Berlin e.V.
-  {\texttt{thiele@wias-berlin.de}}}
-
-\author[2*]{Bruno~Turcksin}
-
-\author[17]{David Wells}
-\affil[17]{Department of Mathematics, University of North Carolina,
+\author[13]{David Wells}
+\affil[13]{Department of Mathematics, University of North Carolina,
   Chapel Hill, NC 27516, USA.
   {\texttt{drwells@email.unc.edu}}}
-
-\author[8, 18]{Vladimir Yushutin}
-\affil[18]{Department of Mathematics, University of Tennessee at Knoxville,
-	 Knoxville TN 37996-1320, USA.
-{\texttt{vyushuti@utk.edu}}}
 
@@
   This paper provides an overview of the new features of the finite element
-  library \dealii, version 9.6.
+  library \dealii, version 9.8.
 \end{abstract}
@@
 
-\dealii version 9.6.0 was released August 11, 2024. This paper provides an
+\dealii version 9.7.0 was released August 11, 2024.
+\todo{Update}
+This paper provides an
 overview of the new features of this release and serves as a citable
-reference for the \dealii software library version 9.6. \dealii is an
+reference for the \dealii software library version 9.7. \dealii is an
 object-oriented finite element library used around the world in the
@@
 \item
-    Substantial performance improvements to the matrix-free and multigrid
-    infrastructure (Section~\ref{sec:mf}).
-  \item
-    Additions to the non-matching infrastructure (see
-    Section~\ref{sec:nonmatching}). In particular, the new class
-    \texttt{FERemoteEvaluation} provides support for evaluating finite
-    element shape functions and solutions on parts of the domain stored by
-    other MPI processes.
-  \item
-    Much work has gone into writing wrappers for \trilinos' \tpetra{} stack
-    of linear algebra classes. \tpetra{} is \trilinos' \kokkos-based
-    replacement for the now-deprecated \epetra{} stack. See
-    Section~\ref{sec:tpetra} for more on this.
-  \item
-    Tool classes \texttt{TaskResult} and \texttt{Lazy} that provide ways to
-    compute values on a separate thread or only when first accessed
-    (Section~\ref{sec:tools}).
-  \item
-    There are six new tutorial programs, on checkpointing simulations
-    (\step{83}), integrating time-dependent solvers with external time
-    stepping libraries (\step{86}, using \petsc's \ts{} library), advanced
-    point evaluation techniques (\step{87}), non-matching grids (\step{89}),
-    and trace-based methods for PDEs on embedded surfaces (\step{90}). See
-    Section~\ref{subsec:steps} for more details.
-  \item
-    In Section~\ref{sec:license} we summarize the motivation and approach
-    behind our relicensing effort to make \dealii available under the terms
-    of the the \emph{Apache License 2.0 with LLVM Exception}.
+  \todo[inline]{Update list}
 \end{itemize}
@@
 \begin{itemize}
-  \item \dealii{} now requires and makes use of the C++17 language standard.
-  \item We have continued to make progress in supporting simplex and
-    mixed meshes -- mesh types that \dealii{} has traditionally not
-    supported at all. Specifically, the current release uses
-    better strategies for refinement of tetrahedra that result in
-    better-shaped child cells. It also contains
-    support for cubic finite elements on simplices.
-   \item The \texttt{FE\_NedelecSZ} class that contains our 
-    implementation of the N\'ed\'elec element using the orientation 
-    scheme of \cite{Zag06} now supports the computation of hanging 
-    node constraints for locally refined, hexahedral meshes. 
-    That is, the implementation of hanging node constraints now
-    correctly addresses the sign-conflict that arises for $hp$ quadrilateral and 
-    hexahedral N\'ed\'elec elements; 
-    see \cite{Kinnewig2024} for details. 
-    The other implementation of N\'ed\'elec elements, in the \texttt{FE\_Nedelec} class, already
-    implements hanging node constraints; therefore, there is no longer
-    a difference for the user between the two classes as far as
-    constraints are concerned. 
-    However, the special case where, in 3D, more than four cells with 
-    different refinement levels share a common edge is not covered yet.
-  \item The \texttt{AffineConstraints} class stores and processes
-    constraints on degrees of freedom in \dealii{}. Such constraints
-    can be of the (homogeneous) form $x_3 = \frac 12 x_{14} + \frac 12 x_{15}$ as is
-    common when using hanging node constraints (here, $x_3$ could
-    represent the value of the solution at a hanging node, and $x_{14}$
-    and $x_{15}$ are the values of the two adjacent degrees of freedom on
-    the parent edge that contains the hanging node); or they can be of
-    the (inhomogeneous) form $x_{12}=42$ as is common when using Dirichlet boundary
-    conditions.
-
-    Since very early in the history of the library, the
-    \texttt{AffineConstraints} class interface required building such
-    constraints in multiple steps: First, one declared a degree of
-    freedom as constrained; then one added the dependencies one after
-    the other (e.g., by adding pairs $(\frac 12, 14)$ and $(\frac 12,
-    15)$ in the hanging node example above); then one added
-    inhomogeneities (by setting it to $42$ in the Dirichlet example
-    above). This piecemeal approach is cumbersome and prevents the
-    library from performing certain error checking steps because a
-    constraint is not known to be completely built at any given point. The new
-    \texttt{add\_constraint()} function now allows defining a
-    constraint in one step.
-  \item We refactored the systems for managing relative line and face
-    orientations to significantly improve consistency across several library
-    modules, including finite element classes, periodicity, and the
-    \pfrst{} interface. \dealii{} has supported using unstructured
-    three-dimensional meshes for a long time. Features like discontinuous
-    Galerkin methods and higher-order elements require that lines and faces of
-    adjacent elements are consistently oriented. For example, consider a line
-    containing two degrees of freedom, which, in reference coordinates, are
-    located at $x_1 = \frac 1 3$ and $x_2 = \frac 2 3$. If the line is read from
-    left-to-right then we get points $(x_1, x_2)$ whereas if it is read from
-    right-to-left we get $(x_2, x_1)$. \dealii{} guarantees that this order will
-    be consistent across all cells sharing that line by storing a flag
-    indicating whether (relative to the vertex enumeration on the present cell)
-    that particular line is in the standard or reversed orientation. Faces in 3d
-    are treated in a similar way. The only significant difference between lines
-    and 3d faces is that, instead of two possible relative orientations,
-    triangular faces have six and quadrilateral faces have eight
-    possible orientations.
-
-    Historically, different places within the library implemented their own
-    ad-hoc orientation logic, data encodings, and default values. We cleaned up
-    this subsystem and instead of using three different booleans
-    (\emph{orientation}, \emph{rotation}, and \emph{flip}), bitsets, or several
-    other encodings, we now use a single \texttt{unsigned char} which we named the
-    \emph{combined orientation}. Since the representation in some library
-    interfaces has changed, most of the old interfaces have been deprecated and
-    a few (such as the low-level interface to periodic boundary conditions) had
-    to be incompatibly altered or removed.
-\item The \texttt{SolverGMRES} class now offers a third orthogonalization
-  method, the classical Gram--Schmidt method with delayed orthogonalization
-  \cite{Bielich2022}. Furthermore, the solver now specifies the maximal basis
-  size of the Arnoldi basis, rather than the number of auxiliary
-  vectors. Some changes have also been made to the GMRES and F-GMRES
-  implementations, making them use the same underlying kernels as much as possible.
-\item The new class \texttt{FE\_Hermite} implements a Hermite interpolation basis of
-  maximum regularity. These bases are always of odd polynomial degree $p$ and have
-  regularity $r=(p-1)/2$.
-
+\item
+\todo[inline]{Update list}
 \end{itemize}
@@
 \href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_5_2_and_9_6_0.html}{changelog}
--- listing more than 180 features and bugfixes --
-contains a complete record of all changes; see \cite{changes96}.
+\todo{Update}
+-- listing more than XXX
+\todo{Update}
+features and bugfixes --
+contains a complete record of all changes; see \cite{changes97}.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Updates to the multigrid and matrix-free algorithms}\label{sec:mf}
-
-We updated a number of significant parts of the multigrid and matrix-free
-infrastructure in \dealii. These changes include:
-\begin{itemize}
-\item Our own implementation of the \texttt{std::experimental::simd}
-  class proposed for inclusion into the C++ standard, called
-  \texttt{VectorizedArray}, now also supports Arm Neon. Arm Neon is an
-  architecture extension of the ARMv7, ARMv8 and ARMv9 architecture families,
-  used, e.g., for the Arm Cortex-A and Arm Cortex-R series of processors or in
-  the Apple 64-bit silicon series (A7--A18, M1--M4 as of 2024).
-  With these instructions, 2 doubles or 4 floats can be processed in one
-  go. Since the matrix-free infrastructure works directly with the \texttt{VectorizedArray}
-  abstraction as data structure, it automatically benefits from this new implementation.
-
-\item The application of the Piola transformation for values and gradients of
-  $H$(div)-conforming Raviart--Thomas elements on non-Cartesian cells has
-  been rewritten for better performance. Together with additional changes in
-  the sum-factorization algorithms, which became more similar to the kernels
-  described in~\cite{KronbichlerKormann2019}, the matrix-free operator
-  evaluation is now 3--5 times faster in these cases.
-
-\item Furthermore, we improved the internal data structures of the tensor-product
-  evaluators as well as the evaluators for simplex elements.  This speeds up
-  the operator evaluation in several scenarios, especially for simplices
-  (around two times higher throughput for operator evaluation) and
-  multi-component systems. The restructuring that led to these optimizations
-  also reduced the compile times and the size of the generated code slightly.
-
-\item In addition, we performed substantial improvements to the global-coarsening
-  multigrid infrastructure: \texttt{MGTransferMF} (previously:
-  \texttt{MGTransferGlobalCoarsening}) and \texttt{MGTwoLevel\-Transfer}. They
-  now allow performing local smoothing, which is a key step towards unifying
-  all transfer operators in \dealii. Finally, \texttt{MGTwoLevelTransfer} can now
-  be set up with an existing \texttt{MatrixFree} object in the case
-  of $p$-multigrid, reducing the setup costs and memory consumption significantly.
-\end{itemize}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Advances in non-matching support}\label{sec:nonmatching}
-
-In the non-matching infrastructure of \dealii, we made the following improvements:
-
-\begin{itemize}
-\item The performance of the non-nested multigrid infrastructure
-(\texttt{MGTwoLevelTransferNon\-Nested}) has been improved significantly by avoiding
-redundant copy operations. Furthermore, support for simplex-shaped cells and
-multiple-component elements has been added.
-\item Several minor performance improvements in \dealii's evaluator class on
-  sets of unstructured points, \texttt{FEPointEvaluation}, have been
-  made. The changes are particularly useful for evaluating several quantities on the
-  same set of points, which is enabled by \texttt{NonMatching::MappingInfo}.
-\item We added the new class \texttt{FERemoteEvaluation}.
-This is a class to access data
-in a distributed matrix-free loop for non-matching discretizations.
-Interfaces are named in analogy to \texttt{FEEvaluation}, in order to seamlessly switch the
-local evaluator functionality at quadrature points of a computation. The key component is the
-underlying MPI communication infrastructure, which is performed via \texttt{RemotePointEvaluation}.
-Tutorial \step{89} has been added to present its usage
-in the context of the application to acoustic conservation equations~\cite{heinz2023high}.
-
-\item The \texttt{FECouplingValues} class provides a powerful tool for
- computing
- integrals of functions that contain finite element fields defined on
- different objects, for example if these fields
- live on separate grids or have different topological dimensions (e.g., cells,
- faces, edges). This is particularly useful in the following scenarios: 
- \begin{itemize}
-  \item Non-local differential operators: Evaluating fractional Laplacian or
-  boundary element methods.  
-  \item Non-matching discretizations: Integrating data between two independent
-  discretization schemes that overlap in some areas. 
-  \item Bulk-surface coupling: Combining data from a bulk discretization and
-  a surface discretization.
- \end{itemize}
-
-The new class enables the combination of degrees of freedom indices, shape
-functions, and quadrature points from two existing \texttt{FEValuesBase}
-objects. The way this combination is performed is controlled by user-provided
-\texttt{DoFCouplingType} and \texttt{QuadratureCouplingType} objects, which
-define how the degrees of freedom and quadrature points are combined.
-
-The flexibility offered by these coupling types allows users to:
-
-\begin{itemize}
- \item Rearrange quadrature points in a tensor product structure for double
-integration.
-\item Reorder quadrature points to easily access shape function values and
-gradients from two different finite element spaces located on the same point.
-\item Identify and integrate over a reordered subset of the quadrature points.
-\end{itemize}
-\end{itemize}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Interface to the \trilinos{} \tpetra{} stack}\label{sec:tpetra}
-
-\dealii{}'s parallel linear algebra facilities (apart from matrix-free capabilities) are largely built on
-wrappers around functionality provided by the \petsc{} and \trilinos{} libraries.
-
-Historically, \trilinos{} has implemented distributed linear algebra classes for vectors and (sparse) matrices in its
-\epetra{} package that uses \MPI{} as its only source of
-parallelism. Building on that, there are multiple interconnected \trilinos{} packages, e.g.\ for (non-)linear solvers 
-and preconditioners, commonly referred to as the 
-``\epetra{} stack''. For sake of brevity we will also refer to the whole stack as \epetra{} in the following.
-\dealii{}'s interfaces to \trilinos{} have traditionally 
-been implemented to use it.
-
-However, several years ago, \trilinos{} also introduced the
-newer \tpetra{} (``templated'' Petra) package that provides additional shared memory
-parallelism and GPU capabilities, referred to as \MPIx, by building on \kokkos{}. 
-In recent years, new features have only been implemented in \tpetra{} as 
-it is slated to replace \epetra{}, and the latter is indeed
-now deprecated with removal slated for 2025. 
-As a consequence, 
-we will eventually have to switch all of our \trilinos{} interfaces to \tpetra{}; 
-the same is true for the need to switch to \tpetra{}-based
-sub-packages (the ``Tpetra stack'') that are intended to replace existing \trilinos{} sub-packages 
-(for example, \texttt{Ifpack2} instead of \texttt{Ifpack}).
-
-In the current release, we have put substantial work into this switch,
-as outlined below. As this is not yet finished, it is important to note that 
-all these features are optional and their individual availability
-depends on the installed \trilinos{} packages. 
-As an example, a \dealii{} build will pass with a \trilinos{} installation 
-without \texttt{Ifpack2}, but its preconditioners will not be included.
-In the following list, we omit the common namespace prefix
-\texttt{LinearAlgebra::} on all mentioned
-\texttt{TpetraWrappers} symbols for readability:
-\begin{itemize}
-  \item The \texttt{TpetraWrappers::Vector} class has been overhauled to also allow specifying
-    a memory space, and the \texttt{TpetraWrappers::BlockVector} class has been added.
-  \item \texttt{TpetraWrappers::SparseMatrix} and \texttt{TpetraWrappers::SparsityPattern} have been implemented mirroring
-    the functionality of the \texttt{TrilinosWrappers::SparseMatrix} \\ and \texttt{TrilinosWrappers::SparsityPattern} classes.
-  \item Various incomplete factorization and relaxation preconditioners (from \texttt{Ifpack2}) have been wrapped,
-    mirroring the existing classes in \texttt{LinearAlgbera::TrilinosWrappers} as closely as possible.
-  \item Two new preconditioner variants \texttt{TpetraWrappers::PreconditionL1Jacobi} and \\
-    \texttt{TpetraWrappers::PreconditionL1GaussSeidel}, based on~\cite{BFKY2011}, have been added.
-  \item \texttt{TpetraWrappers::SolverDirectKLU2} is now available as a direct solver.
-\end{itemize}
-Some functionality in these wrapper classes is still missing, 
-most noticeably wrappers for the algebraic-multigrid preconditioner \texttt{MueLu}
-and the iterative solvers from \texttt{Belos}.
-However, the wrapped \texttt{Ifpack2} preconditioners can already be used with the iterative solvers of \dealii{}.
-To do this you will need to explicitly specify the vector type parameter,
-e.g.\\ \texttt{SolverCG<LinearAlgebra::TpetraWrappers::Vector<double, MemorySpace::Default> >}
-
-The solver and preconditioner classes mentioned above provide reasonable parameter subsets 
-through \texttt{AdditionalData} objects, just as the existing
-\texttt{Epetra} wrappers. 
-Additionally, there are new generic classes \texttt{TpetraWrappers::SolverDirect} 
-and \texttt{TpetraWrappers::PreconditionIfpack} which expose the internal interface through
-a \texttt{Teuchos::ParameterList} and thereby offering the full set of parameters for 
-more experienced \trilinos{} users. This also allows the use of preconditioners or 
-solvers not (yet) wrapped, e.g.\ \texttt{SuperLU\_dist} or \texttt{MUMPS}.
-
-The design goal was to introduce as few changes as possible for the user to allow for an easy transition to the
-\texttt{TpetraWrappers} classes.
-However, based on the difference between \epetra{} and \tpetra{}, the following changes were necessary:
-\begin{itemize}
-  \item The constructor for the class \texttt{TrilinosWrappers::MPI::Vector} only required the \texttt{IndexSet locally\_owned},
-    which describes the set of indices locally owned by the current rank.
-    However, if one wants to create a vector that has read or write access to non-locally owned indices,  the constructor of the
-    \texttt{TpetraWrappers::Vector} requires the \texttt{IndexSet locally\_owned}, the \texttt{IndexSet locally\_relevant}, and
-    a boolean flag, whether to initialize the vector in the read-only or the write-only state.
-    This interface is in fact close to the one of \dealii{}'s own parallel vector class.
-  \item A vector created without providing the \texttt{IndexSet locally\_relevant} is purely local and cannot access non-local indices.
-    Such a vector cannot be copied to a vector that can access non-local indices, as the \texttt{IndexSet locally\_relevant} must
-    be provided at creating the vector object. 
-  \item Some parameters of existing solver and preconditioner are
-    not available in the \tpetra{} wrappers,
-    such that the corresponding \texttt{AdditionalData} objects are not identical. 
-    Since there are many preconditioners we will not list each individual change but instead refer 
-    to our Doxygen documentation. 
-  \item The incomplete Cholesky (IC) factorization preconditioner is not available 
-   in \tpetra{}.
-\end{itemize}
-
-Another goal was increasing the interoperability of \dealii{} and \trilinos{},
-allowing users familiar with both to write `pure' \trilinos{} code within their 
-applications, e.g.\ to test or develop a new feature.
-Therefore, the internal data is stored as it would be in a \trilinos{} 
-application code (as \texttt{Teuchos::RCP}) and can be accessed through member functions.
-
-Finally,
-all \texttt{TpetraWrappers} classes take \texttt{Number} and \texttt{MemorySpace} template arguments,
-in order to control the underlying scalar type and memory space.
-By default, memory is allocated on the CPU, mirroring the behavior of
-the \texttt{TrilinosWrappers}, but choosing a different memory space
-also allows \tpetra{} to utilize GPUs if \kokkos{} is configured
-with a GPU backend.
-Furthermore, to fully leverage the GPU backend and unlock its full potential for users will require further work.
-The support for different number types in vector and matrix classes also enables the usage of automatic differentiation-based tools in
-\trilinos{}, such as \texttt{NOX}.
-To make the automatic differentiation tools of \trilinos{} accessible through the \tpetra{} interface without
-requiring users to have an in-depth understanding of \trilinos{}, additional wrappers for the corresponding
-\trilinos{} packages will be added in future releases.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{More support for advanced programming idioms}\label{sec:tools}
-
-Over the years, \dealii{} has accumulated many classes and functions
-that support modern programming idioms and make it easier to write
-code at higher levels of abstraction.
-
-In the current release, we have added two classes to the list of tools
-of this kind:
-\begin{itemize}
-  \item
-  \texttt{Lazy<T>} is a class that supports the lazy computation and
-  initialization of variables. Its intended use is for member
-  variables of classes that are \emph{sometimes} needed, but perhaps
-  not for all uses of an object. For example, all finite element classes
-  provide interpolation and restriction matrices to support multigrid
-  and other algorithms. One could (i) always compute and store these
-  matrices in the constructor of the class; or one could (ii) re-compute these
-  matrices every time they are requested. The first of these
-  approaches costs memory and compute time even though most places
-  where one creates a finite element object will not actually query
-  these matrices; the second of these approaches is costly in places
-  that \emph{do} query these matrices repeatedly because they are re-computed
-  every time. \texttt{Lazy<T>} provides a middle ground: It provides an
-  abstraction for an object that is initialized upon first use (that
-  is, the first time the value is requested), and then stores the
-  computed value for cheap use later on.
-
-  (C++ provides functionality via \texttt{std::async} with
-  launch policy \texttt{std::launch::deferred} that can achieve
-  similar outcomes. But this functionality is more difficult to use than
-  \texttt{Lazy<T>} because, among other reasons, the code generating the
-  object has to be specified at the place of construction of the
-  object holding the result, rather than at the place of use; and
-  because the holder object -- \texttt{std::future} -- can only be
-  asked \emph{once} for its computed value.)
-
-  \item \texttt{TaskResult<T>} is a class that represents the outcome
-    of a task possibly evaluated on a separate thread. It can be
-    thought of as a ``deferred'' result of a computation in that one
-    wants to state ``This job needs to be done, do it when convenient,
-    and then put the result of the operation into this
-    variable''. Accessing the variable then waits for the operation to
-    complete, if it has not already. \texttt{TaskResult<T>} allows
-    classes to efficiently compute member variables in the background,
-    assuming that they may not be needed right away but only later on.
-
-    (Similar to above, the same effect as \texttt{TaskResult<T>} can
-    be achieved using \texttt{std::async}, this time using the
-    launch policy \texttt{std::launch::async}. This approach suffers
-    from the same issue that one can only query the resulting object
-    once. Moreover, \texttt{std::async} does not integrate with the
-    thread pool that underlies \dealii{}'s approach to parallel
-    processing on modern multi-core machines, whereas
-    \texttt{TaskResult<T>} does.)
-\end{itemize}
+\todo[inline]{Add sections}
 
@@
 Many of the \dealii tutorial programs were revised in a variety of ways
-as part of this release: Around 190 of the more than 1900 (non-merge)
+as part of this release: Around 130 of the more than 1600 (non-merge)
 commits that went into this release touched the tutorial.
@@
   \item
-    \step{83}
-    demonstrates how one can implement
-    checkpoint/restart functionality in \dealii-based programs,
-    using the \boost{} serialization functionality as a
-    foundation. \step{83} was written by Pasquale Africa, Wolfgang
-    Bangerth, and Bruno Blais using \step{19} as its basis.
+    \step{93}
+    demonstrates how one can implement ...
+    \todo{Complete}
+    \step{93} was written by Sam Scheuerman.
   \item
-    \step{86}
-    is a program that solves the heat equation using \petsc's \ts{} (time
-    stepping) framework for the solution of ordinary differential
-    equations. Written by Wolfgang Bangerth, Luca Heltai, and Stefano Zampini
-    (King Abdullah University of Science and Technology), it
-    illustrates how PDE solvers for time-dependent problems can be
-    integrated with existing ODE solver packages to use advanced ODE
-    solver concepts (such as higher-order time integration methods and
-    adaptive time step control), all without sacrificing the things
-    that have traditionally led code authors toward writing their own
-    time stepping routines (like wanting to change the mesh every
-    once in a while, or having to deal with boundary conditions).
-  \item
-    \step{87}~\cite{schreter2023step87} was contributed by Magdalena Schreter-Fleischhacker
-    and Peter Munch. It
-    presents the advanced point-evaluation functionalities of \dealii,
-    which are useful for evaluating finite element solutions at
-    arbitrary points on meshes that can be distributed among
-    processes. The presented mini-examples are motivated by the
-    application to two-phase flow simulations and demonstrate, for
-    example, the evaluation of solution quantities at a surface mesh
-    embedded in a background mesh, as needed in the case in front-tracking.
-  \item
-    \step{89} was contributed by Johannes Heinz (TU Wien),
-    Maximilian Bergbauer (Technical University of Munich),
-    Marco Feder (SISSA), and Peter Munch.
-    It shows how to apply non-matching and/or Chimera methods
-    within matrix-free loops in \dealii.
-  \item
-    \step{90} was contributed by Vladimir Yushutin (Clemson University) and Timo Heister.
-    It implements the trace finite element method (TraceFEM). TraceFEM solves PDEs
-    posed on a, possibly evolving, $(dim-1)$-dimensional surface $\Gamma$ employing
-    a fixed uniform background mesh of a $dim$-dimensional domain in which
-    the surface is embedded. Such surface PDEs arise in problems involving
-    material films with complex properties and in other situations in which
-    a non-trivial condition is imposed on either a stationary or a moving interface.
-    The program considers a steady, complex, non-trivial surface and the prototypical
-    Laplace-Beltrami equation which is a counterpart of
-    the Poisson problem on flat domains.
+    \step{97}
+    is a program that solves ... curl-curl ...
+    \todo{Complete}
+    \step{93} was written by Siarhei Uzunbajakau.
 \end{itemize}
 
-In addition, there are three new programs in the code gallery (a collection of
+In addition, there is one new program in the code gallery (a collection of
 user-contributed programs that often solve more complicated problems
@@
 \begin{itemize}
-  \item \emph{``Crystal growth phase field model''},
-    contributed by Umair Hussain;
-  \item \emph{``Nonlinear heat transfer problem''}, contributed by
-    Narasimhan Swaminathan;
-  \item \emph{``Traveling-wave solutions of a qualitative model for combustion waves''}, contributed by
-    Shamil Magomedov.
+  \item \emph{``Phase field facture model in 3D''},
+    contributed by Wasim Niyaz Munshi,
+    Chandrasekhar Annavarapu,
+    Wolfgang Bangerth, and
+    Marc Fehling.
 \end{itemize}
-
-Furthermore, we added an example to the \texttt{libCEED}
-library~\cite{brown2021libceed}:
-\url{https://github.com/CEED/libCEED/tree/main/examples/deal.II}.
-\texttt{libCEED} is a library that provides matrix-free evaluation routines
-for different hardware. The example shows how to interface the \dealii{}
-data structures with the \texttt{libCEED} ones and solves the BP1-BP6
-benchmarks (scalar/vector Laplace/mass matrix with regular integration and
-over-integration).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Relicensing to Apache License 2.0 with LLVM exception}
-\label{sec:license}
-
-The \dealii project is in the process of relicensing the library under the
-terms of the \emph{Apache License 2.0 with LLVM
-Exception} \cite{Apache-2.0,LLVM-Exception}. To this end we require all new
-code contributions to be dual licensed under the current license (GNU
-Lesser General Public License v2.1 \cite{LGPL-2.1}) and the new license
-(Apache-2.0 with LLVM-exception).
-
-We have chosen to pursue a relicensing effort for deal.II because of some
-long standing problems with the LGPL v2.1 license. Most notably, despite
-our clear intent that the strong copyleft principle should only apply to
-deal.II source code and not user projects (which should be free to choose
-their own license freely) this is not necessarily the case for LGPL v2.1
-due to the language used in the library. This makes potential industry
-partners that evaluate the use of deal.II for their own projects nervous.
-Secondly, our previous choice to use an open source license with strong
-copyleft was motivated by our hope that we receive back code contributions
-from third parties. In practice, however, our observation over the
-last 10 years is that our choice of license has not helped achieve this goal:
-Copyleft only applies when a derived software is sold or published. But almost
-all of our code contributions are voluntary and come from individual contributors
-with an academic background, where these considerations do not apply.
-
-We have thus decided to switch away from the LGPL v2.1 license towards a
-more permissible license. We settled on the Apache License 2.0 with
-LLVM-exception in large parts because it is (a)~a permissible open source
-license with patent clause, (b)~considered to be a verbose, well-written
-license, and (c)~has been recently chosen by other large open source
-projects, notably \llvm{} and \kokkos{}. Our hope going forward is
-that the new license makes it easier for industry partners to use the library.
-
-As a first step we now require all new code contributions to be dual
-licensed under the old and new licenses. As a second step we are in the
-process of contacting the more than 200 past contributors with copyrighted
-contributions to the library. As of September 2024 roughly 80\% of the
-commits and 80\% of the 1.88 million source lines of code have been
-relicensed.
 
@@
 
-The 9.6 release includes
+The 9.7 release includes
 \href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_5_2_and_9_6_0.html}
      {around 40 incompatible changes};
-see \cite{changes96}. Many of these
+     \todo{Update link and number}
+see \cite{changes97}. Many of these
 incompatibilities change internal
@@
 \begin{itemize}
-  \item \dealii{} now requires compilers to support C++17, and has
-    started to extensively use C++17 features.
-  \item The \texttt{CUDAWrappers} namespace and its contents --
-    notably things that enable the usage of \texttt{cuSPARSE} algorithms --
-    have been deprecated and will be removed in the next
-    release. \kokkos{} is now used for device-specific optimizations.
+  \item \todo[inline]{Update}
 \end{itemize}
@@
 \dealii can interface with many other libraries:
+\todo[inline]{Any new interfaces?}
+\todo[inline]{For the references below, if a reference is to a
+  website, we should make sure we update that to the current year (and
+  perhaps author list).}
 \begin{multicols}{3}
@@
 The two previous releases of \dealii can be cited as
-\cite{dealII94,dealII95}.
+\cite{dealII95,dealII96}.
 
@@
 
-% This is up-to-date as of 2024-07-23 11:52PM CEST
+\todo[inline]{Add list of contributors}
 \begin{quote}
-Laryssa      Abdala,
-Mathias      Anselmann,
-Abbas        Ballout,
-Maximilian   Bergbauer,
-Julian       Brotz,
-Marco        Feder,
-Niklas       Fehn,
-Menno        Fraters,
-Quang        Hoang,
-Vladimir     Ivannikov,
-Tao          Jin,
-Yimin        Jin,
-%Sebastian    Kinnewig,
-Paras        Kumar,
-Sébastien    Loriot,
-Nils         Much,
-Abdullah     Mujahid,
-Bob          Myhill,
-Paul A.      Patience,
-Luz          Paz,
-Laura        Prieto Saavedra,
-Sebastian    Proell,
-Hendrik      Ranocha,
-Johannes     Resch,
-Andreas      Ritthaler,
-Malik        Scheifinger,
-David        Schneider,
-%Magdalena    Schreter-Fleischhacker,
-Richard      Schussnig,
-Nils         Schween,
-Kyle         Schwiebert,
-Simranjeet   Singh,
-Simon        Sticko,
-Dominik      Still,
-%Jan Philipp  Thiele,
-Thierry      Thomas,
-Vinayak      Vijay,
-Ivy          Weber,
-Simon        Wiesheier,
-Chengjiang   Yin,
-%Vladimir     Yushutin,
-Stefano      Zampini.
 \end{quote}
@@
 
-P.~C.~Africa was partially supported by the consortium iNEST (Interconnected North-East Innovation Ecosystem),
-Piano Nazionale di Ripresa e Resilienza (PNRR) - Missione 4 Componente 2, Investimento 1.5 - D.D. 1058 23/06/2022,
-ECS00000043, supported by the European Union's NextGenerationEU program.
-
 D.~Arndt and B.~Turcksin: Research sponsored by the Laboratory Directed Research and
@@
 
-W.~Bangerth and M.~Fehling were partially supported by Award OAC-1835673
+W.~Bangerth was partially supported by Award OAC-1835673
 as part of the Cyberinfrastructure for Sustained Scientific Innovation (CSSI)
@@
 
-W.~Bangerth was also partially supported by Award EAR-1925595.
+W.~Bangerth was also partially supported by Award EAR-1925595 and OAC-2410847.
 
@@
 
-T.~Heister and V.~Yushutin were also partiallty supported by NSF Awards OAC-2015848 and EAR-1925575.
+T.~Heister were also partiallty supported by NSF Awards OAC-2015848 and EAR-1925575.
 
@@
 ``Immersed methods for multiscale and multiphysics problems (IMMEDIATE)''.
-
-S.~Kinnewig was supported by the Deutsche Forschungsgemeinschaft (DFG) under Germany’s Excellence
-Strategy within the Cluster of Excellence PhoenixD (EXC 2122, Project ID 390833453)
 
@@
 time on the Palmetto cluster.
+\todo{Still relevant?}
 
@@
 \url{http://www.tacc.utexas.edu}
+\todo{Still relevant?}
 
@@
 See \cite{Boerner2023}.
+\todo{Still relevant?}
 

### 2025-07-03 9.7/paper.tex
Add some of the new features of 9.7.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Work towards C++20 modules}
+\label{subsec:modules}
+
+\dealii{} currently requires a compiler that understands C++17, but
+some additional features are enabled if a compiler supports the C++20
+standard. C++20 also introduces ``modules'', a way by which software
+packages can export their public interface without relying on the
+mechanism of textual inclusion of header files that C++ has inherited from C. The
+end goal of C++20 modules is that one can replace the use of long
+lists of statements such as\\
+\hspace*{1cm}\texttt{\#include <deal.II/grid/tria.h}\\
+by a simple\\
+\hspace*{1cm}\texttt{import dealii;}
+
+A lot of effort -- perhaps 6 weeks of full-time work, and around 150
+pull requests -- has gone into evaluating whether \dealii{} can be
+built using modules (in addition to header files, which will need to
+be provided for a long time to come for backward compatibility). A
+detailed accounting of this effort can be found in 
+\cite{bangerth2025experienceconvertinglargemathematical}. The majority
+of this work addressed features of the \dealii{} code base,
+accumulated over its more than 25 year history, that were acceptable
+in a header-based system but not in a module-based system. To name
+just one of many examples, there can be cycles of header files that mutually
+\texttt{\#include} each other, but this is not allowed for
+modules. Most of the pull requests for this project therefore simply
+cleaned up our code base, though without having to introduce any
+incompatibilities.
+
+As shown in \cite{bangerth2025experienceconvertinglargemathematical},
+the use of C++20 modules
+reduces compile times for the library itself, though the evidence is
+mixed for tutorial programs or applications that build on
+\dealii{}. Given that at the moment few systems have compilers, CMake
+versions, and build systems that are sufficiently new to support
+module builds, it will be several years before this feature will
+become widely available. However, the infrastructure for it is now
+largely in place.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{New and improved tutorials and code gallery programs}
@@
     \step{93}
-    demonstrates how one can implement ...
-    \todo{Complete}
+    demonstrates how one can implement problems in which some of the
+    variables to be solved for are not degrees of freedom of a finite
+    element field (so-called ``non-local'' degrees of freedom, because
+    they have no associated node functional that is tied to a cell or
+    a patch of cells). Specifically, the program solves an
+    optimization problem in which some of the variables are scalars
+    that multiply certain source terms.
     \step{93} was written by Sam Scheuerman.
@@
     \step{97}
-    is a program that solves ... curl-curl ...
-    \todo{Complete}
-    \step{93} was written by Siarhei Uzunbajakau.
+    is a program that solves a curl-curl problem from electromagnetics
+    -- more specifically, from magnetostatics. The program shows a
+    formulation in which two curl-curl problems are solved
+    consecutively to model the magnetic field that results from the
+    current in an electrically conducting coil.
+    \step{97} was written by Siarhei Uzunbajakau.
 \end{itemize}
@@
 
+\todo[inline]{Everyone please update as appropriate.}
 D.~Arndt and B.~Turcksin: Research sponsored by the Laboratory Directed Research and

### 2025-07-06 9.7/paper.tex
dealii-X ack.

@@
 \author[8]{Luca~Heltai}
-\affil[8]{University of Pisa, Italy.}
+\affil[8]{Department of Mathematics, University of Pisa, 
+Via Buonarroti 1/c, 56127 Pisa, Italy.}
 
@@
 
+M.~Kronbichler and L.~Heltai were partially supported by the European Research
+Council (ERC) under the European Union's Horizon 2020 research and innovation
+programme (call HORIZON-EUROHPC-JU-2023-COE-03, grant agreement No. 101172493
+``dealii-X: an Exascale Framework for Digital Twins of the Human Body'').
+
 M.~Maier was partially supported by NSF Award DMS-2045636 and and by the

### 2025-07-06 9.7/paper.tex
More ack.

@@
 
-L.~Heltai was partially supported by the Italian Ministry of
-University and Research (MUR), under the grant MUR PRIN 2022 No. 2022WKWZA8
-``Immersed methods for multiscale and multiphysics problems (IMMEDIATE)''.
+L.~Heltai is a member of Gruppo Nazionale per il Calcolo Scientifico (GNCS) of
+Istituto Nazionale di Alta Matematica (INdAM). LH was partially supported by
+the Italian Ministry of University and Research (MUR), under the grant MUR PRIN
+2022 No. 2022WKWZA8 ``Immersed methods for multiscale and multiphysics problems
+(IMMEDIATE)'', and acknowledges the MIUR Excellence Department Project awarded
+to the Department of Mathematics, University of Pisa, CUP I57G22000700001.
 

### 2025-07-08 9.7/paper.tex
We are at version 9.7
Add question for modules

@@
   This paper provides an overview of the new features of the finite element
-  library \dealii, version 9.8.
+  library \dealii, version 9.7.
 \end{abstract}
@@
 \label{subsec:modules}
+
+\todo[inline]{Given that we haven't merged CMake support, the deal.II module isn't usable yet. Should we still include a discussion here?}
 

### 2025-07-10 9.7/paper.tex
Update funding sources
Mention step-95.
Fix typo and update some years

@@
   \item
+    \step{95} ...
+    \todo[inline]{Complete}
+  \item
     \step{97}
@@
 \begin{itemize}
-  \item \emph{``Phase field facture model in 3D''},
+  \item \emph{``Phase field fracture model in 3D''},
     contributed by Wasim Niyaz Munshi,
@@
 Development Program of Oak Ridge National Laboratory, managed by UT-Battelle,
-LLC, for the U. S. Department of Energy.
+LLC, for the U. S. Department of Energy and supported by the U.S. Department of Energy,
+Office of Science, Office of Advanced Scientific Computing Research,
+Next-Generation Scientific Software Technologies program, under contract
+number DE-AC05-00OR22725.
 

### 2025-07-10 9.7/paper.tex
update contributor list

@@
 \begin{quote}
+%  authors removed:
+%  Daniel       Arndt,
+%  Wolfgang     Bangerth,
+%  Bruno        Blais,
+%  Marc         Fehling,
+%  Rene         Gassmoeller,
+%  Timo         Heister,
+%  Luca         Heltai,
+%  Martin       Kronbichler,
+%  Matthias     Maier,
+%  Peter        Muench,
+%  Jean-Paul    Pelteret,
+%  Bruno        Turcksin,
+%  David        Wells,
+Pasquale     Africa,
+Henry        Arhin,
+Maximilian   Bergbauer,
+Alfredo      Buttari,
+Bruna        Campos,
+Nicholas     Cantrell,
+Xiaoming     Cao,
+Chayapol     Chaoveeraprasit,
+Jerett       Cherry,
+Dario        Coscia,
+John         Coughlin,
+Nikita       Daniliuk,
+Crystal      Farris,
+Marco        Feder,
+Emmanuel     Ferdman,
+Federico     Fernandez,
+Olivier      Gaboriault,
+Mohamad      Ghadban,
+Mahdi        Gharehbaygloo,
+Robin        Görmer,
+Davit        Gyulamiryan,
+Lóránt       Hadnagy,
+Fernando     Herrera,
+Robin        Hiniborch,
+Quang        Hoang,
+Jordan       Hoffart,
+Sascha       Hofstetter,
+Yimin        Jin,
+Yann         Jobic,
+Sean         Johnson,
+Vaishnavi    Kale,
+Sebastian    Kinnewig,
+Andreas      Koch,
+Jason        Landini,
+Zhou         Lei,
+Yingli       Li,
+Nils         Margenberg,
+Oreste       Marquis,
+Ryan         Moulday,
+Nils         Much,
+Tileuzhan    Mukhamet,
+Peter        Munch,
+Wasim~Niyaz  Munshi,
+Natalia      Nebulishvili,
+Luz          Paz,
+David        Pecoraro,
+Davide       Polverino,
+Sanjeeb      Poudel,
+Guilhem      Poy,
+Laura        Prieto~Saavedra,
+Sebastian    Proell,
+Andreas      Ritthaler,
+Mayank       Sabharwal,
+Sam          Scheuerman,
+Magdalena    Schreter,
+Richard      Schussnig,
+Kyle         Schwiebert,
+Marc         Secanell,
+Qingyuan     Shi,
+Wyatt        Smith,
+Simon        Sticko,
+Dominik      Still,
+Edward       Terrell,
+Jan~Philipp  Thiele,
+Xiaochuan    Tian,
+Siarhei      Uzunbajakau,
+Mikael       Vaillant,
+Vinayak      Vijay,
+Stephan      Voss,
+Michał       Wichrowski,
+Simon        Wiesheier,
+Yi-Yung      Yang,
 \end{quote}

### 2025-07-14 9.7/paper.tex
Update affiliation and funding

@@
 
- \author[9,10]{Martin~Kronbichler}
+ \author[9]{Martin~Kronbichler}
  \affil[9]{Faculty of Mathematics, Ruhr University Bochum,
@@
  {\texttt{martin.kronbichler@rub.de}}}
- \affil[10]{Institute of Mathematics,
-   University of Augsburg,
-   Universit\"atsstr.~12a, 86159 Augsburg, Germany.
-   }
 
-\author[11]{Matthias~Maier}
-\affil[11]{Department of Mathematics,
+\author[10]{Matthias~Maier}
+\affil[10]{Department of Mathematics,
   Texas A\&M University,
@@
 
-\author[10,12]{Peter Munch}
- \affil[12]{Uppsala University, Sweden.
+\author[11]{Peter Munch}
+ \affil[11]{Uppsala University, Sweden.
   {\texttt{peter.munch@it.uu.se}}}
@@
 
-\author[13]{David Wells}
-\affil[13]{Department of Mathematics, University of North Carolina,
+\author[12]{David Wells}
+\affil[12]{Department of Mathematics, University of North Carolina,
   Chapel Hill, NC 27516, USA.
@@
 
-M.~Kronbichler and P.~Munch were partially supported by the
-German Ministry of Education and Research, project
+M.~Kronbichler was partially supported by the
+German Federal Ministry of Research, Technology and Space, project
 ``PDExa: Optimized software methods for solving partial differential
-equations on exascale supercomputers'' and the Bayerisches Kompetenznetzwerk
-f\"ur Technisch-Wissen\-schaft\-li\-ches Hoch- und H\"ochstleistungsrechnen
-(KONWIHR), project ``Fast and scalable finite element algorithms for coupled
-multiphysics problems and non-matching grids''.
+equations on exascale supercomputers'', grant agreement No. 16ME0637K.
 

### 2025-07-14 9.7/paper.tex
Describe MappingP1.
It is a nice simplex feature!

@@
 
+\subsection{The \texttt{MappingP1} class}
+\dealii has supported simplex and mixed meshes since the 9.3
+release~\cite{changes93}. That release included \texttt{MappingFE}, a
+\texttt{Mapping} class which works with a variety of different elements by
+delegating most computations to a provided \texttt{FiniteElement} class. This
+release includes a new \texttt{MappingP1} class, which contains an optimized
+implementation of the standard affine simplex mapping (i.e., the mapping from
+the unit to real cell with a constant Jacobian on each cell). This
+implementation is about six times faster than the generic one implemented in
+\texttt{MappingFE}. In addition, this \texttt{Mapping} is now the default linear
+mapping returned by \texttt{ReferenceCell::get\_default\_linear\_mapping()}, so
+all applications which use that function for obtaining a
+\texttt{ReferenceCell}-independent mapping will automatically use the more
+performant class.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2025-07-14 9.7/paper.tex
Add some text for SolutionTransfer.

@@
 
+\subsection{The \texttt{SolutionTransfer} class}
+This release extends the \texttt{SolutionTransfer} class to support mesh
+adaptivity with all \texttt{Triangulation} classes which support refinement.
+Consequently, the \texttt{parallel::distributed::Triangulation} class is no
+longer necessary to use and is deprecated. Some member functions, such as
+\texttt{execute\_pure\_refinement()}, have been removed from the new
+implementation.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2025-07-14 9.7/paper.tex
Add some text for the orientation changes.

@@
   \item \todo[inline]{Update}
+  \item We have continued to unify the various implementations of object
+    orientations in the library. While the three booleans orientation, rotation,
+    and flip have maintained their original meanings (i.e., they have default
+    values of \texttt{true}, \texttt{false}, and \texttt{false}), the default
+    value of the combined orientation (available via
+    \texttt{TriaAccessor::combined\_face\_orientation()} and
+    \texttt{TriaAccessor::line\_orientation()}) is now $0$ rather than $1$. In
+    addition, the return type of these functions is now
+    \texttt{types::geometric\_orientation}, which is a \texttt{typedef} for an
+    8-bit integral type.
 \end{itemize}

### 2025-07-16 9.7/paper.tex
Reshuffle sections.
Fix a typo.
Update a couple of sections.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{The \texttt{MappingP1} class}
+
+Classes derived from the \texttt{Mapping} base class are used to
+represent the transformation from a reference cell (such as the unit
+square or reference triangle) to a concrete cell that is part of a mesh.
+\dealii has supported simplex and mixed meshes since the 9.3
+release~\cite{changes93}. That release included \texttt{MappingFE}, a
+\texttt{Mapping} class that uses the shape functions of a finite
+element to implement this transformation; many common mappings can be
+written in this way, and the class works with a variety of different elements by
+delegating most computations to a provided \texttt{FiniteElement}
+object. For example, the common bi-/trilinear mappings for quadrilaterals and
+hexahedra, as well as the linear mappings on triangles and simplices,
+can be implemented using \texttt{MappingFE} along with the
+\texttt{FE\_Q} or \texttt{FE\_SimplexP} classes.
+
+The current
+release now includes a new \texttt{MappingP1} class, which contains an optimized
+implementation of the standard affine simplex mapping (i.e., the mapping from
+the unit/reference cell to the real cell with a constant Jacobian on each cell). This
+implementation is about six times faster than the generic one implemented in
+\texttt{MappingFE}. In addition, this \texttt{Mapping} is now the default linear
+mapping returned by \texttt{ReferenceCell::get\_default\_linear\_mapping()}, so
+all applications which use that function for obtaining a
+\texttt{ReferenceCell}-independent mapping will automatically use the more
+performant class.
+
+\subsection{The \texttt{SolutionTransfer} class}
+This release extends the \texttt{SolutionTransfer} class to facilitate
+the transfer of solution vectors from one mesh to another in an
+adaptive mesh refinement loop, now to
+all triangulation classes that support mesh refinement.
+Consequently, it has subsumed the functionality of the existing
+\texttt{parallel::distributed::SolutionTransfer} class, which is no
+longer necessary and is now deprecated.
+
+In the process of this update, we have also cleaned up these classes'
+interfaces. In particular, some member functions, such as
+\texttt{execute\_pure\_refinement()}, have been removed from the new
+implementation because they provided functionality for only rarely
+used cases, but substantially complicated the internal design of these
+classes.
+
 \subsection{Work towards C++20 modules}
@@
 \end{itemize}
-
-\subsection{The \texttt{MappingP1} class}
-\dealii has supported simplex and mixed meshes since the 9.3
-release~\cite{changes93}. That release included \texttt{MappingFE}, a
-\texttt{Mapping} class which works with a variety of different elements by
-delegating most computations to a provided \texttt{FiniteElement} class. This
-release includes a new \texttt{MappingP1} class, which contains an optimized
-implementation of the standard affine simplex mapping (i.e., the mapping from
-the unit to real cell with a constant Jacobian on each cell). This
-implementation is about six times faster than the generic one implemented in
-\texttt{MappingFE}. In addition, this \texttt{Mapping} is now the default linear
-mapping returned by \texttt{ReferenceCell::get\_default\_linear\_mapping()}, so
-all applications which use that function for obtaining a
-\texttt{ReferenceCell}-independent mapping will automatically use the more
-performant class.
-
-\subsection{The \texttt{SolutionTransfer} class}
-This release extends the \texttt{SolutionTransfer} class to support mesh
-adaptivity with all \texttt{Triangulation} classes which support refinement.
-Consequently, the \texttt{parallel::distributed::Triangulation} class is no
-longer necessary to use and is deprecated. Some member functions, such as
-\texttt{execute\_pure\_refinement()}, have been removed from the new
-implementation.
 

### 2025-07-19 9.7/paper.tex
We are including modules in the release
Some updates.

@@
 
-\todo[inline]{Given that we haven't merged CMake support, the deal.II module isn't usable yet. Should we still include a discussion here?}
-
 \dealii{} currently requires a compiler that understands C++17, but
@@
 
-A lot of effort -- perhaps 6 weeks of full-time work, and around 150
+A lot of effort -- perhaps 6 weeks of full-time work, and nearly 200
 pull requests -- has gone into evaluating whether \dealii{} can be
@@
 in a header-based system but not in a module-based system. To name
-just one of many examples, there can be cycles of header files that mutually
+just two of many examples: (i) there can be cycles of header files that mutually
 \texttt{\#include} each other, but this is not allowed for
-modules. Most of the pull requests for this project therefore simply
-cleaned up our code base, though without having to introduce any
+modules; and (ii) it is allowed to declare functions in header files
+as \texttt{static} or in anonymous namespaces, but this no longer
+works with modules. In both cases, while allowed, the existing code
+was brittle, inefficient, or hard to understand. Most of the pull requests for this project therefore simply
+cleaned up our code base, without having to introduce any
 incompatibilities.
@@
   \item
-    \step{95} ...
-    \todo[inline]{Complete}
+    \step{95} illustrates the use of matrix-free methods to solve
+    problems with embedded boundaries using the CutFEM method that was
+    previously already introduced in \step{85}. \step{95} was written
+    by Maximilian Bergbauer, with help by Martin Kronbichler and Peter Munch.
   \item

### 2025-07-19 9.7/paper.tex
Update numbers.

@@
 Many of the \dealii tutorial programs were revised in a variety of ways
-as part of this release: Around 130 of the more than 1600 (non-merge)
+as part of this release: Around 125 of the more than 1500 (non-merge)
 commits that went into this release touched the tutorial.
 % data generated using these commands:
-% - tutorial: git log --since 2023/07/07 --until 2024/08/11 --no-merges examples | grep commit | wc -l
-% - total:    git log --since 2023/07/07 --until 2024/08/11 --no-merges | grep commit | wc -l
+% - tutorial: git log --since 2024/08/11 --until 2025/07/19 --no-merges examples | grep commit | wc -l
+% - total:    git log --since 2024/08/11 --until 2025/07/19 --no-merges | grep commit | wc -l
 In addition, there are a number of new tutorial

### 2025-07-28 9.7/paper.tex
Describe the major incompatibilities.

@@
 The 9.7 release includes
-\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_5_2_and_9_6_0.html}
-     {around 40 incompatible changes};
-     \todo{Update link and number}
+\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_6_0_and_9_7_0.html}
+     {around 20 incompatible changes};
 see \cite{changes97}. Many of these
@@
   \item \todo[inline]{Update}
+
+  \item The classes \texttt{Subscriptor} and \texttt{SmartPointer} have been
+    renamed to \texttt{EnableObserverPointer} and \texttt{ObserverPointer}. Many
+    core classes, such as \texttt{DoFHandler} and \texttt{Triangulation},
+    inherit from \texttt{EnableObserverPointer} to enable other classes to store
+    \texttt{ObserverPointer}s pointing to them which, if dereferenced after the
+    pointed-to object goes out of scope, will signal an error. This
+    functionality has been in \dealii since 1998. Since ``smart pointers''
+    nowadays means \texttt{std::unique\_ptr} or \texttt{std::shared\_ptr}, these
+    names were updated to avoid this conflict in terminology. The new names are
+    based on the proposal for \texttt{std::observer\_ptr}~\cite{Brown2014} and
+    \texttt{std::enable\_shared\_from\_this}.
+
+  \item In order to support modules (see Section~\ref{subsec:modules}), a
+    significant number of headers had to be split up to prevent circular
+    inclusions and other minor incompatibilities. For example,
+    \texttt{grid/tria.h} no longer contains declarations of \texttt{CellData} or
+    \texttt{SubCellData}. Those declarations have been moved to
+    \texttt{grid/cell\_data.h}.
+
+  \item The 9.5 release~\cite{dealII95} made \kokkos~\cite{trott2022} a required
+    dependency. Kokkos is a C++ library which provides tools for writing code
+    which can efficiently run on a diverse set of architectures and devices.
+    This release removes the deprecated \texttt{CUDAWrappers} namespace as well
+    as all \texttt{CUDA}-related macros in favor of Kokkos equivalents.
+
   \item We have continued to unify the various implementations of object

### 2025-07-28 9.7/paper.tex
Remove a TODO.
We already updated this list.
Update references.

@@
 \dealii can interface with many other libraries:
-\todo[inline]{Any new interfaces?}
-\todo[inline]{For the references below, if a reference is to a
-  website, we should make sure we update that to the current year (and
-  perhaps author list).}
 \begin{multicols}{3}
@@
     \item ADOL-C \cite{griewank1996adolc}
-    \item ArborX \cite{lebrun2020arborx}
+    \item ArborX \cite{prokopenko2025}
     \item ARPACK \cite{lehoucq1998arpack}
@@
     \item Boost \cite{boost-web-page}
-    \item CGAL \cite{cgal-user-ref}
-    \item cuSOLVER \cite{cusolver-web-page}
-    \item cuSPARSE \cite{cusparse-web-page}
+    \item CGAL \cite{cgal:eb-24b}
     \item Gmsh \cite{geuzaine2009gmsh}
@@
     \item HDF5 \cite{hdf5-web-page}
+    \item Kokkos \cite{trott2022}
+    \item Magic Enum \cite{magic-enum-web-page}
     \item METIS \cite{karypis1998metis}
@@
     \item ScaLAPACK \cite{blackford1997scalapack}
-    \item SLEPc \cite{hernandez2005slepc}
-    \item SUNDIALS \cite{hindmarsh2005sundials}
+    \item SLEPc \cite{roman2023improvements}
+    \item SUNDIALS \cite{gardner2022enabling}
     \item SymEngine \cite{symengine-web-page}
@@
     \item TBB \cite{reinders2007tbb}
-    \item Trilinos \cite{heroux2005trilinos,trilinos-web-page}
+    \item Trilinos \cite{mayr2025trilinos,trilinos-web-page}
     \item UMFPACK \cite{davis2004umfpack}
+    \item VTK \cite{vtk-web-page}
+    \item zlib \cite{zlib-web-page}
   \end{itemize}
@@
 
-\todo[inline]{Add list of contributors}
 \begin{quote}

### 2025-07-28 9.7/paper.tex
add taskflow section
Make some progress.

@@
 \item
-  \todo[inline]{Update list}
+  \todo[inline]{Update list from the subsections below}
 \end{itemize}
@@
 \begin{itemize}
-\item
+\item MPGMRES and GMRES updates
+  \todo[inline]{Matthias and Martin, please update (or make this point
+  into one or two subsections below).}
+  \item FiniteElement::shape\_function\_belongs\_to() 
 \todo[inline]{Update list}
@@
 The
-\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_5_2_and_9_6_0.html}{changelog}
-\todo{Update}
--- listing more than XXX
-\todo{Update}
+\href{https://dealii.org/current/doxygen/deal.II/changes_between_9_6_0_and_9_7_0.html}{changelog}
+-- listing more than 120
 features and bugfixes --
@@
 classes.
+
+\subsection{Support for simplex and mixed meshes}
+
+\todo[inline]{Perhaps talk about this point:
+  GridGenerator::convert\_hypercube\_to\_simplex\_mesh() can now split
+  quadrilaterals into two triangles and hexahedra into six
+  tetrahedra. These anisotropic splits are useful in contexts where it
+  is important to add as few new elements as possible.}
+
+\subsection{Updates of \dealii{}'s interfaces to external libraries}
+\label{subsec:external}
+
+\todo[inline]{Perhaps say something about new interfaces to MUMPS and
+  PSBLAS, and the updated interfaces to ArborX. We no longer bundle
+  TBB and by default build on Taskflow. More interfaces to CGAL. Now
+  bundling Kokkos 4.5.1; are we using something new from Kokkos? magic\_enum}
 
@@
 \href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_6_0_and_9_7_0.html}
-     {around 20 incompatible changes};
+     {around 25 incompatible changes};
 see \cite{changes97}. Many of these
-incompatibilities change internal
+incompatibilities remove previously deprecated functions or classes,
+or require now widely available but newer versions of external
+dependencies than previous \dealii{} versions. Others
+change internal
 interfaces that are not usually used in external
-applications. That said, the following are worth mentioning since they
+applications. The following are worth mentioning since they
 are more broadly visible:
@@
     8-bit integral type.
+  \todo[inline]{I didn't pay attention how the entry above came into
+    this paper. David: Is this an incompatibility for the current
+    release? If so, please just delete this note. Otherwise, please
+    remove the bullet point and this note.}
+  \item The \step{52} tutorial has been removed. This program used to
+    show the use of time stepping functionality in \dealii{}, but
+    this functionality was rudimentary compared to what packages such
+    as SUNDIALS or PETSc TS offer. As a consequence, \step{52} did
+    not show the advanced techniques we would like to promote, and it
+    is now superseded by \step{86} (which builds on PETSc TS).
+  \item The \texttt{parallel::distributed::SolutionTransfer} class has
+    been merged with the \texttt{SolutionTransfer} class, with the
+    former now being deprecated. In the process of cleaning up the
+    interface, we have also removed some rarely used member functions
+    of the latter that prevented us from unifying the two classes.
+  \item The process of removing cycles between header files for the
+    modules project (see Section~\ref{subsec:modules}) has led to a
+    small number of cases where a \dealii{} header file was
+    transitively included into a user program, but now no longer
+    is. This is easily fixed by adding an explicit \texttt{\#include}
+    statement to the user program for the file that is necessary to
+    access declarations the compiler now no longer sees.
+  \item The \texttt{MatrixOut} class, used to create graphical
+    representations of matrices, now defaults to creating sparse
+    output, only showing nonzero entries. This vastly increases the
+    size of matrices that can be visualized.
+  \item The \texttt{FiniteElement} class has a number of functions
+    that allow querying properties of implementations in derived
+    classes. These include asking whether a derived class is able to provide
+    ``(generalized) support points'', i.e., whether the element
+    defines its shape functions via nodal interpolation (or via
+    quadrature-based integrals) and can return an array of these
+    points. These functions returned \texttt{false} for the
+    \texttt{FE\_Nothing} element -- an element that has no degrees of
+    freedom and consequently no shape functions. This answer was not
+    wrong, but missed the point: Codes using these functions want to
+    know whether an element is able to provide arrays with these
+    support points, and for \texttt{FE\_Nothing} the answer should be
+    ``yes'' (i.e., \texttt{true}): The class \textit{can} provide such
+    an array, it will just be empty.
 \end{itemize}

### 2025-07-28 9.7/paper.tex
Outline sections.

@@
 \item FiniteElement::shape\_function\_belongs\_to()
-\todo[inline]{Update list}
+  \todo[inline]{Wolfgang to complete}
+  \item \todo[inline]{Add to this list other points that are not
+    explicitly discussed below}
 \end{itemize}
@@
 which we will discuss in this section.
-
-\todo[inline]{Add sections}
 
@@
 
+\subsection{Matrix-free functionality}
+
+\todo[inline]{There are many entries in the changelog about
+  matrix-free stuff, but it will take someone with more knowledge of
+  the area to discuss them here -- or to remove this subsection.}
+
 \subsection{Support for simplex and mixed meshes}
@@
   is important to add as few new elements as possible.}
+
+\todo[inline]{Are there other noteworthy changes}
 
@@
 \begin{itemize}
-  \item \todo[inline]{Update}
-
   \item The classes \texttt{Subscriptor} and \texttt{SmartPointer} have been
@@
     \texttt{SubCellData}. Those declarations have been moved to
-    \texttt{grid/cell\_data.h}.
+    \texttt{grid/cell\_data.h}. In cases where this leads to errors,
+    the situation is easily (and backward compatibly) fixed by adding an explicit 
+    \texttt{\#include} statement to the user program for the file that is necessary to
+    access declarations the compiler now no longer sees.
 
@@
     which can efficiently run on a diverse set of architectures and devices.
-    This release removes the deprecated \texttt{CUDAWrappers} namespace as well
-    as all \texttt{CUDA}-related macros in favor of Kokkos equivalents.
+    The current release removes the deprecated \texttt{CUDAWrappers} namespace as well
+    as all \texttt{CUDA}-related macros in favor of Kokkos
+    equivalents.
+    \todo[inline]{Perhaps move to Section~\ref{subsec:external}?}
 
@@
     as SUNDIALS or PETSc TS offer. As a consequence, \step{52} did
-    not show the advanced techniques we would like to promote, and it
+    not show the advanced techniques we would like to promote (e.g.,
+    time stepping error control and automatic time step size choice), and it
     is now superseded by \step{86} (which builds on PETSc TS).
@@
     of the latter that prevented us from unifying the two classes.
-  \item The process of removing cycles between header files for the
-    modules project (see Section~\ref{subsec:modules}) has led to a
-    small number of cases where a \dealii{} header file was
-    transitively included into a user program, but now no longer
-    is. This is easily fixed by adding an explicit \texttt{\#include}
-    statement to the user program for the file that is necessary to
-    access declarations the compiler now no longer sees.
   \item The \texttt{MatrixOut} class, used to create graphical
@@
       \cite{GLHPB18}.
+      \item For the design of the video lectures and how they can be
+        used in \dealii{}-based courses: \cite{Zarestky2022}.
           \vfill\null
@@
 
-W.~Bangerth was partially supported by Award OAC-1835673
-as part of the Cyberinfrastructure for Sustained Scientific Innovation (CSSI)
-program.
+W.~Bangerth was partially supported by the National Science Foundation
+under awards OAC-1835673, EAR-1925595, and OAC-2410847.
 
@@
 No.~EAR-2149126 via The University of California -- Davis.
-
-W.~Bangerth was also partially supported by Award EAR-1925595 and OAC-2410847.
 

### 2025-07-28 9.7/paper.tex
Complete outline.

@@
   \todo[inline]{Matthias and Martin, please update (or make this point
-  into one or two subsections below).}
-\item FiniteElement::shape\_function\_belongs\_to()
-  \todo[inline]{Wolfgang to complete}
+  into one or two subsections in Section 2 below).}
+\item A common operation in the finite element codes is to ask whether
+  a specific shape function belongs to a particular solution
+  variable. In the past, this was implemented by using the
+  \texttt{FiniteElement::system\_to\_component\_index()} function that
+  returns, among other information, which vector component a shape
+  function belongs to. One then had to test whether this is the vector
+  component (or one of the vector components) that corresponds to a
+  specific variable. The new
+  \texttt{FiniteElement::shape\_function\_belongs\_to()} makes this
+  easier: It takes as argument a \texttt{FiniteElementExtractor} that
+  corresponds to a scalar, vector-valued, or tensor-valued set of
+  solution variables, and directly returns \texttt{true} or
+  \texttt{false}.
+\item \todo[inline]{Perhaps mention the updated interfaces to ArborX
+  here? (Or in Section~\ref{subsec:external}?)}
 \item \todo[inline]{Add to this list other points that are not
@@
 
-\todo[inline]{Perhaps say something about new interfaces to MUMPS and
-  PSBLAS, and the updated interfaces to ArborX. We no longer bundle
-  TBB and by default build on Taskflow. More interfaces to CGAL. Now
+\dealii{} relies on external software packages for many operations
+that are not within the core functionality of a finite element package
+-- see Section~\ref{sec:cite} for a complete list. As in every
+release, we have continued to expand and revise these interfaces. The
+following sub-sections outline these updates.
+
+\subsubsection{Interfaces to the Threading Building Blocks and to
+  Taskflow}
+
+\todo[inline]{There is text about this elsewhere in this
+  document. Perhaps move here?}
+\cite{huang2021taskflow}
+\cite{reinders2007tbb}
+
+\subsubsection{Kokkos, a performance portability library}
+\label{subsec:external-kokkos}
+
+Kokkos is a C++ library that makes it easy to write codes that can
+portably run on CPUs and GPUs of different vendors, using their
+respectiv underlying programming models.
+The \dealii{} 9.5 release~\cite{dealII95} made \kokkos~\cite{trott2022} a required
+dependency. The current release removes the deprecated \texttt{CUDAWrappers} namespace as well
+as all \texttt{CUDA}-related macros in favor of Kokkos
+equivalents.
+
+\todo[inline]{Now
   bundling Kokkos 4.5.1; are we using something new from Kokkos? magic\_enum}
+
+\subsubsection{MUMPS, a multifrontal parallel direct solver}
+
+\todo[inline]{Luca, want to say something here? If not, please delete
+  the subsubsection.}
+
+\subsubsection{PSBLAS, a XXXXX}
+
+\todo[inline]{Complete section name}
+
+\todo[inline]{Luca, want to say something here? If not, please delete
+  the subsubsection.}
+
+\subsubsection{\texttt{magic\_enum}, a XXXXX}
+
+\todo[inline]{Complete section name}
+\todo[inline]{Luca, want to say something here? If not, please delete
+  the subsubsection.}
+
+\subsubsection{CGAL, a computational geometry engine}
+
+\todo[inline]{Do we want to say something here? If not, please delete
+  the subsubsection.}
 
@@
 
-  \item The 9.5 release~\cite{dealII95} made \kokkos~\cite{trott2022} a required
-    dependency. Kokkos is a C++ library which provides tools for writing code
-    which can efficiently run on a diverse set of architectures and devices.
-    The current release removes the deprecated \texttt{CUDAWrappers} namespace as well
+  \item As already mentioned in
+    Section~\ref{subsec:external-kokkos}, the current release
+    removes the deprecated \texttt{CUDAWrappers} namespace as well 
     as all \texttt{CUDA}-related macros in favor of Kokkos
     equivalents.
-    \todo[inline]{Perhaps move to Section~\ref{subsec:external}?}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{How to cite \dealii}\label{sec:cite}
+\section{How to cite \dealii}
+\label{sec:cite}
 

### 2025-07-28 9.7/paper.tex
Add a couple of fixes.

@@
 \dealii has supported simplex and mixed meshes since the 9.3
-release~\cite{changes93}. That release included \texttt{MappingFE}, a
+release~\cite{dealii93}. That release included \texttt{MappingFE}, a
 \texttt{Mapping} class that uses the shape functions of a finite
@@
 \todo[inline]{Now
-  bundling Kokkos 4.5.1; are we using something new from Kokkos? magic\_enum}
+  bundling Kokkos 4.5.1; are we using something new from Kokkos?}
 

### 2025-07-28 9.7/paper.tex
Write the TBB vs Taskflow paragraph.

@@
 
-\todo[inline]{There is text about this elsewhere in this
-  document. Perhaps move here?}
-\cite{huang2021taskflow}
-\cite{reinders2007tbb}
+A large number of operations within \dealii{} are parallelized,
+including through mechanisms like WorkStream \cite{TKB16} that are
+also available to and useful in application codes. For more than
+15 years, since the 6.3 release, \dealii{} has used the Threading
+Building Blocks (TBB) \cite{reinders2007tbb} as the underlying library to
+provide task- and pipeline-based parallelism. However, the TBB has
+downsides that primarily follow from its use of assembler mnemonics
+and the resulting difficulties with supporting all platforms and
+compilers. As a consequence, starting with the current release,
+\dealii{} builds on the Taskflow library \cite{huang2021taskflow} that
+is entirely implemented in C++, is header-only, and provides an
+interface that matches well with current language standards. \dealii{}
+bundles the Taskflow 3.10 release, and now no longer bundles the TBB,
+though the TBB interfaces are still available if a TBB installation
+is found on a system.
 

### 2025-07-29 9.7/paper.tex
Remove a note.
Also add comments in case someone is curious about what is going on
here.
Add subsubsection for arborx

@@
   \texttt{false}.
-\item \todo[inline]{Perhaps mention the updated interfaces to ArborX
-  here? (Or in Section~\ref{subsec:external}?)}
 \item \todo[inline]{Add to this list other points that are not
@@
 portably run on CPUs and GPUs of different vendors, using their
-respectiv underlying programming models.
+respective underlying programming models.
 The \dealii{} 9.5 release~\cite{dealII95} made \kokkos~\cite{trott2022} a required
@@
   the subsubsection.}
+
+\subsubsection{ArborX, a performance-portable geometric search library}
+ArborX~\cite{prokopenko2025} is a library for geometric search on CPUs and GPUS.
+The new ArborX 2.X series is not backward compatible with the older 1.X series.
+Consequently, this incompatibility required an extensive rewrite of our wrappers to
+support ArborX 2.X. As part of the rewrite, the template parameters had to
+be changed. When using the wrappers with ArborX 1.X, \texttt{ArborX::BVH}
+and \texttt{ArborX::DistributedTree} are templated on both the dimension and
+the number types. When using the wrappers with ArborX 2.X, these classes are
+templated on the geometric objects used for the leaf nodes of the
+\texttt{ArborX::BVH}.
 
@@
 
+    % This is similar to a note in the 9.6 paper, but I (David) made more
+    % orientation changes for 9.7. These were done in #17988 and #18333.
   \item We have continued to unify the various implementations of object
@@
     8-bit integral type.
-  \todo[inline]{I didn't pay attention how the entry above came into
-    this paper. David: Is this an incompatibility for the current
-    release? If so, please just delete this note. Otherwise, please
-    remove the bullet point and this note.}
+
   \item The \step{52} tutorial has been removed. This program used to

### 2025-07-29 9.7/paper.tex
Add a todo
Update Kokkos

@@
 
+\todo[inline]{There were several changes in Portable::MatrixFree: support for
+multiple components, generalization to fe degree less than n q points 1d,
+support for DeviceVector and BlockVector}
+
 \subsection{Support for simplex and mixed meshes}
@@
 
-\subsubsection{Kokkos, a performance portability library}
+\subsubsection{\kokkos, a performance portability library}
 \label{subsec:external-kokkos}
 
-Kokkos is a C++ library that makes it easy to write codes that can
+\kokkos is a C++ library that makes it easy to write codes that can
 portably run on CPUs and GPUs of different vendors, using their
@@
 dependency. The current release removes the deprecated \texttt{CUDAWrappers} namespace as well
-as all \texttt{CUDA}-related macros in favor of Kokkos
-equivalents.
-
-\todo[inline]{Now
-  bundling Kokkos 4.5.1; are we using something new from Kokkos?}
+as all \texttt{CUDA}-related macros in favor of \kokkos
+equivalents. The 9.7 release now bundles \kokkos 4.5.1 but the minimum version required is still 3.4.
 
@@
     removes the deprecated \texttt{CUDAWrappers} namespace as well 
-    as all \texttt{CUDA}-related macros in favor of Kokkos
+    as all \texttt{CUDA}-related macros in favor of \kokkos
     equivalents.

### 2025-07-30 9.7/paper.tex
Update MF acknowledgments.

@@
 
-M.~Fehling was also partially supported by the ERC-CZ grant LL2105
-CONTACT, funded by the Czech Ministry of Education, Youth and Sports.
+M.~Fehling was partially supported by the ERC-CZ grant LL2105 CONTACT,
+funded by the Czech Ministry of Education, Youth and Sports. He was also
+partially supported by the Charles University Research Centre Program No.\@ UNCE/24/SCI/005.
 

### 2025-07-30 9.7/paper.tex
Remove trailing whitespaces.
Add subsubsection for Portable::MatrixFree

@@
 
-\todo[inline]{There were several changes in Portable::MatrixFree: support for
-multiple components, generalization to fe degree less than n q points 1d,
-support for DeviceVector and BlockVector}
+\subsection{Portable::Matrix-free}
+\texttt{Portable::MatrixFree} utilizes \kokkos to provide support for GPUs from
+Nvidia, AMD, and Intel. In this release, we have added support for finite
+elements with multiple components, i.e vector-values finite elements. We have
+also added support for finite elements where the degree is less than the number
+of quadrature points in one direction.
+
+We have also introduced two new classes: \texttt{Portable::DeviceVector} and
+\texttt{Portable::DeviceBlockVector}. \texttt{Portable::DeviceVector} is an
+alias for \texttt{Kokkos::View} that replaces the raw pointer previously used to
+access data on the GPU. \texttt{Portable::DeviceBlockVector} extends
+\texttt{Portable::DeviceVector} in a manner similar to the way
+\texttt{BlockVector} extends \texttt{Vector} on the CPU.
+
+Finally, we have streamlined the interface of the functor invoked by the GPU
+kernel. In the past, this interface included a cell index,
+\texttt{Portable::MatrixFree::Data}, which includes precomputed data such as
+quadrature points, shape functions, and other informations that needed to be
+transferred to the GPU, and \texttt{Portable::MatrixFree::SharedData} which
+contained scratch memory used for computation. Now, all this information has
+been consolidated into a single new structure:
+\texttt{Portable::MatrixFree::Data}, greatly simplifying the interface.
 

### 2025-07-30 9.7/paper.tex
Update the release date.
Add a note on PSBLAS.
Mention the new anisotropic simplex conversions.

@@
 
-\dealii version 9.7.0 was released August 11, 2024.
+\dealii version 9.7.0 was released July 22, 2025.
 \todo{Update}
@@
   \texttt{false}.
+\item Most of the functions in the \texttt{GridGenerator} namespace correspond
+  to simple geometries (such as cubes and spheres) discretized with hypercubes.
+  The function \texttt{GridGenerator::convert\_hypercube\_to\_simplex\_mesh()}
+  converts such discretizations to instead use simplices. That function now
+  supports anisotropic splits, including splitting quadrilaterals into two
+  triangles and hexahedra into six tetrahedra, in addition to the standard
+  isotropic splits. These anisotropic splits significantly reduce the total
+  number of cells and are useful in contexts where it is important to add as few
+  new cells as possible.
+
 \item \todo[inline]{Add to this list other points that are not
@@
 
-\todo[inline]{Perhaps talk about this point:
-  GridGenerator::convert\_hypercube\_to\_simplex\_mesh() can now split
-  quadrilaterals into two triangles and hexahedra into six
-  tetrahedra. These anisotropic splits are useful in contexts where it
-  is important to add as few new elements as possible.}
+\todo[inline]{I moved the `improved anisotropic split' change listed here to the
+  `other changes' section.}
 
@@
   the subsubsection.}
+\todo[inline]{DRW: note that while we support configuring with PSBLAS we don't
+  actually implement any interfaces to it yet (i.e., we have one test but no
+  code in the library which uses \texttt{DEAL\_II\_WITH\_PSBLAS}).}
 

### 2025-08-01 9.7/paper.tex
Mention LTO.
Add some more general changes.
I think these are the last two worth noting here.

@@
   new cells as possible.
-
-\item \todo[inline]{Add to this list other points that are not
-    explicitly discussed below}
+\item The build system now supports interprocedural and link-time optimization
+  (LTO). This can be enabled with the CMake configuration option
+  \texttt{-DDEAL\_II\_WITH\_LTO=ON}. This option enables whole-program
+  optimization at link-time and can make executables significantly faster.
+\item MUMPS~\cite{amestoy2019mumps}, a parallel sparse direct solver library,
+  may now be used with the built-in \texttt{SparseMatrix} and \texttt{Vector} classes
+  via the \texttt{SparseDirectMUMPS} class, in addition to interfaces to MUMPS
+  provided by PETSc and Trilinos.
+\item \texttt{GridIn::read\_vtk()} now loads cell-centered data vectors stored
+  in the input file and makes them available via
+  \texttt{GridIn::get\_cell\_data()}.
 \end{itemize}

### 2025-08-02 9.7/paper.tex
Remove the simplex / mixed-mesh section.
Minor edits.

@@
 \dealii version 9.7.0 was released July 22, 2025.
-\todo{Update}
 This paper provides an
@@
 \item
-  \todo[inline]{Update list from the subsections below}
+  \todo[inline]{Update list from the subsections of Section 2
+    below. Like in the previous release papers, we here basically give
+  a 1-3 sentence summary of the subsections 2.x, with a link to that section.}
 \end{itemize}
@@
 
-\subsection{Support for simplex and mixed meshes}
-
-\todo[inline]{I moved the `improved anisotropic split' change listed here to the
-  `other changes' section.}
-
-\todo[inline]{Are there other noteworthy changes}
+%% The major simplex additions are the changes to
+%% convert_hypercube_to_simplex_mesh() and MappingP1: those are discussed
+%% elsewhere so we don't need a dedicated section for the general topic
 

### 2025-08-02 9.7/paper.tex
Add one more note.
LH contributions.

@@
 
-\todo[inline]{Luca, want to say something here? If not, please delete
-  the subsubsection.}
+MUMPS (MUltifrontal Massively Parallel sparse direct Solver) is a widely-used software library designed for solving large sparse systems of linear equations efficiently on parallel architectures~\cite{amestoy2001mumps}.
 
-\subsubsection{PSBLAS, a XXXXX}
+In the 9.7 release, we have significantly enhanced integration with MUMPS,
+without requiring it to be installed through PETSc or Trilinos. Improvements
+include a simplified user interface for direct solvers, mimicking the existing
+interface we provide for UMFPACK, and enhanced performance optimizations that
+leverage recent MUMPS capabilities. Specifically, the deal.II wrappers now
+transparently support parallel matrix factorization, out-of-core computations,
+Block Low-Rank approximations, and GPU support.
 
-\todo[inline]{Complete section name}
+\subsubsection{PSBLAS, a parallel sparse linear algebra library}
 
-\todo[inline]{Luca, want to say something here? If not, please delete
-  the subsubsection.}
-\todo[inline]{DRW: note that while we support configuring with PSBLAS we don't
-  actually implement any interfaces to it yet (i.e., we have one test but no
-  code in the library which uses \texttt{DEAL\_II\_WITH\_PSBLAS}).}
+PSBLAS is a parallel sparse linear algebra library that provides basic operations for sparse matrices and vectors~\cite{Filippone2000psblas}.
 
-\subsubsection{\texttt{magic\_enum}, a XXXXX}
+The library is specifically designed to leverage modern
+parallel computing architectures, including both CPU and GPU systems, making it well-suited for exascale
+computing environments.
+Release 9.7 includes preliminary support for configuring with PSBLAS, allowing users of both frameworks to take advantage of its capabilities for specific linear algebra tasks.
+This support is still in its early stages, and the library is not yet fully integrated into the \dealii{} ecosystem.
 
-\todo[inline]{Complete section name}
-\todo[inline]{Luca, want to say something here? If not, please delete
-  the subsubsection.}
+\subsubsection{\texttt{magic\_enum}, a static reflection library for enums}
 
-\subsubsection{CGAL, a computational geometry engine}
-
-\todo[inline]{Do we want to say something here? If not, please delete
-  the subsubsection.}
+\texttt{magic\_enum} is a modern, header-only C++17 library designed to provide static reflection capabilities for enums without the need for macros or additional boilerplate code.
+It greatly simplifies working with enum types by enabling conversions between enum values and strings, integer values, and various other operations.
+Version 9.7 uses \texttt{magic\_enum} to provide type safe and efficient conversions between enum values and their string representations in the \texttt{ParameterHandler} class.
 
@@
 lists of statements such as\\
-\hspace*{1cm}\texttt{\#include <deal.II/grid/tria.h}\\
+\hspace*{1cm}\texttt{\#include <deal.II/grid/tria.h>}\\
 by a simple\\
@@
     \item PETSc \cite{petsc-user-ref,petsc-web-page}
+    \item PSBLAS \cite{Filippone2000psblas}
     \item ROL \cite{ridzal2014rol}
@@
 
+This release contains patches written in and contributed from
+Antarctica. As a consequence, \dealii{} now consists of work written
+on all seven continents.
+
 \bigskip

### 2025-08-04 9.7/paper.tex
Remove the duplicated MUMPS section.
This got promoted!
Remove non-principal developers from the authorship list.
We still need to settle on who we will add here.

@@
   pdfauthor={
-    Pasquale Claudio Africa,
     Daniel Arndt,
@@
     Peter Munch,
-    Magdalena Schreter-Fleischhacker,
-    Jan Philipp Thiele,
     Bruno Turcksin,
     David Wells,
-    Vladimir Yushutin
   },
@@
   optimization at link-time and can make executables significantly faster.
-\item MUMPS~\cite{amestoy2019mumps}, a parallel sparse direct solver library,
-  may now be used with the built-in \texttt{SparseMatrix} and \texttt{Vector} classes
-  via the \texttt{SparseDirectMUMPS} class, in addition to interfaces to MUMPS
-  provided by PETSc and Trilinos.
 \item \texttt{GridIn::read\_vtk()} now loads cell-centered data vectors stored

### 2025-08-04 9.7/paper.tex
Make tenses more consistent.

@@
 
-Classes derived from the \texttt{Mapping} base class are used to
-represent the transformation from a reference cell (such as the unit
+Classes derived from the \texttt{Mapping} base class implement
+the transformation from a reference cell (such as the unit
 square or reference triangle) to a concrete cell that is part of a mesh.
@@
 The current
-release now includes a new \texttt{MappingP1} class, which contains an optimized
+release includes a new \texttt{MappingP1} class, which contains an optimized
 implementation of the standard affine simplex mapping (i.e., the mapping from
@@
     % orientation changes for 9.7. These were done in #17988 and #18333.
-  \item We have continued to unify the various implementations of object
+  \item We continued unifying the various implementations of object
     orientations in the library. While the three booleans orientation, rotation,
-    and flip have maintained their original meanings (i.e., they have default
+    and flip have maintained their original definitions (i.e., they have default
     values of \texttt{true}, \texttt{false}, and \texttt{false}), the default
@@
 
-  \item The \step{52} tutorial has been removed. This program used to
-    show the use of time stepping functionality in \dealii{}, but
+  \item The \step{52} tutorial was removed. This program showed
+    the use of time stepping functionality in \dealii{}, but
     this functionality was rudimentary compared to what packages such
@@
     is now superseded by \step{86} (which builds on PETSc TS).
-  \item The \texttt{parallel::distributed::SolutionTransfer} class has
-    been merged with the \texttt{SolutionTransfer} class, with the
+  \item The \texttt{parallel::distributed::SolutionTransfer} class is
+    now merged with the \texttt{SolutionTransfer} class, with the
     former now being deprecated. In the process of cleaning up the
@@
     of the latter that prevented us from unifying the two classes.
-  \item The \texttt{MatrixOut} class, used to create graphical
+  \item The \texttt{MatrixOut} class, which creates graphical
     representations of matrices, now defaults to creating sparse

### 2025-08-04 9.7/paper.tex
Add a cross-reference.

@@
 \subsection{The \texttt{SolutionTransfer} class}
+\label{subsec:solutiontransfer}
 This release extends the \texttt{SolutionTransfer} class to facilitate
@@
     interface, we have also removed some rarely used member functions
-    of the latter that prevented us from unifying the two classes.
+    of the latter that prevented us from unifying the two classes. For more
+    information see Section~\ref{subsec:solutiontransfer}.
   \item The \texttt{MatrixOut} class, which creates graphical

### 2025-08-07 9.7/paper.tex
Update matrix-free and MPGMRES parts

@@
 \begin{itemize}
-\item MPGMRES and GMRES updates
-  \todo[inline]{Matthias and Martin, please update (or make this point
-  into one or two subsections in Section 2 below).}
+\item An implementation of the iterative linear solver, named
+  \texttt{SolverMPGMRES}, has been added. The MPGMRES solver is variant of the
+  widely used generalized minimal residual method (GMRES)
+  \cite{Saad1986gmres}, distinguished by the support of $N$ different
+  preconditioners during the course of the solution process
+  \cite{Greif2016gmres}. Multiple Krylov spaces are constructed by
+  $N$-variate, non-commuting polynomials of the preconditioners and the system
+  matrix applied to a residual of fixed degree. The implementation in \dealii
+  supports two strategies, a full variant that constructs all possible
+  combinations of preconditioner application to the initial residual $r$,
+  which for two preconditioners $P_1, P_2$ looks as
+  \[
+    r, P_1r, P_2r, P_1AP_1r, P_2AP_1r, P_1AP_2r, P_2AP_2r, P_1AP_1AP_1r,
+   P_2AP_1AP_1r, \ldots, P_2AP_2AP_2r, \ldots,
+ \]
+ and a truncated version where no combinations of preconditioners are made, e.g.
+ \[
+    r, P_1r, P_2r, P_1AP_1r, P_2AP_2r, P_1AP_1AP_1r,
+   P_2AP_2AP_2r, \ldots.
+ \]
+ The new solver is beneficial when two different preconditioners can help to
+ more quickly span the space where the solution is expected.
 \item A common operation in the finite element codes is to ask whether
@@
 
-\todo[inline]{There are many entries in the changelog about
-  matrix-free stuff, but it will take someone with more knowledge of
-  the area to discuss them here -- or to remove this subsection.}
+In this release, several extensions, major bug fixes and optimizations of the
+matrix-free infrastructure and related multigrid functionality, such as
+matrix-free geometric and polynomial coarsening, in \dealii have been
+made. The most significant advances were made for the \kokkos-based
+``portable'' infrastructure described below. In addition, improvements where
+made to several areas:
+\begin{itemize}
+\item the \texttt{MatrixFreeTools::compute\_matrix()} and
+  \texttt{MatrixFreeTools::compute\_diagonal()} functions to compute the
+  matrix or only its diagonal from a matrix-free operator were enhanced
+  towards multi-component systems,
+\item support for additional ghost cells in the matrix-free setup was added to
+  support vector access to a wider range operations, such as MPI
+  parallelization of multigrid smoothers based on vertex
+  patches~\cite{Wichrowski2025smoothers},
+\item functionality for identifying degrees of freedom in lexicographic
+  ordering on patches of cells,
+\item support for additional matrices in the
+  \texttt{TensorProductMatrixCreator} for support in fast
+  diagonalization-based inverses.
+\end{itemize}
 
-\subsection{Portable::Matrix-free}
+\subsubsection{Portable::Matrix-free}
 \texttt{Portable::MatrixFree} utilizes \kokkos to provide support for GPUs from
@@
 elements with multiple components, i.e vector-values finite elements. We have
-also added support for finite elements where the degree is less than the number
-of quadrature points in one direction.
+also added support for finite elements where the number
+of quadrature points per direction is different from the size of the
+one-dimensional basis (i.e., polynomial degree plus one), thus supporting more
+cases of evaluation. In addition to the extended functionality, several
+performance optimizations to the evaluation routines have been made, which are
+particularly visible on recent Nvidia GPUs.
 
-We have also introduced two new classes: \texttt{Portable::DeviceVector} and
-\texttt{Portable::DeviceBlockVector}. \texttt{Portable::DeviceVector} is an
-alias for \texttt{Kokkos::View} that replaces the raw pointer previously used to
-access data on the GPU. \texttt{Portable::DeviceBlockVector} extends
+To ensure more uniform handling of temporary data structures necessary for
+storing temporary results in matrix-free evaluations, such as
+sum-factorization algorithms, we have introduced two new classes:
+\texttt{Portable::DeviceVector} and
+\texttt{Portable::DeviceBlockVector}. The class \texttt{Portable::DeviceVector} is an
+alias for \texttt{Kokkos::View} that replaces the raw pointer previously used
+to access data on the GPU. The class \texttt{Portable::DeviceBlockVector} extends
 \texttt{Portable::DeviceVector} in a manner similar to the way
@@
     is now superseded by \step{86} (which builds on PETSc TS).
-  \item The \texttt{parallel::distributed::SolutionTransfer} class is
-    now merged with the \texttt{SolutionTransfer} class, with the
+  \item The \texttt{parallel::distributed::SolutionTransfer} class has been
+    merged with the class \texttt{SolutionTransfer}, with the
     former now being deprecated. In the process of cleaning up the

### 2025-08-07 9.7/paper.tex
Add a TODO.
Fix a line which spilt into the margin.

@@
 \item Most of the functions in the \texttt{GridGenerator} namespace correspond
-  to simple geometries (such as cubes and spheres) discretized with hypercubes.
-  The function \texttt{GridGenerator::convert\_hypercube\_to\_simplex\_mesh()}
-  converts such discretizations to instead use simplices. That function now
-  supports anisotropic splits, including splitting quadrilaterals into two
-  triangles and hexahedra into six tetrahedra, in addition to the standard
-  isotropic splits. These anisotropic splits significantly reduce the total
-  number of cells and are useful in contexts where it is important to add as few
-  new cells as possible.
+  to simple geometries (such as cubes, spheres, and other shapes with analytical
+  formulas) discretized with hypercubes. The function
+  \texttt{GridGenerator::convert\_hypercube\_to\_simplex\_mesh()} converts such
+  discretizations to instead use simplices. That function now supports
+  anisotropic splits, including splitting quadrilaterals into two triangles and
+  hexahedra into six tetrahedra, in addition to the standard isotropic splits.
+  These anisotropic splits significantly reduce the total number of cells and
+  are useful in contexts where it is important to add as few new cells as
+  possible.
 \item The build system now supports interprocedural and link-time optimization
@@
 is found on a system.
+
+\todo[inline]{Timo: Perhaps add the figures from the WorkStream paper
+  updated for TBB vs Taskflow.}
 

### 2025-08-07 9.7/paper.tex
List the major changes with links.

@@
 \item
-  \todo[inline]{Update list from the subsections of Section 2
-    below. Like in the previous release papers, we here basically give
-  a 1-3 sentence summary of the subsections 2.x, with a link to that section.}
+\todo[inline]{Keep this list of major changes up-to-date - the formatting is the
+  same as the 9.6 paper.}
+
+\item \texttt{MappingP1}, a more efficient mapping for simplex elements
+  (Section~\ref{subsec:mappingp1})
+
+\item Improved support for transfering solution vectors during mesh refinement
+  (Section~\ref{subsec:solutiontransfer})
+
+\item Improved matrix-free support for multi-component systems as well as
+  improved portability between GPUs via \texttt{Portable::MatrixFree}
+  (Section~\ref{subsec:matrixfree})
+
+\item Improved support for threading via TaskFlow, as well as support for
+  several new external libraries and improved support for Kokkos and ArborX
+  (Section~\ref{subsec:external})
+
+\item Support for C++20 modules (Section~\ref{subsec:modules})
+
+\item Three new tutorial programs and a new code gallery program
+  (Section~\ref{subsec:steps})
 \end{itemize}
@@
 \subsection{The \texttt{MappingP1} class}
+\label{subsec:mappingp1}
 
@@
 \subsection{Matrix-free functionality}
+\label{subsec:matrixfree}
 

### 2025-08-07 9.7/paper.tex
Fix minor typos

@@
 
-\item Improved support for transfering solution vectors during mesh refinement
+\item Improved support for transferring solution vectors during mesh refinement
   (Section~\ref{subsec:solutiontransfer})
@@
 \subsubsection{ArborX, a performance-portable geometric search library}
-ArborX~\cite{prokopenko2025} is a library for geometric search on CPUs and GPUS.
+ArborX~\cite{prokopenko2025} is a library for geometric search on CPUs and GPUs.
 The new ArborX 2.X series is not backward compatible with the older 1.X series.

### 2025-08-07 9.7/paper.tex
Remove Magdalena's affiliation.
She isn't a coauthor this time.
Fix more minor typos

@@
 
-T.~Heister were also partiallty supported by NSF Awards OAC-2015848 and EAR-1925575.
+T.~Heister were also partially supported by NSF Awards OAC-2015848 and EAR-1925575.
 
@@
 
-M.~Maier was partially supported by NSF Award DMS-2045636 and and by the
+M.~Maier was partially supported by NSF Award DMS-2045636 and by the
 Air Force Office of Scientific Research under grant/contract number
@@
 P.~Munch acknowledges the funding by the Swedish Research Council (VR) under grant 2021-04620 and the strategic funding from the IT department of Uppsala University.
-
-M.~Schreter-Fleischhacker was supported by the Austrian Science Fund (FWF) Schrödinger Fellowship (project number: J4577) and by the European Research Council
-through the ERC Starting Grant ExcelAM (project number: 101117579).
 

### 2025-08-10 9.7/paper.tex
Small changes

@@
 \author[11]{Peter Munch}
- \affil[11]{Uppsala University, Sweden.
-  {\texttt{peter.munch@it.uu.se}}}
+ \affil[11]{Institute of Mathematics, Technical University of Berlin, Germany.
+  {\texttt{muench@math.tu-berlin.de}}}
 
@@
   matrix or only its diagonal from a matrix-free operator were enhanced
-  towards multi-component systems,
+  towards multi-component systems, the DG context (and, in general, to operators
+  with face integrals), and mixed elements,
 \item support for additional ghost cells in the matrix-free setup was added to
@@
 
+\todo[inline]{Peter @ Luca: I think a code snippet would be great here! }
+
 \subsubsection{ArborX, a performance-portable geometric search library}
@@
   \item
-    \step{95} illustrates the use of matrix-free methods to solve
-    problems with embedded boundaries using the CutFEM method that was
-    previously already introduced in \step{85}. \step{95} was written
+    \step{95} extends \step{85} and illustrates the use of matrix-free methods to solve
+    problems with embedded boundaries using the CutFEM method~\cite{bergbauer2025high}. \step{95} was written
     by Maximilian Bergbauer, with help by Martin Kronbichler and Peter Munch.
@@
     Marc Fehling.
+
+\todo[inline]{Peter: A short description would be great here!}
 \end{itemize}
@@
 
-P.~Munch acknowledges the funding by the Swedish Research Council (VR) under grant 2021-04620 and the strategic funding from the IT department of Uppsala University.
-
 D.~Wells was supported by NSF Award OAC-1931516.

### 2025-08-11 9.7/paper.tex
Fix grammar mistake.
Edit for language.

@@
  \affil[2]{Department of Mathematics, Colorado State University, Fort
-   Collins, CO 80523-1874, USA.
+   Collins, CO 80523, USA.
    \texttt{bangerth@colostate.edu}}
@@
 \begin{itemize}
-\item
-\todo[inline]{Keep this list of major changes up-to-date - the formatting is the
-  same as the 9.6 paper.}
+\item \texttt{MappingP1}, a more efficient mapping for simplex elements
+  (Section~\ref{subsec:mappingp1});
 
-\item \texttt{MappingP1}, a more efficient mapping for simplex elements
-  (Section~\ref{subsec:mappingp1})
-
-\item Improved support for transferring solution vectors during mesh refinement
-  (Section~\ref{subsec:solutiontransfer})
+\item Updated support for transferring solution vectors during mesh refinement
+  (Section~\ref{subsec:solutiontransfer});
 
@@
   improved portability between GPUs via \texttt{Portable::MatrixFree}
-  (Section~\ref{subsec:matrixfree})
+  (Section~\ref{subsec:matrixfree});
 
-\item Improved support for threading via TaskFlow, as well as support for
+\item Support for threading via TaskFlow, as well as support for
   several new external libraries and improved support for Kokkos and ArborX
-  (Section~\ref{subsec:external})
+  (Section~\ref{subsec:external});
 
-\item Support for C++20 modules (Section~\ref{subsec:modules})
+\item \dealii{} can now build C++20 modules (Section~\ref{subsec:modules});
 
 \item Three new tutorial programs and a new code gallery program
-  (Section~\ref{subsec:steps})
+  (Section~\ref{subsec:steps}).
 \end{itemize}
 %
+\todo[inline]{Update the list above if we add subsections to section 2.}
 
@@
 \begin{itemize}
-\item An implementation of the iterative linear solver, named
-  \texttt{SolverMPGMRES}, has been added. The MPGMRES solver is variant of the
+\item MPGMRES is a variant of the
   widely used generalized minimal residual method (GMRES)
-  \cite{Saad1986gmres}, distinguished by the support of $N$ different
-  preconditioners during the course of the solution process
-  \cite{Greif2016gmres}. Multiple Krylov spaces are constructed by
+  \cite{Saad1986gmres}, distinguished by the support for multiple
+  preconditioners (thus, \textit{MP}GMRES)  during the course of the solution process
+  \cite{Greif2016gmres}. In our implementation in the
+  \texttt{SolverMPGMRES} class, the use of $N$ preconditioners leads
+  to the construction of multiple Krylov spaces via
   $N$-variate, non-commuting polynomials of the preconditioners and the system
-  matrix applied to a residual of fixed degree. The implementation in \dealii
-  supports two strategies, a full variant that constructs all possible
+  matrix applied to a residual vector. \texttt{SolverMPGMRES}
+  supports two strategies: (i) a full variant that constructs all possible
   combinations of preconditioner application to the initial residual $r$,
@@
     r, P_1r, P_2r, P_1AP_1r, P_2AP_1r, P_1AP_2r, P_2AP_2r, P_1AP_1AP_1r,
-   P_2AP_1AP_1r, \ldots, P_2AP_2AP_2r, \ldots,
+   P_2AP_1AP_1r, \ldots, P_2AP_2AP_2r, \ldots;
  \]
- and a truncated version where no combinations of preconditioners are made, e.g.
+ and (ii) a truncated version where no combinations of preconditioners are considered, i.e.
  \[
@@
  The new solver is beneficial when two different preconditioners can help to
- more quickly span the space where the solution is expected.
-\item A common operation in the finite element codes is to ask whether
+ more quickly span the space in which the solution is expected.
+\item A common operation in finite element codes is to ask whether
   a specific shape function belongs to a particular solution
@@
   These anisotropic splits significantly reduce the total number of cells and
-  are useful in contexts where it is important to add as few new cells as
+  are useful in contexts where one would like to use as few cells as
   possible.
@@
 written in this way, and the class works with a variety of different elements by
-delegating most computations to a provided \texttt{FiniteElement}
+delegating most computations to a \texttt{FiniteElement}
 object. For example, the common bi-/trilinear mappings for quadrilaterals and
@@
 
-\subsection{The \texttt{SolutionTransfer} class}
+\subsection{Updates to the \texttt{SolutionTransfer} class}
 \label{subsec:solutiontransfer}
@@
 
-In this release, several extensions, major bug fixes and optimizations of the
+The current release includes several extensions, major bug fixes, and optimizations of the
 matrix-free infrastructure and related multigrid functionality, such as
-matrix-free geometric and polynomial coarsening, in \dealii have been
-made. The most significant advances were made for the \kokkos-based
+matrix-free geometric and polynomial coarsening. The most significant advances were made for the \kokkos-based
 ``portable'' infrastructure described below. In addition, improvements where
@@
 \begin{itemize}
-\item the \texttt{MatrixFreeTools::compute\_matrix()} and
+\item The \texttt{MatrixFreeTools::compute\_matrix()} and
   \texttt{MatrixFreeTools::compute\_diagonal()} functions to compute the
-  matrix or only its diagonal from a matrix-free operator were enhanced
+  matrix or only its diagonal from a matrix-free operator were extended
   towards multi-component systems, the DG context (and, in general, to operators
-  with face integrals), and mixed elements,
-\item support for additional ghost cells in the matrix-free setup was added to
-  support vector access to a wider range operations, such as MPI
+  with face integrals), and mixed elements.
+\item We added support for additional ghost cells in the matrix-free setup to
+  support vector access to a wider range of operations, such as MPI
   parallelization of multigrid smoothers based on vertex
-  patches~\cite{Wichrowski2025smoothers},
-\item functionality for identifying degrees of freedom in lexicographic
-  ordering on patches of cells,
-\item support for additional matrices in the
-  \texttt{TensorProductMatrixCreator} for support in fast
-  diagonalization-based inverses.
+  patches~\cite{Wichrowski2025smoothers}.
+\item We implemented functionality for identifying degrees of freedom in lexicographic
+  ordering on patches of cells.
+\item Support for additional matrices in the
+  \texttt{TensorProductMatrixCreator} class, supporting fast
+  diagonalization-based inverses, was added.
 \end{itemize}
 
-\subsubsection{Portable::Matrix-free}
-\texttt{Portable::MatrixFree} utilizes \kokkos to provide support for GPUs from
+In a broader context, the
+\texttt{Portable::MatrixFree} class utilizes \kokkos to provide support for GPUs from
 Nvidia, AMD, and Intel. In this release, we have added support for finite
-elements with multiple components, i.e vector-values finite elements. We have
+elements with multiple components, i.e., ``vector-valued'' finite elements. We have
 also added support for finite elements where the number
 of quadrature points per direction is different from the size of the
-one-dimensional basis (i.e., polynomial degree plus one), thus supporting more
+one-dimensional basis (e.g., polynomial degree plus one), thus supporting more
 cases of evaluation. In addition to the extended functionality, several
@@
 
-To ensure more uniform handling of temporary data structures necessary for
+To ensure more uniform handling of data structures necessary for
 storing temporary results in matrix-free evaluations, such as
@@
 \texttt{Portable::DeviceVector} and
-\texttt{Portable::DeviceBlockVector}. The class \texttt{Portable::DeviceVector} is an
+\texttt{Portable::DeviceBlockVector}. The first of these is an
 alias for \texttt{Kokkos::View} that replaces the raw pointer previously used
@@
 kernel. In the past, this interface included a cell index,
-\texttt{Portable::MatrixFree::Data}, which includes precomputed data such as
-quadrature points, shape functions, and other informations that needed to be
-transferred to the GPU, and \texttt{Portable::MatrixFree::SharedData} which
-contained scratch memory used for computation. Now, all this information has
-been consolidated into a single new structure:
+\texttt{Portable::MatrixFree::Data} (which includes precomputed data such as
+quadrature points, shape functions, and other informations that needs to be
+transferred to the GPU), and \texttt{Portable::MatrixFree::SharedData} (which
+contained scratch memory used for computation). Now, all this information has
+been consolidated into a single new class
 \texttt{Portable::MatrixFree::Data}, greatly simplifying the interface.
@@
 is entirely implemented in C++, is header-only, and provides an
-interface that matches well with current language standards. \dealii{}
+interface that matches well with current C++ language standards. \dealii{}
 bundles the Taskflow 3.10 release, and now no longer bundles the TBB,
@@
 
-\kokkos is a C++ library that makes it easy to write codes that can
+\kokkos~\cite{trott2022} is a C++ library that makes it easy to write codes that can
 portably run on CPUs and GPUs of different vendors, using their
 respective underlying programming models.
-The \dealii{} 9.5 release~\cite{dealII95} made \kokkos~\cite{trott2022} a required
+The \dealii{} 9.5 release~\cite{dealII95} made \kokkos a required
 dependency. The current release removes the deprecated \texttt{CUDAWrappers} namespace as well
@@
 
-MUMPS (MUltifrontal Massively Parallel sparse direct Solver) is a widely-used software library designed for solving large sparse systems of linear equations efficiently on parallel architectures~\cite{amestoy2001mumps}.
+MUMPS (MUltifrontal Massively Parallel sparse direct Solver) is a
+widely-used software library designed for efficiently solving large sparse linear systems on parallel architectures~\cite{amestoy2001mumps}.
 
@@
 transparently support parallel matrix factorization, out-of-core computations,
-Block Low-Rank approximations, and GPU support.
+block low-rank approximations, and GPU support.
 
@@
 computing environments.
-Release 9.7 includes preliminary support for configuring with PSBLAS, allowing users of both frameworks to take advantage of its capabilities for specific linear algebra tasks.
+Release 9.7 includes preliminary support for configuring with PSBLAS, allowing users to take advantage of its capabilities for specific linear algebra tasks.
 This support is still in its early stages, and the library is not yet fully integrated into the \dealii{} ecosystem.
@@
 
-T.~Heister were also partially supported by NSF Awards OAC-2015848 and EAR-1925575.
+T.~Heister was also partially supported by NSF Awards OAC-2015848 and EAR-1925575.
 

### 2025-08-11 9.7/paper.tex
Use recent reference for ROL library.
I did not use expanse for the development of deal.II during this term.

@@
     \item PSBLAS \cite{Filippone2000psblas}
-    \item ROL \cite{ridzal2014rol}
+    \item ROL \cite{ROL2022ICCOPT}
     \item ScaLAPACK \cite{blackford1997scalapack}
@@
 
-This work used the \textit{Expanse} HPC system at the San Diego Supercomputer Center (SDSC) at UC San Diego through the CIG Science Gateway and Community Codes for the Geodynamics Community MCA08X011 allocation from the Advanced Cyberinfrastructure Coordination Ecosystem: Services \& Support (ACCESS) program, which is supported by National Science Foundation grants \#2138259, \#2138286, \#2138307, \#2137603, and \#2138296.
-See \cite{Boerner2023}.
-\todo{Still relevant?}
-
 \bibliography{paper}{}

### 2025-08-11 9.7/paper.tex
However, I used the computing cluster of my institute to fix some parallelization bugs.
More editing.

@@
 used cases, but substantially complicated the internal design of these
-classes.
+classes. To maintain prolonged support for these edge cases and to ensure
+a smooth transition to the updated interface, the previous implementation
+remains available in the deprecated \texttt{Legacy::SolutionTransfer} class,
+which is planned for removal in a future release.
 
@@
 
-\subsubsection{\texttt{magic\_enum}, a static reflection library for enums}
+\subsubsection{Magic Enum, a static reflection library for enums}
 
-\texttt{magic\_enum} is a modern, header-only C++17 library designed to provide static reflection capabilities for enums without the need for macros or additional boilerplate code.
+Magic Enum (alternatively known via its in-code spelling \texttt{magic\_enum}) is a modern, header-only C++17 library designed to provide static reflection capabilities for enums without the need for macros or additional boilerplate code.
 It greatly simplifies working with enum types by enabling conversions between enum values and strings, integer values, and various other operations.
@@
 and \texttt{ArborX::DistributedTree} are templated on both the dimension and
-the number types. When using the wrappers with ArborX 2.X, these classes are
+the number type. When using the wrappers with ArborX 2.X, these classes are
 templated on the geometric objects used for the leaf nodes of the
-\texttt{ArborX::BVH}.
+\texttt{ArborX::BVH} type.
 
@@
 module builds, it will be several years before this feature will
-become widely available. However, the infrastructure for it is now
-largely in place.
+become widely usable. However, the infrastructure for it is now
+in place, and one can build the current release into a C++20 module by
+providing CMake with the flag \texttt{-DDEAL\_II\_WITH\_CXX20\_MODULE=ON}.
 
@@
     variables to be solved for are not degrees of freedom of a finite
-    element field (so-called ``non-local'' degrees of freedom, because
+    element field (i.e., they are ``non-local'' degrees of freedom, because
     they have no associated node functional that is tied to a cell or
@@
     Wolfgang Bangerth, and
-    Marc Fehling.
-
-\todo[inline]{Peter: A short description would be great here!}
+    Marc Fehling. The program solves a formulation for fracture
+    propagation in which one uses a separate field that tracks the
+    ``damage'' to the material that has resulted from deformation,
+    with the damage weakening the elastic properties of the deforming solid.
 \end{itemize}
@@
     \texttt{ObserverPointer}s pointing to them which, if dereferenced after the
-    pointed-to object goes out of scope, will signal an error. This
+    pointed-to object goes out of scope, will signal an error that the
+    pointer is now ``dangling''. This
     functionality has been in \dealii since 1998. Since ``smart pointers''
-    nowadays means \texttt{std::unique\_ptr} or \texttt{std::shared\_ptr}, these
+    nowadays typically refer to the \texttt{std::unique\_ptr} or
+    \texttt{std::shared\_ptr} classes, these
     names were updated to avoid this conflict in terminology. The new names are
@@
     inclusions and other minor incompatibilities. For example,
-    \texttt{grid/tria.h} no longer contains declarations of \texttt{CellData} or
-    \texttt{SubCellData}. Those declarations have been moved to
-    \texttt{grid/cell\_data.h}. In cases where this leads to errors,
+    \texttt{deal.II/grid/tria.h} no longer contains declarations of \texttt{CellData} or
+    \texttt{SubCellData}. Those declarations have instead been moved to
+    \texttt{deal.II/grid/cell\_data.h}. In cases where this leads to errors,
     the situation is easily (and backward compatibly) fixed by adding an explicit
@@
 
-  \item As already mentioned in
+  \item As mentioned in
     Section~\ref{subsec:external-kokkos}, the current release
@@
     % orientation changes for 9.7. These were done in #17988 and #18333.
-  \item We continued unifying the various implementations of object
-    orientations in the library. While the three booleans orientation, rotation,
-    and flip have maintained their original definitions (i.e., they have default
+  \item We have continued unifying the various implementations of object
+    orientations in the library, i.e., the description of how cells,
+    faces, or edges are oriented relative to other such objects. While the three booleans `orientation', `rotation',
+    and `flip' have maintained their original definitions (i.e., they have default
     values of \texttt{true}, \texttt{false}, and \texttt{false}), the default
@@
     output, only showing nonzero entries. This vastly increases the
-    size of matrices that can be visualized.
+    size of matrices that can be visualized. Creating dense output
+    remains possible through a flag in the class's
+    \texttt{AdditionalData} structure.
   \item The \texttt{FiniteElement} class has a number of functions
@@
       \item For the design of the video lectures and how they can be
-        used in \dealii{}-based courses: \cite{Zarestky2022}.
+        used in teaching: \cite{Zarestky2022}.
           \vfill\null
@@
 \dealii can interface with many other libraries:
-\begin{multicols}{3}
+\begin{multicols}{4}
   \begin{itemize}[leftmargin=4mm]
@@
     \item Assimp \cite{schulze2021assimp}
-    \item BLAS and LAPACK \cite{anderson1999lapack}
+    \item BLAS, LAPACK \cite{anderson1999lapack}
     \item Boost \cite{boost-web-page}
@@
 
+\todo[inline]{Remove all who we promoted to authors of this paper from
+  the following list.}
+
 \begin{quote}
@@
 %  Matthias     Maier,
-%  Peter        Muench,
+%  Peter        Munch,
 %  Jean-Paul    Pelteret,
@@
 Tileuzhan    Mukhamet,
-Peter        Munch,
 Wasim~Niyaz  Munshi,
@@
 \end{quote}
-Their contributions are much appreciated!
+Their contributions are much appreciated -- this project lives by its
+community of users and contributors!
 
@@
 
-Clemson University is acknowledged for generous allotment of compute
-time on the Palmetto cluster.
-\todo{Still relevant?}
-
-The authors acknowledge the Texas Advanced Computing Center (TACC) at The University of Texas at Austin for providing HPC resources that have contributed to the research results reported within this paper.
-\url{http://www.tacc.utexas.edu}
-\todo{Still relevant?}
+Charles University is acknowledged for providing computing time on the Sn\v{e}hurka cluster.
 

### 2025-08-14 9.7/paper.tex
Mention the Legacy::SolutionTransfer class.
Change a word.
Fix some minor grammar problems.

--- 2025-08-11 9.7/paper.tex
@@
 \texttt{Portable::MatrixFree::Data} (which includes precomputed data such as
-quadrature points, shape functions, and other informations that needs to be
+quadrature points, shape functions, and other information that needs to be
 transferred to the GPU), and \texttt{Portable::MatrixFree::SharedData} (which
@@
     functionality has been in \dealii since 1998. Since ``smart pointers''
-    nowadays typically refer to the \texttt{std::unique\_ptr} or
+    nowadays typically refers to the \texttt{std::unique\_ptr} or
     \texttt{std::shared\_ptr} classes, these

### 2025-08-16 9.7/paper.tex
Make grammar unambiguous in one place :-)
tbb vs taskflow results

@@
 also available to and useful in application codes. For more than
-15 years, since the 6.3 release, \dealii{} has used the Threading
+15 years, since the 6.3 release, \dealii{} has used the Intel Threading
 Building Blocks (TBB) \cite{reinders2007tbb} as the underlying library to
@@
 
-\todo[inline]{Timo: Perhaps add the figures from the WorkStream paper
-  updated for TBB vs Taskflow.}
+In Figure~\ref{fig:tbb-vs-taskflow} we compare the time for assembling
+the temperature matrix and temperature right-hand side in the step-32
+tutorial program, like it was done in \cite{TKB16}. The test was performed
+on a 3d mesh with 393,216 cells and 3,195,010 temperature degrees of freedom
+using a $Q_2$ element. The tests are run
+on AMD EPYC 9654P with 96 cores (and 192 threads). One can clearly see that
+running all copy operations sequentially limits parallel scalability (to less than 10
+threads on the left, where the copy operation is more expensive than the right),
+while using graph coloring (denoted as ``colored'') eliminates this bottleneck.
+It is clearly visible that the Taskflow implementation is better or the same
+in all situations.
+
+\begin{figure}[hb]
+\label{fig:tbb-vs-taskflow}
+\includegraphics{kokkos-vs-tbb/plots-figure1.pdf}\hfill
+\includegraphics{kokkos-vs-tbb/plots-figure2.pdf}
+\caption{Assembly times in a modified step-32 (left: temperature matrix assembly, right: temperature right-hand side assembly) comparing TBB and Taskflow with and without coloring.}
+\end{figure}
 
@@
     pointer is now ``dangling''. This
-    functionality has been in \dealii since 1998. Since ``smart pointers''
+    functionality has been in \dealii since 1998. Since the term ``smart pointers''
     nowadays typically refers to the \texttt{std::unique\_ptr} or

### 2025-08-16 9.7/paper.tex
fix acknowledge

@@
 \begin{figure}[hb]
-\label{fig:tbb-vs-taskflow}
 \includegraphics{kokkos-vs-tbb/plots-figure1.pdf}\hfill
@@
 \caption{Assembly times in a modified step-32 (left: temperature matrix assembly, right: temperature right-hand side assembly) comparing TBB and Taskflow with and without coloring.}
+\label{fig:tbb-vs-taskflow}
 \end{figure}
@@
 
-T.~Heister was also partially supported by NSF Awards OAC-2015848 and EAR-1925575.
-
-T.~Heister was also partially supported by NSF OAC-2410848.
+T.~Heister was also partially supported by NSF Awards OAC-2015848, EAR-1925575, and OAC-2410848.
 
@@
 
+This research used in part resources on the Palmetto Cluster at Clemson University under National Science Foundation awards MRI 1228312, II NEW 1405767, MRI 1725573, and MRI 2018069.
+The views expressed in this article do not necessarily represent the views of NSF or the United States government.
+
 \bibliography{paper}{}

### 2025-08-16 9.7/paper.tex
add iso_Q1 work

@@
   \texttt{GridIn::get\_cell\_data()}.
+\item Sparsity patterns for matrices built with \texttt{FE\_Q\_iso\_Q1} elements
+  now correctly reflect the inherit sparsity on each cell, making the usage of
+  this element usable as a preconditioner for a higher order problems (this is known
+  as low order rediscretization, also see \cite{pazner2023low}).
 \end{itemize}

### 2025-08-18 9.7/paper.tex
Fix grammar.

--- 2025-08-16 9.7/paper.tex
@@
   \texttt{GridIn::get\_cell\_data()}.
+\item Sparsity patterns for matrices built with \texttt{FE\_Q\_iso\_Q1} elements
+  now correctly reflect the inherited sparsity on each cell, making the usage of
+  this element usable as a preconditioner for a higher order problems (this is known
+  as low order rediscretization, see also \cite{pazner2023low}).
 \end{itemize}

### 2025-08-18 9.7/paper.tex
Minor updates.

@@
 running all copy operations sequentially limits parallel scalability (to less than 10
-threads on the left, where the copy operation is more expensive than the right),
-while using graph coloring (denoted as ``colored'') eliminates this bottleneck.
+threads on the left, where the copy operation is more expensive than
+the right), regardless of whether one uses the TBB or Taskflow. On the
+other hand, first ``coloring'' all cells to avoid conflicts in
+the copy operation (labeled as ``colored'' in the figure) eliminates this bottleneck.
 It is clearly visible that the Taskflow implementation is better or the same
@@
 \includegraphics{kokkos-vs-tbb/plots-figure2.pdf}
-\caption{Assembly times in a modified step-32 (left: temperature matrix assembly, right: temperature right-hand side assembly) comparing TBB and Taskflow with and without coloring.}
+\caption{Assembly times in a modified step-32 (left: temperature
+  matrix assembly, right: temperature right-hand side assembly)
+  comparing TBB and Taskflow with and without coloring (corresponding
+  to ``Implementation 2'' and ``Implementation 3'' in \cite{TKB16}).}
 \label{fig:tbb-vs-taskflow}

### 2025-08-18 9.7/paper.tex
Code example for magic enum.

@@
 Version 9.7 uses \texttt{magic\_enum} to provide type safe and efficient conversions between enum values and their string representations in the \texttt{ParameterHandler} class.
-
-\todo[inline]{Peter @ Luca: I think a code snippet would be great here! }
+An example usage of this class would be:
+
+\begin{c++}
+  #include <deal.II/base/parameter_handler.prm>
+
+  enum class SolverType {
+    DirectUMFPACK,
+    DirectMUMPS,
+    Iterative
+  };
+
+  SolverType solver_type = SolverType::DirectUMFPACK;
+  ParameterHandler prm;
+  prm.add_parameter("solver_type", solver_type);
+  prm.parse_input("input.prm");
+\end{c++}
+
+This allows the user to specify in the parameter file \verb|input.prm|, e.g., the type of solver to use:
+\begin{verbatim}
+  set solver_type = DirectUMFPACK
+\end{verbatim}
+
+The library also understands boolean operations between enums, allowing the user to specify, for example:
+\begin{c++}
+  #include <deal.II/base/patterns.h>
+  #include <deal.II/fe/fe_values.h>
+
+  // This is one of the deal.II enums used for the FEValues class
+  UpdateFlags flags = update_values;
+
+  // The following will output "update_values"
+  std::cout << Patterns::Tools::to_string(flags)
+            << std::endl;
+
+  // While the following is equivalent to writing 
+  // flags = update_values|update_gradients;
+  flags = Patterns::Tools::to_value<UpdateFlags>(
+    "update_values|update_gradients");
+
+  // Allowing one to use
+  ParameterHandler prm;
+  prm.add_parameter("update_flags", flags);
+  prm.parse_input("input.prm");
+\end{c++}
+
+In this case, the user can specify in the parameter file \verb|input.prm| the flags to pass to an \verb|FEValues| object: 
+\begin{verbatim}
+  set update_flags = update_values|update_gradients
+\end{verbatim}
 

### 2025-08-19 9.7/taskflow-vs-tbb/step-32-3d.cc
add .cc file

@@ -1,0 +1,3393 @@
+/* ---------------------------------------------------------------------
+ * $Id: step-32.cc 30306 2013-08-13 19:55:44Z heister $
+ *
+ * Copyright (C) 2008 - 2013, 2015 by the deal.II authors
+ *
+ * This file is part of the deal.II library.
+ *
+ * The deal.II library is free software; you can use it, redistribute
+ * it, and/or modify it under the terms of the GNU Lesser General
+ * Public License as published by the Free Software Foundation; either
+ * version 2.1 of the License, or (at your option) any later version.
+ * The full text of the license can be found in the file LICENSE at
+ * the top level of the deal.II distribution.
+ *
+ * ---------------------------------------------------------------------
+
+ *
+ * Authors: Martin Kronbichler, Uppsala University,
+ *          Wolfgang Bangerth, Texas A&M University,
+ *          Timo Heister, University of Goettingen, 2008-2011
+ */
+
+//#define implementation2
+
+// @sect3{Include files}
+
+#include <deal.II/base/quadrature_lib.h>
+#include <deal.II/base/logstream.h>
+#include <deal.II/base/function.h>
+#include <deal.II/base/utilities.h>
+#include <deal.II/base/conditional_ostream.h>
+#include <deal.II/base/work_stream.h>
+#include <deal.II/base/timer.h>
+#include <deal.II/base/parameter_handler.h>
+
+#include <deal.II/lac/full_matrix.h>
+#include <deal.II/lac/solver_bicgstab.h>
+#include <deal.II/lac/solver_cg.h>
+#include <deal.II/lac/solver_gmres.h>
+#include <deal.II/lac/affine_constraints.h>
+#include <deal.II/lac/block_sparsity_pattern.h>
+#include <deal.II/lac/trilinos_parallel_block_vector.h>
+#include <deal.II/lac/trilinos_sparse_matrix.h>
+#include <deal.II/lac/trilinos_block_sparse_matrix.h>
+#include <deal.II/lac/trilinos_precondition.h>
+#include <deal.II/lac/trilinos_solver.h>
+
+#include <deal.II/grid/tria.h>
+#include <deal.II/grid/grid_generator.h>
+#include <deal.II/grid/tria_accessor.h>
+#include <deal.II/grid/tria_iterator.h>
+#include <deal.II/grid/filtered_iterator.h>
[…]

### 2025-08-19 9.7/paper.tex
rename folder
Minor edits.

@@
 \begin{figure}[hb]
-\includegraphics{kokkos-vs-tbb/plots-figure1.pdf}\hfill
-\includegraphics{kokkos-vs-tbb/plots-figure2.pdf}
+\includegraphics{taskflow-vs-tbb/plots-figure1.pdf}\hfill
+\includegraphics{taskflow-vs-tbb/plots-figure2.pdf}
 \caption{Assembly times in a modified step-32 (left: temperature
   matrix assembly, right: temperature right-hand side assembly)
-  comparing TBB and Taskflow with and without coloring (corresponding
-  to ``Implementation 2'' and ``Implementation 3'' in \cite{TKB16}).}
+  comparing TBB and Taskflow with and without coloring. These figures
+  correspond to the middle row of \cite[Fig. 4]{TKB16}. Lines with
+  squares match ``Implementation 2'' in \cite{TKB16}; the lines with
+  circles labeled ``colored'' match ``Implementation 3'' therein.}
 \label{fig:tbb-vs-taskflow}

### 2025-08-19 9.7/paper.tex

@@
     David Wells,
+    Michał Wichrowski,
   },
@@
   {\texttt{drwells@email.unc.edu}}}
+
+  \author[13]{Michał Wichrowski}
+\affil[13]{Interdisciplinary Center for Scientific Computing, Heidelberg University, Heidelberg, Germany.
+  {\texttt{mt.wichrowsk@uw.edu.pl}}}
 
@@
   as low order rediscretization, see also \cite{pazner2023low}).
+  \item The assembly of ghost penalty stabilization matrices, which appear in CutFEM, can now be done 
+    via \texttt{TensorProductMatrixCreator::create\_1d\_ghost\_penalty\_matrix}. For Cartesian cells, the local penalty matrix can be constructed by exploiting the tensor-product structure   
+    of the basis functions. The ghost penalty operator is expressed as 
+    a Kronecker product of precomputed one-dimensional mass and penalty matrices,
+    which avoids the evaluation of high-order derivatives~\cite{wichrowski2025matrix}.
+    The new functionality also handles the necessary lexicographical renumbering 
+    of degrees of freedom across adjacent cells via \texttt{FETools::cell\_to\_face\_patch}.
 \end{itemize}
@@
 Stephan      Voss,
-Michał       Wichrowski,
+% Michał       Wichrowski,
 Simon        Wiesheier,

### 2025-08-19 9.7/paper.tex
A promoted author S.Uzunbajakau is added.

@@
 
-\author[12]{David Wells}
-\affil[12]{Department of Mathematics, University of North Carolina,
+\author[12]{Siarhei Uzunbajakau}
+\affil[12]{CEM Books,
+  Rotterdam, The Netherlands.
+  {\texttt{info@cembooks.nl}}}
+
+\author[13]{David Wells}
+\affil[13]{Department of Mathematics, University of North Carolina,
   Chapel Hill, NC 27516, USA.
@@
 
-  \author[13]{Michał Wichrowski}
-\affil[13]{Interdisciplinary Center for Scientific Computing, Heidelberg University, Heidelberg, Germany.
+  \author[14]{Michał Wichrowski}
+\affil[14]{Interdisciplinary Center for Scientific Computing, Heidelberg University, Heidelberg, Germany.
   {\texttt{mt.wichrowsk@uw.edu.pl}}}
@@
 %  Bruno        Turcksin,
+%  Siarhei      Uzunbajakau,
 %  David        Wells,
+%  Michał       Wichrowski,
 Pasquale     Africa,
@@
 Xiaochuan    Tian,
-Siarhei      Uzunbajakau,
 Mikael       Vaillant,
@@
 Stephan      Voss,
-% Michał       Wichrowski,
 Simon        Wiesheier,

### 2025-08-20 9.7/paper.tex
Help with wrapping lines and pages.

@@
   \item The assembly of ghost penalty stabilization matrices, which appear in CutFEM, can now be done 
-    via \texttt{TensorProductMatrixCreator::create\_1d\_ghost\_penalty\_matrix}. For Cartesian cells, the local penalty matrix can be constructed by exploiting the tensor-product structure   
+    via \texttt{TensorProductMatrixCreator::create\_1d\_ghost\_penalty\_matrix()}. For Cartesian cells, the local penalty matrix can be constructed by exploiting the tensor-product structure   
     of the basis functions. The ghost penalty operator is expressed as 
@@
     The new functionality also handles the necessary lexicographical renumbering 
-    of degrees of freedom across adjacent cells via \texttt{FETools::cell\_to\_face\_patch}.
+    of degrees of freedom across adjacent cells via the
+    \texttt{FETools::cell\_to\_face\_patch()} function.
 \end{itemize}
@@
 -- listing more than 120
-features and bugfixes --
+new features and bugfixes --
 contains a complete record of all changes; see \cite{changes97}.
@@
 
-\begin{figure}[hb]
+\begin{figure}[tb]
 \includegraphics{taskflow-vs-tbb/plots-figure1.pdf}\hfill

### 2025-08-20 9.7/paper.tex
Add author and funding

@@
     Wolfgang Bangerth,
+    Maximilian Bergbauer,
     Bruno Blais,
@@
 
-\author[4]{Bruno Blais}
-\affil[4]{Chemical Engineering High-performance Analysis, Optimization and Simulation (CHAOS) laboratory, Department of Chemical Engineering,
+\author[4]{Maximilian~Bergbauer}
+\affil[4]{Institute for Computational Mechanics, Technical University of
+  Munich, Boltzmannstr. 15, 85748 Garching bei München, Germany.
+  {\texttt{maximilian.bergbauer@tum.de}}}
+
+\author[5]{Bruno Blais}
+\affil[5]{Chemical Engineering High-performance Analysis, Optimization and Simulation (CHAOS) laboratory, Department of Chemical Engineering,
              Polytechnique Montréal,
@@
 
-\author[5]{Marc~Fehling}
- \affil[5]{Department of Mathematical Analysis,
+\author[6]{Marc~Fehling}
+\affil[6]{Department of Mathematical Analysis,
     Faculty of Mathematics and Physics, Charles University,
@@
 
-\author[6]{Rene~Gassm\"{o}ller}
-\affil[6]{GEOMAR Helmholtz Centre for Ocean Research Kiel, 24148 Kiel, Germany}
-
-\author[7]{Timo~Heister}
- \affil[7]{School of Mathematical and Statistical Sciences,
+\author[7]{Rene~Gassm\"{o}ller}
+\affil[7]{GEOMAR Helmholtz Centre for Ocean Research Kiel, 24148 Kiel, Germany}
+
+\author[8]{Timo~Heister}
+\affil[8]{School of Mathematical and Statistical Sciences,
    Clemson University,
@@
 
-\author[8]{Luca~Heltai}
-\affil[8]{Department of Mathematics, University of Pisa,
+\author[9]{Luca~Heltai}
+\affil[9]{Department of Mathematics, University of Pisa,
 Via Buonarroti 1/c, 56127 Pisa, Italy.}
 
- \author[9]{Martin~Kronbichler}
- \affil[9]{Faculty of Mathematics, Ruhr University Bochum,
+\author[10]{Martin~Kronbichler}
+\affil[10]{Faculty of Mathematics, Ruhr University Bochum,
    Universit\"atsstr.~150, 44780 Bochum, Germany.
@@
 
-\author[11]{Peter Munch}
- \affil[11]{Institute of Mathematics, Technical University of Berlin, Germany.
+\author[12]{Peter Munch}
+\affil[12]{Institute of Mathematics, Technical University of Berlin, Germany.
   {\texttt{muench@math.tu-berlin.de}}}
@@
 
-\author[12]{Siarhei Uzunbajakau}
-\affil[12]{CEM Books,
+\author[13]{Siarhei Uzunbajakau}
+\affil[13]{CEM Books,
   Rotterdam, The Netherlands.
@@
 
-\author[13]{David Wells}
-\affil[13]{Department of Mathematics, University of North Carolina,
+\author[14]{David Wells}
+\affil[14]{Department of Mathematics, University of North Carolina,
   Chapel Hill, NC 27516, USA.
@@
 
-  \author[14]{Michał Wichrowski}
-\affil[14]{Interdisciplinary Center for Scientific Computing, Heidelberg University, Heidelberg, Germany.
+\author[15]{Michał Wichrowski}
+\affil[15]{Interdisciplinary Center for Scientific Computing, Heidelberg University, Heidelberg, Germany.
   {\texttt{mt.wichrowsk@uw.edu.pl}}}
@@
 
+M.~Bergbauer was supported by the German Research Foundation (DFG) under the
+project ``High-Performance Cut Discontinuous Galerkin Methods for Flow
+Problems and Surface-Coupled Multiphysics Problems'' Grant Agreement
+No.~456365667.
+
 B.~Blais was supported by the National Science and Engineering Research Council of Canada (NSERC)  through the RGPIN-2020-04510 Discovery Grant and the MMIAOW Canada Research Level 2 in Computer-Assisted Design and Scale-up of Alternative Energy Vectors for Sustainable Chemical Processes.

### 2025-08-20 9.7/paper.tex

@@
     \step{95} extends \step{85} and illustrates the use of matrix-free methods to solve
-    problems with embedded boundaries using the CutFEM method~\cite{bergbauer2025high}. \step{95} was written
-    by Maximilian Bergbauer, with help by Martin Kronbichler and Peter Munch.
+    problems with embedded boundaries using the CutFEM method~\cite{bergbauer2025high}. 
+    It shows the interplay between the matrix-free infrastructure for standard quadrature rules 
+    (\texttt{FEEvaluation} / \texttt{MatrixFree}) and the infrastructure for unstructured
+     quadrature (\texttt{FEPointEvaluation} / \texttt{NonMatching::MappingInfo}) while using explicit 
+     \texttt{SIMD} vectorization. The tutorial implements both continuous and discontinuous Galerkin 
+     methods in 2D and 3D.
+    \step{95} was written by Maximilian Bergbauer, with help by Martin Kronbichler and Peter Munch.
   \item

### 2025-08-20 9.7/paper.tex
Fix a missing author in a list.

@@
     Bruno Turcksin,
+    Siarhei Uzunbajakau,
     David Wells,
-    Michał Wichrowski,
+    Michał Wichrowski
   },

### 2025-08-20 9.7/paper.tex
Address minor spacing issues.

@@
     It shows the interplay between the matrix-free infrastructure for standard quadrature rules 
-    (\texttt{FEEvaluation} / \texttt{MatrixFree}) and the infrastructure for unstructured
-     quadrature (\texttt{FEPointEvaluation} / \texttt{NonMatching::MappingInfo}) while using explicit 
+    (\texttt{FEEvaluation}/\texttt{MatrixFree}) and the infrastructure for unstructured
+     quadrature (\texttt{FEPointEvaluation}/\texttt{NonMatching::MappingInfo}) while using explicit 
      \texttt{SIMD} vectorization. The tutorial implements both continuous and discontinuous Galerkin 

### 2025-08-21 9.7/paper.tex

@@
     Peter Munch,
+    Sam Scheuerman,
     Bruno Turcksin,
@@
    Collins, CO 80523, USA.
-   \texttt{bangerth@colostate.edu}}
+   \texttt{bangerth/sam.scheuerman@colostate.edu}}
  \affil[3]{Department of Geosciences, Colorado State University, Fort
@@
   {\texttt{muench@math.tu-berlin.de}}}
+
+\author[2]{Sam Scheuerman}
 
@@
 %  Jean-Paul    Pelteret,
+%  Sam          Scheuerman,
 %  Bruno        Turcksin,
@@
 Mayank       Sabharwal,
-Sam          Scheuerman,
 Magdalena    Schreter,

### 2025-08-21 9.7/paper.tex
Final changes.

@@
 %
-\todo[inline]{Update the list above if we add subsections to section 2.}
 
@@
     It shows the interplay between the matrix-free infrastructure for standard quadrature rules
-    (\texttt{FEEvaluation}/\texttt{MatrixFree}) and the infrastructure for unstructured
-     quadrature (\texttt{FEPointEvaluation}/\texttt{NonMatching::MappingInfo}) while using explicit
+    (\texttt{FEEvaluation} and \texttt{MatrixFree}) and the infrastructure for unstructured
+     quadrature (\texttt{FEPointEvaluation} and \texttt{NonMatching::MappingInfo}) while using explicit
      \texttt{SIMD} vectorization. The tutorial implements both continuous and discontinuous Galerkin
@@
 contributed code to this release:\\
-
-\todo[inline]{Remove all who we promoted to authors of this paper from
-  the following list.}
 
@@
 
-\todo[inline]{Everyone please update as appropriate.}
 D.~Arndt and B.~Turcksin: Research sponsored by the Laboratory Directed Research and

### 2025-08-21 9.7/paper.tex
Remove from other contributors list

--- 2025-08-20 9.7/paper.tex
@@
     Bruno Turcksin,
-    Siarhei Uzunbajakau,
     David Wells,
-    Michał Wichrowski
+    Michał Wichrowski,
   },
@@
     It shows the interplay between the matrix-free infrastructure for standard quadrature rules 
-    (\texttt{FEEvaluation}/\texttt{MatrixFree}) and the infrastructure for unstructured
-     quadrature (\texttt{FEPointEvaluation}/\texttt{NonMatching::MappingInfo}) while using explicit 
+    (\texttt{FEEvaluation} / \texttt{MatrixFree}) and the infrastructure for unstructured
+     quadrature (\texttt{FEPointEvaluation} / \texttt{NonMatching::MappingInfo}) while using explicit 
      \texttt{SIMD} vectorization. The tutorial implements both continuous and discontinuous Galerkin 
@@
 %  Wolfgang     Bangerth,
+%  Maximilian   Bergbauer,
 %  Bruno        Blais,
@@
 Henry        Arhin,
-Maximilian   Bergbauer,
 Alfredo      Buttari,

### 2025-08-21 9.7/paper.tex
End with a dot

@@
 Simon        Wiesheier,
-Yi-Yung      Yang,
+Yi-Yung      Yang.
 \end{quote}

### 2025-09-01 9.7/paper.tex
small update
clarify sentence

@@
 \affil[9]{Department of Mathematics, University of Pisa,
-Via Buonarroti 1/c, 56127 Pisa, Italy.}
+Via Buonarroti 1/c, 56127 Pisa, Italy. {\texttt{luca.heltai@unipi.it}}}
 
@@
  more quickly span the space in which the solution is expected.
-\item A common operation in finite element codes is to ask whether
-  a specific shape function belongs to a particular solution
-  variable. In the past, this was implemented by using the
+\item A common operation in codes implementing finite element problems with
+  several solution variables is to ask whether
+  a specific shape function belongs to a particular
+  variable, for example the pressure in a Stokes problem. In the past, this was
+  implemented by using the
   \texttt{FiniteElement::system\_to\_component\_index()} function that
